### 2018-12-02 script/Rmd/instruction.Rmd
Upload Rmd for an instruction sheet

---
output:
  pdf_document: default
  html_document: default
  word_document: default
---

---
title: "Instruction"
author: "Atsuko Tominaga"
date: "03/12/2018"
output:
  pdf_document:
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center', fig.lp = 'AAA')
# Install and load packages
if (!require("rmarkdown")) {install.packages("rmarkdown"); require("rmarkdown")}
```

## PARTICIPANT RECRUITMENT:
- More than 10 years experience to play the piano
- Basic English skills

## BEFORE PARTICIPANT ARRIVES:
- Devices
1. Connect the audio interface (Focusrite Scarlett 6i6) to Macbook Pro.
2. Connect a headphone to the audio interface.
3. Connect the piano to the audio interface.
4. Turn on the piano, the audio interface and the screen in front of the piano.
5. Open “Focusrite Control” app ![](focusriteControl.jpg){ width=5% } and click “Load from device”.

```{r 01, out.width = '50%'}
knitr::include_graphics("01.png")
```
```{r 02, out.width = '80%'}
knitr::include_graphics("02.png")
```

6. Go to “Device Settings” and click “LINE” for Analogue 1.

```{r 03, out.width = '80%'}
knitr::include_graphics("03.png")
```

If you don’t see the images above, go to File >> Preset >> select “Analogue + Digital”.
Adjust the volume of the headphone so that it sounds naturally (Later, also ask a participant).

\pagebreak

- Max

1. Go to Desktop >> Atsuko >> ExpertPiano >> open the “mainPatch.maxpat” and the “imageDisplay.maxpat”.
2. Move the window of the “imageDisplay.maxpat” to the right (on the screen in front of the piano”) and press F for a fullscreen mode.
3. Read the instruction for the “mainPatch.maxpat" below.
    + Start a trial with the space key.
    + Stop a trial with the Z key.
    + The blue patcher is used to change an image of the screen and start a metronome manually (mainly used for a practice session and instructions).
    + The green patcher shows ongoing recordings.
    + The yellow patcher shows the name of the current file.
    + The pink patcher (hidden) reads the condition of the current participant.

```{r mainPatch, out.width = '100%'}
knitr::include_graphics("mainPatch_screen.png")
```

4. Press the "Metro" button on the blue patcher to make sure the sound of a metronome comes from the headphone, not from the computer speaker.
5. Also check whether the piano sounds come from the headphone and MIDI inputs are correctly responded on the green patcher (Pitch and Velocity should change according to key presses).
6. If the sound of the metronome and the piano does not come from the computer speaker, go to System Preference >> Sound >> Output >> select “Scarllet 6i6 USB”.
7. Click the “inst_g1” button to show a general instruction on the “imageDisplay.maxpat”.
8. Enter SubNr for the current participant.

\pagebreak

## WHEN PARTICIPANT ARRIVES:

- *Italic*: Instruction

### Introduction

*Thank you for participating in the experiment. The aim of the experiment is to investigate how people play a piece of music with musical expressions. The experiment is going to last about one hour.*

*If you have any questions or concerns, please ask the experimenter at any time about the nature of the study or the methods we are using. Also, you have the right to withdraw from the experiment at any time.*

*Before we start the experiment, please read the informed consent form carefully and fill it out if you agree to take part in the study.*

**[ Collect the informed consent form ]**

### Practice session

**[ Click “inst_g1” ]**

```{r inst_g1, out.width = '80%'}
knitr::include_graphics("inst_g1.png")
```

**[ Click “stim_n” ]**

```{r stim_n, out.width = '80%'}
knitr::include_graphics("stim_n.png")
```

**[ Click “Metro” ]**

**[ Let the participant practise the piece ]**

**[ Click "inst_g2" ]**

```{r inst_g2, out.width = '80%'}
knitr::include_graphics("inst_g2.png")
```

**[ Check whether he or she can produce the sound without pitch errors twice consecutively]**

- If he or she cannot perform the piece within **5 attempts**, he or she cannot continue the experiment.

**[ Click “Black” ]**

### Experiment

- Participants whose SubNr is odd start from the teaching condition whereas those whose SubNr is even start from the performing condition.
- Please see the order of the stimuli on the blue patcher.

#### 1) Teaching condition
##### 1st Stimulus

*You are going to perform the piece with the first musical expressions. I will show sheet music with expressive notations on the screen in front of you. Please read the notations and practise the piece.*

**[ Click 1st Stimulus ]**

```{r stim_a, out.width = '80%', fig.cap = 'Articulation'}
knitr::include_graphics("stim_a.png")
```

**[ Ask the participant to play the piece with the expressions 5 times ]**

Thank you. Do you have any questions?

**[ Click “inst_t” ]**

```{r inst_t, out.width = '80%'}
knitr::include_graphics("inst_t.png")
```

**[ 1 - 8 trials ]**

##### 2nd Stimulus

*Thank you. Now, we are moving on to the second musical expressions. Again, I will show sheet music with expressive notations on the screen in front of you. Please read the notations and practice the piece.*

**[ Click 2nd Stimuli ]**

```{r stim_d, out.width = '80%', fig.cap = 'Dynamics'}
knitr::include_graphics("stim_d.png")
```

**[ Ask the participant to play the piece with the expressions 5 times ]**

*Thank you. Do you have any questions?*

**[ Click “inst_t” ]**

**[ 1 - 8 trials ]**

#####  3rd Stimulus

*Thank you. Now, we are moving on to the third musical expressions. Again, I will show sheet music with expressive notations on the screen in front of you. Please read the notations and practice the piece.*

**[ Click 3rd Stimulus ]**

```{r stim_t, out.width = '80%', fig.cap = 'Tempo Change'}
knitr::include_graphics("stim_t.png")
```

**[ Ask the participant to play the piece with the expressions 5 times ]**

*Thank you. Do you have any questions?*

**[ Click “inst_t” ]**

**[ 1 - 8 trials ]**

*Thank you. Next, you will perform the same piece again but in a different condition.*

#### 2) Performing condition
##### 1st Stimulus

*You are going to perform the piece with the first musical expressions. I will show sheet music with expressive notations on the screen in front of you. Please read the notations and practise the piece twice.*

**[ Click 1st Stimulus ]**

**[ Ask the participant to play the piece with the expressions 5 times ]**

*Thank you. Do you have any questions?*

**[ Click “inst_p” ]**

```{r inst_p, out.width = '80%'}
knitr::include_graphics("inst_p.png")
```

**[ 1 - 8 trials ]**

##### 2nd Stimulus

*Thank you. Now, we are moving on to the second musical expressions. Again, I will show sheet music with expressive notations on the screen in front of you. Please read the notations and practice the piece.*

**[ Click 2nd Stimulus ]**

**[ Ask the participant to play the piece with the expressions 5 times ]**

*Thank you. Do you have any questions?*

**[ Click “inst_p” ]**

**[ 1 - 8 trials ]**

##### 3rd Stimulus

*Thank you. Now, we are moving on to the third musical expressions. Again, I will show sheet music with expressive notations on the screen in front of you. Please read the notations and practice the piece.*

**[ Click 3rd Stimulus ]**

**[ Ask the participant to play the piece with the expressions 5 times ]**

*Thank you. Do you have any questions?*

**[ Click “inst_p” ]**

**[ 1 - 8 trials ]**

### Questionnaire

*Thank you. This is the end of the experiment. Lastly, could you please fill in the short questionnaire?*

**[ Collect the short questionnaire ]**

*Thank you for coming today and please feel free to contact me if you have any questions about my study.*

## AFTER PARTICIPANT LEAVES:
1. Check whether data were recorded correctly. If not, please leave a memo.
2. Quit Max.
3. Turn off the piano, the audio interface and the screen in front of the piano.

### 2018-12-03 script/Rmd/instruction.Rmd
Update Rmd

@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
   word_document: default
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center', fig.lp = 'AAA')
+knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center')
 # Install and load packages
@@
 1. Connect the audio interface (Focusrite Scarlett 6i6) to Macbook Pro.
-2. Connect a headphone to the audio interface.
+2. Connect twp headphones to the audio interface.
 3. Connect the piano to the audio interface.
@@
 If you don’t see the images above, go to File >> Preset >> select “Analogue + Digital”.
-Adjust the volume of the headphone so that it sounds naturally (Later, also ask a participant).
+Adjust the volume of the headphone so that it sounds naturally.
 
@@
 
-- *Italic*: Instruction
+- *Italic*: Verbal instruction
+- **[  ]**: What to do
 
@@
 
-*Thank you for participating in the experiment. The aim of the experiment is to investigate how people play a piece of music with musical expressions. The experiment is going to last about one hour.*
+*Thank you for participating in the experiment. The aim of the experiment is to investigate how people play a piece of music expressively. The experiment is going to last about one hour.*
 
@@
 
+*Please practise until you feel that you confortably perform the piece without pitch errors.
+After practice, we will have a brief test to make sure you can perform the piece.
+Also, you don't have to memorise the piece as you can see sheet music in front of you.
+Let me know when you are ready for the brief test.*
+
 **[ Let the participant practise the piece ]**
@@
 
-*You are going to perform the piece with the first musical expressions.
-I will show sheet music with expressive notations on the screen in front of you.
-Please read the notations and practise the piece.*
+*You are now going to perform the piece in an expressive style indicated by the musical notation.
+I will show sheet music with the musical notation on the screen in front of you.
+Please read the notation and perform the piece 5 times according to the notation.*
 
 **[ Click 1st Stimulus ]**
+**[ Explain the musical notation]**
 
@@
 
-Thank you. Do you have any questions?
+*Thank you. Do you have any questions?*
 
@@
 *Thank you.
-Now, we are moving on to the second musical expressions.
-Again, I will show sheet music with expressive notations on the screen in front of you.
-Please read the notations and practice the piece.*
+Now, we are moving on to the different musical notation.
+Again, I will show sheet music with the musical notation on the screen in front of you.
+Please read the notation and perform the piece 5 times according to the notation.*
 
 **[ Click 2nd Stimuli ]**
+**[ Explain the musical notation]**
 
@@
 
-**[ Ask the participant to play the piece with the expressions 5 times ]**
+**[ Ask the participant to play the piece with the expression 5 times ]**
 
@@
 *Thank you.
-Now, we are moving on to the third musical expressions.
-Again, I will show sheet music with expressive notations on the screen in front of you.
-Please read the notations and practice the piece.*
+Now, we are moving on to the different musical notation.
+Again, I will show sheet music with the musical notation on the screen in front of you.
+Please read the notation and perform the piece 5 times according to the notation.*
 
 **[ Click 3rd Stimulus ]**
+**[ Explain the musical notation]**
 
@@
 
-*You are going to perform the piece with the first musical expressions.
-I will show sheet music with expressive notations on the screen in front of you.
-Please read the notations and practise the piece twice.*
+*You are now going to perform the piece in an expressive style indicated by the musical notation.
+I will show sheet music with the musical notation on the screen in front of you.
+Please read the notation and perform the piece 5 times according to the notation.*
 
 **[ Click 1st Stimulus ]**
+**[ Explain the musical notation]**
 
-**[ Ask the participant to play the piece with the expressions 5 times ]**
+**[ Ask the participant to play the piece with the expression 5 times ]**
 
@@
 *Thank you.
-Now, we are moving on to the second musical expressions.
-Again, I will show sheet music with expressive notations on the screen in front of you.
-Please read the notations and practice the piece.*
+Now, we are moving on to the different musical notation.
+Again, I will show sheet music with the musical notation on the screen in front of you.
+Please read the notation and perform the piece 5 times according to the notation.*
 
@@
 
-**[ Ask the participant to play the piece with the expressions 5 times ]**
+**[ Explain the musical notation]**
+
+**[ Ask the participant to play the piece with the expression 5 times ]**
 
@@
 *Thank you.
-Now, we are moving on to the third musical expressions.
-Again, I will show sheet music with expressive notations on the screen in front of you.
-Please read the notations and practice the piece.*
+Now, we are moving on to the different musical notation.
+Again, I will show sheet music with the musical notation on the screen in front of you.
+Please read the notation and perform the piece 5 times according to the notation.*
 
 **[ Click 3rd Stimulus ]**
+**[ Explain the musical notation]**
 
-**[ Ask the participant to play the piece with the expressions 5 times ]**
+**[ Ask the participant to play the piece with the expression 5 times ]**
 
@@
 
-*Thank you. This is the end of the experiment. Lastly, could you please fill in the short questionnaire?*
+*Thank you. This is the end of the experiment. Lastly, could you please fill in the questionnaire?*
 

### 2018-12-26 script/Rmd/instruction.Rmd
Delete pdf and create html

@@
   html_document: default
-  pdf_document: default
-  word_document: default
 ---
@@
 author: "Atsuko Tominaga"
-date: "03/12/2018"
+date: "26/12/2018"
 output:
-  pdf_document:
-fontsize: 12pt
+  html_document:
+    css: style.css
 ---
@@
 ```
+
+<style type="text/css">
+/* Whole document: */
+body{
+  font-family: Helvetica;
+  font-size: 14pt;
+}
+<!-- /* Headers */ -->
+<!-- h1,h2,h3,h4,h5,h6{ -->
+<!--   font-size: 24pt; -->
+<!-- } -->
+</style>
 
@@
 
-```{r 01, out.width = '50%'}
-knitr::include_graphics("01.png")
-```
-```{r 02, out.width = '80%'}
-knitr::include_graphics("02.png")
-```
+![](01.png){ width=50% }
+![](02.png){ width=80% }
 
@@
 
-```{r 03, out.width = '80%'}
-knitr::include_graphics("03.png")
-```
+![](03.png){ width=80% }
 
@@
 
-```{r mainPatch, out.width = '100%'}
-knitr::include_graphics("mainPatch_screen.png")
-```
+![](mainPatch_screen.png)
 
@@
 
-```{r inst_g1, out.width = '80%'}
-knitr::include_graphics("inst_g1.png")
-```
+![](inst_g1.png){ width=80% }
 
@@
 
-```{r stim_n, out.width = '80%'}
-knitr::include_graphics("stim_n.png")
-```
+![](stim_n.png){ width=80% }
 
@@
 
-```{r inst_g2, out.width = '80%'}
-knitr::include_graphics("inst_g2.png")
-```
+![](inst_g2.png){ width=80% }
 
@@
 #### 1) Teaching condition
-##### 1st Stimulus
+> 1st Stimulus
 
@@
 
-```{r stim_a, out.width = '80%', fig.cap = 'Articulation'}
-knitr::include_graphics("stim_a.png")
-```
+![](stim_a.png){ width=80% }
 
@@
 
-```{r inst_t, out.width = '80%'}
-knitr::include_graphics("inst_t.png")
-```
+![](inst_t.png){ width=80% }
 
@@
 
-##### 2nd Stimulus
+> 2nd Stimulus
 
@@
 
-```{r stim_d, out.width = '80%', fig.cap = 'Dynamics'}
-knitr::include_graphics("stim_d.png")
-```
+![](stim_d.png){ width=80% }
 
@@
 
-#####  3rd Stimulus
+> 3rd Stimulus
 
@@
 
-```{r stim_t, out.width = '80%', fig.cap = 'Tempo Change'}
-knitr::include_graphics("stim_t.png")
-```
+![](stim_t.png){ width=80% }
 
@@
 #### 2) Performing condition
-##### 1st Stimulus
+> 1st Stimulus
 
@@
 
-```{r inst_p, out.width = '80%'}
-knitr::include_graphics("inst_p.png")
-```
+![](inst_p.png){ width=80% }
 
@@
 
-##### 3rd Stimulus
+> 3rd Stimulus
 

### 2018-12-26 script/Rmd/instruction.Rmd
Rewrite for a html file

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center')
+knitr::opts_chunk$set(echo = F, fig.align = 'center')
 # Install and load packages
@@
 }
-<!-- /* Headers */ -->
-<!-- h1,h2,h3,h4,h5,h6{ -->
-<!--   font-size: 24pt; -->
-<!-- } -->
 </style>
@@
 4. Turn on the piano, the audio interface and the screen in front of the piano.
-5. Open “Focusrite Control” app ![](focusriteControl.jpg){ width=5% } and click “Load from device”.
+5. Open “Focusrite Control” app ![](./image/focusriteControl.jpg){ width=5% } and click “Load from device”.
 
-![](01.png){ width=50% }
-![](02.png){ width=80% }
+![](./image/01.png){ width=50% }
+![](./image/02.png){ width=80% }
 
@@
 
-![](03.png){ width=80% }
+![](./image/03.png){ width=80% }
 
@@
 
-\pagebreak
-
 - Max
 
-1. Go to Desktop >> Atsuko >> ExpertPiano >> open the “mainPatch.maxpat” and the “imageDisplay.maxpat”.
+1. Go to Desktop >> Atsuko >> expertpiano >> open the “mainPatch.maxpat” and the “imageDisplay.maxpat”.
 2. Move the window of the “imageDisplay.maxpat” to the right (on the screen in front of the piano”) and press F for a fullscreen mode.
@@
 
-![](mainPatch_screen.png)
+![](./image/mainPatch_screen.png)
 
@@
 8. Enter SubNr for the current participant.
-
-\pagebreak
 
@@
 
-![](inst_g1.png){ width=80% }
+![](./image/inst_g1.png){ width=80% }
 
@@
 
-![](stim_n.png){ width=80% }
+![](./image/stim_n.png){ width=80% }
 
@@
 
-![](inst_g2.png){ width=80% }
+![](./image/inst_g2.png){ width=80% }
 
@@
 
-![](stim_a.png){ width=80% }
+![](./image/stim_a.png){ width=80% }
 
@@
 
-![](inst_t.png){ width=80% }
+![](./image/inst_t.png){ width=80% }
 
@@
 
-![](stim_d.png){ width=80% }
+![](./image/stim_d.png){ width=80% }
 
@@
 
-![](stim_t.png){ width=80% }
+![](./image/stim_t.png){ width=80% }
 
@@
 
-![](inst_p.png){ width=80% }
+![](./image/inst_p.png){ width=80% }
 

### 2019-01-13 script/Rmd/instruction.Rmd
Update instruction

@@
 author: "Atsuko Tominaga"
-date: "26/12/2018"
+date: "13/01/2019"
 output:
@@
 }
+span{
+  color:red
+}
 </style>
@@
 ## PARTICIPANT RECRUITMENT:
+
 - More than 10 years experience to play the piano
@@
 1. Connect the audio interface (Focusrite Scarlett 6i6) to Macbook Pro.
-2. Connect twp headphones to the audio interface.
+2. Connect two headphones to the audio interface.
 3. Connect the piano to the audio interface.
@@
 3. Read the instruction for the “mainPatch.maxpat" below.
-    + Start a trial with the space key.
-    + Stop a trial with the Z key.
-    + The blue patcher is used to change an image of the screen and start a metronome manually (mainly used for a practice session and instructions).
-    + The green patcher shows ongoing recordings.
-    + The yellow patcher shows the name of the current file.
-    + The pink patcher (hidden) reads the condition of the current participant.
-
-![](./image/mainPatch_screen.png)
+    + Start a trial with the A key.
+    + Stop a trial with the D key.
+    + If you want to record the same trial again, use the reset button.
 
@@
 ### Introduction
-
-*Thank you for participating in the experiment. The aim of the experiment is to investigate how people play a piece of music expressively. The experiment is going to last about one hour.*
-
+<span>
+*Thank you for participating in the experiment. The aim of the experiment is to investigate how people play a piece of music expressively. The experiment is going to last about one hour.*<br>
 *If you have any questions or concerns, please ask the experimenter at any time about the nature of the study or the methods we are using.
-Also, you have the right to withdraw from the experiment at any time.*
-
+Also, you have the right to withdraw from the experiment at any time.*<br>
 *Before we start the experiment, please read the informed consent form carefully and fill it out if you agree to take part in the study.*
+</span>
 
@@
 
-**[ Click “inst_g1” ]**
+**[ Click “inst_1” ]**
 
-![](./image/inst_g1.png){ width=80% }
+![](./image/inst_1.png){ width=80% }
 
@@
 
-![](./image/stim_n.png){ width=80% }
+![](./image/prac_stim_n.png){ width=80% }
 
@@
 
+<span>
 *Please practise until you feel that you confortably perform the piece without pitch errors.
@@
 Let me know when you are ready for the brief test.*
+</span>
 
@@
 
-**[ Click "inst_g2" ]**
+**[ Click "inst_2" ]**
 
-![](./image/inst_g2.png){ width=80% }
+![](./image/inst_2.png){ width=80% }
 
@@
 
+<span>
 *You are now going to perform the piece in an expressive style indicated by the musical notation.
 I will show sheet music with the musical notation on the screen in front of you.
-Please read the notation and perform the piece 5 times according to the notation.*
+Please read the notation and if you have any question, please ask me now (Also explain the main information of the expression).
+If you don't have any question, I will give you the sound of the metronome and please perform the piece according to the notation.*
+</span>
+
+**[ Open practicePatch ]**
 
 **[ Click 1st Stimulus ]**
+
 **[ Explain the musical notation]**
 
-![](./image/stim_a.png){ width=80% }
+![](./image/prac_stim_a.png){ width=80% }
 
@@
 
+<span>
 *Thank you. Do you have any questions?*
+</span>
 
@@
 
+<span>
 *Thank you.
@@
 Please read the notation and perform the piece 5 times according to the notation.*
+</span>
 
 **[ Click 2nd Stimuli ]**
+
 **[ Explain the musical notation]**
 
-![](./image/stim_d.png){ width=80% }
+![](./image/prac_stim_d.png){ width=80% }
 
@@
 
-> 3rd Stimulus
-
-*Thank you.
-Now, we are moving on to the different musical notation.
-Again, I will show sheet music with the musical notation on the screen in front of you.
-Please read the notation and perform the piece 5 times according to the notation.*
-
-**[ Click 3rd Stimulus ]**
-**[ Explain the musical notation]**
-
-![](./image/stim_t.png){ width=80% }
-
-**[ Ask the participant to play the piece with the expression 5 times ]**
-
-*Thank you. Do you have any questions?*
-
-**[ Click “inst_t” ]**
-
-**[ 1 - 8 trials ]**
-
+<span>
 *Thank you. Next, you will perform the same piece again but in a different condition.*
+</span>
 
@@
 
+<span>
 *You are now going to perform the piece in an expressive style indicated by the musical notation.
@@
 Please read the notation and perform the piece 5 times according to the notation.*
+</span>
 
 **[ Click 1st Stimulus ]**
+
 **[ Explain the musical notation]**
@@
 Please read the notation and perform the piece 5 times according to the notation.*
+</span>
 
@@
 
+<span>
 *Thank you. Do you have any questions?*
-
-**[ Click “inst_p” ]**
-
-**[ 1 - 8 trials ]**
-
-> 3rd Stimulus
-
-*Thank you.
-Now, we are moving on to the different musical notation.
-Again, I will show sheet music with the musical notation on the screen in front of you.
-Please read the notation and perform the piece 5 times according to the notation.*
-
-**[ Click 3rd Stimulus ]**
-**[ Explain the musical notation]**
-
-**[ Ask the participant to play the piece with the expression 5 times ]**
-
-*Thank you. Do you have any questions?*
+</span>
 
@@
 
+<span>
 *Thank you. This is the end of the experiment. Lastly, could you please fill in the questionnaire?*
+</span>
 
@@
 
+<span>
 *Thank you for coming today and please feel free to contact me if you have any questions about my study.*
+</span>
 

### 2019-01-13 script/Rmd/instruction.Rmd
Update instruction

@@
 - **[  ]**: What to do
+- Please use a checklist to make sure you don't forget each procedure.
 
@@
 ### Practice session
+
+**[ Open practicePatch ]**
 
@@
 
+**[ Click "stim_n"]**
+
 - If he or she cannot perform the piece within **5 attempts**, he or she cannot continue the experiment.
@@
 - Participants whose SubNr is odd start from the teaching condition whereas those whose SubNr is even start from the performing condition.
-- Please see the order of the stimuli on the blue patcher.
+- Please see the order of the stimuli on the practicePatcher.
 
@@
 
-**[ Open practicePatch ]**
-
 **[ Click 1st Stimulus ]**
@@
 
-**[ Ask the participant to play the piece with the expressions 5 times ]**
+**[ Check the participant to play the piece expressively without pitch errors twice consecutively ]**
 
@@
 <span>
-*Thank you.
-Now, we are moving on to the different musical notation.
-Again, I will show sheet music with the musical notation on the screen in front of you.
-Please read the notation and perform the piece 5 times according to the notation.*
+*Thank you. Now, we are moving on to the different musical notation. Again, I will show sheet music with the musical notation on the screen in front of you. Please read the notation.*
 </span>
@@
 
-**[ Ask the participant to play the piece with the expression 5 times ]**
+**[ Check the participant to play the piece expressively without pitch errors twice consecutively ]**
 
@@
 I will show sheet music with the musical notation on the screen in front of you.
-Please read the notation and perform the piece 5 times according to the notation.*
+Please read the notation.*
 </span>
@@
 ## AFTER PARTICIPANT LEAVES:
-1. Check whether data were recorded correctly. If not, please leave a memo.
+1. Check whether data were recorded correctly. If not, leave a memo on the checklist.
 2. Quit Max.

### 2019-01-14 script/Rmd/instruction.Rmd
Update instruction

@@
 author: "Atsuko Tominaga"
-date: "13/01/2019"
+date: "14/01/2019"
 output:
@@
 
-1. Go to Desktop >> Atsuko >> expertpiano >> open the “mainPatch.maxpat” and the “imageDisplay.maxpat”.
+1. Go to Desktop >> Max alias >> open the "mainPatch.maxpat", the “practicePatch.maxpat” and the “imageDisplay.maxpat”.
 2. Move the window of the “imageDisplay.maxpat” to the right (on the screen in front of the piano”) and press F for a fullscreen mode.
-3. Read the instruction for the “mainPatch.maxpat" below.
-    + Start a trial with the A key.
-    + Stop a trial with the D key.
-    + If you want to record the same trial again, use the reset button.
-
-4. Press the "Metro" button on the blue patcher to make sure the sound of a metronome comes from the headphone, not from the computer speaker.
-5. Also check whether the piano sounds come from the headphone and MIDI inputs are correctly responded on the green patcher (Pitch and Velocity should change according to key presses).
+3. Read the instruction on the “mainPatch.maxpat" and the "practicePatch.maxpat". <span>**Both patchers cannot be opende at the same time.**</span>
+4. Press the "Metro" button on the "blue patcher"practicePatcher" to make sure the sound of a metronome comes from the headphone, not from the computer speaker.
+5. Also check whether the piano sounds come from the headphone and MIDI inputs are correctly responded (Pitch and Velocity should change according to key presses on MIDIRecording subpatchers).
 6. If the sound of the metronome and the piano does not come from the computer speaker, go to System Preference >> Sound >> Output >> select “Scarllet 6i6 USB”.
-7. Click the “inst_g1” button to show a general instruction on the “imageDisplay.maxpat”.
+7. Click the “inst_1” button to show a general instruction on the “imageDisplay.maxpat”.
 8. Enter SubNr for the current participant.
@@
 Also, you have the right to withdraw from the experiment at any time.*<br>
-*Before we start the experiment, please read the informed consent form carefully and fill it out if you agree to take part in the study.*
+*Before we start the experiment, please read the information sheet carefully and fill it out if you agree to take part in the study.*
 </span>
@@
 
-**[ Click “Black” ]**
-
 ### Experiment
@@
 - Participants whose SubNr is odd start from the teaching condition whereas those whose SubNr is even start from the performing condition.
-- Please see the order of the stimuli on the practicePatcher.
+- Please see the order of the stimuli (i.e., starting from articulation or dynamics) on the practicePatcher.
 
@@
 I will show sheet music with the musical notation on the screen in front of you.
-Please read the notation and if you have any question, please ask me now (Also explain the main information of the expression).
+Please read the notation and if you have any question, please ask the experimenter now.
 If you don't have any question, I will give you the sound of the metronome and please perform the piece according to the notation.*
@@
 **[ Explain the musical notation]**
+
+Definition (based on Wikipedia)
+
+- Legato: musical notes are played smoothly and connected
+- Staccato: musical notes have shortened duration and are separated from each other
+- Forte: louder sound
+- Piano: quieter sound
 

### 2019-01-18 script/Rmd/instruction.Rmd
Update the instruction 180119

@@
 author: "Atsuko Tominaga"
-date: "14/01/2019"
+date: "18/01/2019"
 output:
@@
 2. Move the window of the “imageDisplay.maxpat” to the right (on the screen in front of the piano”) and press F for a fullscreen mode.
-3. Read the instruction on the “mainPatch.maxpat" and the "practicePatch.maxpat". <span>**Both patchers cannot be opende at the same time.**</span>
-4. Press the "Metro" button on the "blue patcher"practicePatcher" to make sure the sound of a metronome comes from the headphone, not from the computer speaker.
+3. Read the instruction on the “mainPatch.maxpat" and the "practicePatch.maxpat". <span>**Both patchers cannot be opened at the same time.**</span>
+4.
+Press the "Metro" button on the "blue patcher"practicePatcher" to make sure the sound of a metronome comes from the headphone, not from the computer speaker.
+<span>**Please do not skip this procedurea otherwise sometimes the metronome does not work properly only for the first time.** # added on 18/01/19</span>
 5. Also check whether the piano sounds come from the headphone and MIDI inputs are correctly responded (Pitch and Velocity should change according to key presses on MIDIRecording subpatchers).
 6. If the sound of the metronome and the piano does not come from the computer speaker, go to System Preference >> Sound >> Output >> select “Scarllet 6i6 USB”.
-7. Click the “inst_1” button to show a general instruction on the “imageDisplay.maxpat”.
-8. Enter SubNr for the current participant.
+7. If you cannot still solve the problem above, please quit Max and relaunch it again. <span># added on 18/01/19</span>
+8. Click the “inst_1” button to show a general instruction on the “imageDisplay.maxpat”.
+9. Enter SubNr for the current participant.
 
@@
 3. Turn off the piano, the audio interface and the screen in front of the piano.
+
+## Additional instructions after the pilot has started (18/01/19 Updated)
+- If participants ask you whether they should keep the tempo or not, please let them know that they can play the piece as they like, but keep in mind the metronome tells you the original tempo of this piece.
+- If participants ask anything related to detailed information about each skill (e.g., forte should be )
+- If participants start playing before the metronome beats, please record the same trial again.

### 2019-01-22 script/Rmd/instruction.Rmd
Update the instruction sheet

@@
 - If participants ask you whether they should keep the tempo or not, please let them know that they can play the piece as they like, but keep in mind the metronome tells you the original tempo of this piece.
-- If participants ask anything related to detailed information about each skill (e.g., forte should be )
+- If participants ask anything related to detailed information about each skill (e.g., deeper legato? normal/subito forte?), just ask them to play as they like.
 - If participants start playing before the metronome beats, please record the same trial again.
+
+## Memo for patchers (20/01/19 Updated)
+- Change variable names for data storage on practice sub-patchers (data / data_p)

### 2019-01-24 script/Rmd/instruction.Rmd
Update instrution 24/01

@@
 author: "Atsuko Tominaga"
-date: "18/01/2019"
+date: "24/01/2019"
 output:
@@
 Press the "Metro" button on the "blue patcher"practicePatcher" to make sure the sound of a metronome comes from the headphone, not from the computer speaker.
-<span>**Please do not skip this procedurea otherwise sometimes the metronome does not work properly only for the first time.** # added on 18/01/19</span>
+<span>~~**Please do not skip this procedurea otherwise sometimes the metronome does not work properly only for the first time.** # added on 18/01/19~~</span>
 5. Also check whether the piano sounds come from the headphone and MIDI inputs are correctly responded (Pitch and Velocity should change according to key presses on MIDIRecording subpatchers).
@@
 9. Enter SubNr for the current participant.
+
+### Practice Patch <span># added on 24/01/19</span>
+![](./image/praPatch.png)
+
+### Main Patch <span># added on 24/01/19</span>
+![](./image/mainPatch.png)
+
+### Experimental Design <span># added on 24/01/19</span>
+|Condition 1 (teaching / performing)|
+|-----------------------------------|
+|1st Stimulus (articulation / dynamics) - Block 1|
+|2nd Stimulus (dynamics / articulation) - Block 2|
+
+|Condition 2 (performing / teaching)|
+|-----------------------------------|
+|1st Stimulus (articulation / dynamics) - Block 3|
+|2nd Stimulus (dynamics / articulation) - Block 4|
+
+* The order of the conditions is counterbalanced across participants.
+* The order of the stimuli is counterbalanced across participants, but it is fixed within one condition (e.g., if a participant start with articulation and move on dynamics in the first condition, he/she will perform in the same order.)
 
@@
 
+**[ Open practicePatch ]**
+
 **[ Click 1st Stimulus ]**
@@
 
+**[ Close practicePatch ]**
+
+**[ Open mainPatch ]**
+
 **[ 1 - 8 trials ]**
@@
 
+**[ Open practicePatch ]**
+
 **[ Click 2nd Stimuli ]**
 
-**[ Explain the musical notation]**
+**[ Explain the musical notation ]**
 
@@
 **[ Click “inst_t” ]**
+
+**[ Close practicePatch ]**
+
+**[ Open mainPatch ]**
 
@@
 </span>
+
+**[ Open practicePatch ]**
 
@@
 **[ Click “inst_p” ]**
+
+**[ Close practicePatch ]**
+
+**[ Open mainPatch ]**
 
@@
 
-## Memo for patchers (20/01/19 Updated)
-- Change variable names for data storage on practice sub-patchers (data / data_p)
+## Memo for patchers (24/01/19 Updated)
+- Change variable names for data storage on practice sub-patchers (data / data_p) <span># added on 20/01/19</span>
+- Add a tone counter to detect error (if the number of tones is not equal to 150 after performing, there will be an error in performance) <span># added on 24/01/19</span>

### 2019-09-25 script/RNotebook/analysis.Rmd
Add Notebook
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of all analyses for "The Sound of Teaching Music 1".

##### R Notebook:
- Execute a chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.
- Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.
- When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 
- The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

# Outline

```{r setup, include = FALSE}
# Install and load required packages
# data manipulation
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("ez")) {install.packages("ez"); require("ez")}
```

## Interonset intervals (IOIs)
```{r extract, include = FALSE}
df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
df_ioi$SubNr <- as.factor(df_ioi$SubNr)

# Include only articulation
df_ioi_art <- df_ioi %>% dplyr::filter(Skill == "articulation")
# Include only dynamics
df_ioi_dyn <- df_ioi %>% dplyr::filter(Skill == "dynamics")
```

### Articulation
#### 1. Aggregated data for each individual
```{r ioi_art, echo = FALSE}
# Aggregate data
ioi_art <- aggregate(IOI~SubNr*Condition*Skill, data = df_ioi_art,
                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
ioi_art <- cbind(ioi_art[,1:3], as.data.frame(ioi_art[,4]))
# Change colnames
colnames(ioi_art) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
print(ioi_art)
```

#### Plot
- Error bars show the SD of the mean
```{r ioi_art_plot,  echo = FALSE}
# Graph
p_ioi_art <- ggplot(data = ioi_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "Mean IOI (ms)") +
  theme_classic() +
  theme(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"))
plot(p_ioi_art)
```

#### 2. Aggregated data for each condition
```{r ioi_art_all, echo = FALSE}
# Group mean
ioi_art_all <- aggregate(Mean~Condition*Skill, data = ioi_art,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_all <- cbind(ioi_art_all[,1:2], as.data.frame(ioi_art_all[,3]))
# Change colnames
colnames(ioi_art_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
print(ioi_art_all)
```

#### Plot
- Error bars show the SEM of the mean
```{r ioi_art_all_plot, echo = FALSE}
p_ioi_art_all <- ggplot(data = ioi_art_all, aes(x = Skill, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "Mean IOI (ms)") + coord_cartesian(ylim = c(100, 210)) +
  theme_classic() +
  theme(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"))
p_ioi_art_all
```

#### 3. Phrase boundaries for each individual 
```{r ioi_art_ch, echo = FALSE}
# Define phrase boundaries
df_ioi_art$Boundary <- "No"
df_ioi_art$Boundary[df_ioi_art$Interval == 8 | df_ioi_art$Interval == 16 | df_ioi_art$Interval == 24 | df_ioi_art$Interval == 41 | df_ioi_art$Interval == 49 | df_ioi_art$Interval == 57] <- "Yes"

# For each individual
ioi_art_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_art,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_ch <- cbind(ioi_art_ch[,1:4], as.data.frame(ioi_art_ch[,5]))
# Change colnames
colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
print(ioi_art_ch)
```

##### Plot
- Error bars show the SD of the mean
```{r ioi_art_ch_plot, echo = FALSE}
p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "Mean IOI (ms)") + coord_cartesian(ylim = c(100, 280)) +
  theme_classic() +
  theme(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"))
p_ioi_art_ch
```

### Phrase boundaries for each condition
```{r ioi_art_ch_all, echo = FALSE}
# Group mean
ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_art_ch,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_ch_all <- cbind(ioi_art_ch_all[,1:3], as.data.frame(ioi_art_ch_all[,4]))
# Change colnames
colnames(ioi_art_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
print(ioi_art_ch_all)
```

#### Plot
- Error bars show the SEM of the mean
```{r ioi_art_ch_all_plot, echo = FALSE}
p_ioi_art_ch_all <- ggplot(data = ioi_art_ch_all, aes(x = Skill, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  facet_grid(. ~ Boundary) +
  labs(y = "Mean IOI (ms)") + coord_cartesian(ylim = c(100, 250)) +
  theme_classic() +
  theme(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"))
p_ioi_art_ch_all
```

## Key-Overlap time/ratio (KOT/KOR)
```{r}

```

## Key Velocity (KV)

### 2019-09-29 script/RNotebook/analysis.Rmd
Add IOI analyses
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r setup, include = FALSE}
-# Install and load required packages
+# packages
 # data manipulation
@@
 if (!require("ez")) {install.packages("ez"); require("ez")}
+
+# ggplots
+theme_set(theme_classic())
+theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "top")
 ```
 
-## Interonset intervals (IOIs)
+# Interonset intervals (IOIs)
 ```{r extract, include = FALSE}
@@
 
-### Articulation
-#### 1. Aggregated data for each individual
+## Articulation
+### 1. All - individual
 ```{r ioi_art, echo = FALSE}
-# Aggregate data
+# For each individual
 ioi_art <- aggregate(IOI~SubNr*Condition*Skill, data = df_ioi_art,
@@
 #### Plot
-- Error bars show the SD of the mean
-```{r ioi_art_plot,  echo = FALSE}
-# Graph
+##### Bar chart
+- Error bars show SD of the mean
+- A dashed line represents the tempo given by a metronome (80bpm)
+```{r ioi_art_bar,  echo = FALSE}
 p_ioi_art <- ggplot(data = ioi_art, aes(x = SubNr, y = Mean, fill = Condition)) +
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "Mean IOI (ms)") +
-  theme_classic() +
-  theme(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"))
+  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  labs(y = "IOI (ms)")
 plot(p_ioi_art)
@@
 
-#### 2. Aggregated data for each condition
+### 2. All - group
 ```{r ioi_art_all, echo = FALSE}
@@
 #### Plot
-- Error bars show the SEM of the mean
-```{r ioi_art_all_plot, echo = FALSE}
-p_ioi_art_all <- ggplot(data = ioi_art_all, aes(x = Skill, y = Mean, fill = Condition)) +
+##### Bar chart
+- Error bars show SEM of the mean
+- A dashed line represents the tempo given by a metronome (80bpm)
+```{r ioi_art_all_bar, echo = FALSE}
+p_ioi_art_all <- ggplot(data = ioi_art_all, aes(x = Condition, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "Mean IOI (ms)") + coord_cartesian(ylim = c(100, 210)) +
-  theme_classic() +
-  theme(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"))
+  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  labs(y = "IOI (ms)") + coord_cartesian(ylim = c(120, 210)) +
+  theme(legend.position = "None")
 p_ioi_art_all
@@
 
-#### 3. Phrase boundaries for each individual 
+##### Boxplot
+- A dashed line represents the tempo given by a metronome (80bpm)
+```{r ioi_art_all_box, echo = FALSE}
+p_ioi_art_all_box <- ggplot(data = ioi_art, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
+  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  labs(y = "IOI (ms)") + #coord_cartesian(ylim = c(100, 210)) +
+  theme(legend.position = "None")
+p_ioi_art_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality1, echo = FALSE}
+# Normality check
+ioi_art_norm <- by(ioi_art$Mean, list(ioi_art$Condition), shapiro.test)
+print(ioi_art_norm)
+```
+
+##### Two-way ANOVA (ez)
+```{r stats1, echo = FALSE, message = FALSE}
+ioi_art_aov <- ezANOVA(
+  data = df_ioi_art
+  , dv = .(IOI)
+  , wid = .(SubNr)
+  , within = .(Condition)
+  , type = 3
+  , detailed = TRUE
+)
+print(ioi_art_aov)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats2, echo = FALSE, message = FALSE}
+ioi_art_aov_2 <- aov(Mean ~ Condition + Error(SubNr/Condition), data = ioi_art)
+print(summary(ioi_art_aov_2))
+```
+
+### 3. Phrase boundaries - individual
 ```{r ioi_art_ch, echo = FALSE}
@@
 
-##### Plot
-- Error bars show the SD of the mean
-```{r ioi_art_ch_plot, echo = FALSE}
+#### Plot
+##### Bar chart
+- Error bars show SD of the mean
+- A dashed line represents the tempo given by a metronome (80bpm)
+```{r ioi_art_ch_bar, echo = FALSE}
 p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "Mean IOI (ms)") + coord_cartesian(ylim = c(100, 280)) +
-  theme_classic() +
-  theme(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"))
+  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  labs(y = "IOI (ms)") + coord_cartesian(ylim = c(100, 280))
 p_ioi_art_ch
@@
 
-### Phrase boundaries for each condition
+### 4. Phrase boundaries - group
 ```{r ioi_art_ch_all, echo = FALSE}
@@
 #### Plot
-- Error bars show the SEM of the mean
-```{r ioi_art_ch_all_plot, echo = FALSE}
-p_ioi_art_ch_all <- ggplot(data = ioi_art_ch_all, aes(x = Skill, y = Mean, fill = Condition)) +
+- Error bars show SEM of the mean
+- A dashed line represents the tempo given by a metronome (80bpm)
+```{r ioi_art_ch_all_bar, echo = FALSE}
+p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  facet_grid(. ~ Boundary) +
-  labs(y = "Mean IOI (ms)") + coord_cartesian(ylim = c(100, 250)) +
-  theme_classic() +
-  theme(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"))
-p_ioi_art_ch_all
+  #annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #facet_grid(. ~ Boundary) +
+  labs(y = "IOI (ms)") + coord_cartesian(ylim = c(100, 250))
+p_ioi_art_ch_all_bar
 ```
 
-## Key-Overlap time/ratio (KOT/KOR)
-```{r}
-
+##### Boxplot
+- A dashed line represents the tempo given by a metronome (80bpm)
+```{r ioi_art_ch_all_box, echo = FALSE}
+p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y" ,stackdir = "center", dotsize = 1, position=position_dodge(0.75)) +
+  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  labs(y = "IOI (ms)")
+p_ioi_art_ch_box
 ```
 
-## Key Velocity (KV)
+#### Statistics
+##### Normality check
+```{r norrmality2, echo = FALSE}
+ioi_art_ch_norm <- by(ioi_art_ch$Mean, list(ioi_art_ch$Condition, ioi_art_ch$Boundary), shapiro.test)
+print(ioi_art_ch_norm)
+```
+
+##### Two-way ANOVA (ez)
+```{r stats3, echo = FALSE}
+ioi_art_ch_aov <- ezANOVA(
+  data = df_ioi_art
+  , dv = .(IOI)
+  , wid = .(SubNr)
+  , within = .(Condition, Boundary)
+  , type = 3
+  , detailed = TRUE
+)
+print(ioi_art_ch_aov)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats4, echo = FALSE}
+ioi_art_ch_aov_2 <- with(ioi_art_ch,
+  aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary))))
+print(summary(ioi_art_ch_aov_2))
+```
+
+### 5. Trial by trial - individual
+```{r ioi_art_trial, echo = FALSE}
+# For each individual
+ioi_art_trial <- aggregate(IOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_art,
+                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+ioi_art_trial <- cbind(ioi_art_trial[,1:4], as.data.frame(ioi_art_trial[,5]))
+# Change colnames
+colnames(ioi_art_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean", "SD")
+print(ioi_art_trial)
+```
+
+#### Plot
+##### Line graph
+- A dashed line represents the tempo given by a metronome (80bpm)
+```{r ioi_art_trial_line, echo = FALSE, fig.height = 8}
+p_ioi_art_trial_line <- ggplot(data = ioi_art_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  facet_wrap(SubNr ~ .) + 
+  labs(x = "Trial", y = "IOI (ms)") + scale_x_continuous(breaks=seq(1,8,1))
+p_ioi_art_trial_line
+```
+
+### 6. Trial by trial - group
+```{r ioi_art_trial_all, echo = FALSE}
+# Group mean
+ioi_art_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_art_trial,
+                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_art_trial_all <- cbind(ioi_art_trial_all[,1:3], as.data.frame(ioi_art_trial_all[,4]))
+# Change colnames
+colnames(ioi_art_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
+print(ioi_art_trial_all)
+```
+
+#### Plot
+##### Line graph
+```{r ioi_art_trial_all_line, echo = FALSE}
+p_ioi_art_trial_line <- ggplot(data = ioi_art_trial_all, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.05)) + 
+  labs(x = "Trial", y = "IOI (ms)") + scale_x_continuous(breaks=seq(1,8,1))
+p_ioi_art_trial_line
+```
+
+### 7. Variability all - individual
+
+### 8. Variability all - group
+
+### 9. Variability trial-by-trial - individual
+
+### 10. Variability trial-by-trial - group
+
+### Summary (IOI:articulation)
+#### 1. All - individual
+Some participants exhibited a strong tendency for slow demonstration while teaching.
+
+#### 2. All - group
+Participants played slower while teaching.
+
+#### 3. Phrase boundaries - individual
+Some participants played slower particularly at phrase boundaries.
+
+#### 4. Phrase boundaries - group
+Participants played slower while teaching / at phrase boundaries. However, there was no interaction.
+
+#### 5. Trial by trial - individual
+#### 6. Trial by trial - group
+#### 7. Variability all - individual
+#### 8. Variability all - group
+#### 9. Variability trial-by-trial - individual
+#### 10. Variability trial-by-trial - group

### 2019-09-30 script/RNotebook/ioi.Rmd
Rename the file
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of all analyses for "The Sound of Teaching Music 1".

# Outline - Interonset intervals (IOIs)
### [Experimental Design](#design)

### [Articulation](#ioi-articulation)

- Individual results
   + [All](#art_ind_all)
   + [Phrase bondaries](#art_ind_phrase)
   + [Trial by trial](#art_ind_trial)
   + [Variability all](#art_ind_var_all)
   + [Variability trial-by-trial](#art_ind_var_trial)

- Group results and stats (if available)
   + [All](#art_group_all)
   + [Phrase bondaries](#art_group_phrase)
   + [Trial by trial](#art_group_trial)
   + [Variability all](#art_group_var_all)
   + [Variability trial-by-trial](#art_group_var_trial)

- [Summary](#art_summary)

### [Dynamics](#ioi-dynamics)

- Individual results
   + [All](#dyn_ind_all)
   + [Phrase bondaries](#dyn_ind_phrase)
   + [Trial by trial](#dyn_ind_trial)
   + [Variability all](#dyn_ind_var_all)
   + [Variability trial-by-trial](#dyn_ind_var_trial)

- Group results and stats (if available)
   + [All](#dyn_group_all)
   + [Phrase bondaries](#dyn_group_phrase)
   + [Trial by trial](#dyn_group_trial)
   + [Variability all](#dyn_group_var_all)
   + [Variability trial-by-trial](#dyn_group_var_trial)

- [Summary](#dyn_summary)

## Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Skill (2 levels: articulation vs. dynamics)
- Dependent variables: Interonset intervals, variability (SD/mean IOI)

### Main predictions
Participants will play slower while teaching regardless of the skills.

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
df_ioi$SubNr <- as.factor(df_ioi$SubNr)

# Include only articulation
df_ioi_art <- df_ioi %>% dplyr::filter(Skill == "articulation")
# Include only dynamics
df_ioi_dyn <- df_ioi %>% dplyr::filter(Skill == "dynamics")
```

## IOI-Articulation 
### 1. All - individual{#art_ind_all}
- Average IOIs for each participant.

```{r ioi_art, echo = FALSE}
# For each individual
ioi_art <- aggregate(IOI~SubNr*Condition*Skill, data = df_ioi_art,
                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
ioi_art <- cbind(ioi_art[,1:3], as.data.frame(ioi_art[,4]))
# Change colnames
colnames(ioi_art) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
print(ioi_art)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_bar,  echo = FALSE}
p_ioi_art <- ggplot(data = ioi_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation")
p_ioi_art
```

### 2. All - group{#art_group_all}
- Average IOIs for each condition.
- Hypothesis: Participants will play slower when teaching.

```{r ioi_art_all, echo = FALSE}
# Group mean
ioi_art_all <- aggregate(Mean~Condition*Skill, data = ioi_art,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_all <- cbind(ioi_art_all[,1:2], as.data.frame(ioi_art_all[,3]))
# Change colnames
colnames(ioi_art_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
print(ioi_art_all)
```

#### Plot
##### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_all_bar, echo = FALSE}
p_ioi_art_all <- ggplot(data = ioi_art_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) +
  theme(legend.position = "None")
p_ioi_art_all
```

##### Box plot
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_all_box, echo = FALSE}
p_ioi_art_all_box <- ggplot(data = ioi_art, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + #coord_cartesian(ylim = c(100, 210)) +
  theme(legend.position = "None")
p_ioi_art_all_box
```

#### Statistics
##### Normality check
```{r normality1, echo = FALSE}
# Normality check
ioi_art_norm <- by(ioi_art$Mean, list(ioi_art$Condition), shapiro.test)
print(ioi_art_norm)
```

##### Two-way ANOVA (ez)
```{r stats1, echo = FALSE, message = FALSE}
ioi_art_aov <- ezANOVA(
  data = df_ioi_art
  , dv = .(IOI)
  , wid = .(SubNr)
  , within = .(Condition)
  , type = 3
  , detailed = TRUE
)
print(ioi_art_aov)
```

##### Two-way ANOVA (aov)
```{r stats2, echo = FALSE, message = FALSE}
ioi_art_aov_2 <- aov(Mean ~ Condition + Error(SubNr/Condition), data = ioi_art)
print(summary(ioi_art_aov_2))
```

### 3. Phrase boundaries - individual{#art_ind_phrase}
- Compare IOIs at phrase boundaries with others fir each participant.

```{r ioi_art_ch, echo = FALSE}
# Define phrase boundaries
df_ioi_art$Boundary <- "No"
df_ioi_art$Boundary[df_ioi_art$Interval == 8 | df_ioi_art$Interval == 16 | df_ioi_art$Interval == 24 | df_ioi_art$Interval == 41 | df_ioi_art$Interval == 49 | df_ioi_art$Interval == 57] <- "Yes"

# For each individual
ioi_art_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_art,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_ch <- cbind(ioi_art_ch[,1:4], as.data.frame(ioi_art_ch[,5]))
# Change colnames
colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
print(ioi_art_ch)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_ch_bar, echo = FALSE}
p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
p_ioi_art_ch
```

### 4. Phrase boundaries - group{#art_group_phrase}
- Compare IOIs at phrase boundaries with others in the two conditions.
- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations.

```{r ioi_art_ch_all, echo = FALSE}
# Group mean
ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_art_ch,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_ch_all <- cbind(ioi_art_ch_all[,1:3], as.data.frame(ioi_art_ch_all[,4]))
# Change colnames
colnames(ioi_art_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
print(ioi_art_ch_all)
```

#### Plot
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_ch_all_bar, echo = FALSE}
p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  #facet_grid(. ~ Boundary) +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 250))
p_ioi_art_ch_all_bar
```

##### Box plot
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_ch_all_box, echo = FALSE}
p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation")
p_ioi_art_ch_box
```

#### Statistics
##### Normality check
```{r norrmality2, echo = FALSE}
ioi_art_ch_norm <- by(ioi_art_ch$Mean, list(ioi_art_ch$Condition, ioi_art_ch$Boundary), shapiro.test)
print(ioi_art_ch_norm)
```

##### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
ioi_art_ch_aov <- ezANOVA(
  data = df_ioi_art
  , dv = .(IOI)
  , wid = .(SubNr)
  , within = .(Condition, Boundary)
  , type = 3
  , detailed = TRUE
)
print(ioi_art_ch_aov)
```

##### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
ioi_art_ch_aov_2 <- with(ioi_art_ch,
  aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary))))
print(summary(ioi_art_ch_aov_2))
```

### 5. Trial by trial - individual{#art_ind_trial}
- Average IOIs for each trial for each participant.

```{r ioi_art_trial, echo = FALSE}
# For each individual
ioi_art_trial <- aggregate(IOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_art,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_trial <- cbind(ioi_art_trial[,1:4], as.data.frame(ioi_art_trial[,5]))
# Change colnames
colnames(ioi_art_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean", "SD")
print(ioi_art_trial)
```

#### Plot
##### Line graph
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_trial_line, echo = FALSE, fig.height = 8}
p_ioi_art_trial <- ggplot(data = ioi_art_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial
```

### 6. Trial by trial - group{#art_group_trial}
- Average IOIs for each trial for each condition.

```{r ioi_art_trial_all, echo = FALSE}
# Group mean
ioi_art_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_art_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_trial_all <- cbind(ioi_art_trial_all[,1:3], as.data.frame(ioi_art_trial_all[,4]))
# Change colnames
colnames(ioi_art_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
print(ioi_art_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean
```{r ioi_art_trial_all_line, echo = FALSE}
p_ioi_art_trial_all <- ggplot(data = ioi_art_trial_all, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial_all
```

### 7. Variability all - individual{#art_ind_var_all}
- Average variability for each participant.
- IOI variability = SD/mean IOI (per trial)

```{r ioi_art_var, echo = FALSE}
# Create a data frame for variability (SD/)
df_ioi_art_var <- data.frame()
for (subnr in unique(df_ioi_art$SubNr)){
  for (block in unique(df_ioi_art$BlockNr)){
    cond = as.character(unique(df_ioi_art$Condition[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
    skill = as.character(unique(df_ioi_art$Skill[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
    for (trial in unique(df_ioi_art$TrialNr)){
      df_current <- df_ioi_art %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
      if (nrow(df_current) != 0){
        df_ioi_art_var <- rbind(df_ioi_art_var, 
                                data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI)))
      }
    }
  }
}
colnames(df_ioi_art_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability")

# Average variability
# For each individual
ioi_art_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_art_var,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_var <- cbind(ioi_art_var[,1:3], as.data.frame(ioi_art_var[,4]))
# Change colnames
colnames(ioi_art_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
print(ioi_art_var)
```

#### Plot
##### Bar graph
- Error bars show SD of the mean
```{r ioi_art_var_bar, echo = FALSE}
p_ioi_art_var <- ggplot(data = ioi_art_var, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(1)) +
  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation")
p_ioi_art_var
```

### 8. Variability all - group{#art_group_var_all}
- Average variability for each condition.

```{r ioi_art_var_all, echo = FALSE}
# Group mean
ioi_art_var_all <- aggregate(Mean~Condition*Skill, data = ioi_art_var,
                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_var_all <- cbind(ioi_art_var_all[,1:2], as.data.frame(ioi_art_var_all[,3]))
# Change colnames
colnames(ioi_art_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
print(ioi_art_var_all)
```

#### Plot
##### Bar graph
- Error bars show SEM of the mean
```{r ioi_art_var_all_bar, echo = FALSE}
p_ioi_art_var_all <- ggplot(data = ioi_art_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge()) +
  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
  theme(legend.position = "None")
p_ioi_art_var_all
```

##### Box plot
```{r ioi_art_var_all_box, echo = FALSE}
p_ioi_art_var_box <- ggplot(data = ioi_art_var, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
  theme(legend.position = "None")
p_ioi_art_var_box
```

#### Statistics
##### Normality check
```{r norrmality3, echo = FALSE}
ioi_art_var_norm <- by(ioi_art_var$Mean, list(ioi_art_var$Condition), shapiro.test)
print(ioi_art_var_norm)
```

##### Two-way ANOVA (ez)
```{r stats5, echo = FALSE}
ioi_art_var_aov <- ezANOVA(
  data = df_ioi_art_var
  , dv = .(Variability)
  , wid = .(SubNr)
  , within = .(Condition)
  , type = 3
  , detailed = TRUE
)
print(ioi_art_var_aov)
```

##### Two-way ANOVA (aov)
```{r stats6, echo = FALSE}
ioi_art_var_aov_2 <- with(ioi_art_var,
  aov(Mean ~ Condition + Error(SubNr/Condition)))
print(summary(ioi_art_var_aov_2))
```

### 9. Variability trial-by-trial - individual{#art_ind_var_trial}
- Average variability for each trial for each participant.

```{r ioi_art_var_trial, echo = FALSE}
# For each individual
ioi_art_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_art_var,
                         FUN = function(x){c(N = length(x), mean = mean(x))})
ioi_art_var_trial <- cbind(ioi_art_var_trial[,1:4], as.data.frame(ioi_art_var_trial[,5]))
# Change colnames
colnames(ioi_art_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean")
print(ioi_art_var_trial)
```

#### Plot
##### Line graph
```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 8}
p_ioi_art_var_trial <- ggplot(data = ioi_art_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_var_trial
```

### 10. Variability trial-by-trial - group{#art_group_var_trial}
- Average variability for each trial for each condition.

```{r ioi_art_var_trial_all, echo = FALSE}
# Group mean
ioi_art_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_art_var_trial,
                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_var_trial_all <- cbind(ioi_art_var_trial_all[,1:3], as.data.frame(ioi_art_var_trial_all[,4]))
# Change colnames
colnames(ioi_art_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
print(ioi_art_var_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean
```{r ioi_art_var_trial_all_line, echo = FALSE}
p_ioi_art_var_trial_all <- ggplot(data = ioi_art_var_trial_all, 
                                  aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_var_trial_all
```

### Summary (IOI:articulation){#art_summary}
#### 1. All - individual
Some participants exhibited a strong tendency for slow demonstration while teaching.

#### 2. All - group
Participants played slower while teaching.

#### 3. Phrase boundaries - individual
Some participants played slower particularly at phrase boundaries.

#### 4. Phrase boundaries - group
Participants played slower while teaching / at phrase boundaries. However, there was no interaction.

#### 5. Trial by trial - individual
There seems no effect of TrialNr?

#### 6. Trial by trial - group
There seems no effect of TrialNr?

#### 7. Variability all - individual
Some participants showed a strong tendency to fluctuate (in terms of IOIs) in the performing condition.

#### 8. Variability all - group
Participants varied their tempo more in the performing condition.

#### 9. Variability trial-by-trial - individual
There seems no effect of TrialNr?

#### 10. Variability trial-by-trial - group
There seems no effect of TrialNr?

###########
###########
## IOI-Dynamics
### 1. All - individual{#dyn_ind_all}
- Average IOIs for each participant.
```{r ioi_dyn, echo = FALSE}
# For each individual
ioi_dyn <- aggregate(IOI~SubNr*Condition*Skill, data = df_ioi_dyn,
                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
ioi_dyn <- cbind(ioi_dyn[,1:3], as.data.frame(ioi_dyn[,4]))
# Change colnames
colnames(ioi_dyn) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
print(ioi_dyn)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_dyn_bar,  echo = FALSE}
p_ioi_dyn <- ggplot(data = ioi_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics")
p_ioi_dyn
```

### 2. All - group{#dyn_group_all}
- Average IOIs for each condition.
- Hypothesis: Participants will play slower when teaching.

```{r ioi_dyn_all, echo = FALSE}
# Group mean
ioi_dyn_all <- aggregate(Mean~Condition*Skill, data = ioi_dyn,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_dyn_all <- cbind(ioi_dyn_all[,1:2], as.data.frame(ioi_dyn_all[,3]))
# Change colnames
colnames(ioi_dyn_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
print(ioi_dyn_all)
```

#### Plot
##### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_dyn_all_bar, echo = FALSE}
p_ioi_dyn_all <- ggplot(data = ioi_dyn_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(120, 210)) +
  theme(legend.position = "None")
p_ioi_dyn_all
```

##### Box plot
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_dyn_all_box, echo = FALSE}
p_ioi_dyn_all_box <- ggplot(data = ioi_dyn, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + #coord_cartesian(ylim = c(100, 210)) +
  theme(legend.position = "None")
p_ioi_dyn_all_box
```

#### Statistics
##### Normality check
```{r normality4, echo = FALSE}
# Normality check
ioi_dyn_norm <- by(ioi_dyn$Mean, list(ioi_dyn$Condition), shapiro.test)
print(ioi_dyn_norm)
```

##### Two-way ANOVA (ez)
```{r stats7, echo = FALSE, message = FALSE}
ioi_dyn_aov <- ezANOVA(
  data = df_ioi_dyn
  , dv = .(IOI)
  , wid = .(SubNr)
  , within = .(Condition)
  , type = 3
  , detailed = TRUE
)
print(ioi_dyn_aov)
```

##### Two-way ANOVA (aov)
```{r stats8, echo = FALSE, message = FALSE}
ioi_dyn_aov_2 <- aov(Mean ~ Condition + Error(SubNr/Condition), data = ioi_dyn)
print(summary(ioi_dyn_aov_2))
```

### 3. Phrase boundaries - individual{#dyn_ind_phrase}
- Compare IOIs at phrase boundaries with others fir each participant.

```{r ioi_dyn_ch, echo = FALSE}
# Define phrase boundaries
df_ioi_dyn$Boundary <- "No"
df_ioi_dyn$Boundary[df_ioi_dyn$Interval == 8 | df_ioi_dyn$Interval == 16 | df_ioi_dyn$Interval == 24 | df_ioi_dyn$Interval == 41 | df_ioi_dyn$Interval == 49 | df_ioi_dyn$Interval == 57] <- "Yes"

# For each individual
ioi_dyn_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_dyn,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_dyn_ch <- cbind(ioi_dyn_ch[,1:4], as.data.frame(ioi_dyn_ch[,5]))
# Change colnames
colnames(ioi_dyn_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
print(ioi_dyn_ch)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_dyn_ch_bar, echo = FALSE}
p_ioi_dyn_ch <- ggplot(data = ioi_dyn_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
p_ioi_dyn_ch
```

### 4. Phrase boundaries - group{#dyn_group_phrase}
- Compare IOIs at phrase boundaries with others in the two conditions.
- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations.

```{r ioi_dyn_ch_all, echo = FALSE}
# Group mean
ioi_dyn_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_dyn_ch,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_dyn_ch_all <- cbind(ioi_dyn_ch_all[,1:3], as.data.frame(ioi_dyn_ch_all[,4]))
# Change colnames
colnames(ioi_dyn_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
print(ioi_dyn_ch_all)
```

#### Plot
##### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_dyn_ch_all_bar, echo = FALSE}
p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  #facet_grid(. ~ Boundary) +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 250))
p_ioi_dyn_ch_all_bar
```

##### Box plot
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_dyn_ch_all_box, echo = FALSE}
p_ioi_dyn_ch_box <- ggplot(data = ioi_dyn_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics")
p_ioi_dyn_ch_box
```

#### Statistics
##### Normality check
```{r norrmality5, echo = FALSE}
ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test)
print(ioi_dyn_ch_norm)
```

##### Two-way ANOVA (ez)
```{r stats9, echo = FALSE}
ioi_dyn_ch_aov <- ezANOVA(
  data = df_ioi_dyn
  , dv = .(IOI)
  , wid = .(SubNr)
  , within = .(Condition, Boundary)
  , type = 3
  , detailed = TRUE
)
print(ioi_dyn_ch_aov)
```

##### Two-way ANOVA (aov)
```{r stats10, echo = FALSE}
ioi_dyn_ch_aov_2 <- with(ioi_dyn_ch,
  aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary))))
print(summary(ioi_dyn_ch_aov_2))
```

### 5. Trial by trial - individual{#dyn_ind_trial}
- Average IOIs for each trial for each participant.

```{r ioi_dyn_trial, echo = FALSE}
# For each individual
ioi_dyn_trial <- aggregate(IOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_dyn_trial <- cbind(ioi_dyn_trial[,1:4], as.data.frame(ioi_dyn_trial[,5]))
# Change colnames
colnames(ioi_dyn_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean", "SD")
print(ioi_dyn_trial)
```

#### Plot
##### Line graph
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_dyn_trial_line, echo = FALSE, fig.height = 8}
p_ioi_dyn_trial <- ggplot(data = ioi_dyn_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_trial
```

### 6. Trial by trial - group{#dyn_group_trial}
- Average IOIs for each trial for each condition.

```{r ioi_dyn_trial_all, echo = FALSE}
# Group mean
ioi_dyn_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_dyn_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_dyn_trial_all <- cbind(ioi_dyn_trial_all[,1:3], as.data.frame(ioi_dyn_trial_all[,4]))
# Change colnames
colnames(ioi_dyn_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
print(ioi_dyn_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean
```{r ioi_dyn_trial_all_line, echo = FALSE}
p_ioi_dyn_trial_all <- ggplot(data = ioi_dyn_trial_all, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_trial_all
```

### 7. Variability all - individual{#dyn_ind_var_all}
- Average variability for each participant.
- IOI variability = SD/mean IOI (per trial)

```{r ioi_dyn_var, echo = FALSE}
# Create a data frame for variability (SD/)
df_ioi_dyn_var <- data.frame()
for (subnr in unique(df_ioi_dyn$SubNr)){
  for (block in unique(df_ioi_dyn$BlockNr)){
    cond = as.character(unique(df_ioi_dyn$Condition[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
    skill = as.character(unique(df_ioi_dyn$Skill[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
    for (trial in unique(df_ioi_dyn$TrialNr)){
      df_current <- df_ioi_dyn %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
      if (nrow(df_current) != 0){
        df_ioi_dyn_var <- rbind(df_ioi_dyn_var, 
                                data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI)))
      }
    }
  }
}
colnames(df_ioi_dyn_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability")

# Average variability
# For each individual
ioi_dyn_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_dyn_var,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_dyn_var <- cbind(ioi_dyn_var[,1:3], as.data.frame(ioi_dyn_var[,4]))
# Change colnames
colnames(ioi_dyn_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
print(ioi_dyn_var)
```

#### Plot
##### Bar graph
- Error bars show SD of the mean
```{r ioi_dyn_var_bar, echo = FALSE}
p_ioi_dyn_var <- ggplot(data = ioi_dyn_var, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(1)) +
  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics")
p_ioi_dyn_var
```

### 8. Variability all - group{#dyn_group_var_all}
- Average variability for each condition.

```{r ioi_dyn_var_all, echo = FALSE}
# Group mean
ioi_dyn_var_all <- aggregate(Mean~Condition*Skill, data = ioi_dyn_var,
                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_dyn_var_all <- cbind(ioi_dyn_var_all[,1:2], as.data.frame(ioi_dyn_var_all[,3]))
# Change colnames
colnames(ioi_dyn_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
print(ioi_dyn_var_all)
```

#### Plot
##### Bar graph
- Error bars show SEM of the mean
```{r ioi_dyn_var_all_bar, echo = FALSE}
p_ioi_dyn_var_all <- ggplot(data = ioi_dyn_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge()) +
  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
  theme(legend.position = "None")
p_ioi_dyn_var_all
```

##### Box plot
```{r ioi_dyn_var_all_box, echo = FALSE}
p_ioi_dyn_var_box <- ggplot(data = ioi_dyn_var, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
  theme(legend.position = "None")
p_ioi_dyn_var_box
```

#### Statistics
##### Normality check
```{r norrmality6, echo = FALSE}
ioi_dyn_var_norm <- by(ioi_dyn_var$Mean, list(ioi_dyn_var$Condition), shapiro.test)
print(ioi_dyn_var_norm)
```

##### Two-way ANOVA (ez)
```{r stats11, echo = FALSE}
ioi_dyn_var_aov <- ezANOVA(
  data = df_ioi_dyn_var
  , dv = .(Variability)
  , wid = .(SubNr)
  , within = .(Condition)
  , type = 3
  , detailed = TRUE
)
print(ioi_dyn_var_aov)
```

##### Two-way ANOVA (aov)
```{r stats12, echo = FALSE}
ioi_dyn_var_aov_2 <- with(ioi_dyn_var,
  aov(Mean ~ Condition + Error(SubNr/Condition)))
print(summary(ioi_dyn_var_aov_2))
```

### 9. Variability trial-by-trial - individual{#dyn_ind_var_trial}
- Average variability for each trial for each participant.

```{r ioi_dyn_var_trial, echo = FALSE}
# For each individual
ioi_dyn_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn_var,
                         FUN = function(x){c(N = length(x), mean = mean(x))})
ioi_dyn_var_trial <- cbind(ioi_dyn_var_trial[,1:4], as.data.frame(ioi_dyn_var_trial[,5]))
# Change colnames
colnames(ioi_dyn_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean")
print(ioi_dyn_var_trial)
```

#### Plot
##### Line graph
```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 8}
p_ioi_dyn_var_trial <- ggplot(data = ioi_dyn_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_var_trial
```

### 10. Variability trial-by-trial - group{#dyn_group_var_trial}
Average variability for each trial for each condition.

```{r ioi_dyn_var_trial_all, echo = FALSE}
# Group mean
ioi_dyn_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_dyn_var_trial,
                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_dyn_var_trial_all <- cbind(ioi_dyn_var_trial_all[,1:3], as.data.frame(ioi_dyn_var_trial_all[,4]))
# Change colnames
colnames(ioi_dyn_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
print(ioi_dyn_var_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean
```{r ioi_dyn_var_trial_all_line, echo = FALSE}
p_ioi_dyn_var_trial_all <- ggplot(data = ioi_dyn_var_trial_all, 
                                  aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_var_trial_all
```

### Summary (IOI:dynamics){#dyn_summary}
#### 1. All - individual
Some participants exhibited a strong tendency for slow demonstration while teaching.

#### 2. All - group
Participants did not play more slowly during teaching.

#### 3. Phrase boundaries - individual
Some participants played slower particularly at phrase boundaries.

#### 4. Phrase boundaries - group
Participants played slower at phrase boundaries. However, there was no effect of the condition or interaction.

#### 5. Trial by trial - individual
There seems no effect of TrialNr?

#### 6. Trial by trial - group
There seems no effect of TrialNr?

#### 7. Variability all - individual
Some participants showed a strong tendency to fluctuate (in terms of IOIs) in the performing condition.

#### 8. Variability all - group
Participants did not vary their tempo depending on the conditions.

#### 9. Variability trial-by-trial - individual
There seems no effect of TrialNr?

#### 10. Variability trial-by-trial - group
There seems no effect of TrialNr?

### 2019-10-01 script/RNotebook/ioi.Rmd
delete one-way anova and replace them by t-test/cohend
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("stats")) {install.packages("stats"); require("stats")}
+if (!require("effsize")) {install.packages("effsize"); require("effsize")}
 if (!require("ez")) {install.packages("ez"); require("ez")}
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_art_bar,  echo = FALSE}
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_art_all_bar, echo = FALSE}
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_art_all_box, echo = FALSE}
@@
 
-##### Two-way ANOVA (ez)
-```{r stats1, echo = FALSE, message = FALSE}
-ioi_art_aov <- ezANOVA(
-  data = df_ioi_art
-  , dv = .(IOI)
-  , wid = .(SubNr)
-  , within = .(Condition)
-  , type = 3
-  , detailed = TRUE
-)
-print(ioi_art_aov)
+##### Paired t-test
+```{r stats1, echo = FALSE}
+ioi_art_ttest <- t.test(ioi_art$Mean[ioi_art$Condition == "teaching"], ioi_art$Mean[ioi_art$Condition == "performing"], paired = TRUE)
+print(ioi_art_ttest)
 ```
 
-##### Two-way ANOVA (aov)
-```{r stats2, echo = FALSE, message = FALSE}
-ioi_art_aov_2 <- aov(Mean ~ Condition + Error(SubNr/Condition), data = ioi_art)
-print(summary(ioi_art_aov_2))
+##### Effect size
+```{r stats2, echo = FALSE}
+ioi_art_cohend <- cohen.d(ioi_art$Mean, ioi_art$Condition, paired = TRUE)
+print(ioi_art_cohend)
 ```
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_art_ch_bar, echo = FALSE}
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_art_ch_all_bar, echo = FALSE}
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_art_ch_all_box, echo = FALSE}
@@
 )
-print(ioi_art_ch_aov)
+ioi_art_ch_aov$ANOVA$F <- round(ioi_art_ch_aov$ANOVA$F, 4)
+ioi_art_ch_aov$ANOVA$p <- round(ioi_art_ch_aov$ANOVA$p, 4)
+print(ioi_art_ch_aov$ANOVA)
 ```
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_art_trial_line, echo = FALSE, fig.height = 8}
@@
 - Error bars show SEM of the mean
+
 ```{r ioi_art_trial_all_line, echo = FALSE}
@@
 - Error bars show SD of the mean
+
 ```{r ioi_art_var_bar, echo = FALSE}
@@
 - Error bars show SEM of the mean
+
 ```{r ioi_art_var_all_bar, echo = FALSE}
@@
 
-##### Two-way ANOVA (ez)
+##### Paired t-test
 ```{r stats5, echo = FALSE}
-ioi_art_var_aov <- ezANOVA(
-  data = df_ioi_art_var
-  , dv = .(Variability)
-  , wid = .(SubNr)
-  , within = .(Condition)
-  , type = 3
-  , detailed = TRUE
-)
-print(ioi_art_var_aov)
+ioi_art_var_ttest <- t.test(ioi_art_var$Mean[ioi_art_var$Condition == "teaching"], ioi_art_var$Mean[ioi_art_var$Condition == "performing"], paired = TRUE)
+print(ioi_art_var_ttest)
 ```
 
-##### Two-way ANOVA (aov)
+##### Effect size
 ```{r stats6, echo = FALSE}
-ioi_art_var_aov_2 <- with(ioi_art_var,
-  aov(Mean ~ Condition + Error(SubNr/Condition)))
-print(summary(ioi_art_var_aov_2))
+ioi_art_var_cohend <- cohen.d(ioi_art_var$Mean, ioi_art_var$Condition, paired = TRUE)
+print(ioi_art_var_cohend)
 ```
@@
 - Error bars show SEM of the mean
+
 ```{r ioi_art_var_trial_all_line, echo = FALSE}
@@
 #### 10. Variability trial-by-trial - group
-There seems no effect of TrialNr?
+Variability reduced over the trials
 
@@
 ###########
+
 ## IOI-Dynamics
@@
 - Average IOIs for each participant.
+
 ```{r ioi_dyn, echo = FALSE}
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_dyn_bar,  echo = FALSE}
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_dyn_all_bar, echo = FALSE}
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_dyn_all_box, echo = FALSE}
@@
 
-##### Two-way ANOVA (ez)
-```{r stats7, echo = FALSE, message = FALSE}
-ioi_dyn_aov <- ezANOVA(
-  data = df_ioi_dyn
-  , dv = .(IOI)
-  , wid = .(SubNr)
-  , within = .(Condition)
-  , type = 3
-  , detailed = TRUE
-)
-print(ioi_dyn_aov)
+```{r stats7, echo = FALSE}
+ioi_dyn_ttest <- t.test(ioi_dyn$Mean[ioi_dyn$Condition == "teaching"], ioi_dyn$Mean[ioi_dyn$Condition == "performing"], paired = TRUE)
+print(ioi_dyn_ttest)
 ```
 
-##### Two-way ANOVA (aov)
+##### Effect size
 ```{r stats8, echo = FALSE, message = FALSE}
-ioi_dyn_aov_2 <- aov(Mean ~ Condition + Error(SubNr/Condition), data = ioi_dyn)
-print(summary(ioi_dyn_aov_2))
+ioi_dyn_cohend <- cohen.d(ioi_dyn$Mean, ioi_dyn$Condition, paired = TRUE)
+print(ioi_dyn_cohend)
 ```
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_dyn_ch_bar, echo = FALSE}
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_dyn_ch_all_bar, echo = FALSE}
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_dyn_ch_all_box, echo = FALSE}
@@
 )
+ioi_dyn_ch_aov$ANOVA$F <- round(ioi_dyn_ch_aov$ANOVA$F, 4)
+ioi_dyn_ch_aov$ANOVA$p <- round(ioi_dyn_ch_aov$ANOVA$p, 4)
 print(ioi_dyn_ch_aov)
@@
 - A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_dyn_trial_line, echo = FALSE, fig.height = 8}
@@
 - Error bars show SEM of the mean
+
 ```{r ioi_dyn_trial_all_line, echo = FALSE}
@@
 - Error bars show SD of the mean
+
 ```{r ioi_dyn_var_bar, echo = FALSE}
@@
 
-##### Two-way ANOVA (ez)
 ```{r stats11, echo = FALSE}
-ioi_dyn_var_aov <- ezANOVA(
-  data = df_ioi_dyn_var
-  , dv = .(Variability)
-  , wid = .(SubNr)
-  , within = .(Condition)
-  , type = 3
-  , detailed = TRUE
-)
-print(ioi_dyn_var_aov)
+ioi_dyn_var_ttest <- t.test(ioi_dyn_var$Mean[ioi_dyn_var$Condition == "teaching"], ioi_dyn_var$Mean[ioi_dyn_var$Condition == "performing"], paired = TRUE)
+print(ioi_dyn_var_ttest)
 ```
 
-##### Two-way ANOVA (aov)
 ```{r stats12, echo = FALSE}
-ioi_dyn_var_aov_2 <- with(ioi_dyn_var,
-  aov(Mean ~ Condition + Error(SubNr/Condition)))
-print(summary(ioi_dyn_var_aov_2))
+ioi_dyn_var_cohend <- cohen.d(ioi_dyn_var$Mean, ioi_dyn_var$Condition, paired = TRUE)
+print(ioi_dyn_var_cohend)
 ```
@@
 - Error bars show SEM of the mean
+
 ```{r ioi_dyn_var_trial_all_line, echo = FALSE}
@@
 #### 10. Variability trial-by-trial - group
-There seems no effect of TrialNr?
+Variability reduced over the trials

### 2019-10-02 script/RNotebook/ioi.Rmd
correct typo
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-##### Description: This is a summary of all analyses for "The Sound of Teaching Music 1".
+##### Description: This is a summary of all analyses for a tempo.
 
@@
 ##### Normality check
-```{r norrmality2, echo = FALSE}
+```{r normality2, echo = FALSE}
 ioi_art_ch_norm <- by(ioi_art_ch$Mean, list(ioi_art_ch$Condition, ioi_art_ch$Boundary), shapiro.test)
@@
 ##### Normality check
-```{r norrmality3, echo = FALSE}
+```{r normality3, echo = FALSE}
 ioi_art_var_norm <- by(ioi_art_var$Mean, list(ioi_art_var$Condition), shapiro.test)
@@
 ##### Effect size
-```{r stats8, echo = FALSE, message = FALSE}
+```{r stats8, echo = FALSE}
 ioi_dyn_cohend <- cohen.d(ioi_dyn$Mean, ioi_dyn$Condition, paired = TRUE)
@@
 ##### Normality check
-```{r norrmality5, echo = FALSE}
+```{r normality5, echo = FALSE}
 ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test)
@@
 ##### Normality check
-```{r norrmality6, echo = FALSE}
+```{r normality6, echo = FALSE}
 ioi_dyn_var_norm <- by(ioi_dyn_var$Mean, list(ioi_dyn_var$Condition), shapiro.test)

### 2019-10-02 script/RNotebook/articulation.Rmd
update articulation analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of all analyses for articulation.

# Outline - Articulation (KOT, KOR)
### [Experimental Design](#design)

### [Articulation](#kot-articulation)

#### KOT
- Individual results
   + [All](#kot_art_ind_all)
   + [Trial by trial](#kot_art_ind_trial)

- Group results and stats (if available)
   + [All](#kot_art_group_all)
   + [Trial by trial](#kot_art_group_trial)

#### KOR
- Individual results
   + [All](#kor_art_ind_all)
   + [Trial by trial](#kor_art_ind_trial)

- Group results and stats (if available)
   + [All](#kor_art_group_all)
   + [Trial by trial](#kor_art_group_trial)

- [Summary](#art_summary)

### [Dynamics](#kot-dynamics)

#### KOT
- Individual results
   + [All](#kot_dyn_ind_all)
   + [Trial by trial](#kot_dyn_ind_trial)

- Group results and stats (if available)
   + [All](#kot_dyn_group_all)
   + [Trial by trial](#kot_dyn_group_trial)

#### KOR
- Individual results
   + [All](#kor_dyn_ind_all)
   + [Trial by trial](#kor_dyn_ind_trial)

- Group results and stats (if available)
   + [All](#kor_dyn_group_all)
   + [Trial by trial](#kor_dyn_group_trial)

- [Summary](#dyn_summary)

## Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: legato vs. staccato / forte vs. piano)
- Dependent variables: Key-Overlap Time (KOT), Key-Overlap Ratio (KOR)

### Main predictions
Participants will exaggerate legato and staccato more while teaching.
Participants will exaggerate forte and piano more while teaching.

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("effsize")) {install.packages("effsize"); require("effsize")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
df_kot <- read.csv("./trimmed/data_kot.csv", header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
df_ioi$SubNr <- as.factor(df_ioi$SubNr)

# SubNr as a factor
df_ioi$SubNr <- as.factor(df_ioi$SubNr)
df_kot$SubNr <- as.factor(df_kot$SubNr)

# For each individual, each block, each trial
ioi <- aggregate(IOI~SubNr*BlockNr*TrialNr*Condition*Skill, data = df_ioi,
                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
ioi <- cbind(ioi[,1:5], as.data.frame(ioi[,6]))
# Change colnames
colnames(ioi) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "N", "Mean", "SD")

# Calculate KOR for each interval
df_kor <- data.frame()
for (subnr in unique(df_kot$SubNr)){
  for (block in unique(df_kot$BlockNr[df_kot$SubNr == subnr])){
    for (trial in unique(df_kot$TrialNr[df_kot$SubNr == subnr & df_kot$BlockNr == block])){
      df_current <- df_kot %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
      df_current$KOR <- df_current$KOT/ioi$Mean[ioi$SubNr == subnr & ioi$BlockNr == block & ioi$TrialNr == trial]
      df_kor <- rbind(df_kor, df_current)
    }
  }
}

# Include only articulation
df_kot_art <- df_kot %>% dplyr::filter(Skill == "articulation")
df_kor_art <- df_kor %>% dplyr::filter(Skill == "articulation")

# Include only dynamics
df_kot_dyn <- df_kot %>% dplyr::filter(Skill == "dynamics")
df_kor_dyn <- df_kor %>% dplyr::filter(Skill == "dynamics")
```

## Key-Overlap Time (KOT) - Articulation
### 1. All - individual{#kot_art_ind_all}
- Average KOT for each participant.

```{r kot_art, echo = FALSE}
# For each individual
kot_art <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_art <- cbind(kot_art[,1:4], as.data.frame(kot_art[,5]))
# Change colnames
colnames(kot_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kot_art)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

```{r kot_art_bar,  echo = FALSE}
p_kot_art <- ggplot(data = kot_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art
```

### 2. All - group{#kot_art_group_all}
- Average KOT for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kot_art_all, echo = FALSE}
kot_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kot_art,
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kot_art_all <- cbind(kot_art_all[,1:3], as.data.frame(kot_art_all[,4]))
# Change colnames
colnames(kot_art_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(kot_art_all)
```

#### Plot
##### Bar chart
- Error bars show SEM of the mean

```{r kot_art_all_bar, echo = FALSE}　
p_kot_art_all <- ggplot(data = kot_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art_all
```

##### Box plot

```{r kot_art_all_box, echo = FALSE}
p_kot_art_all_box <- ggplot(data = kot_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art_all_box
```

#### Statistics
##### Normality check
```{r normality1, echo = FALSE}
# Reduce unused factors
kot_art$Subcomponent <- factor(kot_art$Subcomponent)
kot_art_norm <- by(kot_art$Mean, list(kot_art$Condition, kot_art$Subcomponent), shapiro.test)
print(kot_art_norm)
```

##### Two-way ANOVA (ez)
```{r stats1, echo = FALSE}
kot_art_aov <- ezANOVA(
  data = subset(df_kot_art, df_kot_art$Subcomponent == "Legato" | df_kot_art$Subcomponent == "Staccato")
  , dv = .(KOT)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kot_art_aov$ANOVA$F <- round(kot_art_aov$ANOVA$F, 4)
kot_art_aov$ANOVA$p <- round(kot_art_aov$ANOVA$p, 4)
print(kot_art_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats2, echo = FALSE}
kot_art_aov_2 <- with(kot_art,
                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
print(summary(kot_art_aov_2))
```

### 3. Trial by trial - individual{#kot_art_ind_trial}
- Average KOT for each trial for each participant.

```{r kot_art_trial, echo = FALSE}
# For each individual
kot_art_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_art_trial <- cbind(kot_art_trial[,1:5], as.data.frame(kot_art_trial[,6]))
# Change colnames
colnames(kot_art_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(kot_art_trial)
```

#### Plot
##### Line graph
```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 8}
p_kot_art_trial_legato <- ggplot(data = subset(kot_art_trial, kot_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_legato
```

```{r kot_art_trial_line_staccato, echo = FALSE, fig.height = 8}
p_kot_art_trial_staccato <- ggplot(data = subset(kot_art_trial, kot_art_trial$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_staccato
```

### 4. Trial by trial - group{#kot_art_group_trial}
- Average KOT for each trial for each condition.

```{r kot_art_trial_all, echo = FALSE}
# Group mean
kot_art_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = kot_art_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kot_art_trial_all <- cbind(kot_art_trial_all[,1:4], as.data.frame(kot_art_trial_all[,5]))
# Change colnames
colnames(kot_art_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(kot_art_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean

```{r kot_art_trial_all_line_legato, echo = FALSE}
p_kot_art_trial_all_legato <- ggplot(data = subset(kot_art_trial_all, kot_art_trial_all$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_all_legato
```

```{r kot_art_trial_all_line_staccato, echo = FALSE}
p_kot_art_trial_all_staccato <- ggplot(data = subset(kot_art_trial_all, kot_art_trial_all$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_all_staccato
```

## Key-Overlap Ratio (KOR) - Articulation
### 1. All - individual{#kor_art_ind_all}
- Average KOR for each participant.

```{r kor_art, echo = FALSE}
# For each individual
kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kor_art <- cbind(kor_art[,1:4], as.data.frame(kor_art[,5]))
# Change colnames
colnames(kor_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kor_art)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

```{r kor_art_bar,  echo = FALSE}
p_kor_art <- ggplot(data = kor_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art
```

### 2. All - group{#kor_art_group_all}
- Average KOR for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kor_art_all, echo = FALSE}
kor_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kor_art,
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kor_art_all <- cbind(kor_art_all[,1:3], as.data.frame(kor_art_all[,4]))
# Change colnames
colnames(kor_art_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(kor_art_all)
```

#### Plot
##### Bar chart
- Error bars show SEM of the mean

```{r kor_art_all_bar, echo = FALSE}　
p_kor_art_all <- ggplot(data = kor_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art_all
```

##### Box plot

```{r kor_art_all_box, echo = FALSE}
p_kor_art_all_box <- ggplot(data = kor_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art_all_box
```

#### Statistics
##### Normality check
```{r normality2, echo = FALSE}
# Reduce unused factors
kor_art$Subcomponent <- factor(kor_art$Subcomponent)
kor_art_norm <- by(kor_art$Mean, list(kor_art$Condition, kor_art$Subcomponent), shapiro.test)
print(kor_art_norm)
```

##### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
kor_art_aov <- ezANOVA(
  data = subset(df_kor_art, df_kor_art$Subcomponent == "Legato" | df_kor_art$Subcomponent == "Staccato")
  , dv = .(KOR)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kor_art_aov$ANOVA$F <- round(kor_art_aov$ANOVA$F, 4)
kor_art_aov$ANOVA$p <- round(kor_art_aov$ANOVA$p, 4)
print(kor_art_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
kor_art_aov_2 <- with(kor_art,
                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
print(summary(kor_art_aov_2))
```

### 3. Trial by trial - individual{#kor_art_ind_trial}
- Average KOR for each trial for each participant.

```{r kor_art_trial, echo = FALSE}
# For each individual
kor_art_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kor_art_trial <- cbind(kor_art_trial[,1:5], as.data.frame(kor_art_trial[,6]))
# Change colnames
colnames(kor_art_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(kor_art_trial)
```

#### Plot
##### Line graph
```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 8}
p_kor_art_trial_legato <- ggplot(data = subset(kor_art_trial, kor_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_legato
```

```{r kor_art_trial_line_staccato, echo = FALSE, fig.height = 8}
p_kor_art_trial_staccato <- ggplot(data = subset(kor_art_trial, kor_art_trial$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_staccato
```

### 4. Trial by trial - group{#kor_art_group_trial}
- Average KOR d for each trial for each condition.

```{r kor_art_trial_all, echo = FALSE}
# Group mean
kor_art_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = kor_art_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kor_art_trial_all <- cbind(kor_art_trial_all[,1:4], as.data.frame(kor_art_trial_all[,5]))
# Change colnames
colnames(kor_art_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(kor_art_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean

```{r kor_art_trial_all_line_legato, echo = FALSE}
p_kor_art_trial_all_legato <- ggplot(data = subset(kor_art_trial_all, kor_art_trial_all$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_all_legato
```

```{r kor_art_trial_all_line_staccato, echo = FALSE}
p_kor_art_trial_all_staccato <- ggplot(data = subset(kor_art_trial_all, kor_art_trial_all$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_all_staccato
```

## Key-Overlap Time (KOT) - Dynamics
### 1. All - individual{#kot_dyn_ind_all}
- Average KOT for each participant.

```{r kot_dyn, echo = FALSE}
# For each individual
kot_dyn <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_dyn <- cbind(kot_dyn[,1:4], as.data.frame(kot_dyn[,5]))
# Change colnames
colnames(kot_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kot_dyn)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

```{r kot_dyn_bar,  echo = FALSE}
p_kot_dyn <- ggplot(data = kot_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_dyn
```

### 2. All - group{#kot_dyn_group_all}
- Average KOT for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kot_dyn_all, echo = FALSE}
kot_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kot_dyn,
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kot_dyn_all <- cbind(kot_dyn_all[,1:3], as.data.frame(kot_dyn_all[,4]))
# Change colnames
colnames(kot_dyn_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(kot_dyn_all)
```

#### Plot
##### Bar chart
- Error bars show SEM of the mean

```{r kot_dyn_all_bar, echo = FALSE}　
p_kot_dyn_all <- ggplot(data = kot_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn_all
```

##### Box plot

```{r kot_dyn_all_box, echo = FALSE}
p_kot_dyn_all_box <- ggplot(data = kot_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn_all_box
```

#### Statistics
##### Normality check
```{r normality3, echo = FALSE}
# Reduce unused factors
kot_dyn$Subcomponent <- factor(kot_dyn$Subcomponent)
kot_dyn_norm <- by(kot_dyn$Mean, list(kot_dyn$Condition, kot_dyn$Subcomponent), shapiro.test)
print(kot_dyn_norm)
```

##### Two-way ANOVA (ez)
```{r stats5, echo = FALSE}
kot_dyn_aov <- ezANOVA(
  data = subset(df_kot_dyn, df_kot_dyn$Subcomponent == "Forte" | df_kot_dyn$Subcomponent == "Piano")
  , dv = .(KOT)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kot_dyn_aov$ANOVA$F <- round(kot_dyn_aov$ANOVA$F, 4)
kot_dyn_aov$ANOVA$p <- round(kot_dyn_aov$ANOVA$p, 4)
print(kot_dyn_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats6, echo = FALSE}
kot_dyn_aov_2 <- with(kot_dyn,
                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
print(summary(kot_dyn_aov_2))
```

### 3. Trial by trial - individual{#kot_dyn_ind_trial}
- Average KOT for each trial for each participant.

```{r kot_dyn_trial, echo = FALSE}
# For each individual
kot_dyn_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_dyn_trial <- cbind(kot_dyn_trial[,1:5], as.data.frame(kot_dyn_trial[,6]))
# Change colnames
colnames(kot_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(kot_dyn_trial)
```

#### Plot
##### Line graph
```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
p_kot_dyn_trial_forte <- ggplot(data = subset(kot_dyn_trial, kot_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_forte
```

```{r kot_dyn_trial_line_piano, echo = FALSE, fig.height = 8}
p_kot_dyn_trial_piano <- ggplot(data = subset(kot_dyn_trial, kot_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_piano
```

### 4. Trial by trial - group{#kot_dyn_group_trial}
- Average KOT for each trial for each condition.

```{r kot_dyn_trial_all, echo = FALSE}
# Group mean
kot_dyn_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = kot_dyn_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kot_dyn_trial_all <- cbind(kot_dyn_trial_all[,1:4], as.data.frame(kot_dyn_trial_all[,5]))
# Change colnames
colnames(kot_dyn_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(kot_dyn_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean

```{r kot_dyn_trial_all_line_forte, echo = FALSE}
p_kot_dyn_trial_all_forte <- ggplot(data = subset(kot_dyn_trial_all, kot_dyn_trial_all$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_all_forte
```

```{r kot_dyn_trial_all_line_piano, echo = FALSE}
p_kot_dyn_trial_all_piano <- ggplot(data = subset(kot_dyn_trial_all, kot_dyn_trial_all$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_all_piano
```

## Key-Overlap Ratio (KOR) - Dynamics
### 1. All - individual{#kor_dyn_ind_all}
- Average KOR for each participant.

```{r kor_dyn, echo = FALSE}
# For each individual
kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kor_dyn <- cbind(kor_dyn[,1:4], as.data.frame(kor_dyn[,5]))
# Change colnames
colnames(kor_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kor_dyn)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

```{r kor_dyn_bar,  echo = FALSE}
p_kor_dyn <- ggplot(data = kor_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn
```

### 2. All - group{#kor_dyn_group_all}
- Average KOR for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kor_dyn_all, echo = FALSE}
kor_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kor_dyn,
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kor_dyn_all <- cbind(kor_dyn_all[,1:3], as.data.frame(kor_dyn_all[,4]))
# Change colnames
colnames(kor_dyn_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(kor_dyn_all)
```

#### Plot
##### Bar chart
- Error bars show SEM of the mean

```{r kor_dyn_all_bar, echo = FALSE}　
p_kor_dyn_all <- ggplot(data = kor_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn_all
```

##### Box plot

```{r kor_dyn_all_box, echo = FALSE}
p_kor_dyn_all_box <- ggplot(data = kor_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn_all_box
```

#### Statistics
##### Normality check
```{r normality4, echo = FALSE}
# Reduce unused factors
kor_dyn$Subcomponent <- factor(kor_dyn$Subcomponent)
kor_dyn_norm <- by(kor_dyn$Mean, list(kor_dyn$Condition, kor_dyn$Subcomponent), shapiro.test)
print(kor_dyn_norm)
```

##### Two-way ANOVA (ez)
```{r stats7, echo = FALSE}
kor_dyn_aov <- ezANOVA(
  data = subset(df_kor_dyn, df_kor_dyn$Subcomponent == "Forte" | df_kor_dyn$Subcomponent == "Piano")
  , dv = .(KOR)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kor_dyn_aov$ANOVA$F <- round(kor_dyn_aov$ANOVA$F, 4)
kor_dyn_aov$ANOVA$p <- round(kor_dyn_aov$ANOVA$p, 4)
print(kor_dyn_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats8, echo = FALSE}
kor_dyn_aov_2 <- with(kor_dyn,
                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
print(summary(kor_dyn_aov_2))
```

### 3. Trial by trial - individual{#kor_dyn_ind_trial}
- Average KOR for each trial for each participant.

```{r kor_dyn_trial, echo = FALSE}
# For each individual
kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kor_dyn_trial <- cbind(kor_dyn_trial[,1:5], as.data.frame(kor_dyn_trial[,6]))
# Change colnames
colnames(kor_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(kor_dyn_trial)
```

#### Plot
##### Line graph
```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
p_kor_dyn_trial_forte <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_forte
```

```{r kor_dyn_trial_line_piano, echo = FALSE, fig.height = 8}
p_kor_dyn_trial_piano <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_piano
```

### 4. Trial by trial - group{#kor_dyn_group_trial}
- Average KOR d for each trial for each condition.

```{r kor_dyn_trial_all, echo = FALSE}
# Group mean
kor_dyn_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = kor_dyn_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kor_dyn_trial_all <- cbind(kor_dyn_trial_all[,1:4], as.data.frame(kor_dyn_trial_all[,5]))
# Change colnames
colnames(kor_dyn_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(kor_dyn_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean

```{r kor_dyn_trial_all_line_forte, echo = FALSE}
p_kor_dyn_trial_all_forte <- ggplot(data = subset(kor_dyn_trial_all, kor_dyn_trial_all$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_all_forte
```

```{r kor_dyn_trial_all_line_piano, echo = FALSE}
p_kor_dyn_trial_all_piano <- ggplot(data = subset(kor_dyn_trial_all, kor_dyn_trial_all$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_all_piano
```

### 2019-10-03 script/RNotebook/ioi.Rmd
remove summaries
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-- [Summary](#art_summary)
-
 ### [Dynamics](#ioi-dynamics)
@@
    + [Variability trial-by-trial](#dyn_group_var_trial)
-
-- [Summary](#dyn_summary)
 
@@
 
-### Summary (IOI:articulation){#art_summary}
-#### 1. All - individual
-Some participants exhibited a strong tendency for slow demonstration while teaching.
-
-#### 2. All - group
-Participants played slower while teaching.
-
-#### 3. Phrase boundaries - individual
-Some participants played slower particularly at phrase boundaries.
-
-#### 4. Phrase boundaries - group
-Participants played slower while teaching / at phrase boundaries. However, there was no interaction.
-
-#### 5. Trial by trial - individual
-There seems no effect of TrialNr?
-
-#### 6. Trial by trial - group
-There seems no effect of TrialNr?
-
-#### 7. Variability all - individual
-Some participants showed a strong tendency to fluctuate (in terms of IOIs) in the performing condition.
-
-#### 8. Variability all - group
-Participants varied their tempo more in the performing condition.
-
-#### 9. Variability trial-by-trial - individual
-There seems no effect of TrialNr?
-
-#### 10. Variability trial-by-trial - group
-Variability reduced over the trials
-
 ###########
@@
 ```
-
-### Summary (IOI:dynamics){#dyn_summary}
-#### 1. All - individual
-Some participants exhibited a strong tendency for slow demonstration while teaching.
-
-#### 2. All - group
-Participants did not play more slowly during teaching.
-
-#### 3. Phrase boundaries - individual
-Some participants played slower particularly at phrase boundaries.
-
-#### 4. Phrase boundaries - group
-Participants played slower at phrase boundaries. However, there was no effect of the condition or interaction.
-
-#### 5. Trial by trial - individual
-There seems no effect of TrialNr?
-
-#### 6. Trial by trial - group
-There seems no effect of TrialNr?
-
-#### 7. Variability all - individual
-Some participants showed a strong tendency to fluctuate (in terms of IOIs) in the performing condition.
-
-#### 8. Variability all - group
-Participants did not vary their tempo depending on the conditions.
-
-#### 9. Variability trial-by-trial - individual
-There seems no effect of TrialNr?
-
-#### 10. Variability trial-by-trial - group
-Variability reduced over the trials

### 2019-10-03 script/RNotebook/articulation.Rmd
remove unnecessary parts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Participants will exaggerate legato and staccato more while teaching.
-Participants will exaggerate forte and piano more while teaching.
 
@@
 df_kot$SubNr <- as.factor(df_kot$SubNr)
-
-# For each individual, each block, each trial
-ioi <- aggregate(IOI~SubNr*BlockNr*TrialNr*Condition*Skill, data = df_ioi,
-                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
-ioi <- cbind(ioi[,1:5], as.data.frame(ioi[,6]))
-# Change colnames
-colnames(ioi) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "N", "Mean", "SD")
 
@@
 
+###########
+###########
+
 ## Key-Overlap Time (KOT) - Dynamics

### 2019-10-03 script/RNotebook/articulation.Rmd
tweak aov formula
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r stats2, echo = FALSE}
-kot_art_aov_2 <- with(kot_art,
-                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
+kot_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
 print(summary(kot_art_aov_2))
@@
 ```{r stats4, echo = FALSE}
-kor_art_aov_2 <- with(kor_art,
-                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
+kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
 print(summary(kor_art_aov_2))
@@
 ```{r stats6, echo = FALSE}
-kot_dyn_aov_2 <- with(kot_dyn,
-                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
+kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
 print(summary(kot_dyn_aov_2))
@@
 ```{r stats8, echo = FALSE}
-kor_dyn_aov_2 <- with(kor_dyn,
-                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
+kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
 print(summary(kor_dyn_aov_2))
@@
 ### 4. Trial by trial - group{#kor_dyn_group_trial}
-- Average KOR d for each trial for each condition.
+- Average KOR for each trial for each condition.
 

### 2019-10-03 script/RNotebook/ioi.Rmd
tweak aov formula
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r stats4, echo = FALSE}
-ioi_art_ch_aov_2 <- with(ioi_art_ch,
-  aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary))))
+ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
 print(summary(ioi_art_ch_aov_2))
@@
 ```{r stats10, echo = FALSE}
-ioi_dyn_ch_aov_2 <- with(ioi_dyn_ch,
-  aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary))))
+ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
 print(summary(ioi_dyn_ch_aov_2))

### 2019-10-03 script/RNotebook/dynamics.Rmd
upload dynamics analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of all analyses for dynamics.

# Outline - Dynamics (Key Velocity)
### [Experimental Design](#design)

### [Dynamics](#vel-dynamics)

#### KV
- Individual results
   + [All](#vel_dyn_ind_all)
   + [Trial by trial](#vel_dyn_ind_trial)

- Group results and stats (if available)
   + [All](#vel_dyn_group_all)
   + [Trial by trial](#vel_dyn_group_trial)

#### KV-Diff
- Individual results
   + [All](#vel_acc_dyn_ind_all)
   + [Trial by trial](#vel_acc_dyn_ind_trial)

- Group results and stats (if available)
   + [All](#vel_acc_dyn_group_all)
   + [Trial by trial](#vel_acc_dyn_group_trial)

### [Articulation](#vel_articulation)
- Individual results
   + [All](#vel_art_ind_all)
   + [Trial by trial](#vel_art_ind_trial)

- Group results and stats (if available)
   + [All](#vel_art_group_all)
   + [Trial by trial](#vel_art_group_trial)

#### KV-Diff
- Individual results
   + [All](#vel_acc_art_ind_all)
   + [Trial by trial](#vel_acc_art_ind_trial)

- Group results and stats (if available)
   + [All](#vel_acc_art_group_all)
   + [Trial by trial](#vel_acc_art_group_trial)

## Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato)
- Dependent variables: Key Velocity (KV), KV difference

### Main predictions
Participants will exaggerate forte and piano more while teaching.

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("effsize")) {install.packages("effsize"); require("effsize")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
df_vel <- read.csv("./trimmed/data_vel.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
df_vel_acc <- read.csv("./trimmed/data_vel_acc.csv", header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
df_vel$SubNr <- as.factor(df_vel$SubNr)
df_vel_acc$SubNr <- as.factor(df_vel_acc$SubNr)

# Include only articulation
df_vel_art <- df_vel %>% dplyr::filter(Skill == "articulation")
df_vel_acc_art <- df_vel_acc %>% dplyr::filter(Skill == "articulation")

# Include only dynamics
df_vel_dyn <- df_vel %>% dplyr::filter(Skill == "dynamics")
df_vel_acc_dyn <- df_vel_acc %>% dplyr::filter(Skill == "dynamics")
```

## Key Velocity (KV) - Dynamics
### 1. All - individual{#vel_dyn_ind_all}
- Average KV for each participant.

```{r vel_dyn, echo = FALSE}
# For each individual
vel_dyn <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent, data = df_vel_dyn,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_dyn <- cbind(vel_dyn[,1:4], as.data.frame(vel_dyn[,5]))
# Change colnames
colnames(vel_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(vel_dyn)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

```{r vel_dyn_bar,  echo = FALSE}
p_vel_dyn <- ggplot(data = vel_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
p_vel_dyn
```

### 2. All - group{#vel_dyn_group_all}
- Average KV for each condition.
- Hypothesis: Participants will exaggerate forte and piano when teaching.

```{r vel_dyn_all, echo = FALSE}
vel_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_dyn,
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_dyn_all <- cbind(vel_dyn_all[,1:3], as.data.frame(vel_dyn_all[,4]))
# Change colnames
colnames(vel_dyn_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(vel_dyn_all)
```

#### Plot
##### Bar chart
- Error bars show SEM of the mean

```{r vel_dyn_all_bar, echo = FALSE}　
p_vel_dyn_all <- ggplot(data = vel_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KV (1-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
p_vel_dyn_all
```

##### Box plot

```{r vel_dyn_all_box, echo = FALSE}
p_vel_dyn_all_box <- ggplot(data = vel_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
p_vel_dyn_all_box
```

#### Statistics
##### Normality check
```{r normality1, echo = FALSE}
# Reduce unused factors
vel_dyn$Subcomponent <- factor(vel_dyn$Subcomponent)
vel_dyn_norm <- by(vel_dyn$Mean, list(vel_dyn$Condition, vel_dyn$Subcomponent), shapiro.test)
print(vel_dyn_norm)
```

##### Two-way ANOVA (ez)
```{r stats1, echo = FALSE}
vel_dyn_aov <- ezANOVA(
  data = df_vel_dyn
  , dv = .(Velocity)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_dyn_aov$ANOVA$F <- round(vel_dyn_aov$ANOVA$F, 4)
vel_dyn_aov$ANOVA$p <- round(vel_dyn_aov$ANOVA$p, 4)
print(vel_dyn_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats2, echo = FALSE}
vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
print(summary(vel_dyn_aov_2))
```

### 3. Trial by trial - individual{#vel_dyn_ind_trial}
- Average KV for each trial for each participant.

```{r vel_dyn_trial, echo = FALSE}
# For each individual
vel_dyn_trial <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*TrialNr, data = df_vel_dyn,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_dyn_trial <- cbind(vel_dyn_trial[,1:5], as.data.frame(vel_dyn_trial[,6]))
# Change colnames
colnames(vel_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(vel_dyn_trial)
```

#### Plot
##### Line graph
```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
p_vel_dyn_trial_forte <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_forte
```

```{r vel_dyn_trial_line_piano, echo = FALSE, fig.height = 8}
p_vel_dyn_trial_piano <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_piano
```

### 4. Trial by trial - group{#vel_dyn_group_trial}
- Average KV for each trial for each condition.

```{r vel_dyn_trial_all, echo = FALSE}
# Group mean
vel_dyn_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_dyn_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_dyn_trial_all <- cbind(vel_dyn_trial_all[,1:4], as.data.frame(vel_dyn_trial_all[,5]))
# Change colnames
colnames(vel_dyn_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(vel_dyn_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean

```{r vel_dyn_trial_all_line_forte, echo = FALSE}
p_vel_dyn_trial_all_forte <- ggplot(data = subset(vel_dyn_trial_all, vel_dyn_trial_all$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_all_forte
```

```{r vel_dyn_trial_all_line_piano, echo = FALSE}
p_vel_dyn_trial_all_piano <- ggplot(data = subset(vel_dyn_trial_all, vel_dyn_trial_all$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_all_piano
```

## Key Velocity Difference (Difference) - Dynamics
### 1. All - individual{#vel_acc_dyn_ind_all}
- Average KV-Diff for each participant.

```{r vel_acc_dyn, echo = FALSE}
# For each individual
vel_acc_dyn <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_acc_dyn <- cbind(vel_acc_dyn[,1:4], as.data.frame(vel_acc_dyn[,5]))
# Change colnames
colnames(vel_acc_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(vel_acc_dyn)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

```{r vel_acc_dyn_bar,  echo = FALSE}
p_vel_acc_dyn <- ggplot(data = vel_acc_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
p_vel_acc_dyn
```

### 2. All - group{#vel_acc_dyn_group_all}
- Average KV-Diff for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r vel_acc_dyn_all, echo = FALSE}
vel_acc_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_acc_dyn,
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_acc_dyn_all <- cbind(vel_acc_dyn_all[,1:3], as.data.frame(vel_acc_dyn_all[,4]))
# Change colnames
colnames(vel_acc_dyn_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(vel_acc_dyn_all)
```

```{r vel_acc_dyn_all_bar,  echo = FALSE}
p_vel_acc_dyn_all <- ggplot(data = vel_acc_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Difference", subtitle = "KV: Dynamics")
p_vel_acc_dyn_all
```

##### Box plot

```{r vel_acc_dyn_all_box, echo = FALSE}
p_vel_acc_dyn_all_box <- ggplot(data = vel_acc_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_acc_dyn_all_box
```

#### Statistics
##### Normality check
```{r normality2, echo = FALSE}
# Reduce unused factors
vel_acc_dyn$Subcomponent <- factor(vel_acc_dyn$Subcomponent)
vel_acc_dyn_norm <- by(vel_acc_dyn$Mean, list(vel_acc_dyn$Condition, vel_acc_dyn$Subcomponent), shapiro.test)
print(vel_acc_dyn_norm)
```

##### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
vel_acc_dyn_aov <- ezANOVA(
  data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF")
  , dv = .(Acc)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_acc_dyn_aov$ANOVA$F <- round(vel_acc_dyn_aov$ANOVA$F, 4)
vel_acc_dyn_aov$ANOVA$p <- round(vel_acc_dyn_aov$ANOVA$p, 4)
print(vel_acc_dyn_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
vel_acc_dyn_aov_2 <- with(vel_acc_dyn,
                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
print(summary(vel_acc_dyn_aov_2))
```

### 3. Trial by trial - individual{#vel_acc_dyn_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel_acc_dyn_trial, echo = FALSE}
# For each individual
vel_acc_dyn_trial <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_acc_dyn_trial <- cbind(vel_acc_dyn_trial[,1:5], as.data.frame(vel_acc_dyn_trial[,6]))
# Change colnames
colnames(vel_acc_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(vel_acc_dyn_trial)
```

#### Plot
##### Line graph
```{r vel_acc_dyn_trial_line_ftop, echo = FALSE, fig.height = 8}
p_vel_acc_dyn_trial_ftop <- ggplot(data = subset(vel_acc_dyn_trial, vel_acc_dyn_trial$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_ftop
```

```{r vel_acc_dyn_trial_line_ptof, echo = FALSE, fig.height = 8}
p_vel_acc_dyn_trial_ptof <- ggplot(data = subset(vel_acc_dyn_trial, vel_acc_dyn_trial$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_ptof
```

### 4. Trial by trial - group{#vel_acc_dyn_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel_acc_dyn_trial_all, echo = FALSE}
# Group mean
vel_acc_dyn_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_acc_dyn_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_acc_dyn_trial_all <- cbind(vel_acc_dyn_trial_all[,1:4], as.data.frame(vel_acc_dyn_trial_all[,5]))
# Change colnames
colnames(vel_acc_dyn_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(vel_acc_dyn_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean

```{r vel_acc_dyn_trial_all_line_ftop, echo = FALSE}
p_vel_acc_dyn_trial_all_ftop <- ggplot(data = subset(vel_acc_dyn_trial_all, vel_acc_dyn_trial_all$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_all_ftop
```

```{r vel_acc_dyn_trial_all_line_ptof, echo = FALSE}
p_vel_acc_dyn_trial_all_ptof <- ggplot(data = subset(vel_acc_dyn_trial_all, vel_acc_dyn_trial_all$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_all_ptof
```

###########
###########

## Key Velocity (KV) - Articulation
### 1. All - individual{#vel_art_ind_all}
- Average KV for each participant.

```{r vel_art, echo = FALSE}
# For each individual
vel_art <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent, data = df_vel_art,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_art <- cbind(vel_art[,1:4], as.data.frame(vel_art[,5]))
# Change colnames
colnames(vel_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(vel_art)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

```{r vel_art_bar,  echo = FALSE}
p_vel_art <- ggplot(data = vel_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
p_vel_art
```

### 2. All - group{#vel_art_group_all}
- Average KV for each condition.

```{r vel_art_all, echo = FALSE}
vel_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_art,
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_art_all <- cbind(vel_art_all[,1:3], as.data.frame(vel_art_all[,4]))
# Change colnames
colnames(vel_art_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(vel_art_all)
```

#### Plot
##### Bar chart
- Error bars show SEM of the mean

```{r vel_art_all_bar, echo = FALSE}　
p_vel_art_all <- ggplot(data = vel_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KV (1-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
p_vel_art_all
```

##### Box plot

```{r vel_art_all_box, echo = FALSE}
p_vel_art_all_box <- ggplot(data = vel_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation")
p_vel_art_all_box
```

#### Statistics
##### Normality check
```{r normality3, echo = FALSE}
# Reduce unused factors
vel_art$Subcomponent <- factor(vel_art$Subcomponent)
vel_art_norm <- by(vel_art$Mean, list(vel_art$Condition, vel_art$Subcomponent), shapiro.test)
print(vel_art_norm)
```

##### Two-way ANOVA (ez)
```{r stats5, echo = FALSE}
vel_art_aov <- ezANOVA(
  data = df_vel_art
  , dv = .(Velocity)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_art_aov$ANOVA$F <- round(vel_art_aov$ANOVA$F, 4)
vel_art_aov$ANOVA$p <- round(vel_art_aov$ANOVA$p, 4)
print(vel_art_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats6, echo = FALSE}
vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
print(summary(vel_art_aov_2))
```

### 3. Trial by trial - individual{#vel_art_ind_trial}
- Average KV for each trial for each participant.

```{r vel_art_trial, echo = FALSE}
# For each individual
vel_art_trial <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*TrialNr, data = df_vel_art,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_art_trial <- cbind(vel_art_trial[,1:5], as.data.frame(vel_art_trial[,6]))
# Change colnames
colnames(vel_art_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(vel_art_trial)
```

#### Plot
##### Line graph
```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 8}
p_vel_art_trial_legato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_legato
```

```{r vel_art_trial_line_piano, echo = FALSE, fig.height = 8}
p_vel_art_trial_staccato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_staccato
```

### 4. Trial by trial - group{#vel_art_group_trial}
- Average KV for each trial for each condition.

```{r vel_art_trial_all, echo = FALSE}
# Group mean
vel_art_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_art_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_art_trial_all <- cbind(vel_art_trial_all[,1:4], as.data.frame(vel_art_trial_all[,5]))
# Change colnames
colnames(vel_art_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(vel_art_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean

```{r vel_art_trial_all_line_legato, echo = FALSE}
p_vel_art_trial_all_legato <- ggplot(data = subset(vel_art_trial_all, vel_art_trial_all$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_all_legato
```

```{r vel_art_trial_all_line_staccato, echo = FALSE}
p_vel_art_trial_all_staccato <- ggplot(data = subset(vel_art_trial_all, vel_art_trial_all$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_all_staccato
```

## Key Velocity Difference (Difference) - Articulation
### 1. All - individual{#vel_acc_art_ind_all}
- Average KV-Diff for each participant.

```{r vel_acc_art, echo = FALSE}
# For each individual
vel_acc_art <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_acc_art <- cbind(vel_acc_art[,1:4], as.data.frame(vel_acc_art[,5]))
# Change colnames
colnames(vel_acc_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(vel_acc_art)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

```{r vel_acc_art_bar,  echo = FALSE}
p_vel_acc_art <- ggplot(data = vel_acc_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulaion")
p_vel_acc_art
```

### 2. All - group{#vel_acc_art_group_all}
- Average KV-Diff for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r vel_acc_art_all, echo = FALSE}
vel_acc_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_acc_art,
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_acc_art_all <- cbind(vel_acc_art_all[,1:3], as.data.frame(vel_acc_art_all[,4]))
# Change colnames
colnames(vel_acc_art_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(vel_acc_art_all)
```

```{r vel_acc_art_all_bar,  echo = FALSE}
p_vel_acc_art_all <- ggplot(data = vel_acc_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Difference", subtitle = "KV: Articulation")
p_vel_acc_art_all
```

##### Box plot

```{r vel_acc_art_all_box, echo = FALSE}
p_vel_acc_art_all_box <- ggplot(data = vel_acc_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_acc_art_all_box
```

#### Statistics
##### Normality check
```{r normality4, echo = FALSE}
# Reduce unused factors
vel_acc_art$Subcomponent <- factor(vel_acc_art$Subcomponent)
vel_acc_art_norm <- by(vel_acc_art$Mean, list(vel_acc_art$Condition, vel_acc_art$Subcomponent), shapiro.test)
print(vel_acc_art_norm)
```

##### Two-way ANOVA (ez)
```{r stats7, echo = FALSE}
vel_acc_art_aov <- ezANOVA(
  data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL")
  , dv = .(Acc)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_acc_art_aov$ANOVA$F <- round(vel_acc_art_aov$ANOVA$F, 4)
vel_acc_art_aov$ANOVA$p <- round(vel_acc_art_aov$ANOVA$p, 4)
print(vel_acc_art_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats8, echo = FALSE}
vel_acc_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_acc_art)
print(summary(vel_acc_art_aov_2))
```

### 3. Trial by trial - individual{#kot_dyn_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel_acc_art_trial, echo = FALSE}
# For each individual
vel_acc_art_trial <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_acc_art_trial <- cbind(vel_acc_art_trial[,1:5], as.data.frame(vel_acc_art_trial[,6]))
# Change colnames
colnames(vel_acc_art_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(vel_acc_art_trial)
```

#### Plot
##### Line graph
```{r vel_acc_art_trial_line_ltos, echo = FALSE, fig.height = 8}
p_vel_acc_art_trial_ltos <- ggplot(data = subset(vel_acc_art_trial, vel_acc_art_trial$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_ltos
```

```{r vel_acc_art_trial_line_stol, echo = FALSE, fig.height = 8}
p_vel_acc_art_trial_stol <- ggplot(data = subset(vel_acc_art_trial, vel_acc_art_trial$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Staccato to Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_stol
```

### 4. Trial by trial - group{#vel_acc_art_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel_acc_art_trial_all, echo = FALSE}
# Group mean
vel_acc_art_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_acc_art_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_acc_art_trial_all <- cbind(vel_acc_art_trial_all[,1:4], as.data.frame(vel_acc_art_trial_all[,5]))
# Change colnames
colnames(vel_acc_art_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(vel_acc_art_trial_all)
```

#### Plot
##### Line graph
- Error bars show SEM of the mean

```{r vel_acc_art_trial_all_line_ltos, echo = FALSE}
p_vel_acc_art_trial_all_ltos <- ggplot(data = subset(vel_acc_art_trial_all, vel_acc_art_trial_all$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_all_ltos
```

```{r vel_acc_art_trial_all_line_stol, echo = FALSE}
p_vel_acc_art_trial_all_stol <- ggplot(data = subset(vel_acc_art_trial_all, vel_acc_art_trial_all$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_all_stol
```

### 2019-10-04 script/RNotebook/ioi.Rmd
add order analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+- Order analysis (teaching/performing first?)
+  + [All](#art_order_group_all)
+
 ### [Dynamics](#ioi-dynamics)
@@
   + [Variability trial-by-trial](#dyn_group_var_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#dyn_order_group_all)
 
@@
 - Error bars show SEM of the mean
+
 ```{r ioi_dyn_var_all_bar, echo = FALSE}
@@
 ```
+
+###########
+###########
+
+## IOI-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+### 1. All{#art_order_group_all}
+```{r ioi_art_order_all, echo = FALSE}
+# For each participant
+ioi_art_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_art,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+ioi_art_order <- cbind(ioi_art_order[,1:4], as.data.frame(ioi_art_order[,5]))
+# Change colnames
+colnames(ioi_art_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD")
+
+# Group mean
+ioi_art_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_art_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_art_order_all <- cbind(ioi_art_order_all[,1:3], as.data.frame(ioi_art_order_all[,4]))
+# Change colnames
+colnames(ioi_art_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM")
+print(ioi_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r ioi_art_order_all_bar, echo = FALSE}
+p_ioi_art_order_all <- ggplot(data = ioi_art_order_all, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge()) +
+  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  facet_wrap(Order ~ .) +
+  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) +
+  theme(legend.position = "None")
+p_ioi_art_order_all
+```
+
+##### Box plot
+```{r ioi_art_order_all_box, echo = FALSE}
+p_ioi_art_order_all_box <- ggplot(data = ioi_art_order, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
+  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  facet_wrap(Order ~ .) +
+  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") +
+  theme(legend.position = "None")
+p_ioi_art_order_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality7, echo = FALSE}
+# Normality check
+ioi_art_order_norm <- by(ioi_art_order$Mean, list(ioi_art_order$Condition, ioi_art_order$Order), shapiro.test)
+print(ioi_art_order_norm)
+```
+
+##### Two-way ANOVA (ez)
+```{r stats13, echo = FALSE}
+ioi_art_order_aov <- ezANOVA(
+  data = df_ioi_art
+  , dv = .(IOI)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition)
+  , type = 3
+  , detailed = TRUE
+)
+ioi_art_order_aov$ANOVA$F <- round(ioi_art_order_aov$ANOVA$F, 4)
+ioi_art_order_aov$ANOVA$p <- round(ioi_art_order_aov$ANOVA$p, 4)
+print(ioi_art_order_aov)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats14, echo = FALSE}
+ioi_art_order_aov_2 <- aov(Mean ~ Condition*Order + Error(SubNr/Condition) + Order, data = ioi_art_order)
+print(summary(ioi_art_order_aov_2))
+```
+
+## IOI-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+### 1. All{#dyn_order_group_all}
+```{r ioi_dyn_order_all, echo = FALSE}
+# For each pdynicipant
+ioi_dyn_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_dyn,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+ioi_dyn_order <- cbind(ioi_dyn_order[,1:4], as.data.frame(ioi_dyn_order[,5]))
+# Change colnames
+colnames(ioi_dyn_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD")
+
+# Group mean
+ioi_dyn_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_dyn_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_dyn_order_all <- cbind(ioi_dyn_order_all[,1:3], as.data.frame(ioi_dyn_order_all[,4]))
+# Change colnames
+colnames(ioi_dyn_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM")
+print(ioi_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r ioi_dyn_order_all_bar, echo = FALSE}
+p_ioi_dyn_order_all <- ggplot(data = ioi_dyn_order_all, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge()) +
+  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  facet_wrap(Order ~ .) +
+  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(120, 210)) +
+  theme(legend.position = "None")
+p_ioi_dyn_order_all
+```
+
+##### Box plot
+```{r ioi_dyn_order_all_box, echo = FALSE}
+p_ioi_dyn_order_all_box <- ggplot(data = ioi_dyn_order, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
+  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  facet_wrap(Order ~ .) +
+  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") +
+  theme(legend.position = "None")
+p_ioi_dyn_order_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality8, echo = FALSE}
+# Normality check
+ioi_dyn_order_norm <- by(ioi_dyn_order$Mean, list(ioi_dyn_order$Condition, ioi_dyn_order$Order), shapiro.test)
+print(ioi_dyn_order_norm)
+```
+
+##### Two-way ANOVA (ez)
+```{r stats15, echo = FALSE}
+ioi_dyn_order_aov <- ezANOVA(
+  data = df_ioi_dyn
+  , dv = .(IOI)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition)
+  , type = 3
+  , detailed = TRUE
+)
+ioi_dyn_order_aov$ANOVA$F <- round(ioi_dyn_order_aov$ANOVA$F, 4)
+ioi_dyn_order_aov$ANOVA$p <- round(ioi_dyn_order_aov$ANOVA$p, 4)
+print(ioi_dyn_order_aov)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats16, echo = FALSE}
+ioi_dyn_order_aov_2 <- aov(Mean ~ Condition*Order + Error(SubNr/Condition) + Order, data = ioi_dyn_order)
+print(summary(ioi_dyn_order_aov_2))
+```

### 2019-10-04 script/RNotebook/articulation.Rmd
add order analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+- Order analysis (teaching/performing first?)
+  + [All](#kot_art_order_group_all)
+
 #### KOR
@@
 
-- [Summary](#art_summary)
+- Order analysis (teaching/performing first?)
+  + [All](#kor_art_order_group_all)
 
@@
 
+- Order analysis (teaching/performing first?)
+  + [All](#kot_dyn_order_group_all)
+
 #### KOR
@@
 
-- [Summary](#dyn_summary)
+- Order analysis (teaching/performing first?)
+  + [All](#kor_dyn_order_group_all)
 
@@
 
+# For each individual, each block, each trial
+ioi <- aggregate(IOI~SubNr*BlockNr*TrialNr*Condition*Skill, data = df_ioi,
+                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
+ioi <- cbind(ioi[,1:5], as.data.frame(ioi[,6]))
+# Change colnames
+colnames(ioi) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "N", "Mean", "SD")
+
 # Calculate KOR for each interval
+
 df_kor <- data.frame()
@@
   facet_grid(Subcomponent ~ .) +
-  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
+  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
 p_kot_dyn
@@
 ```
+
+###########
+###########
+
+## KOT-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+### 1. All{#kot_art_order_group_all}
+```{r kot_art_order_all, echo = FALSE}
+# For each participant
+kot_art_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kot_art_order <- cbind(kot_art_order[,1:5], as.data.frame(kot_art_order[,6]))
+# Change colnames
+colnames(kot_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kot_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_art_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_art_order_all <- cbind(kot_art_order_all[,1:4], as.data.frame(kot_art_order_all[,5]))
+# Change colnames
+colnames(kot_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kot_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kot_art_order_all_bar, echo = FALSE}
+p_kot_art_order_all <- ggplot(data = kot_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
+p_kot_art_order_all
+```
+
+##### Box plot
+```{r kot_art_order_all_box, echo = FALSE}
+p_kot_art_order_all_box <- ggplot(data = kot_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT", subtitle = "KOT: Articulation")
+p_kot_art_order_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality5, echo = FALSE}
+# Normality check
+kot_art_order$Subcomponent <- factor(kot_art_order$Subcomponent)
+kot_art_order_norm <- by(kot_art_order$Mean, list(kot_art_order$Condition, kot_art_order$Subcomponent, kot_art_order$Order), shapiro.test)
+print(kot_art_order_norm)
+```
+
+##### Three-way ANOVA (ez)
+```{r stats9, echo = FALSE}
+kot_art_order_aov <- ezANOVA(
+  data = df_kot_art
+  , dv = .(KOT)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kot_art_order_aov$ANOVA$F <- round(kot_art_order_aov$ANOVA$F, 4)
+kot_art_order_aov$ANOVA$p <- round(kot_art_order_aov$ANOVA$p, 4)
+print(kot_art_order_aov)
+```
+
+## KOR-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+### 1. All{#kor_art_order_group_all}
+```{r kor_art_order_all, echo = FALSE}
+# For each participant
+kor_art_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kor_art_order <- cbind(kor_art_order[,1:5], as.data.frame(kor_art_order[,6]))
+# Change colnames
+colnames(kor_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kor_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_art_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kor_art_order_all <- cbind(kor_art_order_all[,1:4], as.data.frame(kor_art_order_all[,5]))
+# Change colnames
+colnames(kor_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kor_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kor_art_order_all_bar, echo = FALSE}
+p_kor_art_order_all <- ggplot(data = kor_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Articulation")
+p_kor_art_order_all
+```
+
+##### Box plot
+```{r kor_art_order_all_box, echo = FALSE}
+p_kor_art_order_all_box <- ggplot(data = kor_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Articulation")
+p_kor_art_order_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality6, echo = FALSE}
+# Normality check
+kor_art_order$Subcomponent <- factor(kor_art_order$Subcomponent)
+kor_art_order_norm <- by(kor_art_order$Mean, list(kor_art_order$Condition, kor_art_order$Subcomponent, kor_art_order$Order), shapiro.test)
+print(kor_art_order_norm)
+```
+
+##### Three-way ANOVA (ez)
+```{r stats10, echo = FALSE}
+kor_art_order_aov <- ezANOVA(
+  data = df_kor_art
+  , dv = .(KOR)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kor_art_order_aov$ANOVA$F <- round(kor_art_order_aov$ANOVA$F, 4)
+kor_art_order_aov$ANOVA$p <- round(kor_art_order_aov$ANOVA$p, 4)
+print(kor_art_order_aov)
+```
+
+## KOT-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+### 1. All{#kot_dyn_order_group_all}
+```{r kot_dyn_order_all, echo = FALSE}
+# For each participant
+kot_dyn_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kot_dyn_order <- cbind(kot_dyn_order[,1:5], as.data.frame(kot_dyn_order[,6]))
+# Change colnames
+colnames(kot_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kot_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_dyn_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_dyn_order_all <- cbind(kot_dyn_order_all[,1:4], as.data.frame(kot_dyn_order_all[,5]))
+# Change colnames
+colnames(kot_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kot_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kot_dyn_order_all_bar, echo = FALSE}
+p_kot_dyn_order_all <- ggplot(data = kot_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
+p_kot_dyn_order_all
+```
+
+##### Box plot
+```{r kot_dyn_order_all_box, echo = FALSE}
+p_kot_dyn_order_all_box <- ggplot(data = kot_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOY", subtitle = "KOT: Dynamics")
+p_kot_dyn_order_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality7, echo = FALSE}
+# Normality check
+kot_dyn_order$Subcomponent <- factor(kot_dyn_order$Subcomponent)
+kot_dyn_order_norm <- by(kot_dyn_order$Mean, list(kot_dyn_order$Condition, kot_dyn_order$Subcomponent, kot_dyn_order$Order), shapiro.test)
+print(kot_dyn_order_norm)
+```
+
+##### Three-way ANOVA (ez)
+```{r stats11, echo = FALSE}
+kot_dyn_order_aov <- ezANOVA(
+  data = df_kot_dyn
+  , dv = .(KOT)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kot_dyn_order_aov$ANOVA$F <- round(kot_dyn_order_aov$ANOVA$F, 4)
+kot_dyn_order_aov$ANOVA$p <- round(kot_dyn_order_aov$ANOVA$p, 4)
+print(kot_dyn_order_aov)
+```
+
+## KOR-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+### 1. All{#kor_dyn_order_group_all}
+```{r kor_dyn_order_all, echo = FALSE}
+# For each participant
+kor_dyn_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kor_dyn_order <- cbind(kor_dyn_order[,1:5], as.data.frame(kor_dyn_order[,6]))
+# Change colnames
+colnames(kor_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kor_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_dyn_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kor_dyn_order_all <- cbind(kor_dyn_order_all[,1:4], as.data.frame(kor_dyn_order_all[,5]))
+# Change colnames
+colnames(kor_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kor_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kor_dyn_order_all_bar, echo = FALSE}
+p_kor_dyn_order_all <- ggplot(data = kor_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Dynamics")
+p_kor_dyn_order_all
+```
+
+##### Box plot
+```{r kor_dyn_order_all_box, echo = FALSE}
+p_kor_dyn_order_all_box <- ggplot(data = kor_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Dynamics")
+p_kor_dyn_order_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality8, echo = FALSE}
+# Normality check
+kor_dyn_order$Subcomponent <- factor(kor_dyn_order$Subcomponent)
+kor_dyn_order_norm <- by(kor_dyn_order$Mean, list(kor_dyn_order$Condition, kor_dyn_order$Subcomponent, kor_dyn_order$Order), shapiro.test)
+print(kor_dyn_order_norm)
+```
+
+##### Three-way ANOVA (ez)
+```{r stats12, echo = FALSE}
+kor_dyn_order_aov <- ezANOVA(
+  data = df_kor_dyn
+  , dv = .(KOR)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kor_dyn_order_aov$ANOVA$F <- round(kor_dyn_order_aov$ANOVA$F, 4)
+kor_dyn_order_aov$ANOVA$p <- round(kor_dyn_order_aov$ANOVA$p, 4)
+print(kor_dyn_order_aov)
+```

### 2019-10-05 script/RNotebook/dynamics.Rmd
add order analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+- Order analysis (teaching/performing first?)
+  + [All](#vel_dyn_order_group_all)
+
 #### KV-Diff
@@
 
+- Order analysis (teaching/performing first?)
+  + [All](#vel_acc_dyn_order_group_all)
+
 ### [Articulation](#vel_articulation)
+
+#### KV
 - Individual results
@@
   + [Trial by trial](#vel_art_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#vel_art_order_group_all)
 
@@
   + [Trial by trial](#vel_acc_art_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#vel_acc_art_order_group_all)
 
@@
                 width = .2, position = position_dodge(.9)) +
-  labs(y = "KV (1-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
 p_vel_dyn_all
@@
 ```
+
+###########
+###########
+
+## KV-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+### 1. All{#vel_dyn_order_group_all}
+```{r vel_dyn_order_all, echo = all}
+# For each participant
+vel_dyn_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_dyn,
+                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_dyn_order <- cbind(vel_dyn_order[,1:5], as.data.frame(vel_dyn_order[,6]))
+# Change colnames
+colnames(vel_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+vel_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_dyn_order,
+                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_dyn_order_all <- cbind(vel_dyn_order_all[,1:4], as.data.frame(vel_dyn_order_all[,5]))
+# Change colnames
+colnames(vel_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(vel_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r vel_dyn_order_all_bar, echo = FALSE}
+p_vel_dyn_order_all <- ggplot(data = vel_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
+p_vel_dyn_order_all
+```
+
+##### Box plot
+```{r vel_dyn_order_all_box, echo = FALSE}
+p_vel_dyn_order_all_box <- ggplot(data = vel_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
+p_vel_dyn_order_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality5, echo = FALSE}
+# Normality check
+vel_dyn_order$Subcomponent <- factor(vel_dyn_order$Subcomponent)
+vel_dyn_order_norm <- by(vel_dyn_order$Mean, list(vel_dyn_order$Condition, vel_dyn_order$Subcomponent, vel_dyn_order$Order), shapiro.test)
+print(vel_dyn_order_norm)
+```
+
+##### Three-way ANOVA (ez)
+```{r stats9, echo = FALSE}
+vel_dyn_order_aov <- ezANOVA(
+  data = df_vel_dyn
+  , dv = .(Velocity)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_dyn_order_aov$ANOVA$F <- round(vel_dyn_order_aov$ANOVA$F, 4)
+vel_dyn_order_aov$ANOVA$p <- round(vel_dyn_order_aov$ANOVA$p, 4)
+print(vel_dyn_order_aov)
+```
+
+## KV-Diff-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+### 1. All{#vel_acc_dyn_order_group_all}
+```{r vel_acc_dyn_order_all, echo = all}
+# For each participant
+vel_acc_dyn_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
+                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_acc_dyn_order <- cbind(vel_acc_dyn_order[,1:5], as.data.frame(vel_acc_dyn_order[,6]))
+# Change colnames
+colnames(vel_acc_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+vel_acc_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_dyn_order,
+                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_acc_dyn_order_all <- cbind(vel_acc_dyn_order_all[,1:4], as.data.frame(vel_acc_dyn_order_all[,5]))
+# Change colnames
+colnames(vel_acc_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(vel_acc_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r vel_acc_dyn_order_all_bar, echo = FALSE}
+p_vel_acc_dyn_order_all <- ggplot(data = vel_acc_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
+p_vel_acc_dyn_order_all
+```
+
+##### Box plot
+```{r vel_acc_dyn_order_all_box, echo = FALSE}
+p_vel_acc_dyn_order_all_box <- ggplot(data = vel_acc_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
+p_vel_acc_dyn_order_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality6, echo = FALSE}
+# Normality check
+vel_acc_dyn_order$Subcomponent <- factor(vel_acc_dyn_order$Subcomponent)
+vel_acc_dyn_order_norm <- by(vel_acc_dyn_order$Mean, list(vel_acc_dyn_order$Condition, vel_acc_dyn_order$Subcomponent, vel_acc_dyn_order$Order), shapiro.test)
+print(vel_acc_dyn_order_norm)
+```
+
+##### Three-way ANOVA (ez)
+```{r stats10, echo = FALSE}
+vel_acc_dyn_order_aov <- ezANOVA(
+  data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF")
+  , dv = .(Acc)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_acc_dyn_order_aov$ANOVA$F <- round(vel_acc_dyn_order_aov$ANOVA$F, 4)
+vel_acc_dyn_order_aov$ANOVA$p <- round(vel_acc_dyn_order_aov$ANOVA$p, 4)
+print(vel_acc_dyn_order_aov)
+```
+
+## KV-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+### 1. All{#vel_art_order_group_all}
+```{r vel_art_order_all, echo = all}
+# For each participant
+vel_art_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_art,
+                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_art_order <- cbind(vel_art_order[,1:5], as.data.frame(vel_art_order[,6]))
+# Change colnames
+colnames(vel_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+vel_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_art_order,
+                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_art_order_all <- cbind(vel_art_order_all[,1:4], as.data.frame(vel_art_order_all[,5]))
+# Change colnames
+colnames(vel_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(vel_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r vel_art_order_all_bar, echo = FALSE}
+p_vel_art_order_all <- ggplot(data = vel_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
+p_vel_art_order_all
+```
+
+##### Box plot
+```{r vel_art_order_all_box, echo = FALSE}
+p_vel_art_order_all_box <- ggplot(data = vel_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
+p_vel_art_order_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality7, echo = FALSE}
+# Normality check
+vel_art_order$Subcomponent <- factor(vel_art_order$Subcomponent)
+vel_art_order_norm <- by(vel_art_order$Mean, list(vel_art_order$Condition, vel_art_order$Subcomponent, vel_art_order$Order), shapiro.test)
+print(vel_art_order_norm)
+```
+
+##### Three-way ANOVA (ez)
+```{r stats11, echo = FALSE}
+vel_art_order_aov <- ezANOVA(
+  data = df_vel_art
+  , dv = .(Velocity)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_art_order_aov$ANOVA$F <- round(vel_art_order_aov$ANOVA$F, 4)
+vel_art_order_aov$ANOVA$p <- round(vel_art_order_aov$ANOVA$p, 4)
+print(vel_art_order_aov)
+```
+
+## KV-Diff-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+### 1. All{#vel_acc_art_order_group_all}
+```{r vel_acc_art_order_all, echo = all}
+# For each participant
+vel_acc_art_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
+                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_acc_art_order <- cbind(vel_acc_art_order[,1:5], as.data.frame(vel_acc_art_order[,6]))
+# Change colnames
+colnames(vel_acc_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+vel_acc_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_art_order,
+                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_acc_art_order_all <- cbind(vel_acc_art_order_all[,1:4], as.data.frame(vel_acc_art_order_all[,5]))
+# Change colnames
+colnames(vel_acc_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(vel_acc_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r vel_acc_art_order_all_bar, echo = FALSE}
+p_vel_acc_art_order_all <- ggplot(data = vel_acc_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
+p_vel_acc_art_order_all
+```
+
+##### Box plot
+```{r vel_acc_art_order_all_box, echo = FALSE}
+p_vel_acc_art_order_all_box <- ggplot(data = vel_acc_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
+p_vel_acc_art_order_all_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality8, echo = FALSE}
+# Normality check
+vel_acc_art_order$Subcomponent <- factor(vel_acc_art_order$Subcomponent)
+vel_acc_art_order_norm <- by(vel_acc_art_order$Mean, list(vel_acc_art_order$Condition, vel_acc_art_order$Subcomponent, vel_acc_art_order$Order), shapiro.test)
+print(vel_acc_art_order_norm)
+```
+
+##### Three-way ANOVA (ez)
+```{r stats12, echo = FALSE}
+vel_acc_art_order_aov <- ezANOVA(
+  data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL")
+  , dv = .(Acc)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_acc_art_order_aov$ANOVA$F <- round(vel_acc_art_order_aov$ANOVA$F, 4)
+vel_acc_art_order_aov$ANOVA$p <- round(vel_acc_art_order_aov$ANOVA$p, 4)
+print(vel_acc_art_order_aov)
+```

### 2019-10-05 script/RNotebook/questionnaire.Rmd
add summary for a questionnaire
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of demographic information

# Outline - Quesionnaire

```{r read, echo = FALSE}
# Read csv files
df <- read.csv("./questionnaire.csv", header = T, sep = ",", dec = ".")
df_exc <- read.csv("./filtered/data_errorRate.csv", header = T, sep = ",", dec = ".")

# Change colnames
colnames(df) <- c("TimeStamp", "SubNr", "Age", "Gender", "FirstLang", "SecondLang", "DominantHand", "NormalVision", "NormalHearing", "Desc1", "Desc2",
                  "Instruments1", "FormalTraining1", "TotalPractice", "Start1", "End1", "PracticeHours", # Experience in musical instruments
                  "Teaching", "Instruments2", "TotalTeaching1", "FormalTraining2", "Training", "AgeGroup", "Start2", "End2", "Like1", # Experience in teaching instruments
                  "Other", "Subject", "TotalTeaching2", "Like2") # Experinence in teaching in general

# Extract the total number of years of training in piano (input manually based on the original csv file)
df$PianoTotalPractice <- c(14, 20, 27, 21, 10, 16, 9, 13, 25, 14, 13, 11, 6, 4, 6, 8, 21, 15, 18, 7, 11, 12, 12, 18, 12, 5, 5, 10, 16, 10, 8, 22, 10, 5, 4, 14)

print(df)
```

### Descriptive stats for participants' demographics
#### Number of excluded participants
```{r excluded, echo = FALSE}
# Descriptive statistics for participants' demographics
total <- nrow(df)
# Number of excluded participants
exp_error <- length(c(8)) # SubNr 8 - due to an experimental error
pitch_error <- length(df_exc$SubNr[df_exc$LessThan10 == "exclude"])
tempo_dev <- total-exp_error-pitch_error-length(include)
total_exc <- exp_error+pitch_error+tempo_dev

excluded_participants <- data.frame(cbind(total, total_exc, exp_error, pitch_error, tempo_dev))
colnames(excluded_participants) <- c("Total participants", "Total excluded", "Experimental error", "Pitch error", "Tempo deviation")
print(excluded_participants)
```

#### Number of included participants
```{r included, echo = FALSE}
# Data frame with only included participants
include <- df_exc$SubNr[df_exc$Final == "include"]
df_demo <- data.frame()
for (subnr in include){
  df_current <- df %>% dplyr::filter(SubNr == subnr)
  df_demo <- rbind(df_demo, df_current)
}

# Number of included participants
included <- nrow(df_demo)
male <- nrow(subset(df_demo, df_demo$Gender == "Male"))
female <- nrow(subset(df_demo, df_demo$Gender == "Female"))
experience_m <- round(mean(df_demo$PianoTotalPractice), 0)
experience_sd <- round(sd(df_demo$PianoTotalPractice)*sqrt((length(df_demo$PianoTotalPractice)-1)/length(df_demo$PianoTotalPractice)), 2) #biased sd for the current data

included_participants <- data.frame(cbind(total, included, male, female, experience_m, experience_sd))
colnames(included_participants) <- c("Total", "Included participants", "Male", "Female", "Average experience", "SD of Average experience")
print(included_participants)
```

#### Error Rate of excluded participants
```{r errorRate_excluded, echo = FALSE}
# Information about excluded participants due to errorRate (SubNr 16 was excluded but due to not errorRate but tempo deviation)
errorRate_mean_exc <- mean(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
errorRate_sd_exc <- sd(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*sqrt((length(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])-1)/length(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16]))*100 #biased sd for the current data
errorRate_median_exc <- median(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
errorRate_min_exc <- min(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
errorRate_max_exc <- max(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100

errorRate_excluded <- data.frame(cbind(errorRate_mean_exc, errorRate_sd_exc, errorRate_median_exc, errorRate_min_exc, errorRate_max_exc))
colnames(errorRate_excluded) <- c("ER Mean", "ER SD", "ER Median", "ER Min", "ER Max")
print(errorRate_excluded)
```

#### Error Rate: Normality test for excluded participants
```{r normality_for_excluded, echo = FALSE}
# Normality check
errorRate_excluded_norm <- shapiro.test(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])
print(errorRate_excluded_norm)

# Note: The distributions fo error rates were not normal, so we decided to report median values (plus its range) rather than mean values.
```

#### Error Rate of included participants
```{r er_included, echo = FALSE}
# Information about included participants
errorRate_mean_inc <- mean(df_exc$ErrorRate[df_exc$Final == "include"])*100
errorRate_sd_inc <- sd(df_exc$ErrorRate[df_exc$Final == "include"])*sqrt((length(df_exc$ErrorRate[df_exc$Final == "include"])-1)/length(df_exc$ErrorRate[df_exc$Final == "include"]))*100 #biased sd for the current data
errorRate_median_inc <- median(df_exc$ErrorRate[df_exc$Final == "include"])*100
errorRate_min_inc <- min(df_exc$ErrorRate[df_exc$Final == "include"])*100
errorRate_max_inc <- max(df_exc$ErrorRate[df_exc$Final == "include"])*100

errorRate_included <- data.frame(cbind(errorRate_mean_inc, errorRate_sd_inc, errorRate_median_inc, errorRate_min_inc, errorRate_max_inc))
colnames(errorRate_included) <- c("ER Mean", "ER SD", "ER Median", "ER Min", "ER Max")
print(errorRate_included)
```

#### Error Rate: Normality test for included participants
```{r normality_for_included, echo = FALSE}
# Normality check
errorRate_included_norm <- shapiro.test(df_exc$ErrorRate[df_exc$Final == "include"])
print(errorRate_included_norm)

# Note: The distributions fo error rates were not normal, so we decided to report median values (plus its range) rather than mean values.
```

### 2019-10-17 script/RNotebook/dynamics.Rmd
add comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r vel_dyn_all, echo = FALSE}
+# Group mean
 vel_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_dyn,
@@
 ```{r vel_acc_dyn_all, echo = FALSE}
+# Group mean
 vel_acc_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_acc_dyn,
@@
                 width = .2, position = position_dodge(.9)) +
-  labs(y = "Difference", subtitle = "KV: Dynamics")
+  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
 p_vel_acc_dyn_all
@@
 ```{r vel_art_all, echo = FALSE}
+# Group mean
 vel_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_art,
@@
 ```{r vel_acc_art_all, echo = FALSE}
+# Group mean
 vel_acc_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_acc_art,

### 2019-10-18 script/RNotebook/articulation.Rmd
add posthoc comparisons
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+### Post-hoc (direct comparison)
+```{r post-hoc1, echo = FALSE}
+ttest_legato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"], paired = TRUE)
+ttest_staccato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"], paired = TRUE)
+
+print(ttest_legato)
+print(ttest_staccato)
+```
+
+### Post-hoc (pairwise t test)
+```{r post-hoc2, echo = FALSE}
+# Create a group label for a post-hoc test
+kot_art$Group[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"] <- "Performing-Legato" 
+kot_art$Group[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
+kot_art$Group[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"] <- "Teaching-Legato"
+kot_art$Group[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
+# pairwise comparison
+kot_art_pairwise <- pairwise.t.test(kot_art$Mean, kot_art$Group, paired = TRUE, p.adjust.method = "BH")
+kot_art_pairwise$p.value <- round(kot_art_pairwise$p.value, 4)
+print(kot_art_pairwise)
+```
+
 ### 3. Trial by trial - individual{#kot_art_ind_trial}
@@
 print(summary(kor_art_aov_2))
+```
+
+### Post-hoc (direct comparison)
+```{r post-hoc3, echo = FALSE}
+ttest_legato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"], paired = TRUE)
+ttest_staccato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"], paired = TRUE)
+
+print(ttest_legato)
+print(ttest_staccato)
+```
+
+### Post-hoc (pairwise t test)
+```{r post-hoc4, echo = FALSE}
+# Create a group label for a post-hoc test
+kor_art$Group[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"] <- "Performing-Legato" 
+kor_art$Group[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
+kor_art$Group[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"] <- "Teaching-Legato"
+kor_art$Group[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
+# pairwise comparison
+kor_art_pairwise <- pairwise.t.test(kor_art$Mean, kor_art$Group, paired = TRUE, p.adjust.method = "BH")
+kor_art_pairwise$p.value <- round(kor_art_pairwise$p.value, 4)
+print(kor_art_pairwise)
 ```

### 2019-10-20 script/RNotebook/articulation.Rmd
add comments / change ttests variable names
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r kot_art_all, echo = FALSE}
+# Group mean
 kot_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kot_art,
@@
 ```{r post-hoc1, echo = FALSE}
-ttest_legato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"], paired = TRUE)
-ttest_staccato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"], paired = TRUE)
-
-print(ttest_legato)
-print(ttest_staccato)
+kot_ttest_legato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"], paired = TRUE)
+kot_ttest_staccato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"], paired = TRUE)
+
+print(kot_ttest_legato)
+print(kot_ttest_staccato)
 ```
@@
 ```{r kor_art_all, echo = FALSE}
+# Group mean
 kor_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kor_art,
@@
 ```{r post-hoc3, echo = FALSE}
-ttest_legato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"], paired = TRUE)
-ttest_staccato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"], paired = TRUE)
-
-print(ttest_legato)
-print(ttest_staccato)
+kor_ttest_legato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"], paired = TRUE)
+kor_ttest_staccato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"], paired = TRUE)
+
+print(kor_ttest_legato)
+print(kor_ttest_staccato)
 ```
@@
 ```{r kot_dyn_all, echo = FALSE}
+# Group mean
 kot_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kot_dyn,
@@
 ```{r kor_dyn_all, echo = FALSE}
+# Group mean
 kor_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kor_dyn,

### 2019-10-20 script/RNotebook/articulation.Rmd
typo
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   facet_wrap(Order ~ .) +
-  labs(y = "KOY", subtitle = "KOT: Dynamics")
+  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
 p_kot_dyn_order_all_box

### 2019-10-20 script/RNotebook/dynamics.Rmd
add comments / add post-hoc comparisons
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+### Post-hoc (direct comparison)
+```{r post-hoc1, echo = FALSE}
+vel_ttest_forte <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"], paired = TRUE)
+vel_ttest_piano <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"], paired = TRUE)
+
+print(vel_ttest_forte)
+print(vel_ttest_piano)
+```
+
+### Post-hoc (pairwise t test)
+```{r post-hoc2, echo = FALSE}
+# Create a group label for a post-hoc test
+vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"] <- "Performing-Forte" 
+vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"] <- "Performing-Piano"
+vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"] <- "Teaching-Forte"
+vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"] <- "Teaching-Piano"
+# pairwise comparison
+vel_dyn_pairwise <- pairwise.t.test(vel_dyn$Mean, vel_dyn$Group, paired = TRUE, p.adjust.method = "BH")
+vel_dyn_pairwise$p.value <- round(vel_dyn_pairwise$p.value, 4)
+print(vel_dyn_pairwise)
+```
+
 ### 3. Trial by trial - individual{#vel_dyn_ind_trial}
@@
   facet_grid(Subcomponent ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
+  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
 p_vel_acc_dyn
@@
 
+### Post-hoc (direct comparison)
+```{r post-hoc3, echo = FALSE}
+vel_acc_ttest_forte <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "FtoP"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "FtoP"], paired = TRUE)
+vel_acc_ttest_piano <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)
+vel_acc_ttest_piano$p.value <- sprintf('%.6f', vel_acc_ttest_piano$p.value)
+
+print(vel_acc_ttest_forte)
+print(vel_acc_ttest_piano)
+```
+
+### Post-hoc (pairwise t test)
+```{r post-hoc4, echo = FALSE}
+# Create a group label for a post-hoc test
+vel_acc_dyn$Group[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "FtoP"] <- "Performing-FtoP" 
+vel_acc_dyn$Group[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"] <- "Performing-PtoF"
+vel_acc_dyn$Group[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "FtoP"] <- "Teaching-FtoP"
+vel_acc_dyn$Group[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"] <- "Teaching-PtoF"
+# pairwise comparison
+vel_acc_dyn_pairwise <- pairwise.t.test(vel_acc_dyn$Mean, vel_acc_dyn$Group, paired = TRUE, p.adjust.method = "BH")
+vel_acc_dyn_pairwise$p.value <- round(vel_acc_dyn_pairwise$p.value, 4)
+print(vel_acc_dyn_pairwise)
+```
+
 ### 3. Trial by trial - individual{#vel_acc_dyn_ind_trial}
@@
   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
+  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
 p_vel_acc_dyn_trial_all_ptof
@@
                 width = .2, position = position_dodge(.9)) +
-  labs(y = "KV (1-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
+  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
 p_vel_art_all
@@
 
+### Post-hoc (direct comparison)
+```{r post-hoc5, echo = FALSE}
+vel_ttest_legato <- t.test(vel_art$Mean[vel_art$Condition == "performing" & vel_art$Subcomponent == "Legato"], vel_art$Mean[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Legato"], paired = TRUE)
+vel_ttest_staccato <- t.test(vel_art$Mean[vel_art$Condition == "performing" & vel_art$Subcomponent == "Staccato"], vel_art$Mean[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Staccato"], paired = TRUE)
+
+print(vel_ttest_legato)
+print(vel_ttest_staccato)
+```
+
+### Post-hoc (pairwise t test)
+```{r post-hoc6, echo = FALSE}
+# Create a group label for a post-hoc test
+vel_art$Group[vel_art$Condition == "performing" & vel_art$Subcomponent == "Legato"] <- "Performing-Legato" 
+vel_art$Group[vel_art$Condition == "performing" & vel_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
+vel_art$Group[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Legato"] <- "Teaching-Legato"
+vel_art$Group[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
+# pairwise comparison
+vel_art_pairwise <- pairwise.t.test(vel_art$Mean, vel_art$Group, paired = TRUE, p.adjust.method = "BH")
+vel_art_pairwise$p.value <- round(vel_art_pairwise$p.value, 4)
+print(vel_art_pairwise)
+```
+
 ### 3. Trial by trial - individual{#vel_art_ind_trial}
@@
   facet_grid(Subcomponent ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Articulaion")
+  labs(y = "Difference", subtitle = "KV-Diff: Articulaion")
 p_vel_acc_art
@@
                 width = .2, position = position_dodge(.9)) +
-  labs(y = "Difference", subtitle = "KV: Articulation")
+  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
 p_vel_acc_art_all
@@
   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
-  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
+  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
 p_vel_acc_art_all_box
@@
 print(vel_acc_art_aov$ANOVA)
+```
+
+### Post-hoc (direct comparison)
+```{r post-hoc7, echo = FALSE}
+vel_acc_ttest_legato <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "LtoS"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "LtoS"], paired = TRUE)
+vel_acc_ttest_staccato <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "StoL"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "StoL"], paired = TRUE)
+
+print(vel_acc_ttest_legato)
+print(vel_acc_ttest_staccato)
+```
+
+### Post-hoc (pairwise t test)
+```{r post-hoc8, echo = FALSE}
+# Create a group label for a post-hoc test
+vel_acc_art$Group[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "LtoS"] <- "Performing-LtoS" 
+vel_acc_art$Group[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "StoL"] <- "Performing-StoL"
+vel_acc_art$Group[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "LtoS"] <- "Teaching-LtoS"
+vel_acc_art$Group[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "StoL"] <- "Teaching-StoL"
+# pairwise comparison
+vel_acc_art_pairwise <- pairwise.t.test(vel_acc_art$Mean, vel_acc_art$Group, paired = TRUE, p.adjust.method = "BH")
+vel_acc_art_pairwise$p.value <- round(vel_acc_art_pairwise$p.value, 4)
+print(vel_acc_art_pairwise)
 ```
@@
   facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
+  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
 p_vel_acc_dyn_order_all
@@
   facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
+  labs(y = "Difference", subtitle = "KV-Diff: Dynamics") 
 p_vel_acc_dyn_order_all_box
@@
   facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
+  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
 p_vel_art_order_all
@@
   facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
+  labs(y = "KV (0-127)", subtitle = "KV: Articulation") 
 p_vel_art_order_all_box
@@
   facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
+  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
 p_vel_acc_art_order_all
@@
   facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
+  labs(y = "Difference", subtitle = "KV: Articulation") 
 p_vel_acc_art_order_all_box

### 2019-10-21 script/RNotebook/questionnaire.Rmd
correct typo
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
change variable names
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# Outline - Quesionnaire
+# Outline - Questionnaire
 
@@
 df_exc <- read.csv("./filtered/data_errorRate.csv", header = T, sep = ",", dec = ".")
+include <- df_exc$SubNr[df_exc$Final == "include"]
 
@@
 excluded_participants <- data.frame(cbind(total, total_exc, exp_error, pitch_error, tempo_dev))
-colnames(excluded_participants) <- c("Total participants", "Total excluded", "Experimental error", "Pitch error", "Tempo deviation")
+colnames(excluded_participants) <- c("Total", "Excluded participants", "Experimental error", "Pitch error", "Tempo deviation")
 print(excluded_participants)
@@
 # Data frame with only included participants
-include <- df_exc$SubNr[df_exc$Final == "include"]
 df_demo <- data.frame()
@@
 female <- nrow(subset(df_demo, df_demo$Gender == "Female"))
-experience_m <- round(mean(df_demo$PianoTotalPractice), 0)
+experience_m <- round(mean(df_demo$PianoTotalPractice), 1)
 experience_sd <- round(sd(df_demo$PianoTotalPractice)*sqrt((length(df_demo$PianoTotalPractice)-1)/length(df_demo$PianoTotalPractice)), 2) #biased sd for the current data

### 2019-10-21 script/RNotebook/dynamics.Rmd
correct typo
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r post-hoc3, echo = FALSE}
-vel_acc_ttest_forte <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "FtoP"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "FtoP"], paired = TRUE)
-vel_acc_ttest_piano <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)
-vel_acc_ttest_piano$p.value <- sprintf('%.6f', vel_acc_ttest_piano$p.value)
-
-print(vel_acc_ttest_forte)
-print(vel_acc_ttest_piano)
+vel_acc_ttest_ftop <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "FtoP"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "FtoP"], paired = TRUE)
+vel_acc_ttest_ptof <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)
+
+print(vel_acc_ttest_ftop)
+print(vel_acc_ttest_ptof)
 ```

### 2019-10-21 script/RNotebook/articulation.Rmd
add effect size
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-### Post-hoc (direct comparison)
+##### Post-hoc (direct comparison)
 ```{r post-hoc1, echo = FALSE}
@@
 kot_ttest_staccato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"], paired = TRUE)
+# Effect size
+kot_ttest_legato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Legato"], kot_art$Condition[kot_art$Subcomponent == "Legato"], paired = TRUE)
+kot_ttest_staccato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Staccato"], kot_art$Condition[kot_art$Subcomponent == "Staccato"], paired = TRUE)
 
@@
 print(kot_ttest_staccato)
-```
-
-### Post-hoc (pairwise t test)
+print(kot_ttest_legato_cohend)
+print(kot_ttest_staccato_cohend)
+```
+
+##### Post-hoc (pairwise t test)
 ```{r post-hoc2, echo = FALSE}
@@
 
-### Post-hoc (direct comparison)
+##### Post-hoc (direct comparison)
 ```{r post-hoc3, echo = FALSE}
@@
 kor_ttest_staccato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"], paired = TRUE)
+# Effect size
+kor_ttest_legato_cohend <- cohen.d(kor_art$Mean[kor_art$Subcomponent == "Legato"], kor_art$Condition[kor_art$Subcomponent == "Legato"], paired = TRUE)
+kor_ttest_staccato_cohend <- cohen.d(kor_art$Mean[kor_art$Subcomponent == "Staccato"], kor_art$Condition[kor_art$Subcomponent == "Staccato"], paired = TRUE)
 
@@
 print(kor_ttest_staccato)
-```
-
-### Post-hoc (pairwise t test)
+print(kor_ttest_legato_cohend)
+print(kor_ttest_staccato_cohend)
+```
+
+##### Post-hoc (pairwise t test)
 ```{r post-hoc4, echo = FALSE}

### 2019-10-21 script/RNotebook/dynamics.Rmd
add effect size
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 vel_ttest_piano <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"], paired = TRUE)
+# Effect size
+vel_ttest_forte_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Forte"], vel_dyn$Condition[vel_dyn$Subcomponent == "Forte"], paired = TRUE)
+vel_ttest_piano_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Piano"], vel_dyn$Condition[vel_dyn$Subcomponent == "Piano"], paired = TRUE)
 
@@
 print(vel_ttest_piano)
-```
-
-### Post-hoc (pairwise t test)
+print(vel_ttest_forte_cohend)
+print(vel_ttest_piano_cohend)
+```
+
+##### Post-hoc (pairwise t test)
 ```{r post-hoc2, echo = FALSE}
@@
 vel_acc_ttest_ptof <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)
+# Effect size
+vel_acc_ttest_ftop_cohend <- cohen.d(vel_acc_dyn$Mean[vel_acc_dyn$Subcomponent == "FtoP"], vel_acc_dyn$Condition[vel_acc_dyn$Subcomponent == "FtoP"], paired = TRUE)
+vel_acc_ttest_ptof_cohend <- cohen.d(vel_acc_dyn$Mean[vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Condition[vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)
 
@@
 print(vel_acc_ttest_ptof)
-```
-
-### Post-hoc (pairwise t test)
+print(vel_acc_ttest_ftop_cohend)
+print(vel_acc_ttest_ptof_cohend)
+```
+
+##### Post-hoc (pairwise t test)
 ```{r post-hoc4, echo = FALSE}
@@
 
-### Post-hoc (direct comparison)
+##### Post-hoc (direct comparison)
 ```{r post-hoc5, echo = FALSE}
@@
 vel_ttest_staccato <- t.test(vel_art$Mean[vel_art$Condition == "performing" & vel_art$Subcomponent == "Staccato"], vel_art$Mean[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Staccato"], paired = TRUE)
+# Effect size
+vel_ttest_legato_cohend <- cohen.d(vel_art$Mean[vel_art$Subcomponent == "Legato"], vel_art$Condition[vel_art$Subcomponent == "Legato"], paired = TRUE)
+vel_ttest_staccato_cohend <- cohen.d(vel_art$Mean[vel_art$Subcomponent == "Staccato"], vel_art$Condition[vel_art$Subcomponent == "Staccato"], paired = TRUE)
 
@@
 print(vel_ttest_staccato)
-```
-
-### Post-hoc (pairwise t test)
+print(vel_ttest_legato_cohend)
+print(vel_ttest_staccato_cohend)
+```
+
+##### Post-hoc (pairwise t test)
 ```{r post-hoc6, echo = FALSE}
@@
 
-### Post-hoc (direct comparison)
+##### Post-hoc (direct comparison)
 ```{r post-hoc7, echo = FALSE}
-vel_acc_ttest_legato <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "LtoS"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "LtoS"], paired = TRUE)
-vel_acc_ttest_staccato <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "StoL"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "StoL"], paired = TRUE)
-
-print(vel_acc_ttest_legato)
-print(vel_acc_ttest_staccato)
-```
-
-### Post-hoc (pairwise t test)
+vel_acc_ttest_ltos <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "LtoS"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "LtoS"], paired = TRUE)
+vel_acc_ttest_stol <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "StoL"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "StoL"], paired = TRUE)
+# Effect size
+vel_acc_ttest_ltos_cohend <- cohen.d(vel_acc_art$Mean[vel_acc_art$Subcomponent == "LtoS"], vel_acc_art$Condition[vel_acc_art$Subcomponent == "LtoS"], paired = TRUE)
+vel_acc_ttest_stol_cohend <- cohen.d(vel_acc_art$Mean[vel_acc_art$Subcomponent == "StoL"], vel_acc_art$Condition[vel_acc_art$Subcomponent == "StoL"], paired = TRUE)
+
+print(vel_acc_ttest_ltos)
+print(vel_acc_ttest_stol)
+print(vel_acc_ttest_ltos_cohend)
+print(vel_acc_ttest_stol_cohend)
+```
+
+##### Post-hoc (pairwise t test)
 ```{r post-hoc8, echo = FALSE}

### 2019-10-22 script/RNotebook/articulation.Rmd
add sequence plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+- [Sequence plot](#sequence_art)
+
 #### KOR
@@
   + [All](#kot_dyn_order_group_all)
+
+- [Sequence plot](#sequence_dyn)
 
@@
 ```
+
+## Sequence plot
+#### Articulation{#sequence_art}
+```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
+# For each individual
+kot_art_seq <- aggregate(KOT~SubNr*Condition*Skill*Interval, data = df_kot_art,
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+kot_art_seq <- cbind(kot_art_seq[,1:4], as.data.frame(kot_art_seq[,5]))
+# Change colnames
+colnames(kot_art_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+kot_art_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = kot_art_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_art_seq_stats <- cbind(kot_art_seq_stats[,1:3], as.data.frame(kot_art_seq_stats[,4]))
+# Change colnames for each interval
+colnames(kot_art_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_kot_art_seq <- ggplot(data = kot_art_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+p_kot_art_seq
+```
+
+#### Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 15, fig.height = 5, echo = FALSE}
+# For each individual
+kot_dyn_seq <- aggregate(KOT~SubNr*Condition*Skill*Interval, data = df_kot_dyn,
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+kot_dyn_seq <- cbind(kot_dyn_seq[,1:4], as.data.frame(kot_dyn_seq[,5]))
+# Change colnames
+colnames(kot_dyn_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+kot_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = kot_dyn_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_dyn_seq_stats <- cbind(kot_dyn_seq_stats[,1:3], as.data.frame(kot_dyn_seq_stats[,4]))
+# Change colnames for each interval
+colnames(kot_dyn_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_kot_dyn_seq <- ggplot(data = kot_dyn_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+p_kot_dyn_seq
+```

### 2019-10-22 script/RNotebook/dynamics.Rmd
add sequence plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+- [Sequence plot](#sequence_dyn)
+
 ### [Articulation](#vel_articulation)
@@
   + [All](#vel_acc_art_order_group_all)
+
+- [Sequence plot](#sequence_art)
 
@@
 ```
+
+## Sequence plot
+#### Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 15, fig.height = 5, echo = FALSE}
+# Average vel_dynocity for each note
+# For each participant
+vel_dyn_seq <- aggregate(Velocity~SubNr*Condition*Skill*Note, data = df_vel_dyn, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_dyn_seq <- cbind(vel_dyn_seq[,1:4], as.data.frame(vel_dyn_seq[,5]))
+# Change colnames
+colnames(vel_dyn_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD")
+
+# Group mean
+vel_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_dyn_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_dyn_seq_stats <- cbind(vel_dyn_seq_stats[,1:3], as.data.frame(vel_dyn_seq_stats[,4]))
+# Change colnames
+colnames(vel_dyn_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM")
+
+# Average acceleration for each note
+# For each participant
+vel_dyn_acc_seq <- aggregate(Acc~SubNr*Condition*Skill*Interval, data = df_vel_acc_dyn, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_dyn_acc_seq <- cbind(vel_dyn_acc_seq[,1:4], as.data.frame(vel_dyn_acc_seq[,5]))
+# Change colnames
+colnames(vel_dyn_acc_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+vel_dyn_acc_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_dyn_acc_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_dyn_acc_seq_stats <- cbind(vel_dyn_acc_seq_stats[,1:3], as.data.frame(vel_dyn_acc_seq_stats[,4]))
+# Change colnames
+colnames(vel_dyn_acc_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_vel_dyn_seq <- ggplot(data = vel_dyn_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(0.05)) + 
+  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
+p_vel_dyn_seq
+
+p_vel_dyn_acc_seq <- ggplot(data = vel_dyn_acc_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
+p_vel_dyn_acc_seq
+```
+
+#### Articulation{#sequence_art}
+```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
+# Average vel_artocity for each note
+# For each participant
+vel_art_seq <- aggregate(Velocity~SubNr*Condition*Skill*Note, data = df_vel_art, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_art_seq <- cbind(vel_art_seq[,1:4], as.data.frame(vel_art_seq[,5]))
+# Change colnames
+colnames(vel_art_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD")
+
+# Group mean
+vel_art_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_art_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_art_seq_stats <- cbind(vel_art_seq_stats[,1:3], as.data.frame(vel_art_seq_stats[,4]))
+# Change colnames
+colnames(vel_art_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM")
+
+# Average acceleration for each note
+# For each participant
+vel_art_acc_seq <- aggregate(Acc~SubNr*Condition*Skill*Interval, data = df_vel_acc_art, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_art_acc_seq <- cbind(vel_art_acc_seq[,1:4], as.data.frame(vel_art_acc_seq[,5]))
+# Change colnames
+colnames(vel_art_acc_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+vel_art_acc_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_art_acc_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_art_acc_seq_stats <- cbind(vel_art_acc_seq_stats[,1:3], as.data.frame(vel_art_acc_seq_stats[,4]))
+# Change colnames
+colnames(vel_art_acc_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_vel_art_seq <- ggplot(data = vel_art_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(0.05)) + 
+  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
+p_vel_art_seq
+
+p_vel_art_acc_seq <- ggplot(data = vel_art_acc_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
+p_vel_art_acc_seq
+```

### 2019-10-22 script/RNotebook/ioi.Rmd
add sequence plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+- [Sequence plot](#sequence_art)
+
 ### [Dynamics](#ioi-dynamics)
@@
   + [All](#dyn_order_group_all)
+
+- [Sequence plot](#sequence_dyn)
 
@@
 ```
+
+## Sequence plot
+#### Articulation{#sequence_art}
+```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
+# For each individual
+ioi_art_seq <- aggregate(IOI~SubNr*Condition*Skill*Interval, data = df_ioi_art,
+                                 FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+ioi_art_seq <- cbind(ioi_art_seq[,1:4], as.data.frame(ioi_art_seq[,5]))
+# Change colnames
+colnames(ioi_art_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+ioi_art_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = ioi_art_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_art_seq_stats <- cbind(ioi_art_seq_stats[,1:3], as.data.frame(ioi_art_seq_stats[,4]))
+# Change colnames
+colnames(ioi_art_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_ioi_art_seq <- ggplot(data = ioi_art_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  annotate("text", 2, 188, label = "Tempo (80bpm)", vjust = -1, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "IOI (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+p_ioi_art_seq
+```
+
+#### Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 15, fig.height = 5, echo = FALSE}
+# For each individual
+ioi_dyn_seq <- aggregate(IOI~SubNr*Condition*Skill*Interval, data = df_ioi_dyn,
+                                 FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+ioi_dyn_seq <- cbind(ioi_dyn_seq[,1:4], as.data.frame(ioi_dyn_seq[,5]))
+# Change colnames
+colnames(ioi_dyn_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+ioi_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = ioi_dyn_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_dyn_seq_stats <- cbind(ioi_dyn_seq_stats[,1:3], as.data.frame(ioi_dyn_seq_stats[,4]))
+# Change colnames
+colnames(ioi_dyn_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_ioi_dyn_seq <- ggplot(data = ioi_dyn_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  annotate("text", 2, 188, label = "Tempo (80bpm)", vjust = -1, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "IOI (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+p_ioi_dyn_seq
+```

### 2019-11-11 script/RNotebook/manual_filtering.Rmd
developing pitch error remover
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Manual filtering"
output: html_notebook
---

```{r setup, include = FALSE}
# install and load required packages
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("wesanderson")) {install.packages("wesanderson"); require("wesanderson")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")

# read csv
df_all <- read.csv("./filtered/data_all.csv", header = T, sep = ",", dec = ".")
df_onset <- df_all %>% dplyr::filter(Key_OnOff == 1)
df_removed_onset <- read.csv("./filtered/data_removed_onset.csv", header = T, sep = ",", dec = ".") # removed trials by the first filtering

df_ideal <- read.table("./ideal.txt")
colnames(df_ideal) <- "Pitch"
df_ideal$RowNr <- c(1:nrow(df_ideal))
df_ideal <- df_ideal[c(2, 1)]

# create data frame to save data without errors
df_container <- data.frame()
```

```{r manual_filtering, echo = FALSE}
manual_filtering <- function(current_onset, ideal){
  counter = 0
  if (nrow(current_onset) > length(ideal$Pitch) & nrow(current_onset) < length(ideal$Pitch)*1.05){ # if there are more notes than ideal
    df_error$errorType[i] <<- "More"
    for (note in 1:length(ideal$Pitch)){
      # detect onset error
      if (current_onset[note,]$Pitch != ideal$Pitch[note]){
        while (counter == 0) {
          df_error$errorFirst[i] <<- note
          print(sprintf("Onset-Pitch error - RowNr: %i", note))
          counter = counter + 1
        }
      }
    }
  } else if (nrow(current_onset) > length(ideal$Pitch)*0.95 & nrow(current_onset) < length(ideal$Pitch) & nrow(current_onset)){ # if there are less note than ideal
    df_error$errorType[i] <<- "Less"
    for (note in 1:nrow(current_onset)){
      # detect onset error
      if (current_onset[note,]$Pitch != ideal$Pitch[note]){
        while (counter == 0) {
          df_error$errorFirst[i] <<- note
          print(sprintf("Onset-Pitch error - RowNr: %i", note))
          counter = counter + 1
        }
      }
    }
  } else if (nrow(current_onset) == length(ideal$Pitch)){ # there are equal notes to ideal
    for (note in 1:nrow(current_onset)){
      df_error$errorType[i] <<- "Equal"
      # detect onset error
      if (current_onset[note,]$Pitch != ideal$Pitch[note]){
        while (counter == 0) {
          df_error$errorFirst[i] <<- note
          print(sprintf("Onset-Pitch error - RowNr: %i", note))
          counter = counter + 1
        }
      }
    } 
  } else if (nrow(current_onset) >= length(ideal$Pitch)*1.05){ # error more than 5%
    df_error$errorType[i] <<- "Exclude"
    df_error$errorFirst[i] <<- "Exclude the trial (more than 5%)"
  } else if (nrow(current_onset) <= length(ideal$Pitch)*0.95){ # error more than 5%
    df_error$errorType[i] <<- "Exclude"
    df_error$errorFirst[i] <<- "Exclude the trial (less than 5%)"
  } else { # other problems
    df_error$errorType[i] <<- "Other"
    df_error$errorFirst[i] <<- "Check individually"
  }
}
```

# Each participant
# First run
```{r first_error, echo = FALSE}
df_error <- data.frame(df_removed_onset)
for (i in 1:nrow(df_error)){
  print(i)
  current <- df_onset %>% dplyr::filter(SubNr == df_removed_onset[i,1], BlockNr == df_removed_onset[i,2], TrialNr == df_removed_onset[i,3])
  current$RowNr = c(1:nrow(current))
  manual_filtering(current, df_ideal)
}
df_error
```

```{r plot, echo = FALSE}
for (i in 1:nrow(df_removed_onset)){
  print(i)
  current <- df_onset %>% dplyr::filter(SubNr == df_removed_onset[i,1], BlockNr == df_removed_onset[i,2], TrialNr == df_removed_onset[i,3])
  current$RowNr = c(1:nrow(current))

  plot <- ggplot() +
    geom_line(data = current, aes(x = RowNr, y = Pitch), colour = "#F8766D") + 
    geom_line(data = df_ideal, aes(x = RowNr, y = Pitch), colour = "#00BFC4") +
    geom_point(data = current, aes(x = RowNr, y = Pitch), colour = "#F8766D") +
    geom_point(data = df_ideal, aes(x = RowNr, y = Pitch), colour = "#00BFC4") +
    scale_x_continuous("RowNr", current$RowNr) +
    labs(title = sprintf("SubNr: %i, BlockNr: %i, TrialNr: %i", df_removed_onset[i,1], df_removed_onset[i,2], df_removed_onset[i,3]))

  print(plot)
}
```

```{r manual_filtering_2, echo = FALSE}
manual_filtering_2 <- function(current_onset, ideal){
  counter = 0
  if (nrow(current_onset) > length(ideal$Pitch) & nrow(current_onset) < length(ideal$Pitch)*1.05){ # if there are more notes than ideal
    df_error$errorType2[i] <<- "More"
    for (note in 1:length(ideal$Pitch)){
      # detect onset error
      if (current_onset[note,]$Pitch != ideal$Pitch[note]){
        while (counter == 0) {
          df_error$errorSecond[i] <<- note
          print(sprintf("Onset-Pitch error - RowNr: %i", note))
          current_onset <- current_onset[-note,]
          counter = counter + 1
        }
      }
    }
  } else if (nrow(current_onset) > length(ideal$Pitch)*0.95 & nrow(current_onset) < length(ideal$Pitch) & nrow(current_onset)){ # if there are less note than ideal
    df_error$errorType2[i] <<- "Less"
    for (note in 1:nrow(current_onset)){
      # detect onset error
      if (current_onset[note,]$Pitch != ideal$Pitch[note]){
        while (counter == 0) {
          df_error$errorSecond[i] <<- note
          print(sprintf("Onset-Pitch error - RowNr: %i", note))
          current_onset[note,] <- NA
          counter = counter + 1
        }
      }
    }
  } else if (nrow(current_onset) == length(ideal$Pitch)){ # there are equal notes to ideal
    for (note in 1:nrow(current_onset)){
      df_error$errorType2[i] <<- "Equal"
      # detect onset error
      if (current_onset[note,]$Pitch != ideal$Pitch[note]){
        while (counter == 0) {
          df_error$errorSecond[i] <<- note
          print(sprintf("Onset-Pitch error - RowNr: %i", note))
          counter = counter + 1
        }
      }
    } 
  }
}
```

```{r second-run, echo = FALSE}
for (i in 1:nrow(df_error)){
  current <- df_onset %>% dplyr::filter(SubNr == df_error[i,1], BlockNr == df_error[i,2], TrialNr == df_error[i,3])
  current$RowNr = c(1:nrow(current))
  if (df_error$errorType[i] == "More"){
    print(i)
    current <- current[-as.integer(df_error$errorFirst[i]),]
    manual_filtering_2(current, df_ideal)
  } else if (df_error$errorType[i] == "Less"){
    print(i)
    current[as.integer(df_error$errorFirst[i]),] <- NA
    manual_filtering_2(current[-as.integer(df_error$errorFirst[i]),], df_ideal[-as.integer(df_error$errorFirst[i]),])
  }
}
```

# Remove the first error
```{r remove_twomasures, echo = FALSE}
for (i in c(1:length(df_error$errorType))){
  current <- df_onset %>% dplyr::filter(SubNr == df_removed_onset[i,1], BlockNr == df_removed_onset[i,2], TrialNr == df_removed_onset[i,3])
  current$RowNr = c(1:nrow(current))
  if (df_error$errorType[i] == "More"){
    if (df_error$errorFirst[i] <= 8){
      current <- current %>% dplyr::filter(RowNr > 17)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 16))
    } else if (df_error$errorFirst[i] > 8 & df_error$errorFirst[i] <= 16){
      current <- current %>% dplyr::filter(RowNr > 25)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 24))
    } else if (df_error$errorFirst[i] > 16 & df_error$errorFirst[i] <= 24){
      current <- current %>% dplyr::filter(RowNr > 29)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 28))
    } else if (df_error$errorFirst[i] > 24 & df_error$errorFirst[i] <= 28){
      current <- current %>% dplyr::filter(RowNr > 34)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 33))
    } else if (df_error$errorFirst[i] > 28 & df_error$errorFirst[i] <= 33){
      current <- current %>% dplyr::filter(RowNr > 40)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 39))
    } else if (df_error$errorFirst[i] > 33 & df_error$errorFirst[i] <= 39){
      current <- current %>% dplyr::filter(RowNr > 48)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 47))
    } else if (df_error$errorFirst[i] > 39 & df_error$errorFirst[i] <= 47){
      current <- current %>% dplyr::filter(RowNr > 53)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 52))
    } else if (df_error$errorFirst[i] > 47 & df_error$errorFirst[i] <= 52){
      current <- current %>% dplyr::filter(RowNr > 58)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 57))
    } else if (df_error$errorFirst[i] > 52){
    }
  } else if (df_error$errorType[i] == "Less"){
    if (df_error$errorFirst[i] <= 8){
      current <- current %>% dplyr::filter(RowNr > 15)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 16))
    } else if (df_error$errorFirst[i] > 8 & df_error$errorFirst[i] <= 16){
      current <- current %>% dplyr::filter(RowNr > 24)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 24))
    } else if (df_error$errorFirst[i] > 16 & df_error$errorFirst[i] <= 24){
      current <- current %>% dplyr::filter(RowNr > 27)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 28))
    } else if (df_error$errorFirst[i] > 24 & df_error$errorFirst[i] <= 28){
      current <- current %>% dplyr::filter(RowNr > 32)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 33))
    } else if (df_error$errorFirst[i] > 28 & df_error$errorFirst[i] <= 33){
      current <- current %>% dplyr::filter(RowNr > 48)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 39))
    } else if (df_error$errorFirst[i] > 33 & df_error$errorFirst[i] <= 39){
      current <- current %>% dplyr::filter(RowNr > 46)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 47))
    } else if (df_error$errorFirst[i] > 39 & df_error$errorFirst[i] <= 47){
      current <- current %>% dplyr::filter(RowNr > 52)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 52))
    } else if (df_error$errorFirst[i] > 47 & df_error$errorFirst[i] <= 52){
      current <- current %>% dplyr::filter(RowNr > 56)
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 57))
    } else if (df_error$errorFirst[i] > 52){
    }
  } 
}
```

# Second error
```{r manual_filter_2, include = FALSE}
df_error$errorSecond <- NA
manual_filtering_2 <- function(current_onset, ideal){
  counter = 0
    for (note in 1:length(ideal$Pitch)){
      # detect onset error
      if (current_onset[note,]$Pitch != ideal$Pitch[note]){
        while (counter == 0) {
          df_error$errorSecond[i] <<- note
          print(sprintf("Onset-Pitch error - RowNr: %i", note))
          counter = counter + 1
       }
    }
  }
}
```

```{r second_error, echo = FALSE}
for (i in c(1:length(df_error$errorNote))){
  current <- df_onset %>% dplyr::filter(SubNr == df_removed_onset[i,1], BlockNr == df_removed_onset[i,2], TrialNr == df_removed_onset[i,3])
  if (df_error$errorType == "More"){
    if (df_error$errorNote[i] <= 8){
      current_second <- current %>% dplyr::filter(RowNr > df_error$errorFirst[i])
      manual_filtering_2(current, subset(df_ideal, df_ideal$RowNr > 17))
    } else if (df_error$errorFirst[i] <= 16){
      print(a)
    } else if (df_error$errorFirst[i] <= 24){
      print(a)
    } else if (df_error$errorFirst[i] <= 33){
      print(a)
    } else if (df_error$errorFirst[i] <= 39){
      print(a)
    } else if (df_error$errorFirst[i] <= 52){
      print(a)
    } else if (df_error$errorFirst[i] > 53){
    }
  }
}

# after two measures
-
current_6 <- current_6 %>% dplyr::filter(RowNr > 56)
manual_filtering(current_6, subset(df_ideal, df_ideal$RowNr > 57))

+
current_6 <- current_6 %>% dplyr::filter(RowNr > 56)
manual_filtering(current_6, subset(df_ideal, df_ideal$RowNr > 55))

# rescue after RowNr 56
container_6 <- rbind(container_6, current_6)
container_6
```

### 2019-11-28 script/RNotebook/ioi.Rmd
add basic analysis for ioi
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2019-10-03 script/RNotebook/ioi.Rmd
@@
 ---
-title: 'Analysis: The Sound of Teaching Music 1'
+title: 'IOIs Analysis: The Sound of Teaching Music 2'
 output:
@@
 
-##### Description: This is a summary of all analyses for a tempo.
-
-# Outline - Interonset intervals (IOIs)
-### [Experimental Design](#design)
-
-### [Articulation](#ioi-articulation)
-
-- Individual results
-   + [All](#art_ind_all)
-   + [Phrase bondaries](#art_ind_phrase)
-   + [Trial by trial](#art_ind_trial)
-   + [Variability all](#art_ind_var_all)
-   + [Variability trial-by-trial](#art_ind_var_trial)
-
-- Group results and stats (if available)
-   + [All](#art_group_all)
-   + [Phrase bondaries](#art_group_phrase)
-   + [Trial by trial](#art_group_trial)
-   + [Variability all](#art_group_var_all)
-   + [Variability trial-by-trial](#art_group_var_trial)
-
-### [Dynamics](#ioi-dynamics)
-
-- Individual results
-   + [All](#dyn_ind_all)
-   + [Phrase bondaries](#dyn_ind_phrase)
-   + [Trial by trial](#dyn_ind_trial)
-   + [Variability all](#dyn_ind_var_all)
-   + [Variability trial-by-trial](#dyn_ind_var_trial)
-
-- Group results and stats (if available)
-   + [All](#dyn_group_all)
-   + [Phrase bondaries](#dyn_group_phrase)
-   + [Trial by trial](#dyn_group_trial)
-   + [Variability all](#dyn_group_var_all)
-   + [Variability trial-by-trial](#dyn_group_var_trial)
+##### Description: This is a summary of all analyses for a tempo (IOIs).
 
@@
 - Independent variables: Condition (2 levels: teaching vs. performing) x Skill (2 levels: articulation vs. dynamics)
-- Dependent variables: Interonset intervals, variability (SD/mean IOI)
+- Dependent variables: Interonset intervals, (**Not yet**: variability (SD/mean IOI))
 
@@
 df_ioi_dyn <- df_ioi %>% dplyr::filter(Skill == "dynamics")
+
 ```
@@
 ### 1. All - individual{#art_ind_all}
-- Average IOIs for each participant.
+- Average normIOIs for each participant.
 
@@
 # For each individual
-ioi_art <- aggregate(IOI~SubNr*Condition*Skill, data = df_ioi_art,
+ioi_art <- aggregate(normIOI~SubNr*Condition*Skill, data = df_ioi_art,
                  FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
@@
 - Error bars show SD of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the ideal tempo
 
@@
                 width = .2, position = position_dodge(.9)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation")
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "IOI: Articulation")
 p_ioi_art
@@
 ### 2. All - group{#art_group_all}
-- Average IOIs for each condition.
+- Average normIOIs for each condition.
 - Hypothesis: Participants will play slower when teaching.
@@
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the ideal tempo
 
@@
                 width=.2, position = position_dodge(.9)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) +
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(.75, 1.1)) +
   theme(legend.position = "None")
@@
 ##### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the ideal tempo
 
@@
   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + #coord_cartesian(ylim = c(100, 210)) +
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "IOI: Articulation") +
   theme(legend.position = "None")
@@
 
-#### Statistics
-##### Normality check
-```{r normality1, echo = FALSE}
-# Normality check
-ioi_art_norm <- by(ioi_art$Mean, list(ioi_art$Condition), shapiro.test)
-print(ioi_art_norm)
-```
+#### Statistics (TBC)
 
-##### Paired t-test
-```{r stats1, echo = FALSE}
-ioi_art_ttest <- t.test(ioi_art$Mean[ioi_art$Condition == "teaching"], ioi_art$Mean[ioi_art$Condition == "performing"], paired = TRUE)
-print(ioi_art_ttest)
-```
-
-##### Effect size
-```{r stats2, echo = FALSE}
-ioi_art_cohend <- cohen.d(ioi_art$Mean, ioi_art$Condition, paired = TRUE)
-print(ioi_art_cohend)
-```
-
-### 3. Phrase boundaries - individual{#art_ind_phrase}
-- Compare IOIs at phrase boundaries with others fir each participant.
+### 3. Subcomponent change - individual{#art_ind_phrase}
+- Compare normIOIs at which Subcomponent changes (e.g., Legato to Staccato) with others fir each participant.
 
 ```{r ioi_art_ch, echo = FALSE}
-# Define phrase boundaries
-df_ioi_art$Boundary <- "No"
-df_ioi_art$Boundary[df_ioi_art$Interval == 8 | df_ioi_art$Interval == 16 | df_ioi_art$Interval == 24 | df_ioi_art$Interval == 41 | df_ioi_art$Interval == 49 | df_ioi_art$Interval == 57] <- "Yes"
+# Define subcomponent change
+df_ioi_art$Change <- "No"
+df_ioi_art$Change[df_ioi_art$Interval == 5 | df_ioi_art$Interval == 8 | df_ioi_art$Interval == 17 | df_ioi_art$Interval == 20 | df_ioi_art$Interval == 39 | df_ioi_art$Interval == 47] <- "Yes"
 
 # For each individual
-ioi_art_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_art,
+ioi_art_ch <- aggregate(normIOI~SubNr*Condition*Skill*Change, data = df_ioi_art,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 # Change colnames
-colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
+colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Change", "N", "Mean", "SD")
 print(ioi_art_ch)
@@
                 width = .2, position = position_dodge(.9)) +
-  facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
+  facet_grid(Change ~ .) +
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "IOI: Articulation")
 p_ioi_art_ch
@@
 
-### 4. Phrase boundaries - group{#art_group_phrase}
-- Compare IOIs at phrase boundaries with others in the two conditions.
-- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations.
+### 4. Subcomponent change - group{#art_group_phrase}
+- Compare IOIs at which Subcomponent changes (e.g., Legato to Staccato) with others in the two conditions.
+- Hypothesis: Participants will play more slowly where Subcomponent changes compared with other locations.
 
@@
 # Group mean
-ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_art_ch,
+ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Change, data = ioi_art_ch,
                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
@@
 # Change colnames
-colnames(ioi_art_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
+colnames(ioi_art_ch_all) <- c("Condition", "Skill", "Change", "N", "Mean", "SD", "SEM")
 print(ioi_art_ch_all)
@@
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the ideal tempo
 
 ```{r ioi_art_ch_all_bar, echo = FALSE}
-p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
+p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Change, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
                 width=.2, position = position_dodge(.9)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  #facet_grid(. ~ Boundary) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 250))
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(.75, 1.1))
 p_ioi_art_ch_all_bar
@@
 ##### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the ideal tempo
 
 ```{r ioi_art_ch_all_box, echo = FALSE}
-p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
+p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Change, y = Mean, fill = Condition)) +
   geom_boxplot() +
   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation")
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "IOI: Articulation")
 p_ioi_art_ch_box
@@
 
-#### Statistics
-##### Normality check
-```{r normality2, echo = FALSE}
-ioi_art_ch_norm <- by(ioi_art_ch$Mean, list(ioi_art_ch$Condition, ioi_art_ch$Boundary), shapiro.test)
-print(ioi_art_ch_norm)
-```
-
-##### Two-way ANOVA (ez)
-```{r stats3, echo = FALSE}
-ioi_art_ch_aov <- ezANOVA(
-  data = df_ioi_art
-  , dv = .(IOI)
-  , wid = .(SubNr)
-  , within = .(Condition, Boundary)
-  , type = 3
-  , detailed = TRUE
-)
-ioi_art_ch_aov$ANOVA$F <- round(ioi_art_ch_aov$ANOVA$F, 4)
-ioi_art_ch_aov$ANOVA$p <- round(ioi_art_ch_aov$ANOVA$p, 4)
-print(ioi_art_ch_aov$ANOVA)
-```
-
-##### Two-way ANOVA (aov)
-```{r stats4, echo = FALSE}
-ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
-print(summary(ioi_art_ch_aov_2))
-```
+#### Statistics (TBC)
 
 ### 5. Trial by trial - individual{#art_ind_trial}
-- Average IOIs for each trial for each participant.
+- Average normIOIs for each trial for each participant.
 
@@
 # For each individual
-ioi_art_trial <- aggregate(IOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_art,
+ioi_art_trial <- aggregate(normIOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_art,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 ##### Line graph
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the ideal tempo
 
@@
   geom_point() +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
   facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
+  labs(x = "Trial", y = "Normalised IOI", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
 p_ioi_art_trial
@@
 ### 6. Trial by trial - group{#art_group_trial}
-- Average IOIs for each trial for each condition.
+- Average normIOIs for each trial for each condition.
 
@@
   geom_point(position = position_dodge(.2)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
+  labs(x = "Trial", y = "Normalised IOI", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
 p_ioi_art_trial_all
-```
-
-### 7. Variability all - individual{#art_ind_var_all}
-- Average variability for each participant.
-- IOI variability = SD/mean IOI (per trial)
-
-```{r ioi_art_var, echo = FALSE}
-# Create a data frame for variability (SD/)
-df_ioi_art_var <- data.frame()
-for (subnr in unique(df_ioi_art$SubNr)){
-  for (block in unique(df_ioi_art$BlockNr)){
-    cond = as.character(unique(df_ioi_art$Condition[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
-    skill = as.character(unique(df_ioi_art$Skill[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
-    for (trial in unique(df_ioi_art$TrialNr)){
-      df_current <- df_ioi_art %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
-      if (nrow(df_current) != 0){
-        df_ioi_art_var <- rbind(df_ioi_art_var, 
-                                data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI)))
-      }
-    }
-  }
-}
-colnames(df_ioi_art_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability")
-
-# Average variability
-# For each individual
-ioi_art_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_art_var,
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_art_var <- cbind(ioi_art_var[,1:3], as.data.frame(ioi_art_var[,4]))
-# Change colnames
-colnames(ioi_art_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
-print(ioi_art_var)
-```
-
-#### Plot
-##### Bar graph
-- Error bars show SD of the mean
-
-```{r ioi_art_var_bar, echo = FALSE}
-p_ioi_art_var <- ggplot(data = ioi_art_var, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(1)) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation")
-p_ioi_art_var
-```
-
-### 8. Variability all - group{#art_group_var_all}
-- Average variability for each condition.
-
-```{r ioi_art_var_all, echo = FALSE}
-# Group mean
-ioi_art_var_all <- aggregate(Mean~Condition*Skill, data = ioi_art_var,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_var_all <- cbind(ioi_art_var_all[,1:2], as.data.frame(ioi_art_var_all[,3]))
-# Change colnames
-colnames(ioi_art_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
-print(ioi_art_var_all)
-```
-
-#### Plot
-##### Bar graph
-- Error bars show SEM of the mean
-
-```{r ioi_art_var_all_bar, echo = FALSE}
-p_ioi_art_var_all <- ggplot(data = ioi_art_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge()) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
-  theme(legend.position = "None")
-p_ioi_art_var_all
-```
-
-##### Box plot
-```{r ioi_art_var_all_box, echo = FALSE}
-p_ioi_art_var_box <- ggplot(data = ioi_art_var, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
-  theme(legend.position = "None")
-p_ioi_art_var_box
-```
-
-#### Statistics
-##### Normality check
-```{r normality3, echo = FALSE}
-ioi_art_var_norm <- by(ioi_art_var$Mean, list(ioi_art_var$Condition), shapiro.test)
-print(ioi_art_var_norm)
-```
-
-##### Paired t-test
-```{r stats5, echo = FALSE}
-ioi_art_var_ttest <- t.test(ioi_art_var$Mean[ioi_art_var$Condition == "teaching"], ioi_art_var$Mean[ioi_art_var$Condition == "performing"], paired = TRUE)
-print(ioi_art_var_ttest)
-```
-
-##### Effect size
-```{r stats6, echo = FALSE}
-ioi_art_var_cohend <- cohen.d(ioi_art_var$Mean, ioi_art_var$Condition, paired = TRUE)
-print(ioi_art_var_cohend)
-```
-
-### 9. Variability trial-by-trial - individual{#art_ind_var_trial}
-- Average variability for each trial for each participant.
-
-```{r ioi_art_var_trial, echo = FALSE}
-# For each individual
-ioi_art_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_art_var,
-                         FUN = function(x){c(N = length(x), mean = mean(x))})
-ioi_art_var_trial <- cbind(ioi_art_var_trial[,1:4], as.data.frame(ioi_art_var_trial[,5]))
-# Change colnames
-colnames(ioi_art_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean")
-print(ioi_art_var_trial)
-```
-
-#### Plot
-##### Line graph
-```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 8}
-p_ioi_art_var_trial <- ggplot(data = ioi_art_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_art_var_trial
-```
-
-### 10. Variability trial-by-trial - group{#art_group_var_trial}
-- Average variability for each trial for each condition.
-
-```{r ioi_art_var_trial_all, echo = FALSE}
-# Group mean
-ioi_art_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_art_var_trial,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_var_trial_all <- cbind(ioi_art_var_trial_all[,1:3], as.data.frame(ioi_art_var_trial_all[,4]))
-# Change colnames
-colnames(ioi_art_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
-print(ioi_art_var_trial_all)
-```
-
-#### Plot
-##### Line graph
-- Error bars show SEM of the mean
-
-```{r ioi_art_var_trial_all_line, echo = FALSE}
-p_ioi_art_var_trial_all <- ggplot(data = ioi_art_var_trial_all, 
-                                  aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line(position = position_dodge(.2)) +
-  geom_point(position = position_dodge(.2)) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_art_var_trial_all
 ```
@@
 ### 1. All - individual{#dyn_ind_all}
-- Average IOIs for each participant.
+- Average normIOIs for each participant.
 
@@
 # For each individual
-ioi_dyn <- aggregate(IOI~SubNr*Condition*Skill, data = df_ioi_dyn,
+ioi_dyn <- aggregate(normIOI~SubNr*Condition*Skill, data = df_ioi_dyn,
                  FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
@@
                 width = .2, position = position_dodge(.9)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics")
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "IOI: Dynamics")
 p_ioi_dyn
@@
 ### 2. All - group{#dyn_group_all}
-- Average IOIs for each condition.
+- Average normIOIs for each condition.
 - Hypothesis: Participants will play slower when teaching.
@@
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the ideal tempo
 
 ```{r ioi_dyn_all_bar, echo = FALSE}
-p_ioi_dyn_all <- ggplot(data = ioi_dyn_all, aes(x = Condition, y = Mean, fill = Condition)) +
+p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_all, aes(x = Skill, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
                 width=.2, position = position_dodge(.9)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(120, 210)) +
-  theme(legend.position = "None")
-p_ioi_dyn_all
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo + 
+  labs(y = "Normalised IOI", subtitle = "IOI: Dynamics") +  coord_cartesian(ylim = c(.75, 1.1))
+p_ioi_dyn_ch_all_bar
 ```
@@
   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + #coord_cartesian(ylim = c(100, 210)) +
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "IOI: Dynamics") + 
   theme(legend.position = "None")
@@
 
-#### Statistics
-##### Normality check
-```{r normality4, echo = FALSE}
-# Normality check
-ioi_dyn_norm <- by(ioi_dyn$Mean, list(ioi_dyn$Condition), shapiro.test)
-print(ioi_dyn_norm)
-```
+#### Statistics (TBC)
 
-```{r stats7, echo = FALSE}
-ioi_dyn_ttest <- t.test(ioi_dyn$Mean[ioi_dyn$Condition == "teaching"], ioi_dyn$Mean[ioi_dyn$Condition == "performing"], paired = TRUE)
-print(ioi_dyn_ttest)
-```
-
-##### Effect size
-```{r stats8, echo = FALSE}
-ioi_dyn_cohend <- cohen.d(ioi_dyn$Mean, ioi_dyn$Condition, paired = TRUE)
-print(ioi_dyn_cohend)
-```
-
-### 3. Phrase boundaries - individual{#dyn_ind_phrase}
-- Compare IOIs at phrase boundaries with others fir each participant.
+### 3. Subcomponent changes - individual{#dyn_ind_phrase}
+- Compare normIOIs at which Subcomponent changes (e.g., Forte to Piano) with others fir each participant.
 
@@
 # Define phrase boundaries
-df_ioi_dyn$Boundary <- "No"
-df_ioi_dyn$Boundary[df_ioi_dyn$Interval == 8 | df_ioi_dyn$Interval == 16 | df_ioi_dyn$Interval == 24 | df_ioi_dyn$Interval == 41 | df_ioi_dyn$Interval == 49 | df_ioi_dyn$Interval == 57] <- "Yes"
+df_ioi_dyn$Change <- "No"
+df_ioi_dyn$Change[df_ioi_dyn$Interval == 5 | df_ioi_dyn$Interval == 8 | df_ioi_dyn$Interval == 17 | df_ioi_dyn$Interval == 20 | df_ioi_dyn$Interval == 39 | df_ioi_dyn$Interval == 47] <- "Yes"
 
 # For each individual
-ioi_dyn_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_dyn,
+ioi_dyn_ch <- aggregate(normIOI~SubNr*Condition*Skill*Change, data = df_ioi_dyn,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 # Change colnames
-colnames(ioi_dyn_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
+colnames(ioi_dyn_ch) <- c("SubNr", "Condition", "Skill", "Change", "N", "Mean", "SD")
 print(ioi_dyn_ch)
@@
                 width = .2, position = position_dodge(.9)) +
-  facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
+  facet_grid(Change ~ .) +
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "IOI: Dynamics")
 p_ioi_dyn_ch
@@
 
-### 4. Phrase boundaries - group{#dyn_group_phrase}
-- Compare IOIs at phrase boundaries with others in the two conditions.
-- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations.
+### 4. Subcomponent changes - group{#dyn_group_phrase}
+- Compare IOIs at which Subcomponent changes with others in the two conditions.
+- Hypothesis: Participants will play more slowly where Subcomponent changes compared with other locations.
 
@@
 # Group mean
-ioi_dyn_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_dyn_ch,
+ioi_dyn_ch_all <- aggregate(Mean~Condition*Skill*Change, data = ioi_dyn_ch,
                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
@@
 # Change colnames
-colnames(ioi_dyn_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
+colnames(ioi_dyn_ch_all) <- c("Condition", "Skill", "Change", "N", "Mean", "SD", "SEM")
 print(ioi_dyn_ch_all)
@@
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the ideal tempo
 
 ```{r ioi_dyn_ch_all_bar, echo = FALSE}
-p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
+p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_ch_all, aes(x = Change, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
                 width=.2, position = position_dodge(.9)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  #facet_grid(. ~ Boundary) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 250))
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(.75, 1.1))
 p_ioi_dyn_ch_all_bar
@@
 ##### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the ideal tempo
 
 ```{r ioi_dyn_ch_all_box, echo = FALSE}
-p_ioi_dyn_ch_box <- ggplot(data = ioi_dyn_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
+p_ioi_dyn_ch_box <- ggplot(data = ioi_dyn_ch, aes(x = Change, y = Mean, fill = Condition)) +
   geom_boxplot() +
   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics")
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  labs(y = "Normalised IOI", subtitle = "normIOI: Dynamics")
 p_ioi_dyn_ch_box
@@
 
-#### Statistics
-##### Normality check
-```{r normality5, echo = FALSE}
-ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test)
-print(ioi_dyn_ch_norm)
-```
-
-##### Two-way ANOVA (ez)
-```{r stats9, echo = FALSE}
-ioi_dyn_ch_aov <- ezANOVA(
-  data = df_ioi_dyn
-  , dv = .(IOI)
-  , wid = .(SubNr)
-  , within = .(Condition, Boundary)
-  , type = 3
-  , detailed = TRUE
-)
-ioi_dyn_ch_aov$ANOVA$F <- round(ioi_dyn_ch_aov$ANOVA$F, 4)
-ioi_dyn_ch_aov$ANOVA$p <- round(ioi_dyn_ch_aov$ANOVA$p, 4)
-print(ioi_dyn_ch_aov)
-```
-
-##### Two-way ANOVA (aov)
-```{r stats10, echo = FALSE}
-ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
-print(summary(ioi_dyn_ch_aov_2))
-```
+#### Statistics (TBC)
 
 ### 5. Trial by trial - individual{#dyn_ind_trial}
-- Average IOIs for each trial for each participant.
+- Average normIOIs for each trial for each participant.
 
@@
 # For each individual
-ioi_dyn_trial <- aggregate(IOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn,
+ioi_dyn_trial <- aggregate(normIOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
   geom_point() +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
   facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
+  labs(x = "Trial", y = "Normalised IOI", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
 p_ioi_dyn_trial
@@
 ### 6. Trial by trial - group{#dyn_group_trial}
-- Average IOIs for each trial for each condition.
+- Average normIOIs for each trial for each condition.
 
@@
   geom_point(position = position_dodge(.2)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
@@
 ```
-
-### 7. Variability all - individual{#dyn_ind_var_all}
-- Average variability for each participant.
-- IOI variability = SD/mean IOI (per trial)
-
-```{r ioi_dyn_var, echo = FALSE}
-# Create a data frame for variability (SD/)
-df_ioi_dyn_var <- data.frame()
-for (subnr in unique(df_ioi_dyn$SubNr)){
-  for (block in unique(df_ioi_dyn$BlockNr)){
-    cond = as.character(unique(df_ioi_dyn$Condition[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
-    skill = as.character(unique(df_ioi_dyn$Skill[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
-    for (trial in unique(df_ioi_dyn$TrialNr)){
-      df_current <- df_ioi_dyn %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
-      if (nrow(df_current) != 0){
-        df_ioi_dyn_var <- rbind(df_ioi_dyn_var, 
-                                data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI)))
-      }
-    }
-  }
-}
-colnames(df_ioi_dyn_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability")
-
-# Average variability
-# For each individual
-ioi_dyn_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_dyn_var,
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_dyn_var <- cbind(ioi_dyn_var[,1:3], as.data.frame(ioi_dyn_var[,4]))
-# Change colnames
-colnames(ioi_dyn_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
-print(ioi_dyn_var)
-```
-
-#### Plot
-##### Bar graph
-- Error bars show SD of the mean
-
-```{r ioi_dyn_var_bar, echo = FALSE}
-p_ioi_dyn_var <- ggplot(data = ioi_dyn_var, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(1)) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics")
-p_ioi_dyn_var
-```
-
-### 8. Variability all - group{#dyn_group_var_all}
-- Average variability for each condition.
-
-```{r ioi_dyn_var_all, echo = FALSE}
-# Group mean
-ioi_dyn_var_all <- aggregate(Mean~Condition*Skill, data = ioi_dyn_var,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_var_all <- cbind(ioi_dyn_var_all[,1:2], as.data.frame(ioi_dyn_var_all[,3]))
-# Change colnames
-colnames(ioi_dyn_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_var_all)
-```
-
-#### Plot
-##### Bar graph
-- Error bars show SEM of the mean
-```{r ioi_dyn_var_all_bar, echo = FALSE}
-p_ioi_dyn_var_all <- ggplot(data = ioi_dyn_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge()) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
-  theme(legend.position = "None")
-p_ioi_dyn_var_all
-```
-
-##### Box plot
-```{r ioi_dyn_var_all_box, echo = FALSE}
-p_ioi_dyn_var_box <- ggplot(data = ioi_dyn_var, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
-  theme(legend.position = "None")
-p_ioi_dyn_var_box
-```
-
-#### Statistics
-##### Normality check
-```{r normality6, echo = FALSE}
-ioi_dyn_var_norm <- by(ioi_dyn_var$Mean, list(ioi_dyn_var$Condition), shapiro.test)
-print(ioi_dyn_var_norm)
-```
-
-```{r stats11, echo = FALSE}
-ioi_dyn_var_ttest <- t.test(ioi_dyn_var$Mean[ioi_dyn_var$Condition == "teaching"], ioi_dyn_var$Mean[ioi_dyn_var$Condition == "performing"], paired = TRUE)
-print(ioi_dyn_var_ttest)
-```
-
-```{r stats12, echo = FALSE}
-ioi_dyn_var_cohend <- cohen.d(ioi_dyn_var$Mean, ioi_dyn_var$Condition, paired = TRUE)
-print(ioi_dyn_var_cohend)
-```
-
-### 9. Variability trial-by-trial - individual{#dyn_ind_var_trial}
-- Average variability for each trial for each participant.
-
-```{r ioi_dyn_var_trial, echo = FALSE}
-# For each individual
-ioi_dyn_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn_var,
-                         FUN = function(x){c(N = length(x), mean = mean(x))})
-ioi_dyn_var_trial <- cbind(ioi_dyn_var_trial[,1:4], as.data.frame(ioi_dyn_var_trial[,5]))
-# Change colnames
-colnames(ioi_dyn_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean")
-print(ioi_dyn_var_trial)
-```
-
-#### Plot
-##### Line graph
-```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 8}
-p_ioi_dyn_var_trial <- ggplot(data = ioi_dyn_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_dyn_var_trial
-```
-
-### 10. Variability trial-by-trial - group{#dyn_group_var_trial}
-Average variability for each trial for each condition.
-
-```{r ioi_dyn_var_trial_all, echo = FALSE}
-# Group mean
-ioi_dyn_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_dyn_var_trial,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_var_trial_all <- cbind(ioi_dyn_var_trial_all[,1:3], as.data.frame(ioi_dyn_var_trial_all[,4]))
-# Change colnames
-colnames(ioi_dyn_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_var_trial_all)
-```
-
-#### Plot
-##### Line graph
-- Error bars show SEM of the mean
-
-```{r ioi_dyn_var_trial_all_line, echo = FALSE}
-p_ioi_dyn_var_trial_all <- ggplot(data = ioi_dyn_var_trial_all, 
-                                  aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line(position = position_dodge(.2)) +
-  geom_point(position = position_dodge(.2)) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_dyn_var_trial_all
-```

### 2019-11-28 script/RNotebook/dynamics.Rmd
add basic analysis for dynamics
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2019-10-21 script/RNotebook/dynamics.Rmd
@@
 ---
-title: 'Analysis: The Sound of Teaching Music 1'
+title: 'Dynamics: Analysis: The Sound of Teaching Music 2'
 output:
@@
 
-# Outline - Dynamics (Key Velocity)
-### [Experimental Design](#design)
-
-### [Dynamics](#vel-dynamics)
-
-#### KV
-- Individual results
-  + [All](#vel_dyn_ind_all)
-  + [Trial by trial](#vel_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#vel_dyn_group_all)
-  + [Trial by trial](#vel_dyn_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#vel_dyn_order_group_all)
-
-#### KV-Diff
-- Individual results
-  + [All](#vel_acc_dyn_ind_all)
-  + [Trial by trial](#vel_acc_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#vel_acc_dyn_group_all)
-  + [Trial by trial](#vel_acc_dyn_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#vel_acc_dyn_order_group_all)
-
-### [Articulation](#vel_articulation)
-
-#### KV
-- Individual results
-  + [All](#vel_art_ind_all)
-  + [Trial by trial](#vel_art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#vel_art_group_all)
-  + [Trial by trial](#vel_art_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#vel_art_order_group_all)
-
-#### KV-Diff
-- Individual results
-  + [All](#vel_acc_art_ind_all)
-  + [Trial by trial](#vel_acc_art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#vel_acc_art_group_all)
-  + [Trial by trial](#vel_acc_art_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#vel_acc_art_order_group_all)
-
 ## Experimental Design{#design}
 - Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato)
-- Dependent variables: Key Velocity (KV), KV difference
+- Dependent variables: Key Velocity (KV), KV difference (KV-Diff)
 
 ### Main predictions
-Participants will exaggerate forte and piano more while teaching.
+- Participants will exaggerate forte and piano more while teaching.
+- Participants will make a larger contrast between forte and piano.
 
@@
 
-#### Statistics
-##### Normality check
-```{r normality1, echo = FALSE}
-# Reduce unused factors
-vel_dyn$Subcomponent <- factor(vel_dyn$Subcomponent)
-vel_dyn_norm <- by(vel_dyn$Mean, list(vel_dyn$Condition, vel_dyn$Subcomponent), shapiro.test)
-print(vel_dyn_norm)
-```
-
-##### Two-way ANOVA (ez)
-```{r stats1, echo = FALSE}
-vel_dyn_aov <- ezANOVA(
-  data = df_vel_dyn
-  , dv = .(Velocity)
-  , wid = .(SubNr)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_dyn_aov$ANOVA$F <- round(vel_dyn_aov$ANOVA$F, 4)
-vel_dyn_aov$ANOVA$p <- round(vel_dyn_aov$ANOVA$p, 4)
-print(vel_dyn_aov$ANOVA)
-```
-
-##### Two-way ANOVA (aov)
-```{r stats2, echo = FALSE}
-vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
-print(summary(vel_dyn_aov_2))
-```
-
-### Post-hoc (direct comparison)
-```{r post-hoc1, echo = FALSE}
-vel_ttest_forte <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"], paired = TRUE)
-vel_ttest_piano <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"], paired = TRUE)
-
-print(vel_ttest_forte)
-print(vel_ttest_piano)
-```
-
-### Post-hoc (pairwise t test)
-```{r post-hoc2, echo = FALSE}
-# Create a group label for a post-hoc test
-vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"] <- "Performing-Forte" 
-vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"] <- "Performing-Piano"
-vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"] <- "Teaching-Forte"
-vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"] <- "Teaching-Piano"
-# pairwise comparison
-vel_dyn_pairwise <- pairwise.t.test(vel_dyn$Mean, vel_dyn$Group, paired = TRUE, p.adjust.method = "BH")
-vel_dyn_pairwise$p.value <- round(vel_dyn_pairwise$p.value, 4)
-print(vel_dyn_pairwise)
-```
+#### Statistics (TBC)
 
@@
 
-## Key Velocity Difference (Difference) - Dynamics
+## Key Velocity Difference (KV-Diff) - Dynamics
 ### 1. All - individual{#vel_acc_dyn_ind_all}
@@
 - Average KV-Diff for each condition.
-- Hypothesis: Participants will exaggerate legato and staccato when teaching.
+- Hypothesis: Participants will make a larger contrast between forte and piano.
 
@@
 
-#### Statistics
-##### Normality check
-```{r normality2, echo = FALSE}
-# Reduce unused factors
-vel_acc_dyn$Subcomponent <- factor(vel_acc_dyn$Subcomponent)
-vel_acc_dyn_norm <- by(vel_acc_dyn$Mean, list(vel_acc_dyn$Condition, vel_acc_dyn$Subcomponent), shapiro.test)
-print(vel_acc_dyn_norm)
-```
-
-##### Two-way ANOVA (ez)
-```{r stats3, echo = FALSE}
-vel_acc_dyn_aov <- ezANOVA(
-  data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF")
-  , dv = .(Acc)
-  , wid = .(SubNr)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_acc_dyn_aov$ANOVA$F <- round(vel_acc_dyn_aov$ANOVA$F, 4)
-vel_acc_dyn_aov$ANOVA$p <- round(vel_acc_dyn_aov$ANOVA$p, 4)
-print(vel_acc_dyn_aov$ANOVA)
-```
-
-##### Two-way ANOVA (aov)
-```{r stats4, echo = FALSE}
-vel_acc_dyn_aov_2 <- with(vel_acc_dyn,
-                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
-print(summary(vel_acc_dyn_aov_2))
-```
-
-### Post-hoc (direct comparison)
-```{r post-hoc3, echo = FALSE}
-vel_acc_ttest_ftop <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "FtoP"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "FtoP"], paired = TRUE)
-vel_acc_ttest_ptof <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)
-
-print(vel_acc_ttest_ftop)
-print(vel_acc_ttest_ptof)
-```
-
-### Post-hoc (pairwise t test)
-```{r post-hoc4, echo = FALSE}
-# Create a group label for a post-hoc test
-vel_acc_dyn$Group[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "FtoP"] <- "Performing-FtoP" 
-vel_acc_dyn$Group[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"] <- "Performing-PtoF"
-vel_acc_dyn$Group[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "FtoP"] <- "Teaching-FtoP"
-vel_acc_dyn$Group[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"] <- "Teaching-PtoF"
-# pairwise comparison
-vel_acc_dyn_pairwise <- pairwise.t.test(vel_acc_dyn$Mean, vel_acc_dyn$Group, paired = TRUE, p.adjust.method = "BH")
-vel_acc_dyn_pairwise$p.value <- round(vel_acc_dyn_pairwise$p.value, 4)
-print(vel_acc_dyn_pairwise)
-```
+#### Statistics (TBC)
 
@@
 
-#### Statistics
-##### Normality check
-```{r normality3, echo = FALSE}
-# Reduce unused factors
-vel_art$Subcomponent <- factor(vel_art$Subcomponent)
-vel_art_norm <- by(vel_art$Mean, list(vel_art$Condition, vel_art$Subcomponent), shapiro.test)
-print(vel_art_norm)
-```
-
-##### Two-way ANOVA (ez)
-```{r stats5, echo = FALSE}
-vel_art_aov <- ezANOVA(
-  data = df_vel_art
-  , dv = .(Velocity)
-  , wid = .(SubNr)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_art_aov$ANOVA$F <- round(vel_art_aov$ANOVA$F, 4)
-vel_art_aov$ANOVA$p <- round(vel_art_aov$ANOVA$p, 4)
-print(vel_art_aov$ANOVA)
-```
-
-##### Two-way ANOVA (aov)
-```{r stats6, echo = FALSE}
-vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
-print(summary(vel_art_aov_2))
-```
-
-### Post-hoc (direct comparison)
-```{r post-hoc5, echo = FALSE}
-vel_ttest_legato <- t.test(vel_art$Mean[vel_art$Condition == "performing" & vel_art$Subcomponent == "Legato"], vel_art$Mean[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Legato"], paired = TRUE)
-vel_ttest_staccato <- t.test(vel_art$Mean[vel_art$Condition == "performing" & vel_art$Subcomponent == "Staccato"], vel_art$Mean[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Staccato"], paired = TRUE)
-
-print(vel_ttest_legato)
-print(vel_ttest_staccato)
-```
-
-### Post-hoc (pairwise t test)
-```{r post-hoc6, echo = FALSE}
-# Create a group label for a post-hoc test
-vel_art$Group[vel_art$Condition == "performing" & vel_art$Subcomponent == "Legato"] <- "Performing-Legato" 
-vel_art$Group[vel_art$Condition == "performing" & vel_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
-vel_art$Group[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Legato"] <- "Teaching-Legato"
-vel_art$Group[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
-# pairwise comparison
-vel_art_pairwise <- pairwise.t.test(vel_art$Mean, vel_art$Group, paired = TRUE, p.adjust.method = "BH")
-vel_art_pairwise$p.value <- round(vel_art_pairwise$p.value, 4)
-print(vel_art_pairwise)
-```
+#### Statistics (TBC)
 
@@
 
-## Key Velocity Difference (Difference) - Articulation
+## Key Velocity Difference (KV-Diff) - Articulation
 ### 1. All - individual{#vel_acc_art_ind_all}
@@
 - Average KV-Diff for each condition.
-- Hypothesis: Participants will exaggerate legato and staccato when teaching.
 
@@
 ```
+
+##### Bar chart
 
@@
 
-#### Statistics
-##### Normality check
-```{r normality4, echo = FALSE}
-# Reduce unused factors
-vel_acc_art$Subcomponent <- factor(vel_acc_art$Subcomponent)
-vel_acc_art_norm <- by(vel_acc_art$Mean, list(vel_acc_art$Condition, vel_acc_art$Subcomponent), shapiro.test)
-print(vel_acc_art_norm)
-```
-
-##### Two-way ANOVA (ez)
-```{r stats7, echo = FALSE}
-vel_acc_art_aov <- ezANOVA(
-  data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL")
-  , dv = .(Acc)
-  , wid = .(SubNr)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_acc_art_aov$ANOVA$F <- round(vel_acc_art_aov$ANOVA$F, 4)
-vel_acc_art_aov$ANOVA$p <- round(vel_acc_art_aov$ANOVA$p, 4)
-print(vel_acc_art_aov$ANOVA)
-```
-
-### Post-hoc (direct comparison)
-```{r post-hoc7, echo = FALSE}
-vel_acc_ttest_legato <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "LtoS"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "LtoS"], paired = TRUE)
-vel_acc_ttest_staccato <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "StoL"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "StoL"], paired = TRUE)
-
-print(vel_acc_ttest_legato)
-print(vel_acc_ttest_staccato)
-```
-
-### Post-hoc (pairwise t test)
-```{r post-hoc8, echo = FALSE}
-# Create a group label for a post-hoc test
-vel_acc_art$Group[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "LtoS"] <- "Performing-LtoS" 
-vel_acc_art$Group[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "StoL"] <- "Performing-StoL"
-vel_acc_art$Group[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "LtoS"] <- "Teaching-LtoS"
-vel_acc_art$Group[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "StoL"] <- "Teaching-StoL"
-# pairwise comparison
-vel_acc_art_pairwise <- pairwise.t.test(vel_acc_art$Mean, vel_acc_art$Group, paired = TRUE, p.adjust.method = "BH")
-vel_acc_art_pairwise$p.value <- round(vel_acc_art_pairwise$p.value, 4)
-print(vel_acc_art_pairwise)
-```
-
-##### Two-way ANOVA (aov)
-```{r stats8, echo = FALSE}
-vel_acc_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_acc_art)
-print(summary(vel_acc_art_aov_2))
-```
+##### Statistics
 
@@
 ```
-
-###########
-###########
-
-## KV-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-### 1. All{#vel_dyn_order_group_all}
-```{r vel_dyn_order_all, echo = all}
-# For each participant
-vel_dyn_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_dyn,
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_dyn_order <- cbind(vel_dyn_order[,1:5], as.data.frame(vel_dyn_order[,6]))
-# Change colnames
-colnames(vel_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_dyn_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_dyn_order_all <- cbind(vel_dyn_order_all[,1:4], as.data.frame(vel_dyn_order_all[,5]))
-# Change colnames
-colnames(vel_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_dyn_order_all_bar, echo = FALSE}
-p_vel_dyn_order_all <- ggplot(data = vel_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
-p_vel_dyn_order_all
-```
-
-##### Box plot
-```{r vel_dyn_order_all_box, echo = FALSE}
-p_vel_dyn_order_all_box <- ggplot(data = vel_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
-p_vel_dyn_order_all_box
-```
-
-#### Statistics
-##### Normality check
-```{r normality5, echo = FALSE}
-# Normality check
-vel_dyn_order$Subcomponent <- factor(vel_dyn_order$Subcomponent)
-vel_dyn_order_norm <- by(vel_dyn_order$Mean, list(vel_dyn_order$Condition, vel_dyn_order$Subcomponent, vel_dyn_order$Order), shapiro.test)
-print(vel_dyn_order_norm)
-```
-
-##### Three-way ANOVA (ez)
-```{r stats9, echo = FALSE}
-vel_dyn_order_aov <- ezANOVA(
-  data = df_vel_dyn
-  , dv = .(Velocity)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_dyn_order_aov$ANOVA$F <- round(vel_dyn_order_aov$ANOVA$F, 4)
-vel_dyn_order_aov$ANOVA$p <- round(vel_dyn_order_aov$ANOVA$p, 4)
-print(vel_dyn_order_aov)
-```
-
-## KV-Diff-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-### 1. All{#vel_acc_dyn_order_group_all}
-```{r vel_acc_dyn_order_all, echo = all}
-# For each participant
-vel_acc_dyn_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_acc_dyn_order <- cbind(vel_acc_dyn_order[,1:5], as.data.frame(vel_acc_dyn_order[,6]))
-# Change colnames
-colnames(vel_acc_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_acc_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_dyn_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_acc_dyn_order_all <- cbind(vel_acc_dyn_order_all[,1:4], as.data.frame(vel_acc_dyn_order_all[,5]))
-# Change colnames
-colnames(vel_acc_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_acc_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_acc_dyn_order_all_bar, echo = FALSE}
-p_vel_acc_dyn_order_all <- ggplot(data = vel_acc_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
-p_vel_acc_dyn_order_all
-```
-
-##### Box plot
-```{r vel_acc_dyn_order_all_box, echo = FALSE}
-p_vel_acc_dyn_order_all_box <- ggplot(data = vel_acc_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV-Diff: Dynamics") 
-p_vel_acc_dyn_order_all_box
-```
-
-#### Statistics
-##### Normality check
-```{r normality6, echo = FALSE}
-# Normality check
-vel_acc_dyn_order$Subcomponent <- factor(vel_acc_dyn_order$Subcomponent)
-vel_acc_dyn_order_norm <- by(vel_acc_dyn_order$Mean, list(vel_acc_dyn_order$Condition, vel_acc_dyn_order$Subcomponent, vel_acc_dyn_order$Order), shapiro.test)
-print(vel_acc_dyn_order_norm)
-```
-
-##### Three-way ANOVA (ez)
-```{r stats10, echo = FALSE}
-vel_acc_dyn_order_aov <- ezANOVA(
-  data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF")
-  , dv = .(Acc)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_acc_dyn_order_aov$ANOVA$F <- round(vel_acc_dyn_order_aov$ANOVA$F, 4)
-vel_acc_dyn_order_aov$ANOVA$p <- round(vel_acc_dyn_order_aov$ANOVA$p, 4)
-print(vel_acc_dyn_order_aov)
-```
-
-## KV-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-### 1. All{#vel_art_order_group_all}
-```{r vel_art_order_all, echo = all}
-# For each participant
-vel_art_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_art,
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_art_order <- cbind(vel_art_order[,1:5], as.data.frame(vel_art_order[,6]))
-# Change colnames
-colnames(vel_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_art_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_art_order_all <- cbind(vel_art_order_all[,1:4], as.data.frame(vel_art_order_all[,5]))
-# Change colnames
-colnames(vel_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_art_order_all_bar, echo = FALSE}
-p_vel_art_order_all <- ggplot(data = vel_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
-p_vel_art_order_all
-```
-
-##### Box plot
-```{r vel_art_order_all_box, echo = FALSE}
-p_vel_art_order_all_box <- ggplot(data = vel_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Articulation") 
-p_vel_art_order_all_box
-```
-
-#### Statistics
-##### Normality check
-```{r normality7, echo = FALSE}
-# Normality check
-vel_art_order$Subcomponent <- factor(vel_art_order$Subcomponent)
-vel_art_order_norm <- by(vel_art_order$Mean, list(vel_art_order$Condition, vel_art_order$Subcomponent, vel_art_order$Order), shapiro.test)
-print(vel_art_order_norm)
-```
-
-##### Three-way ANOVA (ez)
-```{r stats11, echo = FALSE}
-vel_art_order_aov <- ezANOVA(
-  data = df_vel_art
-  , dv = .(Velocity)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_art_order_aov$ANOVA$F <- round(vel_art_order_aov$ANOVA$F, 4)
-vel_art_order_aov$ANOVA$p <- round(vel_art_order_aov$ANOVA$p, 4)
-print(vel_art_order_aov)
-```
-
-## KV-Diff-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-### 1. All{#vel_acc_art_order_group_all}
-```{r vel_acc_art_order_all, echo = all}
-# For each participant
-vel_acc_art_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_acc_art_order <- cbind(vel_acc_art_order[,1:5], as.data.frame(vel_acc_art_order[,6]))
-# Change colnames
-colnames(vel_acc_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_acc_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_art_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_acc_art_order_all <- cbind(vel_acc_art_order_all[,1:4], as.data.frame(vel_acc_art_order_all[,5]))
-# Change colnames
-colnames(vel_acc_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_acc_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_acc_art_order_all_bar, echo = FALSE}
-p_vel_acc_art_order_all <- ggplot(data = vel_acc_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
-p_vel_acc_art_order_all
-```
-
-##### Box plot
-```{r vel_acc_art_order_all_box, echo = FALSE}
-p_vel_acc_art_order_all_box <- ggplot(data = vel_acc_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV: Articulation") 
-p_vel_acc_art_order_all_box
-```
-
-#### Statistics
-##### Normality check
-```{r normality8, echo = FALSE}
-# Normality check
-vel_acc_art_order$Subcomponent <- factor(vel_acc_art_order$Subcomponent)
-vel_acc_art_order_norm <- by(vel_acc_art_order$Mean, list(vel_acc_art_order$Condition, vel_acc_art_order$Subcomponent, vel_acc_art_order$Order), shapiro.test)
-print(vel_acc_art_order_norm)
-```
-
-##### Three-way ANOVA (ez)
-```{r stats12, echo = FALSE}
-vel_acc_art_order_aov <- ezANOVA(
-  data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL")
-  , dv = .(Acc)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_acc_art_order_aov$ANOVA$F <- round(vel_acc_art_order_aov$ANOVA$F, 4)
-vel_acc_art_order_aov$ANOVA$p <- round(vel_acc_art_order_aov$ANOVA$p, 4)
-print(vel_acc_art_order_aov)
-```

### 2019-11-28 script/RNotebook/dynamics.Rmd
add seq plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```
+
+## Sequence plot
+#### Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 15, fig.height = 5, echo = FALSE}
+# Average vel_dynocity for each note
+# For each participant
+vel_dyn_seq <- aggregate(Velocity~SubNr*Condition*Skill*NoteNr, data = df_vel_dyn, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_dyn_seq <- cbind(vel_dyn_seq[,1:4], as.data.frame(vel_dyn_seq[,5]))
+# Change colnames
+colnames(vel_dyn_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD")
+
+# Group mean
+vel_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_dyn_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_dyn_seq_stats <- cbind(vel_dyn_seq_stats[,1:3], as.data.frame(vel_dyn_seq_stats[,4]))
+# Change colnames
+colnames(vel_dyn_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM")
+
+# Average acceleration for each note
+# For each participant
+vel_dyn_acc_seq <- aggregate(Acc~SubNr*Condition*Skill*Interval, data = df_vel_acc_dyn, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_dyn_acc_seq <- cbind(vel_dyn_acc_seq[,1:4], as.data.frame(vel_dyn_acc_seq[,5]))
+# Change colnames
+colnames(vel_dyn_acc_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+vel_dyn_acc_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_dyn_acc_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_dyn_acc_seq_stats <- cbind(vel_dyn_acc_seq_stats[,1:3], as.data.frame(vel_dyn_acc_seq_stats[,4]))
+# Change colnames
+colnames(vel_dyn_acc_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_vel_dyn_seq <- ggplot(data = vel_dyn_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(0.05)) + 
+  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
+p_vel_dyn_seq
+
+p_vel_dyn_acc_seq <- ggplot(data = vel_dyn_acc_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
+p_vel_dyn_acc_seq
+```
+
+#### Articulation{#sequence_art}
+```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
+# Average vel_artocity for each note
+# For each participant
+vel_art_seq <- aggregate(Velocity~SubNr*Condition*Skill*NoteNr, data = df_vel_art, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_art_seq <- cbind(vel_art_seq[,1:4], as.data.frame(vel_art_seq[,5]))
+# Change colnames
+colnames(vel_art_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD")
+
+# Group mean
+vel_art_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_art_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_art_seq_stats <- cbind(vel_art_seq_stats[,1:3], as.data.frame(vel_art_seq_stats[,4]))
+# Change colnames
+colnames(vel_art_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM")
+
+# Average acceleration for each note
+# For each participant
+vel_art_acc_seq <- aggregate(Acc~SubNr*Condition*Skill*Interval, data = df_vel_acc_art, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_art_acc_seq <- cbind(vel_art_acc_seq[,1:4], as.data.frame(vel_art_acc_seq[,5]))
+# Change colnames
+colnames(vel_art_acc_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+vel_art_acc_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_art_acc_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_art_acc_seq_stats <- cbind(vel_art_acc_seq_stats[,1:3], as.data.frame(vel_art_acc_seq_stats[,4]))
+# Change colnames
+colnames(vel_art_acc_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_vel_art_seq <- ggplot(data = vel_art_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(0.05)) + 
+  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
+p_vel_art_seq
+
+p_vel_art_acc_seq <- ggplot(data = vel_art_acc_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
+p_vel_art_acc_seq
+```

### 2019-11-28 script/RNotebook/ioi.Rmd
add seq plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ##### Line graph
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the ideal tempo.
 
@@
   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
+  labs(x = "Trial", y = "Normalised IOI", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
 p_ioi_dyn_trial_all
 ```
+
+## Sequence plot
+#### Articulation{#sequence_art}
+```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
+# For each individual
+ioi_art_seq <- aggregate(normIOI~SubNr*Condition*Skill*Interval, data = df_ioi_art,
+                                 FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+ioi_art_seq <- cbind(ioi_art_seq[,1:4], as.data.frame(ioi_art_seq[,5]))
+# Change colnames
+colnames(ioi_art_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+ioi_art_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = ioi_art_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_art_seq_stats <- cbind(ioi_art_seq_stats[,1:3], as.data.frame(ioi_art_seq_stats[,4]))
+# Change colnames
+colnames(ioi_art_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_ioi_art_seq <- ggplot(data = ioi_art_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
+p_ioi_art_seq
+```
+
+#### Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 15, fig.height = 5, echo = FALSE}
+# For each individual
+ioi_dyn_seq <- aggregate(normIOI~SubNr*Condition*Skill*Interval, data = df_ioi_dyn,
+                                 FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+ioi_dyn_seq <- cbind(ioi_dyn_seq[,1:4], as.data.frame(ioi_dyn_seq[,5]))
+# Change colnames
+colnames(ioi_dyn_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+ioi_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = ioi_dyn_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_dyn_seq_stats <- cbind(ioi_dyn_seq_stats[,1:3], as.data.frame(ioi_dyn_seq_stats[,4]))
+# Change colnames
+colnames(ioi_dyn_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_ioi_dyn_seq <- ggplot(data = ioi_dyn_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
+p_ioi_dyn_seq
+```

### 2019-11-28 script/RNotebook/ioi.Rmd
tweak comments/description
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2019-10-22 script/RNotebook/ioi.Rmd
@@
 - Error bars show SEM of the mean
+- A dashed line represents the tempo given by a metronome (80bpm)
 
@@
 - Error bars show SD of the mean
+- A dashed line represents the tempo given by a metronome (80bpm)
 
@@
 ##### Box plot
+- A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_art_order_all_box, echo = FALSE}
@@
 ##### Box plot
+- A dashed line represents the tempo given by a metronome (80bpm)
+
 ```{r ioi_dyn_order_all_box, echo = FALSE}

### 2019-12-10 script/RNotebook/articulation.Rmd
tweak comments/description
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
add articulation.Rmd
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2019-10-20 script/RNotebook/articulation.Rmd
@@
 ---
-title: 'Analysis: The Sound of Teaching Music 1'
+title: 'Analysis: The Sound of Teaching Music 2'
 output:
@@
 ##### Description: This is a summary of all analyses for articulation.
-
-# Outline - Articulation (KOT, KOR)
-### [Experimental Design](#design)
-
-### [Articulation](#kot-articulation)
-
-#### KOT
-- Individual results
-  + [All](#kot_art_ind_all)
-  + [Trial by trial](#kot_art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kot_art_group_all)
-  + [Trial by trial](#kot_art_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#kot_art_order_group_all)
-
-#### KOR
-- Individual results
-  + [All](#kor_art_ind_all)
-  + [Trial by trial](#kor_art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kor_art_group_all)
-  + [Trial by trial](#kor_art_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#kor_art_order_group_all)
-
-### [Dynamics](#kot-dynamics)
-
-#### KOT
-- Individual results
-  + [All](#kot_dyn_ind_all)
-  + [Trial by trial](#kot_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kot_dyn_group_all)
-  + [Trial by trial](#kot_dyn_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#kot_dyn_order_group_all)
-
-#### KOR
-- Individual results
-  + [All](#kor_dyn_ind_all)
-  + [Trial by trial](#kor_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kor_dyn_group_all)
-  + [Trial by trial](#kor_dyn_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#kor_dyn_order_group_all)
 
@@
 # For each individual
-kot_art <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57),
+kot_art <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 #### Statistics
-##### Normality check
-```{r normality1, echo = FALSE}
-# Reduce unused factors
-kot_art$Subcomponent <- factor(kot_art$Subcomponent)
-kot_art_norm <- by(kot_art$Mean, list(kot_art$Condition, kot_art$Subcomponent), shapiro.test)
-print(kot_art_norm)
-```
-
-##### Two-way ANOVA (ez)
-```{r stats1, echo = FALSE}
-kot_art_aov <- ezANOVA(
-  data = subset(df_kot_art, df_kot_art$Subcomponent == "Legato" | df_kot_art$Subcomponent == "Staccato")
-  , dv = .(KOT)
-  , wid = .(SubNr)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kot_art_aov$ANOVA$F <- round(kot_art_aov$ANOVA$F, 4)
-kot_art_aov$ANOVA$p <- round(kot_art_aov$ANOVA$p, 4)
-print(kot_art_aov$ANOVA)
-```
-
-##### Two-way ANOVA (aov)
-```{r stats2, echo = FALSE}
-kot_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
-print(summary(kot_art_aov_2))
-```
-
-### Post-hoc (direct comparison)
-```{r post-hoc1, echo = FALSE}
-kot_ttest_legato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"], paired = TRUE)
-kot_ttest_staccato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"], paired = TRUE)
-
-print(kot_ttest_legato)
-print(kot_ttest_staccato)
-```
-
-### Post-hoc (pairwise t test)
-```{r post-hoc2, echo = FALSE}
-# Create a group label for a post-hoc test
-kot_art$Group[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"] <- "Performing-Legato" 
-kot_art$Group[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
-kot_art$Group[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"] <- "Teaching-Legato"
-kot_art$Group[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
-# pairwise comparison
-kot_art_pairwise <- pairwise.t.test(kot_art$Mean, kot_art$Group, paired = TRUE, p.adjust.method = "BH")
-kot_art_pairwise$p.value <- round(kot_art_pairwise$p.value, 4)
-print(kot_art_pairwise)
-```
+##### TBC
 
@@
 # For each individual
-kot_art_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57),
+kot_art_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 # For each individual
-kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57),
+kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 #### Statistics
-##### Normality check
-```{r normality2, echo = FALSE}
-# Reduce unused factors
-kor_art$Subcomponent <- factor(kor_art$Subcomponent)
-kor_art_norm <- by(kor_art$Mean, list(kor_art$Condition, kor_art$Subcomponent), shapiro.test)
-print(kor_art_norm)
-```
-
-##### Two-way ANOVA (ez)
-```{r stats3, echo = FALSE}
-kor_art_aov <- ezANOVA(
-  data = subset(df_kor_art, df_kor_art$Subcomponent == "Legato" | df_kor_art$Subcomponent == "Staccato")
-  , dv = .(KOR)
-  , wid = .(SubNr)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kor_art_aov$ANOVA$F <- round(kor_art_aov$ANOVA$F, 4)
-kor_art_aov$ANOVA$p <- round(kor_art_aov$ANOVA$p, 4)
-print(kor_art_aov$ANOVA)
-```
-
-##### Two-way ANOVA (aov)
-```{r stats4, echo = FALSE}
-kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
-print(summary(kor_art_aov_2))
-```
-
-### Post-hoc (direct comparison)
-```{r post-hoc3, echo = FALSE}
-kor_ttest_legato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"], paired = TRUE)
-kor_ttest_staccato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"], paired = TRUE)
-
-print(kor_ttest_legato)
-print(kor_ttest_staccato)
-```
-
-### Post-hoc (pairwise t test)
-```{r post-hoc4, echo = FALSE}
-# Create a group label for a post-hoc test
-kor_art$Group[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"] <- "Performing-Legato" 
-kor_art$Group[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
-kor_art$Group[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"] <- "Teaching-Legato"
-kor_art$Group[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
-# pairwise comparison
-kor_art_pairwise <- pairwise.t.test(kor_art$Mean, kor_art$Group, paired = TRUE, p.adjust.method = "BH")
-kor_art_pairwise$p.value <- round(kor_art_pairwise$p.value, 4)
-print(kor_art_pairwise)
-```
+##### TBC
 
@@
 # For each individual
-kot_dyn <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57),
+kot_dyn <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 - Average KOT for each condition.
-- Hypothesis: Participants will exaggerate legato and staccato when teaching.
 
@@
 #### Statistics
-##### Normality check
-```{r normality3, echo = FALSE}
-# Reduce unused factors
-kot_dyn$Subcomponent <- factor(kot_dyn$Subcomponent)
-kot_dyn_norm <- by(kot_dyn$Mean, list(kot_dyn$Condition, kot_dyn$Subcomponent), shapiro.test)
-print(kot_dyn_norm)
-```
-
-##### Two-way ANOVA (ez)
-```{r stats5, echo = FALSE}
-kot_dyn_aov <- ezANOVA(
-  data = subset(df_kot_dyn, df_kot_dyn$Subcomponent == "Forte" | df_kot_dyn$Subcomponent == "Piano")
-  , dv = .(KOT)
-  , wid = .(SubNr)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kot_dyn_aov$ANOVA$F <- round(kot_dyn_aov$ANOVA$F, 4)
-kot_dyn_aov$ANOVA$p <- round(kot_dyn_aov$ANOVA$p, 4)
-print(kot_dyn_aov$ANOVA)
-```
-
-##### Two-way ANOVA (aov)
-```{r stats6, echo = FALSE}
-kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
-print(summary(kot_dyn_aov_2))
-```
+##### TBC
 
@@
 # For each individual
-kot_dyn_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57),
+kot_dyn_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 # For each individual
-kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
+kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 - Average KOR for each condition.
-- Hypothesis: Participants will exaggerate legato and staccato when teaching.
 
@@
 #### Statistics
-##### Normality check
-```{r normality4, echo = FALSE}
-# Reduce unused factors
-kor_dyn$Subcomponent <- factor(kor_dyn$Subcomponent)
-kor_dyn_norm <- by(kor_dyn$Mean, list(kor_dyn$Condition, kor_dyn$Subcomponent), shapiro.test)
-print(kor_dyn_norm)
-```
-
-##### Two-way ANOVA (ez)
-```{r stats7, echo = FALSE}
-kor_dyn_aov <- ezANOVA(
-  data = subset(df_kor_dyn, df_kor_dyn$Subcomponent == "Forte" | df_kor_dyn$Subcomponent == "Piano")
-  , dv = .(KOR)
-  , wid = .(SubNr)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kor_dyn_aov$ANOVA$F <- round(kor_dyn_aov$ANOVA$F, 4)
-kor_dyn_aov$ANOVA$p <- round(kor_dyn_aov$ANOVA$p, 4)
-print(kor_dyn_aov$ANOVA)
-```
-
-##### Two-way ANOVA (aov)
-```{r stats8, echo = FALSE}
-kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
-print(summary(kor_dyn_aov_2))
-```
+##### TBC
 
@@
 # For each individual
-kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
+kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 ```
-
-###########
-###########
-
-## KOT-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-### 1. All{#kot_art_order_group_all}
-```{r kot_art_order_all, echo = FALSE}
-# For each participant
-kot_art_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-kot_art_order <- cbind(kot_art_order[,1:5], as.data.frame(kot_art_order[,6]))
-# Change colnames
-colnames(kot_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-kot_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_art_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kot_art_order_all <- cbind(kot_art_order_all[,1:4], as.data.frame(kot_art_order_all[,5]))
-# Change colnames
-colnames(kot_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(kot_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r kot_art_order_all_bar, echo = FALSE}
-p_kot_art_order_all <- ggplot(data = kot_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
-p_kot_art_order_all
-```
-
-##### Box plot
-```{r kot_art_order_all_box, echo = FALSE}
-p_kot_art_order_all_box <- ggplot(data = kot_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOT", subtitle = "KOT: Articulation")
-p_kot_art_order_all_box
-```
-
-#### Statistics
-##### Normality check
-```{r normality5, echo = FALSE}
-# Normality check
-kot_art_order$Subcomponent <- factor(kot_art_order$Subcomponent)
-kot_art_order_norm <- by(kot_art_order$Mean, list(kot_art_order$Condition, kot_art_order$Subcomponent, kot_art_order$Order), shapiro.test)
-print(kot_art_order_norm)
-```
-
-##### Three-way ANOVA (ez)
-```{r stats9, echo = FALSE}
-kot_art_order_aov <- ezANOVA(
-  data = df_kot_art
-  , dv = .(KOT)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kot_art_order_aov$ANOVA$F <- round(kot_art_order_aov$ANOVA$F, 4)
-kot_art_order_aov$ANOVA$p <- round(kot_art_order_aov$ANOVA$p, 4)
-print(kot_art_order_aov)
-```
-
-## KOR-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-### 1. All{#kor_art_order_group_all}
-```{r kor_art_order_all, echo = FALSE}
-# For each participant
-kor_art_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-kor_art_order <- cbind(kor_art_order[,1:5], as.data.frame(kor_art_order[,6]))
-# Change colnames
-colnames(kor_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-kor_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_art_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kor_art_order_all <- cbind(kor_art_order_all[,1:4], as.data.frame(kor_art_order_all[,5]))
-# Change colnames
-colnames(kor_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(kor_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r kor_art_order_all_bar, echo = FALSE}
-p_kor_art_order_all <- ggplot(data = kor_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Articulation")
-p_kor_art_order_all
-```
-
-##### Box plot
-```{r kor_art_order_all_box, echo = FALSE}
-p_kor_art_order_all_box <- ggplot(data = kor_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Articulation")
-p_kor_art_order_all_box
-```
-
-#### Statistics
-##### Normality check
-```{r normality6, echo = FALSE}
-# Normality check
-kor_art_order$Subcomponent <- factor(kor_art_order$Subcomponent)
-kor_art_order_norm <- by(kor_art_order$Mean, list(kor_art_order$Condition, kor_art_order$Subcomponent, kor_art_order$Order), shapiro.test)
-print(kor_art_order_norm)
-```
-
-##### Three-way ANOVA (ez)
-```{r stats10, echo = FALSE}
-kor_art_order_aov <- ezANOVA(
-  data = df_kor_art
-  , dv = .(KOR)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kor_art_order_aov$ANOVA$F <- round(kor_art_order_aov$ANOVA$F, 4)
-kor_art_order_aov$ANOVA$p <- round(kor_art_order_aov$ANOVA$p, 4)
-print(kor_art_order_aov)
-```
-
-## KOT-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-### 1. All{#kot_dyn_order_group_all}
-```{r kot_dyn_order_all, echo = FALSE}
-# For each participant
-kot_dyn_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-kot_dyn_order <- cbind(kot_dyn_order[,1:5], as.data.frame(kot_dyn_order[,6]))
-# Change colnames
-colnames(kot_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-kot_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_dyn_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kot_dyn_order_all <- cbind(kot_dyn_order_all[,1:4], as.data.frame(kot_dyn_order_all[,5]))
-# Change colnames
-colnames(kot_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(kot_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r kot_dyn_order_all_bar, echo = FALSE}
-p_kot_dyn_order_all <- ggplot(data = kot_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
-p_kot_dyn_order_all
-```
-
-##### Box plot
-```{r kot_dyn_order_all_box, echo = FALSE}
-p_kot_dyn_order_all_box <- ggplot(data = kot_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
-p_kot_dyn_order_all_box
-```
-
-#### Statistics
-##### Normality check
-```{r normality7, echo = FALSE}
-# Normality check
-kot_dyn_order$Subcomponent <- factor(kot_dyn_order$Subcomponent)
-kot_dyn_order_norm <- by(kot_dyn_order$Mean, list(kot_dyn_order$Condition, kot_dyn_order$Subcomponent, kot_dyn_order$Order), shapiro.test)
-print(kot_dyn_order_norm)
-```
-
-##### Three-way ANOVA (ez)
-```{r stats11, echo = FALSE}
-kot_dyn_order_aov <- ezANOVA(
-  data = df_kot_dyn
-  , dv = .(KOT)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kot_dyn_order_aov$ANOVA$F <- round(kot_dyn_order_aov$ANOVA$F, 4)
-kot_dyn_order_aov$ANOVA$p <- round(kot_dyn_order_aov$ANOVA$p, 4)
-print(kot_dyn_order_aov)
-```
-
-## KOR-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-### 1. All{#kor_dyn_order_group_all}
-```{r kor_dyn_order_all, echo = FALSE}
-# For each participant
-kor_dyn_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-kor_dyn_order <- cbind(kor_dyn_order[,1:5], as.data.frame(kor_dyn_order[,6]))
-# Change colnames
-colnames(kor_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-kor_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_dyn_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kor_dyn_order_all <- cbind(kor_dyn_order_all[,1:4], as.data.frame(kor_dyn_order_all[,5]))
-# Change colnames
-colnames(kor_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(kor_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r kor_dyn_order_all_bar, echo = FALSE}
-p_kor_dyn_order_all <- ggplot(data = kor_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Dynamics")
-p_kor_dyn_order_all
-```
-
-##### Box plot
-```{r kor_dyn_order_all_box, echo = FALSE}
-p_kor_dyn_order_all_box <- ggplot(data = kor_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Dynamics")
-p_kor_dyn_order_all_box
-```
-
-#### Statistics
-##### Normality check
-```{r normality8, echo = FALSE}
-# Normality check
-kor_dyn_order$Subcomponent <- factor(kor_dyn_order$Subcomponent)
-kor_dyn_order_norm <- by(kor_dyn_order$Mean, list(kor_dyn_order$Condition, kor_dyn_order$Subcomponent, kor_dyn_order$Order), shapiro.test)
-print(kor_dyn_order_norm)
-```
-
-##### Three-way ANOVA (ez)
-```{r stats12, echo = FALSE}
-kor_dyn_order_aov <- ezANOVA(
-  data = df_kor_dyn
-  , dv = .(KOR)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kor_dyn_order_aov$ANOVA$F <- round(kor_dyn_order_aov$ANOVA$F, 4)
-kor_dyn_order_aov$ANOVA$p <- round(kor_dyn_order_aov$ANOVA$p, 4)
-print(kor_dyn_order_aov)
-```

### 2020-01-06 script/RNotebook/ioi_1.Rmd
add stats
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2019-11-28 script/RNotebook/ioi.Rmd
@@
 ##### Description: This is a summary of all analyses for a tempo (IOIs).
+  - data <- data_ioi_1.csv
 
@@
 - Independent variables: Condition (2 levels: teaching vs. performing) x Skill (2 levels: articulation vs. dynamics)
-- Dependent variables: Interonset intervals, (**Not yet**: variability (SD/mean IOI))
+- Dependent variables: Interonset intervals, variability (SD/mean IOI)
 
@@
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_ioi <- read.csv("./trimmed/data_ioi_1.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 
@@
 
-#### Statistics (TBC)
+#### Statistics
+##### Normality check
+```{r normality1, echo = FALSE}
+# Normality check
+ioi_art_norm <- by(ioi_art$Mean, list(ioi_art$Condition), shapiro.test)
+print(ioi_art_norm)
+```
+
+##### Paired t-test
+```{r stats1, echo = FALSE}
+ioi_art_ttest <- t.test(ioi_art$Mean[ioi_art$Condition == "teaching"], ioi_art$Mean[ioi_art$Condition == "performing"], paired = TRUE)
+print(ioi_art_ttest)
+```
+
+##### Effect size
+```{r stats2, echo = FALSE}
+ioi_art_cohend <- cohen.d(ioi_art$Mean, ioi_art$Condition, paired = TRUE)
+print(ioi_art_cohend)
+```
 
@@
 ```{r ioi_art_ch, echo = FALSE}
-# Define subcomponent change
-df_ioi_art$Change <- "No"
-df_ioi_art$Change[df_ioi_art$Interval == 5 | df_ioi_art$Interval == 8 | df_ioi_art$Interval == 17 | df_ioi_art$Interval == 20 | df_ioi_art$Interval == 39 | df_ioi_art$Interval == 47] <- "Yes"
+# Define phrase boundaries
+df_ioi_art$Boundary <- "No"
+df_ioi_art$Boundary[df_ioi_dyn$Subcomponent == "LtoS" | df_ioi_dyn$Subcomponent == "StoL" | df_ioi_dyn$Subcomponent == "FtoP" | df_ioi_art$Subcomponent == "PtoF"] <- "Yes"
 
 # For each individual
-ioi_art_ch <- aggregate(normIOI~SubNr*Condition*Skill*Change, data = df_ioi_art,
+ioi_art_ch <- aggregate(normIOI~SubNr*Condition*Skill*Boundary, data = df_ioi_art,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 # Change colnames
-colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Change", "N", "Mean", "SD")
+colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
 print(ioi_art_ch)
@@
                 width = .2, position = position_dodge(.9)) +
-  facet_grid(Change ~ .) +
+  facet_grid(Boundary ~ .) +
   geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
@@
 # Group mean
-ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Change, data = ioi_art_ch,
+ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_art_ch,
                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
@@
 # Change colnames
-colnames(ioi_art_ch_all) <- c("Condition", "Skill", "Change", "N", "Mean", "SD", "SEM")
+colnames(ioi_art_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
 print(ioi_art_ch_all)
@@
 ```{r ioi_art_ch_all_bar, echo = FALSE}
-p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Change, y = Mean, fill = Condition)) +
+p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
 ```{r ioi_art_ch_all_box, echo = FALSE}
-p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Change, y = Mean, fill = Condition)) +
+p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
   geom_boxplot() +
@@
 
-#### Statistics (TBC)
+#### Statistics
+##### Normality check
+```{r normality2, echo = FALSE}
+ioi_art_ch_norm <- by(ioi_art_ch$Mean, list(ioi_art_ch$Condition, ioi_art_ch$Boundary), shapiro.test)
+print(ioi_art_ch_norm)
+```
+
+##### Two-way ANOVA (ez)
+```{r stats3, echo = FALSE}
+ioi_art_ch_aov <- ezANOVA(
+  data = df_ioi_art
+  , dv = .(normIOI)
+  , wid = .(SubNr)
+  , within = .(Condition, Boundary)
+  , type = 3
+  , detailed = TRUE
+)
+ioi_art_ch_aov$ANOVA$F <- round(ioi_art_ch_aov$ANOVA$F, 4)
+ioi_art_ch_aov$ANOVA$p <- round(ioi_art_ch_aov$ANOVA$p, 4)
+print(ioi_art_ch_aov$ANOVA)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats4, echo = FALSE}
+ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
+print(summary(ioi_art_ch_aov_2))
+```
 
@@
 p_ioi_art_trial_all
+```
+
+### 7. Variability all - individual{#art_ind_var_all}
+- Average variability for each participant.
+- normIOI variability = SD/mean IOI (per trial)
+
+```{r ioi_art_var, echo = FALSE}
+# Create a data frame for variability (SD/)
+df_ioi_art_var <- data.frame()
+for (subnr in unique(df_ioi_art$SubNr)){
+  for (block in unique(df_ioi_art$BlockNr)){
+    cond = as.character(unique(df_ioi_art$Condition[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
+    skill = as.character(unique(df_ioi_art$Skill[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
+    for (trial in unique(df_ioi_art$TrialNr)){
+      df_current <- df_ioi_art %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
+      if (nrow(df_current) != 0){
+        df_ioi_art_var <- rbind(df_ioi_art_var, 
+                                data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI)))
+      }
+    }
+  }
+}
+colnames(df_ioi_art_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability")
+
+# Average variability
+# For each individual
+ioi_art_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_art_var,
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+ioi_art_var <- cbind(ioi_art_var[,1:3], as.data.frame(ioi_art_var[,4]))
+# Change colnames
+colnames(ioi_art_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
+print(ioi_art_var)
+```
+
+#### Plot
+##### Bar graph
+- Error bars show SD of the mean
+- A dashed line represents the tempo given by a metronome (80bpm)
+
+```{r ioi_art_var_bar, echo = FALSE}
+p_ioi_art_var <- ggplot(data = ioi_art_var, aes(x = SubNr, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
+                width = .2, position = position_dodge(1)) +
+  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation")
+p_ioi_art_var
+```
+
+### 8. Variability all - group{#art_group_var_all}
+- Average variability for each condition.
+
+```{r ioi_art_var_all, echo = FALSE}
+# Group mean
+ioi_art_var_all <- aggregate(Mean~Condition*Skill, data = ioi_art_var,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_art_var_all <- cbind(ioi_art_var_all[,1:2], as.data.frame(ioi_art_var_all[,3]))
+# Change colnames
+colnames(ioi_art_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
+print(ioi_art_var_all)
+```
+
+#### Plot
+##### Bar graph
+- Error bars show SEM of the mean
+
+```{r ioi_art_var_all_bar, echo = FALSE}
+p_ioi_art_var_all <- ggplot(data = ioi_art_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge()) +
+  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
+  theme(legend.position = "None")
+p_ioi_art_var_all
+```
+
+##### Box plot
+```{r ioi_art_var_all_box, echo = FALSE}
+p_ioi_art_var_box <- ggplot(data = ioi_art_var, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
+  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
+  theme(legend.position = "None")
+p_ioi_art_var_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality3, echo = FALSE}
+ioi_art_var_norm <- by(ioi_art_var$Mean, list(ioi_art_var$Condition), shapiro.test)
+print(ioi_art_var_norm)
+```
+- Violate the normality assumption
+
+##### Paired t-test
+```{r stats5, echo = FALSE}
+ioi_art_var_ttest <- t.test(ioi_art_var$Mean[ioi_art_var$Condition == "teaching"], ioi_art_var$Mean[ioi_art_var$Condition == "performing"], paired = TRUE)
+print(ioi_art_var_ttest)
+```
+
+##### Effect size
+```{r stats6, echo = FALSE}
+ioi_art_var_cohend <- cohen.d(ioi_art_var$Mean, ioi_art_var$Condition, paired = TRUE)
+print(ioi_art_var_cohend)
+```
+
+### 9. Variability trial-by-trial - individual{#art_ind_var_trial}
+- Average variability for each trial for each participant.
+
+```{r ioi_art_var_trial, echo = FALSE}
+# For each individual
+ioi_art_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_art_var,
+                         FUN = function(x){c(N = length(x), mean = mean(x))})
+ioi_art_var_trial <- cbind(ioi_art_var_trial[,1:4], as.data.frame(ioi_art_var_trial[,5]))
+# Change colnames
+colnames(ioi_art_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean")
+print(ioi_art_var_trial)
+```
+
+#### Plot
+##### Line graph
+```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 8}
+p_ioi_art_var_trial <- ggplot(data = ioi_art_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  facet_wrap(SubNr ~ .) + 
+  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
+p_ioi_art_var_trial
+```
+
+### 10. Variability trial-by-trial - group{#art_group_var_trial}
+- Average variability for each trial for each condition.
+
+```{r ioi_art_var_trial_all, echo = FALSE}
+# Group mean
+ioi_art_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_art_var_trial,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_art_var_trial_all <- cbind(ioi_art_var_trial_all[,1:3], as.data.frame(ioi_art_var_trial_all[,4]))
+# Change colnames
+colnames(ioi_art_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
+print(ioi_art_var_trial_all)
+```
+
+#### Plot
+##### Line graph
+- Error bars show SEM of the mean
+
+```{r ioi_art_var_trial_all_line, echo = FALSE}
+p_ioi_art_var_trial_all <- ggplot(data = ioi_art_var_trial_all, 
+                                  aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+  geom_line(position = position_dodge(.2)) +
+  geom_point(position = position_dodge(.2)) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
+  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
+p_ioi_art_var_trial_all
 ```
@@
 
-#### Statistics (TBC)
+#### Statistics
+##### Normality check
+```{r normality4, echo = FALSE}
+# Normality check
+ioi_dyn_norm <- by(ioi_dyn$Mean, list(ioi_dyn$Condition), shapiro.test)
+print(ioi_dyn_norm)
+```
+
+```{r stats7, echo = FALSE}
+ioi_dyn_ttest <- t.test(ioi_dyn$Mean[ioi_dyn$Condition == "teaching"], ioi_dyn$Mean[ioi_dyn$Condition == "performing"], paired = TRUE)
+print(ioi_dyn_ttest)
+```
+
+##### Effect size
+```{r stats8, echo = FALSE}
+ioi_dyn_cohend <- cohen.d(ioi_dyn$Mean, ioi_dyn$Condition, paired = TRUE)
+print(ioi_dyn_cohend)
+```
 
@@
 ```{r ioi_dyn_ch, echo = FALSE}
-# Define phrase boundaries
-df_ioi_dyn$Change <- "No"
-df_ioi_dyn$Change[df_ioi_dyn$Interval == 5 | df_ioi_dyn$Interval == 8 | df_ioi_dyn$Interval == 17 | df_ioi_dyn$Interval == 20 | df_ioi_dyn$Interval == 39 | df_ioi_dyn$Interval == 47] <- "Yes"
+df_ioi_dyn$Boundary <- "No"
+df_ioi_dyn$Boundary[df_ioi_dyn$Subcomponent == "LtoS" | df_ioi_dyn$Subcomponent == "StoL" | df_ioi_dyn$Subcomponent == "FtoP" | df_ioi_dyn$Subcomponent == "PtoF"] <- "Yes"
 
 # For each individual
-ioi_dyn_ch <- aggregate(normIOI~SubNr*Condition*Skill*Change, data = df_ioi_dyn,
+ioi_dyn_ch <- aggregate(normIOI~SubNr*Condition*Skill*Boundary, data = df_ioi_dyn,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 # Change colnames
-colnames(ioi_dyn_ch) <- c("SubNr", "Condition", "Skill", "Change", "N", "Mean", "SD")
+colnames(ioi_dyn_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
 print(ioi_dyn_ch)
@@
 # Group mean
-ioi_dyn_ch_all <- aggregate(Mean~Condition*Skill*Change, data = ioi_dyn_ch,
+ioi_dyn_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_dyn_ch,
                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
@@
 # Change colnames
-colnames(ioi_dyn_ch_all) <- c("Condition", "Skill", "Change", "N", "Mean", "SD", "SEM")
+colnames(ioi_dyn_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
 print(ioi_dyn_ch_all)
@@
 ```{r ioi_dyn_ch_all_bar, echo = FALSE}
-p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_ch_all, aes(x = Change, y = Mean, fill = Condition)) +
+p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
 ```{r ioi_dyn_ch_all_box, echo = FALSE}
-p_ioi_dyn_ch_box <- ggplot(data = ioi_dyn_ch, aes(x = Change, y = Mean, fill = Condition)) +
+p_ioi_dyn_ch_box <- ggplot(data = ioi_dyn_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
   geom_boxplot() +
@@
 
-#### Statistics (TBC)
+#### Statistics
+##### Normality check
+```{r normality5, echo = FALSE}
+ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test)
+print(ioi_dyn_ch_norm)
+```
+  - Violate the normality assumption
+
+##### Two-way ANOVA (ez)
+```{r stats9, echo = FALSE}
+ioi_dyn_ch_aov <- ezANOVA(
+  data = df_ioi_dyn
+  , dv = .(IOI)
+  , wid = .(SubNr)
+  , within = .(Condition, Boundary)
+  , type = 3
+  , detailed = TRUE
+)
+ioi_dyn_ch_aov$ANOVA$F <- round(ioi_dyn_ch_aov$ANOVA$F, 4)
+ioi_dyn_ch_aov$ANOVA$p <- round(ioi_dyn_ch_aov$ANOVA$p, 4)
+print(ioi_dyn_ch_aov)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats10, echo = FALSE}
+ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
+print(summary(ioi_dyn_ch_aov_2))
+```
 
@@
 
+### 7. Variability all - individual{#dyn_ind_var_all}
+- Average variability for each participant.
+- IOI variability = SD/mean IOI (per trial)
+
+```{r ioi_dyn_var, echo = FALSE}
+# Create a data frame for variability (SD/)
+df_ioi_dyn_var <- data.frame()
+for (subnr in unique(df_ioi_dyn$SubNr)){
+  for (block in unique(df_ioi_dyn$BlockNr)){
+    cond = as.character(unique(df_ioi_dyn$Condition[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
+    skill = as.character(unique(df_ioi_dyn$Skill[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
+    for (trial in unique(df_ioi_dyn$TrialNr)){
+      df_current <- df_ioi_dyn %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
+      if (nrow(df_current) != 0){
+        df_ioi_dyn_var <- rbind(df_ioi_dyn_var, 
+                                data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI)))
+      }
+    }
+  }
+}
+colnames(df_ioi_dyn_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability")
+
+# Average variability
+# For each individual
+ioi_dyn_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_dyn_var,
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+ioi_dyn_var <- cbind(ioi_dyn_var[,1:3], as.data.frame(ioi_dyn_var[,4]))
+# Change colnames
+colnames(ioi_dyn_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
+print(ioi_dyn_var)
+```
+
+#### Plot
+##### Bar graph
+- Error bars show SD of the mean
+
+```{r ioi_dyn_var_bar, echo = FALSE}
+p_ioi_dyn_var <- ggplot(data = ioi_dyn_var, aes(x = SubNr, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
+                width = .2, position = position_dodge(1)) +
+  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics")
+p_ioi_dyn_var
+```
+
+### 8. Variability all - group{#dyn_group_var_all}
+- Average variability for each condition.
+
+```{r ioi_dyn_var_all, echo = FALSE}
+# Group mean
+ioi_dyn_var_all <- aggregate(Mean~Condition*Skill, data = ioi_dyn_var,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_dyn_var_all <- cbind(ioi_dyn_var_all[,1:2], as.data.frame(ioi_dyn_var_all[,3]))
+# Change colnames
+colnames(ioi_dyn_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
+print(ioi_dyn_var_all)
+```
+
+#### Plot
+##### Bar graph
+- Error bars show SEM of the mean
+
+```{r ioi_dyn_var_all_bar, echo = FALSE}
+p_ioi_dyn_var_all <- ggplot(data = ioi_dyn_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge()) +
+  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
+  theme(legend.position = "None")
+p_ioi_dyn_var_all
+```
+
+##### Box plot
+```{r ioi_dyn_var_all_box, echo = FALSE}
+p_ioi_dyn_var_box <- ggplot(data = ioi_dyn_var, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
+  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
+  theme(legend.position = "None")
+p_ioi_dyn_var_box
+```
+
+#### Statistics
+##### Normality check
+```{r normality6, echo = FALSE}
+ioi_dyn_var_norm <- by(ioi_dyn_var$Mean, list(ioi_dyn_var$Condition), shapiro.test)
+print(ioi_dyn_var_norm)
+```
+  - Violate the normality assumption
+
+```{r stats11, echo = FALSE}
+ioi_dyn_var_ttest <- t.test(ioi_dyn_var$Mean[ioi_dyn_var$Condition == "teaching"], ioi_dyn_var$Mean[ioi_dyn_var$Condition == "performing"], paired = TRUE)
+print(ioi_dyn_var_ttest)
+```
+
+```{r stats12, echo = FALSE}
+ioi_dyn_var_cohend <- cohen.d(ioi_dyn_var$Mean, ioi_dyn_var$Condition, paired = TRUE)
+print(ioi_dyn_var_cohend)
+```
+
+### 9. Variability trial-by-trial - individual{#dyn_ind_var_trial}
+- Average variability for each trial for each participant.
+
+```{r ioi_dyn_var_trial, echo = FALSE}
+# For each individual
+ioi_dyn_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn_var,
+                         FUN = function(x){c(N = length(x), mean = mean(x))})
+ioi_dyn_var_trial <- cbind(ioi_dyn_var_trial[,1:4], as.data.frame(ioi_dyn_var_trial[,5]))
+# Change colnames
+colnames(ioi_dyn_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean")
+print(ioi_dyn_var_trial)
+```
+
+#### Plot
+##### Line graph
+```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 8}
+p_ioi_dyn_var_trial <- ggplot(data = ioi_dyn_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  facet_wrap(SubNr ~ .) + 
+  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
+p_ioi_dyn_var_trial
+```
+
+### 10. Variability trial-by-trial - group{#dyn_group_var_trial}
+Average variability for each trial for each condition.
+
+```{r ioi_dyn_var_trial_all, echo = FALSE}
+# Group mean
+ioi_dyn_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_dyn_var_trial,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_dyn_var_trial_all <- cbind(ioi_dyn_var_trial_all[,1:3], as.data.frame(ioi_dyn_var_trial_all[,4]))
+# Change colnames
+colnames(ioi_dyn_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
+print(ioi_dyn_var_trial_all)
+```
+
+#### Plot
+##### Line graph
+- Error bars show SEM of the mean
+
+```{r ioi_dyn_var_trial_all_line, echo = FALSE}
+p_ioi_dyn_var_trial_all <- ggplot(data = ioi_dyn_var_trial_all, 
+                                  aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+  geom_line(position = position_dodge(.2)) +
+  geom_point(position = position_dodge(.2)) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
+  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
+p_ioi_dyn_var_trial_all
+```
+
 ## Sequence plot

### 2020-01-06 script/RNotebook/ioi_3.Rmd
add stats
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-01-06 script/RNotebook/ioi_1.Rmd
@@
 ##### Description: This is a summary of all analyses for a tempo (IOIs).
-  - data <- data_ioi_1.csv
+  - data <- data_ioi_3.csv
 
@@
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi_1.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_ioi <- read.csv("./trimmed/data_ioi_3.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 
@@
 ```{r ioi_art_ch, echo = FALSE}
-# Define phrase boundaries
-df_ioi_art$Boundary <- "No"
-df_ioi_art$Boundary[df_ioi_dyn$Subcomponent == "LtoS" | df_ioi_dyn$Subcomponent == "StoL" | df_ioi_dyn$Subcomponent == "FtoP" | df_ioi_art$Subcomponent == "PtoF"] <- "Yes"
-
 # For each individual
@@
 ```{r ioi_dyn_ch, echo = FALSE}
-df_ioi_dyn$Boundary <- "No"
-df_ioi_dyn$Boundary[df_ioi_dyn$Subcomponent == "LtoS" | df_ioi_dyn$Subcomponent == "StoL" | df_ioi_dyn$Subcomponent == "FtoP" | df_ioi_dyn$Subcomponent == "PtoF"] <- "Yes"
-
 # For each individual

### 2020-01-06 script/RNotebook/ioi_1.Rmd
remove a wrong comment
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 df_ioi_art$Boundary <- "No"
-df_ioi_art$Boundary[df_ioi_dyn$Subcomponent == "LtoS" | df_ioi_dyn$Subcomponent == "StoL" | df_ioi_dyn$Subcomponent == "FtoP" | df_ioi_art$Subcomponent == "PtoF"] <- "Yes"
+df_ioi_art$Boundary[df_ioi_art$Subcomponent == "LtoS" | df_ioi_art$Subcomponent == "StoL" |
+                      df_ioi_art$Subcomponent == "FtoP" | df_ioi_art$Subcomponent == "PtoF"] <- "Yes"
 
@@
 df_ioi_dyn$Boundary <- "No"
-df_ioi_dyn$Boundary[df_ioi_dyn$Subcomponent == "LtoS" | df_ioi_dyn$Subcomponent == "StoL" | df_ioi_dyn$Subcomponent == "FtoP" | df_ioi_dyn$Subcomponent == "PtoF"] <- "Yes"
+df_ioi_dyn$Boundary[df_ioi_dyn$Subcomponent == "LtoS" | df_ioi_dyn$Subcomponent == "StoL" |
+                      df_ioi_dyn$Subcomponent == "FtoP" | df_ioi_dyn$Subcomponent == "PtoF"] <- "Yes"
 
@@
 ```
-  - Violate the normality assumption
 

### 2020-01-06 script/RNotebook/articulation.Rmd
add stats
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 df_kot <- read.csv("./trimmed/data_kot.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_kor <- read.csv("./trimmed/data_kor.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
-df_ioi$SubNr <- as.factor(df_ioi$SubNr)
-
-# SubNr as a factor
-df_ioi$SubNr <- as.factor(df_ioi$SubNr)
 df_kot$SubNr <- as.factor(df_kot$SubNr)
-
-# For each individual, each block, each trial
-ioi <- aggregate(IOI~SubNr*BlockNr*TrialNr*Condition*Skill, data = df_ioi,
-                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
-ioi <- cbind(ioi[,1:5], as.data.frame(ioi[,6]))
-# Change colnames
-colnames(ioi) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "N", "Mean", "SD")
-
-# Calculate KOR for each interval
-
-df_kor <- data.frame()
-for (subnr in unique(df_kot$SubNr)){
-  for (block in unique(df_kot$BlockNr[df_kot$SubNr == subnr])){
-    for (trial in unique(df_kot$TrialNr[df_kot$SubNr == subnr & df_kot$BlockNr == block])){
-      df_current <- df_kot %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
-      df_current$KOR <- df_current$KOT/ioi$Mean[ioi$SubNr == subnr & ioi$BlockNr == block & ioi$TrialNr == trial]
-      df_kor <- rbind(df_kor, df_current)
-    }
-  }
-}
+df_kor$SubNr <- as.factor(df_kor$SubNr)
 
@@
 #### Statistics
-##### TBC
+##### Normality check
+```{r normality1, echo = FALSE}
+# Reduce unused factors
+kot_art$Subcomponent <- factor(kot_art$Subcomponent)
+kot_art_norm <- by(kot_art$Mean, list(kot_art$Condition, kot_art$Subcomponent), shapiro.test)
+print(kot_art_norm)
+```
+
+##### Two-way ANOVA (ez)
+```{r stats1, echo = FALSE}
+kot_art_aov <- ezANOVA(
+  data = subset(df_kot_art, df_kot_art$Subcomponent == "Legato" | df_kot_art$Subcomponent == "Staccato")
+  , dv = .(KOT)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kot_art_aov$ANOVA$F <- round(kot_art_aov$ANOVA$F, 4)
+kot_art_aov$ANOVA$p <- round(kot_art_aov$ANOVA$p, 4)
+print(kot_art_aov$ANOVA)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats2, echo = FALSE}
+kot_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
+print(summary(kot_art_aov_2))
+```
+
+##### Post-hoc (direct comparison)
+```{r post-hoc1, echo = FALSE}
+kot_ttest_legato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"], paired = TRUE)
+kot_ttest_staccato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"], paired = TRUE)
+# Effect size
+kot_ttest_legato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Legato"], kot_art$Condition[kot_art$Subcomponent == "Legato"], paired = TRUE)
+kot_ttest_staccato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Staccato"], kot_art$Condition[kot_art$Subcomponent == "Staccato"], paired = TRUE)
+
+print(kot_ttest_legato)
+print(kot_ttest_staccato)
+print(kot_ttest_legato_cohend)
+print(kot_ttest_staccato_cohend)
+```
+
+##### Post-hoc (pairwise t test)
+```{r post-hoc2, echo = FALSE}
+# Create a group label for a post-hoc test
+kot_art$Group[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"] <- "Performing-Legato" 
+kot_art$Group[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
+kot_art$Group[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"] <- "Teaching-Legato"
+kot_art$Group[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
+# pairwise comparison
+kot_art_pairwise <- pairwise.t.test(kot_art$Mean, kot_art$Group, paired = TRUE, p.adjust.method = "BH")
+kot_art_pairwise$p.value <- round(kot_art_pairwise$p.value, 4)
+print(kot_art_pairwise)
+```
 
@@
 # For each individual
-kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
+kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_art, df_kor_art$Subcomponent != "LtoS" & df_kor_art$Subcomponent != "StoL"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 #### Statistics
-##### TBC
+##### Normality check
+```{r normality2, echo = FALSE}
+# Reduce unused factors
+kor_art$Subcomponent <- factor(kor_art$Subcomponent)
+kor_art_norm <- by(kor_art$Mean, list(kor_art$Condition, kor_art$Subcomponent), shapiro.test)
+print(kor_art_norm)
+```
+
+##### Two-way ANOVA (ez)
+```{r stats3, echo = FALSE}
+kor_art_aov <- ezANOVA(
+  data = subset(df_kor_art, df_kor_art$Subcomponent == "Legato" | df_kor_art$Subcomponent == "Staccato")
+  , dv = .(KOR)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kor_art_aov$ANOVA$F <- round(kor_art_aov$ANOVA$F, 4)
+kor_art_aov$ANOVA$p <- round(kor_art_aov$ANOVA$p, 4)
+print(kor_art_aov$ANOVA)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats4, echo = FALSE}
+kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
+print(summary(kor_art_aov_2))
+```
+
+```{r post-hoc3, echo = FALSE}
+kor_ttest_legato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"], paired = TRUE)
+kor_ttest_staccato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"], paired = TRUE)
+# Effect size
+kor_ttest_legato_cohend <- cohen.d(kor_art$Mean[kor_art$Subcomponent == "Legato"], kor_art$Condition[kor_art$Subcomponent == "Legato"], paired = TRUE)
+kor_ttest_staccato_cohend <- cohen.d(kor_art$Mean[kor_art$Subcomponent == "Staccato"], kor_art$Condition[kor_art$Subcomponent == "Staccato"], paired = TRUE)
+
+print(kor_ttest_legato)
+print(kor_ttest_staccato)
+print(kor_ttest_legato_cohend)
+print(kor_ttest_staccato_cohend)
+```
+
+##### Post-hoc (pairwise t test)
+```{r post-hoc4, echo = FALSE}
+# Create a group label for a post-hoc test
+kor_art$Group[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"] <- "Performing-Legato" 
+kor_art$Group[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
+kor_art$Group[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"] <- "Teaching-Legato"
+kor_art$Group[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
+# pairwise comparison
+kor_art_pairwise <- pairwise.t.test(kor_art$Mean, kor_art$Group, paired = TRUE, p.adjust.method = "BH")
+kor_art_pairwise$p.value <- round(kor_art_pairwise$p.value, 4)
+print(kor_art_pairwise)
+```
 
@@
 # For each individual
-kor_art_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57),
+kor_art_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_art, df_kor_art$Subcomponent != "LtoS" & df_kor_art$Subcomponent != "StoL"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 #### Statistics
-##### TBC
+##### Normality check
+```{r normality3, echo = FALSE}
+# Reduce unused factors
+kot_dyn$Subcomponent <- factor(kot_dyn$Subcomponent)
+kot_dyn_norm <- by(kot_dyn$Mean, list(kot_dyn$Condition, kot_dyn$Subcomponent), shapiro.test)
+print(kot_dyn_norm)
+```
+  - Violate the normality assumption
+
+##### Two-way ANOVA (ez)
+```{r stats5, echo = FALSE}
+kot_dyn_aov <- ezANOVA(
+  data = subset(df_kot_dyn, df_kot_dyn$Subcomponent == "Forte" | df_kot_dyn$Subcomponent == "Piano")
+  , dv = .(KOT)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kot_dyn_aov$ANOVA$F <- round(kot_dyn_aov$ANOVA$F, 4)
+kot_dyn_aov$ANOVA$p <- round(kot_dyn_aov$ANOVA$p, 4)
+print(kot_dyn_aov$ANOVA)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats6, echo = FALSE}
+kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
+print(summary(kot_dyn_aov_2))
+```
 
@@
 # For each individual
-kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
+kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_dyn, df_kor_dyn$Subcomponent != "FtoP" & df_kor_dyn$Subcomponent != "PtoF"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 #### Statistics
-##### TBC
+##### Normality check
+```{r normality4, echo = FALSE}
+# Reduce unused factors
+kor_dyn$Subcomponent <- factor(kor_dyn$Subcomponent)
+kor_dyn_norm <- by(kor_dyn$Mean, list(kor_dyn$Condition, kor_dyn$Subcomponent), shapiro.test)
+print(kor_dyn_norm)
+```
+  - Violate the normality assumtion
+
+##### Two-way ANOVA (ez)
+```{r stats7, echo = FALSE}
+kor_dyn_aov <- ezANOVA(
+  data = subset(df_kor_dyn, df_kor_dyn$Subcomponent == "Forte" | df_kor_dyn$Subcomponent == "Piano")
+  , dv = .(KOR)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kor_dyn_aov$ANOVA$F <- round(kor_dyn_aov$ANOVA$F, 4)
+kor_dyn_aov$ANOVA$p <- round(kor_dyn_aov$ANOVA$p, 4)
+print(kor_dyn_aov$ANOVA)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats8, echo = FALSE}
+kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
+print(summary(kor_dyn_aov_2))
+```
 
@@
 # For each individual
-kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
+kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_dyn, df_kor_dyn$Subcomponent != "FtoP" & df_kor_dyn$Subcomponent != "PtoF"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 ```
+
+## Sequence plot
+#### Articulation{#sequence_art}
+```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
+# For each individual
+kot_art_seq <- aggregate(KOT~SubNr*Condition*Skill*Interval, data = df_kot_art,
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+kot_art_seq <- cbind(kot_art_seq[,1:4], as.data.frame(kot_art_seq[,5]))
+# Change colnames
+colnames(kot_art_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+kot_art_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = kot_art_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_art_seq_stats <- cbind(kot_art_seq_stats[,1:3], as.data.frame(kot_art_seq_stats[,4]))
+# Change colnames for each interval
+colnames(kot_art_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_kot_art_seq <- ggplot(data = kot_art_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+
+# For each individual
+kor_art_seq <- aggregate(KOR~SubNr*Condition*Skill*Interval, data = df_kor_art,
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+kor_art_seq <- cbind(kor_art_seq[,1:4], as.data.frame(kor_art_seq[,5]))
+# Change colnames
+colnames(kor_art_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+kor_art_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = kor_art_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kor_art_seq_stats <- cbind(kor_art_seq_stats[,1:3], as.data.frame(kor_art_seq_stats[,4]))
+# Change colnames for each interval
+colnames(kor_art_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_kor_art_seq <- ggplot(data = kor_art_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "KOR") + scale_x_continuous(breaks=seq(1,66,1))
+
+p_kot_art_seq
+p_kor_art_seq
+```
+
+#### Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 15, fig.height = 5, echo = FALSE}
+# For each individual
+kot_dyn_seq <- aggregate(KOT~SubNr*Condition*Skill*Interval, data = df_kot_dyn,
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+kot_dyn_seq <- cbind(kot_dyn_seq[,1:4], as.data.frame(kot_dyn_seq[,5]))
+# Change colnames
+colnames(kot_dyn_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+kot_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = kot_dyn_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_dyn_seq_stats <- cbind(kot_dyn_seq_stats[,1:3], as.data.frame(kot_dyn_seq_stats[,4]))
+# Change colnames for each interval
+colnames(kot_dyn_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_kot_dyn_seq <- ggplot(data = kot_dyn_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+
+# For each individual
+kor_dyn_seq <- aggregate(KOR~SubNr*Condition*Skill*Interval, data = df_kor_dyn,
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+kor_dyn_seq <- cbind(kor_dyn_seq[,1:4], as.data.frame(kor_dyn_seq[,5]))
+# Change colnames
+colnames(kor_dyn_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+kor_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = kor_dyn_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kor_dyn_seq_stats <- cbind(kor_dyn_seq_stats[,1:3], as.data.frame(kor_dyn_seq_stats[,4]))
+# Change colnames for each interval
+colnames(kor_dyn_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_kor_dyn_seq <- ggplot(data = kor_dyn_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "KOR") + scale_x_continuous(breaks=seq(1,66,1))
+
+p_kot_dyn_seq
+p_kor_dyn_seq
+```

### 2020-01-06 script/RNotebook/dynamics.Rmd
tweak comments/description
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
add stats
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2019-11-28 script/RNotebook/dynamics.Rmd
@@
 df_vel <- read.csv("./trimmed/data_vel.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-df_vel_acc <- read.csv("./trimmed/data_vel_acc.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_vel_diff <- read.csv("./trimmed/data_vel_diff.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 
@@
 df_vel$SubNr <- as.factor(df_vel$SubNr)
-df_vel_acc$SubNr <- as.factor(df_vel_acc$SubNr)
+df_vel_diff$SubNr <- as.factor(df_vel_diff$SubNr)
 
@@
 df_vel_art <- df_vel %>% dplyr::filter(Skill == "articulation")
-df_vel_acc_art <- df_vel_acc %>% dplyr::filter(Skill == "articulation")
+df_vel_diff_art <- df_vel_diff %>% dplyr::filter(Skill == "articulation")
 
@@
 df_vel_dyn <- df_vel %>% dplyr::filter(Skill == "dynamics")
-df_vel_acc_dyn <- df_vel_acc %>% dplyr::filter(Skill == "dynamics")
+df_vel_diff_dyn <- df_vel_diff %>% dplyr::filter(Skill == "dynamics")
 ```
@@
 
-#### Statistics (TBC)
+#### Statistics
+##### Normality check
+```{r normality1, echo = FALSE}
+# Reduce unused factors
+vel_dyn$Subcomponent <- factor(vel_dyn$Subcomponent)
+vel_dyn_norm <- by(vel_dyn$Mean, list(vel_dyn$Condition, vel_dyn$Subcomponent), shapiro.test)
+print(vel_dyn_norm)
+```
+  - Violate the normality assumption
+
+##### Two-way ANOVA (ez)
+```{r stats1, echo = FALSE}
+vel_dyn_aov <- ezANOVA(
+  data = df_vel_dyn
+  , dv = .(Velocity)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_dyn_aov$ANOVA$F <- round(vel_dyn_aov$ANOVA$F, 4)
+vel_dyn_aov$ANOVA$p <- round(vel_dyn_aov$ANOVA$p, 4)
+print(vel_dyn_aov$ANOVA)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats2, echo = FALSE}
+vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
+print(summary(vel_dyn_aov_2))
+```
+
+##### Post-hoc (direct comparison)
+```{r post-hoc1, echo = FALSE}
+vel_ttest_forte <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"], paired = TRUE)
+vel_ttest_piano <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"], paired = TRUE)
+# Effect size
+vel_ttest_forte_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Forte"], vel_dyn$Condition[vel_dyn$Subcomponent == "Forte"], paired = TRUE)
+vel_ttest_piano_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Piano"], vel_dyn$Condition[vel_dyn$Subcomponent == "Piano"], paired = TRUE)
+
+print(vel_ttest_forte)
+print(vel_ttest_piano)
+print(vel_ttest_forte_cohend)
+print(vel_ttest_piano_cohend)
+```
+
+##### Post-hoc (pairwise t test)
+```{r post-hoc2, echo = FALSE}
+# Create a group label for a post-hoc test
+vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"] <- "Performing-Forte" 
+vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"] <- "Performing-Piano"
+vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"] <- "Teaching-Forte"
+vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"] <- "Teaching-Piano"
+# pairwise comparison
+vel_dyn_pairwise <- pairwise.t.test(vel_dyn$Mean, vel_dyn$Group, paired = TRUE, p.adjust.method = "BH")
+vel_dyn_pairwise$p.value <- round(vel_dyn_pairwise$p.value, 4)
+print(vel_dyn_pairwise)
+```
 
@@
 ##### Line graph
-```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
+```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
 p_vel_dyn_trial_forte <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r vel_dyn_trial_line_piano, echo = FALSE, fig.height = 8}
+```{r vel_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
 p_vel_dyn_trial_piano <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ## Key Velocity Difference (KV-Diff) - Dynamics
-### 1. All - individual{#vel_acc_dyn_ind_all}
+### 1. All - individual{#vel_diff_dyn_ind_all}
 - Average KV-Diff for each participant.
 
-```{r vel_acc_dyn, echo = FALSE}
+```{r vel_diff_dyn, echo = FALSE}
 # For each individual
-vel_acc_dyn <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
+vel_diff_dyn <- aggregate(Diff~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_diff_dyn, df_vel_diff_dyn$Subcomponent == "FtoP" | df_vel_diff_dyn$Subcomponent == "PtoF"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_acc_dyn <- cbind(vel_acc_dyn[,1:4], as.data.frame(vel_acc_dyn[,5]))
+vel_diff_dyn <- cbind(vel_diff_dyn[,1:4], as.data.frame(vel_diff_dyn[,5]))
 # Change colnames
-colnames(vel_acc_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
-print(vel_acc_dyn)
+colnames(vel_diff_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
+print(vel_diff_dyn)
 ```
@@
 
-```{r vel_acc_dyn_bar,  echo = FALSE}
-p_vel_acc_dyn <- ggplot(data = vel_acc_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
+```{r vel_diff_dyn_bar,  echo = FALSE}
+p_vel_diff_dyn <- ggplot(data = vel_diff_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
   labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
-p_vel_acc_dyn
+p_vel_diff_dyn
 ```
 
-### 2. All - group{#vel_acc_dyn_group_all}
+### 2. All - group{#vel_diff_dyn_group_all}
 - Average KV-Diff for each condition.
@@
 
-```{r vel_acc_dyn_all, echo = FALSE}
+```{r vel_diff_dyn_all, echo = FALSE}
 # Group mean
-vel_acc_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_acc_dyn,
+vel_diff_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_diff_dyn,
                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_acc_dyn_all <- cbind(vel_acc_dyn_all[,1:3], as.data.frame(vel_acc_dyn_all[,4]))
+vel_diff_dyn_all <- cbind(vel_diff_dyn_all[,1:3], as.data.frame(vel_diff_dyn_all[,4]))
 # Change colnames
-colnames(vel_acc_dyn_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
-print(vel_acc_dyn_all)
+colnames(vel_diff_dyn_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
+print(vel_diff_dyn_all)
 ```
 
-```{r vel_acc_dyn_all_bar,  echo = FALSE}
-p_vel_acc_dyn_all <- ggplot(data = vel_acc_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+```{r vel_diff_dyn_all_bar,  echo = FALSE}
+p_vel_diff_dyn_all <- ggplot(data = vel_diff_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
   labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
-p_vel_acc_dyn_all
+p_vel_diff_dyn_all
 ```
@@
 
-```{r vel_acc_dyn_all_box, echo = FALSE}
-p_vel_acc_dyn_all_box <- ggplot(data = vel_acc_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+```{r vel_diff_dyn_all_box, echo = FALSE}
+p_vel_diff_dyn_all_box <- ggplot(data = vel_diff_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
   geom_boxplot() +
@@
   labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
-p_vel_acc_dyn_all_box
+p_vel_diff_dyn_all_box
 ```
 
-#### Statistics (TBC)
+#### Statistics
+##### Normality check
+```{r normality2, echo = FALSE}
+# Reduce unused factors
+vel_diff_dyn$Subcomponent <- factor(vel_diff_dyn$Subcomponent)
+vel_diff_dyn_norm <- by(vel_diff_dyn$Mean, list(vel_diff_dyn$Condition, vel_diff_dyn$Subcomponent), shapiro.test)
+print(vel_diff_dyn_norm)
+```
 
-### 3. Trial by trial - individual{#vel_acc_dyn_ind_trial}
+##### Two-way ANOVA (ez)
+```{r stats3, echo = FALSE}
+vel_diff_dyn_aov <- ezANOVA(
+  data = subset(df_vel_diff_dyn, df_vel_diff_dyn$Subcomponent == "FtoP" | df_vel_diff_dyn$Subcomponent == "PtoF")
+  , dv = .(Diff)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_diff_dyn_aov$ANOVA$F <- round(vel_diff_dyn_aov$ANOVA$F, 4)
+vel_diff_dyn_aov$ANOVA$p <- round(vel_diff_dyn_aov$ANOVA$p, 4)
+print(vel_diff_dyn_aov$ANOVA)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats4, echo = FALSE}
+vel_diff_dyn_aov_2 <- with(vel_diff_dyn,
+                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
+print(summary(vel_diff_dyn_aov_2))
+```
+
+##### Post-hoc (direct comparison)
+```{r post-hoc3, echo = FALSE}
+vel_diff_ttest_ftop <- t.test(vel_diff_dyn$Mean[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "FtoP"], vel_diff_dyn$Mean[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "FtoP"], paired = TRUE)
+vel_diff_ttest_ptof <- t.test(vel_diff_dyn$Mean[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "PtoF"], vel_diff_dyn$Mean[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "PtoF"], paired = TRUE)
+# Effect size
+vel_diff_ttest_ftop_cohend <- cohen.d(vel_diff_dyn$Mean[vel_diff_dyn$Subcomponent == "FtoP"], vel_diff_dyn$Condition[vel_diff_dyn$Subcomponent == "FtoP"], paired = TRUE)
+vel_diff_ttest_ptof_cohend <- cohen.d(vel_diff_dyn$Mean[vel_diff_dyn$Subcomponent == "PtoF"], vel_diff_dyn$Condition[vel_diff_dyn$Subcomponent == "PtoF"], paired = TRUE)
+
+print(vel_diff_ttest_ftop)
+print(vel_diff_ttest_ptof)
+print(vel_diff_ttest_ftop_cohend)
+print(vel_diff_ttest_ptof_cohend)
+```
+
+##### Post-hoc (pairwise t test)
+```{r post-hoc4, echo = FALSE}
+# Create a group label for a post-hoc test
+vel_diff_dyn$Group[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "FtoP"] <- "Performing-FtoP" 
+vel_diff_dyn$Group[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "PtoF"] <- "Performing-PtoF"
+vel_diff_dyn$Group[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "FtoP"] <- "Teaching-FtoP"
+vel_diff_dyn$Group[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "PtoF"] <- "Teaching-PtoF"
+# pairwise comparison
+vel_diff_dyn_pairwise <- pairwise.t.test(vel_diff_dyn$Mean, vel_diff_dyn$Group, paired = TRUE, p.adjust.method = "BH")
+vel_diff_dyn_pairwise$p.value <- round(vel_diff_dyn_pairwise$p.value, 4)
+print(vel_diff_dyn_pairwise)
+```
+
+### 3. Trial by trial - individual{#vel_diff_dyn_ind_trial}
 - Average KV-Diff for each trial for each participant.
 
-```{r vel_acc_dyn_trial, echo = FALSE}
+```{r vel_diff_dyn_trial, echo = FALSE}
 # For each individual
-vel_acc_dyn_trial <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
+vel_diff_dyn_trial <- aggregate(Diff~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_diff_dyn, df_vel_diff_dyn$Subcomponent == "FtoP" | df_vel_diff_dyn$Subcomponent == "PtoF"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_acc_dyn_trial <- cbind(vel_acc_dyn_trial[,1:5], as.data.frame(vel_acc_dyn_trial[,6]))
+vel_diff_dyn_trial <- cbind(vel_diff_dyn_trial[,1:5], as.data.frame(vel_diff_dyn_trial[,6]))
 # Change colnames
-colnames(vel_acc_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
-print(vel_acc_dyn_trial)
+colnames(vel_diff_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
+print(vel_diff_dyn_trial)
 ```
@@
 ##### Line graph
-```{r vel_acc_dyn_trial_line_ftop, echo = FALSE, fig.height = 8}
-p_vel_acc_dyn_trial_ftop <- ggplot(data = subset(vel_acc_dyn_trial, vel_acc_dyn_trial$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+```{r vel_diff_dyn_trial_line_ftop, echo = FALSE, fig.height = 4}
+p_vel_diff_dyn_trial_ftop <- ggplot(data = subset(vel_diff_dyn_trial, vel_diff_dyn_trial$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
   geom_line() +
@@
   labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
-p_vel_acc_dyn_trial_ftop
+p_vel_diff_dyn_trial_ftop
 ```
 
-```{r vel_acc_dyn_trial_line_ptof, echo = FALSE, fig.height = 8}
-p_vel_acc_dyn_trial_ptof <- ggplot(data = subset(vel_acc_dyn_trial, vel_acc_dyn_trial$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+```{r vel_diff_dyn_trial_line_ptof, echo = FALSE, fig.height = 4}
+p_vel_diff_dyn_trial_ptof <- ggplot(data = subset(vel_diff_dyn_trial, vel_diff_dyn_trial$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
   geom_line() +
@@
   labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
-p_vel_acc_dyn_trial_ptof
+p_vel_diff_dyn_trial_ptof
 ```
 
-### 4. Trial by trial - group{#vel_acc_dyn_group_trial}
+### 4. Trial by trial - group{#vel_diff_dyn_group_trial}
 - Average KV-Diff for each trial for each condition.
 
-```{r vel_acc_dyn_trial_all, echo = FALSE}
+```{r vel_diff_dyn_trial_all, echo = FALSE}
 # Group mean
-vel_acc_dyn_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_acc_dyn_trial,
+vel_diff_dyn_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_diff_dyn_trial,
                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_acc_dyn_trial_all <- cbind(vel_acc_dyn_trial_all[,1:4], as.data.frame(vel_acc_dyn_trial_all[,5]))
+vel_diff_dyn_trial_all <- cbind(vel_diff_dyn_trial_all[,1:4], as.data.frame(vel_diff_dyn_trial_all[,5]))
 # Change colnames
-colnames(vel_acc_dyn_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
-print(vel_acc_dyn_trial_all)
+colnames(vel_diff_dyn_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
+print(vel_diff_dyn_trial_all)
 ```
@@
 
-```{r vel_acc_dyn_trial_all_line_ftop, echo = FALSE}
-p_vel_acc_dyn_trial_all_ftop <- ggplot(data = subset(vel_acc_dyn_trial_all, vel_acc_dyn_trial_all$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+```{r vel_diff_dyn_trial_all_line_ftop, echo = FALSE}
+p_vel_diff_dyn_trial_all_ftop <- ggplot(data = subset(vel_diff_dyn_trial_all, vel_diff_dyn_trial_all$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
   geom_line(position = position_dodge(.2)) +
@@
   labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
-p_vel_acc_dyn_trial_all_ftop
+p_vel_diff_dyn_trial_all_ftop
 ```
 
-```{r vel_acc_dyn_trial_all_line_ptof, echo = FALSE}
-p_vel_acc_dyn_trial_all_ptof <- ggplot(data = subset(vel_acc_dyn_trial_all, vel_acc_dyn_trial_all$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+```{r vel_diff_dyn_trial_all_line_ptof, echo = FALSE}
+p_vel_diff_dyn_trial_all_ptof <- ggplot(data = subset(vel_diff_dyn_trial_all, vel_diff_dyn_trial_all$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
   geom_line(position = position_dodge(.2)) +
@@
   labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
-p_vel_acc_dyn_trial_all_ptof
+p_vel_diff_dyn_trial_all_ptof
 ```
@@
 
-#### Statistics (TBC)
+#### Statistics
+##### Normality check
+```{r normality3, echo = FALSE}
+# Reduce unused factors
+vel_art$Subcomponent <- factor(vel_art$Subcomponent)
+vel_art_norm <- by(vel_art$Mean, list(vel_art$Condition, vel_art$Subcomponent), shapiro.test)
+print(vel_art_norm)
+```
+
+##### Two-way ANOVA (ez)
+```{r stats5, echo = FALSE}
+vel_art_aov <- ezANOVA(
+  data = df_vel_art
+  , dv = .(Velocity)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_art_aov$ANOVA$F <- round(vel_art_aov$ANOVA$F, 4)
+vel_art_aov$ANOVA$p <- round(vel_art_aov$ANOVA$p, 4)
+print(vel_art_aov$ANOVA)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats6, echo = FALSE}
+vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
+print(summary(vel_art_aov_2))
+```
 
@@
 ##### Line graph
-```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 8}
+```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 4}
 p_vel_art_trial_legato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r vel_art_trial_line_piano, echo = FALSE, fig.height = 8}
+```{r vel_art_trial_line_piano, echo = FALSE, fig.height = 4}
 p_vel_art_trial_staccato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ## Key Velocity Difference (KV-Diff) - Articulation
-### 1. All - individual{#vel_acc_art_ind_all}
+### 1. All - individual{#vel_diff_art_ind_all}
 - Average KV-Diff for each participant.
 
-```{r vel_acc_art, echo = FALSE}
+```{r vel_diff_art, echo = FALSE}
 # For each individual
-vel_acc_art <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
+vel_diff_art <- aggregate(Diff~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_diff_art, df_vel_diff_art$Subcomponent == "LtoS" | df_vel_diff_art$Subcomponent == "StoL"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_acc_art <- cbind(vel_acc_art[,1:4], as.data.frame(vel_acc_art[,5]))
+vel_diff_art <- cbind(vel_diff_art[,1:4], as.data.frame(vel_diff_art[,5]))
 # Change colnames
-colnames(vel_acc_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
-print(vel_acc_art)
+colnames(vel_diff_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
+print(vel_diff_art)
 ```
@@
 
-```{r vel_acc_art_bar,  echo = FALSE}
-p_vel_acc_art <- ggplot(data = vel_acc_art, aes(x = SubNr, y = Mean, fill = Condition)) +
+```{r vel_diff_art_bar,  echo = FALSE}
+p_vel_diff_art <- ggplot(data = vel_diff_art, aes(x = SubNr, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
   labs(y = "Difference", subtitle = "KV-Diff: Articulaion")
-p_vel_acc_art
+p_vel_diff_art
 ```
 
-### 2. All - group{#vel_acc_art_group_all}
+### 2. All - group{#vel_diff_art_group_all}
 - Average KV-Diff for each condition.
 
-```{r vel_acc_art_all, echo = FALSE}
+```{r vel_diff_art_all, echo = FALSE}
 # Group mean
-vel_acc_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_acc_art,
+vel_diff_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_diff_art,
                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_acc_art_all <- cbind(vel_acc_art_all[,1:3], as.data.frame(vel_acc_art_all[,4]))
+vel_diff_art_all <- cbind(vel_diff_art_all[,1:3], as.data.frame(vel_diff_art_all[,4]))
 # Change colnames
-colnames(vel_acc_art_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
-print(vel_acc_art_all)
+colnames(vel_diff_art_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
+print(vel_diff_art_all)
 ```
@@
 
-```{r vel_acc_art_all_bar,  echo = FALSE}
-p_vel_acc_art_all <- ggplot(data = vel_acc_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+```{r vel_diff_art_all_bar,  echo = FALSE}
+p_vel_diff_art_all <- ggplot(data = vel_diff_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
   labs(y = "Difference", subtitle = "KV-Diff: Articulation")
-p_vel_acc_art_all
+p_vel_diff_art_all
 ```
@@
 
-```{r vel_acc_art_all_box, echo = FALSE}
-p_vel_acc_art_all_box <- ggplot(data = vel_acc_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+```{r vel_diff_art_all_box, echo = FALSE}
+p_vel_diff_art_all_box <- ggplot(data = vel_diff_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
   geom_boxplot() +
@@
   labs(y = "Difference", subtitle = "KV-Diff: Articulation")
-p_vel_acc_art_all_box
+p_vel_diff_art_all_box
 ```
 
-##### Statistics
+#### Statistics
+##### Normality check
+```{r normality4, echo = FALSE}
+# Reduce unused factors
+vel_diff_art$Subcomponent <- factor(vel_diff_art$Subcomponent)
+vel_diff_art_norm <- by(vel_diff_art$Mean, list(vel_diff_art$Condition, vel_diff_art$Subcomponent), shapiro.test)
+print(vel_diff_art_norm)
+```
+
+##### Two-way ANOVA (ez)
+```{r stats7, echo = FALSE}
+vel_diff_art_aov <- ezANOVA(
+  data = subset(df_vel_diff_art, df_vel_diff_art$Subcomponent == "LtoS" | df_vel_diff_art$Subcomponent == "StoL")
+  , dv = .(Diff)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_diff_art_aov$ANOVA$F <- round(vel_diff_art_aov$ANOVA$F, 4)
+vel_diff_art_aov$ANOVA$p <- round(vel_diff_art_aov$ANOVA$p, 4)
+print(vel_diff_art_aov$ANOVA)
+```
+
+##### Two-way ANOVA (aov)
+```{r stats8, echo = FALSE}
+vel_diff_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_diff_art)
+print(summary(vel_diff_art_aov_2))
+```
 
@@
 
-```{r vel_acc_art_trial, echo = FALSE}
+```{r vel_diff_art_trial, echo = FALSE}
 # For each individual
-vel_acc_art_trial <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
+vel_diff_art_trial <- aggregate(Diff~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_diff_art, df_vel_diff_art$Subcomponent == "LtoS" | df_vel_diff_art$Subcomponent == "StoL"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_acc_art_trial <- cbind(vel_acc_art_trial[,1:5], as.data.frame(vel_acc_art_trial[,6]))
+vel_diff_art_trial <- cbind(vel_diff_art_trial[,1:5], as.data.frame(vel_diff_art_trial[,6]))
 # Change colnames
-colnames(vel_acc_art_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
-print(vel_acc_art_trial)
+colnames(vel_diff_art_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
+print(vel_diff_art_trial)
 ```
@@
 ##### Line graph
-```{r vel_acc_art_trial_line_ltos, echo = FALSE, fig.height = 8}
-p_vel_acc_art_trial_ltos <- ggplot(data = subset(vel_acc_art_trial, vel_acc_art_trial$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+```{r vel_diff_art_trial_line_ltos, echo = FALSE, fig.height = 4}
+p_vel_diff_art_trial_ltos <- ggplot(data = subset(vel_diff_art_trial, vel_diff_art_trial$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
   geom_line() +
@@
   labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
-p_vel_acc_art_trial_ltos
+p_vel_diff_art_trial_ltos
 ```
 
-```{r vel_acc_art_trial_line_stol, echo = FALSE, fig.height = 8}
-p_vel_acc_art_trial_stol <- ggplot(data = subset(vel_acc_art_trial, vel_acc_art_trial$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+```{r vel_diff_art_trial_line_stol, echo = FALSE, fig.height = 4}
+p_vel_diff_art_trial_stol <- ggplot(data = subset(vel_diff_art_trial, vel_diff_art_trial$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
   geom_line() +
@@
   labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Staccato to Legato") + scale_x_continuous(breaks=seq(1,8,1))
-p_vel_acc_art_trial_stol
+p_vel_diff_art_trial_stol
 ```
 
-### 4. Trial by trial - group{#vel_acc_art_group_trial}
+### 4. Trial by trial - group{#vel_diff_art_group_trial}
 - Average KV-Diff for each trial for each condition.
 
-```{r vel_acc_art_trial_all, echo = FALSE}
+```{r vel_diff_art_trial_all, echo = FALSE}
 # Group mean
-vel_acc_art_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_acc_art_trial,
+vel_diff_art_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_diff_art_trial,
                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_acc_art_trial_all <- cbind(vel_acc_art_trial_all[,1:4], as.data.frame(vel_acc_art_trial_all[,5]))
+vel_diff_art_trial_all <- cbind(vel_diff_art_trial_all[,1:4], as.data.frame(vel_diff_art_trial_all[,5]))
 # Change colnames
-colnames(vel_acc_art_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
-print(vel_acc_art_trial_all)
+colnames(vel_diff_art_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
+print(vel_diff_art_trial_all)
 ```
@@
 
-```{r vel_acc_art_trial_all_line_ltos, echo = FALSE}
-p_vel_acc_art_trial_all_ltos <- ggplot(data = subset(vel_acc_art_trial_all, vel_acc_art_trial_all$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+```{r vel_diff_art_trial_all_line_ltos, echo = FALSE}
+p_vel_diff_art_trial_all_ltos <- ggplot(data = subset(vel_diff_art_trial_all, vel_diff_art_trial_all$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
   geom_line(position = position_dodge(.2)) +
@@
   labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
-p_vel_acc_art_trial_all_ltos
+p_vel_diff_art_trial_all_ltos
 ```
 
-```{r vel_acc_art_trial_all_line_stol, echo = FALSE}
-p_vel_acc_art_trial_all_stol <- ggplot(data = subset(vel_acc_art_trial_all, vel_acc_art_trial_all$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
+```{r vel_diff_art_trial_all_line_stol, echo = FALSE}
+p_vel_diff_art_trial_all_stol <- ggplot(data = subset(vel_diff_art_trial_all, vel_diff_art_trial_all$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
   geom_line(position = position_dodge(.2)) +
@@
   labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
-p_vel_acc_art_trial_all_stol
+p_vel_diff_art_trial_all_stol
 ```
+
+## Sequence plot
+#### Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 6, fig.height = 2, echo = FALSE}
+# Average vel_dynocity for each note
+# For each participant
+vel_dyn_seq <- aggregate(Velocity~SubNr*Condition*Skill*NoteNr, data = df_vel_dyn, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_dyn_seq <- cbind(vel_dyn_seq[,1:4], as.data.frame(vel_dyn_seq[,5]))
+# Change colnames
+colnames(vel_dyn_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD")
+
+# Group mean
+vel_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_dyn_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_dyn_seq_stats <- cbind(vel_dyn_seq_stats[,1:3], as.data.frame(vel_dyn_seq_stats[,4]))
+# Change colnames
+colnames(vel_dyn_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM")
+
+# Average acceleration for each note
+# For each participant
+vel_dyn_diff_seq <- aggregate(Diff~SubNr*Condition*Skill*Interval, data = df_vel_diff_dyn, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_dyn_diff_seq <- cbind(vel_dyn_diff_seq[,1:4], as.data.frame(vel_dyn_diff_seq[,5]))
+# Change colnames
+colnames(vel_dyn_diff_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+vel_dyn_diff_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_dyn_diff_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_dyn_diff_seq_stats <- cbind(vel_dyn_diff_seq_stats[,1:3], as.data.frame(vel_dyn_diff_seq_stats[,4]))
+# Change colnames
+colnames(vel_dyn_diff_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_vel_dyn_seq <- ggplot(data = vel_dyn_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(0.05)) + 
+  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
+p_vel_dyn_seq
+
+p_vel_dyn_diff_seq <- ggplot(data = vel_dyn_diff_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
+p_vel_dyn_diff_seq
+```
+
+#### Articulation{#sequence_art}
+```{r seq_art, fig.width = 6, fig.height = 2, echo = FALSE}
+# Average vel_artocity for each note
+# For each participant
+vel_art_seq <- aggregate(Velocity~SubNr*Condition*Skill*NoteNr, data = df_vel_art, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_art_seq <- cbind(vel_art_seq[,1:4], as.data.frame(vel_art_seq[,5]))
+# Change colnames
+colnames(vel_art_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD")
+
+# Group mean
+vel_art_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_art_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_art_seq_stats <- cbind(vel_art_seq_stats[,1:3], as.data.frame(vel_art_seq_stats[,4]))
+# Change colnames
+colnames(vel_art_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM")
+
+# Average acceleration for each note
+# For each participant
+vel_art_diff_seq <- aggregate(Diff~SubNr*Condition*Skill*Interval, data = df_vel_diff_art, 
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_art_diff_seq <- cbind(vel_art_diff_seq[,1:4], as.data.frame(vel_art_diff_seq[,5]))
+# Change colnames
+colnames(vel_art_diff_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+vel_art_diff_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_art_diff_seq,
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_art_diff_seq_stats <- cbind(vel_art_diff_seq_stats[,1:3], as.data.frame(vel_art_diff_seq_stats[,4]))
+# Change colnames
+colnames(vel_art_diff_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_vel_art_seq <- ggplot(data = vel_art_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(0.05)) + 
+  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
+p_vel_art_seq
+
+p_vel_art_diff_seq <- ggplot(data = vel_art_diff_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
+p_vel_art_diff_seq
+```

### 2020-01-06 script/RNotebook/ioi_1.Rmd
modify fig.height/fig.width
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-```{r ioi_art_trial_line, echo = FALSE, fig.height = 8}
+```{r ioi_art_trial_line, echo = FALSE, fig.height = 4}
 p_ioi_art_trial <- ggplot(data = ioi_art_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ##### Line graph
-```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 8}
+```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 4}
 p_ioi_art_var_trial <- ggplot(data = ioi_art_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r ioi_dyn_trial_line, echo = FALSE, fig.height = 8}
+```{r ioi_dyn_trial_line, echo = FALSE, fig.height = 4}
 p_ioi_dyn_trial <- ggplot(data = ioi_dyn_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ##### Line graph
-```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 8}
+```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 4}
 p_ioi_dyn_var_trial <- ggplot(data = ioi_dyn_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Articulation{#sequence_art}
-```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
+```{r seq_art, fig.width = 6, fig.height = 2, echo = FALSE}
 # For each individual
@@
 #### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 15, fig.height = 5, echo = FALSE}
+```{r seq_dyn, fig.width = 6, fig.height = 2, echo = FALSE}
 # For each individual

### 2020-01-06 script/RNotebook/articulation.Rmd
modify fig.height/fig.width
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ##### Line graph
-```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 8}
+```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 4}
 p_kot_art_trial_legato <- ggplot(data = subset(kot_art_trial, kot_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r kot_art_trial_line_staccato, echo = FALSE, fig.height = 8}
+```{r kot_art_trial_line_staccato, echo = FALSE, fig.height = 4}
 p_kot_art_trial_staccato <- ggplot(data = subset(kot_art_trial, kot_art_trial$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ##### Line graph
-```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 8}
+```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 4}
 p_kor_art_trial_legato <- ggplot(data = subset(kor_art_trial, kor_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r kor_art_trial_line_staccato, echo = FALSE, fig.height = 8}
+```{r kor_art_trial_line_staccato, echo = FALSE, fig.height = 4}
 p_kor_art_trial_staccato <- ggplot(data = subset(kor_art_trial, kor_art_trial$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ##### Line graph
-```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
+```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
 p_kot_dyn_trial_forte <- ggplot(data = subset(kot_dyn_trial, kot_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r kot_dyn_trial_line_piano, echo = FALSE, fig.height = 8}
+```{r kot_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
 p_kot_dyn_trial_piano <- ggplot(data = subset(kot_dyn_trial, kot_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ##### Line graph
-```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
+```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
 p_kor_dyn_trial_forte <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r kor_dyn_trial_line_piano, echo = FALSE, fig.height = 8}
+```{r kor_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
 p_kor_dyn_trial_piano <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +

### 2020-01-12 script/RNotebook/ioi_1.Rmd
correct plots, typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r ioi_dyn_all_bar, echo = FALSE}
-p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_all, aes(x = Skill, y = Mean, fill = Condition)) +
+p_ioi_dyn_all_bar <- ggplot(data = ioi_dyn_all, aes(x = Condition, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
   geom_hline(yintercept = 1, linetype = "dashed") + # Tempo + 
-  labs(y = "Normalised IOI", subtitle = "IOI: Dynamics") +  coord_cartesian(ylim = c(.75, 1.1))
-p_ioi_dyn_ch_all_bar
+  labs(y = "Normalised IOI", subtitle = "IOI: Dynamics") +  coord_cartesian(ylim = c(.75, 1.1)) +
+  theme(legend.position = "None")
+p_ioi_dyn_all_bar
 ```

### 2020-02-07 script/RNotebook/ioi.Rmd
update ioi analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: 'Analysis: The Sound of Teaching Music 1'
+title: 'Tempo Analysis: The Sound of Teaching Music 1'
 output:
@@
 
-##### Description: This is a summary of all analyses for a tempo.
+Description: This is a summary of all analyses for a tempo.
+
+- Last checked: 07/02/20
 
 # Outline - Interonset intervals (IOIs)
-### [Experimental Design](#design)
-
-### [Articulation](#ioi-articulation)
+## [Experimental Design](#design)
+
+## [Articulation](#ioi-articulation)
 
@@
   + [Trial by trial](#art_ind_trial)
-  + [Variability all](#art_ind_var_all)
+  + [Variability](#art_ind_var_all)
   + [Variability trial-by-trial](#art_ind_var_trial)
@@
   + [Trial by trial](#art_group_trial)
-  + [Variability all](#art_group_var_all)
+  + [Variability](#art_group_var_all)
   + [Variability trial-by-trial](#art_group_var_trial)
@@
 
-### [Dynamics](#ioi-dynamics)
+## [Dynamics](#ioi-dynamics)
 
@@
 - Independent variables: Condition (2 levels: teaching vs. performing) x Skill (2 levels: articulation vs. dynamics)
-- Dependent variables: Interonset intervals, variability (SD/mean IOI)
-
-### Main predictions
+- Dependent variables: Interonset intervals, Variability (SD/mean IOI)
+
+## Main predictions
 Participants will play slower while teaching regardless of the skills.
@@
 
-## IOI-Articulation 
-### 1. All - individual{#art_ind_all}
+# IOI-Articulation 
+## 1. All - individual{#art_ind_all}
 - Average IOIs for each participant.
@@
 
-#### Plot
-##### Bar chart
+### Plot
+#### Bar chart
 - Error bars show SD of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation")
+  #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
 p_ioi_art
@@
 
-### 2. All - group{#art_group_all}
+## 2. All - group{#art_group_all}
 - Average IOIs for each condition.
@@
 
-#### Plot
-##### Bar chart
+### Plot
+#### Bar chart
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
   labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) +
@@
 
-##### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm)
+#### Box plot
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
   labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + #coord_cartesian(ylim = c(100, 210)) +
@@
 
-#### Statistics
-##### Normality check
+### Statistics
+#### Normality check
 ```{r normality1, echo = FALSE}
@@
 
-##### Paired t-test
+#### Paired t-test
 ```{r stats1, echo = FALSE}
@@
 
-##### Effect size
+#### Effect size
 ```{r stats2, echo = FALSE}
@@
 
-### 3. Phrase boundaries - individual{#art_ind_phrase}
+## 3. Phrase boundaries - individual{#art_ind_phrase}
 - Compare IOIs at phrase boundaries with others fir each participant.
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
   labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
@@
 
-### 4. Phrase boundaries - group{#art_group_phrase}
+## 4. Phrase boundaries - group{#art_group_phrase}
 - Compare IOIs at phrase boundaries with others in the two conditions.
@@
 
-#### Plot
+### Plot
+#### Bar chart
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #annotate("text", 1, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
   #facet_grid(. ~ Boundary) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 250))
+  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 210))
 p_ioi_art_ch_all_bar
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
   labs(y = "IOI (ms)", subtitle = "IOI: Articulation")
@@
 
-#### Statistics
-##### Normality check
+### Statistics
+#### Normality check
 ```{r normality2, echo = FALSE}
@@
 
-##### Two-way ANOVA (ez)
+#### Two-way ANOVA (ez)
 ```{r stats3, echo = FALSE}
@@
 
-##### Two-way ANOVA (aov)
+#### Two-way ANOVA (aov)
 ```{r stats4, echo = FALSE}
@@
 
-### 5. Trial by trial - individual{#art_ind_trial}
+## 5. Trial by trial - individual{#art_ind_trial}
 - Average IOIs for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
-- A dashed line represents the tempo given by a metronome (80bpm)
+### Plot
+#### Line graph
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
@@
 
-### 6. Trial by trial - group{#art_group_trial}
+## 6. Trial by trial - group{#art_group_trial}
 - Average IOIs for each trial for each condition.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
@@
 
-### 7. Variability all - individual{#art_ind_var_all}
+## 7. Variability all - individual{#art_ind_var_all}
 - Average variability for each participant.
@@
 ```{r ioi_art_var, echo = FALSE}
-# Create a data frame for variability (SD/)
+# Create a data frame for Variability (SD/mean IOI)
 df_ioi_art_var <- data.frame()
 for (subnr in unique(df_ioi_art$SubNr)){
-  for (block in unique(df_ioi_art$BlockNr)){
+  for (block in c(1:4)){
     cond = as.character(unique(df_ioi_art$Condition[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
     skill = as.character(unique(df_ioi_art$Skill[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
-    for (trial in unique(df_ioi_art$TrialNr)){
+    for (trial in c(1:8)){
       df_current <- df_ioi_art %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
@@
 
-#### Plot
-##### Bar graph
+### Plot
+#### Bar graph
 - Error bars show SD of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
@@
 
-### 8. Variability all - group{#art_group_var_all}
+## 8. Variability all - group{#art_group_var_all}
 - Average variability for each condition.
@@
 
-#### Plot
-##### Bar graph
+### Plot
+#### Bar graph
 - Error bars show SEM of the mean
@@
 
-##### Box plot
+#### Box plot
 ```{r ioi_art_var_all_box, echo = FALSE}
@@
 
-#### Statistics
-##### Normality check
+### Statistics
+#### Normality check
 ```{r normality3, echo = FALSE}
@@
 
-##### Paired t-test
+#### Paired t-test
 ```{r stats5, echo = FALSE}
@@
 
-##### Effect size
+#### Effect size
 ```{r stats6, echo = FALSE}
@@
 
-### 9. Variability trial-by-trial - individual{#art_ind_var_trial}
+## 9. Variability trial-by-trial - individual{#art_ind_var_trial}
 - Average variability for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 8}
@@
 
-### 10. Variability trial-by-trial - group{#art_group_var_trial}
+## 10. Variability trial-by-trial - group{#art_group_var_trial}
 - Average variability for each trial for each condition.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 - Error bars show SEM of the mean
@@
 
-###########
-###########
-
-## IOI-Dynamics
-### 1. All - individual{#dyn_ind_all}
+# IOI-Dynamics
+## 1. All - individual{#dyn_ind_all}
 - Average IOIs for each participant.
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics")
+  #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
 p_ioi_dyn
@@
 
-### 2. All - group{#dyn_group_all}
+## 2. All - group{#dyn_group_all}
 - Average IOIs for each condition.
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
   labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(120, 210)) +
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
   labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + #coord_cartesian(ylim = c(100, 210)) +
@@
 
-#### Statistics
-##### Normality check
+### Statistics
+#### Normality check
 ```{r normality4, echo = FALSE}
@@
 
-##### Effect size
+#### Effect size
 ```{r stats8, echo = FALSE}
@@
 
-### 3. Phrase boundaries - individual{#dyn_ind_phrase}
+## 3. Phrase boundaries - individual{#dyn_ind_phrase}
 - Compare IOIs at phrase boundaries with others fir each participant.
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
   labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
@@
 
-### 4. Phrase boundaries - group{#dyn_group_phrase}
+## 4. Phrase boundaries - group{#dyn_group_phrase}
 - Compare IOIs at phrase boundaries with others in the two conditions.
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #annotate("text", 1, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
   #facet_grid(. ~ Boundary) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 250))
+  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 210))
 p_ioi_dyn_ch_all_bar
@@
   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
+  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
   labs(y = "IOI (ms)", subtitle = "IOI: Dynamics")
@@
 
-#### Statistics
-##### Normality check
+### Statistics
+#### Normality check
 ```{r normality5, echo = FALSE}
@@
 
-##### Two-way ANOVA (ez)
+#### Two-way ANOVA (ez)
 ```{r stats9, echo = FALSE}
@@
 
-##### Two-way ANOVA (aov)
+#### Two-way ANOVA (aov)
 ```{r stats10, echo = FALSE}
@@
 
-### 5. Trial by trial - individual{#dyn_ind_trial}
+## 5. Trial by trial - individual{#dyn_ind_trial}
 - Average IOIs for each trial for each participant.
@@
 
-### 6. Trial by trial - group{#dyn_group_trial}
+## 6. Trial by trial - group{#dyn_group_trial}
 - Average IOIs for each trial for each condition.
@@
 
-### 7. Variability all - individual{#dyn_ind_var_all}
+## 7. Variability all - individual{#dyn_ind_var_all}
 - Average variability for each participant.
@@
 for (subnr in unique(df_ioi_dyn$SubNr)){
-  for (block in unique(df_ioi_dyn$BlockNr)){
+  for (block in c(1:4)){
     cond = as.character(unique(df_ioi_dyn$Condition[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
     skill = as.character(unique(df_ioi_dyn$Skill[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
-    for (trial in unique(df_ioi_dyn$TrialNr)){
+    for (trial in c(1:8)){
       df_current <- df_ioi_dyn %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
@@
 
-#### Plot
-##### Bar graph
+### Plot
+#### Bar graph
 - Error bars show SD of the mean
@@
 
-### 8. Variability all - group{#dyn_group_var_all}
+## 8. Variability all - group{#dyn_group_var_all}
 - Average variability for each condition.
@@
 
-##### Box plot
+#### Box plot
 ```{r ioi_dyn_var_all_box, echo = FALSE}
@@
 
-#### Statistics
-##### Normality check
+### Statistics
+#### Normality check
 ```{r normality6, echo = FALSE}
@@
 
-### 9. Variability trial-by-trial - individual{#dyn_ind_var_trial}
+## 9. Variability trial-by-trial - individual{#dyn_ind_var_trial}
 - Average variability for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 8}
@@
 
-### 10. Variability trial-by-trial - group{#dyn_group_var_trial}
+## 10. Variability trial-by-trial - group{#dyn_group_var_trial}
 Average variability for each trial for each condition.
@@
 
-###########
-###########
-
-## IOI-Order-Articulation
+# IOI-Order-Articulation
 - Investigate whether there is any order effect of the conditions.
 
-### 1. All{#art_order_group_all}
+## 1. All{#art_order_group_all}
 ```{r ioi_art_order_all, echo = FALSE}
@@
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
@@
 
-#### Statistics
-##### Normality check
+### Statistics
+#### Normality check
 ```{r normality7, echo = FALSE}
@@
 
-##### Two-way ANOVA (ez)
+#### Two-way ANOVA (ez)
 ```{r stats13, echo = FALSE}
@@
 
-##### Two-way ANOVA (aov)
+#### Two-way ANOVA (aov)
 ```{r stats14, echo = FALSE}
@@
 
-## IOI-Order-Dynamics
+# IOI-Order-Dynamics
 - Investigate whether there is any order effect of the conditions.
 
-### 1. All{#dyn_order_group_all}
+## 1. All{#dyn_order_group_all}
 ```{r ioi_dyn_order_all, echo = FALSE}
@@
 
-#### Statistics
-##### Normality check
+### Statistics
+#### Normality check
 ```{r normality8, echo = FALSE}
@@
 
-##### Two-way ANOVA (ez)
+#### Two-way ANOVA (ez)
 ```{r stats15, echo = FALSE}
@@
 
-##### Two-way ANOVA (aov)
+#### Two-way ANOVA (aov)
 ```{r stats16, echo = FALSE}
@@
 
-## Sequence plot
-#### Articulation{#sequence_art}
-```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
+# Sequence plot
+## Articulation{#sequence_art}
+```{r seq_art, fig.width = 20, fig.height = 5, echo = FALSE}
 # For each individual
@@
 
-#### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 15, fig.height = 5, echo = FALSE}
+## Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 20, fig.height = 5, echo = FALSE}
 # For each individual

### 2020-02-07 script/RNotebook/articulation.Rmd
correct plots, typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update kot/kor analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2019-12-10 script/RNotebook/articulation.Rmd
@@
 ---
-title: 'Analysis: The Sound of Teaching Music 2'
+title: 'Articulation Analysis: The Sound of Teaching Music 1'
 output:
@@
 
-##### Description: This is a summary of all analyses for articulation.
+Description: This is a summary of all analyses for articulation.
+
+- Last checked: 07/02/20
+
+# Outline - Articulation (KOT, KOR)
+## [Experimental Design](#design)
+
+## [Articulation](#kot-articulation)
+
+### KOT
+- Individual results
+  + [All](#kot_art_ind_all)
+  + [Trial by trial](#kot_art_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#kot_art_group_all)
+  + [Trial by trial](#kot_art_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#kot_art_order_group_all)
+
+- [Sequence plot](#sequence_art)
+
+### KOR
+- Individual results
+  + [All](#kor_art_ind_all)
+  + [Trial by trial](#kor_art_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#kor_art_group_all)
+  + [Trial by trial](#kor_art_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#kor_art_order_group_all)
+
+## [Dynamics](#kot-dynamics)
+
+### KOT
+- Individual results
+  + [All](#kot_dyn_ind_all)
+  + [Trial by trial](#kot_dyn_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#kot_dyn_group_all)
+  + [Trial by trial](#kot_dyn_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#kot_dyn_order_group_all)
+
+- [Sequence plot](#sequence_dyn)
+
+### KOR
+- Individual results
+  + [All](#kor_dyn_ind_all)
+  + [Trial by trial](#kor_dyn_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#kor_dyn_group_all)
+  + [Trial by trial](#kor_dyn_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#kor_dyn_order_group_all)
 
@@
 
-### Main predictions
+## Main predictions
 Participants will exaggerate legato and staccato more while teaching.
@@
 df_ioi$SubNr <- as.factor(df_ioi$SubNr)
-
-# SubNr as a factor
-df_ioi$SubNr <- as.factor(df_ioi$SubNr)
 df_kot$SubNr <- as.factor(df_kot$SubNr)
@@
 # Calculate KOR for each interval
-
 df_kor <- data.frame()
@@
 
-## Key-Overlap Time (KOT) - Articulation
-### 1. All - individual{#kot_art_ind_all}
+# Key-Overlap Time (KOT) - Articulation{#kot-articulation}
+## 1. All - individual{#kot_art_ind_all}
 - Average KOT for each participant.
@@
 # For each individual
-kot_art <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
+kot_art <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-#### Plot
-##### Bar chart
+### Plot
+#### Bar chart
 - Error bars show SD of the mean
@@
 
-### 2. All - group{#kot_art_group_all}
+## 2. All - group{#kot_art_group_all}
 - Average KOT for each condition.
@@
 
-#### Plot
-##### Bar chart
+### Plot
+#### Bar chart
 - Error bars show SEM of the mean
@@
 
-##### Box plot
-
+#### Box plot
 ```{r kot_art_all_box, echo = FALSE}
@@
 
-#### Statistics
-##### TBC
+### Statistics
+#### Normality check
+```{r normality1, echo = FALSE}
+# Reduce unused factors
+kot_art$Subcomponent <- factor(kot_art$Subcomponent)
+kot_art_norm <- by(kot_art$Mean, list(kot_art$Condition, kot_art$Subcomponent), shapiro.test)
+print(kot_art_norm)
+```
 
-### 3. Trial by trial - individual{#kot_art_ind_trial}
+#### Two-way ANOVA (ez)
+```{r stats1, echo = FALSE}
+kot_art_aov <- ezANOVA(
+  data = subset(df_kot_art, df_kot_art$Subcomponent == "Legato" | df_kot_art$Subcomponent == "Staccato")
+  , dv = .(KOT)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kot_art_aov$ANOVA$F <- round(kot_art_aov$ANOVA$F, 4)
+kot_art_aov$ANOVA$p <- round(kot_art_aov$ANOVA$p, 4)
+print(kot_art_aov$ANOVA)
+```
+
+#### Two-way ANOVA (aov)
+```{r stats2, echo = FALSE}
+kot_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
+print(summary(kot_art_aov_2))
+```
+
+#### Post-hoc (direct comparison)
+```{r post-hoc1, echo = FALSE}
+kot_ttest_legato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"], paired = TRUE)
+kot_ttest_staccato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"], paired = TRUE)
+# Effect size
+kot_ttest_legato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Legato"], kot_art$Condition[kot_art$Subcomponent == "Legato"], paired = TRUE)
+kot_ttest_staccato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Staccato"], kot_art$Condition[kot_art$Subcomponent == "Staccato"], paired = TRUE)
+
+print(kot_ttest_legato)
+print(kot_ttest_staccato)
+print(kot_ttest_legato_cohend)
+print(kot_ttest_staccato_cohend)
+```
+
+#### Post-hoc (pairwise t test)
+```{r post-hoc2, echo = FALSE}
+# Create a group label for a post-hoc test
+kot_art$Group[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"] <- "Performing-Legato" 
+kot_art$Group[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
+kot_art$Group[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"] <- "Teaching-Legato"
+kot_art$Group[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
+# pairwise comparison
+kot_art_pairwise <- pairwise.t.test(kot_art$Mean, kot_art$Group, paired = TRUE, p.adjust.method = "BH")
+kot_art_pairwise$p.value <- round(kot_art_pairwise$p.value, 4)
+print(kot_art_pairwise)
+```
+
+## 3. Trial by trial - individual{#kot_art_ind_trial}
 - Average KOT for each trial for each participant.
@@
 # For each individual
-kot_art_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
+kot_art_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 8}
@@
 
-### 4. Trial by trial - group{#kot_art_group_trial}
+## 4. Trial by trial - group{#kot_art_group_trial}
 - Average KOT for each trial for each condition.
@@
 
-## Key-Overlap Ratio (KOR) - Articulation
-### 1. All - individual{#kor_art_ind_all}
+# Key-Overlap Ratio (KOR) - Articulation
+## 1. All - individual{#kor_art_ind_all}
 - Average KOR for each participant.
@@
 # For each individual
-kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
+kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-### 2. All - group{#kor_art_group_all}
+## 2. All - group{#kor_art_group_all}
 - Average KOR for each condition.
@@
 
-##### Box plot
-
+#### Box plot
 ```{r kor_art_all_box, echo = FALSE}
@@
 
-#### Statistics
-##### TBC
+### Statistics
+#### Normality check
+```{r normality2, echo = FALSE}
+# Reduce unused factors
+kor_art$Subcomponent <- factor(kor_art$Subcomponent)
+kor_art_norm <- by(kor_art$Mean, list(kor_art$Condition, kor_art$Subcomponent), shapiro.test)
+print(kor_art_norm)
+```
 
-### 3. Trial by trial - individual{#kor_art_ind_trial}
+#### Two-way ANOVA (ez)
+```{r stats3, echo = FALSE}
+kor_art_aov <- ezANOVA(
+  data = subset(df_kor_art, df_kor_art$Subcomponent == "Legato" | df_kor_art$Subcomponent == "Staccato")
+  , dv = .(KOR)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kor_art_aov$ANOVA$F <- round(kor_art_aov$ANOVA$F, 4)
+kor_art_aov$ANOVA$p <- round(kor_art_aov$ANOVA$p, 4)
+print(kor_art_aov$ANOVA)
+```
+
+#### Two-way ANOVA (aov)
+```{r stats4, echo = FALSE}
+kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
+print(summary(kor_art_aov_2))
+```
+
+#### Post-hoc (direct comparison)
+```{r post-hoc3, echo = FALSE}
+kor_ttest_legato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"], paired = TRUE)
+kor_ttest_staccato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"], paired = TRUE)
+# Effect size
+kor_ttest_legato_cohend <- cohen.d(kor_art$Mean[kor_art$Subcomponent == "Legato"], kor_art$Condition[kor_art$Subcomponent == "Legato"], paired = TRUE)
+kor_ttest_staccato_cohend <- cohen.d(kor_art$Mean[kor_art$Subcomponent == "Staccato"], kor_art$Condition[kor_art$Subcomponent == "Staccato"], paired = TRUE)
+
+print(kor_ttest_legato)
+print(kor_ttest_staccato)
+print(kor_ttest_legato_cohend)
+print(kor_ttest_staccato_cohend)
+```
+
+#### Post-hoc (pairwise t test)
+```{r post-hoc4, echo = FALSE}
+# Create a group label for a post-hoc test
+kor_art$Group[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"] <- "Performing-Legato" 
+kor_art$Group[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
+kor_art$Group[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"] <- "Teaching-Legato"
+kor_art$Group[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
+# pairwise comparison
+kor_art_pairwise <- pairwise.t.test(kor_art$Mean, kor_art$Group, paired = TRUE, p.adjust.method = "BH")
+kor_art_pairwise$p.value <- round(kor_art_pairwise$p.value, 4)
+print(kor_art_pairwise)
+```
+
+## 3. Trial by trial - individual{#kor_art_ind_trial}
 - Average KOR for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 8}
@@
 
-### 4. Trial by trial - group{#kor_art_group_trial}
+## 4. Trial by trial - group{#kor_art_group_trial}
 - Average KOR d for each trial for each condition.
@@
 
-###########
-###########
-
-## Key-Overlap Time (KOT) - Dynamics
-### 1. All - individual{#kot_dyn_ind_all}
+# Key-Overlap Time (KOT) - Dynamics{#kot-dynamics}
+## 1. All - individual{#kot_dyn_ind_all}
 - Average KOT for each participant.
@@
 # For each individual
-kot_dyn <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
+kot_dyn <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-### 2. All - group{#kot_dyn_group_all}
+## 2. All - group{#kot_dyn_group_all}
 - Average KOT for each condition.
@@
 
-##### Box plot
+#### Box plot
 
@@
 
-#### Statistics
-##### TBC
+### Statistics
+#### Normality check
+```{r normality3, echo = FALSE}
+# Reduce unused factors
+kot_dyn$Subcomponent <- factor(kot_dyn$Subcomponent)
+kot_dyn_norm <- by(kot_dyn$Mean, list(kot_dyn$Condition, kot_dyn$Subcomponent), shapiro.test)
+print(kot_dyn_norm)
+```
 
-### 3. Trial by trial - individual{#kot_dyn_ind_trial}
+#### Two-way ANOVA (ez)
+```{r stats5, echo = FALSE}
+kot_dyn_aov <- ezANOVA(
+  data = subset(df_kot_dyn, df_kot_dyn$Subcomponent == "Forte" | df_kot_dyn$Subcomponent == "Piano")
+  , dv = .(KOT)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kot_dyn_aov$ANOVA$F <- round(kot_dyn_aov$ANOVA$F, 4)
+kot_dyn_aov$ANOVA$p <- round(kot_dyn_aov$ANOVA$p, 4)
+print(kot_dyn_aov$ANOVA)
+```
+
+#### Two-way ANOVA (aov)
+```{r stats6, echo = FALSE}
+kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
+print(summary(kot_dyn_aov_2))
+```
+
+## 3. Trial by trial - individual{#kot_dyn_ind_trial}
 - Average KOT for each trial for each participant.
@@
 # For each individual
-kot_dyn_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
+kot_dyn_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
@@
 
-### 4. Trial by trial - group{#kot_dyn_group_trial}
+## 4. Trial by trial - group{#kot_dyn_group_trial}
 - Average KOT for each trial for each condition.
@@
 
-## Key-Overlap Ratio (KOR) - Dynamics
-### 1. All - individual{#kor_dyn_ind_all}
+# Key-Overlap Ratio (KOR) - Dynamics
+## 1. All - individual{#kor_dyn_ind_all}
 - Average KOR for each participant.
@@
 # For each individual
-kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
+kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-### 2. All - group{#kor_dyn_group_all}
+## 2. All - group{#kor_dyn_group_all}
 - Average KOR for each condition.
@@
 
-##### Box plot
-
+#### Box plot
 ```{r kor_dyn_all_box, echo = FALSE}
@@
 
-#### Statistics
-##### TBC
+### Statistics
+#### Normality check
+```{r normality4, echo = FALSE}
+# Reduce unused factors
+kor_dyn$Subcomponent <- factor(kor_dyn$Subcomponent)
+kor_dyn_norm <- by(kor_dyn$Mean, list(kor_dyn$Condition, kor_dyn$Subcomponent), shapiro.test)
+print(kor_dyn_norm)
+```
 
-### 3. Trial by trial - individual{#kor_dyn_ind_trial}
+#### Two-way ANOVA (ez)
+```{r stats7, echo = FALSE}
+kor_dyn_aov <- ezANOVA(
+  data = subset(df_kor_dyn, df_kor_dyn$Subcomponent == "Forte" | df_kor_dyn$Subcomponent == "Piano")
+  , dv = .(KOR)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kor_dyn_aov$ANOVA$F <- round(kor_dyn_aov$ANOVA$F, 4)
+kor_dyn_aov$ANOVA$p <- round(kor_dyn_aov$ANOVA$p, 4)
+print(kor_dyn_aov$ANOVA)
+```
+
+#### Two-way ANOVA (aov)
+```{r stats8, echo = FALSE}
+kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
+print(summary(kor_dyn_aov_2))
+```
+
+## 3. Trial by trial - individual{#kor_dyn_ind_trial}
 - Average KOR for each trial for each participant.
@@
 # For each individual
-kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
+kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
@@
 
-### 4. Trial by trial - group{#kor_dyn_group_trial}
+## 4. Trial by trial - group{#kor_dyn_group_trial}
 - Average KOR for each trial for each condition.
@@
 ```
+
+# KOT-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#kot_art_order_group_all}
+```{r kot_art_order_all, echo = FALSE}
+# For each participant
+kot_art_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kot_art_order <- cbind(kot_art_order[,1:5], as.data.frame(kot_art_order[,6]))
+# Change colnames
+colnames(kot_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kot_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_art_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_art_order_all <- cbind(kot_art_order_all[,1:4], as.data.frame(kot_art_order_all[,5]))
+# Change colnames
+colnames(kot_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kot_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kot_art_order_all_bar, echo = FALSE}
+p_kot_art_order_all <- ggplot(data = kot_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
+p_kot_art_order_all
+```
+
+#### Box plot
+```{r kot_art_order_all_box, echo = FALSE}
+p_kot_art_order_all_box <- ggplot(data = kot_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT", subtitle = "KOT: Articulation")
+p_kot_art_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality5, echo = FALSE}
+# Normality check
+kot_art_order$Subcomponent <- factor(kot_art_order$Subcomponent)
+kot_art_order_norm <- by(kot_art_order$Mean, list(kot_art_order$Condition, kot_art_order$Subcomponent, kot_art_order$Order), shapiro.test)
+print(kot_art_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats9, echo = FALSE}
+kot_art_order_aov <- ezANOVA(
+  data = df_kot_art
+  , dv = .(KOT)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kot_art_order_aov$ANOVA$F <- round(kot_art_order_aov$ANOVA$F, 4)
+kot_art_order_aov$ANOVA$p <- round(kot_art_order_aov$ANOVA$p, 4)
+print(kot_art_order_aov)
+```
+
+# KOR-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#kor_art_order_group_all}
+```{r kor_art_order_all, echo = FALSE}
+# For each participant
+kor_art_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kor_art_order <- cbind(kor_art_order[,1:5], as.data.frame(kor_art_order[,6]))
+# Change colnames
+colnames(kor_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kor_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_art_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kor_art_order_all <- cbind(kor_art_order_all[,1:4], as.data.frame(kor_art_order_all[,5]))
+# Change colnames
+colnames(kor_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kor_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kor_art_order_all_bar, echo = FALSE}
+p_kor_art_order_all <- ggplot(data = kor_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Articulation")
+p_kor_art_order_all
+```
+
+#### Box plot
+```{r kor_art_order_all_box, echo = FALSE}
+p_kor_art_order_all_box <- ggplot(data = kor_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Articulation")
+p_kor_art_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality6, echo = FALSE}
+# Normality check
+kor_art_order$Subcomponent <- factor(kor_art_order$Subcomponent)
+kor_art_order_norm <- by(kor_art_order$Mean, list(kor_art_order$Condition, kor_art_order$Subcomponent, kor_art_order$Order), shapiro.test)
+print(kor_art_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats10, echo = FALSE}
+kor_art_order_aov <- ezANOVA(
+  data = df_kor_art
+  , dv = .(KOR)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kor_art_order_aov$ANOVA$F <- round(kor_art_order_aov$ANOVA$F, 4)
+kor_art_order_aov$ANOVA$p <- round(kor_art_order_aov$ANOVA$p, 4)
+print(kor_art_order_aov)
+```
+
+# KOT-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#kot_dyn_order_group_all}
+```{r kot_dyn_order_all, echo = FALSE}
+# For each participant
+kot_dyn_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kot_dyn_order <- cbind(kot_dyn_order[,1:5], as.data.frame(kot_dyn_order[,6]))
+# Change colnames
+colnames(kot_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kot_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_dyn_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_dyn_order_all <- cbind(kot_dyn_order_all[,1:4], as.data.frame(kot_dyn_order_all[,5]))
+# Change colnames
+colnames(kot_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kot_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kot_dyn_order_all_bar, echo = FALSE}
+p_kot_dyn_order_all <- ggplot(data = kot_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
+p_kot_dyn_order_all
+```
+
+#### Box plot
+```{r kot_dyn_order_all_box, echo = FALSE}
+p_kot_dyn_order_all_box <- ggplot(data = kot_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
+p_kot_dyn_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality7, echo = FALSE}
+# Normality check
+kot_dyn_order$Subcomponent <- factor(kot_dyn_order$Subcomponent)
+kot_dyn_order_norm <- by(kot_dyn_order$Mean, list(kot_dyn_order$Condition, kot_dyn_order$Subcomponent, kot_dyn_order$Order), shapiro.test)
+print(kot_dyn_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats11, echo = FALSE}
+kot_dyn_order_aov <- ezANOVA(
+  data = df_kot_dyn
+  , dv = .(KOT)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kot_dyn_order_aov$ANOVA$F <- round(kot_dyn_order_aov$ANOVA$F, 4)
+kot_dyn_order_aov$ANOVA$p <- round(kot_dyn_order_aov$ANOVA$p, 4)
+print(kot_dyn_order_aov)
+```
+
+# KOR-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#kor_dyn_order_group_all}
+```{r kor_dyn_order_all, echo = FALSE}
+# For each participant
+kor_dyn_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kor_dyn_order <- cbind(kor_dyn_order[,1:5], as.data.frame(kor_dyn_order[,6]))
+# Change colnames
+colnames(kor_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kor_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_dyn_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kor_dyn_order_all <- cbind(kor_dyn_order_all[,1:4], as.data.frame(kor_dyn_order_all[,5]))
+# Change colnames
+colnames(kor_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kor_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kor_dyn_order_all_bar, echo = FALSE}
+p_kor_dyn_order_all <- ggplot(data = kor_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Dynamics")
+p_kor_dyn_order_all
+```
+
+#### Box plot
+```{r kor_dyn_order_all_box, echo = FALSE}
+p_kor_dyn_order_all_box <- ggplot(data = kor_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Dynamics")
+p_kor_dyn_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality8, echo = FALSE}
+# Normality check
+kor_dyn_order$Subcomponent <- factor(kor_dyn_order$Subcomponent)
+kor_dyn_order_norm <- by(kor_dyn_order$Mean, list(kor_dyn_order$Condition, kor_dyn_order$Subcomponent, kor_dyn_order$Order), shapiro.test)
+print(kor_dyn_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats12, echo = FALSE}
+kor_dyn_order_aov <- ezANOVA(
+  data = df_kor_dyn
+  , dv = .(KOR)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kor_dyn_order_aov$ANOVA$F <- round(kor_dyn_order_aov$ANOVA$F, 4)
+kor_dyn_order_aov$ANOVA$p <- round(kor_dyn_order_aov$ANOVA$p, 4)
+print(kor_dyn_order_aov)
+```
+
+# Sequence plot
+## Articulation{#sequence_art}
+```{r seq_art, fig.width = 20, fig.height = 5, echo = FALSE}
+# For each individual
+kot_art_seq <- aggregate(KOT~SubNr*Condition*Skill*Interval, data = df_kot_art,
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+kot_art_seq <- cbind(kot_art_seq[,1:4], as.data.frame(kot_art_seq[,5]))
+# Change colnames
+colnames(kot_art_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+kot_art_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = kot_art_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_art_seq_stats <- cbind(kot_art_seq_stats[,1:3], as.data.frame(kot_art_seq_stats[,4]))
+# Change colnames for each interval
+colnames(kot_art_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_kot_art_seq <- ggplot(data = kot_art_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+p_kot_art_seq
+```
+
+## Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 20, fig.height = 5, echo = FALSE}
+# For each individual
+kot_dyn_seq <- aggregate(KOT~SubNr*Condition*Skill*Interval, data = df_kot_dyn,
+                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+kot_dyn_seq <- cbind(kot_dyn_seq[,1:4], as.data.frame(kot_dyn_seq[,5]))
+# Change colnames
+colnames(kot_dyn_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+
+# Group mean
+kot_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = kot_dyn_seq,
+                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_dyn_seq_stats <- cbind(kot_dyn_seq_stats[,1:3], as.data.frame(kot_dyn_seq_stats[,4]))
+# Change colnames for each interval
+colnames(kot_dyn_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
+
+p_kot_dyn_seq <- ggplot(data = kot_dyn_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+p_kot_dyn_seq
+```

### 2020-02-08 script/RNotebook/dynamics.Rmd
correct plots, typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update kv/diff analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2019-11-28 script/RNotebook/dynamics.Rmd
@@
 ---
-title: 'Dynamics: Analysis: The Sound of Teaching Music 2'
+title: 'Dynamics Analysis: The Sound of Teaching Music 1'
 output:
@@
 
-##### Description: This is a summary of all analyses for dynamics.
+Description: This is a summary of all analyses for dynamics.
+
+- Last checked: 07/02/20
+
+# Outline - Dynamics (Key Velocity)
+## [Experimental Design](#design)
+
+## [Dynamics](#vel-dynamics)
+
+### KV
+- Individual results
+  + [All](#vel_dyn_ind_all)
+  + [Trial by trial](#vel_dyn_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#vel_dyn_group_all)
+  + [Trial by trial](#vel_dyn_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#vel_dyn_order_group_all)
+
+### KV-Diff
+- Individual results
+  + [All](#vel_acc_dyn_ind_all)
+  + [Trial by trial](#vel_acc_dyn_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#vel_acc_dyn_group_all)
+  + [Trial by trial](#vel_acc_dyn_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#vel_acc_dyn_order_group_all)
+
+- [Sequence plot](#sequence_dyn)
+
+## [Articulation](#vel_articulation)
+
+### KV
+- Individual results
+  + [All](#vel_art_ind_all)
+  + [Trial by trial](#vel_art_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#vel_art_group_all)
+  + [Trial by trial](#vel_art_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#vel_art_order_group_all)
+
+### KV-Diff
+- Individual results
+  + [All](#vel_acc_art_ind_all)
+  + [Trial by trial](#vel_acc_art_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#vel_acc_art_group_all)
+  + [Trial by trial](#vel_acc_art_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#vel_acc_art_order_group_all)
+
+- [Sequence plot](#sequence_art)
 
@@
 - Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato)
-- Dependent variables: Key Velocity (KV), KV difference (KV-Diff)
+- Dependent variables: Key Velocity (KV), KV difference
 
-### Main predictions
-- Participants will exaggerate forte and piano more while teaching.
-- Participants will make a larger contrast between forte and piano.
+## Main predictions
+Participants will exaggerate forte and piano more while teaching.
 
@@
 
-## Key Velocity (KV) - Dynamics
-### 1. All - individual{#vel_dyn_ind_all}
+# Key Velocity (KV) - Dynamics
+## 1. All - individual{#vel_dyn_ind_all}
 - Average KV for each participant.
@@
 
-### 2. All - group{#vel_dyn_group_all}
+## 2. All - group{#vel_dyn_group_all}
 - Average KV for each condition.
@@
 
-##### Box plot
-
+#### Box plot
 ```{r vel_dyn_all_box, echo = FALSE}
@@
 
-#### Statistics (TBC)
+### Statistics
+#### Normality check
+```{r normality1, echo = FALSE}
+# Reduce unused factors
+vel_dyn$Subcomponent <- factor(vel_dyn$Subcomponent)
+vel_dyn_norm <- by(vel_dyn$Mean, list(vel_dyn$Condition, vel_dyn$Subcomponent), shapiro.test)
+print(vel_dyn_norm)
+```
 
-### 3. Trial by trial - individual{#vel_dyn_ind_trial}
+#### Two-way ANOVA (ez)
+```{r stats1, echo = FALSE}
+vel_dyn_aov <- ezANOVA(
+  data = df_vel_dyn
+  , dv = .(Velocity)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_dyn_aov$ANOVA$F <- round(vel_dyn_aov$ANOVA$F, 4)
+vel_dyn_aov$ANOVA$p <- round(vel_dyn_aov$ANOVA$p, 4)
+print(vel_dyn_aov$ANOVA)
+```
+
+#### Two-way ANOVA (aov)
+```{r stats2, echo = FALSE}
+vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
+print(summary(vel_dyn_aov_2))
+```
+
+#### Post-hoc (direct comparison)
+```{r post-hoc1, echo = FALSE}
+vel_ttest_forte <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"], paired = TRUE)
+vel_ttest_piano <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"], paired = TRUE)
+# Effect size
+vel_ttest_forte_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Forte"], vel_dyn$Condition[vel_dyn$Subcomponent == "Forte"], paired = TRUE)
+vel_ttest_piano_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Piano"], vel_dyn$Condition[vel_dyn$Subcomponent == "Piano"], paired = TRUE)
+
+print(vel_ttest_forte)
+print(vel_ttest_piano)
+print(vel_ttest_forte_cohend)
+print(vel_ttest_piano_cohend)
+```
+
+#### Post-hoc (pairwise t test)
+```{r post-hoc2, echo = FALSE}
+# Create a group label for a post-hoc test
+vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"] <- "Performing-Forte" 
+vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"] <- "Performing-Piano"
+vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"] <- "Teaching-Forte"
+vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"] <- "Teaching-Piano"
+# pairwise comparison
+vel_dyn_pairwise <- pairwise.t.test(vel_dyn$Mean, vel_dyn$Group, paired = TRUE, p.adjust.method = "BH")
+vel_dyn_pairwise$p.value <- round(vel_dyn_pairwise$p.value, 4)
+print(vel_dyn_pairwise)
+```
+
+## 3. Trial by trial - individual{#vel_dyn_ind_trial}
 - Average KV for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
@@
 
-### 4. Trial by trial - group{#vel_dyn_group_trial}
+## 4. Trial by trial - group{#vel_dyn_group_trial}
 - Average KV for each trial for each condition.
@@
 
-## Key Velocity Difference (KV-Diff) - Dynamics
-### 1. All - individual{#vel_acc_dyn_ind_all}
+# Key Velocity Difference (Difference) - Dynamics
+## 1. All - individual{#vel_acc_dyn_ind_all}
 - Average KV-Diff for each participant.
@@
 
-### 2. All - group{#vel_acc_dyn_group_all}
+## 2. All - group{#vel_acc_dyn_group_all}
 - Average KV-Diff for each condition.
-- Hypothesis: Participants will make a larger contrast between forte and piano.
 
@@
 
+### Plot
+#### Bar chart
 ```{r vel_acc_dyn_all_bar,  echo = FALSE}
@@
 
-##### Box plot
-
+#### Box plot
 ```{r vel_acc_dyn_all_box, echo = FALSE}
@@
 
-#### Statistics (TBC)
+### Statistics
+#### Normality check
+```{r normality2, echo = FALSE}
+# Reduce unused factors
+vel_acc_dyn$Subcomponent <- factor(vel_acc_dyn$Subcomponent)
+vel_acc_dyn_norm <- by(vel_acc_dyn$Mean, list(vel_acc_dyn$Condition, vel_acc_dyn$Subcomponent), shapiro.test)
+print(vel_acc_dyn_norm)
+```
 
-### 3. Trial by trial - individual{#vel_acc_dyn_ind_trial}
+#### Two-way ANOVA (ez)
+```{r stats3, echo = FALSE}
+vel_acc_dyn_aov <- ezANOVA(
+  data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF")
+  , dv = .(Acc)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_acc_dyn_aov$ANOVA$F <- round(vel_acc_dyn_aov$ANOVA$F, 4)
+vel_acc_dyn_aov$ANOVA$p <- round(vel_acc_dyn_aov$ANOVA$p, 4)
+print(vel_acc_dyn_aov$ANOVA)
+```
+
+#### Two-way ANOVA (aov)
+```{r stats4, echo = FALSE}
+vel_acc_dyn_aov_2 <- with(vel_acc_dyn,
+                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
+print(summary(vel_acc_dyn_aov_2))
+```
+
+#### Post-hoc (direct comparison)
+```{r post-hoc3, echo = FALSE}
+vel_acc_ttest_ftop <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "FtoP"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "FtoP"], paired = TRUE)
+vel_acc_ttest_ptof <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)
+# Effect size
+vel_acc_ttest_ftop_cohend <- cohen.d(vel_acc_dyn$Mean[vel_acc_dyn$Subcomponent == "FtoP"], vel_acc_dyn$Condition[vel_acc_dyn$Subcomponent == "FtoP"], paired = TRUE)
+vel_acc_ttest_ptof_cohend <- cohen.d(vel_acc_dyn$Mean[vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Condition[vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)
+
+print(vel_acc_ttest_ftop)
+print(vel_acc_ttest_ptof)
+print(vel_acc_ttest_ftop_cohend)
+print(vel_acc_ttest_ptof_cohend)
+```
+
+#### Post-hoc (pairwise t test)
+```{r post-hoc4, echo = FALSE}
+# Create a group label for a post-hoc test
+vel_acc_dyn$Group[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "FtoP"] <- "Performing-FtoP" 
+vel_acc_dyn$Group[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"] <- "Performing-PtoF"
+vel_acc_dyn$Group[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "FtoP"] <- "Teaching-FtoP"
+vel_acc_dyn$Group[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"] <- "Teaching-PtoF"
+# pairwise comparison
+vel_acc_dyn_pairwise <- pairwise.t.test(vel_acc_dyn$Mean, vel_acc_dyn$Group, paired = TRUE, p.adjust.method = "BH")
+vel_acc_dyn_pairwise$p.value <- round(vel_acc_dyn_pairwise$p.value, 4)
+print(vel_acc_dyn_pairwise)
+```
+
+## 3. Trial by trial - individual{#vel_acc_dyn_ind_trial}
 - Average KV-Diff for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r vel_acc_dyn_trial_line_ftop, echo = FALSE, fig.height = 8}
@@
 
-### 4. Trial by trial - group{#vel_acc_dyn_group_trial}
+## 4. Trial by trial - group{#vel_acc_dyn_group_trial}
 - Average KV-Diff for each trial for each condition.
@@
 
-###########
-###########
-
-## Key Velocity (KV) - Articulation
-### 1. All - individual{#vel_art_ind_all}
+# Key Velocity (KV) - Articulation
+## 1. All - individual{#vel_art_ind_all}
 - Average KV for each participant.
@@
 
-### 2. All - group{#vel_art_group_all}
+## 2. All - group{#vel_art_group_all}
 - Average KV for each condition.
@@
 
-##### Box plot
-
+#### Box plot
 ```{r vel_art_all_box, echo = FALSE}
@@
 
-#### Statistics (TBC)
+### Statistics
+#### Normality check
+```{r normality3, echo = FALSE}
+# Reduce unused factors
+vel_art$Subcomponent <- factor(vel_art$Subcomponent)
+vel_art_norm <- by(vel_art$Mean, list(vel_art$Condition, vel_art$Subcomponent), shapiro.test)
+print(vel_art_norm)
+```
 
-### 3. Trial by trial - individual{#vel_art_ind_trial}
+#### Two-way ANOVA (ez)
+```{r stats5, echo = FALSE}
+vel_art_aov <- ezANOVA(
+  data = df_vel_art
+  , dv = .(Velocity)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_art_aov$ANOVA$F <- round(vel_art_aov$ANOVA$F, 4)
+vel_art_aov$ANOVA$p <- round(vel_art_aov$ANOVA$p, 4)
+print(vel_art_aov$ANOVA)
+```
+
+#### Two-way ANOVA (aov)
+```{r stats6, echo = FALSE}
+vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
+print(summary(vel_art_aov_2))
+```
+
+#### Post-hoc (direct comparison)
+```{r post-hoc5, echo = FALSE}
+vel_ttest_legato <- t.test(vel_art$Mean[vel_art$Condition == "performing" & vel_art$Subcomponent == "Legato"], vel_art$Mean[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Legato"], paired = TRUE)
+vel_ttest_staccato <- t.test(vel_art$Mean[vel_art$Condition == "performing" & vel_art$Subcomponent == "Staccato"], vel_art$Mean[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Staccato"], paired = TRUE)
+# Effect size
+vel_ttest_legato_cohend <- cohen.d(vel_art$Mean[vel_art$Subcomponent == "Legato"], vel_art$Condition[vel_art$Subcomponent == "Legato"], paired = TRUE)
+vel_ttest_staccato_cohend <- cohen.d(vel_art$Mean[vel_art$Subcomponent == "Staccato"], vel_art$Condition[vel_art$Subcomponent == "Staccato"], paired = TRUE)
+
+print(vel_ttest_legato)
+print(vel_ttest_staccato)
+print(vel_ttest_legato_cohend)
+print(vel_ttest_staccato_cohend)
+```
+
+#### Post-hoc (pairwise t test)
+```{r post-hoc6, echo = FALSE}
+# Create a group label for a post-hoc test
+vel_art$Group[vel_art$Condition == "performing" & vel_art$Subcomponent == "Legato"] <- "Performing-Legato" 
+vel_art$Group[vel_art$Condition == "performing" & vel_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
+vel_art$Group[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Legato"] <- "Teaching-Legato"
+vel_art$Group[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
+# pairwise comparison
+vel_art_pairwise <- pairwise.t.test(vel_art$Mean, vel_art$Group, paired = TRUE, p.adjust.method = "BH")
+vel_art_pairwise$p.value <- round(vel_art_pairwise$p.value, 4)
+print(vel_art_pairwise)
+```
+
+## 3. Trial by trial - individual{#vel_art_ind_trial}
 - Average KV for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 8}
@@
 
-### 4. Trial by trial - group{#vel_art_group_trial}
+## 4. Trial by trial - group{#vel_art_group_trial}
 - Average KV for each trial for each condition.
@@
 
-## Key Velocity Difference (KV-Diff) - Articulation
-### 1. All - individual{#vel_acc_art_ind_all}
+# Key Velocity Difference (Difference) - Articulation
+## 1. All - individual{#vel_acc_art_ind_all}
 - Average KV-Diff for each participant.
@@
 
-### 2. All - group{#vel_acc_art_group_all}
+## 2. All - group{#vel_acc_art_group_all}
 - Average KV-Diff for each condition.
@@
 
-##### Bar chart
-
+### Plot
+#### Bar chart
 ```{r vel_acc_art_all_bar,  echo = FALSE}
@@
 
-##### Box plot
-
+#### Box plot
 ```{r vel_acc_art_all_box, echo = FALSE}
@@
 
-##### Statistics
+### Statistics
+#### Normality check
+```{r normality4, echo = FALSE}
+# Reduce unused factors
+vel_acc_art$Subcomponent <- factor(vel_acc_art$Subcomponent)
+vel_acc_art_norm <- by(vel_acc_art$Mean, list(vel_acc_art$Condition, vel_acc_art$Subcomponent), shapiro.test)
+print(vel_acc_art_norm)
+```
 
-### 3. Trial by trial - individual{#kot_dyn_ind_trial}
+#### Two-way ANOVA (ez)
+```{r stats7, echo = FALSE}
+vel_acc_art_aov <- ezANOVA(
+  data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL")
+  , dv = .(Acc)
+  , wid = .(SubNr)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_acc_art_aov$ANOVA$F <- round(vel_acc_art_aov$ANOVA$F, 4)
+vel_acc_art_aov$ANOVA$p <- round(vel_acc_art_aov$ANOVA$p, 4)
+print(vel_acc_art_aov$ANOVA)
+```
+
+#### Post-hoc (direct comparison)
+```{r post-hoc7, echo = FALSE}
+vel_acc_ttest_ltos <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "LtoS"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "LtoS"], paired = TRUE)
+vel_acc_ttest_stol <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "StoL"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "StoL"], paired = TRUE)
+# Effect size
+vel_acc_ttest_ltos_cohend <- cohen.d(vel_acc_art$Mean[vel_acc_art$Subcomponent == "LtoS"], vel_acc_art$Condition[vel_acc_art$Subcomponent == "LtoS"], paired = TRUE)
+vel_acc_ttest_stol_cohend <- cohen.d(vel_acc_art$Mean[vel_acc_art$Subcomponent == "StoL"], vel_acc_art$Condition[vel_acc_art$Subcomponent == "StoL"], paired = TRUE)
+
+print(vel_acc_ttest_ltos)
+print(vel_acc_ttest_stol)
+print(vel_acc_ttest_ltos_cohend)
+print(vel_acc_ttest_stol_cohend)
+```
+
+#### Post-hoc (pairwise t test)
+```{r post-hoc8, echo = FALSE}
+# Create a group label for a post-hoc test
+vel_acc_art$Group[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "LtoS"] <- "Performing-LtoS" 
+vel_acc_art$Group[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "StoL"] <- "Performing-StoL"
+vel_acc_art$Group[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "LtoS"] <- "Teaching-LtoS"
+vel_acc_art$Group[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "StoL"] <- "Teaching-StoL"
+# pairwise comparison
+vel_acc_art_pairwise <- pairwise.t.test(vel_acc_art$Mean, vel_acc_art$Group, paired = TRUE, p.adjust.method = "BH")
+vel_acc_art_pairwise$p.value <- round(vel_acc_art_pairwise$p.value, 4)
+print(vel_acc_art_pairwise)
+```
+
+#### Two-way ANOVA (aov)
+```{r stats8, echo = FALSE}
+vel_acc_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_acc_art)
+print(summary(vel_acc_art_aov_2))
+```
+
+## 3. Trial by trial - individual{#vel_acc_art_ind_trial}
 - Average KV-Diff for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r vel_acc_art_trial_line_ltos, echo = FALSE, fig.height = 8}
@@
 
-### 4. Trial by trial - group{#vel_acc_art_group_trial}
+## 4. Trial by trial - group{#vel_acc_art_group_trial}
 - Average KV-Diff for each trial for each condition.
@@
 
-## Sequence plot
-#### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 15, fig.height = 5, echo = FALSE}
+# KV-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#vel_dyn_order_group_all}
+```{r vel_dyn_order_all, echo = all}
+# For each participant
+vel_dyn_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_dyn,
+                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_dyn_order <- cbind(vel_dyn_order[,1:5], as.data.frame(vel_dyn_order[,6]))
+# Change colnames
+colnames(vel_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+vel_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_dyn_order,
+                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_dyn_order_all <- cbind(vel_dyn_order_all[,1:4], as.data.frame(vel_dyn_order_all[,5]))
+# Change colnames
+colnames(vel_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(vel_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r vel_dyn_order_all_bar, echo = FALSE}
+p_vel_dyn_order_all <- ggplot(data = vel_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
+p_vel_dyn_order_all
+```
+
+#### Box plot
+```{r vel_dyn_order_all_box, echo = FALSE}
+p_vel_dyn_order_all_box <- ggplot(data = vel_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
+p_vel_dyn_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality5, echo = FALSE}
+# Normality check
+vel_dyn_order$Subcomponent <- factor(vel_dyn_order$Subcomponent)
+vel_dyn_order_norm <- by(vel_dyn_order$Mean, list(vel_dyn_order$Condition, vel_dyn_order$Subcomponent, vel_dyn_order$Order), shapiro.test)
+print(vel_dyn_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats9, echo = FALSE}
+vel_dyn_order_aov <- ezANOVA(
+  data = df_vel_dyn
+  , dv = .(Velocity)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_dyn_order_aov$ANOVA$F <- round(vel_dyn_order_aov$ANOVA$F, 4)
+vel_dyn_order_aov$ANOVA$p <- round(vel_dyn_order_aov$ANOVA$p, 4)
+print(vel_dyn_order_aov)
+```
+
+# KV-Diff-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#vel_acc_dyn_order_group_all}
+```{r vel_acc_dyn_order_all, echo = all}
+# For each participant
+vel_acc_dyn_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
+                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_acc_dyn_order <- cbind(vel_acc_dyn_order[,1:5], as.data.frame(vel_acc_dyn_order[,6]))
+# Change colnames
+colnames(vel_acc_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+vel_acc_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_dyn_order,
+                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_acc_dyn_order_all <- cbind(vel_acc_dyn_order_all[,1:4], as.data.frame(vel_acc_dyn_order_all[,5]))
+# Change colnames
+colnames(vel_acc_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(vel_acc_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r vel_acc_dyn_order_all_bar, echo = FALSE}
+p_vel_acc_dyn_order_all <- ggplot(data = vel_acc_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
+p_vel_acc_dyn_order_all
+```
+
+#### Box plot
+```{r vel_acc_dyn_order_all_box, echo = FALSE}
+p_vel_acc_dyn_order_all_box <- ggplot(data = vel_acc_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "Difference", subtitle = "KV-Diff: Dynamics") 
+p_vel_acc_dyn_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality6, echo = FALSE}
+# Normality check
+vel_acc_dyn_order$Subcomponent <- factor(vel_acc_dyn_order$Subcomponent)
+vel_acc_dyn_order_norm <- by(vel_acc_dyn_order$Mean, list(vel_acc_dyn_order$Condition, vel_acc_dyn_order$Subcomponent, vel_acc_dyn_order$Order), shapiro.test)
+print(vel_acc_dyn_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats10, echo = FALSE}
+vel_acc_dyn_order_aov <- ezANOVA(
+  data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF")
+  , dv = .(Acc)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_acc_dyn_order_aov$ANOVA$F <- round(vel_acc_dyn_order_aov$ANOVA$F, 4)
+vel_acc_dyn_order_aov$ANOVA$p <- round(vel_acc_dyn_order_aov$ANOVA$p, 4)
+print(vel_acc_dyn_order_aov)
+```
+
+# KV-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#vel_art_order_group_all}
+```{r vel_art_order_all, echo = all}
+# For each participant
+vel_art_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_art,
+                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_art_order <- cbind(vel_art_order[,1:5], as.data.frame(vel_art_order[,6]))
+# Change colnames
+colnames(vel_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+vel_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_art_order,
+                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_art_order_all <- cbind(vel_art_order_all[,1:4], as.data.frame(vel_art_order_all[,5]))
+# Change colnames
+colnames(vel_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(vel_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r vel_art_order_all_bar, echo = FALSE}
+p_vel_art_order_all <- ggplot(data = vel_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
+p_vel_art_order_all
+```
+
+#### Box plot
+```{r vel_art_order_all_box, echo = FALSE}
+p_vel_art_order_all_box <- ggplot(data = vel_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KV (0-127)", subtitle = "KV: Articulation") 
+p_vel_art_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality7, echo = FALSE}
+# Normality check
+vel_art_order$Subcomponent <- factor(vel_art_order$Subcomponent)
+vel_art_order_norm <- by(vel_art_order$Mean, list(vel_art_order$Condition, vel_art_order$Subcomponent, vel_art_order$Order), shapiro.test)
+print(vel_art_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats11, echo = FALSE}
+vel_art_order_aov <- ezANOVA(
+  data = df_vel_art
+  , dv = .(Velocity)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_art_order_aov$ANOVA$F <- round(vel_art_order_aov$ANOVA$F, 4)
+vel_art_order_aov$ANOVA$p <- round(vel_art_order_aov$ANOVA$p, 4)
+print(vel_art_order_aov)
+```
+
+# KV-Diff-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#vel_acc_art_order_group_all}
+```{r vel_acc_art_order_all, echo = all}
+# For each participant
+vel_acc_art_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
+                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
+vel_acc_art_order <- cbind(vel_acc_art_order[,1:5], as.data.frame(vel_acc_art_order[,6]))
+# Change colnames
+colnames(vel_acc_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+vel_acc_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_art_order,
+                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+vel_acc_art_order_all <- cbind(vel_acc_art_order_all[,1:4], as.data.frame(vel_acc_art_order_all[,5]))
+# Change colnames
+colnames(vel_acc_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(vel_acc_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r vel_acc_art_order_all_bar, echo = FALSE}
+p_vel_acc_art_order_all <- ggplot(data = vel_acc_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
+p_vel_acc_art_order_all
+```
+
+#### Box plot
+```{r vel_acc_art_order_all_box, echo = FALSE}
+p_vel_acc_art_order_all_box <- ggplot(data = vel_acc_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "Difference", subtitle = "KV: Articulation") 
+p_vel_acc_art_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality8, echo = FALSE}
+# Normality check
+vel_acc_art_order$Subcomponent <- factor(vel_acc_art_order$Subcomponent)
+vel_acc_art_order_norm <- by(vel_acc_art_order$Mean, list(vel_acc_art_order$Condition, vel_acc_art_order$Subcomponent, vel_acc_art_order$Order), shapiro.test)
+print(vel_acc_art_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats12, echo = FALSE}
+vel_acc_art_order_aov <- ezANOVA(
+  data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL")
+  , dv = .(Acc)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+vel_acc_art_order_aov$ANOVA$F <- round(vel_acc_art_order_aov$ANOVA$F, 4)
+vel_acc_art_order_aov$ANOVA$p <- round(vel_acc_art_order_aov$ANOVA$p, 4)
+print(vel_acc_art_order_aov)
+```
+
+# Sequence plot
+### Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 20, fig.height = 5, echo = FALSE}
 # Average vel_dynocity for each note
 # For each participant
-vel_dyn_seq <- aggregate(Velocity~SubNr*Condition*Skill*NoteNr, data = df_vel_dyn, 
+vel_dyn_seq <- aggregate(Velocity~SubNr*Condition*Skill*Note, data = df_vel_dyn, 
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-#### Articulation{#sequence_art}
-```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
+### Articulation{#sequence_art}
+```{r seq_art, fig.width = 20, fig.height = 5, echo = FALSE}
 # Average vel_artocity for each note
 # For each participant
-vel_art_seq <- aggregate(Velocity~SubNr*Condition*Skill*NoteNr, data = df_vel_art, 
+vel_art_seq <- aggregate(Velocity~SubNr*Condition*Skill*Note, data = df_vel_art, 
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})

### 2020-02-09 script/RNotebook/questionnaire.Rmd
update questionnaire
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: 'Analysis: The Sound of Teaching Music 1'
+title: 'Demographics: The Sound of Teaching Music 1'
 output:
@@
 
-##### Description: This is a summary of demographic information
+Description: This is a summary of demographic information
+
+```{r setup, include = FALSE}
+# packages
+# data manipulation
+if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
+```
 
@@
 # Read csv files
-df <- read.csv("./questionnaire.csv", header = T, sep = ",", dec = ".")
+df <- read.csv("./questionnaire.csv", header = T, sep = ",", dec = ".", na.string = "NA")
 df_exc <- read.csv("./filtered/data_errorRate.csv", header = T, sep = ",", dec = ".")
@@
 # Change colnames
-colnames(df) <- c("TimeStamp", "SubNr", "Age", "Gender", "FirstLang", "SecondLang", "DominantHand", "NormalVision", "NormalHearing", "Desc1", "Desc2",
-                  "Instruments1", "FormalTraining1", "TotalPractice", "Start1", "End1", "PracticeHours", # Experience in musical instruments
+colnames(df) <- c("TimeStamp", "SubNr", "Age", "Gender", "FirstLang", "SecondLang", "DominantHand", "NormalVision", "NormalHearing", "Desc1", "Desc2", "Instruments1", "FormalTraining1", "TotalPractice", "Start1", "End1", "PracticeHours", # Experience in musical instruments
                   "Teaching", "Instruments2", "TotalTeaching1", "FormalTraining2", "Training", "AgeGroup", "Start2", "End2", "Like1", # Experience in teaching instruments
@@
 
-### Descriptive stats for participants' demographics
-#### Number of excluded participants
+## Descriptive stats for participants' demographics
+### Number of excluded participants
+
 ```{r excluded, echo = FALSE}
@@
 
-#### Number of included participants
+### Number of included participants
 ```{r included, echo = FALSE}
@@
 
-#### Error Rate of excluded participants
+### Error Rate of excluded participants
 ```{r errorRate_excluded, echo = FALSE}
 # Information about excluded participants due to errorRate (SubNr 16 was excluded but due to not errorRate but tempo deviation)
+# Error Rate = the number of trials which contain any mistake (Not the number of mistakes in a trial)
 errorRate_mean_exc <- mean(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
@@
 
-#### Error Rate: Normality test for excluded participants
+### Error Rate: Normality test for excluded participants
 ```{r normality_for_excluded, echo = FALSE}
@@
 
-#### Error Rate of included participants
+### Error Rate of included participants
 ```{r er_included, echo = FALSE}
 # Information about included participants
+# Error Rate = the number of trials which contain any mistake (Not the number of mistakes in a trial)
 errorRate_mean_inc <- mean(df_exc$ErrorRate[df_exc$Final == "include"])*100
@@
 
-#### Error Rate: Normality test for included participants
+### Error Rate: Normality test for included participants
 ```{r normality_for_included, echo = FALSE}

### 2020-02-10 script/RNotebook/questionnaire.Rmd
correct typo
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ### Error Rate of excluded participants
-```{r errorRate_excluded, echo = FALSE}
+```{r er_excluded, echo = FALSE}
 # Information about excluded participants due to errorRate (SubNr 16 was excluded but due to not errorRate but tempo deviation)

### 2020-02-23 script/RNotebook/ioi_1.Rmd
change fig.width/fig.height
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-01-06 script/RNotebook/ioi_1.Rmd
@@
 #### Articulation{#sequence_art}
-```{r seq_art, fig.width = 15, fig.height = 5, echo = FALSE}
+```{r seq_art, fig.width = 20, fig.height = 5, echo = FALSE}
 # For each individual
@@
 #### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 15, fig.height = 5, echo = FALSE}
+```{r seq_dyn, fig.width = 20, fig.height = 5, echo = FALSE}
 # For each individual

### 2020-03-19 script/RNotebook/ioi_1.Rmd
updated comments/check variability without SubNr 15
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-01-12 script/RNotebook/ioi_1.Rmd
@@
 
-##### Description: This is a summary of all analyses for a tempo (IOIs).
+Description: This is a summary of all analyses for a tempo (IOIs).
   - data <- data_ioi_1.csv
+
+- Last checked: 19/03/20
+
+# Outline - Interonset intervals (IOIs)
+## [Experimental Design](#design)
+
+## [Articulation](#ioi-articulation)
+
+- Individual results
+  + [All](#art_ind_all)
+  + [Phrase bondaries](#art_ind_phrase)
+  + [Trial by trial](#art_ind_trial)
+  + [Variability](#art_ind_var_all)
+  + [Variability trial-by-trial](#art_ind_var_trial)
+
+- Group results and stats (if available)
+  + [All](#art_group_all)
+  + [Phrase bondaries](#art_group_phrase)
+  + [Trial by trial](#art_group_trial)
+  + [Variability](#art_group_var_all)
+  + [Variability trial-by-trial](#art_group_var_trial)
+
+- [Sequence plot](#sequence_art)
+
+## [Dynamics](#ioi-dynamics)
+
+- Individual results
+  + [All](#dyn_ind_all)
+  + [Phrase bondaries](#dyn_ind_phrase)
+  + [Trial by trial](#dyn_ind_trial)
+  + [Variability all](#dyn_ind_var_all)
+  + [Variability trial-by-trial](#dyn_ind_var_trial)
+
+- Group results and stats (if available)
+  + [All](#dyn_group_all)
+  + [Phrase bondaries](#dyn_group_phrase)
+  + [Trial by trial](#dyn_group_trial)
+  + [Variability all](#dyn_group_var_all)
+  + [Variability trial-by-trial](#dyn_group_var_trial)
+
+- [Sequence plot](#sequence_dyn)
 
@@
 
-### Main predictions
+## Main predictions
 Participants will play slower while teaching regardless of the skills.
@@
 df_ioi_dyn <- df_ioi %>% dplyr::filter(Skill == "dynamics")
-
 ```
 
-## IOI-Articulation 
-### 1. All - individual{#art_ind_all}
+# IOI-Articulation 
+## 1. All - individual{#art_ind_all}
 - Average normIOIs for each participant.
@@
 
-### 2. All - group{#art_group_all}
+## 2. All - group{#art_group_all}
 - Average normIOIs for each condition.
@@
 
-##### Box plot
+#### Box plot
 - A dashed line represents the ideal tempo
@@
 
-### 3. Subcomponent change - individual{#art_ind_phrase}
+## 3. Subcomponent change - individual{#art_ind_phrase}
 - Compare normIOIs at which Subcomponent changes (e.g., Legato to Staccato) with others fir each participant.
@@
 
-### 4. Subcomponent change - group{#art_group_phrase}
+## 4. Subcomponent change - group{#art_group_phrase}
 - Compare IOIs at which Subcomponent changes (e.g., Legato to Staccato) with others in the two conditions.
@@
 
-#### Plot
+### Plot
+#### Bar plot
 - Error bars show SEM of the mean
@@
 
-### 5. Trial by trial - individual{#art_ind_trial}
+## 5. Trial by trial - individual{#art_ind_trial}
 - Average normIOIs for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 - A dashed line represents the ideal tempo
@@
 
-### 6. Trial by trial - group{#art_group_trial}
+## 6. Trial by trial - group{#art_group_trial}
 - Average normIOIs for each trial for each condition.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 - Error bars show SEM of the mean
+- A dashed line represents the ideal tempo
 
@@
 
-#### Plot
-##### Bar graph
+### Plot
+#### Bar graph
 - Error bars show SD of the mean
-- A dashed line represents the tempo given by a metronome (80bpm)
 
@@
 
-### 9. Variability trial-by-trial - individual{#art_ind_var_trial}
+### *Except SubNr 15
+```{r ioi_art_var_all_15, echo = FALSE}
+# Group mean
+ioi_art_var_all_15 <- aggregate(Mean~Condition*Skill, data = subset(ioi_art_var, ioi_art_var$SubNr != 15),
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_art_var_all_15 <- cbind(ioi_art_var_all_15[,1:2], as.data.frame(ioi_art_var_all_15[,3]))
+# Change colnames
+colnames(ioi_art_var_all_15) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
+print(ioi_art_var_all_15)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r ioi_art_var_all_bar_15, echo = FALSE}
+p_ioi_art_var_all_15 <- ggplot(data = ioi_art_var_all_15, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge()) +
+  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
+  theme(legend.position = "None")
+p_ioi_art_var_all_15
+```
+
+#### Box plot
+```{r ioi_art_var_all_box_15, echo = FALSE}
+p_ioi_art_var_box_15 <- ggplot(data = subset(ioi_art_var, ioi_art_var$SubNr != 15), aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
+  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
+  theme(legend.position = "None")
+p_ioi_art_var_box_15
+```
+
+### Statistics
+#### Normality check
+```{r normality3_15, echo = FALSE}
+ioi_art_var_norm_15 <- by(subset(ioi_art_var, ioi_art_var$SubNr != 15)$Mean, list(subset(ioi_art_var, ioi_art_var$SubNr != 15)$Condition), shapiro.test)
+print(ioi_art_var_norm_15)
+```
+
+#### Paired t-test
+```{r stats5_15, echo = FALSE}
+ioi_art_var_ttest_15 <- t.test(subset(ioi_art_var, ioi_art_var$SubNr != 15)$Mean[subset(ioi_art_var, ioi_art_var$SubNr != 15)$Condition == "teaching"], subset(ioi_art_var, ioi_art_var$SubNr != 15)$Mean[subset(ioi_art_var, ioi_art_var$SubNr != 15)$Condition == "performing"], paired = TRUE)
+print(ioi_art_var_ttest_15)
+```
+
+#### Effect size
+```{r stats6_15, echo = FALSE}
+ioi_art_var_cohend_15 <- cohen.d(subset(ioi_art_var, ioi_art_var$SubNr != 15)$Mean, subset(ioi_art_var, ioi_art_var$SubNr != 15)$Condition, paired = TRUE)
+print(ioi_art_var_cohend_15)
+```
+
+## 9. Variability trial-by-trial - individual{#art_ind_var_trial}
 - Average variability for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 4}
@@
 
-## IOI-Dynamics
-### 1. All - individual{#dyn_ind_all}
+# IOI-Dynamics
+## 1. All - individual{#dyn_ind_all}
 - Average normIOIs for each participant.
@@
 
-### 2. All - group{#dyn_group_all}
+## 2. All - group{#dyn_group_all}
 - Average normIOIs for each condition.
@@
 
-##### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm)
+#### Box plot
+- A dashed line represents the ideal tempo
 
@@
 
+#### Paired t-test
 ```{r stats7, echo = FALSE}
@@
 
-### 3. Subcomponent changes - individual{#dyn_ind_phrase}
+## 3. Subcomponent changes - individual{#dyn_ind_phrase}
 - Compare normIOIs at which Subcomponent changes (e.g., Forte to Piano) with others fir each participant.
@@
 
-### 4. Subcomponent changes - group{#dyn_group_phrase}
+## 4. Subcomponent changes - group{#dyn_group_phrase}
 - Compare IOIs at which Subcomponent changes with others in the two conditions.
@@
 
-### 5. Trial by trial - individual{#dyn_ind_trial}
+## 5. Trial by trial - individual{#dyn_ind_trial}
 - Average normIOIs for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 - A dashed line represents the ideal tempo.
@@
 
-### 6. Trial by trial - group{#dyn_group_trial}
+## 6. Trial by trial - group{#dyn_group_trial}
 - Average normIOIs for each trial for each condition.
@@
 
+#### Paired t-test
 ```{r stats11, echo = FALSE}
@@
 
+#### Effect size
 ```{r stats12, echo = FALSE}
@@
 
-### 9. Variability trial-by-trial - individual{#dyn_ind_var_trial}
+### * Except SubNr 15
+```{r ioi_dyn_var_all_15, echo = FALSE}
+# Group mean
+ioi_dyn_var_all_15 <- aggregate(Mean~Condition*Skill, data = subset(ioi_dyn_var, ioi_dyn_var$SubNr != 15),
+                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+ioi_dyn_var_all_15 <- cbind(ioi_dyn_var_all_15[,1:2], as.data.frame(ioi_dyn_var_all_15[,3]))
+# Change colnames
+colnames(ioi_dyn_var_all_15) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
+print(ioi_dyn_var_all_15)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r ioi_dyn_var_all_bar_15, echo = FALSE}
+p_ioi_dyn_var_all_15 <- ggplot(data = ioi_dyn_var_all_15, aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge()) +
+  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
+  theme(legend.position = "None")
+p_ioi_dyn_var_all_15
+```
+
+#### Box plot
+```{r ioi_dyn_var_all_box_15, echo = FALSE}
+p_ioi_dyn_var_box_15 <- ggplot(data = subset(ioi_dyn_var, ioi_dyn_var$SubNr != 15), aes(x = Condition, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
+  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
+  theme(legend.position = "None")
+p_ioi_dyn_var_box_15
+```
+
+### Statistics
+#### Normality check
+```{r normality6_15, echo = FALSE}
+ioi_dyn_var_norm_15 <- by(subset(ioi_dyn_var, ioi_dyn_var$SubNr != 15)$Mean, list(subset(ioi_dyn_var, ioi_dyn_var$SubNr != 15)$Condition), shapiro.test)
+print(ioi_dyn_var_norm_15)
+```
+  - Violate the normality assumption
+
+#### Paired t-test
+```{r stats11_15, echo = FALSE}
+ioi_dyn_var_ttest_15 <- t.test(subset(ioi_dyn_var, ioi_dyn_var$SubNr != 15)$Mean[subset(ioi_dyn_var, ioi_dyn_var$SubNr != 15)$Condition == "teaching"], subset(ioi_dyn_var, ioi_dyn_var$SubNr != 15)$Mean[subset(ioi_dyn_var, ioi_dyn_var$SubNr != 15)$Condition == "performing"], paired = TRUE)
+print(ioi_dyn_var_ttest_15)
+```
+
+#### Effect size
+```{r stats12_15, echo = FALSE}
+ioi_dyn_var_cohend_15 <- cohen.d(subset(ioi_dyn_var, ioi_dyn_var$SubNr != 15)$Mean, subset(ioi_dyn_var, ioi_dyn_var$SubNr != 15)$Condition, paired = TRUE)
+print(ioi_dyn_var_cohend_15)
+```
+
+## 9. Variability trial-by-trial - individual{#dyn_ind_var_trial}
 - Average variability for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 4}
@@
 ## Sequence plot
-#### Articulation{#sequence_art}
-```{r seq_art, fig.width = 6, fig.height = 2, echo = FALSE}
+### Articulation{#sequence_art}
+```{r seq_art, fig.width = 10, fig.height = 4, echo = FALSE}
 # For each individual
@@
 
-#### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 6, fig.height = 2, echo = FALSE}
+### Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 10, fig.height = 4, echo = FALSE}
 # For each individual

### 2020-03-20 script/RNotebook/ioi_1.Rmd
?
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ### Articulation{#sequence_art}
-```{r seq_art, fig.width = 10, fig.height = 4, echo = FALSE}
+```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
 # For each individual
@@
 ### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 10, fig.height = 4, echo = FALSE}
+```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
 # For each individual

### 2020-03-20 script/RNotebook/dynamics.Rmd
updated comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-01-06 script/RNotebook/dynamics.Rmd
@@
 ---
-title: 'Dynamics: Analysis: The Sound of Teaching Music 2'
+title: 'Dynamics Analysis: The Sound of Teaching Music 2'
 output:
@@
 
-##### Description: This is a summary of all analyses for dynamics.
+Description: This is a summary of all analyses for dynamics.
+
+- Last checked: 20/03/20
+
+# Outline - Dynamics (Key Velocity)
+## [Experimental Design](#design)
+
+## Dynamics
+
+### [KV](#vel-dynamics)
+- Individual results
+  + [All](#vel_dyn_ind_all)
+  + [Trial by trial](#vel_dyn_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#vel_dyn_group_all)
+  + [Trial by trial](#vel_dyn_group_trial)
+
+### [KV-Diff](#vel-dynamics-diff)
+- Individual results
+  + [All](#vel_diff_dyn_ind_all)
+  + [Trial by trial](#vel_diff_dyn_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#vel_diff_dyn_group_all)
+  + [Trial by trial](#vel_diff_dyn_group_trial)
+
+- [Sequence plot](#sequence_dyn)
+
+## Articulation
+
+### [KV](#vel-articulation)
+- Individual results
+  + [All](#vel_art_ind_all)
+  + [Trial by trial](#vel_art_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#vel_art_group_all)
+  + [Trial by trial](#vel_art_group_trial)
+
+### [KV-Diff](#vel-articulation-diff)
+- Individual results
+  + [All](#vel_diff_art_ind_all)
+  + [Trial by trial](#vel_diff_art_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#vel_diff_art_group_all)
+  + [Trial by trial](#vel_diff_art_group_trial)
+
+- [Sequence plot](#sequence_art)
 
@@
 
-### Main predictions
+## Main predictions
 - Participants will exaggerate forte and piano more while teaching.
@@
 
-## Key Velocity (KV) - Dynamics
-### 1. All - individual{#vel_dyn_ind_all}
+# Key Velocity (KV) - Dynamic{#vel-dynamics}
+## 1. All - individual{#vel_dyn_ind_all}
 - Average KV for each participant.
@@
 
-##### Two-way ANOVA (ez)
+#### Two-way ANOVA (ez)
 ```{r stats1, echo = FALSE}
@@
 
-##### Two-way ANOVA (aov)
+#### Two-way ANOVA (aov)
 ```{r stats2, echo = FALSE}
@@
 
-##### Post-hoc (direct comparison)
+#### Post-hoc (direct comparison)
 ```{r post-hoc1, echo = FALSE}
@@
 vel_ttest_piano <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"], paired = TRUE)
+
+print(vel_ttest_forte)
+print(vel_ttest_piano)
+```
+
+#### Effect size
+```{r effect1, echo = FALSE}
 # Effect size
@@
 
-print(vel_ttest_forte)
-print(vel_ttest_piano)
 print(vel_ttest_forte_cohend)
@@
 
-##### Post-hoc (pairwise t test)
+#### Post-hoc (pairwise t test)
 ```{r post-hoc2, echo = FALSE}
@@
 
-### 3. Trial by trial - individual{#vel_dyn_ind_trial}
+## 3. Trial by trial - individual{#vel_dyn_ind_trial}
 - Average KV for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
@@
 
-## Key Velocity Difference (KV-Diff) - Dynamics
-### 1. All - individual{#vel_diff_dyn_ind_all}
+# Key Velocity Difference (KV-Diff) - Dynamics{#vel-dynamics-diff}
+## 1. All - individual{#vel_diff_dyn_ind_all}
 - Average KV-Diff for each participant.
@@
 
-### 2. All - group{#vel_diff_dyn_group_all}
+## 2. All - group{#vel_diff_dyn_group_all}
 - Average KV-Diff for each condition.
@@
 
+### Plot
+#### Bar chart
 ```{r vel_diff_dyn_all_bar,  echo = FALSE}
@@
 
-##### Box plot
-
+#### Box plot
 ```{r vel_diff_dyn_all_box, echo = FALSE}
@@
 
-##### Post-hoc (direct comparison)
+#### Post-hoc (direct comparison)
 ```{r post-hoc3, echo = FALSE}
@@
 vel_diff_ttest_ptof <- t.test(vel_diff_dyn$Mean[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "PtoF"], vel_diff_dyn$Mean[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "PtoF"], paired = TRUE)
+
+print(vel_diff_ttest_ftop)
+print(vel_diff_ttest_ptof)
+```
+
+#### Effect size
+```{r effect2, echo = FALSE}
 # Effect size
@@
 
-print(vel_diff_ttest_ftop)
-print(vel_diff_ttest_ptof)
 print(vel_diff_ttest_ftop_cohend)
@@
 
-##### Post-hoc (pairwise t test)
+#### Post-hoc (pairwise t test)
 ```{r post-hoc4, echo = FALSE}
@@
 
-### 3. Trial by trial - individual{#vel_diff_dyn_ind_trial}
+## 3. Trial by trial - individual{#vel_diff_dyn_ind_trial}
 - Average KV-Diff for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r vel_diff_dyn_trial_line_ftop, echo = FALSE, fig.height = 4}
@@
 
-### 4. Trial by trial - group{#vel_diff_dyn_group_trial}
+## 4. Trial by trial - group{#vel_diff_dyn_group_trial}
 - Average KV-Diff for each trial for each condition.
@@
 
-## Key Velocity (KV) - Articulation
-### 1. All - individual{#vel_art_ind_all}
+# Key Velocity (KV) - Articulation{#vel-articulation}
+## 1. All - individual{#vel_art_ind_all}
 - Average KV for each participant.
@@
 
-##### Two-way ANOVA (ez)
+#### Two-way ANOVA (ez)
 ```{r stats5, echo = FALSE}
@@
 
-##### Two-way ANOVA (aov)
+#### Two-way ANOVA (aov)
 ```{r stats6, echo = FALSE}
@@
 
-### 3. Trial by trial - individual{#vel_art_ind_trial}
+## 3. Trial by trial - individual{#vel_diff_art_ind_all}
 - Average KV for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 4}
@@
 
-## Key Velocity Difference (KV-Diff) - Articulation
-### 1. All - individual{#vel_diff_art_ind_all}
+# Key Velocity Difference (KV-Diff) - Articulation{#vel-articulation-diff}
+## 1. All - individual{#vel_diff_art_ind_all}
 - Average KV-Diff for each participant.
@@
 
-### 2. All - group{#vel_diff_art_group_all}
+## 2. All - group{#vel_diff_art_group_all}
 - Average KV-Diff for each condition.
@@
 
-##### Bar chart
+### Plot
+#### Bar chart
+- Error bars show SEM of the mean
 
@@
 
-##### Box plot
-
+#### Box plot
 ```{r vel_diff_art_all_box, echo = FALSE}
@@
 
-##### Two-way ANOVA (ez)
+#### Two-way ANOVA (ez)
 ```{r stats7, echo = FALSE}
@@
 
-##### Two-way ANOVA (aov)
+#### Two-way ANOVA (aov)
 ```{r stats8, echo = FALSE}
@@
 
-### 3. Trial by trial - individual{#kot_dyn_ind_trial}
+## 3. Trial by trial - individual{#vel_art_ind_trial}
 - Average KV-Diff for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r vel_diff_art_trial_line_ltos, echo = FALSE, fig.height = 4}
@@
 
-### 4. Trial by trial - group{#vel_diff_art_group_trial}
+## 4. Trial by trial - group{#vel_diff_art_group_trial}
 - Average KV-Diff for each trial for each condition.
@@
 
-## Sequence plot
-#### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 6, fig.height = 2, echo = FALSE}
+# Sequence plot
+## Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
 # Average vel_dynocity for each note
 # For each participant
-vel_dyn_seq <- aggregate(Velocity~SubNr*Condition*Skill*NoteNr, data = df_vel_dyn, 
+vel_dyn_seq <- aggregate(Velocity~SubNr*Condition*Skill*RowNr, data = df_vel_dyn, 
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-#### Articulation{#sequence_art}
-```{r seq_art, fig.width = 6, fig.height = 2, echo = FALSE}
+## Articulation{#sequence_art}
+```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
 # Average vel_artocity for each note
 # For each participant
-vel_art_seq <- aggregate(Velocity~SubNr*Condition*Skill*NoteNr, data = df_vel_art, 
+vel_art_seq <- aggregate(Velocity~SubNr*Condition*Skill*RowNr, data = df_vel_art, 
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})

### 2020-03-26 script/RNotebook/articulation.Rmd
Revert "add outlines"
This reverts commit faea39f12894bc71a3609306f95606bb71627f1b.

--- 2020-01-06 script/RNotebook/articulation.Rmd
@@
 ---
-title: 'Analysis: The Sound of Teaching Music 2'
+title: 'Articulation Analysis: The Sound of Teaching Music 2'
 output:

### 2020-03-26 script/RNotebook/dynamics.Rmd
Revert "add outlines"
This reverts commit faea39f12894bc71a3609306f95606bb71627f1b.

--- 2020-01-06 script/RNotebook/dynamics.Rmd
@@
 ##### Line graph
-```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
+```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
 p_vel_dyn_trial_forte <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r vel_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
+```{r vel_dyn_trial_line_piano, echo = FALSE, fig.height = 8}
 p_vel_dyn_trial_piano <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ##### Line graph
-```{r vel_diff_dyn_trial_line_ftop, echo = FALSE, fig.height = 4}
+```{r vel_diff_dyn_trial_line_ftop, echo = FALSE, fig.height = 8}
 p_vel_diff_dyn_trial_ftop <- ggplot(data = subset(vel_diff_dyn_trial, vel_diff_dyn_trial$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r vel_diff_dyn_trial_line_ptof, echo = FALSE, fig.height = 4}
+```{r vel_diff_dyn_trial_line_ptof, echo = FALSE, fig.height = 8}
 p_vel_diff_dyn_trial_ptof <- ggplot(data = subset(vel_diff_dyn_trial, vel_diff_dyn_trial$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ##### Line graph
-```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 4}
+```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 8}
 p_vel_art_trial_legato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r vel_art_trial_line_piano, echo = FALSE, fig.height = 4}
+```{r vel_art_trial_line_piano, echo = FALSE, fig.height = 8}
 p_vel_art_trial_staccato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ##### Line graph
-```{r vel_diff_art_trial_line_ltos, echo = FALSE, fig.height = 4}
+```{r vel_diff_art_trial_line_ltos, echo = FALSE, fig.height = 8}
 p_vel_diff_art_trial_ltos <- ggplot(data = subset(vel_diff_art_trial, vel_diff_art_trial$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r vel_diff_art_trial_line_stol, echo = FALSE, fig.height = 4}
+```{r vel_diff_art_trial_line_stol, echo = FALSE, fig.height = 8}
 p_vel_diff_art_trial_stol <- ggplot(data = subset(vel_diff_art_trial, vel_diff_art_trial$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 6, fig.height = 2, echo = FALSE}
+```{r seq_dyn, fig.width = 20, fig.height = 5, echo = FALSE}
 # Average vel_dynocity for each note
@@
 #### Articulation{#sequence_art}
-```{r seq_art, fig.width = 6, fig.height = 2, echo = FALSE}
+```{r seq_art, fig.width = 20, fig.height = 5, echo = FALSE}
 # Average vel_artocity for each note

### 2020-03-26 script/RNotebook/articulation.Rmd

--- 2020-01-06 script/RNotebook/articulation.Rmd
@@
 
-##### Description: This is a summary of all analyses for articulation.
+Description: This is a summary of all analyses for articulation.
+
+- Last checked: 19/03/20
+
+# Outline - Articulation (KOT, KOR)
+## [Experimental Design](#design)
+
+## Articulation
+
+### [KOT](#kot_articulation)
+- Individual results
+  + [All](#kot_art_ind_all)
+  + [Trial by trial](#kot_art_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#kot_art_group_all)
+  + [Trial by trial](#kot_art_group_trial)
+
+- [Sequence plot](#sequence_art)
+
+### [KOR](#kor-articulation)
+- Individual results
+  + [All](#kor_art_ind_all)
+  + [Trial by trial](#kor_art_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#kor_art_group_all)
+  + [Trial by trial](#kor_art_group_trial)
+
+- [Sequence plot](#sequence_dyn)
+
+## Dynamics
+
+### [KOT](#kot-dynamics)
+- Individual results
+  + [All](#kot_dyn_ind_all)
+  + [Trial by trial](#kot_dyn_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#kot_dyn_group_all)
+  + [Trial by trial](#kot_dyn_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#kot_dyn_order_group_all)
+
+- [Sequence plot](#sequence_dyn)
+
+### [KOR](#kor-dynamics)
+- Individual results
+  + [All](#kor_dyn_ind_all)
+  + [Trial by trial](#kor_dyn_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#kor_dyn_group_all)
+  + [Trial by trial](#kor_dyn_group_trial)
 
@@
 
-## Key-Overlap Time (KOT) - Articulation
-### 1. All - individual{#kot_art_ind_all}
+# Key-Overlap Time (KOT) - Articulation{#kot_articulation}
+## 1. All - individual{#kot_art_ind_all}
 - Average KOT for each participant.
@@
 
-### 3. Trial by trial - individual{#kot_art_ind_trial}
+## 3. Trial by trial - individual{#kot_art_ind_trial}
 - Average KOT for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 4}
@@
 
-## Key-Overlap Ratio (KOR) - Articulation
-### 1. All - individual{#kor_art_ind_all}
+# Key-Overlap Ratio (KOR) - Articulation{#kor-articulation}
+## 1. All - individual{#kor_art_ind_all}
 - Average KOR for each participant.
@@
 
+### Post-hoc test(direct comparison)
 ```{r post-hoc3, echo = FALSE}
@@
 
-### 3. Trial by trial - individual{#kor_art_ind_trial}
+## 3. Trial by trial - individual{#kor_art_ind_trial}
 - Average KOR for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 4}
@@
 
-### 4. Trial by trial - group{#kor_art_group_trial}
-- Average KOR d for each trial for each condition.
+## 4. Trial by trial - group{#kor_art_group_trial}
+- Average KOR for each trial for each condition.
 
@@
 
-## Key-Overlap Time (KOT) - Dynamics
-### 1. All - individual{#kot_dyn_ind_all}
+# Key-Overlap Time (KOT) - Dynamics{#kot-dynamics}
+## 1. All - individual{#kot_dyn_ind_all}
 - Average KOT for each participant.
@@
 
-##### Box plot
-
+#### Box plot
 ```{r kot_dyn_all_box, echo = FALSE}
@@
 
-### 3. Trial by trial - individual{#kot_dyn_ind_trial}
+## 3. Trial by trial - individual{#kot_dyn_ind_trial}
 - Average KOT for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
@@
 
-## Key-Overlap Ratio (KOR) - Dynamics
-### 1. All - individual{#kor_dyn_ind_all}
+# Key-Overlap Ratio (KOR) - Dynamics{#kor-dynamics}
+## 1. All - individual{#kor_dyn_ind_all}
 - Average KOR for each participant.
@@
 
-### 3. Trial by trial - individual{#kor_dyn_ind_trial}
+## 3. Trial by trial - individual{#kor_dyn_ind_trial}
 - Average KOR for each trial for each participant.
@@
 
-#### Plot
-##### Line graph
+### Plot
+#### Line graph
 ```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
@@
 ## Sequence plot
-#### Articulation{#sequence_art}
-```{r seq_art, fig.width = 6, fig.height = 2, echo = FALSE}
+### Articulation{#sequence_art}
+```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
 # For each individual
@@
 #### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 6, fig.height = 2, echo = FALSE}
+```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
 # For each individual

### 2020-03-26 script/RNotebook/ioi_3.Rmd

--- 2020-01-12 script/RNotebook/ioi_3.Rmd
@@
 
-##### Description: This is a summary of all analyses for a tempo (IOIs).
+Description: This is a summary of all analyses for a tempo (IOIs).
   - data <- data_ioi_3.csv
+
+- Last checked: 19/03/20
+
+# Outline - Interonset intervals (IOIs)
+## [Experimental Design](#design)
+
+## [Articulation](#ioi-articulation)
+
+- Individual results
+  + [All](#art_ind_all)
+  + [Phrase bondaries](#art_ind_phrase)
+  + [Trial by trial](#art_ind_trial)
+  + [Variability](#art_ind_var_all)
+  + [Variability trial-by-trial](#art_ind_var_trial)
+
+- Group results and stats (if available)
+  + [All](#art_group_all)
+  + [Phrase bondaries](#art_group_phrase)
+  + [Trial by trial](#art_group_trial)
+  + [Variability](#art_group_var_all)
+  + [Variability trial-by-trial](#art_group_var_trial)
+
+- [Sequence plot](#sequence_art)
+
+## [Dynamics](#ioi-dynamics)
+
+- Individual results
+  + [All](#dyn_ind_all)
+  + [Phrase bondaries](#dyn_ind_phrase)
+  + [Trial by trial](#dyn_ind_trial)
+  + [Variability all](#dyn_ind_var_all)
+  + [Variability trial-by-trial](#dyn_ind_var_trial)
+
+- Group results and stats (if available)
+  + [All](#dyn_group_all)
+  + [Phrase bondaries](#dyn_group_phrase)
+  + [Trial by trial](#dyn_group_trial)
+  + [Variability all](#dyn_group_var_all)
+  + [Variability trial-by-trial](#dyn_group_var_trial)
+
+- [Sequence plot](#sequence_dyn)
 
@@
 
-#### Plot
+### Plot
+#### Bar graph
 - Error bars show SEM of the mean
@@
 #### Plot
-##### Line graph
+#### Line graph
 - A dashed line represents the ideal tempo
@@
 - Error bars show SEM of the mean
+- A dashed line represents the ideal tempo
 
@@
 
-##### Box plot
+#### Box plot
 - A dashed line represents the tempo given by a metronome (80bpm)
@@
 
-#### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 6, fig.height = 2, echo = FALSE}
+### Dynamics{#sequence_dyn}
+```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
 # For each individual

### 2020-03-26 script/RNotebook/articulation.Rmd
correct comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-### Post-hoc test(direct comparison)
+#### Post-hoc test (direct comparison)
 ```{r post-hoc3, echo = FALSE}

### 2020-04-02 script/RNotebook/additional020420/articulation.Rmd
additional analysis for articulation transitions
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Additional analysis for transition points"
output: html_notebook
---
Description: These are some additional analyses for transition points.

## Articulation
How does it look like articulation transition points?
Posthoc prediciton: Participants would have a larger separation betwenn the end of legato and the beginning of staccato.

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("effsize")) {install.packages("effsize"); require("effsize")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# read csv
df_kot <- fread("./../trimmed/data_kot.csv")

# only transition points for arituclation
df_kot_trans <- df_kot[Skill == "articulation" & Subcomponent != "Legato" & Subcomponent != "Staccato"]
```

```{r stats1, echo = TRUE}
# For each individual
art_trans <- df_kot_trans[, .(N = length(KOT), Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Subcomponent)]
print(art_trans)
# Group
art_trans_all <- art_trans[, .(N = length(Mean), Mean = mean(Mean), SEM = sd(Mean)/sqrt(length(Mean))), by = .(Condition, Subcomponent)]
print(art_trans_all)
# Stats
# normality check
art_trans_norm <- by(art_trans$Mean, list(art_trans$Condition, art_trans$Subcomponent), shapiro.test)
print(art_trans_norm)
# ezAnova
art_trans_aov <- ezANOVA(
  data = df_kot_trans
  , dv = .(KOT)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
print(art_trans_aov)
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r art_trans_bar, echo = FALSE}　
p_art_trans <- ggplot(data = art_trans_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation Transition")
p_art_trans
```

#### Box plot
```{r art_trans_box, echo = FALSE}
p_art_trans_box <- ggplot(data = art_trans, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation Transition")
p_art_trans_box
```

### 2020-04-07 script/RNotebook/articulation.Rmd
rerun all analyses
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-03-26 script/RNotebook/articulation.Rmd
@@
 ---
-title: 'Articulation Analysis: The Sound of Teaching Music 2'
+title: 'Articulation Analysis: The Sound of Teaching Music 1'
 output:
@@
 
-- Last checked: 19/03/20
+- Last checked: 07/02/20
 
@@
 
-## Articulation
-
-### [KOT](#kot_articulation)
+## [Articulation](#kot-articulation)
+
+### KOT
 - Individual results
@@
 
+- Order analysis (teaching/performing first?)
+  + [All](#kot_art_order_group_all)
+
 - [Sequence plot](#sequence_art)
 
-### [KOR](#kor-articulation)
+### KOR
 - Individual results
@@
 
-- [Sequence plot](#sequence_dyn)
-
-## Dynamics
-
-### [KOT](#kot-dynamics)
+- Order analysis (teaching/performing first?)
+  + [All](#kor_art_order_group_all)
+
+## [Dynamics](#kot-dynamics)
+
+### KOT
 - Individual results
@@
 
-### [KOR](#kor-dynamics)
+### KOR
 - Individual results
@@
   + [Trial by trial](#kor_dyn_group_trial)
+
+- Order analysis (teaching/performing first?)
+  + [All](#kor_dyn_order_group_all)
 
@@
 ```{r extract, include = FALSE}
+df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 df_kot <- read.csv("./trimmed/data_kot.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-df_kor <- read.csv("./trimmed/data_kor.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
+df_ioi$SubNr <- as.factor(df_ioi$SubNr)
 df_kot$SubNr <- as.factor(df_kot$SubNr)
-df_kor$SubNr <- as.factor(df_kor$SubNr)
+
+# For each individual, each block, each trial
+ioi <- aggregate(IOI~SubNr*BlockNr*TrialNr*Condition*Skill, data = df_ioi,
+                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
+ioi <- cbind(ioi[,1:5], as.data.frame(ioi[,6]))
+# Change colnames
+colnames(ioi) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "N", "Mean", "SD")
+
+# Calculate KOR for each interval
+df_kor <- data.frame()
+for (subnr in unique(df_kot$SubNr)){
+  for (block in unique(df_kot$BlockNr[df_kot$SubNr == subnr])){
+    for (trial in unique(df_kot$TrialNr[df_kot$SubNr == subnr & df_kot$BlockNr == block])){
+      df_current <- df_kot %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
+      df_current$KOR <- df_current$KOT/ioi$Mean[ioi$SubNr == subnr & ioi$BlockNr == block & ioi$TrialNr == trial]
+      df_kor <- rbind(df_kor, df_current)
+    }
+  }
+}
 
@@
 
-# Key-Overlap Time (KOT) - Articulation{#kot_articulation}
+# Key-Overlap Time (KOT) - Articulation{#kot-articulation}
 ## 1. All - individual{#kot_art_ind_all}
@@
 
-# Key-Overlap Ratio (KOR) - Articulation{#kor-articulation}
+# Key-Overlap Ratio (KOR) - Articulation
 ## 1. All - individual{#kor_art_ind_all}
@@
 # For each individual
-kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_art, df_kor_art$Subcomponent != "LtoS" & df_kor_art$Subcomponent != "StoL"),
+kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-### Post-hoc test(direct comparison)
+#### Post-hoc (direct comparison)
 ```{r post-hoc3, echo = FALSE}
@@
 # For each individual
-kor_art_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_art, df_kor_art$Subcomponent != "LtoS" & df_kor_art$Subcomponent != "StoL"),
+kor_art_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 ## 4. Trial by trial - group{#kor_art_group_trial}
-- Average KOR for each trial for each condition.
+- Average KOR d for each trial for each condition.
 
@@
 
-###########
-###########
-
 # Key-Overlap Time (KOT) - Dynamics{#kot-dynamics}
@@
 #### Box plot
+
 ```{r kot_dyn_all_box, echo = FALSE}
@@
 
-# Key-Overlap Ratio (KOR) - Dynamics{#kor-dynamics}
+# Key-Overlap Ratio (KOR) - Dynamics
 ## 1. All - individual{#kor_dyn_ind_all}
@@
 # For each individual
-kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_dyn, df_kor_dyn$Subcomponent != "FtoP" & df_kor_dyn$Subcomponent != "PtoF"),
+kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 ```
-  - Violate the normality assumtion
 
@@
 # For each individual
-kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_dyn, df_kor_dyn$Subcomponent != "FtoP" & df_kor_dyn$Subcomponent != "PtoF"),
+kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-## Sequence plot
-### Articulation{#sequence_art}
+# KOT-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#kot_art_order_group_all}
+```{r kot_art_order_all, echo = FALSE}
+# For each participant
+kot_art_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kot_art_order <- cbind(kot_art_order[,1:5], as.data.frame(kot_art_order[,6]))
+# Change colnames
+colnames(kot_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kot_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_art_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_art_order_all <- cbind(kot_art_order_all[,1:4], as.data.frame(kot_art_order_all[,5]))
+# Change colnames
+colnames(kot_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kot_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kot_art_order_all_bar, echo = FALSE}
+p_kot_art_order_all <- ggplot(data = kot_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
+p_kot_art_order_all
+```
+
+#### Box plot
+```{r kot_art_order_all_box, echo = FALSE}
+p_kot_art_order_all_box <- ggplot(data = kot_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT", subtitle = "KOT: Articulation")
+p_kot_art_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality5, echo = FALSE}
+# Normality check
+kot_art_order$Subcomponent <- factor(kot_art_order$Subcomponent)
+kot_art_order_norm <- by(kot_art_order$Mean, list(kot_art_order$Condition, kot_art_order$Subcomponent, kot_art_order$Order), shapiro.test)
+print(kot_art_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats9, echo = FALSE}
+kot_art_order_aov <- ezANOVA(
+  data = df_kot_art
+  , dv = .(KOT)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kot_art_order_aov$ANOVA$F <- round(kot_art_order_aov$ANOVA$F, 4)
+kot_art_order_aov$ANOVA$p <- round(kot_art_order_aov$ANOVA$p, 4)
+print(kot_art_order_aov)
+```
+
+# KOR-Order-Articulation
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#kor_art_order_group_all}
+```{r kor_art_order_all, echo = FALSE}
+# For each participant
+kor_art_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kor_art_order <- cbind(kor_art_order[,1:5], as.data.frame(kor_art_order[,6]))
+# Change colnames
+colnames(kor_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kor_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_art_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kor_art_order_all <- cbind(kor_art_order_all[,1:4], as.data.frame(kor_art_order_all[,5]))
+# Change colnames
+colnames(kor_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kor_art_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kor_art_order_all_bar, echo = FALSE}
+p_kor_art_order_all <- ggplot(data = kor_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Articulation")
+p_kor_art_order_all
+```
+
+#### Box plot
+```{r kor_art_order_all_box, echo = FALSE}
+p_kor_art_order_all_box <- ggplot(data = kor_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Articulation")
+p_kor_art_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality6, echo = FALSE}
+# Normality check
+kor_art_order$Subcomponent <- factor(kor_art_order$Subcomponent)
+kor_art_order_norm <- by(kor_art_order$Mean, list(kor_art_order$Condition, kor_art_order$Subcomponent, kor_art_order$Order), shapiro.test)
+print(kor_art_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats10, echo = FALSE}
+kor_art_order_aov <- ezANOVA(
+  data = df_kor_art
+  , dv = .(KOR)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kor_art_order_aov$ANOVA$F <- round(kor_art_order_aov$ANOVA$F, 4)
+kor_art_order_aov$ANOVA$p <- round(kor_art_order_aov$ANOVA$p, 4)
+print(kor_art_order_aov)
+```
+
+# KOT-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#kot_dyn_order_group_all}
+```{r kot_dyn_order_all, echo = FALSE}
+# For each participant
+kot_dyn_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kot_dyn_order <- cbind(kot_dyn_order[,1:5], as.data.frame(kot_dyn_order[,6]))
+# Change colnames
+colnames(kot_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kot_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_dyn_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kot_dyn_order_all <- cbind(kot_dyn_order_all[,1:4], as.data.frame(kot_dyn_order_all[,5]))
+# Change colnames
+colnames(kot_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kot_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kot_dyn_order_all_bar, echo = FALSE}
+p_kot_dyn_order_all <- ggplot(data = kot_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
+p_kot_dyn_order_all
+```
+
+#### Box plot
+```{r kot_dyn_order_all_box, echo = FALSE}
+p_kot_dyn_order_all_box <- ggplot(data = kot_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
+p_kot_dyn_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality7, echo = FALSE}
+# Normality check
+kot_dyn_order$Subcomponent <- factor(kot_dyn_order$Subcomponent)
+kot_dyn_order_norm <- by(kot_dyn_order$Mean, list(kot_dyn_order$Condition, kot_dyn_order$Subcomponent, kot_dyn_order$Order), shapiro.test)
+print(kot_dyn_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats11, echo = FALSE}
+kot_dyn_order_aov <- ezANOVA(
+  data = df_kot_dyn
+  , dv = .(KOT)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kot_dyn_order_aov$ANOVA$F <- round(kot_dyn_order_aov$ANOVA$F, 4)
+kot_dyn_order_aov$ANOVA$p <- round(kot_dyn_order_aov$ANOVA$p, 4)
+print(kot_dyn_order_aov)
+```
+
+# KOR-Order-Dynamics
+- Investigate whether there is any order effect of the conditions.
+
+## 1. All{#kor_dyn_order_group_all}
+```{r kor_dyn_order_all, echo = FALSE}
+# For each participant
+kor_dyn_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
+kor_dyn_order <- cbind(kor_dyn_order[,1:5], as.data.frame(kor_dyn_order[,6]))
+# Change colnames
+colnames(kor_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
+
+# Group mean
+kor_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_dyn_order,
+                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+kor_dyn_order_all <- cbind(kor_dyn_order_all[,1:4], as.data.frame(kor_dyn_order_all[,5]))
+# Change colnames
+colnames(kor_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
+print(kor_dyn_order_all)
+```
+
+### Plot
+#### Bar graph
+- Error bars show SEM of the mean
+
+```{r kor_dyn_order_all_bar, echo = FALSE}
+p_kor_dyn_order_all <- ggplot(data = kor_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
+                width = .2, position = position_dodge(.9)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Dynamics")
+p_kor_dyn_order_all
+```
+
+#### Box plot
+```{r kor_dyn_order_all_box, echo = FALSE}
+p_kor_dyn_order_all_box <- ggplot(data = kor_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
+  facet_wrap(Order ~ .) +
+  labs(y = "KOR", subtitle = "KOR: Dynamics")
+p_kor_dyn_order_all_box
+```
+
+### Statistics
+#### Normality check
+```{r normality8, echo = FALSE}
+# Normality check
+kor_dyn_order$Subcomponent <- factor(kor_dyn_order$Subcomponent)
+kor_dyn_order_norm <- by(kor_dyn_order$Mean, list(kor_dyn_order$Condition, kor_dyn_order$Subcomponent, kor_dyn_order$Order), shapiro.test)
+print(kor_dyn_order_norm)
+```
+
+#### Three-way ANOVA (ez)
+```{r stats12, echo = FALSE}
+kor_dyn_order_aov <- ezANOVA(
+  data = df_kor_dyn
+  , dv = .(KOR)
+  , wid = .(SubNr)
+  , between = .(Order)
+  , within = .(Condition, Subcomponent)
+  , type = 3
+  , detailed = TRUE
+)
+kor_dyn_order_aov$ANOVA$F <- round(kor_dyn_order_aov$ANOVA$F, 4)
+kor_dyn_order_aov$ANOVA$p <- round(kor_dyn_order_aov$ANOVA$p, 4)
+print(kor_dyn_order_aov)
+```
+
+# Sequence plot
+## Articulation{#sequence_art}
 ```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
@@
   labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
-
-# For each individual
-kor_art_seq <- aggregate(KOR~SubNr*Condition*Skill*Interval, data = df_kor_art,
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-kor_art_seq <- cbind(kor_art_seq[,1:4], as.data.frame(kor_art_seq[,5]))
-# Change colnames
-colnames(kor_art_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
-
-# Group mean
-kor_art_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = kor_art_seq,
-                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kor_art_seq_stats <- cbind(kor_art_seq_stats[,1:3], as.data.frame(kor_art_seq_stats[,4]))
-# Change colnames for each interval
-colnames(kor_art_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
-
-p_kor_art_seq <- ggplot(data = kor_art_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
-                position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "KOR") + scale_x_continuous(breaks=seq(1,66,1))
-
 p_kot_art_seq
-p_kor_art_seq
-```
-
-#### Dynamics{#sequence_dyn}
+```
+
+## Dynamics{#sequence_dyn}
 ```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
@@
   labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
-
-# For each individual
-kor_dyn_seq <- aggregate(KOR~SubNr*Condition*Skill*Interval, data = df_kor_dyn,
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-kor_dyn_seq <- cbind(kor_dyn_seq[,1:4], as.data.frame(kor_dyn_seq[,5]))
-# Change colnames
-colnames(kor_dyn_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
-
-# Group mean
-kor_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = kor_dyn_seq,
-                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kor_dyn_seq_stats <- cbind(kor_dyn_seq_stats[,1:3], as.data.frame(kor_dyn_seq_stats[,4]))
-# Change colnames for each interval
-colnames(kor_dyn_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
-
-p_kor_dyn_seq <- ggplot(data = kor_dyn_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
-                position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "KOR") + scale_x_continuous(breaks=seq(1,66,1))
-
 p_kot_dyn_seq
-p_kor_dyn_seq
-```
+```

### 2020-04-07 script/RNotebook/dynamics.Rmd
rerun all analyses
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-02-08 script/RNotebook/dynamics.Rmd
@@
 #### Line graph
-```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
+```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
 p_vel_dyn_trial_forte <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r vel_acc_dyn_trial_line_ftop, echo = FALSE, fig.height = 8}
+```{r vel_acc_dyn_trial_line_ftop, echo = FALSE, fig.height = 4}
 p_vel_acc_dyn_trial_ftop <- ggplot(data = subset(vel_acc_dyn_trial, vel_acc_dyn_trial$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r vel_acc_dyn_trial_line_ptof, echo = FALSE, fig.height = 8}
+```{r vel_acc_dyn_trial_line_ptof, echo = FALSE, fig.height = 4}
 p_vel_acc_dyn_trial_ptof <- ggplot(data = subset(vel_acc_dyn_trial, vel_acc_dyn_trial$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 8}
+```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 4}
 p_vel_art_trial_legato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r vel_acc_art_trial_line_ltos, echo = FALSE, fig.height = 8}
+```{r vel_acc_art_trial_line_ltos, echo = FALSE, fig.height = 4}
 p_vel_acc_art_trial_ltos <- ggplot(data = subset(vel_acc_art_trial, vel_acc_art_trial$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r vel_acc_art_trial_line_stol, echo = FALSE, fig.height = 8}
+```{r vel_acc_art_trial_line_stol, echo = FALSE, fig.height = 4}
 p_vel_acc_art_trial_stol <- ggplot(data = subset(vel_acc_art_trial, vel_acc_art_trial$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 20, fig.height = 5, echo = FALSE}
+```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
 # Average vel_dynocity for each note
@@
 ### Articulation{#sequence_art}
-```{r seq_art, fig.width = 20, fig.height = 5, echo = FALSE}
+```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
 # Average vel_artocity for each note

### 2020-04-07 script/RNotebook/ioi.Rmd
rerun all analyses
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 #### Line graph
-```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 8}
+```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 4}
 p_ioi_art_var_trial <- ggplot(data = ioi_art_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 8}
+```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 4}
 p_ioi_dyn_var_trial <- ggplot(data = ioi_dyn_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ## Articulation{#sequence_art}
-```{r seq_art, fig.width = 20, fig.height = 5, echo = FALSE}
+```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
 # For each individual
@@
 ## Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 20, fig.height = 5, echo = FALSE}
+```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
 # For each individual

### 2020-05-05 script/RNotebook/articulation.Rmd
edit outline
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-03-26 script/RNotebook/articulation.Rmd
@@
 ## [Experimental Design](#design)
+- Main predictions
 
@@
 
-## Main predictions
+### Main predictions
 Participants will exaggerate legato and staccato more while teaching.

### 2020-05-05 script/RNotebook/dynamics.Rmd
edit outline
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-03-20 script/RNotebook/dynamics.Rmd
@@
 
-## Main predictions
+### Main predictions
 - Participants will exaggerate forte and piano more while teaching.

### 2020-05-05 script/RNotebook/ioi_1.Rmd
insert the end line
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
edit outline
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-## Main predictions
+### Main predictions
 Participants will play slower while teaching regardless of the skills.

### 2020-05-19 script/RNotebook/articulation.Rmd
change file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r extract, include = FALSE}
-df_kot <- read.csv("./trimmed/data_kot.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-df_kor <- read.csv("./trimmed/data_kor.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_kot <- read.csv("../R/preprocessor/trimmed/data_kot.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_kor <- read.csv("../R/preprocessor/trimmed/data_kor.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 
@@
 
-#### Dynamics{#sequence_dyn}
+### Dynamics{#sequence_dyn}
 ```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
@@
                 position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,71,1))
 
@@
                 position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "KOR") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "Interval", y = "KOR") + scale_x_continuous(breaks=seq(1,71,1))
 

### 2020-05-19 script/RNotebook/dynamics.Rmd
change file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r extract, include = FALSE}
-df_vel <- read.csv("./trimmed/data_vel.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-df_vel_diff <- read.csv("./trimmed/data_vel_diff.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_vel <- read.csv("../R/preprocessor/trimmed/data_vel.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_vel_diff <- read.csv("../R/preprocessor/trimmed/data_vel_diff.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 
@@
                 position = position_dodge(0.05)) + 
-  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
+  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,72,1))
 p_vel_dyn_seq
@@
                 position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,71,1))
 p_vel_dyn_diff_seq
@@
                 position = position_dodge(0.05)) + 
-  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
+  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,72,1))
 p_vel_art_seq
@@
                 position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,71,1))
 p_vel_art_diff_seq

### 2020-05-19 script/RNotebook/ioi_1.Rmd
change file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi_1.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_ioi <- read.csv("../R/preprocessor/trimmed/data_ioi_1.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 
@@
                 position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,71,1))
 p_ioi_art_seq
@@
                 position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,71,1))
 p_ioi_dyn_seq

### 2020-05-19 script/RNotebook/ioi_3.Rmd
change file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi_3.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_ioi <- read.csv("../R/preprocessor/trimmed/data_ioi_3.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
 

### 2020-06-02 script/RNotebook/questionnaire.Rmd
upload analysis for questionnaire
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: 'Demographics: The Sound of Teaching Music 1'
+title: 'Demographics: The Sound of Teaching Music 2'
 output:
   html_notebook: default
+  html_document:
+    df_print: paged
 editor_options:
@@
 # Read csv files
-df <- read.csv("./questionnaire.csv", header = T, sep = ",", dec = ".", na.string = "NA")
-df_exc <- read.csv("./filtered/data_errorRate.csv", header = T, sep = ",", dec = ".")
-include <- df_exc$SubNr[df_exc$Final == "include"]
+df <- read.csv("./questionnaire_teaching-v2.0.csv", header = T, sep = ",", dec = ".", na.string = "NA")
+df_error_onset <- read.csv("../R/preprocessor/filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".")
+df_error_offset <- read.csv("../R/preprocessor/filtered/data_error_correction_offset.csv", header = T, sep = ",", dec = ".")
+df$Please.specify.the.total.number.of.years.of.formal.training..i.e...lessons.. <- as.numeric(df$Please.specify.the.total.number.of.years.of.formal.training..i.e...lessons..)
 
-# Change colnames
-colnames(df) <- c("TimeStamp", "SubNr", "Age", "Gender", "FirstLang", "SecondLang", "DominantHand", "NormalVision", "NormalHearing", "Desc1", "Desc2", "Instruments1", "FormalTraining1", "TotalPractice", "Start1", "End1", "PracticeHours", # Experience in musical instruments
-                  "Teaching", "Instruments2", "TotalTeaching1", "FormalTraining2", "Training", "AgeGroup", "Start2", "End2", "Like1", # Experience in teaching instruments
-                  "Other", "Subject", "TotalTeaching2", "Like2") # Experinence in teaching in general
+# add SubNr 21 manually
+df[21,2] <- 21
+df[21,3] <- 25
+df[21,4] <- "Male"
+df[21,5] <- "Hungarian"
+df[21,6] <- "English"
+df[21,7] <- "Right"
+df[21,8] <- "Yes"
+df[21,9] <- "Yes"
+df[21,17] <- 11
+df[21,18] <- 12
+df[21,19] <- 13
 
-# Extract the total number of years of training in piano (input manually based on the original csv file)
-df$PianoTotalPractice <- c(14, 20, 27, 21, 10, 16, 9, 13, 25, 14, 13, 11, 6, 4, 6, 8, 21, 15, 18, 7, 11, 12, 12, 18, 12, 5, 5, 10, 16, 10, 8, 22, 10, 5, 4, 14)
+# exclude SubNr 3 due to insufficient motor skills
+include <- df %>% dplyr::filter(Participant.number != 3)
 
-print(df)
+# colnames
+colnames(include) <- c("TimeStamps", "SubNr", "Age", "Gender", "FirstLang", "SecondLang", "DominantHand",
+                       "Vision", "Hearing", "KnowClementi", "PlayClementi", colnames(include)[12],
+                       colnames(include)[13], "School", "Degree", "Major", "TotalTraining", "TotalPlaying",
+                       "AgeStarted", "AgeStopped", "PracticeHours", "Instruments", colnames(include)[23],
+                       colnames(include)[24], colnames(include)[25], colnames(include)[26],
+                       colnames(include)[27], colnames(include)[28], colnames(include)[29],
+                       colnames(include)[30], colnames(include)[31], colnames(include)[32],
+                       colnames(include)[33], colnames(include)[34], colnames(include)[35],
+                       "MusicTeachingExperience", "OtherTeachingExperience", "Subject", "TotalTeaching",
+                       "TeachingLiking")
+
+# correct some gibbrish
+include[13,17] <- 18
+print(include)
 ```
@@
 ## Descriptive stats for participants' demographics
-### Number of excluded participants
-
-```{r excluded, echo = FALSE}
-# Descriptive statistics for participants' demographics
-total <- nrow(df)
-# Number of excluded participants
-exp_error <- length(c(8)) # SubNr 8 - due to an experimental error
-pitch_error <- length(df_exc$SubNr[df_exc$LessThan10 == "exclude"])
-tempo_dev <- total-exp_error-pitch_error-length(include)
-total_exc <- exp_error+pitch_error+tempo_dev
-
-excluded_participants <- data.frame(cbind(total, total_exc, exp_error, pitch_error, tempo_dev))
-colnames(excluded_participants) <- c("Total", "Excluded participants", "Experimental error", "Pitch error", "Tempo deviation")
-print(excluded_participants)
-```
-
-### Number of included participants
-```{r included, echo = FALSE}
-# Data frame with only included participants
-df_demo <- data.frame()
-for (subnr in include){
-  df_current <- df %>% dplyr::filter(SubNr == subnr)
-  df_demo <- rbind(df_demo, df_current)
-}
-
-# Number of included participants
-included <- nrow(df_demo)
-male <- nrow(subset(df_demo, df_demo$Gender == "Male"))
-female <- nrow(subset(df_demo, df_demo$Gender == "Female"))
-experience_m <- round(mean(df_demo$PianoTotalPractice), 1)
-experience_sd <- round(sd(df_demo$PianoTotalPractice)*sqrt((length(df_demo$PianoTotalPractice)-1)/length(df_demo$PianoTotalPractice)), 2) #biased sd for the current data
-
-included_participants <- data.frame(cbind(total, included, male, female, experience_m, experience_sd))
-colnames(included_participants) <- c("Total", "Included participants", "Male", "Female", "Average experience", "SD of Average experience")
-print(included_participants)
-```
-
-### Error Rate of excluded participants
-```{r er_excluded, echo = FALSE}
-# Information about excluded participants due to errorRate (SubNr 16 was excluded but due to not errorRate but tempo deviation)
-# Error Rate = the number of trials which contain any mistake (Not the number of mistakes in a trial)
-errorRate_mean_exc <- mean(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
-errorRate_sd_exc <- sd(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*sqrt((length(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])-1)/length(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16]))*100 #biased sd for the current data
-errorRate_median_exc <- median(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
-errorRate_min_exc <- min(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
-errorRate_max_exc <- max(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
-
-errorRate_excluded <- data.frame(cbind(errorRate_mean_exc, errorRate_sd_exc, errorRate_median_exc, errorRate_min_exc, errorRate_max_exc))
-colnames(errorRate_excluded) <- c("ER Mean", "ER SD", "ER Median", "ER Min", "ER Max")
-print(errorRate_excluded)
-```
-
-### Error Rate: Normality test for excluded participants
-```{r normality_for_excluded, echo = FALSE}
-# Normality check
-errorRate_excluded_norm <- shapiro.test(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])
-print(errorRate_excluded_norm)
-
-# Note: The distributions fo error rates were not normal, so we decided to report median values (plus its range) rather than mean values.
-```
-
-### Error Rate of included participants
-```{r er_included, echo = FALSE}
-# Information about included participants
-# Error Rate = the number of trials which contain any mistake (Not the number of mistakes in a trial)
-errorRate_mean_inc <- mean(df_exc$ErrorRate[df_exc$Final == "include"])*100
-errorRate_sd_inc <- sd(df_exc$ErrorRate[df_exc$Final == "include"])*sqrt((length(df_exc$ErrorRate[df_exc$Final == "include"])-1)/length(df_exc$ErrorRate[df_exc$Final == "include"]))*100 #biased sd for the current data
-errorRate_median_inc <- median(df_exc$ErrorRate[df_exc$Final == "include"])*100
-errorRate_min_inc <- min(df_exc$ErrorRate[df_exc$Final == "include"])*100
-errorRate_max_inc <- max(df_exc$ErrorRate[df_exc$Final == "include"])*100
-
-errorRate_included <- data.frame(cbind(errorRate_mean_inc, errorRate_sd_inc, errorRate_median_inc, errorRate_min_inc, errorRate_max_inc))
-colnames(errorRate_included) <- c("ER Mean", "ER SD", "ER Median", "ER Min", "ER Max")
-print(errorRate_included)
-```
-
-### Error Rate: Normality test for included participants
-```{r normality_for_included, echo = FALSE}
-# Normality check
-errorRate_included_norm <- shapiro.test(df_exc$ErrorRate[df_exc$Final == "include"])
-print(errorRate_included_norm)
-
-# Note: The distributions fo error rates were not normal, so we decided to report median values (plus its range) rather than mean values.
-```
+- Number of excluded participants: `r nrow(df)-nrow(include)`
+- Number of included participants: `r nrow(include)`
+- Male: `r nrow(subset(include, include$Gender == "Male"))`
+- Female: `r nrow(subset(include, include$Gender == "Female"))`
+- Dominant hand: Right - `r nrow(subset(include, include$DominantHand == "Right"))`, Left - `r nrow(subset(include, include$DominantHand == "Left"))`
+- Average: the total number of years of formal learning: `r mean(include$TotalTraining)`
+- Average: the total number of years of playing: `r mean(include$TotalPlaying)`

### 2020-07-24 script/RNotebook/additional240720/ioi.Rmd
update ioi.Rmd - still checking
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-07 script/RNotebook/ioi.Rmd
@@
 
-- Last checked: 07/02/20
+- Last checked: 24/07/20
 
@@
   #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
+  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") #+ coord_cartesian(ylim = c(100, 280))
 p_ioi_art
@@
 ## 3. Phrase boundaries - individual{#art_ind_phrase}
-- Compare IOIs at phrase boundaries with others fir each participant.
+- Compare IOIs at phrase boundaries with others for each participant.
 
@@
 ```{r ioi_art_ch_all, echo = FALSE}
-# Group mean
-ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_art_ch,
+# Exclude SubNr 1 as data for the teaching condition were missing
+# Group mean
+ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = subset(ioi_art_ch, ioi_art_ch$SubNr != 1),
                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
+
 ioi_art_ch_all <- cbind(ioi_art_ch_all[,1:3], as.data.frame(ioi_art_ch_all[,4]))
@@
 ioi_art_ch_aov <- ezANOVA(
-  data = df_ioi_art
+  data = subset(df_ioi_art, df_ioi_art$SubNr != 1)
   , dv = .(IOI)
@@
 ```{r stats4, echo = FALSE}
-ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
+ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = subset(ioi_art_ch, ioi_art_ch$SubNr != 1))
 print(summary(ioi_art_ch_aov_2))
@@
 ```{r ioi_dyn_all, echo = FALSE}
-# Group mean
-ioi_dyn_all <- aggregate(Mean~Condition*Skill, data = ioi_dyn,
+# Exclude SubNr 3 and 14 as data were missing for one of the conditions
+# Group mean
+ioi_dyn_all <- aggregate(Mean~Condition*Skill, data = subset(ioi_dyn, ioi_dyn$SubNr != 3 & ioi_dyn$SubNr != 14),
                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
@@
 ```{r ioi_dyn_ch_all, echo = FALSE}
-# Group mean
-ioi_dyn_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_dyn_ch,
+# Exclude SubNr 3, 14 as data were missing for one of the conditions.
+# Group mean
+ioi_dyn_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = subset(ioi_dyn_ch, ioi_dyn_ch$SubNr != 3 & ioi_dyn_ch$SubNr != 14),
                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
@@
 ioi_dyn_ch_aov <- ezANOVA(
-  data = df_ioi_dyn
+  data = subset(df_ioi_dyn, df_ioi_dyn$SubNr != 3 & df_ioi_dyn$SubNr != 14)
   , dv = .(IOI)
@@
 ```{r stats10, echo = FALSE}
-ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
+ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = subset(ioi_dyn_ch, ioi_dyn_ch$SubNr != 3 & ioi_dyn_ch$SubNr != 14))
 print(summary(ioi_dyn_ch_aov_2))

### 2020-07-28 script/RNotebook/ioi.Rmd
update analysis with the latest trimmed
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-02-07 script/RNotebook/ioi.Rmd
@@
 
-- Last checked: 07/02/20
+- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
 
@@
 ## Articulation{#sequence_art}
-```{r seq_art, fig.width = 20, fig.height = 5, echo = FALSE}
+```{r seq_art, fig.width = 20, fig.height = 4, echo = FALSE}
 # For each individual
@@
 ## Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 20, fig.height = 5, echo = FALSE}
+```{r seq_dyn, fig.width = 20, fig.height = 4, echo = FALSE}
 # For each individual

### 2020-07-28 script/RNotebook/articulation.Rmd
update analysis with the latest trimmed
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-07 script/RNotebook/articulation.Rmd
@@
 #### Line graph
-```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 4}
+```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 8}
 p_kot_art_trial_legato <- ggplot(data = subset(kot_art_trial, kot_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r kot_art_trial_line_staccato, echo = FALSE, fig.height = 4}
+```{r kot_art_trial_line_staccato, echo = FALSE, fig.height = 8}
 p_kot_art_trial_staccato <- ggplot(data = subset(kot_art_trial, kot_art_trial$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 4}
+```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 8}
 p_kor_art_trial_legato <- ggplot(data = subset(kor_art_trial, kor_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r kor_art_trial_line_staccato, echo = FALSE, fig.height = 4}
+```{r kor_art_trial_line_staccato, echo = FALSE, fig.height = 8}
 p_kor_art_trial_staccato <- ggplot(data = subset(kor_art_trial, kor_art_trial$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
+```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
 p_kot_dyn_trial_forte <- ggplot(data = subset(kot_dyn_trial, kot_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r kot_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
+```{r kot_dyn_trial_line_piano, echo = FALSE, fig.height = 8}
 p_kot_dyn_trial_piano <- ggplot(data = subset(kot_dyn_trial, kot_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
+```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
 p_kor_dyn_trial_forte <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r kor_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
+```{r kor_dyn_trial_line_piano, echo = FALSE, fig.height = 8}
 p_kor_dyn_trial_piano <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ## Articulation{#sequence_art}
-```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
+```{r seq_art, fig.width = 20, fig.height = 4, echo = FALSE}
 # For each individual
@@
 ## Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
+```{r seq_dyn, fig.width = 20, fig.height = 4, echo = FALSE}
 # For each individual

### 2020-07-28 script/RNotebook/dynamics.Rmd
update analysis with the latest trimmed
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-07 script/RNotebook/dynamics.Rmd
@@
 #### Line graph
-```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
+```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
 p_vel_dyn_trial_forte <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r vel_acc_dyn_trial_line_ftop, echo = FALSE, fig.height = 4}
+```{r vel_acc_dyn_trial_line_ftop, echo = FALSE, fig.height = 8}
 p_vel_acc_dyn_trial_ftop <- ggplot(data = subset(vel_acc_dyn_trial, vel_acc_dyn_trial$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r vel_acc_dyn_trial_line_ptof, echo = FALSE, fig.height = 4}
+```{r vel_acc_dyn_trial_line_ptof, echo = FALSE, fig.height = 8}
 p_vel_acc_dyn_trial_ptof <- ggplot(data = subset(vel_acc_dyn_trial, vel_acc_dyn_trial$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 4}
+```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 8}
 p_vel_art_trial_legato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r vel_acc_art_trial_line_ltos, echo = FALSE, fig.height = 4}
+```{r vel_acc_art_trial_line_ltos, echo = FALSE, fig.height = 8}
 p_vel_acc_art_trial_ltos <- ggplot(data = subset(vel_acc_art_trial, vel_acc_art_trial$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 
-```{r vel_acc_art_trial_line_stol, echo = FALSE, fig.height = 4}
+```{r vel_acc_art_trial_line_stol, echo = FALSE, fig.height = 8}
 p_vel_acc_art_trial_stol <- ggplot(data = subset(vel_acc_art_trial, vel_acc_art_trial$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
+```{r seq_dyn, fig.width = 20, fig.height = 4, echo = FALSE}
 # Average vel_dynocity for each note
@@
 ### Articulation{#sequence_art}
-```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
+```{r seq_art, fig.width = 20, fig.height = 4, echo = FALSE}
 # Average vel_artocity for each note

### 2020-07-28 script/RNotebook/articulation.Rmd
change fig.height - .Rhistory problem...
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 #### Line graph
-```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 8}
+```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 4}
 p_kot_art_trial_legato <- ggplot(data = subset(kot_art_trial, kot_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 8}
+```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 4}
 p_kor_art_trial_legato <- ggplot(data = subset(kor_art_trial, kor_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
+```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
 p_kot_dyn_trial_forte <- ggplot(data = subset(kot_dyn_trial, kot_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 #### Line graph
-```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
+```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
 p_kor_dyn_trial_forte <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
@@
 ## Articulation{#sequence_art}
-```{r seq_art, fig.width = 20, fig.height = 4, echo = FALSE}
+```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
 # For each individual
@@
 ## Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 20, fig.height = 4, echo = FALSE}
+```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
 # For each individual

### 2020-07-28 script/RNotebook/dynamics.Rmd
change fig.height - .Rhistory problem...
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 20, fig.height = 4, echo = FALSE}
+```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
 # Average vel_dynocity for each note
@@
 ### Articulation{#sequence_art}
-```{r seq_art, fig.width = 20, fig.height = 4, echo = FALSE}
+```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
 # Average vel_artocity for each note

### 2020-07-28 script/RNotebook/ioi.Rmd
change fig.height - .Rhistory problem...
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-07 script/RNotebook/ioi.Rmd
@@
 
-```{r ioi_dyn_trial_line, echo = FALSE, fig.height = 4}
+```{r ioi_dyn_trial_line, echo = FALSE, fig.height = 8}
 p_ioi_dyn_trial <- ggplot(data = ioi_dyn_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +

### 2020-07-29 script/RNotebook/additional020420/articulation.Rmd
update analysis with the new trimmed
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: "Additional analysis for transition points"
+title: "Additional analysis for articulation transition points"
 output: html_notebook
 ---
-Description: These are some additional analyses for transition points.
+Description: These are some additional analyses for articulation transition points.
 
@@
 How does it look like articulation transition points?
-Posthoc prediciton: Participants would have a larger separation betwenn the end of legato and the beginning of staccato.
+
+Post-hoc prediction: Participants would have **a larger separation** between the end of legato and the beginning of staccato.
 
@@
 # read csv
-df_kot <- fread("./../trimmed/data_kot.csv")
+dt_kot <- fread("./../trimmed/data_kot.csv")
 
 # only transition points for arituclation
-df_kot_trans <- df_kot[Skill == "articulation" & Subcomponent != "Legato" & Subcomponent != "Staccato"]
+dt_kot_trans <- dt_kot[Skill == "articulation" & Subcomponent != "Legato" & Subcomponent != "Staccato"]
 ```
@@
 # For each individual
-art_trans <- df_kot_trans[, .(N = length(KOT), Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Subcomponent)]
+art_trans <- dt_kot_trans[, .(N = length(KOT), Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Subcomponent)]
 print(art_trans)
@@
 print(art_trans_norm)
+
+# Converting variables to Factor
+dt_kot_trans$SubNr <- factor(dt_kot_trans$SubNr)
+dt_kot_trans$Condition <- factor(dt_kot_trans$Condition)
+dt_kot_trans$Subcomponent <- factor(dt_kot_trans$Subcomponent)
+
 # ezAnova
 art_trans_aov <- ezANOVA(
-  data = df_kot_trans
+  data = dt_kot_trans
   , dv = .(KOT)

### 2020-07-29 script/RNotebook/questionnaire.Rmd
add plot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-02-10 script/RNotebook/questionnaire.Rmd
@@
 if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
+# plot
+if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
 ```
@@
 ```
+
+### Error Rate: Plot
+```{r plot, echo = FALSE}
+p_hist <- ggplot(subset(df_exc, df_exc$Final == "include"), aes(x = ErrorRate)) +
+  geom_histogram(position = "identity", alpha = .5, binwidth = .01)
+plot(p_hist)
+```

### 2020-07-29 script/RNotebook/articulation.Rmd
define filenames separately
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+```{r file, include = FALSE}
+filename_ioi = "../R/trimmed/data_ioi.csv"
+filename_kot = "../R/trimmed/data_kot.csv"
+```
+
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-df_kot <- read.csv("./trimmed/data_kot.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_ioi <- read.csv(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
+df_kot <- read.csv(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed csv
 

### 2020-07-29 script/RNotebook/dynamics.Rmd
define filenames separately
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+```{r file, include = FALSE}
+filename_vel = "../R/trimmed/data_vel.csv"
+filename_vel_acc = "../R/trimmed/data_vel_acc.csv"
+```
+
 ```{r extract, include = FALSE}
-df_vel <- read.csv("./trimmed/data_vel.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-df_vel_acc <- read.csv("./trimmed/data_vel_acc.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_vel <- read.csv(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed csv
+df_vel_acc <- read.csv(filename_vel_acc, header = T, sep = ",", dec = ".") # read a trimmed csv
 

### 2020-07-29 script/RNotebook/ioi.Rmd
define filenames separately
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+```{r file, include = FALSE}
+filename_ioi = "../R/trimmed/data_ioi.csv"
+```
+
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_ioi <- read.csv(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 

### 2020-07-29 script/RNotebook/questionnaire.Rmd
define filenames separately
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+```{r file, include = FALSE}
+filename_que = "./questionnaire.csv"
+filename_exc = "../R/filtered/data_errorRate.csv"
+```
+
 ```{r read, echo = FALSE}
 # Read csv files
-df <- read.csv("./questionnaire.csv", header = T, sep = ",", dec = ".", na.string = "NA")
-df_exc <- read.csv("./filtered/data_errorRate.csv", header = T, sep = ",", dec = ".")
+df <- read.csv(filename_que, header = T, sep = ",", dec = ".", na.string = "NA")
+df_exc <- read.csv(filename_exc, header = T, sep = ",", dec = ".")
 include <- df_exc$SubNr[df_exc$Final == "include"]

### 2020-07-29 script/RNotebook/additional240720/ioi.Rmd
updata analysis + non-parametric test
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-07 script/RNotebook/ioi.Rmd
@@
 
-Description: This is a summary of all analyses for a tempo.
-
-- Last checked: 07/02/20
+Description: This is a summary of all analyses for a tempo. **Add excluded 14 participants.**
+
+- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
 
@@
   + [All](#art_ind_all)
-  + [Phrase bondaries](#art_ind_phrase)
+  <!-- + [Phrase bondaries](#art_ind_phrase) -->
   + [Trial by trial](#art_ind_trial)
-  + [Variability](#art_ind_var_all)
-  + [Variability trial-by-trial](#art_ind_var_trial)
+  <!-- + [Variability](#art_ind_var_all) -->
+  <!-- + [Variability trial-by-trial](#art_ind_var_trial) -->
 
@@
   + [All](#art_group_all)
-  + [Phrase bondaries](#art_group_phrase)
+  <!-- + [Phrase bondaries](#art_group_phrase) -->
   + [Trial by trial](#art_group_trial)
-  + [Variability](#art_group_var_all)
-  + [Variability trial-by-trial](#art_group_var_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#art_order_group_all)
+  <!-- + [Variability](#art_group_var_all) -->
+  <!-- + [Variability trial-by-trial](#art_group_var_trial) -->
+
+<!-- - Order analysis (teaching/performing first?) -->
+<!--   + [All](#art_order_group_all) -->
 
@@
   + [All](#dyn_ind_all)
-  + [Phrase bondaries](#dyn_ind_phrase)
+  <!-- + [Phrase bondaries](#dyn_ind_phrase) -->
   + [Trial by trial](#dyn_ind_trial)
-  + [Variability all](#dyn_ind_var_all)
-  + [Variability trial-by-trial](#dyn_ind_var_trial)
+  <!-- + [Variability all](#dyn_ind_var_all) -->
+  <!-- + [Variability trial-by-trial](#dyn_ind_var_trial) -->
 
@@
   + [All](#dyn_group_all)
-  + [Phrase bondaries](#dyn_group_phrase)
+  <!-- + [Phrase bondaries](#dyn_group_phrase) -->
   + [Trial by trial](#dyn_group_trial)
-  + [Variability all](#dyn_group_var_all)
-  + [Variability trial-by-trial](#dyn_group_var_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#dyn_order_group_all)
+  <!-- + [Variability all](#dyn_group_var_all) -->
+  <!-- + [Variability trial-by-trial](#dyn_group_var_trial) -->
+
+<!-- - Order analysis (teaching/performing first?) -->
+<!--   + [All](#dyn_order_group_all) -->
 
@@
 
+```{r file, include = FALSE}
+filename_ioi = "./trimmed/data_ioi.csv"
+```
+
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_ioi <- read.csv(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
@@
 
+#### Transformation
+- 1/x transformation
+
+```{r trans_art, echo = FALSE}
+# transformation
+ioi_art_trans <- ioi_art
+ioi_art_trans$Mean <- 1/(ioi_art$Mean)
+
+# normality check
+ioi_art_trans_norm <- by(ioi_art_trans$Mean, list(ioi_art_trans$Condition), shapiro.test)
+print(ioi_art_trans_norm)
+
+# use transformed data 
+ioi_art <- ioi_art_trans
+```
+
 #### Paired t-test
@@
 ```
-
-## 3. Phrase boundaries - individual{#art_ind_phrase}
-- Compare IOIs at phrase boundaries with others fir each participant.
-
-```{r ioi_art_ch, echo = FALSE}
-# Define phrase boundaries
-df_ioi_art$Boundary <- "No"
-df_ioi_art$Boundary[df_ioi_art$Interval == 8 | df_ioi_art$Interval == 16 | df_ioi_art$Interval == 24 | df_ioi_art$Interval == 41 | df_ioi_art$Interval == 49 | df_ioi_art$Interval == 57] <- "Yes"
-
+#### Non-parametric test (Wilcoxon-rank)
+```{r non-para_art, echo = FALSE}
 # For each individual
-ioi_art_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_art,
-                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_art_ch <- cbind(ioi_art_ch[,1:4], as.data.frame(ioi_art_ch[,5]))
-# Change colnames
-colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
-print(ioi_art_ch)
-```
-
-### Plot
-#### Bar chart
-- Error bars show SD of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_art_ch_bar, echo = FALSE}
-p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
-  facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
-p_ioi_art_ch
-```
-
-## 4. Phrase boundaries - group{#art_group_phrase}
-- Compare IOIs at phrase boundaries with others in the two conditions.
-- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations.
-
-```{r ioi_art_ch_all, echo = FALSE}
-# Group mean
-ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_art_ch,
-                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_ch_all <- cbind(ioi_art_ch_all[,1:3], as.data.frame(ioi_art_ch_all[,4]))
-# Change colnames
-colnames(ioi_art_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
-print(ioi_art_ch_all)
-```
-
-### Plot
-#### Bar chart
-- Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_art_ch_all_bar, echo = FALSE}
-p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width=.2, position = position_dodge(.9)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  #facet_grid(. ~ Boundary) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 210))
-p_ioi_art_ch_all_bar
-```
-
-#### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_art_ch_all_box, echo = FALSE}
-p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation")
-p_ioi_art_ch_box
-```
-
-### Statistics
-#### Normality check
-```{r normality2, echo = FALSE}
-ioi_art_ch_norm <- by(ioi_art_ch$Mean, list(ioi_art_ch$Condition, ioi_art_ch$Boundary), shapiro.test)
-print(ioi_art_ch_norm)
-```
-
-#### Two-way ANOVA (ez)
-```{r stats3, echo = FALSE}
-ioi_art_ch_aov <- ezANOVA(
-  data = df_ioi_art
-  , dv = .(IOI)
-  , wid = .(SubNr)
-  , within = .(Condition, Boundary)
-  , type = 3
-  , detailed = TRUE
-)
-ioi_art_ch_aov$ANOVA$F <- round(ioi_art_ch_aov$ANOVA$F, 4)
-ioi_art_ch_aov$ANOVA$p <- round(ioi_art_ch_aov$ANOVA$p, 4)
-print(ioi_art_ch_aov$ANOVA)
-```
-
-#### Two-way ANOVA (aov)
-```{r stats4, echo = FALSE}
-ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
-print(summary(ioi_art_ch_aov_2))
-```
+n_ioi_art <- aggregate(IOI~SubNr*Condition*Skill, data = df_ioi_art,
+                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
+n_ioi_art <- cbind(n_ioi_art[,1:3], as.data.frame(n_ioi_art[,4]))
+# Change colnames
+colnames(n_ioi_art) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
+
+nonpara_art <- wilcox.test(Mean ~ Condition, data = n_ioi_art, paired = TRUE)
+print(nonpara_art)
+```
+
+<!-- ## 3. Phrase boundaries - individual{#art_ind_phrase} -->
+<!-- - Compare IOIs at phrase boundaries with others fir each participant. -->
+
+<!-- ```{r ioi_art_ch, echo = FALSE} -->
+<!-- # Define phrase boundaries -->
+<!-- df_ioi_art$Boundary <- "No" -->
+<!-- df_ioi_art$Boundary[df_ioi_art$Interval == 8 | df_ioi_art$Interval == 16 | df_ioi_art$Interval == 24 | df_ioi_art$Interval == 41 | df_ioi_art$Interval == 49 | df_ioi_art$Interval == 57] <- "Yes" -->
+
+<!-- # For each individual -->
+<!-- ioi_art_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_art, -->
+<!--                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- ioi_art_ch <- cbind(ioi_art_ch[,1:4], as.data.frame(ioi_art_ch[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD") -->
+<!-- print(ioi_art_ch) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar chart -->
+<!-- - Error bars show SD of the mean -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_art_ch_bar, echo = FALSE} -->
+<!-- p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_grid(Boundary ~ .) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo -->
+<!--   #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") + -->
+<!--   labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280)) -->
+<!-- p_ioi_art_ch -->
+<!-- ``` -->
+
+<!-- ## 4. Phrase boundaries - group{#art_group_phrase} -->
+<!-- - Compare IOIs at phrase boundaries with others in the two conditions. -->
+<!-- - Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations. -->
+
+<!-- ```{r ioi_art_ch_all, echo = FALSE} -->
+<!-- # Group mean -->
+<!-- ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_art_ch, -->
+<!--                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- ioi_art_ch_all <- cbind(ioi_art_ch_all[,1:3], as.data.frame(ioi_art_ch_all[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_art_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM") -->
+<!-- print(ioi_art_ch_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar chart -->
+<!-- - Error bars show SEM of the mean -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_art_ch_all_bar, echo = FALSE} -->
+<!-- p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width=.2, position = position_dodge(.9)) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo -->
+<!--   #annotate("text", 1, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") + -->
+<!--   #facet_grid(. ~ Boundary) + -->
+<!--   labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 210)) -->
+<!-- p_ioi_art_ch_all_bar -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_art_ch_all_box, echo = FALSE} -->
+<!-- p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Boundary, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo -->
+<!--   #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") + -->
+<!--   labs(y = "IOI (ms)", subtitle = "IOI: Articulation") -->
+<!-- p_ioi_art_ch_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality2, echo = FALSE} -->
+<!-- ioi_art_ch_norm <- by(ioi_art_ch$Mean, list(ioi_art_ch$Condition, ioi_art_ch$Boundary), shapiro.test) -->
+<!-- print(ioi_art_ch_norm) -->
+<!-- ``` -->
+
+<!-- #### Two-way ANOVA (ez) -->
+<!-- ```{r stats3, echo = FALSE} -->
+<!-- ioi_art_ch_aov <- ezANOVA( -->
+<!--   data = df_ioi_art -->
+<!--   , dv = .(IOI) -->
+<!--   , wid = .(SubNr) -->
+<!--   , within = .(Condition, Boundary) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- ioi_art_ch_aov$ANOVA$F <- round(ioi_art_ch_aov$ANOVA$F, 4) -->
+<!-- ioi_art_ch_aov$ANOVA$p <- round(ioi_art_ch_aov$ANOVA$p, 4) -->
+<!-- print(ioi_art_ch_aov$ANOVA) -->
+<!-- ``` -->
+
+<!-- #### Two-way ANOVA (aov) -->
+<!-- ```{r stats4, echo = FALSE} -->
+<!-- ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch) -->
+<!-- print(summary(ioi_art_ch_aov_2)) -->
+<!-- ``` -->
 
@@
 
-## 7. Variability all - individual{#art_ind_var_all}
-- Average variability for each participant.
-- IOI variability = SD/mean IOI (per trial)
-
-```{r ioi_art_var, echo = FALSE}
-# Create a data frame for Variability (SD/mean IOI)
-df_ioi_art_var <- data.frame()
-for (subnr in unique(df_ioi_art$SubNr)){
-  for (block in c(1:4)){
-    cond = as.character(unique(df_ioi_art$Condition[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
-    skill = as.character(unique(df_ioi_art$Skill[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
-    for (trial in c(1:8)){
-      df_current <- df_ioi_art %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
-      if (nrow(df_current) != 0){
-        df_ioi_art_var <- rbind(df_ioi_art_var, 
-                                data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI)))
-      }
-    }
-  }
-}
-colnames(df_ioi_art_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability")
-
-# Average variability
-# For each individual
-ioi_art_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_art_var,
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_art_var <- cbind(ioi_art_var[,1:3], as.data.frame(ioi_art_var[,4]))
-# Change colnames
-colnames(ioi_art_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
-print(ioi_art_var)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SD of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_art_var_bar, echo = FALSE}
-p_ioi_art_var <- ggplot(data = ioi_art_var, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(1)) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation")
-p_ioi_art_var
-```
-
-## 8. Variability all - group{#art_group_var_all}
-- Average variability for each condition.
-
-```{r ioi_art_var_all, echo = FALSE}
-# Group mean
-ioi_art_var_all <- aggregate(Mean~Condition*Skill, data = ioi_art_var,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_var_all <- cbind(ioi_art_var_all[,1:2], as.data.frame(ioi_art_var_all[,3]))
-# Change colnames
-colnames(ioi_art_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
-print(ioi_art_var_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r ioi_art_var_all_bar, echo = FALSE}
-p_ioi_art_var_all <- ggplot(data = ioi_art_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge()) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
-  theme(legend.position = "None")
-p_ioi_art_var_all
-```
-
-#### Box plot
-```{r ioi_art_var_all_box, echo = FALSE}
-p_ioi_art_var_box <- ggplot(data = ioi_art_var, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
-  theme(legend.position = "None")
-p_ioi_art_var_box
-```
-
-### Statistics
-#### Normality check
-```{r normality3, echo = FALSE}
-ioi_art_var_norm <- by(ioi_art_var$Mean, list(ioi_art_var$Condition), shapiro.test)
-print(ioi_art_var_norm)
-```
-
-#### Paired t-test
-```{r stats5, echo = FALSE}
-ioi_art_var_ttest <- t.test(ioi_art_var$Mean[ioi_art_var$Condition == "teaching"], ioi_art_var$Mean[ioi_art_var$Condition == "performing"], paired = TRUE)
-print(ioi_art_var_ttest)
-```
-
-#### Effect size
-```{r stats6, echo = FALSE}
-ioi_art_var_cohend <- cohen.d(ioi_art_var$Mean, ioi_art_var$Condition, paired = TRUE)
-print(ioi_art_var_cohend)
-```
-
-## 9. Variability trial-by-trial - individual{#art_ind_var_trial}
-- Average variability for each trial for each participant.
-
-```{r ioi_art_var_trial, echo = FALSE}
-# For each individual
-ioi_art_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_art_var,
-                         FUN = function(x){c(N = length(x), mean = mean(x))})
-ioi_art_var_trial <- cbind(ioi_art_var_trial[,1:4], as.data.frame(ioi_art_var_trial[,5]))
-# Change colnames
-colnames(ioi_art_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean")
-print(ioi_art_var_trial)
-```
-
-### Plot
-#### Line graph
-```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 4}
-p_ioi_art_var_trial <- ggplot(data = ioi_art_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_art_var_trial
-```
-
-## 10. Variability trial-by-trial - group{#art_group_var_trial}
-- Average variability for each trial for each condition.
-
-```{r ioi_art_var_trial_all, echo = FALSE}
-# Group mean
-ioi_art_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_art_var_trial,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_var_trial_all <- cbind(ioi_art_var_trial_all[,1:3], as.data.frame(ioi_art_var_trial_all[,4]))
-# Change colnames
-colnames(ioi_art_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
-print(ioi_art_var_trial_all)
-```
-
-### Plot
-#### Line graph
-- Error bars show SEM of the mean
-
-```{r ioi_art_var_trial_all_line, echo = FALSE}
-p_ioi_art_var_trial_all <- ggplot(data = ioi_art_var_trial_all, 
-                                  aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line(position = position_dodge(.2)) +
-  geom_point(position = position_dodge(.2)) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_art_var_trial_all
-```
+<!-- ## 7. Variability all - individual{#art_ind_var_all} -->
+<!-- - Average variability for each participant. -->
+<!-- - IOI variability = SD/mean IOI (per trial) -->
+
+<!-- ```{r ioi_art_var, echo = FALSE} -->
+<!-- # Create a data frame for Variability (SD/mean IOI) -->
+<!-- df_ioi_art_var <- data.frame() -->
+<!-- for (subnr in unique(df_ioi_art$SubNr)){ -->
+<!--   for (block in c(1:4)){ -->
+<!--     cond = as.character(unique(df_ioi_art$Condition[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block])) -->
+<!--     skill = as.character(unique(df_ioi_art$Skill[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block])) -->
+<!--     for (trial in c(1:8)){ -->
+<!--       df_current <- df_ioi_art %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial) -->
+<!--       if (nrow(df_current) != 0){ -->
+<!--         df_ioi_art_var <- rbind(df_ioi_art_var,  -->
+<!--                                 data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI))) -->
+<!--       } -->
+<!--     } -->
+<!--   } -->
+<!-- } -->
+<!-- colnames(df_ioi_art_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability") -->
+
+<!-- # Average variability -->
+<!-- # For each individual -->
+<!-- ioi_art_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_art_var, -->
+<!--                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- ioi_art_var <- cbind(ioi_art_var[,1:3], as.data.frame(ioi_art_var[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_art_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD") -->
+<!-- print(ioi_art_var) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SD of the mean -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_art_var_bar, echo = FALSE} -->
+<!-- p_ioi_art_var <- ggplot(data = ioi_art_var, aes(x = SubNr, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), -->
+<!--                 width = .2, position = position_dodge(1)) + -->
+<!--   labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") -->
+<!-- p_ioi_art_var -->
+<!-- ``` -->
+
+<!-- ## 8. Variability all - group{#art_group_var_all} -->
+<!-- - Average variability for each condition. -->
+
+<!-- ```{r ioi_art_var_all, echo = FALSE} -->
+<!-- # Group mean -->
+<!-- ioi_art_var_all <- aggregate(Mean~Condition*Skill, data = ioi_art_var, -->
+<!--                            FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- ioi_art_var_all <- cbind(ioi_art_var_all[,1:2], as.data.frame(ioi_art_var_all[,3])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_art_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM") -->
+<!-- print(ioi_art_var_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r ioi_art_var_all_bar, echo = FALSE} -->
+<!-- p_ioi_art_var_all <- ggplot(data = ioi_art_var_all, aes(x = Condition, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge()) + -->
+<!--   labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + -->
+<!--   theme(legend.position = "None") -->
+<!-- p_ioi_art_var_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r ioi_art_var_all_box, echo = FALSE} -->
+<!-- p_ioi_art_var_box <- ggplot(data = ioi_art_var, aes(x = Condition, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) + -->
+<!--   labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + -->
+<!--   theme(legend.position = "None") -->
+<!-- p_ioi_art_var_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality3, echo = FALSE} -->
+<!-- ioi_art_var_norm <- by(ioi_art_var$Mean, list(ioi_art_var$Condition), shapiro.test) -->
+<!-- print(ioi_art_var_norm) -->
+<!-- ``` -->
+
+<!-- #### Paired t-test -->
+<!-- ```{r stats5, echo = FALSE} -->
+<!-- ioi_art_var_ttest <- t.test(ioi_art_var$Mean[ioi_art_var$Condition == "teaching"], ioi_art_var$Mean[ioi_art_var$Condition == "performing"], paired = TRUE) -->
+<!-- print(ioi_art_var_ttest) -->
+<!-- ``` -->
+
+<!-- #### Effect size -->
+<!-- ```{r stats6, echo = FALSE} -->
+<!-- ioi_art_var_cohend <- cohen.d(ioi_art_var$Mean, ioi_art_var$Condition, paired = TRUE) -->
+<!-- print(ioi_art_var_cohend) -->
+<!-- ``` -->
+
+<!-- ## 9. Variability trial-by-trial - individual{#art_ind_var_trial} -->
+<!-- - Average variability for each trial for each participant. -->
+
+<!-- ```{r ioi_art_var_trial, echo = FALSE} -->
+<!-- # For each individual -->
+<!-- ioi_art_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_art_var, -->
+<!--                          FUN = function(x){c(N = length(x), mean = mean(x))}) -->
+<!-- ioi_art_var_trial <- cbind(ioi_art_var_trial[,1:4], as.data.frame(ioi_art_var_trial[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_art_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean") -->
+<!-- print(ioi_art_var_trial) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Line graph -->
+<!-- ```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 4} -->
+<!-- p_ioi_art_var_trial <- ggplot(data = ioi_art_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) + -->
+<!--   geom_line() + -->
+<!--   geom_point() + -->
+<!--   facet_wrap(SubNr ~ .) +  -->
+<!--   labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1)) -->
+<!-- p_ioi_art_var_trial -->
+<!-- ``` -->
+
+<!-- ## 10. Variability trial-by-trial - group{#art_group_var_trial} -->
+<!-- - Average variability for each trial for each condition. -->
+
+<!-- ```{r ioi_art_var_trial_all, echo = FALSE} -->
+<!-- # Group mean -->
+<!-- ioi_art_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_art_var_trial, -->
+<!--                            FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- ioi_art_var_trial_all <- cbind(ioi_art_var_trial_all[,1:3], as.data.frame(ioi_art_var_trial_all[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_art_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM") -->
+<!-- print(ioi_art_var_trial_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Line graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r ioi_art_var_trial_all_line, echo = FALSE} -->
+<!-- p_ioi_art_var_trial_all <- ggplot(data = ioi_art_var_trial_all,  -->
+<!--                                   aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) + -->
+<!--   geom_line(position = position_dodge(.2)) + -->
+<!--   geom_point(position = position_dodge(.2)) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) +  -->
+<!--   labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1)) -->
+<!-- p_ioi_art_var_trial_all -->
+<!-- ``` -->
 
@@
 # Group mean
-ioi_dyn_all <- aggregate(Mean~Condition*Skill, data = ioi_dyn,
+ioi_dyn_all <- aggregate(Mean~Condition*Skill, data = subset(ioi_dyn, ioi_dyn$SubNr != 3 & ioi_dyn$SubNr != 14), # exclude SubNr 3 & 14 as they missed one of the conditions completely.
                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
@@
 
+#### Transformation
+- 1/x transformation
+
+```{r trans_dyn, echo = FALSE}
+# transformation
+ioi_dyn_trans <- ioi_dyn
+ioi_dyn_trans$Mean <- 1/(ioi_dyn$Mean)
+
+# normality check
+ioi_dyn_trans_norm <- by(ioi_dyn_trans$Mean, list(ioi_dyn_trans$Condition), shapiro.test)
+print(ioi_dyn_trans_norm)
+
+# use transformed data 
+ioi_dyn <- ioi_dyn_trans
+```
+
 ```{r stats7, echo = FALSE}
@@
 
-## 3. Phrase boundaries - individual{#dyn_ind_phrase}
-- Compare IOIs at phrase boundaries with others fir each participant.
-
-```{r ioi_dyn_ch, echo = FALSE}
-# Define phrase boundaries
-df_ioi_dyn$Boundary <- "No"
-df_ioi_dyn$Boundary[df_ioi_dyn$Interval == 8 | df_ioi_dyn$Interval == 16 | df_ioi_dyn$Interval == 24 | df_ioi_dyn$Interval == 41 | df_ioi_dyn$Interval == 49 | df_ioi_dyn$Interval == 57] <- "Yes"
-
+#### Non-parametric test (Wilcoxon-rank)
+```{r non-para_dyn, echo = FALSE}
 # For each individual
-ioi_dyn_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_dyn,
-                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_dyn_ch <- cbind(ioi_dyn_ch[,1:4], as.data.frame(ioi_dyn_ch[,5]))
-# Change colnames
-colnames(ioi_dyn_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
-print(ioi_dyn_ch)
-```
-
-### Plot
-#### Bar chart
-- Error bars show SD of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_dyn_ch_bar, echo = FALSE}
-p_ioi_dyn_ch <- ggplot(data = ioi_dyn_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
-  facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
-p_ioi_dyn_ch
-```
-
-## 4. Phrase boundaries - group{#dyn_group_phrase}
-- Compare IOIs at phrase boundaries with others in the two conditions.
-- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations.
-
-```{r ioi_dyn_ch_all, echo = FALSE}
-# Group mean
-ioi_dyn_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_dyn_ch,
-                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_ch_all <- cbind(ioi_dyn_ch_all[,1:3], as.data.frame(ioi_dyn_ch_all[,4]))
-# Change colnames
-colnames(ioi_dyn_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_ch_all)
-```
-
-### Plot
-#### Bar chart
-- Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_dyn_ch_all_bar, echo = FALSE}
-p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width=.2, position = position_dodge(.9)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  #facet_grid(. ~ Boundary) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 210))
-p_ioi_dyn_ch_all_bar
-```
-
-#### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_dyn_ch_all_box, echo = FALSE}
-p_ioi_dyn_ch_box <- ggplot(data = ioi_dyn_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics")
-p_ioi_dyn_ch_box
-```
-
-### Statistics
-#### Normality check
-```{r normality5, echo = FALSE}
-ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test)
-print(ioi_dyn_ch_norm)
-```
-
-#### Two-way ANOVA (ez)
-```{r stats9, echo = FALSE}
-ioi_dyn_ch_aov <- ezANOVA(
-  data = df_ioi_dyn
-  , dv = .(IOI)
-  , wid = .(SubNr)
-  , within = .(Condition, Boundary)
-  , type = 3
-  , detailed = TRUE
-)
-ioi_dyn_ch_aov$ANOVA$F <- round(ioi_dyn_ch_aov$ANOVA$F, 4)
-ioi_dyn_ch_aov$ANOVA$p <- round(ioi_dyn_ch_aov$ANOVA$p, 4)
-print(ioi_dyn_ch_aov)
-```
-
-#### Two-way ANOVA (aov)
-```{r stats10, echo = FALSE}
-ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
-print(summary(ioi_dyn_ch_aov_2))
-```
+n_ioi_dyn <- aggregate(IOI~SubNr*Condition*Skill, data = subset(df_ioi_dyn, df_ioi_dyn$SubNr != 3 & df_ioi_dyn$SubNr != 14),
+                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
+n_ioi_dyn <- cbind(n_ioi_dyn[,1:3], as.data.frame(n_ioi_dyn[,4]))
+# Change colnames
+colnames(n_ioi_dyn) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
+
+nonpara_dyn <- wilcox.test(Mean ~ Condition, data = n_ioi_dyn, paired = TRUE)
+print(nonpara_dyn)
+```
+<!-- ## 3. Phrase boundaries - individual{#dyn_ind_phrase} -->
+<!-- - Compare IOIs at phrase boundaries with others fir each participant. -->
+
+<!-- ```{r ioi_dyn_ch, echo = FALSE} -->
+<!-- # Define phrase boundaries -->
+<!-- df_ioi_dyn$Boundary <- "No" -->
+<!-- df_ioi_dyn$Boundary[df_ioi_dyn$Interval == 8 | df_ioi_dyn$Interval == 16 | df_ioi_dyn$Interval == 24 | df_ioi_dyn$Interval == 41 | df_ioi_dyn$Interval == 49 | df_ioi_dyn$Interval == 57] <- "Yes" -->
+
+<!-- # For each individual -->
+<!-- ioi_dyn_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_dyn, -->
+<!--                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- ioi_dyn_ch <- cbind(ioi_dyn_ch[,1:4], as.data.frame(ioi_dyn_ch[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_dyn_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD") -->
+<!-- print(ioi_dyn_ch) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar chart -->
+<!-- - Error bars show SD of the mean -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_dyn_ch_bar, echo = FALSE} -->
+<!-- p_ioi_dyn_ch <- ggplot(data = ioi_dyn_ch, aes(x = SubNr, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_grid(Boundary ~ .) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo -->
+<!--   #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") + -->
+<!--   labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280)) -->
+<!-- p_ioi_dyn_ch -->
+<!-- ``` -->
+
+<!-- ## 4. Phrase boundaries - group{#dyn_group_phrase} -->
+<!-- - Compare IOIs at phrase boundaries with others in the two conditions. -->
+<!-- - Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations. -->
+
+<!-- ```{r ioi_dyn_ch_all, echo = FALSE} -->
+<!-- # Group mean -->
+<!-- ioi_dyn_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_dyn_ch, -->
+<!--                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- ioi_dyn_ch_all <- cbind(ioi_dyn_ch_all[,1:3], as.data.frame(ioi_dyn_ch_all[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_dyn_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM") -->
+<!-- print(ioi_dyn_ch_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar chart -->
+<!-- - Error bars show SEM of the mean -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_dyn_ch_all_bar, echo = FALSE} -->
+<!-- p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width=.2, position = position_dodge(.9)) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo -->
+<!--   #annotate("text", 1, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") + -->
+<!--   #facet_grid(. ~ Boundary) + -->
+<!--   labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 210)) -->
+<!-- p_ioi_dyn_ch_all_bar -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_dyn_ch_all_box, echo = FALSE} -->
+<!-- p_ioi_dyn_ch_box <- ggplot(data = ioi_dyn_ch, aes(x = Boundary, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo -->
+<!--   #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") + -->
+<!--   labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") -->
+<!-- p_ioi_dyn_ch_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality5, echo = FALSE} -->
+<!-- ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test) -->
+<!-- print(ioi_dyn_ch_norm) -->
+<!-- ``` -->
+
+<!-- #### Two-way ANOVA (ez) -->
+<!-- ```{r stats9, echo = FALSE} -->
+<!-- ioi_dyn_ch_aov <- ezANOVA( -->
+<!--   data = df_ioi_dyn -->
+<!--   , dv = .(IOI) -->
+<!--   , wid = .(SubNr) -->
+<!--   , within = .(Condition, Boundary) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- ioi_dyn_ch_aov$ANOVA$F <- round(ioi_dyn_ch_aov$ANOVA$F, 4) -->
+<!-- ioi_dyn_ch_aov$ANOVA$p <- round(ioi_dyn_ch_aov$ANOVA$p, 4) -->
+<!-- print(ioi_dyn_ch_aov) -->
+<!-- ``` -->
+
+<!-- #### Two-way ANOVA (aov) -->
+<!-- ```{r stats10, echo = FALSE} -->
+<!-- ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch) -->
+<!-- print(summary(ioi_dyn_ch_aov_2)) -->
+<!-- ``` -->
 
@@
 
-## 7. Variability all - individual{#dyn_ind_var_all}
-- Average variability for each participant.
-- IOI variability = SD/mean IOI (per trial)
-
-```{r ioi_dyn_var, echo = FALSE}
-# Create a data frame for variability (SD/)
-df_ioi_dyn_var <- data.frame()
-for (subnr in unique(df_ioi_dyn$SubNr)){
-  for (block in c(1:4)){
-    cond = as.character(unique(df_ioi_dyn$Condition[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
-    skill = as.character(unique(df_ioi_dyn$Skill[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
-    for (trial in c(1:8)){
-      df_current <- df_ioi_dyn %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
-      if (nrow(df_current) != 0){
-        df_ioi_dyn_var <- rbind(df_ioi_dyn_var, 
-                                data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI)))
-      }
-    }
-  }
-}
-colnames(df_ioi_dyn_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability")
-
-# Average variability
-# For each individual
-ioi_dyn_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_dyn_var,
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_dyn_var <- cbind(ioi_dyn_var[,1:3], as.data.frame(ioi_dyn_var[,4]))
-# Change colnames
-colnames(ioi_dyn_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
-print(ioi_dyn_var)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SD of the mean
-
-```{r ioi_dyn_var_bar, echo = FALSE}
-p_ioi_dyn_var <- ggplot(data = ioi_dyn_var, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(1)) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics")
-p_ioi_dyn_var
-```
-
-## 8. Variability all - group{#dyn_group_var_all}
-- Average variability for each condition.
-
-```{r ioi_dyn_var_all, echo = FALSE}
-# Group mean
-ioi_dyn_var_all <- aggregate(Mean~Condition*Skill, data = ioi_dyn_var,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_var_all <- cbind(ioi_dyn_var_all[,1:2], as.data.frame(ioi_dyn_var_all[,3]))
-# Change colnames
-colnames(ioi_dyn_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_var_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r ioi_dyn_var_all_bar, echo = FALSE}
-p_ioi_dyn_var_all <- ggplot(data = ioi_dyn_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge()) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
-  theme(legend.position = "None")
-p_ioi_dyn_var_all
-```
-
-#### Box plot
-```{r ioi_dyn_var_all_box, echo = FALSE}
-p_ioi_dyn_var_box <- ggplot(data = ioi_dyn_var, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
-  theme(legend.position = "None")
-p_ioi_dyn_var_box
-```
-
-### Statistics
-#### Normality check
-```{r normality6, echo = FALSE}
-ioi_dyn_var_norm <- by(ioi_dyn_var$Mean, list(ioi_dyn_var$Condition), shapiro.test)
-print(ioi_dyn_var_norm)
-```
-
-```{r stats11, echo = FALSE}
-ioi_dyn_var_ttest <- t.test(ioi_dyn_var$Mean[ioi_dyn_var$Condition == "teaching"], ioi_dyn_var$Mean[ioi_dyn_var$Condition == "performing"], paired = TRUE)
-print(ioi_dyn_var_ttest)
-```
-
-```{r stats12, echo = FALSE}
-ioi_dyn_var_cohend <- cohen.d(ioi_dyn_var$Mean, ioi_dyn_var$Condition, paired = TRUE)
-print(ioi_dyn_var_cohend)
-```
-
-## 9. Variability trial-by-trial - individual{#dyn_ind_var_trial}
-- Average variability for each trial for each participant.
-
-```{r ioi_dyn_var_trial, echo = FALSE}
-# For each individual
-ioi_dyn_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn_var,
-                         FUN = function(x){c(N = length(x), mean = mean(x))})
-ioi_dyn_var_trial <- cbind(ioi_dyn_var_trial[,1:4], as.data.frame(ioi_dyn_var_trial[,5]))
-# Change colnames
-colnames(ioi_dyn_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean")
-print(ioi_dyn_var_trial)
-```
-
-### Plot
-#### Line graph
-```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 4}
-p_ioi_dyn_var_trial <- ggplot(data = ioi_dyn_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_dyn_var_trial
-```
-
-## 10. Variability trial-by-trial - group{#dyn_group_var_trial}
-Average variability for each trial for each condition.
-
-```{r ioi_dyn_var_trial_all, echo = FALSE}
-# Group mean
-ioi_dyn_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_dyn_var_trial,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_var_trial_all <- cbind(ioi_dyn_var_trial_all[,1:3], as.data.frame(ioi_dyn_var_trial_all[,4]))
-# Change colnames
-colnames(ioi_dyn_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_var_trial_all)
-```
-
-### Plot
-#### Line graph
-- Error bars show SEM of the mean
-
-```{r ioi_dyn_var_trial_all_line, echo = FALSE}
-p_ioi_dyn_var_trial_all <- ggplot(data = ioi_dyn_var_trial_all, 
-                                  aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line(position = position_dodge(.2)) +
-  geom_point(position = position_dodge(.2)) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_dyn_var_trial_all
-```
-
-# IOI-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#art_order_group_all}
-```{r ioi_art_order_all, echo = FALSE}
-# For each participant
-ioi_art_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_art,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-ioi_art_order <- cbind(ioi_art_order[,1:4], as.data.frame(ioi_art_order[,5]))
-# Change colnames
-colnames(ioi_art_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD")
-
-# Group mean
-ioi_art_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_art_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_order_all <- cbind(ioi_art_order_all[,1:3], as.data.frame(ioi_art_order_all[,4]))
-# Change colnames
-colnames(ioi_art_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM")
-print(ioi_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_art_order_all_bar, echo = FALSE}
-p_ioi_art_order_all <- ggplot(data = ioi_art_order_all, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge()) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  facet_wrap(Order ~ .) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) +
-  theme(legend.position = "None")
-p_ioi_art_order_all
-```
-
-#### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_art_order_all_box, echo = FALSE}
-p_ioi_art_order_all_box <- ggplot(data = ioi_art_order, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  facet_wrap(Order ~ .) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") +
-  theme(legend.position = "None")
-p_ioi_art_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality7, echo = FALSE}
-# Normality check
-ioi_art_order_norm <- by(ioi_art_order$Mean, list(ioi_art_order$Condition, ioi_art_order$Order), shapiro.test)
-print(ioi_art_order_norm)
-```
-
-#### Two-way ANOVA (ez)
-```{r stats13, echo = FALSE}
-ioi_art_order_aov <- ezANOVA(
-  data = df_ioi_art
-  , dv = .(IOI)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition)
-  , type = 3
-  , detailed = TRUE
-)
-ioi_art_order_aov$ANOVA$F <- round(ioi_art_order_aov$ANOVA$F, 4)
-ioi_art_order_aov$ANOVA$p <- round(ioi_art_order_aov$ANOVA$p, 4)
-print(ioi_art_order_aov)
-```
-
-#### Two-way ANOVA (aov)
-```{r stats14, echo = FALSE}
-ioi_art_order_aov_2 <- aov(Mean ~ Condition*Order + Error(SubNr/Condition) + Order, data = ioi_art_order)
-print(summary(ioi_art_order_aov_2))
-```
-
-# IOI-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#dyn_order_group_all}
-```{r ioi_dyn_order_all, echo = FALSE}
-# For each pdynicipant
-ioi_dyn_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_dyn,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-ioi_dyn_order <- cbind(ioi_dyn_order[,1:4], as.data.frame(ioi_dyn_order[,5]))
-# Change colnames
-colnames(ioi_dyn_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD")
-
-# Group mean
-ioi_dyn_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_dyn_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_order_all <- cbind(ioi_dyn_order_all[,1:3], as.data.frame(ioi_dyn_order_all[,4]))
-# Change colnames
-colnames(ioi_dyn_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_dyn_order_all_bar, echo = FALSE}
-p_ioi_dyn_order_all <- ggplot(data = ioi_dyn_order_all, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge()) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  facet_wrap(Order ~ .) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(120, 210)) +
-  theme(legend.position = "None")
-p_ioi_dyn_order_all
-```
-
-#### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_dyn_order_all_box, echo = FALSE}
-p_ioi_dyn_order_all_box <- ggplot(data = ioi_dyn_order, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  facet_wrap(Order ~ .) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") +
-  theme(legend.position = "None")
-p_ioi_dyn_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality8, echo = FALSE}
-# Normality check
-ioi_dyn_order_norm <- by(ioi_dyn_order$Mean, list(ioi_dyn_order$Condition, ioi_dyn_order$Order), shapiro.test)
-print(ioi_dyn_order_norm)
-```
-
-#### Two-way ANOVA (ez)
-```{r stats15, echo = FALSE}
-ioi_dyn_order_aov <- ezANOVA(
-  data = df_ioi_dyn
-  , dv = .(IOI)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition)
-  , type = 3
-  , detailed = TRUE
-)
-ioi_dyn_order_aov$ANOVA$F <- round(ioi_dyn_order_aov$ANOVA$F, 4)
-ioi_dyn_order_aov$ANOVA$p <- round(ioi_dyn_order_aov$ANOVA$p, 4)
-print(ioi_dyn_order_aov)
-```
-
-#### Two-way ANOVA (aov)
-```{r stats16, echo = FALSE}
-ioi_dyn_order_aov_2 <- aov(Mean ~ Condition*Order + Error(SubNr/Condition) + Order, data = ioi_dyn_order)
-print(summary(ioi_dyn_order_aov_2))
-```
+<!-- ## 7. Variability all - individual{#dyn_ind_var_all} -->
+<!-- - Average variability for each participant. -->
+<!-- - IOI variability = SD/mean IOI (per trial) -->
+
+<!-- ```{r ioi_dyn_var, echo = FALSE} -->
+<!-- # Create a data frame for variability (SD/) -->
+<!-- df_ioi_dyn_var <- data.frame() -->
+<!-- for (subnr in unique(df_ioi_dyn$SubNr)){ -->
+<!--   for (block in c(1:4)){ -->
+<!--     cond = as.character(unique(df_ioi_dyn$Condition[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block])) -->
+<!--     skill = as.character(unique(df_ioi_dyn$Skill[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block])) -->
+<!--     for (trial in c(1:8)){ -->
+<!--       df_current <- df_ioi_dyn %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial) -->
+<!--       if (nrow(df_current) != 0){ -->
+<!--         df_ioi_dyn_var <- rbind(df_ioi_dyn_var,  -->
+<!--                                 data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI))) -->
+<!--       } -->
+<!--     } -->
+<!--   } -->
+<!-- } -->
+<!-- colnames(df_ioi_dyn_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability") -->
+
+<!-- # Average variability -->
+<!-- # For each individual -->
+<!-- ioi_dyn_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_dyn_var, -->
+<!--                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- ioi_dyn_var <- cbind(ioi_dyn_var[,1:3], as.data.frame(ioi_dyn_var[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_dyn_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD") -->
+<!-- print(ioi_dyn_var) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SD of the mean -->
+
+<!-- ```{r ioi_dyn_var_bar, echo = FALSE} -->
+<!-- p_ioi_dyn_var <- ggplot(data = ioi_dyn_var, aes(x = SubNr, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), -->
+<!--                 width = .2, position = position_dodge(1)) + -->
+<!--   labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") -->
+<!-- p_ioi_dyn_var -->
+<!-- ``` -->
+
+<!-- ## 8. Variability all - group{#dyn_group_var_all} -->
+<!-- - Average variability for each condition. -->
+
+<!-- ```{r ioi_dyn_var_all, echo = FALSE} -->
+<!-- # Group mean -->
+<!-- ioi_dyn_var_all <- aggregate(Mean~Condition*Skill, data = ioi_dyn_var, -->
+<!--                            FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- ioi_dyn_var_all <- cbind(ioi_dyn_var_all[,1:2], as.data.frame(ioi_dyn_var_all[,3])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_dyn_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM") -->
+<!-- print(ioi_dyn_var_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r ioi_dyn_var_all_bar, echo = FALSE} -->
+<!-- p_ioi_dyn_var_all <- ggplot(data = ioi_dyn_var_all, aes(x = Condition, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge()) + -->
+<!--   labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + -->
+<!--   theme(legend.position = "None") -->
+<!-- p_ioi_dyn_var_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r ioi_dyn_var_all_box, echo = FALSE} -->
+<!-- p_ioi_dyn_var_box <- ggplot(data = ioi_dyn_var, aes(x = Condition, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) + -->
+<!--   labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + -->
+<!--   theme(legend.position = "None") -->
+<!-- p_ioi_dyn_var_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality6, echo = FALSE} -->
+<!-- ioi_dyn_var_norm <- by(ioi_dyn_var$Mean, list(ioi_dyn_var$Condition), shapiro.test) -->
+<!-- print(ioi_dyn_var_norm) -->
+<!-- ``` -->
+
+<!-- ```{r stats11, echo = FALSE} -->
+<!-- ioi_dyn_var_ttest <- t.test(ioi_dyn_var$Mean[ioi_dyn_var$Condition == "teaching"], ioi_dyn_var$Mean[ioi_dyn_var$Condition == "performing"], paired = TRUE) -->
+<!-- print(ioi_dyn_var_ttest) -->
+<!-- ``` -->
+
+<!-- ```{r stats12, echo = FALSE} -->
+<!-- ioi_dyn_var_cohend <- cohen.d(ioi_dyn_var$Mean, ioi_dyn_var$Condition, paired = TRUE) -->
+<!-- print(ioi_dyn_var_cohend) -->
+<!-- ``` -->
+
+<!-- ## 9. Variability trial-by-trial - individual{#dyn_ind_var_trial} -->
+<!-- - Average variability for each trial for each participant. -->
+
+<!-- ```{r ioi_dyn_var_trial, echo = FALSE} -->
+<!-- # For each individual -->
+<!-- ioi_dyn_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn_var, -->
+<!--                          FUN = function(x){c(N = length(x), mean = mean(x))}) -->
+<!-- ioi_dyn_var_trial <- cbind(ioi_dyn_var_trial[,1:4], as.data.frame(ioi_dyn_var_trial[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_dyn_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean") -->
+<!-- print(ioi_dyn_var_trial) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Line graph -->
+<!-- ```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 4} -->
+<!-- p_ioi_dyn_var_trial <- ggplot(data = ioi_dyn_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) + -->
+<!--   geom_line() + -->
+<!--   geom_point() + -->
+<!--   facet_wrap(SubNr ~ .) +  -->
+<!--   labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1)) -->
+<!-- p_ioi_dyn_var_trial -->
+<!-- ``` -->
+
+<!-- ## 10. Variability trial-by-trial - group{#dyn_group_var_trial} -->
+<!-- Average variability for each trial for each condition. -->
+
+<!-- ```{r ioi_dyn_var_trial_all, echo = FALSE} -->
+<!-- # Group mean -->
+<!-- ioi_dyn_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_dyn_var_trial, -->
+<!--                            FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- ioi_dyn_var_trial_all <- cbind(ioi_dyn_var_trial_all[,1:3], as.data.frame(ioi_dyn_var_trial_all[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_dyn_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM") -->
+<!-- print(ioi_dyn_var_trial_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Line graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r ioi_dyn_var_trial_all_line, echo = FALSE} -->
+<!-- p_ioi_dyn_var_trial_all <- ggplot(data = ioi_dyn_var_trial_all,  -->
+<!--                                   aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) + -->
+<!--   geom_line(position = position_dodge(.2)) + -->
+<!--   geom_point(position = position_dodge(.2)) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) +  -->
+<!--   labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1)) -->
+<!-- p_ioi_dyn_var_trial_all -->
+<!-- ``` -->
+
+<!-- # IOI-Order-Articulation -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#art_order_group_all} -->
+<!-- ```{r ioi_art_order_all, echo = FALSE} -->
+<!-- # For each participant -->
+<!-- ioi_art_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_art, -->
+<!--                                      FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)}) -->
+<!-- ioi_art_order <- cbind(ioi_art_order[,1:4], as.data.frame(ioi_art_order[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_art_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- ioi_art_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_art_order, -->
+<!--                                      FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- ioi_art_order_all <- cbind(ioi_art_order_all[,1:3], as.data.frame(ioi_art_order_all[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_art_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(ioi_art_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_art_order_all_bar, echo = FALSE} -->
+<!-- p_ioi_art_order_all <- ggplot(data = ioi_art_order_all, aes(x = Condition, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge()) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) + -->
+<!--   theme(legend.position = "None") -->
+<!-- p_ioi_art_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_art_order_all_box, echo = FALSE} -->
+<!-- p_ioi_art_order_all_box <- ggplot(data = ioi_art_order, aes(x = Condition, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + -->
+<!--   theme(legend.position = "None") -->
+<!-- p_ioi_art_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality7, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- ioi_art_order_norm <- by(ioi_art_order$Mean, list(ioi_art_order$Condition, ioi_art_order$Order), shapiro.test) -->
+<!-- print(ioi_art_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Two-way ANOVA (ez) -->
+<!-- ```{r stats13, echo = FALSE} -->
+<!-- ioi_art_order_aov <- ezANOVA( -->
+<!--   data = df_ioi_art -->
+<!--   , dv = .(IOI) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- ioi_art_order_aov$ANOVA$F <- round(ioi_art_order_aov$ANOVA$F, 4) -->
+<!-- ioi_art_order_aov$ANOVA$p <- round(ioi_art_order_aov$ANOVA$p, 4) -->
+<!-- print(ioi_art_order_aov) -->
+<!-- ``` -->
+
+<!-- #### Two-way ANOVA (aov) -->
+<!-- ```{r stats14, echo = FALSE} -->
+<!-- ioi_art_order_aov_2 <- aov(Mean ~ Condition*Order + Error(SubNr/Condition) + Order, data = ioi_art_order) -->
+<!-- print(summary(ioi_art_order_aov_2)) -->
+<!-- ``` -->
+
+<!-- # IOI-Order-Dynamics -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#dyn_order_group_all} -->
+<!-- ```{r ioi_dyn_order_all, echo = FALSE} -->
+<!-- # For each pdynicipant -->
+<!-- ioi_dyn_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_dyn, -->
+<!--                                      FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)}) -->
+<!-- ioi_dyn_order <- cbind(ioi_dyn_order[,1:4], as.data.frame(ioi_dyn_order[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_dyn_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- ioi_dyn_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_dyn_order, -->
+<!--                                      FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- ioi_dyn_order_all <- cbind(ioi_dyn_order_all[,1:3], as.data.frame(ioi_dyn_order_all[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(ioi_dyn_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(ioi_dyn_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_dyn_order_all_bar, echo = FALSE} -->
+<!-- p_ioi_dyn_order_all <- ggplot(data = ioi_dyn_order_all, aes(x = Condition, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge()) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(120, 210)) + -->
+<!--   theme(legend.position = "None") -->
+<!-- p_ioi_dyn_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms) -->
+
+<!-- ```{r ioi_dyn_order_all_box, echo = FALSE} -->
+<!-- p_ioi_dyn_order_all_box <- ggplot(data = ioi_dyn_order, aes(x = Condition, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dashed") + # Tempo -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + -->
+<!--   theme(legend.position = "None") -->
+<!-- p_ioi_dyn_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality8, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- ioi_dyn_order_norm <- by(ioi_dyn_order$Mean, list(ioi_dyn_order$Condition, ioi_dyn_order$Order), shapiro.test) -->
+<!-- print(ioi_dyn_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Two-way ANOVA (ez) -->
+<!-- ```{r stats15, echo = FALSE} -->
+<!-- ioi_dyn_order_aov <- ezANOVA( -->
+<!--   data = df_ioi_dyn -->
+<!--   , dv = .(IOI) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- ioi_dyn_order_aov$ANOVA$F <- round(ioi_dyn_order_aov$ANOVA$F, 4) -->
+<!-- ioi_dyn_order_aov$ANOVA$p <- round(ioi_dyn_order_aov$ANOVA$p, 4) -->
+<!-- print(ioi_dyn_order_aov) -->
+<!-- ``` -->
+
+<!-- #### Two-way ANOVA (aov) -->
+<!-- ```{r stats16, echo = FALSE} -->
+<!-- ioi_dyn_order_aov_2 <- aov(Mean ~ Condition*Order + Error(SubNr/Condition) + Order, data = ioi_dyn_order) -->
+<!-- print(summary(ioi_dyn_order_aov_2)) -->
+<!-- ``` -->
 

### 2020-07-29 script/RNotebook/additional240720/articulation.Rmd
upload analysis with new trimmed
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-07-29 script/RNotebook/articulation.Rmd
@@
 
-- Order analysis (teaching/performing first?)
-  + [All](#kot_art_order_group_all)
+<!-- - Order analysis (teaching/performing first?) -->
+<!--   + [All](#kot_art_order_group_all) -->
 
@@
 
-- Order analysis (teaching/performing first?)
-  + [All](#kor_art_order_group_all)
+<!-- - Order analysis (teaching/performing first?) -->
+<!--   + [All](#kor_art_order_group_all) -->
 
@@
 
-- Order analysis (teaching/performing first?)
-  + [All](#kot_dyn_order_group_all)
+<!-- - Order analysis (teaching/performing first?) -->
+<!--   + [All](#kot_dyn_order_group_all) -->
 
@@
 
-### KOR
-- Individual results
-  + [All](#kor_dyn_ind_all)
-  + [Trial by trial](#kor_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kor_dyn_group_all)
-  + [Trial by trial](#kor_dyn_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#kor_dyn_order_group_all)
+<!-- ### KOR -->
+<!-- - Individual results -->
+<!--   + [All](#kor_dyn_ind_all) -->
+<!--   + [Trial by trial](#kor_dyn_ind_trial) -->
+
+<!-- - Group results and stats (if available) -->
+<!--   + [All](#kor_dyn_group_all) -->
+<!--   + [Trial by trial](#kor_dyn_group_trial) -->
+
+<!-- - Order analysis (teaching/performing first?) -->
+<!--   + [All](#kor_dyn_order_group_all) -->
 
@@
 ```{r file, include = FALSE}
-filename_ioi = "../R/trimmed/data_ioi.csv"
-filename_kot = "../R/trimmed/data_kot.csv"
+filename_ioi = "./trimmed/data_ioi.csv"
+filename_kot = "./trimmed/data_kot.csv"
 ```
@@
     for (trial in unique(df_kot$TrialNr[df_kot$SubNr == subnr & df_kot$BlockNr == block])){
+      if (length(ioi$Mean[ioi$SubNr == subnr & ioi$BlockNr == block & ioi$TrialNr == trial]) != 0){ # exclude if mean IOI does not exist
         df_current <- df_kot %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
@@
         df_kor <- rbind(df_kor, df_current)
+      }
     }
@@
 # Group mean
-kot_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kot_dyn,
+kot_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = subset(kot_dyn, kot_dyn$SubNr != 3 & kot_dyn$SubNr != 14 & kot_dyn$SubNr != 31), # exclude SubNr 3 & 14 & 31 as one of the conditions is completely missing
                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
@@
 ```{r stats5, echo = FALSE}
+# exclude SubNr 3 & 14 & 31
+df_kot_dyn_exc <- subset(df_kot_dyn, df_kot_dyn$SubNr != 3 & df_kot_dyn$SubNr != 14 & df_kot_dyn$SubNr != 31)
 kot_dyn_aov <- ezANOVA(
-  data = subset(df_kot_dyn, df_kot_dyn$Subcomponent == "Forte" | df_kot_dyn$Subcomponent == "Piano")
+  data = subset(df_kot_dyn_exc, df_kot_dyn_exc$Subcomponent == "Forte" | df_kot_dyn_exc$Subcomponent == "Piano")
   , dv = .(KOT)
@@
 ```{r stats6, echo = FALSE}
-kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
+kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = subset(kot_dyn, kot_dyn$SubNr != 3 & kot_dyn$SubNr != 14 & kot_dyn$SubNr != 31))
 print(summary(kot_dyn_aov_2))
@@
 
-# Key-Overlap Ratio (KOR) - Dynamics
-## 1. All - individual{#kor_dyn_ind_all}
-- Average KOR for each participant.
-
-```{r kor_dyn, echo = FALSE}
-# For each individual
-kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-kor_dyn <- cbind(kor_dyn[,1:4], as.data.frame(kor_dyn[,5]))
-# Change colnames
-colnames(kor_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
-print(kor_dyn)
-```
-
-### Plot
-#### Bar chart
-- Error bars show SD of the mean
-
-```{r kor_dyn_bar,  echo = FALSE}
-p_kor_dyn <- ggplot(data = kor_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
-  facet_grid(Subcomponent ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Dynamics")
-p_kor_dyn
-```
-
-## 2. All - group{#kor_dyn_group_all}
-- Average KOR for each condition.
-
-```{r kor_dyn_all, echo = FALSE}
-# Group mean
-kor_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kor_dyn,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kor_dyn_all <- cbind(kor_dyn_all[,1:3], as.data.frame(kor_dyn_all[,4]))
-# Change colnames
-colnames(kor_dyn_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
-print(kor_dyn_all)
-```
-
-### Plot
-#### Bar chart
-- Error bars show SEM of the mean
-
-```{r kor_dyn_all_bar, echo = FALSE}　
-p_kor_dyn_all <- ggplot(data = kor_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  labs(y = "KOR", subtitle = "KOR: Dynamics")
-p_kor_dyn_all
-```
-
-#### Box plot
-```{r kor_dyn_all_box, echo = FALSE}
-p_kor_dyn_all_box <- ggplot(data = kor_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
-  labs(y = "KOR", subtitle = "KOR: Dynamics")
-p_kor_dyn_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality4, echo = FALSE}
-# Reduce unused factors
-kor_dyn$Subcomponent <- factor(kor_dyn$Subcomponent)
-kor_dyn_norm <- by(kor_dyn$Mean, list(kor_dyn$Condition, kor_dyn$Subcomponent), shapiro.test)
-print(kor_dyn_norm)
-```
-
-#### Two-way ANOVA (ez)
-```{r stats7, echo = FALSE}
-kor_dyn_aov <- ezANOVA(
-  data = subset(df_kor_dyn, df_kor_dyn$Subcomponent == "Forte" | df_kor_dyn$Subcomponent == "Piano")
-  , dv = .(KOR)
-  , wid = .(SubNr)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kor_dyn_aov$ANOVA$F <- round(kor_dyn_aov$ANOVA$F, 4)
-kor_dyn_aov$ANOVA$p <- round(kor_dyn_aov$ANOVA$p, 4)
-print(kor_dyn_aov$ANOVA)
-```
-
-#### Two-way ANOVA (aov)
-```{r stats8, echo = FALSE}
-kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
-print(summary(kor_dyn_aov_2))
-```
-
-## 3. Trial by trial - individual{#kor_dyn_ind_trial}
-- Average KOR for each trial for each participant.
-
-```{r kor_dyn_trial, echo = FALSE}
-# For each individual
-kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-kor_dyn_trial <- cbind(kor_dyn_trial[,1:5], as.data.frame(kor_dyn_trial[,6]))
-# Change colnames
-colnames(kor_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
-print(kor_dyn_trial)
-```
-
-### Plot
-#### Line graph
-```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
-p_kor_dyn_trial_forte <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
-p_kor_dyn_trial_forte
-```
-
-```{r kor_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
-p_kor_dyn_trial_piano <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
-p_kor_dyn_trial_piano
-```
-
-## 4. Trial by trial - group{#kor_dyn_group_trial}
-- Average KOR for each trial for each condition.
-
-```{r kor_dyn_trial_all, echo = FALSE}
-# Group mean
-kor_dyn_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = kor_dyn_trial,
-                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kor_dyn_trial_all <- cbind(kor_dyn_trial_all[,1:4], as.data.frame(kor_dyn_trial_all[,5]))
-# Change colnames
-colnames(kor_dyn_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
-print(kor_dyn_trial_all)
-```
-
-### Plot
-#### Line graph
-- Error bars show SEM of the mean
-
-```{r kor_dyn_trial_all_line_forte, echo = FALSE}
-p_kor_dyn_trial_all_forte <- ggplot(data = subset(kor_dyn_trial_all, kor_dyn_trial_all$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line(position = position_dodge(.2)) +
-  geom_point(position = position_dodge(.2)) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
-p_kor_dyn_trial_all_forte
-```
-
-```{r kor_dyn_trial_all_line_piano, echo = FALSE}
-p_kor_dyn_trial_all_piano <- ggplot(data = subset(kor_dyn_trial_all, kor_dyn_trial_all$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line(position = position_dodge(.2)) +
-  geom_point(position = position_dodge(.2)) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
-p_kor_dyn_trial_all_piano
-```
-
-# KOT-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#kot_art_order_group_all}
-```{r kot_art_order_all, echo = FALSE}
-# For each participant
-kot_art_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-kot_art_order <- cbind(kot_art_order[,1:5], as.data.frame(kot_art_order[,6]))
-# Change colnames
-colnames(kot_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-kot_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_art_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kot_art_order_all <- cbind(kot_art_order_all[,1:4], as.data.frame(kot_art_order_all[,5]))
-# Change colnames
-colnames(kot_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(kot_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r kot_art_order_all_bar, echo = FALSE}
-p_kot_art_order_all <- ggplot(data = kot_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
-p_kot_art_order_all
-```
-
-#### Box plot
-```{r kot_art_order_all_box, echo = FALSE}
-p_kot_art_order_all_box <- ggplot(data = kot_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOT", subtitle = "KOT: Articulation")
-p_kot_art_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality5, echo = FALSE}
-# Normality check
-kot_art_order$Subcomponent <- factor(kot_art_order$Subcomponent)
-kot_art_order_norm <- by(kot_art_order$Mean, list(kot_art_order$Condition, kot_art_order$Subcomponent, kot_art_order$Order), shapiro.test)
-print(kot_art_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats9, echo = FALSE}
-kot_art_order_aov <- ezANOVA(
-  data = df_kot_art
-  , dv = .(KOT)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kot_art_order_aov$ANOVA$F <- round(kot_art_order_aov$ANOVA$F, 4)
-kot_art_order_aov$ANOVA$p <- round(kot_art_order_aov$ANOVA$p, 4)
-print(kot_art_order_aov)
-```
-
-# KOR-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#kor_art_order_group_all}
-```{r kor_art_order_all, echo = FALSE}
-# For each participant
-kor_art_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-kor_art_order <- cbind(kor_art_order[,1:5], as.data.frame(kor_art_order[,6]))
-# Change colnames
-colnames(kor_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-kor_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_art_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kor_art_order_all <- cbind(kor_art_order_all[,1:4], as.data.frame(kor_art_order_all[,5]))
-# Change colnames
-colnames(kor_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(kor_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r kor_art_order_all_bar, echo = FALSE}
-p_kor_art_order_all <- ggplot(data = kor_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Articulation")
-p_kor_art_order_all
-```
-
-#### Box plot
-```{r kor_art_order_all_box, echo = FALSE}
-p_kor_art_order_all_box <- ggplot(data = kor_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Articulation")
-p_kor_art_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality6, echo = FALSE}
-# Normality check
-kor_art_order$Subcomponent <- factor(kor_art_order$Subcomponent)
-kor_art_order_norm <- by(kor_art_order$Mean, list(kor_art_order$Condition, kor_art_order$Subcomponent, kor_art_order$Order), shapiro.test)
-print(kor_art_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats10, echo = FALSE}
-kor_art_order_aov <- ezANOVA(
-  data = df_kor_art
-  , dv = .(KOR)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kor_art_order_aov$ANOVA$F <- round(kor_art_order_aov$ANOVA$F, 4)
-kor_art_order_aov$ANOVA$p <- round(kor_art_order_aov$ANOVA$p, 4)
-print(kor_art_order_aov)
-```
-
-# KOT-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#kot_dyn_order_group_all}
-```{r kot_dyn_order_all, echo = FALSE}
-# For each participant
-kot_dyn_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-kot_dyn_order <- cbind(kot_dyn_order[,1:5], as.data.frame(kot_dyn_order[,6]))
-# Change colnames
-colnames(kot_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-kot_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_dyn_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kot_dyn_order_all <- cbind(kot_dyn_order_all[,1:4], as.data.frame(kot_dyn_order_all[,5]))
-# Change colnames
-colnames(kot_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(kot_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r kot_dyn_order_all_bar, echo = FALSE}
-p_kot_dyn_order_all <- ggplot(data = kot_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
-p_kot_dyn_order_all
-```
-
-#### Box plot
-```{r kot_dyn_order_all_box, echo = FALSE}
-p_kot_dyn_order_all_box <- ggplot(data = kot_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
-p_kot_dyn_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality7, echo = FALSE}
-# Normality check
-kot_dyn_order$Subcomponent <- factor(kot_dyn_order$Subcomponent)
-kot_dyn_order_norm <- by(kot_dyn_order$Mean, list(kot_dyn_order$Condition, kot_dyn_order$Subcomponent, kot_dyn_order$Order), shapiro.test)
-print(kot_dyn_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats11, echo = FALSE}
-kot_dyn_order_aov <- ezANOVA(
-  data = df_kot_dyn
-  , dv = .(KOT)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kot_dyn_order_aov$ANOVA$F <- round(kot_dyn_order_aov$ANOVA$F, 4)
-kot_dyn_order_aov$ANOVA$p <- round(kot_dyn_order_aov$ANOVA$p, 4)
-print(kot_dyn_order_aov)
-```
-
-# KOR-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#kor_dyn_order_group_all}
-```{r kor_dyn_order_all, echo = FALSE}
-# For each participant
-kor_dyn_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-kor_dyn_order <- cbind(kor_dyn_order[,1:5], as.data.frame(kor_dyn_order[,6]))
-# Change colnames
-colnames(kor_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-kor_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_dyn_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kor_dyn_order_all <- cbind(kor_dyn_order_all[,1:4], as.data.frame(kor_dyn_order_all[,5]))
-# Change colnames
-colnames(kor_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(kor_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r kor_dyn_order_all_bar, echo = FALSE}
-p_kor_dyn_order_all <- ggplot(data = kor_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Dynamics")
-p_kor_dyn_order_all
-```
-
-#### Box plot
-```{r kor_dyn_order_all_box, echo = FALSE}
-p_kor_dyn_order_all_box <- ggplot(data = kor_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Dynamics")
-p_kor_dyn_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality8, echo = FALSE}
-# Normality check
-kor_dyn_order$Subcomponent <- factor(kor_dyn_order$Subcomponent)
-kor_dyn_order_norm <- by(kor_dyn_order$Mean, list(kor_dyn_order$Condition, kor_dyn_order$Subcomponent, kor_dyn_order$Order), shapiro.test)
-print(kor_dyn_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats12, echo = FALSE}
-kor_dyn_order_aov <- ezANOVA(
-  data = df_kor_dyn
-  , dv = .(KOR)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kor_dyn_order_aov$ANOVA$F <- round(kor_dyn_order_aov$ANOVA$F, 4)
-kor_dyn_order_aov$ANOVA$p <- round(kor_dyn_order_aov$ANOVA$p, 4)
-print(kor_dyn_order_aov)
-```
+<!-- # Key-Overlap Ratio (KOR) - Dynamics -->
+<!-- ## 1. All - individual{#kor_dyn_ind_all} -->
+<!-- - Average KOR for each participant. -->
+
+<!-- ```{r kor_dyn, echo = FALSE} -->
+<!-- # For each individual -->
+<!-- kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57), -->
+<!--                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- kor_dyn <- cbind(kor_dyn[,1:4], as.data.frame(kor_dyn[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(kor_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD") -->
+<!-- print(kor_dyn) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar chart -->
+<!-- - Error bars show SD of the mean -->
+
+<!-- ```{r kor_dyn_bar,  echo = FALSE} -->
+<!-- p_kor_dyn <- ggplot(data = kor_dyn, aes(x = SubNr, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_grid(Subcomponent ~ .) + -->
+<!--   labs(y = "KOR", subtitle = "KOR: Dynamics") -->
+<!-- p_kor_dyn -->
+<!-- ``` -->
+
+<!-- ## 2. All - group{#kor_dyn_group_all} -->
+<!-- - Average KOR for each condition. -->
+
+<!-- ```{r kor_dyn_all, echo = FALSE} -->
+<!-- # Group mean -->
+<!-- kor_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = kor_dyn, -->
+<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- kor_dyn_all <- cbind(kor_dyn_all[,1:3], as.data.frame(kor_dyn_all[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(kor_dyn_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM") -->
+<!-- print(kor_dyn_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar chart -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r kor_dyn_all_bar, echo = FALSE}　 -->
+<!-- p_kor_dyn_all <- ggplot(data = kor_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   labs(y = "KOR", subtitle = "KOR: Dynamics") -->
+<!-- p_kor_dyn_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r kor_dyn_all_box, echo = FALSE} -->
+<!-- p_kor_dyn_all_box <- ggplot(data = kor_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) + -->
+<!--   labs(y = "KOR", subtitle = "KOR: Dynamics") -->
+<!-- p_kor_dyn_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality4, echo = FALSE} -->
+<!-- # Reduce unused factors -->
+<!-- kor_dyn$Subcomponent <- factor(kor_dyn$Subcomponent) -->
+<!-- kor_dyn_norm <- by(kor_dyn$Mean, list(kor_dyn$Condition, kor_dyn$Subcomponent), shapiro.test) -->
+<!-- print(kor_dyn_norm) -->
+<!-- ``` -->
+
+<!-- #### Two-way ANOVA (ez) -->
+<!-- ```{r stats7, echo = FALSE} -->
+<!-- kor_dyn_aov <- ezANOVA( -->
+<!--   data = subset(df_kor_dyn, df_kor_dyn$Subcomponent == "Forte" | df_kor_dyn$Subcomponent == "Piano") -->
+<!--   , dv = .(KOR) -->
+<!--   , wid = .(SubNr) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- kor_dyn_aov$ANOVA$F <- round(kor_dyn_aov$ANOVA$F, 4) -->
+<!-- kor_dyn_aov$ANOVA$p <- round(kor_dyn_aov$ANOVA$p, 4) -->
+<!-- print(kor_dyn_aov$ANOVA) -->
+<!-- ``` -->
+
+<!-- #### Two-way ANOVA (aov) -->
+<!-- ```{r stats8, echo = FALSE} -->
+<!-- kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn) -->
+<!-- print(summary(kor_dyn_aov_2)) -->
+<!-- ``` -->
+
+<!-- ## 3. Trial by trial - individual{#kor_dyn_ind_trial} -->
+<!-- - Average KOR for each trial for each participant. -->
+
+<!-- ```{r kor_dyn_trial, echo = FALSE} -->
+<!-- # For each individual -->
+<!-- kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57), -->
+<!--                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- kor_dyn_trial <- cbind(kor_dyn_trial[,1:5], as.data.frame(kor_dyn_trial[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(kor_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD") -->
+<!-- print(kor_dyn_trial) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Line graph -->
+<!-- ```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 4} -->
+<!-- p_kor_dyn_trial_forte <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) + -->
+<!--   geom_line() + -->
+<!--   geom_point() + -->
+<!--   facet_wrap(SubNr ~ .) +  -->
+<!--   labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1)) -->
+<!-- p_kor_dyn_trial_forte -->
+<!-- ``` -->
+
+<!-- ```{r kor_dyn_trial_line_piano, echo = FALSE, fig.height = 4} -->
+<!-- p_kor_dyn_trial_piano <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) + -->
+<!--   geom_line() + -->
+<!--   geom_point() + -->
+<!--   facet_wrap(SubNr ~ .) +  -->
+<!--   labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1)) -->
+<!-- p_kor_dyn_trial_piano -->
+<!-- ``` -->
+
+<!-- ## 4. Trial by trial - group{#kor_dyn_group_trial} -->
+<!-- - Average KOR for each trial for each condition. -->
+
+<!-- ```{r kor_dyn_trial_all, echo = FALSE} -->
+<!-- # Group mean -->
+<!-- kor_dyn_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = kor_dyn_trial, -->
+<!--                        FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- kor_dyn_trial_all <- cbind(kor_dyn_trial_all[,1:4], as.data.frame(kor_dyn_trial_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(kor_dyn_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM") -->
+<!-- print(kor_dyn_trial_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Line graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r kor_dyn_trial_all_line_forte, echo = FALSE} -->
+<!-- p_kor_dyn_trial_all_forte <- ggplot(data = subset(kor_dyn_trial_all, kor_dyn_trial_all$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) + -->
+<!--   geom_line(position = position_dodge(.2)) + -->
+<!--   geom_point(position = position_dodge(.2)) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) +  -->
+<!--   labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1)) -->
+<!-- p_kor_dyn_trial_all_forte -->
+<!-- ``` -->
+
+<!-- ```{r kor_dyn_trial_all_line_piano, echo = FALSE} -->
+<!-- p_kor_dyn_trial_all_piano <- ggplot(data = subset(kor_dyn_trial_all, kor_dyn_trial_all$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) + -->
+<!--   geom_line(position = position_dodge(.2)) + -->
+<!--   geom_point(position = position_dodge(.2)) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) +  -->
+<!--   labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1)) -->
+<!-- p_kor_dyn_trial_all_piano -->
+<!-- ``` -->
+
+<!-- # KOT-Order-Articulation -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#kot_art_order_group_all} -->
+<!-- ```{r kot_art_order_all, echo = FALSE} -->
+<!-- # For each participant -->
+<!-- kot_art_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)}) -->
+<!-- kot_art_order <- cbind(kot_art_order[,1:5], as.data.frame(kot_art_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(kot_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- kot_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_art_order, -->
+<!--                                      FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- kot_art_order_all <- cbind(kot_art_order_all[,1:4], as.data.frame(kot_art_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(kot_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(kot_art_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r kot_art_order_all_bar, echo = FALSE} -->
+<!-- p_kot_art_order_all <- ggplot(data = kot_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KOT (ms)", subtitle = "KOT: Articulation") -->
+<!-- p_kot_art_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r kot_art_order_all_box, echo = FALSE} -->
+<!-- p_kot_art_order_all_box <- ggplot(data = kot_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KOT", subtitle = "KOT: Articulation") -->
+<!-- p_kot_art_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality5, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- kot_art_order$Subcomponent <- factor(kot_art_order$Subcomponent) -->
+<!-- kot_art_order_norm <- by(kot_art_order$Mean, list(kot_art_order$Condition, kot_art_order$Subcomponent, kot_art_order$Order), shapiro.test) -->
+<!-- print(kot_art_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats9, echo = FALSE} -->
+<!-- kot_art_order_aov <- ezANOVA( -->
+<!--   data = df_kot_art -->
+<!--   , dv = .(KOT) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- kot_art_order_aov$ANOVA$F <- round(kot_art_order_aov$ANOVA$F, 4) -->
+<!-- kot_art_order_aov$ANOVA$p <- round(kot_art_order_aov$ANOVA$p, 4) -->
+<!-- print(kot_art_order_aov) -->
+<!-- ``` -->
+
+<!-- # KOR-Order-Articulation -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#kor_art_order_group_all} -->
+<!-- ```{r kor_art_order_all, echo = FALSE} -->
+<!-- # For each participant -->
+<!-- kor_art_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)}) -->
+<!-- kor_art_order <- cbind(kor_art_order[,1:5], as.data.frame(kor_art_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(kor_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- kor_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_art_order, -->
+<!--                                      FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- kor_art_order_all <- cbind(kor_art_order_all[,1:4], as.data.frame(kor_art_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(kor_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(kor_art_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r kor_art_order_all_bar, echo = FALSE} -->
+<!-- p_kor_art_order_all <- ggplot(data = kor_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KOR", subtitle = "KOR: Articulation") -->
+<!-- p_kor_art_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r kor_art_order_all_box, echo = FALSE} -->
+<!-- p_kor_art_order_all_box <- ggplot(data = kor_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KOR", subtitle = "KOR: Articulation") -->
+<!-- p_kor_art_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality6, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- kor_art_order$Subcomponent <- factor(kor_art_order$Subcomponent) -->
+<!-- kor_art_order_norm <- by(kor_art_order$Mean, list(kor_art_order$Condition, kor_art_order$Subcomponent, kor_art_order$Order), shapiro.test) -->
+<!-- print(kor_art_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats10, echo = FALSE} -->
+<!-- kor_art_order_aov <- ezANOVA( -->
+<!--   data = df_kor_art -->
+<!--   , dv = .(KOR) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- kor_art_order_aov$ANOVA$F <- round(kor_art_order_aov$ANOVA$F, 4) -->
+<!-- kor_art_order_aov$ANOVA$p <- round(kor_art_order_aov$ANOVA$p, 4) -->
+<!-- print(kor_art_order_aov) -->
+<!-- ``` -->
+
+<!-- # KOT-Order-Dynamics -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#kot_dyn_order_group_all} -->
+<!-- ```{r kot_dyn_order_all, echo = FALSE} -->
+<!-- # For each participant -->
+<!-- kot_dyn_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)}) -->
+<!-- kot_dyn_order <- cbind(kot_dyn_order[,1:5], as.data.frame(kot_dyn_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(kot_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- kot_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_dyn_order, -->
+<!--                                      FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- kot_dyn_order_all <- cbind(kot_dyn_order_all[,1:4], as.data.frame(kot_dyn_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(kot_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(kot_dyn_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r kot_dyn_order_all_bar, echo = FALSE} -->
+<!-- p_kot_dyn_order_all <- ggplot(data = kot_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KOT (ms)", subtitle = "KOT: Dynamics") -->
+<!-- p_kot_dyn_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r kot_dyn_order_all_box, echo = FALSE} -->
+<!-- p_kot_dyn_order_all_box <- ggplot(data = kot_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KOT (ms)", subtitle = "KOT: Dynamics") -->
+<!-- p_kot_dyn_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality7, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- kot_dyn_order$Subcomponent <- factor(kot_dyn_order$Subcomponent) -->
+<!-- kot_dyn_order_norm <- by(kot_dyn_order$Mean, list(kot_dyn_order$Condition, kot_dyn_order$Subcomponent, kot_dyn_order$Order), shapiro.test) -->
+<!-- print(kot_dyn_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats11, echo = FALSE} -->
+<!-- kot_dyn_order_aov <- ezANOVA( -->
+<!--   data = df_kot_dyn -->
+<!--   , dv = .(KOT) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- kot_dyn_order_aov$ANOVA$F <- round(kot_dyn_order_aov$ANOVA$F, 4) -->
+<!-- kot_dyn_order_aov$ANOVA$p <- round(kot_dyn_order_aov$ANOVA$p, 4) -->
+<!-- print(kot_dyn_order_aov) -->
+<!-- ``` -->
+
+<!-- # KOR-Order-Dynamics -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#kor_dyn_order_group_all} -->
+<!-- ```{r kor_dyn_order_all, echo = FALSE} -->
+<!-- # For each participant -->
+<!-- kor_dyn_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)}) -->
+<!-- kor_dyn_order <- cbind(kor_dyn_order[,1:5], as.data.frame(kor_dyn_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(kor_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- kor_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_dyn_order, -->
+<!--                                      FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- kor_dyn_order_all <- cbind(kor_dyn_order_all[,1:4], as.data.frame(kor_dyn_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(kor_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(kor_dyn_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r kor_dyn_order_all_bar, echo = FALSE} -->
+<!-- p_kor_dyn_order_all <- ggplot(data = kor_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KOR", subtitle = "KOR: Dynamics") -->
+<!-- p_kor_dyn_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r kor_dyn_order_all_box, echo = FALSE} -->
+<!-- p_kor_dyn_order_all_box <- ggplot(data = kor_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KOR", subtitle = "KOR: Dynamics") -->
+<!-- p_kor_dyn_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality8, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- kor_dyn_order$Subcomponent <- factor(kor_dyn_order$Subcomponent) -->
+<!-- kor_dyn_order_norm <- by(kor_dyn_order$Mean, list(kor_dyn_order$Condition, kor_dyn_order$Subcomponent, kor_dyn_order$Order), shapiro.test) -->
+<!-- print(kor_dyn_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats12, echo = FALSE} -->
+<!-- kor_dyn_order_aov <- ezANOVA( -->
+<!--   data = df_kor_dyn -->
+<!--   , dv = .(KOR) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- kor_dyn_order_aov$ANOVA$F <- round(kor_dyn_order_aov$ANOVA$F, 4) -->
+<!-- kor_dyn_order_aov$ANOVA$p <- round(kor_dyn_order_aov$ANOVA$p, 4) -->
+<!-- print(kor_dyn_order_aov) -->
+<!-- ``` -->
 

### 2020-07-29 script/RNotebook/additional240720/dynamics.Rmd
upload analysis with new trimmed
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-07-29 script/RNotebook/dynamics.Rmd
@@
 
-Description: This is a summary of all analyses for dynamics.
+Description: This is a summary of all analyses for dynamics. **Add excluded 14 participants.**
 
@@
 
-- Order analysis (teaching/performing first?)
-  + [All](#vel_dyn_order_group_all)
+<!-- - Order analysis (teaching/performing first?) -->
+<!--   + [All](#vel_dyn_order_group_all) -->
 
@@
 
-- Order analysis (teaching/performing first?)
-  + [All](#vel_acc_dyn_order_group_all)
+<!-- - Order analysis (teaching/performing first?) -->
+<!--   + [All](#vel_acc_dyn_order_group_all) -->
 
@@
 
-- Order analysis (teaching/performing first?)
-  + [All](#vel_art_order_group_all)
+<!-- - Order analysis (teaching/performing first?) -->
+<!--   + [All](#vel_art_order_group_all) -->
 
@@
 
-- Order analysis (teaching/performing first?)
-  + [All](#vel_acc_art_order_group_all)
+<!-- - Order analysis (teaching/performing first?) -->
+<!--   + [All](#vel_acc_art_order_group_all) -->
 
@@
 ```{r file, include = FALSE}
-filename_vel = "../R/trimmed/data_vel.csv"
-filename_vel_acc = "../R/trimmed/data_vel_acc.csv"
+filename_vel = "./trimmed/data_vel.csv"
+filename_vel_acc = "./trimmed/data_vel_acc.csv"
 ```
@@
 # Group mean
-vel_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_dyn,
+vel_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = subset(vel_dyn, vel_dyn$SubNr != 3 & vel_dyn$SubNr != 14), # exclude SubNr 3 & 14 as one of the conditions is completely missing
                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
@@
 ```{r vel_dyn_all_box, echo = FALSE}
-p_vel_dyn_all_box <- ggplot(data = vel_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+p_vel_dyn_all_box <- ggplot(data = subset(vel_dyn, vel_dyn$SubNr != 3 & vel_dyn$SubNr != 14), aes(x = Subcomponent, y = Mean, fill = Condition)) +
   geom_boxplot() +
@@
 ```{r normality1, echo = FALSE}
+# exclude SubNr 3 & 14
+vel_dyn <- subset(vel_dyn, vel_dyn$SubNr != 3 & vel_dyn$SubNr != 14)
 # Reduce unused factors
@@
 vel_dyn_aov <- ezANOVA(
-  data = df_vel_dyn
+  data = subset(df_vel_dyn, df_vel_dyn$SubNr != 3 & df_vel_dyn$SubNr != 14)
   , dv = .(Velocity)
@@
 # Group mean
-vel_acc_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_acc_dyn,
+vel_acc_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = subset(vel_acc_dyn, vel_acc_dyn$SubNr != 3 & vel_acc_dyn$SubNr != 14), # exclude SubNr 3 & 14 as one of the conditions is completely missing
                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
@@
 ```{r vel_acc_dyn_all_box, echo = FALSE}
-p_vel_acc_dyn_all_box <- ggplot(data = vel_acc_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
+p_vel_acc_dyn_all_box <- ggplot(data = subset(vel_acc_dyn, vel_acc_dyn$SubNr != 3 & vel_acc_dyn$SubNr != 14), aes(x = Subcomponent, y = Mean, fill = Condition)) +
   geom_boxplot() +
@@
 ```{r normality2, echo = FALSE}
+# exclude SubNr 3 & 14
+vel_acc_dyn <- subset(vel_acc_dyn, vel_acc_dyn$SubNr != 3 & vel_acc_dyn$SubNr != 14)
 # Reduce unused factors
@@
 ```{r stats3, echo = FALSE}
+df_vel_acc_dyn_exc <- subset(df_vel_acc_dyn, df_vel_acc_dyn$SubNr != 3 & df_vel_acc_dyn$SubNr != 14)
 vel_acc_dyn_aov <- ezANOVA(
-  data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF")
+  data = subset(df_vel_acc_dyn_exc, df_vel_acc_dyn_exc$Subcomponent == "FtoP" | df_vel_acc_dyn_exc$Subcomponent == "PtoF")
   , dv = .(Acc)
@@
 
-# KV-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#vel_dyn_order_group_all}
-```{r vel_dyn_order_all, echo = all}
-# For each participant
-vel_dyn_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_dyn,
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_dyn_order <- cbind(vel_dyn_order[,1:5], as.data.frame(vel_dyn_order[,6]))
-# Change colnames
-colnames(vel_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_dyn_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_dyn_order_all <- cbind(vel_dyn_order_all[,1:4], as.data.frame(vel_dyn_order_all[,5]))
-# Change colnames
-colnames(vel_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_dyn_order_all_bar, echo = FALSE}
-p_vel_dyn_order_all <- ggplot(data = vel_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
-p_vel_dyn_order_all
-```
-
-#### Box plot
-```{r vel_dyn_order_all_box, echo = FALSE}
-p_vel_dyn_order_all_box <- ggplot(data = vel_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
-p_vel_dyn_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality5, echo = FALSE}
-# Normality check
-vel_dyn_order$Subcomponent <- factor(vel_dyn_order$Subcomponent)
-vel_dyn_order_norm <- by(vel_dyn_order$Mean, list(vel_dyn_order$Condition, vel_dyn_order$Subcomponent, vel_dyn_order$Order), shapiro.test)
-print(vel_dyn_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats9, echo = FALSE}
-vel_dyn_order_aov <- ezANOVA(
-  data = df_vel_dyn
-  , dv = .(Velocity)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_dyn_order_aov$ANOVA$F <- round(vel_dyn_order_aov$ANOVA$F, 4)
-vel_dyn_order_aov$ANOVA$p <- round(vel_dyn_order_aov$ANOVA$p, 4)
-print(vel_dyn_order_aov)
-```
-
-# KV-Diff-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#vel_acc_dyn_order_group_all}
-```{r vel_acc_dyn_order_all, echo = all}
-# For each participant
-vel_acc_dyn_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_acc_dyn_order <- cbind(vel_acc_dyn_order[,1:5], as.data.frame(vel_acc_dyn_order[,6]))
-# Change colnames
-colnames(vel_acc_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_acc_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_dyn_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_acc_dyn_order_all <- cbind(vel_acc_dyn_order_all[,1:4], as.data.frame(vel_acc_dyn_order_all[,5]))
-# Change colnames
-colnames(vel_acc_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_acc_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_acc_dyn_order_all_bar, echo = FALSE}
-p_vel_acc_dyn_order_all <- ggplot(data = vel_acc_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
-p_vel_acc_dyn_order_all
-```
-
-#### Box plot
-```{r vel_acc_dyn_order_all_box, echo = FALSE}
-p_vel_acc_dyn_order_all_box <- ggplot(data = vel_acc_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV-Diff: Dynamics") 
-p_vel_acc_dyn_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality6, echo = FALSE}
-# Normality check
-vel_acc_dyn_order$Subcomponent <- factor(vel_acc_dyn_order$Subcomponent)
-vel_acc_dyn_order_norm <- by(vel_acc_dyn_order$Mean, list(vel_acc_dyn_order$Condition, vel_acc_dyn_order$Subcomponent, vel_acc_dyn_order$Order), shapiro.test)
-print(vel_acc_dyn_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats10, echo = FALSE}
-vel_acc_dyn_order_aov <- ezANOVA(
-  data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF")
-  , dv = .(Acc)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_acc_dyn_order_aov$ANOVA$F <- round(vel_acc_dyn_order_aov$ANOVA$F, 4)
-vel_acc_dyn_order_aov$ANOVA$p <- round(vel_acc_dyn_order_aov$ANOVA$p, 4)
-print(vel_acc_dyn_order_aov)
-```
-
-# KV-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#vel_art_order_group_all}
-```{r vel_art_order_all, echo = all}
-# For each participant
-vel_art_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_art,
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_art_order <- cbind(vel_art_order[,1:5], as.data.frame(vel_art_order[,6]))
-# Change colnames
-colnames(vel_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_art_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_art_order_all <- cbind(vel_art_order_all[,1:4], as.data.frame(vel_art_order_all[,5]))
-# Change colnames
-colnames(vel_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_art_order_all_bar, echo = FALSE}
-p_vel_art_order_all <- ggplot(data = vel_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
-p_vel_art_order_all
-```
-
-#### Box plot
-```{r vel_art_order_all_box, echo = FALSE}
-p_vel_art_order_all_box <- ggplot(data = vel_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Articulation") 
-p_vel_art_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality7, echo = FALSE}
-# Normality check
-vel_art_order$Subcomponent <- factor(vel_art_order$Subcomponent)
-vel_art_order_norm <- by(vel_art_order$Mean, list(vel_art_order$Condition, vel_art_order$Subcomponent, vel_art_order$Order), shapiro.test)
-print(vel_art_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats11, echo = FALSE}
-vel_art_order_aov <- ezANOVA(
-  data = df_vel_art
-  , dv = .(Velocity)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_art_order_aov$ANOVA$F <- round(vel_art_order_aov$ANOVA$F, 4)
-vel_art_order_aov$ANOVA$p <- round(vel_art_order_aov$ANOVA$p, 4)
-print(vel_art_order_aov)
-```
-
-# KV-Diff-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#vel_acc_art_order_group_all}
-```{r vel_acc_art_order_all, echo = all}
-# For each participant
-vel_acc_art_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_acc_art_order <- cbind(vel_acc_art_order[,1:5], as.data.frame(vel_acc_art_order[,6]))
-# Change colnames
-colnames(vel_acc_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_acc_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_art_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_acc_art_order_all <- cbind(vel_acc_art_order_all[,1:4], as.data.frame(vel_acc_art_order_all[,5]))
-# Change colnames
-colnames(vel_acc_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_acc_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_acc_art_order_all_bar, echo = FALSE}
-p_vel_acc_art_order_all <- ggplot(data = vel_acc_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
-p_vel_acc_art_order_all
-```
-
-#### Box plot
-```{r vel_acc_art_order_all_box, echo = FALSE}
-p_vel_acc_art_order_all_box <- ggplot(data = vel_acc_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV: Articulation") 
-p_vel_acc_art_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality8, echo = FALSE}
-# Normality check
-vel_acc_art_order$Subcomponent <- factor(vel_acc_art_order$Subcomponent)
-vel_acc_art_order_norm <- by(vel_acc_art_order$Mean, list(vel_acc_art_order$Condition, vel_acc_art_order$Subcomponent, vel_acc_art_order$Order), shapiro.test)
-print(vel_acc_art_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats12, echo = FALSE}
-vel_acc_art_order_aov <- ezANOVA(
-  data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL")
-  , dv = .(Acc)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_acc_art_order_aov$ANOVA$F <- round(vel_acc_art_order_aov$ANOVA$F, 4)
-vel_acc_art_order_aov$ANOVA$p <- round(vel_acc_art_order_aov$ANOVA$p, 4)
-print(vel_acc_art_order_aov)
-```
+<!-- # KV-Order-Dynamics -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#vel_dyn_order_group_all} -->
+<!-- ```{r vel_dyn_order_all, echo = all} -->
+<!-- # For each participant -->
+<!-- vel_dyn_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_dyn, -->
+<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_dyn_order <- cbind(vel_dyn_order[,1:5], as.data.frame(vel_dyn_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_dyn_order, -->
+<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_dyn_order_all <- cbind(vel_dyn_order_all[,1:4], as.data.frame(vel_dyn_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(vel_dyn_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r vel_dyn_order_all_bar, echo = FALSE} -->
+<!-- p_vel_dyn_order_all <- ggplot(data = vel_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100)) -->
+<!-- p_vel_dyn_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r vel_dyn_order_all_box, echo = FALSE} -->
+<!-- p_vel_dyn_order_all_box <- ggplot(data = vel_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KV (0-127)", subtitle = "KV: Dynamics")  -->
+<!-- p_vel_dyn_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality5, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- vel_dyn_order$Subcomponent <- factor(vel_dyn_order$Subcomponent) -->
+<!-- vel_dyn_order_norm <- by(vel_dyn_order$Mean, list(vel_dyn_order$Condition, vel_dyn_order$Subcomponent, vel_dyn_order$Order), shapiro.test) -->
+<!-- print(vel_dyn_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats9, echo = FALSE} -->
+<!-- vel_dyn_order_aov <- ezANOVA( -->
+<!--   data = df_vel_dyn -->
+<!--   , dv = .(Velocity) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- vel_dyn_order_aov$ANOVA$F <- round(vel_dyn_order_aov$ANOVA$F, 4) -->
+<!-- vel_dyn_order_aov$ANOVA$p <- round(vel_dyn_order_aov$ANOVA$p, 4) -->
+<!-- print(vel_dyn_order_aov) -->
+<!-- ``` -->
+
+<!-- # KV-Diff-Order-Dynamics -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#vel_acc_dyn_order_group_all} -->
+<!-- ```{r vel_acc_dyn_order_all, echo = all} -->
+<!-- # For each participant -->
+<!-- vel_acc_dyn_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"), -->
+<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_acc_dyn_order <- cbind(vel_acc_dyn_order[,1:5], as.data.frame(vel_acc_dyn_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_acc_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_acc_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_dyn_order, -->
+<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_acc_dyn_order_all <- cbind(vel_acc_dyn_order_all[,1:4], as.data.frame(vel_acc_dyn_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_acc_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(vel_acc_dyn_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r vel_acc_dyn_order_all_bar, echo = FALSE} -->
+<!-- p_vel_acc_dyn_order_all <- ggplot(data = vel_acc_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "Difference", subtitle = "KV-Diff: Dynamics") -->
+<!-- p_vel_acc_dyn_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r vel_acc_dyn_order_all_box, echo = FALSE} -->
+<!-- p_vel_acc_dyn_order_all_box <- ggplot(data = vel_acc_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "Difference", subtitle = "KV-Diff: Dynamics")  -->
+<!-- p_vel_acc_dyn_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality6, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- vel_acc_dyn_order$Subcomponent <- factor(vel_acc_dyn_order$Subcomponent) -->
+<!-- vel_acc_dyn_order_norm <- by(vel_acc_dyn_order$Mean, list(vel_acc_dyn_order$Condition, vel_acc_dyn_order$Subcomponent, vel_acc_dyn_order$Order), shapiro.test) -->
+<!-- print(vel_acc_dyn_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats10, echo = FALSE} -->
+<!-- vel_acc_dyn_order_aov <- ezANOVA( -->
+<!--   data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF") -->
+<!--   , dv = .(Acc) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- vel_acc_dyn_order_aov$ANOVA$F <- round(vel_acc_dyn_order_aov$ANOVA$F, 4) -->
+<!-- vel_acc_dyn_order_aov$ANOVA$p <- round(vel_acc_dyn_order_aov$ANOVA$p, 4) -->
+<!-- print(vel_acc_dyn_order_aov) -->
+<!-- ``` -->
+
+<!-- # KV-Order-Articulation -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#vel_art_order_group_all} -->
+<!-- ```{r vel_art_order_all, echo = all} -->
+<!-- # For each participant -->
+<!-- vel_art_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_art, -->
+<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_art_order <- cbind(vel_art_order[,1:5], as.data.frame(vel_art_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_art_order, -->
+<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_art_order_all <- cbind(vel_art_order_all[,1:4], as.data.frame(vel_art_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(vel_art_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r vel_art_order_all_bar, echo = FALSE} -->
+<!-- p_vel_art_order_all <- ggplot(data = vel_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100)) -->
+<!-- p_vel_art_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r vel_art_order_all_box, echo = FALSE} -->
+<!-- p_vel_art_order_all_box <- ggplot(data = vel_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KV (0-127)", subtitle = "KV: Articulation")  -->
+<!-- p_vel_art_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality7, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- vel_art_order$Subcomponent <- factor(vel_art_order$Subcomponent) -->
+<!-- vel_art_order_norm <- by(vel_art_order$Mean, list(vel_art_order$Condition, vel_art_order$Subcomponent, vel_art_order$Order), shapiro.test) -->
+<!-- print(vel_art_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats11, echo = FALSE} -->
+<!-- vel_art_order_aov <- ezANOVA( -->
+<!--   data = df_vel_art -->
+<!--   , dv = .(Velocity) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- vel_art_order_aov$ANOVA$F <- round(vel_art_order_aov$ANOVA$F, 4) -->
+<!-- vel_art_order_aov$ANOVA$p <- round(vel_art_order_aov$ANOVA$p, 4) -->
+<!-- print(vel_art_order_aov) -->
+<!-- ``` -->
+
+<!-- # KV-Diff-Order-Articulation -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#vel_acc_art_order_group_all} -->
+<!-- ```{r vel_acc_art_order_all, echo = all} -->
+<!-- # For each participant -->
+<!-- vel_acc_art_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"), -->
+<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_acc_art_order <- cbind(vel_acc_art_order[,1:5], as.data.frame(vel_acc_art_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_acc_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_acc_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_art_order, -->
+<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_acc_art_order_all <- cbind(vel_acc_art_order_all[,1:4], as.data.frame(vel_acc_art_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_acc_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(vel_acc_art_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r vel_acc_art_order_all_bar, echo = FALSE} -->
+<!-- p_vel_acc_art_order_all <- ggplot(data = vel_acc_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "Difference", subtitle = "KV-Diff: Articulation") -->
+<!-- p_vel_acc_art_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r vel_acc_art_order_all_box, echo = FALSE} -->
+<!-- p_vel_acc_art_order_all_box <- ggplot(data = vel_acc_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "Difference", subtitle = "KV: Articulation")  -->
+<!-- p_vel_acc_art_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality8, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- vel_acc_art_order$Subcomponent <- factor(vel_acc_art_order$Subcomponent) -->
+<!-- vel_acc_art_order_norm <- by(vel_acc_art_order$Mean, list(vel_acc_art_order$Condition, vel_acc_art_order$Subcomponent, vel_acc_art_order$Order), shapiro.test) -->
+<!-- print(vel_acc_art_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats12, echo = FALSE} -->
+<!-- vel_acc_art_order_aov <- ezANOVA( -->
+<!--   data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL") -->
+<!--   , dv = .(Acc) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- vel_acc_art_order_aov$ANOVA$F <- round(vel_acc_art_order_aov$ANOVA$F, 4) -->
+<!-- vel_acc_art_order_aov$ANOVA$p <- round(vel_acc_art_order_aov$ANOVA$p, 4) -->
+<!-- print(vel_acc_art_order_aov) -->
+<!-- ``` -->
 

### 2020-09-25 script/RNotebook/articulation.Rmd
update analysis scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-07-28 script/RNotebook/articulation.Rmd
@@
 
+```{r file, include = FALSE}
+filename_ioi = "../R/preprocessor/trimmed/data_ioi_1.txt"
+filename_kot = "../R/preprocessor/trimmed/data_kot.txt"
+```
+
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-df_kot <- read.csv("./trimmed/data_kot.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_ioi <- read.csv(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
+df_kot <- read.csv(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed csv
 
@@
 df_kot$SubNr <- as.factor(df_kot$SubNr)
+
+# exclude participants
+df_ioi <- df_ioi %>% dplyr::filter(SubNr != 1 & SubNr != 3 & SubNr != 14 & SubNr != 16)
+df_kot <- df_kot %>% dplyr::filter(SubNr != 1 & SubNr != 3 & SubNr != 14 & SubNr != 16)
 
@@
 
-# KOT-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#kot_art_order_group_all}
-```{r kot_art_order_all, echo = FALSE}
-# For each participant
-kot_art_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-kot_art_order <- cbind(kot_art_order[,1:5], as.data.frame(kot_art_order[,6]))
-# Change colnames
-colnames(kot_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-kot_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_art_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-kot_art_order_all <- cbind(kot_art_order_all[,1:4], as.data.frame(kot_art_order_all[,5]))
-# Change colnames
-colnames(kot_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(kot_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r kot_art_order_all_bar, echo = FALSE}
-p_kot_art_order_all <- ggplot(data = kot_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
-p_kot_art_order_all
-```
-
-#### Box plot
-```{r kot_art_order_all_box, echo = FALSE}
-p_kot_art_order_all_box <- ggplot(data = kot_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KOT", subtitle = "KOT: Articulation")
-p_kot_art_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality5, echo = FALSE}
-# Normality check
-kot_art_order$Subcomponent <- factor(kot_art_order$Subcomponent)
-kot_art_order_norm <- by(kot_art_order$Mean, list(kot_art_order$Condition, kot_art_order$Subcomponent, kot_art_order$Order), shapiro.test)
-print(kot_art_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats9, echo = FALSE}
-kot_art_order_aov <- ezANOVA(
-  data = df_kot_art
-  , dv = .(KOT)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-kot_art_order_aov$ANOVA$F <- round(kot_art_order_aov$ANOVA$F, 4)
-kot_art_order_aov$ANOVA$p <- round(kot_art_order_aov$ANOVA$p, 4)
-print(kot_art_order_aov)
-```
+<!-- # KOT-Order-Articulation -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#kot_art_order_group_all} -->
+<!-- ```{r kot_art_order_all, echo = FALSE} -->
+<!-- # For each participant -->
+<!-- kot_art_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)}) -->
+<!-- kot_art_order <- cbind(kot_art_order[,1:5], as.data.frame(kot_art_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(kot_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- kot_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_art_order, -->
+<!--                                      FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- kot_art_order_all <- cbind(kot_art_order_all[,1:4], as.data.frame(kot_art_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(kot_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(kot_art_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r kot_art_order_all_bar, echo = FALSE} -->
+<!-- p_kot_art_order_all <- ggplot(data = kot_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KOT (ms)", subtitle = "KOT: Articulation") -->
+<!-- p_kot_art_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r kot_art_order_all_box, echo = FALSE} -->
+<!-- p_kot_art_order_all_box <- ggplot(data = kot_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KOT", subtitle = "KOT: Articulation") -->
+<!-- p_kot_art_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality5, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- kot_art_order$Subcomponent <- factor(kot_art_order$Subcomponent) -->
+<!-- kot_art_order_norm <- by(kot_art_order$Mean, list(kot_art_order$Condition, kot_art_order$Subcomponent, kot_art_order$Order), shapiro.test) -->
+<!-- print(kot_art_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats9, echo = FALSE} -->
+<!-- kot_art_order_aov <- ezANOVA( -->
+<!--   data = df_kot_art -->
+<!--   , dv = .(KOT) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- kot_art_order_aov$ANOVA$F <- round(kot_art_order_aov$ANOVA$F, 4) -->
+<!-- kot_art_order_aov$ANOVA$p <- round(kot_art_order_aov$ANOVA$p, 4) -->
+<!-- print(kot_art_order_aov) -->
+<!-- ``` -->
 

### 2020-09-25 script/RNotebook/dynamics.Rmd
update analysis scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-07-28 script/RNotebook/dynamics.Rmd
@@
 
+```{r file, include = FALSE}
+filename_vel = "../R/preprocessor/trimmed/data_vel.txt"
+filename_vel_acc = "../R/preprocessor/trimmed/data_vel_diff.txt"
+```
+
 ```{r extract, include = FALSE}
-df_vel <- read.csv("./trimmed/data_vel.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-df_vel_acc <- read.csv("./trimmed/data_vel_acc.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_vel <- read.csv(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed csv
+df_vel_acc <- read.csv(filename_vel_acc, header = T, sep = ",", dec = ".") # read a trimmed csv
 
@@
 df_vel_acc$SubNr <- as.factor(df_vel_acc$SubNr)
+
+# exclude participants
+df_vel <- df_vel %>% dplyr::filter(SubNr != 1 & SubNr != 3 & SubNr != 14 & SubNr != 16)
+df_vel_acc <- df_vel_acc %>% dplyr::filter(SubNr != 1 & SubNr != 3 & SubNr != 14 & SubNr != 16)
 
@@
 # For each individual
-vel_acc_dyn <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
+vel_acc_dyn <- aggregate(Diff~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
   data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF")
-  , dv = .(Acc)
+  , dv = .(Diff)
   , wid = .(SubNr)
@@
 # For each individual
-vel_acc_dyn_trial <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
+vel_acc_dyn_trial <- aggregate(Diff~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 # For each individual
-vel_acc_art <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
+vel_acc_art <- aggregate(Diff~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
   data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL")
-  , dv = .(Acc)
+  , dv = .(Diff)
   , wid = .(SubNr)
@@
 # For each individual
-vel_acc_art_trial <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
+vel_acc_art_trial <- aggregate(Diff~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
@@
 
-# KV-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#vel_dyn_order_group_all}
-```{r vel_dyn_order_all, echo = all}
-# For each participant
-vel_dyn_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_dyn,
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_dyn_order <- cbind(vel_dyn_order[,1:5], as.data.frame(vel_dyn_order[,6]))
-# Change colnames
-colnames(vel_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_dyn_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_dyn_order_all <- cbind(vel_dyn_order_all[,1:4], as.data.frame(vel_dyn_order_all[,5]))
-# Change colnames
-colnames(vel_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_dyn_order_all_bar, echo = FALSE}
-p_vel_dyn_order_all <- ggplot(data = vel_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
-p_vel_dyn_order_all
-```
-
-#### Box plot
-```{r vel_dyn_order_all_box, echo = FALSE}
-p_vel_dyn_order_all_box <- ggplot(data = vel_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") 
-p_vel_dyn_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality5, echo = FALSE}
-# Normality check
-vel_dyn_order$Subcomponent <- factor(vel_dyn_order$Subcomponent)
-vel_dyn_order_norm <- by(vel_dyn_order$Mean, list(vel_dyn_order$Condition, vel_dyn_order$Subcomponent, vel_dyn_order$Order), shapiro.test)
-print(vel_dyn_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats9, echo = FALSE}
-vel_dyn_order_aov <- ezANOVA(
-  data = df_vel_dyn
-  , dv = .(Velocity)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_dyn_order_aov$ANOVA$F <- round(vel_dyn_order_aov$ANOVA$F, 4)
-vel_dyn_order_aov$ANOVA$p <- round(vel_dyn_order_aov$ANOVA$p, 4)
-print(vel_dyn_order_aov)
-```
-
-# KV-Diff-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#vel_acc_dyn_order_group_all}
-```{r vel_acc_dyn_order_all, echo = all}
-# For each participant
-vel_acc_dyn_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_acc_dyn_order <- cbind(vel_acc_dyn_order[,1:5], as.data.frame(vel_acc_dyn_order[,6]))
-# Change colnames
-colnames(vel_acc_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_acc_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_dyn_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_acc_dyn_order_all <- cbind(vel_acc_dyn_order_all[,1:4], as.data.frame(vel_acc_dyn_order_all[,5]))
-# Change colnames
-colnames(vel_acc_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_acc_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_acc_dyn_order_all_bar, echo = FALSE}
-p_vel_acc_dyn_order_all <- ggplot(data = vel_acc_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
-p_vel_acc_dyn_order_all
-```
-
-#### Box plot
-```{r vel_acc_dyn_order_all_box, echo = FALSE}
-p_vel_acc_dyn_order_all_box <- ggplot(data = vel_acc_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV-Diff: Dynamics") 
-p_vel_acc_dyn_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality6, echo = FALSE}
-# Normality check
-vel_acc_dyn_order$Subcomponent <- factor(vel_acc_dyn_order$Subcomponent)
-vel_acc_dyn_order_norm <- by(vel_acc_dyn_order$Mean, list(vel_acc_dyn_order$Condition, vel_acc_dyn_order$Subcomponent, vel_acc_dyn_order$Order), shapiro.test)
-print(vel_acc_dyn_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats10, echo = FALSE}
-vel_acc_dyn_order_aov <- ezANOVA(
-  data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF")
-  , dv = .(Acc)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_acc_dyn_order_aov$ANOVA$F <- round(vel_acc_dyn_order_aov$ANOVA$F, 4)
-vel_acc_dyn_order_aov$ANOVA$p <- round(vel_acc_dyn_order_aov$ANOVA$p, 4)
-print(vel_acc_dyn_order_aov)
-```
-
-# KV-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#vel_art_order_group_all}
-```{r vel_art_order_all, echo = all}
-# For each participant
-vel_art_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_art,
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_art_order <- cbind(vel_art_order[,1:5], as.data.frame(vel_art_order[,6]))
-# Change colnames
-colnames(vel_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_art_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_art_order_all <- cbind(vel_art_order_all[,1:4], as.data.frame(vel_art_order_all[,5]))
-# Change colnames
-colnames(vel_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_art_order_all_bar, echo = FALSE}
-p_vel_art_order_all <- ggplot(data = vel_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
-p_vel_art_order_all
-```
-
-#### Box plot
-```{r vel_art_order_all_box, echo = FALSE}
-p_vel_art_order_all_box <- ggplot(data = vel_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "KV (0-127)", subtitle = "KV: Articulation") 
-p_vel_art_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality7, echo = FALSE}
-# Normality check
-vel_art_order$Subcomponent <- factor(vel_art_order$Subcomponent)
-vel_art_order_norm <- by(vel_art_order$Mean, list(vel_art_order$Condition, vel_art_order$Subcomponent, vel_art_order$Order), shapiro.test)
-print(vel_art_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats11, echo = FALSE}
-vel_art_order_aov <- ezANOVA(
-  data = df_vel_art
-  , dv = .(Velocity)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_art_order_aov$ANOVA$F <- round(vel_art_order_aov$ANOVA$F, 4)
-vel_art_order_aov$ANOVA$p <- round(vel_art_order_aov$ANOVA$p, 4)
-print(vel_art_order_aov)
-```
-
-# KV-Diff-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#vel_acc_art_order_group_all}
-```{r vel_acc_art_order_all, echo = all}
-# For each participant
-vel_acc_art_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
-                               FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_acc_art_order <- cbind(vel_acc_art_order[,1:5], as.data.frame(vel_acc_art_order[,6]))
-# Change colnames
-colnames(vel_acc_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")
-
-# Group mean
-vel_acc_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_art_order,
-                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_acc_art_order_all <- cbind(vel_acc_art_order_all[,1:4], as.data.frame(vel_acc_art_order_all[,5]))
-# Change colnames
-colnames(vel_acc_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
-print(vel_acc_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r vel_acc_art_order_all_bar, echo = FALSE}
-p_vel_acc_art_order_all <- ggplot(data = vel_acc_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge(.9)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
-p_vel_acc_art_order_all
-```
-
-#### Box plot
-```{r vel_acc_art_order_all_box, echo = FALSE}
-p_vel_acc_art_order_all_box <- ggplot(data = vel_acc_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
-  facet_wrap(Order ~ .) +
-  labs(y = "Difference", subtitle = "KV: Articulation") 
-p_vel_acc_art_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality8, echo = FALSE}
-# Normality check
-vel_acc_art_order$Subcomponent <- factor(vel_acc_art_order$Subcomponent)
-vel_acc_art_order_norm <- by(vel_acc_art_order$Mean, list(vel_acc_art_order$Condition, vel_acc_art_order$Subcomponent, vel_acc_art_order$Order), shapiro.test)
-print(vel_acc_art_order_norm)
-```
-
-#### Three-way ANOVA (ez)
-```{r stats12, echo = FALSE}
-vel_acc_art_order_aov <- ezANOVA(
-  data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL")
-  , dv = .(Acc)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition, Subcomponent)
-  , type = 3
-  , detailed = TRUE
-)
-vel_acc_art_order_aov$ANOVA$F <- round(vel_acc_art_order_aov$ANOVA$F, 4)
-vel_acc_art_order_aov$ANOVA$p <- round(vel_acc_art_order_aov$ANOVA$p, 4)
-print(vel_acc_art_order_aov)
-```
-
-# Sequence plot
-### Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 20, fig.height = 4, echo = FALSE}
-# Average vel_dynocity for each note
-# For each participant
-vel_dyn_seq <- aggregate(Velocity~SubNr*Condition*Skill*Note, data = df_vel_dyn, 
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_dyn_seq <- cbind(vel_dyn_seq[,1:4], as.data.frame(vel_dyn_seq[,5]))
-# Change colnames
-colnames(vel_dyn_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD")
-
-# Group mean
-vel_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_dyn_seq,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_dyn_seq_stats <- cbind(vel_dyn_seq_stats[,1:3], as.data.frame(vel_dyn_seq_stats[,4]))
-# Change colnames
-colnames(vel_dyn_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM")
-
-# Average acceleration for each note
-# For each participant
-vel_dyn_acc_seq <- aggregate(Acc~SubNr*Condition*Skill*Interval, data = df_vel_acc_dyn, 
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_dyn_acc_seq <- cbind(vel_dyn_acc_seq[,1:4], as.data.frame(vel_dyn_acc_seq[,5]))
-# Change colnames
-colnames(vel_dyn_acc_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
-
-# Group mean
-vel_dyn_acc_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_dyn_acc_seq,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_dyn_acc_seq_stats <- cbind(vel_dyn_acc_seq_stats[,1:3], as.data.frame(vel_dyn_acc_seq_stats[,4]))
-# Change colnames
-colnames(vel_dyn_acc_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
-
-p_vel_dyn_seq <- ggplot(data = vel_dyn_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
-                position = position_dodge(0.05)) + 
-  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
-p_vel_dyn_seq
-
-p_vel_dyn_acc_seq <- ggplot(data = vel_dyn_acc_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
-                position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
-p_vel_dyn_acc_seq
-```
-
-### Articulation{#sequence_art}
-```{r seq_art, fig.width = 20, fig.height = 4, echo = FALSE}
-# Average vel_artocity for each note
-# For each participant
-vel_art_seq <- aggregate(Velocity~SubNr*Condition*Skill*Note, data = df_vel_art, 
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_art_seq <- cbind(vel_art_seq[,1:4], as.data.frame(vel_art_seq[,5]))
-# Change colnames
-colnames(vel_art_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD")
-
-# Group mean
-vel_art_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_art_seq,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_art_seq_stats <- cbind(vel_art_seq_stats[,1:3], as.data.frame(vel_art_seq_stats[,4]))
-# Change colnames
-colnames(vel_art_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM")
-
-# Average acceleration for each note
-# For each participant
-vel_art_acc_seq <- aggregate(Acc~SubNr*Condition*Skill*Interval, data = df_vel_acc_art, 
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-vel_art_acc_seq <- cbind(vel_art_acc_seq[,1:4], as.data.frame(vel_art_acc_seq[,5]))
-# Change colnames
-colnames(vel_art_acc_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
-
-# Group mean
-vel_art_acc_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_art_acc_seq,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-vel_art_acc_seq_stats <- cbind(vel_art_acc_seq_stats[,1:3], as.data.frame(vel_art_acc_seq_stats[,4]))
-# Change colnames
-colnames(vel_art_acc_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
-
-p_vel_art_seq <- ggplot(data = vel_art_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
-                position = position_dodge(0.05)) + 
-  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
-p_vel_art_seq
-
-p_vel_art_acc_seq <- ggplot(data = vel_art_acc_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
-                position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
-p_vel_art_acc_seq
-```
+<!-- # KV-Order-Dynamics -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#vel_dyn_order_group_all} -->
+<!-- ```{r vel_dyn_order_all, echo = all} -->
+<!-- # For each participant -->
+<!-- vel_dyn_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_dyn, -->
+<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_dyn_order <- cbind(vel_dyn_order[,1:5], as.data.frame(vel_dyn_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_dyn_order, -->
+<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_dyn_order_all <- cbind(vel_dyn_order_all[,1:4], as.data.frame(vel_dyn_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(vel_dyn_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r vel_dyn_order_all_bar, echo = FALSE} -->
+<!-- p_vel_dyn_order_all <- ggplot(data = vel_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100)) -->
+<!-- p_vel_dyn_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r vel_dyn_order_all_box, echo = FALSE} -->
+<!-- p_vel_dyn_order_all_box <- ggplot(data = vel_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KV (0-127)", subtitle = "KV: Dynamics")  -->
+<!-- p_vel_dyn_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality5, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- vel_dyn_order$Subcomponent <- factor(vel_dyn_order$Subcomponent) -->
+<!-- vel_dyn_order_norm <- by(vel_dyn_order$Mean, list(vel_dyn_order$Condition, vel_dyn_order$Subcomponent, vel_dyn_order$Order), shapiro.test) -->
+<!-- print(vel_dyn_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats9, echo = FALSE} -->
+<!-- vel_dyn_order_aov <- ezANOVA( -->
+<!--   data = df_vel_dyn -->
+<!--   , dv = .(Velocity) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- vel_dyn_order_aov$ANOVA$F <- round(vel_dyn_order_aov$ANOVA$F, 4) -->
+<!-- vel_dyn_order_aov$ANOVA$p <- round(vel_dyn_order_aov$ANOVA$p, 4) -->
+<!-- print(vel_dyn_order_aov) -->
+<!-- ``` -->
+
+<!-- # KV-Diff-Order-Dynamics -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#vel_acc_dyn_order_group_all} -->
+<!-- ```{r vel_acc_dyn_order_all, echo = all} -->
+<!-- # For each participant -->
+<!-- vel_acc_dyn_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"), -->
+<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_acc_dyn_order <- cbind(vel_acc_dyn_order[,1:5], as.data.frame(vel_acc_dyn_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_acc_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_acc_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_dyn_order, -->
+<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_acc_dyn_order_all <- cbind(vel_acc_dyn_order_all[,1:4], as.data.frame(vel_acc_dyn_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_acc_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(vel_acc_dyn_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r vel_acc_dyn_order_all_bar, echo = FALSE} -->
+<!-- p_vel_acc_dyn_order_all <- ggplot(data = vel_acc_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "Difference", subtitle = "KV-Diff: Dynamics") -->
+<!-- p_vel_acc_dyn_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r vel_acc_dyn_order_all_box, echo = FALSE} -->
+<!-- p_vel_acc_dyn_order_all_box <- ggplot(data = vel_acc_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "Difference", subtitle = "KV-Diff: Dynamics")  -->
+<!-- p_vel_acc_dyn_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality6, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- vel_acc_dyn_order$Subcomponent <- factor(vel_acc_dyn_order$Subcomponent) -->
+<!-- vel_acc_dyn_order_norm <- by(vel_acc_dyn_order$Mean, list(vel_acc_dyn_order$Condition, vel_acc_dyn_order$Subcomponent, vel_acc_dyn_order$Order), shapiro.test) -->
+<!-- print(vel_acc_dyn_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats10, echo = FALSE} -->
+<!-- vel_acc_dyn_order_aov <- ezANOVA( -->
+<!--   data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF") -->
+<!--   , dv = .(Acc) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- vel_acc_dyn_order_aov$ANOVA$F <- round(vel_acc_dyn_order_aov$ANOVA$F, 4) -->
+<!-- vel_acc_dyn_order_aov$ANOVA$p <- round(vel_acc_dyn_order_aov$ANOVA$p, 4) -->
+<!-- print(vel_acc_dyn_order_aov) -->
+<!-- ``` -->
+
+<!-- # KV-Order-Articulation -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#vel_art_order_group_all} -->
+<!-- ```{r vel_art_order_all, echo = all} -->
+<!-- # For each participant -->
+<!-- vel_art_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_art, -->
+<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_art_order <- cbind(vel_art_order[,1:5], as.data.frame(vel_art_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_art_order, -->
+<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_art_order_all <- cbind(vel_art_order_all[,1:4], as.data.frame(vel_art_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(vel_art_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r vel_art_order_all_bar, echo = FALSE} -->
+<!-- p_vel_art_order_all <- ggplot(data = vel_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100)) -->
+<!-- p_vel_art_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r vel_art_order_all_box, echo = FALSE} -->
+<!-- p_vel_art_order_all_box <- ggplot(data = vel_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "KV (0-127)", subtitle = "KV: Articulation")  -->
+<!-- p_vel_art_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality7, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- vel_art_order$Subcomponent <- factor(vel_art_order$Subcomponent) -->
+<!-- vel_art_order_norm <- by(vel_art_order$Mean, list(vel_art_order$Condition, vel_art_order$Subcomponent, vel_art_order$Order), shapiro.test) -->
+<!-- print(vel_art_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats11, echo = FALSE} -->
+<!-- vel_art_order_aov <- ezANOVA( -->
+<!--   data = df_vel_art -->
+<!--   , dv = .(Velocity) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- vel_art_order_aov$ANOVA$F <- round(vel_art_order_aov$ANOVA$F, 4) -->
+<!-- vel_art_order_aov$ANOVA$p <- round(vel_art_order_aov$ANOVA$p, 4) -->
+<!-- print(vel_art_order_aov) -->
+<!-- ``` -->
+
+<!-- # KV-Diff-Order-Articulation -->
+<!-- - Investigate whether there is any order effect of the conditions. -->
+
+<!-- ## 1. All{#vel_acc_art_order_group_all} -->
+<!-- ```{r vel_acc_art_order_all, echo = all} -->
+<!-- # For each participant -->
+<!-- vel_acc_art_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"), -->
+<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_acc_art_order <- cbind(vel_acc_art_order[,1:5], as.data.frame(vel_acc_art_order[,6])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_acc_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_acc_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_art_order, -->
+<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_acc_art_order_all <- cbind(vel_acc_art_order_all[,1:4], as.data.frame(vel_acc_art_order_all[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_acc_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
+<!-- print(vel_acc_art_order_all) -->
+<!-- ``` -->
+
+<!-- ### Plot -->
+<!-- #### Bar graph -->
+<!-- - Error bars show SEM of the mean -->
+
+<!-- ```{r vel_acc_art_order_all_bar, echo = FALSE} -->
+<!-- p_vel_acc_art_order_all <- ggplot(data = vel_acc_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
+<!--                 width = .2, position = position_dodge(.9)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "Difference", subtitle = "KV-Diff: Articulation") -->
+<!-- p_vel_acc_art_order_all -->
+<!-- ``` -->
+
+<!-- #### Box plot -->
+<!-- ```{r vel_acc_art_order_all_box, echo = FALSE} -->
+<!-- p_vel_acc_art_order_all_box <- ggplot(data = vel_acc_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
+<!--   geom_boxplot() + -->
+<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
+<!--   facet_wrap(Order ~ .) + -->
+<!--   labs(y = "Difference", subtitle = "KV: Articulation")  -->
+<!-- p_vel_acc_art_order_all_box -->
+<!-- ``` -->
+
+<!-- ### Statistics -->
+<!-- #### Normality check -->
+<!-- ```{r normality8, echo = FALSE} -->
+<!-- # Normality check -->
+<!-- vel_acc_art_order$Subcomponent <- factor(vel_acc_art_order$Subcomponent) -->
+<!-- vel_acc_art_order_norm <- by(vel_acc_art_order$Mean, list(vel_acc_art_order$Condition, vel_acc_art_order$Subcomponent, vel_acc_art_order$Order), shapiro.test) -->
+<!-- print(vel_acc_art_order_norm) -->
+<!-- ``` -->
+
+<!-- #### Three-way ANOVA (ez) -->
+<!-- ```{r stats12, echo = FALSE} -->
+<!-- vel_acc_art_order_aov <- ezANOVA( -->
+<!--   data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL") -->
+<!--   , dv = .(Acc) -->
+<!--   , wid = .(SubNr) -->
+<!--   , between = .(Order) -->
+<!--   , within = .(Condition, Subcomponent) -->
+<!--   , type = 3 -->
+<!--   , detailed = TRUE -->
+<!-- ) -->
+<!-- vel_acc_art_order_aov$ANOVA$F <- round(vel_acc_art_order_aov$ANOVA$F, 4) -->
+<!-- vel_acc_art_order_aov$ANOVA$p <- round(vel_acc_art_order_aov$ANOVA$p, 4) -->
+<!-- print(vel_acc_art_order_aov) -->
+<!-- ``` -->
+
+<!-- # Sequence plot -->
+<!-- ### Dynamics{#sequence_dyn} -->
+<!-- ```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE} -->
+<!-- # Average vel_dynocity for each note -->
+<!-- # For each participant -->
+<!-- vel_dyn_seq <- aggregate(Velocity~SubNr*Condition*Skill*Note, data = df_vel_dyn,  -->
+<!--                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_dyn_seq <- cbind(vel_dyn_seq[,1:4], as.data.frame(vel_dyn_seq[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_dyn_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_dyn_seq, -->
+<!--                            FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_dyn_seq_stats <- cbind(vel_dyn_seq_stats[,1:3], as.data.frame(vel_dyn_seq_stats[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_dyn_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM") -->
+
+<!-- # Average acceleration for each note -->
+<!-- # For each participant -->
+<!-- vel_dyn_acc_seq <- aggregate(Acc~SubNr*Condition*Skill*Interval, data = df_vel_acc_dyn,  -->
+<!--                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_dyn_acc_seq <- cbind(vel_dyn_acc_seq[,1:4], as.data.frame(vel_dyn_acc_seq[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_dyn_acc_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_dyn_acc_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_dyn_acc_seq, -->
+<!--                            FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_dyn_acc_seq_stats <- cbind(vel_dyn_acc_seq_stats[,1:3], as.data.frame(vel_dyn_acc_seq_stats[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_dyn_acc_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM") -->
+
+<!-- p_vel_dyn_seq <- ggplot(data = vel_dyn_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) + -->
+<!--   geom_line() + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2, -->
+<!--                 position = position_dodge(0.05)) +  -->
+<!--   labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1)) -->
+<!-- p_vel_dyn_seq -->
+
+<!-- p_vel_dyn_acc_seq <- ggplot(data = vel_dyn_acc_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) + -->
+<!--   geom_line() + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2, -->
+<!--                 position = position_dodge(.05)) +  -->
+<!--   labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1)) -->
+<!-- p_vel_dyn_acc_seq -->
+<!-- ``` -->
+
+<!-- ### Articulation{#sequence_art} -->
+<!-- ```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE} -->
+<!-- # Average vel_artocity for each note -->
+<!-- # For each participant -->
+<!-- vel_art_seq <- aggregate(Velocity~SubNr*Condition*Skill*Note, data = df_vel_art,  -->
+<!--                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_art_seq <- cbind(vel_art_seq[,1:4], as.data.frame(vel_art_seq[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_art_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_art_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_art_seq, -->
+<!--                            FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_art_seq_stats <- cbind(vel_art_seq_stats[,1:3], as.data.frame(vel_art_seq_stats[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_art_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM") -->
+
+<!-- # Average acceleration for each note -->
+<!-- # For each participant -->
+<!-- vel_art_acc_seq <- aggregate(Acc~SubNr*Condition*Skill*Interval, data = df_vel_acc_art,  -->
+<!--                      FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
+<!-- vel_art_acc_seq <- cbind(vel_art_acc_seq[,1:4], as.data.frame(vel_art_acc_seq[,5])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_art_acc_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD") -->
+
+<!-- # Group mean -->
+<!-- vel_art_acc_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_art_acc_seq, -->
+<!--                            FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
+<!-- vel_art_acc_seq_stats <- cbind(vel_art_acc_seq_stats[,1:3], as.data.frame(vel_art_acc_seq_stats[,4])) -->
+<!-- # Change colnames -->
+<!-- colnames(vel_art_acc_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM") -->
+
+<!-- p_vel_art_seq <- ggplot(data = vel_art_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) + -->
+<!--   geom_line() + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2, -->
+<!--                 position = position_dodge(0.05)) +  -->
+<!--   labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1)) -->
+<!-- p_vel_art_seq -->
+
+<!-- p_vel_art_acc_seq <- ggplot(data = vel_art_acc_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) + -->
+<!--   geom_line() + -->
+<!--   geom_point() + -->
+<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2, -->
+<!--                 position = position_dodge(.05)) +  -->
+<!--   labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1)) -->
+<!-- p_vel_art_acc_seq -->
+<!-- ``` -->

### 2020-09-25 script/RNotebook/ioi.Rmd
update analysis scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-07-28 script/RNotebook/ioi.Rmd
@@
 
+```{r file, include = FALSE}
+filename_ioi = "../R/preprocessor/trimmed/data_ioi_1.txt"
+```
+
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+df_ioi <- read.csv(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
@@
 df_ioi$SubNr <- as.factor(df_ioi$SubNr)
+
+# exclude participants
+df_ioi <- df_ioi %>% dplyr::filter(SubNr != 1 & SubNr != 3 & SubNr != 14 & SubNr != 16)
 

### 2020-11-16 script/R/preprocessor/error_summary.Rmd
error summary
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Error Summary (teaching-v1.0)"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("effsize")) {install.packages("effsize"); require("effsize")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# function
source("./function.R")

# read a text file for ideal performance
dt_ideal <- read.table("./ideal.txt")
colnames(dt_ideal) <- "Pitch"
dt_ideal$RowNr <- 1:nrow(dt_ideal)
setcolorder(dt_ideal, c(2, 1))
# 
dt_onset <- fread(file = "./filtered/dt_correct_onset.txt", sep = ",")
dt_offset <- fread(file = "./filtered/dt_correct_offset.txt", sep = ",")
error_onset <- fread(file = "./filtered/error_onset.txt", sep = ",")
error_offset <- fread(file = "./filtered/error_offset.txt", sep = ",")
```

- Each performance (1 trial) contains 72 responses for onsets and 72 responses for offsets (i.e., if there is no mistake, 144 responses in total)
- Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)

# Outline
1. [Onset reports](#onset)
2. [Offset reports](#offset)

## ONSET
### Corrected Error Onset (All)
```{r onset_all, echo = FALSE}
error_onset_all <- error_onset[CorrectionNr != "Exclude",]
error_onset_all
```

### Corrected Error Onset (Summary)
```{r onset_summary, echo = FALSE}
error_onset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
error_onset_summary <- error_onset_all[, .(N = .N, Sum = sum(as.numeric(CorrectionNr))), by = .(Reason)]
error_onset_summary
```

### Missing/Excluded Trials Onset (All)
```{r na_onset_all, echo = FALSE}
missing_onset <- checker(dt_onset, dt_ideal)
missing_onset
```

1. **2 participants should be removed from the entire data analysis because they missed at least one block (16 trials)**

- SubNr 3: experimental error (8 trials)
- SubNr 14: played in octave higher (8 trials)

2. **some performances were removed because 1) participants didn't follow sheet music or 2) participants didn't finish till the end (5 trials)**

### Missing/Excluded Trials Onset (Summary)
```{r na_onset_summary, echo = FALSE}
missing_onset_summary <- missing_onset[, .(N = .N), by = .(Reason)]
missing_onset_summary
```

## OFFSET
### Corrected Error Offset (All)
```{r offset_all, echo = FALSE}
error_offset_all <- error_offset[CorrectionNr != "Exclude",]

# separate errors depending on reasons
# SubNr 11, BlockNr 4, TrialNr 8
error_offset_all <- rbind(error_offset_all, error_offset_all[4])
error_offset_all$Reason[4] <- "Extra Notes"
error_offset_all$CorrectionNr[4] <- 1
error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 2
# SubNr 27, BlockNr 3, TrialNr 6
error_offset_all <- rbind(error_offset_all, error_offset_all[5])
error_offset_all$Reason[5] <- "Extra Notes"
error_offset_all$CorrectionNr[5] <- 1
error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 6

error_offset_all
```

### Corrected Error Offset (Summary)
```{r offset_summary, echo = FALSE}
error_offset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
error_offset_summary <- error_offset_all[, .(N = .N, Sum = sum(as.numeric(CorrectionNr))), by = .(Reason)]
error_offset_summary
```

### Missing/Excluded Trials Offset (All)
```{r na_offset_all, echo = FALSE}
missing_offset <- checker(dt_offset, dt_ideal)
missing_offset
```

1. **2 participants should be removed from the entire data analysis because they missed at least one block (16 trials)**

- SubNr 3: experimental error (8 trials)
- SubNr 14: played in octave higher (8 trials)

2. **some performances were removed because 1) participants didn't follow sheet music or 2) participants didn't finish till the end (5 trials)**

### Missing/Excluded Trials Offset (Summary)
```{r na_offset_summary, echo = FALSE}
missing_offset_summary <- missing_offset[, .(N = .N), by = .(Reason)]
missing_offset_summary
```

### 2020-11-16 script/R/preprocessor/error_summary.Rmd
72 responses to 67 responses
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-- Each performance (1 trial) contains 72 responses for onsets and 72 responses for offsets (i.e., if there is no mistake, 144 responses in total)
+- Each performance (1 trial) contains 67 responses for onsets and 67 responses for offsets (i.e., if there is no mistake, 134 responses in total)
 - Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)

### 2020-11-16 script/R/preprocessor/error_summary.Rmd
upload summary
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-11-16 script/R/preprocessor/error_summary.Rmd
@@
 ---
-title: "Error Summary (teaching-v1.0)"
+title: "Error Summary (teaching-v2.0)"
 output: html_notebook
@@
 error_onset_all <- error_onset[CorrectionNr != "Exclude",]
+
+# separate errors depending on reasons
+# SubNr 10, BlockNr 1, TrialNr 6
+error_onset_all <- rbind(error_onset_all, error_onset_all[2])
+error_onset_all$Reason[2] <- "Extra Notes"
+error_onset_all$CorrectionNr[2] <- 2
+error_onset_all$Reason[nrow(error_onset_all)] <- "Substituted"
+error_onset_all$CorrectionNr[nrow(error_onset_all)] <- 1
+
 error_onset_all
@@
 
-1. **2 participants should be removed from the entire data analysis because they missed at least one block (16 trials)**
-
-- SubNr 3: experimental error (8 trials)
-- SubNr 14: played in octave higher (8 trials)
-
-2. **some performances were removed because 1) participants didn't follow sheet music or 2) participants didn't finish till the end (5 trials)**
-
 ### Missing/Excluded Trials Onset (Summary)
@@
 # separate errors depending on reasons
-# SubNr 11, BlockNr 4, TrialNr 8
-error_offset_all <- rbind(error_offset_all, error_offset_all[4])
-error_offset_all$Reason[4] <- "Extra Notes"
-error_offset_all$CorrectionNr[4] <- 1
+# SubNr 1, BlockNr 2, TrialNr 3
+error_offset_all <- rbind(error_offset_all, error_offset_all[1])
+error_offset_all$Reason[1] <- "Extra Notes"
+error_offset_all$CorrectionNr[1] <- 1
 error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
-error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 2
-# SubNr 27, BlockNr 3, TrialNr 6
+error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 4
+# SubNr 8, BlockNr 4, TrialNr 2
 error_offset_all <- rbind(error_offset_all, error_offset_all[5])
@@
 error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
-error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 6
+error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 2
+# SubNr 10, BlockNr 1, TrialNr 6
+error_offset_all <- rbind(error_offset_all, error_offset_all[6])
+error_offset_all$Reason[6] <- "Extra Notes"
+error_offset_all$CorrectionNr[6] <- 2
+error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
+error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 1
+# SubNr 15, BlockNr 3, TrialNr 5
+error_offset_all <- rbind(error_offset_all, error_offset_all[7])
+error_offset_all$Reason[7] <- "Extra Notes"
+error_offset_all$CorrectionNr[7] <- 1
+error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
+error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 2
 
@@
 
-1. **2 participants should be removed from the entire data analysis because they missed at least one block (16 trials)**
-
-- SubNr 3: experimental error (8 trials)
-- SubNr 14: played in octave higher (8 trials)
-
-2. **some performances were removed because 1) participants didn't follow sheet music or 2) participants didn't finish till the end (5 trials)**
-
 ### Missing/Excluded Trials Offset (Summary)

### 2020-11-18 script/RNotebook_2/ioi.Rmd
stats for ioi
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-07-28 script/RNotebook/ioi.Rmd
@@
   + [Trial by trial](#art_ind_trial)
-  + [Variability](#art_ind_var_all)
-  + [Variability trial-by-trial](#art_ind_var_trial)
 
@@
   + [Trial by trial](#art_group_trial)
-  + [Variability](#art_group_var_all)
-  + [Variability trial-by-trial](#art_group_var_trial)
 
@@
   + [Trial by trial](#dyn_ind_trial)
-  + [Variability all](#dyn_ind_var_all)
-  + [Variability trial-by-trial](#dyn_ind_var_trial)
 
@@
   + [Trial by trial](#dyn_group_trial)
-  + [Variability all](#dyn_group_var_all)
-  + [Variability trial-by-trial](#dyn_group_var_trial)
 
@@
 
+## Excluded participnats
+- SubNr 1: Most of responses (more than 90%) were removed due to outliers
+- SubNr 3: Experimental error
+- SubNr 8: Experimental error
+- SubNr 14: Experimental error
+
 ```{r setup, include = FALSE}
@@
 # data manipulation
-if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 # plot
@@
 
+```{r file, include = FALSE}
+filename_ioi = "../R/preprocessor/trimmed/data_ioi_1.txt"
+```
+
 ```{r extract, include = FALSE}
-df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
+dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
-df_ioi$SubNr <- as.factor(df_ioi$SubNr)
+dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
+
+# Exclude SubNr 1 because most responses were removed because of outliers
+dt_ioi <- dt_ioi[SubNr != 1]
 
 # Include only articulation
-df_ioi_art <- df_ioi %>% dplyr::filter(Skill == "articulation")
+dt_ioi_art <- dt_ioi[Skill == "articulation"]
 # Include only dynamics
-df_ioi_dyn <- df_ioi %>% dplyr::filter(Skill == "dynamics")
+dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 ```
@@
 # For each individual
-ioi_art <- aggregate(IOI~SubNr*Condition*Skill, data = df_ioi_art,
-                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
-ioi_art <- cbind(ioi_art[,1:3], as.data.frame(ioi_art[,4]))
-# Change colnames
-colnames(ioi_art) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
-print(ioi_art)
+ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
+ioi_art
 ```
@@
 # Group mean
-ioi_art_all <- aggregate(Mean~Condition*Skill, data = ioi_art,
-                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_all <- cbind(ioi_art_all[,1:2], as.data.frame(ioi_art_all[,3]))
-# Change colnames
-colnames(ioi_art_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
-print(ioi_art_all)
+ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art)/2)), by = .(Condition, Skill)]
+ioi_art_all
 ```
@@
 #### Normality check
+- Violate the normality
 ```{r normality1, echo = FALSE}
@@
 ```{r stats1, echo = FALSE}
-ioi_art_ttest <- t.test(ioi_art$Mean[ioi_art$Condition == "teaching"], ioi_art$Mean[ioi_art$Condition == "performing"], paired = TRUE)
+ioi_art_ttest <- t.test(ioi_art[Condition == "teaching"]$Mean, ioi_art[Condition == "performing"]$Mean, paired = TRUE)
 print(ioi_art_ttest)
@@
 
-#### Effect size
+#### Wilcoxon test
 ```{r stats2, echo = FALSE}
-ioi_art_cohend <- cohen.d(ioi_art$Mean, ioi_art$Condition, paired = TRUE)
-print(ioi_art_cohend)
+ioi_art_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_art, paired = TRUE)
+ioi_art_wilcoxon
 ```
@@
 # Define phrase boundaries
-df_ioi_art$Boundary <- "No"
-df_ioi_art$Boundary[df_ioi_art$Interval == 8 | df_ioi_art$Interval == 16 | df_ioi_art$Interval == 24 | df_ioi_art$Interval == 41 | df_ioi_art$Interval == 49 | df_ioi_art$Interval == 57] <- "Yes"
+dt_ioi_art$Boundary <- "No"
+dt_ioi_art[Interval == 8 | Interval == 16 | Interval == 24 | Interval == 41 | Interval == 49 | Interval == 57]$Boundary <- "Yes"
 
 # For each individual
-ioi_art_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_art,
-                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_art_ch <- cbind(ioi_art_ch[,1:4], as.data.frame(ioi_art_ch[,5]))
-# Change colnames
-colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
-print(ioi_art_ch)
+ioi_art_ch <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
+ioi_art_ch
 ```
@@
 # Group mean
-ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_art_ch,
-                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_ch_all <- cbind(ioi_art_ch_all[,1:3], as.data.frame(ioi_art_ch_all[,4]))
-# Change colnames
-colnames(ioi_art_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
-print(ioi_art_ch_all)
+ioi_art_ch_all <- ioi_art_ch[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art_ch)/2)), by = .(Condition, Skill, Boundary)]
+ioi_art_ch_all
 ```
@@
 #### Normality check
+- Violate the normality
 ```{r normality2, echo = FALSE}
@@
 ```
-
 #### Two-way ANOVA (ez)
@@
 ioi_art_ch_aov <- ezANOVA(
-  data = df_ioi_art
+  data = dt_ioi_art
   , dv = .(IOI)
@@
 # For each individual
-ioi_art_trial <- aggregate(IOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_art,
-                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_art_trial <- cbind(ioi_art_trial[,1:4], as.data.frame(ioi_art_trial[,5]))
-# Change colnames
-colnames(ioi_art_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean", "SD")
-print(ioi_art_trial)
+ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
+ioi_art_trial
 ```
@@
 # Group mean
-ioi_art_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_art_trial,
-                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_trial_all <- cbind(ioi_art_trial_all[,1:3], as.data.frame(ioi_art_trial_all[,4]))
-# Change colnames
-colnames(ioi_art_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
-print(ioi_art_trial_all)
+ioi_art_trial_all <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art_trial)/2)), by = .(Condition, Skill, TrialNr)]
+ioi_art_trial_all
 ```
@@
 
-## 7. Variability all - individual{#art_ind_var_all}
-- Average variability for each participant.
-- IOI variability = SD/mean IOI (per trial)
-
-```{r ioi_art_var, echo = FALSE}
-# Create a data frame for Variability (SD/mean IOI)
-df_ioi_art_var <- data.frame()
-for (subnr in unique(df_ioi_art$SubNr)){
-  for (block in c(1:4)){
-    cond = as.character(unique(df_ioi_art$Condition[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
-    skill = as.character(unique(df_ioi_art$Skill[df_ioi_art$SubNr == subnr & df_ioi_art$BlockNr == block]))
-    for (trial in c(1:8)){
-      df_current <- df_ioi_art %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
-      if (nrow(df_current) != 0){
-        df_ioi_art_var <- rbind(df_ioi_art_var, 
-                                data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI)))
-      }
-    }
-  }
-}
-colnames(df_ioi_art_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability")
-
-# Average variability
-# For each individual
-ioi_art_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_art_var,
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_art_var <- cbind(ioi_art_var[,1:3], as.data.frame(ioi_art_var[,4]))
-# Change colnames
-colnames(ioi_art_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
-print(ioi_art_var)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SD of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_art_var_bar, echo = FALSE}
-p_ioi_art_var <- ggplot(data = ioi_art_var, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(1)) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation")
-p_ioi_art_var
-```
-
-## 8. Variability all - group{#art_group_var_all}
-- Average variability for each condition.
-
-```{r ioi_art_var_all, echo = FALSE}
-# Group mean
-ioi_art_var_all <- aggregate(Mean~Condition*Skill, data = ioi_art_var,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_var_all <- cbind(ioi_art_var_all[,1:2], as.data.frame(ioi_art_var_all[,3]))
-# Change colnames
-colnames(ioi_art_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
-print(ioi_art_var_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r ioi_art_var_all_bar, echo = FALSE}
-p_ioi_art_var_all <- ggplot(data = ioi_art_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge()) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
-  theme(legend.position = "None")
-p_ioi_art_var_all
-```
-
-#### Box plot
-```{r ioi_art_var_all_box, echo = FALSE}
-p_ioi_art_var_box <- ggplot(data = ioi_art_var, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") +
-  theme(legend.position = "None")
-p_ioi_art_var_box
-```
-
-### Statistics
-#### Normality check
-```{r normality3, echo = FALSE}
-ioi_art_var_norm <- by(ioi_art_var$Mean, list(ioi_art_var$Condition), shapiro.test)
-print(ioi_art_var_norm)
-```
-
-#### Paired t-test
-```{r stats5, echo = FALSE}
-ioi_art_var_ttest <- t.test(ioi_art_var$Mean[ioi_art_var$Condition == "teaching"], ioi_art_var$Mean[ioi_art_var$Condition == "performing"], paired = TRUE)
-print(ioi_art_var_ttest)
-```
-
-#### Effect size
-```{r stats6, echo = FALSE}
-ioi_art_var_cohend <- cohen.d(ioi_art_var$Mean, ioi_art_var$Condition, paired = TRUE)
-print(ioi_art_var_cohend)
-```
-
-## 9. Variability trial-by-trial - individual{#art_ind_var_trial}
-- Average variability for each trial for each participant.
-
-```{r ioi_art_var_trial, echo = FALSE}
-# For each individual
-ioi_art_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_art_var,
-                         FUN = function(x){c(N = length(x), mean = mean(x))})
-ioi_art_var_trial <- cbind(ioi_art_var_trial[,1:4], as.data.frame(ioi_art_var_trial[,5]))
-# Change colnames
-colnames(ioi_art_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean")
-print(ioi_art_var_trial)
-```
-
-### Plot
-#### Line graph
-```{r ioi_art_var_trial_line, echo = FALSE, fig.height = 4}
-p_ioi_art_var_trial <- ggplot(data = ioi_art_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_art_var_trial
-```
-
-## 10. Variability trial-by-trial - group{#art_group_var_trial}
-- Average variability for each trial for each condition.
-
-```{r ioi_art_var_trial_all, echo = FALSE}
-# Group mean
-ioi_art_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_art_var_trial,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_var_trial_all <- cbind(ioi_art_var_trial_all[,1:3], as.data.frame(ioi_art_var_trial_all[,4]))
-# Change colnames
-colnames(ioi_art_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
-print(ioi_art_var_trial_all)
-```
-
-### Plot
-#### Line graph
-- Error bars show SEM of the mean
-
-```{r ioi_art_var_trial_all_line, echo = FALSE}
-p_ioi_art_var_trial_all <- ggplot(data = ioi_art_var_trial_all, 
-                                  aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line(position = position_dodge(.2)) +
-  geom_point(position = position_dodge(.2)) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_art_var_trial_all
-```
-
 # IOI-Dynamics
@@
 # For each individual
-ioi_dyn <- aggregate(IOI~SubNr*Condition*Skill, data = df_ioi_dyn,
-                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
-ioi_dyn <- cbind(ioi_dyn[,1:3], as.data.frame(ioi_dyn[,4]))
-# Change colnames
-colnames(ioi_dyn) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
-print(ioi_dyn)
+ioi_dyn <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
+ioi_dyn
 ```
@@
 # Group mean
-ioi_dyn_all <- aggregate(Mean~Condition*Skill, data = ioi_dyn,
-                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_all <- cbind(ioi_dyn_all[,1:2], as.data.frame(ioi_dyn_all[,3]))
-# Change colnames
-colnames(ioi_dyn_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_all)
+ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn)/2)), by = .(Condition, Skill)]
+ioi_dyn_all
 ```
@@
 #### Normality check
-```{r normality4, echo = FALSE}
+- Violate the normality
+```{r normality3, echo = FALSE}
 # Normality check
@@
 
-```{r stats7, echo = FALSE}
+#### Paired t-test
+```{r stats5, echo = FALSE}
 ioi_dyn_ttest <- t.test(ioi_dyn$Mean[ioi_dyn$Condition == "teaching"], ioi_dyn$Mean[ioi_dyn$Condition == "performing"], paired = TRUE)
@@
 
-#### Effect size
-```{r stats8, echo = FALSE}
-ioi_dyn_cohend <- cohen.d(ioi_dyn$Mean, ioi_dyn$Condition, paired = TRUE)
-print(ioi_dyn_cohend)
+#### Wilcoxon test
+```{r stats6, echo = FALSE}
+ioi_dyn_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_dyn, paired = TRUE)
+ioi_dyn_wilcoxon
 ```
@@
 # Define phrase boundaries
-df_ioi_dyn$Boundary <- "No"
-df_ioi_dyn$Boundary[df_ioi_dyn$Interval == 8 | df_ioi_dyn$Interval == 16 | df_ioi_dyn$Interval == 24 | df_ioi_dyn$Interval == 41 | df_ioi_dyn$Interval == 49 | df_ioi_dyn$Interval == 57] <- "Yes"
+dt_ioi_dyn$Boundary <- "No"
+dt_ioi_dyn[Interval == 8 | Interval == 16 | Interval == 24 | Interval == 41 | Interval == 49 | Interval == 57]$Boundary <- "Yes"
 
 # For each individual
-ioi_dyn_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_dyn,
-                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_dyn_ch <- cbind(ioi_dyn_ch[,1:4], as.data.frame(ioi_dyn_ch[,5]))
-# Change colnames
-colnames(ioi_dyn_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
-print(ioi_dyn_ch)
+ioi_dyn_ch <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
+ioi_dyn_ch
 ```
@@
 ### Plot
-#### Bar chart
+#### Bar chdyn
 - Error bars show SD of the mean
@@
 # Group mean
-ioi_dyn_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_dyn_ch,
-                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_ch_all <- cbind(ioi_dyn_ch_all[,1:3], as.data.frame(ioi_dyn_ch_all[,4]))
-# Change colnames
-colnames(ioi_dyn_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_ch_all)
+ioi_dyn_ch_all <- ioi_dyn_ch[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn_ch)/2)), by = .(Condition, Skill, Boundary)]
+ioi_dyn_ch_all
 ```
@@
 ### Plot
-#### Bar chart
+#### Bar chdyn
 - Error bars show SEM of the mean
@@
 #### Normality check
-```{r normality5, echo = FALSE}
+- Violate the normality
+```{r normality4, echo = FALSE}
 ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test)
@@
 ```
-
 #### Two-way ANOVA (ez)
-```{r stats9, echo = FALSE}
+```{r stats7, echo = FALSE}
 ioi_dyn_ch_aov <- ezANOVA(
-  data = df_ioi_dyn
+  data = dt_ioi_dyn
   , dv = .(IOI)
@@
 ioi_dyn_ch_aov$ANOVA$p <- round(ioi_dyn_ch_aov$ANOVA$p, 4)
-print(ioi_dyn_ch_aov)
+print(ioi_dyn_ch_aov$ANOVA)
 ```
@@
 #### Two-way ANOVA (aov)
-```{r stats10, echo = FALSE}
+```{r stats8, echo = FALSE}
 ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
@@
 # For each individual
-ioi_dyn_trial <- aggregate(IOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn,
-                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_dyn_trial <- cbind(ioi_dyn_trial[,1:4], as.data.frame(ioi_dyn_trial[,5]))
-# Change colnames
-colnames(ioi_dyn_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean", "SD")
-print(ioi_dyn_trial)
+ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
+ioi_dyn_trial
 ```
@@
 # Group mean
-ioi_dyn_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_dyn_trial,
-                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_trial_all <- cbind(ioi_dyn_trial_all[,1:3], as.data.frame(ioi_dyn_trial_all[,4]))
-# Change colnames
-colnames(ioi_dyn_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_trial_all)
+ioi_dyn_trial_all <- ioi_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn_trial)/2)), by = .(Condition, Skill, TrialNr)]
+ioi_dyn_trial_all
 ```
@@
 ```
-
-## 7. Variability all - individual{#dyn_ind_var_all}
-- Average variability for each participant.
-- IOI variability = SD/mean IOI (per trial)
-
-```{r ioi_dyn_var, echo = FALSE}
-# Create a data frame for variability (SD/)
-df_ioi_dyn_var <- data.frame()
-for (subnr in unique(df_ioi_dyn$SubNr)){
-  for (block in c(1:4)){
-    cond = as.character(unique(df_ioi_dyn$Condition[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
-    skill = as.character(unique(df_ioi_dyn$Skill[df_ioi_dyn$SubNr == subnr & df_ioi_dyn$BlockNr == block]))
-    for (trial in c(1:8)){
-      df_current <- df_ioi_dyn %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
-      if (nrow(df_current) != 0){
-        df_ioi_dyn_var <- rbind(df_ioi_dyn_var, 
-                                data.frame(subnr, block, trial, cond, skill, sd(df_current$IOI)/mean(df_current$IOI)))
-      }
-    }
-  }
-}
-colnames(df_ioi_dyn_var) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "Variability")
-
-# Average variability
-# For each individual
-ioi_dyn_var <- aggregate(Variability~SubNr*Condition*Skill, data = df_ioi_dyn_var,
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_dyn_var <- cbind(ioi_dyn_var[,1:3], as.data.frame(ioi_dyn_var[,4]))
-# Change colnames
-colnames(ioi_dyn_var) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
-print(ioi_dyn_var)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SD of the mean
-
-```{r ioi_dyn_var_bar, echo = FALSE}
-p_ioi_dyn_var <- ggplot(data = ioi_dyn_var, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(1)) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics")
-p_ioi_dyn_var
-```
-
-## 8. Variability all - group{#dyn_group_var_all}
-- Average variability for each condition.
-
-```{r ioi_dyn_var_all, echo = FALSE}
-# Group mean
-ioi_dyn_var_all <- aggregate(Mean~Condition*Skill, data = ioi_dyn_var,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_var_all <- cbind(ioi_dyn_var_all[,1:2], as.data.frame(ioi_dyn_var_all[,3]))
-# Change colnames
-colnames(ioi_dyn_var_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_var_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-
-```{r ioi_dyn_var_all_bar, echo = FALSE}
-p_ioi_dyn_var_all <- ggplot(data = ioi_dyn_var_all, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge()) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
-  theme(legend.position = "None")
-p_ioi_dyn_var_all
-```
-
-#### Box plot
-```{r ioi_dyn_var_all_box, echo = FALSE}
-p_ioi_dyn_var_box <- ggplot(data = ioi_dyn_var, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  labs(y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") +
-  theme(legend.position = "None")
-p_ioi_dyn_var_box
-```
-
-### Statistics
-#### Normality check
-```{r normality6, echo = FALSE}
-ioi_dyn_var_norm <- by(ioi_dyn_var$Mean, list(ioi_dyn_var$Condition), shapiro.test)
-print(ioi_dyn_var_norm)
-```
-
-```{r stats11, echo = FALSE}
-ioi_dyn_var_ttest <- t.test(ioi_dyn_var$Mean[ioi_dyn_var$Condition == "teaching"], ioi_dyn_var$Mean[ioi_dyn_var$Condition == "performing"], paired = TRUE)
-print(ioi_dyn_var_ttest)
-```
-
-```{r stats12, echo = FALSE}
-ioi_dyn_var_cohend <- cohen.d(ioi_dyn_var$Mean, ioi_dyn_var$Condition, paired = TRUE)
-print(ioi_dyn_var_cohend)
-```
-
-## 9. Variability trial-by-trial - individual{#dyn_ind_var_trial}
-- Average variability for each trial for each participant.
-
-```{r ioi_dyn_var_trial, echo = FALSE}
-# For each individual
-ioi_dyn_var_trial <- aggregate(Variability~SubNr*Condition*Skill*TrialNr, data = df_ioi_dyn_var,
-                         FUN = function(x){c(N = length(x), mean = mean(x))})
-ioi_dyn_var_trial <- cbind(ioi_dyn_var_trial[,1:4], as.data.frame(ioi_dyn_var_trial[,5]))
-# Change colnames
-colnames(ioi_dyn_var_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean")
-print(ioi_dyn_var_trial)
-```
-
-### Plot
-#### Line graph
-```{r ioi_dyn_var_trial_line, echo = FALSE, fig.height = 4}
-p_ioi_dyn_var_trial <- ggplot(data = ioi_dyn_var_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_dyn_var_trial
-```
-
-## 10. Variability trial-by-trial - group{#dyn_group_var_trial}
-Average variability for each trial for each condition.
-
-```{r ioi_dyn_var_trial_all, echo = FALSE}
-# Group mean
-ioi_dyn_var_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_dyn_var_trial,
-                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_var_trial_all <- cbind(ioi_dyn_var_trial_all[,1:3], as.data.frame(ioi_dyn_var_trial_all[,4]))
-# Change colnames
-colnames(ioi_dyn_var_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_var_trial_all)
-```
-
-### Plot
-#### Line graph
-- Error bars show SEM of the mean
-
-```{r ioi_dyn_var_trial_all_line, echo = FALSE}
-p_ioi_dyn_var_trial_all <- ggplot(data = ioi_dyn_var_trial_all, 
-                                  aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line(position = position_dodge(.2)) +
-  geom_point(position = position_dodge(.2)) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
-  labs(x = "Trial", y = "Variability (SD/mean IOI)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_dyn_var_trial_all
-```
-
-# IOI-Order-Articulation
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#art_order_group_all}
-```{r ioi_art_order_all, echo = FALSE}
-# For each participant
-ioi_art_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_art,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-ioi_art_order <- cbind(ioi_art_order[,1:4], as.data.frame(ioi_art_order[,5]))
-# Change colnames
-colnames(ioi_art_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD")
-
-# Group mean
-ioi_art_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_art_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_order_all <- cbind(ioi_art_order_all[,1:3], as.data.frame(ioi_art_order_all[,4]))
-# Change colnames
-colnames(ioi_art_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM")
-print(ioi_art_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_art_order_all_bar, echo = FALSE}
-p_ioi_art_order_all <- ggplot(data = ioi_art_order_all, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge()) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  facet_wrap(Order ~ .) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) +
-  theme(legend.position = "None")
-p_ioi_art_order_all
-```
-
-#### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_art_order_all_box, echo = FALSE}
-p_ioi_art_order_all_box <- ggplot(data = ioi_art_order, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  facet_wrap(Order ~ .) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") +
-  theme(legend.position = "None")
-p_ioi_art_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality7, echo = FALSE}
-# Normality check
-ioi_art_order_norm <- by(ioi_art_order$Mean, list(ioi_art_order$Condition, ioi_art_order$Order), shapiro.test)
-print(ioi_art_order_norm)
-```
-
-#### Two-way ANOVA (ez)
-```{r stats13, echo = FALSE}
-ioi_art_order_aov <- ezANOVA(
-  data = df_ioi_art
-  , dv = .(IOI)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition)
-  , type = 3
-  , detailed = TRUE
-)
-ioi_art_order_aov$ANOVA$F <- round(ioi_art_order_aov$ANOVA$F, 4)
-ioi_art_order_aov$ANOVA$p <- round(ioi_art_order_aov$ANOVA$p, 4)
-print(ioi_art_order_aov)
-```
-
-#### Two-way ANOVA (aov)
-```{r stats14, echo = FALSE}
-ioi_art_order_aov_2 <- aov(Mean ~ Condition*Order + Error(SubNr/Condition) + Order, data = ioi_art_order)
-print(summary(ioi_art_order_aov_2))
-```
-
-# IOI-Order-Dynamics
-- Investigate whether there is any order effect of the conditions.
-
-## 1. All{#dyn_order_group_all}
-```{r ioi_dyn_order_all, echo = FALSE}
-# For each pdynicipant
-ioi_dyn_order <- aggregate(IOI~SubNr*Condition*Skill*Order, data = df_ioi_dyn,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
-ioi_dyn_order <- cbind(ioi_dyn_order[,1:4], as.data.frame(ioi_dyn_order[,5]))
-# Change colnames
-colnames(ioi_dyn_order) <- c("SubNr", "Condition", "Skill", "Order", "N", "Mean", "SD")
-
-# Group mean
-ioi_dyn_order_all <- aggregate(Mean~Condition*Skill*Order, data = ioi_dyn_order,
-                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_order_all <- cbind(ioi_dyn_order_all[,1:3], as.data.frame(ioi_dyn_order_all[,4]))
-# Change colnames
-colnames(ioi_dyn_order_all) <- c("Condition", "Skill", "Order", "N", "Mean", "SD", "SEM")
-print(ioi_dyn_order_all)
-```
-
-### Plot
-#### Bar graph
-- Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_dyn_order_all_bar, echo = FALSE}
-p_ioi_dyn_order_all <- ggplot(data = ioi_dyn_order_all, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
-                width = .2, position = position_dodge()) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  facet_wrap(Order ~ .) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(120, 210)) +
-  theme(legend.position = "None")
-p_ioi_dyn_order_all
-```
-
-#### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
-
-```{r ioi_dyn_order_all_box, echo = FALSE}
-p_ioi_dyn_order_all_box <- ggplot(data = ioi_dyn_order, aes(x = Condition, y = Mean, fill = Condition)) +
-  geom_boxplot() +
-  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  facet_wrap(Order ~ .) +
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") +
-  theme(legend.position = "None")
-p_ioi_dyn_order_all_box
-```
-
-### Statistics
-#### Normality check
-```{r normality8, echo = FALSE}
-# Normality check
-ioi_dyn_order_norm <- by(ioi_dyn_order$Mean, list(ioi_dyn_order$Condition, ioi_dyn_order$Order), shapiro.test)
-print(ioi_dyn_order_norm)
-```
-
-#### Two-way ANOVA (ez)
-```{r stats15, echo = FALSE}
-ioi_dyn_order_aov <- ezANOVA(
-  data = df_ioi_dyn
-  , dv = .(IOI)
-  , wid = .(SubNr)
-  , between = .(Order)
-  , within = .(Condition)
-  , type = 3
-  , detailed = TRUE
-)
-ioi_dyn_order_aov$ANOVA$F <- round(ioi_dyn_order_aov$ANOVA$F, 4)
-ioi_dyn_order_aov$ANOVA$p <- round(ioi_dyn_order_aov$ANOVA$p, 4)
-print(ioi_dyn_order_aov)
-```
-
-#### Two-way ANOVA (aov)
-```{r stats16, echo = FALSE}
-ioi_dyn_order_aov_2 <- aov(Mean ~ Condition*Order + Error(SubNr/Condition) + Order, data = ioi_dyn_order)
-print(summary(ioi_dyn_order_aov_2))
-```
-
 # Sequence plot
 ## Articulation{#sequence_art}
-```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
+```{r seq_art, fig.width = 9, fig.height = 3, echo = FALSE}
 # For each individual
-ioi_art_seq <- aggregate(IOI~SubNr*Condition*Skill*Interval, data = df_ioi_art,
-                                 FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_art_seq <- cbind(ioi_art_seq[,1:4], as.data.frame(ioi_art_seq[,5]))
-# Change colnames
-colnames(ioi_art_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+ioi_art_seq <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+# Group mean
+ioi_art_seq_all <- ioi_art_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art_seq)/2)), by = .(Condition, Skill, Interval)]
 
-# Group mean
-ioi_art_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = ioi_art_seq,
-                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_art_seq_stats <- cbind(ioi_art_seq_stats[,1:3], as.data.frame(ioi_art_seq_stats[,4]))
-# Change colnames
-colnames(ioi_art_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
-
-p_ioi_art_seq <- ggplot(data = ioi_art_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+p_ioi_art_seq <- ggplot(data = ioi_art_seq_all, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
   geom_line() +
@@
                 position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "IOI (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "Interval", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,66,1))
 p_ioi_art_seq
@@
 ## Dynamics{#sequence_dyn}
-```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
+```{r seq_dyn, fig.width = 9, fig.height = 3, echo = FALSE}
 # For each individual
-ioi_dyn_seq <- aggregate(IOI~SubNr*Condition*Skill*Interval, data = df_ioi_dyn,
-                                 FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-ioi_dyn_seq <- cbind(ioi_dyn_seq[,1:4], as.data.frame(ioi_dyn_seq[,5]))
-# Change colnames
-colnames(ioi_dyn_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")
+ioi_dyn_seq <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+# Group mean
+ioi_dyn_seq_all <- ioi_dyn_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn_seq)/2)), by = .(Condition, Skill, Interval)]
 
-# Group mean
-ioi_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = ioi_dyn_seq,
-                           FUN = function(x){round(c(length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
-ioi_dyn_seq_stats <- cbind(ioi_dyn_seq_stats[,1:3], as.data.frame(ioi_dyn_seq_stats[,4]))
-# Change colnames
-colnames(ioi_dyn_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")
-
-p_ioi_dyn_seq <- ggplot(data = ioi_dyn_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
+p_ioi_dyn_seq <- ggplot(data = ioi_dyn_seq_all, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
   geom_line() +
@@
                 position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "IOI (ms)") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "Interval", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,66,1))
 p_ioi_dyn_seq

### 2020-11-18 script/RNotebook_2/ioi.Rmd
remove print()
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ioi_art_norm <- by(ioi_art$Mean, list(ioi_art$Condition), shapiro.test)
-print(ioi_art_norm)
+ioi_art_norm
 ```
@@
 ioi_art_ttest <- t.test(ioi_art[Condition == "teaching"]$Mean, ioi_art[Condition == "performing"]$Mean, paired = TRUE)
-print(ioi_art_ttest)
+ioi_art_ttest
 ```
@@
 ioi_art_ch_norm <- by(ioi_art_ch$Mean, list(ioi_art_ch$Condition, ioi_art_ch$Boundary), shapiro.test)
-print(ioi_art_ch_norm)
+ioi_art_ch_norm
 ```
@@
 ioi_art_ch_aov$ANOVA$p <- round(ioi_art_ch_aov$ANOVA$p, 4)
-print(ioi_art_ch_aov$ANOVA)
+ioi_art_ch_aov$ANOVA
 ```
@@
 ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
-print(summary(ioi_art_ch_aov_2))
+summary(ioi_art_ch_aov_2)
 ```
@@
 ioi_dyn_norm <- by(ioi_dyn$Mean, list(ioi_dyn$Condition), shapiro.test)
-print(ioi_dyn_norm)
+ioi_dyn_norm
 ```
@@
 ioi_dyn_ttest <- t.test(ioi_dyn$Mean[ioi_dyn$Condition == "teaching"], ioi_dyn$Mean[ioi_dyn$Condition == "performing"], paired = TRUE)
-print(ioi_dyn_ttest)
+ioi_dyn_ttest
 ```
@@
 ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test)
-print(ioi_dyn_ch_norm)
+ioi_dyn_ch_norm
 ```
@@
 ioi_dyn_ch_aov$ANOVA$p <- round(ioi_dyn_ch_aov$ANOVA$p, 4)
-print(ioi_dyn_ch_aov$ANOVA)
+ioi_dyn_ch_aov$ANOVA
 ```
@@
 ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
-print(summary(ioi_dyn_ch_aov_2))
+summary(ioi_dyn_ch_aov_2)
 ```

### 2020-11-18 script/RNotebook_2/articulation.Rmd
update analysis scripts (check later)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Articulation Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for articulation.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Outline - Articulation (KOT, KOR)
## [Experimental Design](#design)

## [Articulation](#kot-articulation)

### KOT
- Individual results
  + [All](#kot_art_ind_all)
  + [Trial by trial](#kot_art_ind_trial)

- Group results and stats (if available)
  + [All](#kot_art_group_all)
  + [Trial by trial](#kot_art_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kot_art_order_group_all)

- [Sequence plot](#sequence_art)

### KOR
- Individual results
  + [All](#kor_art_ind_all)
  + [Trial by trial](#kor_art_ind_trial)

- Group results and stats (if available)
  + [All](#kor_art_group_all)
  + [Trial by trial](#kor_art_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kor_art_order_group_all)

## [Dynamics](#kot-dynamics)

### KOT
- Individual results
  + [All](#kot_dyn_ind_all)
  + [Trial by trial](#kot_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#kot_dyn_group_all)
  + [Trial by trial](#kot_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kot_dyn_order_group_all)

- [Sequence plot](#sequence_dyn)

### KOR
- Individual results
  + [All](#kor_dyn_ind_all)
  + [Trial by trial](#kor_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#kor_dyn_group_all)
  + [Trial by trial](#kor_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kor_dyn_order_group_all)

## Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: legato vs. staccato / forte vs. piano)
- Dependent variables: Key-Overlap Time (KOT), Key-Overlap Ratio (KOR)

## Main predictions
Participants will exaggerate legato and staccato more while teaching.

## Excluded participnats
- SubNr 1: Most of responses (more than 90%) were removed due to outliers
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("effsize")) {install.packages("effsize"); require("effsize")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r file, include = FALSE}
filename_ioi = "../R/preprocessor/trimmed/data_ioi_1.txt"
filename_kot = "../R/preprocessor/trimmed/data_kot.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
dt_kot$SubNr <- as.factor(dt_kot$SubNr)

# Exclude SubNr 1 because most responses were removed because of outliers
dt_ioi <- dt_ioi[SubNr != 1]
dt_kot <- dt_kot[SubNr != 1]

# For each individual, each block, each trial
ioi <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]

# Calculate KOR for each interval (outliers were determined based on KOT)
dt_kor <- data.table()
for (subnr in unique(dt_kot$SubNr)){
  for (block in unique(dt_kot[SubNr == subnr]$BlockNr)){
    for (trial in unique(dt_kot[SubNr == subnr & BlockNr == block]$TrialNr)){
      dt_current <- dt_kot[SubNr == subnr & BlockNr == block & TrialNr == trial]
      dt_current$KOR <- dt_current$KOT/ioi[SubNr == subnr & ioi$BlockNr == block & TrialNr == trial]$Mean
      dt_kor <- rbind(dt_kor, dt_current)
    }
  }
}

# Include only articulation
dt_kot_art <- dt_kot[Skill == "articulation"]
dt_kor_art <- dt_kor[Skill == "articulation"]

# Include only dynamics
dt_kot_dyn <- dt_kot[Skill == "dynamics"]
dt_kor_dyn <- dt_kor[Skill == "dynamics"]
```

# Key-Overlap Time (KOT) - Articulation{#kot-articulation}
## 1. All - individual{#kot_art_ind_all}
- Average KOT for each participant.

```{r kot_art, echo = FALSE}
# For each individual
kot_art <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
kot_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kot_art_bar,  echo = FALSE}
p_kot_art <- ggplot(data = kot_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art
```

## 2. All - group{#kot_art_group_all}
- Average KOT for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kot_art_all, echo = FALSE}
# Group mean
kot_art_all <- kot_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kot_art_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r kot_art_all_bar, echo = FALSE}　
p_kot_art_all <- ggplot(data = kot_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art_all
```

#### Box plot
```{r kot_art_all_box, echo = FALSE}
p_kot_art_all_box <- ggplot(data = kot_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art_all_box
```

### Statistics
#### Normality check
```{r normality1, echo = FALSE}
# Reduce unused factors
kot_art$Subcomponent <- factor(kot_art$Subcomponent)
kot_art_norm <- by(kot_art$Mean, list(kot_art$Condition, kot_art$Subcomponent), shapiro.test)
kot_art_norm
```

#### Two-way ANOVA (ez)
```{r stats1, echo = FALSE}
kot_art_aov <- ezANOVA(
  data = dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato"]
  , dv = .(KOT)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kot_art_aov$ANOVA$F <- round(kot_art_aov$ANOVA$F, 4)
kot_art_aov$ANOVA$p <- round(kot_art_aov$ANOVA$p, 4)
kot_art_aov$ANOVA
```

#### Two-way ANOVA (aov)
```{r stats2, echo = FALSE}
kot_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
summary(kot_art_aov_2)
```

#### Post-hoc (direct comparison)
```{r post-hoc1, echo = FALSE}
kot_ttest_legato <- t.test(kot_art[Condition == "performing" & Subcomponent == "Legato"]$Mean, kot_art[Condition == "teaching" & Subcomponent == "Legato"]$Mean, paired = TRUE)
kot_ttest_staccato <- t.test(kot_art[Condition == "performing" & Subcomponent == "Staccato"]$Mean, kot_art[Condition == "teaching" & Subcomponent == "Staccato"]$Mean, paired = TRUE)
# # Effect size
# kot_ttest_legato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Legato"], kot_art$Condition[kot_art$Subcomponent == "Legato"], paired = TRUE)
# kot_ttest_staccato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Staccato"], kot_art$Condition[kot_art$Subcomponent == "Staccato"], paired = TRUE)

kot_ttest_legato
kot_ttest_staccato
#kot_ttest_legato_cohend
#kot_ttest_staccato_cohend
```

#### Post-hoc (pairwise t test)
```{r post-hoc2, echo = FALSE}
# Create a group label for a post-hoc test
kot_art$Group <- "NA"
kot_art[Condition == "performing" & Subcomponent == "Legato"]$Group <- "Performing-Legato" 
kot_art[Condition == "performing" & Subcomponent == "Staccato"]$Group <- "Performing-Staccato"
kot_art[Condition == "teaching" & Subcomponent == "Legato"]$Group <- "Teaching-Legato"
kot_art[Condition == "teaching" & Subcomponent == "Staccato"]$Group <- "Teaching-Staccato"
# pairwise comparison
kot_art_pairwise <- pairwise.t.test(kot_art$Mean, kot_art$Group, paired = TRUE, p.adjust.method = "BH")
kot_art_pairwise$p.value <- round(kot_art_pairwise$p.value, 4)
kot_art_pairwise
```

## 3. Trial by trial - individual{#kot_art_ind_trial}
- Average KOT for each trial for each participant.

```{r kot_art_trial, echo = FALSE}
# For each individual
kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kot_art_trial
```

### Plot
#### Line graph
```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 4}
p_kot_art_trial_legato <- ggplot(data = kot_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_legato
```

```{r kot_art_trial_line_staccato, echo = FALSE, fig.height = 4}
p_kot_art_trial_staccato <- ggplot(data = kot_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_staccato
```

## 4. Trial by trial - group{#kot_art_group_trial}
- Average KOT for each trial for each condition.

```{r kot_art_trial_all, echo = FALSE}
# Group mean
kot_art_trial_all <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kot_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kot_art_trial_all_line_legato, echo = FALSE}
p_kot_art_trial_all_legato <- ggplot(data = kot_art_trial_all[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_all_legato
```

```{r kot_art_trial_all_line_staccato, echo = FALSE}
p_kot_art_trial_all_staccato <- ggplot(data = kot_art_trial_all[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_all_staccato
```

# Key-Overlap Ratio (KOR) - Articulation
## 1. All - individual{#kor_art_ind_all}
- Average KOR for each participant.

```{r kor_art, echo = FALSE}
# For each individual
kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
kor_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kor_art_bar,  echo = FALSE}
p_kor_art <- ggplot(data = kor_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art
```

## 2. All - group{#kor_art_group_all}
- Average KOR for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kor_art_all, echo = FALSE}
# Group mean
kor_art_all <- kor_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kor_art_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r kor_art_all_bar, echo = FALSE}　
p_kor_art_all <- ggplot(data = kor_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art_all
```

#### Box plot
```{r kor_art_all_box, echo = FALSE}
p_kor_art_all_box <- ggplot(data = kor_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art_all_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality2, echo = FALSE}
# Reduce unused factors
kor_art$Subcomponent <- factor(kor_art$Subcomponent)
kor_art_norm <- by(kor_art$Mean, list(kor_art$Condition, kor_art$Subcomponent), shapiro.test)
kor_art_norm
```

#### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
kor_art_aov <- ezANOVA(
  data = dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato"]
  , dv = .(KOR)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kor_art_aov$ANOVA$F <- round(kor_art_aov$ANOVA$F, 4)
kor_art_aov$ANOVA$p <- round(kor_art_aov$ANOVA$p, 4)
print(kor_art_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
summary(kor_art_aov_2)    
```

#### Post-hoc (direct comparison)
```{r post-hoc3, echo = FALSE}
kor_ttest_legato <- t.test(kor_art[Condition == "performing" & Subcomponent == "Legato"]$Mean, kor_art[Condition == "teaching" & Subcomponent == "Legato"]$Mean, paired = TRUE)
kor_ttest_staccato <- t.test(kor_art[Condition == "performing" & Subcomponent == "Staccato"]$Mean, kor_art[Condition == "teaching" & Subcomponent == "Staccato"]$Mean, paired = TRUE)
# # Effect size
# kor_ttest_legato_cohend <- cohen.d(kor_art[Subcomponent == "Legato"]$Mean, kor_art[Subcomponent == "Legato"]$Condition, paired = TRUE)
# kor_ttest_staccato_cohend <- cohen.d(kor_art[Subcomponent == "Staccato"]$Mean, kor_art[Subcomponent == "Staccato"]$Condition, paired = TRUE)

kor_ttest_legato
kor_ttest_staccato
# kor_ttest_legato_cohend
# kor_ttest_staccato_cohend
```

#### Post-hoc (pairwise t test)
```{r post-hoc4, echo = FALSE}
# Create a group label for a post-hoc test
kor_art$Group <- "NA"
kor_art[Condition == "performing" & Subcomponent == "Legato"]$Group <- "Performing-Legato" 
kor_art[Condition == "performing" & Subcomponent == "Staccato"]$Group <- "Performing-Staccato"
kor_art[Condition == "teaching" & Subcomponent == "Legato"]$Group <- "Teaching-Legato"
kor_art[Condition == "teaching" & Subcomponent == "Staccato"]$Group <- "Teaching-Staccato"
# pairwise comparison
kor_art_pairwise <- pairwise.t.test(kor_art$Mean, kor_art$Group, paired = TRUE, p.adjust.method = "BH")
kor_art_pairwise$p.value <- round(kor_art_pairwise$p.value, 4)
kor_art_pairwise
```

## 3. Trial by trial - individual{#kor_art_ind_trial}
- Average KOR for each trial for each participant.

```{r kor_art_trial, echo = FALSE}
# For each individual
kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kor_art_trial
```

### Plot
#### Line graph
```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 4}
p_kor_art_trial_legato <- ggplot(data = kor_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_legato
```

```{r kor_art_trial_line_staccato, echo = FALSE, fig.height = 4}
p_kor_art_trial_staccato <- ggplot(data = kor_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_staccato
```

## 4. Trial by trial - group{#kor_art_group_trial}
- Average KOR d for each trial for each condition.

```{r kor_art_trial_all, echo = FALSE}
# Group mean
kor_art_trial_all <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kor_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kor_art_trial_all_line_legato, echo = FALSE}
p_kor_art_trial_all_legato <- ggplot(data = kor_art_trial_all[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_all_legato
```

```{r kor_art_trial_all_line_staccato, echo = FALSE}
p_kor_art_trial_all_staccato <- ggplot(data = kor_art_trial_all[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_all_staccato
```

# Key-Overlap Time (KOT) - Dynamics{#kot-dynamics}
## 1. All - individual{#kot_dyn_ind_all}
- Average KOT for each participant.

```{r kot_dyn, echo = FALSE}
# For each individual
kot_dyn <- dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
kot_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kot_dyn_bar,  echo = FALSE}
p_kot_dyn <- ggplot(data = kot_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn
```

## 2. All - group{#kot_dyn_group_all}
- Average KOT for each condition.

```{r kot_dyn_all, echo = FALSE}
# Group mean
kot_dyn_all <- kot_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kot_dyn_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r kot_dyn_all_bar, echo = FALSE}　
p_kot_dyn_all <- ggplot(data = kot_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn_all
```

#### Box plot

```{r kot_dyn_all_box, echo = FALSE}
p_kot_dyn_all_box <- ggplot(data = kot_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn_all_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality3, echo = FALSE}
# Reduce unused factors
kot_dyn$Subcomponent <- factor(kot_dyn$Subcomponent)
kot_dyn_norm <- by(kot_dyn$Mean, list(kot_dyn$Condition, kot_dyn$Subcomponent), shapiro.test)
print(kot_dyn_norm)
```

#### Two-way ANOVA (ez)
```{r stats5, echo = FALSE}
kot_dyn_aov <- ezANOVA(
  data = dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano"]
  , dv = .(KOT)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kot_dyn_aov$ANOVA$F <- round(kot_dyn_aov$ANOVA$F, 4)
kot_dyn_aov$ANOVA$p <- round(kot_dyn_aov$ANOVA$p, 4)
print(kot_dyn_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats6, echo = FALSE}
kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
summary(kot_dyn_aov_2)
```

## 3. Trial by trial - individual{#kot_dyn_ind_trial}
- Average KOT for each trial for each participant.

```{r kot_dyn_trial, echo = FALSE}
# For each individual
kot_dyn_trial <- dt_kot_dyn[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kot_dyn_trial
```

### Plot
#### Line graph
```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
p_kot_dyn_trial_forte <- ggplot(data = kot_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_forte
```

```{r kot_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
p_kot_dyn_trial_piano <- ggplot(data = kot_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_piano
```

## 4. Trial by trial - group{#kot_dyn_group_trial}
- Average KOT for each trial for each condition.

```{r kot_dyn_trial_all, echo = FALSE}
# Group mean
kot_dyn_trial_all <- kot_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kot_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kot_dyn_trial_all_line_forte, echo = FALSE}
p_kot_dyn_trial_all_forte <- ggplot(data = kot_dyn_trial_all[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_all_forte
```

```{r kot_dyn_trial_all_line_piano, echo = FALSE}
p_kot_dyn_trial_all_piano <- ggplot(data = kot_dyn_trial_all[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_all_piano
```

# Key-Overlap Ratio (KOR) - Dynamics
## 1. All - individual{#kor_dyn_ind_all}
- Average KOR for each participant.

```{r kor_dyn, echo = FALSE}
# For each individual
kor_dyn <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
kor_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kor_dyn_bar,  echo = FALSE}
p_kor_dyn <- ggplot(data = kor_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn
```

## 2. All - group{#kor_dyn_group_all}
- Average KOR for each condition.

```{r kor_dyn_all, echo = FALSE}
# Group mean
kor_dyn_all <- kor_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kor_dyn_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r kor_dyn_all_bar, echo = FALSE}　
p_kor_dyn_all <- ggplot(data = kor_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn_all
```

#### Box plot
```{r kor_dyn_all_box, echo = FALSE}
p_kor_dyn_all_box <- ggplot(data = kor_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn_all_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality4, echo = FALSE}
# Reduce unused factors
kor_dyn$Subcomponent <- factor(kor_dyn$Subcomponent)
kor_dyn_norm <- by(kor_dyn$Mean, list(kor_dyn$Condition, kor_dyn$Subcomponent), shapiro.test)
print(kor_dyn_norm)
```

#### Two-way ANOVA (ez)
```{r stats7, echo = FALSE}
kor_dyn_aov <- ezANOVA(
  data = dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano"]
  , dv = .(KOR)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kor_dyn_aov$ANOVA$F <- round(kor_dyn_aov$ANOVA$F, 4)
kor_dyn_aov$ANOVA$p <- round(kor_dyn_aov$ANOVA$p, 4)
print(kor_dyn_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats8, echo = FALSE}
kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
summary(kor_dyn_aov_2)
```

## 3. Trial by trial - individual{#kor_dyn_ind_trial}
- Average KOR for each trial for each participant.

```{r kor_dyn_trial, echo = FALSE}
# For each individual
kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kor_dyn_trial
```

### Plot
#### Line graph
```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
p_kor_dyn_trial_forte <- ggplot(data = kor_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_forte
```

```{r kor_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
p_kor_dyn_trial_piano <- ggplot(data = kor_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_piano
```

## 4. Trial by trial - group{#kor_dyn_group_trial}
- Average KOR for each trial for each condition.

```{r kor_dyn_trial_all, echo = FALSE}
# Group mean
kor_dyn_trial_all <- kor_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent,TrialNr)]
kor_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kor_dyn_trial_all_line_forte, echo = FALSE}
p_kor_dyn_trial_all_forte <- ggplot(data = kor_dyn_trial_all[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_all_forte
```

```{r kor_dyn_trial_all_line_piano, echo = FALSE}
p_kor_dyn_trial_all_piano <- ggplot(data = kor_dyn_trial_all[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_all_piano
```

# Sequence plot
## Articulation{#sequence_art}
```{r seq_art, fig.width = 9, fig.height = 3, echo = FALSE}
# For each individual
kot_art_seq <- dt_kot_art[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Interval)]

# Group mean
kot_art_seq_all <- kot_art_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Interval)]

p_kot_art_seq <- ggplot(data = kot_art_seq_all, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
p_kot_art_seq
```

## Dynamics{#sequence_dyn}
```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
kot_dyn_seq <- dt_kot_dyn[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Interval)]

# Group mean
kot_dyn_seq_all <- kot_dyn_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Interval)]

p_kot_dyn_seq <- ggplot(data = kot_dyn_seq_all, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
p_kot_dyn_seq
```

### 2020-11-18 script/RNotebook_2/dynamics.Rmd
update analysis scripts (check later)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Dynamics Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for dynamics.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Outline - Dynamics (Key Velocity)
## [Experimental Design](#design)

## [Dynamics](#vel-dynamics)

### KV
- Individual results
  + [All](#vel_dyn_ind_all)
  + [Trial by trial](#vel_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_dyn_group_all)
  + [Trial by trial](#vel_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#vel_dyn_order_group_all)

### KV-Diff
- Individual results
  + [All](#vel_acc_dyn_ind_all)
  + [Trial by trial](#vel_acc_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_acc_dyn_group_all)
  + [Trial by trial](#vel_acc_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#vel_acc_dyn_order_group_all)

- [Sequence plot](#sequence_dyn)

## [Articulation](#vel_articulation)

### KV
- Individual results
  + [All](#vel_art_ind_all)
  + [Trial by trial](#vel_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_art_group_all)
  + [Trial by trial](#vel_art_group_trial)

### KV-Diff
- Individual results
  + [All](#vel_acc_art_ind_all)
  + [Trial by trial](#vel_acc_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_acc_art_group_all)
  + [Trial by trial](#vel_acc_art_group_trial)

- [Sequence plot](#sequence_art)

## Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato)
- Dependent variables: Key Velocity (KV), KV difference

## Main predictions
Participants will exaggerate forte and piano more while teaching.

## Excluded participnats
- SubNr 1: Most of responses (more than 90%) were removed due to outliers
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("effsize")) {install.packages("effsize"); require("effsize")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r file, include = FALSE}
filename_vel = "../R/preprocessor/trimmed/data_vel.txt"
filename_vel_acc = "../R/preprocessor/trimmed/data_vel_diff.txt"
```

```{r extract, include = FALSE}
dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_vel_acc <- fread(filename_vel_acc, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_vel$SubNr <- as.factor(dt_vel$SubNr)
dt_vel_acc$SubNr <- as.factor(dt_vel_acc$SubNr)

# Exclude SubNr 1 because most responses were removed because of outliers
dt_vel <- dt_vel[SubNr != 1]
dt_vel_acc <- dt_vel_acc[SubNr != 1]

# Include only articulation
dt_vel_art <- dt_vel[Skill == "articulation"]
dt_vel_acc_art <- dt_vel_acc[Skill == "articulation"]

# Include only dynamics
dt_vel_dyn <- dt_vel[Skill == "dynamics"]
dt_vel_acc_dyn <- dt_vel_acc[Skill == "dynamics"]
```

# Key Velocity (KV) - Dynamics
## 1. All - individual{#vel_dyn_ind_all}
- Average KV for each participant.

```{r vel_dyn, echo = FALSE}
# For each individual
vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel_dyn_bar,  echo = FALSE}
p_vel_dyn <- ggplot(data = vel_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
p_vel_dyn
```

## 2. All - group{#vel_dyn_group_all}
- Average KV for each condition.
- Hypothesis: Participants will exaggerate forte and piano when teaching.

```{r vel_dyn_all, echo = FALSE}
# Group mean
vel_dyn_all <- vel_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_dyn_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r vel_dyn_all_bar, echo = FALSE}　
p_vel_dyn_all <- ggplot(data = vel_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
p_vel_dyn_all
```

#### Box plot
```{r vel_dyn_all_box, echo = FALSE}
p_vel_dyn_all_box <- ggplot(data = vel_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
p_vel_dyn_all_box
```

### Statistics
#### Normality check
```{r normality1, echo = FALSE}
# Reduce unused factors
vel_dyn$Subcomponent <- factor(vel_dyn$Subcomponent)
vel_dyn_norm <- by(vel_dyn$Mean, list(vel_dyn$Condition, vel_dyn$Subcomponent), shapiro.test)
print(vel_dyn_norm)
```

#### Two-way ANOVA (ez)
```{r stats1, echo = FALSE}
vel_dyn_aov <- ezANOVA(
  data = dt_vel_dyn
  , dv = .(Velocity)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_dyn_aov$ANOVA$F <- round(vel_dyn_aov$ANOVA$F, 4)
vel_dyn_aov$ANOVA$p <- round(vel_dyn_aov$ANOVA$p, 4)
vel_dyn_aov$ANOVA
```

#### Two-way ANOVA (aov)
```{r stats2, echo = FALSE}
vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
summary(vel_dyn_aov_2)
```

#### Post-hoc (direct comparison)
```{r post-hoc1, echo = FALSE}
vel_ttest_forte <- t.test(vel_dyn[Condition == "performing" & Subcomponent == "Forte"]$Mean, vel_dyn[Condition == "teaching" & Subcomponent == "Forte"]$Mean, paired = TRUE)
vel_ttest_piano <- t.test(vel_dyn[Condition == "performing" & Subcomponent == "Piano"]$Mean, vel_dyn[Condition == "teaching" & Subcomponent == "Piano"]$Mean, paired = TRUE)
# # Effect size
# vel_ttest_forte_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Forte"], vel_dyn$Condition[vel_dyn$Subcomponent == "Forte"], paired = TRUE)
# vel_ttest_piano_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Piano"], vel_dyn$Condition[vel_dyn$Subcomponent == "Piano"], paired = TRUE)

vel_ttest_forte
vel_ttest_piano
# vel_ttest_forte_cohend
# vel_ttest_piano_cohend
```

#### Post-hoc (pairwise t test)
```{r post-hoc2, echo = FALSE}
# Create a group label for a post-hoc test
vel_dyn$Group <- "NA"
vel_dyn[Condition == "performing" & Subcomponent == "Forte"]$Group <- "Performing-Forte" 
vel_dyn[Condition == "performing" & Subcomponent == "Piano"]$Group <- "Performing-Piano"
vel_dyn[Condition == "teaching" & Subcomponent == "Forte"]$Group <- "Teaching-Forte"
vel_dyn[Condition == "teaching" & Subcomponent == "Piano"]$Group <- "Teaching-Piano"
# pairwise comparison
vel_dyn_pairwise <- pairwise.t.test(vel_dyn$Mean, vel_dyn$Group, paired = TRUE, p.adjust.method = "BH")
vel_dyn_pairwise$p.value <- round(vel_dyn_pairwise$p.value, 4)
vel_dyn_pairwise
```

## 3. Trial by trial - individual{#vel_dyn_ind_trial}
- Average KV for each trial for each participant.

```{r vel_dyn_trial, echo = FALSE}
# For each individual
vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_dyn_trial
```

### Plot
#### Line graph
```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
p_vel_dyn_trial_forte <- ggplot(data = vel_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_forte
```

```{r vel_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
p_vel_dyn_trial_piano <- ggplot(data = vel_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_piano
```

## 4. Trial by trial - group{#vel_dyn_group_trial}
- Average KV for each trial for each condition.

```{r vel_dyn_trial_all, echo = FALSE}
# Group mean
vel_dyn_trial_all <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel_dyn_trial_all_line_forte, echo = FALSE}
p_vel_dyn_trial_all_forte <- ggplot(data = vel_dyn_trial_all[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_all_forte
```

```{r vel_dyn_trial_all_line_piano, echo = FALSE}
p_vel_dyn_trial_all_piano <- ggplot(data = vel_dyn_trial_all[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_all_piano
```

# Key Velocity Difference (Difference) - Dynamics
## 1. All - individual{#vel_acc_dyn_ind_all}
- Average KV-Diff for each participant.

```{r vel_acc_dyn, echo = FALSE}
# For each individual
vel_acc_dyn <- dt_vel_acc_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_acc_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel_acc_dyn_bar,  echo = FALSE}
p_vel_acc_dyn <- ggplot(data = vel_acc_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_acc_dyn
```

## 2. All - group{#vel_acc_dyn_group_all}
- Average KV-Diff for each condition.

```{r vel_acc_dyn_all, echo = FALSE}
# Group mean
vel_acc_dyn_all <- vel_acc_dyn[, .(N = .N, Mean = mean(Mean), SD = mean(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_acc_dyn_all
```

### Plot
#### Bar chart
```{r vel_acc_dyn_all_bar,  echo = FALSE}
p_vel_acc_dyn_all <- ggplot(data = vel_acc_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_acc_dyn_all
```

#### Box plot
```{r vel_acc_dyn_all_box, echo = FALSE}
p_vel_acc_dyn_all_box <- ggplot(data = vel_acc_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_acc_dyn_all_box
```

### Statistics
#### Normality check
```{r normality2, echo = FALSE}
# Reduce unused factors
vel_acc_dyn$Subcomponent <- factor(vel_acc_dyn$Subcomponent)
vel_acc_dyn_norm <- by(vel_acc_dyn$Mean, list(vel_acc_dyn$Condition, vel_acc_dyn$Subcomponent), shapiro.test)
vel_acc_dyn_norm
```

#### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
vel_acc_dyn_aov <- ezANOVA(
  data = dt_vel_acc_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF"]
  , dv = .(Diff)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_acc_dyn_aov$ANOVA$F <- round(vel_acc_dyn_aov$ANOVA$F, 4)
vel_acc_dyn_aov$ANOVA$p <- round(vel_acc_dyn_aov$ANOVA$p, 4)
vel_acc_dyn_aov$ANOVA
```

#### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
vel_acc_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_acc_dyn)
summary(vel_acc_dyn_aov_2)
```

#### Post-hoc (direct comparison)
```{r post-hoc3, echo = FALSE}
vel_acc_ttest_ftop <- t.test(vel_acc_dyn[Condition == "performing" & Subcomponent == "FtoP"]$Mean, vel_acc_dyn[Condition == "teaching" & Subcomponent == "FtoP"]$Mean, paired = TRUE)
vel_acc_ttest_ptof <- t.test(vel_acc_dyn[Condition == "performing" & Subcomponent == "PtoF"]$Mean, vel_acc_dyn[Condition == "teaching" & Subcomponent == "PtoF"]$Mean, paired = TRUE)
# # Effect size
# vel_acc_ttest_ftop_cohend <- cohen.d(vel_acc_dyn$Mean[vel_acc_dyn$Subcomponent == "FtoP"], vel_acc_dyn$Condition[vel_acc_dyn$Subcomponent == "FtoP"], paired = TRUE)
# vel_acc_ttest_ptof_cohend <- cohen.d(vel_acc_dyn$Mean[vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Condition[vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)

vel_acc_ttest_ftop
vel_acc_ttest_ptof
# vel_acc_ttest_ftop_cohend
# vel_acc_ttest_ptof_cohend
```

#### Post-hoc (pairwise t test)
```{r post-hoc4, echo = FALSE}
# Create a group label for a post-hoc test
vel_acc_dyn$Group <- "NA"
vel_acc_dyn[Condition == "performing" & Subcomponent == "FtoP"]$Group <- "Performing-FtoP" 
vel_acc_dyn[Condition == "performing" & Subcomponent == "PtoF"]$Group <- "Performing-PtoF"
vel_acc_dyn[Condition == "teaching" & Subcomponent == "FtoP"]$Group <- "Teaching-FtoP"
vel_acc_dyn[Condition == "teaching" & Subcomponent == "PtoF"]$Group <- "Teaching-PtoF"
# pairwise comparison
vel_acc_dyn_pairwise <- pairwise.t.test(vel_acc_dyn$Mean, vel_acc_dyn$Group, paired = TRUE, p.adjust.method = "BH")
vel_acc_dyn_pairwise$p.value <- round(vel_acc_dyn_pairwise$p.value, 4)
vel_acc_dyn_pairwise
```

## 3. Trial by trial - individual{#vel_acc_dyn_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel_acc_dyn_trial, echo = FALSE}
# For each individual
vel_acc_dyn_trial <- dt_vel_acc_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = mean(Diff)), by = .(SubNr, Condition, Skill, Subcomponent,TrialNr)]
vel_acc_dyn_trial
```

### Plot
#### Line graph
```{r vel_acc_dyn_trial_line_ftop, echo = FALSE, fig.height = 4}
p_vel_acc_dyn_trial_ftop <- ggplot(data = vel_acc_dyn_trial[Subcomponent == "FtoP"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_ftop
```

```{r vel_acc_dyn_trial_line_ptof, echo = FALSE, fig.height = 4}
p_vel_acc_dyn_trial_ptof <- ggplot(data = vel_acc_dyn_trial[Subcomponent == "PtoF"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_ptof
```

## 4. Trial by trial - group{#vel_acc_dyn_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel_acc_dyn_trial_all, echo = FALSE}
# Group mean
vel_acc_dyn_trial_all <- vel_acc_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = mean(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_acc_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel_acc_dyn_trial_all_line_ftop, echo = FALSE}
p_vel_acc_dyn_trial_all_ftop <- ggplot(data = vel_acc_dyn_trial_all[Subcomponent == "FtoP"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_all_ftop
```

```{r vel_acc_dyn_trial_all_line_ptof, echo = FALSE}
p_vel_acc_dyn_trial_all_ptof <- ggplot(data = vel_acc_dyn_trial_all[Subcomponent == "PtoF"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_all_ptof
```

# Key Velocity (KV) - Articulation
## 1. All - individual{#vel_art_ind_all}
- Average KV for each participant.

```{r vel_art, echo = FALSE}
# For each individual
vel_art <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel_art_bar,  echo = FALSE}
p_vel_art <- ggplot(data = vel_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
p_vel_art
```

## 2. All - group{#vel_art_group_all}
- Average KV for each condition.

```{r vel_art_all, echo = FALSE}
# Group mean
vel_art_all <- vel_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_art_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r vel_art_all_bar, echo = FALSE}　
p_vel_art_all <- ggplot(data = vel_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
p_vel_art_all
```

#### Box plot
```{r vel_art_all_box, echo = FALSE}
p_vel_art_all_box <- ggplot(data = vel_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation")
p_vel_art_all_box
```

### Statistics
#### Normality check
```{r normality3, echo = FALSE}
# Reduce unused factors
vel_art$Subcomponent <- factor(vel_art$Subcomponent)
vel_art_norm <- by(vel_art$Mean, list(vel_art$Condition, vel_art$Subcomponent), shapiro.test)
vel_art_norm
```

#### Two-way ANOVA (ez)
```{r stats5, echo = FALSE}
vel_art_aov <- ezANOVA(
  data = dt_vel_art
  , dv = .(Velocity)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_art_aov$ANOVA$F <- round(vel_art_aov$ANOVA$F, 4)
vel_art_aov$ANOVA$p <- round(vel_art_aov$ANOVA$p, 4)
vel_art_aov$ANOVA
```

#### Two-way ANOVA (aov)
```{r stats6, echo = FALSE}
vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
summary(vel_art_aov_2)
```

#### Post-hoc (direct comparison)
```{r post-hoc5, echo = FALSE}
vel_ttest_legato <- t.test(vel_art[Condition == "performing" & Subcomponent == "Legato"]$Mean, vel_art[Condition == "teaching" & Subcomponent == "Legato"]$Mean, paired = TRUE)
vel_ttest_staccato <- t.test(vel_art[Condition == "performing" & Subcomponent == "Staccato"]$Mean, vel_art[Condition == "teaching" & Subcomponent == "Staccato"]$Mean, paired = TRUE)
# # Effect size
# vel_ttest_legato_cohend <- cohen.d(vel_art$Mean[vel_art$Subcomponent == "Legato"], vel_art$Condition[vel_art$Subcomponent == "Legato"], paired = TRUE)
# vel_ttest_staccato_cohend <- cohen.d(vel_art$Mean[vel_art$Subcomponent == "Staccato"], vel_art$Condition[vel_art$Subcomponent == "Staccato"], paired = TRUE)

vel_ttest_legato
vel_ttest_staccato
# vel_ttest_legato_cohend
# vel_ttest_staccato_cohend
```

#### Post-hoc (pairwise t test)
```{r post-hoc6, echo = FALSE}
# Create a group label for a post-hoc test
vel_art$Group <- "NA"
vel_art[Condition == "performing" & Subcomponent == "Legato"]$Group <- "Performing-Legato" 
vel_art[Condition == "performing" & Subcomponent == "Staccato"]$Group <- "Performing-Staccato"
vel_art[Condition == "teaching" & Subcomponent == "Legato"]$Group <- "Teaching-Legato"
vel_art[Condition == "teaching" & Subcomponent == "Staccato"]$Group <- "Teaching-Staccato"
# pairwise comparison
vel_art_pairwise <- pairwise.t.test(vel_art$Mean, vel_art$Group, paired = TRUE, p.adjust.method = "BH")
vel_art_pairwise$p.value <- round(vel_art_pairwise$p.value, 4)
vel_art_pairwise
```

## 3. Trial by trial - individual{#vel_art_ind_trial}
- Average KV for each trial for each participant.

```{r vel_art_trial, echo = FALSE}
# For each individual
vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_art_trial
```

### Plot
#### Line graph
```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 4}
p_vel_art_trial_legato <- ggplot(data = vel_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_legato
```

```{r vel_art_trial_line_piano, echo = FALSE, fig.height = 4}
p_vel_art_trial_staccato <- ggplot(data = vel_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_staccato
```

## 4. Trial by trial - group{#vel_art_group_trial}
- Average KV for each trial for each condition.

```{r vel_art_trial_all, echo = FALSE}
# Group mean
vel_art_trial_all <- vel_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel_art_trial_all_line_legato, echo = FALSE}
p_vel_art_trial_all_legato <- ggplot(data = vel_art_trial_all[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_all_legato
```

```{r vel_art_trial_all_line_staccato, echo = FALSE}
p_vel_art_trial_all_staccato <- ggplot(data = vel_art_trial_all[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_all_staccato
```

# Key Velocity Difference (Difference) - Articulation
## 1. All - individual{#vel_acc_art_ind_all}
- Average KV-Diff for each participant.

```{r vel_acc_art, echo = FALSE}
# For each individual
vel_acc_art <- dt_vel_acc_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_acc_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel_acc_art_bar,  echo = FALSE}
p_vel_acc_art <- ggplot(data = vel_acc_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Difference", subtitle = "KV-Diff: Articulaion")
p_vel_acc_art
```

## 2. All - group{#vel_acc_art_group_all}
- Average KV-Diff for each condition.

```{r vel_acc_art_all, echo = FALSE}
# Group mean
vel_acc_art_all <- vel_acc_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_acc_art_all
```

### Plot
#### Bar chart
```{r vel_acc_art_all_bar,  echo = FALSE}
p_vel_acc_art_all <- ggplot(data = vel_acc_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
p_vel_acc_art_all
```

#### Box plot
```{r vel_acc_art_all_box, echo = FALSE}
p_vel_acc_art_all_box <- ggplot(data = vel_acc_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
p_vel_acc_art_all_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality4, echo = FALSE}
# Reduce unused factors
vel_acc_art$Subcomponent <- factor(vel_acc_art$Subcomponent)
vel_acc_art_norm <- by(vel_acc_art$Mean, list(vel_acc_art$Condition, vel_acc_art$Subcomponent), shapiro.test)
vel_acc_art_norm
```

#### Two-way ANOVA (ez)
```{r stats7, echo = FALSE}
vel_acc_art_aov <- ezANOVA(
  data = dt_vel_acc_art[Subcomponent == "LtoS" | Subcomponent == "StoL"]
  , dv = .(Diff)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_acc_art_aov$ANOVA$F <- round(vel_acc_art_aov$ANOVA$F, 4)
vel_acc_art_aov$ANOVA$p <- round(vel_acc_art_aov$ANOVA$p, 4)
print(vel_acc_art_aov$ANOVA)
```

#### Post-hoc (direct comparison)
```{r post-hoc7, echo = FALSE}
vel_acc_ttest_ltos <- t.test(vel_acc_art[Condition == "performing" & Subcomponent == "LtoS"]$Mean, vel_acc_art[Condition == "teaching" & Subcomponent == "LtoS"]$Mean, paired = TRUE)
vel_acc_ttest_stol <- t.test(vel_acc_art[Condition == "performing" & Subcomponent == "StoL"]$Mean, vel_acc_art[Condition == "teaching" & Subcomponent == "StoL"]$Mean, paired = TRUE)
# # Effect size
# vel_acc_ttest_ltos_cohend <- cohen.d(vel_acc_art$Mean[vel_acc_art$Subcomponent == "LtoS"], vel_acc_art$Condition[vel_acc_art$Subcomponent == "LtoS"], paired = TRUE)
# vel_acc_ttest_stol_cohend <- cohen.d(vel_acc_art$Mean[vel_acc_art$Subcomponent == "StoL"], vel_acc_art$Condition[vel_acc_art$Subcomponent == "StoL"], paired = TRUE)

vel_acc_ttest_ltos
vel_acc_ttest_stol
# vel_acc_ttest_ltos_cohend
# vel_acc_ttest_stol_cohend
```

#### Post-hoc (pairwise t test)
```{r post-hoc8, echo = FALSE}
# Create a group label for a post-hoc test
vel_acc_art$Group <- "NA"
vel_acc_art[Condition == "performing" & Subcomponent == "LtoS"]$Group <- "Performing-LtoS" 
vel_acc_art[Condition == "performing" & Subcomponent == "StoL"]$Group <- "Performing-StoL"
vel_acc_art[Condition == "teaching" & Subcomponent == "LtoS"]$Group <- "Teaching-LtoS"
vel_acc_art[Condition == "teaching" & Subcomponent == "StoL"]$Group <- "Teaching-StoL"
# pairwise comparison
vel_acc_art_pairwise <- pairwise.t.test(vel_acc_art$Mean, vel_acc_art$Group, paired = TRUE, p.adjust.method = "BH")
vel_acc_art_pairwise$p.value <- round(vel_acc_art_pairwise$p.value, 4)
vel_acc_art_pairwise
```

#### Two-way ANOVA (aov)
```{r stats8, echo = FALSE}
vel_acc_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_acc_art)
summary(vel_acc_art_aov_2)
```

## 3. Trial by trial - individual{#vel_acc_art_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel_acc_art_trial, echo = FALSE}
# For each individual
vel_acc_art_trial <- dt_vel_acc_art[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_acc_art_trial
```

### Plot
#### Line graph
```{r vel_acc_art_trial_line_ltos, echo = FALSE, fig.height = 4}
p_vel_acc_art_trial_ltos <- ggplot(data = vel_acc_art_trial[Subcomponent == "LtoS"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_ltos
```

```{r vel_acc_art_trial_line_stol, echo = FALSE, fig.height = 4}
p_vel_acc_art_trial_stol <- ggplot(data = vel_acc_art_trial[Subcomponent == "StoL"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Staccato to Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_stol
```

## 4. Trial by trial - group{#vel_acc_art_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel_acc_art_trial_all, echo = FALSE}
# Group mean
vel_acc_art_trial_all <- vel_acc_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_acc_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel_acc_art_trial_all_line_ltos, echo = FALSE}
p_vel_acc_art_trial_all_ltos <- ggplot(data = vel_acc_art_trial_all[Subcomponent == "LtoS"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_all_ltos
```

```{r vel_acc_art_trial_all_line_stol, echo = FALSE}
p_vel_acc_art_trial_all_stol <- ggplot(data = vel_acc_art_trial_all[Subcomponent == "StoL"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_all_stol
```

### 2020-11-18 script/RNotebook_2/ioi.Rmd
update analysis scripts (check later)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # Group mean
-ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art)/2)), by = .(Condition, Skill)]
+ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
 ioi_art_all
@@
 # Group mean
-ioi_art_trial_all <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art_trial)/2)), by = .(Condition, Skill, TrialNr)]
+ioi_art_trial_all <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, TrialNr)]
 ioi_art_trial_all
@@
 # Group mean
-ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn)/2)), by = .(Condition, Skill)]
+ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean),  SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
 ioi_dyn_all
@@
 # Group mean
-ioi_dyn_ch_all <- ioi_dyn_ch[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn_ch)/2)), by = .(Condition, Skill, Boundary)]
+ioi_dyn_ch_all <- ioi_dyn_ch[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Boundary)]
 ioi_dyn_ch_all
@@
 # Group mean
-ioi_dyn_trial_all <- ioi_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn_trial)/2)), by = .(Condition, Skill, TrialNr)]
+ioi_dyn_trial_all <- ioi_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean),  SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, TrialNr)]
 ioi_dyn_trial_all
@@
 # Group mean
-ioi_art_seq_all <- ioi_art_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art_seq)/2)), by = .(Condition, Skill, Interval)]
+ioi_art_seq_all <- ioi_art_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Interval)]
 
@@
 # Group mean
-ioi_dyn_seq_all <- ioi_dyn_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn_seq)/2)), by = .(Condition, Skill, Interval)]
+ioi_dyn_seq_all <- ioi_dyn_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Interval)]
 

### 2020-11-24 script/R/preprocessor/filtered/error_summary.Rmd
move a file into filtered
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Error Summary (teaching-v1.0)"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}

# function
source("./function.R")

# read a text file for ideal performance
dt_ideal <- read.table("./ideal.txt")
colnames(dt_ideal) <- "Pitch"
dt_ideal$RowNr <- 1:nrow(dt_ideal)
setcolorder(dt_ideal, c(2, 1))
# 
dt_onset <- fread(file = "./dt_correct_onset.txt", sep = ",")
dt_offset <- fread(file = "./dt_correct_offset.txt", sep = ",")
error_onset <- fread(file = "./error_onset.txt", sep = ",")
error_offset <- fread(file = "./error_offset.txt", sep = ",")
```

- Each performance (1 trial) contains 67 responses for onsets and 67 responses for offsets (i.e., if there is no mistake, 134 responses in total)
- Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
- Note: SubNr 8 was already excluded due to an experimental error

# Outline
1. [Onset reports](#onset)
2. [Offset reports](#offset)

## ONSET
### Corrected Error Onset (All)
```{r onset-all, echo = FALSE}
error_onset_all <- error_onset[CorrectionNr != "Exclude",]
error_onset_all
```

### Corrected Error Onset (Summary)
```{r onset-summary, echo = FALSE}
error_onset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
error_onset_summary <- error_onset_all[, .(N = .N, Sum = sum(as.numeric(CorrectionNr))), by = .(Reason)]
error_onset_summary
```

### Missing/Excluded Trials Onset (All)
```{r na-onset-all, echo = FALSE}
missing_onset <- checker(dt_onset, dt_ideal)
missing_onset
```

1. **2 participants should be removed from the entire data analysis because they missed at least one block (16 trials)**

- SubNr 3: experimental error (8 trials)
- SubNr 14: played in octave higher (8 trials)

2. **some performances were removed because 1) participants didn't follow sheet music or 2) participants didn't finish till the end (5 trials)**

### Missing/Excluded Trials Onset (Summary)
```{r na-onset-summary, echo = FALSE}
missing_onset_summary <- missing_onset[, .(N = .N), by = .(Reason)]
missing_onset_summary
```

## OFFSET
### Corrected Error Offset (All)
```{r offset-all, echo = FALSE}
error_offset_all <- error_offset[CorrectionNr != "Exclude",]

# separate errors depending on reasons
# SubNr 11, BlockNr 4, TrialNr 8
error_offset_all <- rbind(error_offset_all, error_offset_all[4])
error_offset_all$Reason[4] <- "Extra Notes"
error_offset_all$CorrectionNr[4] <- 1
error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 2
# SubNr 27, BlockNr 3, TrialNr 6
error_offset_all <- rbind(error_offset_all, error_offset_all[5])
error_offset_all$Reason[5] <- "Extra Notes"
error_offset_all$CorrectionNr[5] <- 1
error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 6

error_offset_all
```

### Corrected Error Offset (Summary)
```{r offset-summary, echo = FALSE}
error_offset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
error_offset_summary <- error_offset_all[, .(N = .N, Sum = sum(as.numeric(CorrectionNr))), by = .(Reason)]
error_offset_summary
```

### Missing/Excluded Trials Offset (All)
```{r na-offset-all, echo = FALSE}
missing_offset <- checker(dt_offset, dt_ideal)
missing_offset
```

1. **2 participants should be removed from the entire data analysis because they missed at least one block (16 trials)**

- SubNr 3: experimental error (8 trials)
- SubNr 14: played in octave higher (8 trials)

2. **some performances were removed because 1) participants didn't follow sheet music or 2) participants didn't finish till the end (5 trials)**

### Missing/Excluded Trials Offset (Summary)
```{r na-offset-summary, echo = FALSE}
missing_offset_summary <- missing_offset[, .(N = .N), by = .(Reason)]
missing_offset_summary
```

### 2020-11-24 script/R/preprocessor/filtered/error_summary.Rmd
change paths
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # function
-source("./function.R")
+source("../function.R")
 
 # read a text file for ideal performance
-dt_ideal <- read.table("./ideal.txt")
+dt_ideal <- read.table("../ideal.txt")
 colnames(dt_ideal) <- "Pitch"

### 2020-11-24 script/R/preprocessor/trimmed/included_summary.Rmd
add info about included trials
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Included  Summary (teaching-v1.0)"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}

# read txt files
dt_ioi <- fread(file = "./data_ioi_1.txt", sep = ",")
dt_kot <- fread(file = "./data_kot.txt", sep = ",")
dt_vel <- fread(file = "./data_vel.txt", sep = ",")
dt_vel_diff <- fread(file = "./data_vel_diff.txt", sep = ",")
```

- Each performance (1 trial) contains 67 responses for onsets and 67 responses for offsets (i.e., if there is no mistake, 134 responses in total)
- Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
- Note: SubNr 8 was already excluded due to an experimental error, SubNr 3 & 14 were excluded due to an experimental error

### IOIs (All)
```{r ioi, echo = FALSE}
dt_ioi
```

### IOIs (Summary)
```{r ioi-summary, echo = FALSE}
dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
```
- SubNr 1 is excluded from data analysis because most of the responses were excluded due to outliers.

### KOT (All)
```{r kot, echo = FALSE}
dt_kot[SubNr != 1]
```

### KOT (Summary)
```{r kot-summary, echo = FALSE}
dt_kot[SubNr != 1, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

### Key Velocity (All)
```{r vel, echo = FALSE}
dt_vel[SubNr != 1]
```

### Key Velocity (Individual)
```{r vel-summary, echo = FALSE}
dt_vel[SubNr != 1, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

### KV Difference (All)
```{r vel-diff, echo = FALSE}
dt_vel_diff[SubNr != 1]
```

### KV Difference (Individual)
```{r vel-diff-summary, echo = FALSE}
dt_vel_diff[SubNr != 1, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

## Excluded participnats
- SubNr 1: Most of responses (more than 90%) were removed due to outliers
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

### Export only included participants
data_analysis~.txt are used for data analysis.
```{r export, echo = TRUE}
fwrite(dt_ioi[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./data_analysis_ioi.txt", sep = ",")
fwrite(dt_kot[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./data_analysis_kot.txt", sep = ",")
fwrite(dt_vel[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./data_analysis_vel.txt", sep = ",")
fwrite(dt_vel_diff[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./data_analysis_vel_diff.txt", sep = ",")
```

### 2020-11-24 script/R/analysis/ioi.Rmd
new analysis for ioi
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Tempo Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of tempo analysis.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Skill (2 levels: articulation vs. dynamics)
- Dependent variables: Interonset intervals

1. Main predictions
- Participants will play slower while teaching regardless of the skills.

2. Excluded participants

- SubNr 1: Most of responses (more than 90%) were removed due to outliers
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

3. Stats packages
- [afex](https://github.com/singmann/afex) for ANOVAs

# Outline - Tempo (IOIs)
## [Articulation](#ioi-articulation)

- Individual results
  + [All](#art_ind_all)
  + [Phrase bondaries](#art_ind_phrase)
  + [Trial by trial](#art_ind_trial)

- Group results and stats (if available)
  + [All](#art_group_all)
  + [Phrase bondaries](#art_group_phrase)
  + [Trial by trial](#art_group_trial)

- [Sequence plot](#sequence_art)

## [Dynamics](#ioi-dynamics)

- Individual results
  + [All](#dyn_ind_all)
  + [Phrase bondaries](#dyn_ind_phrase)
  + [Trial by trial](#dyn_ind_trial)

- Group results and stats (if available)
  + [All](#dyn_group_all)
  + [Phrase bondaries](#dyn_group_phrase)
  + [Trial by trial](#dyn_group_trial)

- [Sequence plot](#sequence_dyn)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# statistics
if (!require("afex")) {install.packages("afex"); require("afex")}

# ggplots
theme_set(theme_classic()) 
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r file, include = FALSE}
filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)

# Include only articulation
dt_ioi_art <- dt_ioi[Skill == "articulation"]
# Include only dynamics
dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
```

# IOI-Articulation 
## 1. All - individual{#art_ind_all}
- Average IOIs for each participant.

```{r ioi-art, echo = FALSE}
# For each individual
ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
ioi_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-bar,  echo = FALSE}
p_ioi_art <- ggplot(data = ioi_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
p_ioi_art
```

## 2. All - group{#art_group_all}
- Average IOIs for each condition.
- Hypothesis: Participants will play slower when teaching.

```{r ioi-art-all, echo = FALSE}
# Group mean
ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
ioi_art_all
```

### Plot
#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-plot, echo = FALSE}
ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean)

ggpaired(ioi_art_plot, cond1 = "teaching", cond2 = "performing", color = "condition", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "none")
```

### Statistics
#### Normality check
- Violate the normality for differences between the two conditions

```{r normality1, echo = FALSE}
# Normality check
ioi_art_diff <- ioi_art[Condition == "teaching"]$Mean-ioi_art[Condition == "performing"]$Mean
ioi_art_norm <- shapiro.test(ioi_art_diff)
ioi_art_norm
```

#### Wilcoxon signed-rank test
```{r wilcoxon-art, echo = FALSE}
ioi_art_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_art, paired = TRUE, alternative = "two.sided")
ioi_art_wilcoxon
```

## 3. Phrase boundaries - individual{#art_ind_phrase}
- Compare IOIs at phrase boundaries with others for each participant.

```{r ioi-art-ch, echo = FALSE}
# Define phrase boundaries
dt_ioi_art$Boundary <- "No"
dt_ioi_art[Subcomponent == "LtoS" | Subcomponent == "StoL"]$Boundary <- "Yes"

# For each individual
ioi_art_ch <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
ioi_art_ch
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-ch-bar, echo = FALSE}
p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
p_ioi_art_ch
```

## 4. Phrase boundaries - group{#art_group_phrase}
- Compare IOIs at phrase boundaries with others in the two conditions.
- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations (especially for teaching purposes).

```{r ioi-art-ch-all, echo = FALSE}
# Group mean
ioi_art_ch_all <- ioi_art_ch[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art_ch)/2)), by = .(Condition, Skill, Boundary)]
ioi_art_ch_all
```

### Plot
#### Box plot
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-ch-all-bar, echo = FALSE}
ioi_art_ch_plot <- data.table(subject = ioi_art_ch[Condition == "teaching"]$SubNr, teaching = ioi_art_ch[Condition == "teaching"]$Mean, performing = ioi_art_ch[Condition == "performing"]$Mean, boundary = ioi_art_ch[Condition == "teaching"]$Boundary)

ggpaired(ioi_art_ch_plot, cond1 = "teaching", cond2 = "performing", color = "condition", legend = "none", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-art-boundary, echo = FALSE}
ioi_art_ch_aov <- aov_car(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
summary(ioi_art_ch_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-boundary-2, echo = FALSE}
ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
summary(ioi_art_ch_aov_2)
```

## 5. Trial by trial - individual{#art_ind_trial}
- Average IOIs for each trial for each participant.

```{r ioi-art-trial, echo = FALSE}
# For each individual
ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
ioi_art_trial
```

### Plot
#### Line graph
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-trial-line, echo = FALSE, fig.height = 4}
p_ioi_art_trial <- ggplot(data = ioi_art_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial
```

## 6. Trial by trial - group{#art_group_trial}
- Average IOIs for each trial for each condition.

```{r ioi-art-trial-all, echo = FALSE}
# Group mean
ioi_art_trial_all <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, TrialNr)]
ioi_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-trial-all-line, echo = FALSE}
ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "jco", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# IOI-Dynamics
## 1. All - individual{#dyn_ind_all}
- Average IOIs for each participant.

```{r ioi-dyn, echo = FALSE}
# For each individual
ioi_dyn <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
ioi_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-bar,  echo = FALSE}
p_ioi_dyn <- ggplot(data = ioi_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
p_ioi_dyn
```

## 2. All - group{#dyn_group_all}
- Average IOIs for each condition.
- Hypothesis: Participants will play slower when teaching.

```{r ioi-dyn-all, echo = FALSE}
# Group mean
ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean),  SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
ioi_dyn_all
```

### Plot
#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-all-box, echo = FALSE}
ioi_dyn_plot <- data.table(subject = ioi_dyn[Condition == "teaching"]$SubNr, teaching = ioi_dyn[Condition == "teaching"]$Mean, performing = ioi_dyn[Condition == "performing"]$Mean)

ggpaired(ioi_dyn_plot, cond1 = "teaching", cond2 = "performing", color = "condition", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "none")
```

### Statistics
#### Normality check
- Violate the normality for differences between the two conditions

```{r normality2, echo = FALSE}
# Normality check
ioi_dyn_diff <- ioi_dyn[Condition == "teaching"]$Mean-ioi_dyn[Condition == "performing"]$Mean
ioi_dyn_norm <- shapiro.test(ioi_dyn_diff)
ioi_dyn_norm
```

#### Wilcoxon signed-rank test
```{r wilcoxon-dyn, echo = FALSE}
ioi_dyn_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_dyn, paired = TRUE, alternative = "two.sided")
ioi_dyn_wilcoxon
```

## 3. Phrase boundaries - individual{#dyn_ind_phrase}
- Compare IOIs at phrase boundaries with others for each participant.

```{r ioi-dyn-ch, echo = FALSE}
# Define phrase boundaries
dt_ioi_dyn$Boundary <- "No"
dt_ioi_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF"]$Boundary <- "Yes"

# For each individual
ioi_dyn_ch <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
ioi_dyn_ch
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-ch-bar, echo = FALSE}
p_ioi_dyn_ch <- ggplot(data = ioi_dyn_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
p_ioi_dyn_ch
```

## 4. Phrase boundaries - group{#dyn_group_phrase}
- Compare IOIs at phrase boundaries with others in the two conditions.
- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations (especially for teaching purposes).

```{r ioi-dyn-ch-all, echo = FALSE}
# Group mean
ioi_dyn_ch_all <- ioi_dyn_ch[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Boundary)]
ioi_dyn_ch_all
```

### Plot
#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-ch-all-bar, echo = FALSE}
ioi_dyn_ch_plot <- data.table(subject = ioi_dyn_ch[Condition == "teaching"]$SubNr, teaching = ioi_dyn_ch[Condition == "teaching"]$Mean, performing = ioi_dyn_ch[Condition == "performing"]$Mean, boundary = ioi_dyn_ch[Condition == "teaching"]$Boundary)

ggpaired(ioi_dyn_ch_plot, cond1 = "teaching", cond2 = "performing", color = "condition", legend = "none", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-dyn-boundary, echo = FALSE}
ioi_dyn_ch_aov <- aov_car(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
summary(ioi_dyn_ch_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-boundary2, echo = FALSE}
ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
summary(ioi_dyn_ch_aov_2)
```

## 5. Trial by trial - individual{#dyn_ind_trial}
- Average IOIs for each trial for each participant.

```{r ioi-dyn-trial, echo = FALSE}
# For each individual
ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
ioi_dyn_trial
```

### Plot
#### Line graph
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-trial-line, echo = FALSE, fig.height = 4}
p_ioi_dyn_trial <- ggplot(data = ioi_dyn_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_trial
```

## 6. Trial by trial - group{#dyn_group_trial}
- Average IOIs for each trial for each condition.

```{r ioi-dyn-trial-all, echo = FALSE}
# Group mean
ioi_dyn_trial_all <- ioi_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean),  SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, TrialNr)]
ioi_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_dyn_trial_all_line, echo = FALSE}
ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "jco", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```
# Sequence plot
## Articulation{#sequence_art}
```{r seq-art, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
ioi_art_seq <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]

ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "jco", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  scale_x_continuous(breaks=seq(1,66,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

## Dynamics{#sequence_dyn}
```{r seq-dyn, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
ioi_dyn_seq <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]

ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "jco", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  scale_x_continuous(breaks=seq(1,66,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

### 2020-11-24 script/R/analysis/ioi.Rmd
correct some comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-ggpaired(ioi_art_ch_plot, cond1 = "teaching", cond2 = "performing", color = "condition", legend = "none", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggpaired(ioi_art_ch_plot, cond1 = "teaching", cond2 = "performing", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 
-ggpaired(ioi_dyn_ch_plot, cond1 = "teaching", cond2 = "performing", color = "condition", legend = "none", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggpaired(ioi_dyn_ch_plot, cond1 = "teaching", cond2 = "performing", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 
-```{r ioi_dyn_trial_all_line, echo = FALSE}
+```{r ioi-dyn-trial-all-line, echo = FALSE}
 ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "jco", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +

### 2020-11-25 script/R/analysis/articulation.Rmd
new analysis for kot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Articulation Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for articulation.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: legato vs. staccato / forte vs. piano)
- Dependent variables: Key-Overlap Time (KOT), Key-Overlap Ratio (KOR)

1. Main predictions
Participants will exaggerate legato and staccato more while teaching.

2. Excluded participants

- SubNr 1: Most of responses (more than 90%) were removed due to outliers
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

3. Stats packages
- [afex](https://github.com/singmann/afex) for ANOVAs
- [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons

# Outline - Articulation (KOT, KOR)
## [Articulation](#kot-articulation)

### KOT
- Individual results
  + [All](#kot_art_ind_all)
  + [Trial by trial](#kot_art_ind_trial)

- Group results and stats (if available)
  + [All](#kot_art_group_all)
  + [Trial by trial](#kot_art_group_trial)

- [Sequence plot](#sequence_art)

### KOR
- Individual results
  + [All](#kor_art_ind_all)
  + [Trial by trial](#kor_art_ind_trial)

- Group results and stats (if available)
  + [All](#kor_art_group_all)
  + [Trial by trial](#kor_art_group_trial)

## [Dynamics](#kot-dynamics)

### KOT
- Individual results
  + [All](#kot_dyn_ind_all)
  + [Trial by trial](#kot_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#kot_dyn_group_all)
  + [Trial by trial](#kot_dyn_group_trial)

- [Sequence plot](#sequence_dyn)

### KOR
- Individual results
  + [All](#kor_dyn_ind_all)
  + [Trial by trial](#kor_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#kor_dyn_group_all)
  + [Trial by trial](#kor_dyn_group_trial)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# statistics
if (!require("afex")) {install.packages("afex"); require("afex")}
if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}

# ggplots
theme_set(theme_classic()) 
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r file, include = FALSE}
filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
dt_kot$SubNr <- as.factor(dt_kot$SubNr)

# For each individual, each block, each trial
ioi <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]

# Calculate KOR for each interval (outliers were already determined based on KOT)
dt_kor <- data.table()
for (subnr in unique(dt_kot$SubNr)){
  for (block in unique(dt_kot[SubNr == subnr]$BlockNr)){
    for (trial in unique(dt_kot[SubNr == subnr & BlockNr == block]$TrialNr)){
      dt_current <- dt_kot[SubNr == subnr & BlockNr == block & TrialNr == trial]
      dt_current$KOR <- dt_current$KOT/ioi[SubNr == subnr & ioi$BlockNr == block & TrialNr == trial]$Mean
      dt_kor <- rbind(dt_kor, dt_current)
    }
  }
}

# Include only articulation
dt_kot_art <- dt_kot[Skill == "articulation"]
dt_kor_art <- dt_kor[Skill == "articulation"]

# Include only dynamics
dt_kot_dyn <- dt_kot[Skill == "dynamics"]
dt_kor_dyn <- dt_kor[Skill == "dynamics"]
```

# Key-Overlap Time (KOT) - Articulation{#kot-articulation}
## 1. All - individual{#kot_art_ind_all}
- Average KOT for each participant.

```{r kot-art, echo = FALSE}
# For each individual
kot_art <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
kot_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kot-art-bar,  echo = FALSE}
p_kot_art <- ggplot(data = kot_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art
```

## 2. All - group{#kot_art_group_all}
- Average KOT for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kot-art-all, echo = FALSE}
# Group mean
kot_art_all <- kot_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kot_art_all
```

### Plot
#### Box plot
```{r kot-art-all-box, echo = FALSE}
kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent)

ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (car_aov)
```{r anova-art-kot-1, echo = FALSE}
kot_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
summary(kot_art_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-kot-2, echo = FALSE}
kot_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
summary(kot_art_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-1, echo = FALSE}
test(pairs(emmeans(kot_art_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(kot_art_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#kot_art_ind_trial}
- Average KOT for each trial for each participant.

```{r kot-art-trial, echo = FALSE}
# For each individual
kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kot_art_trial
```

### Plot
#### Line graph
```{r kot-art-trial-line-legato, echo = FALSE, fig.height = 4}
p_kot_art_trial_legato <- ggplot(data = kot_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_legato
```

```{r kot-art-trial-line-staccato, echo = FALSE, fig.height = 4}
p_kot_art_trial_staccato <- ggplot(data = kot_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_staccato
```

## 4. Trial by trial - group{#kot_art_group_trial}
- Average KOT for each trial for each condition.

```{r kot-art-trial-all, echo = FALSE}
# Group mean
kot_art_trial_all <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kot_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kot-art-trial-all-line-legato, echo = FALSE}
ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r kot-art-trial-all-line-staccato, echo = FALSE}
ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Key-Overlap Ratio (KOR) - Articulation
## 1. All - individual{#kor_art_ind_all}
- Average KOR for each participant.

```{r kor-art, echo = FALSE}
# For each individual
kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
kor_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kor-art-bar,  echo = FALSE}
p_kor_art <- ggplot(data = kor_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art
```

## 2. All - group{#kor_art_group_all}
- Average KOR for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kor-art-all, echo = FALSE}
# Group mean
kor_art_all <- kor_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kor_art_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r kor-art-all-bar, echo = FALSE}
kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent)

ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (car_aov)
```{r anova-art-kor-1, echo = FALSE}
kor_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
summary(kor_art_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-kor-2, echo = FALSE}
kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
summary(kor_art_aov_2)    
```

#### Post-hoc (emmeans)
```{r post-hoc-2, echo = FALSE}
test(pairs(emmeans(kor_art_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(kor_art_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#kor_art_ind_trial}
- Average KOR for each trial for each participant.

```{r kor-art-trial, echo = FALSE}
# For each individual
kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kor_art_trial
```

### Plot
#### Line graph
```{r kor-art-trial-line-legato, echo = FALSE, fig.height = 4}
p_kor_art_trial_legato <- ggplot(data = kor_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_legato
```

```{r kor-art-trial-line-staccato, echo = FALSE, fig.height = 4}
p_kor_art_trial_staccato <- ggplot(data = kor_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_staccato
```

## 4. Trial by trial - group{#kor_art_group_trial}
- Average KOR d for each trial for each condition.

```{r kor-art-trial-all, echo = FALSE}
# Group mean
kor_art_trial_all <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kor_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kor-art-trial-all-line-legato, echo = FALSE}
ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r kor-art-trial-all-line-staccato, echo = FALSE}
ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Key-Overlap Time (KOT) - Dynamics{#kot-dynamics}
## 1. All - individual{#kot_dyn_ind_all}
- Average KOT for each participant.

```{r kot-dyn, echo = FALSE}
# For each individual
kot_dyn <- dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
kot_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kot-dyn-bar,  echo = FALSE}
p_kot_dyn <- ggplot(data = kot_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn
```

## 2. All - group{#kot_dyn_group_all}
- Average KOT for each condition.

```{r kot-dyn-all, echo = FALSE}
# Group mean
kot_dyn_all <- kot_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kot_dyn_all
```

### Plot
#### Box plot

```{r kot-dyn-all-box, echo = FALSE}
kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (car_aov)
```{r anova-dyn-kot-1, echo = FALSE}
kot_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
summary(kot_dyn_aov)
```

#### Two-way ANOVA (aov)
```{r anova-dyn-kot-2, echo = FALSE}
kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
summary(kot_dyn_aov_2)
```

## 3. Trial by trial - individual{#kot_dyn_ind_trial}
- Average KOT for each trial for each participant.

```{r kot-dyn-trial, echo = FALSE}
# For each individual
kot_dyn_trial <- dt_kot_dyn[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kot_dyn_trial
```

### Plot
#### Line graph
```{r kot-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
p_kot_dyn_trial_forte <- ggplot(data = kot_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_forte
```

```{r kot-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
p_kot_dyn_trial_piano <- ggplot(data = kot_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_piano
```

## 4. Trial by trial - group{#kot_dyn_group_trial}
- Average KOT for each trial for each condition.

```{r kot-dyn-trial-all, echo = FALSE}
# Group mean
kot_dyn_trial_all <- kot_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kot_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kot-dyn-trial-all-line-forte, echo = FALSE}
ggline(kot_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r kot-dyn-trial-all-line-piano, echo = FALSE}
ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Key-Overlap Ratio (KOR) - Dynamics
## 1. All - individual{#kor_dyn_ind_all}
- Average KOR for each participant.

```{r kor-dyn, echo = FALSE}
# For each individual
kor_dyn <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
kor_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kor-dyn-bar,  echo = FALSE}
p_kor_dyn <- ggplot(data = kor_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn
```

## 2. All - group{#kor_dyn_group_all}
- Average KOR for each condition.

```{r kor-dyn-all, echo = FALSE}
# Group mean
kor_dyn_all <- kor_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kor_dyn_all
```

### Plot
#### Box plot
```{r kor-dyn-all-box, echo = FALSE}
kor_dyn_plot <- data.table(subject = kor_dyn[Condition == "teaching"]$SubNr, teaching = kor_dyn[Condition == "teaching"]$Mean, performing = kor_dyn[Condition == "performing"]$Mean, subcomponent = kor_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (car_aov)
```{r anova-dyn-kor-1, echo = FALSE}
kor_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
summary(kor_dyn_aov)
```

#### Two-way ANOVA (aov)
```{r anova-dyn-kor-2, echo = FALSE}
kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
summary(kor_dyn_aov_2)
```

## 3. Trial by trial - individual{#kor_dyn_ind_trial}
- Average KOR for each trial for each participant.

```{r kor-dyn-trial, echo = FALSE}
# For each individual
kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kor_dyn_trial
```

### Plot
#### Line graph
```{r kor-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
p_kor_dyn_trial_forte <- ggplot(data = kor_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_forte
```

```{r kor-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
p_kor_dyn_trial_piano <- ggplot(data = kor_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_piano
```

## 4. Trial by trial - group{#kor_dyn_group_trial}
- Average KOR for each trial for each condition.

```{r kor-dyn-trial-all, echo = FALSE}
# Group mean
kor_dyn_trial_all <- kor_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent,TrialNr)]
kor_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kor-dyn-trial-all-line-forte, echo = FALSE}
ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Forte") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r kor_dyn_trial_all_line_piano, echo = FALSE}
ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Sequence plot
## Articulation{#sequence_art}
```{r seq-art, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
kot_art_seq <- dt_kot_art[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Interval)]

ggline(kot_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Articulation") +
  scale_x_continuous(breaks=seq(1,66,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

## Dynamics{#sequence_dyn}
```{r seq-dyn, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
kot_dyn_seq <- dt_kot_dyn[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Interval)]

ggline(kot_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Dynamics") +
  scale_x_continuous(breaks=seq(1,66,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

### 2020-11-25 script/R/analysis/dynamics.Rmd
new analysis for dynamics
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Dynamics Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of dynamics analysis.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato)
- Dependent variables: Key Velocity (KV), KV difference

1. Main predictions
Participants will exaggerate forte and piano more while teaching.

2. Excluded participants

- SubNr 1: Most of responses (more than 90%) were removed due to outliers
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

3. Stats packages
- [afex](https://github.com/singmann/afex) for ANOVAs
- [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons

# Outline - Dynamics (Key Velocity)

## [Dynamics](#vel-dynamics)

### KV
- Individual results
  + [All](#vel_dyn_ind_all)
  + [Trial by trial](#vel_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_dyn_group_all)
  + [Trial by trial](#vel_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#vel_dyn_order_group_all)

### KV-Diff
- Individual results
  + [All](#vel_diff_dyn_ind_all)
  + [Trial by trial](#vel_diff_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_diff_dyn_group_all)
  + [Trial by trial](#vel_diff_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#vel_diff_dyn_order_group_all)

- [Sequence plot](#sequence_dyn)

## [Articulation](#vel_articulation)

### KV
- Individual results
  + [All](#vel_art_ind_all)
  + [Trial by trial](#vel_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_art_group_all)
  + [Trial by trial](#vel_art_group_trial)

### KV-Diff
- Individual results
  + [All](#vel_diff_art_ind_all)
  + [Trial by trial](#vel_diff_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_diff_art_group_all)
  + [Trial by trial](#vel_diff_art_group_trial)

- [Sequence plot](#sequence_art)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# statistics
if (!require("afex")) {install.packages("afex"); require("afex")}
if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}

# ggplots
theme_set(theme_classic()) 
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r file, include = FALSE}
filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
```

```{r extract, include = FALSE}
dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_vel$SubNr <- as.factor(dt_vel$SubNr)
dt_vel_diff$SubNr <- as.factor(dt_vel_diff$SubNr)

# Include only articulation
dt_vel_art <- dt_vel[Skill == "articulation"]
dt_vel_diff_art <- dt_vel_diff[Skill == "articulation"]

# Include only dynamics
dt_vel_dyn <- dt_vel[Skill == "dynamics"]
dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
```

# Key Velocity (KV) - Dynamics
## 1. All - individual{#vel_dyn_ind_all}
- Average KV for each participant.

```{r vel-dyn, echo = FALSE}
# For each individual
vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel-dyn-bar,  echo = FALSE}
p_vel_dyn <- ggplot(data = vel_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
p_vel_dyn
```

## 2. All - group{#vel_dyn_group_all}
- Average KV for each condition.
- Hypothesis: Participants will exaggerate forte and piano when teaching.

```{r vel-dyn-all, echo = FALSE}
# Group mean
vel_dyn_all <- vel_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_dyn_all
```

### Plot
#### Box plot
```{r vel-dyn-all-box, echo = FALSE}
vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-dyn-vel-1, echo = FALSE}
vel_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
summary(vel_dyn_aov)
```

#### Two-way ANOVA (aov)
```{r anova-dyn-vel-2, echo = FALSE}
vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
summary(vel_dyn_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-1, echo = FALSE}
test(pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#vel_dyn_ind_trial}
- Average KV for each trial for each participant.

```{r vel-dyn-trial, echo = FALSE}
# For each individual
vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_dyn_trial
```

### Plot
#### Line graph
```{r vel-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
p_vel_dyn_trial_forte <- ggplot(data = vel_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_forte
```

```{r vel-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
p_vel_dyn_trial_piano <- ggplot(data = vel_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_piano
```

## 4. Trial by trial - group{#vel_dyn_group_trial}
- Average KV for each trial for each condition.

```{r vel-dyn-trial-all, echo = FALSE}
# Group mean
vel_dyn_trial_all <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel-dyn-trial-all-line-forte, echo = FALSE}
ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r vel-dyn-trial-all-line-piano, echo = FALSE}
ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Key Velocity Difference (Difference) - Dynamics
## 1. All - individual{#vel_diff_dyn_ind_all}
- Average KV-Diff for each participant.

```{r vel-acc-dyn, echo = FALSE}
# For each individual
vel_diff_dyn <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_diff_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel-acc-dyn-bar,  echo = FALSE}
p_vel_diff_dyn <- ggplot(data = vel_diff_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_diff_dyn
```

## 2. All - group{#vel_diff_dyn_group_all}
- Average KV-Diff for each condition.

```{r vel-acc-dyn-all, echo = FALSE}
# Group mean
vel_diff_dyn_all <- vel_diff_dyn[, .(N = .N, Mean = mean(Mean), SD = mean(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_diff_dyn_all
```

### Plot
#### Box plot
```{r vel-acc-dyn-all-box, echo = FALSE}
vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-dyn-vel-diff-1, echo = FALSE}
vel_diff_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
summary(vel_diff_dyn_aov)
```

#### Two-way ANOVA (aov)
```{r anova-dyn-vel-diff-2, echo = FALSE}
vel_diff_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
summary(vel_diff_dyn_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-2, echo = FALSE}
test(pairs(emmeans(vel_diff_dyn_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(vel_diff_dyn_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#vel_diff_dyn_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel-diff-dyn-trial, echo = FALSE}
# For each individual
vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = mean(Diff)), by = .(SubNr, Condition, Skill, Subcomponent,TrialNr)]
vel_diff_dyn_trial
```

### Plot
#### Line graph
```{r vel-diff-dyn-trial-line-ftop, echo = FALSE, fig.height = 4}
p_vel_diff_dyn_trial_ftop <- ggplot(data = vel_diff_dyn_trial[Subcomponent == "FtoP"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_diff_dyn_trial_ftop
```

```{r vel-diff-dyn-trial-line-ptof, echo = FALSE, fig.height = 4}
p_vel_diff_dyn_trial_ptof <- ggplot(data = vel_diff_dyn_trial[Subcomponent == "PtoF"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_diff_dyn_trial_ptof
```

## 4. Trial by trial - group{#vel_diff_dyn_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel-diff-dyn-trial-all, echo = FALSE}
# Group mean
vel_diff_dyn_trial_all <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = mean(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_diff_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel-diff-dyn-trial-all-line-ftop, echo = FALSE}
ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r vel-diff-dyn-trial-all-line-ptof, echo = FALSE}
ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Key Velocity (KV) - Articulation
## 1. All - individual{#vel_art_ind_all}
- Average KV for each participant.

```{r vel-art, echo = FALSE}
# For each individual
vel_art <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel-art-bar,  echo = FALSE}
p_vel_art <- ggplot(data = vel_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
p_vel_art
```

## 2. All - group{#vel_art_group_all}
- Average KV for each condition.

```{r vel-art-all, echo = FALSE}
# Group mean
vel_art_all <- vel_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_art_all
```

### Plot
#### Box plot
```{r vel-art-all-box, echo = FALSE}
vel_art_plot <- data.table(subject = vel_art[Condition == "teaching"]$SubNr, teaching = vel_art[Condition == "teaching"]$Mean, performing = vel_art[Condition == "performing"]$Mean, subcomponent = vel_art[Condition == "teaching"]$Subcomponent)

ggpaired(vel_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-art-vel-1, echo = FALSE}
vel_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
summary(vel_art_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-vel-2, echo = FALSE}
vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
summary(vel_art_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-3, echo = FALSE}
test(pairs(emmeans(vel_art_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(vel_art_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#vel_art_ind_trial}
- Average KV for each trial for each participant.

```{r vel-art-trial, echo = FALSE}
# For each individual
vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_art_trial
```

### Plot
#### Line graph
```{r vel-art-trial-line-legato, echo = FALSE, fig.height = 4}
p_vel_art_trial_legato <- ggplot(data = vel_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_legato
```

```{r vel-art-trial-line-staccato, echo = FALSE, fig.height = 4}
p_vel_art_trial_staccato <- ggplot(data = vel_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_staccato
```

## 4. Trial by trial - group{#vel_art_group_trial}
- Average KV for each trial for each condition.

```{r vel-art-trial-all, echo = FALSE}
# Group mean
vel_art_trial_all <- vel_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel-art-trial-all-line-legato, echo = FALSE}
ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r vel-art-trial-all-line-staccato, echo = FALSE}
ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Key Velocity Difference (Difference) - Articulation
## 1. All - individual{#vel_diff_art_ind_all}
- Average KV-Diff for each participant.

```{r vel-diff-art, echo = FALSE}
# For each individual
vel_diff_art <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_diff_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel-diff-art-bar,  echo = FALSE}
p_vel_diff_art <- ggplot(data = vel_diff_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Difference", subtitle = "KV-Diff: Articulaion")
p_vel_diff_art
```

## 2. All - group{#vel_diff_art_group_all}
- Average KV-Diff for each condition.

```{r vel-diff-art-all, echo = FALSE}
# Group mean
vel_diff_art_all <- vel_diff_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_diff_art_all
```

### Plot
#### Box plot
```{r vel-diff-art-all-box, echo = FALSE}
vel_diff_art_plot <- data.table(subject = vel_diff_art[Condition == "teaching"]$SubNr, teaching = vel_diff_art[Condition == "teaching"]$Mean, performing = vel_diff_art[Condition == "performing"]$Mean, subcomponent = vel_diff_art[Condition == "teaching"]$Subcomponent)

ggpaired(vel_diff_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV-Diff: Articulation") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-art-vel-diff-1, echo = FALSE}
vel_diff_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_art)
summary(vel_diff_art_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-vel-diff-2, echo = FALSE}
vel_diff_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_art)
summary(vel_diff_art_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-4, echo = FALSE}
test(pairs(emmeans(vel_diff_art_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(vel_diff_art_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#vel_diff_art_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel-diff-art-trial, echo = FALSE}
# For each individual
vel_diff_art_trial <- dt_vel_diff_art[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_diff_art_trial
```

### Plot
#### Line graph
```{r vel-diff-artt-rial-line-ltos, echo = FALSE, fig.height = 4}
p_vel_diff_art_trial_ltos <- ggplot(data = vel_diff_art_trial[Subcomponent == "LtoS"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_diff_art_trial_ltos
```

```{r vel-diff-art-trial-line-stol, echo = FALSE, fig.height = 4}
p_vel_diff_art_trial_stol <- ggplot(data = vel_diff_art_trial[Subcomponent == "StoL"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Staccato to Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_diff_art_trial_stol
```

## 4. Trial by trial - group{#vel_diff_art_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel-diff-art-trial-all, echo = FALSE}
# Group mean
vel_diff_art_trial_all <- vel_diff_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_diff_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel-diff-art-trial-all-line-ltos, echo = FALSE}
ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r vel-diff-art-trial-all-line-stol, echo = FALSE}
ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

## Dynamics{#sequence_dyn}
```{r seq-dyn, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
vel_dyn_seq <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, RowNr)]

ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Note", ylab = "", title = "KV: Dynamics") +
  scale_x_continuous(breaks=seq(1,67,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")

vel_diff_dyn_seq <- dt_vel_diff_dyn[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Interval)]

ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") +
  scale_x_continuous(breaks=seq(1,66,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

## Articulation{#sequence_art}
```{r seq-art, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
vel_art_seq <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, RowNr)]

ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Note", ylab = "", title = "KV: Articulation") +
  scale_x_continuous(breaks=seq(1,67,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")

vel_diff_art_seq <- dt_vel_diff_art[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Interval)]

ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") +
  scale_x_continuous(breaks=seq(1,66,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

### 2020-11-25 script/R/analysis/ioi.Rmd
correct some comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # Experimental Design{#design}
-- Independent variables: Condition (2 levels: teaching vs. performing) x Skill (2 levels: articulation vs. dynamics)
-- Dependent variables: Interonset intervals
+- Independent variables: Condition (2 levels: teaching vs. performing) x Technique (2 levels: articulation vs. dynamics)
+- Dependent variables: Interonset intervals (IOIs)
 
 1. Main predictions
-- Participants will play slower while teaching regardless of the skills.
+- Participants will play slower while teaching regardless of the techniques.
 
@@
 
-ggpaired(ioi_art_plot, cond1 = "teaching", cond2 = "performing", color = "condition", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 
-ggpaired(ioi_art_ch_plot, cond1 = "teaching", cond2 = "performing", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggpaired(ioi_art_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 ```{r ioi-art-trial-all-line, echo = FALSE}
-ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "jco", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 
-ggpaired(ioi_dyn_plot, cond1 = "teaching", cond2 = "performing", color = "condition", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggpaired(ioi_dyn_plot, cond1 = "teaching", cond2 = "performing", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 
-ggpaired(ioi_dyn_ch_plot, cond1 = "teaching", cond2 = "performing", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "jco", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 ```{r ioi-dyn-trial-all-line, echo = FALSE}
-ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "jco", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 
-ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "jco", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 
-ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "jco", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   geom_hline(yintercept = 188, linetype = "dashed") +

### 2020-12-01 script/R/analysis/ioi.Rmd
modify figures and comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("afex")) {install.packages("afex"); require("afex")}
-
-# ggplots
-theme_set(theme_classic()) 
-theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
 ```
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Boxplot
 - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
-```{r ioi-art-bar,  echo = FALSE}
-p_ioi_art <- ggplot(data = ioi_art, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
-p_ioi_art
+```{r ioi-art-box,  echo = FALSE}
+ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 
-```{r ioi-art-plot, echo = FALSE}
+```{r ioi-art-all-box, echo = FALSE}
 ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean)
@@
   geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "none")
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
-```{r ioi-art-ch-bar, echo = FALSE}
-p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+```{r ioi-art-ch-box, echo = FALSE}
+ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
-p_ioi_art_ch
+  geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 #### Box plot
-- Error bars show SEM of the mean
 - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
-```{r ioi-art-ch-all-bar, echo = FALSE}
+```{r ioi-art-ch-all-box, echo = FALSE}
 ioi_art_ch_plot <- data.table(subject = ioi_art_ch[Condition == "teaching"]$SubNr, teaching = ioi_art_ch[Condition == "teaching"]$Mean, performing = ioi_art_ch[Condition == "performing"]$Mean, boundary = ioi_art_ch[Condition == "teaching"]$Boundary)
@@
   geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
-  theme(legend.position = "none")
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ```{r ioi-art-trial-line, echo = FALSE, fig.height = 4}
-p_ioi_art_trial <- ggplot(data = ioi_art_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+ggline(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   facet_wrap(SubNr ~ .) +
-  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_art_trial
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1)) +
+  geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 ```{r ioi-art-trial-all-line, echo = FALSE}
-ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   geom_hline(yintercept = 188, linetype = "dashed") +
-  scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  scale_x_continuous(breaks = seq(1,8,1)) +
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
-```{r ioi-dyn-bar,  echo = FALSE}
-p_ioi_dyn <- ggplot(data = ioi_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
-p_ioi_dyn
+```{r ioi-dyn-box,  echo = FALSE}
+ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 
-ggpaired(ioi_dyn_plot, cond1 = "teaching", cond2 = "performing", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "none")
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
-```{r ioi-dyn-ch-bar, echo = FALSE}
-p_ioi_dyn_ch <- ggplot(data = ioi_dyn_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+```{r ioi-dyn-ch-box, echo = FALSE}
+ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
-  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
-p_ioi_dyn_ch
+  geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 
-```{r ioi-dyn-ch-all-bar, echo = FALSE}
+```{r ioi-dyn-ch-all-box, echo = FALSE}
 ioi_dyn_ch_plot <- data.table(subject = ioi_dyn_ch[Condition == "teaching"]$SubNr, teaching = ioi_dyn_ch[Condition == "teaching"]$Mean, performing = ioi_dyn_ch[Condition == "performing"]$Mean, boundary = ioi_dyn_ch[Condition == "teaching"]$Boundary)
@@
 ```{r ioi-dyn-trial-line, echo = FALSE, fig.height = 4}
-p_ioi_dyn_trial <- ggplot(data = ioi_dyn_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
+ggline(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   facet_wrap(SubNr ~ .) +
-  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
-p_ioi_dyn_trial
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1)) +
+  geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 ```{r ioi-dyn-trial-all-line, echo = FALSE}
-ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   geom_hline(yintercept = 188, linetype = "dashed") +
-  scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  scale_x_continuous(breaks = seq(1,8,1)) +
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
   geom_hline(yintercept = 188, linetype = "dashed") +
-  scale_x_continuous(breaks=seq(1,66,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  scale_x_continuous(breaks = seq(1,66,1)) +
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```

### 2020-12-01 script/R/analysis/articulation.Rmd
correct comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
modify figures/comments (TBC)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a summary of all analyses for articulation.
+Description: This is a summary of articulation analysis.
 
@@
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
-
-# ggplots
-theme_set(theme_classic()) 
-theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
 ```
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 
-```{r kot-art-bar,  echo = FALSE}
-p_kot_art <- ggplot(data = kot_art, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+```{r kot-art-box,  echo = FALSE}
+ggboxplot(dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") +
   facet_grid(Subcomponent ~ .) +
-  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
-p_kot_art
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ### Plot
+#### Scatter plot
+```{r kot-art-all-scatter, echo = FALSE}
+ggerrorplot(kot_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+```
+
 #### Box plot
@@
 ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
-  theme(legend.position = "none")
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ```{r kot-art-trial-line-legato, echo = FALSE, fig.height = 4}
-p_kot_art_trial_legato <- ggplot(data = kot_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
-p_kot_art_trial_legato
+ggline(dt_kot_art[Subcomponent == "Legato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-p_kot_art_trial_staccato <- ggplot(data = kot_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
-p_kot_art_trial_staccato
+ggline(dt_kot_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-art-trial-all-line-legato, echo = FALSE}
-ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
+ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
   scale_x_continuous(breaks=seq(1,8,1)) +
@@
 ```{r kot-art-trial-all-line-staccato, echo = FALSE}
-ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
+ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
   scale_x_continuous(breaks=seq(1,8,1)) +
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 
-```{r kor-art-bar,  echo = FALSE}
-p_kor_art <- ggplot(data = kor_art, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+```{r kor-art-box,  echo = FALSE}
+ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
   facet_grid(Subcomponent ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Articulation")
-p_kor_art
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ### Plot
-#### Bar chart
+#### Scatter plot
 - Error bars show SEM of the mean
+
+```{r kor-art-all-scatter, echo = FALSE}
+ggerrorplot(kor_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+```
 
@@
 ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
-  theme(legend.position = "none")
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ```{r kor-art-trial-line-legato, echo = FALSE, fig.height = 4}
-p_kor_art_trial_legato <- ggplot(data = kor_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
-p_kor_art_trial_legato
+ggline(dt_kor_art[Subcomponent == "Legato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Legato") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kor-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-p_kor_art_trial_staccato <- ggplot(data = kor_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
-p_kor_art_trial_staccato
+ggline(dt_kor_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Staccato") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kor-art-trial-all-line-legato, echo = FALSE}
-ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
+ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
   scale_x_continuous(breaks=seq(1,8,1)) +
@@
 ```{r kor-art-trial-all-line-staccato, echo = FALSE}
-ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
+ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
   scale_x_continuous(breaks=seq(1,8,1)) +
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 
-```{r kot-dyn-bar,  echo = FALSE}
-p_kot_dyn <- ggplot(data = kot_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+```{r kot-dyn-box,  echo = FALSE}
+ggboxplot(dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Dynamics") +
   facet_grid(Subcomponent ~ .) +
-  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
-p_kot_dyn
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics") +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
-  theme(legend.position = "none")
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ```{r kot-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-p_kot_dyn_trial_forte <- ggplot(data = kot_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
-p_kot_dyn_trial_forte
+ggline(dt_kot_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-p_kot_dyn_trial_piano <- ggplot(data = kot_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
-p_kot_dyn_trial_piano
+ggline(dt_kot_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-dyn-trial-all-line-forte, echo = FALSE}
-ggline(kot_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
+ggline(kot_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
   scale_x_continuous(breaks=seq(1,8,1)) +
@@
 ```{r kot-dyn-trial-all-line-piano, echo = FALSE}
-ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
+ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
   scale_x_continuous(breaks=seq(1,8,1)) +
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
-
-```{r kor-dyn-bar,  echo = FALSE}
-p_kor_dyn <- ggplot(data = kor_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+#### Box plot
+```{r kor-dyn-box,  echo = FALSE}
+ggboxplot(dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Dynamics") +
   facet_grid(Subcomponent ~ .) +
-  labs(y = "KOR", subtitle = "KOR: Dynamics")
-p_kor_dyn
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
-  theme(legend.position = "none")
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ```{r kor-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-p_kor_dyn_trial_forte <- ggplot(data = kor_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
-p_kor_dyn_trial_forte
+ggline(dt_kor_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Forte") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kor-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-p_kor_dyn_trial_piano <- ggplot(data = kor_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
-  geom_line() +
-  geom_point() +
-  facet_wrap(SubNr ~ .) + 
-  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
-p_kor_dyn_trial_piano
+ggline(dt_kor_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Piano") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
   scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ```{r kor_dyn_trial_all_line_piano, echo = FALSE}
-ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") +
+ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") +
   scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
   scale_x_continuous(breaks=seq(1,66,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```

### 2020-12-01 script/R/analysis/ioi.Rmd
modify figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   geom_hline(yintercept = 188, linetype = "dashed") +
-  scale_x_continuous(breaks = seq(1,8,1)) +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
   geom_hline(yintercept = 188, linetype = "dashed") +
-  scale_x_continuous(breaks = seq(1,66,1)) +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,66,1))
 ```
@@
   geom_hline(yintercept = 188, linetype = "dashed") +
-  scale_x_continuous(breaks = seq(1,66,1)) +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+    scale_x_continuous(breaks = seq(1,66,1))
 ```

### 2020-12-01 script/R/analysis/articulation.Rmd
modify figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 #### Scatter plot
+- Error bars show SEM of the mean
+
 ```{r kot-art-all-scatter, echo = FALSE}
@@
 ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
-  scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  scale_x_continuous(breaks = seq(1,8,1)) +
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
-  scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  scale_x_continuous(breaks = seq(1,8,1)) +
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
-  scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
-  scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ### Plot
+#### Scatter plot
+- Error bars show SEM of the mean
+
+```{r kot-dyn-all-scatter, echo = FALSE}
+ggerrorplot(kot_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+```
+
 #### Box plot
-
 ```{r kot-dyn-all-box, echo = FALSE}
@@
 ggline(kot_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
-  scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
-  scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
-  theme(legend.position = "bottom")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ### Plot
+#### Scatter plot
+- Error bars show SEM of the mean
+
+```{r kor-dyn-all-scatter, echo = FALSE}
+ggerrorplot(kor_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+```
+
 #### Box plot
@@
 ```{r kor-dyn-trial-all-line-forte, echo = FALSE}
-ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Forte") +
-  scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Forte") +
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") +
-  scale_x_continuous(breaks=seq(1,8,1)) +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,66,1))
 ```
@@
 
-ggline(kot_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  scale_x_continuous(breaks=seq(1,66,1)) +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggline(kot_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Articulation") +
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,66,1))
 ```
@@
 
-ggline(kot_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Dynamics") +
-  scale_x_continuous(breaks=seq(1,66,1)) +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggline(kot_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Dynamics")+
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,66,1)) 
 ```

### 2020-12-02 script/R/analysis/ioi.Rmd
modfiy fig size
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ## Articulation{#sequence_art}
-```{r seq-art, fig.width = 10, fig.height = 2, echo = FALSE}
+```{r seq-art, fig.width = 9, fig.height = 3, echo = FALSE}
 # For each individual
@@
   geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,66,1))
@@
 ## Dynamics{#sequence_dyn}
-```{r seq-dyn, fig.width = 10, fig.height = 2, echo = FALSE}
+```{r seq-dyn, fig.width = 9, fig.height = 3, echo = FALSE}
 # For each individual
@@
   geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
     scale_x_continuous(breaks = seq(1,66,1))

### 2020-12-02 script/R/demographics/questionnaire.Rmd
update demographics
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Demographics: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of demographic information

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

# Questionnaire

```{r file, include = FALSE}
filename_q = "./questionnaire.csv"
```

```{r extract, echo = FALSE}
# read csv files
dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")

# change some colnames
colnames(dt)[2] <- "SubNr"
colnames(dt)[3] <- "Age"
colnames(dt)[4] <- "Gender"
colnames(dt)[7] <- "Handedness"

# change some characteristics
dt$Age <- as.numeric(dt$Age)

# extract the total number of years of training in piano (input manually based on the original csv file)
dt$PianoTotalPractice <- c(14, 20, 27, 21, 10, 16, 9, 13, 25, 14, 13, 11, 6, 4, 6, 8, 21, 15, 18, 7, 11, 12, 12, 18, 12, 5, 5, 10, 16, 10, 8, 22, 10, 5, 4, 14)

# exclude participants
dt_included <- dt[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16]

print(dt_included)
```

## Descriptive stats for included participants
- Number of included participants: `r nrow(dt_included)` (mean age: `r round(mean(dt_included$Age), 2)`, sd: `r round(sd(dt_included$Age), 2)`)
- Male: `r nrow(dt_included[Gender == "Male"])` (mean age: `r round(mean(dt_included[Gender == "Male"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Male"]$Age), 2)`)
- Female: `r nrow(dt_included[Gender == "Female"])` (mean age: `r round(mean(dt_included[Gender == "Female"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Female"]$Age), 2)`)
- Average training years: `r round(mean(dt_included$PianoTotalPractice), 2)` years
- Handedness: Right (`r nrow(dt_included[Handedness == "Right"])`), Left (`r nrow(dt_included[Handedness == "Left"])`), Both (`r nrow(dt_included[Handedness == "Both"])`)

## Some descriptive questions
1. Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?

```{r 1, echo = FALSE}
data.table("Answer" = dt_included$`Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?`)
```

2. Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.

```{r 2, echo = FALSE}
data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
```

### 2020-12-02 script/analysis/preprocessor/error_summary.Rmd
rename folders
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-11-24 script/R/preprocessor/filtered/error_summary.Rmd
@@
 
+# Workflow
+1. filter with `filtering.R`
+2. check the details of errors and remove some participants if needed (**this script**: `error_summary.Rmd`)
+3. trim data (i.e., remove outliers) with `trimming.R`
+4. check trimmed data and remove some participants if needed (`included_summary.Rmd`)
+
 ```{r setup, include = FALSE}
@@
 if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
+if (!require("here")) {install.packages("here"); require("here")}
+
+# set working directory
+here::i_am("error_summary.Rmd")
 
 # function
-source("./function.R")
+source(here("function.R"))
 
 # read a text file for ideal performance
-dt_ideal <- read.table("./ideal.txt")
+dt_ideal <- read.table(here("ideal.txt"))
 colnames(dt_ideal) <- "Pitch"
@@
 # 
-dt_onset <- fread(file = "./dt_correct_onset.txt", sep = ",")
-dt_offset <- fread(file = "./dt_correct_offset.txt", sep = ",")
-error_onset <- fread(file = "./error_onset.txt", sep = ",")
-error_offset <- fread(file = "./error_offset.txt", sep = ",")
+dt_onset <- fread(file = here("filtered", "dt_correct_onset.txt"), sep = ",")
+dt_offset <- fread(file = here("filtered", "dt_correct_offset.txt"), sep = ",")
+error_onset <- fread(file = here("filtered", "error_onset.txt"), sep = ",")
+error_offset <- fread(file = here("filtered", "error_offset.txt"), sep = ",")
 ```
 
+# Design
 - Each performance (1 trial) contains 67 responses for onsets and 67 responses for offsets (i.e., if there is no mistake, 134 responses in total)
 - Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
-- Note: SubNr 8 was already excluded due to an experimental error
+- Note: SubNr 8 was excluded due to an experimental error
 

### 2020-12-02 script/analysis/preprocessor/included_summary.Rmd
rename folders
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-11-24 script/R/preprocessor/trimmed/included_summary.Rmd
@@
 
+# Workflow
+1. filter with `filtering.R`
+2. check the details of errors and remove some participants if needed (`error_summary.Rmd`)
+3. trim data (i.e., remove outliers) with `trimming.R`
+4. check trimmed data and remove some participants if needed (**this script**: `included_summary.Rmd`)
+
 ```{r setup, include = FALSE}
@@
 if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
+if (!require("here")) {install.packages("here"); require("here")}
+
+# set working directory
+here::i_am("included_summary.Rmd")
 
 # read txt files
-dt_ioi <- fread(file = "./data_ioi_1.txt", sep = ",")
-dt_kot <- fread(file = "./data_kot.txt", sep = ",")
-dt_vel <- fread(file = "./data_vel.txt", sep = ",")
-dt_vel_diff <- fread(file = "./data_vel_diff.txt", sep = ",")
+dt_ioi <- fread(file = here("trimmed", "data_ioi_1.txt"), sep = ",")
+dt_kot <- fread(file = here("trimmed", "data_kot.txt"), sep = ",")
+dt_vel <- fread(file = here("trimmed", "data_vel.txt"), sep = ",")
+dt_vel_diff <- fread(file = here("trimmed", "data_vel_diff.txt"), sep = ",")
 ```
 
+# Design
 - Each performance (1 trial) contains 67 responses for onsets and 67 responses for offsets (i.e., if there is no mistake, 134 responses in total)
 - Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
-- Note: SubNr 8 was already excluded due to an experimental error, SubNr 3 & 14 were excluded due to an experimental error
+- Note: SubNr 3, 8, 14 were already excluded due to an experimental error. SubNr 16 was removed due to a deviated tempo
 
@@
 ```{r export, echo = TRUE}
-fwrite(dt_ioi[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./data_analysis_ioi.txt", sep = ",")
-fwrite(dt_kot[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./data_analysis_kot.txt", sep = ",")
-fwrite(dt_vel[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./data_analysis_vel.txt", sep = ",")
-fwrite(dt_vel_diff[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./data_analysis_vel_diff.txt", sep = ",")
+fwrite(dt_ioi[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_ioi.txt"), sep = ",")
+fwrite(dt_kot[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_kot.txt"), sep = ",")
+fwrite(dt_vel[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_vel.txt"), sep = ",")
+fwrite(dt_vel_diff[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_vel_diff.txt"), sep = ",")
 ```

### 2020-12-02 script/analysis/stats/articulation.Rmd
rename folders
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-01 script/R/analysis/articulation.Rmd
@@
 
-```{r kor-art-all-bar, echo = FALSE}
+```{r kor-art-all-box, echo = FALSE}
 kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent)
@@
 ggline(kot_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,66,1))
@@
 ggline(kot_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Dynamics")+
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,66,1)) 

### 2020-12-02 script/R/preprocessor/error_summary.Rmd
add descriptions about excluded participants
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
removed old analysis and update error_summary
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
-# statistics
-if (!require("stats")) {install.packages("stats"); require("stats")}
-if (!require("effsize")) {install.packages("effsize"); require("effsize")}
-if (!require("ez")) {install.packages("ez"); require("ez")}
+if (!require("here")) {install.packages("here"); require("here")}
+
+# set working directory
+here::i_am("error_summary.Rmd")
 
 # function
-source("./function.R")
+source(here("function.R"))
 
 # read a text file for ideal performance
-dt_ideal <- read.table("./ideal.txt")
+dt_ideal <- read.table(here("ideal.txt"))
 colnames(dt_ideal) <- "Pitch"
@@
 # 
-dt_onset <- fread(file = "./filtered/dt_correct_onset.txt", sep = ",")
-dt_offset <- fread(file = "./filtered/dt_correct_offset.txt", sep = ",")
-error_onset <- fread(file = "./filtered/error_onset.txt", sep = ",")
-error_offset <- fread(file = "./filtered/error_offset.txt", sep = ",")
+dt_onset <- fread(file = here("filtered", "dt_correct_onset.txt"), sep = ",")
+dt_offset <- fread(file = here("filtered", "dt_correct_offset.txt"), sep = ",")
+error_onset <- fread(file = here("filtered", "error_onset.txt"), sep = ",")
+error_offset <- fread(file = here("filtered", "error_offset.txt"), sep = ",")
 ```
 
+# Design
 - Each performance (1 trial) contains 72 responses for onsets and 72 responses for offsets (i.e., if there is no mistake, 144 responses in total)
 - Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
+- Note: SubNr 3 was excluded due to insufficient motor skills
 
@@
 ### Corrected Error Onset (All)
-```{r onset_all, echo = FALSE}
-error_onset_all <- error_onset[CorrectionNr != "Exclude",]
+```{r onset-all, echo = FALSE}
+error_onset_all <- error_onset[CorrectionNr != "Exclude"]
 
@@
 ### Corrected Error Onset (Summary)
-```{r onset_summary, echo = FALSE}
+```{r onset-summary, echo = FALSE}
 error_onset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
@@
 ### Missing/Excluded Trials Onset (All)
-```{r na_onset_all, echo = FALSE}
+```{r na-onset-all, echo = FALSE}
 missing_onset <- checker(dt_onset, dt_ideal)
@@
 ### Missing/Excluded Trials Onset (Summary)
-```{r na_onset_summary, echo = FALSE}
+```{r na-onset-summary, echo = FALSE}
 missing_onset_summary <- missing_onset[, .(N = .N), by = .(Reason)]
@@
 ### Corrected Error Offset (All)
-```{r offset_all, echo = FALSE}
+```{r offset-all, echo = FALSE}
 error_offset_all <- error_offset[CorrectionNr != "Exclude",]
@@
 ### Corrected Error Offset (Summary)
-```{r offset_summary, echo = FALSE}
+```{r offset-summary, echo = FALSE}
 error_offset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
@@
 ### Missing/Excluded Trials Offset (All)
-```{r na_offset_all, echo = FALSE}
+```{r na-offset-all, echo = FALSE}
 missing_offset <- checker(dt_offset, dt_ideal)
@@
 ### Missing/Excluded Trials Offset (Summary)
-```{r na_offset_summary, echo = FALSE}
+```{r na-offset-summary, echo = FALSE}
 missing_offset_summary <- missing_offset[, .(N = .N), by = .(Reason)]

### 2020-12-02 script/R/analysis/articulation.R

@@ -1,693 +1,0 @@
-## ----setup, include = FALSE-----------------------------------------------------------------------------------------------
-# packages
-# data manipulation
-if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
-# plot
-if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
-# statistics
-if (!require("stats")) {install.packages("stats"); require("stats")}
-if (!require("effsize")) {install.packages("effsize"); require("effsize")}
-if (!require("ez")) {install.packages("ez"); require("ez")}
-
-# ggplots
-theme_set(theme_classic())
-theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
-
-
-## ----extract, include = FALSE---------------------------------------------------------------------------------------------
-df_kot <- read.csv("../R/preprocessor/trimmed/data_kot.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-df_kor <- read.csv("../R/preprocessor/trimmed/data_kor.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-
-# SubNr as a factor
-df_kot$SubNr <- as.factor(df_kot$SubNr)
-df_kor$SubNr <- as.factor(df_kor$SubNr)
-
-# Include only articulation
-df_kot_art <- df_kot %>% dplyr::filter(Skill == "articulation")
-df_kor_art <- df_kor %>% dplyr::filter(Skill == "articulation")
-
-# Include only dynamics
-df_kot_dyn <- df_kot %>% dplyr::filter(Skill == "dynamics")
-df_kor_dyn <- df_kor %>% dplyr::filter(Skill == "dynamics")
-
-
-## ----kot_art, echo = FALSE------------------------------------------------------------------------------------------------
-# For each individual
-kot_art <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
-                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
-kot_art <- cbind(kot_art[,1:4], as.data.frame(kot_art[,5]))
[…]

### 2020-12-02 script/R/analysis/export.R

@@ -1,6 +1,0 @@
-# export R scripts from Rmd
-knitr::purl("../../RNotebook/ioi_1.Rmd")
-knitr::purl("../../RNotebook/ioi_3.Rmd")
-knitr::purl("../../RNotebook/articulation.Rmd")
-knitr::purl("../../RNotebook/dynamics.Rmd")
-knitr::purl("../../RNotebook/questionnaire.Rmd")

### 2020-12-02 script/R/analysis/ioi_1.R

@@ -1,777 +1,0 @@
-## ----setup, include = FALSE-----------------------------------------------------------------------------------------------
-# packages
-# data manipulation
-if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
-# plot
-if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
-# statistics
-if (!require("stats")) {install.packages("stats"); require("stats")}
-if (!require("effsize")) {install.packages("effsize"); require("effsize")}
-if (!require("ez")) {install.packages("ez"); require("ez")}
-
-# ggplots
-theme_set(theme_classic())
-theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
-
-
-## ----extract, include = FALSE---------------------------------------------------------------------------------------------
-df_ioi <- read.csv("../R/preprocessor/trimmed/data_ioi_1.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-
-# SubNr as a factor
-df_ioi$SubNr <- as.factor(df_ioi$SubNr)
-
-# Include only articulation
-df_ioi_art <- df_ioi %>% dplyr::filter(Skill == "articulation")
-# Include only dynamics
-df_ioi_dyn <- df_ioi %>% dplyr::filter(Skill == "dynamics")
-
-
-## ----ioi_art, echo = FALSE------------------------------------------------------------------------------------------------
-# For each individual
-ioi_art <- aggregate(normIOI~SubNr*Condition*Skill, data = df_ioi_art,
-                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
-ioi_art <- cbind(ioi_art[,1:3], as.data.frame(ioi_art[,4]))
-# Change colnames
-colnames(ioi_art) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
-print(ioi_art)
-
-
-## ----ioi_art_bar,  echo = FALSE-------------------------------------------------------------------------------------------
-p_ioi_art <- ggplot(data = ioi_art, aes(x = SubNr, y = Mean, fill = Condition)) +
[…]

### 2020-12-02 script/R/analysis/ioi_3.R

@@ -1,768 +1,0 @@
-## ----setup, include = FALSE-----------------------------------------------------------------------------------------------
-# packages
-# data manipulation
-if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
-# plot
-if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
-# statistics
-if (!require("stats")) {install.packages("stats"); require("stats")}
-if (!require("effsize")) {install.packages("effsize"); require("effsize")}
-if (!require("ez")) {install.packages("ez"); require("ez")}
-
-# ggplots
-theme_set(theme_classic())
-theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
-
-
-## ----extract, include = FALSE---------------------------------------------------------------------------------------------
-df_ioi <- read.csv("../R/preprocessor/trimmed/data_ioi_3.csv", header = T, sep = ",", dec = ".") # read a trimmed csv
-
-# SubNr as a factor
-df_ioi$SubNr <- as.factor(df_ioi$SubNr)
-
-# Include only articulation
-df_ioi_art <- df_ioi %>% dplyr::filter(Skill == "articulation")
-# Include only dynamics
-df_ioi_dyn <- df_ioi %>% dplyr::filter(Skill == "dynamics")
-
-
-## ----ioi_art, echo = FALSE------------------------------------------------------------------------------------------------
-# For each individual
-ioi_art <- aggregate(normIOI~SubNr*Condition*Skill, data = df_ioi_art,
-                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
-ioi_art <- cbind(ioi_art[,1:3], as.data.frame(ioi_art[,4]))
-# Change colnames
-colnames(ioi_art) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
-print(ioi_art)
-
-
-## ----ioi_art_bar,  echo = FALSE-------------------------------------------------------------------------------------------
-p_ioi_art <- ggplot(data = ioi_art, aes(x = SubNr, y = Mean, fill = Condition)) +
[…]

### 2020-12-02 script/R/analysis/questionnaire.R

@@ -1,45 +1,0 @@
-## ----setup, include = FALSE-----------------------------------------------------------------------------------------------
-# packages
-# data manipulation
-if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
-
-
-## ----read, echo = FALSE---------------------------------------------------------------------------------------------------
-# Read csv files
-df <- read.csv("./questionnaire_teaching-v2.0.csv", header = T, sep = ",", dec = ".", na.string = "NA")
-df_error_onset <- read.csv("../R/preprocessor/filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".")
-df_error_offset <- read.csv("../R/preprocessor/filtered/data_error_correction_offset.csv", header = T, sep = ",", dec = ".")
-df$Please.specify.the.total.number.of.years.of.formal.training..i.e...lessons.. <- as.numeric(df$Please.specify.the.total.number.of.years.of.formal.training..i.e...lessons..)
-
-# add SubNr 21 manually
-df[21,2] <- 21
-df[21,3] <- 25
-df[21,4] <- "Male"
-df[21,5] <- "Hungarian"
-df[21,6] <- "English"
-df[21,7] <- "Right"
-df[21,8] <- "Yes"
-df[21,9] <- "Yes"
-df[21,17] <- 11
-df[21,18] <- 12
-df[21,19] <- 13
-
-# exclude SubNr 3 due to insufficient motor skills
-include <- df %>% dplyr::filter(Participant.number != 3)
-
-# colnames
-colnames(include) <- c("TimeStamps", "SubNr", "Age", "Gender", "FirstLang", "SecondLang", "DominantHand",
-                       "Vision", "Hearing", "KnowClementi", "PlayClementi", colnames(include)[12],
-                       colnames(include)[13], "School", "Degree", "Major", "TotalTraining", "TotalPlaying",
-                       "AgeStarted", "AgeStopped", "PracticeHours", "Instruments", colnames(include)[23],
-                       colnames(include)[24], colnames(include)[25], colnames(include)[26],
-                       colnames(include)[27], colnames(include)[28], colnames(include)[29],
-                       colnames(include)[30], colnames(include)[31], colnames(include)[32],
-                       colnames(include)[33], colnames(include)[34], colnames(include)[35],
-                       "MusicTeachingExperience", "OtherTeachingExperience", "Subject", "TotalTeaching",
-                       "TeachingLiking")
-
-# correct some gibbrish
-include[13,17] <- 18
-print(include)
-

### 2020-12-02 analysis/preprocessor/error_summary.Rmd
update readme and error_summary
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 script/analysis/preprocessor/error_summary.Rmd
@@
 2. check the details of errors and remove some participants if needed (**this script**: `error_summary.Rmd`)
-3. trim data (i.e., remove outliers) with `trimming.R`
+3. trim data (i.e., remove outliers) with `trimming_ioi.R`, `trimming_kot.R`, `trimming_vel.R`
 4. check trimmed data and remove some participants if needed (`included_summary.Rmd`)
@@
 ```{r onset-all, echo = FALSE}
-error_onset_all <- error_onset[CorrectionNr != "Exclude",]
+error_onset_all <- error_onset[CorrectionNr != "Exclude"]
 error_onset_all

### 2020-12-02 analysis/stats/ioi.Rmd
add commas
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 script/R/analysis/ioi.Rmd
@@
 #### Boxplot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
 
@@
 #### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
 
@@
 #### Line graph
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
 
@@
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
 

### 2020-12-02 analysis/stats/articulation.Rmd
upload articulation analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 script/analysis/stats/articulation.Rmd
@@
 
+#### Box plot
 ```{r kor-art-all-box, echo = FALSE}
@@
   theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  scale_x_continuous(breaks = seq(1,66,1))
+  scale_x_continuous(breaks = seq(1,8,1))
 ```

### 2020-12-02 script/R/stats/articulation.Rmd
upload velocity analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 analysis/stats/articulation.Rmd
@@
 ---
-title: 'Articulation Analysis: The Sound of Teaching Music 1'
+title: 'Articulation Analysis: The Sound of Teaching Music 2'
 output:
@@
 - Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: legato vs. staccato / forte vs. piano)
-- Dependent variables: Key-Overlap Time (KOT), Key-Overlap Ratio (KOR)
+- Dependent variables: Key-Overlap Ratio (KOR), Key-Overlap Time (KOT)
 
@@
 2. Excluded participants
-
-- SubNr 1: Most of responses (more than 90%) were removed due to outliers
-- SubNr 3: Experimental error
-- SubNr 8: Experimental error
-- SubNr 14: Experimental error
-- SubNr 16: A deviated tempo
+- SubNr 3: Insufficient motor skills
 
@@
 
-# Outline - Articulation (KOT, KOR)
-## [Articulation](#kot-articulation)
+# Outline - Articulation (KOR, KOT)
+## [Articulation](#kor-articulation)
+
+### KOR
+- Individual results
+  + [All](#kor_art_ind_all)
+  + [Trial by trial](#kor_art_ind_trial)
+
+- Group results and stats (if available)
+  + [All](#kor_art_group_all)
+  + [Trial by trial](#kor_art_group_trial)
 
@@
 
+## [Dynamics](#kor-dynamics)
+
 ### KOR
 - Individual results
-  + [All](#kor_art_ind_all)
-  + [Trial by trial](#kor_art_ind_trial)
+  + [All](#kor_dyn_ind_all)
+  + [Trial by trial](#kor_dyn_ind_trial)
 
 - Group results and stats (if available)
-  + [All](#kor_art_group_all)
-  + [Trial by trial](#kor_art_group_trial)
-
-## [Dynamics](#kot-dynamics)
+  + [All](#kor_dyn_group_all)
+  + [Trial by trial](#kor_dyn_group_trial)
 
@@
 - [Sequence plot](#sequence_dyn)
-
-### KOR
-- Individual results
-  + [All](#kor_dyn_ind_all)
-  + [Trial by trial](#kor_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kor_dyn_group_all)
-  + [Trial by trial](#kor_dyn_group_trial)
 
@@
 ```{r file, include = FALSE}
-filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
+filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
@@
 ```{r extract, include = FALSE}
-dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
+dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
@@
 # SubNr as a factor
-dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
+dt_kor$SubNr <- as.factor(dt_kor$SubNr)
 dt_kot$SubNr <- as.factor(dt_kot$SubNr)
 
-# For each individual, each block, each trial
-ioi <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]
-
-# Calculate KOR for each interval (outliers were already determined based on KOT)
-dt_kor <- data.table()
-for (subnr in unique(dt_kot$SubNr)){
-  for (block in unique(dt_kot[SubNr == subnr]$BlockNr)){
-    for (trial in unique(dt_kot[SubNr == subnr & BlockNr == block]$TrialNr)){
-      dt_current <- dt_kot[SubNr == subnr & BlockNr == block & TrialNr == trial]
-      dt_current$KOR <- dt_current$KOT/ioi[SubNr == subnr & ioi$BlockNr == block & TrialNr == trial]$Mean
-      dt_kor <- rbind(dt_kor, dt_current)
-    }
-  }
-}
-
 # Include only articulation
+dt_kor_art <- dt_kor[Skill == "articulation"]
 dt_kot_art <- dt_kot[Skill == "articulation"]
-dt_kor_art <- dt_kor[Skill == "articulation"]
 
 # Include only dynamics
+dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
-dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 ```
 
-# Key-Overlap Time (KOT) - Articulation{#kot-articulation}
+# Key-Overlap Ratio (KOR) - Articulation{#kor-articulation}
+## 1. All - individual{#kor_art_ind_all}
+- Average KOR for each participant.
+
+```{r kor-art, echo = FALSE}
+# For each individual
+kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
+kor_art
+```
+
+### Plot
+#### Box plot
+
+```{r kor-art-box,  echo = FALSE}
+ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
+  facet_grid(Subcomponent ~ .) +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+```
+
+## 2. All - group{#kor_art_group_all}
+- Average KOR for each condition.
+- Hypothesis: Participants will exaggerate legato and staccato when teaching.
+
+```{r kor-art-all, echo = FALSE}
+# Group mean
+kor_art_all <- kor_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
+kor_art_all
+```
+
+### Plot
+#### Scatter plot
+- Error bars show SEM of the mean
+
+```{r kor-art-all-scatter, echo = FALSE}
+ggerrorplot(kor_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+```
+
+#### Box plot
+```{r kor-art-all-box, echo = FALSE}
+kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent)
+
+ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+```
+
+### Statistics
+#### Two-way ANOVA (car_aov)
+```{r anova-art-kor-1, echo = FALSE}
+kor_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
+summary(kor_art_aov)
+```
+
+#### Two-way ANOVA (aov)
+```{r anova-art-kor-2, echo = FALSE}
+kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
+summary(kor_art_aov_2)    
+```
+
+#### Post-hoc (emmeans)
+```{r post-hoc-2, echo = FALSE}
+test(pairs(emmeans(kor_art_aov,~Condition|Subcomponent)), joint = TRUE)
+pairs(emmeans(kor_art_aov,~Condition|Subcomponent), adjust = "tukey")
+```
+
+## 3. Trial by trial - individual{#kor_art_ind_trial}
+- Average KOR for each trial for each participant.
+
+```{r kor-art-trial, echo = FALSE}
+# For each individual
+kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
+kor_art_trial
+```
+
+### Plot
+#### Line graph
+```{r kor-art-trial-line-legato, echo = FALSE, fig.height = 4}
+ggline(dt_kor_art[Subcomponent == "Legato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Legato") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
+```
+
+```{r kor-art-trial-line-staccato, echo = FALSE, fig.height = 4}
+ggline(dt_kor_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Staccato") +
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
+```
+
+## 4. Trial by trial - group{#kor_art_group_trial}
+- Average KOR d for each trial for each condition.
+
+```{r kor-art-trial-all, echo = FALSE}
+# Group mean
+kor_art_trial_all <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
+kor_art_trial_all
+```
+
+### Plot
+#### Line graph
+- Error bars show SEM of the mean
+
+```{r kor-art-trial-all-line-legato, echo = FALSE}
+ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
+```
+
+```{r kor-art-trial-all-line-staccato, echo = FALSE}
+ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
+  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  scale_x_continuous(breaks = seq(1,8,1))
+```
+
+# Key-Overlap Time (KOT) - Articulation
 ## 1. All - individual{#kot_art_ind_all}
@@
 
-# Key-Overlap Ratio (KOR) - Articulation
-## 1. All - individual{#kor_art_ind_all}
+# Key-Overlap Ratio (KOR) - Dynamics{#kor-dynamics}
+## 1. All - individual{#kor_dyn_ind_all}
 - Average KOR for each participant.
 
-```{r kor-art, echo = FALSE}
+```{r kor-dyn, echo = FALSE}
 # For each individual
-kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
-kor_art
+kor_dyn <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
+kor_dyn
 ```
@@
 #### Box plot
-
-```{r kor-art-box,  echo = FALSE}
-ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
+```{r kor-dyn-box,  echo = FALSE}
+ggboxplot(dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Dynamics") +
   facet_grid(Subcomponent ~ .) +
@@
 
-## 2. All - group{#kor_art_group_all}
+## 2. All - group{#kor_dyn_group_all}
 - Average KOR for each condition.
-- Hypothesis: Participants will exaggerate legato and staccato when teaching.
 
-```{r kor-art-all, echo = FALSE}
+```{r kor-dyn-all, echo = FALSE}
 # Group mean
-kor_art_all <- kor_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
-kor_art_all
+kor_dyn_all <- kor_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
+kor_dyn_all
 ```
@@
 
-```{r kor-art-all-scatter, echo = FALSE}
-ggerrorplot(kor_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
+```{r kor-dyn-all-scatter, echo = FALSE}
+ggerrorplot(kor_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
@@
 #### Box plot
-```{r kor-art-all-box, echo = FALSE}
-kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent)
+```{r kor-dyn-all-box, echo = FALSE}
+kor_dyn_plot <- data.table(subject = kor_dyn[Condition == "teaching"]$SubNr, teaching = kor_dyn[Condition == "teaching"]$Mean, performing = kor_dyn[Condition == "performing"]$Mean, subcomponent = kor_dyn[Condition == "teaching"]$Subcomponent)
 
-ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
+ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
@@
 #### Two-way ANOVA (car_aov)
-```{r anova-art-kor-1, echo = FALSE}
-kor_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
-summary(kor_art_aov)
+```{r anova-dyn-kor-1, echo = FALSE}
+kor_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
+summary(kor_dyn_aov)
 ```
@@
 #### Two-way ANOVA (aov)
-```{r anova-art-kor-2, echo = FALSE}
-kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
-summary(kor_art_aov_2)    
+```{r anova-dyn-kor-2, echo = FALSE}
+kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
+summary(kor_dyn_aov_2)
 ```
 
-#### Post-hoc (emmeans)
-```{r post-hoc-2, echo = FALSE}
-test(pairs(emmeans(kor_art_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(kor_art_aov,~Condition|Subcomponent), adjust = "tukey")
-```
-
-## 3. Trial by trial - individual{#kor_art_ind_trial}
+## 3. Trial by trial - individual{#kor_dyn_ind_trial}
 - Average KOR for each trial for each participant.
 
-```{r kor-art-trial, echo = FALSE}
+```{r kor-dyn-trial, echo = FALSE}
 # For each individual
-kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
-kor_art_trial
+kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
+kor_dyn_trial
 ```
@@
 #### Line graph
-```{r kor-art-trial-line-legato, echo = FALSE, fig.height = 4}
-ggline(dt_kor_art[Subcomponent == "Legato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Legato") +
+```{r kor-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
+ggline(dt_kor_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Forte") +
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
@@
 
-```{r kor-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(dt_kor_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Staccato") +
+```{r kor-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
+ggline(dt_kor_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Piano") +
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
@@
 
-## 4. Trial by trial - group{#kor_art_group_trial}
-- Average KOR d for each trial for each condition.
+## 4. Trial by trial - group{#kor_dyn_group_trial}
+- Average KOR for each trial for each condition.
 
-```{r kor-art-trial-all, echo = FALSE}
+```{r kor-dyn-trial-all, echo = FALSE}
 # Group mean
-kor_art_trial_all <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
-kor_art_trial_all
+kor_dyn_trial_all <- kor_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent,TrialNr)]
+kor_dyn_trial_all
 ```
@@
 
-```{r kor-art-trial-all-line-legato, echo = FALSE}
-ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
+```{r kor-dyn-trial-all-line-forte, echo = FALSE}
+ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Forte") +
   theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
@@
 
-```{r kor-art-trial-all-line-staccato, echo = FALSE}
-ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
+```{r kor_dyn_trial_all_line_piano, echo = FALSE}
+ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") +
   theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
@@
 
-# Key-Overlap Time (KOT) - Dynamics{#kot-dynamics}
+# Key-Overlap Time (KOT) - Dynamics
 ## 1. All - individual{#kot_dyn_ind_all}
@@
 
-# Key-Overlap Ratio (KOR) - Dynamics
-## 1. All - individual{#kor_dyn_ind_all}
-- Average KOR for each participant.
-
-```{r kor-dyn, echo = FALSE}
-# For each individual
-kor_dyn <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
-kor_dyn
-```
-
-### Plot
-#### Box plot
-```{r kor-dyn-box,  echo = FALSE}
-ggboxplot(dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Dynamics") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
-```
-
-## 2. All - group{#kor_dyn_group_all}
-- Average KOR for each condition.
-
-```{r kor-dyn-all, echo = FALSE}
-# Group mean
-kor_dyn_all <- kor_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
-kor_dyn_all
-```
-
-### Plot
-#### Scatter plot
-- Error bars show SEM of the mean
-
-```{r kor-dyn-all-scatter, echo = FALSE}
-ggerrorplot(kor_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
-```
-
-#### Box plot
-```{r kor-dyn-all-box, echo = FALSE}
-kor_dyn_plot <- data.table(subject = kor_dyn[Condition == "teaching"]$SubNr, teaching = kor_dyn[Condition == "teaching"]$Mean, performing = kor_dyn[Condition == "performing"]$Mean, subcomponent = kor_dyn[Condition == "teaching"]$Subcomponent)
-
-ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
-```
-
-### Statistics
-#### Two-way ANOVA (car_aov)
-```{r anova-dyn-kor-1, echo = FALSE}
-kor_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
-summary(kor_dyn_aov)
-```
-
-#### Two-way ANOVA (aov)
-```{r anova-dyn-kor-2, echo = FALSE}
-kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
-summary(kor_dyn_aov_2)
-```
-
-## 3. Trial by trial - individual{#kor_dyn_ind_trial}
-- Average KOR for each trial for each participant.
-
-```{r kor-dyn-trial, echo = FALSE}
-# For each individual
-kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
-kor_dyn_trial
-```
-
-### Plot
-#### Line graph
-```{r kor-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-ggline(dt_kor_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Forte") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  scale_x_continuous(breaks = seq(1,8,1))
-```
-
-```{r kor-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-ggline(dt_kor_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Piano") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  scale_x_continuous(breaks = seq(1,8,1))
-```
-
-## 4. Trial by trial - group{#kor_dyn_group_trial}
-- Average KOR for each trial for each condition.
-
-```{r kor-dyn-trial-all, echo = FALSE}
-# Group mean
-kor_dyn_trial_all <- kor_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent,TrialNr)]
-kor_dyn_trial_all
-```
-
-### Plot
-#### Line graph
-- Error bars show SEM of the mean
-
-```{r kor-dyn-trial-all-line-forte, echo = FALSE}
-ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Forte") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  scale_x_continuous(breaks = seq(1,8,1))
-```
-
-```{r kor_dyn_trial_all_line_piano, echo = FALSE}
-ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  scale_x_continuous(breaks = seq(1,8,1))
-```
-
 # Sequence plot
@@
 # For each individual
-kot_art_seq <- dt_kot_art[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Interval)]
+kor_art_seq <- dt_kor_art[, .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Interval)]
 
-ggline(kot_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Articulation") +
+ggline(kor_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOR", title = "KOT: Articulation") +
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  scale_x_continuous(breaks = seq(1,66,1))
+  scale_x_continuous(breaks = seq(1,71,1))
 ```
@@
 # For each individual
-kot_dyn_seq <- dt_kot_dyn[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Interval)]
+kor_dyn_seq <- dt_kor_dyn[, .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Interval)]
 
-ggline(kot_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Dynamics")+
+ggline(kor_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOR", title = "KOT: Dynamics")+
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  scale_x_continuous(breaks = seq(1,66,1)) 
+  scale_x_continuous(breaks = seq(1,71,1)) 
 ```

### 2020-12-02 analysis/preprocessor/included_summary.Rmd
rename folders
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Included  Summary (teaching-v2.0)"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Workflow
1. filter with `filtering.R`
2. check the details of errors and remove some participants if needed (`error_summary.Rmd`)
3. trim data (i.e., remove outliers) with `trimming_ioi.R`, `trimming_kot_kor.R`, `trimming_vel.R`
4. check trimmed data and remove some participants if needed (**this script**: `included_summary.Rmd`)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("here")) {install.packages("here"); require("here")}

# set working directory
here::i_am("included_summary.Rmd")

# read txt files
dt_ioi <- fread(file = here("trimmed", "data_ioi_1.txt"), sep = ",")
dt_kor <- fread(file = here("trimmed", "data_kor.txt"), sep = ",")
dt_kot <- fread(file = here("trimmed", "data_kot.txt"), sep = ",")
dt_vel <- fread(file = here("trimmed", "data_vel.txt"), sep = ",")
dt_vel_diff <- fread(file = here("trimmed", "data_vel_diff.txt"), sep = ",")
```

# Design
- Each performance (1 trial) contains 72 responses for onsets and 72 responses for offsets (i.e., if there is no mistake, 144 responses in total)
- Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
- Note: SubNr 3 was excluded due to insufficient motor skills

### IOIs (All)
```{r ioi, echo = FALSE}
dt_ioi
```

### IOIs (Summary)
```{r ioi-summary, echo = FALSE}
dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
```

### KOR (All)
```{r kor, echo = FALSE}
dt_kor
```

### KOR (Summary)
```{r kor-summary, echo = FALSE}
dt_kor[, .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

### KOT (All)
```{r kot, echo = FALSE}
dt_kot
```

### KOT (Summary)
```{r kot-summary, echo = FALSE}
dt_kot[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

### Key Velocity (All)
```{r vel, echo = FALSE}
dt_vel
```

### Key Velocity (Summary)
```{r vel-summary, echo = FALSE}
dt_vel[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

### KV Difference (All)
```{r vel-diff, echo = FALSE}
dt_vel_diff
```

### KV Difference (Summary)
```{r vel-diff-summary, echo = FALSE}
dt_vel_diff[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

## Excluded participnats
- SubNr 3: Insufficient motor skills

### Export only included participants
data_analysis~.txt are used for data analysis.
```{r export, echo = TRUE}
fwrite(dt_ioi[SubNr != 3], file = here("trimmed", "data_analysis_ioi.txt"), sep = ",")
fwrite(dt_kor[SubNr != 3], file = here("trimmed", "data_analysis_kor.txt"), sep = ",")
fwrite(dt_kot[SubNr != 3], file = here("trimmed", "data_analysis_kot.txt"), sep = ",")
fwrite(dt_vel[SubNr != 3], file = here("trimmed", "data_analysis_vel.txt"), sep = ",")
fwrite(dt_vel_diff[SubNr != 3], file = here("trimmed", "data_analysis_vel_diff.txt"), sep = ",")
```

### 2020-12-02 analysis/stats/dynamics.Rmd
rename folders
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Dynamics Analysis: The Sound of Teaching Music 2'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of dynamics analysis.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato)
- Dependent variables: Key Velocity (KV), KV difference

1. Main predictions
- Participants will exaggerate forte and piano more while teaching.

2. Excluded participants
- SubNr 3: Insufficient motor skills

3. Stats packages
- [afex](https://github.com/singmann/afex) for ANOVAs
- [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons

# Outline - Dynamics (Key Velocity)

## [Dynamics](#vel-dynamics)

### KV
- Individual results
  + [All](#vel_dyn_ind_all)
  + [Trial by trial](#vel_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_dyn_group_all)
  + [Trial by trial](#vel_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#vel_dyn_order_group_all)

### KV-Diff
- Individual results
  + [All](#vel_diff_dyn_ind_all)
  + [Trial by trial](#vel_diff_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_diff_dyn_group_all)
  + [Trial by trial](#vel_diff_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#vel_diff_dyn_order_group_all)

- [Sequence plot](#sequence_dyn)

## [Articulation](#vel_articulation)

### KV
- Individual results
  + [All](#vel_art_ind_all)
  + [Trial by trial](#vel_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_art_group_all)
  + [Trial by trial](#vel_art_group_trial)

### KV-Diff
- Individual results
  + [All](#vel_diff_art_ind_all)
  + [Trial by trial](#vel_diff_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_diff_art_group_all)
  + [Trial by trial](#vel_diff_art_group_trial)

- [Sequence plot](#sequence_art)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# statistics
if (!require("afex")) {install.packages("afex"); require("afex")}
if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
```

```{r file, include = FALSE}
filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
```

```{r extract, include = FALSE}
dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_vel$SubNr <- as.factor(dt_vel$SubNr)
dt_vel_diff$SubNr <- as.factor(dt_vel_diff$SubNr)

# Include only articulation
dt_vel_art <- dt_vel[Skill == "articulation"]
dt_vel_diff_art <- dt_vel_diff[Skill == "articulation"]

# Include only dynamics
dt_vel_dyn <- dt_vel[Skill == "dynamics"]
dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
```

# Key Velocity (KV) - Dynamics
## 1. All - individual{#vel_dyn_ind_all}
- Average KV for each participant.

```{r vel-dyn, echo = FALSE}
# For each individual
vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_dyn
```

### Plot
#### Box plot
```{r vel-dyn-box,  echo = FALSE}
ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
  facet_grid(Subcomponent ~ .) +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
```

## 2. All - group{#vel_dyn_group_all}
- Average KV for each condition.
- Hypothesis: Participants will exaggerate forte and piano when teaching.

```{r vel-dyn-all, echo = FALSE}
# Group mean
vel_dyn_all <- vel_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_dyn_all
```

### Plot
#### Scatter plot
- Error bars show SEM of the mean

```{r vel-dyn-all-scatter, echo = FALSE}
ggerrorplot(vel_dyn, x = "Condition", y = "Mean", add = c("mean", "mean_se"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
```

#### Box plot
```{r vel-dyn-all-box, echo = FALSE}
vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-dyn-vel-1, echo = FALSE}
vel_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
summary(vel_dyn_aov)
```

#### Two-way ANOVA (aov)
```{r anova-dyn-vel-2, echo = FALSE}
vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
summary(vel_dyn_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-1, echo = FALSE}
test(pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#vel_dyn_ind_trial}
- Average KV for each trial for each participant.

```{r vel-dyn-trial, echo = FALSE}
# For each individual
vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_dyn_trial
```

### Plot
#### Line graph
```{r vel-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
ggline(dt_vel_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
ggline(dt_vel_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

## 4. Trial by trial - group{#vel_dyn_group_trial}
- Average KV for each trial for each condition.

```{r vel-dyn-trial-all, echo = FALSE}
# Group mean
vel_dyn_trial_all <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel-dyn-trial-all-line-forte, echo = FALSE}
ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
    scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-dyn-trial-all-line-piano, echo = FALSE}
ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
    scale_x_continuous(breaks = seq(1,8,1))
```

# Key Velocity Difference (Difference) - Dynamics
## 1. All - individual{#vel_diff_dyn_ind_all}
- Average KV-Diff for each participant.

```{r vel-diff-dyn, echo = FALSE}
# For each individual
vel_diff_dyn <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_diff_dyn
```

### Plot
#### Box plot
```{r vel-diff-dyn-box,  echo = FALSE}
ggboxplot(dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Dynamics") +
  facet_grid(Subcomponent ~ .) +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
```

## 2. All - group{#vel_diff_dyn_group_all}
- Average KV-Diff for each condition.

```{r vel-diff-dyn-all, echo = FALSE}
# Group mean
vel_diff_dyn_all <- vel_diff_dyn[, .(N = .N, Mean = mean(Mean), SD = mean(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_diff_dyn_all
```

### Plot
#### Scatter plot
- Error bars show SEM of the mean

```{r vel-diff-dyn-all-scatter, echo = FALSE}
ggerrorplot(vel_diff_dyn, x = "Condition", y = "Mean", add = c("mean", "mean_se"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics") +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
```

#### Box plot
```{r vel-diff-dyn-all-box, echo = FALSE}
vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics") +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed", )
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-dyn-vel-diff-1, echo = FALSE}
vel_diff_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
summary(vel_diff_dyn_aov)
```

#### Two-way ANOVA (aov)
```{r anova-dyn-vel-diff-2, echo = FALSE}
vel_diff_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
summary(vel_diff_dyn_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-2, echo = FALSE}
test(pairs(emmeans(vel_diff_dyn_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(vel_diff_dyn_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#vel_diff_dyn_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel-diff-dyn-trial, echo = FALSE}
# For each individual
vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = mean(Diff)), by = .(SubNr, Condition, Skill, Subcomponent,TrialNr)]
vel_diff_dyn_trial
```

### Plot
#### Line graph
```{r vel-diff-dyn-trial-line-ftop, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_dyn[Subcomponent == "FtoP", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-diff-dyn-trial-line-ptof, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_dyn[Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

## 4. Trial by trial - group{#vel_diff_dyn_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel-diff-dyn-trial-all, echo = FALSE}
# Group mean
vel_diff_dyn_trial_all <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = mean(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_diff_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel-diff-dyn-trial-all-line-ftop, echo = FALSE}
ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-diff-dyn-trial-all-line-ptof, echo = FALSE}
ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

# Key Velocity (KV) - Articulation
## 1. All - individual{#vel_art_ind_all}
- Average KV for each participant.

```{r vel-art, echo = FALSE}
# For each individual
vel_art <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_art
```

### Plot
#### Box plot
```{r vel-art-box,  echo = FALSE}
ggboxplot(dt_vel_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") +
  facet_grid(Subcomponent ~ .) +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
```

## 2. All - group{#vel_art_group_all}
- Average KV for each condition.

```{r vel-art-all, echo = FALSE}
# Group mean
vel_art_all <- vel_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_art_all
```

### Plot
#### Scatter plot
- Error bars show SEM of the mean

```{r vel-art-all-scatter, echo = FALSE}
ggerrorplot(vel_art, x = "Condition", y = "Mean", add = c("mean", "mean_se"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation") +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
```

#### Box plot
```{r vel-art-all-box, echo = FALSE}
vel_art_plot <- data.table(subject = vel_art[Condition == "teaching"]$SubNr, teaching = vel_art[Condition == "teaching"]$Mean, performing = vel_art[Condition == "performing"]$Mean, subcomponent = vel_art[Condition == "teaching"]$Subcomponent)

ggpaired(vel_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation") +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-art-vel-1, echo = FALSE}
vel_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
summary(vel_art_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-vel-2, echo = FALSE}
vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
summary(vel_art_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-3, echo = FALSE}
test(pairs(emmeans(vel_art_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(vel_art_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#vel_art_ind_trial}
- Average KV for each trial for each participant.

```{r vel-art-trial, echo = FALSE}
# For each individual
vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_art_trial
```

### Plot
#### Line graph
```{r vel-art-trial-line-legato, echo = FALSE, fig.height = 4}
ggline(dt_vel_art[Subcomponent == "Legato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-art-trial-line-staccato, echo = FALSE, fig.height = 4}
ggline(dt_vel_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

## 4. Trial by trial - group{#vel_art_group_trial}
- Average KV for each trial for each condition.

```{r vel-art-trial-all, echo = FALSE}
# Group mean
vel_art_trial_all <- vel_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel-art-trial-all-line-legato, echo = FALSE}
ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-art-trial-all-line-staccato, echo = FALSE}
ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

# Key Velocity Difference (Difference) - Articulation{#vel_articulation}
## 1. All - individual{#vel_diff_art_ind_all}
- Average KV-Diff for each participant.

```{r vel-diff-art, echo = FALSE}
# For each individual
vel_diff_art <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_diff_art
```

### Plot
#### Box plot
```{r vel-diff-art-box,  echo = FALSE}
ggboxplot(dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
  facet_grid(Subcomponent ~ .) +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
```

## 2. All - group{#vel_diff_art_group_all}
- Average KV-Diff for each condition.

```{r vel-diff-art-all, echo = FALSE}
# Group mean
vel_diff_art_all <- vel_diff_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_diff_art_all
```

### Plot
#### Scatter plot
- Error bars show SEM of the mean

```{r vel-diff_art-all-scatter, echo = FALSE}
ggerrorplot(vel_diff_art, x = "Condition", y = "Mean", add = c("mean", "mean_se"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Articulation") +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
```

#### Box plot
```{r vel-diff-art-all-box, echo = FALSE}
vel_diff_art_plot <- data.table(subject = vel_diff_art[Condition == "teaching"]$SubNr, teaching = vel_diff_art[Condition == "teaching"]$Mean, performing = vel_diff_art[Condition == "performing"]$Mean, subcomponent = vel_diff_art[Condition == "teaching"]$Subcomponent)

ggpaired(vel_diff_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV-Diff: Articulation") +
  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-art-vel-diff-1, echo = FALSE}
vel_diff_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_art)
summary(vel_diff_art_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-vel-diff-2, echo = FALSE}
vel_diff_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_art)
summary(vel_diff_art_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-4, echo = FALSE}
test(pairs(emmeans(vel_diff_art_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(vel_diff_art_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#vel_diff_art_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel-diff-art-trial, echo = FALSE}
# For each individual
vel_diff_art_trial <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_diff_art_trial
```

### Plot
#### Line graph
```{r vel-diff-artt-rial-line-ltos, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_art[Subcomponent == "LtoS", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-diff-art-trial-line-stol, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_art[Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,8,1))
```

## 4. Trial by trial - group{#vel_diff_art_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel-diff-art-trial-all, echo = FALSE}
# Group mean
vel_diff_art_trial_all <- vel_diff_art_trial[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_diff_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel-diff-art-trial-all-line-ltos, echo = FALSE}
ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks=seq(1,8,1))
```

```{r vel-diff-art-trial-all-line-stol, echo = FALSE}
ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks=seq(1,8,1))
```

## Dynamics{#sequence_dyn}
```{r seq-dyn, fig.width = 9, fig.height = 3, echo = FALSE}
# For each individual
vel_dyn_seq <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, RowNr)]

ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Note", ylab = "", title = "KV: Dynamics") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,72,1))

vel_diff_dyn_seq <- dt_vel_diff_dyn[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Interval)]

ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,71,1))
```

## Articulation{#sequence_art}
```{r seq-art, fig.width = 9, fig.height = 3, echo = FALSE}
# For each individual
vel_art_seq <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, RowNr)]

ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Note", ylab = "", title = "KV: Articulation") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,72,1))

vel_diff_art_seq <- dt_vel_diff_art[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Interval)]

ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") +
  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
  scale_x_continuous(breaks = seq(1,71,1))
```

### 2020-12-03 analysis/stats/ioi.Rmd

@@
 #### Two-way ANOVA (aov)
-```{r anova-art-boundary2, echo = FALSE}
+```{r anova-dyn-boundary2, echo = FALSE}
 ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)

### 2020-12-03 experiment-1/analysis/preprocessor/included_summary.Rmd
move stuff to experiment-1
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 script/analysis/preprocessor/included_summary.Rmd
@@
 2. check the details of errors and remove some participants if needed (`error_summary.Rmd`)
-3. trim data (i.e., remove outliers) with `trimming.R`
+3. trim data (i.e., remove outliers) with `trimming_ioi.R`, `trimming_kot.R`, `trimming_vel.R`
 4. check trimmed data and remove some participants if needed (**this script**: `included_summary.Rmd`)
@@
 
-### Key Velocity (Individual)
+### Key Velocity (Summary)
 ```{r vel-summary, echo = FALSE}
@@
 
-### KV Difference (Individual)
+### KV Difference (Summary)
 ```{r vel-diff-summary, echo = FALSE}

### 2020-12-03 experiment-1/analysis/stats/dynamics.Rmd
move stuff to experiment-1
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 analysis/stats/dynamics.Rmd
@@
 ---
-title: 'Dynamics Analysis: The Sound of Teaching Music 2'
+title: 'Dynamics Analysis: The Sound of Teaching Music 1'
 output:
@@
 2. Excluded participants
-- SubNr 3: Insufficient motor skills
+
+- SubNr 1: Most of responses (more than 90%) were removed due to outliers
+- SubNr 3: Experimental error
+- SubNr 8: Experimental error
+- SubNr 14: Experimental error
+- SubNr 16: A deviated tempo
 
@@
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  scale_x_continuous(breaks = seq(1,72,1))
+  scale_x_continuous(breaks = seq(1,67,1))
 
@@
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  scale_x_continuous(breaks = seq(1,71,1))
+  scale_x_continuous(breaks = seq(1,66,1))
 ```

### 2020-12-03 analysis/demographics/questionnaire.Rmd

---
title: 'Demographics: The Sound of Teaching Music 2'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of demographic information

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

# Questionnaire

```{r file, include = FALSE}
filename_q = "./questionnaire.csv"
```

```{r extract, echo = FALSE}
# read csv files
dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
dt <- dt[!21]

# change some colnames
colnames(dt)[2] <- "SubNr"
colnames(dt)[3] <- "Age"
colnames(dt)[4] <- "Gender"
colnames(dt)[7] <- "Handedness"
colnames(dt)[18] <- "PianoTotalPractice"

# extract the total number of years of training in piano (input manually based on the original csv file)
dt$PianoTotalPractice <- c(13, 15, 15, 16, 9, 10, 24, 8, 19, 6, 12, 10, 21, 23, 15, 22, 22, 24, 17, 15, 12)

# change some characteristics
dt$Age <- as.numeric(dt$Age)
dt$PianoTotalPractice <- as.numeric(dt$PianoTotalPractice)

# exclude participants
dt_included <- dt[SubNr != 3]

print(dt_included)
```

## Descriptive stats for included participants
Note: A different questionnaire was used for only SubNr 21 and adjusted for necessary columns (but not others; see original csv for the precise information)

- Number of included participants: `r nrow(dt_included)` (mean age: `r round(mean(dt_included$Age), 2)`, sd: `r round(sd(dt_included$Age), 2)`)
- Male: `r nrow(dt_included[Gender == "Male"])` (mean age: `r round(mean(dt_included[Gender == "Male"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Male"]$Age), 2)`)
- Female: `r nrow(dt_included[Gender == "Female"])` (mean age: `r round(mean(dt_included[Gender == "Female"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Female"]$Age), 2)`)
- Average training years: `r round(mean(dt_included$PianoTotalPractice), 2)` years
- Handedness: Right (`r nrow(dt_included[Handedness == "Right"])`), Left (`r nrow(dt_included[Handedness == "Left"])`), Both (`r nrow(dt_included[Handedness == "Both"])`)

## Some descriptive questions
1. Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?

```{r 1, echo = FALSE}
data.table("Answer" = dt_included$`Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?`)
```

2. Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.

```{r 2, echo = FALSE}
data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
```

### 2020-12-03 analysis/stats/ioi.Rmd

@@
 ---
-title: 'Tempo Analysis: The Sound of Teaching Music 1'
+title: 'Tempo Analysis: The Sound of Teaching Music 2'
 output:
@@
 ## 1. All - individual{#art_ind_all}
-- Average IOIs for each participant.
+- Average normIOIs for each participant.
 
@@
 # For each individual
-ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
+ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
 ioi_art
@@
 #### Boxplot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+- A dashed line represents the tempo given by a metronome.
 
 ```{r ioi-art-box,  echo = FALSE}
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  geom_hline(yintercept = 188, linetype = "dashed")
+  geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ## 2. All - group{#art_group_all}
-- Average IOIs for each condition.
+- Average normIOIs for each condition.
 - Hypothesis: Participants will play slower when teaching.
@@
 #### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+- A dashed line represents the tempo given by a metronome.
 
@@
 
-ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
+ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
@@
 #### Normality check
-- Violate the normality for differences between the two conditions
 
@@
 ioi_art_norm
+```
+
+#### Paired *t*-test
+
+```{r paired-art, echo = FALSE}
+ioi_art_ttest <- t.test(ioi_art[Condition == "teaching"]$Mean, ioi_art[Condition == "performing"]$Mean, paired = TRUE, alternative = "two.sided")
+ioi_art_ttest
 ```
@@
 ## 3. Phrase boundaries - individual{#art_ind_phrase}
-- Compare IOIs at phrase boundaries with others for each participant.
+- Compare normIOIs at phrase boundaries with others for each participant.
 
@@
 # For each individual
-ioi_art_ch <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
+ioi_art_ch <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Boundary)]
 ioi_art_ch
@@
 #### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+- A dashed line represents the tempo given by a metronome.
 
 ```{r ioi-art-ch-box, echo = FALSE}
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 188, linetype = "dashed")
+  geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ## 4. Phrase boundaries - group{#art_group_phrase}
-- Compare IOIs at phrase boundaries with others in the two conditions.
+- Compare normIOIs at phrase boundaries with others in the two conditions.
 - Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations (especially for teaching purposes).
@@
 #### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
 
@@
 
-ggpaired(ioi_art_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
+ggpaired(ioi_art_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
@@
 ## 5. Trial by trial - individual{#art_ind_trial}
-- Average IOIs for each trial for each participant.
+- Average normIOIs for each trial for each participant.
 
@@
 # For each individual
-ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
+ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, TrialNr)]
 ioi_art_trial
@@
 #### Line graph
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+- A dashed line represents the tempo given by a metronome.
 
 ```{r ioi-art-trial-line, echo = FALSE, fig.height = 4}
-ggline(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggline(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
   facet_wrap(SubNr ~ .) +
@@
   scale_x_continuous(breaks = seq(1,8,1)) +
-  geom_hline(yintercept = 188, linetype = "dashed")
+  geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ## 6. Trial by trial - group{#art_group_trial}
-- Average IOIs for each trial for each condition.
+- Average normIOIs for each trial for each condition.
 
@@
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+- A dashed line represents the tempo given by a metronome.
 
 ```{r ioi-art-trial-all-line, echo = FALSE}
-ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
+ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
@@
 ## 1. All - individual{#dyn_ind_all}
-- Average IOIs for each participant.
+- Average normIOIs for each participant.
 
@@
 # For each individual
-ioi_dyn <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
+ioi_dyn <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
 ioi_dyn
@@
 #### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+- A dashed line represents the tempo given by a metronome.
 
 ```{r ioi-dyn-box,  echo = FALSE}
-ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  geom_hline(yintercept = 188, linetype = "dashed")
+  geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ## 2. All - group{#dyn_group_all}
-- Average IOIs for each condition.
+- Average normIOIs for each condition.
 - Hypothesis: Participants will play slower when teaching.
@@
 
-ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
+ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
@@
 ioi_dyn_norm
+```
+
+#### Paired *t*-test
+
+```{r paired-dyn, echo = FALSE}
+ioi_dyn_ttest <- t.test(ioi_dyn[Condition == "teaching"]$Mean, ioi_dyn[Condition == "performing"]$Mean, paired = TRUE, alternative = "two.sided")
+ioi_dyn_ttest
 ```
@@
 ## 3. Phrase boundaries - individual{#dyn_ind_phrase}
-- Compare IOIs at phrase boundaries with others for each participant.
+- Compare normIOIs at phrase boundaries with others for each participant.
 
@@
 # For each individual
-ioi_dyn_ch <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
+ioi_dyn_ch <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Boundary)]
 ioi_dyn_ch
@@
 #### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+- A dashed line represents the tempo given by a metronome.
 
 ```{r ioi-dyn-ch-box, echo = FALSE}
-ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 188, linetype = "dashed")
+  geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ## 4. Phrase boundaries - group{#dyn_group_phrase}
-- Compare IOIs at phrase boundaries with others in the two conditions.
+- Compare normIOIs at phrase boundaries with others in the two conditions.
 - Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations (especially for teaching purposes).
@@
 
-ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
+ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
@@
 ## 5. Trial by trial - individual{#dyn_ind_trial}
-- Average IOIs for each trial for each participant.
+- Average normIOIs for each trial for each participant.
 
@@
 # For each individual
-ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
+ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, TrialNr)]
 ioi_dyn_trial
@@
 #### Line graph
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+- A dashed line represents the tempo given by a metronome.
 
 ```{r ioi-dyn-trial-line, echo = FALSE, fig.height = 4}
-ggline(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggline(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
   facet_wrap(SubNr ~ .) +
@@
 ## 6. Trial by trial - group{#dyn_group_trial}
-- Average IOIs for each trial for each condition.
+- Average normIOIs for each trial for each condition.
 
@@
 - Error bars show SEM of the mean
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+- A dashed line represents the tempo given by a metronome.
 
 ```{r ioi-dyn-trial-all-line, echo = FALSE}
-ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
+ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
@@
 # For each individual
-ioi_art_seq <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+ioi_art_seq <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Interval)]
 
-ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
+ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-  scale_x_continuous(breaks = seq(1,66,1))
+  scale_x_continuous(breaks = seq(1,71,1))
 ```
@@
 # For each individual
-ioi_dyn_seq <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+ioi_dyn_seq <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Interval)]
 
-ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
+ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
-    scale_x_continuous(breaks = seq(1,66,1))
+    scale_x_continuous(breaks = seq(1,71,1))
 ```

### 2020-12-03 paper/manuscript.Rmd
manuscript init
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title             : "The Sound of Teaching Music: Expert Pianists' Sound Modulation"
shorttitle        : "The Sound of Teaching Music"

author: 
  - name          : "Atsuko Tominaga"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2020-12-03 paper/manuscript.tex
manuscript init
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The Sound of Teaching Music: Expert Pianists' Sound Modulation},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{The Sound of Teaching Music}
\author{Atsuko Tominaga\textsuperscript{1}\ \& Ernst-August Doelle\textsuperscript{1,2}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Wilhelm-Wundt-University\\\textsuperscript{2} Konstanz Business School}
\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

Correspondence concerning this article should be addressed to Atsuko Tominaga, Postal address. E-mail: my@email.com}
\keywords{keywords\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\title{The Sound of Teaching Music: Expert Pianists' Sound Modulation}

\date{}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
}

\begin{document}
\maketitle

\hypertarget{methods}{%
\section{Methods}\label{methods}}

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

\hypertarget{material}{%
\subsection{Material}\label{material}}

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

We used R (Version 3.6.3; R Core Team, 2020) for all our analyses.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
\end{cslreferences}

\endgroup

\end{document}

### 2020-12-04 experiment-1/analysis/preprocessor/included_summary.Rmd
export r chunks as R scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 data_analysis~.txt are used for data analysis.
-```{r export, echo = TRUE}
+```{r export-csv, echo = TRUE}
 fwrite(dt_ioi[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_ioi.txt"), sep = ",")
@@
 ```
+
+```{r export, include = FALSE}
+knitr::purl("included_summary.Rmd")
+```

### 2020-12-04 experiment-1/analysis/demographics/questionnaire.R

@@ -1,0 +1,45 @@
+## ----setup, include = FALSE---------------------------------------------------------
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+
+
+## ----file, include = FALSE----------------------------------------------------------
+filename_q = "./questionnaire.csv"
+
+
+## ----extract, echo = FALSE----------------------------------------------------------
+# read csv files
+dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
+
+# change some colnames
+colnames(dt)[2] <- "SubNr"
+colnames(dt)[3] <- "Age"
+colnames(dt)[4] <- "Gender"
+colnames(dt)[7] <- "Handedness"
+
+# change some characteristics
+dt$Age <- as.numeric(dt$Age)
+
+# extract the total number of years of training in piano (input manually based on the original csv file)
+dt$PianoTotalPractice <- c(14, 20, 27, 21, 10, 16, 9, 13, 25, 14, 13, 11, 6, 4, 6, 8, 21, 15, 18, 7, 11, 12, 12, 18, 12, 5, 5, 10, 16, 10, 8, 22, 10, 5, 4, 14)
+
+# exclude participants
+dt_included <- dt[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16]
+
+print(dt_included)
+
+
+## ----1, echo = FALSE----------------------------------------------------------------
+data.table("Answer" = dt_included$`Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?`)
+
+
+## ----2, echo = FALSE----------------------------------------------------------------
+data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
+
+
+## ----export, include = FALSE--------------------------------------------------------
+knitr::purl("questionnaire.Rmd")
+

### 2020-12-04 experiment-1/analysis/preprocessor/error_summary.R

@@ -1,17 +1,4 @@
----
-title: "Error Summary (teaching-v1.0)"
-output: html_notebook
----
-
-- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
-
-# Workflow
-1. filter with `filtering.R`
-2. check the details of errors and remove some participants if needed (**this script**: `error_summary.Rmd`)
-3. trim data (i.e., remove outliers) with `trimming_ioi.R`, `trimming_kot.R`, `trimming_vel.R`
-4. check trimmed data and remove some participants if needed (`included_summary.Rmd`)
-
-```{r setup, include = FALSE}
+## ----setup, include = FALSE--------------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -35,53 +22,30 @@ dt_onset <- fread(file = here("filtered", "dt_correct_onset.txt"), sep = ",")
 dt_offset <- fread(file = here("filtered", "dt_correct_offset.txt"), sep = ",")
 error_onset <- fread(file = here("filtered", "error_onset.txt"), sep = ",")
 error_offset <- fread(file = here("filtered", "error_offset.txt"), sep = ",")
-```
 
-# Design
-- Each performance (1 trial) contains 67 responses for onsets and 67 responses for offsets (i.e., if there is no mistake, 134 responses in total)
-- Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
-- Note: SubNr 8 was excluded due to an experimental error
 
-# Outline
-1. [Onset reports](#onset)
-2. [Offset reports](#offset)
-
-## ONSET
-### Corrected Error Onset (All)
-```{r onset-all, echo = FALSE}
+## ----onset-all, echo = FALSE-------------------------------------------------------
 error_onset_all <- error_onset[CorrectionNr != "Exclude"]
 error_onset_all
-```
 
-### Corrected Error Onset (Summary)
-```{r onset-summary, echo = FALSE}
+
+## ----onset-summary, echo = FALSE---------------------------------------------------
 error_onset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
 error_onset_summary <- error_onset_all[, .(N = .N, Sum = sum(as.numeric(CorrectionNr))), by = .(Reason)]
 error_onset_summary
-```
 
-### Missing/Excluded Trials Onset (All)
-```{r na-onset-all, echo = FALSE}
+
[…]

### 2020-12-04 experiment-1/analysis/preprocessor/included_summary.R

@@ -1,0 +1,60 @@
+## ----setup, include = FALSE---------------------------------------------------------
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
+if (!require("here")) {install.packages("here"); require("here")}
+
+# set working directory
+here::i_am("included_summary.Rmd")
+
+# read txt files
+dt_ioi <- fread(file = here("trimmed", "data_ioi_1.txt"), sep = ",")
+dt_kot <- fread(file = here("trimmed", "data_kot.txt"), sep = ",")
+dt_vel <- fread(file = here("trimmed", "data_vel.txt"), sep = ",")
+dt_vel_diff <- fread(file = here("trimmed", "data_vel_diff.txt"), sep = ",")
+
+
+## ----ioi, echo = FALSE--------------------------------------------------------------
+dt_ioi
+
+
+## ----ioi-summary, echo = FALSE------------------------------------------------------
+dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
+
+
+## ----kot, echo = FALSE--------------------------------------------------------------
+dt_kot[SubNr != 1]
+
+
+## ----kot-summary, echo = FALSE------------------------------------------------------
+dt_kot[SubNr != 1, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
+
+
+## ----vel, echo = FALSE--------------------------------------------------------------
+dt_vel[SubNr != 1]
+
+
+## ----vel-summary, echo = FALSE------------------------------------------------------
+dt_vel[SubNr != 1, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
+
+
+## ----vel-diff, echo = FALSE---------------------------------------------------------
+dt_vel_diff[SubNr != 1]
+
+
+## ----vel-diff-summary, echo = FALSE-------------------------------------------------
+dt_vel_diff[SubNr != 1, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
+
+
[…]

### 2020-12-04 experiment-1/analysis/stats/articulation.R

@@ -1,80 +1,4 @@
----
-title: 'Articulation Analysis: The Sound of Teaching Music 1'
-output:
-  html_notebook: default
-editor_options: 
-  chunk_output_type: inline
----
-
-Description: This is a summary of articulation analysis.
-
-- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
-
-# Experimental Design{#design}
-- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: legato vs. staccato / forte vs. piano)
-- Dependent variables: Key-Overlap Time (KOT), Key-Overlap Ratio (KOR)
-
-1. Main predictions
-Participants will exaggerate legato and staccato more while teaching.
-
-2. Excluded participants
-
-- SubNr 1: Most of responses (more than 90%) were removed due to outliers
-- SubNr 3: Experimental error
-- SubNr 8: Experimental error
-- SubNr 14: Experimental error
-- SubNr 16: A deviated tempo
-
-3. Stats packages
-- [afex](https://github.com/singmann/afex) for ANOVAs
-- [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons
-
-# Outline - Articulation (KOT, KOR)
-## [Articulation](#kot-articulation)
-
-### KOT
-- Individual results
-  + [All](#kot_art_ind_all)
-  + [Trial by trial](#kot_art_ind_trial)
- 
-- Group results and stats (if available)
-  + [All](#kot_art_group_all)
-  + [Trial by trial](#kot_art_group_trial)
-  
-- [Sequence plot](#sequence_art)
-
-### KOR
-- Individual results
-  + [All](#kor_art_ind_all)
-  + [Trial by trial](#kor_art_ind_trial)
- 
-- Group results and stats (if available)
-  + [All](#kor_art_group_all)
-  + [Trial by trial](#kor_art_group_trial)
-
-## [Dynamics](#kot-dynamics)
-
-### KOT
-- Individual results
-  + [All](#kot_dyn_ind_all)
-  + [Trial by trial](#kot_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kot_dyn_group_all)
-  + [Trial by trial](#kot_dyn_group_trial)
-
-- [Sequence plot](#sequence_dyn)
-
-### KOR
-- Individual results
-  + [All](#kor_dyn_ind_all)
-  + [Trial by trial](#kor_dyn_ind_trial)
[…]

### 2020-12-04 experiment-1/analysis/stats/dynamics.R

@@ -1,87 +1,4 @@
----
-title: 'Dynamics Analysis: The Sound of Teaching Music 1'
-output:
-  html_notebook: default
-editor_options: 
-  chunk_output_type: inline
----
-
-Description: This is a summary of dynamics analysis.
-
-- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
-
-# Experimental Design{#design}
-- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato)
-- Dependent variables: Key Velocity (KV), KV difference
-
-1. Main predictions
-- Participants will exaggerate forte and piano more while teaching.
-
-2. Excluded participants
-
-- SubNr 1: Most of responses (more than 90%) were removed due to outliers
-- SubNr 3: Experimental error
-- SubNr 8: Experimental error
-- SubNr 14: Experimental error
-- SubNr 16: A deviated tempo
-
-3. Stats packages
-- [afex](https://github.com/singmann/afex) for ANOVAs
-- [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons
-
-# Outline - Dynamics (Key Velocity)
-
-## [Dynamics](#vel-dynamics)
-
-### KV
-- Individual results
-  + [All](#vel_dyn_ind_all)
-  + [Trial by trial](#vel_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#vel_dyn_group_all)
-  + [Trial by trial](#vel_dyn_group_trial)
-  
-- Order analysis (teaching/performing first?)
-  + [All](#vel_dyn_order_group_all)
-
-### KV-Diff
-- Individual results
-  + [All](#vel_diff_dyn_ind_all)
-  + [Trial by trial](#vel_diff_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#vel_diff_dyn_group_all)
-  + [Trial by trial](#vel_diff_dyn_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#vel_diff_dyn_order_group_all)
-  
-- [Sequence plot](#sequence_dyn)
-
-## [Articulation](#vel_articulation)
-
-### KV
-- Individual results
-  + [All](#vel_art_ind_all)
-  + [Trial by trial](#vel_art_ind_trial)
-
[…]

### 2020-12-04 experiment-1/analysis/stats/ioi.R

@@ -1,63 +1,4 @@
----
-title: 'Tempo Analysis: The Sound of Teaching Music 1'
-output:
-  html_notebook: default
-editor_options: 
-  chunk_output_type: inline
----
-
-Description: This is a summary of tempo analysis.
-
-- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
-
-# Experimental Design{#design}
-- Independent variables: Condition (2 levels: teaching vs. performing) x Technique (2 levels: articulation vs. dynamics)
-- Dependent variables: Interonset intervals (IOIs)
-
-1. Main predictions
-- Participants will play slower while teaching regardless of the techniques.
-
-2. Excluded participants
-
-- SubNr 1: Most of responses (more than 90%) were removed due to outliers
-- SubNr 3: Experimental error
-- SubNr 8: Experimental error
-- SubNr 14: Experimental error
-- SubNr 16: A deviated tempo
-
-3. Stats packages
-- [afex](https://github.com/singmann/afex) for ANOVAs
-
-# Outline - Tempo (IOIs)
-## [Articulation](#ioi-articulation)
-
-- Individual results
-  + [All](#art_ind_all)
-  + [Phrase bondaries](#art_ind_phrase)
-  + [Trial by trial](#art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#art_group_all)
-  + [Phrase bondaries](#art_group_phrase)
-  + [Trial by trial](#art_group_trial)
-  
-- [Sequence plot](#sequence_art)
-
-## [Dynamics](#ioi-dynamics)
-
-- Individual results
-  + [All](#dyn_ind_all)
-  + [Phrase bondaries](#dyn_ind_phrase)
-  + [Trial by trial](#dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#dyn_group_all)
-  + [Phrase bondaries](#dyn_group_phrase)
-  + [Trial by trial](#dyn_group_trial)
-  
-- [Sequence plot](#sequence_dyn)
-
-```{r setup, include = FALSE}
+## ----setup, include = FALSE--------------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
[…]

### 2020-12-04 paper/manuscript.Rmd
write outlines
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title             : "The Sound of Teaching Music: Expert Pianists' Sound Modulation"
+title             : "The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices"
 shorttitle        : "The Sound of Teaching Music"
@@
     corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
-    email         : "my@email.com"
-  - name          : "Ernst-August Doelle"
-    affiliation   : "1,2"
+    address       : "Quellenstraße 51, 1100 Vienna, Austria"
+    email         : "Tominaga_Atsuko@phd.ceu.edu"
+
+  - name          : "Günther Knoblich"
+    affiliation   : "1"
+
+  - name          : "Natalie Sebanz"
+    affiliation   : "1"
 
@@
   - id            : "1"
-    institution   : "Wilhelm-Wundt-University"
-  - id            : "2"
-    institution   : "Konstanz Business School"
+    institution   : "Department of Cognitive Science, Central European University"
 
 authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
 
@@
 
-keywords          : "keywords"
+keywords          : "teaching, expertise, skill transmission, artistic expression, music"
 wordcount         : "X"
 
-bibliography      : ["r-references.bib"]
+bibliography      : [""]
 
-floatsintext      : no
+floatsintext      : yes
 figurelist        : no
@@
 draft             : no
+noextraspace      : yes
 
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output:
+  papaja::apa6_pdf:
+    latex_engine: xelatex
 ---
+```{r setup, include = FALSE}
+if (!require("papaja")) {install.packages("papaja"); require("papaja")}
+if (!require("here")) {install.packages("here"); require("here")}
 
-```{r setup, include = FALSE}
-library("papaja")
+# create image folder
+if (!file.exists(here("paper/image"))){
+  dir.create(here("paper/image"))
+}
 ```
 
-```{r analysis-preferences}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+# 1. Introduction
+
+# 2. Experiment 1
+
+## 2.1. Method
+```{r questionnaire-1, include = FALSE}
+source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
 ```
 
-# Methods
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
+### 2.1.1. Participants
+### 2.1.2. Apparatus and stimuli
+### 2.1.3. Procedure
+### 2.1.4. Design and data analysis
 
-## Participants
+## 2.2. Results (Articulation)
+### 2.2.1. Interonset Intervals (IOIs)
+### 2.2.2. Key-Overlap Time (KOT)
+### 2.2.3. Key Velocity (KV)
+### 2.2.4. KV Transition Points
 
-## Material
+## 2.3. Results (Dynamics)
+### 2.3.1. Interonset Intervals (IOIs)
+### 2.3.2. Key Velocity (KV)
+### 2.3.3. KV Transition Points
+### 2.3.4. Key-Overlap Time (KOT)
 
-## Procedure
+## 2.4. Discussion
 
-## Data analysis
-We used `r cite_r("r-references.bib")` for all our analyses.
+\newpage
 
-# Results
+# 3. Experiment 2
+## 3.1. Method
+```{r questionnaire-2, include = FALSE}
+#source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+```
 
-# Discussion
+### 3.1.1. Participants
+### 3.1.2. Apparatus and stimuli
+### 3.1.3. Procedure
+### 3.1.4. Design and data analysis
+
+## 3.2. Results (Articulation)
+### 3.2.1. Interonset Intervals (IOIs)
+### 3.2.2. Key-Overlap Time (KOT)
+### 3.2.3. Key Velocity (KV)
+### 3.2.4. KV Transition Points
+
+## 3.3. Results (Dynamics)
+### 3.3.1. Interonset Intervals (IOIs)
+### 3.3.2. Key Velocity (KV)
+### 3.3.3. KV Transition Points
+### 3.3.4. Key-Overlap Time (KOT)
+
+## 3.4. Discussion
+
+# 4. General Discussion
+
+# Acknowledgements
+We thank Dávid Csűrös for his help with data collection.
+This research was supported by the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013)/ERC (European Research Council) grant agreement no.
+609819, SOMICS, and by ERC grant agreement no.
+616072, JAXPERTISE.
+
+# Declarations of interest
+All authors have no conflict of interest.
 
@@
 # References
-```{r create_r-references}
-r_refs(file = "r-references.bib")
-```
 
@@
 \endgroup
+\raggedbottom

### 2020-12-04 paper/manuscript.tex
write outlines
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   english,
-  man]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{lmodern}
@@
 \hypersetup{
-  pdftitle={The Sound of Teaching Music: Expert Pianists' Sound Modulation},
+  pdftitle={The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices},
   pdflang={en-EN},
-  pdfkeywords={keywords},
+  pdfkeywords={teaching, expertise, skill transmission, artistic expression, music},
   hidelinks,
@@
 \shorttitle{The Sound of Teaching Music}
-\author{Atsuko Tominaga\textsuperscript{1}\ \& Ernst-August Doelle\textsuperscript{1,2}}
+\author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
 \affiliation{
 \vspace{0.5cm}
-\textsuperscript{1} Wilhelm-Wundt-University\\\textsuperscript{2} Konstanz Business School}
-\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+\textsuperscript{1} Department of Cognitive Science, Central European University}
+\authornote{
 
-Enter author note here.
-
-Correspondence concerning this article should be addressed to Atsuko Tominaga, Postal address. E-mail: my@email.com}
-\keywords{keywords\newline\indent Word count: X}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
+Correspondence concerning this article should be addressed to Atsuko Tominaga, Quellenstraße 51, 1100 Vienna, Austria. E-mail: Tominaga\_Atsuko@phd.ceu.edu}
+\keywords{teaching, expertise, skill transmission, artistic expression, music\newline\indent Word count: X}
 \usepackage{lineno}
@@
 \newenvironment{cslreferences}%
-  {\setlength{\parindent}{0pt}%
-  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {}%
   {\par}
 
-\title{The Sound of Teaching Music: Expert Pianists' Sound Modulation}
+\title{The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices}
 
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\hypertarget{introduction}{%
+\section{1. Introduction}\label{introduction}}
 
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
+\hypertarget{experiment-1}{%
+\section{2. Experiment 1}\label{experiment-1}}
+
+\hypertarget{method}{%
+\subsection{2.1. Method}\label{method}}
 
 \hypertarget{participants}{%
-\subsection{Participants}\label{participants}}
+\subsubsection{2.1.1. Participants}\label{participants}}
 
-\hypertarget{material}{%
-\subsection{Material}\label{material}}
+\hypertarget{apparatus-and-stimuli}{%
+\subsubsection{2.1.2. Apparatus and stimuli}\label{apparatus-and-stimuli}}
 
 \hypertarget{procedure}{%
-\subsection{Procedure}\label{procedure}}
+\subsubsection{2.1.3. Procedure}\label{procedure}}
 
-\hypertarget{data-analysis}{%
-\subsection{Data analysis}\label{data-analysis}}
+\hypertarget{design-and-data-analysis}{%
+\subsubsection{2.1.4. Design and data analysis}\label{design-and-data-analysis}}
 
-We used R (Version 3.6.3; R Core Team, 2020) for all our analyses.
+\hypertarget{results-articulation}{%
+\subsection{2.2. Results (Articulation)}\label{results-articulation}}
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\hypertarget{interonset-intervals-iois}{%
+\subsubsection{2.2.1. Interonset Intervals (IOIs)}\label{interonset-intervals-iois}}
+
+\hypertarget{key-overlap-time-kot}{%
+\subsubsection{2.2.2. Key-Overlap Time (KOT)}\label{key-overlap-time-kot}}
+
+\hypertarget{key-velocity-kv}{%
+\subsubsection{2.2.3. Key Velocity (KV)}\label{key-velocity-kv}}
+
+\hypertarget{kv-transition-points}{%
+\subsubsection{2.2.4. KV Transition Points}\label{kv-transition-points}}
+
+\hypertarget{results-dynamics}{%
+\subsection{2.3. Results (Dynamics)}\label{results-dynamics}}
+
+\hypertarget{interonset-intervals-iois-1}{%
+\subsubsection{2.3.1. Interonset Intervals (IOIs)}\label{interonset-intervals-iois-1}}
+
+\hypertarget{key-velocity-kv-1}{%
+\subsubsection{2.3.2. Key Velocity (KV)}\label{key-velocity-kv-1}}
+
+\hypertarget{kv-transition-points-1}{%
+\subsubsection{2.3.3. KV Transition Points}\label{kv-transition-points-1}}
+
+\hypertarget{key-overlap-time-kot-1}{%
+\subsubsection{2.3.4. Key-Overlap Time (KOT)}\label{key-overlap-time-kot-1}}
 
 \hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\subsection{2.4. Discussion}\label{discussion}}
+
+\newpage
+
+\hypertarget{experiment-2}{%
+\section{3. Experiment 2}\label{experiment-2}}
+
+\hypertarget{method-1}{%
+\subsection{3.1. Method}\label{method-1}}
+
+\hypertarget{participants-1}{%
+\subsubsection{3.1.1. Participants}\label{participants-1}}
+
+\hypertarget{apparatus-and-stimuli-1}{%
+\subsubsection{3.1.2. Apparatus and stimuli}\label{apparatus-and-stimuli-1}}
+
+\hypertarget{procedure-1}{%
+\subsubsection{3.1.3. Procedure}\label{procedure-1}}
+
+\hypertarget{design-and-data-analysis-1}{%
+\subsubsection{3.1.4. Design and data analysis}\label{design-and-data-analysis-1}}
+
+\hypertarget{results-articulation-1}{%
+\subsection{3.2. Results (Articulation)}\label{results-articulation-1}}
+
+\hypertarget{interonset-intervals-iois-2}{%
+\subsubsection{3.2.1. Interonset Intervals (IOIs)}\label{interonset-intervals-iois-2}}
+
+\hypertarget{key-overlap-time-kot-2}{%
+\subsubsection{3.2.2. Key-Overlap Time (KOT)}\label{key-overlap-time-kot-2}}
+
+\hypertarget{key-velocity-kv-2}{%
+\subsubsection{3.2.3. Key Velocity (KV)}\label{key-velocity-kv-2}}
+
+\hypertarget{kv-transition-points-2}{%
+\subsubsection{3.2.4. KV Transition Points}\label{kv-transition-points-2}}
+
+\hypertarget{results-dynamics-1}{%
+\subsection{3.3. Results (Dynamics)}\label{results-dynamics-1}}
+
+\hypertarget{interonset-intervals-iois-3}{%
+\subsubsection{3.3.1. Interonset Intervals (IOIs)}\label{interonset-intervals-iois-3}}
+
+\hypertarget{key-velocity-kv-3}{%
+\subsubsection{3.3.2. Key Velocity (KV)}\label{key-velocity-kv-3}}
+
+\hypertarget{kv-transition-points-3}{%
+\subsubsection{3.3.3. KV Transition Points}\label{kv-transition-points-3}}
+
+\hypertarget{key-overlap-time-kot-3}{%
+\subsubsection{3.3.4. Key-Overlap Time (KOT)}\label{key-overlap-time-kot-3}}
+
+\hypertarget{discussion-1}{%
+\subsection{3.4. Discussion}\label{discussion-1}}
+
+\hypertarget{general-discussion}{%
+\section{4. General Discussion}\label{general-discussion}}
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+We thank Dávid Csűrös for his help with data collection.
+This research was supported by the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013)/ERC (European Research Council) grant agreement no.
+609819, SOMICS, and by ERC grant agreement no.
+616072, JAXPERTISE.
+
+\hypertarget{declarations-of-interest}{%
+\section{Declarations of interest}\label{declarations-of-interest}}
+
+All authors have no conflict of interest.
 
@@
 \begin{cslreferences}
-\leavevmode\hypertarget{ref-R-base}{}%
-R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
 \end{cslreferences}
@@
 \endgroup
+\raggedbottom
 

### 2020-12-04 experiment-1/analysis/demographics/questionnaire.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------------
+## ----setup, include = FALSE--------------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -6,11 +6,11 @@ if (!require("data.table")) {install.packages("data.table"); require("data.table
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----file, include = FALSE----------------------------------------------------------
+## ----file, include = FALSE---------------------------------------------------------
 filename_q = "./questionnaire.csv"
 
 
-## ----extract, echo = FALSE----------------------------------------------------------
+## ----extract, echo = FALSE---------------------------------------------------------
 # read csv files
 dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
 
@@ -32,14 +32,14 @@ dt_included <- dt[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 
 print(dt_included)
 
 
-## ----1, echo = FALSE----------------------------------------------------------------
+## ----1, echo = FALSE---------------------------------------------------------------
 data.table("Answer" = dt_included$`Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?`)
 
 
-## ----2, echo = FALSE----------------------------------------------------------------
+## ----2, echo = FALSE---------------------------------------------------------------
 data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
 
 
-## ----export, include = FALSE--------------------------------------------------------
+## ----export, include = FALSE-------------------------------------------------------
 knitr::purl("questionnaire.Rmd")
 

### 2020-12-10 experiment-2/analysis/demographics/questionnaire.Rmd
change a variable name dt >> dt_2
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Demographics: The Sound of Teaching Music 2'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of demographic information

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

# Questionnaire

```{r file, include = FALSE}
filename_q = "./questionnaire.csv"
```

```{r extract, echo = FALSE}
# read csv files
dt_2 <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
dt_2 <- dt_2[!21]

# change some colnames
colnames(dt_2)[2] <- "SubNr"
colnames(dt_2)[3] <- "Age"
colnames(dt_2)[4] <- "Gender"
colnames(dt_2)[7] <- "Handedness"
colnames(dt_2)[18] <- "PianoTotalPractice"

# extract the total number of years of training in piano (input manually based on the original csv file)
dt_2$PianoTotalPractice <- c(13, 15, 15, 16, 9, 10, 24, 8, 19, 6, 12, 10, 21, 23, 15, 22, 22, 24, 17, 15, 12)

# change some characteristics
dt_2$Age <- as.numeric(dt_2$Age)
dt_2$PianoTotalPractice <- as.numeric(dt_2$PianoTotalPractice)

# exclude participants
dt_2_included <- dt_2[SubNr != 3]

print(dt_2_included)
```

## Descriptive stats for included participants
Note: A different questionnaire was used for only SubNr 21 and adjusted for necessary columns (but not others; see original csv for the precise information)

- Number of included participants: `r nrow(dt_2_included)` (mean age: `r round(mean(dt_2_included$Age), 2)`, sd: `r round(sd(dt_2_included$Age), 2)`)
- Male: `r nrow(dt_2_included[Gender == "Male"])` (mean age: `r round(mean(dt_2_included[Gender == "Male"]$Age), 2)`, sd: `r round(sd(dt_2_included[Gender == "Male"]$Age), 2)`)
- Female: `r nrow(dt_2_included[Gender == "Female"])` (mean age: `r round(mean(dt_2_included[Gender == "Female"]$Age), 2)`, sd: `r round(sd(dt_2_included[Gender == "Female"]$Age), 2)`)
- Average training years: `r round(mean(dt_2_included$PianoTotalPractice), 2)` years
- Handedness: Right (`r nrow(dt_2_included[Handedness == "Right"])`), Left (`r nrow(dt_2_included[Handedness == "Left"])`), Both (`r nrow(dt_2_included[Handedness == "Both"])`)

## Some descriptive questions
1. Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?

```{r 1, echo = FALSE}
data.table("Answer" = dt_2_included$`Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?`)
```

2. Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.

```{r 2, echo = FALSE}
data.table("Answer" = dt_2_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
```

### 2020-12-10 paper/manuscript.Rmd
add methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("here")) {install.packages("here"); require("here")}
+if (!require("magick")) {install.packages("magick"); require("magick")}
 
@@
 ### 2.1.1. Participants
+
+We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment.
+For data analysis, we excluded 5 participants due to an experimental error (3), an overall deviated tempo (2).
+Thirty-one participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis.
+They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`).
+All participants gave their informed consent before the experiment started and received vouchers for their participation.
+The study was approved by the United Ethical REview Committee for Research in Psychology (EPKEB) in Hungary.
+
 ### 2.1.2. Apparatus and stimuli
+
+A weighted Yamaha MIDI digital piano was used to record participants' performance via Max/MSP on a MacBook Pro with MacOS X Mojave 10.14.3.
+The laptop and piano were connected to a high-fidelity soundcard (Focusrite Scarlett 6i6) to deliver a metronome and piano sound.
+All auditory feedback was given to participants through headphones (Audio-Technica ATH-M50X).
+Sheet music was displayed on a computer monitor in front of the participants.
+The pitch, onset and offset time of each note, and key velocity profiles were obtained from MIDI data using Max/MSP patchers.
+
+One musical excerpt was used as a stimulus.
+The excerpt was taken from ``A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Brunam and modified for the experiment.
+It consisted of a 6-measure isochronous melody noted in 4/4 meter.
+The stimulus was composed in C major to be played with the right hand only.
+Non-expressive sheet music (i.e., sheet music without musical expression; see *Figure \ref{fig:stimuli}, A*) was used for the purpose of practice.
+Expressive notations were added to the non-expressive sheet music for the experiment.
+They referred to either articulation or dynamics (*Figure \ref{fig:stimuli}, B-C*).
+Articulation was notated as either legato or staccato.
+Legato indicates that musical notes are to be connected and should sound smooth.
+Staccato requires producing musical notes with shortened duration, keeping them separate from each other.
+Dynamics was notated in terms of forte and piano.
+Forte indicates that musical notes should be played loudly whereas piano indicates that musical notes should be played softly.
+The notation did not include any indication of fingering (i.e., the positioning of the fingers when playing the piano) because the piece was simple and pilot testing had shown that specifying fingering was not required.
+
 ### 2.1.3. Procedure
+
+In order to ensure that participants had sufficient motor skills to perform the piece, we asked them to practise the piece with the non-expressive sheet music (*Figure \ref{fig:stimuli}, A*).
+Once they could perform the piece without pitch errors twice consecutively, the experiment started.
+
+First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt the excerpt with notated expressions of either articulation or dynamics (*Figure \ref{fig:stimuli}, B-C*).
+As soon as they had produced the excerpt with notated expressions without pitch errors twice consecutively, the test trials began.
+
+In the teaching condition, participants were instructed to play the excerpt of music as if they were teaching it to students.
+It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with notated expressions by listening to the participant's performance.
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience.
+Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
+The order of the conditions was counterbalanced across participants.
+The order of the expressive techniques within each condition was also counterbalanced across participants.
+A leading metronome (80 beats per minute, 8 beats) indicated the target tempo before each trial.
+
+At the end of the experiment, participants filled in a questionnaire asking about their demographic information and experience in piano performance/teaching.
+
 ### 2.1.4. Design and data analysis
+
+The experiment had two factors; Condition (teaching vs.
+performing) and Expressive Technique (articulation vs.
+dynamics).
+Additionally, the factor Subcomponent was used when analysing data for each subcomponent for each technique separately (i.e., LS; legato vs.
+staccato, FP; forte vs.
+piano).
+
+Three dependent variables were computed for data analysis.
+Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a tempo of tempo.
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and measures the smoothness of musical sequences (CITE).
+A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
+Tone intensity is assessed by key velocity (KV) and measures the loudness of a musical note.
+The value of KV in MIDI varies between 0 (minimum) and 127 (maximum).
+
+Pitch errors were identified by comparing the participants' performance with the sheet music. (TBC)
+
+```{r stim-1, out.width = "100%", fig.cap = "\\label{fig:stimuli}(A)Non-expressive sheet music.
+(B)Articulation.
+The curved line (slur) indicates legato and the dots indicate staccato.
+(C)Dynamics.
+The symbol `f' denotes forte and the symbol `p' denotes piano.
+For data analysis, only the 16th notes were included.", echo = FALSE}
+stim_a <- image_border(image_read(here("paper/image/sheetmusic-1", "stim_a.png")), "#FFFFFF", "120x120")
+stim_d <- image_border(image_read(here("paper/image/sheetmusic-1", "stim_d.png")), "#FFFFFF", "120x120")
+stim_n <- image_border(image_read(here("paper/image/sheetmusic-1", "stim_n.png")), "#FFFFFF", "120x120")
+
+# Combine
+img_stimuli <- image_append(c(stim_n, stim_a, stim_d), stack = TRUE)
+img_stimuli <- image_annotate(img_stimuli, "(A)", size = 80, location = "+100+10", color = "black", font = "helvetica")
+img_stimuli <- image_annotate(img_stimuli, "(B)", size = 80, location = "+100+660", color = "black", font = "helvetica")
+img_stimuli <- image_annotate(img_stimuli, "(C)", size = 80, location = "+100+1310", color = "black", font = "helvetica")
+img_stimuli
+```
 
@@
 ```{r questionnaire-2, include = FALSE}
-#source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
 ```
@@
 ### 3.1.1. Participants
+
+We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying piano performance at a music school at the time of recruitment.
+One participant was excluded due to insufficient motor skills.
+Twenty participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in data analysis.
+They had `r round(mean(dt_2_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_2_included$PianoTotalPractice), 2)`).
+
 ### 3.1.2. Apparatus and stimuli
+
+The same apparatus as Experiment 1 was used.
+We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expression (i.e., articulation, dynamics) and relatively simple in terms of motor skills.
+The first 12 measures of the original piece and modified it so that it had the most equal number of data points for each dependent variable (see details in *[Data analysis](#analysis-2)*).
+The stimulus consisted of a 12-measure isochronous melody notated in 4/4 meter.
+Only the right hand was used to play.
+Non-expressive sheet was used for the purpose of practice (*Figure \ref{fig:stimuli-2}, A*).
+Expressive notations were added to the non-expressive sheet music for the experiment (*Figure \ref{fig:stimuli-2}, B-C*).
+These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
+The fingering was also assigned and confirmed by the same doctoral student.
+
 ### 3.1.3. Procedure
+
 ### 3.1.4. Design and data analysis
+
+```{r stim-2, out.width = "100%", fig.cap = "\\label{fig:stimuli-2}(A)Non-expressive sheet music.
+(B)Articulation.
+The curved line (slur) indicates legato and the dots indicate staccato.
+(C)Dynamics.
+The symbol `f' denotes forte and the symbol `p' denotes piano.
+For data analysis, only the 8th notes were included.", echo = FALSE}
+stim_a_2 <- image_border(image_read(here("paper/image/sheetmusic-2", "stim_a.png")), "#FFFFFF", "120x120")
+stim_d_2 <- image_border(image_read(here("paper/image/sheetmusic-2", "stim_d.png")), "#FFFFFF", "120x120")
+stim_n_2 <- image_border(image_read(here("paper/image/sheetmusic-2", "stim_n.png")), "#FFFFFF", "120x120")
+
+# Combine
+img_stimuli_2 <- image_append(c(stim_n_2, stim_a_2, stim_d_2), stack = TRUE)
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(A)", size = 80, location = "+100+20", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(B)", size = 80, location = "+100+1075", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(C)", size = 80, location = "+100+2100", color = "black", font = "helvetica")
+img_stimuli_2
+```
 
@@
 ## 3.4. Discussion
+
+\newpage
 

### 2020-12-10 paper/manuscript.tex
add methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment.
+For data analysis, we excluded 5 participants due to an experimental error (3), an overall deviated tempo (2).
+Thirty-one participants (15 female) were included in data analysis.
+They had 12.45 years of practice on average (\emph{SD} = 5.66).
+All participants gave their informed consent before the experiment started and received vouchers for their participation.
+The study was approved by the United Ethical REview Committee for Research in Psychology (EPKEB) in Hungary.
+
 \hypertarget{apparatus-and-stimuli}{%
 \subsubsection{2.1.2. Apparatus and stimuli}\label{apparatus-and-stimuli}}
+
+A weighted Yamaha MIDI digital piano was used to record participants' performance via Max/MSP on a MacBook Pro with MacOS X Mojave 10.14.3.
+The laptop and piano were connected to a high-fidelity soundcard (Focusrite Scarlett 6i6) to deliver a metronome and piano sound.
+All auditory feedback was given to participants through headphones (Audio-Technica ATH-M50X).
+Sheet music was displayed on a computer monitor in front of the participants.
+The pitch, onset and offset time of each note, and key velocity profiles were obtained from MIDI data using Max/MSP patchers.
+
+One musical excerpt was used as a stimulus.
+The excerpt was taken from ``A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Brunam and modified for the experiment.
+It consisted of a 6-measure isochronous melody noted in 4/4 meter.
+The stimulus was composed in C major to be played with the right hand only.
+Non-expressive sheet music (i.e., sheet music without musical expression; see \emph{Figure \ref{fig:stimuli}, A}) was used for the purpose of practice.
+Expressive notations were added to the non-expressive sheet music for the experiment.
+They referred to either articulation or dynamics (\emph{Figure \ref{fig:stimuli}, B-C}).
+Articulation was notated as either legato or staccato.
+Legato indicates that musical notes are to be connected and should sound smooth.
+Staccato requires producing musical notes with shortened duration, keeping them separate from each other.
+Dynamics was notated in terms of forte and piano.
+Forte indicates that musical notes should be played loudly whereas piano indicates that musical notes should be played softly.
+The notation did not include any indication of fingering (i.e., the positioning of the fingers when playing the piano) because the piece was simple and pilot testing had shown that specifying fingering was not required.
 
@@
 
+In order to ensure that participants had sufficient motor skills to perform the piece, we asked them to practise the piece with the non-expressive sheet music (\emph{Figure \ref{fig:stimuli}, A}).
+Once they could perform the piece without pitch errors twice consecutively, the experiment started.
+
+First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt the excerpt with notated expressions of either articulation or dynamics (\emph{Figure \ref{fig:stimuli}, B-C}).
+As soon as they had produced the excerpt with notated expressions without pitch errors twice consecutively, the test trials began.
+
+In the teaching condition, participants were instructed to play the excerpt of music as if they were teaching it to students.
+It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with notated expressions by listening to the participant's performance.
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience.
+Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
+The order of the conditions was counterbalanced across participants.
+The order of the expressive techniques within each condition was also counterbalanced across participants.
+A leading metronome (80 beats per minute, 8 beats) indicated the target tempo before each trial.
+
+At the end of the experiment, participants filled in a questionnaire asking about their demographic information and experience in piano performance/teaching.
+
 \hypertarget{design-and-data-analysis}{%
 \subsubsection{2.1.4. Design and data analysis}\label{design-and-data-analysis}}
+
+The experiment had two factors; Condition (teaching vs.~performing) and Expressive Technique (articulation vs.~dynamics).
+Additionally, the factor Subcomponent was used when analysing data for each subcomponent for each technique separately (i.e., LS; legato vs.~staccato, FP; forte vs.~piano).
+
+Three dependent variables were computed for data analysis.
+Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a tempo of tempo.
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and measures the smoothness of musical sequences (CITE).
+A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
+Tone intensity is assessed by key velocity (KV) and measures the loudness of a musical note.
+The value of KV in MIDI varies between 0 (minimum) and 127 (maximum).
+
+Pitch errors were identified by comparing the participants' performance with the sheet music. (TBC)
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stim-1-1} \caption{\label{fig:stimuli}(A)Non-expressive sheet music.
+(B)Articulation.
+The curved line (slur) indicates legato and the dots indicate staccato.
+(C)Dynamics.
+The symbol `f' denotes forte and the symbol `p' denotes piano.
+For data analysis, only the 16th notes were included.}\label{fig:stim-1}
+\end{figure}
 
@@
 
+We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying piano performance at a music school at the time of recruitment.
+One participant was excluded due to insufficient motor skills.
+Twenty participants (9 female) were included in data analysis.
+They had 15.65 years of practice on average (\emph{SD} = 5.67).
+
 \hypertarget{apparatus-and-stimuli-1}{%
 \subsubsection{3.1.2. Apparatus and stimuli}\label{apparatus-and-stimuli-1}}
+
+The same apparatus as Experiment 1 was used.
+We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expression (i.e., articulation, dynamics) and relatively simple in terms of motor skills.
+The first 12 measures of the original piece and modified it so that it had the most equal number of data points for each dependent variable (see details in \emph{\protect\hyperlink{analysis-2}{Data analysis}}).
+The stimulus consisted of a 12-measure isochronous melody notated in 4/4 meter.
+Only the right hand was used to play.
+Non-expressive sheet was used for the purpose of practice (\emph{Figure \ref{fig:stimuli-2}, A}).
+Expressive notations were added to the non-expressive sheet music for the experiment (\emph{Figure \ref{fig:stimuli-2}, B-C}).
+These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
+The fingering was also assigned and confirmed by the same doctoral student.
 
@@
 \subsubsection{3.1.4. Design and data analysis}\label{design-and-data-analysis-1}}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stim-2-1} \caption{\label{fig:stimuli-2}(A)Non-expressive sheet music.
+(B)Articulation.
+The curved line (slur) indicates legato and the dots indicate staccato.
+(C)Dynamics.
+The symbol `f' denotes forte and the symbol `p' denotes piano.
+For data analysis, only the 16th notes were included.}\label{fig:stim-2}
+\end{figure}
 
@@
 
+\newpage
+
 \hypertarget{general-discussion}{%

### 2020-12-11 experiment-1/analysis/preprocessor/error_summary.Rmd
remove here()
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 analysis/preprocessor/error_summary.Rmd
@@
 if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("here")) {install.packages("here"); require("here")}
-
-# set working directory
-here::i_am("error_summary.Rmd")
 
 # function
-source(here("function.R"))
+source("function.R")
 
 # read a text file for ideal performance
-dt_ideal <- read.table(here("ideal.txt"))
+dt_ideal <- read.table("ideal.txt")
 colnames(dt_ideal) <- "Pitch"
@@
 # 
-dt_onset <- fread(file = here("filtered", "dt_correct_onset.txt"), sep = ",")
-dt_offset <- fread(file = here("filtered", "dt_correct_offset.txt"), sep = ",")
-error_onset <- fread(file = here("filtered", "error_onset.txt"), sep = ",")
-error_offset <- fread(file = here("filtered", "error_offset.txt"), sep = ",")
+dt_onset <- fread(file = "filtered/dt_correct_onset.txt", sep = ",")
+dt_offset <- fread(file = "filtered/dt_correct_offset.txt", sep = ",")
+error_onset <- fread(file = "filtered/error_onset.txt", sep = ",")
+error_offset <- fread(file = "filtered/error_offset.txt", sep = ",")
 ```
@@
 ```
+
+```{r export, include = FALSE}
+knitr::purl("error_summary.Rmd")
+```

### 2020-12-11 experiment-1/analysis/preprocessor/error_summary.R

@@ -1,51 +1,47 @@
-## ----setup, include = FALSE--------------------------------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 # plot
 if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("here")) {install.packages("here"); require("here")}
-
-# set working directory
-here::i_am("error_summary.Rmd")
 
 # function
-source(here("function.R"))
+source("function.R")
 
 # read a text file for ideal performance
-dt_ideal <- read.table(here("ideal.txt"))
+dt_ideal <- read.table("ideal.txt")
 colnames(dt_ideal) <- "Pitch"
 dt_ideal$RowNr <- 1:nrow(dt_ideal)
 setcolorder(dt_ideal, c(2, 1))
 # 
-dt_onset <- fread(file = here("filtered", "dt_correct_onset.txt"), sep = ",")
-dt_offset <- fread(file = here("filtered", "dt_correct_offset.txt"), sep = ",")
-error_onset <- fread(file = here("filtered", "error_onset.txt"), sep = ",")
-error_offset <- fread(file = here("filtered", "error_offset.txt"), sep = ",")
+dt_onset <- fread(file = "filtered/dt_correct_onset.txt", sep = ",")
+dt_offset <- fread(file = "filtered/dt_correct_offset.txt", sep = ",")
+error_onset <- fread(file = "filtered/error_onset.txt", sep = ",")
+error_offset <- fread(file = "filtered/error_offset.txt", sep = ",")
 
 
-## ----onset-all, echo = FALSE-------------------------------------------------------
+## ----onset-all, echo = FALSE--------------------------------------------------
 error_onset_all <- error_onset[CorrectionNr != "Exclude"]
 error_onset_all
 
 
-## ----onset-summary, echo = FALSE---------------------------------------------------
+## ----onset-summary, echo = FALSE----------------------------------------------
 error_onset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
 error_onset_summary <- error_onset_all[, .(N = .N, Sum = sum(as.numeric(CorrectionNr))), by = .(Reason)]
 error_onset_summary
 
 
-## ----na-onset-all, echo = FALSE----------------------------------------------------
+## ----na-onset-all, echo = FALSE-----------------------------------------------
 missing_onset <- checker(dt_onset, dt_ideal)
 missing_onset
 
 
[…]

### 2020-12-11 paper/manuscript.Rmd
add methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-bibliography      : [""]
+bibliography      : ["references.bib"]
 
@@
 ---
-```{r setup, include = FALSE}
+```{r set-up, include = FALSE}
 if (!require("papaja")) {install.packages("papaja"); require("papaja")}
@@
 ## 2.1. Method
-```{r questionnaire-1, include = FALSE}
+```{r methods-1, include = FALSE}
 source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+source(here("experiment-1/analysis/preprocessor", "error_summary.R"), chdir = TRUE)
 ```
@@
 We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 5 participants due to an experimental error (3), an overall deviated tempo (2).
-Thirty-one participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis.
+For data analysis, we excluded 5 participants due to experimental errors (3; performances were not recorded for one or more blocks, or participants played the piece in an octave higher), an overall deviated tempo (2).
+As a result, 31 participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis.
+Most participants were right handed (left; `r nrow(dt_included[Handedness == "Left"])`, both; `r nrow(dt_included[Handedness == "Both"])`).
 They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`).
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
-The study was approved by the United Ethical REview Committee for Research in Psychology (EPKEB) in Hungary.
+The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.
 
@@
 One musical excerpt was used as a stimulus.
-The excerpt was taken from ``A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Brunam and modified for the experiment.
+The excerpt was taken from ``A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Burnam and modified for the experiment.
 It consisted of a 6-measure isochronous melody noted in 4/4 meter.
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without musical expression; see *Figure \ref{fig:stimuli}, A*) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations; see *Figure \ref{fig:stimuli}, A*) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
@@
 In order to ensure that participants had sufficient motor skills to perform the piece, we asked them to practise the piece with the non-expressive sheet music (*Figure \ref{fig:stimuli}, A*).
-Once they could perform the piece without pitch errors twice consecutively, the experiment started.
+Once they could perform the piece without pitch errors twice consecutively within 5 attempts, the experiment started.
 
@@
 Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a tempo of tempo.
-Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and measures the smoothness of musical sequences (CITE).
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and measures the smoothness of musical sequences [@bresin_2000].
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
@@
 
-Pitch errors were identified by comparing the participants' performance with the sheet music. (TBC)
+Data cleaning, preprocessing and statistical analysis were performed in R version 3.6.3.
+Pitch errors were identified separately for onsets and offsets by comparing the participants' performance with the sheet music.
+Three trials were entirely excluded from data analysis because participants did not follow sheet music or complete performing till the end.
+We manually removed accidental pitch errors produced by participants.
+Pitch errors included either extra, missing or substituted tones.
+For onsets, `r round(sum(error_onset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_onset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_onset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_onset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
+For offsets, `r round(sum(error_offset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
+After we removed pitch errors, outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and were excluded from data analysis.
+Removed outliers were less than 5% of overall responses for each dependent variable.
+
+We performed separate analyses for the two expressive techniques (i.e., articulation, dynamics).
+Paired-sample *t*-tests were performed to compare the mean IOIs in the teaching and performing condition.
+For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.
+performing) and Subcomponent (LS; legato vs.
+staccato or FP; forte vs.
+piano, respectively).
+The ``aov_car" function in the afex package was used for a repeated-measures ANOVA.
+For post-hoc comparisons, we used emmeans package to investigate interactions fomr the ANOVA.
 
@@
 
+\newpage
+
 ## 2.2. Results (Articulation)
+
 ### 2.2.1. Interonset Intervals (IOIs)
@@
 ## 3.1. Method
-```{r questionnaire-2, include = FALSE}
+```{r methods-2, include = FALSE}
 source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+source(here("experiment-2/analysis/preprocessor", "error_summary.R"), chdir = TRUE)
 ```
@@
 We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying piano performance at a music school at the time of recruitment.
-One participant was excluded due to insufficient motor skills.
+One participant was excluded due to insufficient motor skills (see details in *[Procedure](#procedure-2)*).
 Twenty participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in data analysis.
+Most participants were right handed (left; `r nrow(dt_2_included[Handedness == "Left"])`).
 They had `r round(mean(dt_2_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_2_included$PianoTotalPractice), 2)`).
@@
 The same apparatus as Experiment 1 was used.
-We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expression (i.e., articulation, dynamics) and relatively simple in terms of motor skills.
-The first 12 measures of the original piece and modified it so that it had the most equal number of data points for each dependent variable (see details in *[Data analysis](#analysis-2)*).
+We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expressions (i.e., articulation, dynamics) and was relatively simple in terms of motor skills.
+The first 12 measures of the original piece was used and modified so that it had the almost equal number of data points for each dependent variable (see details in *[Data analysis](#analysis-2)*).
 The stimulus consisted of a 12-measure isochronous melody notated in 4/4 meter.
-Only the right hand was used to play.
 Non-expressive sheet was used for the purpose of practice (*Figure \ref{fig:stimuli-2}, A*).
@@
 
-### 3.1.3. Procedure
+### 3.1.3. Procedure{#procedure-2}
+
+We employed the similar procedure as Experiment 1.
+Unlike Experiment 1, all participants were required to memorise the piece without expressive notations (i.e., *Figure \ref{fig:stimuli-2}, A*) prior to the experiment.
+Those who could not perform the piece twice consecutively within 5 attempts were considered not to have sufficient motor skills to implement musical expressions and excluded from data analysis.
+Also, we modified the wording of instructions for the teaching and performing condition so that both had the same focus on expressive notations (see details in *[Supplemental Materials](#supplemental)*).
 
 ### 3.1.4. Design and data analysis
+
+Data cleaning, preprocessing and statistical analysis were identical as Experiment 1.
+For statistical analysis, only 8th notes with expressive notations were included.
+As a result, only one 8th note in the 4th measure without any expression was not included to calculate dependent variables.
+Three trials were entirely excluded from data analysis because participants did not follow sheet music.
+With the same method as Experiment 1, pitch errors were removed manually.
+For onsets, `r round(sum(error_onset_summary_2$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_onset_summary_2[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_onset_summary_2[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_onset_summary_2[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
+For offsets, `r round(sum(error_offset_summary_2$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_offset_summary_2[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_offset_summary_2[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_offset_summary_2[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
+Removed outliers (i.e., responses outside 3 standard deviations from the mean) were less than 5% of overall responses for each dependent variable.
 

### 2020-12-11 paper/manuscript.tex
add methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \newenvironment{cslreferences}%
-  {}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
   {\par}
@@
 We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 5 participants due to an experimental error (3), an overall deviated tempo (2).
-Thirty-one participants (15 female) were included in data analysis.
+For data analysis, we excluded 5 participants due to experimental errors (3; performances were not recorded for one or more blocks, or participants played the piece in an octave higher), an overall deviated tempo (2).
+As a result, 31 participants (15 female) were included in data analysis.
+Most participants were right handed (left; 2, both; 2).
 They had 12.45 years of practice on average (\emph{SD} = 5.66).
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
-The study was approved by the United Ethical REview Committee for Research in Psychology (EPKEB) in Hungary.
+The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.
 
@@
 One musical excerpt was used as a stimulus.
-The excerpt was taken from ``A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Brunam and modified for the experiment.
+The excerpt was taken from ``A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Burnam and modified for the experiment.
 It consisted of a 6-measure isochronous melody noted in 4/4 meter.
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without musical expression; see \emph{Figure \ref{fig:stimuli}, A}) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations; see \emph{Figure \ref{fig:stimuli}, A}) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
@@
 In order to ensure that participants had sufficient motor skills to perform the piece, we asked them to practise the piece with the non-expressive sheet music (\emph{Figure \ref{fig:stimuli}, A}).
-Once they could perform the piece without pitch errors twice consecutively, the experiment started.
+Once they could perform the piece without pitch errors twice consecutively within 5 attempts, the experiment started.
 
@@
 Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a tempo of tempo.
-Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and measures the smoothness of musical sequences (CITE).
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and measures the smoothness of musical sequences (Bresin \& Battel, 2000).
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
@@
 
-Pitch errors were identified by comparing the participants' performance with the sheet music. (TBC)
+Data cleaning, preprocessing and statistical analysis were performed in R version 3.6.3.
+Pitch errors were identified separately for onsets and offsets by comparing the participants' performance with the sheet music.
+Three trials were entirely excluded from data analysis because participants did not follow sheet music or complete performing till the end.
+We manually removed accidental pitch errors produced by participants.
+Pitch errors included either extra, missing or substituted tones.
+For onsets, 5.16\% of the trials contained at least one pitch error (extra notes; 2.78\%, missing notes; 2.24\%, substituted notes; 0.13\%).
+For offsets, 6.24\% of the trials contained at least one pitch error (extra notes; 2.78\%, missing notes; 2.29\%, substituted notes; 1.17\%).
+After we removed pitch errors, outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and were excluded from data analysis.
+Removed outliers were less than 5\% of overall responses for each dependent variable.
+
+We performed separate analyses for the two expressive techniques (i.e., articulation, dynamics).
+Paired-sample \emph{t}-tests were performed to compare the mean IOIs in the teaching and performing condition.
+For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.~performing) and Subcomponent (LS; legato vs.~staccato or FP; forte vs.~piano, respectively).
+The ``aov\_car" function in the afex package was used for a repeated-measures ANOVA.
+For post-hoc comparisons, we used emmeans package to investigate interactions fomr the ANOVA.
 
@@
 \end{figure}
+
+\newpage
 
@@
 We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying piano performance at a music school at the time of recruitment.
-One participant was excluded due to insufficient motor skills.
+One participant was excluded due to insufficient motor skills (see details in \emph{\protect\hyperlink{procedure-2}{Procedure}}).
 Twenty participants (9 female) were included in data analysis.
+Most participants were right handed (left; 2).
 They had 15.65 years of practice on average (\emph{SD} = 5.67).
@@
 The same apparatus as Experiment 1 was used.
-We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expression (i.e., articulation, dynamics) and relatively simple in terms of motor skills.
-The first 12 measures of the original piece and modified it so that it had the most equal number of data points for each dependent variable (see details in \emph{\protect\hyperlink{analysis-2}{Data analysis}}).
+We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expressions (i.e., articulation, dynamics) and was relatively simple in terms of motor skills.
+The first 12 measures of the original piece was used and modified so that it had the almost equal number of data points for each dependent variable (see details in \emph{\protect\hyperlink{analysis-2}{Data analysis}}).
 The stimulus consisted of a 12-measure isochronous melody notated in 4/4 meter.
-Only the right hand was used to play.
 Non-expressive sheet was used for the purpose of practice (\emph{Figure \ref{fig:stimuli-2}, A}).
@@
 
-\hypertarget{procedure-1}{%
-\subsubsection{3.1.3. Procedure}\label{procedure-1}}
+\hypertarget{procedure-2}{%
+\subsubsection{3.1.3. Procedure}\label{procedure-2}}
+
+We employed the similar procedure as Experiment 1.
+Unlike Experiment 1, all participants were required to memorise the piece without expressive notations (i.e., \emph{Figure \ref{fig:stimuli-2}, A}) prior to the experiment.
+Those who could not perform the piece twice consecutively within 5 attempts were considered not to have sufficient motor skills to implement musical expressions and excluded from data analysis.
+Also, we modified the wording of instructions for the teaching and performing condition so that both had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplemental}{Supplemental Materials}}).
 
@@
 \subsubsection{3.1.4. Design and data analysis}\label{design-and-data-analysis-1}}
+
+Data cleaning, preprocessing and statistical analysis were identical as Experiment 1.
+For statistical analysis, only 8th notes with expressive notations were included.
+As a result, only one 8th note in the 4th measure without any expression was not included to calculate dependent variables.
+Three trials were entirely excluded from data analysis because participants did not follow sheet music.
+With the same method as Experiment 1, pitch errors were removed manually.
+For onsets, 3.32\% of the trials contained at least one pitch error (extra notes; 1.66\%, missing notes; 1.57\%, substituted notes; 0.09\%).
+For offsets, 5.11\% of the trials contained at least one pitch error (extra notes; 1.66\%, missing notes; 1.57\%, substituted notes; 1.88\%).
+Removed outliers (i.e., responses outside 3 standard deviations from the mean) were less than 5\% of overall responses for each dependent variable.
 
@@
 The symbol `f' denotes forte and the symbol `p' denotes piano.
-For data analysis, only the 16th notes were included.}\label{fig:stim-2}
+For data analysis, only the 8th notes were included.}\label{fig:stim-2}
 \end{figure}
@@
 \begin{cslreferences}
+\leavevmode\hypertarget{ref-bresin_2000}{}%
+Bresin, R., \& Battel, G.
+U.
+(2000).
+Articulation Strategies in Expressive Piano Performance Analysis of Legato, Staccato, and Repeated Notes in Performances of the Andante Movement of Mozart's Sonata in G Major (K 545).
+\emph{Journal of New Music Research}, \emph{29}(3), 211--224.
+\url{https://doi.org/10.1076/jnmr.29.3.211.3092}
 \end{cslreferences}

### 2020-12-14 experiment-1/analysis/stats/ioi.R
export as R scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE--------------------------------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -8,11 +8,11 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 
 
-## ----file, include = FALSE---------------------------------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE------------------------------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -24,25 +24,25 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE---------------------------------------------------------
+## ----ioi-art, echo = FALSE----------------------------------------------------
 # For each individual
 ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE----------------------------------------------------
+## ----ioi-art-box,  echo = FALSE-----------------------------------------------
 ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + 
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   geom_hline(yintercept = 188, linetype = "dashed")
 
 
[…]

### 2020-12-14 experiment-2/analysis/stats/articulation.R

@@ -1,0 +1,368 @@
+## ----setup, include = FALSE------------------------------------------------------
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# statistics
+if (!require("afex")) {install.packages("afex"); require("afex")}
+if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
+
+
+## ----file, include = FALSE-------------------------------------------------------
+filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
+filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
+
+
+## ----extract, include = FALSE----------------------------------------------------
+dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
+dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
+
+# SubNr as a factor
+dt_kor$SubNr <- as.factor(dt_kor$SubNr)
+dt_kot$SubNr <- as.factor(dt_kot$SubNr)
+
+# Include only articulation
+dt_kor_art <- dt_kor[Skill == "articulation"]
+dt_kot_art <- dt_kot[Skill == "articulation"]
+
+# Include only dynamics
+dt_kor_dyn <- dt_kor[Skill == "dynamics"]
+dt_kot_dyn <- dt_kot[Skill == "dynamics"]
+
+
+## ----kor-art, echo = FALSE-------------------------------------------------------
+# For each individual
+kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
+kor_art
+
+
+## ----kor-art-box,  echo = FALSE--------------------------------------------------
+ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
[…]

### 2020-12-14 experiment-2/analysis/stats/dynamics.R

@@ -1,0 +1,391 @@
+## ----setup, include = FALSE------------------------------------------------------
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# statistics
+if (!require("afex")) {install.packages("afex"); require("afex")}
+if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
+
+
+## ----file, include = FALSE-------------------------------------------------------
+filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
+filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
+
+
+## ----extract, include = FALSE----------------------------------------------------
+dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
+dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
+
+# SubNr as a factor
+dt_vel$SubNr <- as.factor(dt_vel$SubNr)
+dt_vel_diff$SubNr <- as.factor(dt_vel_diff$SubNr)
+
+# Include only articulation
+dt_vel_art <- dt_vel[Skill == "articulation"]
+dt_vel_diff_art <- dt_vel_diff[Skill == "articulation"]
+
+# Include only dynamics
+dt_vel_dyn <- dt_vel[Skill == "dynamics"]
+dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
+
+
+## ----vel-dyn, echo = FALSE-------------------------------------------------------
+# For each individual
+vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
+vel_dyn
+
+
+## ----vel-dyn-box,  echo = FALSE--------------------------------------------------
[…]

### 2020-12-14 experiment-2/analysis/stats/ioi.R

@@ -1,0 +1,271 @@
+## ----setup, include = FALSE------------------------------------------------------
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# statistics
+if (!require("afex")) {install.packages("afex"); require("afex")}
+
+
+## ----file, include = FALSE-------------------------------------------------------
+filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
+
+
+## ----extract, include = FALSE----------------------------------------------------
+dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
+
+# SubNr as a factor
+dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
+
+# Include only articulation
+dt_ioi_art <- dt_ioi[Skill == "articulation"]
+# Include only dynamics
+dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
+
+
+## ----ioi-art, echo = FALSE-------------------------------------------------------
+# For each individual
+ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
+ioi_art
+
+
+## ----ioi-art-box,  echo = FALSE--------------------------------------------------
+ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + 
+  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+  geom_hline(yintercept = 1, linetype = "dashed")
+
+
+## ----ioi-art-all, echo = FALSE---------------------------------------------------
+# Group mean
+ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
[…]

### 2020-12-14 paper/manuscript.Rmd
completed methods (for now)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 footnotelist      : no
-linenumbers       : yes
+linenumbers       : no
 mask              : no
 draft             : no
-noextraspace      : yes
 
@@
 We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 5 participants due to experimental errors (3; performances were not recorded for one or more blocks, or participants played the piece in an octave higher), an overall deviated tempo (2).
-As a result, 31 participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis.
-Most participants were right handed (left; `r nrow(dt_included[Handedness == "Left"])`, both; `r nrow(dt_included[Handedness == "Both"])`).
+For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2).
+Thirty-one participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis.
+Most participants were right-handed (left; `r nrow(dt_included[Handedness == "Left"])`, both; `r nrow(dt_included[Handedness == "Both"])`).
 They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`).
@@
 Staccato requires producing musical notes with shortened duration, keeping them separate from each other.
-Dynamics was notated in terms of forte and piano.
+Dynamics was notated as either forte and piano.
 Forte indicates that musical notes should be played loudly whereas piano indicates that musical notes should be played softly.
@@
 
-First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt the excerpt with notated expressions of either articulation or dynamics (*Figure \ref{fig:stimuli}, B-C*).
+First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with notated expressions of either articulation or dynamics (*Figure \ref{fig:stimuli}, B-C*).
 As soon as they had produced the excerpt with notated expressions without pitch errors twice consecutively, the test trials began.
@@
 In the teaching condition, participants were instructed to play the excerpt of music as if they were teaching it to students.
-It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with notated expressions by listening to the participant's performance.
+It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with a notated expression by listening to the participant's performance.
 In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience.
@@
 Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a tempo of tempo.
-Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and measures the smoothness of musical sequences [@bresin_2000].
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone, and measures the smoothness of musical sequences [@bresin_2000].
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
@@
 Data cleaning, preprocessing and statistical analysis were performed in R version 3.6.3.
-Pitch errors were identified separately for onsets and offsets by comparing the participants' performance with the sheet music.
+For statistical analysis, only 16th notes with expressive notations were included.
 Three trials were entirely excluded from data analysis because participants did not follow sheet music or complete performing till the end.
-We manually removed accidental pitch errors produced by participants.
-Pitch errors included either extra, missing or substituted tones.
+Pitch errors were identified by comparing a sequence of musical notes produced by a participant with the sequence of musical notes according to the sheet music.
+Pitch errors included either extra, missing or substituted tones and were manually removed by using editData package.
 For onsets, `r round(sum(error_onset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_onset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_onset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_onset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
 For offsets, `r round(sum(error_offset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
-After we removed pitch errors, outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and were excluded from data analysis.
+After we removed pitch errors, outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
 Removed outliers were less than 5% of overall responses for each dependent variable.
@@
 We performed separate analyses for the two expressive techniques (i.e., articulation, dynamics).
-Paired-sample *t*-tests were performed to compare the mean IOIs in the teaching and performing condition.
+A paired-sample *t*-test or a Wilcoxon signed-rank tests (a non-parametric alternative to a paired *t*-test) was performed to compare the mean IOIs in the teaching and performing condition.
 For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.
@@
 The ``aov_car" function in the afex package was used for a repeated-measures ANOVA.
-For post-hoc comparisons, we used emmeans package to investigate interactions fomr the ANOVA.
+For post-hoc comparisons, we used emmeans package to investigate interactions from the ANOVA.
 
@@
 
+```{r results-1, include = FALSE}
+source(here("experiment-1/analysis/stats", "ioi.R"), chdir = TRUE)
+source(here("experiment-1/analysis/stats", "articulation.R"), chdir = TRUE)
+source(here("experiment-1/analysis/stats", "dynamics.R"), chdir = TRUE)
+```
+
+All effects are reported as significant at *p* < .05.
+We first report results for performance the piece with the notated expression of articulation (*Figure \ref{fig:stimuli}, B*).
+followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli}, C*).
+
 ### 2.2.1. Interonset Intervals (IOIs)
+
+In order to compare the mean IOIs between the teaching and performing condition, we used a Wiscoxon Signed-rank test, instead of a paired *t*-test, because the Shapiro-Wilk test showed that the difference...
+was significantly different from the normal distribution.
+
+The Wisconxin Signed-rank test revealed....
+
+```{r ioi-art-1, echo = FALSE}
+ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean)
+
+ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+  geom_hline(yintercept = 188, linetype = "dashed") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+```
+
 ### 2.2.2. Key-Overlap Time (KOT)
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-LS (legato vs.
+staccato) revealed that there was a significant main effect of Subcomponent-LS (F) and a significant interation betweenC Condition and Subcomponent-LS (F).
+Post-hoc comparisons showed that participants produced shorter staccato in the teaching condition than in the performing condition (t) whereas there was no significant difference in legato between the teaching and performing condition (t).
+
+We also calculated KOR...
+
 ### 2.2.3. Key Velocity (KV)
+
 ### 2.2.4. KV Transition Points
@@
 Twenty participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in data analysis.
-Most participants were right handed (left; `r nrow(dt_2_included[Handedness == "Left"])`).
+Most participants were right-handed (left; `r nrow(dt_2_included[Handedness == "Left"])`).
 They had `r round(mean(dt_2_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_2_included$PianoTotalPractice), 2)`).
@@
 We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expressions (i.e., articulation, dynamics) and was relatively simple in terms of motor skills.
-The first 12 measures of the original piece was used and modified so that it had the almost equal number of data points for each dependent variable (see details in *[Data analysis](#analysis-2)*).
+The first 12 measures of the original piece was used and modified so that it had the almost equal number of data points for each dependent variable.
 The stimulus consisted of a 12-measure isochronous melody notated in 4/4 meter.
@@
 Unlike Experiment 1, all participants were required to memorise the piece without expressive notations (i.e., *Figure \ref{fig:stimuli-2}, A*) prior to the experiment.
-Those who could not perform the piece twice consecutively within 5 attempts were considered not to have sufficient motor skills to implement musical expressions and excluded from data analysis.
 Also, we modified the wording of instructions for the teaching and performing condition so that both had the same focus on expressive notations (see details in *[Supplemental Materials](#supplemental)*).
+As Experiment 1, we asked participants to perform the piece without looking at the non-expressive sheet music in order to make sure that participants memorise the piece and had sufficient motor skills and memorised.
+Those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis.
 
@@
 img_stimuli_2 <- image_append(c(stim_n_2, stim_a_2, stim_d_2), stack = TRUE)
-img_stimuli_2 <- image_annotate(img_stimuli_2, "(A)", size = 80, location = "+100+20", color = "black", font = "helvetica")
-img_stimuli_2 <- image_annotate(img_stimuli_2, "(B)", size = 80, location = "+100+1075", color = "black", font = "helvetica")
-img_stimuli_2 <- image_annotate(img_stimuli_2, "(C)", size = 80, location = "+100+2100", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(A)", size = 80, location = "+100+30", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(B)", size = 80, location = "+100+1085", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(C)", size = 80, location = "+100+2110", color = "black", font = "helvetica")
 img_stimuli_2

### 2020-12-14 paper/manuscript.tex
completed methods (for now)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \keywords{teaching, expertise, skill transmission, artistic expression, music\newline\indent Word count: X}
-\usepackage{lineno}
-
-\linenumbers
 \usepackage{csquotes}
@@
 We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 5 participants due to experimental errors (3; performances were not recorded for one or more blocks, or participants played the piece in an octave higher), an overall deviated tempo (2).
+For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2).
 As a result, 31 participants (15 female) were included in data analysis.
@@
 
-First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt the excerpt with notated expressions of either articulation or dynamics (\emph{Figure \ref{fig:stimuli}, B-C}).
+First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with notated expressions of either articulation or dynamics (\emph{Figure \ref{fig:stimuli}, B-C}).
 As soon as they had produced the excerpt with notated expressions without pitch errors twice consecutively, the test trials began.
@@
 Data cleaning, preprocessing and statistical analysis were performed in R version 3.6.3.
-Pitch errors were identified separately for onsets and offsets by comparing the participants' performance with the sheet music.
 Three trials were entirely excluded from data analysis because participants did not follow sheet music or complete performing till the end.
-We manually removed accidental pitch errors produced by participants.
-Pitch errors included either extra, missing or substituted tones.
+Pitch errors were identified by comparing a sequence of musical notes produced by a participant with the sequence of musical notes according to the sheet music.
+Pitch errors included either extra, missing or substituted tones and were manually removed.
 For onsets, 5.16\% of the trials contained at least one pitch error (extra notes; 2.78\%, missing notes; 2.24\%, substituted notes; 0.13\%).
 For offsets, 6.24\% of the trials contained at least one pitch error (extra notes; 2.78\%, missing notes; 2.29\%, substituted notes; 1.17\%).
-After we removed pitch errors, outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and were excluded from data analysis.
+After we removed pitch errors, outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
 Removed outliers were less than 5\% of overall responses for each dependent variable.
@@
 
+All effects are reported as significant at \emph{p} \textless{} .05.
+We first report results for performance the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli}, B}).
+followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli}, C}).
+
 \hypertarget{interonset-intervals-iois}{%
@@
 
+In order to compare the mean IOIs between the teaching and performing condition, we used a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because the Shapiro-Wilk test showed that the difference\ldots{} was significantly different from the normal distribution.
+
+The Wisconxin Signed-rank test revealed\ldots.
+
+\includegraphics{manuscript_files/figure-latex/ioi-art-1-1.pdf}
+
 \hypertarget{key-overlap-time-kot}{%
 \subsubsection{2.2.2. Key-Overlap Time (KOT)}\label{key-overlap-time-kot}}
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (F) and a significant interation betweenC Condition and Subcomponent-LS (F).
+Post-hoc comparisons showed that participants produced shorter staccato in the teaching condition than in the performing condition (t) whereas there was no significant difference in legato between the teaching and performing condition (t).
+
+We also calculated KOR\ldots{}
 

### 2020-12-14 experiment-1/analysis/stats/ioi.Rmd
add median/IQR
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 script/R/analysis/ioi.Rmd
@@
 # Group mean
-ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
+ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Condition, Skill)]
 ioi_art_all
@@
 # Group mean
-ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean),  SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
+ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Condition, Skill)]
 ioi_dyn_all
@@
 ```
+
+```{r export, include = FALSE}
+knitr::purl("ioi.Rmd")
+```

### 2020-12-14 experiment-1/analysis/stats/ioi.R

@@ -38,7 +38,7 @@ ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, 
 
 ## ----ioi-art-all, echo = FALSE------------------------------------------------
 # Group mean
-ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
+ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Condition, Skill)]
 ioi_art_all
 
 
@@ -144,7 +144,7 @@ ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, 
 
 ## ----ioi-dyn-all, echo = FALSE------------------------------------------------
 # Group mean
-ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean),  SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
+ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Condition, Skill)]
 ioi_dyn_all
 
 

### 2020-12-14 paper/manuscript.Rmd
add some methods and ioi analysis for exp1 and 2
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with a notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience.
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplemental Materials](#supplemental)*).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
@@
 img_stimuli <- image_append(c(stim_n, stim_a, stim_d), stack = TRUE)
-img_stimuli <- image_annotate(img_stimuli, "(A)", size = 80, location = "+100+10", color = "black", font = "helvetica")
-img_stimuli <- image_annotate(img_stimuli, "(B)", size = 80, location = "+100+660", color = "black", font = "helvetica")
-img_stimuli <- image_annotate(img_stimuli, "(C)", size = 80, location = "+100+1310", color = "black", font = "helvetica")
+img_stimuli <- image_annotate(img_stimuli, "(A)", size = 80, location = "+100+0", color = "black", font = "helvetica")
+img_stimuli <- image_annotate(img_stimuli, "(B)", size = 80, location = "+100+650", color = "black", font = "helvetica")
+img_stimuli <- image_annotate(img_stimuli, "(C)", size = 80, location = "+100+1300", color = "black", font = "helvetica")
 img_stimuli
@@
 
-In order to compare the mean IOIs between the teaching and performing condition, we used a Wiscoxon Signed-rank test, instead of a paired *t*-test, because the Shapiro-Wilk test showed that the difference...
-was significantly different from the normal distribution.
+In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution.
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
 
-The Wisconxin Signed-rank test revealed....
-
-```{r ioi-art-1, echo = FALSE}
+```{r ioi-art-1, include = FALSE}
 ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean)
 
-ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ioi_art_1 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+  ylim(150, 260) +
   geom_hline(yintercept = 188, linetype = "dashed") +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 200, paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_art_1.png"), plot = ioi_art_1, dpi = 600, width = 5, height = 5)
+img_ioi_art_1 <- image_read(here("paper/image", "ioi_art_1.png"))
 ```
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) revealed that there was a significant main effect of Subcomponent-LS (F) and a significant interation betweenC Condition and Subcomponent-LS (F).
+staccato) revealed that there was a significant main effect of Subcomponent-LS (F) and a significant interaction between Condition and Subcomponent-LS (F).
 Post-hoc comparisons showed that participants produced shorter staccato in the teaching condition than in the performing condition (t) whereas there was no significant difference in legato between the teaching and performing condition (t).
@@
 ### 2.3.1. Interonset Intervals (IOIs)
+
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
+
+```{r ioi-dyn-1, include = FALSE}
+ioi_dyn_plot <- data.table(subject = ioi_dyn[Condition == "teaching"]$SubNr, teaching = ioi_dyn[Condition == "teaching"]$Mean, performing = ioi_dyn[Condition == "performing"]$Mean)
+
+ioi_dyn_1 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+  ylim(150, 260) +
+  geom_hline(yintercept = 188, linetype = "dashed") +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 200, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_dyn_1.png"), plot = ioi_dyn_1, dpi = 600, width = 5, height = 5)
+img_ioi_dyn_1 <- image_read(here("paper/image", "ioi_dyn_1.png"))
+```
+
 ### 2.3.2. Key Velocity (KV)
@@
 ### 2.3.4. Key-Overlap Time (KOT)
+
+```{r plot-ioi-1, out.width = "100%", fig.cap = "\\label{fig:ioi-1}Experiment 1: IOIs (ms) when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+# combine
+img_ioi_1 <- image_append(c(img_ioi_art_1, img_ioi_dyn_1))
+image_write(img_ioi_1, path = here("paper/image", "img_ioi_1.png"), format = "png")
+img_ioi_1
+```
 
@@
 
-We employed the similar procedure as Experiment 1.
-Unlike Experiment 1, all participants were required to memorise the piece without expressive notations (i.e., *Figure \ref{fig:stimuli-2}, A*) prior to the experiment.
-Also, we modified the wording of instructions for the teaching and performing condition so that both had the same focus on expressive notations (see details in *[Supplemental Materials](#supplemental)*).
-As Experiment 1, we asked participants to perform the piece without looking at the non-expressive sheet music in order to make sure that participants memorise the piece and had sufficient motor skills and memorised.
-Those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis.
+We employed the same procedure as Experiment 1 with several modifications.
+First, all participants were required to memorise the piece without expressive notations (i.e., *Figure \ref{fig:stimuli-2}, A*) prior to the experiment.
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[Supplemental Materials](#supplemental)*).
+Third, participants could choose their preferred tempo from one of the 3 options (100, 110 and 120 beats per minute) as a leading metronome.
+In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music.
+As Experiment 1, those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis.
+The rest of the procedure was identical to Experiment 1.
 
@@
 As a result, only one 8th note in the 4th measure without any expression was not included to calculate dependent variables.
+Also, IOIs were normalised by their preferred tempo because each participant chose different tempi from the 3 options.
 Three trials were entirely excluded from data analysis because participants did not follow sheet music.
@@
 img_stimuli_2 <- image_append(c(stim_n_2, stim_a_2, stim_d_2), stack = TRUE)
-img_stimuli_2 <- image_annotate(img_stimuli_2, "(A)", size = 80, location = "+100+30", color = "black", font = "helvetica")
-img_stimuli_2 <- image_annotate(img_stimuli_2, "(B)", size = 80, location = "+100+1085", color = "black", font = "helvetica")
-img_stimuli_2 <- image_annotate(img_stimuli_2, "(C)", size = 80, location = "+100+2110", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(A)", size = 80, location = "+100+70", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(B)", size = 80, location = "+100+1125", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(C)", size = 80, location = "+100+2150", color = "black", font = "helvetica")
 img_stimuli_2
@@
 ## 3.2. Results (Articulation)
+
+```{r results-2, include = FALSE}
+source(here("experiment-2/analysis/stats", "ioi.R"), chdir = TRUE)
+source(here("experiment-2/analysis/stats", "articulation.R"), chdir = TRUE)
+source(here("experiment-2/analysis/stats", "dynamics.R"), chdir = TRUE)
+```
+
+As Experiment 1, we first report results for performance the piece with the notated expression of articulation (*Figure \ref{fig:stimuli-2}, B*).
+followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli-2}, C*).
+
 ### 3.2.1. Interonset Intervals (IOIs)
+
+A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`
+, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
+A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`).
+
+```{r ioi-art-2, echo = FALSE}
+ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean)
+
+ioi_art_2 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+  ylim(0.8, 1.1) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 1.05, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_art_2.png"), plot = ioi_art_2, dpi = 600, width = 5, height = 5)
+img_ioi_art_2 <- image_read(here("paper/image", "ioi_art_2.png"))
+```
+
 ### 3.2.2. Key-Overlap Time (KOT)
@@
 ### 3.3.1. Interonset Intervals (IOIs)
+
+A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with notated articulation (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
+A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_wilcoxon$p.value))}`).
+
+```{r ioi-dyn-2, echo = FALSE}
+ioi_dyn_plot <- data.table(subject = ioi_dyn[Condition == "teaching"]$SubNr, teaching = ioi_dyn[Condition == "teaching"]$Mean, performing = ioi_dyn[Condition == "performing"]$Mean)
+
+ioi_dyn_2 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+  ylim(0.8, 1.1) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 1.05, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_dyn_2.png"), plot = ioi_dyn_2, dpi = 600, width = 5, height = 5)
+img_ioi_dyn_2 <- image_read(here("paper/image", "ioi_dyn_2.png"))
+```
+
+```{r plot-ioi-2, out.width = "100%", fig.cap = "\\label{fig:ioi-2}Experiment 2: Normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+img_ioi_2 <- image_append(c(img_ioi_art_2, img_ioi_dyn_2))
+image_write(img_ioi_2, path = here("paper/image", "img_ioi_2.png"), format = "png")
+img_ioi_2
+```
+
 ### 3.3.2. Key Velocity (KV)

### 2020-12-14 paper/manuscript.tex
add some methods and ioi analysis for exp1 and 2
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2).
-As a result, 31 participants (15 female) were included in data analysis.
-Most participants were right handed (left; 2, both; 2).
+Thirty-one participants (15 female) were included in data analysis.
+Most participants were right-handed (left; 2, both; 2).
 They had 12.45 years of practice on average (\emph{SD} = 5.66).
@@
 In the teaching condition, participants were instructed to play the excerpt of music as if they were teaching it to students.
-It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with notated expressions by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience.
+It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with a notated expression by listening to the participant's performance.
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplemental}{Supplemental Materials}}).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
@@
 Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a tempo of tempo.
-Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and measures the smoothness of musical sequences (Bresin \& Battel, 2000).
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone, and measures the smoothness of musical sequences (Bresin \& Battel, 2000).
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
@@
 Data cleaning, preprocessing and statistical analysis were performed in R version 3.6.3.
+For statistical analysis, only 16th notes with expressive notations were included.
 Three trials were entirely excluded from data analysis because participants did not follow sheet music or complete performing till the end.
 Pitch errors were identified by comparing a sequence of musical notes produced by a participant with the sequence of musical notes according to the sheet music.
-Pitch errors included either extra, missing or substituted tones and were manually removed.
+Pitch errors included either extra, missing or substituted tones and were manually removed by using editData package.
 For onsets, 5.16\% of the trials contained at least one pitch error (extra notes; 2.78\%, missing notes; 2.24\%, substituted notes; 0.13\%).
@@
 We performed separate analyses for the two expressive techniques (i.e., articulation, dynamics).
-Paired-sample \emph{t}-tests were performed to compare the mean IOIs in the teaching and performing condition.
+A paired-sample \emph{t}-test or a Wilcoxon signed-rank tests (a non-parametric alternative to a paired \emph{t}-test) was performed to compare the mean IOIs in the teaching and performing condition.
 For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.~performing) and Subcomponent (LS; legato vs.~staccato or FP; forte vs.~piano, respectively).
 The ``aov\_car" function in the afex package was used for a repeated-measures ANOVA.
-For post-hoc comparisons, we used emmeans package to investigate interactions fomr the ANOVA.
+For post-hoc comparisons, we used emmeans package to investigate interactions from the ANOVA.
 
@@
 
-In order to compare the mean IOIs between the teaching and performing condition, we used a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because the Shapiro-Wilk test showed that the difference\ldots{} was significantly different from the normal distribution.
-
-The Wisconxin Signed-rank test revealed\ldots.
-
-\includegraphics{manuscript_files/figure-latex/ioi-art-1-1.pdf}
+In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution.
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.21{]} while playing the piece with notated articulation (\emph{p} = 0.017; \emph{Figure \ref{fig:ioi-1}}).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (F) and a significant interation betweenC Condition and Subcomponent-LS (F).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (F) and a significant interaction between Condition and Subcomponent-LS (F).
 Post-hoc comparisons showed that participants produced shorter staccato in the teaching condition than in the performing condition (t) whereas there was no significant difference in legato between the teaching and performing condition (t).
@@
 
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.68 (ms), \emph{IQR} = 21.99{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 20.04{]} while playing the piece with notated dynamics (\emph{p} = 0.120; \emph{Figure \ref{fig:ioi-1}}).
+
 \hypertarget{key-velocity-kv-1}{%
@@
 \subsubsection{2.3.4. Key-Overlap Time (KOT)}\label{key-overlap-time-kot-1}}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-1-1} \caption{\label{fig:ioi-1}Experiment 1: IOIs (ms) when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-ioi-1}
+\end{figure}
 
@@
 Twenty participants (9 female) were included in data analysis.
-Most participants were right handed (left; 2).
+Most participants were right-handed (left; 2).
 They had 15.65 years of practice on average (\emph{SD} = 5.67).
@@
 We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expressions (i.e., articulation, dynamics) and was relatively simple in terms of motor skills.
-The first 12 measures of the original piece was used and modified so that it had the almost equal number of data points for each dependent variable (see details in \emph{\protect\hyperlink{analysis-2}{Data analysis}}).
+The first 12 measures of the original piece was used and modified so that it had the almost equal number of data points for each dependent variable.
 The stimulus consisted of a 12-measure isochronous melody notated in 4/4 meter.
@@
 
-We employed the similar procedure as Experiment 1.
-Unlike Experiment 1, all participants were required to memorise the piece without expressive notations (i.e., \emph{Figure \ref{fig:stimuli-2}, A}) prior to the experiment.
-Those who could not perform the piece twice consecutively within 5 attempts were considered not to have sufficient motor skills to implement musical expressions and excluded from data analysis.
-Also, we modified the wording of instructions for the teaching and performing condition so that both had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplemental}{Supplemental Materials}}).
+We employed the same procedure as Experiment 1 with several modifications.
+First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Figure \ref{fig:stimuli-2}, A}) prior to the experiment.
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplemental}{Supplemental Materials}}).
+Third, participants could choose their preferred tempo from one of the 3 options (100, 110 and 120 beats per minute) as a leading metronome.
+In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music.
+As Experiment 1, those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis.
+The rest of the procedure was identical to Experiment 1.
 
@@
 
+As Experiment 1, we first report results for performance the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli-2}, B}).
+followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli-2}, C}).
+
 \hypertarget{interonset-intervals-iois-2}{%
 \subsubsection{3.2.1. Interonset Intervals (IOIs)}\label{interonset-intervals-iois-2}}
+
+A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.966, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.953, \emph{SD} = 0.040{]} while playing the piece with notated articulation (\emph{t}(19) = 2.47
+, \emph{p} = 0.023; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.033).
 
@@
 \subsubsection{3.3.1. Interonset Intervals (IOIs)}\label{interonset-intervals-iois-3}}
+
+A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.957, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.956, \emph{SD} = 0.043{]} while playing the piece with notated articulation (\emph{t}(19) = 0.21, \emph{p} = 0.837; \emph{Figure \ref{fig:ioi-2}}).
+A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (\emph{p} = 0.956).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Experiment 2: Normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-ioi-2}
+\end{figure}
 

### 2020-12-14 experiment-2/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE------------------------------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE-------------------------------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE----------------------------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,321 +31,321 @@ dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 
 
-## ----kor-art, echo = FALSE-------------------------------------------------------
+## ----kor-art, echo = FALSE----------------------------------------------------
 # For each individual
 kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kor_art
 
 
-## ----kor-art-box,  echo = FALSE--------------------------------------------------
+## ----kor-art-box,  echo = FALSE-----------------------------------------------
 ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
   facet_grid(Subcomponent ~ .) +
[…]

### 2020-12-14 paper/manuscript.Rmd
add previous intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # 1. Introduction
+
+Deliberate teaching has supported human skill transmission over generations and provides a key route for learning [@tomasello_2016; @whiten_2017].
+Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
+In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners [@brand_2002; @fernald_1985].
+These modulations include slowing down and exaggerating sounds and actions.
+Similar findings were obtained in studies with adult learners, where exaggerations were observed when native British English speakers were talking to first language English learners [@uther_2007] and when skilled adults were teaching xylophone melodies to novices [@mcellin_2017].
+The observed modulations are thought to provide communicative signals that can facilitate learning by affecting attention and memory [@csibra_2009].
+
+Previous research has shown that demonstrators not only adjust their gestures [@campisi_2013] and actions [@fukuyama_2015] to learners' skills but engage in specific action modulations to highlight particular aspects of demonstrated actions.
+For example, @schaik_2019 showed that adults used specific action modulations for demonstrating different action effects of objects to infants.
+@NIPS2016_6413 found that demonstrators chose costly movement paths at structurally important points to disambiguate one goal over other possibilities.
+
+Importantly, in some domains teaching requires not only demonstrating *what* to do, but *how* to perform actions.
+In artistic contexts, novices need to learn how exactly to perform actions by relying on particular techniques [@sloboda_2000].
+For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques [@juslin_2003].
+This raises the question of whether and how experts modulate their actions during demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
+Studying action modulations that serve the teaching of artistic techniques is of broader theoretical importance: It can show whether action modulations can be used to selectively highlight subtle aspects of actions that define expert performance.
+
+In the present study, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, articulation, loudness) can be modulated at the same time.
+Specific and subtle aspects of the performance need to be highlighted to teach a respective technique.
+We asked whether skilled pianists use sound to communicate didactic intentions.
+Pianists were asked to perform a piece in order to demonstrate to a learner how to implement the notated expressions (teaching condition) and to play the same piece for an audience (performing condition).
+Two basic expressive techniques in piano performance were used: articulation (legato and staccato) and dynamics (forte and piano).
+
+If experts rely on generic action modulations, it can be expected that they will play more slowly during demonstration, regardless of the kind of expressive techniques to be taught.
+To the extent that experts use action modulations to support their teaching of specific expressive techniques, they should produce longer legato and shorter staccato when teaching articulation whereas they should produce louder forte and softer piano when teaching dynamics.
+Furthermore, one could speculate that they might produce modulations specifically at structurally important points that best highlight the technique to be taught.
+Importantly, they should avoid modulating irrelevant properties of expression, e.g., the smoothness of the sound while teaching dynamics.
+
+\clearpage
 

### 2020-12-14 paper/manuscript.tex
add previous intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \section{1. Introduction}\label{introduction}}
+
+Deliberate teaching has supported human skill transmission over generations and provides a key route for learning (Tomasello, 2016; Whiten, 2017).
+Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
+In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners (Brand, Baldwin, \& Ashburn, 2002; Fernald, 1985).
+These modulations include slowing down and exaggerating sounds and actions.
+Similar findings were obtained in studies with adult learners, where exaggerations were observed when native British English speakers were talking to first language English learners (Uther, Knoll, \& Burnham, 2007) and when skilled adults were teaching xylophone melodies to novices (McEllin, Knoblich, \& Sebanz, 2017).
+The observed modulations are thought to provide communicative signals that can facilitate learning by affecting attention and memory (Csibra \& Gergely, 2009).
+
+Previous research has shown that demonstrators not only adjust their gestures (Campisi \& Özyürek, 2013) and actions (Fukuyama et al., 2015) to learners' skills but engage in specific action modulations to highlight particular aspects of demonstrated actions.
+For example, Schaik, Meyer, Ham, and Hunnius (2019) showed that adults used specific action modulations for demonstrating different action effects of objects to infants.
+Ho, Littman, MacGlashan, Cushman, and Austerweil (2016) found that demonstrators chose costly movement paths at structurally important points to disambiguate one goal over other possibilities.
+
+Importantly, in some domains teaching requires not only demonstrating \emph{what} to do, but \emph{how} to perform actions.
+In artistic contexts, novices need to learn how exactly to perform actions by relying on particular techniques (Sloboda, 2000).
+For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques (Juslin, 2003).
+This raises the question of whether and how experts modulate their actions during demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
+Studying action modulations that serve the teaching of artistic techniques is of broader theoretical importance: It can show whether action modulations can be used to selectively highlight subtle aspects of actions that define expert performance.
+
+In the present study, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, articulation, loudness) can be modulated at the same time.
+Specific and subtle aspects of the performance need to be highlighted to teach a respective technique.
+We asked whether skilled pianists use sound to communicate didactic intentions.
+Pianists were asked to perform a piece in order to demonstrate to a learner how to implement the notated expressions (teaching condition) and to play the same piece for an audience (performing condition).
+Two basic expressive techniques in piano performance were used: articulation (legato and staccato) and dynamics (forte and piano).
+
+If experts rely on generic action modulations, it can be expected that they will play more slowly during demonstration, regardless of the kind of expressive techniques to be taught.
+To the extent that experts use action modulations to support their teaching of specific expressive techniques, they should produce longer legato and shorter staccato when teaching articulation whereas they should produce louder forte and softer piano when teaching dynamics.
+Furthermore, one could speculate that they might produce modulations specifically at structurally important points that best highlight the technique to be taught.
+Importantly, they should avoid modulating irrelevant properties of expression, e.g., the smoothness of the sound while teaching dynamics.
+
+\clearpage
 
@@
 \begin{cslreferences}
+\leavevmode\hypertarget{ref-brand_2002}{}%
+Brand, R.
+J., Baldwin, D.
+A., \& Ashburn, L.
+A.
+(2002).
+Evidence for ``motionese'': Modifications in mothers' infant-directed action.
+\emph{Developmental Science}, \emph{5}(1), 72--83.
+\url{https://doi.org/10.1111/1467-7687.00211}
+
 \leavevmode\hypertarget{ref-bresin_2000}{}%
@@
 \url{https://doi.org/10.1076/jnmr.29.3.211.3092}
+
+\leavevmode\hypertarget{ref-campisi_2013}{}%
+Campisi, E., \& Özyürek, A.
+(2013).
+Iconicity as a communicative strategy: Recipient design in multimodal demonstrations for adults and children.
+\emph{Journal of Pragmatics}, \emph{47}(1), 14--27.
+\url{https://doi.org/10.1016/j.pragma.2012.12.007}
+
+\leavevmode\hypertarget{ref-csibra_2009}{}%
+Csibra, G., \& Gergely, G. (2009). Natural pedagogy. \emph{Trends in Cognitive Sciences}, \emph{13}(4), 148--153. \url{https://doi.org/10.1016/j.tics.2009.01.005}
+
+\leavevmode\hypertarget{ref-fernald_1985}{}%
+Fernald, A. (1985). Four-month-old infants prefer to listen to motherese. \emph{Infant Behavior and Development}, \emph{8}(2), 181--195. \url{https://doi.org/10.1016/S0163-6383(85)80005-9}
+
+\leavevmode\hypertarget{ref-fukuyama_2015}{}%
+Fukuyama, H., Qin, S., Kanakogi, Y., Nagai, Y., Asada, M., \& Myowa-Yamakoshi, M.
+(2015).
+Infant's action skill dynamically modulates parental action demonstration in the dyadic interaction.
+\emph{Developmental Science}, \emph{18}(6), 1006--1013.
+\url{https://doi.org/10.1111/desc.12270}
+
+\leavevmode\hypertarget{ref-NIPS2016_6413}{}%
+Ho, M.
+K., Littman, M., MacGlashan, J., Cushman, F., \& Austerweil, J.
+L.
+(2016).
+Showing versus doing: Teaching by demonstration.
+In D.
+D.
+Lee, M.
+Sugiyama, U.
+V.
+Luxburg, I.
+Guyon, \& R.
+Garnett (Eds.), \emph{Advances in neural information processing systems 29} (pp.
+3027--3035).
+Curran Associates, Inc.
+
+\leavevmode\hypertarget{ref-juslin_2003}{}%
+Juslin, P.
+N.
+(2003).
+Five Facets of Musical Expression: A Psychologist's Perspective on Music Performance.
+\emph{Psychology of Music}, \emph{31}(3), 273--302.
+\url{https://doi.org/10.1177/03057356030313003}
+
+\leavevmode\hypertarget{ref-mcellin_2017}{}%
+McEllin, L., Knoblich, G., \& Sebanz, N.
+(2017).
+Distinct kinematic markers of demonstration and joint action coordination?
+Evidence from virtual xylophone playing.
+\emph{Journal of Experimental Psychology: Human Perception and Performance}, \emph{44}(6), 885.
+\url{https://doi.org/10.1037/xhp0000505}
+
+\leavevmode\hypertarget{ref-schaik_2019}{}%
+Schaik, J.
+E.
+van, Meyer, M., Ham, C.
+R.
+van, \& Hunnius, S.
+(2019).
+Motion tracking of parents' infant- versus adult-directed actions reveals general and action-specific modulations.
+\emph{Developmental Science}, \emph{0}(0), e12869.
+\url{https://doi.org/10.1111/desc.12869}
+
+\leavevmode\hypertarget{ref-sloboda_2000}{}%
+Sloboda, J. A. (2000). Individual differences in music performance. \emph{Trends in Cognitive Sciences}, \emph{4}(10), 397--403. \url{https://doi.org/10.1016/S1364-6613(00)01531-X}
+
+\leavevmode\hypertarget{ref-tomasello_2016}{}%
+Tomasello, M. (2016). The ontogeny of cultural learning. \emph{Current Opinion in Psychology}, \emph{8}, 1--4. \url{https://doi.org/10.1016/j.copsyc.2015.09.008}
+
+\leavevmode\hypertarget{ref-uther_2007}{}%
+Uther, M., Knoll, M.
+A., \& Burnham, D.
+(2007).
+Do you speak E-NG-L-I-SH?
+A comparison of foreigner- and infant-directed speech.
+\emph{Speech Communication}, \emph{49}(1), 2--7.
+\url{https://doi.org/10.1016/j.specom.2006.10.003}
+
+\leavevmode\hypertarget{ref-whiten_2017}{}%
+Whiten, A. (2017). Social Learning and Culture in Child and Chimpanzee. \emph{Annual Review of Psychology}, \emph{68}(1), 129--154. \url{https://doi.org/10.1146/annurev-psych-010416-044108}
 \end{cslreferences}

### 2020-12-14 experiment-1/analysis/preprocessor/included_summary.Rmd
update included_summary
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("here")) {install.packages("here"); require("here")}
-
-# set working directory
-here::i_am("included_summary.Rmd")
 
 # read txt files
-dt_ioi <- fread(file = here("trimmed", "data_ioi_1.txt"), sep = ",")
-dt_kot <- fread(file = here("trimmed", "data_kot.txt"), sep = ",")
-dt_vel <- fread(file = here("trimmed", "data_vel.txt"), sep = ",")
-dt_vel_diff <- fread(file = here("trimmed", "data_vel_diff.txt"), sep = ",")
+dt_ioi <- fread(file = "./trimmed/data_ioi_1.txt", sep = ",")
+dt_kot <- fread(file = "./trimmed/data_kot.txt", sep = ",")
+dt_vel <- fread(file = "./trimmed/data_vel.txt", sep = ",")
+dt_vel_diff <- fread(file = "./trimmed/data_vel_diff.txt", sep = ",")
 ```
@@
 ```
-- SubNr 1 is excluded from data analysis because most of the responses were excluded due to outliers.
 
@@
 ## Excluded participnats
-- SubNr 1: Most of responses (more than 90%) were removed due to outliers
+- SubNr 1: A deviated tempo
 - SubNr 3: Experimental error

### 2020-12-14 experiment-1/analysis/preprocessor/included_summary.R

@@ -1,60 +1,56 @@
-## ----setup, include = FALSE---------------------------------------------------------
+## ----setup, include = FALSE-------------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 # plot
 if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("here")) {install.packages("here"); require("here")}
-
-# set working directory
-here::i_am("included_summary.Rmd")
 
 # read txt files
-dt_ioi <- fread(file = here("trimmed", "data_ioi_1.txt"), sep = ",")
-dt_kot <- fread(file = here("trimmed", "data_kot.txt"), sep = ",")
-dt_vel <- fread(file = here("trimmed", "data_vel.txt"), sep = ",")
-dt_vel_diff <- fread(file = here("trimmed", "data_vel_diff.txt"), sep = ",")
+dt_ioi <- fread(file = "./trimmed/data_ioi_1.txt", sep = ",")
+dt_kot <- fread(file = "./trimmed/data_kot.txt", sep = ",")
+dt_vel <- fread(file = "./trimmed/data_vel.txt", sep = ",")
+dt_vel_diff <- fread(file = "./trimmed/data_vel_diff.txt", sep = ",")
 
 
-## ----ioi, echo = FALSE--------------------------------------------------------------
+## ----ioi, echo = FALSE------------------------------------------------------------
 dt_ioi
 
 
-## ----ioi-summary, echo = FALSE------------------------------------------------------
+## ----ioi-summary, echo = FALSE----------------------------------------------------
 dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
 
 
-## ----kot, echo = FALSE--------------------------------------------------------------
+## ----kot, echo = FALSE------------------------------------------------------------
 dt_kot[SubNr != 1]
 
 
-## ----kot-summary, echo = FALSE------------------------------------------------------
+## ----kot-summary, echo = FALSE----------------------------------------------------
 dt_kot[SubNr != 1, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
 
 
-## ----vel, echo = FALSE--------------------------------------------------------------
[…]

### 2020-12-14 experiment-1/analysis/preprocessor/included_summary.Rmd
update included_summary
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r export-csv, echo = TRUE}
-fwrite(dt_ioi[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_ioi.txt"), sep = ",")
-fwrite(dt_kot[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_kot.txt"), sep = ",")
-fwrite(dt_vel[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_vel.txt"), sep = ",")
-fwrite(dt_vel_diff[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_vel_diff.txt"), sep = ",")
+fwrite(dt_ioi[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./trimmed/data_analysis_ioi.txt", sep = ",")
+fwrite(dt_kot[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./trimmed/data_analysis_kot.txt", sep = ",")
+fwrite(dt_vel[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./trimmed/data_analysis_vel.txt", sep = ",")
+fwrite(dt_vel_diff[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./trimmed/data_analysis_vel_diff.txt", sep = ",")
 ```

### 2020-12-14 experiment-1/analysis/preprocessor/included_summary.R

@@ -45,10 +45,10 @@ dt_vel_diff[SubNr != 1, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubN
 
 
 ## ----export-csv, echo = TRUE------------------------------------------------------
-fwrite(dt_ioi[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_ioi.txt"), sep = ",")
-fwrite(dt_kot[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_kot.txt"), sep = ",")
-fwrite(dt_vel[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_vel.txt"), sep = ",")
-fwrite(dt_vel_diff[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = here("trimmed", "data_analysis_vel_diff.txt"), sep = ",")
+fwrite(dt_ioi[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./trimmed/data_analysis_ioi.txt", sep = ",")
+fwrite(dt_kot[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./trimmed/data_analysis_kot.txt", sep = ",")
+fwrite(dt_vel[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./trimmed/data_analysis_vel.txt", sep = ",")
+fwrite(dt_vel_diff[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16], file = "./trimmed/data_analysis_vel_diff.txt", sep = ",")
 
 
 ## ----export, include = FALSE------------------------------------------------------

### 2020-12-14 experiment-1/analysis/stats/articulation.Rmd
update analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 analysis/stats/articulation.Rmd
@@
 
-- SubNr 1: Most of responses (more than 90%) were removed due to outliers
+- SubNr 1: A deviated tempo
 - SubNr 3: Experimental error
@@
 ```
+
+```{r export, include = FALSE}
+knitr::purl("articulation.Rmd")
+```

### 2020-12-14 experiment-1/analysis/stats/dynamics.Rmd
update analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```
+
+```{r export, include = FALSE}
+knitr::purl("dynamics.Rmd")
+```

### 2020-12-14 experiment-1/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE------------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE-------------------------------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE----------------------------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,30 +31,30 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE----------------------------------------------------
+## ----vel-dyn, echo = FALSE-------------------------------------------------------
 # For each individual
 vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE-----------------------------------------------
+## ----vel-dyn-box,  echo = FALSE--------------------------------------------------
[…]

### 2020-12-15 experiment-2/analysis/preprocessor/included_summary.Rmd
update included_summary
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 analysis/preprocessor/included_summary.Rmd
@@
 if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-if (!require("here")) {install.packages("here"); require("here")}
-
-# set working directory
-here::i_am("included_summary.Rmd")
 
 # read txt files
-dt_ioi <- fread(file = here("trimmed", "data_ioi_1.txt"), sep = ",")
-dt_kor <- fread(file = here("trimmed", "data_kor.txt"), sep = ",")
-dt_kot <- fread(file = here("trimmed", "data_kot.txt"), sep = ",")
-dt_vel <- fread(file = here("trimmed", "data_vel.txt"), sep = ",")
-dt_vel_diff <- fread(file = here("trimmed", "data_vel_diff.txt"), sep = ",")
+dt_ioi <- fread(file = "./trimmed/data_ioi_1.txt", sep = ",")
+dt_kor <- fread(file = "./trimmed/data_kor.txt", sep = ",")
+dt_kot <- fread(file = "./trimmed/data_kot.txt", sep = ",")
+dt_vel <- fread(file = "./trimmed/data_vel.txt", sep = ",")
+dt_vel_diff <- fread(file = "./trimmed/data_vel_diff.txt", sep = ",")
 ```
@@
 - Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
-- Note: SubNr 3 was excluded due to insufficient motor skills
+- Note: SubNr 3 was excluded due to insufficient motor skills. SubNr 2 was excluded due to a deviated tempo.
 
@@
 ```{r export, echo = TRUE}
-fwrite(dt_ioi[SubNr != 3], file = here("trimmed", "data_analysis_ioi.txt"), sep = ",")
-fwrite(dt_kor[SubNr != 3], file = here("trimmed", "data_analysis_kor.txt"), sep = ",")
-fwrite(dt_kot[SubNr != 3], file = here("trimmed", "data_analysis_kot.txt"), sep = ",")
-fwrite(dt_vel[SubNr != 3], file = here("trimmed", "data_analysis_vel.txt"), sep = ",")
-fwrite(dt_vel_diff[SubNr != 3], file = here("trimmed", "data_analysis_vel_diff.txt"), sep = ",")
+fwrite(dt_ioi[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_ioi.txt", sep = ",")
+fwrite(dt_kor[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_kor.txt", sep = ",")
+fwrite(dt_kot[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_kot.txt", sep = ",")
+fwrite(dt_vel[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_vel.txt", sep = ",")
+fwrite(dt_vel_diff[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_vel_diff.txt", sep = ",")
 ```

### 2020-12-15 experiment-2/analysis/stats/articulation.R
update analysis without subnr 2
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE------------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE-------------------------------------------------------
 filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE----------------------------------------------------
 dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,321 +31,321 @@ dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 
 
-## ----kor-art, echo = FALSE----------------------------------------------------
+## ----kor-art, echo = FALSE-------------------------------------------------------
 # For each individual
 kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kor_art
 
 
-## ----kor-art-box,  echo = FALSE-----------------------------------------------
+## ----kor-art-box,  echo = FALSE--------------------------------------------------
 ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
   facet_grid(Subcomponent ~ .) +
[…]

### 2020-12-15 experiment-2/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE------------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -8,11 +8,11 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE-------------------------------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE----------------------------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -24,25 +24,25 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE----------------------------------------------------
+## ----ioi-art, echo = FALSE-------------------------------------------------------
 # For each individual
 ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE-----------------------------------------------
+## ----ioi-art-box,  echo = FALSE--------------------------------------------------
 ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + 
   theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   geom_hline(yintercept = 1, linetype = "dashed")
 
 
[…]

### 2020-12-15 paper/manuscript.Rmd
add some figs
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
-
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-
-  <!-- https://tinyurl.com/ybremelq -->
+  aaaa
 
@@
 We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2).
+For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2; outside 2 standard deviations from the average tempo across participants).
 Thirty-one participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis.
@@
 
-# Combine
+# combine
 img_stimuli <- image_append(c(stim_n, stim_a, stim_d), stack = TRUE)
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-LS (legato vs.
-staccato) revealed that there was a significant main effect of Subcomponent-LS (F) and a significant interaction between Condition and Subcomponent-LS (F).
-Post-hoc comparisons showed that participants produced shorter staccato in the teaching condition than in the performing condition (t) whereas there was no significant difference in legato between the teaching and performing condition (t).
+Blah
 
-We also calculated KOR...
+```{r kot-art-1, include = FALSE}
+kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent)
+
+kot_art_1 <- ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation") +
+  ylim(-170, 105) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -50, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "kot_art_1.png"), plot = kot_art_1, dpi = 600, width = 7, height = 5)
+img_kot_art_1 <- image_read(here("paper/image", "kot_art_1.png"))
+```
 
@@
 
+Blah
+
 ### 2.2.4. KV Transition Points
+
+Blah
 
@@
 ### 2.3.2. Key Velocity (KV)
+
+Blah
+
 ### 2.3.3. KV Transition Points
+
+Blah
+
 ### 2.3.4. Key-Overlap Time (KOT)
+
+Blah
+
+```{r kot-dyn-1, include = FALSE}
+kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent)
+
+kot_dyn_1 <- ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics") +
+  ylim(-170, 105) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 0, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "kot_dyn_1.png"), plot = kot_dyn_1, dpi = 600, width = 7, height = 5)
+img_kot_dyn_1 <- image_read(here("paper/image", "kot_dyn_1.png"))
+```
 
@@
 img_ioi_1
+```
+
+```{r plot-kot-1, out.width = "100%", fig.cap = "\\label{fig:kot-1}Experiment 1: KOT (ms) when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+# combine
+img_kot_1 <- image_append(c(img_kot_art_1, img_kot_dyn_1))
+image_write(img_kot_1, path = here("paper/image", "img_kot_1.png"), format = "png")
+img_kot_1
 ```
@@
 
-# Combine
+# combine
 img_stimuli_2 <- image_append(c(stim_n_2, stim_a_2, stim_d_2), stack = TRUE)
@@
 
-### 3.2.2. Key-Overlap Time (KOT)
+### 3.2.2. Key-Overlap Ratio (KOR)
+
+Blah
+
+```{r kot-art-2, include = FALSE}
+kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent)
+
+kot_art_2 <- ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
+  ylim(-1, 0.4) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -0.5, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "kot_art_2.png"), plot = kot_art_1, dpi = 600, width = 7, height = 5)
+img_kot_art_2 <- image_read(here("paper/image", "kot_art_2.png"))
+```
+
 ### 3.2.3. Key Velocity (KV)
+
+Blah
+
 ### 3.2.4. KV Transition Points
+
+Blah
 
@@
 ### 3.3.2. Key Velocity (KV)
+
+Blah
+
 ### 3.3.3. KV Transition Points
-### 3.3.4. Key-Overlap Time (KOT)
+
+Blah
+
+### 3.3.4. Key-Overlap Ratio (KOR)
+
+Blah
+
+```{r kot-dyn-2, include = FALSE}
+kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent)
+
+kot_dyn_2 <- ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
+  ylim(-1, 0.4) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -0.5, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "kot_dyn_2.png"), plot = kot_dyn_1, dpi = 600, width = 7, height = 5)
+img_kot_dyn_2 <- image_read(here("paper/image", "kot_dyn_2.png"))
+```
 
@@
 \begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
+\setlength{\parindent}{-0in}
+\setlength{\leftskip}{0in}
 

### 2020-12-15 paper/manuscript.tex
add some figs
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \abstract{
-One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.
-
-Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.
-
-One sentence clearly stating the \textbf{general problem} being addressed by this particular study.
-
-One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).
-
-Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
-
-One or two sentences to put the results into a more \textbf{general context}.
-
-Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
+aaaa
 }
@@
 We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2).
+For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2; outside 2 standard deviations from the average tempo across participants).
 Thirty-one participants (15 female) were included in data analysis.
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution.
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.21{]} while playing the piece with notated articulation (\emph{p} = 0.017; \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with notated articulation (\emph{p} = 0.021; \emph{Figure \ref{fig:ioi-1}}).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (F) and a significant interaction between Condition and Subcomponent-LS (F).
-Post-hoc comparisons showed that participants produced shorter staccato in the teaching condition than in the performing condition (t) whereas there was no significant difference in legato between the teaching and performing condition (t).
-
-We also calculated KOR\ldots{}
+Blah
 
@@
 
+Blah
+
 \hypertarget{kv-transition-points}{%
 \subsubsection{2.2.4. KV Transition Points}\label{kv-transition-points}}
+
+Blah
 
@@
 
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.68 (ms), \emph{IQR} = 21.99{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 20.04{]} while playing the piece with notated dynamics (\emph{p} = 0.120; \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with notated dynamics (\emph{p} = 0.111; \emph{Figure \ref{fig:ioi-1}}).
 
@@
 
+Blah
+
 \hypertarget{kv-transition-points-1}{%
@@
 
+Blah
+
 \hypertarget{key-overlap-time-kot-1}{%
 \subsubsection{2.3.4. Key-Overlap Time (KOT)}\label{key-overlap-time-kot-1}}
+
+Blah
 
@@
 Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-ioi-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-1-1} \caption{\label{fig:kot-1}Experiment 1: KOT (ms) when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-kot-1}
 \end{figure}
@@
 
-A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.966, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.953, \emph{SD} = 0.040{]} while playing the piece with notated articulation (\emph{t}(19) = 2.47
-, \emph{p} = 0.023; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.033).
+A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.972, \emph{SD} = 0.042{]} than in the performing condition {[}\emph{M} = 0.958, \emph{SD} = 0.034{]} while playing the piece with notated articulation (\emph{t}(18) = 2.64
+, \emph{p} = 0.017; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.026).
 
-\hypertarget{key-overlap-time-kot-2}{%
-\subsubsection{3.2.2. Key-Overlap Time (KOT)}\label{key-overlap-time-kot-2}}
+\hypertarget{key-overlap-ratio-kor}{%
+\subsubsection{3.2.2. Key-Overlap Ratio (KOR)}\label{key-overlap-ratio-kor}}
+
+Blah
 
@@
 
+Blah
+
 \hypertarget{kv-transition-points-2}{%
 \subsubsection{3.2.4. KV Transition Points}\label{kv-transition-points-2}}
+
+Blah
 
@@
 
-A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.957, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.956, \emph{SD} = 0.043{]} while playing the piece with notated articulation (\emph{t}(19) = 0.21, \emph{p} = 0.837; \emph{Figure \ref{fig:ioi-2}}).
-A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (\emph{p} = 0.956).
+A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.962, \emph{SD} = 0.041{]} and the performing condition {[}\emph{M} = 0.961, \emph{SD} = 0.037{]} while playing the piece with notated articulation (\emph{t}(18) = 0.37, \emph{p} = 0.719; \emph{Figure \ref{fig:ioi-2}}).
+A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (\emph{p} = 0.922).
 
@@
 
+Blah
+
 \hypertarget{kv-transition-points-3}{%
@@
 
-\hypertarget{key-overlap-time-kot-3}{%
-\subsubsection{3.3.4. Key-Overlap Time (KOT)}\label{key-overlap-time-kot-3}}
+Blah
+
+\hypertarget{key-overlap-ratio-kor-1}{%
+\subsubsection{3.3.4. Key-Overlap Ratio (KOR)}\label{key-overlap-ratio-kor-1}}
+
+Blah
 

### 2020-12-15 paper/manuscript.Rmd
add kot/kor plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying piano performance at a music school at the time of recruitment.
-One participant was excluded due to insufficient motor skills (see details in *[Procedure](#procedure-2)*).
-Twenty participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in data analysis.
+For data analysis, we excluded 1 participants due to insufficient motor skills (1) and an overall deviated tempo (1).
+Nineteen participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in data analysis.
 Most participants were right-handed (left; `r nrow(dt_2_included[Handedness == "Left"])`).
@@
 
-```{r kot-art-2, include = FALSE}
-kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent)
+```{r kor-art-2, include = FALSE}
+kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent)
 
-kot_art_2 <- ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
+kor_art_2 <- ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
   ylim(-1, 0.4) +
@@
 # save plots
-ggsave(here("paper/image", "kot_art_2.png"), plot = kot_art_1, dpi = 600, width = 7, height = 5)
-img_kot_art_2 <- image_read(here("paper/image", "kot_art_2.png"))
+ggsave(here("paper/image", "kor_art_2.png"), plot = kor_art_2, dpi = 600, width = 7, height = 5)
+img_kor_art_2 <- image_read(here("paper/image", "kor_art_2.png"))
 ```
@@
 
-```{r plot-ioi-2, out.width = "100%", fig.cap = "\\label{fig:ioi-2}Experiment 2: Normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
-img_ioi_2 <- image_append(c(img_ioi_art_2, img_ioi_dyn_2))
-image_write(img_ioi_2, path = here("paper/image", "img_ioi_2.png"), format = "png")
-img_ioi_2
-```
-
 ### 3.3.2. Key Velocity (KV)
@@
 
-```{r kot-dyn-2, include = FALSE}
-kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent)
+```{r kor-dyn-2, include = FALSE}
+kor_dyn_plot <- data.table(subject = kor_dyn[Condition == "teaching"]$SubNr, teaching = kor_dyn[Condition == "teaching"]$Mean, performing = kor_dyn[Condition == "performing"]$Mean, subcomponent = kor_dyn[Condition == "teaching"]$Subcomponent)
 
-kot_dyn_2 <- ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
+kor_dyn_2 <- ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
   ylim(-1, 0.4) +
@@
 # save plots
-ggsave(here("paper/image", "kot_dyn_2.png"), plot = kot_dyn_1, dpi = 600, width = 7, height = 5)
-img_kot_dyn_2 <- image_read(here("paper/image", "kot_dyn_2.png"))
+ggsave(here("paper/image", "kor_dyn_2.png"), plot = kor_dyn_2, dpi = 600, width = 7, height = 5)
+img_kor_dyn_2 <- image_read(here("paper/image", "kor_dyn_2.png"))
 ```
 
+```{r plot-ioi-2, out.width = "100%", fig.cap = "\\label{fig:ioi-2}Experiment 2: Normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+img_ioi_2 <- image_append(c(img_ioi_art_2, img_ioi_dyn_2))
+image_write(img_ioi_2, path = here("paper/image", "img_ioi_2.png"), format = "png")
+img_ioi_2
+```
+
+```{r plot-kot-2, out.width = "100%", fig.cap = "\\label{fig:kot-2}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+# combine
+img_kor_2 <- image_append(c(img_kor_art_2, img_kor_dyn_2))
+image_write(img_kor_2, path = here("paper/image", "img_kor_2.png"), format = "png")
+img_kor_2
+```
 ## 3.4. Discussion

### 2020-12-15 paper/manuscript.tex
add kot/kor plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying piano performance at a music school at the time of recruitment.
-One participant was excluded due to insufficient motor skills (see details in \emph{\protect\hyperlink{procedure-2}{Procedure}}).
-Twenty participants (9 female) were included in data analysis.
+For data analysis, we excluded 1 participants due to insufficient motor skills (1) and an overall deviated tempo (1).
+Nineteen participants (9 female) were included in data analysis.
 Most participants were right-handed (left; 2).
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Experiment 2: Normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-ioi-2}
-\end{figure}
-
 \hypertarget{key-velocity-kv-3}{%
@@
 Blah
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Experiment 2: Normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-ioi-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-1}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-kot-2}
+\end{figure}
 

### 2020-12-15 experiment-1/analysis/stats/articulation.Rmd
update articulation (posthoc tests)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ### Statistics
-#### Two-way ANOVA (car_aov)
+#### Two-way ANOVA (aov_car)
 ```{r anova-art-kot-1, echo = FALSE}
@@
 ```{r post-hoc-1, echo = FALSE}
-test(pairs(emmeans(kot_art_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(kot_art_aov,~Condition|Subcomponent), adjust = "tukey")
+kot_art_posthoc <- test(pairs(emmeans(kot_art_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+kot_art_posthoc
 ```
@@
 ### Statistics
-#### Two-way ANOVA (car_aov)
+#### Two-way ANOVA (aov_car)
 ```{r anova-art-kor-1, echo = FALSE}
@@
 ```{r post-hoc-2, echo = FALSE}
-test(pairs(emmeans(kor_art_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(kor_art_aov,~Condition|Subcomponent), adjust = "tukey")
+kor_art_posthoc <- test(pairs(emmeans(kor_art_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+kor_art_posthoc
 ```
@@
 ### Statistics
-#### Two-way ANOVA (car_aov)
+#### Two-way ANOVA (aov_car)
 ```{r anova-dyn-kot-1, echo = FALSE}
@@
 ### Statistics
-#### Two-way ANOVA (car_aov)
+#### Two-way ANOVA (aov_car)
 ```{r anova-dyn-kor-1, echo = FALSE}

### 2020-12-15 experiment-2/analysis/stats/articulation.Rmd
export as R scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update analysis without subnr 2
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update articulation (posthoc tests)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 script/R/stats/articulation.Rmd
@@
 2. Excluded participants
+- SubNr 2: A deviated tempo
 - SubNr 3: Insufficient motor skills

### 2020-12-15 experiment-1/analysis/stats/articulation.R

@@ -87,8 +87,8 @@ summary(kot_art_aov_2)
 
 
 ## ----post-hoc-1, echo = FALSE-------------------------------------------------
-test(pairs(emmeans(kot_art_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(kot_art_aov,~Condition|Subcomponent), adjust = "tukey")
+kot_art_posthoc <- test(pairs(emmeans(kot_art_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+kot_art_posthoc
 
 
 ## ----kot-art-trial, echo = FALSE----------------------------------------------
@@ -168,8 +168,8 @@ summary(kor_art_aov_2)
 
 
 ## ----post-hoc-2, echo = FALSE-------------------------------------------------
-test(pairs(emmeans(kor_art_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(kor_art_aov,~Condition|Subcomponent), adjust = "tukey")
+kor_art_posthoc <- test(pairs(emmeans(kor_art_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+kor_art_posthoc
 
 
 ## ----kor-art-trial, echo = FALSE----------------------------------------------

### 2020-12-15 experiment-2/analysis/stats/articulation.R

@@ -72,8 +72,8 @@ summary(kor_art_aov_2)
 
 
 ## ----post-hoc-2, echo = FALSE-------------------------------------------------
-test(pairs(emmeans(kor_art_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(kor_art_aov,~Condition|Subcomponent), adjust = "tukey")
+kor_art_posthoc <- test(pairs(emmeans(kor_art_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+kor_art_posthoc
 
 
 ## ----kor-art-trial, echo = FALSE----------------------------------------------
@@ -153,8 +153,8 @@ summary(kot_art_aov_2)
 
 
 ## ----post-hoc-1, echo = FALSE-------------------------------------------------
-test(pairs(emmeans(kot_art_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(kot_art_aov,~Condition|Subcomponent), adjust = "tukey")
+kot_art_posthoc <- test(pairs(emmeans(kot_art_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+kot_art_posthoc
 
 
 ## ----kot-art-trial, echo = FALSE----------------------------------------------

### 2020-12-15 paper/manuscript.Rmd
add stats for KOT/KOR
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-LS (legato vs.
+staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[3,5])`).
+Post-hoc comparisons based on the estimated marginal  means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching and performing condition (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-FP (forte vs.
+piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[3,5])`).
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-LS (legato vs.
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`).
+Post-hoc comparisons based on the estimated marginal  means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching and performing condition (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-FP (forte vs.
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,5])`).
 

### 2020-12-15 paper/manuscript.tex
add stats for KOT/KOR
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.01).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (\emph{p} = 0.004) while there was no significant difference in legato between the teaching and performing condition (\emph{p} = 0.32).
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (\emph{F}(1,30) = 0.423, \emph{p} = 0.520, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.000, \emph{p} = 0.987, \(\eta_G^2\) = 0.00).
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 2324, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 6.90, \emph{p} = 0.017, \(\eta_G^2\) = 0.01).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (\emph{p} = 0.001) while there was no significant difference in legato between the teaching and performing condition (\emph{p} = 0.85).
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 559, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.93), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (\emph{F}(1,18) = 0.016, \emph{p} = 0.902, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 0.576, \emph{p} = 0.458, \(\eta_G^2\) = 0.00).
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-1}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right).
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-2}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.

### 2020-12-15 experiment-1/analysis/stats/dynamics.Rmd
some corrections for plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
-  theme(legend.position = "none")
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ```{r post-hoc-1, echo = FALSE}
-test(pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent), adjust = "tukey")
+vel_dyn_posthoc <- test(pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+vel_dyn_posthoc
 ```
@@
 ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed", )
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
 ```
@@
 ```{r post-hoc-2, echo = FALSE}
-test(pairs(emmeans(vel_diff_dyn_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(vel_diff_dyn_aov,~Condition|Subcomponent), adjust = "tukey")
+vel_diff_dyn_posthoc <- test(pairs(emmeans(vel_diff_dyn_aov,~Condition|Subcomponent), adjust = "turkey"), joint = TRUE)
+vel_diff_dyn_posthoc
 ```
@@
 ```{r post-hoc-3, echo = FALSE}
-test(pairs(emmeans(vel_art_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(vel_art_aov,~Condition|Subcomponent), adjust = "tukey")
+vel_art_posthoc <- test(pairs(emmeans(vel_art_aov,~Condition|Subcomponent), adjust = "turkey"), joint = TRUE)
+vel_art_posthoc
 ```
@@
 ```{r post-hoc-4, echo = FALSE}
-test(pairs(emmeans(vel_diff_art_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(vel_diff_art_aov,~Condition|Subcomponent), adjust = "tukey")
+vel_diff_art_posthoc <- test(pairs(emmeans(vel_diff_art_aov,~Condition|Subcomponent), adjust = "turkey"), joint = TRUE)
+vel_diff_art_posthoc
 ```

### 2020-12-15 paper/manuscript.Rmd
some corrections for plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The ``aov_car" function in the afex package was used for a repeated-measures ANOVA.
-For post-hoc comparisons, we used emmeans package to investigate interactions from the ANOVA.
+For post-hoc comparisons, we used emmeans package.
 
@@
 ioi_art_1 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  ylim(150, 260) +
+  ylim(145, 260) +
   geom_hline(yintercept = 188, linetype = "dashed") +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 200, paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 230, paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-LS (legato vs.
+staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[3,5])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not show a significant difference between the teaching and performing condition for each subcomponent (Legato; *p* = `r sprintf('%.3f', vel_art_posthoc$p.value[1])`, Staccato; *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
 
-### 2.2.4. KV Transition Points
+```{r vel-art-1, echo = FALSE}
+vel_art_plot <- data.table(subject = vel_art[Condition == "teaching"]$SubNr, teaching = vel_art[Condition == "teaching"]$Mean, performing = vel_art[Condition == "performing"]$Mean, subcomponent = vel_art[Condition == "teaching"]$Subcomponent)
 
-Blah
+vel_art_1 <- ggpaired(vel_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation") +
+  ylim(50, 100) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_art_1.png"), plot = vel_art_1, dpi = 600, width = 7, height = 5)
+img_vel_art_1 <- image_read(here("paper/image", "vel_art_1.png"))
+```
 
@@
 ioi_dyn_1 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  ylim(150, 260) +
+  ylim(145, 260) +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-FP (forte vs.
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[3,5])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition than in the performing condition (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) while there was no significant difference in piano between the teaching and performing condition (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`).
+
+```{r vel-dyn-1, echo = FALSE}
+vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent)
+
+vel_dyn_1 <- ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
+  ylim(50, 100) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_dyn_1.png"), plot = vel_dyn_1, dpi = 600, width = 7, height = 5)
+img_vel_dyn_1 <- image_read(here("paper/image", "vel_dyn_1.png"))
+```
 
@@
 
-Blah
+Additionally, we conducted an exploratory analysis by focusing on specific points at which each subcomponent changed from one to the other (i.e., forte to piano; FtoP, piano to forte; PtoF).
+These points could be structurally important in order to make a contrast between forte and piano.
+We calculated a KV difference for each interval subtracting the KV value of the current note from that of the following note.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-FP (FtoP vs.
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[3,5])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`.
+
+```{r vel-diff-dyn-1, echo = FALSE}
+vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)
+
+vel_diff_dyn_1 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics") +
+  ylim(-40, 45) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_diff_dyn_1.png"), plot = vel_diff_dyn_1, dpi = 600, width = 7, height = 5)
+img_vel_diff_dyn_1 <- image_read(here("paper/image", "vel_diff_dyn_1.png"))
+```
 
@@
   ylim(-170, 105) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 0, paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -50, paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 
+```{r plot-vel-1, out.width = "100%", fig.cap = "\\label{fig:vel-1}Experiment 1: KV (0-127) when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+# combine
+img_vel_1 <- image_append(c(img_vel_art_1, img_vel_dyn_1))
+image_write(img_vel_1, path = here("paper/image", "img_vel_1.png"), format = "png")
+img_vel_1
+```
+
+```{r plot-vel-diff-1, out.width = "100%", fig.cap = "\\label{fig:vel-diff-1}Experiment 1: KV Difference at transition points when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+img_vel_diff_dyn_1
+```
+
 ## 2.4. Discussion
+
+Blah
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-LS (legato vs.
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[2,5])`), reflecting louder sound for legato notes compared to staccato notes.
+Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[3,5])`).
 
-### 3.2.4. KV Transition Points
+```{r vel-art-2, echo = FALSE}
+vel_art_plot <- data.table(subject = vel_art[Condition == "teaching"]$SubNr, teaching = vel_art[Condition == "teaching"]$Mean, performing = vel_art[Condition == "performing"]$Mean, subcomponent = vel_art[Condition == "teaching"]$Subcomponent)
 
-Blah
+vel_art_2 <- ggpaired(vel_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation") +
+  ylim(c(50, 100)) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_art_2.png"), plot = vel_art_2, dpi = 600, width = 7, height = 5)
+img_vel_art_2 <- image_read(here("paper/image", "vel_art_2.png"))
+```
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-FP (forte vs.
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[3,5])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition than in the performing condition (Forte; *p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`, Piano; *p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`).
+
+```{r vel-dyn-2, echo = FALSE}
+vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent)
+
+vel_dyn_2 <- ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
+  ylim(c(50, 100)) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_dyn_2.png"), plot = vel_dyn_2, dpi = 600, width = 7, height = 5)
+img_vel_dyn_2 <- image_read(here("paper/image", "vel_dyn_2.png"))
+```
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Subcomponent-FP (FtoP vs.
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[3,5])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`.
+
+```{r vel-diff-dyn-2, echo = FALSE}
+vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)
+
+vel_diff_dyn_2 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics") +
+  ylim(-40, 45) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_diff_dyn_2.png"), plot = vel_diff_dyn_2, dpi = 600, width = 7, height = 5)
+img_vel_diff_dyn_2 <- image_read(here("paper/image", "vel_diff_dyn_2.png"))
+```
 
@@
 ```
+
+```{r plot-vel-2, out.width = "100%", fig.cap = "\\label{fig:vel-2}Experiment 2: KV (0-127) when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+# combine
+img_vel_2 <- image_append(c(img_vel_art_2, img_vel_dyn_2))
+image_write(img_vel_2, path = here("paper/image", "img_vel_2.png"), format = "png")
+img_vel_2
+```
+
+```{r plot-vel-diff-2, out.width = "100%", fig.cap = "\\label{fig:vel-diff-1}Experiment 2: KV Difference at transition points when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+img_vel_diff_dyn_2
+```
 ## 3.4. Discussion
+
+Blah
 

### 2020-12-15 paper/manuscript.tex
some corrections for plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The ``aov\_car" function in the afex package was used for a repeated-measures ANOVA.
-For post-hoc comparisons, we used emmeans package to investigate interactions from the ANOVA.
+For post-hoc comparisons, we used emmeans package.
 
@@
 
-Blah
-
-\hypertarget{kv-transition-points}{%
-\subsubsection{2.2.4. KV Transition Points}\label{kv-transition-points}}
-
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.0044).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not show a significant difference between the teaching and performing condition for each subcomponent (Legato; \emph{p} = 0.696, Staccato; \emph{p} = 0.07).
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 369.54, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7367) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.0034).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition than in the performing condition (\emph{p} = 0.021) while there was no significant difference in piano between the teaching and performing condition (\emph{p} = 0.448).
 
-\hypertarget{kv-transition-points-1}{%
-\subsubsection{2.3.3. KV Transition Points}\label{kv-transition-points-1}}
+\hypertarget{kv-transition-points}{%
+\subsubsection{2.3.3. KV Transition Points}\label{kv-transition-points}}
 
-Blah
+Additionally, we conducted an exploratory analysis by focusing on specific points at which each subcomponent changed from one to the other (i.e., forte to piano; FtoP, piano to forte; PtoF).
+These points could be structurally important in order to make a contrast between forte and piano.
+We calculated a KV difference for each interval subtracting the KV value of the current note from that of the following note.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 223.89, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8344) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.0321).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.002, PtoF; \emph{p} \textless{} 0.001.
+
+\begin{verbatim}
+## Warning: Removed 2 rows containing missing values (geom_text).
+\end{verbatim}
 
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-1-1} \caption{\label{fig:vel-1}Experiment 1: KV (0-127) when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-1-1} \caption{\label{fig:vel-diff-1}Experiment 1: KV Difference at transition points when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff-1}
+\end{figure}
+
 \hypertarget{discussion}{%
 \subsection{2.4. Discussion}\label{discussion}}
+
+Blah
 
@@
 
-Blah
-
-\hypertarget{kv-transition-points-2}{%
-\subsubsection{3.2.4. KV Transition Points}\label{kv-transition-points-2}}
-
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.011, \(\eta_G^2\) = 0.0400), reflecting louder sound for legato notes compared to staccato notes.
+Neither the main effect of Condition (\emph{F}(1,18) = 0.46, \emph{p} = 0.506, \(\eta_G^2\) = 0.0008) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.799, \(\eta_G^2\) = 0.0001).
 
@@
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126.10, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7190) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.0217).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition than in the performing condition (Forte; \emph{p} = 0.036, Piano; \emph{p} = 0.028).
 
-\hypertarget{kv-transition-points-3}{%
-\subsubsection{3.3.3. KV Transition Points}\label{kv-transition-points-3}}
+\hypertarget{kv-transition-points-1}{%
+\subsubsection{3.3.3. KV Transition Points}\label{kv-transition-points-1}}
 
-Blah
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 110.64, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8167) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.0641).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.017, PtoF; \emph{p} = 0.003.
+
+\begin{verbatim}
+## Warning: Removed 2 rows containing missing values (geom_text).
+\end{verbatim}
 
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-2-1} \caption{\label{fig:vel-2}Experiment 2: KV (0-127) when playing the piece with either articulation (left) or dynamics (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-2}
+\end{figure}
+
 \hypertarget{discussion-1}{%
 \subsection{3.4. Discussion}\label{discussion-1}}
+
+Blah
 

### 2020-12-15 experiment-1/analysis/stats/dynamics.R

@@ -58,8 +58,7 @@ ggerrorplot(vel_dyn, x = "Condition", y = "Mean", add = c("mean", "mean_se"), er
 vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent)
 
 ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
-  theme(legend.position = "none")
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
 
 
 ## ----anova-dyn-vel-1, echo = FALSE--------------------------------------------
@@ -73,8 +72,8 @@ summary(vel_dyn_aov_2)
 
 
 ## ----post-hoc-1, echo = FALSE-------------------------------------------------
-test(pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent)), joint = TRUE)
-pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent), adjust = "tukey")
+vel_dyn_posthoc <- test(pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+vel_dyn_posthoc
 
 
 ## ----vel-dyn-trial, echo = FALSE----------------------------------------------
@@ -140,7 +139,7 @@ ggerrorplot(vel_diff_dyn, x = "Condition", y = "Mean", add = c("mean", "mean_se"
 vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)
 
[…]

### 2020-12-15 experiment-2/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE------------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE-------------------------------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE----------------------------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,332 +31,331 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE----------------------------------------------------
+## ----vel-dyn, echo = FALSE-------------------------------------------------------
 # For each individual
 vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE-----------------------------------------------
+## ----vel-dyn-box,  echo = FALSE--------------------------------------------------
[…]

### 2020-12-16 paper/manuscript.Rmd
update intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Deliberate teaching has supported human skill transmission over generations and provides a key route for learning [@tomasello_2016; @whiten_2017].
+Deliberate teaching has supported human skill transmission over generations and provides a key route for learning [@thornton_2008; @tomasello_2016].
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
@@
 
-In the present study, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, articulation, loudness) can be modulated at the same time.
+In the present experiments, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, articulation, loudness) can be modulated at the same time.
 Specific and subtle aspects of the performance need to be highlighted to teach a respective technique.
@@
 
-If experts rely on generic action modulations, it can be expected that they will play more slowly during demonstration, regardless of the kind of expressive techniques to be taught.
+If experts rely on generic action modulations, it can be expected that they will play more slowly during teaching demonstration, regardless of the kind of expressive techniques to be taught.
 To the extent that experts use action modulations to support their teaching of specific expressive techniques, they should produce longer legato and shorter staccato when teaching articulation whereas they should produce louder forte and softer piano when teaching dynamics.
@@
 Importantly, they should avoid modulating irrelevant properties of expression, e.g., the smoothness of the sound while teaching dynamics.
+
+In Experiment 1, we employed a simple musical scale to examine whether and how skilled pianists varied their performances depending on which expressive techniques (i.e., either articulation or dynamics) they were supposed to teach.
+Experiment 2 was conducted to replicate our findings from Experiment 1 with a more realistic piece of music.
 
@@
 
-Blah
+The findings from Experiment 1 indicated that skilled pianists were able to modify their performance for teaching purposes.
+For IOIs (tempo), although there was a general trend of slowing down, we found a small but significant difference only when playing the piece with the notated articulation.
+This could be due to the leading metronome, which was given prior to each performance.
+As participants tried to follow the ideal tempo through their performances, it was possible that there was only a tiny difference between the two conditions.
+Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.
+to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics.
+With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspect of the technique.
+Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
+Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
+These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
+
+One of the limitations of Experiment 1 is that we employed a simple musical scale, which was good to gain an experimental control but not a realistic and natural stimulus to implement artistic expressions.
+In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment.
+We predict that participants will played more slowly regardless of the techniques to be taught or only when teaching articulation based on the finding from Experiment 1.
+Also, we expect that participants will be able to highlight only the relevant aspects of the techniques to be taught.
+More precisely, they should be able to produce longer legato and shorter staccato while teaching articulation.
+When teaching dynamics, they should be able to produce louder forte and softer piano.
+Given the findings we observed in Experiment 1, participants will make a larger contrast between forte and piano at transition points.
 
@@
 
-```{r plot-vel-diff-2, out.width = "100%", fig.cap = "\\label{fig:vel-diff-1}Experiment 2: KV Difference at transition points when playing the piece with dynamics.
+```{r plot-vel-diff-2, out.width = "100%", fig.cap = "\\label{fig:vel-diff-2}Experiment 2: KV Difference at transition points when playing the piece with dynamics.
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.

### 2020-12-16 paper/manuscript.tex
update intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Deliberate teaching has supported human skill transmission over generations and provides a key route for learning (Tomasello, 2016; Whiten, 2017).
+Deliberate teaching has supported human skill transmission over generations and provides a key route for learning (Thornton \& Raihani, 2008; Tomasello, 2016).
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
@@
 Importantly, they should avoid modulating irrelevant properties of expression, e.g., the smoothness of the sound while teaching dynamics.
+
+In Experiment 1, we employed a simple musical scale to examine whether and how skilled pianists varied depending on which expressive techniques (i.e., either articulation or dynamics) was taught.
+Experiment 2 was conducted to replicate our findings from Experiment 1 with a more realistic piece of music.
 
@@
 
-Blah
+The findings from Experiment 1 indicated that skilled pianists were able to modify their performance for teaching purposes.
+For IOIs (tempo), although there was a general trend of slowing down, we found a small but significant difference only when playing the piece with the notated articulation.
+This could be due to the leading metronome, which was given prior to each performance.
+As participants tried to follow the ideal tempo through their performances, it was possible that there was only a tiny difference between the two conditions.
+Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.~to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics.
+With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspect of the technique.
+Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
+Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
+These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
+
+One of the limitations of Experiment 1 is that we employed a simple musical scale, which was good to gain an experimental control but not a realistic and natural stimulus to implement artistic expressions.
+In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment.
+We predict that participants will played more slowly regardless of the techniques to be taught or only when teaching articulation based on the finding from Experiment 1.
+Also, we expect that participants will be able to highlight only the relevant aspects of the techniques to be taught.
+More precisely, they should be able to produce longer legato and shorter staccato while teaching articulation.
+When teaching dynamics, they should be able to produce louder forte and softer piano.
+Given the findings we observed in Experiment 1, participants will make a larger contrast between forte and piano at transition points.
 
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Experiment 2: KV Difference at transition points when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff-2}
+\end{figure}
+
 \hypertarget{discussion-1}{%
@@
 
+\leavevmode\hypertarget{ref-thornton_2008}{}%
+Thornton, A., \& Raihani, N. J. (2008). The evolution of teaching. \emph{Animal Behaviour}, \emph{75}(6), 1823--1836. \url{https://doi.org/10.1016/j.anbehav.2007.12.014}
+
 \leavevmode\hypertarget{ref-tomasello_2016}{}%
@@
 \url{https://doi.org/10.1016/j.specom.2006.10.003}
-
-\leavevmode\hypertarget{ref-whiten_2017}{}%
-Whiten, A. (2017). Social Learning and Culture in Child and Chimpanzee. \emph{Annual Review of Psychology}, \emph{68}(1), 129--154. \url{https://doi.org/10.1146/annurev-psych-010416-044108}
 \end{cslreferences}

### 2020-12-16 experiment-2/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE------------------------------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE-------------------------------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE----------------------------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,331 +31,331 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE-------------------------------------------------------
+## ----vel-dyn, echo = FALSE----------------------------------------------------
 # For each individual
 vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE--------------------------------------------------
+## ----vel-dyn-box,  echo = FALSE-----------------------------------------------
[…]

### 2020-12-16 paper/manuscript.Rmd
modified discussion/correct some stats report
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 abstract: |
-  aaaa
+  Experts modulate their performance of actions for teaching purposes, performing slower and exaggerated movements when demonstrating novel actions to novices.
+The present study asked whether such modulations are also used to teach techniques of artistic expression.
+One domain where subtle movement modulations are crucial for achieving artistic expression is music.
+We investigated how skilled pianists modulate their playing to demonstrate to students the techniques required for conveying piece-related musical expressions, compared to performing the piece without didactic intentions.
+Expressions in the piece concerned either articulation (i.e., legato and staccato) or dynamics (i.e., forte and piano).
+The pianists played either with the goal to perform the piece to an audience or with the goal to teach the respective techniques to novices.When intending to teach articulation, skilled pianists produced more exaggerated staccato.
+When intending to teach dynamics, they created a larger contrast between forte and piano at structurally important points.
+We found consistent results with a simple musical scale (Experiment 1) and a more naturalistic piece of music (Experiment 2).
+These findings suggest that action modulations are performed not only to attract learners' attention but to highlight the most relevant aspects of the actions to be learnt.
 
@@
 For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques [@juslin_2003].
-This raises the question of whether and how experts modulate their actions during demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
+This raises the question of whether and how experts modulate their actions during didactic demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
 Studying action modulations that serve the teaching of artistic techniques is of broader theoretical importance: It can show whether action modulations can be used to selectively highlight subtle aspects of actions that define expert performance.
@@
 
-If experts rely on generic action modulations, it can be expected that they will play more slowly during teaching demonstration, regardless of the kind of expressive techniques to be taught.
+If experts rely on generic action modulations, it can be expected that they will play more slowly during didactic demonstration, regardless of the kind of expressive techniques to be taught.
 To the extent that experts use action modulations to support their teaching of specific expressive techniques, they should produce longer legato and shorter staccato when teaching articulation whereas they should produce louder forte and softer piano when teaching dynamics.
@@
 
-In Experiment 1, we employed a simple musical scale to examine whether and how skilled pianists varied their performances depending on which expressive techniques (i.e., either articulation or dynamics) they were supposed to teach.
-Experiment 2 was conducted to replicate our findings from Experiment 1 with a more realistic piece of music.
-
-\clearpage
+In Experiment 1, we employed a simple musical scale to examine whether and how skilled pianists vary their performance depending on which expressive techniques (i.e., either articulation or dynamics) they were teaching.
+Experiment 2 was conducted to replicate our findings from Experiment 1 with a more naturalistic piece of music.
 
@@
 
-In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution.
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
+In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* < 0.001).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
 
@@
 ioi_art_1 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  ylim(145, 260) +
+  ylim(150, 260) +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[3,5])`).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching and performing condition (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
+staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`; *Figure \ref{fig:kot-1}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition[*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
 
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[3,5])`).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not show a significant difference between the teaching and performing condition for each subcomponent (Legato; *p* = `r sprintf('%.3f', vel_art_posthoc$p.value[1])`, Staccato; *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
+staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`; *Figure \ref{fig:vel-1}*).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment failed to find a significant difference between the teaching [Legato; *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`, Staccato; *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] and performing condition [Legato; *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`, Staccato; *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] for each subcomponent (Legato; *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[1])`, Staccato; *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
 
@@
 
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
 
@@
 ioi_dyn_1 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  ylim(145, 260) +
+  ylim(150, 260) +
   geom_hline(yintercept = 188, linetype = "dashed") +
@@
 performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[3,5])`).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition than in the performing condition (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) while there was no significant difference in piano between the teaching and performing condition (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`).
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-1}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) while there was no significant difference in piano between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition[*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`).
 
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[3,5])`).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`.
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-diff}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`] (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`).
 
@@
 
-vel_diff_dyn_1 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics") +
+vel_diff_dyn_1 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Exp 1") +
   ylim(-40, 45) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 45, paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 performing) and Subcomponent-FP (forte vs.
-piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[3,5])`).
+piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`; *Figure \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`).
 
@@
 ```
+
+## 2.4. Discussion
+
+The findings from Experiment 1 indicated that skilled pianists were able to modify their performance for teaching purposes.
+For IOIs (tempo), we found a small but significant difference only when playing the piece with the notated articulation.
+This could be due to the leading metronome, which was given prior to each performance.
+As participants tried to follow the ideal tempo through their performances, it was possible that there was only a tiny difference between the two conditions.
+Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.
+to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics.
+With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspect of the technique.
+Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
+Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
+These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
+
+One of the limitations of Experiment 1 is that we employed a simple musical scale, which was good to gain an experimental control but not a naturalistic stimulus to implement artistic expressions.
+In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment.
+We predict that participants will played more slowly regardless of the techniques to be taught or only when teaching articulation based on the finding from Experiment 1.
+Also, we expect that participants will be able to highlight only the relevant aspects of the techniques to be taught.
+More precisely, they should be able to produce longer legato and shorter staccato while teaching articulation.
+When teaching dynamics, they should be able to produce louder forte and softer piano.
+Given the findings we observed in Experiment 1, participants will make a larger contrast between forte and piano at transition points.
 
@@
 ```
-
-```{r plot-vel-diff-1, out.width = "100%", fig.cap = "\\label{fig:vel-diff-1}Experiment 1: KV Difference at transition points when playing the piece with dynamics.
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
-img_vel_diff_dyn_1
-```
-
-## 2.4. Discussion
-
-The findings from Experiment 1 indicated that skilled pianists were able to modify their performance for teaching purposes.
-For IOIs (tempo), although there was a general trend of slowing down, we found a small but significant difference only when playing the piece with the notated articulation.
-This could be due to the leading metronome, which was given prior to each performance.
-As participants tried to follow the ideal tempo through their performances, it was possible that there was only a tiny difference between the two conditions.
-Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.
-to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics.
-With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspect of the technique.
-Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
-Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
-These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
-
-One of the limitations of Experiment 1 is that we employed a simple musical scale, which was good to gain an experimental control but not a realistic and natural stimulus to implement artistic expressions.
-In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment.
-We predict that participants will played more slowly regardless of the techniques to be taught or only when teaching articulation based on the finding from Experiment 1.
-Also, we expect that participants will be able to highlight only the relevant aspects of the techniques to be taught.
-More precisely, they should be able to produce longer legato and shorter staccato while teaching articulation.
-When teaching dynamics, they should be able to produce louder forte and softer piano.
-Given the findings we observed in Experiment 1, participants will make a larger contrast between forte and piano at transition points.
 
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`).
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`; *Figure \ref{fig:kor-2}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching and performing condition (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[2,5])`), reflecting louder sound for legato notes compared to staccato notes.
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[2,5])`; *Figure \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
 Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[3,5])`).
@@
 performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[3,5])`).
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-2}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition than in the performing condition (Forte; *p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`, Piano; *p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`).
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[3,5])`).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`.
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-diff}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`).
 
@@
 
-vel_diff_dyn_2 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics") +
+vel_diff_dyn_2 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Exp 2") +
   ylim(-40, 45) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 45, paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,5])`).
+Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:kor-2}*).
 
@@
 
+## 3.4. Discussion
+
+The results from Experiment 2 demonstrated that we successfully replicated our earlier findings that skilled pianists were able to modulate their performance for teaching purposes.
+As Experiment 1, we observed slower performance only while teaching the notated articulation.
+With regard to KOR (articulation), participants highlighted articulation by producing shorter staccato.
+Again, there was no significant difference in legato between the two conditions as Experiment 1.
+In term of KV (dynamics), participants produced louder forte and softer piano, which showed a clearer pattern of exaggeration compared to Experiment 1 where only softer piano was observed.
+Furthermore, we again found that participants made a larger contrast between forte and piano bidirectionally at transition points.
+To sum, Experiment 2 confirmed that participants modified their performance in systematic and fine-grained ways in order to teach expressive techniques with a naturalistic piece of music.
+
 ```{r plot-ioi-2, out.width = "100%", fig.cap = "\\label{fig:ioi-2}Experiment 2: Normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
@@
 
-```{r plot-kot-2, out.width = "100%", fig.cap = "\\label{fig:kot-2}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right).
+```{r plot-kor-2, out.width = "100%", fig.cap = "\\label{fig:kor-2}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
@@
 
-```{r plot-vel-diff-2, out.width = "100%", fig.cap = "\\label{fig:vel-diff-2}Experiment 2: KV Difference at transition points when playing the piece with dynamics.
+```{r plot-vel-diff, out.width = "100%", fig.cap = "\\label{fig:vel-diff}Experiment 1, 2: KV Difference at transition points when playing the piece with dynamics (left; Experiment 1, right; Experiment 2).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
 Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
-img_vel_diff_dyn_2
+# combine
+img_vel_diff <- image_append(c(img_vel_diff_dyn_1, img_vel_diff_dyn_2))
+image_write(img_vel_diff, path = here("paper/image", "img_vel_diff.png"), format = "png")
+img_vel_diff
 ```
-## 3.4. Discussion
-
-Blah
 
@@
 # 4. General Discussion
+
+The aim of the study was to investigate how expert pianists could modulate their performance in order to teach expressive artistic techniques.
+Overall, the findings of the two experiments showed that pianists performed one and the same piece differently depending on whether they played with the intention to teach expressive techniques or with the intention to perform the piece for an audience.
+When playing with the intention to teach, pianists selectively highlighted relevant aspects of artistic expressions.
+In particular, they exaggerated staccato when playing with the intention to teach expressions concerning articulation, and they exaggerated changes between forte and piano (in both directions) when playing with the intention to teach expressions concerning dynamics.
+
+Although participants tended to play slower in the teaching condition than in the performing condition in general, we found a significant difference only for articulation, not for dynamics.
+As we mentioned in Experiment 1, it could be that the metronome beats given prior to each performance discouraged participants from deviating from the prescribed tempo.
+Another possibility is that expression is actually best taught when leaving the tempo unchanged when it is irrelevant.
+This would imply that general didactic modulations like slowing down, reported in research on parent-infant interaction, might be less useful in the context of teaching expressive artistic techniques where timing itself can be used to add an expression to a performance.
+
+Participants exaggerated articulation for teaching purposes by highlighting staccato.
+The lack of a difference for legato might stem from a ceiling effect.
+When participants played the piece with dynamics, there was no difference between the two conditions with regard to KOT/KOR, which indicates that participants did not modulate irrelevant properties of the sound.
+However, participants made a larger contrast between forte and piano for teaching purposes to highlight the transition of dynamics.
+We expected there would be an overall exaggeration in KV while teaching dynamics and indeed found that participants produced louder forte and softer piano, but only in Experiment 2.
+The finding of making a larger contrast in dynamics at transition points suggests an effective way to highlight the technique as the loudness is determined relatively.
+Taken together, these findings demonstrate that expert pianists systematically modulated their sound depending on the specific technique they were trying to convey.
+
+Musicians and music teachers tend to consider expressive skills as a performer’s most important skills [@lindstrom_2003].
+In the present study, experts focused on the teaching of expressions that were notated.
+However, expression also has other facets [@juslin_2003] that are not only piece-related but performer- or context-related.
+An interesting topic for future research is how experts teach not only expressive techniques as such but convey possibilities of interpretation.
+
+One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students.
+Given that teachers modulate their demonstration throughout online interactions with learners [@okazaki_2019], future studies are needed to investigate how experts dynamically adapt their performance to their student’s skill level and demonstrated abilities.
+
+The present findings extend earlier research on teaching-related action modulations.
+First, our study sheds light on the teaching of expressive techniques that are an integral part of skill acquisition in artistic domains.
+We showed that compared to an expressive performance baseline, experts made specific modulations in order to teach particular techniques.
+Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention.
+Rather, expert demonstrators seem to follow principles of relevance in communication [@sperber_1995], highlighting only those aspects they are intending to demonstrate.
+How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher's intentions explicitly adds to the learning success are important questions for future research.
 
@@
 \endgroup
-\raggedbottom
+
+\clearpage
+
+# Supplementary Materials{#supplementary}
+## Instructions
+### Experiment 1
+In the teaching condition, participants were shown with the following instruction in front of the monitor (Italic sentences were highlighted with a yellow colour on a black background.); Now, play what you practised as if you were teaching it to students.
+Students already know how to produce the sequence of the  tones and now are trying to *learn how to perform the piece expressively* by listening to your performance.
+*Do your best as a teacher to produce the piece according to the notation that you just practised*.
+
+In the performing condition, participants were displayed with the following instruction in front of the monitor; Now, play what you practised as if you were performing it to an audience.
+*Do your best as a performer to produce the piece according to the notation that you just practised*.
+
+### Experiment 2
+
+In the teaching condition, we gave participants the exact same instruction as Experiment 1.
+
+In the performing condition, participants were displayed with the following instruction in front of the monitor (Italic sentences were highlighted with a yellow colour on a black background.); Now, play what you practised as if you were performing it to an audience.
+*Perform the piece expressively with your interpretation*.
+*Do your best as a performer to produce the piece according to the notation that you just practised*.

### 2020-12-16 paper/manuscript.tex
modified discussion/correct some stats report
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \abstract{
-aaaa
+Experts modulate their performance of actions for teaching purposes, performing slower and exaggerated movements when demonstrating novel actions to novices.
+The present study asked whether such modulations are also used to teach techniques of artistic expression.
+One domain where subtle movement modulations are crucial for achieving artistic expression is music.
+We investigated how skilled pianists modulate their playing to demonstrate to students the techniques required for conveying piece-related musical expressions, compared to performing the piece without didactic intentions.
+Expressions in the piece concerned either articulation (i.e., legato and staccato) or dynamics (i.e., forte and piano).
+The pianists played either with the goal to perform the piece to an audience or with the goal to teach the respective techniques to novices.When intending to teach articulation, skilled pianists produced more exaggerated staccato.
+When intending to teach dynamics, they created a larger contrast between forte and piano at structurally important points.
+We found consistent results with a simple musical scale (Experiment 1) and a more naturalistic piece of music (Experiment 2).
+These findings suggest that action modulations are performed not only to attract learners' attention but to highlight the most relevant aspects of the actions to be learnt.
 }
@@
 For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques (Juslin, 2003).
-This raises the question of whether and how experts modulate their actions during demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
+This raises the question of whether and how experts modulate their actions during didactic demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
 Studying action modulations that serve the teaching of artistic techniques is of broader theoretical importance: It can show whether action modulations can be used to selectively highlight subtle aspects of actions that define expert performance.
@@
 
-In Experiment 1, we employed a simple musical scale to examine whether and how skilled pianists varied depending on which expressive techniques (i.e., either articulation or dynamics) was taught.
-Experiment 2 was conducted to replicate our findings from Experiment 1 with a more realistic piece of music.
-
-\clearpage
+In Experiment 1, we employed a simple musical scale to examine whether and how skilled pianists vary their performance depending on which expressive techniques (i.e., either articulation or dynamics) they were teaching.
+Experiment 2 was conducted to replicate our findings from Experiment 1 with a more naturalistic piece of music.
 
@@
 
-In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution.
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with notated articulation (\emph{p} = 0.021; \emph{Figure \ref{fig:ioi-1}}).
+In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021; \emph{Figure \ref{fig:ioi-1}}).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.01).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (\emph{p} = 0.004) while there was no significant difference in legato between the teaching and performing condition (\emph{p} = 0.32).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011; \emph{Figure \ref{fig:kot-1}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.004) while there was no significant difference in legato between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition{[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.32).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.0044).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not show a significant difference between the teaching and performing condition for each subcomponent (Legato; \emph{p} = 0.696, Staccato; \emph{p} = 0.07).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004; \emph{Figure \ref{fig:vel-1}}).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment failed to find a significant difference between the teaching {[}Legato; \emph{M} = 70.34, \emph{SD} = 6.45, Staccato; \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato; \emph{M} = 70.68, \emph{SD} = 5.84, Staccato; \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato; \emph{p} = 0.70, Staccato; \emph{p} = 0.07).
 
@@
 
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with notated dynamics (\emph{p} = 0.111; \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.111; \emph{Figure \ref{fig:ioi-1}}).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 369.54, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7367) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.0034).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition than in the performing condition (\emph{p} = 0.021) while there was no significant difference in piano between the teaching and performing condition (\emph{p} = 0.448).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003; \emph{Figure \ref{fig:vel-1}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.021) while there was no significant difference in piano between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition{[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.45).
 
@@
 We calculated a KV difference for each interval subtracting the KV value of the current note from that of the following note.
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 223.89, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8344) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.0321).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.002, PtoF; \emph{p} \textless{} 0.001.
-
-\begin{verbatim}
-## Warning: Removed 2 rows containing missing values (geom_text).
-\end{verbatim}
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032; \emph{Figure \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition {[}FtoP; \emph{M} = -14.06, \emph{SD} = 7.20, PtoF; \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP; \emph{M} = -11.69, \emph{SD} = 5.97, PtoF; \emph{M} = 20.88, \emph{SD} = 8.67{]} (FtoP; \emph{p} = 0.002, PtoF; \emph{p} \textless{} 0.001).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,30) = 0.423, \emph{p} = 0.520, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.000, \emph{p} = 0.987, \(\eta_G^2\) = 0.00).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31; \emph{Figure \ref{fig:kot-1}}), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.520, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.000, \emph{p} = 0.99, \(\eta_G^2\) = 0.000).
+
+\hypertarget{discussion}{%
+\subsection{2.4. Discussion}\label{discussion}}
+
+The findings from Experiment 1 indicated that skilled pianists were able to modify their performance for teaching purposes.
+For IOIs (tempo), we found a small but significant difference only when playing the piece with the notated articulation.
+This could be due to the leading metronome, which was given prior to each performance.
+As participants tried to follow the ideal tempo through their performances, it was possible that there was only a tiny difference between the two conditions.
+Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.~to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics.
+With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspect of the technique.
+Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
+Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
+These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
+
+One of the limitations of Experiment 1 is that we employed a simple musical scale, which was good to gain an experimental control but not a naturalistic stimulus to implement artistic expressions.
+In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment.
+We predict that participants will played more slowly regardless of the techniques to be taught or only when teaching articulation based on the finding from Experiment 1.
+Also, we expect that participants will be able to highlight only the relevant aspects of the techniques to be taught.
+More precisely, they should be able to produce longer legato and shorter staccato while teaching articulation.
+When teaching dynamics, they should be able to produce louder forte and softer piano.
+Given the findings we observed in Experiment 1, participants will make a larger contrast between forte and piano at transition points.
 
@@
 \end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-1-1} \caption{\label{fig:vel-diff-1}Experiment 1: KV Difference at transition points when playing the piece with dynamics.
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff-1}
-\end{figure}
-
-\hypertarget{discussion}{%
-\subsection{2.4. Discussion}\label{discussion}}
-
-The findings from Experiment 1 indicated that skilled pianists were able to modify their performance for teaching purposes.
-For IOIs (tempo), although there was a general trend of slowing down, we found a small but significant difference only when playing the piece with the notated articulation.
-This could be due to the leading metronome, which was given prior to each performance.
-As participants tried to follow the ideal tempo through their performances, it was possible that there was only a tiny difference between the two conditions.
-Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.~to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics.
-With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspect of the technique.
-Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
-Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
-These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
-
-One of the limitations of Experiment 1 is that we employed a simple musical scale, which was good to gain an experimental control but not a realistic and natural stimulus to implement artistic expressions.
-In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment.
-We predict that participants will played more slowly regardless of the techniques to be taught or only when teaching articulation based on the finding from Experiment 1.
-Also, we expect that participants will be able to highlight only the relevant aspects of the techniques to be taught.
-More precisely, they should be able to produce longer legato and shorter staccato while teaching articulation.
-When teaching dynamics, they should be able to produce louder forte and softer piano.
-Given the findings we observed in Experiment 1, participants will make a larger contrast between forte and piano at transition points.
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 2324, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 6.90, \emph{p} = 0.017, \(\eta_G^2\) = 0.01).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 2324, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 6.90, \emph{p} = 0.017, \(\eta_G^2\) = 0.01; \emph{Figure \ref{fig:kor-2}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (\emph{p} = 0.001) while there was no significant difference in legato between the teaching and performing condition (\emph{p} = 0.85).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.011, \(\eta_G^2\) = 0.0400), reflecting louder sound for legato notes compared to staccato notes.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.011, \(\eta_G^2\) = 0.0400; \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
 Neither the main effect of Condition (\emph{F}(1,18) = 0.46, \emph{p} = 0.506, \(\eta_G^2\) = 0.0008) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.799, \(\eta_G^2\) = 0.0001).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126.10, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7190) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.0217).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126.10, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7190) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.0217; \emph{Figure \ref{fig:vel-2}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition than in the performing condition (Forte; \emph{p} = 0.036, Piano; \emph{p} = 0.028).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 110.64, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8167) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.0641).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.017, PtoF; \emph{p} = 0.003.
-
-\begin{verbatim}
-## Warning: Removed 2 rows containing missing values (geom_text).
-\end{verbatim}
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 110.64, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8167) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.0641; \emph{Figure \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.017, PtoF; \emph{p} = 0.003).
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 559, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.93), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,18) = 0.016, \emph{p} = 0.902, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 0.576, \emph{p} = 0.458, \(\eta_G^2\) = 0.00).
+Neither the main effect of Condition (\emph{F}(1,18) = 0.016, \emph{p} = 0.902, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 0.576, \emph{p} = 0.458, \(\eta_G^2\) = 0.00; \emph{Figure \ref{fig:kor-2}}).
+
+\hypertarget{discussion-1}{%
+\subsection{3.4. Discussion}\label{discussion-1}}
+
+The results from Experiment 2 demonstrated that we successfully replicated our earlier findings that skilled pianists were able to modulate their performance for teaching purposes.
+As Experiment 1, we observed slower performance only while teaching the notated articulation.
+With regard to KOR (articulation), participants highlighted articulation by producing shorter staccato.
+Again, there was no significant difference in legato between the two conditions as Experiment 1.
+In term of KV (dynamics), participants produced louder forte and softer piano, which showed a clearer pattern of exaggeration compared to Experiment 1 where only softer piano was observed.
+Furthermore, we again found that participants made a larger contrast between forte and piano bidirectionally at transition points.
+To sum, Experiment 2 confirmed that participants modified their performance in systematic and fine-grained ways in order to teach expressive techniques with a naturalistic piece of music.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-2}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right).
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kor-2-1} \caption{\label{fig:kor-2}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-kot-2}
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-kor-2}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Experiment 2: KV Difference at transition points when playing the piece with dynamics.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-1} \caption{\label{fig:vel-diff}Experiment 1, 2: KV Difference at transition points when playing the piece with dynamics (left; Experiment 1, right; Experiment 2).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff-2}
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff}
 \end{figure}
-
-\hypertarget{discussion-1}{%
-\subsection{3.4. Discussion}\label{discussion-1}}
-
-Blah
 
@@
 \section{4. General Discussion}\label{general-discussion}}
+
+The aim of the study was to investigate how expert pianists could modulate their performance in order to teach expressive artistic techniques.
+Overall, the findings of the two experiments showed that pianists performed one and the same piece differently depending on whether they played with the intention to teach expressive techniques or with the intention to perform the piece for an audience.
+When playing with the intention to teach, pianists selectively highlighted relevant aspects of artistic expressions.
+In particular, they exaggerated staccato when playing with the intention to teach expressions concerning articulation, and they exaggerated changes between forte and piano (in both directions) when playing with the intention to teach expressions concerning dynamics.
+
+Although participants tended to play slower in the teaching condition than in the performing condition in general, we found a significant difference only for articulation, not for dynamics.
+As we mentioned in Experiment 1, it could be that the metronome beats given prior to each performance discouraged participants from deviating from the prescribed tempo.
+Another possibility is that expression is actually best taught when leaving the tempo unchanged when it is irrelevant.
+This would imply that general didactic modulations like slowing down, reported in research on parent-infant interaction, might be less useful in the context of teaching expressive artistic techniques where timing itself can be used to add an expression to a performance.
+
+Participants exaggerated articulation for teaching purposes by highlighting staccato.
+The lack of a difference for legato might stem from a ceiling effect.
+When participants played the piece with dynamics, there was no difference between the two conditions with regard to KOT/KOR, which indicates that participants did not modulate irrelevant properties of the sound.
+However, participants made a larger contrast between forte and piano for teaching purposes to highlight the transition of dynamics.
+We expected there would be an overall exaggeration in KV while teaching dynamics and indeed found that participants produced louder forte and softer piano, but only in Experiment 2.
+The finding of making a larger contrast in dynamics at transition points suggests an effective way to highlight the technique as the loudness is determined relatively.
+Taken together, these findings demonstrate that expert pianists systematically modulated their sound depending on the specific technique they were trying to convey.
+
+Musicians and music teachers tend to consider expressive skills as a performer's most important skills (Lindström, Juslin, Bresin, \& Williamon, 2003).
+In the present study, experts focused on the teaching of expressions that were notated.
+However, expression also has other facets (Juslin, 2003) that are not only piece-related but performer- or context-related.
+An interesting topic for future research is how experts teach not only expressive techniques as such but convey possibilities of interpretation.
+
+One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students.
+Given that teachers modulate their demonstration throughout online interactions with learners (Okazaki, Muraoka, \& Osu, 2019), future studies are needed to investigate how experts dynamically adapt their performance to their student's skill level and demonstrated abilities.
+
+The present findings extend earlier research on teaching-related action modulations.
+First, our study sheds light on the teaching of expressive techniques that are an integral part of skill acquisition in artistic domains.
+We showed that compared to an expressive performance baseline, experts made specific modulations in order to teach particular techniques.
+Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention.
+Rather, expert demonstrators seem to follow principles of relevance in communication (Sperber \& Wilson, 1995), highlighting only those aspects they are intending to demonstrate.
+How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher's intentions explicitly adds to the learning success are important questions for future research.
 
@@
 
+\leavevmode\hypertarget{ref-lindstrom_2003}{}%
+Lindström, E., Juslin, P.
+N., Bresin, R., \& Williamon, A.
+(2003).
+``Expressivity comes from within your soul'': A questionnaire study of music students' perspectives on expressivity.
+\emph{Research Studies in Music Education}, \emph{20}(1), 23--47.
+\url{https://doi.org/10.1177/1321103X030200010201}
+
 \leavevmode\hypertarget{ref-mcellin_2017}{}%
@@
 \url{https://doi.org/10.1037/xhp0000505}
+
+\leavevmode\hypertarget{ref-okazaki_2019}{}%
+Okazaki, S., Muraoka, Y., \& Osu, R.
+(2019).
+Teacher-learner interaction quantifies scaffolding behaviour in imitation learning.
+\emph{Scientific Reports}, \emph{9}(1), 1--13.
+\url{https://doi.org/10.1038/s41598-019-44049-x}
 
@@
 Sloboda, J. A. (2000). Individual differences in music performance. \emph{Trends in Cognitive Sciences}, \emph{4}(10), 397--403. \url{https://doi.org/10.1016/S1364-6613(00)01531-X}
+
+\leavevmode\hypertarget{ref-sperber_1995}{}%
+Sperber, D., \& Wilson, D. (1995). \emph{Relevance: Communication and cognition, 2nd ed}. Malden: Blackwell Publishing.
 
@@
 \endgroup
-\raggedbottom
+
+\clearpage
+
+\hypertarget{supplementary}{%
+\section{Supplementary Materials}\label{supplementary}}
+
+\hypertarget{instructions}{%
+\subsection{Instructions}\label{instructions}}
+
+\hypertarget{experiment-1-1}{%
+\subsubsection{Experiment 1}\label{experiment-1-1}}
+
+In the teaching condition, participants were shown with the following instruction in front of the monitor (Italic sentences were highlighted with a yellow colour on a black background.); Now, play what you practised as if you were teaching it to students.
+Students already know how to produce the sequence of the tones and now are trying to \emph{learn how to perform the piece expressively} by listening to your performance.
+\emph{Do your best as a teacher to produce the piece according to the notation that you just practised}.
+
+In the performing condition, participants were displayed with the following instruction in front of the monitor; Now, play what you practised as if you were performing it to an audience.
+\emph{Do your best as a performer to produce the piece according to the notation that you just practised}.
+
+\hypertarget{experiment-2-1}{%
+\subsubsection{Experiment 2}\label{experiment-2-1}}
+
+In the teaching condition, we gave participants the exact same instruction as Experiment 1.
+
+In the performing condition, participants were displayed with the following instruction in front of the monitor (Italic sentences were highlighted with a yellow colour on a black background.); Now, play what you practised as if you were performing it to an audience.
+\emph{Perform the piece expressively with your interpretation}.
+\emph{Do your best as a performer to produce the piece according to the notation that you just practised}.
 

### 2020-12-16 experiment-1/analysis/stats/dynamics.Rmd
correct SD = mean to SD = sd
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # Group mean
-vel_diff_dyn_all <- vel_diff_dyn[, .(N = .N, Mean = mean(Mean), SD = mean(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
+vel_diff_dyn_all <- vel_diff_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
 vel_diff_dyn_all
@@
 # For each individual
-vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = mean(Diff)), by = .(SubNr, Condition, Skill, Subcomponent,TrialNr)]
+vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent,TrialNr)]
 vel_diff_dyn_trial
@@
 # Group mean
-vel_diff_dyn_trial_all <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = mean(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
+vel_diff_dyn_trial_all <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
 vel_diff_dyn_trial_all

### 2020-12-16 experiment-1/analysis/stats/dynamics.R

@@ -126,7 +126,7 @@ ggboxplot(dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N =
 
 ## ----vel-diff-dyn-all, echo = FALSE-------------------------------------------
 # Group mean
-vel_diff_dyn_all <- vel_diff_dyn[, .(N = .N, Mean = mean(Mean), SD = mean(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
+vel_diff_dyn_all <- vel_diff_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
 vel_diff_dyn_all
 
 
@@ -159,7 +159,7 @@ vel_diff_dyn_posthoc
 
 ## ----vel-diff-dyn-trial, echo = FALSE-----------------------------------------
 # For each individual
-vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = mean(Diff)), by = .(SubNr, Condition, Skill, Subcomponent,TrialNr)]
+vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent,TrialNr)]
 vel_diff_dyn_trial
 
 
@@ -177,7 +177,7 @@ ggline(dt_vel_diff_dyn[Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD =
 
 ## ----vel-diff-dyn-trial-all, echo = FALSE-------------------------------------
 # Group mean
-vel_diff_dyn_trial_all <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = mean(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
+vel_diff_dyn_trial_all <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
 vel_diff_dyn_trial_all
 
 

### 2020-12-16 paper/manuscript.Rmd
correct some stats reports
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 All effects are reported as significant at *p* < .05.
-We first report results for performance the piece with the notated expression of articulation (*Figure \ref{fig:stimuli}, B*).
-followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli}, C*).
+We first report results for performance the piece with the notated expression of articulation (*Figure \ref{fig:stimuli}, B*), followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli}, C*).
 
@@
 These points could be structurally important in order to make a contrast between forte and piano.
-We calculated a KV difference for each interval subtracting the KV value of the current note from that of the following note.
+We calculated a KV difference for each interval by subtracting the KV value of the current note from that of the following note.
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
@@
 vel_diff_dyn_1 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Exp 1") +
-  ylim(-40, 45) +
+  ylim(-40, 50) +
   stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 45, paired = TRUE) +
@@
 to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics.
-With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspect of the technique.
+With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspects of the techniques.
 Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
 Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
-These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
+These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques in a simple musical scale.
 
-One of the limitations of Experiment 1 is that we employed a simple musical scale, which was good to gain an experimental control but not a naturalistic stimulus to implement artistic expressions.
+One of the limitations of Experiment 1 is that we employed a simple musical scale, which was suitable to gain an experimental control.
+However the stimulus was not a naturalistic piece to implement artistic expressions.
 In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment.
@@
 As a result, only one 8th note in the 4th measure without any expression was not included to calculate dependent variables.
-Also, IOIs were normalised by their preferred tempo because each participant chose different tempi from the 3 options.
+IOIs were normalised by their preferred tempo because each participant chose different tempi from the 3 options.
+Key-overlap Ratio (KOR) was calculated by dividing KOT by the mean IOI of each performance in order to normalise KOT due to a different tempo for each participant.
 Three trials were entirely excluded from data analysis because participants did not follow sheet music.
@@
 
+\clearpage
+
 ## 3.2. Results (Articulation)
@@
 
-As Experiment 1, we first report results for performance the piece with the notated expression of articulation (*Figure \ref{fig:stimuli-2}, B*).
-followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli-2}, C*).
+As Experiment 1, we first report results for performance of the piece with the notated expression of articulation (*Figure \ref{fig:stimuli-2}, B*), followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli-2}, C*).
 
@@
 
-A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`
+A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`
 , *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
@@
 staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`; *Figure \ref{fig:kor-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching and performing condition (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition[*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
 
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[2,5])`; *Figure \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
-Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_art_aov$anova_table[3,5])`).
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_art_aov$anova_table[2,5])`; *Figure \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
+Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`).
 
@@
 
-A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with notated articulation (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
-A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_wilcoxon$p.value))}`).
+A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with notated articulation (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
+A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`).
 
@@
 performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition than in the performing condition (Forte; *p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`, Piano; *p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`).
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-2}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition [Forte; *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano; *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte; *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano; *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition (Forte; *p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`, Piano; *p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`).
 
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-diff}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`).
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-diff}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`] (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`).
 
@@
 vel_diff_dyn_2 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Exp 2") +
-  ylim(-40, 45) +
+  ylim(-40, 50) +
   stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 45, paired = TRUE) +
@@
 piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:kor-2}*).
+Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:kor-2}*).
 
@@
 Again, there was no significant difference in legato between the two conditions as Experiment 1.
-In term of KV (dynamics), participants produced louder forte and softer piano, which showed a clearer pattern of exaggeration compared to Experiment 1 where only softer piano was observed.
-Furthermore, we again found that participants made a larger contrast between forte and piano bidirectionally at transition points.
-To sum, Experiment 2 confirmed that participants modified their performance in systematic and fine-grained ways in order to teach expressive techniques with a naturalistic piece of music.
+In term of KV (dynamics), participants produced louder forte and softer piano, which showed a clearer pattern of exaggeration compared to Experiment 1, where only softer piano was observed.
+Furthermore, we found that participants made a larger contrast between forte and piano bidirectionally at transition points.
+To sum, Experiment 2 confirmed that participants modified their performance in systematic and fine-grained ways in order to teach expressive techniques in a naturalistic piece of music.
 

### 2020-12-16 paper/manuscript.tex
correct some stats reports
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 All effects are reported as significant at \emph{p} \textless{} .05.
-We first report results for performance the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli}, B}).
-followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli}, C}).
+We first report results for performance the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli}, B}), followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli}, C}).
 
@@
 These points could be structurally important in order to make a contrast between forte and piano.
-We calculated a KV difference for each interval subtracting the KV value of the current note from that of the following note.
+We calculated a KV difference for each interval by subtracting the KV value of the current note from that of the following note.
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032; \emph{Figure \ref{fig:vel-diff}}).
@@
 Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.~to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics.
-With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspect of the technique.
+With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspects of the techniques.
 Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
 Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
-These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
+These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques in a simple musical scale.
 
-One of the limitations of Experiment 1 is that we employed a simple musical scale, which was good to gain an experimental control but not a naturalistic stimulus to implement artistic expressions.
+One of the limitations of Experiment 1 is that we employed a simple musical scale, which was suitable to gain an experimental control.
+However the stimulus was not a naturalistic piece to implement artistic expressions.
 In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment.
@@
 
+\clearpage
+
 \hypertarget{results-articulation-1}{%
@@
 
-As Experiment 1, we first report results for performance the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli-2}, B}).
-followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli-2}, C}).
+As Experiment 1, we first report results for performance of the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli-2}, B}), followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli-2}, C}).
 
@@
 
-A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.972, \emph{SD} = 0.042{]} than in the performing condition {[}\emph{M} = 0.958, \emph{SD} = 0.034{]} while playing the piece with notated articulation (\emph{t}(18) = 2.64
+A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.97, \emph{SD} = 0.042{]} than in the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.034{]} while playing the piece with notated articulation (\emph{t}(18) = 2.64
 , \emph{p} = 0.017; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.026).
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 2324, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 6.90, \emph{p} = 0.017, \(\eta_G^2\) = 0.01; \emph{Figure \ref{fig:kor-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition than in the performing condition (\emph{p} = 0.001) while there was no significant difference in legato between the teaching and performing condition (\emph{p} = 0.85).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -0.77, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.001) while there was no significant difference in legato between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition{[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.85).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.011, \(\eta_G^2\) = 0.0400; \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
-Neither the main effect of Condition (\emph{F}(1,18) = 0.46, \emph{p} = 0.506, \(\eta_G^2\) = 0.0008) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.799, \(\eta_G^2\) = 0.0001).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.011, \(\eta_G^2\) = 0.04; \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
+Neither the main effect of Condition (\emph{F}(1,18) = 0.46, \emph{p} = 0.51, \(\eta_G^2\) = 0.001) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.80, \(\eta_G^2\) = 0.000).
 
@@
 
-A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.962, \emph{SD} = 0.041{]} and the performing condition {[}\emph{M} = 0.961, \emph{SD} = 0.037{]} while playing the piece with notated articulation (\emph{t}(18) = 0.37, \emph{p} = 0.719; \emph{Figure \ref{fig:ioi-2}}).
-A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (\emph{p} = 0.922).
+A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.041{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.037{]} while playing the piece with notated articulation (\emph{t}(18) = 0.37, \emph{p} = 0.72; \emph{Figure \ref{fig:ioi-2}}).
+A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (\emph{p} = 0.92).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126.10, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7190) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.0217; \emph{Figure \ref{fig:vel-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition than in the performing condition (Forte; \emph{p} = 0.036, Piano; \emph{p} = 0.028).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7190) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.022; \emph{Figure \ref{fig:vel-2}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition {[}Forte; \emph{M} = 82.55, \emph{SD} = 8.57, Piano; \emph{M} = 60.41, \emph{SD} = 3.83{]} than in the performing condition {[}Forte; \emph{M} = 80.71, \emph{SD} = 7.53, Piano; \emph{M} = 62.34, \emph{SD} = 4.93{]} than in the performing condition (Forte; \emph{p} = 0.036, Piano; \emph{p} = 0.028).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 110.64, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8167) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.0641; \emph{Figure \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.017, PtoF; \emph{p} = 0.003).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 111, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.0641; \emph{Figure \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition {[}FtoP; \emph{M} = -17.35, \emph{SD} = 8.46, PtoF; \emph{M} = 20.60, \emph{SD} = 9.10{]} than in the performing condition {[}FtoP; \emph{M} = -13.71, \emph{SD} = 7.73, PtoF; \emph{M} = 15.87, \emph{SD} = 7.47{]} (FtoP; \emph{p} = 0.017, PtoF; \emph{p} = 0.003).
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 559, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.93), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,18) = 0.016, \emph{p} = 0.902, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 0.576, \emph{p} = 0.458, \(\eta_G^2\) = 0.00; \emph{Figure \ref{fig:kor-2}}).
+Neither the main effect of Condition (\emph{F}(1,18) = 0.016, \emph{p} = 0.90, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 0.58, \emph{p} = 0.46, \(\eta_G^2\) = 0.001; \emph{Figure \ref{fig:kor-2}}).
 

### 2020-12-16 paper/manuscript.Rmd
first draft
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 keywords          : "teaching, expertise, skill transmission, artistic expression, music"
-wordcount         : "X"
+wordcount         : "5022"
 
@@
 classoption       : "man"
-output:
-  papaja::apa6_pdf:
-    latex_engine: xelatex
+output: papaja::apa6_docx
 ---
@@
 
-In the present experiments, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, articulation, loudness) can be modulated at the same time.
+In the present study, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, articulation, loudness) can be modulated at the same time.
 Specific and subtle aspects of the performance need to be highlighted to teach a respective technique.
@@
 Once they could perform the piece without pitch errors twice consecutively within 5 attempts, the experiment started.
+Those who could not meet this requirement were excluded from data analysis (as a result, none of the participants were excluded).
 
 First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with notated expressions of either articulation or dynamics (*Figure \ref{fig:stimuli}, B-C*).
-As soon as they had produced the excerpt with notated expressions without pitch errors twice consecutively, the test trials began.
+As soon as they had produced the excerpt with a notated expression without pitch errors twice consecutively, the test trials began.
 
@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with a notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplemental Materials](#supplemental)*).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplemental Materials](#supplementary)*).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
@@
 The ``aov_car" function in the afex package was used for a repeated-measures ANOVA.
-For post-hoc comparisons, we used emmeans package.
+For post-hoc comparisons on the estimated marginal means, we used emmeans package.
 
@@
   ylim(-170, 105) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -50, paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -50, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   ylim(50, 100) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 80, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 We calculated a KV difference for each interval by subtracting the KV value of the current note from that of the following note.
+Outliers were also removed with the same criteria for other dependent variables and less than 5 % were excluded from data analysis.
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
@@
 vel_diff_dyn_1 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Exp 1") +
-  ylim(-40, 50) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 45, paired = TRUE) +
+  ylim(-40, 60) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 45, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
+Importantly, participants did not modulate their performance in term of irrelevant aspects of the techniques for teaching purposes (e.g., modulating the smoothness of sound while teaching dynamics).
 These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques in a simple musical scale.
@@
 Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+# kot exaggeration
+kot_exaggeration_1 <- data.table("Legato" == kot_art_all$Mean)
+
 # combine
@@
 We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 1 participants due to insufficient motor skills (1) and an overall deviated tempo (1).
+For data analysis, we excluded 1 participants due to insufficient motor skills (1) and an overall deviated tempo (1; outside 2 standard deviations from the average tempo across participants).
 Nineteen participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in data analysis.
@@
 In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music.
-As Experiment 1, those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis.
+As Experiment 1, those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis (as a result, one participant was excluded).
 The rest of the procedure was identical to Experiment 1.
@@
 
-Data cleaning, preprocessing and statistical analysis were identical as Experiment 1.
+Data cleaning, preprocessing and statistical analysis were almost identical as Experiment 1.
 For statistical analysis, only 8th notes with expressive notations were included.
@@
 Key-overlap Ratio (KOR) was calculated by dividing KOT by the mean IOI of each performance in order to normalise KOT due to a different tempo for each participant.
+Additionally, we include KV difference (i.e., KV difference for each interval) at transition points (i.e., forte to piano or piano to forte) as a dependent variable based on the findings of Experiment 1.
 Three trials were entirely excluded from data analysis because participants did not follow sheet music.
@@
   geom_hline(yintercept = 1, linetype = "dashed") +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 1.05, paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 1.05, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   ylim(-1, 0.4) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -0.5, paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -0.5, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   ylim(c(50, 100)) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-2}*).
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-2}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition [Forte; *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano; *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte; *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano; *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition (Forte; *p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`, Piano; *p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`).
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.4f', vel_diff_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-diff}*).
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-diff}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`] (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`).
@@
 vel_diff_dyn_2 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Exp 2") +
-  ylim(-40, 50) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 45, paired = TRUE) +
+  ylim(-40, 60) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 45, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 Furthermore, we found that participants made a larger contrast between forte and piano bidirectionally at transition points.
-To sum, Experiment 2 confirmed that participants modified their performance in systematic and fine-grained ways in order to teach expressive techniques in a naturalistic piece of music.
+Importantly, participants did not modulate their performance in term of irrelevant aspects of the techniques for teaching purposes as we found in Experiment 1.
+To sum, Experiment 2 demonstrated consistent results of systematic and fine-grained ways of didactic performance even in a naturalistic piece of music.
 
@@
 When participants played the piece with dynamics, there was no difference between the two conditions with regard to KOT/KOR, which indicates that participants did not modulate irrelevant properties of the sound.
-However, participants made a larger contrast between forte and piano for teaching purposes to highlight the transition of dynamics.
+Also, participants made a larger contrast between forte and piano for teaching purposes to highlight the transition of dynamics.
 We expected there would be an overall exaggeration in KV while teaching dynamics and indeed found that participants produced louder forte and softer piano, but only in Experiment 2.
@@
 ### Experiment 1
-In the teaching condition, participants were shown with the following instruction in front of the monitor (Italic sentences were highlighted with a yellow colour on a black background.); Now, play what you practised as if you were teaching it to students.
+In the teaching condition, participants were shown with the following instruction on a computer monitor (Italic sentences were highlighted with a yellow colour on a black background); Now, play what you practised as if you were teaching it to students.
 Students already know how to produce the sequence of the  tones and now are trying to *learn how to perform the piece expressively* by listening to your performance.
@@
 
-In the performing condition, participants were displayed with the following instruction in front of the monitor; Now, play what you practised as if you were performing it to an audience.
+In the performing condition, participants were displayed with the following instruction on a computer monitor monitor; Now, play what you practised as if you were performing it to an audience.
 *Do your best as a performer to produce the piece according to the notation that you just practised*.
@@
 
-In the performing condition, participants were displayed with the following instruction in front of the monitor (Italic sentences were highlighted with a yellow colour on a black background.); Now, play what you practised as if you were performing it to an audience.
+In the performing condition, participants were displayed with the following instruction on a computer monitor; Now, play what you practised as if you were performing it to an audience.
 *Perform the piece expressively with your interpretation*.

### 2020-12-16 paper/manuscript.tex
first draft
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Correspondence concerning this article should be addressed to Atsuko Tominaga, Quellenstraße 51, 1100 Vienna, Austria. E-mail: Tominaga\_Atsuko@phd.ceu.edu}
-\keywords{teaching, expertise, skill transmission, artistic expression, music\newline\indent Word count: X}
+\keywords{teaching, expertise, skill transmission, artistic expression, music\newline\indent Word count: 5220}
 \usepackage{csquotes}
@@
 Once they could perform the piece without pitch errors twice consecutively within 5 attempts, the experiment started.
+Those who could not meet this requirement were excluded from data analysis (as a result, none of the participants were excluded).
 
 First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with notated expressions of either articulation or dynamics (\emph{Figure \ref{fig:stimuli}, B-C}).
-As soon as they had produced the excerpt with notated expressions without pitch errors twice consecutively, the test trials began.
+As soon as they had produced the excerpt with a notated expression without pitch errors twice consecutively, the test trials began.
 
@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with a notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplemental}{Supplemental Materials}}).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplemental Materials}}).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
@@
 The ``aov\_car" function in the afex package was used for a repeated-measures ANOVA.
-For post-hoc comparisons, we used emmeans package.
+For post-hoc comparisons on the estimated marginal means, we used emmeans package.
 
@@
 We calculated a KV difference for each interval by subtracting the KV value of the current note from that of the following note.
+Outliers were also removed with the same criteria for other dependent variables and less than 5 \% were excluded from data analysis.
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032; \emph{Figure \ref{fig:vel-diff}}).
@@
 We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 1 participants due to insufficient motor skills (1) and an overall deviated tempo (1).
+For data analysis, we excluded 1 participants due to insufficient motor skills (1) and an overall deviated tempo (1; outside 2 standard deviations from the average tempo across participants).
 Nineteen participants (9 female) were included in data analysis.
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7190) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.022; \emph{Figure \ref{fig:vel-2}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.022; \emph{Figure \ref{fig:vel-2}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition {[}Forte; \emph{M} = 82.55, \emph{SD} = 8.57, Piano; \emph{M} = 60.41, \emph{SD} = 3.83{]} than in the performing condition {[}Forte; \emph{M} = 80.71, \emph{SD} = 7.53, Piano; \emph{M} = 62.34, \emph{SD} = 4.93{]} than in the performing condition (Forte; \emph{p} = 0.036, Piano; \emph{p} = 0.028).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 111, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.0641; \emph{Figure \ref{fig:vel-diff}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 111, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.064; \emph{Figure \ref{fig:vel-diff}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition {[}FtoP; \emph{M} = -17.35, \emph{SD} = 8.46, PtoF; \emph{M} = 20.60, \emph{SD} = 9.10{]} than in the performing condition {[}FtoP; \emph{M} = -13.71, \emph{SD} = 7.73, PtoF; \emph{M} = 15.87, \emph{SD} = 7.47{]} (FtoP; \emph{p} = 0.017, PtoF; \emph{p} = 0.003).
@@
 
-In the teaching condition, participants were shown with the following instruction in front of the monitor (Italic sentences were highlighted with a yellow colour on a black background.); Now, play what you practised as if you were teaching it to students.
+In the teaching condition, participants were shown with the following instruction on a computer monitor (Italic sentences were highlighted with a yellow colour on a black background); Now, play what you practised as if you were teaching it to students.
 Students already know how to produce the sequence of the tones and now are trying to \emph{learn how to perform the piece expressively} by listening to your performance.
@@
 
-In the performing condition, participants were displayed with the following instruction in front of the monitor; Now, play what you practised as if you were performing it to an audience.
+In the performing condition, participants were displayed with the following instruction on a computer monitor monitor; Now, play what you practised as if you were performing it to an audience.
 \emph{Do your best as a performer to produce the piece according to the notation that you just practised}.
@@
 
-In the performing condition, participants were displayed with the following instruction in front of the monitor (Italic sentences were highlighted with a yellow colour on a black background.); Now, play what you practised as if you were performing it to an audience.
+In the performing condition, participants were displayed with the following instruction on a computer monitor; Now, play what you practised as if you were performing it to an audience.
 \emph{Perform the piece expressively with your interpretation}.

### 2021-01-12 paper/manuscript.Rmd
delete irrelevant folders
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 output: papaja::apa6_docx
+
 ---

### 2021-06-17 paper/manuscript.Rmd
update intro and method (exp1)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 classoption       : "man"
-output: papaja::apa6_docx
+output:
+  papaja::apa6_pdf:
+    latex_engine: xelatex
 
@@
 This raises the question of whether and how experts modulate their actions during didactic demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
-Studying action modulations that serve the teaching of artistic techniques is of broader theoretical importance: It can show whether action modulations can be used to selectively highlight subtle aspects of actions that define expert performance.
+Studying action modulations that serve the teaching of artistic techniques is also of broader theoretical importance: It can show whether action modulations can be used to selectively highlight subtle aspects of actions that define expert performance.
 
 In the present study, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, articulation, loudness) can be modulated at the same time.
-Specific and subtle aspects of the performance need to be highlighted to teach a respective technique.
-We asked whether skilled pianists use sound to communicate didactic intentions.
+Specific and subtle aspects of the performance may need to be highlighted to teach a respective technique.
 Pianists were asked to perform a piece in order to demonstrate to a learner how to implement the notated expressions (teaching condition) and to play the same piece for an audience (performing condition).
@@
 
-In Experiment 1, we employed a simple musical scale to examine whether and how skilled pianists vary their performance depending on which expressive techniques (i.e., either articulation or dynamics) they were teaching.
+In Experiment 1, we employed a simple musical scale to examine whether and how skilled pianists vary their performance depending on which expressive techniques (i.e., either articulation or dynamics) they are teaching.
 Experiment 2 was conducted to replicate our findings from Experiment 1 with a more naturalistic piece of music.
@@
 
-We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2; outside 2 standard deviations from the average tempo across participants).
+We recruited 36 piano experts who played the piano for at least the past 10 years or were studying advanced piano performance at a music school at the time of recruitment.
+In order to ensure that participants had sufficient motor skills to perform the piece, we asked them to practise the piece with the non-expressive sheet music (*Figure \ref{fig:stimuli}, A*) in the lab.
+Once they could perform the piece without pitch errors twice consecutively within 5 attempts, the experiment started.
+Those who could not meet this requirement were excluded from data analysis (as a result, none of the participants were excluded based on this criterion).
+For data analysis, we excluded three participants due to experimental errors, and two participants because they deviated substantially from the prescribed tempo (outside 2 standard deviations from the average tempo across participants).
 Thirty-one participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis.
-Most participants were right-handed (left; `r nrow(dt_included[Handedness == "Left"])`, both; `r nrow(dt_included[Handedness == "Both"])`).
+Most participants were right-handed (left; `r nrow(dt_included[Handedness == "Left"])`, ambidextrous; `r nrow(dt_included[Handedness == "Both"])`).
 They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`).
@@
 Forte indicates that musical notes should be played loudly whereas piano indicates that musical notes should be played softly.
-The notation did not include any indication of fingering (i.e., the positioning of the fingers when playing the piano) because the piece was simple and pilot testing had shown that specifying fingering was not required.
+The notation did not include any indication of fingering (i.e., the positioning of the fingers when playing the piano) because the piece was simple and pilot testing had shown that specifying fingering was not necessary.
 
@@
 
-In order to ensure that participants had sufficient motor skills to perform the piece, we asked them to practise the piece with the non-expressive sheet music (*Figure \ref{fig:stimuli}, A*).
-Once they could perform the piece without pitch errors twice consecutively within 5 attempts, the experiment started.
-Those who could not meet this requirement were excluded from data analysis (as a result, none of the participants were excluded).
-
 First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with notated expressions of either articulation or dynamics (*Figure \ref{fig:stimuli}, B-C*).
-As soon as they had produced the excerpt with a notated expression without pitch errors twice consecutively, the test trials began.
+As soon as they had produced the excerpt with the notated expression without pitch errors twice consecutively, the test trials began.
 
 In the teaching condition, participants were instructed to play the excerpt of music as if they were teaching it to students.
-It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with a notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplemental Materials](#supplementary)*).
+It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplementary Materials](#supplementary)*).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
-The order of the conditions was counterbalanced across participants.
-The order of the expressive techniques within each condition was also counterbalanced across participants.
+The order of the conditions was blocked and counterbalanced across participants.
+The order of the expressive techniques within each condition was also blocked and counterbalanced across participants.
 A leading metronome (80 beats per minute, 8 beats) indicated the target tempo before each trial.
@@
 
-### 2.1.4. Design and data analysis
-
-The experiment had two factors; Condition (teaching vs.
-performing) and Expressive Technique (articulation vs.
-dynamics).
-Additionally, the factor Subcomponent was used when analysing data for each subcomponent for each technique separately (i.e., LS; legato vs.
-staccato, FP; forte vs.
-piano).
+### 2.1.4. Data analysis
 
 Three dependent variables were computed for data analysis.
-Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a tempo of tempo.
+Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
 Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone, and measures the smoothness of musical sequences [@bresin_2000].
@@
 
-Data cleaning, preprocessing and statistical analysis were performed in R version 3.6.3.
+Data cleaning, preprocessing and statistical analysis were performed in R version 4.0.5.
 For statistical analysis, only 16th notes with expressive notations were included.
-Three trials were entirely excluded from data analysis because participants did not follow sheet music or complete performing till the end.
-Pitch errors were identified by comparing a sequence of musical notes produced by a participant with the sequence of musical notes according to the sheet music.
-Pitch errors included either extra, missing or substituted tones and were manually removed by using editData package.
+Overall, three trials were excluded from data analysis because participants did not follow the sheet music or stopped performing before the end.
+Pitch errors were identified by comparing the sequence of musical notes produced by a participant with the sequence of musical notes according to the sheet music.
+Pitch errors included either extra, missing or substituted tones and were manually removed by using the *editData* R package.
 For onsets, `r round(sum(error_onset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_onset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_onset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_onset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
 For offsets, `r round(sum(error_offset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
-After we removed pitch errors, outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
-Removed outliers were less than 5% of overall responses for each dependent variable.
+In addition to pitch errors, we removed outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
+For each dependent variable, this resulted in less than 5% of overall responses being removed as outliers.
 
@@
 piano, respectively).
-The ``aov_car" function in the afex package was used for a repeated-measures ANOVA.
-For post-hoc comparisons on the estimated marginal means, we used emmeans package.
+The *afex* R package was used for a repeated-measures ANOVA.
+For post-hoc comparisons on the estimated marginal means, we used the *emmeans* R package.
 

### 2021-06-17 paper/manuscript.tex
update intro and method (exp1)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   man,floatsintext]{apa6}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
   pdftitle={The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices},
+  pdfauthor={Atsuko Tominaga1, Günther Knoblich1, \& Natalie Sebanz1},
   pdflang={en-EN},
@@
   {}{\typeout{Failed to patch abstract.}}
+\patchcmd{\HyOrg@maketitle}
+  {\section{\protect\normalfont{\@title}}}
+  {\section*{\protect\normalfont{\@title}}}
+  {}{\typeout{Failed to patch title.}}
 \makeatother
 \shorttitle{The Sound of Teaching Music}
-\author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
-\affiliation{
-\vspace{0.5cm}
-\textsuperscript{1} Department of Cognitive Science, Central European University}
-\authornote{
-
-Correspondence concerning this article should be addressed to Atsuko Tominaga, Quellenstraße 51, 1100 Vienna, Austria. E-mail: Tominaga\_Atsuko@phd.ceu.edu}
-\keywords{teaching, expertise, skill transmission, artistic expression, music\newline\indent Word count: 5220}
+\keywords{teaching, expertise, skill transmission, artistic expression, music\newline\indent Word count: 5022}
 \usepackage{csquotes}
-\ifxetex
+\ifXeTeX
   % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
@@
 \else
-  \usepackage[shorthands=off,main=english]{babel}
+  \usepackage[main=english]{babel}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 \fi
-\ifluatex
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
@@
 \setlength{\cslhangindent}{1.5em}
-\newenvironment{cslreferences}%
-  {\setlength{\parindent}{0pt}%
-  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
-  {\par}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
 \title{The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices}
+\author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
+\date{}
 
-\date{}
+\authornote{
+
+Correspondence concerning this article should be addressed to Atsuko Tominaga, Quellenstraße 51, 1100 Vienna, Austria.
+E-mail: \href{mailto:Tominaga_Atsuko@phd.ceu.edu}{\nolinkurl{Tominaga\_Atsuko@phd.ceu.edu}}
+
+}
+
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Cognitive Science, Central European University}
 
@@
 
-We recruited 36 piano experts who played the piano for at least the past 10 years or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 5 participants due to experimental errors (3), an overall deviated tempo (2; outside 2 standard deviations from the average tempo across participants).
+We recruited 36 piano experts who played the piano for at least the past 10 years or were studying advanced piano performance at a music school at the time of recruitment.
+In order to ensure that participants had sufficient motor skills to perform the piece, we asked them to practise the piece with the non-expressive sheet music (\emph{Figure \ref{fig:stimuli}, A}) in the lab.
+Once they could perform the piece without pitch errors twice consecutively within 5 attempts, the experiment started.
+Those who could not meet this requirement were excluded from data analysis (as a result, none of the participants were excluded based on this criterion).
+For data analysis, we excluded three participants due to experimental errors, and two participants because they deviated substantially from the prescribed tempo (outside 2 standard deviations from the average tempo across participants).
 Thirty-one participants (15 female) were included in data analysis.
-Most participants were right-handed (left; 2, both; 2).
+Most participants were right-handed (left; 2, ambidextrous; 2).
 They had 12.45 years of practice on average (\emph{SD} = 5.66).
@@
 
-In order to ensure that participants had sufficient motor skills to perform the piece, we asked them to practise the piece with the non-expressive sheet music (\emph{Figure \ref{fig:stimuli}, A}).
-Once they could perform the piece without pitch errors twice consecutively within 5 attempts, the experiment started.
-Those who could not meet this requirement were excluded from data analysis (as a result, none of the participants were excluded).
-
 First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with notated expressions of either articulation or dynamics (\emph{Figure \ref{fig:stimuli}, B-C}).
-As soon as they had produced the excerpt with a notated expression without pitch errors twice consecutively, the test trials began.
+As soon as they had produced the excerpt with the notated expression without pitch errors twice consecutively, the test trials began.
 
 In the teaching condition, participants were instructed to play the excerpt of music as if they were teaching it to students.
-It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with a notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplemental Materials}}).
+It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary Materials}}).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
-The order of the conditions was counterbalanced across participants.
-The order of the expressive techniques within each condition was also counterbalanced across participants.
+The order of the conditions was blocked and counterbalanced across participants.
+The order of the expressive techniques within each condition was also blocked and counterbalanced across participants.
 A leading metronome (80 beats per minute, 8 beats) indicated the target tempo before each trial.
@@
 
-\hypertarget{design-and-data-analysis}{%
-\subsubsection{2.1.4. Design and data analysis}\label{design-and-data-analysis}}
-
-The experiment had two factors; Condition (teaching vs.~performing) and Expressive Technique (articulation vs.~dynamics).
-Additionally, the factor Subcomponent was used when analysing data for each subcomponent for each technique separately (i.e., LS; legato vs.~staccato, FP; forte vs.~piano).
+\hypertarget{data-analysis}{%
+\subsubsection{2.1.4. Data analysis}\label{data-analysis}}
 
 Three dependent variables were computed for data analysis.
-Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a tempo of tempo.
+Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
 Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone, and measures the smoothness of musical sequences (Bresin \& Battel, 2000).
@@
 
-Data cleaning, preprocessing and statistical analysis were performed in R version 3.6.3.
+Data cleaning, preprocessing and statistical analysis were performed in R version 4.0.5.
 For statistical analysis, only 16th notes with expressive notations were included.
-Three trials were entirely excluded from data analysis because participants did not follow sheet music or complete performing till the end.
-Pitch errors were identified by comparing a sequence of musical notes produced by a participant with the sequence of musical notes according to the sheet music.
-Pitch errors included either extra, missing or substituted tones and were manually removed by using editData package.
+Overall, three trials were excluded from data analysis because participants did not follow the sheet music or stopped performing before the end.
+Pitch errors were identified by comparing the sequence of musical notes produced by a participant with the sequence of musical notes according to the sheet music.
+Pitch errors included either extra, missing or substituted tones and were manually removed by using the \emph{editData} R package.
 For onsets, 5.16\% of the trials contained at least one pitch error (extra notes; 2.78\%, missing notes; 2.24\%, substituted notes; 0.13\%).
 For offsets, 6.24\% of the trials contained at least one pitch error (extra notes; 2.78\%, missing notes; 2.29\%, substituted notes; 1.17\%).
-After we removed pitch errors, outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
-Removed outliers were less than 5\% of overall responses for each dependent variable.
+In addition to pitch errors, we removed outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
+For each dependent variable, this resulted in less than 5\% of overall responses being removed as outliers.
 
@@
 For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.~performing) and Subcomponent (LS; legato vs.~staccato or FP; forte vs.~piano, respectively).
-The ``aov\_car" function in the afex package was used for a repeated-measures ANOVA.
-For post-hoc comparisons on the estimated marginal means, we used emmeans package.
+The \emph{afex} R package was used for a repeated-measures ANOVA.
+For post-hoc comparisons on the estimated marginal means, we used the \emph{emmeans} R package.
 
@@
 
-\hypertarget{design-and-data-analysis-1}{%
-\subsubsection{3.1.4. Design and data analysis}\label{design-and-data-analysis-1}}
+\hypertarget{design-and-data-analysis}{%
+\subsubsection{3.1.4. Design and data analysis}\label{design-and-data-analysis}}
 
@@
 \hypertarget{refs}{}
-\begin{cslreferences}
-\leavevmode\hypertarget{ref-brand_2002}{}%
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-brand_2002}{}}%
 Brand, R.
@@
 (2002).
-Evidence for ``motionese'': Modifications in mothers' infant-directed action.
+Evidence for {`motionese'}: Modifications in mothers' infant-directed action.
 \emph{Developmental Science}, \emph{5}(1), 72--83.
@@
 
-\leavevmode\hypertarget{ref-bresin_2000}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bresin_2000}{}}%
 Bresin, R., \& Battel, G.
@@
 (2000).
-Articulation Strategies in Expressive Piano Performance Analysis of Legato, Staccato, and Repeated Notes in Performances of the Andante Movement of Mozart's Sonata in G Major (K 545).
+Articulation {Strategies} in {Expressive Piano Performance Analysis} of {Legato}, {Staccato}, and {Repeated Notes} in {Performances} of the {Andante Movement} of {Mozart}'s {Sonata} in {G Major} ({K} 545).
 \emph{Journal of New Music Research}, \emph{29}(3), 211--224.
@@
 
-\leavevmode\hypertarget{ref-campisi_2013}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-campisi_2013}{}}%
 Campisi, E., \& Özyürek, A.
 (2013).
-Iconicity as a communicative strategy: Recipient design in multimodal demonstrations for adults and children.
+Iconicity as a communicative strategy: {Recipient} design in multimodal demonstrations for adults and children.
 \emph{Journal of Pragmatics}, \emph{47}(1), 14--27.
@@
 
-\leavevmode\hypertarget{ref-csibra_2009}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-csibra_2009}{}}%
 Csibra, G., \& Gergely, G. (2009). Natural pedagogy. \emph{Trends in Cognitive Sciences}, \emph{13}(4), 148--153. \url{https://doi.org/10.1016/j.tics.2009.01.005}
 
-\leavevmode\hypertarget{ref-fernald_1985}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-fernald_1985}{}}%
 Fernald, A. (1985). Four-month-old infants prefer to listen to motherese. \emph{Infant Behavior and Development}, \emph{8}(2), 181--195. \url{https://doi.org/10.1016/S0163-6383(85)80005-9}
 
-\leavevmode\hypertarget{ref-fukuyama_2015}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-fukuyama_2015}{}}%
 Fukuyama, H., Qin, S., Kanakogi, Y., Nagai, Y., Asada, M., \& Myowa-Yamakoshi, M.
@@
 
-\leavevmode\hypertarget{ref-NIPS2016_6413}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-NIPS2016_6413}{}}%
 Ho, M.
@@
 (2016).
-Showing versus doing: Teaching by demonstration.
+Showing versus doing: {Teaching} by demonstration.
 In D.
@@
 3027--3035).
-Curran Associates, Inc.
+{Curran Associates, Inc.}
 
-\leavevmode\hypertarget{ref-juslin_2003}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-juslin_2003}{}}%
 Juslin, P.
@@
 (2003).
-Five Facets of Musical Expression: A Psychologist's Perspective on Music Performance.
+Five {Facets} of {Musical Expression}: {A Psychologist}'s {Perspective} on {Music Performance}.
 \emph{Psychology of Music}, \emph{31}(3), 273--302.
@@
 
-\leavevmode\hypertarget{ref-lindstrom_2003}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-lindstrom_2003}{}}%
 Lindström, E., Juslin, P.
@@
 (2003).
-``Expressivity comes from within your soul'': A questionnaire study of music students' perspectives on expressivity.
+{``{Expressivity} comes from within your soul''}: {A} questionnaire study of music students' perspectives on expressivity.
 \emph{Research Studies in Music Education}, \emph{20}(1), 23--47.
@@
 
-\leavevmode\hypertarget{ref-mcellin_2017}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-mcellin_2017}{}}%
 McEllin, L., Knoblich, G., \& Sebanz, N.
@@
 Distinct kinematic markers of demonstration and joint action coordination?
-Evidence from virtual xylophone playing.
+{Evidence} from virtual xylophone playing.
 \emph{Journal of Experimental Psychology: Human Perception and Performance}, \emph{44}(6), 885.
@@
 
-\leavevmode\hypertarget{ref-okazaki_2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-okazaki_2019}{}}%
 Okazaki, S., Muraoka, Y., \& Osu, R.
@@
 
-\leavevmode\hypertarget{ref-schaik_2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-schaik_2019}{}}%
 Schaik, J.
@@
 
-\leavevmode\hypertarget{ref-sloboda_2000}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-sloboda_2000}{}}%
 Sloboda, J. A. (2000). Individual differences in music performance. \emph{Trends in Cognitive Sciences}, \emph{4}(10), 397--403. \url{https://doi.org/10.1016/S1364-6613(00)01531-X}
 
-\leavevmode\hypertarget{ref-sperber_1995}{}%
-Sperber, D., \& Wilson, D. (1995). \emph{Relevance: Communication and cognition, 2nd ed}. Malden: Blackwell Publishing.
+\leavevmode\vadjust pre{\hypertarget{ref-sperber_1995}{}}%
+Sperber, D., \& Wilson, D. (1995). \emph{Relevance: {Communication} and cognition, 2nd ed}. {Malden}: {Blackwell Publishing}.
 
-\leavevmode\hypertarget{ref-thornton_2008}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-thornton_2008}{}}%
 Thornton, A., \& Raihani, N. J. (2008). The evolution of teaching. \emph{Animal Behaviour}, \emph{75}(6), 1823--1836. \url{https://doi.org/10.1016/j.anbehav.2007.12.014}
 
-\leavevmode\hypertarget{ref-tomasello_2016}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-tomasello_2016}{}}%
 Tomasello, M. (2016). The ontogeny of cultural learning. \emph{Current Opinion in Psychology}, \emph{8}, 1--4. \url{https://doi.org/10.1016/j.copsyc.2015.09.008}
 
-\leavevmode\hypertarget{ref-uther_2007}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-uther_2007}{}}%
 Uther, M., Knoll, M.
@@
 (2007).
-Do you speak E-NG-L-I-SH?
-A comparison of foreigner- and infant-directed speech.
+Do you speak {E}-{NG}-{L}-{I}-{SH}?
+{A} comparison of foreigner- and infant-directed speech.
 \emph{Speech Communication}, \emph{49}(1), 2--7.
 \url{https://doi.org/10.1016/j.specom.2006.10.003}
-\end{cslreferences}
+
+\end{CSLReferences}
 

### 2021-06-17 paper/manuscript.Rmd
don't save combined images
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 img_ioi_1 <- image_append(c(img_ioi_art_1, img_ioi_dyn_1))
-image_write(img_ioi_1, path = here("paper/image", "img_ioi_1.png"), format = "png")
+#image_write(img_ioi_1, path = here("paper/image", "img_ioi_1.png"), format = "png")
 img_ioi_1
@@
 img_kot_1 <- image_append(c(img_kot_art_1, img_kot_dyn_1))
-image_write(img_kot_1, path = here("paper/image", "img_kot_1.png"), format = "png")
+#image_write(img_kot_1, path = here("paper/image", "img_kot_1.png"), format = "png")
 img_kot_1
@@
 img_vel_1 <- image_append(c(img_vel_art_1, img_vel_dyn_1))
-image_write(img_vel_1, path = here("paper/image", "img_vel_1.png"), format = "png")
+#image_write(img_vel_1, path = here("paper/image", "img_vel_1.png"), format = "png")
 img_vel_1
@@
 img_ioi_2 <- image_append(c(img_ioi_art_2, img_ioi_dyn_2))
-image_write(img_ioi_2, path = here("paper/image", "img_ioi_2.png"), format = "png")
+#image_write(img_ioi_2, path = here("paper/image", "img_ioi_2.png"), format = "png")
 img_ioi_2
@@
 img_kor_2 <- image_append(c(img_kor_art_2, img_kor_dyn_2))
-image_write(img_kor_2, path = here("paper/image", "img_kor_2.png"), format = "png")
+#image_write(img_kor_2, path = here("paper/image", "img_kor_2.png"), format = "png")
 img_kor_2
@@
 img_vel_2 <- image_append(c(img_vel_art_2, img_vel_dyn_2))
-image_write(img_vel_2, path = here("paper/image", "img_vel_2.png"), format = "png")
+#image_write(img_vel_2, path = here("paper/image", "img_vel_2.png"), format = "png")
 img_vel_2
@@
 img_vel_diff <- image_append(c(img_vel_diff_dyn_1, img_vel_diff_dyn_2))
-image_write(img_vel_diff, path = here("paper/image", "img_vel_diff.png"), format = "png")
+#image_write(img_vel_diff, path = here("paper/image", "img_vel_diff.png"), format = "png")
 img_vel_diff

### 2021-06-17 paper/manuscript.Rmd
update general discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-The aim of the study was to investigate how expert pianists could modulate their performance in order to teach expressive artistic techniques.
+The aim of the study was to investigate how expert pianists modulate their performance in order to teach expressive techniques.
 Overall, the findings of the two experiments showed that pianists performed one and the same piece differently depending on whether they played with the intention to teach expressive techniques or with the intention to perform the piece for an audience.
 When playing with the intention to teach, pianists selectively highlighted relevant aspects of artistic expressions.
-In particular, they exaggerated staccato when playing with the intention to teach expressions concerning articulation, and they exaggerated changes between forte and piano (in both directions) when playing with the intention to teach expressions concerning dynamics.
+Across two different pieces, they exaggerated staccato when playing with the intention to teach expressions concerning articulation.
+The lack of a difference for legato when teaching articulation might stem from a ceiling effect.
+When playing with the intention to teach expressions concerning dynamics, participants exaggerated changes between forte and piano (in both directions).
+This constitutes an effective way to highlight the technique as loudness is determined relatively.
+Taken together, our findings demonstrate that expert pianists systematically modulated their sound depending on the specific technique they were trying to convey.
 
 Although participants tended to play slower in the teaching condition than in the performing condition in general, we found a significant difference only for articulation, not for dynamics.
-As we mentioned in Experiment 1, it could be that the metronome beats given prior to each performance discouraged participants from deviating from the prescribed tempo.
+As we mentioned in the discussion of Experiment 1, it could be that the metronome beats given prior to each performance discouraged participants from deviating from the prescribed tempo.
 Another possibility is that expression is actually best taught when leaving the tempo unchanged when it is irrelevant.
-This would imply that general didactic modulations like slowing down, reported in research on parent-infant interaction, might be less useful in the context of teaching expressive artistic techniques where timing itself can be used to add an expression to a performance.
-
-Participants exaggerated articulation for teaching purposes by highlighting staccato.
-The lack of a difference for legato might stem from a ceiling effect.
-When participants played the piece with dynamics, there was no difference between the two conditions with regard to KOT/KOR, which indicates that participants did not modulate irrelevant properties of the sound.
-Also, participants made a larger contrast between forte and piano for teaching purposes to highlight the transition of dynamics.
-We expected there would be an overall exaggeration in KV while teaching dynamics and indeed found that participants produced louder forte and softer piano, but only in Experiment 2.
-The finding of making a larger contrast in dynamics at transition points suggests an effective way to highlight the technique as the loudness is determined relatively.
-Taken together, these findings demonstrate that expert pianists systematically modulated their sound depending on the specific technique they were trying to convey.
+This would imply that general didactic modulations like slowing down might be less useful in the context of teaching expressive artistic techniques where timing itself can be used to add an expression to a performance.
 
@@
 One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students.
-Given that teachers modulate their demonstration throughout online interactions with learners [@okazaki_2019], future studies are needed to investigate how experts dynamically adapt their performance to their student’s skill level and demonstrated abilities.
+Given that teachers modulate their demonstration throughout online interactions with learners [@okazaki_2019], future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities.
 
@@
 # Declarations of interest
-All authors have no conflict of interest.
+The authors have no conflict of interest.
 

### 2021-06-17 paper/manuscript.tex
update general discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students.
-Given that teachers modulate their demonstration throughout online interactions with learners (Okazaki, Muraoka, \& Osu, 2019), future studies are needed to investigate how experts dynamically adapt their performance to their student's skill level and demonstrated abilities.
+Given that teachers modulate their demonstration throughout online interactions with learners (Okazaki, Muraoka, \& Osu, 2019), future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities.
 
@@
 
-All authors have no conflict of interest.
+The authors have no conflict of interest.
 

### 2021-06-17 paper/manuscript.Rmd
updated methods (exp2)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Thirty-one participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis.
-Most participants were right-handed (left; `r nrow(dt_included[Handedness == "Left"])`, ambidextrous; `r nrow(dt_included[Handedness == "Both"])`).
+Most participants were right-handed (left: `r nrow(dt_included[Handedness == "Left"])`, ambidextrous: `r nrow(dt_included[Handedness == "Both"])`).
 They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`).
@@
 
-We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 1 participants due to insufficient motor skills (1) and an overall deviated tempo (1; outside 2 standard deviations from the average tempo across participants).
-Nineteen participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in data analysis.
-Most participants were right-handed (left; `r nrow(dt_2_included[Handedness == "Left"])`).
+We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school at the time of recruitment.
+For data analysis, we excluded one participants due to insufficient motor skills, and one because of large tempo deviations (outside 2 standard deviations from the average tempo across participants).
+Nineteen participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in the data analysis.
+Most participants were right-handed (left: `r nrow(dt_2_included[Handedness == "Left"])`).
 They had `r round(mean(dt_2_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_2_included$PianoTotalPractice), 2)`).
@@
 
-The same apparatus as Experiment 1 was used.
+The same apparatus as in Experiment 1 was used.
 We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expressions (i.e., articulation, dynamics) and was relatively simple in terms of motor skills.
-The first 12 measures of the original piece was used and modified so that it had the almost equal number of data points for each dependent variable.
-The stimulus consisted of a 12-measure isochronous melody notated in 4/4 meter.
-Non-expressive sheet was used for the purpose of practice (*Figure \ref{fig:stimuli-2}, A*).
+The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
+The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter.
+Non-expressive sheet music was used for the purpose of practice (*Figure \ref{fig:stimuli-2}, A*).
 Expressive notations were added to the non-expressive sheet music for the experiment (*Figure \ref{fig:stimuli-2}, B-C*).
@@
 
-We employed the same procedure as Experiment 1 with several modifications.
+We employed the same procedure as in Experiment 1 with several modifications.
 First, all participants were required to memorise the piece without expressive notations (i.e., *Figure \ref{fig:stimuli-2}, A*) prior to the experiment.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[Supplemental Materials](#supplemental)*).
-Third, participants could choose their preferred tempo from one of the 3 options (100, 110 and 120 beats per minute) as a leading metronome.
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[Supplementary Materials](#supplemental)*).
+Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 beats per minute).
+The chosen tempo was again cued by a leading metronome.
 In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music.
@@
 
-Data cleaning, preprocessing and statistical analysis were almost identical as Experiment 1.
+Data cleaning, preprocessing and statistical analysis were almost identical to Experiment 1.
 For statistical analysis, only 8th notes with expressive notations were included.
-As a result, only one 8th note in the 4th measure without any expression was not included to calculate dependent variables.
-IOIs were normalised by their preferred tempo because each participant chose different tempi from the 3 options.
-Key-overlap Ratio (KOR) was calculated by dividing KOT by the mean IOI of each performance in order to normalise KOT due to a different tempo for each participant.
+As a result, only one 8th note in the 4th measure without any expression was not included.
+IOIs were normalised by their preferred tempo because each participant chose a tempo from the three options.
+Given the different tempi, a key-overlap ratio (KOR) was calculated by dividing KOT by the mean IOI of each performance in order to normalise KOT.
 Additionally, we include KV difference (i.e., KV difference for each interval) at transition points (i.e., forte to piano or piano to forte) as a dependent variable based on the findings of Experiment 1.
-Three trials were entirely excluded from data analysis because participants did not follow sheet music.
-With the same method as Experiment 1, pitch errors were removed manually.
+Three trials were entirely excluded from data analysis because participants did not follow the sheet music.
+Using the same approach as in Experiment 1, pitch errors were removed manually.
 For onsets, `r round(sum(error_onset_summary_2$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_onset_summary_2[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_onset_summary_2[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_onset_summary_2[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
 For offsets, `r round(sum(error_offset_summary_2$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_offset_summary_2[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_offset_summary_2[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_offset_summary_2[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
-Removed outliers (i.e., responses outside 3 standard deviations from the mean) were less than 5% of overall responses for each dependent variable.
+For each dependent variable, removing outliers (i.e., responses outside 3 standard deviations from the mean) resulted in less than 5% of overall responses being removed.
 

### 2021-06-17 paper/manuscript.tex
updated methods (exp2)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Thirty-one participants (15 female) were included in data analysis.
-Most participants were right-handed (left; 2, ambidextrous; 2).
+Most participants were right-handed (left: 2, ambidextrous: 2).
 They had 12.45 years of practice on average (\emph{SD} = 5.66).
@@
 
-We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying piano performance at a music school at the time of recruitment.
-For data analysis, we excluded 1 participants due to insufficient motor skills (1) and an overall deviated tempo (1; outside 2 standard deviations from the average tempo across participants).
-Nineteen participants (9 female) were included in data analysis.
-Most participants were right-handed (left; 2).
+We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school at the time of recruitment.
+For data analysis, we excluded one participants due to insufficient motor skills, and one because of large tempo deviations (outside 2 standard deviations from the average tempo across participants).
+Nineteen participants (9 female) were included in the data analysis.
+Most participants were right-handed (left: 2).
 They had 15.65 years of practice on average (\emph{SD} = 5.67).
@@
 
-The same apparatus as Experiment 1 was used.
+The same apparatus as in Experiment 1 was used.
 We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expressions (i.e., articulation, dynamics) and was relatively simple in terms of motor skills.
-The first 12 measures of the original piece was used and modified so that it had the almost equal number of data points for each dependent variable.
-The stimulus consisted of a 12-measure isochronous melody notated in 4/4 meter.
-Non-expressive sheet was used for the purpose of practice (\emph{Figure \ref{fig:stimuli-2}, A}).
+The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
+The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter.
+Non-expressive sheet music was used for the purpose of practice (\emph{Figure \ref{fig:stimuli-2}, A}).
 Expressive notations were added to the non-expressive sheet music for the experiment (\emph{Figure \ref{fig:stimuli-2}, B-C}).
@@
 
-We employed the same procedure as Experiment 1 with several modifications.
+We employed the same procedure as in Experiment 1 with several modifications.
 First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Figure \ref{fig:stimuli-2}, A}) prior to the experiment.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplemental}{Supplemental Materials}}).
-Third, participants could choose their preferred tempo from one of the 3 options (100, 110 and 120 beats per minute) as a leading metronome.
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplemental}{Supplementary Materials}}).
+Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 beats per minute).
+The chosen tempo was again cued by a leading metronome.
 In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music.
@@
 
-Data cleaning, preprocessing and statistical analysis were almost identical as Experiment 1.
+Data cleaning, preprocessing and statistical analysis were almost identical to Experiment 1.
 For statistical analysis, only 8th notes with expressive notations were included.
-As a result, only one 8th note in the 4th measure without any expression was not included to calculate dependent variables.
-IOIs were normalised by their preferred tempo because each participant chose different tempi from the 3 options.
-Key-overlap Ratio (KOR) was calculated by dividing KOT by the mean IOI of each performance in order to normalise KOT due to a different tempo for each participant.
+As a result, only one 8th note in the 4th measure without any expression was not included.
+IOIs were normalised by their preferred tempo because each participant chose a tempo from the three options.
+Given the different tempi, a key-overlap ratio (KOR) was calculated by dividing KOT by the mean IOI of each performance in order to normalise KOT.
 Additionally, we include KV difference (i.e., KV difference for each interval) at transition points (i.e., forte to piano or piano to forte) as a dependent variable based on the findings of Experiment 1.
-Three trials were entirely excluded from data analysis because participants did not follow sheet music.
-With the same method as Experiment 1, pitch errors were removed manually.
+Three trials were entirely excluded from data analysis because participants did not follow the sheet music.
+Using the same approach as in Experiment 1, pitch errors were removed manually.
 For onsets, 3.32\% of the trials contained at least one pitch error (extra notes; 1.66\%, missing notes; 1.57\%, substituted notes; 0.09\%).
 For offsets, 5.11\% of the trials contained at least one pitch error (extra notes; 1.66\%, missing notes; 1.57\%, substituted notes; 1.88\%).
-Removed outliers (i.e., responses outside 3 standard deviations from the mean) were less than 5\% of overall responses for each dependent variable.
+For each dependent variable, removing outliers (i.e., responses outside 3 standard deviations from the mean) resulted in less than 5\% of overall responses being removed.
 

### 2021-06-18 paper/manuscript.Rmd
updated results and discussions
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-## 2.2. Results (Articulation)
+## 2.2. Results
 
@@
 All effects are reported as significant at *p* < .05.
-We first report results for performance the piece with the notated expression of articulation (*Figure \ref{fig:stimuli}, B*), followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli}, C*).
+We first report results for performance of the piece with the notated expression of articulation (*Figure \ref{fig:stimuli}, B*), followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli}, C*).
+To recall our predictions, if participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time should reveal exaggerated legato/staccato when teaching articulation, and key velocity should reveal amplificaions of forte and piano when teaching dynamics.
 
-### 2.2.1. Interonset Intervals (IOIs)
+### 2.2.1. Articulation: Interonset Intervals (IOIs)
 
@@
 
-### 2.2.2. Key-Overlap Time (KOT)
+### 2.2.2. Articulation: Key-Overlap Time (KOT)
 
@@
 
-### 2.2.3. Key Velocity (KV)
+### 2.2.3. Articulation: Key Velocity (KV)
 
@@
 staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`; *Figure \ref{fig:vel-1}*).
-However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment failed to find a significant difference between the teaching [Legato; *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`, Staccato; *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] and performing condition [Legato; *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`, Staccato; *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] for each subcomponent (Legato; *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[1])`, Staccato; *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching [Legato; *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`, Staccato; *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] and performing condition [Legato; *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`, Staccato; *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] for each subcomponent (Legato; *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[1])`, Staccato; *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
 
@@
 
-## 2.3. Results (Dynamics)
-### 2.3.1. Interonset Intervals (IOIs)
+### 2.2.4. Dynamics: Interonset Intervals (IOIs)
 
@@
 
-### 2.3.2. Key Velocity (KV)
+### 2.2.5. Dynamics: Key Velocity (KV)
 
@@
 
-### 2.3.3. KV Transition Points
+### 2.2.6. Dynamics: KV Transition Points
 
@@
 These points could be structurally important in order to make a contrast between forte and piano.
-We calculated a KV difference for each interval by subtracting the KV value of the current note from that of the following note.
-Outliers were also removed with the same criteria for other dependent variables and less than 5 % were excluded from data analysis.
+We calculated the KV difference for each interval by subtracting the KV value of the current note from that of the following note.
+Outliers were removed using the same criteria as for the other dependent variables and less than 5 % of the data were excluded from data analysis.
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
@@
 PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-diff}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`] (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
 
@@
 
-### 2.3.4. Key-Overlap Time (KOT)
+### 2.2.7. Dynamics: Key-Overlap Time (KOT)
 
@@
 piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`; *Figure \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`).
+Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`) was significant.
 
@@
 
-## 2.4. Discussion
+## 2.3. Discussion
 
-The findings from Experiment 1 indicated that skilled pianists were able to modify their performance for teaching purposes.
-For IOIs (tempo), we found a small but significant difference only when playing the piece with the notated articulation.
-This could be due to the leading metronome, which was given prior to each performance.
-As participants tried to follow the ideal tempo through their performances, it was possible that there was only a tiny difference between the two conditions.
-Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.
-to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics.
-With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspects of the techniques.
+The findings from Experiment 1 indicated that skilled pianists modified their performance for teaching purposes.
+For IOIs (tempo), we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
+This finding is in line with earlier studies that found slower performance of actions in a teaching context [e.g., @brand_2002; @mcellin_2017].
+However, we did not find a significant difference in tempo when participants were teaching dynamics.
+It could be that trying to keep the prescribed tempo (conveyed throught the leading metronome) limited the extent to which participants slowed down their performance during teaching.
+Another possibility is that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.
+to what extent two notes overlap) and was therefore employed when teaching articulation, whereas slower performance might not help when teaching dynamics.
+
+The results for KOT (articulation) and KV (dynamics) showed that participants successfully highlighted the relevant aspects of articulation and dynamics.
 Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
 Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
-Importantly, participants did not modulate their performance in term of irrelevant aspects of the techniques for teaching purposes (e.g., modulating the smoothness of sound while teaching dynamics).
+Importantly, participants did not modulate their performance in terms of irrelevant aspects of the techniques for teaching purposes (e.g., modulating the smoothness of sound while teaching dynamics).
 These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques in a simple musical scale.
 
-One of the limitations of Experiment 1 is that we employed a simple musical scale, which was suitable to gain an experimental control.
-However the stimulus was not a naturalistic piece to implement artistic expressions.
-In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment.
-We predict that participants will played more slowly regardless of the techniques to be taught or only when teaching articulation based on the finding from Experiment 1.
-Also, we expect that participants will be able to highlight only the relevant aspects of the techniques to be taught.
-More precisely, they should be able to produce longer legato and shorter staccato while teaching articulation.
+In Experiment 1, we employed a simple musical scale in order to maximize experimental control.
+To test whether our results generalise to a more naturalistic piece, in Experiment 2, we chose an actual piano pieces and modified it for the purpose of the experiment.
+If pianists selectively highlight only the relevant aspects of the techniques to be taught also in a more naturalistic piece containing more opportunities for expression, they should again produce longer legato and shorter staccato while teaching articulation.
 When teaching dynamics, they should be able to produce louder forte and softer piano.
-Given the findings we observed in Experiment 1, participants will make a larger contrast between forte and piano at transition points.
+Given the findings we observed in Experiment 1, we also predicted that participants would make a larger contrast between forte and piano at transition points, and that they might play more slowly when teaching, especially when teaching articulation.
 
@@
 
-## 3.2. Results (Articulation)
+## 3.2. Results
 
@@
 
-### 3.2.1. Interonset Intervals (IOIs)
+### 3.2.1. Articulation: Interonset Intervals (IOIs)
 
@@
 
-### 3.2.2. Key-Overlap Ratio (KOR)
+### 3.2.2. Articulation: Key-Overlap Ratio (KOR)
 
@@
 
-### 3.2.3. Key Velocity (KV)
+### 3.2.3. Articulation: Key Velocity (KV)
 
@@
 staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_art_aov$anova_table[2,5])`; *Figure \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
-Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`).
+Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`) was significant.
 
@@
 
-## 3.3. Results (Dynamics)
-### 3.3.1. Interonset Intervals (IOIs)
+### 3.2.4. Dynamics: Interonset Intervals (IOIs)
 
-A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with notated articulation (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
-A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`).
+A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
+A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`).
 
@@
 
-### 3.3.2. Key Velocity (KV)
+### 3.2.5. Dynamics: Key Velocity (KV)
 
@@
 piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition [Forte; *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano; *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte; *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano; *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition (Forte; *p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`, Piano; *p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) and softer piano (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`)in the teaching condition [Forte; *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano; *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte; *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano; *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition.
 
@@
 
-### 3.3.3. KV Transition Points
+### 3.2.6. KV Transition Points
 
@@
 
-### 3.3.4. Key-Overlap Ratio (KOR)
+### 3.2.7. Dynamics: Key-Overlap Ratio (KOR)
 
@@
 piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:kor-2}*).
+Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:kor-2}*) was significant.
 
@@
 
-## 3.4. Discussion
+## 3.3. Discussion
 
-The results from Experiment 2 demonstrated that we successfully replicated our earlier findings that skilled pianists were able to modulate their performance for teaching purposes.
-As Experiment 1, we observed slower performance only while teaching the notated articulation.
-With regard to KOR (articulation), participants highlighted articulation by producing shorter staccato.
-Again, there was no significant difference in legato between the two conditions as Experiment 1.
-In term of KV (dynamics), participants produced louder forte and softer piano, which showed a clearer pattern of exaggeration compared to Experiment 1, where only softer piano was observed.
-Furthermore, we found that participants made a larger contrast between forte and piano bidirectionally at transition points.
-Importantly, participants did not modulate their performance in term of irrelevant aspects of the techniques for teaching purposes as we found in Experiment 1.
-To sum, Experiment 2 demonstrated consistent results of systematic and fine-grained ways of didactic performance even in a naturalistic piece of music.
+The results from Experiment 2 replicated our earlier findings and provide further evidence that skilled pianists were able to modulate their performance for teaching purposes.
+As in Experiment 1, we observed slower performance only while teaching the notated articulation.
+Again, KOR showed that when teaching articulation, participants produced shorter staccato, while there was no significant difference in legato between the two conditions as Experiment 1.
+While teaching dynamics, KV showed that participants produced louder forte and softer piano.
+This pattern of exaggeration was more pronounced than in Experiment 1, where only louder forte was observed.
+Furthermore, we again found that when teaching dynamics, participants made a larger contrast between forte and piano bidirectionally at transition points.
+Importantly, participants did not modulate performance aspects that were irrelevant with regard to the techniques to be taught (i.e., no modulation of dynamics when teaching articulation and vice versa).
+Overall, Experiment 2 demonstrated systematic and fine-grained didactic performance modulations for a naturalistic piece of music.
 

### 2021-06-18 paper/manuscript.tex
updated results and discussions
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-\hypertarget{results-articulation}{%
-\subsection{2.2. Results (Articulation)}\label{results-articulation}}
+\hypertarget{results}{%
+\subsection{2.2. Results}\label{results}}
 
 All effects are reported as significant at \emph{p} \textless{} .05.
-We first report results for performance the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli}, B}), followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli}, C}).
+We first report results for performance of the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli}, B}), followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli}, C}).
+To recall our predictions, if participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time should reveal exaggerated legato/staccato when teaching articulation, and key velocity should reveal amplificaions of forte and piano when teaching dynamics.
 
-\hypertarget{interonset-intervals-iois}{%
-\subsubsection{2.2.1. Interonset Intervals (IOIs)}\label{interonset-intervals-iois}}
+\hypertarget{articulation-interonset-intervals-iois}{%
+\subsubsection{2.2.1. Articulation: Interonset Intervals (IOIs)}\label{articulation-interonset-intervals-iois}}
 
@@
 
-\hypertarget{key-overlap-time-kot}{%
-\subsubsection{2.2.2. Key-Overlap Time (KOT)}\label{key-overlap-time-kot}}
+\hypertarget{articulation-key-overlap-time-kot}{%
+\subsubsection{2.2.2. Articulation: Key-Overlap Time (KOT)}\label{articulation-key-overlap-time-kot}}
 
@@
 
-\hypertarget{key-velocity-kv}{%
-\subsubsection{2.2.3. Key Velocity (KV)}\label{key-velocity-kv}}
+\hypertarget{articulation-key-velocity-kv}{%
+\subsubsection{2.2.3. Articulation: Key Velocity (KV)}\label{articulation-key-velocity-kv}}
 
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004; \emph{Figure \ref{fig:vel-1}}).
-However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment failed to find a significant difference between the teaching {[}Legato; \emph{M} = 70.34, \emph{SD} = 6.45, Staccato; \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato; \emph{M} = 70.68, \emph{SD} = 5.84, Staccato; \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato; \emph{p} = 0.70, Staccato; \emph{p} = 0.07).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Legato; \emph{M} = 70.34, \emph{SD} = 6.45, Staccato; \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato; \emph{M} = 70.68, \emph{SD} = 5.84, Staccato; \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato; \emph{p} = 0.70, Staccato; \emph{p} = 0.07).
 
-\hypertarget{results-dynamics}{%
-\subsection{2.3. Results (Dynamics)}\label{results-dynamics}}
-
-\hypertarget{interonset-intervals-iois-1}{%
-\subsubsection{2.3.1. Interonset Intervals (IOIs)}\label{interonset-intervals-iois-1}}
+\hypertarget{dynamics-interonset-intervals-iois}{%
+\subsubsection{2.2.4. Dynamics: Interonset Intervals (IOIs)}\label{dynamics-interonset-intervals-iois}}
 
@@
 
-\hypertarget{key-velocity-kv-1}{%
-\subsubsection{2.3.2. Key Velocity (KV)}\label{key-velocity-kv-1}}
+\hypertarget{dynamics-key-velocity-kv}{%
+\subsubsection{2.2.5. Dynamics: Key Velocity (KV)}\label{dynamics-key-velocity-kv}}
 
@@
 
-\hypertarget{kv-transition-points}{%
-\subsubsection{2.3.3. KV Transition Points}\label{kv-transition-points}}
+\hypertarget{dynamics-kv-transition-points}{%
+\subsubsection{2.2.6. Dynamics: KV Transition Points}\label{dynamics-kv-transition-points}}
 
@@
 These points could be structurally important in order to make a contrast between forte and piano.
-We calculated a KV difference for each interval by subtracting the KV value of the current note from that of the following note.
-Outliers were also removed with the same criteria for other dependent variables and less than 5 \% were excluded from data analysis.
+We calculated the KV difference for each interval by subtracting the KV value of the current note from that of the following note.
+Outliers were removed using the same criteria as for the other dependent variables and less than 5 \% of the data were excluded from data analysis.
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032; \emph{Figure \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition {[}FtoP; \emph{M} = -14.06, \emph{SD} = 7.20, PtoF; \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP; \emph{M} = -11.69, \emph{SD} = 5.97, PtoF; \emph{M} = 20.88, \emph{SD} = 8.67{]} (FtoP; \emph{p} = 0.002, PtoF; \emph{p} \textless{} 0.001).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP; \emph{p} = 0.002) and from piano to forte (PtoF; \emph{p} \textless{} 0.001) in the teaching condition {[}FtoP; \emph{M} = -14.06, \emph{SD} = 7.20, PtoF; \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP; \emph{M} = -11.69, \emph{SD} = 5.97, PtoF; \emph{M} = 20.88, \emph{SD} = 8.67{]}.
 
-\hypertarget{key-overlap-time-kot-1}{%
-\subsubsection{2.3.4. Key-Overlap Time (KOT)}\label{key-overlap-time-kot-1}}
+\hypertarget{dynamics-key-overlap-time-kot}{%
+\subsubsection{2.2.7. Dynamics: Key-Overlap Time (KOT)}\label{dynamics-key-overlap-time-kot}}
 
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31; \emph{Figure \ref{fig:kot-1}}), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.520, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.000, \emph{p} = 0.99, \(\eta_G^2\) = 0.000).
+Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.520, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.000, \emph{p} = 0.99, \(\eta_G^2\) = 0.000) was significant.
 
 \hypertarget{discussion}{%
-\subsection{2.4. Discussion}\label{discussion}}
+\subsection{2.3. Discussion}\label{discussion}}
 
-The findings from Experiment 1 indicated that skilled pianists were able to modify their performance for teaching purposes.
-For IOIs (tempo), we found a small but significant difference only when playing the piece with the notated articulation.
-This could be due to the leading metronome, which was given prior to each performance.
-As participants tried to follow the ideal tempo through their performances, it was possible that there was only a tiny difference between the two conditions.
-Another possibility could be that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.~to what extent two notes overlap) when teaching articulation whereas such relation was irrelevant when teaching dynamics.
-With regard to KOT (articulation) and KV (dynamics), participants successfully highlight the relevant aspects of the techniques.
+The findings from Experiment 1 indicated that skilled pianists modified their performance for teaching purposes.
+For IOIs (tempo), we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
+This finding is in line with earlier studies that found slower performance of actions in a teaching context (e.g., Brand et al., 2002; McEllin et al., 2017).
+However, we did not find a significant difference in tempo when participants were teaching dynamics.
+It could be that trying to keep the prescribed tempo (conveyed throught the leading metronome) limited the extent to which participants slowed down their performance during teaching.
+Another possibility is that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.~to what extent two notes overlap) and was therefore employed when teaching articulation, whereas slower performance might not help when teaching dynamics.
+
+The results for KOT (articulation) and KV (dynamics) showed that participants successfully highlighted the relevant aspects of articulation and dynamics.
 Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
 Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
-Importantly, participants did not modulate their performance in term of irrelevant aspects of the techniques for teaching purposes (e.g., modulating the smoothness of sound while teaching dynamics).
+Importantly, participants did not modulate their performance in terms of irrelevant aspects of the techniques for teaching purposes (e.g., modulating the smoothness of sound while teaching dynamics).
 These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques in a simple musical scale.
 
-One of the limitations of Experiment 1 is that we employed a simple musical scale, which was suitable to gain an experimental control.
-However the stimulus was not a naturalistic piece to implement artistic expressions.
-In Experiment 2, we chose one piece of music from actual piano pieces and modified it for the purpose of the experiment.
-We predict that participants will played more slowly regardless of the techniques to be taught or only when teaching articulation based on the finding from Experiment 1.
-Also, we expect that participants will be able to highlight only the relevant aspects of the techniques to be taught.
-More precisely, they should be able to produce longer legato and shorter staccato while teaching articulation.
+In Experiment 1, we employed a simple musical scale in order to maximize experimental control.
+To test whether our results generalise to a more naturalistic piece, in Experiment 2, we chose an actual piano pieces and modified it for the purpose of the experiment.
+If pianists selectively highlight only the relevant aspects of the techniques to be taught also in a more naturalistic piece containing more opportunities for expression, they should again produce longer legato and shorter staccato while teaching articulation.
 When teaching dynamics, they should be able to produce louder forte and softer piano.
-Given the findings we observed in Experiment 1, participants will make a larger contrast between forte and piano at transition points.
+Given the findings we observed in Experiment 1, we also predicted that participants would make a larger contrast between forte and piano at transition points, and that they might play more slowly when teaching, especially when teaching articulation.
 
@@
 
-\hypertarget{results-articulation-1}{%
-\subsection{3.2. Results (Articulation)}\label{results-articulation-1}}
+\hypertarget{results-1}{%
+\subsection{3.2. Results}\label{results-1}}
 
@@
 
-\hypertarget{interonset-intervals-iois-2}{%
-\subsubsection{3.2.1. Interonset Intervals (IOIs)}\label{interonset-intervals-iois-2}}
+\hypertarget{articulation-interonset-intervals-iois-1}{%
+\subsubsection{3.2.1. Articulation: Interonset Intervals (IOIs)}\label{articulation-interonset-intervals-iois-1}}
 
@@
 
-\hypertarget{key-overlap-ratio-kor}{%
-\subsubsection{3.2.2. Key-Overlap Ratio (KOR)}\label{key-overlap-ratio-kor}}
+\hypertarget{articulation-key-overlap-ratio-kor}{%
+\subsubsection{3.2.2. Articulation: Key-Overlap Ratio (KOR)}\label{articulation-key-overlap-ratio-kor}}
 
@@
 
-\hypertarget{key-velocity-kv-2}{%
-\subsubsection{3.2.3. Key Velocity (KV)}\label{key-velocity-kv-2}}
+\hypertarget{articulation-key-velocity-kv-1}{%
+\subsubsection{3.2.3. Articulation: Key Velocity (KV)}\label{articulation-key-velocity-kv-1}}
 
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.011, \(\eta_G^2\) = 0.04; \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
-Neither the main effect of Condition (\emph{F}(1,18) = 0.46, \emph{p} = 0.51, \(\eta_G^2\) = 0.001) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.80, \(\eta_G^2\) = 0.000).
+Neither the main effect of Condition (\emph{F}(1,18) = 0.46, \emph{p} = 0.51, \(\eta_G^2\) = 0.001) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.80, \(\eta_G^2\) = 0.000) was significant.
 
-\hypertarget{results-dynamics-1}{%
-\subsection{3.3. Results (Dynamics)}\label{results-dynamics-1}}
+\hypertarget{dynamics-interonset-intervals-iois-1}{%
+\subsubsection{3.2.4. Dynamics: Interonset Intervals (IOIs)}\label{dynamics-interonset-intervals-iois-1}}
 
-\hypertarget{interonset-intervals-iois-3}{%
-\subsubsection{3.3.1. Interonset Intervals (IOIs)}\label{interonset-intervals-iois-3}}
+A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.041{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.037{]} while playing the piece with the notated dynamics (\emph{t}(18) = 0.37, \emph{p} = 0.72; \emph{Figure \ref{fig:ioi-2}}).
+A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (\emph{p} = 0.92).
 
-A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.041{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.037{]} while playing the piece with notated articulation (\emph{t}(18) = 0.37, \emph{p} = 0.72; \emph{Figure \ref{fig:ioi-2}}).
-A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two condition (\emph{p} = 0.92).
-
-\hypertarget{key-velocity-kv-3}{%
-\subsubsection{3.3.2. Key Velocity (KV)}\label{key-velocity-kv-3}}
+\hypertarget{dynamics-key-velocity-kv-1}{%
+\subsubsection{3.2.5. Dynamics: Key Velocity (KV)}\label{dynamics-key-velocity-kv-1}}
 
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.022; \emph{Figure \ref{fig:vel-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition {[}Forte; \emph{M} = 82.55, \emph{SD} = 8.57, Piano; \emph{M} = 60.41, \emph{SD} = 3.83{]} than in the performing condition {[}Forte; \emph{M} = 80.71, \emph{SD} = 7.53, Piano; \emph{M} = 62.34, \emph{SD} = 4.93{]} than in the performing condition (Forte; \emph{p} = 0.036, Piano; \emph{p} = 0.028).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (\emph{p} = 0.036) and softer piano (\emph{p} = 0.028)in the teaching condition {[}Forte; \emph{M} = 82.55, \emph{SD} = 8.57, Piano; \emph{M} = 60.41, \emph{SD} = 3.83{]} than in the performing condition {[}Forte; \emph{M} = 80.71, \emph{SD} = 7.53, Piano; \emph{M} = 62.34, \emph{SD} = 4.93{]} than in the performing condition.
 
-\hypertarget{kv-transition-points-1}{%
-\subsubsection{3.3.3. KV Transition Points}\label{kv-transition-points-1}}
+\hypertarget{kv-transition-points}{%
+\subsubsection{3.2.6. KV Transition Points}\label{kv-transition-points}}
 
@@
 
-\hypertarget{key-overlap-ratio-kor-1}{%
-\subsubsection{3.3.4. Key-Overlap Ratio (KOR)}\label{key-overlap-ratio-kor-1}}
+\hypertarget{dynamics-key-overlap-ratio-kor}{%
+\subsubsection{3.2.7. Dynamics: Key-Overlap Ratio (KOR)}\label{dynamics-key-overlap-ratio-kor}}
 
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 559, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.93), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,18) = 0.016, \emph{p} = 0.90, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 0.58, \emph{p} = 0.46, \(\eta_G^2\) = 0.001; \emph{Figure \ref{fig:kor-2}}).
+Neither the main effect of Condition (\emph{F}(1,18) = 0.016, \emph{p} = 0.90, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 0.58, \emph{p} = 0.46, \(\eta_G^2\) = 0.001; \emph{Figure \ref{fig:kor-2}}) was significant.
 
 \hypertarget{discussion-1}{%
-\subsection{3.4. Discussion}\label{discussion-1}}
+\subsection{3.3. Discussion}\label{discussion-1}}
 
-The results from Experiment 2 demonstrated that we successfully replicated our earlier findings that skilled pianists were able to modulate their performance for teaching purposes.
-As Experiment 1, we observed slower performance only while teaching the notated articulation.
-With regard to KOR (articulation), participants highlighted articulation by producing shorter staccato.
-Again, there was no significant difference in legato between the two conditions as Experiment 1.
-In term of KV (dynamics), participants produced louder forte and softer piano, which showed a clearer pattern of exaggeration compared to Experiment 1, where only softer piano was observed.
-Furthermore, we found that participants made a larger contrast between forte and piano bidirectionally at transition points.
-Importantly, participants did not modulate their performance in term of irrelevant aspects of the techniques for teaching purposes as we found in Experiment 1.
-To sum, Experiment 2 demonstrated consistent results of systematic and fine-grained ways of didactic performance even in a naturalistic piece of music.
+The results from Experiment 2 replicated our earlier findings and provide further evidence that skilled pianists were able to modulate their performance for teaching purposes.
+As in Experiment 1, we observed slower performance only while teaching the notated articulation.
+Again, KOR showed that when teaching articulation, participants produced shorter staccato, while there was no significant difference in legato between the two conditions as Experiment 1.
+While teaching dynamics, KV showed that participants produced louder forte and softer piano.
+This pattern of exaggeration was more pronounced than in Experiment 1, where only louder forte was observed.
+Furthermore, we again found that when teaching dynamics, participants made a larger contrast between forte and piano bidirectionally at transition points.
+Importantly, participants did not modulate performance aspects that were irrelevant with regard to the techniques to be taught (i.e., no modulation of dynamics when teaching articulation and vice versa).
+Overall, Experiment 2 demonstrated systematic and fine-grained didactic performance modulations for a naturalistic piece of music.
 

### 2021-06-18 paper/manuscript.Rmd
tweak intro/exp1 and update the reference
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
-In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners [@brand_2002; @fernald_1985].
+In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners [@brand_2002; @saint-georges_2013].
 These modulations include slowing down and exaggerating sounds and actions.
@@
 
-In the present study, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, articulation, loudness) can be modulated at the same time.
+In the present study, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, smoothness, loudness) can be modulated at the same time.
 Specific and subtle aspects of the performance may need to be highlighted to teach a respective technique.
@@
 We recruited 36 piano experts who played the piano for at least the past 10 years or were studying advanced piano performance at a music school at the time of recruitment.
-In order to ensure that participants had sufficient motor skills to perform the piece, we asked them to practise the piece with the non-expressive sheet music (*Figure \ref{fig:stimuli}, A*) in the lab.
-Once they could perform the piece without pitch errors twice consecutively within 5 attempts, the experiment started.
-Those who could not meet this requirement were excluded from data analysis (as a result, none of the participants were excluded based on this criterion).
 For data analysis, we excluded three participants due to experimental errors, and two participants because they deviated substantially from the prescribed tempo (outside 2 standard deviations from the average tempo across participants).
 Thirty-one participants (`r nrow(dt_included[Gender == "Female"])` female) were included in data analysis.
-Most participants were right-handed (left: `r nrow(dt_included[Handedness == "Left"])`, ambidextrous: `r nrow(dt_included[Handedness == "Both"])`).
+Most participants were right-handed (left: `r nrow(dt_included[Handedness == "Left"])`, ambidextrous: `r nrow(dt_included[Handedness == "Both"])`) with a mean age of `r round(mean(dt_included$Age), 2)` (*SD* = `r round(sd(dt_included$Age), 2)`).
 They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`).
@@
 Pitch errors included either extra, missing or substituted tones and were manually removed by using the *editData* R package.
-For onsets, `r round(sum(error_onset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_onset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_onset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_onset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
-For offsets, `r round(sum(error_offset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
-In addition to pitch errors, we removed outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
+For onsets, `r round(sum(error_onset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_onset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes: `r round(error_onset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes: `r round(error_onset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
+For offsets, `r round(sum(error_offset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes: `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes: `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
+In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
 For each dependent variable, this resulted in less than 5% of overall responses being removed as outliers.
@@
 piano, respectively).
-The *afex* R package was used for a repeated-measures ANOVA.
+The *t.test* function in the *stats* R package was used for a *t*-test.
+The *aov_car* function in the *afex* R package was used for a repeated-measures ANOVA.
 For post-hoc comparisons on the estimated marginal means, we used the *emmeans* R package.
@@
 We first report results for performance of the piece with the notated expression of articulation (*Figure \ref{fig:stimuli}, B*), followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli}, C*).
-To recall our predictions, if participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time should reveal exaggerated legato/staccato when teaching articulation, and key velocity should reveal amplificaions of forte and piano when teaching dynamics.
+To recall our predictions, if participants play slowly when they are teaching in general, interonset intervals (IOIs) should be larger when teaching.
+If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should reveal exaggerated legato/staccato when teaching articulation, and key velocity (KV) should reveal amplifications of forte and piano when teaching dynamics.
 
@@
 staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`; *Figure \ref{fig:kot-1}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition[*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
 
@@
 
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
 
@@
 piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-1}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) while there was no significant difference in piano between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition[*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) while there was no significant difference in piano between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`).
 
@@
 piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`; *Figure \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`) was significant.
+Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[3,5])`) was significant.
 
@@
 The findings from Experiment 1 indicated that skilled pianists modified their performance for teaching purposes.
-For IOIs (tempo), we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
+For IOIs, we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
 This finding is in line with earlier studies that found slower performance of actions in a teaching context [e.g., @brand_2002; @mcellin_2017].
 However, we did not find a significant difference in tempo when participants were teaching dynamics.
-It could be that trying to keep the prescribed tempo (conveyed throught the leading metronome) limited the extent to which participants slowed down their performance during teaching.
-Another possibility is that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.
-to what extent two notes overlap) and was therefore employed when teaching articulation, whereas slower performance might not help when teaching dynamics.
+It could be that trying to keep the prescribed tempo (conveyed through the leading metronome) limited the extent to which participants slowed down their performance during teaching.
+Another possibility is that slower performance may be beneficial to highlight the relation between two notes (i.e., to what extent two notes overlap) and was therefore employed when teaching articulation, whereas slower performance might not help when teaching dynamics.
 
-The results for KOT (articulation) and KV (dynamics) showed that participants successfully highlighted the relevant aspects of articulation and dynamics.
+The results for KOT and KV showed that participants successfully highlighted the relevant aspects of articulation and dynamics.
 Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
@@
 Importantly, participants did not modulate their performance in terms of irrelevant aspects of the techniques for teaching purposes (e.g., modulating the smoothness of sound while teaching dynamics).
-These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques in a simple musical scale.
+These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
 
-In Experiment 1, we employed a simple musical scale in order to maximize experimental control.
-To test whether our results generalise to a more naturalistic piece, in Experiment 2, we chose an actual piano pieces and modified it for the purpose of the experiment.
-If pianists selectively highlight only the relevant aspects of the techniques to be taught also in a more naturalistic piece containing more opportunities for expression, they should again produce longer legato and shorter staccato while teaching articulation.
-When teaching dynamics, they should be able to produce louder forte and softer piano.
+In Experiment 1, we employed a simple musical scale in order to maximise experimental control.
+To test whether our results generalise to a more naturalistic piece, in Experiment 2, we chose an actual piano piece and modified it for the purpose of the experiment.
+If pianists selectively highlight the relevant aspects of the techniques to be taught also in a more naturalistic piece containing more opportunities for expression, they should again produce longer legato and shorter staccato while teaching articulation.
+When teaching dynamics, they should produce louder forte and softer piano.
 Given the findings we observed in Experiment 1, we also predicted that participants would make a larger contrast between forte and piano at transition points, and that they might play more slowly when teaching, especially when teaching articulation.
@@
 Nineteen participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in the data analysis.
-Most participants were right-handed (left: `r nrow(dt_2_included[Handedness == "Left"])`).
+Most participants were right-handed (left: `r nrow(dt_2_included[Handedness == "Left"])`) with a mean age of `r round(mean(dt_2_included$Age), 2)` (*SD* = `r round(sd(dt_2_included$Age), 2)`).
 They had `r round(mean(dt_2_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_2_included$PianoTotalPractice), 2)`).
@@
 staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`; *Figure \ref{fig:kor-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition[*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
 

### 2021-06-18 paper/manuscript.tex
tweak intro/exp1 and update the reference
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
-In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners (Brand, Baldwin, \& Ashburn, 2002; Fernald, 1985).
+In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners (Brand, Baldwin, \& Ashburn, 2002; Saint-Georges et al., 2013).
 These modulations include slowing down and exaggerating sounds and actions.
@@
 We recruited 36 piano experts who played the piano for at least the past 10 years or were studying advanced piano performance at a music school at the time of recruitment.
-In order to ensure that participants had sufficient motor skills to perform the piece, we asked them to practise the piece with the non-expressive sheet music (\emph{Figure \ref{fig:stimuli}, A}) in the lab.
-Once they could perform the piece without pitch errors twice consecutively within 5 attempts, the experiment started.
-Those who could not meet this requirement were excluded from data analysis (as a result, none of the participants were excluded based on this criterion).
 For data analysis, we excluded three participants due to experimental errors, and two participants because they deviated substantially from the prescribed tempo (outside 2 standard deviations from the average tempo across participants).
 Thirty-one participants (15 female) were included in data analysis.
-Most participants were right-handed (left: 2, ambidextrous: 2).
+Most participants were right-handed (left: 2, ambidextrous: 2) with a mean age of 24.16 (\emph{SD} = 4.26).
 They had 12.45 years of practice on average (\emph{SD} = 5.66).
@@
 Pitch errors included either extra, missing or substituted tones and were manually removed by using the \emph{editData} R package.
-For onsets, 5.16\% of the trials contained at least one pitch error (extra notes; 2.78\%, missing notes; 2.24\%, substituted notes; 0.13\%).
-For offsets, 6.24\% of the trials contained at least one pitch error (extra notes; 2.78\%, missing notes; 2.29\%, substituted notes; 1.17\%).
-In addition to pitch errors, we removed outliers for IOIs, KOT and KV were defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
+For onsets, 5.16\% of the trials contained at least one pitch error (extra notes: 2.78\%, missing notes: 2.24\%, substituted notes: 0.13\%).
+For offsets, 6.24\% of the trials contained at least one pitch error (extra notes: 2.78\%, missing notes: 2.29\%, substituted notes: 1.17\%).
+In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
 For each dependent variable, this resulted in less than 5\% of overall responses being removed as outliers.
@@
 For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.~performing) and Subcomponent (LS; legato vs.~staccato or FP; forte vs.~piano, respectively).
-The \emph{afex} R package was used for a repeated-measures ANOVA.
+The \emph{t.test} function in the \emph{stats} R package was used for a \emph{t}-test.
+The \emph{aov\_car} function in the \emph{afex} R package was used for a repeated-measures ANOVA.
 For post-hoc comparisons on the estimated marginal means, we used the \emph{emmeans} R package.
@@
 We first report results for performance of the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli}, B}), followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli}, C}).
-To recall our predictions, if participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time should reveal exaggerated legato/staccato when teaching articulation, and key velocity should reveal amplificaions of forte and piano when teaching dynamics.
+To recall our predictions, if participants play slowly when they are teaching in general, interonset intervals (IOIs) should be larger when teaching.
+If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should reveal exaggerated legato/staccato when teaching articulation, and key velocity (KV) should reveal amplifications of forte and piano when teaching dynamics.
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011; \emph{Figure \ref{fig:kot-1}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.004) while there was no significant difference in legato between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition{[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.32).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.004) while there was no significant difference in legato between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition {[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.32).
 
@@
 
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.111; \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11; \emph{Figure \ref{fig:ioi-1}}).
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003; \emph{Figure \ref{fig:vel-1}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.021) while there was no significant difference in piano between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition{[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.45).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.021) while there was no significant difference in piano between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.45).
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31; \emph{Figure \ref{fig:kot-1}}), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.520, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.000, \emph{p} = 0.99, \(\eta_G^2\) = 0.000) was significant.
+Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.52, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.00, \emph{p} = 0.99, \(\eta_G^2\) = 0.00) was significant.
 
@@
 The findings from Experiment 1 indicated that skilled pianists modified their performance for teaching purposes.
-For IOIs (tempo), we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
+For IOIs, we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
 This finding is in line with earlier studies that found slower performance of actions in a teaching context (e.g., Brand et al., 2002; McEllin et al., 2017).
 However, we did not find a significant difference in tempo when participants were teaching dynamics.
-It could be that trying to keep the prescribed tempo (conveyed throught the leading metronome) limited the extent to which participants slowed down their performance during teaching.
-Another possibility is that slower performance may be more or only beneficial to highlight the relation between two notes (i.e.~to what extent two notes overlap) and was therefore employed when teaching articulation, whereas slower performance might not help when teaching dynamics.
+It could be that trying to keep the prescribed tempo (conveyed through the leading metronome) limited the extent to which participants slowed down their performance during teaching.
+Another possibility is that slower performance may be beneficial to highlight the relation between two notes (i.e., to what extent two notes overlap) and was therefore employed when teaching articulation, whereas slower performance might not help when teaching dynamics.
 
-The results for KOT (articulation) and KV (dynamics) showed that participants successfully highlighted the relevant aspects of articulation and dynamics.
+The results for KOT and KV showed that participants successfully highlighted the relevant aspects of articulation and dynamics.
 Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
@@
 Nineteen participants (9 female) were included in the data analysis.
-Most participants were right-handed (left: 2).
+Most participants were right-handed (left: 2) with a mean age of 25.90 (\emph{SD} = 4.68).
 They had 15.65 years of practice on average (\emph{SD} = 5.67).
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 2324, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 6.90, \emph{p} = 0.017, \(\eta_G^2\) = 0.01; \emph{Figure \ref{fig:kor-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -0.77, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.001) while there was no significant difference in legato between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition{[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.85).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -0.77, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.001) while there was no significant difference in legato between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.85).
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-fernald_1985}{}}%
-Fernald, A. (1985). Four-month-old infants prefer to listen to motherese. \emph{Infant Behavior and Development}, \emph{8}(2), 181--195. \url{https://doi.org/10.1016/S0163-6383(85)80005-9}
-
 \leavevmode\vadjust pre{\hypertarget{ref-fukuyama_2015}{}}%
@@
 Teacher-learner interaction quantifies scaffolding behaviour in imitation learning.
-\emph{Scientific Reports}, \emph{9}(1), 1--13.
+\emph{Sci Rep}, \emph{9}(1), 1--13.
 \url{https://doi.org/10.1038/s41598-019-44049-x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-saint-georges_2013}{}}%
+Saint-Georges, C., Chetouani, M., Cassel, R., Apicella, F., Mahdhaoui, A., Muratori, F., \ldots{} Cohen, D.
+(2013).
+Motherese in {Interaction}: {At} the {Cross}-{Road} of {Emotion} and {Cognition}?
+({A Systematic Review}).
+\emph{PLOS ONE}, \emph{8}(10), e78103.
+\url{https://doi.org/10.1371/journal.pone.0078103}
 

### 2021-06-18 experiment-1/analysis/preprocessor/included_summary.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
remove SubNr != 1
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 - Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
-- Note: SubNr 3, 8, 14 were already excluded due to an experimental error. SubNr 16 was removed due to a deviated tempo
+- Note: SubNr 3, 8, 14 were already excluded due to an experimental error. SubNr 1 & 16 was removed due to a deviated tempo
 
@@
 ```{r kot, echo = FALSE}
-dt_kot[SubNr != 1]
+dt_kot
 ```
@@
 ```{r kot-summary, echo = FALSE}
-dt_kot[SubNr != 1, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
+dt_kot[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
 ```
@@
 ```{r vel, echo = FALSE}
-dt_vel[SubNr != 1]
+dt_vel
 ```
@@
 ```{r vel-summary, echo = FALSE}
-dt_vel[SubNr != 1, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
+dt_vel[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
 ```
@@
 ```{r vel-diff, echo = FALSE}
-dt_vel_diff[SubNr != 1]
+dt_vel_diff
 ```
@@
 ```{r vel-diff-summary, echo = FALSE}
-dt_vel_diff[SubNr != 1, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
+dt_vel_diff[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
 ```

### 2021-06-19 experiment-1/analysis/stats/ioi.Rmd
removed styles from plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r ioi-art-box,  echo = FALSE}
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   geom_hline(yintercept = 188, linetype = "dashed")
@@
 
-ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+  geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 ```{r ioi-art-ch-box, echo = FALSE}
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   facet_grid(Boundary ~ .) +
@@
 
-ggpaired(ioi_art_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(ioi_art_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+  geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 ```{r ioi-art-trial-line, echo = FALSE, fig.height = 4}
-ggline(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggline(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   facet_wrap(SubNr ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,8,1)) +
@@
 ```{r ioi-art-trial-all-line, echo = FALSE}
-ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r ioi-dyn-box,  echo = FALSE}
-ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   geom_hline(yintercept = 188, linetype = "dashed")
@@
 
-ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+  geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 ```{r ioi-dyn-ch-box, echo = FALSE}
-ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   facet_grid(Boundary ~ .) +
@@
 
-ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+  geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 ```{r ioi-dyn-trial-line, echo = FALSE, fig.height = 4}
-ggline(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggline(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   facet_wrap(SubNr ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,8,1)) +
@@
 ```{r ioi-dyn-trial-all-line, echo = FALSE}
-ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 
-ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,66,1))
@@
 
-ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
   geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
     scale_x_continuous(breaks = seq(1,66,1))

### 2021-06-19 experiment-1/analysis/stats/ioi.R

@@ -31,8 +31,7 @@ ioi_art
 
 
 ## ----ioi-art-box,  echo = FALSE-----------------------------------------------
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + 
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + 
   geom_hline(yintercept = 188, linetype = "dashed")
 
 
@@ -45,9 +44,8 @@ ioi_art_all
 ## ----ioi-art-all-box, echo = FALSE--------------------------------------------
 ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean)
 
-ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+  geom_hline(yintercept = 188, linetype = "dashed")
 
 
 ## ----normality1, echo = FALSE-------------------------------------------------
@@ -73,8 +71,7 @@ ioi_art_ch
 
 
[…]

### 2021-06-19 experiment-1/analysis/stats/articulation.Rmd
removed styles from plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 #### Box plot
-
 ```{r kot-art-box,  echo = FALSE}
-ggboxplot(dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggboxplot(dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") +
+  facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r kot-art-all-scatter, echo = FALSE}
-ggerrorplot(kot_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggerrorplot(kot_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation")
 ```
@@
 
-ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation")
 ```
@@
 ```{r kot-art-trial-line-legato, echo = FALSE, fig.height = 4}
-ggline(dt_kot_art[Subcomponent == "Legato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_kot_art[Subcomponent == "Legato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kot-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(dt_kot_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_kot_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kot-art-trial-all-line-legato, echo = FALSE}
-ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
-  scale_x_continuous(breaks = seq(1,8,1)) +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-art-trial-all-line-staccato, echo = FALSE}
-ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
-  scale_x_continuous(breaks = seq(1,8,1)) +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
+  scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kor-art-box,  echo = FALSE}
-ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
+  facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r kor-art-all-scatter, echo = FALSE}
-ggerrorplot(kor_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggerrorplot(kor_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation")
 ```
@@
 
-ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation")
 ```
@@
 ```{r kor-art-trial-line-legato, echo = FALSE, fig.height = 4}
-ggline(dt_kor_art[Subcomponent == "Legato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Legato") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_kor_art[Subcomponent == "Legato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Legato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kor-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(dt_kor_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Staccato") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_kor_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Staccato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kor-art-trial-all-line-legato, echo = FALSE}
-ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kor-art-trial-all-line-staccato, echo = FALSE}
-ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Staccato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kot-dyn-box,  echo = FALSE}
-ggboxplot(dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Dynamics") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggboxplot(dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Dynamics") +
+  facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r kot-dyn-all-scatter, echo = FALSE}
-ggerrorplot(kot_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggerrorplot(kot_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics")
 ```
@@
 
-ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics")
 ```
@@
 ```{r kot-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-ggline(dt_kot_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_kot_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kot-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-ggline(dt_kot_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_kot_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kot-dyn-trial-all-line-forte, echo = FALSE}
-ggline(kot_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kot_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kot-dyn-trial-all-line-piano, echo = FALSE}
-ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kor-dyn-box,  echo = FALSE}
-ggboxplot(dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Dynamics") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggboxplot(dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Dynamics") +
+  facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r kor-dyn-all-scatter, echo = FALSE}
-ggerrorplot(kor_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggerrorplot(kor_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics")
 ```
@@
 
-ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics")
 ```
@@
 ```{r kor-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-ggline(dt_kor_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Forte") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_kor_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Forte") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kor-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-ggline(dt_kor_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Piano") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_kor_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Piano") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kor-dyn-trial-all-line-forte, echo = FALSE}
-ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Forte") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Forte") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kor_dyn_trial_all_line_piano, echo = FALSE}
-ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 
-ggline(kot_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kot_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Articulation") +
   scale_x_continuous(breaks = seq(1,66,1))
@@
 
-ggline(kot_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Dynamics")+
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kot_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Dynamics") +
   scale_x_continuous(breaks = seq(1,66,1)) 

### 2021-06-19 experiment-1/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -46,338 +46,308 @@ dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 
 
-## ----kot-art, echo = FALSE----------------------------------------------------
+## ----kot-art, echo = FALSE----------------------------------
 # For each individual
 kot_art <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kot_art
 
 
-## ----kot-art-box,  echo = FALSE-----------------------------------------------
-ggboxplot(dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  facet_grid(Subcomponent ~ .) +
[…]

### 2021-06-19 experiment-1/analysis/stats/dynamics.Rmd
removed styles from plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r vel-dyn-box,  echo = FALSE}
-ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
+  facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r vel-dyn-all-scatter, echo = FALSE}
-ggerrorplot(vel_dyn, x = "Condition", y = "Mean", add = c("mean", "mean_se"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggerrorplot(vel_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics")
 ```
@@
 
-ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics")
 ```
@@
 ```{r vel-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-ggline(dt_vel_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_vel_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-ggline(dt_vel_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_vel_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-dyn-trial-all-line-forte, echo = FALSE}
-ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
     scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-dyn-trial-all-line-piano, echo = FALSE}
-ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
     scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-dyn-box,  echo = FALSE}
-ggboxplot(dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Dynamics") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggboxplot(dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Dynamics") +
+  facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r vel-diff-dyn-all-scatter, echo = FALSE}
-ggerrorplot(vel_diff_dyn, x = "Condition", y = "Mean", add = c("mean", "mean_se"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggerrorplot(vel_diff_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics")
 ```
@@
 
-ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics")
 ```
@@
 ```{r post-hoc-2, echo = FALSE}
-vel_diff_dyn_posthoc <- test(pairs(emmeans(vel_diff_dyn_aov,~Condition|Subcomponent), adjust = "turkey"), joint = TRUE)
+vel_diff_dyn_posthoc <- test(pairs(emmeans(vel_diff_dyn_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
 vel_diff_dyn_posthoc
@@
 ```{r vel-diff-dyn-trial-line-ftop, echo = FALSE, fig.height = 4}
-ggline(dt_vel_diff_dyn[Subcomponent == "FtoP", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_vel_diff_dyn[Subcomponent == "FtoP", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-dyn-trial-line-ptof, echo = FALSE, fig.height = 4}
-ggline(dt_vel_diff_dyn[Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_vel_diff_dyn[Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-dyn-trial-all-line-ftop, echo = FALSE}
-ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-dyn-trial-all-line-ptof, echo = FALSE}
-ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-art-box,  echo = FALSE}
-ggboxplot(dt_vel_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggboxplot(dt_vel_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") +
+  facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r vel-art-all-scatter, echo = FALSE}
-ggerrorplot(vel_art, x = "Condition", y = "Mean", add = c("mean", "mean_se"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggerrorplot(vel_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation")
 ```
@@
 
-ggpaired(vel_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(vel_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation")
 ```
@@
 ```{r post-hoc-3, echo = FALSE}
-vel_art_posthoc <- test(pairs(emmeans(vel_art_aov,~Condition|Subcomponent), adjust = "turkey"), joint = TRUE)
+vel_art_posthoc <- test(pairs(emmeans(vel_art_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
 vel_art_posthoc
@@
 ```{r vel-art-trial-line-legato, echo = FALSE, fig.height = 4}
-ggline(dt_vel_art[Subcomponent == "Legato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_vel_art[Subcomponent == "Legato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(dt_vel_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_vel_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-art-trial-all-line-legato, echo = FALSE}
-ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-art-trial-all-line-staccato, echo = FALSE}
-ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-art-box,  echo = FALSE}
-ggboxplot(dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggboxplot(dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
+  facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r vel-diff_art-all-scatter, echo = FALSE}
-ggerrorplot(vel_diff_art, x = "Condition", y = "Mean", add = c("mean", "mean_se"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Articulation") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggerrorplot(vel_diff_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Articulation")
 ```
@@
 
-ggpaired(vel_diff_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV-Diff: Articulation") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(vel_diff_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV-Diff: Articulation")
 ```
@@
 ```{r post-hoc-4, echo = FALSE}
-vel_diff_art_posthoc <- test(pairs(emmeans(vel_diff_art_aov,~Condition|Subcomponent), adjust = "turkey"), joint = TRUE)
+vel_diff_art_posthoc <- test(pairs(emmeans(vel_diff_art_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
 vel_diff_art_posthoc
@@
 ```{r vel-diff-artt-rial-line-ltos, echo = FALSE, fig.height = 4}
-ggline(dt_vel_diff_art[Subcomponent == "LtoS", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_vel_diff_art[Subcomponent == "LtoS", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-art-trial-line-stol, echo = FALSE, fig.height = 4}
-ggline(dt_vel_diff_art[Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(dt_vel_diff_art[Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-art-trial-all-line-ltos, echo = FALSE}
-ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
   scale_x_continuous(breaks=seq(1,8,1))
@@
 ```{r vel-diff-art-trial-all-line-stol, echo = FALSE}
-ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
   scale_x_continuous(breaks=seq(1,8,1))
@@
 
-ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Note", ylab = "", title = "KV: Dynamics") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Dynamics") +
   scale_x_continuous(breaks = seq(1,67,1))
@@
 
-ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") +
   scale_x_continuous(breaks = seq(1,66,1))
@@
 
-ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Note", ylab = "", title = "KV: Articulation") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Articulation") +
   scale_x_continuous(breaks = seq(1,67,1))
@@
 
-ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") +
   scale_x_continuous(breaks = seq(1,66,1))

### 2021-06-19 experiment-1/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,331 +31,321 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE----------------------------------------------------
+## ----vel-dyn, echo = FALSE----------------------------------
 # For each individual
 vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE-----------------------------------------------
-ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  facet_grid(Subcomponent ~ .) +
[…]

### 2021-06-19 experiment-2/analysis/preprocessor/included_summary.Rmd
add comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
export included_summary.Rmd
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ## Excluded participnats
+- SubNr 2: A deviated tempo
 - SubNr 3: Insufficient motor skills
@@
 data_analysis~.txt are used for data analysis.
-```{r export, echo = TRUE}
+```{r export-csv, echo = TRUE}
 fwrite(dt_ioi[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_ioi.txt", sep = ",")

### 2021-06-19 experiment-2/analysis/preprocessor/included_summary.R

@@ -1,0 +1,66 @@
+## ----setup, include = FALSE---------------------------------
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
+
+# read txt files
+dt_ioi <- fread(file = "./trimmed/data_ioi_1.txt", sep = ",")
+dt_kor <- fread(file = "./trimmed/data_kor.txt", sep = ",")
+dt_kot <- fread(file = "./trimmed/data_kot.txt", sep = ",")
+dt_vel <- fread(file = "./trimmed/data_vel.txt", sep = ",")
+dt_vel_diff <- fread(file = "./trimmed/data_vel_diff.txt", sep = ",")
+
+
+## ----ioi, echo = FALSE--------------------------------------
+dt_ioi
+
+
+## ----ioi-summary, echo = FALSE------------------------------
+dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
+
+
+## ----kor, echo = FALSE--------------------------------------
+dt_kor
+
+
+## ----kor-summary, echo = FALSE------------------------------
+dt_kor[, .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
+
+
+## ----kot, echo = FALSE--------------------------------------
+dt_kot
+
+
+## ----kot-summary, echo = FALSE------------------------------
+dt_kot[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
+
+
+## ----vel, echo = FALSE--------------------------------------
+dt_vel
+
+
+## ----vel-summary, echo = FALSE------------------------------
+dt_vel[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
+
+
+## ----vel-diff, echo = FALSE---------------------------------
+dt_vel_diff
+
+
+## ----vel-diff-summary, echo = FALSE-------------------------
+dt_vel_diff[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
+
+
[…]

### 2021-06-19 experiment-2/analysis/preprocessor/included_summary.Rmd
corrected trimming (don't remove SubNr 2)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 - Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
-- Note: SubNr 3 was excluded due to insufficient motor skills. SubNr 2 was excluded due to a deviated tempo.
+- Note: SubNr 3 was excluded due to insufficient motor skills.
 
@@
 ```{r export-csv, echo = TRUE}
-fwrite(dt_ioi[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_ioi.txt", sep = ",")
-fwrite(dt_kor[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_kor.txt", sep = ",")
-fwrite(dt_kot[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_kot.txt", sep = ",")
-fwrite(dt_vel[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_vel.txt", sep = ",")
-fwrite(dt_vel_diff[SubNr != 2 & SubNr != 3], file = "./trimmed/data_analysis_vel_diff.txt", sep = ",")
+fwrite(dt_ioi[SubNr != 3], file = "./trimmed/data_analysis_ioi.txt", sep = ",")
+fwrite(dt_kor[SubNr != 3], file = "./trimmed/data_analysis_kor.txt", sep = ",")
+fwrite(dt_kot[SubNr != 3], file = "./trimmed/data_analysis_kot.txt", sep = ",")
+fwrite(dt_vel[SubNr != 3], file = "./trimmed/data_analysis_vel.txt", sep = ",")
+fwrite(dt_vel_diff[SubNr != 3], file = "./trimmed/data_analysis_vel_diff.txt", sep = ",")
 ```

### 2021-06-19 experiment-2/analysis/preprocessor/trimming_ioi.R

@@ -155,13 +155,16 @@ plot(p_ioi_box)
 ioi_summary <- dt_ioi_subset_all[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr)]
 # exclude tempo deviated participants
 ioi_summary$Include <- "No"
-ioi_summary[Mean < mean(ioi_summary$Mean)+2*sd(ioi_summary$Mean) & Mean > mean(ioi_summary$Mean)-2*sd(ioi_summary$Mean)]$Include <- "Yes"
+ioi_summary[Mean < mean(ioi_summary$Mean)+3*sd(ioi_summary$Mean) & Mean > mean(ioi_summary$Mean)-3*sd(ioi_summary$Mean)]$Include <- "Yes"
 # exclude participants due to deviated tempi
 dt_ioi_subset <- data.table()
 dt_ioi_subset_all$Exclude <- 0
-for (i in 1:nrow(ioi_summary[Include == "No"])){
-  dt_ioi_subset_all[SubNr == ioi_summary[Include == "No"]$SubNr[i]]$Exclude <- 1
+if (nrow(ioi_summary[Include == "No"]) > 0){
+  for (i in 1:nrow(ioi_summary[Include == "No"])){
+    dt_ioi_subset_all[SubNr == ioi_summary[Include == "No"]$SubNr[i]]$Exclude <- 1
+  }
 }
+
 dt_ioi_subset <- dt_ioi_subset_all[Exclude == 0]
 
 # as a result, SubNr == 2 were excluded due to deviated tempi

### 2021-06-19 experiment-2/analysis/preprocessor/trimming_kot_kor.R

@@ -74,8 +74,8 @@ dt_offset <- dt_kot_offset_all[order(SubNr, BlockNr, TrialNr, NoteNr)]
 # (see error_summary.Rmd
 # and trimming_ioi.R) 
 ####################################
-dt_kot_onset <- dt_onset[SubNr != 2]
-dt_kot_offset <- dt_offset[SubNr != 2]
+dt_kot_onset <- dt_onset
+dt_kot_offset <- dt_offset
 
 ####################################
 # Define Subcomponents

### 2021-06-19 experiment-2/analysis/stats/ioi.Rmd
export as R scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update analysis without subnr 2
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
removed styles from plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-03 analysis/stats/ioi.Rmd
@@
 ```{r ioi-art-box,  echo = FALSE}
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
   geom_hline(yintercept = 1, linetype = "dashed")
@@
 
-ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") +
-  geom_hline(yintercept = 1, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+  geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 #### Normality check
-
 ```{r normality1, echo = FALSE}
@@
 #### Paired *t*-test
-
 ```{r paired-art, echo = FALSE}
@@
 ```{r ioi-art-ch-box, echo = FALSE}
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
   facet_grid(Boundary ~ .) +
@@
 
-ggpaired(ioi_art_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") +
-  geom_hline(yintercept = 1, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(ioi_art_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+  geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ```{r ioi-art-trial-line, echo = FALSE, fig.height = 4}
-ggline(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+ggline(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
   facet_wrap(SubNr ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,8,1)) +
@@
 ```{r ioi-art-trial-all-line, echo = FALSE}
-ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Normalised IOIs", title = "IOI: Articulation") +
   geom_hline(yintercept = 1, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r ioi-dyn-box,  echo = FALSE}
-ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
   geom_hline(yintercept = 1, linetype = "dashed")
@@
 
-ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 1, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+  geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ```{r ioi-dyn-ch-box, echo = FALSE}
-ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
   facet_grid(Boundary ~ .) +
@@
 
-ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 1, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+  geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ```{r ioi-dyn-trial-line, echo = FALSE, fig.height = 4}
-ggline(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", palette = "aaas", xlab = "TrialNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+ggline(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
   facet_wrap(SubNr ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,8,1)) +
@@
 ```{r ioi-dyn-trial-all-line, echo = FALSE}
-ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
   geom_hline(yintercept = 1, linetype = "dashed") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 
-ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "Normalised IOIs", title = "IOI: Articulation") +
   geom_hline(yintercept = 1, linetype = "dashed") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
   scale_x_continuous(breaks = seq(1,71,1))
@@
 
-ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
   geom_hline(yintercept = 1, linetype = "dashed") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
     scale_x_continuous(breaks = seq(1,71,1))
 ```
+
+```{r export, include = FALSE}
+knitr::purl("ioi.Rmd")
+```

### 2021-06-19 experiment-2/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -8,11 +8,11 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -24,50 +24,48 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE----------------------------------------------------
+## ----ioi-art, echo = FALSE----------------------------------
 # For each individual
 ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE-----------------------------------------------
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + 
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+## ----ioi-art-box,  echo = FALSE-----------------------------
[…]

### 2021-06-19 experiment-2/analysis/stats/articulation.Rmd
removed styles from plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 #### Box plot
-
 ```{r kor-art-box,  echo = FALSE}
-ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
+  facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r kor-art-trial-all-line-staccato, echo = FALSE}
-ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
-  theme_pubr(base_size = 20, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kot-art-box,  echo = FALSE}
-ggboxplot(dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  facet_grid(Subcomponent ~ .) +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed")
+ggboxplot(dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") +
+  facet_grid(Subcomponent ~ .)
 ```
@@
 
-ggline(kor_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOR", title = "KOT: Articulation") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kor_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOR", title = "KOT: Articulation") +
   scale_x_continuous(breaks = seq(1,71,1))
@@
 
-ggline(kor_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOR", title = "KOT: Dynamics")+
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(kor_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOR", title = "KOT: Dynamics")+
   scale_x_continuous(breaks = seq(1,71,1)) 

### 2021-06-19 experiment-2/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,338 +31,308 @@ dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 
 
-## ----kor-art, echo = FALSE----------------------------------------------------
+## ----kor-art, echo = FALSE----------------------------------
 # For each individual
 kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kor_art
 
 
-## ----kor-art-box,  echo = FALSE-----------------------------------------------
-ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
-  facet_grid(Subcomponent ~ .) +
[…]

### 2021-06-19 experiment-2/analysis/stats/dynamics.Rmd
removed styles from plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Note", ylab = "", title = "KV: Dynamics") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Dynamics") +
   scale_x_continuous(breaks = seq(1,72,1))
@@
 
-ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") +
   scale_x_continuous(breaks = seq(1,71,1))
@@
 
-ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Note", ylab = "", title = "KV: Articulation") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Articulation") +
   scale_x_continuous(breaks = seq(1,72,1))
@@
 
-ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") +
-  theme_pubr(base_size = 12, legend = "bottom", base_family = "Helvetica Neue LT Std 57 Condensed") +
+ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") +
   scale_x_continuous(breaks = seq(1,71,1))

### 2021-06-19 experiment-2/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,360 +31,328 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE----------------------------------------------------
+## ----vel-dyn, echo = FALSE----------------------------------
 # For each individual
 vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE-----------------------------------------------
-ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", palette = "aaas", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  facet_grid(Subcomponent ~ .) +
[…]

### 2021-06-19 paper/manuscript.Rmd
updated analyses for exp2
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Pitch errors included either extra, missing or substituted tones and were manually removed by using the *editData* R package.
-For onsets, `r round(sum(error_onset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_onset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes: `r round(error_onset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes: `r round(error_onset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
-For offsets, `r round(sum(error_offset_summary$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes: `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes: `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
+For onsets, `r round(sum(error_onset_summary$N)/((31*4*8)-5)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_onset_summary[Reason == "Extra Notes"]$N/((31*4*8)-5)*100, 2)`%, missing notes: `r round(error_onset_summary[Reason == "Missing Notes"]$N/((31*4*8)-5)*100, 2)`%, substituted notes: `r round(error_onset_summary[Reason == "Substituted"]$N/((31*4*8)-5)*100, 2)`%).
+For offsets, `r round(sum(error_offset_summary$N)/((31*4*8)-5)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*8)-5)*100, 2)`%, missing notes: `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*8)-5)*100, 2)`%, substituted notes: `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*8)-5)*100, 2)`%).
+As a result, less than 1 % of the total responses (i.e., piano keystrokes) were corrected.
 In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
@@
 
-In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* < 0.001).
+In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* < `r if(ioi_art_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_norm$p.value))}`).
 The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
@@
 
+In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_norm$p.value))}`).
 The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
@@
 img_vel_dyn_1 <- image_read(here("paper/image", "vel_dyn_1.png"))
+
+# annotate * for forte
+img_vel_dyn_1 <- image_annotate(img_vel_dyn_1, "*", size = 280, color = "black", location = "+1430+800")
 ```
@@
 For data analysis, we excluded one participants due to insufficient motor skills, and one because of large tempo deviations (outside 2 standard deviations from the average tempo across participants).
-Nineteen participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in the data analysis.
+Twenty participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in the data analysis.
 Most participants were right-handed (left: `r nrow(dt_2_included[Handedness == "Left"])`) with a mean age of `r round(mean(dt_2_included$Age), 2)` (*SD* = `r round(sd(dt_2_included$Age), 2)`).
@@
 
-### 3.1.4. Design and data analysis
+### 3.1.4. Data analysis
 
@@
 Using the same approach as in Experiment 1, pitch errors were removed manually.
-For onsets, `r round(sum(error_onset_summary_2$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_onset_summary_2[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_onset_summary_2[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_onset_summary_2[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
-For offsets, `r round(sum(error_offset_summary_2$N)/((31*4*18)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_offset_summary_2[Reason == "Extra Notes"]$N/((31*4*18)-3)*100, 2)`%, missing notes; `r round(error_offset_summary_2[Reason == "Missing Notes"]$N/((31*4*18)-3)*100, 2)`%, substituted notes; `r round(error_offset_summary_2[Reason == "Substituted"]$N/((31*4*18)-3)*100, 2)`%).
+For onsets, `r round(sum(error_onset_summary_2$N)/((20*4*8)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_onset_summary_2[Reason == "Extra Notes"]$N/((20*4*8)-3)*100, 2)`%, missing notes; `r round(error_onset_summary_2[Reason == "Missing Notes"]$N/((20*4*8)-3)*100, 2)`%, substituted notes; `r round(error_onset_summary_2[Reason == "Substituted"]$N/((20*4*8)-3)*100, 2)`%).
+For offsets, `r round(sum(error_offset_summary_2$N)/((20*4*8)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_offset_summary_2[Reason == "Extra Notes"]$N/((20*4*8)-3)*100, 2)`%, missing notes; `r round(error_offset_summary_2[Reason == "Missing Notes"]$N/((20*4*8)-3)*100, 2)`%, substituted notes; `r round(error_offset_summary_2[Reason == "Substituted"]$N/((20*4*8)-3)*100, 2)`%).
 For each dependent variable, removing outliers (i.e., responses outside 3 standard deviations from the mean) resulted in less than 5% of overall responses being removed.

### 2021-06-19 paper/manuscript.tex
updated analyses for exp2
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Pitch errors included either extra, missing or substituted tones and were manually removed by using the \emph{editData} R package.
-For onsets, 5.16\% of the trials contained at least one pitch error (extra notes: 2.78\%, missing notes: 2.24\%, substituted notes: 0.13\%).
-For offsets, 6.24\% of the trials contained at least one pitch error (extra notes: 2.78\%, missing notes: 2.29\%, substituted notes: 1.17\%).
+For onsets, 11.65\% of the trials contained at least one pitch error (extra notes: 6.28\%, missing notes: 5.07\%, substituted notes: 0.30\%).
+For offsets, 14.08\% of the trials contained at least one pitch error (extra notes: 6.28\%, missing notes: 5.17\%, substituted notes: 2.63\%).
+As a result, less than 1 \% of the total responses (i.e., piano keystrokes) were corrected.
 In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
@@
 
-In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
+In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} \textless{} 0.001).
 The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021; \emph{Figure \ref{fig:ioi-1}}).
@@
 
+In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
 The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11; \emph{Figure \ref{fig:ioi-1}}).
@@
 For data analysis, we excluded one participants due to insufficient motor skills, and one because of large tempo deviations (outside 2 standard deviations from the average tempo across participants).
-Nineteen participants (9 female) were included in the data analysis.
+Twenty participants (9 female) were included in the data analysis.
 Most participants were right-handed (left: 2) with a mean age of 25.90 (\emph{SD} = 4.68).
@@
 
-\hypertarget{design-and-data-analysis}{%
-\subsubsection{3.1.4. Design and data analysis}\label{design-and-data-analysis}}
+\hypertarget{data-analysis-1}{%
+\subsubsection{3.1.4. Data analysis}\label{data-analysis-1}}
 
@@
 Using the same approach as in Experiment 1, pitch errors were removed manually.
-For onsets, 3.32\% of the trials contained at least one pitch error (extra notes; 1.66\%, missing notes; 1.57\%, substituted notes; 0.09\%).
-For offsets, 5.11\% of the trials contained at least one pitch error (extra notes; 1.66\%, missing notes; 1.57\%, substituted notes; 1.88\%).
+For onsets, 11.62\% of the trials contained at least one pitch error (extra notes; 5.81\%, missing notes; 5.49\%, substituted notes; 0.31\%).
+For offsets, 17.90\% of the trials contained at least one pitch error (extra notes; 5.81\%, missing notes; 5.49\%, substituted notes; 6.59\%).
 For each dependent variable, removing outliers (i.e., responses outside 3 standard deviations from the mean) resulted in less than 5\% of overall responses being removed.
@@
 
-A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.97, \emph{SD} = 0.042{]} than in the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.034{]} while playing the piece with notated articulation (\emph{t}(18) = 2.64
-, \emph{p} = 0.017; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.026).
+A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.97, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.95, \emph{SD} = 0.040{]} while playing the piece with notated articulation (\emph{t}(19) = 2.47
+, \emph{p} = 0.023; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.033).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 2324, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 6.90, \emph{p} = 0.017, \(\eta_G^2\) = 0.01; \emph{Figure \ref{fig:kor-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -0.77, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.001) while there was no significant difference in legato between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.85).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01; \emph{Figure \ref{fig:kor-2}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.002) while there was no significant difference in legato between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.56).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.011, \(\eta_G^2\) = 0.04; \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
-Neither the main effect of Condition (\emph{F}(1,18) = 0.46, \emph{p} = 0.51, \(\eta_G^2\) = 0.001) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.80, \(\eta_G^2\) = 0.000) was significant.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 7.91, \emph{p} = 0.011, \(\eta_G^2\) = 0.03; \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
+Neither the main effect of Condition (\emph{F}(1,19) = 1.33, \emph{p} = 0.26, \(\eta_G^2\) = 0.003) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 7.91, \emph{p} = 0.71, \(\eta_G^2\) = 0.000) was significant.
 
@@
 
-A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.041{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.037{]} while playing the piece with the notated dynamics (\emph{t}(18) = 0.37, \emph{p} = 0.72; \emph{Figure \ref{fig:ioi-2}}).
-A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (\emph{p} = 0.92).
+A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.043{]} while playing the piece with the notated dynamics (\emph{t}(19) = 0.21, \emph{p} = 0.84; \emph{Figure \ref{fig:ioi-2}}).
+A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (\emph{p} = 0.96).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.022; \emph{Figure \ref{fig:vel-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (\emph{p} = 0.036) and softer piano (\emph{p} = 0.028)in the teaching condition {[}Forte; \emph{M} = 82.55, \emph{SD} = 8.57, Piano; \emph{M} = 60.41, \emph{SD} = 3.83{]} than in the performing condition {[}Forte; \emph{M} = 80.71, \emph{SD} = 7.53, Piano; \emph{M} = 62.34, \emph{SD} = 4.93{]} than in the performing condition.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 132, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 6.12, \emph{p} = 0.023, \(\eta_G^2\) = 0.018; \emph{Figure \ref{fig:vel-2}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (\emph{p} = 0.040) and softer piano (\emph{p} = 0.024)in the teaching condition {[}Forte; \emph{M} = 83.20, \emph{SD} = 8.84, Piano; \emph{M} = 60.53, \emph{SD} = 3.76{]} than in the performing condition {[}Forte; \emph{M} = 81.49, \emph{SD} = 8.12, Piano; \emph{M} = 62.42, \emph{SD} = 4.82{]} than in the performing condition.
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 111, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.064; \emph{Figure \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition {[}FtoP; \emph{M} = -17.35, \emph{SD} = 8.46, PtoF; \emph{M} = 20.60, \emph{SD} = 9.10{]} than in the performing condition {[}FtoP; \emph{M} = -13.71, \emph{SD} = 7.73, PtoF; \emph{M} = 15.87, \emph{SD} = 7.47{]} (FtoP; \emph{p} = 0.017, PtoF; \emph{p} = 0.003).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059; \emph{Figure \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition {[}FtoP; \emph{M} = -17.23, \emph{SD} = 8.26, PtoF; \emph{M} = 20.71, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP; \emph{M} = -14.04, \emph{SD} = 7.67, PtoF; \emph{M} = 16.05, \emph{SD} = 7.32{]} (FtoP; \emph{p} = 0.029, PtoF; \emph{p} = 0.002).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 559, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.93), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,18) = 0.016, \emph{p} = 0.90, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 0.58, \emph{p} = 0.46, \(\eta_G^2\) = 0.001; \emph{Figure \ref{fig:kor-2}}) was significant.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 630, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (\emph{F}(1,19) = 0.081, \emph{p} = 0.78, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 0.35, \emph{p} = 0.56, \(\eta_G^2\) = 0.001; \emph{Figure \ref{fig:kor-2}}) was significant.
 

### 2021-06-19 experiment-1/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -46,293 +46,293 @@ dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 
 
-## ----kot-art, echo = FALSE----------------------------------
+## ----kot-art, echo = FALSE----------------------------------------------------
 # For each individual
 kot_art <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kot_art
 
 
-## ----kot-art-box,  echo = FALSE-----------------------------
+## ----kot-art-box,  echo = FALSE-----------------------------------------------
 ggboxplot(dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") +
   facet_grid(Subcomponent ~ .)
 
 
[…]

### 2021-06-19 experiment-1/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------
+## ----setup, include = FALSE--------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------
+## ----file, include = FALSE---------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE-------------------------------
+## ----extract, include = FALSE------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,49 +31,49 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE----------------------------------
+## ----vel-dyn, echo = FALSE---------------------------------
 # For each individual
 vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE-----------------------------
+## ----vel-dyn-box,  echo = FALSE----------------------------
 ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
[…]

### 2021-06-19 experiment-2/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE-------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,303 +31,303 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE----------------------------------
+## ----vel-dyn, echo = FALSE----------------------------------------------------
 # For each individual
 vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE-----------------------------
+## ----vel-dyn-box,  echo = FALSE-----------------------------------------------
[…]

### 2021-06-19 experiment-2/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -8,11 +8,11 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 
 
-## ----file, include = FALSE----------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE-------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -24,48 +24,48 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE----------------------------------
+## ----ioi-art, echo = FALSE----------------------------------------------------
 # For each individual
 ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE-----------------------------
+## ----ioi-art-box,  echo = FALSE-----------------------------------------------
 ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + 
   geom_hline(yintercept = 1, linetype = "dashed")
 
 
-## ----ioi-art-all, echo = FALSE------------------------------
+## ----ioi-art-all, echo = FALSE------------------------------------------------
 # Group mean
[…]

### 2021-06-19 paper/manuscript.Rmd
tweak plots for significance
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   ylim(-170, 105) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -50, method = "t.test", paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.45, label.y = -50, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   ylim(-40, 60) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 45, method = "t.test", paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.4, label.y = 45, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   ylim(-1, 0.4) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = -0.5, method = "t.test", paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.4, label.y = -0.5, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 
-### 3.2.6. KV Transition Points
+### 3.2.6. Dynamics: KV Transition Points
 
@@
   ylim(-40, 60) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 45, method = "t.test", paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.45, label.y = 45, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

### 2021-06-20 experiment-1/analysis/stats/dynamics.Rmd
tweak notes
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-- Order analysis (teaching/performing first?)
-  + [All](#vel_dyn_order_group_all)
-
 ### KV-Diff
@@
   + [Trial by trial](#vel_diff_dyn_group_trial)
-
-- Order analysis (teaching/performing first?)
-  + [All](#vel_diff_dyn_order_group_all)
 
@@
 #### Scatter plot
-- Error bars show SEM of the mean
+- Error bars show SEM of the mean.
 
@@
 #### Line graph
-- Error bars show SEM of the mean
+- Error bars show SEM of the mean.
 

### 2021-06-20 experiment-2/analysis/stats/articulation.Rmd
tweak notes
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 1. Main predictions
-Participants will exaggerate legato and staccato more while teaching.
+- Participants will exaggerate legato and staccato more while teaching.
 

### 2021-06-20 experiment-2/analysis/stats/dynamics.Rmd
tweak notes
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 2. Excluded participants
-- SubNr 2: A deviated tempo
 - SubNr 3: Insufficient motor skills

### 2021-06-20 experiment-2/analysis/stats/ioi.Rmd
tweak notes
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 #### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)
+- A dashed line represents the tempo given by a metronome.
 
@@
 #### Line graph
-- Error bars show SEM of the mean
+- Error bars show SEM of the mean.
 - A dashed line represents the tempo given by a metronome.

### 2021-06-20 paper/manuscript.Rmd
update abstract
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Expressions in the piece concerned either articulation (i.e., legato and staccato) or dynamics (i.e., forte and piano).
-The pianists played either with the goal to perform the piece to an audience or with the goal to teach the respective techniques to novices.When intending to teach articulation, skilled pianists produced more exaggerated staccato.
+The pianists played either with the goal to perform the piece to an audience or with the goal to teach the respective techniques to novices.
+When intending to teach articulation, skilled pianists produced more exaggerated staccato.
 When intending to teach dynamics, they created a larger contrast between forte and piano at structurally important points.

### 2021-06-20 paper/manuscript.tex
tweak plots for significance
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update abstract
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-\hypertarget{kv-transition-points}{%
-\subsubsection{3.2.6. KV Transition Points}\label{kv-transition-points}}
+\hypertarget{dynamics-kv-transition-points-1}{%
+\subsubsection{3.2.6. Dynamics: KV Transition Points}\label{dynamics-kv-transition-points-1}}
 

### 2021-06-21 experiment-1/analysis/stats/ioi.Rmd
add effect size
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("afex")) {install.packages("afex"); require("afex")}
+if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 ```
@@
 ioi_art_wilcoxon
+```
+#### Effect size
+```{r effect-art, echo = FALSE}
+wilcoxeff_art <- ioi_art %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
+wilcoxeff_art
 ```
@@
 
+#### Effect size
+```{r effect-dyn, echo = FALSE}
+wilcoxeff_dyn <- ioi_dyn %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
+wilcoxeff_dyn
+```
+
 ## 3. Phrase boundaries - individual{#dyn_ind_phrase}

### 2021-06-21 experiment-2/analysis/stats/ioi.Rmd
add effect size
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```
+#### Effect size
+```{r effect-cohen-art, echo = FALSE}
+coheneff_art <- ioi_art %>% cohens_d(Mean ~ Condition, paired = TRUE)
+coheneff_art
+```
 
@@
 ioi_art_wilcoxon
+```
+
+#### Effect size
+```{r effect-wilcox-art, echo = FALSE}
+wilcoxeff_art <- ioi_art %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
+wilcoxeff_art
 ```
@@
 ```
+#### Effect size
+```{r effect-cohen-dyn, echo = FALSE}
+coheneff_dyn <- ioi_dyn %>% cohens_d(Mean ~ Condition, paired = TRUE)
+coheneff_dyn
+```
 
@@
 ioi_dyn_wilcoxon
+```
+
+#### Effect size
+```{r effect-wilcox-dyn, echo = FALSE}
+wilcoxeff_dyn <- ioi_dyn %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
+wilcoxeff_dyn
 ```

### 2021-06-21 experiment-2/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -6,13 +6,14 @@ if (!require("data.table")) {install.packages("data.table"); require("data.table
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 # statistics
 if (!require("afex")) {install.packages("afex"); require("afex")}
+if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -24,48 +25,58 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE----------------------------------------------------
+## ----ioi-art, echo = FALSE----------------------------------
 # For each individual
 ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE-----------------------------------------------
+## ----ioi-art-box,  echo = FALSE-----------------------------
 ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + 
   geom_hline(yintercept = 1, linetype = "dashed")
 
 
[…]

### 2021-06-21 experiment-1/analysis/stats/ioi.Rmd
tweak notes
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 - [afex](https://github.com/singmann/afex) for ANOVAs
+- [rstatix](https://github.com/kassambara/rstatix) for effect sizes
 
@@
 #### Line graph
-- Error bars show SEM of the mean
+- Error bars show SEM of the mean.
 - A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).

### 2021-06-21 experiment-1/analysis/stats/ioi.R

@@ -62,7 +62,8 @@ ioi_art_wilcoxon
 
 
 ## ----effect-art, echo = FALSE-------------------------------
-ioi_art %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
+wilcoxeff_art <- ioi_art %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
+wilcoxeff_art
 
 
 ## ----ioi-art-ch, echo = FALSE-------------------------------
@@ -166,7 +167,8 @@ ioi_dyn_wilcoxon
 
 
 ## ----effect-dyn, echo = FALSE-------------------------------
-ioi_dyn %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
+wilcoxeff_dyn <- ioi_dyn %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
+wilcoxeff_dyn
 
 
 ## ----ioi-dyn-ch, echo = FALSE-------------------------------

### 2021-06-21 paper/manuscript.Rmd
add effect size
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The *t.test* function in the *stats* R package was used for a *t*-test.
+For calculating an effect size, we used the *rstatix* R package.
 The *aov_car* function in the *afex* R package was used for a repeated-measures ANOVA.
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* < `r if(ioi_art_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)``; *Figure \ref{fig:ioi-1}*).
 
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`; *Figure \ref{fig:ioi-1}*).
 
@@
 We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school at the time of recruitment.
-For data analysis, we excluded one participants due to insufficient motor skills, and one because of large tempo deviations (outside 2 standard deviations from the average tempo across participants).
+For data analysis, we excluded one participants due to insufficient motor skills.
 Twenty participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in the data analysis.
@@
 A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`
-, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
-A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`).
+, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(coheneff_art$effsize, 2)`; *Figure \ref{fig:ioi-2}*).
+A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`).
 
@@
 
-A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
+A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`, Cohen's *d* = `r round(coheneff_dyn$effsize, 2)`; *Figure \ref{fig:ioi-2}*).
 A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`).

### 2021-06-21 paper/manuscript.tex
add effect size
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The \emph{t.test} function in the \emph{stats} R package was used for a \emph{t}-test.
+For calculating an effect size, we used the \emph{rstatix} R package.
 The \emph{aov\_car} function in the \emph{afex} R package was used for a repeated-measures ANOVA.
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} \textless{} 0.001).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021; \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41`; \emph{Figure \ref{fig:ioi-1}}).
 
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11; \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29; \emph{Figure \ref{fig:ioi-1}}).
 
@@
 We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school at the time of recruitment.
-For data analysis, we excluded one participants due to insufficient motor skills, and one because of large tempo deviations (outside 2 standard deviations from the average tempo across participants).
+For data analysis, we excluded one participants due to insufficient motor skills.
 Twenty participants (9 female) were included in the data analysis.
@@
 A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.97, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.95, \emph{SD} = 0.040{]} while playing the piece with notated articulation (\emph{t}(19) = 2.47
-, \emph{p} = 0.023; \emph{Figure \ref{fig:ioi-2}}). A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.033).
+, \emph{p} = 0.023, Cohen's \emph{d} = -0.55; \emph{Figure \ref{fig:ioi-2}}).
+A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.033, \emph{r} = 0.48).
 
@@
 
-A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.043{]} while playing the piece with the notated dynamics (\emph{t}(19) = 0.21, \emph{p} = 0.84; \emph{Figure \ref{fig:ioi-2}}).
+A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.043{]} while playing the piece with the notated dynamics (\emph{t}(19) = 0.21, \emph{p} = 0.84, Cohen's \emph{d} = -0.05; \emph{Figure \ref{fig:ioi-2}}).
 A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (\emph{p} = 0.96).

### 2021-06-21 experiment-1/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,11 +9,11 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 
 
-## ----file, include = FALSE----------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE-------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -25,48 +25,48 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE----------------------------------
+## ----ioi-art, echo = FALSE----------------------------------------------------
 # For each individual
 ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE-----------------------------
+## ----ioi-art-box,  echo = FALSE-----------------------------------------------
 ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + 
   geom_hline(yintercept = 188, linetype = "dashed")
 
 
-## ----ioi-art-all, echo = FALSE------------------------------
+## ----ioi-art-all, echo = FALSE------------------------------------------------
 # Group mean
[…]

### 2021-06-21 experiment-2/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,293 +31,293 @@ dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 
 
-## ----kor-art, echo = FALSE----------------------------------
+## ----kor-art, echo = FALSE----------------------------------------------------
 # For each individual
 kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kor_art
 
 
-## ----kor-art-box,  echo = FALSE-----------------------------
+## ----kor-art-box,  echo = FALSE-----------------------------------------------
 ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
   facet_grid(Subcomponent ~ .)
 
 
[…]

### 2021-06-21 experiment-2/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,11 +9,11 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 
 
-## ----file, include = FALSE----------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE-------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -25,58 +25,58 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE----------------------------------
+## ----ioi-art, echo = FALSE----------------------------------------------------
 # For each individual
 ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE-----------------------------
+## ----ioi-art-box,  echo = FALSE-----------------------------------------------
 ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + 
   geom_hline(yintercept = 1, linetype = "dashed")
 
 
-## ----ioi-art-all, echo = FALSE------------------------------
+## ----ioi-art-all, echo = FALSE------------------------------------------------
 # Group mean
[…]

### 2021-06-21 paper/manuscript.Rmd
tweak some wordings
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-06-20 paper/manuscript.Rmd
@@
 ```{r set-up, include = FALSE}
-if (!require("papaja")) {install.packages("papaja"); require("papaja")}
-if (!require("here")) {install.packages("here"); require("here")}
-if (!require("magick")) {install.packages("magick"); require("magick")}
+if (!require("papaja")) {install.packages("papaja"): require("papaja")}
+if (!require("here")) {install.packages("here"): require("here")}
+if (!require("magick")) {install.packages("magick"): require("magick")}
 
@@
 
-Deliberate teaching has supported human skill transmission over generations and provides a key route for learning [@thornton_2008; @tomasello_2016].
+Deliberate teaching has supported human skill transmission over generations and provides a key route for learning [@thornton_2008: @tomasello_2016].
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
-In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners [@brand_2002; @saint-georges_2013].
+In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners [@brand_2002: @saint-georges_2013].
 These modulations include slowing down and exaggerating sounds and actions.
@@
 Furthermore, one could speculate that they might produce modulations specifically at structurally important points that best highlight the technique to be taught.
-Importantly, they should avoid modulating irrelevant properties of expression, e.g., the smoothness of the sound while teaching dynamics.
+Importantly, they should avoid modulating irrelevant properties of expression, e.g., the smoothness of sound while teaching dynamics.
 
@@
 
-A weighted Yamaha MIDI digital piano was used to record participants' performance via Max/MSP on a MacBook Pro with MacOS X Mojave 10.14.3.
+A weighted Yamaha MIDI digital piano was used to record participants' performance via Max/MSP ([https://cycling74.com/products/max](https://cycling74.com/products/max)) on a MacBook Pro with Mac OS X Mojave 10.14.3.
 The laptop and piano were connected to a high-fidelity soundcard (Focusrite Scarlett 6i6) to deliver a metronome and piano sound.
@@
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations; see *Figure \ref{fig:stimuli}, A*) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations: see *Figure \ref{fig:stimuli}, A*) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
@@
 The order of the expressive techniques within each condition was also blocked and counterbalanced across participants.
-A leading metronome (80 beats per minute, 8 beats) indicated the target tempo before each trial.
+A leading metronome (80 quarter-beats per minute, 8 beats) indicated the target tempo before each trial.
 
@@
 For offsets, `r round(sum(error_offset_summary$N)/((31*4*8)-5)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*8)-5)*100, 2)`%, missing notes: `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*8)-5)*100, 2)`%, substituted notes: `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*8)-5)*100, 2)`%).
-As a result, less than 1 % of the total responses (i.e., piano keystrokes) were corrected.
+As a result, less than 1 % of the total responses were corrected.
 In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
@@
 We performed separate analyses for the two expressive techniques (i.e., articulation, dynamics).
-A paired-sample *t*-test or a Wilcoxon signed-rank tests (a non-parametric alternative to a paired *t*-test) was performed to compare the mean IOIs in the teaching and performing condition.
+A paired-sample *t*-test or a Wilcoxon Signed-rank test (a non-parametric alternative to a paired *t*-test) was performed to compare the mean IOIs in the teaching and performing condition.
 For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.
-performing) and Subcomponent (LS; legato vs.
-staccato or FP; forte vs.
+performing) and Subcomponent (LS: legato vs.
+staccato or FP: forte vs.
 piano, respectively).
-The *t.test* function in the *stats* R package was used for a *t*-test.
+The *t.test* or *wilcox.test* function in the *stats* R package was used for a *t*-test or a Wilcoxon Signed-rank test.
+For calculating an effect size, we used the *rstatix* R package.
 The *aov_car* function in the *afex* R package was used for a repeated-measures ANOVA.
@@
 We first report results for performance of the piece with the notated expression of articulation (*Figure \ref{fig:stimuli}, B*), followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli}, C*).
-To recall our predictions, if participants play slowly when they are teaching in general, interonset intervals (IOIs) should be larger when teaching.
+To recall our predictions, if participants play more slowly when they are trying to teach, interonset intervals (IOIs) should be larger when teaching.
 If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should reveal exaggerated legato/staccato when teaching articulation, and key velocity (KV) should reveal amplifications of forte and piano when teaching dynamics.
@@
 
-In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* < `r if(ioi_art_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
+In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_norm$p.value))}`).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`: *Figure \ref{fig:ioi-1}*).
 
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`; *Figure \ref{fig:kot-1}*).
+staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`: *Figure \ref{fig:kot-1}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`; *Figure \ref{fig:vel-1}*).
-However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching [Legato; *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`, Staccato; *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] and performing condition [Legato; *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`, Staccato; *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] for each subcomponent (Legato; *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[1])`, Staccato; *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
+staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`: *Figure \ref{fig:vel-1}*).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching [Legato: *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`, Staccato: *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] and performing condition [Legato: *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`, Staccato: *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] for each subcomponent (Legato: *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[1])`, Staccato: *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
 
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`; *Figure \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`: *Figure \ref{fig:ioi-1}*).
 
@@
 performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-1}*).
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`: *Figure \ref{fig:vel-1}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) while there was no significant difference in piano between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`).
@@
 
-Additionally, we conducted an exploratory analysis by focusing on specific points at which each subcomponent changed from one to the other (i.e., forte to piano; FtoP, piano to forte; PtoF).
+Additionally, we conducted an exploratory analysis by focusing on specific points at which each subcomponent changed from one to the other (i.e., forte to piano: FtoP, piano to forte: PtoF).
 These points could be structurally important in order to make a contrast between forte and piano.
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-diff}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`: *Figure \ref{fig:vel-diff}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
 
@@
 performing) and Subcomponent-FP (forte vs.
-piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`; *Figure \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[3,5])`) was significant.
+piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`: *Figure \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`) was significant.
 
@@
 For IOIs, we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
-This finding is in line with earlier studies that found slower performance of actions in a teaching context [e.g., @brand_2002; @mcellin_2017].
+This finding is in line with earlier studies that found slower performance of actions in a teaching context [e.g., @brand_2002: @mcellin_2017].
 However, we did not find a significant difference in tempo when participants were teaching dynamics.
@@
 
-We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school at the time of recruitment.
-For data analysis, we excluded one participants due to insufficient motor skills, and one because of large tempo deviations (outside 2 standard deviations from the average tempo across participants).
+We recruited 21 piano experts who already had a degree (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school at the time of recruitment.
+For data analysis, we excluded one participant due to insufficient motor skills.
 Twenty participants (`r nrow(dt_2_included[Gender == "Female"])` female) were included in the data analysis.
@@
 Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[Supplementary Materials](#supplemental)*).
-Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 beats per minute).
+Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter-beats per minute).
 The chosen tempo was again cued by a leading metronome.
 In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music.
-As Experiment 1, those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis (as a result, one participant was excluded).
+Those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis (as a result, one participant was excluded).
 The rest of the procedure was identical to Experiment 1.
@@
 Using the same approach as in Experiment 1, pitch errors were removed manually.
-For onsets, `r round(sum(error_onset_summary_2$N)/((20*4*8)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_onset_summary_2[Reason == "Extra Notes"]$N/((20*4*8)-3)*100, 2)`%, missing notes; `r round(error_onset_summary_2[Reason == "Missing Notes"]$N/((20*4*8)-3)*100, 2)`%, substituted notes; `r round(error_onset_summary_2[Reason == "Substituted"]$N/((20*4*8)-3)*100, 2)`%).
-For offsets, `r round(sum(error_offset_summary_2$N)/((20*4*8)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes; `r round(error_offset_summary_2[Reason == "Extra Notes"]$N/((20*4*8)-3)*100, 2)`%, missing notes; `r round(error_offset_summary_2[Reason == "Missing Notes"]$N/((20*4*8)-3)*100, 2)`%, substituted notes; `r round(error_offset_summary_2[Reason == "Substituted"]$N/((20*4*8)-3)*100, 2)`%).
+For onsets, `r round(sum(error_onset_summary_2$N)/((20*4*8)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_onset_summary_2[Reason == "Extra Notes"]$N/((20*4*8)-3)*100, 2)`%, missing notes: `r round(error_onset_summary_2[Reason == "Missing Notes"]$N/((20*4*8)-3)*100, 2)`%, substituted notes: `r round(error_onset_summary_2[Reason == "Substituted"]$N/((20*4*8)-3)*100, 2)`%).
+For offsets, `r round(sum(error_offset_summary_2$N)/((20*4*8)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_offset_summary_2[Reason == "Extra Notes"]$N/((20*4*8)-3)*100, 2)`%, missing notes: `r round(error_offset_summary_2[Reason == "Missing Notes"]$N/((20*4*8)-3)*100, 2)`%, substituted notes: `r round(error_offset_summary_2[Reason == "Substituted"]$N/((20*4*8)-3)*100, 2)`%).
+As a result, less than 1 % of the total responses were corrected.
 For each dependent variable, removing outliers (i.e., responses outside 3 standard deviations from the mean) resulted in less than 5% of overall responses being removed.
@@
 
-A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`
-, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
-A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`).
+A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`
+, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art$effsize), 3)`: *Figure \ref{fig:ioi-2}*).
+A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`).
 
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`; *Figure \ref{fig:kor-2}*).
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`: *Figure \ref{fig:kor-2}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_art_aov$anova_table[2,5])`; *Figure \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_art_aov$anova_table[2,5])`: *Figure \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
 Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`) was significant.
@@
   ylim(c(50, 100)) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, method = "t.test", paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, method = "t.test", paired = TRUE, symnum.args = list(cutpoints = c(0.001, 0.01, 0.05, 1), symbols = c("**", "**", "*", "ns"))) + # adjust significance level according to the actual stats
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 
-A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`; *Figure \ref{fig:ioi-2}*).
+A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn$effsize), 3)`: *Figure \ref{fig:ioi-2}*).
 A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`).
@@
 performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) and softer piano (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`)in the teaching condition [Forte; *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano; *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte; *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano; *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition.
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`: *Figure \ref{fig:vel-2}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) and softer piano (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) in the teaching condition [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition.
 
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:vel-diff}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF; *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`] (FtoP; *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`, PtoF; *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`).
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`: *Figure \ref{fig:vel-diff}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
 
@@
 piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`; *Figure \ref{fig:kor-2}*) was significant.
+Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`: *Figure \ref{fig:kor-2}*) was significant.
 
@@
 
-```{r plot-vel-diff, out.width = "100%", fig.cap = "\\label{fig:vel-diff}Experiment 1, 2: KV Difference at transition points when playing the piece with dynamics (left; Experiment 1, right; Experiment 2).
+```{r plot-vel-diff, out.width = "100%", fig.cap = "\\label{fig:vel-diff}Experiment 1, 2: KV Difference at transition points when playing the piece with dynamics (left: Experiment 1, right: Experiment 2).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
@@
 
-Musicians and music teachers tend to consider expressive skills as a performer’s most important skills [@lindstrom_2003].
+Musicians and music teachers tend to consider expressive skills as a performer’s most important skill [@lindstrom_2003].
 In the present study, experts focused on the teaching of expressions that were notated.
@@
 ### Experiment 1
-In the teaching condition, participants were shown with the following instruction on a computer monitor (Italic sentences were highlighted with a yellow colour on a black background); Now, play what you practised as if you were teaching it to students.
+In the teaching condition, participants were shown with the following instruction on a computer monitor (Italic sentences were highlighted with a yellow colour on a black background): Now, play what you practised as if you were teaching it to students.
 Students already know how to produce the sequence of the  tones and now are trying to *learn how to perform the piece expressively* by listening to your performance.
@@
 
-In the performing condition, participants were displayed with the following instruction on a computer monitor monitor; Now, play what you practised as if you were performing it to an audience.
+In the performing condition, participants were displayed with the following instruction on a computer monitor monitor: Now, play what you practised as if you were performing it to an audience.
 *Do your best as a performer to produce the piece according to the notation that you just practised*.
@@
 
-In the performing condition, participants were displayed with the following instruction on a computer monitor; Now, play what you practised as if you were performing it to an audience.
+In the performing condition, participants were displayed with the following instruction on a computer monitor: Now, play what you practised as if you were performing it to an audience.
 *Perform the piece expressively with your interpretation*.

### 2021-06-21 paper/manuscript.tex
tweak some wordings
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Deliberate teaching has supported human skill transmission over generations and provides a key route for learning (Thornton \& Raihani, 2008; Tomasello, 2016).
+Deliberate teaching has supported human skill transmission over generations and provides a key route for learning Tomasello (2016).
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
-In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners (Brand, Baldwin, \& Ashburn, 2002; Saint-Georges et al., 2013).
+In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners Saint-Georges et al.
+(2013).
 These modulations include slowing down and exaggerating sounds and actions.
@@
 
-A weighted Yamaha MIDI digital piano was used to record participants' performance via Max/MSP on a MacBook Pro with MacOS X Mojave 10.14.3.
+A weighted Yamaha MIDI digital piano was used to record participants' performance via Max/MSP (\url{https://cycling74.com/products/max}) on a MacBook Pro with Mac OS X Mojave 10.14.3.
 The laptop and piano were connected to a high-fidelity soundcard (Focusrite Scarlett 6i6) to deliver a metronome and piano sound.
@@
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations; see \emph{Figure \ref{fig:stimuli}, A}) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations: see \emph{Figure \ref{fig:stimuli}, A}) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
@@
 For offsets, 14.08\% of the trials contained at least one pitch error (extra notes: 6.28\%, missing notes: 5.17\%, substituted notes: 2.63\%).
-As a result, less than 1 \% of the total responses (i.e., piano keystrokes) were corrected.
+As a result, less than 1 \% of the total responses were corrected.
 In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
@@
 We performed separate analyses for the two expressive techniques (i.e., articulation, dynamics).
-A paired-sample \emph{t}-test or a Wilcoxon signed-rank tests (a non-parametric alternative to a paired \emph{t}-test) was performed to compare the mean IOIs in the teaching and performing condition.
-For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.~performing) and Subcomponent (LS; legato vs.~staccato or FP; forte vs.~piano, respectively).
-The \emph{t.test} function in the \emph{stats} R package was used for a \emph{t}-test.
+A paired-sample \emph{t}-test or a Wilcoxon Signed-rank test (a non-parametric alternative to a paired \emph{t}-test) was performed to compare the mean IOIs in the teaching and performing condition.
+For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.~performing) and Subcomponent (LS: legato vs.~staccato or FP: forte vs.~piano, respectively).
+The \emph{t.test} or \emph{wilcox.test} function in the \emph{stats} R package was used for a \emph{t}-test or a Wilcoxon Signed-rank test.
 For calculating an effect size, we used the \emph{rstatix} R package.
@@
 We first report results for performance of the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli}, B}), followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli}, C}).
-To recall our predictions, if participants play slowly when they are teaching in general, interonset intervals (IOIs) should be larger when teaching.
+To recall our predictions, if participants play more slowly when they are trying to teach, interonset intervals (IOIs) should be larger when teaching.
 If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should reveal exaggerated legato/staccato when teaching articulation, and key velocity (KV) should reveal amplifications of forte and piano when teaching dynamics.
@@
 
-In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} \textless{} 0.001).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41`; \emph{Figure \ref{fig:ioi-1}}).
+In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41: \emph{Figure \ref{fig:ioi-1}}).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011; \emph{Figure \ref{fig:kot-1}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011: \emph{Figure \ref{fig:kot-1}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.004) while there was no significant difference in legato between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition {[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.32).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004; \emph{Figure \ref{fig:vel-1}}).
-However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Legato; \emph{M} = 70.34, \emph{SD} = 6.45, Staccato; \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato; \emph{M} = 70.68, \emph{SD} = 5.84, Staccato; \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato; \emph{p} = 0.70, Staccato; \emph{p} = 0.07).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004: \emph{Figure \ref{fig:vel-1}}).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Legato: \emph{M} = 70.34, \emph{SD} = 6.45, Staccato: \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato: \emph{M} = 70.68, \emph{SD} = 5.84, Staccato: \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato: \emph{p} = 0.70, Staccato: \emph{p} = 0.07).
 
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29; \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29: \emph{Figure \ref{fig:ioi-1}}).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003; \emph{Figure \ref{fig:vel-1}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003: \emph{Figure \ref{fig:vel-1}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.021) while there was no significant difference in piano between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.45).
@@
 Outliers were removed using the same criteria as for the other dependent variables and less than 5 \% of the data were excluded from data analysis.
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032; \emph{Figure \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP; \emph{p} = 0.002) and from piano to forte (PtoF; \emph{p} \textless{} 0.001) in the teaching condition {[}FtoP; \emph{M} = -14.06, \emph{SD} = 7.20, PtoF; \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP; \emph{M} = -11.69, \emph{SD} = 5.97, PtoF; \emph{M} = 20.88, \emph{SD} = 8.67{]}.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032: \emph{Figure \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: \emph{p} = 0.002) and from piano to forte (PtoF: \emph{p} \textless{} 0.001) in the teaching condition {[}FtoP: \emph{M} = -14.06, \emph{SD} = 7.20, PtoF: \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP: \emph{M} = -11.69, \emph{SD} = 5.97, PtoF: \emph{M} = 20.88, \emph{SD} = 8.67{]}.
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31; \emph{Figure \ref{fig:kot-1}}), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.52, \(\eta_G^2\) = 0.00) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.00, \emph{p} = 0.99, \(\eta_G^2\) = 0.00) was significant.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31: \emph{Figure \ref{fig:kot-1}}), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.52, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.00, \emph{p} = 0.99, \(\eta_G^2\) = 0.000) was significant.
 
@@
 For IOIs, we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
-This finding is in line with earlier studies that found slower performance of actions in a teaching context (e.g., Brand et al., 2002; McEllin et al., 2017).
+This finding is in line with earlier studies that found slower performance of actions in a teaching context McEllin et al.
+(2017).
 However, we did not find a significant difference in tempo when participants were teaching dynamics.
@@
 
-We recruited 21 piano experts who already had a degree in (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school at the time of recruitment.
-For data analysis, we excluded one participants due to insufficient motor skills.
+We recruited 21 piano experts who already had a degree (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school at the time of recruitment.
+For data analysis, we excluded one participant due to insufficient motor skills.
 Twenty participants (9 female) were included in the data analysis.
@@
 Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplemental}{Supplementary Materials}}).
-Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 beats per minute).
+Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter-beats per minute).
 The chosen tempo was again cued by a leading metronome.
 In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music.
-As Experiment 1, those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis (as a result, one participant was excluded).
+Those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis (as a result, one participant was excluded).
 The rest of the procedure was identical to Experiment 1.
@@
 Using the same approach as in Experiment 1, pitch errors were removed manually.
-For onsets, 11.62\% of the trials contained at least one pitch error (extra notes; 5.81\%, missing notes; 5.49\%, substituted notes; 0.31\%).
-For offsets, 17.90\% of the trials contained at least one pitch error (extra notes; 5.81\%, missing notes; 5.49\%, substituted notes; 6.59\%).
+For onsets, 11.62\% of the trials contained at least one pitch error (extra notes: 5.81\%, missing notes: 5.49\%, substituted notes: 0.31\%).
+For offsets, 17.90\% of the trials contained at least one pitch error (extra notes: 5.81\%, missing notes: 5.49\%, substituted notes: 6.59\%).
+As a result, less than 1 \% of the total responses were corrected.
 For each dependent variable, removing outliers (i.e., responses outside 3 standard deviations from the mean) resulted in less than 5\% of overall responses being removed.
@@
 
-A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.97, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.95, \emph{SD} = 0.040{]} while playing the piece with notated articulation (\emph{t}(19) = 2.47
-, \emph{p} = 0.023, Cohen's \emph{d} = -0.55; \emph{Figure \ref{fig:ioi-2}}).
-A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two condition (\emph{p} = 0.033, \emph{r} = 0.48).
+A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.97, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.95, \emph{SD} = 0.040{]} while playing the piece with the notated articulation (\emph{t}(19) = 2.47
+, \emph{p} = 0.023, Cohen's \emph{d} = 0.55: \emph{Figure \ref{fig:ioi-2}}).
+A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (\emph{p} = 0.033, \emph{r} = 0.48).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01; \emph{Figure \ref{fig:kor-2}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01: \emph{Figure \ref{fig:kor-2}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.002) while there was no significant difference in legato between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.56).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 7.91, \emph{p} = 0.011, \(\eta_G^2\) = 0.03; \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 7.91, \emph{p} = 0.011, \(\eta_G^2\) = 0.03: \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
 Neither the main effect of Condition (\emph{F}(1,19) = 1.33, \emph{p} = 0.26, \(\eta_G^2\) = 0.003) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 7.91, \emph{p} = 0.71, \(\eta_G^2\) = 0.000) was significant.
+
+\begin{verbatim}
+## Warning: Computation failed in `stat_compare_means()`:
+## 'symbols' must be unique, but are = **|**|*|ns
+
+## Warning: Computation failed in `stat_compare_means()`:
+## 'symbols' must be unique, but are = **|**|*|ns
+\end{verbatim}
 
@@
 
-A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.043{]} while playing the piece with the notated dynamics (\emph{t}(19) = 0.21, \emph{p} = 0.84, Cohen's \emph{d} = -0.05; \emph{Figure \ref{fig:ioi-2}}).
+A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.043{]} while playing the piece with the notated dynamics (\emph{t}(19) = 0.21, \emph{p} = 0.84, Cohen's \emph{d} = 0.05: \emph{Figure \ref{fig:ioi-2}}).
 A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (\emph{p} = 0.96).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 132, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 6.12, \emph{p} = 0.023, \(\eta_G^2\) = 0.018; \emph{Figure \ref{fig:vel-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (\emph{p} = 0.040) and softer piano (\emph{p} = 0.024)in the teaching condition {[}Forte; \emph{M} = 83.20, \emph{SD} = 8.84, Piano; \emph{M} = 60.53, \emph{SD} = 3.76{]} than in the performing condition {[}Forte; \emph{M} = 81.49, \emph{SD} = 8.12, Piano; \emph{M} = 62.42, \emph{SD} = 4.82{]} than in the performing condition.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 132, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 6.12, \emph{p} = 0.023, \(\eta_G^2\) = 0.018: \emph{Figure \ref{fig:vel-2}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (\emph{p} = 0.040) and softer piano (\emph{p} = 0.024) in the teaching condition {[}Forte: \emph{M} = 83.20, \emph{SD} = 8.84, Piano: \emph{M} = 60.53, \emph{SD} = 3.76{]} than in the performing condition {[}Forte: \emph{M} = 81.49, \emph{SD} = 8.12, Piano: \emph{M} = 62.42, \emph{SD} = 4.82{]} than in the performing condition.
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059; \emph{Figure \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition {[}FtoP; \emph{M} = -17.23, \emph{SD} = 8.26, PtoF; \emph{M} = 20.71, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP; \emph{M} = -14.04, \emph{SD} = 7.67, PtoF; \emph{M} = 16.05, \emph{SD} = 7.32{]} (FtoP; \emph{p} = 0.029, PtoF; \emph{p} = 0.002).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059: \emph{Figure \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: \emph{p} = 0.029) and from piano to forte (PtoF: \emph{p} = 0.002) in the teaching condition {[}FtoP: \emph{M} = -17.23, \emph{SD} = 8.26, PtoF: \emph{M} = 20.71, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP: \emph{M} = -14.04, \emph{SD} = 7.67, PtoF: \emph{M} = 16.05, \emph{SD} = 7.32{]}.
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 630, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,19) = 0.081, \emph{p} = 0.78, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 0.35, \emph{p} = 0.56, \(\eta_G^2\) = 0.001; \emph{Figure \ref{fig:kor-2}}) was significant.
+Neither the main effect of Condition (\emph{F}(1,19) = 0.081, \emph{p} = 0.78, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 0.35, \emph{p} = 0.56, \(\eta_G^2\) = 0.001: \emph{Figure \ref{fig:kor-2}}) was significant.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-1} \caption{\label{fig:vel-diff}Experiment 1, 2: KV Difference at transition points when playing the piece with dynamics (left; Experiment 1, right; Experiment 2).
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-1} \caption{\label{fig:vel-diff}Experiment 1, 2: KV Difference at transition points when playing the piece with dynamics (left: Experiment 1, right: Experiment 2).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
@@
 
-Musicians and music teachers tend to consider expressive skills as a performer's most important skills (Lindström, Juslin, Bresin, \& Williamon, 2003).
+Musicians and music teachers tend to consider expressive skills as a performer's most important skill (Lindström, Juslin, Bresin, \& Williamon, 2003).
 In the present study, experts focused on the teaching of expressions that were notated.
@@
 
-In the teaching condition, participants were shown with the following instruction on a computer monitor (Italic sentences were highlighted with a yellow colour on a black background); Now, play what you practised as if you were teaching it to students.
+In the teaching condition, participants were shown with the following instruction on a computer monitor (Italic sentences were highlighted with a yellow colour on a black background): Now, play what you practised as if you were teaching it to students.
 Students already know how to produce the sequence of the tones and now are trying to \emph{learn how to perform the piece expressively} by listening to your performance.
@@
 
-In the performing condition, participants were displayed with the following instruction on a computer monitor monitor; Now, play what you practised as if you were performing it to an audience.
+In the performing condition, participants were displayed with the following instruction on a computer monitor monitor: Now, play what you practised as if you were performing it to an audience.
 \emph{Do your best as a performer to produce the piece according to the notation that you just practised}.
@@
 
-In the performing condition, participants were displayed with the following instruction on a computer monitor; Now, play what you practised as if you were performing it to an audience.
+In the performing condition, participants were displayed with the following instruction on a computer monitor: Now, play what you practised as if you were performing it to an audience.
 \emph{Perform the piece expressively with your interpretation}.

### 2021-06-21 experiment-2/analysis/stats/articulation.R
car_aov > aov_car
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,293 +31,293 @@ dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 
 
-## ----kor-art, echo = FALSE----------------------------------------------------
+## ----kor-art, echo = FALSE----------------------------------
 # For each individual
 kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kor_art
 
 
-## ----kor-art-box,  echo = FALSE-----------------------------------------------
+## ----kor-art-box,  echo = FALSE-----------------------------
 ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
   facet_grid(Subcomponent ~ .)
 
 
[…]

### 2021-06-21 paper/manuscript.Rmd
correct stats and figure (kor_art_2)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 keywords          : "teaching, expertise, skill transmission, artistic expression, music"
-wordcount         : "5022"
+wordcount         : "around 5500 (except references)"
 
@@
 
-Deliberate teaching has supported human skill transmission over generations and provides a key route for learning [@thornton_2008: @tomasello_2016].
+Deliberate teaching has supported human skill transmission over generations and provides a key route for learning [@thornton_2008; @tomasello_2016].
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
-In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners [@brand_2002: @saint-georges_2013].
+In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners [@brand_2002; @saint-georges_2013].
 These modulations include slowing down and exaggerating sounds and actions.
@@
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations: see *Figure \ref{fig:stimuli}, A*) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations, see *Figure \ref{fig:stimuli}, A*) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
@@
 For statistical analysis, only 16th notes with expressive notations were included.
-Overall, three trials were excluded from data analysis because participants did not follow the sheet music or stopped performing before the end.
+Overall, five trials were excluded from data analysis because participants did not follow the sheet music or stopped performing before the end.
 Pitch errors were identified by comparing the sequence of musical notes produced by a participant with the sequence of musical notes according to the sheet music.
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`: *Figure \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, *Figure \ref{fig:ioi-1}*).
 
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`: *Figure \ref{fig:kot-1}*).
+staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`, *Figure \ref{fig:kot-1}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`: *Figure \ref{fig:vel-1}*).
-However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching [Legato: *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`, Staccato: *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] and performing condition [Legato: *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`, Staccato: *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] for each subcomponent (Legato: *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[1])`, Staccato: *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
+staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`, *Figure \ref{fig:vel-1}*).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching [Legato: *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`; Staccato: *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] and performing condition [Legato: *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`, Staccato: *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] for each subcomponent (Legato: *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[1])`, Staccato: *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
 
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`: *Figure \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`, *Figure \ref{fig:ioi-1}*).
 
@@
 performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`: *Figure \ref{fig:vel-1}*).
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-1}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) while there was no significant difference in piano between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`).
@@
 We calculated the KV difference for each interval by subtracting the KV value of the current note from that of the following note.
-Outliers were removed using the same criteria as for the other dependent variables and less than 5 % of the data were excluded from data analysis.
+Outliers were removed using the same criteria as for the other dependent variables and less than 5 % of the data were excluded from analysis.
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`: *Figure \ref{fig:vel-diff}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-diff}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`; PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
 
@@
 performing) and Subcomponent-FP (forte vs.
-piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`: *Figure \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`) was significant.
+piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`, *Figure \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`) was significant.
 
@@
 For IOIs, we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
-This finding is in line with earlier studies that found slower performance of actions in a teaching context [e.g., @brand_2002: @mcellin_2017].
+This finding is in line with earlier studies that found slower performance of actions in a teaching context [e.g., @brand_2002; @mcellin_2017].
 However, we did not find a significant difference in tempo when participants were teaching dynamics.
@@
 
-The results for KOT and KV showed that participants successfully highlighted the relevant aspects of articulation and dynamics.
+The results for KOT and KV showed that participants successfully highlighted relevant aspects of articulation and dynamics.
 Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
-Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points where one of two subcomponents moved to the other (i.e., forte to piano or piano to forte).
+Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points (i.e., forte to piano or piano to forte).
 Importantly, participants did not modulate their performance in terms of irrelevant aspects of the techniques for teaching purposes (e.g., modulating the smoothness of sound while teaching dynamics).
@@
 A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`
-, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art$effsize), 3)`: *Figure \ref{fig:ioi-2}*).
+, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art$effsize), 3)`, *Figure \ref{fig:ioi-2}*).
 A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`).
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`: *Figure \ref{fig:kor-2}*).
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`, *Figure \ref{fig:kor-2}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
@@
   ylim(-1, 0.4) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.4, label.y = -0.5, method = "t.test", paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.45, label.y = -0.5, method = "t.test", paired = TRUE, symnum.args = list(cutpoints = c(0, 0.0001, 0.01, 0.05, 1), symbols = c("***", "**", "*", "ns"))) + # adjust significance level according to the actual stats
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_art_aov$anova_table[2,5])`: *Figure \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
-Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`) was significant.
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_art_aov$anova_table[2,5])`, *Figure \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
+Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`) was significant.
 
@@
   ylim(c(50, 100)) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, method = "t.test", paired = TRUE, symnum.args = list(cutpoints = c(0.001, 0.01, 0.05, 1), symbols = c("**", "**", "*", "ns"))) + # adjust significance level according to the actual stats
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 90, method = "t.test", paired = TRUE) + 
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 
-A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn$effsize), 3)`: *Figure \ref{fig:ioi-2}*).
+A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn$effsize), 3)`, *Figure \ref{fig:ioi-2}*).
 A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`).
@@
 performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`: *Figure \ref{fig:vel-2}*).
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-2}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) and softer piano (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) in the teaching condition [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition.
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`: *Figure \ref{fig:vel-diff}*).
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-diff}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
@@
 piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`: *Figure \ref{fig:kor-2}*) was significant.
+Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:kor-2}*) was significant.
 

### 2021-06-21 paper/manuscript.tex
correct stats and figure (kor_art_2)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \shorttitle{The Sound of Teaching Music}
-\keywords{teaching, expertise, skill transmission, artistic expression, music\newline\indent Word count: 5022}
+\keywords{teaching, expertise, skill transmission, artistic expression, music\newline\indent Word count: around 5500 (except references)}
 \usepackage{csquotes}
@@
 
-Deliberate teaching has supported human skill transmission over generations and provides a key route for learning Tomasello (2016).
+Deliberate teaching has supported human skill transmission over generations and provides a key route for learning (Thornton \& Raihani, 2008; Tomasello, 2016).
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
-In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners Saint-Georges et al.
-(2013).
+In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners (Brand, Baldwin, \& Ashburn, 2002; Saint-Georges et al., 2013).
 These modulations include slowing down and exaggerating sounds and actions.
@@
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations: see \emph{Figure \ref{fig:stimuli}, A}) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations, see \emph{Figure \ref{fig:stimuli}, A}) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41: \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41, \emph{Figure \ref{fig:ioi-1}}).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011: \emph{Figure \ref{fig:kot-1}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011, \emph{Figure \ref{fig:kot-1}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.004) while there was no significant difference in legato between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition {[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.32).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004: \emph{Figure \ref{fig:vel-1}}).
-However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Legato: \emph{M} = 70.34, \emph{SD} = 6.45, Staccato: \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato: \emph{M} = 70.68, \emph{SD} = 5.84, Staccato: \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato: \emph{p} = 0.70, Staccato: \emph{p} = 0.07).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004, \emph{Figure \ref{fig:vel-1}}).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Legato: \emph{M} = 70.34, \emph{SD} = 6.45; Staccato: \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato: \emph{M} = 70.68, \emph{SD} = 5.84, Staccato: \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato: \emph{p} = 0.70, Staccato: \emph{p} = 0.07).
 
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29: \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29, \emph{Figure \ref{fig:ioi-1}}).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003: \emph{Figure \ref{fig:vel-1}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003, \emph{Figure \ref{fig:vel-1}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.021) while there was no significant difference in piano between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.45).
@@
 We calculated the KV difference for each interval by subtracting the KV value of the current note from that of the following note.
-Outliers were removed using the same criteria as for the other dependent variables and less than 5 \% of the data were excluded from data analysis.
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032: \emph{Figure \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: \emph{p} = 0.002) and from piano to forte (PtoF: \emph{p} \textless{} 0.001) in the teaching condition {[}FtoP: \emph{M} = -14.06, \emph{SD} = 7.20, PtoF: \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP: \emph{M} = -11.69, \emph{SD} = 5.97, PtoF: \emph{M} = 20.88, \emph{SD} = 8.67{]}.
+Outliers were removed using the same criteria as for the other dependent variables and less than 5 \% of the data were excluded from analysis.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032, \emph{Figure \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: \emph{p} = 0.002) and from piano to forte (PtoF: \emph{p} \textless{} 0.001) in the teaching condition {[}FtoP: \emph{M} = -14.06, \emph{SD} = 7.20; PtoF: \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP: \emph{M} = -11.69, \emph{SD} = 5.97, PtoF: \emph{M} = 20.88, \emph{SD} = 8.67{]}.
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31: \emph{Figure \ref{fig:kot-1}}), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.52, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.00, \emph{p} = 0.99, \(\eta_G^2\) = 0.000) was significant.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31, \emph{Figure \ref{fig:kot-1}}), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.52, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.000, \emph{p} = 0.99, \(\eta_G^2\) = 0.000) was significant.
 
@@
 For IOIs, we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
-This finding is in line with earlier studies that found slower performance of actions in a teaching context McEllin et al.
-(2017).
+This finding is in line with earlier studies that found slower performance of actions in a teaching context (e.g., Brand et al., 2002; McEllin et al., 2017).
 However, we did not find a significant difference in tempo when participants were teaching dynamics.
@@
 A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.97, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.95, \emph{SD} = 0.040{]} while playing the piece with the notated articulation (\emph{t}(19) = 2.47
-, \emph{p} = 0.023, Cohen's \emph{d} = 0.55: \emph{Figure \ref{fig:ioi-2}}).
+, \emph{p} = 0.023, Cohen's \emph{d} = 0.55, \emph{Figure \ref{fig:ioi-2}}).
 A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (\emph{p} = 0.033, \emph{r} = 0.48).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01: \emph{Figure \ref{fig:kor-2}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01, \emph{Figure \ref{fig:kor-2}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.002) while there was no significant difference in legato between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.56).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 7.91, \emph{p} = 0.011, \(\eta_G^2\) = 0.03: \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
-Neither the main effect of Condition (\emph{F}(1,19) = 1.33, \emph{p} = 0.26, \(\eta_G^2\) = 0.003) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 7.91, \emph{p} = 0.71, \(\eta_G^2\) = 0.000) was significant.
-
-\begin{verbatim}
-## Warning: Computation failed in `stat_compare_means()`:
-## 'symbols' must be unique, but are = **|**|*|ns
-
-## Warning: Computation failed in `stat_compare_means()`:
-## 'symbols' must be unique, but are = **|**|*|ns
-\end{verbatim}
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 7.91, \emph{p} = 0.011, \(\eta_G^2\) = 0.03, \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
+Neither the main effect of Condition (\emph{F}(1,19) = 1.33, \emph{p} = 0.26, \(\eta_G^2\) = 0.003) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 0.14, \emph{p} = 0.71, \(\eta_G^2\) = 0.000) was significant.
 
@@
 
-A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.043{]} while playing the piece with the notated dynamics (\emph{t}(19) = 0.21, \emph{p} = 0.84, Cohen's \emph{d} = 0.05: \emph{Figure \ref{fig:ioi-2}}).
+A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.043{]} while playing the piece with the notated dynamics (\emph{t}(19) = 0.21, \emph{p} = 0.84, Cohen's \emph{d} = 0.05, \emph{Figure \ref{fig:ioi-2}}).
 A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (\emph{p} = 0.96).
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 132, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 6.12, \emph{p} = 0.023, \(\eta_G^2\) = 0.018: \emph{Figure \ref{fig:vel-2}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 132, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 6.12, \emph{p} = 0.023, \(\eta_G^2\) = 0.018, \emph{Figure \ref{fig:vel-2}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (\emph{p} = 0.040) and softer piano (\emph{p} = 0.024) in the teaching condition {[}Forte: \emph{M} = 83.20, \emph{SD} = 8.84, Piano: \emph{M} = 60.53, \emph{SD} = 3.76{]} than in the performing condition {[}Forte: \emph{M} = 81.49, \emph{SD} = 8.12, Piano: \emph{M} = 62.42, \emph{SD} = 4.82{]} than in the performing condition.
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059: \emph{Figure \ref{fig:vel-diff}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059, \emph{Figure \ref{fig:vel-diff}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: \emph{p} = 0.029) and from piano to forte (PtoF: \emph{p} = 0.002) in the teaching condition {[}FtoP: \emph{M} = -17.23, \emph{SD} = 8.26, PtoF: \emph{M} = 20.71, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP: \emph{M} = -14.04, \emph{SD} = 7.67, PtoF: \emph{M} = 16.05, \emph{SD} = 7.32{]}.
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 630, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,19) = 0.081, \emph{p} = 0.78, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 0.35, \emph{p} = 0.56, \(\eta_G^2\) = 0.001: \emph{Figure \ref{fig:kor-2}}) was significant.
+Neither the main effect of Condition (\emph{F}(1,19) = 0.081, \emph{p} = 0.78, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 0.35, \emph{p} = 0.56, \(\eta_G^2\) = 0.001, \emph{Figure \ref{fig:kor-2}}) was significant.
 

### 2021-06-21 paper/manuscript.Rmd
correction from feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-The results from Experiment 2 replicated our earlier findings and provide further evidence that skilled pianists were able to modulate their performance for teaching purposes.
+The results from Experiment 2 replicate our earlier findings and provide further evidence that skilled pianists were able to modulate their performance for teaching purposes.
 As in Experiment 1, we observed slower performance only while teaching the notated articulation.
@@
 Although participants tended to play slower in the teaching condition than in the performing condition in general, we found a significant difference only for articulation, not for dynamics.
-As we mentioned in the discussion of Experiment 1, it could be that the metronome beats given prior to each performance discouraged participants from deviating from the prescribed tempo.
+As mentioned in the discussion of Experiment 1, it could be that the metronome beats given prior to each performance discouraged participants from deviating from the prescribed tempo.
 Another possibility is that expression is actually best taught when leaving the tempo unchanged when it is irrelevant.
-This would imply that general didactic modulations like slowing down might be less useful in the context of teaching expressive artistic techniques where timing itself can be used to add an expression to a performance.
+This would imply that general didactic modulations like slowing down might be less useful in the context of teaching expressive techniques where timing itself can be used to add an expression to a performance.
 

### 2021-06-21 paper/manuscript.Rmd
correction after the first feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 classoption       : "man"
-output:
-  papaja::apa6_pdf:
-    latex_engine: xelatex
+output: papaja::apa6_word
 
 ---
+
 ```{r set-up, include = FALSE}
@@
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations, see *Figure \ref{fig:stimuli}, A*) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations, *Figure \ref{fig:stimuli}, A*) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
@@
 
-First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with notated expressions of either articulation or dynamics (*Figure \ref{fig:stimuli}, B-C*).
+First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with the notated expression of either articulation or dynamics (*Figure \ref{fig:stimuli}, B-C*).
 As soon as they had produced the excerpt with the notated expression without pitch errors twice consecutively, the test trials began.
@@
 For offsets, `r round(sum(error_offset_summary$N)/((31*4*8)-5)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*8)-5)*100, 2)`%, missing notes: `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*8)-5)*100, 2)`%, substituted notes: `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*8)-5)*100, 2)`%).
-As a result, less than 1 % of the total responses were corrected.
+As a result, less than 1 % of total responses were corrected.
 In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-diff}*).
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-diff}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`; PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
@@
 The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
-The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter.
+The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
 Non-expressive sheet music was used for the purpose of practice (*Figure \ref{fig:stimuli-2}, A*).
@@
 In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music.
-Those who could not perform the piece twice consecutively within 5 attempts were excluded from data analysis (as a result, one participant was excluded).
+Those who could not perform the piece twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded).
 The rest of the procedure was identical to Experiment 1.
@@
 For offsets, `r round(sum(error_offset_summary_2$N)/((20*4*8)-3)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_offset_summary_2[Reason == "Extra Notes"]$N/((20*4*8)-3)*100, 2)`%, missing notes: `r round(error_offset_summary_2[Reason == "Missing Notes"]$N/((20*4*8)-3)*100, 2)`%, substituted notes: `r round(error_offset_summary_2[Reason == "Substituted"]$N/((20*4*8)-3)*100, 2)`%).
-As a result, less than 1 % of the total responses were corrected.
+As a result, less than 1 % of total responses were corrected.
 For each dependent variable, removing outliers (i.e., responses outside 3 standard deviations from the mean) resulted in less than 5% of overall responses being removed.
@@
 ```
-
-\clearpage
 
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction of Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-diff}*).
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-diff}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
@@
 As in Experiment 1, we observed slower performance only while teaching the notated articulation.
-Again, KOR showed that when teaching articulation, participants produced shorter staccato, while there was no significant difference in legato between the two conditions as Experiment 1.
+Again, KOR showed that when teaching articulation, participants produced shorter staccato, while there was no significant difference in legato between the two conditions.
 While teaching dynamics, KV showed that participants produced louder forte and softer piano.
@@
 ### Experiment 1
-In the teaching condition, participants were shown with the following instruction on a computer monitor (Italic sentences were highlighted with a yellow colour on a black background): Now, play what you practised as if you were teaching it to students.
+In the teaching condition, participants were shown the following instruction on a computer monitor (Italic sentences were highlighted in yellow colour on a black background): "Now, play what you practised as if you were teaching it to students.
 Students already know how to produce the sequence of the tones and now are trying to *learn how to perform the piece expressively* by listening to your performance.
-*Do your best as a teacher to produce the piece according to the notation that you just practised*.
+*Do your best as a teacher to produce the piece according to the notation that you just practised.*"
 
-In the performing condition, participants were displayed with the following instruction on a computer monitor monitor: Now, play what you practised as if you were performing it to an audience.
-*Do your best as a performer to produce the piece according to the notation that you just practised*.
+In the performing condition, participants were shown the following instruction on a computer monitor monitor: "Now, play what you practised as if you were performing it to an audience.
+*Do your best as a performer to produce the piece according to the notation that you just practised.*"
 
@@
 
-In the teaching condition, we gave participants the exact same instruction as Experiment 1.
-
-In the performing condition, participants were displayed with the following instruction on a computer monitor: Now, play what you practised as if you were performing it to an audience.
-*Perform the piece expressively with your interpretation*.
-*Do your best as a performer to produce the piece according to the notation that you just practised*.
+In the teaching condition of Experiment 2, we gave participants the exact same instruction as in Experiment 1.
+In the performing condition, participants were given the following instruction on a computer monitor: "Now, play what you practised as if you were performing it to an audience.
+*Perform the piece expressively with your interpretation.
+Do your best as a performer to produce the piece according to the notation that you just practised.*"

### 2021-06-21 paper/manuscript.tex
correction after the first feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations, see \emph{Figure \ref{fig:stimuli}, A}) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations, \emph{Figure \ref{fig:stimuli}, A}) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
@@
 
-First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with notated expressions of either articulation or dynamics (\emph{Figure \ref{fig:stimuli}, B-C}).
+First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with the notated expression of either articulation or dynamics (\emph{Figure \ref{fig:stimuli}, B-C}).
 As soon as they had produced the excerpt with the notated expression without pitch errors twice consecutively, the test trials began.
@@
 For offsets, 14.08\% of the trials contained at least one pitch error (extra notes: 6.28\%, missing notes: 5.17\%, substituted notes: 2.63\%).
-As a result, less than 1 \% of the total responses were corrected.
+As a result, less than 1 \% of total responses were corrected.
 In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
@@
 Outliers were removed using the same criteria as for the other dependent variables and less than 5 \% of the data were excluded from analysis.
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032, \emph{Figure \ref{fig:vel-diff}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032, \emph{Figure \ref{fig:vel-diff}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: \emph{p} = 0.002) and from piano to forte (PtoF: \emph{p} \textless{} 0.001) in the teaching condition {[}FtoP: \emph{M} = -14.06, \emph{SD} = 7.20; PtoF: \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP: \emph{M} = -11.69, \emph{SD} = 5.97, PtoF: \emph{M} = 20.88, \emph{SD} = 8.67{]}.
@@
 The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
-The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter.
+The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
 Non-expressive sheet music was used for the purpose of practice (\emph{Figure \ref{fig:stimuli-2}, A}).
@@
 For offsets, 17.90\% of the trials contained at least one pitch error (extra notes: 5.81\%, missing notes: 5.49\%, substituted notes: 6.59\%).
-As a result, less than 1 \% of the total responses were corrected.
+As a result, less than 1 \% of total responses were corrected.
 For each dependent variable, removing outliers (i.e., responses outside 3 standard deviations from the mean) resulted in less than 5\% of overall responses being removed.
@@
 \end{figure}
-
-\clearpage
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059, \emph{Figure \ref{fig:vel-diff}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059, \emph{Figure \ref{fig:vel-diff}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: \emph{p} = 0.029) and from piano to forte (PtoF: \emph{p} = 0.002) in the teaching condition {[}FtoP: \emph{M} = -17.23, \emph{SD} = 8.26, PtoF: \emph{M} = 20.71, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP: \emph{M} = -14.04, \emph{SD} = 7.67, PtoF: \emph{M} = 16.05, \emph{SD} = 7.32{]}.
@@
 
-In the teaching condition, participants were shown with the following instruction on a computer monitor (Italic sentences were highlighted with a yellow colour on a black background): Now, play what you practised as if you were teaching it to students.
+In the teaching condition, participants were shown the following instruction on a computer monitor (Italic sentences were highlighted in yellow colour on a black background): ``Now, play what you practised as if you were teaching it to students.
 Students already know how to produce the sequence of the tones and now are trying to \emph{learn how to perform the piece expressively} by listening to your performance.
-\emph{Do your best as a teacher to produce the piece according to the notation that you just practised}.
+\emph{Do your best as a teacher to produce the piece according to the notation that you just practised.}''
 
-In the performing condition, participants were displayed with the following instruction on a computer monitor monitor: Now, play what you practised as if you were performing it to an audience.
-\emph{Do your best as a performer to produce the piece according to the notation that you just practised}.
+In the performing condition, participants were shown the following instruction on a computer monitor monitor: ``Now, play what you practised as if you were performing it to an audience.
+\emph{Do your best as a performer to produce the piece according to the notation that you just practised.}''
 
@@
 
-In the teaching condition, we gave participants the exact same instruction as Experiment 1.
-
-In the performing condition, participants were displayed with the following instruction on a computer monitor: Now, play what you practised as if you were performing it to an audience.
-\emph{Perform the piece expressively with your interpretation}.
-\emph{Do your best as a performer to produce the piece according to the notation that you just practised}.
+In the teaching condition of Experiment 2, we gave participants the exact same instruction as in Experiment 1.
+In the performing condition, participants were given the following instruction on a computer monitor: ``Now, play what you practised as if you were performing it to an audience.
+\emph{Perform the piece expressively with your interpretation.
+Do your best as a performer to produce the piece according to the notation that you just practised.}''
 

### 2021-07-11 paper/manuscript.Rmd
remove latex/pdf
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-bibliography      : ["references.bib"]
+bibliography      : ["reference.bib"]
+csl: plos.csl
 
-floatsintext      : yes
+floatsintext      : no
 figurelist        : no
@@
 footnotelist      : no
-linenumbers       : no
+linenumbers       : yes
 mask              : no
@@
 
-documentclass     : "apa6"
-classoption       : "man"
 output: papaja::apa6_word
-
 ---
@@
 
-# 1. Introduction
+# Introduction
 
@@
 
-# 2. Experiment 1
+# Experiment 1
 
-## 2.1. Method
+## Materials and Methods
 ```{r methods-1, include = FALSE}
@@
 
-### 2.1.1. Participants
+### Participants
 
@@
 
-### 2.1.2. Apparatus and stimuli
+### Apparatus and stimuli
 
@@
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations, *Figure \ref{fig:stimuli}, A*) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations, *Fig \ref{fig:stimuli}, A*) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
-They referred to either articulation or dynamics (*Figure \ref{fig:stimuli}, B-C*).
+They referred to either articulation or dynamics (*Fig \ref{fig:stimuli}, B-C*).
 Articulation was notated as either legato or staccato.
@@
 
-### 2.1.3. Procedure
+### Procedure
 
-First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with the notated expression of either articulation or dynamics (*Figure \ref{fig:stimuli}, B-C*).
+First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with the notated expression of either articulation or dynamics (*Fig \ref{fig:stimuli}, B-C*).
 As soon as they had produced the excerpt with the notated expression without pitch errors twice consecutively, the test trials began.
@@
 
-### 2.1.4. Data analysis
+### Data analysis
 
@@
 For data analysis, only the 16th notes were included.", echo = FALSE}
-stim_a <- image_border(image_read(here("paper/image/sheetmusic-1", "stim_a.png")), "#FFFFFF", "120x120")
-stim_d <- image_border(image_read(here("paper/image/sheetmusic-1", "stim_d.png")), "#FFFFFF", "120x120")
-stim_n <- image_border(image_read(here("paper/image/sheetmusic-1", "stim_n.png")), "#FFFFFF", "120x120")
+stim_a <- image_border(image_read(here("paper/sheetmusic/1", "stim_a.png")), "#FFFFFF", "120x120")
+stim_d <- image_border(image_read(here("paper/sheetmusic/1", "stim_d.png")), "#FFFFFF", "120x120")
+stim_n <- image_border(image_read(here("paper/sheetmusic/1", "stim_n.png")), "#FFFFFF", "120x120")
 
@@
 img_stimuli
+
+# save
+image_write(img_stimuli, path = here("paper/image/", "stimuli.tiff"), format = "tiff")
 ```
@@
 
-## 2.2. Results
+## Results
 
@@
 All effects are reported as significant at *p* < .05.
-We first report results for performance of the piece with the notated expression of articulation (*Figure \ref{fig:stimuli}, B*), followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli}, C*).
+We first report results for performance of the piece with the notated expression of articulation (*Fig \ref{fig:stimuli}, B*), followed by performance of the piece with the notated expression of dynamics (*Fig \ref{fig:stimuli}, C*).
 To recall our predictions, if participants play more slowly when they are trying to teach, interonset intervals (IOIs) should be larger when teaching.
@@
 
-### 2.2.1. Articulation: Interonset Intervals (IOIs)
+## Articulation
+### Interonset Intervals (IOIs)
 
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, *Figure \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, *Fig \ref{fig:ioi-1}*).
 
@@
 
-### 2.2.2. Articulation: Key-Overlap Time (KOT)
+### Articulation: Key-Overlap Time (KOT)
 
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`, *Figure \ref{fig:kot-1}*).
+staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`, *Fig \ref{fig:kot-1}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
@@
 
-### 2.2.3. Articulation: Key Velocity (KV)
+### Key Velocity (KV)
 
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`, *Figure \ref{fig:vel-1}*).
+staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`, *Fig \ref{fig:vel-1}*).
 However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching [Legato: *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`; Staccato: *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] and performing condition [Legato: *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`, Staccato: *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] for each subcomponent (Legato: *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[1])`, Staccato: *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
@@
 
-### 2.2.4. Dynamics: Interonset Intervals (IOIs)
+## Dynamics
+### Interonset Intervals (IOIs)
 
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`, *Figure \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`, *Fig \ref{fig:ioi-1}*).
 
@@
 
-### 2.2.5. Dynamics: Key Velocity (KV)
+### Key Velocity (KV)
 
@@
 performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-1}*).
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-1}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) while there was no significant difference in piano between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`).
@@
 
-### 2.2.6. Dynamics: KV Transition Points
+### KV Transition Points
 
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-diff}*).
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`; PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
@@
 
-### 2.2.7. Dynamics: Key-Overlap Time (KOT)
+### Key-Overlap Time (KOT)
 
@@
 performing) and Subcomponent-FP (forte vs.
-piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`, *Figure \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
+piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`, *Fig \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
 Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`) was significant.
@@
 
-## 2.3. Discussion
+## Discussion
 
@@
 img_ioi_1 <- image_append(c(img_ioi_art_1, img_ioi_dyn_1))
-#image_write(img_ioi_1, path = here("paper/image", "img_ioi_1.png"), format = "png")
+image_write(img_ioi_1, path = here("paper/image", "img_ioi_1.tiff"), format = "tiff")
 img_ioi_1
@@
 Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
-# kot exaggeration
-kot_exaggeration_1 <- data.table("Legato" == kot_art_all$Mean)
-
 # combine
 img_kot_1 <- image_append(c(img_kot_art_1, img_kot_dyn_1))
-#image_write(img_kot_1, path = here("paper/image", "img_kot_1.png"), format = "png")
+image_write(img_kot_1, path = here("paper/image", "img_kot_1.tiff"), format = "tiff")
 img_kot_1
@@
 img_vel_1 <- image_append(c(img_vel_art_1, img_vel_dyn_1))
-#image_write(img_vel_1, path = here("paper/image", "img_vel_1.png"), format = "png")
+image_write(img_vel_1, path = here("paper/image", "img_vel_1.tiff"), format = "tiff")
 img_vel_1
@@
 
-# 3. Experiment 2
-## 3.1. Method
+# Experiment 2
+## Materials and Methods
 ```{r methods-2, include = FALSE}
@@
 
-### 3.1.1. Participants
+### Participants
 
@@
 
-### 3.1.2. Apparatus and stimuli
+### Apparatus and stimuli
 
@@
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Non-expressive sheet music was used for the purpose of practice (*Figure \ref{fig:stimuli-2}, A*).
-Expressive notations were added to the non-expressive sheet music for the experiment (*Figure \ref{fig:stimuli-2}, B-C*).
+Non-expressive sheet music was used for the purpose of practice (*Fig \ref{fig:stimuli-2}, A*).
+Expressive notations were added to the non-expressive sheet music for the experiment (*Fig \ref{fig:stimuli-2}, B-C*).
 These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
@@
 
-### 3.1.3. Procedure{#procedure-2}
+### Procedure{#procedure-2}
 
 We employed the same procedure as in Experiment 1 with several modifications.
-First, all participants were required to memorise the piece without expressive notations (i.e., *Figure \ref{fig:stimuli-2}, A*) prior to the experiment.
+First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2}, A*) prior to the experiment.
 Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[Supplementary Materials](#supplemental)*).
@@
 
-### 3.1.4. Data analysis
+### Data analysis
 
@@
 For data analysis, only the 8th notes were included.", echo = FALSE}
-stim_a_2 <- image_border(image_read(here("paper/image/sheetmusic-2", "stim_a.png")), "#FFFFFF", "120x120")
-stim_d_2 <- image_border(image_read(here("paper/image/sheetmusic-2", "stim_d.png")), "#FFFFFF", "120x120")
-stim_n_2 <- image_border(image_read(here("paper/image/sheetmusic-2", "stim_n.png")), "#FFFFFF", "120x120")
+stim_a_2 <- image_border(image_read(here("paper/sheetmusic/2", "stim_a.png")), "#FFFFFF", "120x120")
+stim_d_2 <- image_border(image_read(here("paper/sheetmusic/2", "stim_d.png")), "#FFFFFF", "120x120")
+stim_n_2 <- image_border(image_read(here("paper/sheetmusic/2", "stim_n.png")), "#FFFFFF", "120x120")
 
@@
 img_stimuli_2
+
+# save
+image_write(img_stimuli_2, path = here("paper/image/", "stimuli_2.tiff"), format = "tiff")
 ```
 
-## 3.2. Results
+## Results
 
@@
 
-As Experiment 1, we first report results for performance of the piece with the notated expression of articulation (*Figure \ref{fig:stimuli-2}, B*), followed by performance of the piece with the notated expression of dynamics (*Figure \ref{fig:stimuli-2}, C*).
+As Experiment 1, we first report results for performance of the piece with the notated expression of articulation (*Fig \ref{fig:stimuli-2}, B*), followed by performance of the piece with the notated expression of dynamics (*Fig \ref{fig:stimuli-2}, C*).
 
-### 3.2.1. Articulation: Interonset Intervals (IOIs)
+## Articulation
+### Interonset Intervals (IOIs)
 
 A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`
-, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art$effsize), 3)`, *Figure \ref{fig:ioi-2}*).
+, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art$effsize), 3)`, *Fig \ref{fig:ioi-2}*).
 A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`).
@@
 
-### 3.2.2. Articulation: Key-Overlap Ratio (KOR)
+### Key-Overlap Ratio (KOR)
 
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`, *Figure \ref{fig:kor-2}*).
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
@@
 
-### 3.2.3. Articulation: Key Velocity (KV)
+### Key Velocity (KV)
 
@@
 performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_art_aov$anova_table[2,5])`, *Figure \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
+staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_art_aov$anova_table[2,5])`, *Fig \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
 Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`) was significant.
@@
 
-### 3.2.4. Dynamics: Interonset Intervals (IOIs)
+## Dynamics
+### Interonset Intervals (IOIs)
 
-A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn$effsize), 3)`, *Figure \ref{fig:ioi-2}*).
+A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn$effsize), 3)`, *Fig \ref{fig:ioi-2}*).
 A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`).
@@
 
-### 3.2.5. Dynamics: Key Velocity (KV)
+### Key Velocity (KV)
 
@@
 performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-2}*).
+piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-2}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) and softer piano (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) in the teaching condition [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition.
@@
 
-### 3.2.6. Dynamics: KV Transition Points
+### KV Transition Points
 
@@
 performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:vel-diff}*).
+PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
@@
 
-### 3.2.7. Dynamics: Key-Overlap Ratio (KOR)
+### Key-Overlap Ratio (KOR)
 
@@
 piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`, *Figure \ref{fig:kor-2}*) was significant.
+Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*) was significant.
 
@@
 
-## 3.3. Discussion
+## Discussion
 
@@
 img_ioi_2 <- image_append(c(img_ioi_art_2, img_ioi_dyn_2))
-#image_write(img_ioi_2, path = here("paper/image", "img_ioi_2.png"), format = "png")
+image_write(img_ioi_2, path = here("paper/image", "img_ioi_2.tiff"), format = "tiff")
 img_ioi_2
@@
 img_kor_2 <- image_append(c(img_kor_art_2, img_kor_dyn_2))
-#image_write(img_kor_2, path = here("paper/image", "img_kor_2.png"), format = "png")
+image_write(img_kor_2, path = here("paper/image", "img_kor_2.tiff"), format = "tiff")
 img_kor_2
@@
 img_vel_2 <- image_append(c(img_vel_art_2, img_vel_dyn_2))
-#image_write(img_vel_2, path = here("paper/image", "img_vel_2.png"), format = "png")
+image_write(img_vel_2, path = here("paper/image", "img_vel_2.tiff"), format = "tiff")
 img_vel_2
@@
 img_vel_diff <- image_append(c(img_vel_diff_dyn_1, img_vel_diff_dyn_2))
-#image_write(img_vel_diff, path = here("paper/image", "img_vel_diff.png"), format = "png")
+image_write(img_vel_diff, path = here("paper/image", "img_vel_diff.tiff"), format = "tiff")
 img_vel_diff
@@
 
-# 4. General Discussion
+# General Discussion
 

### 2021-07-12 paper/manuscript.Rmd
reflect feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title             : "The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices"
-shorttitle        : "The Sound of Teaching Music"
+title             : "The sound of teaching music: Expert pianists' performance modulations for novices"
+shorttitle        : "The sound of teaching music"
 
@@
   - id            : "1"
-    institution   : "Department of Cognitive Science, Central European University"
+    institution   : "Department of Cognitive Science, Central European University, Vienna, Austria"
 
@@
 keywords          : "teaching, expertise, skill transmission, artistic expression, music"
-wordcount         : "around 5500 (except references)"
+#wordcount         : ""
 
@@
 
-## Materials and Methods
+## Materials and methods
 ```{r methods-1, include = FALSE}
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-LS (legato vs.
-staccato) revealed that there was a significant main effect of Subcomponent-LS (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`, *Fig \ref{fig:kot-1}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
+performing) and Articulation (legato vs.
+staccato) revealed that there was a significant main effect of Articulation (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Articulation (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`, *Fig \ref{fig:kot-1}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato key overlap time was more negative in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`).
+For legato there was no significant difference in key overlap time between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`, *Fig \ref{fig:vel-1}*).
+performing) and Articulation (legato vs.
+staccato) showed that there was a significant interaction between Condition and Articulation (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`, *Fig \ref{fig:vel-1}*).
 However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching [Legato: *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`; Staccato: *M* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] and performing condition [Legato: *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`, Staccato: *M* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r vel_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] for each subcomponent (Legato: *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[1])`, Staccato: *p* = `r sprintf('%.2f', vel_art_posthoc$p.value[2])`).
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-1}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) while there was no significant difference in piano between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`).
+performing) and Dynamics (forte vs.
+piano) showed that there was a significant main effect of Dynamics (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Dynamics (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-1}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced higher key velocity in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) when instructed to play forte.
+There was no significant difference in key velocity between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`) when instructed to play piano.
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`; PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
+performing) and Transition Type (FtoP vs.
+PtoF) showed that there was a significant main effect of Dynamics (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`; PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-FP (forte vs.
-piano) revealed that there was a significant main effect of Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`, *Fig \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`) was significant.
+performing) and Dynamics (forte vs.
+piano) revealed that there was a significant main effect of Dynamics (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`, *Fig \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Dynamics (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`) was significant.
 
@@
 These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
-
-In Experiment 1, we employed a simple musical scale in order to maximise experimental control.
-To test whether our results generalise to a more naturalistic piece, in Experiment 2, we chose an actual piano piece and modified it for the purpose of the experiment.
-If pianists selectively highlight the relevant aspects of the techniques to be taught also in a more naturalistic piece containing more opportunities for expression, they should again produce longer legato and shorter staccato while teaching articulation.
-When teaching dynamics, they should produce louder forte and softer piano.
-Given the findings we observed in Experiment 1, we also predicted that participants would make a larger contrast between forte and piano at transition points, and that they might play more slowly when teaching, especially when teaching articulation.
 
@@
 # Experiment 2
-## Materials and Methods
+
+In Experiment 1, we employed a simple musical scale in order to maximise experimental control.
+To test whether our results generalise to a more naturalistic piece, in Experiment 2, we chose an actual piano piece and modified it for the purpose of the experiment.
+If pianists selectively highlight the relevant aspects of the techniques to be taught also in a more naturalistic piece containing more opportunities for expression, they should again produce longer legato and shorter staccato while teaching articulation.
+When teaching dynamics, they should produce louder forte and softer piano.
+Given the findings we observed in Experiment 1, we also predicted that participants would make a larger contrast between forte and piano at transition points, and that they might play more slowly when teaching, especially when teaching articulation.
+
+## Materials and methods
 ```{r methods-2, include = FALSE}
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-LS (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`) while there was no significant difference in legato between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
+performing) and Articulation (legato vs.
+staccato) showed that there was a significant main effect of Articulation (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Articulation (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato key overlap time was more negative in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`).
+For letago here was no significant difference in key overlap time between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-LS (legato vs.
-staccato) showed that there was a significant main effect of Subcomponent-LS (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_art_aov$anova_table[2,5])`, *Fig \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
-Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-LS (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`) was significant.
+performing) and Articulation (legato vs.
+staccato) showed that there was a significant main effect of Articulation (*F*(`r vel_art_aov$anova_table[2,1]`,`r vel_art_aov$anova_table[2,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[2,4])`, *p* `r if(vel_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_art_aov$anova_table[2,5])`, *Fig \ref{fig:vel-2}*), reflecting louder sound for legato notes compared to staccato notes.
+Neither the main effect of Condition (*F*(`r vel_art_aov$anova_table[1,1]`,`r vel_art_aov$anova_table[1,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[1,4])`, *p* `r if(vel_art_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[1,5])`) nor the interaction between Condition and Articulation (*F*(`r vel_art_aov$anova_table[3,1]`,`r vel_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_art_aov$anova_table[3,4])`, *p* `r if(vel_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_art_aov$anova_table[3,5])`) was significant.
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) and softer piano (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) in the teaching condition [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition.
+performing) and Dynamics (forte vs.
+piano) showed that there was a significant main effect of Dynamics (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Dynamics (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-2}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced higher key velocity (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) when instructed to play forte.
+Also, participants produced lower key velocity (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) in the teaching condition [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition when instructed to play piano.
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-FP (FtoP vs.
-PtoF) showed that there was a significant main effect of Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Subcomponent-FP (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
+performing) and Transition Type (FtoP vs.
+PtoF) showed that there was a significant main effect of Dynamics (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Subcomponent-FP (forte vs.
-piano) showed that there was a significant main effect of Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Subcomponent-FP (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*) was significant.
+performing) and Dynamics (forte vs.
+piano) showed that there was a significant main effect of Dynamics (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Dynamics (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*) was significant.
 
@@
 We thank Dávid Csűrös for his help with data collection.
-This research was supported by the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013)/ERC (European Research Council) grant agreement no.
-609819, SOMICS, and by ERC grant agreement no.
-616072, JAXPERTISE.
-
-# Declarations of interest
-The authors have no conflict of interest.
 
@@
 
-# Supplementary Materials{#supplementary}
-## Instructions
-### Experiment 1
+# Supporting information{#supplementary}
+
+# S1 File. Instructions
+## Experiment 1
+
 In the teaching condition, participants were shown the following instruction on a computer monitor (Italic sentences were highlighted in yellow colour on a black background): "Now, play what you practised as if you were teaching it to students.
@@
 
-### Experiment 2
+## Experiment 2
 
@@
 Do your best as a performer to produce the piece according to the notation that you just practised.*"
+
+# General information
+## Section/Category
+Social and Behavioral Sciences
+
+## Opposed reviewers
+N/A?
+
+## Required statements
+
+- This manuscript adheres to the appropriate reporting guidelines and community standards for data availability.
+- Any related manuscripts currently in press or under consideration elsewhere are mentioned in the cover letter and will be uploaded as part of your submission as a related manuscript.
+- Any persons named in the Acknowledgments section of the manuscript, or referred to as the source of a personal communication, have agreed to being so named.
+- All authors have read, and confirm that they meet, ICMJE criteria for authorship.
+- All contributing authors are aware of and agree to the submission of this manuscript.
+
+Agreed to all?
+
+## United States Government Employee
+No authors are employees of the U.S. government.
+
+## Financial Disclosure
+This research was supported by the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013)/ERC (European Research Council) grant agreement no.
+609819, SOMICS, awarded to Günther Knoblich.
+and by ERC grant agreement no.
+616072, JAXPERTISE, awarded to Natalie Sebanz.
+
+## Competing Interests
+The authors have declared that no competing interests exist.
+
+## Ethics Statement
+The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary (2018-124). The written form of consent was obtained.
+
+## Previous Interactions
+
+- I have had previous interactions about this manuscript with an Academic Editor of this journal.
+- This manuscript was previously submitted to a different PLOS journal as either a presubmission inquiry or a full submission.
+- One or more of the authors (including myself) currently serve, or have previously served, as an Academic Editor or Guest Editor for this journal.
+
+None of them?
+
+## Group Authorship
+No
+
+## Request Editor
+- Stefan Koelsch
+- Any suggestion?
+
+## Data Availability
+- Yes - all data are fully available without restriction
+
+## Describe where the data may be found in full sentences.
+All data files are available from the Open Science Framework repository (DOI:10.17605/OSF.IO/8NBJH).
+
+## Immediate Preprint Posting
+Yes?
+
+## Publication Fees
+- I am ineligible for any Open Access Funding Support and will pay the full fee upon article acceptance
+
+## Author Contributions
+- Conceptualization: Atsuko Tominaga, Günther Knoblich, Natalie Sebanz.
+- Data curation: Atsuko Tominaga.
+- -Formal analysis: Atsuko Tominaga, Günther Knoblich, Natalie Sebanz.
+Funding acquisition: Günther Knoblich, Natalie Sebanz.
+- Methodology: Atsuko Tominaga, Günther Knoblich, Natalie Sebanz.
+- Project administration: Natalie Sebanz.
+- Supervision: Günther Knoblich, Natalie Sebanz.
+- Visualization: Atsuko Tominaga.
+- Writing – original draft: Atsuko Tominaga.
+- Writing – review & editing: Atsuko Tominaga, Günther Knoblich, - Natalie Sebanz.

### 2021-07-13 paper/manuscript.Rmd
export figures for another folder
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   dir.create(here("paper/image"))
+}
+if (!file.exists(here("paper/figure"))){
+  dir.create(here("paper/figure"))
 }
@@
 # save
-image_write(img_stimuli, path = here("paper/image/", "stimuli.tiff"), format = "tiff")
+image_write(img_stimuli, path = here("paper/figure", "Fig1.tiff"), format = "tiff")
 ```
@@
 To recall our predictions, if participants play more slowly when they are trying to teach, interonset intervals (IOIs) should be larger when teaching.
-If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should reveal exaggerated legato/staccato when teaching articulation, and key velocity (KV) should reveal amplifications of forte and piano when teaching dynamics.
+If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should reveal be more positive for legato and more negative for staccato when teaching articulation, and key velocity (KV) should reveal amplifications of forte and piano when teaching dynamics.
 
@@
 performing) and Transition Type (FtoP vs.
-PtoF) showed that there was a significant main effect of Dynamics (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
+PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`; PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
@@
 The results for KOT and KV showed that participants successfully highlighted relevant aspects of articulation and dynamics.
-Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
+Specifically, key-overlap time was more negative when teaching articulation and key velocity for forte was higher when teaching dynamics.
 Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points (i.e., forte to piano or piano to forte).
@@
 img_ioi_1 <- image_append(c(img_ioi_art_1, img_ioi_dyn_1))
-image_write(img_ioi_1, path = here("paper/image", "img_ioi_1.tiff"), format = "tiff")
+image_write(img_ioi_1, path = here("paper/figure", "Fig1.tiff"), format = "tiff")
 img_ioi_1
@@
 img_kot_1 <- image_append(c(img_kot_art_1, img_kot_dyn_1))
-image_write(img_kot_1, path = here("paper/image", "img_kot_1.tiff"), format = "tiff")
+image_write(img_kot_1, path = here("paper/figure", "Fig3.tiff"), format = "tiff")
 img_kot_1
@@
 img_vel_1 <- image_append(c(img_vel_art_1, img_vel_dyn_1))
-image_write(img_vel_1, path = here("paper/image", "img_vel_1.tiff"), format = "tiff")
+image_write(img_vel_1, path = here("paper/figure", "Fig4.tiff"), format = "tiff")
 img_vel_1
@@
 # save
-image_write(img_stimuli_2, path = here("paper/image/", "stimuli_2.tiff"), format = "tiff")
+image_write(img_stimuli_2, path = here("paper/figure", "Fig5.tiff"), format = "tiff")
 ```
@@
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced higher key velocity (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) when instructed to play forte.
-Also, participants produced lower key velocity (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) in the teaching condition [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] than in the performing condition when instructed to play piano.
+Also, participants produced lower key velocity (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) in the teaching condition [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] when instructed to play piano.
 
@@
 performing) and Transition Type (FtoP vs.
-PtoF) showed that there was a significant main effect of Dynamics (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
+PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
@@
 As in Experiment 1, we observed slower performance only while teaching the notated articulation.
-Again, KOR showed that when teaching articulation, participants produced shorter staccato, while there was no significant difference in legato between the two conditions.
-While teaching dynamics, KV showed that participants produced louder forte and softer piano.
-This pattern of exaggeration was more pronounced than in Experiment 1, where only louder forte was observed.
+Again, KOR showed that when teaching articulation, key-overlap time was more negative for staccato, while there was no significant difference in legato between the two conditions.
+While teaching dynamics, KV showed that participants produced higher key velocity for forte and lower key velocity for piano.
+This pattern of exaggeration was more pronounced than in Experiment 1, where only higher key velocity for forte was observed.
 Furthermore, we again found that when teaching dynamics, participants made a larger contrast between forte and piano bidirectionally at transition points.
@@
 img_ioi_2 <- image_append(c(img_ioi_art_2, img_ioi_dyn_2))
-image_write(img_ioi_2, path = here("paper/image", "img_ioi_2.tiff"), format = "tiff")
+image_write(img_ioi_2, path = here("paper/figure", "Fig6.tiff"), format = "tiff")
 img_ioi_2
@@
 img_kor_2 <- image_append(c(img_kor_art_2, img_kor_dyn_2))
-image_write(img_kor_2, path = here("paper/image", "img_kor_2.tiff"), format = "tiff")
+image_write(img_kor_2, path = here("paper/figure", "Fig7.tiff"), format = "tiff")
 img_kor_2
@@
 img_vel_2 <- image_append(c(img_vel_art_2, img_vel_dyn_2))
-image_write(img_vel_2, path = here("paper/image", "img_vel_2.tiff"), format = "tiff")
+image_write(img_vel_2, path = here("paper/figure", "Fig8.tiff"), format = "tiff")
 img_vel_2
@@
 img_vel_diff <- image_append(c(img_vel_diff_dyn_1, img_vel_diff_dyn_2))
-image_write(img_vel_diff, path = here("paper/image", "img_vel_diff.tiff"), format = "tiff")
+image_write(img_vel_diff, path = here("paper/figure", "Fig9.tiff"), format = "tiff")
 img_vel_diff
@@
 Do your best as a performer to produce the piece according to the notation that you just practised.*"
-
-# General information
-## Section/Category
-Social and Behavioral Sciences
-
-## Opposed reviewers
-N/A?
-
-## Required statements
-
-- This manuscript adheres to the appropriate reporting guidelines and community standards for data availability.
-- Any related manuscripts currently in press or under consideration elsewhere are mentioned in the cover letter and will be uploaded as part of your submission as a related manuscript.
-- Any persons named in the Acknowledgments section of the manuscript, or referred to as the source of a personal communication, have agreed to being so named.
-- All authors have read, and confirm that they meet, ICMJE criteria for authorship.
-- All contributing authors are aware of and agree to the submission of this manuscript.
-
-Agreed to all?
-
-## United States Government Employee
-No authors are employees of the U.S. government.
-
-## Financial Disclosure
-This research was supported by the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013)/ERC (European Research Council) grant agreement no.
-609819, SOMICS, awarded to Günther Knoblich.
-and by ERC grant agreement no.
-616072, JAXPERTISE, awarded to Natalie Sebanz.
-
-## Competing Interests
-The authors have declared that no competing interests exist.
-
-## Ethics Statement
-The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary (2018-124). The written form of consent was obtained.
-
-## Previous Interactions
-
-- I have had previous interactions about this manuscript with an Academic Editor of this journal.
-- This manuscript was previously submitted to a different PLOS journal as either a presubmission inquiry or a full submission.
-- One or more of the authors (including myself) currently serve, or have previously served, as an Academic Editor or Guest Editor for this journal.
-
-None of them?
-
-## Group Authorship
-No
-
-## Request Editor
-- Stefan Koelsch
-- Any suggestion?
-
-## Data Availability
-- Yes - all data are fully available without restriction
-
-## Describe where the data may be found in full sentences.
-All data files are available from the Open Science Framework repository (DOI:10.17605/OSF.IO/8NBJH).
-
-## Immediate Preprint Posting
-Yes?
-
-## Publication Fees
-- I am ineligible for any Open Access Funding Support and will pay the full fee upon article acceptance
-
-## Author Contributions
-- Conceptualization: Atsuko Tominaga, Günther Knoblich, Natalie Sebanz.
-- Data curation: Atsuko Tominaga.
-- -Formal analysis: Atsuko Tominaga, Günther Knoblich, Natalie Sebanz.
-Funding acquisition: Günther Knoblich, Natalie Sebanz.
-- Methodology: Atsuko Tominaga, Günther Knoblich, Natalie Sebanz.
-- Project administration: Natalie Sebanz.
-- Supervision: Günther Knoblich, Natalie Sebanz.
-- Visualization: Atsuko Tominaga.
-- Writing – original draft: Atsuko Tominaga.
-- Writing – review & editing: Atsuko Tominaga, Günther Knoblich, - Natalie Sebanz.

### 2021-07-13 paper/manuscript.Rmd
change dpt and tweak fig4
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # save plots
-ggsave(here("paper/image", "ioi_art_1.png"), plot = ioi_art_1, dpi = 600, width = 5, height = 5)
+ggsave(here("paper/image", "ioi_art_1.png"), plot = ioi_art_1, dpi = 300, width = 4, height = 4)
 img_ioi_art_1 <- image_read(here("paper/image", "ioi_art_1.png"))
@@
 # save plots
-ggsave(here("paper/image", "kot_art_1.png"), plot = kot_art_1, dpi = 600, width = 7, height = 5)
+ggsave(here("paper/image", "kot_art_1.png"), plot = kot_art_1, dpi = 300, width = 7, height = 5)
 img_kot_art_1 <- image_read(here("paper/image", "kot_art_1.png"))
@@
 # save plots
-ggsave(here("paper/image", "vel_art_1.png"), plot = vel_art_1, dpi = 600, width = 7, height = 5)
+ggsave(here("paper/image", "vel_art_1.png"), plot = vel_art_1, dpi = 300, width = 7, height = 5)
 img_vel_art_1 <- image_read(here("paper/image", "vel_art_1.png"))
@@
 # save plots
-ggsave(here("paper/image", "ioi_dyn_1.png"), plot = ioi_dyn_1, dpi = 600, width = 5, height = 5)
+ggsave(here("paper/image", "ioi_dyn_1.png"), plot = ioi_dyn_1, dpi = 300, width = 4, height = 4)
 img_ioi_dyn_1 <- image_read(here("paper/image", "ioi_dyn_1.png"))
@@
 # save plots
-ggsave(here("paper/image", "vel_dyn_1.png"), plot = vel_dyn_1, dpi = 600, width = 7, height = 5)
+ggsave(here("paper/image", "vel_dyn_1.png"), plot = vel_dyn_1, dpi = 300, width = 7, height = 5)
 img_vel_dyn_1 <- image_read(here("paper/image", "vel_dyn_1.png"))
@@
 # annotate * for forte
-img_vel_dyn_1 <- image_annotate(img_vel_dyn_1, "*", size = 280, color = "black", location = "+1430+800")
+img_vel_dyn_1 <- image_annotate(img_vel_dyn_1, "*", size = 130, color = "black", location = "+725+400")
 ```
@@
 # save plots
-ggsave(here("paper/image", "vel_diff_dyn_1.png"), plot = vel_diff_dyn_1, dpi = 600, width = 7, height = 5)
+ggsave(here("paper/image", "vel_diff_dyn_1.png"), plot = vel_diff_dyn_1, dpi = 300, width = 7, height = 5)
 img_vel_diff_dyn_1 <- image_read(here("paper/image", "vel_diff_dyn_1.png"))
@@
 # save plots
-ggsave(here("paper/image", "kot_dyn_1.png"), plot = kot_dyn_1, dpi = 600, width = 7, height = 5)
+ggsave(here("paper/image", "kot_dyn_1.png"), plot = kot_dyn_1, dpi = 300, width = 7, height = 5)
 img_kot_dyn_1 <- image_read(here("paper/image", "kot_dyn_1.png"))
@@
 img_ioi_1 <- image_append(c(img_ioi_art_1, img_ioi_dyn_1))
-image_write(img_ioi_1, path = here("paper/figure", "Fig1.tiff"), format = "tiff")
+image_write(img_ioi_1, path = here("paper/figure", "Fig2.tiff"), format = "tiff")
 img_ioi_1
@@
 # save plots
-ggsave(here("paper/image", "ioi_art_2.png"), plot = ioi_art_2, dpi = 600, width = 5, height = 5)
+ggsave(here("paper/image", "ioi_art_2.png"), plot = ioi_art_2, dpi = 300, width = 4, height = 4)
 img_ioi_art_2 <- image_read(here("paper/image", "ioi_art_2.png"))
@@
 # save plots
-ggsave(here("paper/image", "kor_art_2.png"), plot = kor_art_2, dpi = 600, width = 7, height = 5)
+ggsave(here("paper/image", "kor_art_2.png"), plot = kor_art_2, dpi = 300, width = 7, height = 5)
 img_kor_art_2 <- image_read(here("paper/image", "kor_art_2.png"))
@@
 # save plots
-ggsave(here("paper/image", "vel_art_2.png"), plot = vel_art_2, dpi = 600, width = 7, height = 5)
+ggsave(here("paper/image", "vel_art_2.png"), plot = vel_art_2, dpi = 300, width = 7, height = 5)
 img_vel_art_2 <- image_read(here("paper/image", "vel_art_2.png"))
@@
 # save plots
-ggsave(here("paper/image", "ioi_dyn_2.png"), plot = ioi_dyn_2, dpi = 600, width = 5, height = 5)
+ggsave(here("paper/image", "ioi_dyn_2.png"), plot = ioi_dyn_2, dpi = 300, width = 4, height = 4)
 img_ioi_dyn_2 <- image_read(here("paper/image", "ioi_dyn_2.png"))
@@
 # save plots
-ggsave(here("paper/image", "vel_dyn_2.png"), plot = vel_dyn_2, dpi = 600, width = 7, height = 5)
+ggsave(here("paper/image", "vel_dyn_2.png"), plot = vel_dyn_2, dpi = 300, width = 7, height = 5)
 img_vel_dyn_2 <- image_read(here("paper/image", "vel_dyn_2.png"))
@@
 # save plots
-ggsave(here("paper/image", "vel_diff_dyn_2.png"), plot = vel_diff_dyn_2, dpi = 600, width = 7, height = 5)
+ggsave(here("paper/image", "vel_diff_dyn_2.png"), plot = vel_diff_dyn_2, dpi = 300, width = 7, height = 5)
 img_vel_diff_dyn_2 <- image_read(here("paper/image", "vel_diff_dyn_2.png"))
@@
 # save plots
-ggsave(here("paper/image", "kor_dyn_2.png"), plot = kor_dyn_2, dpi = 600, width = 7, height = 5)
+ggsave(here("paper/image", "kor_dyn_2.png"), plot = kor_dyn_2, dpi = 300, width = 7, height = 5)
 img_kor_dyn_2 <- image_read(here("paper/image", "kor_dyn_2.png"))
@@
 
-# Supporting information{#supplementary}
-
 # S1 File. Instructions

### 2021-07-15 paper/manuscript.Rmd
tweak results, figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-floatsintext      : no
+floatsintext      : yes
 figurelist        : no
@@
 footnotelist      : no
-linenumbers       : yes
+linenumbers       : no
 mask              : no
@@
 
-output: papaja::apa6_word
+#output: papaja::apa6_word
+output:
+  papaja::apa6_pdf:
+    latex_engine: xelatex
 ---
@@
 Previous research has shown that demonstrators not only adjust their gestures [@campisi_2013] and actions [@fukuyama_2015] to learners' skills but engage in specific action modulations to highlight particular aspects of demonstrated actions.
-For example, @schaik_2019 showed that adults used specific action modulations for demonstrating different action effects of objects to infants.
-@NIPS2016_6413 found that demonstrators chose costly movement paths at structurally important points to disambiguate one goal over other possibilities.
+For example, Schaik and colleagues [@schaik_2019] showed that adults used specific action modulations for demonstrating different action effects of objects to infants.
+Ho and colleagues [@NIPS2016_6413] found that demonstrators chose costly movement paths at structurally important points to disambiguate one goal over other possibilities.
 
@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplementary Materials](#supplementary)*).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[S1 File](#supplementary)*).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
@@
 The order of the expressive techniques within each condition was also blocked and counterbalanced across participants.
-A leading metronome (80 quarter-beats per minute, 8 beats) indicated the target tempo before each trial.
+A leading metronome (80 quarter beats per minute, 8 beats) indicated the target tempo before each trial.
 
@@
 For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.
-performing) and Subcomponent (LS: legato vs.
-staccato or FP: forte vs.
+performing) and Subcomponent (Articulation: legato vs.
+staccato or Dynamics: forte vs.
 piano, respectively).
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, *Fig \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, two-tailed, *Fig \ref{fig:ioi-1}*).
 
@@
 staccato) revealed that there was a significant main effect of Articulation (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Articulation (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`, *Fig \ref{fig:kot-1}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato key overlap time was more negative in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`).
-For legato there was no significant difference in key overlap time between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato key-overlap time was more negative in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`).
+For legato there was no significant difference in key-overlap time between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
 
@@
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`, *Fig \ref{fig:ioi-1}*).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`, two-tailed, *Fig \ref{fig:ioi-1}*).
 
@@
 performing) and Dynamics (forte vs.
-piano) revealed that there was a significant main effect of Dynamics (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`, *Fig \ref{fig:kot-1}*), reflecting more key overlap for forte notes compared to piano notes.
+piano) revealed that there was a significant main effect of Dynamics (*F*(`r kot_dyn_aov$anova_table[2,1]`,`r kot_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_dyn_aov$anova_table[2,4])`, *p* `r if(kot_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_dyn_aov$anova_table[2,5])`, *Fig \ref{fig:kot-1}*), reflecting more key-overlap for forte notes compared to piano notes.
 Neither the main effect of Condition (*F*(`r kot_dyn_aov$anova_table[1,1]`,`r kot_dyn_aov$anova_table[1,2]`) = `r sprintf('%.2f', kot_dyn_aov$anova_table[1,4])`, *p* `r if(kot_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Dynamics (*F*(`r kot_dyn_aov$anova_table[3,1]`,`r kot_dyn_aov$anova_table[3,2]`) = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,4])`, *p* `r if(kot_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kot_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_dyn_aov$anova_table[3,5])`) was significant.
@@
 To test whether our results generalise to a more naturalistic piece, in Experiment 2, we chose an actual piano piece and modified it for the purpose of the experiment.
-If pianists selectively highlight the relevant aspects of the techniques to be taught also in a more naturalistic piece containing more opportunities for expression, they should again produce longer legato and shorter staccato while teaching articulation.
-When teaching dynamics, they should produce louder forte and softer piano.
+If pianists selectively highlight the relevant aspects of the techniques to be taught also in a more naturalistic piece containing more opportunities for expression, key-overlap time should be again more positive for legato and more negative for staccato while teaching articulation.
+When teaching dynamics, participants should produce higher key velocity for forte and lower key velocity for piano.
 Given the findings we observed in Experiment 1, we also predicted that participants would make a larger contrast between forte and piano at transition points, and that they might play more slowly when teaching, especially when teaching articulation.
@@
 They had `r round(mean(dt_2_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_2_included$PianoTotalPractice), 2)`).
+All participants gave their informed consent before the experiment started and received vouchers for their participation.
+The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.
 
@@
 First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2}, A*) prior to the experiment.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[Supplementary Materials](#supplemental)*).
-Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter-beats per minute).
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[S1 File](#supplemental)*).
+Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute).
 The chosen tempo was again cued by a leading metronome.
@@
 A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`
-, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art$effsize), 3)`, *Fig \ref{fig:ioi-2}*).
-A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`).
+, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art$effsize), 3)`, two-tailed, *Fig \ref{fig:ioi-2}*).
+A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, two-tailed).
 
@@
 staccato) showed that there was a significant main effect of Articulation (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Articulation (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato key overlap time was more negative in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`).
-For letago here was no significant difference in key overlap time between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato key-overlap time was more negative in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`).
+For letago here was no significant difference in key-overlap time between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
 
@@
 
-A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn$effsize), 3)`, *Fig \ref{fig:ioi-2}*).
-A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`).
+A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn$effsize), 3)`, two-tailed, *Fig \ref{fig:ioi-2}*).
+A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, two-tailed).
 
@@
 piano) showed that there was a significant main effect of Dynamics (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Dynamics (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced higher key velocity (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) when instructed to play forte.
-Also, participants produced lower key velocity (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) in the teaching condition [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] when instructed to play piano.
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced higher key velocity in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`](*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) when instructed to play forte.
+Also, participants produced lower key velocity in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`](*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) when instructed to play piano.
 
@@
 performing) and Dynamics (forte vs.
-piano) showed that there was a significant main effect of Dynamics (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key overlap for forte notes compared to piano notes.
+piano) showed that there was a significant main effect of Dynamics (*F*(`r kor_dyn_aov$anova_table[2,1]`,`r kor_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_dyn_aov$anova_table[2,4])`, *p* `r if(kor_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_dyn_aov$anova_table[2,5])`), reflecting more key-overlap for forte notes compared to piano notes.
 Neither the main effect of Condition (*F*(`r kor_dyn_aov$anova_table[1,1]`,`r kor_dyn_aov$anova_table[1,2]`) = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,4])`, *p* `r if(kor_dyn_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[1,5])`) nor the interaction between Condition and Dynamics (*F*(`r kor_dyn_aov$anova_table[3,1]`,`r kor_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_dyn_aov$anova_table[3,4])`, *p* `r if(kor_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', kor_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kor_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*) was significant.
@@
 When playing with the intention to teach, pianists selectively highlighted relevant aspects of artistic expressions.
-Across two different pieces, they exaggerated staccato when playing with the intention to teach expressions concerning articulation.
+Across two different pieces, key-overlap time was more negative for staccato when playing with the intention to teach expressions concerning articulation.
 The lack of a difference for legato when teaching articulation might stem from a ceiling effect.
-When playing with the intention to teach expressions concerning dynamics, participants exaggerated changes between forte and piano (in both directions).
+When playing with the intention to teach expressions concerning dynamics, participants made larger key velocity changes between forte and piano (in both directions).
 This constitutes an effective way to highlight the technique as loudness is determined relatively.
@@
 We thank Dávid Csűrös for his help with data collection.
-
-\newpage
 

### 2021-07-16 paper/manuscript.Rmd
latest version / fix figures and some wordings
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-#output: papaja::apa6_word
-output:
-  papaja::apa6_pdf:
-    latex_engine: xelatex
+output: papaja::apa6_word
+# output:
+#   papaja::apa6_pdf:
+#     latex_engine: xelatex
 ---
@@
 Two basic expressive techniques in piano performance were used: articulation (legato and staccato) and dynamics (forte and piano).
+Articulation was assessed by key-overlap time between two consecutive notes.
+Positive key-overlap time indicates legato styles whereas negative key-overlap time indicates staccato styles.
+Dynamics was evaluated by key velocity of each key press.
+Higher key velocity means forte styles while lower key velocity means piano styles.
 
 If experts rely on generic action modulations, it can be expected that they will play more slowly during didactic demonstration, regardless of the kind of expressive techniques to be taught.
-To the extent that experts use action modulations to support their teaching of specific expressive techniques, they should produce longer legato and shorter staccato when teaching articulation whereas they should produce louder forte and softer piano when teaching dynamics.
+To the extent that experts use action modulations to support their teaching of specific expressive techniques, they should exaggerate legato (i.e., key-overlap time should be more positive) and staccato (i.e., key-overlap time should be more negative) when teaching articulation whereas they should exaggerate forte (i.e., key velocity should be higher) and piano (i.e., key velocity should be lower) when teaching dynamics.
 Furthermore, one could speculate that they might produce modulations specifically at structurally important points that best highlight the technique to be taught.
@@
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
-The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.
+The study (No.
+2018-124) was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.
 
@@
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations, *Fig \ref{fig:stimuli}, A*) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations, *Fig \ref{fig:stimuli} A*) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
-They referred to either articulation or dynamics (*Fig \ref{fig:stimuli}, B-C*).
+They referred to either articulation or dynamics (*Fig \ref{fig:stimuli} B-C*).
 Articulation was notated as either legato or staccato.
@@
 
-First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with the notated expression of either articulation or dynamics (*Fig \ref{fig:stimuli}, B-C*).
+First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with the notated expression of either articulation or dynamics (*Fig \ref{fig:stimuli} B-C*).
 As soon as they had produced the excerpt with the notated expression without pitch errors twice consecutively, the test trials began.
@@
 Tone intensity is assessed by key velocity (KV) and measures the loudness of a musical note.
+A higher value indicates forte styles whereas a lower value indicates piano styles.
 The value of KV in MIDI varies between 0 (minimum) and 127 (maximum).
@@
 For data analysis, only the 16th notes were included.", echo = FALSE}
-stim_a <- image_border(image_read(here("paper/sheetmusic/1", "stim_a.png")), "#FFFFFF", "120x120")
-stim_d <- image_border(image_read(here("paper/sheetmusic/1", "stim_d.png")), "#FFFFFF", "120x120")
-stim_n <- image_border(image_read(here("paper/sheetmusic/1", "stim_n.png")), "#FFFFFF", "120x120")
+stim_a <- image_border(image_read(here("paper/sheetmusic/1", "stim_a.png")), "#FFFFFF", "140x140")
+stim_d <- image_border(image_read(here("paper/sheetmusic/1", "stim_d.png")), "#FFFFFF", "140x140")
+stim_n <- image_border(image_read(here("paper/sheetmusic/1", "stim_n.png")), "#FFFFFF", "140x140")
 
@@
 img_stimuli <- image_append(c(stim_n, stim_a, stim_d), stack = TRUE)
-img_stimuli <- image_annotate(img_stimuli, "(A)", size = 80, location = "+100+0", color = "black", font = "helvetica")
-img_stimuli <- image_annotate(img_stimuli, "(B)", size = 80, location = "+100+650", color = "black", font = "helvetica")
-img_stimuli <- image_annotate(img_stimuli, "(C)", size = 80, location = "+100+1300", color = "black", font = "helvetica")
+img_stimuli <- image_annotate(img_stimuli, "(A)", size = 80, location = "+120+70", color = "black", font = "helvetica")
+img_stimuli <- image_annotate(img_stimuli, "(B)", size = 80, location = "+120+810", color = "black", font = "helvetica")
+img_stimuli <- image_annotate(img_stimuli, "(C)", size = 80, location = "+120+1570", color = "black", font = "helvetica")
 img_stimuli
@@
 # save
+img_stimuli <- image_scale(img_stimuli, "2200")
 image_write(img_stimuli, path = here("paper/figure", "Fig1.tiff"), format = "tiff")
@@
 All effects are reported as significant at *p* < .05.
-We first report results for performance of the piece with the notated expression of articulation (*Fig \ref{fig:stimuli}, B*), followed by performance of the piece with the notated expression of dynamics (*Fig \ref{fig:stimuli}, C*).
+We first report results for performance of the piece with the notated expression of articulation (*Fig \ref{fig:stimuli} B*), followed by performance of the piece with the notated expression of dynamics (*Fig \ref{fig:stimuli} C*).
 To recall our predictions, if participants play more slowly when they are trying to teach, interonset intervals (IOIs) should be larger when teaching.
-If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should reveal be more positive for legato and more negative for staccato when teaching articulation, and key velocity (KV) should reveal amplifications of forte and piano when teaching dynamics.
+If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should be more positive for legato and more negative for staccato when teaching articulation, and key velocity (KV) should be higher for forte and lower for piano when teaching dynamics.
 
@@
 # save plots
-ggsave(here("paper/image", "ioi_art_1.png"), plot = ioi_art_1, dpi = 300, width = 4, height = 4)
-img_ioi_art_1 <- image_read(here("paper/image", "ioi_art_1.png"))
+ggsave(here("paper/image", "ioi_art_1.png"), plot = ioi_art_1, dpi = 300, width = 5, height = 5)
+img_ioi_art_1 <- image_read(here("paper/image", "ioi_art_1.png"), density = 300)
 ```
@@
 staccato) revealed that there was a significant main effect of Articulation (*F*(`r kot_art_aov$anova_table[2,1]`,`r kot_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kot_art_aov$anova_table[2,4])`, *p* `r if(kot_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kot_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Articulation (*F*(`r kot_art_aov$anova_table[3,1]`,`r kot_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kot_art_aov$anova_table[3,4])`, *p* `r if(kot_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kot_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', kot_art_aov$anova_table[3,5])`, *Fig \ref{fig:kot-1}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato key-overlap time was more negative in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`).
-For legato there was no significant difference in key-overlap time between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KOT was more negative in the teaching condition [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kot_art_posthoc$p.value[2])`).
+For legato there was no significant difference in KOT between the teaching [*M* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean` (ms), *SD* = `r kot_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kot_art_posthoc$p.value[1])`).
 
@@
 ggsave(here("paper/image", "kot_art_1.png"), plot = kot_art_1, dpi = 300, width = 7, height = 5)
-img_kot_art_1 <- image_read(here("paper/image", "kot_art_1.png"))
+img_kot_art_1 <- image_read(here("paper/image", "kot_art_1.png"), density = 300)
 ```
@@
 ggsave(here("paper/image", "vel_art_1.png"), plot = vel_art_1, dpi = 300, width = 7, height = 5)
-img_vel_art_1 <- image_read(here("paper/image", "vel_art_1.png"))
+img_vel_art_1 <- image_read(here("paper/image", "vel_art_1.png"), density = 300)
 ```
@@
 # save plots
-ggsave(here("paper/image", "ioi_dyn_1.png"), plot = ioi_dyn_1, dpi = 300, width = 4, height = 4)
-img_ioi_dyn_1 <- image_read(here("paper/image", "ioi_dyn_1.png"))
+ggsave(here("paper/image", "ioi_dyn_1.png"), plot = ioi_dyn_1, dpi = 300, width = 5, height = 5)
+img_ioi_dyn_1 <- image_read(here("paper/image", "ioi_dyn_1.png"), density = 300)
 ```
@@
 piano) showed that there was a significant main effect of Dynamics (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Dynamics (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-1}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced higher key velocity in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) when instructed to play forte.
-There was no significant difference in key velocity between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`) when instructed to play piano.
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) when instructed to play forte.
+There was no significant difference in KV between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`) when instructed to play piano.
 
@@
 ggsave(here("paper/image", "vel_dyn_1.png"), plot = vel_dyn_1, dpi = 300, width = 7, height = 5)
-img_vel_dyn_1 <- image_read(here("paper/image", "vel_dyn_1.png"))
+img_vel_dyn_1 <- image_read(here("paper/image", "vel_dyn_1.png"), density = 300)
 
@@
 PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`; PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (*p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (*p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`; PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
 
@@
 ggsave(here("paper/image", "vel_diff_dyn_1.png"), plot = vel_diff_dyn_1, dpi = 300, width = 7, height = 5)
-img_vel_diff_dyn_1 <- image_read(here("paper/image", "vel_diff_dyn_1.png"))
+img_vel_diff_dyn_1 <- image_read(here("paper/image", "vel_diff_dyn_1.png"), density = 300)
 ```
@@
 ggsave(here("paper/image", "kot_dyn_1.png"), plot = kot_dyn_1, dpi = 300, width = 7, height = 5)
-img_kot_dyn_1 <- image_read(here("paper/image", "kot_dyn_1.png"))
+img_kot_dyn_1 <- image_read(here("paper/image", "kot_dyn_1.png"), density = 300)
 ```
@@
 The results for KOT and KV showed that participants successfully highlighted relevant aspects of articulation and dynamics.
-Specifically, key-overlap time was more negative when teaching articulation and key velocity for forte was higher when teaching dynamics.
+Specifically, participants exaggerated staccato when teaching articulation and forte when teaching dynamics.
 Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points (i.e., forte to piano or piano to forte).
@@
 img_ioi_1 <- image_append(c(img_ioi_art_1, img_ioi_dyn_1))
+img_ioi_1 <- image_scale(img_ioi_1, "2200")
 image_write(img_ioi_1, path = here("paper/figure", "Fig2.tiff"), format = "tiff")
@@
 img_kot_1 <- image_append(c(img_kot_art_1, img_kot_dyn_1))
+img_kot_1 <- image_scale(img_kot_1, "2200")
 image_write(img_kot_1, path = here("paper/figure", "Fig3.tiff"), format = "tiff")
@@
 img_vel_1 <- image_append(c(img_vel_art_1, img_vel_dyn_1))
+img_vel_1 <- image_scale(img_vel_1, "2200")
 image_write(img_vel_1, path = here("paper/figure", "Fig4.tiff"), format = "tiff")
@@
 If pianists selectively highlight the relevant aspects of the techniques to be taught also in a more naturalistic piece containing more opportunities for expression, key-overlap time should be again more positive for legato and more negative for staccato while teaching articulation.
-When teaching dynamics, participants should produce higher key velocity for forte and lower key velocity for piano.
-Given the findings we observed in Experiment 1, we also predicted that participants would make a larger contrast between forte and piano at transition points, and that they might play more slowly when teaching, especially when teaching articulation.
+When teaching dynamics, key velocity should be higher for forte and lower for piano.
+Given the findings we observed in Experiment 1, we also predicted that participants would make a larger key velocity contrast between forte and piano at transition points, and that they might play more slowly when teaching, especially when teaching articulation.
 
@@
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Non-expressive sheet music was used for the purpose of practice (*Fig \ref{fig:stimuli-2}, A*).
-Expressive notations were added to the non-expressive sheet music for the experiment (*Fig \ref{fig:stimuli-2}, B-C*).
+Non-expressive sheet music was used for the purpose of practice (*Fig \ref{fig:stimuli-2} A*).
+Expressive notations were added to the non-expressive sheet music for the experiment (*Fig \ref{fig:stimuli-2} B-C*).
 These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
@@
 We employed the same procedure as in Experiment 1 with several modifications.
-First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2}, A*) prior to the experiment.
+First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2} A*) prior to the experiment.
 Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[S1 File](#supplemental)*).
@@
 The chosen tempo was again cued by a leading metronome.
-In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music.
-Those who could not perform the piece twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded).
+In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music (*Fig \ref{fig:stimuli-2} A*).
+Those who could not perform the piece without pitch errors twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded).
 The rest of the procedure was identical to Experiment 1.
@@
 img_stimuli_2 <- image_append(c(stim_n_2, stim_a_2, stim_d_2), stack = TRUE)
-img_stimuli_2 <- image_annotate(img_stimuli_2, "(A)", size = 80, location = "+100+70", color = "black", font = "helvetica")
-img_stimuli_2 <- image_annotate(img_stimuli_2, "(B)", size = 80, location = "+100+1125", color = "black", font = "helvetica")
-img_stimuli_2 <- image_annotate(img_stimuli_2, "(C)", size = 80, location = "+100+2150", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(A)", size = 80, location = "+100+100", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(B)", size = 80, location = "+100+1110", color = "black", font = "helvetica")
+img_stimuli_2 <- image_annotate(img_stimuli_2, "(C)", size = 80, location = "+100+2110", color = "black", font = "helvetica")
 img_stimuli_2
@@
 # save
+img_stimuli_2 <- image_scale(img_stimuli_2, "x2600")
 image_write(img_stimuli_2, path = here("paper/figure", "Fig5.tiff"), format = "tiff")
@@
 
-As Experiment 1, we first report results for performance of the piece with the notated expression of articulation (*Fig \ref{fig:stimuli-2}, B*), followed by performance of the piece with the notated expression of dynamics (*Fig \ref{fig:stimuli-2}, C*).
+As Experiment 1, we first report results for performance of the piece with the notated expression of articulation (*Fig \ref{fig:stimuli-2} B*), followed by performance of the piece with the notated expression of dynamics (*Fig \ref{fig:stimuli-2} C*).
 
@@
 # save plots
-ggsave(here("paper/image", "ioi_art_2.png"), plot = ioi_art_2, dpi = 300, width = 4, height = 4)
-img_ioi_art_2 <- image_read(here("paper/image", "ioi_art_2.png"))
+ggsave(here("paper/image", "ioi_art_2.png"), plot = ioi_art_2, dpi = 300, width = 5, height = 5)
+img_ioi_art_2 <- image_read(here("paper/image", "ioi_art_2.png"), density = 300)
 ```
@@
 staccato) showed that there was a significant main effect of Articulation (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Articulation (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato key-overlap time was more negative in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`).
-For letago here was no significant difference in key-overlap time between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KOR was more negative in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`).
+For legato there was no significant difference in KOR between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
 
@@
 ggsave(here("paper/image", "kor_art_2.png"), plot = kor_art_2, dpi = 300, width = 7, height = 5)
-img_kor_art_2 <- image_read(here("paper/image", "kor_art_2.png"))
+img_kor_art_2 <- image_read(here("paper/image", "kor_art_2.png"), density = 300)
 ```
@@
 ggsave(here("paper/image", "vel_art_2.png"), plot = vel_art_2, dpi = 300, width = 7, height = 5)
-img_vel_art_2 <- image_read(here("paper/image", "vel_art_2.png"))
+img_vel_art_2 <- image_read(here("paper/image", "vel_art_2.png"), density = 300)
 ```
@@
 # save plots
-ggsave(here("paper/image", "ioi_dyn_2.png"), plot = ioi_dyn_2, dpi = 300, width = 4, height = 4)
-img_ioi_dyn_2 <- image_read(here("paper/image", "ioi_dyn_2.png"))
+ggsave(here("paper/image", "ioi_dyn_2.png"), plot = ioi_dyn_2, dpi = 300, width = 5, height = 5)
+img_ioi_dyn_2 <- image_read(here("paper/image", "ioi_dyn_2.png"), density = 300)
 ```
@@
 piano) showed that there was a significant main effect of Dynamics (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Dynamics (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced higher key velocity in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`](*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) when instructed to play forte.
-Also, participants produced lower key velocity in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`](*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) when instructed to play piano.
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) when instructed to play forte.
+Also, KV was lower in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[2]))}`) when instructed to play piano.
 
@@
 ggsave(here("paper/image", "vel_dyn_2.png"), plot = vel_dyn_2, dpi = 300, width = 7, height = 5)
-img_vel_dyn_2 <- image_read(here("paper/image", "vel_dyn_2.png"))
+img_vel_dyn_2 <- image_read(here("paper/image", "vel_dyn_2.png"), density = 300)
 ```
@@
 PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[2,1]`,`r vel_diff_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov$anova_table[3,1]`,`r vel_diff_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-diff}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (FtoP: *p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (PtoF: *p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (*p* `r if(vel_diff_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[1]))}`) and from piano to forte (*p* `r if(vel_diff_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc$p.value[2]))}`) in the teaching condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] than in the performing condition [FtoP: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
 
@@
 ggsave(here("paper/image", "vel_diff_dyn_2.png"), plot = vel_diff_dyn_2, dpi = 300, width = 7, height = 5)
-img_vel_diff_dyn_2 <- image_read(here("paper/image", "vel_diff_dyn_2.png"))
+img_vel_diff_dyn_2 <- image_read(here("paper/image", "vel_diff_dyn_2.png"), density = 300)
 ```
@@
 ggsave(here("paper/image", "kor_dyn_2.png"), plot = kor_dyn_2, dpi = 300, width = 7, height = 5)
-img_kor_dyn_2 <- image_read(here("paper/image", "kor_dyn_2.png"))
+img_kor_dyn_2 <- image_read(here("paper/image", "kor_dyn_2.png"), density = 300)
 ```
@@
 As in Experiment 1, we observed slower performance only while teaching the notated articulation.
-Again, KOR showed that when teaching articulation, key-overlap time was more negative for staccato, while there was no significant difference in legato between the two conditions.
-While teaching dynamics, KV showed that participants produced higher key velocity for forte and lower key velocity for piano.
-This pattern of exaggeration was more pronounced than in Experiment 1, where only higher key velocity for forte was observed.
-Furthermore, we again found that when teaching dynamics, participants made a larger contrast between forte and piano bidirectionally at transition points.
+Again, KOR showed that when teaching articulation, participants exaggerated staccato while there was no significant difference in legato between the two conditions.
+While teaching dynamics, KV showed that participants exaggerated forte and piano.
+This pattern of exaggeration was more pronounced than in Experiment 1, where only higher KV for forte was observed.
+Furthermore, we again found that when teaching dynamics, participants made a larger dynamics contrast between forte and piano bidirectionally at transition points.
 Importantly, participants did not modulate performance aspects that were irrelevant with regard to the techniques to be taught (i.e., no modulation of dynamics when teaching articulation and vice versa).
@@
 img_ioi_2 <- image_append(c(img_ioi_art_2, img_ioi_dyn_2))
+img_ioi_2 <- image_scale(img_ioi_2, "2200")
 image_write(img_ioi_2, path = here("paper/figure", "Fig6.tiff"), format = "tiff")
@@
 img_kor_2 <- image_append(c(img_kor_art_2, img_kor_dyn_2))
+img_kor_2 <- image_scale(img_kor_2, "2200")
 image_write(img_kor_2, path = here("paper/figure", "Fig7.tiff"), format = "tiff")
@@
 img_vel_2 <- image_append(c(img_vel_art_2, img_vel_dyn_2))
+img_vel_2 <- image_scale(img_vel_2, "2200")
 image_write(img_vel_2, path = here("paper/figure", "Fig8.tiff"), format = "tiff")
@@
 img_vel_diff <- image_append(c(img_vel_diff_dyn_1, img_vel_diff_dyn_2))
+img_vel_diff <- image_scale(img_vel_diff, "2200")
 image_write(img_vel_diff, path = here("paper/figure", "Fig9.tiff"), format = "tiff")
@@
 When playing with the intention to teach, pianists selectively highlighted relevant aspects of artistic expressions.
-Across two different pieces, key-overlap time was more negative for staccato when playing with the intention to teach expressions concerning articulation.
+Across two different pieces, participants exaggerated staccato when playing with the intention to teach expressions concerning articulation.
 The lack of a difference for legato when teaching articulation might stem from a ceiling effect.
-When playing with the intention to teach expressions concerning dynamics, participants made larger key velocity changes between forte and piano (in both directions).
+When playing with the intention to teach expressions concerning dynamics, participants made larger dynamics changes between forte and piano (in both directions).
 This constitutes an effective way to highlight the technique as loudness is determined relatively.

### 2021-07-16 paper/manuscript.tex
latest version / fix figures and some wordings
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   man,floatsintext]{apa6}
+\title{The sound of teaching music: Expert pianists' performance modulations for novices}
+\author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
+\date{}
+
 \usepackage{amsmath,amssymb}
@@
 \hypersetup{
-  pdftitle={The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices},
+  pdftitle={The sound of teaching music: Expert pianists' performance modulations for novices},
   pdfauthor={Atsuko Tominaga1, Günther Knoblich1, \& Natalie Sebanz1},
@@
 \fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 % Manuscript styling
@@
 \makeatother
-\shorttitle{The Sound of Teaching Music}
-\keywords{teaching, expertise, skill transmission, artistic expression, music\newline\indent Word count: around 5500 (except references)}
+\shorttitle{The sound of teaching music}
+\keywords{teaching, expertise, skill transmission, artistic expression, music}
 \usepackage{csquotes}
@@
 \fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-\title{The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices}
-\author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
-\date{}
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Cognitive Science, Central European University}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Cognitive Science, Central European University, Vienna, Austria}
 
@@
 \hypertarget{introduction}{%
-\section{1. Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}}
 
-Deliberate teaching has supported human skill transmission over generations and provides a key route for learning (Thornton \& Raihani, 2008; Tomasello, 2016).
+Deliberate teaching has supported human skill transmission over generations and provides a key route for learning {[}1,2{]}.
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
-In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners (Brand, Baldwin, \& Ashburn, 2002; Saint-Georges et al., 2013).
+In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners {[}3,4{]}.
 These modulations include slowing down and exaggerating sounds and actions.
-Similar findings were obtained in studies with adult learners, where exaggerations were observed when native British English speakers were talking to first language English learners (Uther, Knoll, \& Burnham, 2007) and when skilled adults were teaching xylophone melodies to novices (McEllin, Knoblich, \& Sebanz, 2017).
-The observed modulations are thought to provide communicative signals that can facilitate learning by affecting attention and memory (Csibra \& Gergely, 2009).
+Similar findings were obtained in studies with adult learners, where exaggerations were observed when native British English speakers were talking to first language English learners {[}5{]} and when skilled adults were teaching xylophone melodies to novices {[}6{]}.
+The observed modulations are thought to provide communicative signals that can facilitate learning by affecting attention and memory {[}7{]}.
 
-Previous research has shown that demonstrators not only adjust their gestures (Campisi \& Özyürek, 2013) and actions (Fukuyama et al., 2015) to learners' skills but engage in specific action modulations to highlight particular aspects of demonstrated actions.
-For example, Schaik, Meyer, Ham, and Hunnius (2019) showed that adults used specific action modulations for demonstrating different action effects of objects to infants.
-Ho, Littman, MacGlashan, Cushman, and Austerweil (2016) found that demonstrators chose costly movement paths at structurally important points to disambiguate one goal over other possibilities.
+Previous research has shown that demonstrators not only adjust their gestures {[}8{]} and actions {[}9{]} to learners' skills but engage in specific action modulations to highlight particular aspects of demonstrated actions.
+For example, Schaik and colleagues {[}10{]} showed that adults used specific action modulations for demonstrating different action effects of objects to infants.
+Ho and colleagues {[}11{]} found that demonstrators chose costly movement paths at structurally important points to disambiguate one goal over other possibilities.
 
 Importantly, in some domains teaching requires not only demonstrating \emph{what} to do, but \emph{how} to perform actions.
-In artistic contexts, novices need to learn how exactly to perform actions by relying on particular techniques (Sloboda, 2000).
-For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques (Juslin, 2003).
+In artistic contexts, novices need to learn how exactly to perform actions by relying on particular techniques {[}12{]}.
+For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques {[}13{]}.
 This raises the question of whether and how experts modulate their actions during didactic demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
@@
 \hypertarget{experiment-1}{%
-\section{2. Experiment 1}\label{experiment-1}}
+\section{Experiment 1}\label{experiment-1}}
 
-\hypertarget{method}{%
-\subsection{2.1. Method}\label{method}}
+\hypertarget{materials-and-methods}{%
+\subsection{Materials and methods}\label{materials-and-methods}}
 
 \hypertarget{participants}{%
-\subsubsection{2.1.1. Participants}\label{participants}}
+\subsubsection{Participants}\label{participants}}
 
@@
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
-The study was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.
+The study (No.~2018-124) was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.
 
 \hypertarget{apparatus-and-stimuli}{%
-\subsubsection{2.1.2. Apparatus and stimuli}\label{apparatus-and-stimuli}}
+\subsubsection{Apparatus and stimuli}\label{apparatus-and-stimuli}}
 
@@
 One musical excerpt was used as a stimulus.
-The excerpt was taken from ``A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Burnam and modified for the experiment.
+The excerpt was taken from ``A Dozen a Day - Play with Ease in Many Keys'' by Edna-Mae Burnam and modified for the experiment.
 It consisted of a 6-measure isochronous melody noted in 4/4 meter.
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations, \emph{Figure \ref{fig:stimuli}, A}) was used for the purpose of practice.
+Non-expressive sheet music (i.e., sheet music without expressive notations, \emph{Fig \ref{fig:stimuli} A}) was used for the purpose of practice.
 Expressive notations were added to the non-expressive sheet music for the experiment.
-They referred to either articulation or dynamics (\emph{Figure \ref{fig:stimuli}, B-C}).
+They referred to either articulation or dynamics (\emph{Fig \ref{fig:stimuli} B-C}).
 Articulation was notated as either legato or staccato.
@@
 \hypertarget{procedure}{%
-\subsubsection{2.1.3. Procedure}\label{procedure}}
+\subsubsection{Procedure}\label{procedure}}
 
-First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with the notated expression of either articulation or dynamics (\emph{Figure \ref{fig:stimuli}, B-C}).
+First, participants were allocated to either the teaching or performing condition and asked to practise the excerpt with the notated expression of either articulation or dynamics (\emph{Fig \ref{fig:stimuli} B-C}).
 As soon as they had produced the excerpt with the notated expression without pitch errors twice consecutively, the test trials began.
@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary Materials}}).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{S1 File}}).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
@@
 \hypertarget{data-analysis}{%
-\subsubsection{2.1.4. Data analysis}\label{data-analysis}}
+\subsubsection{Data analysis}\label{data-analysis}}
 
@@
 Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
-Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone, and measures the smoothness of musical sequences (Bresin \& Battel, 2000).
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone, and measures the smoothness of musical sequences {[}14{]}.
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
 Tone intensity is assessed by key velocity (KV) and measures the loudness of a musical note.
+A higher value indicates forte styles whereas a lower value indicates piano styles.
 The value of KV in MIDI varies between 0 (minimum) and 127 (maximum).
@@
 A paired-sample \emph{t}-test or a Wilcoxon Signed-rank test (a non-parametric alternative to a paired \emph{t}-test) was performed to compare the mean IOIs in the teaching and performing condition.
-For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.~performing) and Subcomponent (LS: legato vs.~staccato or FP: forte vs.~piano, respectively).
+For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.~performing) and Subcomponent (Articulation: legato vs.~staccato or Dynamics: forte vs.~piano, respectively).
 The \emph{t.test} or \emph{wilcox.test} function in the \emph{stats} R package was used for a \emph{t}-test or a Wilcoxon Signed-rank test.
@@
 \hypertarget{results}{%
-\subsection{2.2. Results}\label{results}}
+\subsection{Results}\label{results}}
 
 All effects are reported as significant at \emph{p} \textless{} .05.
-We first report results for performance of the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli}, B}), followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli}, C}).
+We first report results for performance of the piece with the notated expression of articulation (\emph{Fig \ref{fig:stimuli} B}), followed by performance of the piece with the notated expression of dynamics (\emph{Fig \ref{fig:stimuli} C}).
 To recall our predictions, if participants play more slowly when they are trying to teach, interonset intervals (IOIs) should be larger when teaching.
-If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should reveal exaggerated legato/staccato when teaching articulation, and key velocity (KV) should reveal amplifications of forte and piano when teaching dynamics.
+If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should be more positive for legato and more negative for staccato when teaching articulation, and key velocity (KV) should be higher for forte and lower for piano when teaching dynamics.
 
-\hypertarget{articulation-interonset-intervals-iois}{%
-\subsubsection{2.2.1. Articulation: Interonset Intervals (IOIs)}\label{articulation-interonset-intervals-iois}}
+\hypertarget{articulation}{%
+\subsection{Articulation}\label{articulation}}
+
+\hypertarget{interonset-intervals-iois}{%
+\subsubsection{Interonset Intervals (IOIs)}\label{interonset-intervals-iois}}
 
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41, \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
 
 \hypertarget{articulation-key-overlap-time-kot}{%
-\subsubsection{2.2.2. Articulation: Key-Overlap Time (KOT)}\label{articulation-key-overlap-time-kot}}
+\subsubsection{Articulation: Key-Overlap Time (KOT)}\label{articulation-key-overlap-time-kot}}
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) revealed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011, \emph{Figure \ref{fig:kot-1}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.004) while there was no significant difference in legato between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition {[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.32).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) revealed that there was a significant main effect of Articulation (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Articulation (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011, \emph{Fig \ref{fig:kot-1}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KOT was more negative in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.004).
+For legato there was no significant difference in KOT between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition {[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.32).
 
-\hypertarget{articulation-key-velocity-kv}{%
-\subsubsection{2.2.3. Articulation: Key Velocity (KV)}\label{articulation-key-velocity-kv}}
+\hypertarget{key-velocity-kv}{%
+\subsubsection{Key Velocity (KV)}\label{key-velocity-kv}}
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004, \emph{Figure \ref{fig:vel-1}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant interaction between Condition and Articulation (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004, \emph{Fig \ref{fig:vel-1}}).
 However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Legato: \emph{M} = 70.34, \emph{SD} = 6.45; Staccato: \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato: \emph{M} = 70.68, \emph{SD} = 5.84, Staccato: \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato: \emph{p} = 0.70, Staccato: \emph{p} = 0.07).
 
-\hypertarget{dynamics-interonset-intervals-iois}{%
-\subsubsection{2.2.4. Dynamics: Interonset Intervals (IOIs)}\label{dynamics-interonset-intervals-iois}}
+\hypertarget{dynamics}{%
+\subsection{Dynamics}\label{dynamics}}
+
+\hypertarget{interonset-intervals-iois-1}{%
+\subsubsection{Interonset Intervals (IOIs)}\label{interonset-intervals-iois-1}}
 
 In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29, \emph{Figure \ref{fig:ioi-1}}).
+The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
 
-\hypertarget{dynamics-key-velocity-kv}{%
-\subsubsection{2.2.5. Dynamics: Key Velocity (KV)}\label{dynamics-key-velocity-kv}}
+\hypertarget{key-velocity-kv-1}{%
+\subsubsection{Key Velocity (KV)}\label{key-velocity-kv-1}}
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003, \emph{Figure \ref{fig:vel-1}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.021) while there was no significant difference in piano between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.45).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Dynamics (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003, \emph{Fig \ref{fig:vel-1}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.021) when instructed to play forte.
+There was no significant difference in KV between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.45) when instructed to play piano.
 
-\hypertarget{dynamics-kv-transition-points}{%
-\subsubsection{2.2.6. Dynamics: KV Transition Points}\label{dynamics-kv-transition-points}}
+\hypertarget{kv-transition-points}{%
+\subsubsection{KV Transition Points}\label{kv-transition-points}}
 
@@
 Outliers were removed using the same criteria as for the other dependent variables and less than 5 \% of the data were excluded from analysis.
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032, \emph{Figure \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: \emph{p} = 0.002) and from piano to forte (PtoF: \emph{p} \textless{} 0.001) in the teaching condition {[}FtoP: \emph{M} = -14.06, \emph{SD} = 7.20; PtoF: \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP: \emph{M} = -11.69, \emph{SD} = 5.97, PtoF: \emph{M} = 20.88, \emph{SD} = 8.67{]}.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction between Condition and Transition Type (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032, \emph{Fig \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (\emph{p} = 0.002) and from piano to forte (\emph{p} \textless{} 0.001) in the teaching condition {[}FtoP: \emph{M} = -14.06, \emph{SD} = 7.20; PtoF: \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP: \emph{M} = -11.69, \emph{SD} = 5.97, PtoF: \emph{M} = 20.88, \emph{SD} = 8.67{]}.
 
-\hypertarget{dynamics-key-overlap-time-kot}{%
-\subsubsection{2.2.7. Dynamics: Key-Overlap Time (KOT)}\label{dynamics-key-overlap-time-kot}}
+\hypertarget{key-overlap-time-kot}{%
+\subsubsection{Key-Overlap Time (KOT)}\label{key-overlap-time-kot}}
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) revealed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31, \emph{Figure \ref{fig:kot-1}}), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.52, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 0.000, \emph{p} = 0.99, \(\eta_G^2\) = 0.000) was significant.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) revealed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31, \emph{Fig \ref{fig:kot-1}}), reflecting more key-overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.52, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Dynamics (\emph{F}(1,30) = 0.000, \emph{p} = 0.99, \(\eta_G^2\) = 0.000) was significant.
 
 \hypertarget{discussion}{%
-\subsection{2.3. Discussion}\label{discussion}}
+\subsection{Discussion}\label{discussion}}
 
@@
 For IOIs, we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
-This finding is in line with earlier studies that found slower performance of actions in a teaching context (e.g., Brand et al., 2002; McEllin et al., 2017).
+This finding is in line with earlier studies that found slower performance of actions in a teaching context {[}e.g., 3,6{]}.
 However, we did not find a significant difference in tempo when participants were teaching dynamics.
@@
 The results for KOT and KV showed that participants successfully highlighted relevant aspects of articulation and dynamics.
-Specifically, they produced shorter staccato when teaching articulation and produced louder forte when teaching dynamics.
+Specifically, participants exaggerated staccato when teaching articulation and forte when teaching dynamics.
 Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points (i.e., forte to piano or piano to forte).
@@
 \hypertarget{experiment-2}{%
-\section{3. Experiment 2}\label{experiment-2}}
+\section{Experiment 2}\label{experiment-2}}
 
-\hypertarget{method-1}{%
-\subsection{3.1. Method}\label{method-1}}
+In Experiment 1, we employed a simple musical scale in order to maximise experimental control.
+To test whether our results generalise to a more naturalistic piece, in Experiment 2, we chose an actual piano piece and modified it for the purpose of the experiment.
+If pianists selectively highlight the relevant aspects of the techniques to be taught also in a more naturalistic piece containing more opportunities for expression, key-overlap time should be again more positive for legato and more negative for staccato while teaching articulation.
+When teaching dynamics, key velocity should be higher for forte and lower for piano.
+Given the findings we observed in Experiment 1, we also predicted that participants would make a larger key velocity contrast between forte and piano at transition points, and that they might play more slowly when teaching, especially when teaching articulation.
+
+\hypertarget{materials-and-methods-1}{%
+\subsection{Materials and methods}\label{materials-and-methods-1}}
 
 \hypertarget{participants-1}{%
-\subsubsection{3.1.1. Participants}\label{participants-1}}
+\subsubsection{Participants}\label{participants-1}}
 
@@
 They had 15.65 years of practice on average (\emph{SD} = 5.67).
+All participants gave their informed consent before the experiment started and received vouchers for their participation.
+The study (No.~2018-124) was approved by the United Ethical Review Committee for Research in Psychology (EPKEB) in Hungary.
 
 \hypertarget{apparatus-and-stimuli-1}{%
-\subsubsection{3.1.2. Apparatus and stimuli}\label{apparatus-and-stimuli-1}}
+\subsubsection{Apparatus and stimuli}\label{apparatus-and-stimuli-1}}
 
@@
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Non-expressive sheet music was used for the purpose of practice (\emph{Figure \ref{fig:stimuli-2}, A}).
-Expressive notations were added to the non-expressive sheet music for the experiment (\emph{Figure \ref{fig:stimuli-2}, B-C}).
+Non-expressive sheet music was used for the purpose of practice (\emph{Fig \ref{fig:stimuli-2} A}).
+Expressive notations were added to the non-expressive sheet music for the experiment (\emph{Fig \ref{fig:stimuli-2} B-C}).
 These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
@@
 \hypertarget{procedure-2}{%
-\subsubsection{3.1.3. Procedure}\label{procedure-2}}
+\subsubsection{Procedure}\label{procedure-2}}
 
 We employed the same procedure as in Experiment 1 with several modifications.
-First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Figure \ref{fig:stimuli-2}, A}) prior to the experiment.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplemental}{Supplementary Materials}}).
-Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter-beats per minute).
+First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Fig \ref{fig:stimuli-2} A}) prior to the experiment.
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplemental}{S1 File}}).
+Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute).
 The chosen tempo was again cued by a leading metronome.
-In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music.
-Those who could not perform the piece twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded).
+In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music (\emph{Fig \ref{fig:stimuli-2} A}).
+Those who could not perform the piece without pitch errors twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded).
 The rest of the procedure was identical to Experiment 1.
@@
 \hypertarget{data-analysis-1}{%
-\subsubsection{3.1.4. Data analysis}\label{data-analysis-1}}
+\subsubsection{Data analysis}\label{data-analysis-1}}
 
@@
 \hypertarget{results-1}{%
-\subsection{3.2. Results}\label{results-1}}
+\subsection{Results}\label{results-1}}
 
-As Experiment 1, we first report results for performance of the piece with the notated expression of articulation (\emph{Figure \ref{fig:stimuli-2}, B}), followed by performance of the piece with the notated expression of dynamics (\emph{Figure \ref{fig:stimuli-2}, C}).
+As Experiment 1, we first report results for performance of the piece with the notated expression of articulation (\emph{Fig \ref{fig:stimuli-2} B}), followed by performance of the piece with the notated expression of dynamics (\emph{Fig \ref{fig:stimuli-2} C}).
 
-\hypertarget{articulation-interonset-intervals-iois-1}{%
-\subsubsection{3.2.1. Articulation: Interonset Intervals (IOIs)}\label{articulation-interonset-intervals-iois-1}}
+\hypertarget{articulation-1}{%
+\subsection{Articulation}\label{articulation-1}}
+
+\hypertarget{interonset-intervals-iois-2}{%
+\subsubsection{Interonset Intervals (IOIs)}\label{interonset-intervals-iois-2}}
 
 A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.97, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.95, \emph{SD} = 0.040{]} while playing the piece with the notated articulation (\emph{t}(19) = 2.47
-, \emph{p} = 0.023, Cohen's \emph{d} = 0.55, \emph{Figure \ref{fig:ioi-2}}).
-A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (\emph{p} = 0.033, \emph{r} = 0.48).
+, \emph{p} = 0.023, Cohen's \emph{d} = 0.55, two-tailed, \emph{Fig \ref{fig:ioi-2}}).
+A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (\emph{p} = 0.033, \emph{r} = 0.48, two-tailed).
 
-\hypertarget{articulation-key-overlap-ratio-kor}{%
-\subsubsection{3.2.2. Articulation: Key-Overlap Ratio (KOR)}\label{articulation-key-overlap-ratio-kor}}
+\hypertarget{key-overlap-ratio-kor}{%
+\subsubsection{Key-Overlap Ratio (KOR)}\label{key-overlap-ratio-kor}}
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01, \emph{Figure \ref{fig:kor-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that participants produced shorter staccato in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.002) while there was no significant difference in legato between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.56).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant main effect of Articulation (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Articulation (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01, \emph{Fig \ref{fig:kor-2}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KOR was more negative in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.002).
+For legato there was no significant difference in KOR between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.56).
 
-\hypertarget{articulation-key-velocity-kv-1}{%
-\subsubsection{3.2.3. Articulation: Key Velocity (KV)}\label{articulation-key-velocity-kv-1}}
+\hypertarget{key-velocity-kv-2}{%
+\subsubsection{Key Velocity (KV)}\label{key-velocity-kv-2}}
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,19) = 7.91, \emph{p} = 0.011, \(\eta_G^2\) = 0.03, \emph{Figure \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
-Neither the main effect of Condition (\emph{F}(1,19) = 1.33, \emph{p} = 0.26, \(\eta_G^2\) = 0.003) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,19) = 0.14, \emph{p} = 0.71, \(\eta_G^2\) = 0.000) was significant.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant main effect of Articulation (\emph{F}(1,19) = 7.91, \emph{p} = 0.011, \(\eta_G^2\) = 0.03, \emph{Fig \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
+Neither the main effect of Condition (\emph{F}(1,19) = 1.33, \emph{p} = 0.26, \(\eta_G^2\) = 0.003) nor the interaction between Condition and Articulation (\emph{F}(1,19) = 0.14, \emph{p} = 0.71, \(\eta_G^2\) = 0.000) was significant.
 
-\hypertarget{dynamics-interonset-intervals-iois-1}{%
-\subsubsection{3.2.4. Dynamics: Interonset Intervals (IOIs)}\label{dynamics-interonset-intervals-iois-1}}
+\hypertarget{dynamics-1}{%
+\subsection{Dynamics}\label{dynamics-1}}
 
-A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.043{]} while playing the piece with the notated dynamics (\emph{t}(19) = 0.21, \emph{p} = 0.84, Cohen's \emph{d} = 0.05, \emph{Figure \ref{fig:ioi-2}}).
-A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (\emph{p} = 0.96).
+\hypertarget{interonset-intervals-iois-3}{%
+\subsubsection{Interonset Intervals (IOIs)}\label{interonset-intervals-iois-3}}
 
-\hypertarget{dynamics-key-velocity-kv-1}{%
-\subsubsection{3.2.5. Dynamics: Key Velocity (KV)}\label{dynamics-key-velocity-kv-1}}
+A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.043{]} while playing the piece with the notated dynamics (\emph{t}(19) = 0.21, \emph{p} = 0.84, Cohen's \emph{d} = 0.05, two-tailed, \emph{Fig \ref{fig:ioi-2}}).
+A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (\emph{p} = 0.96, two-tailed).
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 132, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 6.12, \emph{p} = 0.023, \(\eta_G^2\) = 0.018, \emph{Figure \ref{fig:vel-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte (\emph{p} = 0.040) and softer piano (\emph{p} = 0.024) in the teaching condition {[}Forte: \emph{M} = 83.20, \emph{SD} = 8.84, Piano: \emph{M} = 60.53, \emph{SD} = 3.76{]} than in the performing condition {[}Forte: \emph{M} = 81.49, \emph{SD} = 8.12, Piano: \emph{M} = 62.42, \emph{SD} = 4.82{]} than in the performing condition.
+\hypertarget{key-velocity-kv-3}{%
+\subsubsection{Key Velocity (KV)}\label{key-velocity-kv-3}}
 
-\hypertarget{dynamics-kv-transition-points-1}{%
-\subsubsection{3.2.6. Dynamics: KV Transition Points}\label{dynamics-kv-transition-points-1}}
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,19) = 132, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Dynamics (\emph{F}(1,19) = 6.12, \emph{p} = 0.023, \(\eta_G^2\) = 0.018, \emph{Fig \ref{fig:vel-2}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 83.20, \emph{SD} = 8.84{]} than in the performing condition {[}\emph{M} = 81.49, \emph{SD} = 8.12{]} (\emph{p} = 0.040) when instructed to play forte.
+Also, KV was lower in the teaching condition {[}\emph{M} = 60.53, \emph{SD} = 3.76{]} than in the performing condition {[}\emph{M} = 62.42, \emph{SD} = 4.82{]} (\emph{p} = 0.024) when instructed to play piano.
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059, \emph{Figure \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano (FtoP: \emph{p} = 0.029) and from piano to forte (PtoF: \emph{p} = 0.002) in the teaching condition {[}FtoP: \emph{M} = -17.23, \emph{SD} = 8.26, PtoF: \emph{M} = 20.71, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP: \emph{M} = -14.04, \emph{SD} = 7.67, PtoF: \emph{M} = 16.05, \emph{SD} = 7.32{]}.
+\hypertarget{kv-transition-points-1}{%
+\subsubsection{KV Transition Points}\label{kv-transition-points-1}}
 
-\hypertarget{dynamics-key-overlap-ratio-kor}{%
-\subsubsection{3.2.7. Dynamics: Key-Overlap Ratio (KOR)}\label{dynamics-key-overlap-ratio-kor}}
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction between Condition and Transition Type (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059, \emph{Fig \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (\emph{p} = 0.029) and from piano to forte (\emph{p} = 0.002) in the teaching condition {[}FtoP: \emph{M} = -17.23, \emph{SD} = 8.26, PtoF: \emph{M} = 20.71, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP: \emph{M} = -14.04, \emph{SD} = 7.67, PtoF: \emph{M} = 16.05, \emph{SD} = 7.32{]}.
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,19) = 630, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94), reflecting more key overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,19) = 0.081, \emph{p} = 0.78, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Subcomponent-FP (\emph{F}(1,19) = 0.35, \emph{p} = 0.56, \(\eta_G^2\) = 0.001, \emph{Figure \ref{fig:kor-2}}) was significant.
+\hypertarget{key-overlap-ratio-kor-1}{%
+\subsubsection{Key-Overlap Ratio (KOR)}\label{key-overlap-ratio-kor-1}}
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,19) = 630, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94), reflecting more key-overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (\emph{F}(1,19) = 0.081, \emph{p} = 0.78, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Dynamics (\emph{F}(1,19) = 0.35, \emph{p} = 0.56, \(\eta_G^2\) = 0.001, \emph{Fig \ref{fig:kor-2}}) was significant.
 
 \hypertarget{discussion-1}{%
-\subsection{3.3. Discussion}\label{discussion-1}}
+\subsection{Discussion}\label{discussion-1}}
 
@@
 As in Experiment 1, we observed slower performance only while teaching the notated articulation.
-Again, KOR showed that when teaching articulation, participants produced shorter staccato, while there was no significant difference in legato between the two conditions.
-While teaching dynamics, KV showed that participants produced louder forte and softer piano.
-This pattern of exaggeration was more pronounced than in Experiment 1, where only louder forte was observed.
-Furthermore, we again found that when teaching dynamics, participants made a larger contrast between forte and piano bidirectionally at transition points.
+Again, KOR showed that when teaching articulation, participants exaggerated staccato while there was no significant difference in legato between the two conditions.
+While teaching dynamics, KV showed that participants exaggerated forte and piano.
+This pattern of exaggeration was more pronounced than in Experiment 1, where only higher KV for forte was observed.
+Furthermore, we again found that when teaching dynamics, participants made a larger dynamics contrast between forte and piano bidirectionally at transition points.
 Importantly, participants did not modulate performance aspects that were irrelevant with regard to the techniques to be taught (i.e., no modulation of dynamics when teaching articulation and vice versa).
@@
 \hypertarget{general-discussion}{%
-\section{4. General Discussion}\label{general-discussion}}
+\section{General Discussion}\label{general-discussion}}
 
@@
 When playing with the intention to teach, pianists selectively highlighted relevant aspects of artistic expressions.
-Across two different pieces, they exaggerated staccato when playing with the intention to teach expressions concerning articulation.
+Across two different pieces, participants exaggerated staccato when playing with the intention to teach expressions concerning articulation.
 The lack of a difference for legato when teaching articulation might stem from a ceiling effect.
-When playing with the intention to teach expressions concerning dynamics, participants exaggerated changes between forte and piano (in both directions).
+When playing with the intention to teach expressions concerning dynamics, participants made larger dynamics changes between forte and piano (in both directions).
 This constitutes an effective way to highlight the technique as loudness is determined relatively.
@@
 
-Musicians and music teachers tend to consider expressive skills as a performer's most important skill (Lindström, Juslin, Bresin, \& Williamon, 2003).
+Musicians and music teachers tend to consider expressive skills as a performer's most important skill {[}15{]}.
 In the present study, experts focused on the teaching of expressions that were notated.
-However, expression also has other facets (Juslin, 2003) that are not only piece-related but performer- or context-related.
+However, expression also has other facets {[}13{]} that are not only piece-related but performer- or context-related.
 An interesting topic for future research is how experts teach not only expressive techniques as such but convey possibilities of interpretation.
@@
 One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students.
-Given that teachers modulate their demonstration throughout online interactions with learners (Okazaki, Muraoka, \& Osu, 2019), future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities.
+Given that teachers modulate their demonstration throughout online interactions with learners {[}16{]}, future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities.
 
@@
 Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention.
-Rather, expert demonstrators seem to follow principles of relevance in communication (Sperber \& Wilson, 1995), highlighting only those aspects they are intending to demonstrate.
+Rather, expert demonstrators seem to follow principles of relevance in communication {[}17{]}, highlighting only those aspects they are intending to demonstrate.
 How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher's intentions explicitly adds to the learning success are important questions for future research.
@@
 We thank Dávid Csűrös for his help with data collection.
-This research was supported by the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013)/ERC (European Research Council) grant agreement no.
-609819, SOMICS, and by ERC grant agreement no.
-616072, JAXPERTISE.
-
-\hypertarget{declarations-of-interest}{%
-\section{Declarations of interest}\label{declarations-of-interest}}
-
-The authors have no conflict of interest.
-
-\newpage
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
+\begin{CSLReferences}{0}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-thornton_2008}{}}%
+\CSLLeftMargin{1. }
+\CSLRightInline{Thornton A, Raihani NJ. The evolution of teaching. Animal Behaviour. 2008;75: 1823--1836. doi:\href{https://doi.org/10.1016/j.anbehav.2007.12.014}{10.1016/j.anbehav.2007.12.014}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tomasello_2016}{}}%
+\CSLLeftMargin{2. }
+\CSLRightInline{Tomasello M. The ontogeny of cultural learning. Current Opinion in Psychology. 2016;8: 1--4. doi:\href{https://doi.org/10.1016/j.copsyc.2015.09.008}{10.1016/j.copsyc.2015.09.008}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-brand_2002}{}}%
-Brand, R.
-J., Baldwin, D.
-A., \& Ashburn, L.
-A.
-(2002).
+\CSLLeftMargin{3. }
+\CSLRightInline{Brand RJ, Baldwin DA, Ashburn LA.
 Evidence for {`motionese'}: Modifications in mothers' infant-directed action.
-\emph{Developmental Science}, \emph{5}(1), 72--83.
-\url{https://doi.org/10.1111/1467-7687.00211}
+Developmental Science.
+2002;5: 72--83.
+doi:\href{https://doi.org/10.1111/1467-7687.00211}{10.1111/1467-7687.00211}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-saint-georges_2013}{}}%
+\CSLLeftMargin{4. }
+\CSLRightInline{Saint-Georges C, Chetouani M, Cassel R, Apicella F, Mahdhaoui A, Muratori F, et al.
+Motherese in {Interaction}: {At} the {Cross}-{Road} of {Emotion} and {Cognition}?
+({A Systematic Review}).
+PLOS ONE.
+{Public Library of Science}; 2013;8: e78103.
+doi:\href{https://doi.org/10.1371/journal.pone.0078103}{10.1371/journal.pone.0078103}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-uther_2007}{}}%
+\CSLLeftMargin{5. }
+\CSLRightInline{Uther M, Knoll MA, Burnham D.
+Do you speak {E}-{NG}-{L}-{I}-{SH}?
+{A} comparison of foreigner- and infant-directed speech.
+Speech Communication.
+2007;49: 2--7.
+doi:\href{https://doi.org/10.1016/j.specom.2006.10.003}{10.1016/j.specom.2006.10.003}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcellin_2017}{}}%
+\CSLLeftMargin{6. }
+\CSLRightInline{McEllin L, Knoblich G, Sebanz N.
+Distinct kinematic markers of demonstration and joint action coordination?
+{Evidence} from virtual xylophone playing.
+Journal of Experimental Psychology: Human Perception and Performance.
+2017;44: 885.
+doi:\href{https://doi.org/10.1037/xhp0000505}{10.1037/xhp0000505}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-csibra_2009}{}}%
+\CSLLeftMargin{7. }
+\CSLRightInline{Csibra G, Gergely G. Natural pedagogy. Trends in Cognitive Sciences. 2009;13: 148--153. doi:\href{https://doi.org/10.1016/j.tics.2009.01.005}{10.1016/j.tics.2009.01.005}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-campisi_2013}{}}%
+\CSLLeftMargin{8. }
+\CSLRightInline{Campisi E, Özyürek A.
+Iconicity as a communicative strategy: {Recipient} design in multimodal demonstrations for adults and children.
+Journal of Pragmatics.
+2013;47: 14--27.
+doi:\href{https://doi.org/10.1016/j.pragma.2012.12.007}{10.1016/j.pragma.2012.12.007}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-fukuyama_2015}{}}%
+\CSLLeftMargin{9. }
+\CSLRightInline{Fukuyama H, Qin S, Kanakogi Y, Nagai Y, Asada M, Myowa-Yamakoshi M.
+Infant's action skill dynamically modulates parental action demonstration in the dyadic interaction.
+Developmental Science.
+2015;18: 1006--1013.
+doi:\href{https://doi.org/10.1111/desc.12270}{10.1111/desc.12270}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schaik_2019}{}}%
+\CSLLeftMargin{10. }
+\CSLRightInline{Schaik JE van, Meyer M, Ham CR van, Hunnius S.
+Motion tracking of parents' infant- versus adult-directed actions reveals general and action-specific modulations.
+Developmental Science.
+2019;0: e12869.
+doi:\href{https://doi.org/10.1111/desc.12869}{10.1111/desc.12869}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-NIPS2016_6413}{}}%
+\CSLLeftMargin{11. }
+\CSLRightInline{Ho MK, Littman M, MacGlashan J, Cushman F, Austerweil JL.
+Showing versus doing: {Teaching} by demonstration.
+In: Lee DD, Sugiyama M, Luxburg UV, Guyon I, Garnett R, editors.
+Advances in neural information processing systems 29.
+{Curran Associates, Inc.}; 2016.
+pp.
+3027--3035.
+}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sloboda_2000}{}}%
+\CSLLeftMargin{12. }
+\CSLRightInline{Sloboda JA.
+Individual differences in music performance.
+Trends in Cognitive Sciences.
+2000;4: 397--403.
+doi:\href{https://doi.org/10.1016/S1364-6613(00)01531-X}{10.1016/S1364-6613(00)01531-X}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-juslin_2003}{}}%
+\CSLLeftMargin{13. }
+\CSLRightInline{Juslin PN.
+Five {Facets} of {Musical Expression}: {A Psychologist}'s {Perspective} on {Music Performance}.
+Psychology of Music.
+{SAGE Publications Ltd}; 2003;31: 273--302.
+doi:\href{https://doi.org/10.1177/03057356030313003}{10.1177/03057356030313003}}
 
 \leavevmode\vadjust pre{\hypertarget{ref-bresin_2000}{}}%
-Bresin, R., \& Battel, G.
-U.
-(2000).
+\CSLLeftMargin{14. }
+\CSLRightInline{Bresin R, Battel GU.
 Articulation {Strategies} in {Expressive Piano Performance Analysis} of {Legato}, {Staccato}, and {Repeated Notes} in {Performances} of the {Andante Movement} of {Mozart}'s {Sonata} in {G Major} ({K} 545).
-\emph{Journal of New Music Research}, \emph{29}(3), 211--224.
-\url{https://doi.org/10.1076/jnmr.29.3.211.3092}
-
-\leavevmode\vadjust pre{\hypertarget{ref-campisi_2013}{}}%
-Campisi, E., \& Özyürek, A.
-(2013).
-Iconicity as a communicative strategy: {Recipient} design in multimodal demonstrations for adults and children.
-\emph{Journal of Pragmatics}, \emph{47}(1), 14--27.
-\url{https://doi.org/10.1016/j.pragma.2012.12.007}
-
-\leavevmode\vadjust pre{\hypertarget{ref-csibra_2009}{}}%
-Csibra, G., \& Gergely, G. (2009). Natural pedagogy. \emph{Trends in Cognitive Sciences}, \emph{13}(4), 148--153. \url{https://doi.org/10.1016/j.tics.2009.01.005}
-
-\leavevmode\vadjust pre{\hypertarget{ref-fukuyama_2015}{}}%
-Fukuyama, H., Qin, S., Kanakogi, Y., Nagai, Y., Asada, M., \& Myowa-Yamakoshi, M.
-(2015).
-Infant's action skill dynamically modulates parental action demonstration in the dyadic interaction.
-\emph{Developmental Science}, \emph{18}(6), 1006--1013.
-\url{https://doi.org/10.1111/desc.12270}
-
-\leavevmode\vadjust pre{\hypertarget{ref-NIPS2016_6413}{}}%
-Ho, M.
-K., Littman, M., MacGlashan, J., Cushman, F., \& Austerweil, J.
-L.
-(2016).
-Showing versus doing: {Teaching} by demonstration.
-In D.
-D.
-Lee, M.
-Sugiyama, U.
-V.
-Luxburg, I.
-Guyon, \& R.
-Garnett (Eds.), \emph{Advances in neural information processing systems 29} (pp.
-3027--3035).
-{Curran Associates, Inc.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-juslin_2003}{}}%
-Juslin, P.
-N.
-(2003).
-Five {Facets} of {Musical Expression}: {A Psychologist}'s {Perspective} on {Music Performance}.
-\emph{Psychology of Music}, \emph{31}(3), 273--302.
-\url{https://doi.org/10.1177/03057356030313003}
+Journal of New Music Research.
+2000;29: 211--224.
+doi:\href{https://doi.org/10.1076/jnmr.29.3.211.3092}{10.1076/jnmr.29.3.211.3092}}
 
 \leavevmode\vadjust pre{\hypertarget{ref-lindstrom_2003}{}}%
-Lindström, E., Juslin, P.
-N., Bresin, R., \& Williamon, A.
-(2003).
+\CSLLeftMargin{15. }
+\CSLRightInline{Lindström E, Juslin PN, Bresin R, Williamon A.
 {``{Expressivity} comes from within your soul''}: {A} questionnaire study of music students' perspectives on expressivity.
-\emph{Research Studies in Music Education}, \emph{20}(1), 23--47.
-\url{https://doi.org/10.1177/1321103X030200010201}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mcellin_2017}{}}%
-McEllin, L., Knoblich, G., \& Sebanz, N.
-(2017).
-Distinct kinematic markers of demonstration and joint action coordination?
-{Evidence} from virtual xylophone playing.
-\emph{Journal of Experimental Psychology: Human Perception and Performance}, \emph{44}(6), 885.
-\url{https://doi.org/10.1037/xhp0000505}
+Research Studies in Music Education.
+{SAGE Publications Ltd}; 2003;20: 23--47.
+doi:\href{https://doi.org/10.1177/1321103X030200010201}{10.1177/1321103X030200010201}}
 
 \leavevmode\vadjust pre{\hypertarget{ref-okazaki_2019}{}}%
-Okazaki, S., Muraoka, Y., \& Osu, R.
-(2019).
+\CSLLeftMargin{16. }
+\CSLRightInline{Okazaki S, Muraoka Y, Osu R.
 Teacher-learner interaction quantifies scaffolding behaviour in imitation learning.
-\emph{Sci Rep}, \emph{9}(1), 1--13.
-\url{https://doi.org/10.1038/s41598-019-44049-x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-saint-georges_2013}{}}%
-Saint-Georges, C., Chetouani, M., Cassel, R., Apicella, F., Mahdhaoui, A., Muratori, F., \ldots{} Cohen, D.
-(2013).
-Motherese in {Interaction}: {At} the {Cross}-{Road} of {Emotion} and {Cognition}?
-({A Systematic Review}).
-\emph{PLOS ONE}, \emph{8}(10), e78103.
-\url{https://doi.org/10.1371/journal.pone.0078103}
-
-\leavevmode\vadjust pre{\hypertarget{ref-schaik_2019}{}}%
-Schaik, J.
-E.
-van, Meyer, M., Ham, C.
-R.
-van, \& Hunnius, S.
-(2019).
-Motion tracking of parents' infant- versus adult-directed actions reveals general and action-specific modulations.
-\emph{Developmental Science}, \emph{0}(0), e12869.
-\url{https://doi.org/10.1111/desc.12869}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sloboda_2000}{}}%
-Sloboda, J. A. (2000). Individual differences in music performance. \emph{Trends in Cognitive Sciences}, \emph{4}(10), 397--403. \url{https://doi.org/10.1016/S1364-6613(00)01531-X}
+Sci Rep.
+2019;9: 1--13.
+doi:\href{https://doi.org/10.1038/s41598-019-44049-x}{10.1038/s41598-019-44049-x}}
 
 \leavevmode\vadjust pre{\hypertarget{ref-sperber_1995}{}}%
-Sperber, D., \& Wilson, D. (1995). \emph{Relevance: {Communication} and cognition, 2nd ed}. {Malden}: {Blackwell Publishing}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-thornton_2008}{}}%
-Thornton, A., \& Raihani, N. J. (2008). The evolution of teaching. \emph{Animal Behaviour}, \emph{75}(6), 1823--1836. \url{https://doi.org/10.1016/j.anbehav.2007.12.014}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tomasello_2016}{}}%
-Tomasello, M. (2016). The ontogeny of cultural learning. \emph{Current Opinion in Psychology}, \emph{8}, 1--4. \url{https://doi.org/10.1016/j.copsyc.2015.09.008}
-
-\leavevmode\vadjust pre{\hypertarget{ref-uther_2007}{}}%
-Uther, M., Knoll, M.
-A., \& Burnham, D.
-(2007).
-Do you speak {E}-{NG}-{L}-{I}-{SH}?
-{A} comparison of foreigner- and infant-directed speech.
-\emph{Speech Communication}, \emph{49}(1), 2--7.
-\url{https://doi.org/10.1016/j.specom.2006.10.003}
+\CSLLeftMargin{17. }
+\CSLRightInline{Sperber D, Wilson D. Relevance: {Communication} and cognition, 2nd ed. {Malden}: {Blackwell Publishing}; 1995. }
 
@@
 
-\hypertarget{supplementary}{%
-\section{Supplementary Materials}\label{supplementary}}
-
-\hypertarget{instructions}{%
-\subsection{Instructions}\label{instructions}}
+\hypertarget{s1-file.-instructions}{%
+\section{S1 File. Instructions}\label{s1-file.-instructions}}
 
 \hypertarget{experiment-1-1}{%
-\subsubsection{Experiment 1}\label{experiment-1-1}}
+\subsection{Experiment 1}\label{experiment-1-1}}
 
@@
 \hypertarget{experiment-2-1}{%
-\subsubsection{Experiment 2}\label{experiment-2-1}}
+\subsection{Experiment 2}\label{experiment-2-1}}
 

### 2021-07-27 paper/manuscript.Rmd
finalise the manuscript for preprint
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-output: papaja::apa6_word
-# output:
-#   papaja::apa6_pdf:
-#     latex_engine: xelatex
+#output: papaja::apa6_word
+output:
+  papaja::apa6_pdf:
+    latex_engine: xelatex
 ---
@@
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
-In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners [@brand_2002; @saint-georges_2013].
+For instance, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners [@brand_2002; @saint-georges_2013].
 These modulations include slowing down and exaggerating sounds and actions.
@@
 
-Previous research has shown that demonstrators not only adjust their gestures [@campisi_2013] and actions [@fukuyama_2015] to learners' skills but engage in specific action modulations to highlight particular aspects of demonstrated actions.
+Previous research has shown that demonstrators not only adjust their gestures [@campisi_2013] and actions [@fukuyama_2015] to learners' skills but engage in specific action modulations to highlight certain aspects of demonstrated actions.
 For example, Schaik and colleagues [@schaik_2019] showed that adults used specific action modulations for demonstrating different action effects of objects to infants.
@@
 Importantly, in some domains teaching requires not only demonstrating *what* to do, but *how* to perform actions.
-In artistic contexts, novices need to learn how exactly to perform actions by relying on particular techniques [@sloboda_2000].
+In artistic contexts, novices need to learn how exactly to perform actions by relying on specific techniques [@sloboda_2000].
 For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques [@juslin_2003].
@@
 Specific and subtle aspects of the performance may need to be highlighted to teach a respective technique.
-Pianists were asked to perform a piece in order to demonstrate to a learner how to implement the notated expressions (teaching condition) and to play the same piece for an audience (performing condition).
+Pianists were asked to perform a piece to demonstrate to a learner how to implement the notated expressions (teaching condition) and to play the same piece for an audience (performing condition).
 Two basic expressive techniques in piano performance were used: articulation (legato and staccato) and dynamics (forte and piano).
@@
 Three dependent variables were computed for data analysis.
-Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
-Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone, and measures the smoothness of musical sequences [@bresin_2000].
+Inter-onset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences [@bresin_2000].
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
@@
 As a result, less than 1 % of total responses were corrected.
-In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
+In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable.
 For each dependent variable, this resulted in less than 5% of overall responses being removed as outliers.
 
-We performed separate analyses for the two expressive techniques (i.e., articulation, dynamics).
+We performed separate analyses for the two expressive techniques (i.e., articulation and dynamics).
 A paired-sample *t*-test or a Wilcoxon Signed-rank test (a non-parametric alternative to a paired *t*-test) was performed to compare the mean IOIs in the teaching and performing condition.
@@
 We first report results for performance of the piece with the notated expression of articulation (*Fig \ref{fig:stimuli} B*), followed by performance of the piece with the notated expression of dynamics (*Fig \ref{fig:stimuli} C*).
-To recall our predictions, if participants play more slowly when they are trying to teach, interonset intervals (IOIs) should be larger when teaching.
+To recall our predictions, if participants play more slowly when they are trying to teach, inter-onset intervals (IOIs) should be larger when teaching.
 If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should be more positive for legato and more negative for staccato when teaching articulation, and key velocity (KV) should be higher for forte and lower for piano when teaching dynamics.
@@
 ## Articulation
-### Interonset Intervals (IOIs)
+### Inter-onset Intervals (IOIs)
 
-In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, two-tailed, *Fig \ref{fig:ioi-1}*).
+To compare the mean IOIs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_norm$p.value))}`).
+The Wilcoxon Signed-rank test revealed that participants played more slowly in the teaching condition [*Mdn* = `r ioi_art_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "teaching"]$IQR`] than in the performing condition [*Mdn* = `r ioi_art_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_art_all[Condition == "performing"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, two-tailed, *Fig \ref{fig:ioi-1}*).
 
@@
 
-### Articulation: Key-Overlap Time (KOT)
+### Key-Overlap Time (KOT)
 
@@
 ## Dynamics
-### Interonset Intervals (IOIs)
+### Inter-onset Intervals (IOIs)
 
-In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_norm$p.value))}`).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`, two-tailed, *Fig \ref{fig:ioi-1}*).
+To compare the mean IOIs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_norm$p.value))}`).
+The Wilcoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`, two-tailed, *Fig \ref{fig:ioi-1}*).
 
@@
 Additionally, we conducted an exploratory analysis by focusing on specific points at which each subcomponent changed from one to the other (i.e., forte to piano: FtoP, piano to forte: PtoF).
-These points could be structurally important in order to make a contrast between forte and piano.
+These points could be structurally important to make a contrast between forte and piano.
 We calculated the KV difference for each interval by subtracting the KV value of the current note from that of the following note.
@@
 Importantly, participants did not modulate their performance in terms of irrelevant aspects of the techniques for teaching purposes (e.g., modulating the smoothness of sound while teaching dynamics).
-These findings confirmed that participants modulated their performance in systematic and fine-grained ways in order to teach expressive techniques.
+These findings confirmed that participants modulated their performance in systematic and fine-grained ways to teach expressive techniques.
 
@@
 
-In Experiment 1, we employed a simple musical scale in order to maximise experimental control.
+In Experiment 1, we employed a simple musical scale to maximise experimental control.
 To test whether our results generalise to a more naturalistic piece, in Experiment 2, we chose an actual piano piece and modified it for the purpose of the experiment.
@@
 The same apparatus as in Experiment 1 was used.
-We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expressions (i.e., articulation, dynamics) and was relatively simple in terms of motor skills.
+We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expressions (i.e., articulation and dynamics) and was relatively simple in terms of motor skills.
 The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
@@
 The chosen tempo was again cued by a leading metronome.
-In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music (*Fig \ref{fig:stimuli-2} A*).
+To make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music (*Fig \ref{fig:stimuli-2} A*).
 Those who could not perform the piece without pitch errors twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded).
@@
 IOIs were normalised by their preferred tempo because each participant chose a tempo from the three options.
-Given the different tempi, a key-overlap ratio (KOR) was calculated by dividing KOT by the mean IOI of each performance in order to normalise KOT.
+Given the different tempi, key-overlap ratios (KORs) were calculated by dividing KOT by the mean IOI of each performance to normalise KOT.
 Additionally, we include KV difference (i.e., KV difference for each interval) at transition points (i.e., forte to piano or piano to forte) as a dependent variable based on the findings of Experiment 1.
@@
 ## Articulation
-### Interonset Intervals (IOIs)
+### Inter-onset Intervals (IOIs)
 
@@
 , *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art$effsize), 3)`, two-tailed, *Fig \ref{fig:ioi-2}*).
-A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, two-tailed).
+A Wilcoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, two-tailed).
 
@@
 
-### Key-Overlap Ratio (KOR)
+### Key-Overlap Ratios (KORs)
 
@@
 staccato) showed that there was a significant main effect of Articulation (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Articulation (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KOR was more negative in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`).
-For legato there was no significant difference in KOR between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KORs were more negative in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`).
+For legato there was no significant difference in KORs between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
 
@@
 
-kor_art_2 <- ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
+kor_art_2 <- ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KORs", title = "KORs: Articulation") +
   ylim(-1, 0.4) +
@@
 ## Dynamics
-### Interonset Intervals (IOIs)
+### Inter-onset Intervals (IOIs)
 
 A paired-sample *t*-test showed no significant difference between the teaching condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "teaching"]$SD)`] and the performing condition [*M* = `r sprintf('%.2f', ioi_dyn_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_dyn_all[Condition == "performing"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_ttest$statistic))`, *p* `r if(ioi_dyn_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn$effsize), 3)`, two-tailed, *Fig \ref{fig:ioi-2}*).
-A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, two-tailed).
+A Wilcoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`, two-tailed).
 
@@
 
-kor_dyn_2 <- ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
+kor_dyn_2 <- ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KORs", title = "KORs: Dynamics") +
   ylim(-1, 0.4) +
@@
 As in Experiment 1, we observed slower performance only while teaching the notated articulation.
-Again, KOR showed that when teaching articulation, participants exaggerated staccato while there was no significant difference in legato between the two conditions.
+Again, KORs showed that when teaching articulation, participants exaggerated staccato while there was no significant difference in legato between the two conditions.
 While teaching dynamics, KV showed that participants exaggerated forte and piano.
 This pattern of exaggeration was more pronounced than in Experiment 1, where only higher KV for forte was observed.
-Furthermore, we again found that when teaching dynamics, participants made a larger dynamics contrast between forte and piano bidirectionally at transition points.
-Importantly, participants did not modulate performance aspects that were irrelevant with regard to the techniques to be taught (i.e., no modulation of dynamics when teaching articulation and vice versa).
+Furthermore, we again found that when teaching dynamics, at transition points participants produced a larger contrast in dynamics between forte and piano, bidirectionally.
+Importantly, participants did not modulate performance aspects that were irrelevant for the techniques to be taught (i.e., no modulation of dynamics when teaching articulation and vice versa).
 Overall, Experiment 2 demonstrated systematic and fine-grained didactic performance modulations for a naturalistic piece of music.
@@
 
-```{r plot-kor-2, out.width = "100%", fig.cap = "\\label{fig:kor-2}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right).
+```{r plot-kor-2, out.width = "100%", fig.cap = "\\label{fig:kor-2}Experiment 2: KORs when playing the piece with either articulation (left) or dynamics (right).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
@@
 
-The aim of the study was to investigate how expert pianists modulate their performance in order to teach expressive techniques.
+The aim of the study was to investigate how expert pianists modulate their performance to teach expressive techniques.
 Overall, the findings of the two experiments showed that pianists performed one and the same piece differently depending on whether they played with the intention to teach expressive techniques or with the intention to perform the piece for an audience.
@@
 As mentioned in the discussion of Experiment 1, it could be that the metronome beats given prior to each performance discouraged participants from deviating from the prescribed tempo.
-Another possibility is that expression is actually best taught when leaving the tempo unchanged when it is irrelevant.
+Another possibility is that expression is best taught when leaving the tempo unchanged when it is irrelevant.
 This would imply that general didactic modulations like slowing down might be less useful in the context of teaching expressive techniques where timing itself can be used to add an expression to a performance.
@@
 First, our study sheds light on the teaching of expressive techniques that are an integral part of skill acquisition in artistic domains.
-We showed that compared to an expressive performance baseline, experts made specific modulations in order to teach particular techniques.
+We showed that compared to an expressive performance baseline, experts made specific modulations to teach particular techniques.
 Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention.

### 2021-07-27 paper/manuscript.tex
finalise the manuscript for preprint
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
-In particular, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners {[}3,4{]}.
+For instance, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners {[}3,4{]}.
 These modulations include slowing down and exaggerating sounds and actions.
@@
 
-Previous research has shown that demonstrators not only adjust their gestures {[}8{]} and actions {[}9{]} to learners' skills but engage in specific action modulations to highlight particular aspects of demonstrated actions.
+Previous research has shown that demonstrators not only adjust their gestures {[}8{]} and actions {[}9{]} to learners' skills but engage in specific action modulations to highlight certain aspects of demonstrated actions.
 For example, Schaik and colleagues {[}10{]} showed that adults used specific action modulations for demonstrating different action effects of objects to infants.
@@
 Importantly, in some domains teaching requires not only demonstrating \emph{what} to do, but \emph{how} to perform actions.
-In artistic contexts, novices need to learn how exactly to perform actions by relying on particular techniques {[}12{]}.
+In artistic contexts, novices need to learn how exactly to perform actions by relying on specific techniques {[}12{]}.
 For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques {[}13{]}.
@@
 Three dependent variables were computed for data analysis.
-Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
-Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone, and measures the smoothness of musical sequences {[}14{]}.
+Inter-onset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences {[}14{]}.
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
@@
 As a result, less than 1 \% of total responses were corrected.
-In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable and excluded from data analysis.
+In addition to pitch errors, we removed outliers for IOIs, KOT and KV, defined as values more than 3 standard deviations from the mean of each dependent variable.
 For each dependent variable, this resulted in less than 5\% of overall responses being removed as outliers.
 
-We performed separate analyses for the two expressive techniques (i.e., articulation, dynamics).
+We performed separate analyses for the two expressive techniques (i.e., articulation and dynamics).
 A paired-sample \emph{t}-test or a Wilcoxon Signed-rank test (a non-parametric alternative to a paired \emph{t}-test) was performed to compare the mean IOIs in the teaching and performing condition.
@@
 We first report results for performance of the piece with the notated expression of articulation (\emph{Fig \ref{fig:stimuli} B}), followed by performance of the piece with the notated expression of dynamics (\emph{Fig \ref{fig:stimuli} C}).
-To recall our predictions, if participants play more slowly when they are trying to teach, interonset intervals (IOIs) should be larger when teaching.
+To recall our predictions, if participants play more slowly when they are trying to teach, inter-onset intervals (IOIs) should be larger when teaching.
 If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should be more positive for legato and more negative for staccato when teaching articulation, and key velocity (KV) should be higher for forte and lower for piano when teaching dynamics.
@@
 
-\hypertarget{interonset-intervals-iois}{%
-\subsubsection{Interonset Intervals (IOIs)}\label{interonset-intervals-iois}}
+\hypertarget{inter-onset-intervals-iois}{%
+\subsubsection{Inter-onset Intervals (IOIs)}\label{inter-onset-intervals-iois}}
 
-In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wiscoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
+To compare the mean IOIs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
+The Wilcoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
 
-\hypertarget{articulation-key-overlap-time-kot}{%
-\subsubsection{Articulation: Key-Overlap Time (KOT)}\label{articulation-key-overlap-time-kot}}
+\hypertarget{key-overlap-time-kot}{%
+\subsubsection{Key-Overlap Time (KOT)}\label{key-overlap-time-kot}}
 
@@
 
-\hypertarget{interonset-intervals-iois-1}{%
-\subsubsection{Interonset Intervals (IOIs)}\label{interonset-intervals-iois-1}}
+\hypertarget{inter-onset-intervals-iois-1}{%
+\subsubsection{Inter-onset Intervals (IOIs)}\label{inter-onset-intervals-iois-1}}
 
-In order to compare the mean IOIs between the teaching and performing condition, we conducted a Wiscoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wiscoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
+To compare the mean IOIs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
+The Wilcoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
 
@@
 
-\hypertarget{key-overlap-time-kot}{%
-\subsubsection{Key-Overlap Time (KOT)}\label{key-overlap-time-kot}}
+\hypertarget{key-overlap-time-kot-1}{%
+\subsubsection{Key-Overlap Time (KOT)}\label{key-overlap-time-kot-1}}
 
@@
 The chosen tempo was again cued by a leading metronome.
-In order to make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music (\emph{Fig \ref{fig:stimuli-2} A}).
+To make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music (\emph{Fig \ref{fig:stimuli-2} A}).
 Those who could not perform the piece without pitch errors twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded).
@@
 
-\hypertarget{interonset-intervals-iois-2}{%
-\subsubsection{Interonset Intervals (IOIs)}\label{interonset-intervals-iois-2}}
+\hypertarget{inter-onset-intervals-iois-2}{%
+\subsubsection{Inter-onset Intervals (IOIs)}\label{inter-onset-intervals-iois-2}}
 
@@
 , \emph{p} = 0.023, Cohen's \emph{d} = 0.55, two-tailed, \emph{Fig \ref{fig:ioi-2}}).
-A Wiscoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (\emph{p} = 0.033, \emph{r} = 0.48, two-tailed).
+A Wilcoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (\emph{p} = 0.033, \emph{r} = 0.48, two-tailed).
 
-\hypertarget{key-overlap-ratio-kor}{%
-\subsubsection{Key-Overlap Ratio (KOR)}\label{key-overlap-ratio-kor}}
+\hypertarget{key-overlap-ratios-kors}{%
+\subsubsection{Key-Overlap Ratios (KORs)}\label{key-overlap-ratios-kors}}
 
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant main effect of Articulation (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Articulation (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01, \emph{Fig \ref{fig:kor-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KOR was more negative in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.002).
-For legato there was no significant difference in KOR between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.56).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KORs were more negative in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.002).
+For legato there was no significant difference in KORs between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.56).
 
@@
 
-\hypertarget{interonset-intervals-iois-3}{%
-\subsubsection{Interonset Intervals (IOIs)}\label{interonset-intervals-iois-3}}
+\hypertarget{inter-onset-intervals-iois-3}{%
+\subsubsection{Inter-onset Intervals (IOIs)}\label{inter-onset-intervals-iois-3}}
 
 A paired-sample \emph{t}-test showed no significant difference between the teaching condition {[}\emph{M} = 0.96, \emph{SD} = 0.048{]} and the performing condition {[}\emph{M} = 0.96, \emph{SD} = 0.043{]} while playing the piece with the notated dynamics (\emph{t}(19) = 0.21, \emph{p} = 0.84, Cohen's \emph{d} = 0.05, two-tailed, \emph{Fig \ref{fig:ioi-2}}).
-A Wiscoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (\emph{p} = 0.96, two-tailed).
+A Wilcoxon Signed-rank test also confirmed that there was no significant difference between the two conditions (\emph{p} = 0.96, \emph{r} = 0.02, two-tailed).
 
@@
 
-\hypertarget{key-overlap-ratio-kor-1}{%
-\subsubsection{Key-Overlap Ratio (KOR)}\label{key-overlap-ratio-kor-1}}
+\hypertarget{key-overlap-ratios-kors-1}{%
+\subsubsection{Key-Overlap Ratios (KORs)}\label{key-overlap-ratios-kors-1}}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kor-2-1} \caption{\label{fig:kor-2}Experiment 2: KOR when playing the piece with either articulation (left) or dynamics (right).
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kor-2-1} \caption{\label{fig:kor-2}Experiment 2: KORs when playing the piece with either articulation (left) or dynamics (right).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.

### 2022-01-14 paper/manuscript.Rmd
updated manuscript for resubmission
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   - name          : "Atsuko Tominaga"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Quellenstraße 51, 1100 Vienna, Austria"
-    email         : "Tominaga_Atsuko@phd.ceu.edu"
+    affiliation   : "1,*"
+#    corresponding : yes    # Define only one corresponding author
+#    address       : "Quellenstraße 51, 1100 Vienna, Austria"
+#    email         : "Tominaga_Atsuko@phd.ceu.edu"
 
@@
   - id            : "1"
-    institution   : "Department of Cognitive Science, Central European University, Vienna, Austria"
+    institution   : "Department of Cognitive Science, Central European University, Quellenstraße 51, 1100 Vienna, Austria"
+  - id            : "*"
+    institution   : "Corresponding author: Tominaga_Atsuko@phd.ceu.edu"
 
@@
 bibliography      : ["reference.bib"]
-csl: plos.csl
+csl: nature.csl
 
@@
 
-#output: papaja::apa6_word
-output:
-  papaja::apa6_pdf:
-    latex_engine: xelatex
+output: papaja::apa6_word
+#output:
+#  papaja::apa6_pdf:
+#    latex_engine: xelatex
 ---
@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[S1 File](#supplementary)*).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplementary Information](#supplementary)*).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
@@
 First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2} A*) prior to the experiment.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[S1 File](#supplemental)*).
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[Supplementary Information](#supplementary)*).
 Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute).
@@
 
-# Acknowledgements
-We thank Dávid Csűrös for his help with data collection.
-
 # References
@@
 
+# Acknowledgements
+This research was supported by the European Research Council under the European Union’s Seventh Frame- work Program (FP7/2007–2013)/ERC Grant agreement no.
+609819, SOMICS, and by ERC Grant agreement no.
+616072, JAXPERTISE.
+We thank Dávid Csűrös for his help with data collection.
+
+# Author contributions
+A.T, G.K and N.S developed the study concept and designed the experiments.
+Stimuli and procedure were created by A.T.
+Data collection and analyses were performed by A.T.
+under the supervision of G.K.
+and N.S.
+A.T.
+drafted the manuscript, and N.S.
+and G.K.
+provided critical revisions.
+All authors approved the final version of the manuscript for submission.
+
+# Data availability statement
+All data is available at [https://osf.io/8nbjh/](https://osf.io/8nbjh/).
+
+# Competing interests
+The authors declare no competing interests.
+
 \clearpage
 
-# S1 File. Instructions
-## Experiment 1
+# Supplementary Information{#supplementary}
+## Instructions (Experiment 1)
 
@@
 
-## Experiment 2
+## Instructions (Experiment 2)
 

### 2022-01-14 paper/manuscript.tex
updated manuscript for resubmission
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \documentclass[
-  english,
   man,floatsintext]{apa6}
-\title{The sound of teaching music: Expert pianists' performance modulations for novices}
-\author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
-\date{}
-
 \usepackage{amsmath,amssymb}
@@
   pdftitle={The sound of teaching music: Expert pianists' performance modulations for novices},
-  pdfauthor={Atsuko Tominaga1, Günther Knoblich1, \& Natalie Sebanz1},
+  pdfauthor={Atsuko Tominaga1,*, Günther Knoblich1, \& Natalie Sebanz1},
   pdflang={en-EN},
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\ifLuaTeX
+\usepackage[bidi=basic]{babel}
+\else
+\usepackage[bidi=default]{babel}
+\fi
+\babelprovide[main,import]{english}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 % Manuscript styling
@@
 
+% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
+% See https://github.com/crsh/papaja/issues/292
+\makeatletter
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries\itshape\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-\z@\relax}%
+  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+\makeatother
+
 % \usepackage{etoolbox}
@@
 \makeatother
-\shorttitle{The sound of teaching music}
+
+\usepackage{xpatch}
+\makeatletter
+\xapptocmd\appendix
+  {\xapptocmd\section
+    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
+    {}{\InnerPatchFailed}%
+  }
+{}{\PatchFailed}
 \keywords{teaching, expertise, skill transmission, artistic expression, music}
 \usepackage{csquotes}
-\ifXeTeX
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
 \ifLuaTeX
@@
 
-\authornote{
+\title{The sound of teaching music: Expert pianists' performance modulations for novices}
+\author{Atsuko Tominaga\textsuperscript{1,*}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
+\date{}
 
-Correspondence concerning this article should be addressed to Atsuko Tominaga, Quellenstraße 51, 1100 Vienna, Austria.
-E-mail: \href{mailto:Tominaga_Atsuko@phd.ceu.edu}{\nolinkurl{Tominaga\_Atsuko@phd.ceu.edu}}
+\shorttitle{The sound of teaching music}
 
-}
-
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Cognitive Science, Central European University, Vienna, Austria}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Cognitive Science, Central European University, Quellenstraße 51, 1100 Vienna, Austria\\\textsuperscript{*} Corresponding author: \href{mailto:Tominaga_Atsuko@phd.ceu.edu}{\nolinkurl{Tominaga\_Atsuko@phd.ceu.edu}}}
 
@@
 
-Deliberate teaching has supported human skill transmission over generations and provides a key route for learning {[}1,2{]}.
+Deliberate teaching has supported human skill transmission over generations and provides a key route for learning\textsuperscript{1,2}.
 Experts modulate their behaviour so that novices can extract relevant information to learn a novel skill.
-For instance, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners {[}3,4{]}.
+For instance, adults modulate their speech (motherese) and actions (motionese) when demonstrating a skill to infant learners\textsuperscript{3,4}.
 These modulations include slowing down and exaggerating sounds and actions.
-Similar findings were obtained in studies with adult learners, where exaggerations were observed when native British English speakers were talking to first language English learners {[}5{]} and when skilled adults were teaching xylophone melodies to novices {[}6{]}.
-The observed modulations are thought to provide communicative signals that can facilitate learning by affecting attention and memory {[}7{]}.
+Similar findings were obtained in studies with adult learners, where exaggerations were observed when native British English speakers were talking to first language English learners\textsuperscript{5} and when skilled adults were teaching xylophone melodies to novices\textsuperscript{6}.
+The observed modulations are thought to provide communicative signals that can facilitate learning by affecting attention and memory\textsuperscript{7}.
 
-Previous research has shown that demonstrators not only adjust their gestures {[}8{]} and actions {[}9{]} to learners' skills but engage in specific action modulations to highlight certain aspects of demonstrated actions.
-For example, Schaik and colleagues {[}10{]} showed that adults used specific action modulations for demonstrating different action effects of objects to infants.
-Ho and colleagues {[}11{]} found that demonstrators chose costly movement paths at structurally important points to disambiguate one goal over other possibilities.
+Previous research has shown that demonstrators not only adjust their gestures\textsuperscript{8} and actions\textsuperscript{9} to learners' skills but engage in specific action modulations to highlight certain aspects of demonstrated actions.
+For example, Schaik and colleagues\textsuperscript{10} showed that adults used specific action modulations for demonstrating different action effects of objects to infants.
+Ho and colleagues\textsuperscript{11} found that demonstrators chose costly movement paths at structurally important points to disambiguate one goal over other possibilities.
 
 Importantly, in some domains teaching requires not only demonstrating \emph{what} to do, but \emph{how} to perform actions.
-In artistic contexts, novices need to learn how exactly to perform actions by relying on specific techniques {[}12{]}.
-For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques {[}13{]}.
+In artistic contexts, novices need to learn how exactly to perform actions by relying on specific techniques\textsuperscript{12}.
+For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques\textsuperscript{13}.
 This raises the question of whether and how experts modulate their actions during didactic demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{S1 File}}).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary Information}}).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
@@
 Inter-onset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
-Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences {[}14{]}.
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences\textsuperscript{14}.
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) revealed that there was a significant main effect of Articulation (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Articulation (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011, \emph{Fig \ref{fig:kot-1}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KOT was more negative in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.004).
-For legato there was no significant difference in KOT between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition {[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.32).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KOT was more negative in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.007).
+For legato there was no significant difference in KOT between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition {[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.31).
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant interaction between Condition and Articulation (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004, \emph{Fig \ref{fig:vel-1}}).
-However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Legato: \emph{M} = 70.34, \emph{SD} = 6.45; Staccato: \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato: \emph{M} = 70.68, \emph{SD} = 5.84, Staccato: \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato: \emph{p} = 0.70, Staccato: \emph{p} = 0.07).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Legato: \emph{M} = 70.34, \emph{SD} = 6.45; Staccato: \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato: \emph{M} = 70.68, \emph{SD} = 5.84, Staccato: \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato: \emph{p} = 0.68, Staccato: \emph{p} = 0.09).
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Dynamics (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003, \emph{Fig \ref{fig:vel-1}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.021) when instructed to play forte.
-There was no significant difference in KV between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.45) when instructed to play piano.
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.069) when instructed to play forte.
+There was no significant difference in KV between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.24) when instructed to play piano.
 
@@
 For IOIs, we found a small but significant slowing down during teaching specifically when playing the piece with the notated articulation.
-This finding is in line with earlier studies that found slower performance of actions in a teaching context {[}e.g., 3,6{]}.
+This finding is in line with earlier studies that found slower performance of actions in a teaching context\textsuperscript{e.g., 3,6}.
 However, we did not find a significant difference in tempo when participants were teaching dynamics.
@@
 First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Fig \ref{fig:stimuli-2} A}) prior to the experiment.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplemental}{S1 File}}).
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplementary}{Supplementary Information}}).
 Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute).
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant main effect of Articulation (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Articulation (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01, \emph{Fig \ref{fig:kor-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KORs were more negative in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.002).
-For legato there was no significant difference in KORs between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.56).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KORs were more negative in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.000).
+For legato there was no significant difference in KORs between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.63).
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,19) = 132, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Dynamics (\emph{F}(1,19) = 6.12, \emph{p} = 0.023, \(\eta_G^2\) = 0.018, \emph{Fig \ref{fig:vel-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 83.20, \emph{SD} = 8.84{]} than in the performing condition {[}\emph{M} = 81.49, \emph{SD} = 8.12{]} (\emph{p} = 0.040) when instructed to play forte.
-Also, KV was lower in the teaching condition {[}\emph{M} = 60.53, \emph{SD} = 3.76{]} than in the performing condition {[}\emph{M} = 62.42, \emph{SD} = 4.82{]} (\emph{p} = 0.024) when instructed to play piano.
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 83.20, \emph{SD} = 8.84{]} than in the performing condition {[}\emph{M} = 81.49, \emph{SD} = 8.12{]} (\emph{p} = 0.020) when instructed to play forte.
+Also, KV was lower in the teaching condition {[}\emph{M} = 60.53, \emph{SD} = 3.76{]} than in the performing condition {[}\emph{M} = 62.42, \emph{SD} = 4.82{]} (\emph{p} = 0.046) when instructed to play piano.
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction between Condition and Transition Type (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059, \emph{Fig \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (\emph{p} = 0.029) and from piano to forte (\emph{p} = 0.002) in the teaching condition {[}FtoP: \emph{M} = -17.23, \emph{SD} = 8.26, PtoF: \emph{M} = 20.71, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP: \emph{M} = -14.04, \emph{SD} = 7.67, PtoF: \emph{M} = 16.05, \emph{SD} = 7.32{]}.
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (\emph{p} = 0.029) and from piano to forte (\emph{p} = 0.003) in the teaching condition {[}FtoP: \emph{M} = -17.23, \emph{SD} = 8.26, PtoF: \emph{M} = 20.71, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP: \emph{M} = -14.04, \emph{SD} = 7.67, PtoF: \emph{M} = 16.05, \emph{SD} = 7.32{]}.
 
@@
 
-Musicians and music teachers tend to consider expressive skills as a performer's most important skill {[}15{]}.
+Musicians and music teachers tend to consider expressive skills as a performer's most important skill\textsuperscript{15}.
 In the present study, experts focused on the teaching of expressions that were notated.
-However, expression also has other facets {[}13{]} that are not only piece-related but performer- or context-related.
+However, expression also has other facets\textsuperscript{13} that are not only piece-related but performer- or context-related.
 An interesting topic for future research is how experts teach not only expressive techniques as such but convey possibilities of interpretation.
@@
 One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students.
-Given that teachers modulate their demonstration throughout online interactions with learners {[}16{]}, future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities.
+Given that teachers modulate their demonstration throughout online interactions with learners\textsuperscript{16}, future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities.
 
@@
 Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention.
-Rather, expert demonstrators seem to follow principles of relevance in communication {[}17{]}, highlighting only those aspects they are intending to demonstrate.
+Rather, expert demonstrators seem to follow principles of relevance in communication\textsuperscript{17}, highlighting only those aspects they are intending to demonstrate.
 How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher's intentions explicitly adds to the learning success are important questions for future research.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-We thank Dávid Csűrös for his help with data collection.
 
@@
 \CSLLeftMargin{1. }
-\CSLRightInline{Thornton A, Raihani NJ. The evolution of teaching. Animal Behaviour. 2008;75: 1823--1836. doi:\href{https://doi.org/10.1016/j.anbehav.2007.12.014}{10.1016/j.anbehav.2007.12.014}}
+\CSLRightInline{Thornton, A. \& Raihani, N. J. \href{https://doi.org/10.1016/j.anbehav.2007.12.014}{The evolution of teaching}. \emph{Animal Behaviour} \textbf{75}, 1823--1836 (2008).}
 
@@
 \CSLLeftMargin{2. }
-\CSLRightInline{Tomasello M. The ontogeny of cultural learning. Current Opinion in Psychology. 2016;8: 1--4. doi:\href{https://doi.org/10.1016/j.copsyc.2015.09.008}{10.1016/j.copsyc.2015.09.008}}
+\CSLRightInline{Tomasello, M. \href{https://doi.org/10.1016/j.copsyc.2015.09.008}{The ontogeny of cultural learning}. \emph{Current Opinion in Psychology} \textbf{8}, 1--4 (2016).}
 
@@
 \CSLLeftMargin{3. }
-\CSLRightInline{Brand RJ, Baldwin DA, Ashburn LA.
-Evidence for {`motionese'}: Modifications in mothers' infant-directed action.
-Developmental Science.
-2002;5: 72--83.
-doi:\href{https://doi.org/10.1111/1467-7687.00211}{10.1111/1467-7687.00211}}
+\CSLRightInline{Brand, R.
+J., Baldwin, D.
+A.
+\& Ashburn, L.
+A.
+\href{https://doi.org/10.1111/1467-7687.00211}{Evidence for {`motionese'}: Modifications in mothers' infant-directed action}.
+\emph{Developmental Science} \textbf{5}, 72--83 (2002).}
 
@@
 \CSLLeftMargin{4. }
-\CSLRightInline{Saint-Georges C, Chetouani M, Cassel R, Apicella F, Mahdhaoui A, Muratori F, et al.
-Motherese in {Interaction}: {At} the {Cross}-{Road} of {Emotion} and {Cognition}?
-({A Systematic Review}).
-PLOS ONE.
-{Public Library of Science}; 2013;8: e78103.
-doi:\href{https://doi.org/10.1371/journal.pone.0078103}{10.1371/journal.pone.0078103}}
+\CSLRightInline{Saint-Georges, C.
+\emph{et al.} \href{https://doi.org/10.1371/journal.pone.0078103}{Motherese in {Interaction}: {At} the {Cross-Road} of {Emotion} and {Cognition}?
+({A Systematic Review})}.
+\emph{PLOS ONE} \textbf{8}, e78103 (2013).}
 
@@
 \CSLLeftMargin{5. }
-\CSLRightInline{Uther M, Knoll MA, Burnham D.
-Do you speak {E}-{NG}-{L}-{I}-{SH}?
-{A} comparison of foreigner- and infant-directed speech.
-Speech Communication.
-2007;49: 2--7.
-doi:\href{https://doi.org/10.1016/j.specom.2006.10.003}{10.1016/j.specom.2006.10.003}}
+\CSLRightInline{Uther, M., Knoll, M.
+A.
+\& Burnham, D.
+\href{https://doi.org/10.1016/j.specom.2006.10.003}{Do you speak {E-NG-L-I-SH}?
+{A} comparison of foreigner- and infant-directed speech}.
+\emph{Speech Communication} \textbf{49}, 2--7 (2007).}
 
@@
 \CSLLeftMargin{6. }
-\CSLRightInline{McEllin L, Knoblich G, Sebanz N.
-Distinct kinematic markers of demonstration and joint action coordination?
-{Evidence} from virtual xylophone playing.
-Journal of Experimental Psychology: Human Perception and Performance.
-2017;44: 885.
-doi:\href{https://doi.org/10.1037/xhp0000505}{10.1037/xhp0000505}}
+\CSLRightInline{McEllin, L., Knoblich, G.
+\& Sebanz, N.
+\href{https://doi.org/10.1037/xhp0000505}{Distinct kinematic markers of demonstration and joint action coordination?
+{Evidence} from virtual xylophone playing.} \emph{Journal of Experimental Psychology: Human Perception and Performance} \textbf{44}, 885 (2017).}
 
@@
 \CSLLeftMargin{7. }
-\CSLRightInline{Csibra G, Gergely G. Natural pedagogy. Trends in Cognitive Sciences. 2009;13: 148--153. doi:\href{https://doi.org/10.1016/j.tics.2009.01.005}{10.1016/j.tics.2009.01.005}}
+\CSLRightInline{Csibra, G. \& Gergely, G. \href{https://doi.org/10.1016/j.tics.2009.01.005}{Natural pedagogy}. \emph{Trends in Cognitive Sciences} \textbf{13}, 148--153 (2009).}
 
@@
 \CSLLeftMargin{8. }
-\CSLRightInline{Campisi E, Özyürek A.
-Iconicity as a communicative strategy: {Recipient} design in multimodal demonstrations for adults and children.
-Journal of Pragmatics.
-2013;47: 14--27.
-doi:\href{https://doi.org/10.1016/j.pragma.2012.12.007}{10.1016/j.pragma.2012.12.007}}
+\CSLRightInline{Campisi, E.
+\& Özyürek, A.
+\href{https://doi.org/10.1016/j.pragma.2012.12.007}{Iconicity as a communicative strategy: {Recipient} design in multimodal demonstrations for adults and children}.
+\emph{Journal of Pragmatics} \textbf{47}, 14--27 (2013).}
 
@@
 \CSLLeftMargin{9. }
-\CSLRightInline{Fukuyama H, Qin S, Kanakogi Y, Nagai Y, Asada M, Myowa-Yamakoshi M.
-Infant's action skill dynamically modulates parental action demonstration in the dyadic interaction.
-Developmental Science.
-2015;18: 1006--1013.
-doi:\href{https://doi.org/10.1111/desc.12270}{10.1111/desc.12270}}
+\CSLRightInline{Fukuyama, H.
+\emph{et al.} \href{https://doi.org/10.1111/desc.12270}{Infant's action skill dynamically modulates parental action demonstration in the dyadic interaction}.
+\emph{Developmental Science} \textbf{18}, 1006--1013 (2015).}
 
@@
 \CSLLeftMargin{10. }
-\CSLRightInline{Schaik JE van, Meyer M, Ham CR van, Hunnius S.
-Motion tracking of parents' infant- versus adult-directed actions reveals general and action-specific modulations.
-Developmental Science.
-2019;0: e12869.
-doi:\href{https://doi.org/10.1111/desc.12869}{10.1111/desc.12869}}
+\CSLRightInline{Schaik, J.
+E.
+van, Meyer, M., Ham, C.
+R.
+van \& Hunnius, S.
+\href{https://doi.org/10.1111/desc.12869}{Motion tracking of parents' infant- versus adult-directed actions reveals general and action-specific modulations}.
+\emph{Developmental Science} \textbf{0}, e12869 (2019).}
 
@@
 \CSLLeftMargin{11. }
-\CSLRightInline{Ho MK, Littman M, MacGlashan J, Cushman F, Austerweil JL.
+\CSLRightInline{Ho, M.
+K., Littman, M., MacGlashan, J., Cushman, F.
+\& Austerweil, J.
+L.
 Showing versus doing: {Teaching} by demonstration.
-In: Lee DD, Sugiyama M, Luxburg UV, Guyon I, Garnett R, editors.
-Advances in neural information processing systems 29.
-{Curran Associates, Inc.}; 2016.
-pp.
-3027--3035.
-}
+in \emph{Advances in neural information processing systems 29} (eds.
+Lee, D.
+D., Sugiyama, M., Luxburg, U.
+V., Guyon, I.
+\& Garnett, R.) 3027--3035 ({Curran Associates, Inc.}, 2016).}
 
@@
 \CSLLeftMargin{12. }
-\CSLRightInline{Sloboda JA.
-Individual differences in music performance.
-Trends in Cognitive Sciences.
-2000;4: 397--403.
-doi:\href{https://doi.org/10.1016/S1364-6613(00)01531-X}{10.1016/S1364-6613(00)01531-X}}
+\CSLRightInline{Sloboda, J. A. \href{https://doi.org/10.1016/S1364-6613(00)01531-X}{Individual differences in music performance}. \emph{Trends in Cognitive Sciences} \textbf{4}, 397--403 (2000).}
 
@@
 \CSLLeftMargin{13. }
-\CSLRightInline{Juslin PN.
-Five {Facets} of {Musical Expression}: {A Psychologist}'s {Perspective} on {Music Performance}.
-Psychology of Music.
-{SAGE Publications Ltd}; 2003;31: 273--302.
-doi:\href{https://doi.org/10.1177/03057356030313003}{10.1177/03057356030313003}}
+\CSLRightInline{Juslin, P.
+N.
+\href{https://doi.org/10.1177/03057356030313003}{Five {Facets} of {Musical Expression}: {A Psychologist}'s {Perspective} on {Music Performance}}.
+\emph{Psychology of Music} \textbf{31}, 273--302 (2003).}
 
@@
 \CSLLeftMargin{14. }
-\CSLRightInline{Bresin R, Battel GU.
-Articulation {Strategies} in {Expressive Piano Performance Analysis} of {Legato}, {Staccato}, and {Repeated Notes} in {Performances} of the {Andante Movement} of {Mozart}'s {Sonata} in {G Major} ({K} 545).
-Journal of New Music Research.
-2000;29: 211--224.
-doi:\href{https://doi.org/10.1076/jnmr.29.3.211.3092}{10.1076/jnmr.29.3.211.3092}}
+\CSLRightInline{Bresin, R.
+\& Battel, G.
+U.
+\href{https://doi.org/10.1076/jnmr.29.3.211.3092}{Articulation {Strategies} in {Expressive Piano Performance Analysis} of {Legato}, {Staccato}, and {Repeated Notes} in {Performances} of the {Andante Movement} of {Mozart}'s {Sonata} in {G Major} ({K} 545)}.
+\emph{Journal of New Music Research} \textbf{29}, 211--224 (2000).}
 
@@
 \CSLLeftMargin{15. }
-\CSLRightInline{Lindström E, Juslin PN, Bresin R, Williamon A.
-{``{Expressivity} comes from within your soul''}: {A} questionnaire study of music students' perspectives on expressivity.
-Research Studies in Music Education.
-{SAGE Publications Ltd}; 2003;20: 23--47.
-doi:\href{https://doi.org/10.1177/1321103X030200010201}{10.1177/1321103X030200010201}}
+\CSLRightInline{Lindström, E., Juslin, P.
+N., Bresin, R.
+\& Williamon, A.
+\href{https://doi.org/10.1177/1321103X030200010201}{{`{Expressivity} comes from within your soul'}: {A} questionnaire study of music students' perspectives on expressivity}.
+\emph{Research Studies in Music Education} \textbf{20}, 23--47 (2003).}
 
@@
 \CSLLeftMargin{16. }
-\CSLRightInline{Okazaki S, Muraoka Y, Osu R.
-Teacher-learner interaction quantifies scaffolding behaviour in imitation learning.
-Sci Rep.
-2019;9: 1--13.
-doi:\href{https://doi.org/10.1038/s41598-019-44049-x}{10.1038/s41598-019-44049-x}}
+\CSLRightInline{Okazaki, S., Muraoka, Y.
+\& Osu, R.
+\href{https://doi.org/10.1038/s41598-019-44049-x}{Teacher-learner interaction quantifies scaffolding behaviour in imitation learning}.
+\emph{Sci Rep} \textbf{9}, 1--13 (2019).}
 
@@
 \CSLLeftMargin{17. }
-\CSLRightInline{Sperber D, Wilson D. Relevance: {Communication} and cognition, 2nd ed. {Malden}: {Blackwell Publishing}; 1995. }
+\CSLRightInline{Sperber, D. \& Wilson, D. \emph{Relevance: {Communication} and cognition, 2nd ed}. ({Blackwell Publishing}, 1995).}
 
@@
 
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+This research was supported by the European Research Council under the European Union's Seventh Frame- work Program (FP7/2007--2013)/ERC Grant agreement no.
+609819, SOMICS, and by ERC Grant agreement no.
+616072, JAXPERTISE.
+We thank Dávid Csűrös for his help with data collection.
+
+\hypertarget{author-contributions}{%
+\section{Author contributions}\label{author-contributions}}
+
+A.T, G.K and N.S developed the study concept and designed the experiments.
+Stimuli and procedure were created by A.T.
+Data collection and analyses were performed by A.T.
+under the supervision of G.K.
+and N.S.
+A.T.
+drafted the manuscript, and N.S.
+and G.K.
+provided critical revisions.
+All authors approved the final version of the manuscript for submission.
+
+\hypertarget{data-availability-statement}{%
+\section{Data availability statement}\label{data-availability-statement}}
+
+All data is available at \url{https://osf.io/8nbjh/}.
+
+\hypertarget{competing-interests}{%
+\section{Competing interests}\label{competing-interests}}
+
+The authors declare no competing interests.
+
 \clearpage
 
-\hypertarget{s1-file.-instructions}{%
-\section{S1 File. Instructions}\label{s1-file.-instructions}}
+\hypertarget{supplementary}{%
+\section{Supplementary Information}\label{supplementary}}
 
-\hypertarget{experiment-1-1}{%
-\subsection{Experiment 1}\label{experiment-1-1}}
+\hypertarget{instructions-experiment-1}{%
+\subsection{Instructions (Experiment 1)}\label{instructions-experiment-1}}
 
@@
 
-\hypertarget{experiment-2-1}{%
-\subsection{Experiment 2}\label{experiment-2-1}}
+\hypertarget{instructions-experiment-2}{%
+\subsection{Instructions (Experiment 2)}\label{instructions-experiment-2}}
 

### 2022-01-20 paper/manuscript.Rmd
update manuscript, supplementary material
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
     institution   : "Corresponding author: Tominaga_Atsuko@phd.ceu.edu"
-
-authornote: |
 
@@
 
-output: papaja::apa6_word
-#output:
-#  papaja::apa6_pdf:
-#    latex_engine: xelatex
+# output: papaja::apa6_word
+output:
+ papaja::apa6_pdf:
+   latex_engine: xelatex
 ---
@@
 # Acknowledgements
-This research was supported by the European Research Council under the European Union’s Seventh Frame- work Program (FP7/2007–2013)/ERC Grant agreement no.
+This research was supported by the European Research Council under the European Union’s Seventh Framework Program (FP7/2007–2013)/ERC Grant agreement no.
 609819, SOMICS, and by ERC Grant agreement no.
@@
 
-In the performing condition, participants were shown the following instruction on a computer monitor monitor: "Now, play what you practised as if you were performing it to an audience.
+In the performing condition, participants were shown the following instruction on a computer monitor: "Now, play what you practised as if you were performing it to an audience.
 *Do your best as a performer to produce the piece according to the notation that you just practised.*"

### 2022-01-20 paper/manuscript.tex
update manuscript, supplementary material
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-This research was supported by the European Research Council under the European Union's Seventh Frame- work Program (FP7/2007--2013)/ERC Grant agreement no.
+This research was supported by the European Research Council under the European Union's Seventh Framework Program (FP7/2007--2013)/ERC Grant agreement no.
 609819, SOMICS, and by ERC Grant agreement no.
@@
 
-In the performing condition, participants were shown the following instruction on a computer monitor monitor: ``Now, play what you practised as if you were performing it to an audience.
+In the performing condition, participants were shown the following instruction on a computer monitor: ``Now, play what you practised as if you were performing it to an audience.
 \emph{Do your best as a performer to produce the piece according to the notation that you just practised.}''

### 2022-01-20 paper/manuscript.Rmd
updated manuscript
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# output: papaja::apa6_word
-output:
- papaja::apa6_pdf:
-   latex_engine: xelatex
+output: papaja::apa6_word
+# output:
+#  papaja::apa6_pdf:
+#    latex_engine: xelatex
 ---
@@
 Staccato requires producing musical notes with shortened duration, keeping them separate from each other.
-Dynamics was notated as either forte and piano.
+Dynamics was notated as either forte or piano.
 Forte indicates that musical notes should be played loudly whereas piano indicates that musical notes should be played softly.
@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplementary Information](#supplementary)*).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplementary material](#supplementary)*).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
@@
 
-ioi_art_1 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ioi_art_1 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
   ylim(150, 260) +
@@
 
-ioi_dyn_1 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ioi_dyn_1 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Dynamics") +
   ylim(150, 260) +
@@
 ```{r plot-ioi-1, out.width = "100%", fig.cap = "\\label{fig:ioi-1}Experiment 1: IOIs (ms) when playing the piece with either articulation (left) or dynamics (right).
+A dashed line represents the tempo given by a metronome.
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
@@
 First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2} A*) prior to the experiment.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[Supplementary Information](#supplementary)*).
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[Supplementary material](#supplementary)*).
 Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute).
@@
 
-ioi_art_2 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+ioi_art_2 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOIs: Articulation") +
   ylim(0.8, 1.1) +
@@
 staccato) showed that there was a significant main effect of Articulation (*F*(`r kor_art_aov$anova_table[2,1]`,`r kor_art_aov$anova_table[2,2]`) = `r sprintf('%.0f', kor_art_aov$anova_table[2,4])`, *p* `r if(kor_art_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[2,5])`) and a significant interaction between Condition and Articulation (*F*(`r kor_art_aov$anova_table[3,1]`,`r kor_art_aov$anova_table[3,2]`) = `r sprintf('%.2f', kor_art_aov$anova_table[3,4])`, *p* `r if(kor_art_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', kor_art_aov$anova_table[3,5])`, *Fig \ref{fig:kor-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KORs were more negative in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r sprintf('%.3f', kor_art_posthoc$p.value[2])`).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KORs were more negative in the teaching condition [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Staccato"]$SD`] than in the performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Staccato"]$SD`] (*p* = `r if(kor_art_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', kor_art_posthoc$p.value[2]))}`).
 For legato there was no significant difference in KORs between the teaching [*M* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "teaching" & Subcomponent == "Legato"]$SD`] and performing condition [*M* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$Mean`, *SD* = `r kor_art_all[Condition == "performing" & Subcomponent == "Legato"]$SD`] (*p* = `r sprintf('%.2f', kor_art_posthoc$p.value[1])`).
@@
 
-ioi_dyn_2 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+ioi_dyn_2 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOIs: Dynamics") +
   ylim(0.8, 1.1) +
@@
 ```{r plot-ioi-2, out.width = "100%", fig.cap = "\\label{fig:ioi-2}Experiment 2: Normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
+A dashed line represents the tempo given by a metronome.
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
@@
 
-# Supplementary Information{#supplementary}
+# Supplementary material{#supplementary}
 ## Instructions (Experiment 1)

### 2022-01-20 paper/manuscript.tex
updated manuscript
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary Information}}).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}}).
 Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
@@
 \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-1-1} \caption{\label{fig:ioi-1}Experiment 1: IOIs (ms) when playing the piece with either articulation (left) or dynamics (right).
+A dashed line represents the tempo given by a metronome.
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
@@
 First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Fig \ref{fig:stimuli-2} A}) prior to the experiment.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplementary}{Supplementary Information}}).
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}}).
 Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute).
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant main effect of Articulation (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Articulation (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01, \emph{Fig \ref{fig:kor-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KORs were more negative in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = 0.000).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KORs were more negative in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = \textless{} 0.001).
 For legato there was no significant difference in KORs between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.63).
@@
 \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Experiment 2: Normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
+A dashed line represents the tempo given by a metronome.
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
@@
 \hypertarget{supplementary}{%
-\section{Supplementary Information}\label{supplementary}}
+\section{Supplementary material}\label{supplementary}}
 

### 2022-06-15 experiment-1/analysis/stats/variability.Rmd
start variabilityy analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Variability analysis"
output: html_notebook
---

### 2022-06-17 experiment-1/analysis/stats/variability.Rmd
upload variability analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 title: "Variability analysis"
-output: html_notebook
+output:
+  html_notebook: default
+editor_options: 
+  chunk_output_type: inline
 ---
+
+Description: This is a summary of variability analysis. Added after the first review from the journal.
+
+- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
+
+```{r setup, include = FALSE}
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# statistics
+if (!require("afex")) {install.packages("afex"); require("afex")}
+if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
+
+filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
+dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".")
+
+# SubNr as a factor
+dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
+
+# Include only articulation
+dt_ioi_art <- dt_ioi[Skill == "articulation"]
+# per participant
+ioi_art_seq_each <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+
+# Include only dynamics
+dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
+# per participant
+ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+```
+
+## Sequence Plots
+- Error bars show the standard error.
+- Dotted lines show the target tempo (IOI = 188ms).
+- Twodashed lines show transition points (i.e., from legato to staccato, from staccato to legato, from forte to piano, from piano to forte).
+
+### Overall
+
+```{r, echo = FALSE, fig.width = 6, fig.height = 1.2}
+# Plot the whole sequence
+ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+  geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
+  geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points
+  scale_x_continuous(breaks = seq(1,66,1))
+
+ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+  geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
+  geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points
+  scale_x_continuous(breaks = seq(1,66,1))
+```
+
+<!-- ### Each participant -->
+<!-- ```{r, echo = FALSE, fig.width = 5, fig.height = 5} -->
+<!-- # Plot the whole sequence for each participant -->
+<!-- ggplot(ioi_art_seq_each, aes(x = Interval, y = Mean, color = SubNr)) + -->
+<!--   geom_line() + geom_point() + facet_grid(Condition ~ .) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dotted") + # target tempo -->
+<!--   geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points -->
+<!--   scale_x_discrete(breaks = seq(1,66,3)) + labs(x = "Interval", y = "IOIs (ms)", title = "IOI: Articulation") + theme_pubr() + theme(legend.position = "right") -->
+
+<!-- ioi_dyn_seq <- ioi_dyn_seq_each[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Interval)] -->
+<!-- ggplot(ioi_dyn_seq_each, aes(x = Interval, y = Mean, color = SubNr)) + -->
+<!--   geom_line() + geom_point() + facet_grid(Condition ~ .) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dotted") + # target tempo -->
+<!--   geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points -->
+<!--   scale_x_discrete(breaks = seq(1,66,3)) + labs(x = "Interval", y = "IOIs (ms)", title = "IOI: Dynamics") + theme_pubr() + theme(legend.position = "right") -->
+<!-- ``` -->
+
+## Coefficient of Variation
+
+```{r, echo = FALSE}
+ioi_cv_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+ioi_cv_subject <- ioi_cv_trial[, .(N = .N, MeanCV = mean(CV), SD = sd(CV)), by = .(SubNr, Condition, Skill)]
+ioi_cv_subject <- ioi_cv_subject[order(SubNr, Condition, Skill)] # ordering
+
+ggplot(ioi_cv_subject, aes(x = Condition, y = MeanCV, color = Condition)) +
+  geom_boxplot() +
+  geom_point(aes(color = Condition)) +
+  geom_line(aes(group = SubNr), size = 0.4, color = "black", alpha = 0.2) +
+  facet_grid(. ~ Skill) +
+  theme_pubr()
+```
+
+### Stats
+- With Participant 2 (outside +/- 3SD)
+
+```{r, echo = FALSE}
+# descriptive stats
+ioi_cv_subject[, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
+
+aov_ez(id = "SubNr", dv = "MeanCV",
+       within = c("Condition", "Skill"),
+       data = ioi_cv_subject)
+```
+
+- Without Participant 2 (outside +/- 3SD)
+
+```{r, echo = FALSE}
+# descriptive stats
+ioi_cv_subject[SubNr != 2, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
+
+aov_ez(id = "SubNr", dv = "MeanCV",
+       within = c("Condition", "Skill"),
+       data = ioi_cv_subject[SubNr != 2])
+```
+
+## IOIs at transition points
+
+```{r, echo = FALSE}
+ioi_tra <- dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
+ioi_tra <- ioi_tra[order(SubNr, Condition, Skill)] # ordering
+
+ggplot(ioi_tra, aes(x = Condition, y = Mean, color = Condition)) +
+  geom_boxplot() +
+  geom_point(aes(color = Condition)) +
+  geom_line(aes(group = SubNr), size = 0.4, color = "black", alpha = 0.2) +
+  facet_grid(. ~ Skill) +
+  theme_pubr()
+```
+
+```{r, echo = FALSE}
+# descriptive stats
+ioi_tra[, .(.N, MeanCV = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill)]
+
+aov_ez(id = "SubNr", dv = "Mean",
+       within = c("Condition", "Skill"),
+       data = ioi_tra)
+```
+
+## IOIs at transition points vs. others
+Caution: the number of points included in each transition group is considerably different
+
+```{r, echo = FALSE}
+dt_ioi$Transition <- "No"
+dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF"]$Transition <- "Yes"
+
+# transition points vs. others
+ioi_tra_others <- dt_ioi[, .(.N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Transition)]
+ioi_tra_others <- ioi_tra_others[order(SubNr, Condition, Skill)] # ordering
+
+# descriptive stats
+ioi_tra_others[, .(.N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Transition)]
+
+ggplot(ioi_tra_others, aes(x = Condition, y = Mean, color = Condition)) +
+  geom_boxplot() +
+  geom_point(aes(color = Condition)) +
+  geom_line(aes(group = SubNr), size = 0.4, color = "black", alpha = 0.2) +
+  facet_grid(Skill ~ Transition) +
+  theme_pubr()
+```
+
+```{r, echo = FALSE}
+aov_ez(id = "SubNr", dv = "Mean",
+       within = c("Condition", "Skill", "Transition"),
+       data = ioi_tra_others)
+```
+
+add some postdoc comparisons...

### 2022-06-21 experiment-2/analysis/stats/variability.Rmd
upload variability for exp2 / export R scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-06-17 experiment-1/analysis/stats/variability.Rmd
@@
 # Plot the whole sequence
-ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
-  geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points
+ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
+  geom_vline(xintercept = c(5, 17, 47, 8, 20, 39), linetype = "twodash") + # transition points
   scale_x_continuous(breaks = seq(1,66,1))
 
-ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
-  geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points
+ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Dynamics") +
+  geom_vline(xintercept = c(5, 17, 47, 8, 20, 39), linetype = "twodash") + # transition points
   scale_x_continuous(breaks = seq(1,66,1))
@@
 <!--   geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points -->
-<!--   scale_x_discrete(breaks = seq(1,66,3)) + labs(x = "Interval", y = "IOIs (ms)", title = "IOI: Articulation") + theme_pubr() + theme(legend.position = "right") -->
+<!--   scale_x_discrete(breaks = seq(1,66,3)) + labs(x = "Interval", y = "IOIs (ms)", title = "IOIs: Articulation") + theme_pubr() + theme(legend.position = "right") -->
 
@@
 <!--   geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points -->
-<!--   scale_x_discrete(breaks = seq(1,66,3)) + labs(x = "Interval", y = "IOIs (ms)", title = "IOI: Dynamics") + theme_pubr() + theme(legend.position = "right") -->
+<!--   scale_x_discrete(breaks = seq(1,66,3)) + labs(x = "Interval", y = "IOIs (ms)", title = "IOIs: Dynamics") + theme_pubr() + theme(legend.position = "right") -->
 <!-- ``` -->
@@
   geom_boxplot() +
+  geom_line(aes(group = SubNr), size = 0.4, color = "gray") +
   geom_point(aes(color = Condition)) +
-  geom_line(aes(group = SubNr), size = 0.4, color = "black", alpha = 0.2) +
   facet_grid(. ~ Skill) +
@@
 ### Stats
-- With Participant 2 (outside +/- 3SD)
+- With Participant 15 (outside +/- 3SD)
 
@@
 
-- Without Participant 2 (outside +/- 3SD)
+- Without Participant 15 (outside +/- 3SD)
 
@@
 # descriptive stats
-ioi_cv_subject[SubNr != 2, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
+ioi_cv_subject[SubNr != 15, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
 
@@
        within = c("Condition", "Skill"),
-       data = ioi_cv_subject[SubNr != 2])
+       data = ioi_cv_subject[SubNr != 15])
 ```
@@
   geom_boxplot() +
+  geom_line(aes(group = SubNr), size = 0.4, color = "gray") +
   geom_point(aes(color = Condition)) +
-  geom_line(aes(group = SubNr), size = 0.4, color = "black", alpha = 0.2) +
   facet_grid(Skill ~ Transition) +
@@
 
-add some postdoc comparisons...
+```{r export, include = FALSE}
+knitr::purl("variability.Rmd")
+```

### 2022-06-21 experiment-1/analysis/stats/variability.R

@@ -1,16 +1,4 @@
----
-title: "Variability analysis"
-output:
-  html_notebook: default
-editor_options: 
-  chunk_output_type: inline
----
-
-Description: This is a summary of variability analysis. Added after the first review from the journal.
-
-- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
-
-```{r setup, include = FALSE}
+## ----setup, include = FALSE--------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -35,16 +23,9 @@ ioi_art_seq_each <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by =
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 # per participant
 ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
-```
 
-## Sequence Plots
-- Error bars show the standard error.
-- Dotted lines show the target tempo (IOI = 188ms).
-- Twodashed lines show transition points (i.e., from legato to staccato, from staccato to legato, from forte to piano, from piano to forte).
 
-### Overall
-
-```{r, echo = FALSE, fig.width = 6, fig.height = 1.2}
+## ---- echo = FALSE, fig.width = 6, fig.height = 1.2--------
 # Plot the whole sequence
 ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
   geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
@@ -55,29 +36,9 @@ ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "C
   geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
   geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points
   scale_x_continuous(breaks = seq(1,66,1))
-```
-
-<!-- ### Each participant -->
[…]

### 2022-06-21 experiment-2/analysis/stats/variability.R

@@ -1,16 +1,4 @@
----
-title: "Variability analysis"
-output:
-  html_notebook: default
-editor_options: 
-  chunk_output_type: inline
----
-
-Description: This is a summary of variability analysis. Added after the first review from the journal.
-
-- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
-
-```{r setup, include = FALSE}
+## ----setup, include = FALSE--------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -35,49 +23,20 @@ ioi_art_seq_each <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by =
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 # per participant
 ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
-```
 
-## Sequence Plots
-- Error bars show the standard error.
-- Dotted lines show the target tempo (IOI = 188ms).
-- Twodashed lines show transition points (i.e., from legato to staccato, from staccato to legato, from forte to piano, from piano to forte).
 
-### Overall
-
-```{r, echo = FALSE, fig.width = 6, fig.height = 1.2}
+## ---- echo = FALSE, fig.width = 6, fig.height = 1.2--------
 # Plot the whole sequence
 ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
-  geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points
+  geom_vline(xintercept = c(5, 17, 47, 8, 20, 39), linetype = "twodash") + # transition points
   scale_x_continuous(breaks = seq(1,66,1))
 
[…]

### 2022-06-23 experiment-1/analysis/demographics/questionnaire.Rmd
export r chunks as R scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update demographic data
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-12-02 script/R/demographics/questionnaire.Rmd
@@
 colnames(dt)[7] <- "Handedness"
+colnames(dt)[18] <- "TeachingMusic"
 
@@
 dt$PianoTotalPractice <- c(14, 20, 27, 21, 10, 16, 9, 13, 25, 14, 13, 11, 6, 4, 6, 8, 21, 15, 18, 7, 11, 12, 12, 18, 12, 5, 5, 10, 16, 10, 8, 22, 10, 5, 4, 14)
+dt$TeachingPianoYears <- c(NA, 4, NA, 12, NA, 3, NA, NA, 3, NA, NA, NA, NA, NA, 5, NA, 5, NA, 0.17, NA, NA, NA, NA, 0.17, NA, NA, NA, NA, 2, NA, NA, 0.5, NA, NA, NA, NA)
+dt$TeachingPiano <- "No"
+dt[!is.na(TeachingPianoYears)]$TeachingPiano <- "Yes"
 
@@
 ## Descriptive stats for included participants
+
 - Number of included participants: `r nrow(dt_included)` (mean age: `r round(mean(dt_included$Age), 2)`, sd: `r round(sd(dt_included$Age), 2)`)
@@
 - Handedness: Right (`r nrow(dt_included[Handedness == "Right"])`), Left (`r nrow(dt_included[Handedness == "Left"])`), Both (`r nrow(dt_included[Handedness == "Both"])`)
+- Teaching experience (any instrument): Yes (`r nrow(dt_included[TeachingMusic == "Yes"])`), No (`r nrow(dt_included[TeachingMusic == "No"])`)
+- Teaching experience (piano): Yes (`r nrow(dt_included[TeachingPiano == "Yes"])`, mean years of teaching: `r round( mean(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`, sd: `r round(sd(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`), No (`r nrow(dt_included[TeachingPiano == "No"])`)
 
 ## Some descriptive questions
+
 1. Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?
@@
 ```
+
+```{r export, include = FALSE}
+knitr::purl("questionnaire.Rmd")
+```

### 2022-06-23 experiment-2/analysis/demographics/questionnaire.Rmd
export Rmd as R
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update demographic data
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 colnames(dt_2)[18] <- "PianoTotalPractice"
+colnames(dt_2)[28] <- "TeachingMusic"
 
@@
 dt_2$PianoTotalPractice <- c(13, 15, 15, 16, 9, 10, 24, 8, 19, 6, 12, 10, 21, 23, 15, 22, 22, 24, 17, 15, 12)
+dt_2$TeachingPianoYears <- c(NA, 4, NA, 2, NA, NA, NA, NA, 5, NA, NA, NA, 0.5, 9, 11, 2, NA, 8, 2, 1, 1)
+dt_2$TeachingPiano <- "No"
+dt_2[!is.na(TeachingPianoYears)]$TeachingPiano <- "Yes"
 
@@
 - Handedness: Right (`r nrow(dt_2_included[Handedness == "Right"])`), Left (`r nrow(dt_2_included[Handedness == "Left"])`), Both (`r nrow(dt_2_included[Handedness == "Both"])`)
+- Teaching experience (any instrument): Yes (`r nrow(dt_2_included[TeachingMusic == "Yes"])`), No (`r nrow(dt_2_included[TeachingMusic == "No"])`)
+- Teaching experience (piano): Yes (`r nrow(dt_2_included[TeachingPiano == "Yes"])`, mean years of teaching: `r round( mean(dt_2_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`, sd: `r round(sd(dt_2_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`), No (`r nrow(dt_2_included[TeachingPiano == "No"])`)
 

### 2022-06-23 experiment-1/analysis/demographics/questionnaire.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE--------------------------------------------------------
+## ----setup, include = FALSE----------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -6,11 +6,11 @@ if (!require("data.table")) {install.packages("data.table"); require("data.table
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----file, include = FALSE---------------------------------------------------------
+## ----file, include = FALSE-----------------------------------
 filename_q = "./questionnaire.csv"
 
 
-## ----extract, echo = FALSE---------------------------------------------------------
+## ----extract, echo = FALSE-----------------------------------
 # read csv files
 dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
 
@@ -19,12 +19,16 @@ colnames(dt)[2] <- "SubNr"
 colnames(dt)[3] <- "Age"
 colnames(dt)[4] <- "Gender"
 colnames(dt)[7] <- "Handedness"
+colnames(dt)[18] <- "TeachingMusic"
 
 # change some characteristics
 dt$Age <- as.numeric(dt$Age)
 
 # extract the total number of years of training in piano (input manually based on the original csv file)
 dt$PianoTotalPractice <- c(14, 20, 27, 21, 10, 16, 9, 13, 25, 14, 13, 11, 6, 4, 6, 8, 21, 15, 18, 7, 11, 12, 12, 18, 12, 5, 5, 10, 16, 10, 8, 22, 10, 5, 4, 14)
+dt$TeachingPianoYears <- c(NA, 4, NA, 12, NA, 3, NA, NA, 3, NA, NA, NA, NA, NA, 5, NA, 5, NA, 0.17, NA, NA, NA, NA, 0.17, NA, NA, NA, NA, 2, NA, NA, 0.5, NA, NA, NA, NA)
+dt$TeachingPiano <- "No"
+dt[!is.na(TeachingPianoYears)]$TeachingPiano <- "Yes"
 
 # exclude participants
 dt_included <- dt[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16]
@@ -32,14 +36,14 @@ dt_included <- dt[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 
 print(dt_included)
 
 
[…]

### 2022-06-23 experiment-2/analysis/demographics/questionnaire.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE----------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -6,11 +6,11 @@ if (!require("data.table")) {install.packages("data.table"); require("data.table
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE-----------------------------------
 filename_q = "./questionnaire.csv"
 
 
-## ----extract, echo = FALSE----------------------------------------------------
+## ----extract, echo = FALSE-----------------------------------
 # read csv files
 dt_2 <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
 dt_2 <- dt_2[!21]
@@ -21,9 +21,13 @@ colnames(dt_2)[3] <- "Age"
 colnames(dt_2)[4] <- "Gender"
 colnames(dt_2)[7] <- "Handedness"
 colnames(dt_2)[18] <- "PianoTotalPractice"
+colnames(dt_2)[28] <- "TeachingMusic"
 
 # extract the total number of years of training in piano (input manually based on the original csv file)
 dt_2$PianoTotalPractice <- c(13, 15, 15, 16, 9, 10, 24, 8, 19, 6, 12, 10, 21, 23, 15, 22, 22, 24, 17, 15, 12)
+dt_2$TeachingPianoYears <- c(NA, 4, NA, 2, NA, NA, NA, NA, 5, NA, NA, NA, 0.5, 9, 11, 2, NA, 8, 2, 1, 1)
+dt_2$TeachingPiano <- "No"
+dt_2[!is.na(TeachingPianoYears)]$TeachingPiano <- "Yes"
 
 # change some characteristics
 dt_2$Age <- as.numeric(dt_2$Age)
@@ -35,14 +39,14 @@ dt_2_included <- dt_2[SubNr != 3]
 print(dt_2_included)
 
 
-## ----1, echo = FALSE----------------------------------------------------------
+## ----1, echo = FALSE-----------------------------------------
 data.table("Answer" = dt_2_included$`Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?`)
 
 
-## ----2, echo = FALSE----------------------------------------------------------
+## ----2, echo = FALSE-----------------------------------------
[…]

### 2022-06-24 paper/manuscript.Rmd
implement variability analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
add teaching experiences for demographic info
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-output: papaja::apa6_word
-# output:
-#  papaja::apa6_pdf:
-#    latex_engine: xelatex
+# output: papaja::apa6_word
+output:
+  papaja::apa6_pdf:
+    latex_engine: xelatex
 ---
@@
 They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`).
+`r nrow(dt_included[TeachingPiano == "Yes"])` participants had an experience in teaching the piano (*M* = `r round( mean(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)` years, *SD* = `r round( sd(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`).
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
@@
 They had `r round(mean(dt_2_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_2_included$PianoTotalPractice), 2)`).
+`r nrow(dt_2_included[TeachingPiano == "Yes"])` participants had an experience in teaching the piano (*M* = `r round( mean(dt_2_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)` years, *SD* = `r round( sd(dt_2_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`).
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
@@
 # Supplementary material{#supplementary}
-## Instructions (Experiment 1)
+# Instructions
+
+## Experiment 1
 
@@
 
-## Instructions (Experiment 2)
+## Experiment 2
 
@@
 Do your best as a performer to produce the piece according to the notation that you just practised.*"
+
+\clearpage
+
+# Variability
+
+## Data analysis
+
+## Results (Experiment 1)
+```{r variability-1, include = FALSE}
+source(here("experiment-1/analysis/stats", "variability.R"), chdir = TRUE)
+```
+
+## Results (Experiment 2)
+```{r variability-2, include = FALSE}
+source(here("experiment-2/analysis/stats", "variability.R"), chdir = TRUE)
+```

### 2022-06-26 experiment-1/analysis/stats/dynamics.Rmd
add some additional analysis for nomalised KV diff
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 - [Sequence plot](#sequence_art)
+
+- Additional analysis based on reviewers' comments
 
@@
 
+# Additional analysis
+- Normalise KV difference by the range of KV value between forte and piano for each participant.
+
+```{r, echo = FALSE}
+# normalise diff by KV range
+vel_diff_dyn$NormalisedDiff <- 0
+for (i in unique(vel_diff_dyn$SubNr)){
+  vel_diff_dyn[SubNr == i & Condition == "teaching"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
+  vel_diff_dyn[SubNr == i & Condition == "performing"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
+}
+vel_diff_dyn
+
+# group
+vel_diff_dyn_all_2 <- vel_diff_dyn[, .(N = .N, NormalisedMean = mean(NormalisedDiff), SD = sd(NormalisedDiff), SEM = sd(NormalisedDiff)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
+vel_diff_dyn_all_2
+```
+
+## Plot
+```{r, echo = FALSE}
+vel_diff_dyn_plot_2 <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$NormalisedDiff, performing = vel_diff_dyn[Condition == "performing"]$NormalisedDiff, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)
+
+ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
+```
+
+## Stats
+```{r, echo = FALSE}
+# stats
+vel_diff_dyn_aov_2 <- aov_car(NormalisedDiff ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
+summary(vel_diff_dyn_aov_2)
+```
+
 ```{r export, include = FALSE}

### 2022-06-26 experiment-1/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE--------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE---------------------------------
+## ----file, include = FALSE----------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,49 +31,49 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE---------------------------------
+## ----vel-dyn, echo = FALSE----------------------------------
 # For each individual
 vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE----------------------------
+## ----vel-dyn-box,  echo = FALSE-----------------------------
 ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
   facet_grid(Subcomponent ~ .)
 
 
[…]

### 2022-06-27 experiment-1/analysis/stats/dynamics.Rmd
update additional analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-06-20 experiment-1/analysis/stats/dynamics.Rmd
@@
 
-- [Sequence plot](#sequence_dyn)
-
 ## [Articulation](#vel_articulation)
@@
 
-- [Sequence plot](#sequence_art)
+### [Sequence plot](#sequence)
+
+- [Additional analysis based on reviewers' comments](#additional)
 
@@
 # For each individual
-vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
+vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+vel_dyn <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
@@
 ```{r vel-dyn-box,  echo = FALSE}
-ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
+ggboxplot(vel_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
   facet_grid(Subcomponent ~ .)
@@
 ### Plot
-#### Scatter plot
-- Error bars show SEM of the mean.
-
-```{r vel-dyn-all-scatter, echo = FALSE}
-ggerrorplot(vel_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics")
-```
-
 #### Box plot
@@
 # For each individual
-vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
 vel_dyn_trial
@@
 # For each individual
-vel_diff_dyn <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
+vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+vel_diff_dyn <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_diff_dyn
@@
 ```{r vel-diff-dyn-box,  echo = FALSE}
-ggboxplot(dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Dynamics") +
+ggboxplot(vel_diff_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Dynamics") +
   facet_grid(Subcomponent ~ .)
@@
 ### Plot
-#### Scatter plot
-- Error bars show SEM of the mean.
-
-```{r vel-diff-dyn-all-scatter, echo = FALSE}
-ggerrorplot(vel_diff_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics")
-```
-
 #### Box plot
@@
 # For each individual
-vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent,TrialNr)]
 vel_diff_dyn_trial
@@
 ### Plot
-#### Scatter plot
-- Error bars show SEM of the mean.
-
-```{r vel-art-all-scatter, echo = FALSE}
-ggerrorplot(vel_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation")
-```
-
 #### Box plot
@@
 ### Plot
-#### Scatter plot
-- Error bars show SEM of the mean.
-
-```{r vel-diff_art-all-scatter, echo = FALSE}
-ggerrorplot(vel_diff_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Articulation")
-```
-
 #### Box plot
@@
 
+# Sequence plot{#sequence}
 ## Dynamics{#sequence_dyn}
@@
 
+# Additional analysis{#additional}
+- Normalise KV difference by the range of KV value between forte and piano for each participant.
+
+```{r, echo = FALSE}
+# normalise diff by KV range
+vel_diff_dyn_trial$NormalisedDiff <- 0
+for (i in unique(vel_diff_dyn_trial$SubNr)){
+  for (m in unique(vel_diff_dyn_trial$TrialNr)){
+
+    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))
+
+    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))
+
+  }
+}
+
+# each participant
+vel_diff_dyn_normalised <- vel_diff_dyn_trial[, .(N = .N, NormalisedMean = mean(NormalisedDiff), SD = sd(NormalisedDiff), SEM = sd(NormalisedDiff)/sqrt(.N)), by = .(SubNr, Condition, Skill, Subcomponent)]
+
+# group
+vel_diff_dyn_normalised_all <- vel_diff_dyn_normalised[, .(N = .N, Mean = mean(NormalisedMean), SD = sd(NormalisedMean), SEM = sd(NormalisedMean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
+vel_diff_dyn_normalised_all
+```
+
+## Plot
+```{r, echo = FALSE}
+vel_diff_dyn_plot_2 <- data.table(subject = vel_diff_dyn_normalised[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn_normalised[Condition == "teaching"]$NormalisedMean, performing = vel_diff_dyn_normalised[Condition == "performing"]$NormalisedMean, subcomponent = vel_diff_dyn_normalised[Condition == "teaching"]$Subcomponent)
+
+ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
+```
+
+## Stats
+```{r, echo = FALSE}
+# stats
+vel_diff_dyn_aov_2 <- aov_car(NormalisedMean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn_normalised)
+summary(vel_diff_dyn_aov_2)
+```
+
+```{r, echo = FALSE}
+vel_diff_dyn_posthoc_2 <- test(pairs(emmeans(vel_diff_dyn_aov_2,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+vel_diff_dyn_posthoc_2
+```
+
 ```{r export, include = FALSE}

### 2022-06-27 experiment-1/analysis/stats/dynamics.R

@@ -33,12 +33,14 @@ dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 ## ----vel-dyn, echo = FALSE----------------------------------
 # For each individual
-vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
+vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+vel_dyn <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
 ## ----vel-dyn-box,  echo = FALSE-----------------------------
-ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
+ggboxplot(vel_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
   facet_grid(Subcomponent ~ .)
 
 
@@ -46,10 +48,6 @@ ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N
 # Group mean
 vel_dyn_all <- vel_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
 vel_dyn_all
-
-
-## ----vel-dyn-all-scatter, echo = FALSE----------------------
-ggerrorplot(vel_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics")
 
 
 ## ----vel-dyn-all-box, echo = FALSE--------------------------
[…]

### 2022-06-27 experiment-2/analysis/stats/dynamics.Rmd
update additional analysis (normalised per trial / participant)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-06-20 experiment-2/analysis/stats/dynamics.Rmd
@@
 ```{r vel-dyn-box,  echo = FALSE}
-ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(vel_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") + facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r vel-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-ggline(dt_vel_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
+ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-ggline(dt_vel_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
+ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-dyn-trial-line-ftop, echo = FALSE, fig.height = 4}
-ggline(dt_vel_diff_dyn[Subcomponent == "FtoP", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
+ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-dyn-trial-line-ptof, echo = FALSE, fig.height = 4}
-ggline(dt_vel_diff_dyn[Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
+ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 # For each individual
-vel_art <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
+vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+vel_art <- vel_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_art
@@
 ```{r vel-art-box,  echo = FALSE}
-ggboxplot(dt_vel_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") +
+ggboxplot(vel_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") +
   facet_grid(Subcomponent ~ .)
@@
 # For each individual
-vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
 vel_art_trial
@@
 ```{r vel-art-trial-line-legato, echo = FALSE, fig.height = 4}
-ggline(dt_vel_art[Subcomponent == "Legato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
+ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(dt_vel_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
+ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-artt-rial-line-ltos, echo = FALSE, fig.height = 4}
-ggline(dt_vel_diff_art[Subcomponent == "LtoS", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
+ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r vel-diff-art-trial-line-stol, echo = FALSE, fig.height = 4}
-ggline(dt_vel_diff_art[Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
+ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 # Group mean
-vel_diff_art_trial_all <- vel_diff_art_trial[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
+vel_diff_art_trial_all <- vel_diff_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
 vel_diff_art_trial_all
@@
 
+# Additional analysis{#additional}
+- Normalise KV difference by the range of KV value between forte and piano for each participant.
+
+## Normalised KV difference per trial
+
+```{r, echo = FALSE, message = FALSE}
+# normalise diff by KV range
+vel_diff_dyn_trial$NormalisedDiff <- 0
+for (i in unique(vel_diff_dyn_trial$SubNr)){
+  for (m in unique(vel_diff_dyn_trial[SubNr == i]$TrialNr)){
+    # teaching
+    # FtoP
+    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))
+    # PtoF
+    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))
+
+    # performing
+    # FtoP
+    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))
+    # PtoF
+    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))
+
+    # error messages for non-existing trials but it's not reflected on vel_diff_dyn_trial anyway
+  }
+}
+
+# each participant
+vel_diff_dyn_normalised <- vel_diff_dyn_trial[, .(N = .N, NormalisedMean = mean(NormalisedDiff), SD = sd(NormalisedDiff), SEM = sd(NormalisedDiff)/sqrt(.N)), by = .(SubNr, Condition, Skill, Subcomponent)]
+
+# group
+vel_diff_dyn_normalised_all <- vel_diff_dyn_normalised[, .(N = .N, Mean = mean(NormalisedMean), SD = sd(NormalisedMean), SEM = sd(NormalisedMean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
+vel_diff_dyn_normalised_all
+```
+
+### Plot
+```{r, echo = FALSE}
+vel_diff_dyn_plot_2 <- data.table(subject = vel_diff_dyn_normalised[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn_normalised[Condition == "teaching"]$NormalisedMean, performing = vel_diff_dyn_normalised[Condition == "performing"]$NormalisedMean, subcomponent = vel_diff_dyn_normalised[Condition == "teaching"]$Subcomponent)
+
+ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
+```
+
+### Stats
+```{r, echo = FALSE}
+# stats
+vel_diff_dyn_aov_3 <- aov_car(NormalisedMean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn_normalised)
+summary(vel_diff_dyn_aov_3)
+```
+
+## Normalised KV difference per participant
+```{r, echo = FALSE}
+vel_diff_dyn$NormalisedDiff <- 0
+for (i in unique(vel_diff_dyn_trial$SubNr)){
+  # teaching
+  # FtoP
+  vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
+  # PtoF
+  vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
+  # performing
+  # FtoP
+  vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
+  # PtoF
+  vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
+}
+
+# group
+vel_diff_dyn_normalised_all_2 <- vel_diff_dyn[, .(N = .N, Mean = mean(NormalisedDiff), SD = sd(NormalisedDiff), SEM = sd(NormalisedDiff)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
+vel_diff_dyn_normalised_all_2
+```
+
+### Plot
+```{r, echo = FALSE}
+vel_diff_dyn_plot_3 <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$NormalisedDiff, performing = vel_diff_dyn[Condition == "performing"]$NormalisedDiff, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)
+
+ggpaired(vel_diff_dyn_plot_3, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
+```
+
+### Stats
+```{r, echo = FALSE}
+# stats
+vel_diff_dyn_aov_4 <- aov_car(NormalisedDiff ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
+summary(vel_diff_dyn_aov_4)
+```
+
 ```{r export, include = FALSE}

### 2022-06-27 experiment-2/analysis/stats/dynamics.R

@@ -33,23 +33,20 @@ dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 ## ----vel-dyn, echo = FALSE---------------------------------
 # For each individual
-vel_dyn <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
+vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+vel_dyn <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
 ## ----vel-dyn-box,  echo = FALSE----------------------------
-ggboxplot(dt_vel_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(vel_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") + facet_grid(Subcomponent ~ .)
 
 
 ## ----vel-dyn-all, echo = FALSE-----------------------------
 # Group mean
 vel_dyn_all <- vel_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
 vel_dyn_all
-
-
-## ----vel-dyn-all-scatter, echo = FALSE---------------------
-ggerrorplot(vel_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics")
 
 
 ## ----vel-dyn-all-box, echo = FALSE-------------------------
@@ -75,17 +72,16 @@ vel_dyn_posthoc
[…]

### 2022-06-27 experiment-2/analysis/stats/dynamics.Rmd
add warking = FALSE
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-```{r, echo = FALSE, message = FALSE}
+```{r, echo = FALSE, warning = FALSE}
 # normalise diff by KV range

### 2022-06-27 experiment-2/analysis/stats/dynamics.R

@@ -339,7 +339,7 @@ ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position =
   scale_x_continuous(breaks = seq(1,71,1))
 
 
-## ---- echo = FALSE, message = FALSE------------------------
+## ---- echo = FALSE, warning = FALSE------------------------
 # normalise diff by KV range
 vel_diff_dyn_trial$NormalisedDiff <- 0
 for (i in unique(vel_diff_dyn_trial$SubNr)){

### 2022-06-27 experiment-1/analysis/stats/dynamics.Rmd
update additional analysis (per trial/participant)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r vel-art-trial-line-legato, echo = FALSE, fig.height = 4}
-ggline(dt_vel_art[Subcomponent == "Legato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(dt_vel_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 # For each individual
-vel_diff_art <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
+vel_diff_art_trial <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+vel_diff_art <- vel_diff_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_diff_art
@@
 ```{r vel-diff-art-box,  echo = FALSE}
-ggboxplot(dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
+ggboxplot(vel_diff_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
   facet_grid(Subcomponent ~ .)
@@
 # For each individual
-vel_diff_art_trial <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
 vel_diff_art_trial
@@
 ```{r vel-diff-art-trial-all-line-ltos, echo = FALSE}
-ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
-  scale_x_continuous(breaks=seq(1,8,1))
+ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
 ```
@@
 ```{r vel-diff-art-trial-all-line-stol, echo = FALSE}
-ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
-  scale_x_continuous(breaks=seq(1,8,1))
+ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") + scale_x_continuous(breaks=seq(1,8,1))
 ```
@@
 
-```{r, echo = FALSE}
+## Normalised KV difference per trial
+
+```{r, echo = FALSE, warning = FALSE}
 # normalise diff by KV range
@@
 for (i in unique(vel_diff_dyn_trial$SubNr)){
-  for (m in unique(vel_diff_dyn_trial$TrialNr)){
+  for (m in unique(vel_diff_dyn_trial[SubNr == i]$TrialNr)){
+    # teaching
+    # FtoP
+    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))
+    # PtoF
+    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))
 
-    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))
+    # performing
+    # FtoP
+    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))
+    # PtoF
+    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))
 
-    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))
-
+    # error messages for non-existing trials but it's not reflected on vel_diff_dyn_trial anyway
   }
@@
 
-## Plot
+### Plot
 ```{r, echo = FALSE}
@@
 
-## Stats
+### Stats
 ```{r, echo = FALSE}
 # stats
-vel_diff_dyn_aov_2 <- aov_car(NormalisedMean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn_normalised)
-summary(vel_diff_dyn_aov_2)
+vel_diff_dyn_aov_3 <- aov_car(NormalisedMean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn_normalised)
+summary(vel_diff_dyn_aov_3)
+```
+#### Post-hoc (emmeans)
+```{r, echo = FALSE}
+vel_diff_dyn_posthoc_2 <- test(pairs(emmeans(vel_diff_dyn_aov_3,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+vel_diff_dyn_posthoc_2
 ```
 
+## Normalised KV difference per participant
 ```{r, echo = FALSE}
-vel_diff_dyn_posthoc_2 <- test(pairs(emmeans(vel_diff_dyn_aov_2,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
-vel_diff_dyn_posthoc_2
+vel_diff_dyn$NormalisedDiff <- 0
+for (i in unique(vel_diff_dyn_trial$SubNr)){
+  # teaching
+  # FtoP
+  vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
+  # PtoF
+  vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
+  # performing
+  # FtoP
+  vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
+  # PtoF
+  vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
+}
+
+# group
+vel_diff_dyn_normalised_all_2 <- vel_diff_dyn[, .(N = .N, Mean = mean(NormalisedDiff), SD = sd(NormalisedDiff), SEM = sd(NormalisedDiff)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
+vel_diff_dyn_normalised_all_2
+```
+
+### Plot
+```{r, echo = FALSE}
+vel_diff_dyn_plot_3 <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$NormalisedDiff, performing = vel_diff_dyn[Condition == "performing"]$NormalisedDiff, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)
+
+ggpaired(vel_diff_dyn_plot_3, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
+```
+
+### Stats
+```{r, echo = FALSE}
+# stats
+vel_diff_dyn_aov_4 <- aov_car(NormalisedDiff ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
+summary(vel_diff_dyn_aov_4)
+```
+
+#### Post-hoc (emmeans)
+```{r, echo = FALSE}
+vel_diff_dyn_posthoc_3 <- test(pairs(emmeans(vel_diff_dyn_aov_4,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+vel_diff_dyn_posthoc_3
 ```

### 2022-06-27 experiment-1/analysis/stats/dynamics.R

@@ -148,12 +148,12 @@ vel_diff_dyn_trial
 
 
 ## ----vel-diff-dyn-trial-line-ftop, echo = FALSE, fig.height = 4----
-ggline(dt_vel_diff_dyn[Subcomponent == "FtoP", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
+ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
   scale_x_continuous(breaks = seq(1,8,1))
 
 
 ## ----vel-diff-dyn-trial-line-ptof, echo = FALSE, fig.height = 4----
-ggline(dt_vel_diff_dyn[Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
+ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
   scale_x_continuous(breaks = seq(1,8,1))
 
 
@@ -175,12 +175,14 @@ ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", ad
 
 ## ----vel-art, echo = FALSE----------------------------------
 # For each individual
-vel_art <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
[…]

### 2022-06-27 experiment-2/analysis/stats/dynamics.Rmd
update additional analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r vel-art-box,  echo = FALSE}
-ggboxplot(vel_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(vel_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") + facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r vel-diff-art-box,  echo = FALSE}
-ggboxplot(dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
+ggboxplot(vel_diff_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
   facet_grid(Subcomponent ~ .)

### 2022-06-27 experiment-2/analysis/stats/dynamics.R

@@ -181,8 +181,7 @@ vel_art
 
 
 ## ----vel-art-box,  echo = FALSE----------------------------
-ggboxplot(vel_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(vel_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") + facet_grid(Subcomponent ~ .)
 
 
 ## ----vel-art-all, echo = FALSE-----------------------------

### 2022-06-28 experiment-1/analysis/stats/articulation.Rmd
update articulation analysis (average per trial first)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-- [Sequence plot](#sequence_art)
-
 ### KOR
@@
 
-- [Sequence plot](#sequence_dyn)
-
 ### KOR
@@
   + [Trial by trial](#kor_dyn_group_trial)
+
+## [Sequence plot](#sequence)
 
@@
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
+
+# chunk option
+knitr::opts_chunk$set(warning = FALSE)
 ```
@@
 # For each individual
-kot_art <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
+kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+kot_art <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kot_art
@@
 ```{r kot-art-box,  echo = FALSE}
-ggboxplot(dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(kot_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") + facet_grid(Subcomponent ~ .)
 ```
@@
 ### Plot
-#### Scatter plot
-- Error bars show SEM of the mean.
-
-```{r kot-art-all-scatter, echo = FALSE}
-ggerrorplot(kot_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation")
-```
-
 #### Box plot
@@
 # For each individual
-kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
 kot_art_trial
@@
 ```{r kot-art-trial-line-legato, echo = FALSE, fig.height = 4}
-ggline(dt_kot_art[Subcomponent == "Legato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(dt_kot_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
+ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 # For each individual
-kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
+kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+kor_art <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kor_art
@@
 ```{r kor-art-box,  echo = FALSE}
-ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(kor_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") + facet_grid(Subcomponent ~ .)
 ```
@@
 ### Plot
-#### Scatter plot
-- Error bars show SEM of the mean.
-
-```{r kor-art-all-scatter, echo = FALSE}
-ggerrorplot(kor_art, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation")
-```
-
 #### Box plot
@@
 # For each individual
-kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
 kor_art_trial
@@
 ```{r kor-art-trial-line-legato, echo = FALSE, fig.height = 4}
-ggline(dt_kor_art[Subcomponent == "Legato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Legato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Legato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kor-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(dt_kor_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Staccato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Staccato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kor-art-trial-all-line-legato, echo = FALSE}
-ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kor-art-trial-all-line-staccato, echo = FALSE}
-ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Staccato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Staccato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 # For each individual
-kot_dyn <- dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
+kot_dyn_trial <- dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+kot_dyn <- kot_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kot_dyn
@@
 ```{r kot-dyn-box,  echo = FALSE}
-ggboxplot(dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Dynamics") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(kot_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Dynamics") + facet_grid(Subcomponent ~ .)
 ```
@@
 ### Plot
-#### Scatter plot
-- Error bars show SEM of the mean.
-
-```{r kot-dyn-all-scatter, echo = FALSE}
-ggerrorplot(kot_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics")
-```
-
 #### Box plot
@@
 # For each individual
-kot_dyn_trial <- dt_kot_dyn[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
 kot_dyn_trial
@@
 ```{r kot-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-ggline(dt_kot_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-ggline(dt_kot_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-dyn-trial-all-line-forte, echo = FALSE}
-ggline(kot_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-dyn-trial-all-line-piano, echo = FALSE}
-ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 # For each individual
-kor_dyn <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
+kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+kor_dyn <- kor_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kor_dyn
@@
 ```{r kor-dyn-box,  echo = FALSE}
-ggboxplot(dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Dynamics") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(kor_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Dynamics") + facet_grid(Subcomponent ~ .)
 ```
@@
 ### Plot
-#### Scatter plot
-- Error bars show SEM of the mean.
-
-```{r kor-dyn-all-scatter, echo = FALSE}
-ggerrorplot(kor_dyn, x = "Condition", y = "Mean", add = "jitter", add.params = list(color = "darkgray"), error.plot = "errorbar", color = "Condition", facet.by = "Subcomponent", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics")
-```
-
 #### Box plot
@@
 # For each individual
-kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
 kor_dyn_trial
@@
 ```{r kor-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-ggline(dt_kor_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Forte") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Forte") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kor-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-ggline(dt_kor_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Piano") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Dynamics/Piano") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kor-dyn-trial-all-line-forte, echo = FALSE}
-ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Forte") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Forte") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kor_dyn_trial_all_line_piano, echo = FALSE}
-ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") + scale_x_continuous(breaks = seq(1,8,1))
 ```
 
-# Sequence plot
+# Sequence plot{#sequence}
 ## Articulation{#sequence_art}

### 2022-06-28 experiment-1/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE--------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE---------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -46,293 +46,264 @@ dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 
 
-## ----kot-art, echo = FALSE----------------------------------------------------
+## ----kot-art, echo = FALSE---------------------------------
 # For each individual
-kot_art <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
+kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+kot_art <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kot_art
 
 
-## ----kot-art-box,  echo = FALSE-----------------------------------------------
[…]

### 2022-06-28 experiment-1/analysis/stats/dynamics.Rmd
update dynamics (chunk option)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-### [Sequence plot](#sequence)
+## [Sequence plot](#sequence)
 
-- [Additional analysis based on reviewers' comments](#additional)
+## [Additional analysis](#additional)
 
@@
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
-# statistics
+# statistic
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
+
+# chunk option
+knitr::opts_chunk$set(warning = FALSE)
 ```

### 2022-06-28 experiment-1/analysis/stats/dynamics.R

@@ -4,9 +4,12 @@
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 # plot
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
-# statistics
+# statistic
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
+
+# chunk option
+knitr::opts_chunk$set(warning = FALSE)
 
 
 ## ----file, include = FALSE----------------------------------
@@ -379,6 +382,11 @@ summary(vel_diff_dyn_aov_3)
 
 
 ## ---- echo = FALSE------------------------------------------
+vel_diff_dyn_posthoc_2 <- test(pairs(emmeans(vel_diff_dyn_aov_3,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+vel_diff_dyn_posthoc_2
+
+
+## ---- echo = FALSE------------------------------------------
 vel_diff_dyn$NormalisedDiff <- 0
 for (i in unique(vel_diff_dyn_trial$SubNr)){
   # teaching
@@ -408,6 +416,11 @@ ggpaired(vel_diff_dyn_plot_3, cond = "performing", cond2 = "teaching", color = "
 # stats
 vel_diff_dyn_aov_4 <- aov_car(NormalisedDiff ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
 summary(vel_diff_dyn_aov_4)
+
+
+## ---- echo = FALSE------------------------------------------
+vel_diff_dyn_posthoc_3 <- test(pairs(emmeans(vel_diff_dyn_aov_4,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
+vel_diff_dyn_posthoc_3
 
 
 ## ----export, include = FALSE--------------------------------

### 2022-06-28 experiment-2/analysis/stats/articulation.Rmd
update analysis (average per trial first)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-- [Sequence plot](#sequence_art)
-
 ## [Dynamics](#kor-dynamics)
@@
 
-- [Sequence plot](#sequence_dyn)
+## [Sequence plot](#sequence)
 
@@
 ```{r kor-art-box,  echo = FALSE}
-ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOR", title = "KOR: Articulation") + facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r kor-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(dt_kor_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Staccato") +
+ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOR", title = "KOR: Articulation/Staccato") +
   scale_x_continuous(breaks = seq(1,8,1))
@@
 ```{r kor-art-trial-all-line-staccato, echo = FALSE}
-ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-dyn-trial-all-line-piano, echo = FALSE}
-ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") + scale_x_continuous(breaks = seq(1,8,1))
 ```
 
-# Sequence plot
+# Sequence plot{#sequence}
 ## Articulation{#sequence_art}

### 2022-06-28 experiment-2/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE--------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE---------------------------------
 filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE------------------------------
 dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,293 +31,266 @@ dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 
 
-## ----kor-art, echo = FALSE----------------------------------------------------
+## ----kor-art, echo = FALSE---------------------------------
 # For each individual
-kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
+kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+kor_art <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kor_art
 
 
-## ----kor-art-box,  echo = FALSE-----------------------------------------------
[…]

### 2022-06-28 experiment-2/analysis/stats/dynamics.R
update analysis (chunk option)

@@ -8,6 +8,9 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
+# chunk option
+knitr::opts_chunk$set(warning = FALSE)
+
 
 ## ----file, include = FALSE---------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
@@ -244,12 +247,14 @@ ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add
 
 ## ----vel-diff-art, echo = FALSE----------------------------
 # For each individual
-vel_diff_art <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
+vel_diff_art_trial <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
+vel_diff_art <- vel_diff_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_diff_art
 
 
 ## ----vel-diff-art-box,  echo = FALSE-----------------------
-ggboxplot(dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
+ggboxplot(vel_diff_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
   facet_grid(Subcomponent ~ .)
 
 
[…]

### 2022-07-01 experiment-1/analysis/stats/ioi.Rmd
update ioi (average per trial)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-- [Sequence plot](#sequence_art)
-
 ## [Dynamics](#ioi-dynamics)
@@
 # For each individual
-ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
+ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+
+ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
 ioi_art
@@
 ```{r ioi-art-box,  echo = FALSE}
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed")
+ggboxplot(ioi_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 # For each individual
-ioi_art_ch <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
+ioi_art_ch_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)]
+
+ioi_art_ch <- ioi_art_ch_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Boundary)]
 ioi_art_ch
@@
 ```{r ioi-art-ch-box, echo = FALSE}
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 188, linetype = "dashed")
+ggboxplot(ioi_art_ch_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + facet_grid(Boundary ~ .) + geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 ```{r ioi-art-trial, echo = FALSE}
-# For each individual
-ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
 ioi_art_trial
@@
 ```{r ioi-art-trial-line, echo = FALSE, fig.height = 4}
-ggline(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
+ggline(ioi_art_trial, x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "IOIs (ms)", title = "IOI: Articulation") +
   facet_wrap(SubNr ~ .) +
@@
 # For each individual
-ioi_dyn <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
+ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+
+ioi_dyn <- ioi_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
 ioi_dyn
@@
 ```{r ioi-dyn-box,  echo = FALSE}
-ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggboxplot(ioi_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") + 
   geom_hline(yintercept = 188, linetype = "dashed")
@@
 # For each individual
-ioi_dyn_ch <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
+ioi_dyn_ch_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)]
+
+ioi_dyn_ch <- ioi_dyn_ch_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Boundary)]
 ioi_dyn_ch
@@
 ```{r ioi-dyn-ch-box, echo = FALSE}
-ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
+ggboxplot(ioi_dyn_ch_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") + 
   facet_grid(Boundary ~ .) +
@@
 ```
-# Sequence plot
+
+# Sequence plot{#sequence}
 ## Articulation{#sequence_art}

### 2022-07-01 experiment-1/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE--------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,11 +9,11 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE---------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -25,48 +25,48 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE----------------------------------------------------
+## ----ioi-art, echo = FALSE---------------------------------
 # For each individual
 ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE-----------------------------------------------
+## ----ioi-art-box,  echo = FALSE----------------------------
 ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + 
   geom_hline(yintercept = 188, linetype = "dashed")
 
 
-## ----ioi-art-all, echo = FALSE------------------------------------------------
+## ----ioi-art-all, echo = FALSE-----------------------------
 # Group mean
[…]

### 2022-07-01 experiment-2/analysis/stats/ioi.Rmd
tweak notes
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update ioi (average per trial)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # For each individual
-ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
+ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+
+ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
 ioi_art
@@
 ```{r ioi-art-box,  echo = FALSE}
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+ggboxplot(ioi_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + 
   geom_hline(yintercept = 1, linetype = "dashed")
@@
 
-ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") +
-  geom_hline(yintercept = 1, linetype = "dashed")
+ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") + geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 # For each individual
-ioi_art_ch <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Boundary)]
+ioi_art_ch_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)]
+
+ioi_art_ch <- ioi_art_ch_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Boundary)]
 ioi_art_ch
@@
 ```{r ioi-art-ch-box, echo = FALSE}
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
-  facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 1, linetype = "dashed")
+ggboxplot(ioi_art_ch_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + facet_grid(Boundary ~ .) + geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 # For each individual
-ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, TrialNr)]
 ioi_art_trial
@@
 # For each individual
-ioi_dyn <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
+ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+
+ioi_dyn <- ioi_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
 ioi_dyn
@@
 ```{r ioi-dyn-box,  echo = FALSE}
-ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 1, linetype = "dashed")
+ggboxplot(ioi_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") + geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 #### Normality check
-
 ```{r normality2, echo = FALSE}
@@
 #### Paired *t*-test
-
 ```{r paired-dyn, echo = FALSE}
@@
 # For each individual
-ioi_dyn_ch <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Boundary)]
+ioi_dyn_ch_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)]
+
+ioi_dyn_ch <- ioi_dyn_ch_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Boundary)]
 ioi_dyn_ch
@@
 ```{r ioi-dyn-ch-box, echo = FALSE}
-ggboxplot(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
-  facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 1, linetype = "dashed")
+ggboxplot(ioi_dyn_ch_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") + facet_grid(Boundary ~ .) + geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 # For each individual
-ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, TrialNr)]
 ioi_dyn_trial
@@
 ```{r ioi-dyn-trial-line, echo = FALSE, fig.height = 4}
-ggline(dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
+ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
   facet_wrap(SubNr ~ .) +

### 2022-07-01 experiment-2/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE--------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,11 +9,11 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE---------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -25,237 +25,237 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE----------------------------------------------------
+## ----ioi-art, echo = FALSE---------------------------------
 # For each individual
-ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill)]
+ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+
+ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE-----------------------------------------------
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + 
[…]

### 2022-07-03 experiment-1/analysis/stats/ioi.Rmd
correct typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
+
+# chunk option
+knitr::opts_chunk$set(warning = FALSE)
 ```
@@
 
-ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed")
+ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") + geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 #### Normality check
-- Violate the normality for differences between the two conditions
+- The Shapiro-Wilk test showed that the normality for differences between the two conditions was violated.
 
@@
 ```{r ioi-art-trial, echo = FALSE}
+# For each individual
 ioi_art_trial
@@
 ```{r ioi-art-trial-all-line, echo = FALSE}
-ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") + geom_hline(yintercept = 188, linetype = "dashed") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r ioi-dyn-box,  echo = FALSE}
-ggboxplot(ioi_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") + 
-  geom_hline(yintercept = 188, linetype = "dashed")
+ggboxplot(ioi_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") + geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 
-ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dashed")
+ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") + geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 ```{r ioi-dyn-ch-box, echo = FALSE}
-ggboxplot(ioi_dyn_ch_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") + 
-  facet_grid(Boundary ~ .) +
-  geom_hline(yintercept = 188, linetype = "dashed")
+ggboxplot(ioi_dyn_ch_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Dynamics") + facet_grid(Boundary ~ .) + geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 
-ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dashed")
+ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") + geom_hline(yintercept = 188, linetype = "dashed")
 ```
@@
 ```{r ioi-dyn-trial-all-line, echo = FALSE}
-ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") + geom_hline(yintercept = 188, linetype = "dashed") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 
-ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
-  scale_x_continuous(breaks = seq(1,66,1))
+ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") + geom_hline(yintercept = 188, linetype = "dashed") + scale_x_continuous(breaks = seq(1,66,1))
 ```
@@
 
-ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
-    scale_x_continuous(breaks = seq(1,66,1))
+ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") + geom_hline(yintercept = 188, linetype = "dashed") + scale_x_continuous(breaks = seq(1,66,1))
 ```

### 2022-07-03 experiment-1/analysis/stats/ioi.R

@@ -8,6 +8,9 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 
+# chunk option
+knitr::opts_chunk$set(warning = FALSE)
+
 
 ## ----file, include = FALSE---------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
@@ -27,13 +30,14 @@ dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 ## ----ioi-art, echo = FALSE---------------------------------
 # For each individual
-ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
+ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+
+ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
 ## ----ioi-art-box,  echo = FALSE----------------------------
-ggboxplot(dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + 
-  geom_hline(yintercept = 188, linetype = "dashed")
+ggboxplot(ioi_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + geom_hline(yintercept = 188, linetype = "dashed")
 
 
 ## ----ioi-art-all, echo = FALSE-----------------------------
@@ -45,8 +49,7 @@ ioi_art_all
 ## ----ioi-art-all-box, echo = FALSE-------------------------
[…]

### 2022-07-03 experiment-1/analysis/stats/dynamics.R
correct typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -43,8 +43,7 @@ vel_dyn
 
 
 ## ----vel-dyn-box,  echo = FALSE-----------------------------
-ggboxplot(vel_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(vel_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") + facet_grid(Subcomponent ~ .)
 
 
 ## ----vel-dyn-all, echo = FALSE------------------------------
@@ -80,13 +79,11 @@ vel_dyn_trial
 
 
 ## ----vel-dyn-trial-line-forte, echo = FALSE, fig.height = 4----
-ggline(dt_vel_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(dt_vel_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") + scale_x_continuous(breaks = seq(1,8,1))
 
 
 ## ----vel-dyn-trial-line-piano, echo = FALSE, fig.height = 4----
-ggline(dt_vel_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
-  scale_x_continuous(breaks = seq(1,8,1))
[…]

### 2022-07-03 experiment-1/analysis/stats/articulation.Rmd
correct typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r kot-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-art-trial-all-line-legato, echo = FALSE}
-ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r kot-art-trial-all-line-staccato, echo = FALSE}
-ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 #### Box plot
-
 ```{r kot-dyn-box,  echo = FALSE}
@@
 
-ggline(kot_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Articulation") +
-  scale_x_continuous(breaks = seq(1,66,1))
+ggline(kot_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Articulation") + scale_x_continuous(breaks = seq(1,66,1))
 ```
@@
 
-ggline(kot_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Dynamics") +
-  scale_x_continuous(breaks = seq(1,66,1)) 
+ggline(kot_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Dynamics") + scale_x_continuous(breaks = seq(1,66,1))
 ```

### 2022-07-03 experiment-1/analysis/stats/articulation.R

@@ -8,6 +8,9 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
+# chunk option
+knitr::opts_chunk$set(warning = FALSE)
+
 
 ## ----file, include = FALSE---------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
@@ -106,13 +109,11 @@ kot_art_trial_all
 
 
 ## ----kot-art-trial-all-line-legato, echo = FALSE-----------
-ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") + scale_x_continuous(breaks = seq(1,8,1))
 
 
 ## ----kot-art-trial-all-line-staccato, echo = FALSE---------
-ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") + scale_x_continuous(breaks = seq(1,8,1))
 
 
 ## ----kor-art, echo = FALSE---------------------------------

### 2022-07-04 experiment-1/analysis/stats/variability.Rmd
update analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-06-15 experiment-1/analysis/stats/variability.Rmd
@@
 ---
-title: "Variability analysis"
-output: html_notebook
+title: "Variability analysis 1"
+output:
+  html_notebook: default
+editor_options: 
+  chunk_output_type: inline
 ---
+
+Description: This is a summary of variability analysis. Added after the first review from the journal.
+
+- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
+
+```{r setup, include = FALSE}
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# statistics
+if (!require("afex")) {install.packages("afex"); require("afex")}
+if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
+
+filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
+dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".")
+
+# SubNr as a factor
+dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
+
+# Include only articulation
+dt_ioi_art <- dt_ioi[Skill == "articulation"]
+# per participant
+ioi_art_seq_each <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+
+# Include only dynamics
+dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
+# per participant
+ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+```
+
+## Sequence Plots
+- Error bars show the standard error.
+- Dotted lines show the target tempo (IOI = 188ms).
+- Twodashed lines show transition points (i.e., from legato to staccato, from staccato to legato, from forte to piano, from piano to forte).
+
+### Overall
+```{r, echo = FALSE, fig.width = 6, fig.height = 1.2}
+# Plot the whole sequence
+ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
+  geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
+  geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points
+  scale_x_continuous(breaks = seq(1,66,1))
+
+ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Dynamics") +
+  geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
+  geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points
+  scale_x_continuous(breaks = seq(1,66,1))
+```
+
+<!-- ### Each participant -->
+<!-- ```{r, echo = FALSE, fig.width = 5, fig.height = 5} -->
+<!-- # Plot the whole sequence for each participant -->
+<!-- ggplot(ioi_art_seq_each, aes(x = Interval, y = Mean, color = SubNr)) + -->
+<!--   geom_line() + geom_point() + facet_grid(Condition ~ .) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dotted") + # target tempo -->
+<!--   geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points -->
+<!--   scale_x_discrete(breaks = seq(1,66,3)) + labs(x = "Interval", y = "IOIs (ms)", title = "IOIs: Articulation") + theme_pubr() + theme(legend.position = "right") -->
+
+<!-- ioi_dyn_seq <- ioi_dyn_seq_each[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Interval)] -->
+<!-- ggplot(ioi_dyn_seq_each, aes(x = Interval, y = Mean, color = SubNr)) + -->
+<!--   geom_line() + geom_point() + facet_grid(Condition ~ .) + -->
+<!--   geom_hline(yintercept = 188, linetype = "dotted") + # target tempo -->
+<!--   geom_vline(xintercept = c(8, 24, 49, 16, 41, 57), linetype = "twodash") + # transition points -->
+<!--   scale_x_discrete(breaks = seq(1,66,3)) + labs(x = "Interval", y = "IOIs (ms)", title = "IOIs: Dynamics") + theme_pubr() + theme(legend.position = "right") -->
+<!-- ``` -->
+
+## Coefficient of Variation
+```{r, echo = FALSE}
+ioi_cv_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+ioi_cv_subject <- ioi_cv_trial[, .(N = .N, MeanCV = mean(CV), SD = sd(CV)), by = .(SubNr, Condition, Skill)]
+ioi_cv_subject <- ioi_cv_subject[order(SubNr, Condition, Skill)] # ordering
+ioi_cv_subject
+
+ggplot(ioi_cv_subject, aes(x = Condition, y = MeanCV, color = Condition)) +
+  geom_boxplot() +
+  geom_line(aes(group = SubNr), size = 0.4, color = "gray") +
+  geom_point(aes(color = Condition)) +
+  facet_grid(. ~ Skill) +
+  theme_pubr()
+```
+
+### Stats
+- With Participant 2 (outside +/- 3SD)
+
+```{r, echo = FALSE}
+# descriptive stats
+ioi_cv_subject[, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
+
+aov_ez(id = "SubNr", dv = "MeanCV",
+       within = c("Condition", "Skill"),
+       data = ioi_cv_subject)
+```
+
+- Without Participant 2 (outside +/- 3SD)
+
+```{r, echo = FALSE}
+# descriptive stats
+ioi_cv_subject[SubNr != 2, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
+
+aov_ez(id = "SubNr", dv = "MeanCV",
+       within = c("Condition", "Skill"),
+       data = ioi_cv_subject[SubNr != 2])
+```
+
+## IOIs at transition points
+```{r, echo = FALSE}
+ioi_tra_trial <- dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+ioi_tra_subject <- ioi_tra_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
+ioi_tra_subject <- ioi_tra_subject[order(SubNr, Condition, Skill)] # ordering
+
+ggplot(ioi_tra_subject, aes(x = Condition, y = Mean, color = Condition)) +
+  geom_boxplot() +
+  geom_line(aes(group = SubNr), size = 0.4, color = "gray") +
+  geom_point(aes(color = Condition)) +
+  facet_grid(. ~ Skill) +
+  theme_pubr()
+```
+
+```{r, echo = FALSE}
+# descriptive stats
+ioi_tra_subject[, .(.N, MeanCV = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill)]
+
+aov_ez(id = "SubNr", dv = "Mean",
+       within = c("Condition", "Skill"),
+       data = ioi_tra_subject)
+```
+
+## IOIs at transition points vs. others
+Caution: the number of points included in each transition group is considerably different.
+
+```{r, echo = FALSE}
+dt_ioi$Transition <- "No"
+dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF"]$Transition <- "Yes"
+
+# transition points vs. others
+ioi_tra_others_trial <- dt_ioi[, .(.N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Transition)]
+ioi_tra_others_subject <- ioi_tra_others_trial[, .(.N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Transition)]
+ioi_tra_others_subject <- ioi_tra_others_subject[order(SubNr, Condition, Skill)] # ordering
+
+# descriptive stats
+ioi_tra_others_subject[, .(.N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Transition)]
+
+ggplot(ioi_tra_others_subject, aes(x = Condition, y = Mean, color = Condition)) +
+  geom_boxplot() +
+  geom_line(aes(group = SubNr), size = 0.4, color = "gray") +
+  geom_point(aes(color = Condition)) +
+  facet_grid(Skill ~ Transition) +
+  theme_pubr()
+```
+
+```{r, echo = FALSE}
+aov_ez(id = "SubNr", dv = "Mean",
+       within = c("Condition", "Skill", "Transition"),
+       data = ioi_tra_others_subject)
+```
+
+```{r export, include = FALSE}
+knitr::purl("variability.Rmd")
+```

### 2022-07-04 experiment-1/analysis/stats/articulation.R
remove unnessesary tags
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE--------------------------------
+## ----setup, include = FALSE-------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -12,12 +12,12 @@ if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 knitr::opts_chunk$set(warning = FALSE)
 
 
-## ----file, include = FALSE---------------------------------
+## ----file, include = FALSE--------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE------------------------------
+## ----extract, include = FALSE-----------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -49,7 +49,7 @@ dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 
 
-## ----kot-art, echo = FALSE---------------------------------
+## ----kot-art, echo = FALSE--------------------------------
 # For each individual
 kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
 
@@ -57,38 +57,38 @@ kot_art <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(S
 kot_art
 
 
-## ----kot-art-box,  echo = FALSE----------------------------
+## ----kot-art-box,  echo = FALSE---------------------------
 ggboxplot(kot_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") + facet_grid(Subcomponent ~ .)
 
 
-## ----kot-art-all, echo = FALSE-----------------------------
[…]

### 2022-07-04 experiment-1/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE--------------------------------
+## ----setup, include = FALSE-------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -12,11 +12,11 @@ if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 knitr::opts_chunk$set(warning = FALSE)
 
 
-## ----file, include = FALSE---------------------------------
+## ----file, include = FALSE--------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE------------------------------
+## ----extract, include = FALSE-----------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -28,7 +28,7 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE---------------------------------
+## ----ioi-art, echo = FALSE--------------------------------
 # For each individual
 ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
 
@@ -36,40 +36,40 @@ ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(S
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE----------------------------
+## ----ioi-art-box,  echo = FALSE---------------------------
 ggboxplot(ioi_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + geom_hline(yintercept = 188, linetype = "dashed")
 
 
-## ----ioi-art-all, echo = FALSE-----------------------------
+## ----ioi-art-all, echo = FALSE----------------------------
 # Group mean
[…]

### 2022-07-04 experiment-1/analysis/stats/ioi.Rmd
remove unnessesary tags
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
correct typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# Experimental Design{#design}
+# Experimental Design
 - Independent variables: Condition (2 levels: teaching vs. performing) x Technique (2 levels: articulation vs. dynamics)
@@
   + [All](#art_ind_all)
-  + [Phrase bondaries](#art_ind_phrase)
+  + [Phrase boundaries](#art_ind_phrase)
   + [Trial by trial](#art_ind_trial)
@@
   + [All](#art_group_all)
-  + [Phrase bondaries](#art_group_phrase)
+  + [Phrase boundaries](#art_group_phrase)
   + [Trial by trial](#art_group_trial)
@@
   + [All](#dyn_ind_all)
-  + [Phrase bondaries](#dyn_ind_phrase)
+  + [Phrase boundaries](#dyn_ind_phrase)
   + [Trial by trial](#dyn_ind_trial)
@@
   + [All](#dyn_group_all)
-  + [Phrase bondaries](#dyn_group_phrase)
+  + [Phrase boundaries](#dyn_group_phrase)
   + [Trial by trial](#dyn_group_trial)

### 2022-07-04 experiment-2/analysis/stats/articulation.Rmd
update analysis / correct typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# Experimental Design{#design}
+# Experimental Design
 - Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: legato vs. staccato / forte vs. piano)

### 2022-07-04 experiment-2/analysis/stats/dynamics.Rmd
update analysis / correct typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# Experimental Design{#design}
+# Experimental Design
 - Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato)
@@
 ```{r vel-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-dyn-trial-all-line-forte, echo = FALSE}
-ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
-    scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-dyn-trial-all-line-piano, echo = FALSE}
-ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
-    scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-art-trial-line-legato, echo = FALSE, fig.height = 4}
-ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-art-trial-line-staccato, echo = FALSE, fig.height = 4}
-ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-art-trial-all-line-legato, echo = FALSE}
-ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Legato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-art-trial-all-line-staccato, echo = FALSE}
-ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-diff-art-box,  echo = FALSE}
-ggboxplot(vel_diff_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(vel_diff_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") + facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r vel-diff-artt-rial-line-ltos, echo = FALSE, fig.height = 4}
-ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-diff-art-trial-line-stol, echo = FALSE, fig.height = 4}
-ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 
-ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Dynamics") +
-  scale_x_continuous(breaks = seq(1,72,1))
+ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Dynamics") + scale_x_continuous(breaks = seq(1,72,1))
 
@@
 
-ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") +
-  scale_x_continuous(breaks = seq(1,71,1))
+ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") + scale_x_continuous(breaks = seq(1,71,1))
 ```
@@
 
-ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Articulation") +
-  scale_x_continuous(breaks = seq(1,72,1))
+ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Articulation") + scale_x_continuous(breaks = seq(1,72,1))
 
@@
 
-ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") +
-  scale_x_continuous(breaks = seq(1,71,1))
+ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") + scale_x_continuous(breaks = seq(1,71,1))
 ```
@@
 ## Normalised KV difference per trial
-
 ```{r, echo = FALSE, warning = FALSE}

### 2022-07-04 experiment-2/analysis/stats/ioi.Rmd
update analysis / correct typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r ioi-art-box,  echo = FALSE}
-ggboxplot(ioi_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + 
-  geom_hline(yintercept = 1, linetype = "dashed")
+ggboxplot(ioi_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ```
+
 #### Effect size
@@
 
-ggpaired(ioi_art_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") +
-  geom_hline(yintercept = 1, linetype = "dashed")
+ggpaired(ioi_art_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") + geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ```{r ioi-art-trial-line, echo = FALSE, fig.height = 4}
-ggline(dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
+ggline(ioi_art_trial, x = "TrialNr", y = "Mean", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Normalised IOIs", title = "IOI: Articulation") +
   facet_wrap(SubNr ~ .) +
@@
 ```{r ioi-art-trial-all-line, echo = FALSE}
-ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Normalised IOIs", title = "IOI: Articulation") +
-  geom_hline(yintercept = 1, linetype = "dashed") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Normalised IOIs", title = "IOI: Articulation") + geom_hline(yintercept = 1, linetype = "dashed") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 
-ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 1, linetype = "dashed")
+ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") + geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 
-ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 1, linetype = "dashed")
+ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") + geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 ```{r ioi-dyn-trial-all-line, echo = FALSE}
-ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 1, linetype = "dashed") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Normalised IOIs", title = "IOI: Dynamics") + geom_hline(yintercept = 1, linetype = "dashed") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 
-ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "Normalised IOIs", title = "IOI: Articulation") +
-  geom_hline(yintercept = 1, linetype = "dashed") +
-  scale_x_continuous(breaks = seq(1,71,1))
+ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "Normalised IOIs", title = "IOI: Articulation") + geom_hline(yintercept = 1, linetype = "dashed") + scale_x_continuous(breaks = seq(1,71,1))
 ```
@@
 
-ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "Normalised IOIs", title = "IOI: Dynamics") +
-  geom_hline(yintercept = 1, linetype = "dashed") +
-    scale_x_continuous(breaks = seq(1,71,1))
+ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "Normalised IOIs", title = "IOI: Dynamics") + geom_hline(yintercept = 1, linetype = "dashed") +　scale_x_continuous(breaks = seq(1,71,1))
 ```

### 2022-07-04 experiment-2/analysis/stats/variability.Rmd
update analysis / correct typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: "Variability analysis"
+title: "Variability analysis 2"
 output:
@@
 # transition points vs. others
-ioi_tra_others <- dt_ioi[, .(.N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Transition)]
-ioi_tra_others <- ioi_tra_others[order(SubNr, Condition, Skill)] # ordering
+ioi_tra_others_trial <- dt_ioi[, .(.N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Transition)]
+ioi_tra_others_subject <- ioi_tra_others_trial[, .(.N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Transition)]
+ioi_tra_others_subject <- ioi_tra_others_subject[order(SubNr, Condition, Skill)] # ordering
 
 # descriptive stats
-ioi_tra_others[, .(.N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Transition)]
+ioi_tra_others_subject[, .(.N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Transition)]
 
-ggplot(ioi_tra_others, aes(x = Condition, y = Mean, color = Condition)) +
+ggplot(ioi_tra_others_subject, aes(x = Condition, y = Mean, color = Condition)) +
   geom_boxplot() +
@@
        within = c("Condition", "Skill", "Transition"),
-       data = ioi_tra_others)
+       data = ioi_tra_others_subject)
 ```

### 2022-07-04 experiment-2/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE--------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -8,13 +8,16 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
+# chunk option
+knitr::opts_chunk$set(warning = FALSE)
 
-## ----file, include = FALSE---------------------------------
+
+## ----file, include = FALSE----------------------------------------------------
 filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,7 +34,7 @@ dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 
 
-## ----kor-art, echo = FALSE---------------------------------
+## ----kor-art, echo = FALSE----------------------------------------------------
 # For each individual
 kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
 
@@ -39,66 +42,66 @@ kor_art <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(S
 kor_art
 
 
-## ----kor-art-box,  echo = FALSE----------------------------
+## ----kor-art-box,  echo = FALSE-----------------------------------------------
[…]

### 2022-07-04 experiment-2/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE--------------------------------
+## ----setup, include = FALSE-------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -8,12 +8,15 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 
+# chunk option
+knitr::opts_chunk$set(warning = FALSE)
 
-## ----file, include = FALSE---------------------------------
+
+## ----file, include = FALSE--------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE------------------------------
+## ----extract, include = FALSE-----------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -25,7 +28,7 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE---------------------------------
+## ----ioi-art, echo = FALSE--------------------------------
 # For each individual
 ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
 
@@ -33,51 +36,50 @@ ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(S
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE----------------------------
-ggboxplot(ioi_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + 
-  geom_hline(yintercept = 1, linetype = "dashed")
+## ----ioi-art-box,  echo = FALSE---------------------------
[…]

### 2022-07-04 paper/manuscript.tex
add teaching experiences for demographic info
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
updated stats
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={The sound of teaching music: Expert pianists' performance modulations for novices},
-  pdfauthor={Atsuko Tominaga1,*, Günther Knoblich1, \& Natalie Sebanz1},
-  pdflang={en-EN},
-  pdfkeywords={teaching, expertise, skill transmission, artistic expression, music},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx}
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={The sound of teaching music: Expert pianists' performance modulations for novices},
+  pdfauthor={Atsuko Tominaga1,*, Günther Knoblich1, \& Natalie Sebanz1},
+  pdflang={en-EN},
+  pdfkeywords={teaching, expertise, skill transmission, artistic expression, music},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 They had 12.45 years of practice on average (\emph{SD} = 5.66).
+10 participants had an experience in teaching the piano (\emph{M} = 3.48 years, \emph{SD} = 3.51).
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
@@
 To compare the mean IOIs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wilcoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.24 (ms), \emph{IQR} = 19.32{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
+The Wilcoxon Signed-rank test revealed that participants played more slowly in the teaching condition {[}\emph{Mdn} = 189.55 (ms), \emph{IQR} = 26.74{]} than in the performing condition {[}\emph{Mdn} = 185.23 (ms), \emph{IQR} = 19.30{]} while playing the piece with the notated articulation (\emph{p} = 0.021, \emph{r} = 0.41, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) revealed that there was a significant main effect of Articulation (\emph{F}(1,30) = 1158, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Articulation (\emph{F}(1,30) = 8.31, \emph{p} = 0.007, \(\eta_G^2\) = 0.011, \emph{Fig \ref{fig:kot-1}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KOT was more negative in the teaching condition {[}\emph{M} = -134.03 (ms), \emph{SD} = 19.86{]} than in the performing condition {[}\emph{M} = -128.35 (ms), \emph{SD} = 20.80{]} (\emph{p} = 0.007).
-For legato there was no significant difference in KOT between the teaching {[}\emph{M} = 15.91 (ms), \emph{SD} = 15.49{]} and performing condition {[}\emph{M} = 14.00 (ms), \emph{SD} = 17.01{]} (\emph{p} = 0.31).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) revealed that there was a significant main effect of Articulation (\emph{F}(1,30) = 1149, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94) and a significant interaction between Condition and Articulation (\emph{F}(1,30) = 8.52, \emph{p} = 0.007, \(\eta_G^2\) = 0.010, \emph{Fig \ref{fig:kot-1}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KOT was more negative in the teaching condition {[}\emph{M} = -133.79 (ms), \emph{SD} = 20.13{]} than in the performing condition {[}\emph{M} = -128.34 (ms), \emph{SD} = 20.86{]} (\emph{p} = 0.005).
+For legato there was no significant difference in KOT between the teaching {[}\emph{M} = 15.90 (ms), \emph{SD} = 15.51{]} and performing condition {[}\emph{M} = 14.01 (ms), \emph{SD} = 17.02{]} (\emph{p} = 0.32).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant interaction between Condition and Articulation (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.004, \emph{Fig \ref{fig:vel-1}}).
-However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Legato: \emph{M} = 70.34, \emph{SD} = 6.45; Staccato: \emph{M} = 72.01, \emph{SD} = 8.74{]} and performing condition {[}Legato: \emph{M} = 70.68, \emph{SD} = 5.84, Staccato: \emph{M} = 70.42, \emph{SD} = 7.96{]} for each subcomponent (Legato: \emph{p} = 0.68, Staccato: \emph{p} = 0.09).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant interaction between Condition and Articulation (\emph{F}(1,30) = 7.45, \emph{p} = 0.011, \(\eta_G^2\) = 0.004, \emph{Fig \ref{fig:vel-1}}).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Legato: \emph{M} = 70.34, \emph{SD} = 6.46; Staccato: \emph{M} = 72.00, \emph{SD} = 8.74{]} and performing condition {[}Legato: \emph{M} = 70.68, \emph{SD} = 5.86, Staccato: \emph{M} = 70.41, \emph{SD} = 7.97{]} for each subcomponent (Legato: \emph{p} = 0.68, Staccato: \emph{p} = 0.09).
 
@@
 To compare the mean IOIs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wilcoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.29 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.58{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
+The Wilcoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.28 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.59{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 370, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Dynamics (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.003, \emph{Fig \ref{fig:vel-1}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.30{]} than in the performing condition {[}\emph{M} = 81.66, \emph{SD} = 7.43{]} (\emph{p} = 0.069) when instructed to play forte.
-There was no significant difference in KV between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.07{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.40{]} (\emph{p} = 0.24) when instructed to play piano.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 369, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Dynamics (\emph{F}(1,30) = 6.83, \emph{p} = 0.014, \(\eta_G^2\) = 0.003, \emph{Fig \ref{fig:vel-1}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.31{]} than in the performing condition {[}\emph{M} = 81.65, \emph{SD} = 7.43{]} (\emph{p} = 0.068) when instructed to play forte.
+There was no significant difference in KV between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.08{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.41{]} (\emph{p} = 0.24) when instructed to play piano.
 
@@
 Outliers were removed using the same criteria as for the other dependent variables and less than 5 \% of the data were excluded from analysis.
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction between Condition and Transition Type (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.032, \emph{Fig \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (\emph{p} = 0.002) and from piano to forte (\emph{p} \textless{} 0.001) in the teaching condition {[}FtoP: \emph{M} = -14.06, \emph{SD} = 7.20; PtoF: \emph{M} = 24.26, \emph{SD} = 9.75{]} than in the performing condition {[}FtoP: \emph{M} = -11.69, \emph{SD} = 5.97, PtoF: \emph{M} = 20.88, \emph{SD} = 8.67{]}.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,30) = 224, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction between Condition and Transition Type (\emph{F}(1,30) = 25.99, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.033, \emph{Fig \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (\emph{p} = 0.002) and from piano to forte (\emph{p} \textless{} 0.001) in the teaching condition {[}FtoP: \emph{M} = -14.11, \emph{SD} = 7.17; PtoF: \emph{M} = 24.27, \emph{SD} = 9.77{]} than in the performing condition {[}FtoP: \emph{M} = -11.70, \emph{SD} = 5.99, PtoF: \emph{M} = 20.87, \emph{SD} = 8.67{]}.
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) revealed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 377, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.31, \emph{Fig \ref{fig:kot-1}}), reflecting more key-overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,30) = 0.42, \emph{p} = 0.52, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Dynamics (\emph{F}(1,30) = 0.000, \emph{p} = 0.99, \(\eta_G^2\) = 0.000) was significant.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) revealed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 308, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.30, \emph{Fig \ref{fig:kot-1}}), reflecting more key-overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (\emph{F}(1,30) = 0.60, \emph{p} = 0.44, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Dynamics (\emph{F}(1,30) = 0.053, \emph{p} = 0.82, \(\eta_G^2\) = 0.000) was significant.
 
@@
 They had 15.65 years of practice on average (\emph{SD} = 5.67).
+11 participants had an experience in teaching the piano (\emph{M} = 4.14 years, \emph{SD} = 3.65).
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant main effect of Articulation (\emph{F}(1,19) = 2573, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Articulation (\emph{F}(1,19) = 8.91, \emph{p} = 0.008, \(\eta_G^2\) = 0.01, \emph{Fig \ref{fig:kor-2}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant main effect of Articulation (\emph{F}(1,19) = 2566, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.98) and a significant interaction between Condition and Articulation (\emph{F}(1,19) = 8.75, \emph{p} = 0.008, \(\eta_G^2\) = 0.01, \emph{Fig \ref{fig:kor-2}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that for staccato KORs were more negative in the teaching condition {[}\emph{M} = -0.76, \emph{SD} = 0.06{]} than in the performing condition {[}\emph{M} = -0.74, \emph{SD} = 0.04{]} (\emph{p} = \textless{} 0.001).
-For legato there was no significant difference in KORs between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.63).
+For legato there was no significant difference in KORs between the teaching {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} and performing condition {[}\emph{M} = 0.12, \emph{SD} = 0.07{]} (\emph{p} = 0.64).
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant main effect of Articulation (\emph{F}(1,19) = 7.91, \emph{p} = 0.011, \(\eta_G^2\) = 0.03, \emph{Fig \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
-Neither the main effect of Condition (\emph{F}(1,19) = 1.33, \emph{p} = 0.26, \(\eta_G^2\) = 0.003) nor the interaction between Condition and Articulation (\emph{F}(1,19) = 0.14, \emph{p} = 0.71, \(\eta_G^2\) = 0.000) was significant.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Articulation (legato vs.~staccato) showed that there was a significant main effect of Articulation (\emph{F}(1,19) = 7.85, \emph{p} = 0.011, \(\eta_G^2\) = 0.03, \emph{Fig \ref{fig:vel-2}}), reflecting louder sound for legato notes compared to staccato notes.
+Neither the main effect of Condition (\emph{F}(1,19) = 1.34, \emph{p} = 0.26, \(\eta_G^2\) = 0.003) nor the interaction between Condition and Articulation (\emph{F}(1,19) = 0.15, \emph{p} = 0.71, \(\eta_G^2\) = 0.000) was significant.
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,19) = 132, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Dynamics (\emph{F}(1,19) = 6.12, \emph{p} = 0.023, \(\eta_G^2\) = 0.018, \emph{Fig \ref{fig:vel-2}}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,19) = 131, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.72) and a significant interaction between Condition and Dynamics (\emph{F}(1,19) = 6.11, \emph{p} = 0.023, \(\eta_G^2\) = 0.018, \emph{Fig \ref{fig:vel-2}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 83.20, \emph{SD} = 8.84{]} than in the performing condition {[}\emph{M} = 81.49, \emph{SD} = 8.12{]} (\emph{p} = 0.020) when instructed to play forte.
-Also, KV was lower in the teaching condition {[}\emph{M} = 60.53, \emph{SD} = 3.76{]} than in the performing condition {[}\emph{M} = 62.42, \emph{SD} = 4.82{]} (\emph{p} = 0.046) when instructed to play piano.
+Also, KV was lower in the teaching condition {[}\emph{M} = 60.54, \emph{SD} = 3.76{]} than in the performing condition {[}\emph{M} = 62.43, \emph{SD} = 4.83{]} (\emph{p} = 0.046) when instructed to play piano.
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,19) = 124, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.82) and a significant interaction between Condition and Transition Type (\emph{F}(1,19) = 9.15, \emph{p} = 0.007, \(\eta_G^2\) = 0.059, \emph{Fig \ref{fig:vel-diff}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (\emph{p} = 0.029) and from piano to forte (\emph{p} = 0.003) in the teaching condition {[}FtoP: \emph{M} = -17.23, \emph{SD} = 8.26, PtoF: \emph{M} = 20.71, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP: \emph{M} = -14.04, \emph{SD} = 7.67, PtoF: \emph{M} = 16.05, \emph{SD} = 7.32{]}.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,19) = 126, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.83) and a significant interaction between Condition and Transition Type (\emph{F}(1,19) = 9.13, \emph{p} = 0.007, \(\eta_G^2\) = 0.059, \emph{Fig \ref{fig:vel-diff}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when changing from forte to piano (\emph{p} = 0.028) and from piano to forte (\emph{p} = 0.004) in the teaching condition {[}FtoP: \emph{M} = -17.28, \emph{SD} = 8.17, PtoF: \emph{M} = 20.72, \emph{SD} = 8.87{]} than in the performing condition {[}FtoP: \emph{M} = -14.08, \emph{SD} = 7.62, PtoF: \emph{M} = 16.08, \emph{SD} = 7.32{]}.
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,19) = 630, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94), reflecting more key-overlap for forte notes compared to piano notes.
-Neither the main effect of Condition (\emph{F}(1,19) = 0.081, \emph{p} = 0.78, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Dynamics (\emph{F}(1,19) = 0.35, \emph{p} = 0.56, \(\eta_G^2\) = 0.001, \emph{Fig \ref{fig:kor-2}}) was significant.
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,19) = 629, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.94), reflecting more key-overlap for forte notes compared to piano notes.
+Neither the main effect of Condition (\emph{F}(1,19) = 0.084, \emph{p} = 0.77, \(\eta_G^2\) = 0.000) nor the interaction between Condition and Dynamics (\emph{F}(1,19) = 0.45, \emph{p} = 0.51, \(\eta_G^2\) = 0.001, \emph{Fig \ref{fig:kor-2}}) was significant.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-thornton_2008}{}}%
-\CSLLeftMargin{1. }
+\CSLLeftMargin{1. }%
 \CSLRightInline{Thornton, A. \& Raihani, N. J. \href{https://doi.org/10.1016/j.anbehav.2007.12.014}{The evolution of teaching}. \emph{Animal Behaviour} \textbf{75}, 1823--1836 (2008).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-tomasello_2016}{}}%
-\CSLLeftMargin{2. }
+\CSLLeftMargin{2. }%
 \CSLRightInline{Tomasello, M. \href{https://doi.org/10.1016/j.copsyc.2015.09.008}{The ontogeny of cultural learning}. \emph{Current Opinion in Psychology} \textbf{8}, 1--4 (2016).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-brand_2002}{}}%
-\CSLLeftMargin{3. }
+\CSLLeftMargin{3. }%
 \CSLRightInline{Brand, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-saint-georges_2013}{}}%
-\CSLLeftMargin{4. }
+\CSLLeftMargin{4. }%
 \CSLRightInline{Saint-Georges, C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-uther_2007}{}}%
-\CSLLeftMargin{5. }
+\CSLLeftMargin{5. }%
 \CSLRightInline{Uther, M., Knoll, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mcellin_2017}{}}%
-\CSLLeftMargin{6. }
+\CSLLeftMargin{6. }%
 \CSLRightInline{McEllin, L., Knoblich, G.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-csibra_2009}{}}%
-\CSLLeftMargin{7. }
+\CSLLeftMargin{7. }%
 \CSLRightInline{Csibra, G. \& Gergely, G. \href{https://doi.org/10.1016/j.tics.2009.01.005}{Natural pedagogy}. \emph{Trends in Cognitive Sciences} \textbf{13}, 148--153 (2009).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-campisi_2013}{}}%
-\CSLLeftMargin{8. }
+\CSLLeftMargin{8. }%
 \CSLRightInline{Campisi, E.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-fukuyama_2015}{}}%
-\CSLLeftMargin{9. }
+\CSLLeftMargin{9. }%
 \CSLRightInline{Fukuyama, H.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-schaik_2019}{}}%
-\CSLLeftMargin{10. }
+\CSLLeftMargin{10. }%
 \CSLRightInline{Schaik, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-NIPS2016_6413}{}}%
-\CSLLeftMargin{11. }
+\CSLLeftMargin{11. }%
 \CSLRightInline{Ho, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sloboda_2000}{}}%
-\CSLLeftMargin{12. }
+\CSLLeftMargin{12. }%
 \CSLRightInline{Sloboda, J. A. \href{https://doi.org/10.1016/S1364-6613(00)01531-X}{Individual differences in music performance}. \emph{Trends in Cognitive Sciences} \textbf{4}, 397--403 (2000).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-juslin_2003}{}}%
-\CSLLeftMargin{13. }
+\CSLLeftMargin{13. }%
 \CSLRightInline{Juslin, P.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bresin_2000}{}}%
-\CSLLeftMargin{14. }
+\CSLLeftMargin{14. }%
 \CSLRightInline{Bresin, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lindstrom_2003}{}}%
-\CSLLeftMargin{15. }
+\CSLLeftMargin{15. }%
 \CSLRightInline{Lindström, E., Juslin, P.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-okazaki_2019}{}}%
-\CSLLeftMargin{16. }
+\CSLLeftMargin{16. }%
 \CSLRightInline{Okazaki, S., Muraoka, Y.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-sperber_1995}{}}%
-\CSLLeftMargin{17. }
+\CSLLeftMargin{17. }%
 \CSLRightInline{Sperber, D. \& Wilson, D. \emph{Relevance: {Communication} and cognition, 2nd ed}. ({Blackwell Publishing}, 1995).}
@@
 
-\hypertarget{instructions-experiment-1}{%
-\subsection{Instructions (Experiment 1)}\label{instructions-experiment-1}}
+\hypertarget{instructions}{%
+\section{Instructions}\label{instructions}}
+
+\hypertarget{experiment-1-1}{%
+\subsection{Experiment 1}\label{experiment-1-1}}
 
@@
 
-\hypertarget{instructions-experiment-2}{%
-\subsection{Instructions (Experiment 2)}\label{instructions-experiment-2}}
+\hypertarget{experiment-2-1}{%
+\subsection{Experiment 2}\label{experiment-2-1}}
 
@@
 
+\clearpage
+
+\hypertarget{variability}{%
+\section{Variability}\label{variability}}
+
+\hypertarget{data-analysis-2}{%
+\subsection{Data analysis}\label{data-analysis-2}}
+
+\hypertarget{results-experiment-1}{%
+\subsection{Results (Experiment 1)}\label{results-experiment-1}}
+
+\hypertarget{results-experiment-2}{%
+\subsection{Results (Experiment 2)}\label{results-experiment-2}}
+
 \end{document}

### 2022-07-07 experiment-1/analysis/stats/variability.Rmd
update variability analysis (separate stats for each technique)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ioi_cv_subject <- ioi_cv_subject[order(SubNr, Condition, Skill)] # ordering
-ioi_cv_subject
+ioi_cv <- ioi_cv_subject[, .(N = .N, MeanCV = mean(MeanCV), MedianCV = median(MeanCV), SD = sd(MeanCV), IQR = IQR(MeanCV)), by = .(Condition, Skill)]
+ioi_cv
 
@@
 ### Stats
-- With Participant 2 (outside +/- 3SD)
+1) Articulation
 
 ```{r, echo = FALSE}
-# descriptive stats
-ioi_cv_subject[, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
+# normality check
+diff <- ioi_cv_subject[Condition == "teaching" & Skill == "articulation"]$MeanCV-ioi_cv_subject[Condition == "performing" & Skill == "articulation"]$MeanCV
+ioi_art_cv_norm <- shapiro.test(diff)
+ioi_art_cv_norm
 
-aov_ez(id = "SubNr", dv = "MeanCV",
-       within = c("Condition", "Skill"),
-       data = ioi_cv_subject)
+# wilcoxon test
+ioi_art_cv_wilcoxon <- wilcox.test(MeanCV ~ Condition, data = ioi_cv_subject[Skill == "articulation"], paired = TRUE, alternative = "two.sided")
+ioi_art_cv_wilcoxon
+
+# effect size
+wilcoxeff_art_cv <- ioi_cv_subject[Skill == "articulation"] %>% wilcox_effsize(MeanCV ~ Condition, paired = TRUE)
+wilcoxeff_art_cv
 ```
 
-- Without Participant 2 (outside +/- 3SD)
+2) Dynamics
 
 ```{r, echo = FALSE}
-# descriptive stats
-ioi_cv_subject[SubNr != 2, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
+# normality check
+diff2 <- ioi_cv_subject[Condition == "teaching" & Skill == "dynamics"]$MeanCV-ioi_cv_subject[Condition == "performing" & Skill == "dynamics"]$MeanCV
+ioi_dyn_cv_norm <- shapiro.test(diff2)
+ioi_dyn_cv_norm
 
-aov_ez(id = "SubNr", dv = "MeanCV",
-       within = c("Condition", "Skill"),
-       data = ioi_cv_subject[SubNr != 2])
+# wilcoxon test
+ioi_dyn_cv_wilcoxon <- wilcox.test(MeanCV ~ Condition, data = ioi_cv_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
+ioi_dyn_cv_wilcoxon
+
+# effect size
+wilcoxeff_dyn_cv <- ioi_cv_subject[Skill == "dynamics"] %>% wilcox_effsize(MeanCV ~ Condition, paired = TRUE)
+wilcoxeff_dyn_cv
 ```
@@
 ioi_tra_subject <- ioi_tra_subject[order(SubNr, Condition, Skill)] # ordering
+ioi_tra <- ioi_tra_subject[, .(N = .N, Mean = mean(Mean), Median = median(Mean), SD = sd(Mean), IQR = IQR(Mean)), by = .(Condition, Skill)]
+ioi_tra
 
@@
 
+### Stats
+1) Articulation
+
 ```{r, echo = FALSE}
-# descriptive stats
-ioi_tra_subject[, .(.N, MeanCV = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill)]
+# normality check
+diff3 <- ioi_tra_subject[Condition == "teaching" & Skill == "articulation"]$Mean-ioi_tra_subject[Condition == "performing" & Skill == "articulation"]$Mean
+ioi_art_tra_norm <- shapiro.test(diff3)
+ioi_art_tra_norm
 
-aov_ez(id = "SubNr", dv = "Mean",
-       within = c("Condition", "Skill"),
-       data = ioi_tra_subject)
+# wilcoxon test
+ioi_art_tra_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "articulation"], paired = TRUE, alternative = "two.sided")
+ioi_art_tra_wilcoxon
+
+# effect size
+wilcoxeff_art_tra <- ioi_tra_subject[Skill == "articulation"] %>% wilcox_d(Mean ~ Condition, paired = TRUE)
+wilcoxeff_art_tra
 ```
 
-## IOIs at transition points vs. others
-Caution: the number of points included in each transition group is considerably different.
+2) Dynamics
 
 ```{r, echo = FALSE}
-dt_ioi$Transition <- "No"
-dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF"]$Transition <- "Yes"
+# normality check
+diff4 <- ioi_tra_subject[Condition == "teaching" & Skill == "dynamics"]$Mean-ioi_tra_subject[Condition == "performing" & Skill == "dynamics"]$Mean
+ioi_dyn_tra_norm <- shapiro.test(diff4)
+ioi_dyn_tra_norm
 
-# transition points vs. others
-ioi_tra_others_trial <- dt_ioi[, .(.N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Transition)]
-ioi_tra_others_subject <- ioi_tra_others_trial[, .(.N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Transition)]
-ioi_tra_others_subject <- ioi_tra_others_subject[order(SubNr, Condition, Skill)] # ordering
+# t test
+ioi_dyn_tra_t_test <- t.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
+ioi_dyn_tra_t_test
 
-# descriptive stats
-ioi_tra_others_subject[, .(.N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Transition)]
+# wilcoxon test
+ioi_dyn_tra_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
+ioi_dyn_tra_wilcoxon
 
-ggplot(ioi_tra_others_subject, aes(x = Condition, y = Mean, color = Condition)) +
-  geom_boxplot() +
-  geom_line(aes(group = SubNr), size = 0.4, color = "gray") +
-  geom_point(aes(color = Condition)) +
-  facet_grid(Skill ~ Transition) +
-  theme_pubr()
-```
-
-```{r, echo = FALSE}
-aov_ez(id = "SubNr", dv = "Mean",
-       within = c("Condition", "Skill", "Transition"),
-       data = ioi_tra_others_subject)
+# effect size
+wilcoxeff_dyn_tra <- ioi_tra_subject[Skill == "dynamics"] %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
+wilcoxeff_dyn_tra
 ```

### 2022-07-07 experiment-1/analysis/stats/variability.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE-------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -25,7 +25,7 @@ dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
 
 
-## ---- echo = FALSE, fig.width = 6, fig.height = 1.2---------------------------
+## ---- echo = FALSE, fig.width = 6, fig.height = 1.2-------
 # Plot the whole sequence
 ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
   geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
@@ -38,7 +38,7 @@ ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "C
   scale_x_continuous(breaks = seq(1,66,1))
 
 
-## ---- echo = FALSE------------------------------------------------------------
+## ---- echo = FALSE----------------------------------------
 ioi_cv_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
 ioi_cv_subject <- ioi_cv_trial[, .(N = .N, MeanCV = mean(CV), SD = sd(CV)), by = .(SubNr, Condition, Skill)]
 ioi_cv_subject <- ioi_cv_subject[order(SubNr, Condition, Skill)] # ordering
@@ -52,25 +52,27 @@ ggplot(ioi_cv_subject, aes(x = Condition, y = MeanCV, color = Condition)) +
   theme_pubr()
 
 
-## ---- echo = FALSE------------------------------------------------------------
-# descriptive stats
-ioi_cv_subject[, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
[…]

### 2022-07-07 experiment-1/analysis/stats/variability.Rmd
edit x/y axis labels
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   geom_point(aes(color = Condition)) +
+  labs(y = "CVs") +
   facet_grid(. ~ Skill) +
@@
   geom_point(aes(color = Condition)) +
+  labs(y = "IOIs (ms)") +
   facet_grid(. ~ Skill) +
@@
 # effect size
-wilcoxeff_art_tra <- ioi_tra_subject[Skill == "articulation"] %>% wilcox_d(Mean ~ Condition, paired = TRUE)
+wilcoxeff_art_tra <- ioi_tra_subject[Skill == "articulation"] %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
 wilcoxeff_art_tra

### 2022-07-07 experiment-1/analysis/stats/variability.R

@@ -42,12 +42,14 @@ ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "C
 ioi_cv_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
 ioi_cv_subject <- ioi_cv_trial[, .(N = .N, MeanCV = mean(CV), SD = sd(CV)), by = .(SubNr, Condition, Skill)]
 ioi_cv_subject <- ioi_cv_subject[order(SubNr, Condition, Skill)] # ordering
-ioi_cv_subject
+ioi_cv <- ioi_cv_subject[, .(N = .N, MeanCV = mean(MeanCV), MedianCV = median(MeanCV), SD = sd(MeanCV), IQR = IQR(MeanCV)), by = .(Condition, Skill)]
+ioi_cv
 
 ggplot(ioi_cv_subject, aes(x = Condition, y = MeanCV, color = Condition)) +
   geom_boxplot() +
   geom_line(aes(group = SubNr), size = 0.4, color = "gray") +
   geom_point(aes(color = Condition)) +
+  labs(y = "CVs") +
   facet_grid(. ~ Skill) +
   theme_pubr()
 
@@ -58,8 +60,13 @@ diff <- ioi_cv_subject[Condition == "teaching" & Skill == "articulation"]$MeanCV
 ioi_art_cv_norm <- shapiro.test(diff)
 ioi_art_cv_norm
 
+# wilcoxon test
 ioi_art_cv_wilcoxon <- wilcox.test(MeanCV ~ Condition, data = ioi_cv_subject[Skill == "articulation"], paired = TRUE, alternative = "two.sided")
 ioi_art_cv_wilcoxon
+
+# effect size
+wilcoxeff_art_cv <- ioi_cv_subject[Skill == "articulation"] %>% wilcox_effsize(MeanCV ~ Condition, paired = TRUE)
+wilcoxeff_art_cv
 
 
 ## ---- echo = FALSE----------------------------------------
@@ -68,19 +75,27 @@ diff2 <- ioi_cv_subject[Condition == "teaching" & Skill == "dynamics"]$MeanCV-io
 ioi_dyn_cv_norm <- shapiro.test(diff2)
 ioi_dyn_cv_norm
 
+# wilcoxon test
[…]

### 2022-07-07 experiment-2/analysis/stats/variability.Rmd
use normIOIs / edit x/y axis labels
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # per participant
-ioi_art_seq_each <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+ioi_art_seq_each <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Interval)]
 
@@
 # per participant
-ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Interval)]
 ```
@@
 ### Overall
-
 ```{r, echo = FALSE, fig.width = 6, fig.height = 1.2}
 # Plot the whole sequence
-ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
+ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "Normalised IOIs", title = "IOIs: Articulation") +
   geom_vline(xintercept = c(5, 17, 47, 8, 20, 39), linetype = "twodash") + # transition points
@@
 
-ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Dynamics") +
+ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "Normalised IOIs", title = "IOIs: Dynamics") +
   geom_vline(xintercept = c(5, 17, 47, 8, 20, 39), linetype = "twodash") + # transition points
@@
 ## Coefficient of Variation
-
 ```{r, echo = FALSE}
-ioi_cv_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+ioi_cv_trial <- dt_ioi[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI), CV = sd(normIOI)/mean(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
 ioi_cv_subject <- ioi_cv_trial[, .(N = .N, MeanCV = mean(CV), SD = sd(CV)), by = .(SubNr, Condition, Skill)]
 ioi_cv_subject <- ioi_cv_subject[order(SubNr, Condition, Skill)] # ordering
+ioi_cv <- ioi_cv_subject[, .(N = .N, MeanCV = mean(MeanCV), MedianCV = median(MeanCV), SD = sd(MeanCV), IQR = IQR(MeanCV)), by = .(Condition, Skill)]
+ioi_cv
 
@@
 ### Stats
-- With Participant 15 (outside +/- 3SD)
+1) Articulation
 
 ```{r, echo = FALSE}
-# descriptive stats
-ioi_cv_subject[, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
+# normality check
+diff <- ioi_cv_subject[Condition == "teaching" & Skill == "articulation"]$MeanCV-ioi_cv_subject[Condition == "performing" & Skill == "articulation"]$MeanCV
+ioi_art_cv_norm <- shapiro.test(diff)
+ioi_art_cv_norm
 
-aov_ez(id = "SubNr", dv = "MeanCV",
-       within = c("Condition", "Skill"),
-       data = ioi_cv_subject)
+# wilcoxon test
+ioi_art_cv_wilcoxon <- wilcox.test(MeanCV ~ Condition, data = ioi_cv_subject[Skill == "articulation"], paired = TRUE, alternative = "two.sided")
+ioi_art_cv_wilcoxon
+
+# effect size
+wilcoxeff_art_cv <- ioi_cv_subject[Skill == "articulation"] %>% wilcox_effsize(MeanCV ~ Condition, paired = TRUE)
+wilcoxeff_art_cv
 ```
 
-- Without Participant 15 (outside +/- 3SD)
+2) Dynamics
 
 ```{r, echo = FALSE}
-# descriptive stats
-ioi_cv_subject[SubNr != 15, .(.N, MeanCV = mean(MeanCV), SDCV = sd(MeanCV)), by = .(Condition, Skill)]
+# normality check
+diff2 <- ioi_cv_subject[Condition == "teaching" & Skill == "dynamics"]$MeanCV-ioi_cv_subject[Condition == "performing" & Skill == "dynamics"]$MeanCV
+ioi_dyn_cv_norm <- shapiro.test(diff2)
+ioi_dyn_cv_norm
 
-aov_ez(id = "SubNr", dv = "MeanCV",
-       within = c("Condition", "Skill"),
-       data = ioi_cv_subject[SubNr != 15])
+# t test
+ioi_dyn_cv_ttest <- t.test(MeanCV ~ Condition, data = ioi_cv_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
+ioi_dyn_cv_ttest
+
+# wilcoxon test
+ioi_dyn_cv_wilcoxon <- wilcox.test(MeanCV ~ Condition, data = ioi_cv_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
+ioi_dyn_cv_wilcoxon
+
+# effect size
+coheneff_dyn_cv <- ioi_cv_subject[Skill == "dynamics"] %>% cohens_d(MeanCV ~ Condition, paired = TRUE)
+coheneff_dyn_cv
 ```
@@
 ## IOIs at transition points
+```{r, echo = FALSE}
+ioi_tra_trial <- dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+ioi_tra_subject <- ioi_tra_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
+ioi_tra_subject <- ioi_tra_subject[order(SubNr, Condition, Skill)] # ordering
+ioi_tra <- ioi_tra_subject[, .(N = .N, Mean = mean(Mean), Median = median(Mean), SD = sd(Mean), IQR = IQR(Mean)), by = .(Condition, Skill)]
+ioi_tra
 
-```{r, echo = FALSE}
-ioi_tra <- dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
-ioi_tra <- ioi_tra[order(SubNr, Condition, Skill)] # ordering
-
-ggplot(ioi_tra, aes(x = Condition, y = Mean, color = Condition)) +
+ggplot(ioi_tra_subject, aes(x = Condition, y = Mean, color = Condition)) +
   geom_boxplot() +
@@
   geom_point(aes(color = Condition)) +
+  labs(y = "Normalised IOIs") +
   facet_grid(. ~ Skill) +
@@
 
+### Stats
+1) Articulation
+
 ```{r, echo = FALSE}
-# descriptive stats
-ioi_tra[, .(.N, MeanCV = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill)]
+# normality check
+diff3 <- ioi_tra_subject[Condition == "teaching" & Skill == "articulation"]$Mean-ioi_tra_subject[Condition == "performing" & Skill == "articulation"]$Mean
+ioi_art_tra_norm <- shapiro.test(diff3)
+ioi_art_tra_norm
 
-aov_ez(id = "SubNr", dv = "Mean",
-       within = c("Condition", "Skill"),
-       data = ioi_tra)
+# t test
+ioi_art_tra_ttest <- t.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "articulation"], paired = TRUE, alternative = "two.sided")
+ioi_art_tra_ttest
+
+# wilcoxon test
+ioi_art_tra_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "articulation"], paired = TRUE, alternative = "two.sided")
+ioi_art_tra_wilcoxon
+
+# effect size
+coheneff_art_tra <- ioi_tra_subject[Skill == "articulation"] %>% cohens_d(Mean ~ Condition, paired = TRUE)
+coheneff_art_tra
 ```
 
-## IOIs at transition points vs. others
-Caution: the number of points included in each transition group is considerably different
+2) Dynamics
 
 ```{r, echo = FALSE}
-dt_ioi$Transition <- "No"
-dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF"]$Transition <- "Yes"
+# normality check
+diff4 <- ioi_tra_subject[Condition == "teaching" & Skill == "dynamics"]$Mean-ioi_tra_subject[Condition == "performing" & Skill == "dynamics"]$Mean
+ioi_dyn_tra_norm <- shapiro.test(diff4)
+ioi_dyn_tra_norm
 
-# transition points vs. others
-ioi_tra_others_trial <- dt_ioi[, .(.N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Transition)]
-ioi_tra_others_subject <- ioi_tra_others_trial[, .(.N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Transition)]
-ioi_tra_others_subject <- ioi_tra_others_subject[order(SubNr, Condition, Skill)] # ordering
+# wilcoxon test
+ioi_dyn_tra_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
+ioi_dyn_tra_wilcoxon
 
-# descriptive stats
-ioi_tra_others_subject[, .(.N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Transition)]
-
-ggplot(ioi_tra_others_subject, aes(x = Condition, y = Mean, color = Condition)) +
-  geom_boxplot() +
-  geom_line(aes(group = SubNr), size = 0.4, color = "gray") +
-  geom_point(aes(color = Condition)) +
-  facet_grid(Skill ~ Transition) +
-  theme_pubr()
-```
-
-```{r, echo = FALSE}
-aov_ez(id = "SubNr", dv = "Mean",
-       within = c("Condition", "Skill", "Transition"),
-       data = ioi_tra_others_subject)
+# effect size
+wilcoxeff_dyn_tra <- ioi_tra_subject[Skill == "dynamics"] %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
+wilcoxeff_dyn_tra
 ```

### 2022-07-07 experiment-2/analysis/stats/variability.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE-------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -17,15 +17,15 @@ dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
 # Include only articulation
 dt_ioi_art <- dt_ioi[Skill == "articulation"]
 # per participant
-ioi_art_seq_each <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+ioi_art_seq_each <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Interval)]
 
 # Include only dynamics
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 # per participant
-ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
+ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Interval)]
 
 
-## ---- echo = FALSE, fig.width = 6, fig.height = 1.2---------------------------
+## ---- echo = FALSE, fig.width = 6, fig.height = 1.2-------
 # Plot the whole sequence
 ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
   geom_vline(xintercept = c(5, 17, 47, 8, 20, 39), linetype = "twodash") + # transition points
@@ -36,10 +36,12 @@ ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "C
   scale_x_continuous(breaks = seq(1,66,1))
 
 
-## ---- echo = FALSE------------------------------------------------------------
[…]

### 2022-07-07 paper/manuscript.Rmd
add some of variability results (tbc)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ## Data analysis
+The tempo variability of performance was assessed with the coefficient of variation (CV) of the IOIs to in investigate the tempo stability of performance between the teaching and performing condition.
+CVs were calculated by dividing the standard deviation by the mean of IOIs.
 
@@
 
+### Articulation
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_cv_norm$p.value))}`).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not significantly differ between the teaching [*Mdn* = `r ioi_cv[Condition == "teaching" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_cv[Condition == "teaching" & Skill == "articulation"]$IQR`] and performing condition [*Mdn* = `r ioi_cv[Condition == "performing" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_cv[Condition == "performing" & Skill == "articulation"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_cv$effsize, 2)`, two-tailed).
+
+### Dynamics
+
 ## Results (Experiment 2)
@@
 ```
+
+# Temporal grouping

### 2022-07-07 paper/manuscript.tex
add some of variability results (tbc)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+The tempo variability of performance was assessed with the coefficient of variation (CV) of the IOIs to in investigate the tempo stability of performance between the teaching and performing condition.
+CVs were calculated by dividing the standard deviation by the mean of IOIs.
+
 \hypertarget{results-experiment-1}{%
 \subsection{Results (Experiment 1)}\label{results-experiment-1}}
+
+\hypertarget{articulation-2}{%
+\subsubsection{Articulation}\label{articulation-2}}
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.007).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not significantly differ between the teaching {[}\emph{Mdn} = 0.06 (ms), \emph{IQR} = 0.02{]} and performing condition {[}\emph{Mdn} = 0.07 (ms), \emph{IQR} = 0.02{]} while playing the piece with the notated articulation (\emph{p} = 0.058, \emph{r} = 0.34, two-tailed).
+
+\hypertarget{dynamics-2}{%
+\subsubsection{Dynamics}\label{dynamics-2}}
 
@@
 
+\hypertarget{temporal-grouping}{%
+\section{Temporal grouping}\label{temporal-grouping}}
+
 \end{document}

### 2022-07-07 experiment-1/analysis/stats/variability.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE-------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -25,7 +25,7 @@ dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
 
 
-## ---- echo = FALSE, fig.width = 6, fig.height = 1.2-------
+## ---- echo = FALSE, fig.width = 6, fig.height = 1.2---------------------------
 # Plot the whole sequence
 ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
   geom_hline(yintercept = 188, linetype = "dotted") + # target tempo
@@ -38,7 +38,7 @@ ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "C
   scale_x_continuous(breaks = seq(1,66,1))
 
 
-## ---- echo = FALSE----------------------------------------
+## ---- echo = FALSE------------------------------------------------------------
 ioi_cv_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
 ioi_cv_subject <- ioi_cv_trial[, .(N = .N, MeanCV = mean(CV), SD = sd(CV)), by = .(SubNr, Condition, Skill)]
 ioi_cv_subject <- ioi_cv_subject[order(SubNr, Condition, Skill)] # ordering
@@ -54,7 +54,7 @@ ggplot(ioi_cv_subject, aes(x = Condition, y = MeanCV, color = Condition)) +
   theme_pubr()
 
 
-## ---- echo = FALSE----------------------------------------
+## ---- echo = FALSE------------------------------------------------------------
 # normality check
[…]

### 2022-07-07 experiment-2/analysis/stats/variability.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE-------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -25,18 +25,18 @@ dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 ioi_dyn_seq_each <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, Interval)]
 
 
-## ---- echo = FALSE, fig.width = 6, fig.height = 1.2-------
+## ---- echo = FALSE, fig.width = 6, fig.height = 1.2---------------------------
 # Plot the whole sequence
-ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
+ggline(ioi_art_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "Normalised IOIs", title = "IOIs: Articulation") +
   geom_vline(xintercept = c(5, 17, 47, 8, 20, 39), linetype = "twodash") + # transition points
   scale_x_continuous(breaks = seq(1,66,1))
 
-ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "IOIs (ms)", title = "IOIs: Dynamics") +
+ggline(ioi_dyn_seq_each, x = "Interval", y = "Mean", add = "mean_se", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "Normalised IOIs", title = "IOIs: Dynamics") +
   geom_vline(xintercept = c(5, 17, 47, 8, 20, 39), linetype = "twodash") + # transition points
   scale_x_continuous(breaks = seq(1,66,1))
 
 
-## ---- echo = FALSE----------------------------------------
+## ---- echo = FALSE------------------------------------------------------------
[…]

### 2022-07-08 paper/manuscript.Rmd
add variability analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# Variability
+# Tempo Variability
 
@@
 To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_cv_norm$p.value))}`).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not significantly differ between the teaching [*Mdn* = `r ioi_cv[Condition == "teaching" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_cv[Condition == "teaching" & Skill == "articulation"]$IQR`] and performing condition [*Mdn* = `r ioi_cv[Condition == "performing" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_cv[Condition == "performing" & Skill == "articulation"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_cv$effsize, 2)`, two-tailed).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$IQR)`] while playing the piece with the notated articulation (*p* `r if(ioi_art_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_cv$effsize, 2)`, two-tailed).
 
 ### Dynamics
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_cv_norm$p.value))}`).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "dynamics"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "dynamics"]$IQR)`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn_cv$effsize, 2)`, two-tailed).
 
@@
 
+### Articulation
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_cv_norm$p.value))}`).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$IQR)`] while playing the piece with the notated articulation (*p* `r if(ioi_art_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_cv$effsize, 2)`, two-tailed).
+
+### Dynamics
+
+A paired-sample *t*-test showed that the tempo variability of performance did not differ significantly between the teaching [*M* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$SD)`] and the performing condition [*M* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_cv_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_cv_ttest$statistic))`, *p* `r if(ioi_dyn_cv_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_cv_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn_cv$effsize), 3)`, two-tailed, *Fig \ref{fig:ioi-2}*).
+
 # Temporal grouping

### 2022-07-08 paper/manuscript.tex
add variability analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-\hypertarget{variability}{%
-\section{Variability}\label{variability}}
+\hypertarget{tempo-variability}{%
+\section{Tempo Variability}\label{tempo-variability}}
 
@@
 To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.007).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not significantly differ between the teaching {[}\emph{Mdn} = 0.06 (ms), \emph{IQR} = 0.02{]} and performing condition {[}\emph{Mdn} = 0.07 (ms), \emph{IQR} = 0.02{]} while playing the piece with the notated articulation (\emph{p} = 0.058, \emph{r} = 0.34, two-tailed).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.064, \emph{IQR} = 0.024{]} and performing condition {[}\emph{Mdn} = 0.066, \emph{IQR} = 0.016{]} while playing the piece with the notated articulation (\emph{p} = 0.06, \emph{r} = 0.34, two-tailed).
 
@@
 
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.024).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.070, \emph{IQR} = 0.013{]} and performing condition {[}\emph{Mdn} = 0.074, \emph{IQR} = 0.026{]} while playing the piece with the notated dynamics (\emph{p} = 0.22, \emph{r} = 0.22, two-tailed).
+
 \hypertarget{results-experiment-2}{%
 \subsection{Results (Experiment 2)}\label{results-experiment-2}}
+
+\hypertarget{articulation-3}{%
+\subsubsection{Articulation}\label{articulation-3}}
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.017).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.044, \emph{IQR} = 0.012{]} and performing condition {[}\emph{Mdn} = 0.047, \emph{IQR} = 0.014{]} while playing the piece with the notated articulation (\emph{p} = 0.07, \emph{r} = 0.41, two-tailed).
+
+\hypertarget{dynamics-3}{%
+\subsubsection{Dynamics}\label{dynamics-3}}
+
+A paired-sample \emph{t}-test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{M} = 0.048, \emph{SD} = 0.011{]} and the performing condition {[}\emph{M} = 0.048, \emph{SD} = 0.011{]} while playing the piece with the notated dynamics (\emph{t}(19) = 1.83, \emph{p} = 0.08, Cohen's \emph{d} = 0.41, two-tailed, \emph{Fig \ref{fig:ioi-2}}).
 

### 2022-07-08 paper/manuscript.Rmd
add variability plots / correct significance *** on kv-diff plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   ylim(-40, 60) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.4, label.y = 45, method = "t.test", paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.4, label.y = 45, method = "t.test", paired = TRUE, symnum.args = list(cutpoints = c(0, 0.0001, 0.01, 0.05, 1), symbols = c("***", "**", "*", "ns"))) + # adjust significance level according to the actual stats
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+# combine
 img_ioi_2 <- image_append(c(img_ioi_art_2, img_ioi_dyn_2))
@@
 
+```{r, echo = FALSE}
+ioi_cv_plot <- data.table(subject = ioi_cv_subject[Condition == "teaching"]$SubNr, teaching = ioi_cv_subject[Condition == "teaching"]$MeanCV, performing = ioi_cv_subject[Condition == "performing"]$MeanCV, skill = ioi_cv_subject[Condition == "teaching"]$Skill)
+
+ioi_cv_1 <- ggpaired(ioi_cv_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "CVs", title = "CVs: Exp 1") +
+  ylim(0.02, 0.20) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_cv_1.png"), plot = ioi_cv_1, dpi = 300, width = 7, height = 5)
+img_ioi_cv_1 <- image_read(here("paper/image", "ioi_cv_1.png"), density = 300)
+```
+
 ## Results (Experiment 2)
@@
 
+```{r, echo = FALSE}
+ioi_cv_plot <- data.table(subject = ioi_cv_subject[Condition == "teaching"]$SubNr, teaching = ioi_cv_subject[Condition == "teaching"]$MeanCV, performing = ioi_cv_subject[Condition == "performing"]$MeanCV, skill = ioi_cv_subject[Condition == "teaching"]$Skill)
+
+ioi_cv_2 <- ggpaired(ioi_cv_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "CVs", title = "CVs: Exp 2") +
+  ylim(0.02, 0.20) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_cv_2.png"), plot = ioi_cv_2, dpi = 300, width = 7, height = 5)
+img_ioi_cv_2 <- image_read(here("paper/image", "ioi_cv_2.png"), density = 300)
+```
+
+```{r plot-cv-1, out.width = "100%", fig.cap = "\\label{fig:vel-2}Experiment 1, 2: CVs when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+img_ioi_cv <- image_append(c(img_ioi_cv_1, img_ioi_cv_2))
+img_ioi_cv <- image_scale(img_ioi_cv, "2200")
+image_write(img_ioi_cv, path = here("paper/figure", "Fig10.tiff"), format = "tiff")
+img_ioi_cv
+```
+
 # Temporal grouping

### 2022-07-08 experiment-1/analysis/stats/variability.Rmd
correct typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # t test
-ioi_dyn_tra_t_test <- t.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
-ioi_dyn_tra_t_test
+ioi_dyn_tra_ttest <- t.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
+ioi_dyn_tra_ttest
 
@@
 # effect size
-wilcoxeff_dyn_tra <- ioi_tra_subject[Skill == "dynamics"] %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
-wilcoxeff_dyn_tra
+coheneff_dyn_tra <- ioi_tra_subject[Skill == "dynamics"] %>% cohens_d(Mean ~ Condition, paired = TRUE)
+coheneff_dyn_tra
 ```

### 2022-07-08 experiment-1/analysis/stats/variability.R

@@ -122,16 +122,16 @@ ioi_dyn_tra_norm <- shapiro.test(diff4)
 ioi_dyn_tra_norm
 
 # t test
-ioi_dyn_tra_t_test <- t.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
-ioi_dyn_tra_t_test
+ioi_dyn_tra_ttest <- t.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
+ioi_dyn_tra_ttest
 
 # wilcoxon test
 ioi_dyn_tra_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_tra_subject[Skill == "dynamics"], paired = TRUE, alternative = "two.sided")
 ioi_dyn_tra_wilcoxon
 
 # effect size
-wilcoxeff_dyn_tra <- ioi_tra_subject[Skill == "dynamics"] %>% wilcox_effsize(Mean ~ Condition, paired = TRUE)
-wilcoxeff_dyn_tra
+coheneff_dyn_tra <- ioi_tra_subject[Skill == "dynamics"] %>% cohens_d(Mean ~ Condition, paired = TRUE)
+coheneff_dyn_tra
 
 
 ## ----export, include = FALSE--------------------------------------------------

### 2022-07-08 paper/manuscript.Rmd
add tempo grouping analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-## Data analysis
-The tempo variability of performance was assessed with the coefficient of variation (CV) of the IOIs to in investigate the tempo stability of performance between the teaching and performing condition.
+The tempo variability of performance was assessed with the coefficient of variation (CV) of the IOIs to investigate the tempo stability of performance between the teaching and performing condition.
 CVs were calculated by dividing the standard deviation by the mean of IOIs.
 
-## Results (Experiment 1)
+## Results
+## Experiment 1
 ```{r variability-1, include = FALSE}
@@
 To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_cv_norm$p.value))}`).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$IQR)`] while playing the piece with the notated articulation (*p* `r if(ioi_art_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_cv$effsize, 2)`, two-tailed).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$IQR)`] while playing the piece with the notated articulation (*p* `r if(ioi_art_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_cv$effsize, 2)`, two-tailed, *Fig \ref{fig:cv}*).
 
@@
 To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_cv_norm$p.value))}`).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "dynamics"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "dynamics"]$IQR)`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn_cv$effsize, 2)`, two-tailed).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "dynamics"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "dynamics"]$IQR)`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn_cv$effsize, 2)`, two-tailed, *Fig \ref{fig:cv}*).
 
-```{r, echo = FALSE}
+```{r ioi-cv-1, echo = FALSE}
 ioi_cv_plot <- data.table(subject = ioi_cv_subject[Condition == "teaching"]$SubNr, teaching = ioi_cv_subject[Condition == "teaching"]$MeanCV, performing = ioi_cv_subject[Condition == "performing"]$MeanCV, skill = ioi_cv_subject[Condition == "teaching"]$Skill)
@@
 
-## Results (Experiment 2)
+## Experiment 2
 ```{r variability-2, include = FALSE}
@@
 
-A paired-sample *t*-test showed that the tempo variability of performance did not differ significantly between the teaching [*M* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$SD)`] and the performing condition [*M* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_cv_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_cv_ttest$statistic))`, *p* `r if(ioi_dyn_cv_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_cv_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn_cv$effsize), 3)`, two-tailed, *Fig \ref{fig:ioi-2}*).
+A paired-sample *t*-test showed that the tempo variability of performance did not differ significantly between the teaching [*M* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$SD)`] and the performing condition [*M* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_cv_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_cv_ttest$statistic))`, *p* `r if(ioi_dyn_cv_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_cv_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn_cv$effsize), 3)`, two-tailed, *Fig \ref{fig:cv}*).
 
-```{r, echo = FALSE}
+```{r ioi-cv-2, echo = FALSE}
 ioi_cv_plot <- data.table(subject = ioi_cv_subject[Condition == "teaching"]$SubNr, teaching = ioi_cv_subject[Condition == "teaching"]$MeanCV, performing = ioi_cv_subject[Condition == "performing"]$MeanCV, skill = ioi_cv_subject[Condition == "teaching"]$Skill)
@@
 
-```{r plot-cv-1, out.width = "100%", fig.cap = "\\label{fig:vel-2}Experiment 1, 2: CVs when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+```{r plot-cv, out.width = "100%", fig.cap = "\\label{fig:cv}Experiment 1, 2: CVs when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
@@
 # Temporal grouping
+
+In order to investigated if participants used temporal grouping to make a boundary between each subcomponent (e.g., between legato and staccato, between forte and piano) clearer for teaching purposes, we looked at IOIs only at transition points.
+
+## Results
+## Experiment 1
+```{r variability-1-1, include = FALSE}
+source(here("experiment-1/analysis/stats", "variability.R"), chdir = TRUE)
+```
+
+### Articulation
+
+To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_tra_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_tra_norm$p.value))}`).
+The Wilcoxon Signed-rank test revealed that IOIs at transition points were larger for the teaching condition [*Mdn* = `r ioi_tra[Condition == "teaching" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_tra[Condition == "teaching" & Skill == "articulation"]$IQR`] than in the performing condition [*Mdn* = `r ioi_tra[Condition == "performing" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_tra[Condition == "performing" & Skill == "articulation"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_tra_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_tra_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_tra$effsize, 2)`, two-tailed, *Fig \ref{fig:tra}*).
+
+### Dynamics
+
+A paired-sample *t*-test showed that IOIs at transition points were larger for the teaching condition [*M* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$Mean` (ms), *SD* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$SD`] than the performing condition [*M* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$Mean` (ms), *SD* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$SD`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_tra_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_tra_ttest$statistic))`, *p* `r if(ioi_dyn_tra_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_tra_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn_tra$effsize), 3)`, two-tailed, *Fig \ref{fig:tra}*).
+
+```{r ioi-tra-1, echo = FALSE}
+ioi_tra_plot <- data.table(subject = ioi_tra_subject[Condition == "teaching"]$SubNr, teaching = ioi_tra_subject[Condition == "teaching"]$Mean, performing = ioi_tra_subject[Condition == "performing"]$Mean, skill = ioi_tra_subject[Condition == "teaching"]$Skill)
+
+ioi_tra_1 <- ggpaired(ioi_tra_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs at transition: Exp 1") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_tra_1.png"), plot = ioi_tra_1, dpi = 300, width = 7, height = 5)
+img_ioi_tra_1 <- image_read(here("paper/image", "ioi_tra_1.png"), density = 300)
+```
+
+## Experiment 2
+### Articulation
+```{r variability-2-2, include = FALSE}
+source(here("experiment-2/analysis/stats", "variability.R"), chdir = TRUE)
+```
+
+A paired-sample *t*-test showed that IOIs at transition points were larger for the teaching condition [*M* = `r sprintf('%.2f', ioi_tra[Condition == "teaching" & Skill == "articulation"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_tra[Condition == "teaching" & Skill == "articulation"]$SD)`] than the performing condition [*M* = `r sprintf('%.2f', ioi_tra[Condition == "performing" & Skill == "articulation"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_tra[Condition == "performing" & Skill == "articulation"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_tra_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_tra_ttest$statistic))`, *p* `r if(ioi_art_tra_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_tra_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art_tra$effsize), 3)`, two-tailed, *Fig \ref{fig:tra}*).
+
+### Dynamics
+
+To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_tra_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_tra_norm$p.value))}`).
+The Wilcoxon Signed-rank test revealed that IOIs at transition points did not differ significantly between the teaching [*Mdn* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$Median`, *IQR* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$IQR`] and the performing condition [*Mdn* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$Median`, *IQR* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_tra_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_tra_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn_tra$effsize, 2)`, two-tailed, *Fig \ref{fig:tra}*).
+
+```{r ioi-tra-2, echo = FALSE}
+ioi_tra_plot <- data.table(subject = ioi_tra_subject[Condition == "teaching"]$SubNr, teaching = ioi_tra_subject[Condition == "teaching"]$Mean, performing = ioi_tra_subject[Condition == "performing"]$Mean, skill = ioi_tra_subject[Condition == "teaching"]$Skill)
+
+ioi_tra_2 <- ggpaired(ioi_tra_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOIs at transition: Exp 2") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_tra_2.png"), plot = ioi_tra_2, dpi = 300, width = 7, height = 5)
+img_ioi_tra_2 <- image_read(here("paper/image", "ioi_tra_2.png"), density = 300)
+```
+
+```{r plot-tra, out.width = "100%", fig.cap = "\\label{fig:tra}Experiment 1, 2: IOIs at transition points when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+img_ioi_tra <- image_append(c(img_ioi_tra_1, img_ioi_tra_2))
+img_ioi_tra <- image_scale(img_ioi_tra, "2200")
+image_write(img_ioi_tra, path = here("paper/figure", "Fig10.tiff"), format = "tiff")
+img_ioi_tra
+```

### 2022-07-08 paper/manuscript.tex
add variability plots / correct significance *** on kv-diff plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
add tempo grouping analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-07-07 paper/manuscript.tex
@@
 
-\hypertarget{variability}{%
-\section{Variability}\label{variability}}
+\hypertarget{tempo-variability}{%
+\section{Tempo Variability}\label{tempo-variability}}
 
-\hypertarget{data-analysis-2}{%
-\subsection{Data analysis}\label{data-analysis-2}}
-
-The tempo variability of performance was assessed with the coefficient of variation (CV) of the IOIs to in investigate the tempo stability of performance between the teaching and performing condition.
+The tempo variability of performance was assessed with the coefficient of variation (CV) of the IOIs to investigate the tempo stability of performance between the teaching and performing condition.
 CVs were calculated by dividing the standard deviation by the mean of IOIs.
 
-\hypertarget{results-experiment-1}{%
-\subsection{Results (Experiment 1)}\label{results-experiment-1}}
+\hypertarget{results-2}{%
+\subsection{Results}\label{results-2}}
+
+\hypertarget{experiment-1-2}{%
+\subsection{Experiment 1}\label{experiment-1-2}}
 
@@
 To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.007).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not significantly differ between the teaching {[}\emph{Mdn} = 0.06 (ms), \emph{IQR} = 0.02{]} and performing condition {[}\emph{Mdn} = 0.07 (ms), \emph{IQR} = 0.02{]} while playing the piece with the notated articulation (\emph{p} = 0.058, \emph{r} = 0.34, two-tailed).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.064, \emph{IQR} = 0.024{]} and performing condition {[}\emph{Mdn} = 0.066, \emph{IQR} = 0.016{]} while playing the piece with the notated articulation (\emph{p} = 0.06, \emph{r} = 0.34, two-tailed, \emph{Fig \ref{fig:cv}}).
 
@@
 
-\hypertarget{results-experiment-2}{%
-\subsection{Results (Experiment 2)}\label{results-experiment-2}}
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.024).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.070, \emph{IQR} = 0.013{]} and performing condition {[}\emph{Mdn} = 0.074, \emph{IQR} = 0.026{]} while playing the piece with the notated dynamics (\emph{p} = 0.22, \emph{r} = 0.22, two-tailed, \emph{Fig \ref{fig:cv}}).
+
+\hypertarget{experiment-2-2}{%
+\subsection{Experiment 2}\label{experiment-2-2}}
+
+\hypertarget{articulation-3}{%
+\subsubsection{Articulation}\label{articulation-3}}
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.017).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.044, \emph{IQR} = 0.012{]} and performing condition {[}\emph{Mdn} = 0.047, \emph{IQR} = 0.014{]} while playing the piece with the notated articulation (\emph{p} = 0.07, \emph{r} = 0.41, two-tailed, \emph{Fig \ref{fig:cv}}).
+
+\hypertarget{dynamics-3}{%
+\subsubsection{Dynamics}\label{dynamics-3}}
+
+A paired-sample \emph{t}-test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{M} = 0.048, \emph{SD} = 0.011{]} and the performing condition {[}\emph{M} = 0.048, \emph{SD} = 0.011{]} while playing the piece with the notated dynamics (\emph{t}(19) = 1.83, \emph{p} = 0.08, Cohen's \emph{d} = 0.41, two-tailed, \emph{Fig \ref{fig:cv}}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-cv-1} \caption{\label{fig:cv}Experiment 1, 2: CVs when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-cv}
+\end{figure}
 
@@
 
+In order to investigated if participants used temporal grouping to make a boundary between each subcomponent (e.g., between legato and staccato, between forte and piano) clearer for teaching purposes, we looked at IOIs only at transition points.
+
+\hypertarget{results-3}{%
+\subsection{Results}\label{results-3}}
+
+\hypertarget{experiment-1-3}{%
+\subsection{Experiment 1}\label{experiment-1-3}}
+
+\hypertarget{articulation-4}{%
+\subsubsection{Articulation}\label{articulation-4}}
+
+To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
+The Wilcoxon Signed-rank test revealed that IOIs at transition points were larger for the teaching condition {[}\emph{Mdn} = 195.75 (ms), \emph{IQR} = 25.75{]} than in the performing condition {[}\emph{Mdn} = 188.77 (ms), \emph{IQR} = 23.53{]} while playing the piece with the notated articulation (\emph{p} = 0.02, \emph{r} = 0.42, two-tailed, \emph{Fig \ref{fig:tra}}).
+
+\hypertarget{dynamics-4}{%
+\subsubsection{Dynamics}\label{dynamics-4}}
+
+A paired-sample \emph{t}-test showed that IOIs at transition points were larger for the teaching condition {[}\emph{M} = 200.90 (ms), \emph{SD} = 20.96{]} than the performing condition {[}\emph{M} = 197.63 (ms), \emph{SD} = 19.30{]} while playing the piece with the notated dynamics (\emph{t}(30) = 2.10, \emph{p} = 0.04, Cohen's \emph{d} = 0.38, two-tailed, \emph{Fig \ref{fig:tra}}).
+
+\hypertarget{experiment-2-3}{%
+\subsection{Experiment 2}\label{experiment-2-3}}
+
+\hypertarget{articulation-5}{%
+\subsubsection{Articulation}\label{articulation-5}}
+
+A paired-sample \emph{t}-test showed that IOIs at transition points were larger for the teaching condition {[}\emph{M} = 1.00, \emph{SD} = 0.067{]} than the performing condition {[}\emph{M} = 0.98, \emph{SD} = 0.049{]} while playing the piece with the notated articulation (\emph{t}(19) = 2.27, \emph{p} = 0.03, Cohen's \emph{d} = 0.51, two-tailed, \emph{Fig \ref{fig:tra}}).
+
+\hypertarget{dynamics-5}{%
+\subsubsection{Dynamics}\label{dynamics-5}}
+
+To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.004).
+The Wilcoxon Signed-rank test revealed that IOIs at transition points did not differ significantly between the teaching {[}\emph{Mdn} = 1.00, \emph{IQR} = 0.06{]} and the performing condition {[}\emph{Mdn} = 0.99, \emph{IQR} = 0.05{]} while playing the piece with the notated dynamics (\emph{p} = 0.60, \emph{r} = 0.13, two-tailed, \emph{Fig \ref{fig:tra}}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-tra-1} \caption{\label{fig:tra}Experiment 1, 2: IOIs at transition points when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-tra}
+\end{figure}
+
 \end{document}

### 2022-07-10 paper/manuscript.Rmd
non-expressive >> original / update ioi_tra plot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations, *Fig \ref{fig:stimuli} A*) was used for the purpose of practice.
-Expressive notations were added to the non-expressive sheet music for the experiment.
+Original sheet music (i.e., sheet music without expressive notations, *Fig \ref{fig:stimuli} A*) was used for the purpose of practice.
+Expressive notations were added to the original sheet music for the experiment.
 They referred to either articulation or dynamics (*Fig \ref{fig:stimuli} B-C*).
@@
 
-```{r stim-1, out.width = "100%", fig.cap = "\\label{fig:stimuli}(A)Non-expressive sheet music.
+```{r stim-1, out.width = "100%", fig.cap = "\\label{fig:stimuli}(A)Original sheet music.
 (B)Articulation.
@@
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Non-expressive sheet music was used for the purpose of practice (*Fig \ref{fig:stimuli-2} A*).
-Expressive notations were added to the non-expressive sheet music for the experiment (*Fig \ref{fig:stimuli-2} B-C*).
+Original sheet music was used for the purpose of practice (*Fig \ref{fig:stimuli-2} A*).
+Expressive notations were added to the original sheet music for the experiment (*Fig \ref{fig:stimuli-2} B-C*).
 These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
@@
 The chosen tempo was again cued by a leading metronome.
-To make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music (*Fig \ref{fig:stimuli-2} A*).
+To make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the original sheet music (*Fig \ref{fig:stimuli-2} A*).
 Those who could not perform the piece without pitch errors twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded).
@@
 
-```{r stim-2, out.width = "100%", fig.cap = "\\label{fig:stimuli-2}(A)Non-expressive sheet music.
+```{r stim-2, out.width = "100%", fig.cap = "\\label{fig:stimuli-2}(A)Original sheet music.
 (B)Articulation.
@@
 ioi_tra_1 <- ggpaired(ioi_tra_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs at transition: Exp 1") +
+  geom_hline(yintercept = 188, linetype = "dashed") +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 ioi_tra_2 <- ggpaired(ioi_tra_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOIs at transition: Exp 2") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 ```{r plot-tra, out.width = "100%", fig.cap = "\\label{fig:tra}Experiment 1, 2: IOIs at transition points when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+A dashed line represents the tempo given by a metronome.
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.

### 2022-07-10 paper/manuscript.tex
non-expressive >> original / update ioi_tra plot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The stimulus was composed in C major to be played with the right hand only.
-Non-expressive sheet music (i.e., sheet music without expressive notations, \emph{Fig \ref{fig:stimuli} A}) was used for the purpose of practice.
-Expressive notations were added to the non-expressive sheet music for the experiment.
+Original sheet music (i.e., sheet music without expressive notations, \emph{Fig \ref{fig:stimuli} A}) was used for the purpose of practice.
+Expressive notations were added to the original sheet music for the experiment.
 They referred to either articulation or dynamics (\emph{Fig \ref{fig:stimuli} B-C}).
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stim-1-1} \caption{\label{fig:stimuli}(A)Non-expressive sheet music.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stim-1-1} \caption{\label{fig:stimuli}(A)Original sheet music.
 (B)Articulation.
@@
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Non-expressive sheet music was used for the purpose of practice (\emph{Fig \ref{fig:stimuli-2} A}).
-Expressive notations were added to the non-expressive sheet music for the experiment (\emph{Fig \ref{fig:stimuli-2} B-C}).
+Original sheet music was used for the purpose of practice (\emph{Fig \ref{fig:stimuli-2} A}).
+Expressive notations were added to the original sheet music for the experiment (\emph{Fig \ref{fig:stimuli-2} B-C}).
 These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
@@
 The chosen tempo was again cued by a leading metronome.
-To make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the non-expressive sheet music (\emph{Fig \ref{fig:stimuli-2} A}).
+To make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the original sheet music (\emph{Fig \ref{fig:stimuli-2} A}).
 Those who could not perform the piece without pitch errors twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded).
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stim-2-1} \caption{\label{fig:stimuli-2}(A)Non-expressive sheet music.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stim-2-1} \caption{\label{fig:stimuli-2}(A)Original sheet music.
 (B)Articulation.
@@
 \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-tra-1} \caption{\label{fig:tra}Experiment 1, 2: IOIs at transition points when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+A dashed line represents the tempo given by a metronome.
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.

### 2022-07-27 experiment-1/analysis/stats/dynamics.Rmd
correct typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
remove unnessesary tags
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update dynamics analysis to remove warnings
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
-
-# chunk option
-knitr::opts_chunk$set(warning = FALSE)
 ```
@@
 ```{r vel-dyn-box,  echo = FALSE}
-ggboxplot(vel_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(vel_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") + facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r vel-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
-ggline(dt_vel_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(dt_vel_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
-ggline(dt_vel_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(dt_vel_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-diff-dyn-box,  echo = FALSE}
-ggboxplot(vel_diff_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Dynamics") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(vel_diff_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Dynamics") + facet_grid(Subcomponent ~ .)
 ```
@@
 ```{r vel-diff-dyn-trial-line-ftop, echo = FALSE, fig.height = 4}
-ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-diff-dyn-trial-line-ptof, echo = FALSE, fig.height = 4}
-ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-diff-dyn-trial-all-line-ftop, echo = FALSE}
-ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-diff-dyn-trial-all-line-ptof, echo = FALSE}
-ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
-  scale_x_continuous(breaks = seq(1,8,1))
+ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks = seq(1,8,1))
 ```
@@
 ```{r vel-diff-art-box,  echo = FALSE}
-ggboxplot(vel_diff_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
-  facet_grid(Subcomponent ~ .)
+ggboxplot(vel_diff_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") + facet_grid(Subcomponent ~ .)
 ```
@@
 
-ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Dynamics") +
-  scale_x_continuous(breaks = seq(1,67,1))
+ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Dynamics") + scale_x_continuous(breaks = seq(1,67,1))
 
@@
 
-ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") +
-  scale_x_continuous(breaks = seq(1,66,1))
+ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") + scale_x_continuous(breaks = seq(1,66,1))
 ```
@@
 
-ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Articulation") +
-  scale_x_continuous(breaks = seq(1,67,1))
+ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Articulation") + scale_x_continuous(breaks = seq(1,67,1))
 
@@
 
-ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") +
-  scale_x_continuous(breaks = seq(1,66,1))
+ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") + scale_x_continuous(breaks = seq(1,66,1))
 ```
@@
 
-```{r, echo = FALSE, warning = FALSE}
+```{r, echo = FALSE}
 # normalise diff by KV range
@@
 for (i in unique(vel_diff_dyn_trial$SubNr)){
-  for (m in unique(vel_diff_dyn_trial[SubNr == i]$TrialNr)){
-    # teaching
+  for (m in unique(vel_diff_dyn_trial[SubNr == i]$BlockNr)){
+    for (n in unique(vel_diff_dyn_trial[SubNr == i & BlockNr == m]$TrialNr)){
       # FtoP
-    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))
+      vel_diff_dyn_trial[SubNr == i & BlockNr == m & TrialNr == n & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & BlockNr == m & TrialNr == n & Subcomponent == "FtoP"]$Mean/(max(vel_dyn_trial[SubNr == i & BlockNr == m & TrialNr == n]$Mean)-min(vel_dyn_trial[SubNr == i & BlockNr == m & TrialNr == n]$Mean))
       # PtoF
-    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))
-
-    # performing
-    # FtoP
-    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))
-    # PtoF
-    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))
-
-    # error messages for non-existing trials but it's not reflected on vel_diff_dyn_trial anyway
+      vel_diff_dyn_trial[SubNr == i & BlockNr == m & TrialNr == n & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & BlockNr == m & TrialNr == n & Subcomponent == "PtoF"]$Mean/(max(vel_dyn_trial[SubNr == i & BlockNr == m & TrialNr == n]$Mean)-min(vel_dyn_trial[SubNr == i & BlockNr == m & TrialNr == n]$Mean))
+    }
   }

### 2022-07-27 experiment-2/analysis/stats/dynamics.Rmd
update dynamics analysis to remove warnings
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-06-27 experiment-2/analysis/stats/dynamics.Rmd
@@
 ## Normalised KV difference per trial
-
-```{r, echo = FALSE, warning = FALSE}
+```{r, echo = FALSE}
 # normalise diff by KV range

### 2022-07-27 experiment-2/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE-------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -12,12 +12,12 @@ if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 knitr::opts_chunk$set(warning = FALSE)
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE--------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-----------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -34,7 +34,7 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE----------------------------------------------------
+## ----vel-dyn, echo = FALSE--------------------------------
 # For each individual
 vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
 
@@ -42,65 +42,65 @@ vel_dyn <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(S
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE-----------------------------------------------
+## ----vel-dyn-box,  echo = FALSE---------------------------
 ggboxplot(vel_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") + facet_grid(Subcomponent ~ .)
 
 
[…]

### 2022-07-27 experiment-1/analysis/stats/ioi.Rmd
remove chunk setting to suppress warnings
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
-
-# chunk option
-knitr::opts_chunk$set(warning = FALSE)
 ```

### 2022-07-27 experiment-1/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE-------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -8,16 +8,13 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
-# chunk option
-knitr::opts_chunk$set(warning = FALSE)
 
-
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE--------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-----------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -49,7 +46,7 @@ dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 
 
-## ----kot-art, echo = FALSE----------------------------------------------------
+## ----kot-art, echo = FALSE--------------------------------
 # For each individual
 kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
 
@@ -57,65 +54,65 @@ kot_art <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(S
 kot_art
 
 
-## ----kot-art-box,  echo = FALSE-----------------------------------------------
+## ----kot-art-box,  echo = FALSE---------------------------
[…]

### 2022-07-27 paper/manuscript.Rmd
edit .gitignore
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
add additional analysis for dynamics difference
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+\clearpage
+
 # Temporal grouping
@@
 ```
+
+\clearpage
+
+# Additional analysis for dynamics difference
+## Experiment 1
+```{r additional-dynamics-1, include = FALSE}
+source(here("experiment-1/analysis/stats", "dynamics.R"), chdir = TRUE)
+```
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Transition Type (FtoP vs.
+PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig TBC*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (*p* `r if(vel_diff_dyn_posthoc_2$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[1]))}`) and from piano to forte (*p* `r if(vel_diff_dyn_posthoc_2$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[2]))}`) between the teaching condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] and the performing condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
+
+## Experiment 2
+```{r additional-dynamics-2, include = FALSE}
+source(here("experiment-2/analysis/stats", "dynamics.R"), chdir = TRUE)
+```
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Transition Type (FtoP vs.
+PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`).
+However, there was no significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig TBC*).

### 2022-07-27 paper/manuscript.tex
add additional analysis for dynamics difference
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+\hypertarget{additional-analysis-for-dynamics-difference}{%
+\section{Additional analysis for dynamics difference}\label{additional-analysis-for-dynamics-difference}}
+
+\hypertarget{experiment-1-4}{%
+\subsection{Experiment 1}\label{experiment-1-4}}
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,30) = 741, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.88) and a significant interaction between Condition and Transition Type (\emph{F}(1,30) = 7.87, \emph{p} = 0.009, \(\eta_G^2\) = 0.009, \emph{Fig TBC}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (\emph{p} = 0.085) and from piano to forte (\emph{p} = 0.075) between the teaching condition {[}FtoP: \emph{M} = -0.62, \emph{SD} = 0.24, PtoF: \emph{M} = 1.10, \emph{SD} = 0.34{]} and the performing condition {[}FtoP: \emph{M} = -0.56, \emph{SD} = 0.25, PtoF: \emph{M} = 1.03, \emph{SD} = 0.40{]}.
+
+\hypertarget{experiment-2-4}{%
+\subsection{Experiment 2}\label{experiment-2-4}}
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,19) = 42, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.63).
+However, there was no significant interaction between Condition and Transition Type (\emph{F}(1,19) = 1.10, \emph{p} = 0.306, \(\eta_G^2\) = 0.008, \emph{Fig TBC}).
+
 \end{document}

### 2022-07-27 experiment-1/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE-------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE--------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-----------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -46,7 +46,7 @@ dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 
 
-## ----kot-art, echo = FALSE--------------------------------
+## ----kot-art, echo = FALSE----------------------------------------------------
 # For each individual
 kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
 
@@ -54,65 +54,65 @@ kot_art <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(S
 kot_art
 
 
-## ----kot-art-box,  echo = FALSE---------------------------
+## ----kot-art-box,  echo = FALSE-----------------------------------------------
 ggboxplot(kot_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KOT (ms)", title = "KOT: Articulation") + facet_grid(Subcomponent ~ .)
 
 
-## ----kot-art-all, echo = FALSE----------------------------
[…]

### 2022-07-27 experiment-1/analysis/stats/ioi.R

@@ -8,9 +8,6 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 
-# chunk option
-knitr::opts_chunk$set(warning = FALSE)
-
 
 ## ----file, include = FALSE----------------------------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"

### 2022-07-27 experiment-2/analysis/stats/articulation.R

@@ -8,9 +8,6 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
-# chunk option
-knitr::opts_chunk$set(warning = FALSE)
-
 
 ## ----file, include = FALSE----------------------------------------------------
 filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"

### 2022-07-27 experiment-2/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE-------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -8,16 +8,13 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
-# chunk option
-knitr::opts_chunk$set(warning = FALSE)
 
-
-## ----file, include = FALSE--------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE-----------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -34,7 +31,7 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE--------------------------------
+## ----vel-dyn, echo = FALSE----------------------------------------------------
 # For each individual
 vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
 
@@ -42,65 +39,65 @@ vel_dyn <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(S
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE---------------------------
+## ----vel-dyn-box,  echo = FALSE-----------------------------------------------
[…]

### 2022-07-27 paper/manuscript.Rmd
update additional analysis for dynamis difference
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r ioi-art-1, include = FALSE}
-ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean)
-
 ioi_art_1 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
@@
 ```{r kot-art-1, include = FALSE}
-kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent)
-
 kot_art_1 <- ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation") +
@@
 ```{r vel-art-1, echo = FALSE}
-vel_art_plot <- data.table(subject = vel_art[Condition == "teaching"]$SubNr, teaching = vel_art[Condition == "teaching"]$Mean, performing = vel_art[Condition == "performing"]$Mean, subcomponent = vel_art[Condition == "teaching"]$Subcomponent)
-
 vel_art_1 <- ggpaired(vel_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation") +
@@
 ```{r ioi-dyn-1, include = FALSE}
-ioi_dyn_plot <- data.table(subject = ioi_dyn[Condition == "teaching"]$SubNr, teaching = ioi_dyn[Condition == "teaching"]$Mean, performing = ioi_dyn[Condition == "performing"]$Mean)
-
 ioi_dyn_1 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Dynamics") +
@@
 ```{r vel-dyn-1, echo = FALSE}
-vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent)
-
 vel_dyn_1 <- ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
@@
 ```{r vel-diff-dyn-1, echo = FALSE}
-vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)
-
 vel_diff_dyn_1 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Exp 1") +
@@
 ```{r kot-dyn-1, include = FALSE}
-kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent)
-
 kot_dyn_1 <- ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics") +
@@
 ```{r ioi-art-2, echo = FALSE}
-ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean)
-
 ioi_art_2 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOIs: Articulation") +
@@
 ```{r kor-art-2, include = FALSE}
-kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent)
-
 kor_art_2 <- ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KORs", title = "KORs: Articulation") +
@@
 ```{r vel-art-2, echo = FALSE}
-vel_art_plot <- data.table(subject = vel_art[Condition == "teaching"]$SubNr, teaching = vel_art[Condition == "teaching"]$Mean, performing = vel_art[Condition == "performing"]$Mean, subcomponent = vel_art[Condition == "teaching"]$Subcomponent)
-
 vel_art_2 <- ggpaired(vel_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation") +
@@
 ```{r ioi-dyn-2, echo = FALSE}
-ioi_dyn_plot <- data.table(subject = ioi_dyn[Condition == "teaching"]$SubNr, teaching = ioi_dyn[Condition == "teaching"]$Mean, performing = ioi_dyn[Condition == "performing"]$Mean)
-
 ioi_dyn_2 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOIs: Dynamics") +
@@
 ```{r vel-dyn-2, echo = FALSE}
-vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent)
-
 vel_dyn_2 <- ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics") +
@@
 ```{r vel-diff-dyn-2, echo = FALSE}
-vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)
-
 vel_diff_dyn_2 <- ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Exp 2") +
@@
 ```{r kor-dyn-2, include = FALSE}
-kor_dyn_plot <- data.table(subject = kor_dyn[Condition == "teaching"]$SubNr, teaching = kor_dyn[Condition == "teaching"]$Mean, performing = kor_dyn[Condition == "performing"]$Mean, subcomponent = kor_dyn[Condition == "teaching"]$Subcomponent)
-
 kor_dyn_2 <- ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KORs", title = "KORs: Dynamics") +
@@
 img_ioi_cv <- image_scale(img_ioi_cv, "2200")
-image_write(img_ioi_cv, path = here("paper/figure", "Fig10.tiff"), format = "tiff")
+image_write(img_ioi_cv, path = here("paper/figure", "Fig12.tiff"), format = "tiff")
 img_ioi_cv
@@
 # Additional analysis for dynamics difference
+
+## Results
 ## Experiment 1
@@
 performing) and Transition Type (FtoP vs.
-PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig TBC*).
+PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig \ref{fig:vel-diff-2}*).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (*p* `r if(vel_diff_dyn_posthoc_2$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[1]))}`) and from piano to forte (*p* `r if(vel_diff_dyn_posthoc_2$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[2]))}`) between the teaching condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] and the performing condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
+
+```{r vel-diff-dyn-normalised-1, echo = FALSE}
+vel_diff_dyn_normalised_1 <- ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas",xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff: Dynamics")
+
+# save plots
+ggsave(here("paper/image", "vel_diff_dyn_normalised_1.png"), plot = vel_diff_dyn_normalised_1, dpi = 300, width = 7, height = 5)
+img_vel_diff_dyn_normalised_1 <- image_read(here("paper/image", "vel_diff_dyn_normalised_1.png"), density = 300)
+```
 
@@
 PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`).
-However, there was no significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig TBC*).
+However, there was no significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig \ref{fig:vel-diff-2}*).
+
+```{r vel-diff-dyn-normalised-2, echo = FALSE}
+vel_diff_dyn_normalised_2 <- ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff: Dynamics")
+
+# save plots
+ggsave(here("paper/image", "vel_diff_dyn_normalised_2.png"), plot = vel_diff_dyn_normalised_2, dpi = 300, width = 7, height = 5)
+img_vel_diff_dyn_normalised_2 <- image_read(here("paper/image", "vel_diff_dyn_normalised_2.png"), density = 300)
+```
+
+```{r plot-vel-diff-2, out.width = "100%", fig.cap = "\\label{fig:vel-diff-2}Experiment 1, 2: Normalised KV Difference at transition points when playing the piece with dynamics (left: Experiment 1, right: Experiment 2).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+img_vel_diff_dyn_normalised <- image_append(c(img_vel_diff_dyn_normalised_1, img_vel_diff_dyn_normalised_2))
+img_vel_diff_dyn_normalised <- image_scale(img_vel_diff_dyn_normalised, "2200")
+image_write(img_vel_diff_dyn_normalised, path = here("paper/figure", "Fig11.tiff"), format = "tiff")
+img_vel_diff_dyn_normalised
+```

### 2022-07-27 paper/manuscript.tex
update additional analysis for dynamis difference
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+\clearpage
+
 \hypertarget{temporal-grouping}{%
@@
 
+\clearpage
+
 \hypertarget{additional-analysis-for-dynamics-difference}{%
 \section{Additional analysis for dynamics difference}\label{additional-analysis-for-dynamics-difference}}
+
+\hypertarget{results-4}{%
+\subsection{Results}\label{results-4}}
 
@@
 
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,30) = 741, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.88) and a significant interaction between Condition and Transition Type (\emph{F}(1,30) = 7.87, \emph{p} = 0.009, \(\eta_G^2\) = 0.009, \emph{Fig TBC}).
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,30) = 741, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.88) and a significant interaction between Condition and Transition Type (\emph{F}(1,30) = 7.87, \emph{p} = 0.009, \(\eta_G^2\) = 0.009, \emph{Fig \ref{fig:vel-diff-2}}).
 Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (\emph{p} = 0.085) and from piano to forte (\emph{p} = 0.075) between the teaching condition {[}FtoP: \emph{M} = -0.62, \emph{SD} = 0.24, PtoF: \emph{M} = 1.10, \emph{SD} = 0.34{]} and the performing condition {[}FtoP: \emph{M} = -0.56, \emph{SD} = 0.25, PtoF: \emph{M} = 1.03, \emph{SD} = 0.40{]}.
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,19) = 42, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.63).
-However, there was no significant interaction between Condition and Transition Type (\emph{F}(1,19) = 1.10, \emph{p} = 0.306, \(\eta_G^2\) = 0.008, \emph{Fig TBC}).
+However, there was no significant interaction between Condition and Transition Type (\emph{F}(1,19) = 1.10, \emph{p} = 0.306, \(\eta_G^2\) = 0.008, \emph{Fig \ref{fig:vel-diff-2}}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Experiment 1, 2: Normalised KV Difference at transition points when playing the piece with dynamics (left: Experiment 1, right: Experiment 2).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff-2}
+\end{figure}
 

### 2022-07-28 paper/manuscript.Rmd
update supplementary material and reflected minor issues
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# output: papaja::apa6_word
-output:
-  papaja::apa6_pdf:
-    latex_engine: xelatex
+output: papaja::apa6_word
+# output:
+#   papaja::apa6_pdf:
+#     latex_engine: xelatex
 ---
@@
 For offsets, `r round(sum(error_offset_summary$N)/((31*4*8)-5)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*8)-5)*100, 2)`%, missing notes: `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*8)-5)*100, 2)`%, substituted notes: `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*8)-5)*100, 2)`%).
+We found that some participants did not follow sheet music (e.g., hold some notes longer than notated), therefore the order of offsets did not correspond to that of onsets.
+We counted these as errors and removed the erroneous notes even if the order of onsets was correct.
 As a result, less than 1 % of total responses were corrected.
@@
 We employed the same procedure as in Experiment 1 with several modifications.
-First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2} A*) prior to the experiment.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in *[Supplementary material](#supplementary)*).
-Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute).
+First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2} A*) prior to the experiment so that they had enough time to familiarise themselves with the piece and perform it without pitch errors while performing it with different notated expressions in the lab.
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there was less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in *[Supplementary material](#supplementary)*).
+Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability [@zamm_2018].
 The chosen tempo was again cued by a leading metronome.
@@
 # Supplementary material{#supplementary}
+
 # Instructions
@@
 
-# Tempo Variability
+# Temporal Variability
 
@@
 ioi_cv_1 <- ggpaired(ioi_cv_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "CVs", title = "CVs: Exp 1") +
-  ylim(0.02, 0.20) +
+  ylim(0, 0.20) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 0.08, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 ioi_cv_2 <- ggpaired(ioi_cv_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "CVs", title = "CVs: Exp 2") +
-  ylim(0.02, 0.20) +
+  ylim(0, 0.20) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 0.08, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   geom_hline(yintercept = 188, linetype = "dashed") +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 150, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   geom_hline(yintercept = 1, linetype = "dashed") +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 1.05, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 
+In order to examine if the dynamics difference between forte and piano at transition points is larger when the difference between forte and piano in overall performance is considered, we normalised the KV difference at transition points by dividing it by the average range between forte and piano per performance (per trial).
+
 ## Results
@@
 ```{r vel-diff-dyn-normalised-1, echo = FALSE}
-vel_diff_dyn_normalised_1 <- ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas",xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff: Dynamics")
+vel_diff_dyn_normalised_1 <- ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas",xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff: Exp 1") + ylim(-6, 5) + stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 3, method = "t.test", paired = TRUE) + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
 
@@
 ```{r vel-diff-dyn-normalised-2, echo = FALSE}
-vel_diff_dyn_normalised_2 <- ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff: Dynamics")
+vel_diff_dyn_normalised_2 <- ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff: Exp 2") + ylim(-6, 5) + stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 3, method = "t.test", paired = TRUE) + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
 

### 2022-07-28 paper/manuscript.tex
update supplementary material and reflected minor issues
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 % \newenvironment{ltable}
-%   {\begin{landscape}\begin{center}\begin{threeparttable}}
-%   {\end{threeparttable}\end{center}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
+%   {\begin{landscape}\centering\begin{threeparttable}}
+%   {\end{threeparttable}\end{landscape}}
+\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
 
@@
 
-\abstract{
+\abstract{%
 Experts modulate their performance of actions for teaching purposes, performing slower and exaggerated movements when demonstrating novel actions to novices.
@@
 
+In order to examine if the dynamics difference between forte and piano at transition points is larger when the difference between forte and piano in overall performance is considered, we normalised the KV difference at transition points by dividing it by the average range between forte and piano per performance (per trial).
+
 \hypertarget{results-4}{%

### 2022-08-19 paper/manuscript.Rmd
update intro and citations
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   Experts modulate their performance of actions for teaching purposes, performing slower and exaggerated movements when demonstrating novel actions to novices.
-The present study asked whether such modulations are also used to teach techniques of artistic expression.
-One domain where subtle movement modulations are crucial for achieving artistic expression is music.
-We investigated how skilled pianists modulate their playing to demonstrate to students the techniques required for conveying piece-related musical expressions, compared to performing the piece without didactic intentions.
+The present study asked whether such modulations also occur during teaching performance of a music instrument, where subtle movement modulations are crucial for achieving artistic expression.
+While exaggerating performances of goal-directed actions outside of artistic contexts may be straightforward, it is an open question whether and how exaggeration for the purpose of teaching operates for actions that are expressive even when performed outside of a teaching context.
+Pianists were asked to demonstrate to students the techniques required for implementing notated expressions, compared to performing the piece without didactic intentions.
 Expressions in the piece concerned either articulation (i.e., legato and staccato) or dynamics (i.e., forte and piano).
@@
 When intending to teach articulation, skilled pianists produced more exaggerated staccato.
-When intending to teach dynamics, they created a larger contrast between forte and piano at structurally important points.
-We found consistent results with a simple musical scale (Experiment 1) and a more naturalistic piece of music (Experiment 2).
-These findings suggest that action modulations are performed not only to attract learners' attention but to highlight the most relevant aspects of the actions to be learnt.
+When intending to teach dynamics, they created a larger contrast between forte and piano.
+We found consistent results across a simple musical scale (Experiment 1) and a more naturalistic piece of music (Experiment 2).
+These findings show that teaching-specific action modulations generalize to expressive actions and suggest that action modulations serve to highlight the most relevant aspects of the actions to be learnt.
 
-keywords          : "teaching, expertise, skill transmission, artistic expression, music"
+keywords          : "teaching, skill transmission, sensorimotor communication, artistic expression, music"
 #wordcount         : ""
@@
 
-Importantly, in some domains teaching requires not only demonstrating *what* to do, but *how* to perform actions.
-In artistic contexts, novices need to learn how exactly to perform actions by relying on specific techniques [@sloboda_2000].
-For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques [@juslin_2003].
-This raises the question of whether and how experts modulate their actions during didactic demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
-Studying action modulations that serve the teaching of artistic techniques is also of broader theoretical importance: It can show whether action modulations can be used to selectively highlight subtle aspects of actions that define expert performance.
+Performing actions with exaggeration is straightforward for actions that are normally performed in the most efficient way possible [@pezzulo_2013].
+However, how can particular aspects of actions be highlighted when the actions themselves are expressive even outside of a teaching context?
+This is the case in music performance, where pieces are played with expression.
+Expressivity is a vital component of performance and typically the main focus of music teaching [@laukka_2004].
+Expressive skills are generally considered to be separate from technical skills, however, fine motor control is required to implement subtle sound modulations in expressive performance [@sloboda_2000].
+In some music genres such as Western classical music, it is crucial to acquire the motor skills needed to perform a piece expressively.
+For example, pianists translate their own interpretations of music and convey their emotions by modulating specific parameters such as timing, smoothness and loudness of sound.
+This raises the question of whether and how musicians modulate their actions during didactic demonstration of expressive techniques.
 
-In the present study, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, smoothness, loudness) can be modulated at the same time.
-Specific and subtle aspects of the performance may need to be highlighted to teach a respective technique.
+In naturalistic teaching settings, teachers have many possibilities for how to convey to a learner how to play a piece expressively.
+They may use verbal communication [@woody_2000, woody_2006] and non-verbal bodily cues [@simones_2015] as well as modulating the sounds that they are producing.
+In the present study, we focused entirely on the teaching of expressive techniques through sound in order to determine whether expert musicians would systematically exaggerate the actions necessary to implement a particular technique for musical expression.
 Pianists were asked to perform a piece to demonstrate to a learner how to implement the notated expressions (teaching condition) and to play the same piece for an audience (performing condition).
@@
 
-If experts rely on generic action modulations, it can be expected that they will play more slowly during didactic demonstration, regardless of the kind of expressive techniques to be taught.
-To the extent that experts use action modulations to support their teaching of specific expressive techniques, they should exaggerate legato (i.e., key-overlap time should be more positive) and staccato (i.e., key-overlap time should be more negative) when teaching articulation whereas they should exaggerate forte (i.e., key velocity should be higher) and piano (i.e., key velocity should be lower) when teaching dynamics.
+If experts rely on generic action modulations, it can be expected that they will play more slowly during didactic demonstration, regardless of the kind of techniques to be taught.
+To the extent that experts use action modulations to support their teaching of specific techniques, they should exaggerate legato (i.e., key-overlap time should be more positive) and staccato (i.e., key-overlap time should be more negative) when teaching articulation whereas they should exaggerate forte (i.e., key velocity should be higher) and piano (i.e., key velocity should be lower) when teaching dynamics.
 Furthermore, one could speculate that they might produce modulations specifically at structurally important points that best highlight the technique to be taught.

### 2022-08-20 paper/manuscript.Rmd
update references and after intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Furthermore, one could speculate that they might produce modulations specifically at structurally important points that best highlight the technique to be taught.
-Importantly, they should avoid modulating irrelevant properties of expression, e.g., the smoothness of sound while teaching dynamics.
+Importantly, they should avoid modulating irrelevant properties of expressio (e.g., the smoothness of sound while teaching dynamics).
 
@@
 In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplementary material](#supplementary)*).
-Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
+Participants played the piece 8 times per technique per condition, so there were 32 trials in total (2 conditions x 2 techniques x 8 trials).
 The order of the conditions was blocked and counterbalanced across participants.
-The order of the expressive techniques within each condition was also blocked and counterbalanced across participants.
+The order of the techniques within each condition was also blocked and counterbalanced across participants.
 A leading metronome (80 quarter beats per minute, 8 beats) indicated the target tempo before each trial.
@@
 For offsets, `r round(sum(error_offset_summary$N)/((31*4*8)-5)*100, 2)`% of the trials contained at least one pitch error (extra notes: `r round(error_offset_summary[Reason == "Extra Notes"]$N/((31*4*8)-5)*100, 2)`%, missing notes: `r round(error_offset_summary[Reason == "Missing Notes"]$N/((31*4*8)-5)*100, 2)`%, substituted notes: `r round(error_offset_summary[Reason == "Substituted"]$N/((31*4*8)-5)*100, 2)`%).
-We found that some participants did not follow sheet music (e.g., hold some notes longer than notated), therefore the order of offsets did not correspond to that of onsets.
+We found that some participants did not precisely follow the sheet music (e.g., they held some notes longer than notated), therefore the order of offsets did not correspond to that of onsets.
 We counted these as errors and removed the erroneous notes even if the order of onsets was correct.
@@
 
-We performed separate analyses for the two expressive techniques (i.e., articulation and dynamics).
+We performed separate analyses for the two techniques (i.e., articulation and dynamics).
 A paired-sample *t*-test or a Wilcoxon Signed-rank test (a non-parametric alternative to a paired *t*-test) was performed to compare the mean IOIs in the teaching and performing condition.
@@
 To recall our predictions, if participants play more slowly when they are trying to teach, inter-onset intervals (IOIs) should be larger when teaching.
-If participants specifically modulate relevant aspects of the expressive techniques they are trying to teach, key-overlap time (KOT) should be more positive for legato and more negative for staccato when teaching articulation, and key velocity (KV) should be higher for forte and lower for piano when teaching dynamics.
+If participants specifically modulate relevant aspects of the techniques they are trying to teach, key-overlap time (KOT) should be more positive for legato and more negative for staccato when teaching articulation, and key velocity (KV) should be higher for forte and lower for piano when teaching dynamics.
 
@@
 Importantly, participants did not modulate their performance in terms of irrelevant aspects of the techniques for teaching purposes (e.g., modulating the smoothness of sound while teaching dynamics).
-These findings confirmed that participants modulated their performance in systematic and fine-grained ways to teach expressive techniques.
+These findings confirmed that participants modulated their performance in systematic and fine-grained ways to teach particular techniques.
 
@@
 The same apparatus as in Experiment 1 was used.
-We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contained our targeted expressions (i.e., articulation and dynamics) and was relatively simple in terms of motor skills.
+We selected Clementi's Sonatina Op.36 (No.3) in C major as a stimulus because it contains our targeted expressions (i.e., articulation and dynamics) and is relatively simple in terms of motor skills.
 The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
@@
 First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2} A*) prior to the experiment so that they had enough time to familiarise themselves with the piece and perform it without pitch errors while performing it with different notated expressions in the lab.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there was less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in *[Supplementary material](#supplementary)*).
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in *[Supplementary material](#supplementary)*).
 Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability [@zamm_2018].
@@
 Given the different tempi, key-overlap ratios (KORs) were calculated by dividing KOT by the mean IOI of each performance to normalise KOT.
-Additionally, we include KV difference (i.e., KV difference for each interval) at transition points (i.e., forte to piano or piano to forte) as a dependent variable based on the findings of Experiment 1.
-Three trials were entirely excluded from data analysis because participants did not follow the sheet music.
+Additionally, we included KV difference (i.e., KV difference for each interval) at transition points (i.e., forte to piano or piano to forte) as a dependent variable based on the findings of Experiment 1.
+Three trials were entirely excluded from data analysis because participants did not follow the sheet music closely enough.
 Using the same approach as in Experiment 1, pitch errors were removed manually.
@@
 
-The results from Experiment 2 replicate our earlier findings and provide further evidence that skilled pianists were able to modulate their performance for teaching purposes.
+The results from Experiment 2 replicate our earlier findings and provide further evidence that skilled pianists modulated their performance in specific ways for teaching purposes.
 As in Experiment 1, we observed slower performance only while teaching the notated articulation.
@@
 
-The aim of the study was to investigate how expert pianists modulate their performance to teach expressive techniques.
+The aim of this study was to investigate how expert pianists modulate their performance to teach expressive techniques.
 Overall, the findings of the two experiments showed that pianists performed one and the same piece differently depending on whether they played with the intention to teach expressive techniques or with the intention to perform the piece for an audience.
@@
 Across two different pieces, participants exaggerated staccato when playing with the intention to teach expressions concerning articulation.
-The lack of a difference for legato when teaching articulation might stem from a ceiling effect.
+The lack of a difference for legato when teaching articulation might stem from a ceiling effect because there might not be enough room for exaggeration without significantly dropping the tempo.
 When playing with the intention to teach expressions concerning dynamics, participants made larger dynamics changes between forte and piano (in both directions).
@@
 Another possibility is that expression is best taught when leaving the tempo unchanged when it is irrelevant.
-This would imply that general didactic modulations like slowing down might be less useful in the context of teaching expressive techniques where timing itself can be used to add an expression to a performance.
-
-Musicians and music teachers tend to consider expressive skills as a performer’s most important skill [@lindstrom_2003].
-In the present study, experts focused on the teaching of expressions that were notated.
-However, expression also has other facets [@juslin_2003] that are not only piece-related but performer- or context-related.
-An interesting topic for future research is how experts teach not only expressive techniques as such but convey possibilities of interpretation.
+This would imply that general didactic modulations like slowing down might be less useful in the context of teaching techniques where timing itself can be used to add an expression to a performance.
 
 One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students.
-Given that teachers modulate their demonstration throughout online interactions with learners [@okazaki_2019], future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities.
+It can be expected that teachers would also make didactic modulations in some domains other than acoustic properties, such as gestures if they are physically present in front of students.
+Moreover, given that teachers modulate their demonstration throughout ongoing interactions with learners [@okazaki_2019], future studies are needed to investigate how experts dynamically adapt their performance to their students’ skill level and demonstrated abilities.
+Also, some of the participants in the current study did not have teaching experience at all and even those who had teaching experience had taught for only a few years.
+It would be important to investigate if our findings would be observed in participants with extensive experience teaching the piano.
+
+Another limitation of the study is that it is not certain how the results might generalise to performances of more complex pieces and pieces with fewer notations.
+In more complex pieces of music, multiple expressive notations could be assigned to one note or phrase.
+Also, it is rare that almost all the notes are assigned to one expressive notation as in our stimuli.
+Future studies are needed to determine whether pianists would exaggerate specific parts of a piece or prioritise one aspect over the others given multiple concurrent expressive notations and more possibilities for giving the music different interpretations.
+
+Musicians and music teachers tend to consider expressive skills as performers’ most important skills [@lindstrom_2003].
+In the present study, experts focused on the teaching of expressions that were notated.
+However, expression also has other facets [@juslin_2003] that are not only piece-related (e.g., notated expressions) but performer- or context-related (e.g., performer’s interpretation, listening contexts such as recordings or concerts).
+The ultimate goal of expressive performance is considered to lie in developing performers’ own interpretations of music and convey affects and emotions by using expressive tools such as articulation, dynamics, tempo and timing [@meissner_2021].
+The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings [e.g., @woody_2000].
+It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.
 
 The present findings extend earlier research on teaching-related action modulations.
-First, our study sheds light on the teaching of expressive techniques that are an integral part of skill acquisition in artistic domains.
+First, our study sheds light on the teaching of technical skills that are an integral part of skill acquisition in artistic domains.
 We showed that compared to an expressive performance baseline, experts made specific modulations to teach particular techniques.
-Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention.
+Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners’ attention.
 Rather, expert demonstrators seem to follow principles of relevance in communication [@sperber_1995], highlighting only those aspects they are intending to demonstrate.

### 2022-08-21 paper/manuscript.Rmd
update supplementary material
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-In order to investigated if participants used temporal grouping to make a boundary between each subcomponent (e.g., between legato and staccato, between forte and piano) clearer for teaching purposes, we looked at IOIs only at transition points.
+In order to investigate if participants used temporal grouping to make boundaries between subcomponents (i.e., between legato and staccato, between forte and piano) clearer for teaching purposes, we looked at IOIs only at transition points.
 
@@
 To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_tra_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_tra_norm$p.value))}`).
-The Wilcoxon Signed-rank test revealed that IOIs at transition points were larger for the teaching condition [*Mdn* = `r ioi_tra[Condition == "teaching" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_tra[Condition == "teaching" & Skill == "articulation"]$IQR`] than in the performing condition [*Mdn* = `r ioi_tra[Condition == "performing" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_tra[Condition == "performing" & Skill == "articulation"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_tra_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_tra_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_tra$effsize, 2)`, two-tailed, *Fig \ref{fig:tra}*).
+The Wilcoxon Signed-rank test revealed that IOIs at transition points were larger in the teaching condition [*Mdn* = `r ioi_tra[Condition == "teaching" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_tra[Condition == "teaching" & Skill == "articulation"]$IQR`] than in the performing condition [*Mdn* = `r ioi_tra[Condition == "performing" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_tra[Condition == "performing" & Skill == "articulation"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_tra_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_tra_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_tra$effsize, 2)`, two-tailed, *Fig \ref{fig:tra}*).
 
@@
 
-A paired-sample *t*-test showed that IOIs at transition points were larger for the teaching condition [*M* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$Mean` (ms), *SD* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$SD`] than the performing condition [*M* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$Mean` (ms), *SD* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$SD`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_tra_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_tra_ttest$statistic))`, *p* `r if(ioi_dyn_tra_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_tra_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn_tra$effsize), 3)`, two-tailed, *Fig \ref{fig:tra}*).
+A paired-sample *t*-test showed that IOIs at transition points were larger in the teaching condition [*M* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$Mean` (ms), *SD* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$SD`] than in the performing condition [*M* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$Mean` (ms), *SD* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$SD`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_tra_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_tra_ttest$statistic))`, *p* `r if(ioi_dyn_tra_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_tra_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn_tra$effsize), 3)`, two-tailed, *Fig \ref{fig:tra}*).
 
@@
 
-A paired-sample *t*-test showed that IOIs at transition points were larger for the teaching condition [*M* = `r sprintf('%.2f', ioi_tra[Condition == "teaching" & Skill == "articulation"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_tra[Condition == "teaching" & Skill == "articulation"]$SD)`] than the performing condition [*M* = `r sprintf('%.2f', ioi_tra[Condition == "performing" & Skill == "articulation"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_tra[Condition == "performing" & Skill == "articulation"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_tra_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_tra_ttest$statistic))`, *p* `r if(ioi_art_tra_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_tra_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art_tra$effsize), 3)`, two-tailed, *Fig \ref{fig:tra}*).
+A paired-sample *t*-test showed that IOIs at transition points were larger in the teaching condition [*M* = `r sprintf('%.2f', ioi_tra[Condition == "teaching" & Skill == "articulation"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_tra[Condition == "teaching" & Skill == "articulation"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_tra[Condition == "performing" & Skill == "articulation"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_tra[Condition == "performing" & Skill == "articulation"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_tra_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_tra_ttest$statistic))`, *p* `r if(ioi_art_tra_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_tra_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art_tra$effsize), 3)`, two-tailed, *Fig \ref{fig:tra}*).
 

### 2022-08-22 paper/manuscript.Rmd
update reference number
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 In naturalistic teaching settings, teachers have many possibilities for how to convey to a learner how to play a piece expressively.
-They may use verbal communication [@woody_2000, woody_2006] and non-verbal bodily cues [@simones_2015] as well as modulating the sounds that they are producing.
+They may use verbal communication [@woody_2000; @woody_2006] and non-verbal bodily cues [@simones_2015] as well as modulating the sounds that they are producing.
 In the present study, we focused entirely on the teaching of expressive techniques through sound in order to determine whether expert musicians would systematically exaggerate the actions necessary to implement a particular technique for musical expression.

### 2022-08-30 experiment-1/analysis/stats/articulation.Rmd
use setcolorder to sort data
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 - [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons
-
-# Outline - Articulation (KOT, KOR)
-## [Articulation](#kot-articulation)
-
-### KOT
-- Individual results
-  + [All](#kot_art_ind_all)
-  + [Trial by trial](#kot_art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kot_art_group_all)
-  + [Trial by trial](#kot_art_group_trial)
-
-### KOR
-- Individual results
-  + [All](#kor_art_ind_all)
-  + [Trial by trial](#kor_art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kor_art_group_all)
-  + [Trial by trial](#kor_art_group_trial)
-
-## [Dynamics](#kot-dynamics)
-
-### KOT
-- Individual results
-  + [All](#kot_dyn_ind_all)
-  + [Trial by trial](#kot_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kot_dyn_group_all)
-  + [Trial by trial](#kot_dyn_group_trial)
-
-### KOR
-- Individual results
-  + [All](#kor_dyn_ind_all)
-  + [Trial by trial](#kor_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kor_dyn_group_all)
-  + [Trial by trial](#kor_dyn_group_trial)
-
-## [Sequence plot](#sequence)
 
@@
 
-# Key-Overlap Time (KOT) - Articulation{#kot-articulation}
-## 1. All - individual{#kot_art_ind_all}
+# Key-Overlap Time (KOT) - Articulation
+## 1. All - individual
 - Average KOT for each participant.
@@
 kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(kot_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
@@
 
-## 2. All - group{#kot_art_group_all}
+## 2. All - group
 - Average KOT for each condition.
@@
 
-## 3. Trial by trial - individual{#kot_art_ind_trial}
+## 3. Trial by trial - individual
 - Average KOT for each trial for each participant.
@@
 
-## 4. Trial by trial - group{#kot_art_group_trial}
+## 4. Trial by trial - group
 - Average KOT for each trial for each condition.
@@
 # Key-Overlap Ratio (KOR) - Articulation
-## 1. All - individual{#kor_art_ind_all}
+## 1. All - individual
 - Average KOR for each participant.
@@
 kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(kor_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
@@
 #### Box plot
-
 ```{r kor-art-box,  echo = FALSE}
@@
 
-## 2. All - group{#kor_art_group_all}
+## 2. All - group
 - Average KOR for each condition.
@@
 
-## 3. Trial by trial - individual{#kor_art_ind_trial}
+## 3. Trial by trial - individual
 - Average KOR for each trial for each participant.
@@
 
-## 4. Trial by trial - group{#kor_art_group_trial}
+## 4. Trial by trial - group
 - Average KOR d for each trial for each condition.
@@
 
-# Key-Overlap Time (KOT) - Dynamics{#kot-dynamics}
-## 1. All - individual{#kot_dyn_ind_all}
+# Key-Overlap Time (KOT) - Dynamics
+## 1. All - individual
 - Average KOT for each participant.
@@
 kot_dyn_trial <- dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(kot_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
@@
 
-## 2. All - group{#kot_dyn_group_all}
+## 2. All - group
 - Average KOT for each condition.
@@
 
-## 3. Trial by trial - individual{#kot_dyn_ind_trial}
+## 3. Trial by trial - individual
 - Average KOT for each trial for each participant.
@@
 
-## 4. Trial by trial - group{#kot_dyn_group_trial}
+## 4. Trial by trial - group
 - Average KOT for each trial for each condition.
@@
 # Key-Overlap Ratio (KOR) - Dynamics
-## 1. All - individual{#kor_dyn_ind_all}
+## 1. All - individual
 - Average KOR for each participant.
@@
 kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(kor_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
@@
 
-## 2. All - group{#kor_dyn_group_all}
+## 2. All - group
 - Average KOR for each condition.
@@
 
-## 3. Trial by trial - individual{#kor_dyn_ind_trial}
+## 3. Trial by trial - individual
 - Average KOR for each trial for each participant.
@@
 
-## 4. Trial by trial - group{#kor_dyn_group_trial}
+## 4. Trial by trial - group
 - Average KOR for each trial for each condition.
@@
 
-# Sequence plot{#sequence}
-## Articulation{#sequence_art}
+# Sequence plot
+## Articulation
 ```{r seq-art, fig.width = 9, fig.height = 3, echo = FALSE}
@@
 
-## Dynamics{#sequence_dyn}
+## Dynamics
 ```{r seq-dyn, fig.width = 9, fig.height = 3, echo = FALSE}

### 2022-08-30 experiment-1/analysis/stats/dynamics.Rmd
use setcolorder to sort data
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 - [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons
-
-# Outline - Dynamics (Key Velocity)
-
-## [Dynamics](#vel-dynamics)
-
-### KV
-- Individual results
-  + [All](#vel_dyn_ind_all)
-  + [Trial by trial](#vel_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#vel_dyn_group_all)
-  + [Trial by trial](#vel_dyn_group_trial)
-
-### KV-Diff
-- Individual results
-  + [All](#vel_diff_dyn_ind_all)
-  + [Trial by trial](#vel_diff_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#vel_diff_dyn_group_all)
-  + [Trial by trial](#vel_diff_dyn_group_trial)
-
-## [Articulation](#vel_articulation)
-
-### KV
-- Individual results
-  + [All](#vel_art_ind_all)
-  + [Trial by trial](#vel_art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#vel_art_group_all)
-  + [Trial by trial](#vel_art_group_trial)
-
-### KV-Diff
-- Individual results
-  + [All](#vel_diff_art_ind_all)
-  + [Trial by trial](#vel_diff_art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#vel_diff_art_group_all)
-  + [Trial by trial](#vel_diff_art_group_trial)
-
-## [Sequence plot](#sequence)
-
-## [Additional analysis](#additional)
 
@@
 # Key Velocity (KV) - Dynamics
-## 1. All - individual{#vel_dyn_ind_all}
+## 1. All - individual
 - Average KV for each participant.
@@
 vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(vel_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
@@
 
-## 2. All - group{#vel_dyn_group_all}
+## 2. All - group
 - Average KV for each condition.
@@
 
-## 3. Trial by trial - individual{#vel_dyn_ind_trial}
+## 3. Trial by trial - individual
 - Average KV for each trial for each participant.
@@
 
-## 4. Trial by trial - group{#vel_dyn_group_trial}
+## 4. Trial by trial - group
 - Average KV for each trial for each condition.
@@
 # Key Velocity Difference (Difference) - Dynamics
-## 1. All - individual{#vel_diff_dyn_ind_all}
+## 1. All - individual
 - Average KV-Diff for each participant.
@@
 vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(vel_diff_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
@@
 
-## 2. All - group{#vel_diff_dyn_group_all}
+## 2. All - group
 - Average KV-Diff for each condition.
@@
 
-## 3. Trial by trial - individual{#vel_diff_dyn_ind_trial}
+## 3. Trial by trial - individual
 - Average KV-Diff for each trial for each participant.
@@
 
-## 4. Trial by trial - group{#vel_diff_dyn_group_trial}
+## 4. Trial by trial - group
 - Average KV-Diff for each trial for each condition.
@@
 # Key Velocity (KV) - Articulation
-## 1. All - individual{#vel_art_ind_all}
+## 1. All - individual
 - Average KV for each participant.
@@
 vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(vel_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
@@
 
-## 2. All - group{#vel_art_group_all}
+## 2. All - group
 - Average KV for each condition.
@@
 
-## 3. Trial by trial - individual{#vel_art_ind_trial}
+## 3. Trial by trial - individual
 - Average KV for each trial for each participant.
@@
 
-## 4. Trial by trial - group{#vel_art_group_trial}
+## 4. Trial by trial - group
 - Average KV for each trial for each condition.
@@
 
-# Key Velocity Difference (Difference) - Articulation{#vel_articulation}
-## 1. All - individual{#vel_diff_art_ind_all}
+# Key Velocity Difference (Difference) - Articulation
+## 1. All - individual
 - Average KV-Diff for each participant.
@@
 vel_diff_art_trial <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(vel_diff_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
@@
 
-## 2. All - group{#vel_diff_art_group_all}
+## 2. All - group
 - Average KV-Diff for each condition.
@@
 
-## 3. Trial by trial - individual{#vel_diff_art_ind_trial}
+## 3. Trial by trial - individual
 - Average KV-Diff for each trial for each participant.
@@
 
-## 4. Trial by trial - group{#vel_diff_art_group_trial}
+## 4. Trial by trial - group
 - Average KV-Diff for each trial for each condition.
@@
 
-# Sequence plot{#sequence}
-## Dynamics{#sequence_dyn}
+# Sequence plot
+## Dynamics
 ```{r seq-dyn, fig.width = 9, fig.height = 3, echo = FALSE}
@@
 
-## Articulation{#sequence_art}
+## Articulation
 ```{r seq-art, fig.width = 9, fig.height = 3, echo = FALSE}
@@
 
-# Additional analysis{#additional}
+# Additional analysis
 - Normalise KV difference by the range of KV value between forte and piano for each participant.
@@
 ## Normalised KV difference per trial
-
 ```{r, echo = FALSE}

### 2022-08-30 experiment-1/analysis/stats/ioi.Rmd
use setcolorder to sort data
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# Outline - Tempo (IOIs)
-## [Articulation](#ioi-articulation)
-
-- Individual results
-  + [All](#art_ind_all)
-  + [Phrase boundaries](#art_ind_phrase)
-  + [Trial by trial](#art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#art_group_all)
-  + [Phrase boundaries](#art_group_phrase)
-  + [Trial by trial](#art_group_trial)
-
-## [Dynamics](#ioi-dynamics)
-
-- Individual results
-  + [All](#dyn_ind_all)
-  + [Phrase boundaries](#dyn_ind_phrase)
-  + [Trial by trial](#dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#dyn_group_all)
-  + [Phrase boundaries](#dyn_group_phrase)
-  + [Trial by trial](#dyn_group_trial)
-
-## [Sequence plot](#sequence)
-
 ```{r setup, include = FALSE}
@@
 # IOI-Articulation 
-## 1. All - individual{#art_ind_all}
+## 1. All - individual
 - Average IOIs for each participant.
@@
 ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+setorder(ioi_art_trial, "SubNr", "Condition", "Skill")
 
@@
 
-## 2. All - group{#art_group_all}
+## 2. All - group
 - Average IOIs for each condition.
@@
 
-## 3. Phrase boundaries - individual{#art_ind_phrase}
+## 3. Phrase boundaries - individual
 - Compare IOIs at phrase boundaries with others for each participant.
@@
 ioi_art_ch_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)]
+setorder(ioi_art_ch_trial, "SubNr", "Condition", "Skill", "Boundary")
 
@@
 
-## 4. Phrase boundaries - group{#art_group_phrase}
+## 4. Phrase boundaries - group
 - Compare IOIs at phrase boundaries with others in the two conditions.
@@
 
-## 5. Trial by trial - individual{#art_ind_trial}
+## 5. Trial by trial - individual
 - Average IOIs for each trial for each participant.
@@
 
-## 6. Trial by trial - group{#art_group_trial}
+## 6. Trial by trial - group
 - Average IOIs for each trial for each condition.
@@
 # IOI-Dynamics
-## 1. All - individual{#dyn_ind_all}
+## 1. All - individual
 - Average IOIs for each participant.
@@
 ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+setorder(ioi_dyn_trial, "SubNr", "Condition", "Skill")
 
@@
 
-## 2. All - group{#dyn_group_all}
+## 2. All - group
 - Average IOIs for each condition.
@@
 
-## 3. Phrase boundaries - individual{#dyn_ind_phrase}
+## 3. Phrase boundaries - individual
 - Compare IOIs at phrase boundaries with others for each participant.
@@
 ioi_dyn_ch_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)]
+setorder(ioi_dyn_ch_trial, "SubNr", "Condition", "Skill", "Boundary")
 
@@
 
-## 4. Phrase boundaries - group{#dyn_group_phrase}
+## 4. Phrase boundaries - group
 - Compare IOIs at phrase boundaries with others in the two conditions.
@@
 
-## 5. Trial by trial - individual{#dyn_ind_trial}
+## 5. Trial by trial - individual
 - Average IOIs for each trial for each participant.
@@
 # For each individual
-ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
 ioi_dyn_trial
@@
 
-## 6. Trial by trial - group{#dyn_group_trial}
+## 6. Trial by trial - group
 - Average IOIs for each trial for each condition.

### 2022-08-30 experiment-2/analysis/stats/articulation.Rmd
use setorder to sort data
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 - [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons
-
-# Outline - Articulation (KOR, KOT)
-## [Articulation](#kor-articulation)
-
-### KOR
-- Individual results
-  + [All](#kor_art_ind_all)
-  + [Trial by trial](#kor_art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kor_art_group_all)
-  + [Trial by trial](#kor_art_group_trial)
-
-### KOT
-- Individual results
-  + [All](#kot_art_ind_all)
-  + [Trial by trial](#kot_art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kot_art_group_all)
-  + [Trial by trial](#kot_art_group_trial)
-
-## [Dynamics](#kor-dynamics)
-
-### KOR
-- Individual results
-  + [All](#kor_dyn_ind_all)
-  + [Trial by trial](#kor_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kor_dyn_group_all)
-  + [Trial by trial](#kor_dyn_group_trial)
-
-### KOT
-- Individual results
-  + [All](#kot_dyn_ind_all)
-  + [Trial by trial](#kot_dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#kot_dyn_group_all)
-  + [Trial by trial](#kot_dyn_group_trial)
-
-## [Sequence plot](#sequence)
 
@@
 
-# Key-Overlap Ratio (KOR) - Articulation{#kor-articulation}
-## 1. All - individual{#kor_art_ind_all}
+# Key-Overlap Ratio (KOR) - Articulation
+## 1. All - individual
 - Average KOR for each participant.
@@
 # Key-Overlap Time (KOT) - Articulation
-## 1. All - individual{#kot_art_ind_all}
+## 1. All - individual
 - Average KOT for each participant.
@@
 #### Box plot
-
 ```{r kot-art-box,  echo = FALSE}
@@
 
-# Key-Overlap Ratio (KOR) - Dynamics{#kor-dynamics}
-## 1. All - individual{#kor_dyn_ind_all}
+# Key-Overlap Ratio (KOR) - Dynamics
+## 1. All - individual
 - Average KOR for each participant.
@@
 # Key-Overlap Time (KOT) - Dynamics
-## 1. All - individual{#kot_dyn_ind_all}
+## 1. All - individual
 - Average KOT for each participant.

### 2022-08-30 experiment-2/analysis/stats/ioi.Rmd
use setorder to sort data
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 - [rstatix](https://github.com/kassambara/rstatix) for effect sizes
-
-# Outline - Tempo (IOIs)
-## [Articulation](#ioi-articulation)
-
-- Individual results
-  + [All](#art_ind_all)
-  + [Phrase boundaries](#art_ind_phrase)
-  + [Trial by trial](#art_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#art_group_all)
-  + [Phrase boundaries](#art_group_phrase)
-  + [Trial by trial](#art_group_trial)
-
-## [Dynamics](#ioi-dynamics)
-
-- Individual results
-  + [All](#dyn_ind_all)
-  + [Phrase boundaries](#dyn_ind_phrase)
-  + [Trial by trial](#dyn_ind_trial)
-
-- Group results and stats (if available)
-  + [All](#dyn_group_all)
-  + [Phrase boundaries](#dyn_group_phrase)
-  + [Trial by trial](#dyn_group_trial)
-
-## [Sequence plot](#sequence)
 
@@
 # IOI-Articulation 
-## 1. All - individual{#art_ind_all}
+## 1. All - individual
 - Average normIOIs for each participant.
@@
 ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+setorder(ioi_art_trial, "SubNr", "Condition", "Skill")
 
@@
 
-## 2. All - group{#art_group_all}
+## 2. All - group
 - Average normIOIs for each condition.
@@
 
-## 3. Phrase boundaries - individual{#art_ind_phrase}
+## 3. Phrase boundaries - individual
 - Compare normIOIs at phrase boundaries with others for each participant.
@@
 ioi_art_ch_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)]
+setorder(ioi_art_ch_trial, "SubNr", "Condition", "Skill", "Boundary")
 
@@
 
-## 4. Phrase boundaries - group{#art_group_phrase}
+## 4. Phrase boundaries - group
 - Compare normIOIs at phrase boundaries with others in the two conditions.
@@
 
-## 5. Trial by trial - individual{#art_ind_trial}
+## 5. Trial by trial - individual
 - Average normIOIs for each trial for each participant.
@@
 
-## 6. Trial by trial - group{#art_group_trial}
+## 6. Trial by trial - group
 - Average normIOIs for each trial for each condition.
@@
 # IOI-Dynamics
-## 1. All - individual{#dyn_ind_all}
+## 1. All - individual
 - Average normIOIs for each participant.
@@
 ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+setorder(ioi_dyn_trial, "SubNr", "Condition", "Skill")
 
@@
 
-## 2. All - group{#dyn_group_all}
+## 2. All - group
 - Average normIOIs for each condition.
@@
 
-## 3. Phrase boundaries - individual{#dyn_ind_phrase}
+## 3. Phrase boundaries - individual
 - Compare normIOIs at phrase boundaries with others for each participant.
@@
 ioi_dyn_ch_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Boundary)]
+setorder(ioi_dyn_ch_trial, "SubNr", "Condition", "Skill", "Boundary")
 
@@
 
-## 4. Phrase boundaries - group{#dyn_group_phrase}
+## 4. Phrase boundaries - group
 - Compare normIOIs at phrase boundaries with others in the two conditions.
@@
 
-## 5. Trial by trial - individual{#dyn_ind_trial}
+## 5. Trial by trial - individual
 - Average normIOIs for each trial for each participant.
@@
 
-## 6. Trial by trial - group{#dyn_group_trial}
+## 6. Trial by trial - group
 - Average normIOIs for each trial for each condition.

### 2022-08-30 experiment-1/analysis/stats/articulation.Rmd
change fig width/height / remove unnecessary analysis from variability
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ## Articulation
-```{r seq-art, fig.width = 9, fig.height = 3, echo = FALSE}
+```{r seq-art, fig.width = 6, fig.height = 1.2, echo = FALSE}
 # For each individual
@@
 ## Dynamics
-```{r seq-dyn, fig.width = 9, fig.height = 3, echo = FALSE}
+```{r seq-dyn, fig.width = 6, fig.height = 1.2, echo = FALSE}
 # For each individual

### 2022-08-30 experiment-1/analysis/stats/dynamics.Rmd
change fig width/height / remove unnecessary analysis from variability
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
+ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff (normalised): Dynamics")
 ```
@@
 
-## Normalised KV difference per participant
-```{r, echo = FALSE}
-vel_diff_dyn$NormalisedDiff <- 0
-for (i in unique(vel_diff_dyn_trial$SubNr)){
-  # teaching
-  # FtoP
-  vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
-  # PtoF
-  vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
-  # performing
-  # FtoP
-  vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
-  # PtoF
-  vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
-}
-
-# group
-vel_diff_dyn_normalised_all_2 <- vel_diff_dyn[, .(N = .N, Mean = mean(NormalisedDiff), SD = sd(NormalisedDiff), SEM = sd(NormalisedDiff)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
-vel_diff_dyn_normalised_all_2
-```
-
-### Plot
-```{r, echo = FALSE}
-vel_diff_dyn_plot_3 <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$NormalisedDiff, performing = vel_diff_dyn[Condition == "performing"]$NormalisedDiff, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)
-
-ggpaired(vel_diff_dyn_plot_3, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
-```
-
-### Stats
-```{r, echo = FALSE}
-# stats
-vel_diff_dyn_aov_4 <- aov_car(NormalisedDiff ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
-summary(vel_diff_dyn_aov_4)
-```
-
-#### Post-hoc (emmeans)
-```{r, echo = FALSE}
-vel_diff_dyn_posthoc_3 <- test(pairs(emmeans(vel_diff_dyn_aov_4,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
-vel_diff_dyn_posthoc_3
-```
-
 ```{r export, include = FALSE}

### 2022-08-30 experiment-1/analysis/stats/variability.Rmd
change the title of the analysis page
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
change fig width/height / remove unnecessary analysis from variability
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: "Variability analysis 1"
+title: "Variability analysis: The Sound of Teaching Music 1"
 output:
@@
 ioi_cv_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+setorder(ioi_cv_trial, "SubNr", "Condition", "Skill")
+
 ioi_cv_subject <- ioi_cv_trial[, .(N = .N, MeanCV = mean(CV), SD = sd(CV)), by = .(SubNr, Condition, Skill)]
@@
   theme_pubr()
+
+ioi_cv_trial
 ```

### 2022-08-30 experiment-1/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -46,274 +46,278 @@ dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 
 
-## ----kot-art, echo = FALSE----------------------------------------------------
+## ----kot-art, echo = FALSE----------------------------------
 # For each individual
 kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(kot_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
 kot_art <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kot_art
 
 
-## ----kot-art-box,  echo = FALSE-----------------------------------------------
+## ----kot-art-box,  echo = FALSE-----------------------------
[…]

### 2022-08-30 experiment-1/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------
+## ----setup, include = FALSE----------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------
+## ----file, include = FALSE-----------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE-------------------------------
+## ----extract, include = FALSE--------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,46 +31,47 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE----------------------------------
+## ----vel-dyn, echo = FALSE-----------------------------------
 # For each individual
 vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(vel_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
 vel_dyn <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE-----------------------------
+## ----vel-dyn-box,  echo = FALSE------------------------------
[…]

### 2022-08-30 experiment-1/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,11 +9,11 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -25,224 +25,227 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE----------------------------------------------------
+## ----ioi-art, echo = FALSE----------------------------------
 # For each individual
 ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+setorder(ioi_art_trial, "SubNr", "Condition", "Skill")
 
 ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE-----------------------------------------------
+## ----ioi-art-box,  echo = FALSE-----------------------------
 ggboxplot(ioi_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "IOIs (ms)", title = "IOI: Articulation") + geom_hline(yintercept = 188, linetype = "dashed")
 
 
-## ----ioi-art-all, echo = FALSE------------------------------------------------
[…]

### 2022-08-30 experiment-1/analysis/teaching/teaching_art.Rmd
upload analysis between those who had teaching experience vs those who didn't
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Articulation Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of articulation analysis  when comparing those who had teaching experiences and those who don't.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: legato vs. staccato / forte vs. piano) x Teaching (2 levels: Yes vs. No)
- Dependent variables: Key-Overlap Time (KOT), Key-Overlap Ratio (KOR)

1. Main predictions
- Participants will exaggerate legato and staccato more while teaching.

2. Excluded participants

- SubNr 1: A deviated tempo
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

<!-- 3. Stats packages -->
<!-- - [afex](https://github.com/singmann/afex) for ANOVAs -->
<!-- - [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons -->

```{r setup, include = FALSE}
# packages
if (!require("here")) {install.packages("here"): require("here")}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# # statistics
# if (!require("afex")) {install.packages("afex"); require("afex")}
# if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}

# read info about teaching experience
source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
```

```{r file, include = FALSE}
filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
dt_kot$SubNr <- as.factor(dt_kot$SubNr)

# For each individual, each block, each trial
ioi <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]

# Calculate KOR for each interval (outliers were already determined based on KOT)
dt_kor <- data.table()
for (subnr in unique(dt_kot$SubNr)){
  for (block in unique(dt_kot[SubNr == subnr]$BlockNr)){
    for (trial in unique(dt_kot[SubNr == subnr & BlockNr == block]$TrialNr)){
      dt_current <- dt_kot[SubNr == subnr & BlockNr == block & TrialNr == trial]
      dt_current$KOR <- dt_current$KOT/ioi[SubNr == subnr & ioi$BlockNr == block & TrialNr == trial]$Mean
      dt_kor <- rbind(dt_kor, dt_current)
    }
  }
}

# Include only articulation
dt_kot_art <- dt_kot[Skill == "articulation"]
dt_kor_art <- dt_kor[Skill == "articulation"]

# Include only dynamics
dt_kot_dyn <- dt_kot[Skill == "dynamics"]
dt_kor_dyn <- dt_kor[Skill == "dynamics"]
```

# Key-Overlap Time (KOT) - Articulation
## 1. All - individual
- Average KOT for each participant.

```{r kot-art, echo = FALSE}
# For each individual
kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
setorder(kot_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")

kot_art <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
kot_art$Teaching <- rep(dt_included$TeachingPiano, each = 4)
kot_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
kot_art
```

## 2. All - group
- Average KOT for each condition.

```{r kot-art-all, echo = FALSE}
# Group mean
kot_art_all <- kot_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
kot_art_all
```

### Plot
#### Box plot
```{r kot-art-all-box, echo = FALSE}
kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(kot_art_plot[subcomponent == "Legato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation-Legato")

ggpaired(kot_art_plot[subcomponent == "Staccato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation-Staccato")
```

# Key-Overlap Ratio (KOR) - Articulation
## 1. All - individual
- Average KOR for each participant.

```{r kor-art, echo = FALSE}
# For each individual
kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
setorder(kor_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")

kor_art <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
kor_art$Teaching <- rep(dt_included$TeachingPiano, each = 4)
kor_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
kor_art
```

## 2. All - group
- Average KOR for each condition.

```{r kor-art-all, echo = FALSE}
# Group mean
kor_art_all <- kor_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
kor_art_all
```

### Plot
#### Box plot
```{r kor-art-all-box, echo = FALSE}
kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(kor_art_plot[subcomponent == "Legato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation-Legato")

ggpaired(kor_art_plot[subcomponent == "Staccato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation-Staccato")
```

# Key-Overlap Time (KOT) - Dynamics
## 1. All - individual
- Average KOT for each participant.

```{r kot-dyn, echo = FALSE}
# For each individual
kot_dyn_trial <- dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
setorder(kot_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")

kot_dyn <- kot_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
kot_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 4)
kot_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
kot_dyn
```

## 2. All - group
- Average KOT for each condition.

```{r kot-dyn-all, echo = FALSE}
# Group mean
kot_dyn_all <- kot_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
kot_dyn_all
```

### Plot
#### Box plot
```{r kot-dyn-all-box, echo = FALSE}
kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(kot_dyn_plot[subcomponent == "Forte"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics-Forte")

ggpaired(kot_dyn_plot[subcomponent == "Piano"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics-Piano")
```

# Key-Overlap Ratio (KOR) - Dynamics
## 1. All - individual
- Average KOR for each participant.

```{r kor-dyn, echo = FALSE}
# For each individual
kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
setorder(kor_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")

kor_dyn <- kor_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
kor_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 4)
kor_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
kor_dyn
```

## 2. All - group
- Average KOR for each condition.

```{r kor-dyn-all, echo = FALSE}
# Group mean
kor_dyn_all <- kor_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
kor_dyn_all
```

### Plot
#### Box plot
```{r kor-dyn-all-box, echo = FALSE}
kor_dyn_plot <- data.table(subject = kor_dyn[Condition == "teaching"]$SubNr, teaching = kor_dyn[Condition == "teaching"]$Mean, performing = kor_dyn[Condition == "performing"]$Mean, subcomponent = kor_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(kor_dyn_plot[subcomponent == "Forte"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics-Forte")

ggpaired(kor_dyn_plot[subcomponent == "Piano"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics-Piano")
```

```{r export, include = FALSE}
knitr::purl("teaching_art.Rmd")
```

### 2022-08-30 experiment-1/analysis/teaching/teaching_dyn.Rmd
upload analysis between those who had teaching experience vs those who didn't
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Dynamics Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of dynamics analysis when comparing those who had teaching experiences and those who don't.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato) x Teaching (2 levels: Yes vs. No)
- Dependent variables: Key Velocity (KV), KV difference

1. Main predictions
- Participants will exaggerate forte and piano more while teaching.

2. Excluded participants

- SubNr 1: A deviated tempo
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

<!-- 3. Stats packages -->
<!-- - [afex](https://github.com/singmann/afex) for ANOVAs -->
<!-- - [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons -->

```{r setup, include = FALSE}
# packages
if (!require("here")) {install.packages("here"): require("here")}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# # statistic
# if (!require("afex")) {install.packages("afex"); require("afex")}
# if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}

# read info about teaching experience
source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
```

```{r file, include = FALSE}
filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
```

```{r extract, include = FALSE}
dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_vel$SubNr <- as.factor(dt_vel$SubNr)
dt_vel_diff$SubNr <- as.factor(dt_vel_diff$SubNr)

# Include only articulation
dt_vel_art <- dt_vel[Skill == "articulation"]
dt_vel_diff_art <- dt_vel_diff[Skill == "articulation"]

# Include only dynamics
dt_vel_dyn <- dt_vel[Skill == "dynamics"]
dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
```

# Key Velocity (KV) - Dynamics
## 1. All - individual
- Average KV for each participant.

```{r vel-dyn, echo = FALSE}
# For each individual
vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
setorder(vel_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")

vel_dyn <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
vel_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 4)
vel_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
vel_dyn
```

## 2. All - group
- Average KV for each condition.
- Hypothesis: Participants will exaggerate forte and piano when teaching.

```{r vel-dyn-all, echo = FALSE}
# Group mean
vel_dyn_all <- vel_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
vel_dyn_all
```

### Plot
#### Box plot
```{r vel-dyn-all-box, echo = FALSE}
vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(vel_dyn_plot[subcomponent == "Forte"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics-Forte")

ggpaired(vel_dyn_plot[subcomponent == "Piano"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics-Piano")
```

# Key Velocity Difference (Difference) - Dynamics
## 1. All - individual
- Average KV-Diff for each participant.

```{r vel-diff-dyn, echo = FALSE}
# For each individual
vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]

vel_diff_dyn <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
setorder(vel_diff_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")

vel_diff_dyn <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
vel_diff_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 4)
vel_diff_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
vel_diff_dyn
```

## 2. All - group
- Average KV-Diff for each condition.

```{r vel-diff-dyn-all, echo = FALSE}
# Group mean
vel_diff_dyn_all <- vel_diff_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
vel_diff_dyn_all
```

### Plot
#### Box plot
```{r vel-diff-dyn-all-box, echo = FALSE}
vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(vel_diff_dyn_plot[subcomponent == "FtoP"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics-FtoP")

ggpaired(vel_diff_dyn_plot[subcomponent == "PtoF"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics-PtoF")
```

# Key Velocity (KV) - Articulation
## 1. All - individual
- Average KV for each participant.

```{r vel-art, echo = FALSE}
# For each individual
vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]

vel_art <- vel_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
setorder(vel_art, "SubNr", "Condition", "Skill", "Subcomponent")

# add teaching info
vel_art$Teaching <- rep(dt_included$TeachingPiano, each = 4)
vel_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)

vel_art
```

## 2. All - group
- Average KV for each condition.

```{r vel-art-all, echo = FALSE}
# Group mean
vel_art_all <- vel_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
vel_art_all
```

### Plot
#### Box plot
```{r vel-art-all-box, echo = FALSE}
vel_art_plot <- data.table(subject = vel_art[Condition == "teaching"]$SubNr, teaching = vel_art[Condition == "teaching"]$Mean, performing = vel_art[Condition == "performing"]$Mean, subcomponent = vel_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(vel_art_plot[subcomponent == "Legato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation-Legato")

ggpaired(vel_art_plot[subcomponent == "Staccato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation-Staccato")
```

# Key Velocity Difference (Difference) - Articulation
## 1. All - individual
- Average KV-Diff for each participant.

```{r vel-diff-art, echo = FALSE}
# For each individual
vel_diff_art_trial <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
setorder(vel_diff_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")

vel_diff_art <- vel_diff_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
vel_diff_art$Teaching <- rep(dt_included$TeachingPiano, each = 4)
vel_diff_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
vel_diff_art
```

## 2. All - group
- Average KV-Diff for each condition.

```{r vel-diff-art-all, echo = FALSE}
# Group mean
vel_diff_art_all <- vel_diff_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
vel_diff_art_all
```

### Plot
#### Box plot
```{r vel-diff-art-all-box, echo = FALSE}
vel_diff_art_plot <- data.table(subject = vel_diff_art[Condition == "teaching"]$SubNr, teaching = vel_diff_art[Condition == "teaching"]$Mean, performing = vel_diff_art[Condition == "performing"]$Mean, subcomponent = vel_diff_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(vel_diff_art_plot[subcomponent == "LtoS"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Articulation-LtoS")

ggpaired(vel_diff_art_plot[subcomponent == "StoL"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Articulation-StoL")
```

```{r export, include = FALSE}
knitr::purl("teaching_dyn.Rmd")
```

### 2022-08-30 experiment-1/analysis/teaching/teaching_ioi.Rmd
upload analysis between those who had teaching experience vs those who didn't
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Tempo Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of tempo analysis when comparing those who had teaching experiences and those who don't.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design
- Independent variables: Condition (2 levels: teaching vs. performing) x Technique (2 levels: articulation vs. dynamics) x Teaching (2 levels: Yes vs. No)
- Dependent variables: Interonset intervals (IOIs)

1. Main predictions
- Participants will play slower while teaching regardless of the techniques.

2. Excluded participants

- SubNr 1: A deviated tempo
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

<!-- 3. Stats packages -->
<!-- - [afex](https://github.com/singmann/afex) for ANOVAs -->
<!-- - [rstatix](https://github.com/kassambara/rstatix) for effect sizes -->

```{r setup, include = FALSE}
# packages
if (!require("here")) {install.packages("here"): require("here")}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# # statistics
# if (!require("afex")) {install.packages("afex"); require("afex")}
# if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}

# read info about teaching experience
source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
```

```{r file, include = FALSE}
filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)

# Include only articulation
dt_ioi_art <- dt_ioi[Skill == "articulation"]
# Include only dynamics
dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
```

# IOI-Articulation 
## 1. All - individual
- Average IOIs for each participant.

```{r ioi-art, echo = FALSE}
# For each individual
ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
setorder(ioi_art_trial, "SubNr", "Condition", "Skill")

ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]

# add teaching info
ioi_art$Teaching <- rep(dt_included$TeachingPiano, each = 2)
ioi_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 2)
ioi_art
```

## 2. All - group
- Average IOIs for each condition.

```{r ioi-art-all, echo = FALSE}
# Group mean
ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Condition, Skill, Teaching)]
ioi_art_all
```

### Plot
#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).

```{r ioi-art-all-box, echo = FALSE}
ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean, teachingexp = dt_included$TeachingPiano, teachingyears = dt_included$TeachingPianoYears)

ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") + geom_hline(yintercept = 188, linetype = "dashed")
```

# IOI-Dynamics
## 1. All - individual
- Average IOIs for each participant.

```{r ioi-dyn, echo = FALSE}
# For each individual
ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
setorder(ioi_dyn_trial, "SubNr", "Condition", "Skill")

ioi_dyn <- ioi_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]

# add teaching info
ioi_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 2)
ioi_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 2)
ioi_dyn
```

## 2. All - group
- Average IOIs for each condition.

```{r ioi-dyn-all, echo = FALSE}
# Group mean
ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Condition, Skill, Teaching)]
ioi_dyn_all
```

### Plot
#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).

```{r ioi-dyn-all-box, echo = FALSE}
ioi_dyn_plot <- data.table(subject = ioi_dyn[Condition == "teaching"]$SubNr, teaching = ioi_dyn[Condition == "teaching"]$Mean, performing = ioi_dyn[Condition == "performing"]$Mean, teachingexp = dt_included$TeachingPiano, teachingyears = dt_included$TeachingPianoYears)

ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") + geom_hline(yintercept = 188, linetype = "dashed")
```

```{r export, include = FALSE}
knitr::purl("teaching_ioi.Rmd")
```

### 2022-08-30 experiment-1/analysis/teaching/teaching_art.R

@@ -1,0 +1,166 @@
+## ----setup, include = FALSE-----------------------------------------------------------------------------------
+# packages
+if (!require("here")) {install.packages("here"): require("here")}
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# # statistics
+# if (!require("afex")) {install.packages("afex"); require("afex")}
+# if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
+
+# read info about teaching experience
+source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+
+
+## ----file, include = FALSE------------------------------------------------------------------------------------
+filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
+filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
+
+
+## ----extract, include = FALSE---------------------------------------------------------------------------------
+dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
+dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
+
+# SubNr as a factor
+dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
+dt_kot$SubNr <- as.factor(dt_kot$SubNr)
+
+# For each individual, each block, each trial
+ioi <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]
+
+# Calculate KOR for each interval (outliers were already determined based on KOT)
+dt_kor <- data.table()
+for (subnr in unique(dt_kot$SubNr)){
+  for (block in unique(dt_kot[SubNr == subnr]$BlockNr)){
+    for (trial in unique(dt_kot[SubNr == subnr & BlockNr == block]$TrialNr)){
+      dt_current <- dt_kot[SubNr == subnr & BlockNr == block & TrialNr == trial]
+      dt_current$KOR <- dt_current$KOT/ioi[SubNr == subnr & ioi$BlockNr == block & TrialNr == trial]$Mean
[…]

### 2022-08-30 experiment-1/analysis/teaching/teaching_dyn.R

@@ -1,0 +1,154 @@
+## ----setup, include = FALSE-----------------------------------------------------------------------------------
+# packages
+if (!require("here")) {install.packages("here"): require("here")}
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# # statistic
+# if (!require("afex")) {install.packages("afex"); require("afex")}
+# if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
+
+# read info about teaching experience
+source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+
+
+## ----file, include = FALSE------------------------------------------------------------------------------------
+filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
+filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
+
+
+## ----extract, include = FALSE---------------------------------------------------------------------------------
+dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
+dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
+
+# SubNr as a factor
+dt_vel$SubNr <- as.factor(dt_vel$SubNr)
+dt_vel_diff$SubNr <- as.factor(dt_vel_diff$SubNr)
+
+# Include only articulation
+dt_vel_art <- dt_vel[Skill == "articulation"]
+dt_vel_diff_art <- dt_vel_diff[Skill == "articulation"]
+
+# Include only dynamics
+dt_vel_dyn <- dt_vel[Skill == "dynamics"]
+dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
+
+
+## ----vel-dyn, echo = FALSE------------------------------------------------------------------------------------
+# For each individual
+vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
[…]

### 2022-08-30 experiment-1/analysis/teaching/teaching_ioi.R

@@ -1,0 +1,86 @@
+## ----setup, include = FALSE-----------------------------------------------------------------------------------
+# packages
+if (!require("here")) {install.packages("here"): require("here")}
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# # statistics
+# if (!require("afex")) {install.packages("afex"); require("afex")}
+# if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
+
+# read info about teaching experience
+source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+
+
+## ----file, include = FALSE------------------------------------------------------------------------------------
+filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
+
+
+## ----extract, include = FALSE---------------------------------------------------------------------------------
+dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
+
+# SubNr as a factor
+dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
+
+# Include only articulation
+dt_ioi_art <- dt_ioi[Skill == "articulation"]
+# Include only dynamics
+dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
+
+
+## ----ioi-art, echo = FALSE------------------------------------------------------------------------------------
+# For each individual
+ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+
+ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
+setorder(ioi_art, "SubNr", "Condition", "Skill")
+
+# add teaching info
+ioi_art$Teaching <- rep(dt_included$TeachingPiano, each = 2)
+ioi_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 2)
+
+ioi_art
+
+
+## ----ioi-art-all, echo = FALSE--------------------------------------------------------------------------------
+# Group mean
[…]

### 2022-08-30 experiment-2/analysis/teaching/teaching_art.Rmd
upload analysis between those who had teaching experience vs those who didn't
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-08-30 experiment-1/analysis/teaching/teaching_art.Rmd
@@
 ---
-title: 'Articulation Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 1'
+title: 'Articulation Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 2'
 output:
@@
 2. Excluded participants
-
-- SubNr 1: A deviated tempo
-- SubNr 3: Experimental error
-- SubNr 8: Experimental error
-- SubNr 14: Experimental error
-- SubNr 16: A deviated tempo
+- SubNr 3: Insufficient motor skills
 
@@
 # read info about teaching experience
-source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
 ```
@@
 # SubNr as a factor
-dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
+dt_kor$SubNr <- as.factor(dt_kor$SubNr)
 dt_kot$SubNr <- as.factor(dt_kot$SubNr)
 
-# For each individual, each block, each trial
-ioi <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]
-
-# Calculate KOR for each interval (outliers were already determined based on KOT)
-dt_kor <- data.table()
-for (subnr in unique(dt_kot$SubNr)){
-  for (block in unique(dt_kot[SubNr == subnr]$BlockNr)){
-    for (trial in unique(dt_kot[SubNr == subnr & BlockNr == block]$TrialNr)){
-      dt_current <- dt_kot[SubNr == subnr & BlockNr == block & TrialNr == trial]
-      dt_current$KOR <- dt_current$KOT/ioi[SubNr == subnr & ioi$BlockNr == block & TrialNr == trial]$Mean
-      dt_kor <- rbind(dt_kor, dt_current)
-    }
-  }
-}
-
 # Include only articulation
+dt_kor_art <- dt_kor[Skill == "articulation"]
 dt_kot_art <- dt_kot[Skill == "articulation"]
-dt_kor_art <- dt_kor[Skill == "articulation"]
 
 # Include only dynamics
+dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
-dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 ```
@@
 # add teaching info
-kot_art$Teaching <- rep(dt_included$TeachingPiano, each = 4)
-kot_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
+kot_art$Teaching <- rep(dt_2_included$TeachingPiano, each = 4)
+kot_art$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 4)
 kot_art
@@
 ```{r kot-art-all-box, echo = FALSE}
-kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))
+kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_2_included$TeachingPiano, each = 2), teachingyears = rep(dt_2_included$TeachingPianoYears, each = 2))
 
@@
 # add teaching info
-kor_art$Teaching <- rep(dt_included$TeachingPiano, each = 4)
-kor_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
+kor_art$Teaching <- rep(dt_2_included$TeachingPiano, each = 4)
+kor_art$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 4)
+
 kor_art
@@
 ```{r kor-art-all-box, echo = FALSE}
-kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))
+kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_2_included$TeachingPiano, each = 2), teachingyears = rep(dt_2_included$TeachingPianoYears, each = 2))
 
@@
 # add teaching info
-kot_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 4)
-kot_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
+kot_dyn$Teaching <- rep(dt_2_included$TeachingPiano, each = 4)
+kot_dyn$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 4)
 kot_dyn
@@
 ```{r kot-dyn-all-box, echo = FALSE}
-kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))
+kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_2_included$TeachingPiano, each = 2), teachingyears = rep(dt_2_included$TeachingPianoYears, each = 2))
 
@@
 # add teaching info
-kor_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 4)
-kor_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
+kor_dyn$Teaching <- rep(dt_2_included$TeachingPiano, each = 4)
+kor_dyn$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 4)
 kor_dyn
@@
 ```{r kor-dyn-all-box, echo = FALSE}
-kor_dyn_plot <- data.table(subject = kor_dyn[Condition == "teaching"]$SubNr, teaching = kor_dyn[Condition == "teaching"]$Mean, performing = kor_dyn[Condition == "performing"]$Mean, subcomponent = kor_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))
+kor_dyn_plot <- data.table(subject = kor_dyn[Condition == "teaching"]$SubNr, teaching = kor_dyn[Condition == "teaching"]$Mean, performing = kor_dyn[Condition == "performing"]$Mean, subcomponent = kor_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_2_included$TeachingPiano, each = 2), teachingyears = rep(dt_2_included$TeachingPianoYears, each = 2))
 

### 2022-08-30 experiment-2/analysis/teaching/teaching_dyn.Rmd
upload analysis between those who had teaching experience vs those who didn't
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-08-30 experiment-1/analysis/teaching/teaching_dyn.Rmd
@@
 ---
-title: 'Dynamics Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 1'
+title: 'Dynamics Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 2'
 output:
@@
 # add teaching info
-vel_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 4)
-vel_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
+vel_dyn$Teaching <- rep(dt_2_included$TeachingPiano, each = 4)
+vel_dyn$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 4)
 vel_dyn
@@
 ```{r vel-dyn-all-box, echo = FALSE}
-vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))
+vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_2_included$TeachingPiano, each = 2), teachingyears = rep(dt_2_included$TeachingPianoYears, each = 2))
 
@@
 vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
-
-vel_diff_dyn <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 setorder(vel_diff_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")
@@
 # add teaching info
-vel_diff_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 4)
-vel_diff_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
+vel_diff_dyn$Teaching <- rep(dt_2_included$TeachingPiano, each = 4)
+vel_diff_dyn$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 4)
 vel_diff_dyn
@@
 ```{r vel-diff-dyn-all-box, echo = FALSE}
-vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))
+vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_2_included$TeachingPiano, each = 2), teachingyears = rep(dt_2_included$TeachingPianoYears, each = 2))
 
@@
 vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(vel_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
 vel_art <- vel_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
-setorder(vel_art, "SubNr", "Condition", "Skill", "Subcomponent")
 
 # add teaching info
-vel_art$Teaching <- rep(dt_included$TeachingPiano, each = 4)
-vel_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
-
+vel_art$Teaching <- rep(dt_2_included$TeachingPiano, each = 4)
+vel_art$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 4)
 vel_art
@@
 ```{r vel-art-all-box, echo = FALSE}
-vel_art_plot <- data.table(subject = vel_art[Condition == "teaching"]$SubNr, teaching = vel_art[Condition == "teaching"]$Mean, performing = vel_art[Condition == "performing"]$Mean, subcomponent = vel_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))
+vel_art_plot <- data.table(subject = vel_art[Condition == "teaching"]$SubNr, teaching = vel_art[Condition == "teaching"]$Mean, performing = vel_art[Condition == "performing"]$Mean, subcomponent = vel_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_2_included$TeachingPiano, each = 2), teachingyears = rep(dt_2_included$TeachingPianoYears, each = 2))
 
@@
 # add teaching info
-vel_diff_art$Teaching <- rep(dt_included$TeachingPiano, each = 4)
-vel_diff_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
+vel_diff_art$Teaching <- rep(dt_2_included$TeachingPiano, each = 4)
+vel_diff_art$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 4)
 vel_diff_art
@@
 ```{r vel-diff-art-all-box, echo = FALSE}
-vel_diff_art_plot <- data.table(subject = vel_diff_art[Condition == "teaching"]$SubNr, teaching = vel_diff_art[Condition == "teaching"]$Mean, performing = vel_diff_art[Condition == "performing"]$Mean, subcomponent = vel_diff_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))
+vel_diff_art_plot <- data.table(subject = vel_diff_art[Condition == "teaching"]$SubNr, teaching = vel_diff_art[Condition == "teaching"]$Mean, performing = vel_diff_art[Condition == "performing"]$Mean, subcomponent = vel_diff_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_2_included$TeachingPiano, each = 2), teachingyears = rep(dt_2_included$TeachingPianoYears, each = 2))
 

### 2022-08-30 experiment-2/analysis/teaching/teaching_ioi.Rmd
upload analysis between those who had teaching experience vs those who didn't
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-08-30 experiment-1/analysis/teaching/teaching_ioi.Rmd
@@
 ---
-title: 'Tempo Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 1'
+title: 'Tempo Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 2'
 output:
@@
 # add teaching info
-ioi_art$Teaching <- rep(dt_included$TeachingPiano, each = 2)
-ioi_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 2)
+ioi_art$Teaching <- rep(dt_2_included$TeachingPiano, each = 2)
+ioi_art$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 2)
 ioi_art
@@
 ```{r ioi-art-all-box, echo = FALSE}
-ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean, teachingexp = dt_included$TeachingPiano, teachingyears = dt_included$TeachingPianoYears)
+ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean, teachingexp = dt_2_included$TeachingPiano, teachingyears = dt_2_included$TeachingPianoYears)
 
-ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") + geom_hline(yintercept = 188, linetype = "dashed")
+ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation")
 ```
@@
 # add teaching info
-ioi_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 2)
-ioi_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 2)
+ioi_dyn$Teaching <- rep(dt_2_included$TeachingPiano, each = 2)
+ioi_dyn$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 2)
 ioi_dyn
@@
 ```{r ioi-dyn-all-box, echo = FALSE}
-ioi_dyn_plot <- data.table(subject = ioi_dyn[Condition == "teaching"]$SubNr, teaching = ioi_dyn[Condition == "teaching"]$Mean, performing = ioi_dyn[Condition == "performing"]$Mean, teachingexp = dt_included$TeachingPiano, teachingyears = dt_included$TeachingPianoYears)
+ioi_dyn_plot <- data.table(subject = ioi_dyn[Condition == "teaching"]$SubNr, teaching = ioi_dyn[Condition == "teaching"]$Mean, performing = ioi_dyn[Condition == "performing"]$Mean, teachingexp = dt_2_included$TeachingPiano, teachingyears = dt_2_included$TeachingPianoYears)
 
-ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") + geom_hline(yintercept = 188, linetype = "dashed")
+ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics")
 ```

### 2022-08-30 experiment-2/analysis/teaching/teaching_art.R

@@ -1,0 +1,152 @@
+## ----setup, include = FALSE-----------------------------------------------------------------------------------
+# packages
+if (!require("here")) {install.packages("here"): require("here")}
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# # statistics
+# if (!require("afex")) {install.packages("afex"); require("afex")}
+# if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
+
+# read info about teaching experience
+source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+
+
+## ----file, include = FALSE------------------------------------------------------------------------------------
+filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
+filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
+
+
+## ----extract, include = FALSE---------------------------------------------------------------------------------
+dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
+dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
+
+# SubNr as a factor
+dt_kor$SubNr <- as.factor(dt_kor$SubNr)
+dt_kot$SubNr <- as.factor(dt_kot$SubNr)
+
+# Include only articulation
+dt_kor_art <- dt_kor[Skill == "articulation"]
+dt_kot_art <- dt_kot[Skill == "articulation"]
+
+# Include only dynamics
+dt_kor_dyn <- dt_kor[Skill == "dynamics"]
+dt_kot_dyn <- dt_kot[Skill == "dynamics"]
+
+
+## ----kot-art, echo = FALSE------------------------------------------------------------------------------------
+# For each individual
+kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
[…]

### 2022-08-30 experiment-2/analysis/teaching/teaching_dyn.R

@@ -1,0 +1,154 @@
+## ----setup, include = FALSE-----------------------------------------------------------------------------------
+# packages
+if (!require("here")) {install.packages("here"): require("here")}
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# # statistic
+# if (!require("afex")) {install.packages("afex"); require("afex")}
+# if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
+
+# read info about teaching experience
+source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+
+
+## ----file, include = FALSE------------------------------------------------------------------------------------
+filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
+filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
+
+
+## ----extract, include = FALSE---------------------------------------------------------------------------------
+dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
+dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
+
+# SubNr as a factor
+dt_vel$SubNr <- as.factor(dt_vel$SubNr)
+dt_vel_diff$SubNr <- as.factor(dt_vel_diff$SubNr)
+
+# Include only articulation
+dt_vel_art <- dt_vel[Skill == "articulation"]
+dt_vel_diff_art <- dt_vel_diff[Skill == "articulation"]
+
+# Include only dynamics
+dt_vel_dyn <- dt_vel[Skill == "dynamics"]
+dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
+
+
+## ----vel-dyn, echo = FALSE------------------------------------------------------------------------------------
+# For each individual
+vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+
[…]

### 2022-08-30 experiment-2/analysis/teaching/teaching_ioi.R

@@ -1,0 +1,86 @@
+## ----setup, include = FALSE-----------------------------------------------------------------------------------
+# packages
+if (!require("here")) {install.packages("here"): require("here")}
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+# # statistics
+# if (!require("afex")) {install.packages("afex"); require("afex")}
+# if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
+
+# read info about teaching experience
+source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
+
+
+## ----file, include = FALSE------------------------------------------------------------------------------------
+filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
+
+
+## ----extract, include = FALSE---------------------------------------------------------------------------------
+dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
+
+# SubNr as a factor
+dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
+
+# Include only articulation
+dt_ioi_art <- dt_ioi[Skill == "articulation"]
+# Include only dynamics
+dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
+
+
+## ----ioi-art, echo = FALSE------------------------------------------------------------------------------------
+# For each individual
+ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+
+ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
+setorder(ioi_art, "SubNr", "Condition", "Skill")
+
+# add teaching info
+ioi_art$Teaching <- rep(dt_2_included$TeachingPiano, each = 2)
+ioi_art$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 2)
+
+ioi_art
+
+
+## ----ioi-art-all, echo = FALSE--------------------------------------------------------------------------------
+# Group mean
[…]

### 2022-08-30 experiment-2/analysis/stats/articulation.Rmd
change fig width/height / remove unnecessary analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-ggline(kor_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOR", title = "KOT: Articulation") +
+ggline(kor_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOR", title = "KOR: Articulation") +
   scale_x_continuous(breaks = seq(1,71,1))
@@
 
-ggline(kor_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOR", title = "KOT: Dynamics")+
+ggline(kor_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Interval", ylab = "KOR", title = "KOR: Dynamics")+
   scale_x_continuous(breaks = seq(1,71,1)) 

### 2022-08-30 experiment-2/analysis/stats/dynamics.Rmd
change fig width/height / remove unnecessary analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-## Normalised KV difference per participant
-```{r, echo = FALSE}
-vel_diff_dyn$NormalisedDiff <- 0
-for (i in unique(vel_diff_dyn_trial$SubNr)){
-  # teaching
-  # FtoP
-  vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
-  # PtoF
-  vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "teaching" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn[SubNr == i & Condition == "teaching"]$Mean)-min(vel_dyn[SubNr == i & Condition == "teaching"]$Mean))
-  # performing
-  # FtoP
-  vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "FtoP"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "FtoP"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
-  # PtoF
-  vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "PtoF"]$NormalisedDiff <- vel_diff_dyn[SubNr == i & Condition == "performing" & Subcomponent == "PtoF"]$Mean/(max(vel_dyn[SubNr == i & Condition == "performing"]$Mean)-min(vel_dyn[SubNr == i & Condition == "performing"]$Mean))
-}
-
-# group
-vel_diff_dyn_normalised_all_2 <- vel_diff_dyn[, .(N = .N, Mean = mean(NormalisedDiff), SD = sd(NormalisedDiff), SEM = sd(NormalisedDiff)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
-vel_diff_dyn_normalised_all_2
-```
-
-### Plot
-```{r, echo = FALSE}
-vel_diff_dyn_plot_3 <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$NormalisedDiff, performing = vel_diff_dyn[Condition == "performing"]$NormalisedDiff, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)
-
-ggpaired(vel_diff_dyn_plot_3, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
-```
-
-### Stats
-```{r, echo = FALSE}
-# stats
-vel_diff_dyn_aov_4 <- aov_car(NormalisedDiff ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
-summary(vel_diff_dyn_aov_4)
-```
-
 ```{r export, include = FALSE}

### 2022-08-30 experiment-2/analysis/stats/variability.Rmd
change the title of the analysis page
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
change fig width/height / remove unnecessary analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: "Variability analysis 2"
+title: "Variability analysis: The Sound of Teaching Music 2"
 output:
@@
 ioi_cv_trial <- dt_ioi[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI), CV = sd(normIOI)/mean(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+setorder(ioi_cv_trial, "SubNr", "Condition", "Skill")
+
 ioi_cv_subject <- ioi_cv_trial[, .(N = .N, MeanCV = mean(CV), SD = sd(CV)), by = .(SubNr, Condition, Skill)]

### 2022-08-30 experiment-2/analysis/stats/articulation.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,266 +31,270 @@ dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 
 
-## ----kor-art, echo = FALSE----------------------------------------------------
+## ----kor-art, echo = FALSE----------------------------------
 # For each individual
 kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(kor_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
 kor_art <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 kor_art
 
 
-## ----kor-art-box,  echo = FALSE-----------------------------------------------
+## ----kor-art-box,  echo = FALSE-----------------------------
[…]

### 2022-08-30 experiment-2/analysis/stats/dynamics.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE-----------------------------------------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,12 +9,12 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE------------------------------------------------------------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE---------------------------------------------------------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -31,276 +31,280 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE----------------------------------------------------
+## ----vel-dyn, echo = FALSE------------------------------------------------------------------------------------
 # For each individual
 vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(vel_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
 vel_dyn <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
 vel_dyn
 
 
-## ----vel-dyn-box,  echo = FALSE-----------------------------------------------
[…]

### 2022-08-30 experiment-2/analysis/stats/ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
@@ -9,11 +9,11 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE-------------------------------------------------
+## ----extract, include = FALSE-------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -25,242 +25,246 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE----------------------------------------------------
+## ----ioi-art, echo = FALSE----------------------------------
 # For each individual
 ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+setorder(ioi_art_trial, "SubNr", "Condition", "Skill")
 
 ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
 ioi_art
 
 
-## ----ioi-art-box,  echo = FALSE-----------------------------------------------
+## ----ioi-art-box,  echo = FALSE-----------------------------
 ggboxplot(ioi_art_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Normalised IOIs", title = "IOI: Articulation") + geom_hline(yintercept = 1, linetype = "dashed")
 
 
-## ----ioi-art-all, echo = FALSE------------------------------------------------
[…]

### 2022-08-31 experiment-2/analysis/teaching/teaching_ioi.Rmd
use normIOIs / draw a dashed line to 1
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # For each individual
-ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
 setorder(ioi_art_trial, "SubNr", "Condition", "Skill")
@@
 #### Box plot
-- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).
+A dashed line represents the tempo given by a metronome.
 
@@
 
-ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation")
+ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Articulation") + geom_hline(yintercept = 1, linetype = "dashed")
 ```
@@
 # For each individual
-ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
 setorder(ioi_dyn_trial, "SubNr", "Condition", "Skill")
@@
 
-ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics")
+ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOI: Dynamics") + geom_hline(yintercept = 1, linetype = "dashed")
 ```

### 2022-08-31 experiment-2/analysis/teaching/teaching_art.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE-----------------------------------------------------------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 if (!require("here")) {install.packages("here"): require("here")}
 # data manipulation
@@ -13,12 +13,12 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
 
 
-## ----file, include = FALSE------------------------------------------------------------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_kor = "../preprocessor/trimmed/data_analysis_kor.txt"
 filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
 
 
-## ----extract, include = FALSE---------------------------------------------------------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_kor <- fread(filename_kor, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -35,27 +35,26 @@ dt_kor_dyn <- dt_kor[Skill == "dynamics"]
 dt_kot_dyn <- dt_kot[Skill == "dynamics"]
 
 
-## ----kot-art, echo = FALSE------------------------------------------------------------------------------------
+## ----kot-art, echo = FALSE----------------------------------------------------
 # For each individual
 kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(kot_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
 kot_art <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
-setorder(kot_art, "SubNr", "Condition", "Skill", "Subcomponent")
 
 # add teaching info
 kot_art$Teaching <- rep(dt_2_included$TeachingPiano, each = 4)
[…]

### 2022-08-31 experiment-2/analysis/teaching/teaching_dyn.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE-----------------------------------------------------------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 if (!require("here")) {install.packages("here"): require("here")}
 # data manipulation
@@ -13,12 +13,12 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
 
 
-## ----file, include = FALSE------------------------------------------------------------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
 filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
 
 
-## ----extract, include = FALSE---------------------------------------------------------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
 dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt
 
@@ -35,27 +35,26 @@ dt_vel_dyn <- dt_vel[Skill == "dynamics"]
 dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
 
 
-## ----vel-dyn, echo = FALSE------------------------------------------------------------------------------------
+## ----vel-dyn, echo = FALSE----------------------------------------------------
 # For each individual
 vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
+setorder(vel_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")
 
 vel_dyn <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
-setorder(vel_dyn, "SubNr", "Condition", "Skill", "Subcomponent")
 
 # add teaching info
[…]

### 2022-08-31 experiment-2/analysis/teaching/teaching_ioi.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE-----------------------------------------------------------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 if (!require("here")) {install.packages("here"): require("here")}
 # data manipulation
@@ -13,11 +13,11 @@ if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
 
 
-## ----file, include = FALSE------------------------------------------------------------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
 
 
-## ----extract, include = FALSE---------------------------------------------------------------------------------
+## ----extract, include = FALSE-------------------------------------------------
 dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
@@ -29,58 +29,56 @@ dt_ioi_art <- dt_ioi[Skill == "articulation"]
 dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
 
 
-## ----ioi-art, echo = FALSE------------------------------------------------------------------------------------
+## ----ioi-art, echo = FALSE----------------------------------------------------
 # For each individual
-ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
+setorder(ioi_art_trial, "SubNr", "Condition", "Skill")
 
 ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]
-setorder(ioi_art, "SubNr", "Condition", "Skill")
 
 # add teaching info
 ioi_art$Teaching <- rep(dt_2_included$TeachingPiano, each = 2)
 ioi_art$TeachingYears <- rep(dt_2_included$TeachingPianoYears, each = 2)
-
 ioi_art
 
 
[…]

### 2022-08-31 paper/manuscript.Rmd
upload analysis for teaching experience
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# Instructions
+# 1. Instructions
 
@@
 
-# Temporal Variability
+# 2. Temporal Variability
 
@@
 
-# Temporal grouping
+# 3. Temporal grouping
 
@@
 
-# Additional analysis for dynamics difference
+# 4. Additional analysis for dynamics difference
 
@@
 ```
+
+\clearpage
+
+# 5. Descriptive statistics depending on participants' teaching experience
+
+## Experiment 1
+```{r teaching-1, include = FALSE}
+source(here("experiment-1/analysis/teaching", "teaching_ioi.R"), chdir = TRUE)
+source(here("experiment-1/analysis/teaching", "teaching_art.R"), chdir = TRUE)
+source(here("experiment-1/analysis/teaching", "teaching_dyn.R"), chdir = TRUE)
+```
+
+```{r ioi-teaching-1, include = FALSE}
+# articulation
+ioi_art_plot$teachingexp <- factor(ioi_art_plot$teachingexp, levels = c("Yes", "No"))
+
+ioi_art_teaching_1 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Articulation") + geom_hline(yintercept = 188, linetype = "dashed") +
+  ylim(150, 260) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_art_teaching_1.png"), plot = ioi_art_teaching_1, dpi = 300, width = 7, height = 5)
+img_ioi_art_teaching_1 <- image_read(here("paper/image", "ioi_art_teaching_1.png"), density = 300)
+
+# dynamics
+ioi_dyn_plot$teachingexp <- factor(ioi_dyn_plot$teachingexp, levels = c("Yes", "No"))
+
+ioi_dyn_teaching_1 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Dynamics") + geom_hline(yintercept = 188, linetype = "dashed") +
+  ylim(150, 260) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_dyn_teaching_1.png"), plot = ioi_dyn_teaching_1, dpi = 300, width = 7, height = 5)
+img_ioi_dyn_teaching_1 <- image_read(here("paper/image", "ioi_dyn_teaching_1.png"), density = 300)
+```
+
+```{r art-teaching-1, include = FALSE}
+kot_art_plot$teachingexp <- factor(kot_art_plot$teachingexp, levels = c("Yes", "No"))
+
+art_leg_teaching_1 <- ggpaired(kot_art_plot[subcomponent == "Legato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation-Legato") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "art_leg_teaching_1.png"), plot = art_leg_teaching_1, dpi = 300, width = 7, height = 5)
+img_art_leg_teaching_1 <- image_read(here("paper/image", "art_leg_teaching_1.png"), density = 300)
+
+art_sta_teaching_1 <- ggpaired(kot_art_plot[subcomponent == "Staccato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation-Staccato") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "art_sta_teaching_1.png"), plot = art_sta_teaching_1, dpi = 300, width = 7, height = 5)
+img_art_sta_teaching_1 <- image_read(here("paper/image", "art_sta_teaching_1.png"), density = 300)
+```
+
+```{r dyn-teaching-1, include = FALSE}
+vel_dyn_plot$teachingexp <- factor(vel_dyn_plot$teachingexp, levels = c("Yes", "No"))
+
+vel_for_teaching_1 <- ggpaired(vel_dyn_plot[subcomponent == "Forte"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics-Forte") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_for_teaching_1.png"), plot = vel_for_teaching_1, dpi = 300, width = 7, height = 5)
+img_vel_for_teaching_1 <- image_read(here("paper/image", "vel_for_teaching_1.png"), density = 300)
+
+vel_pia_teaching_1 <- ggpaired(vel_dyn_plot[subcomponent == "Piano"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics-Piano") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_pia_teaching_1.png"), plot = vel_pia_teaching_1, dpi = 300, width = 7, height = 5)
+img_vel_pia_teaching_1 <- image_read(here("paper/image", "vel_pia_teaching_1.png"), density = 300)
+```
+
+```{r dyn-diff-teaching-1, include = FALSE}
+vel_diff_dyn_plot$teachingexp <- factor(vel_diff_dyn_plot$teachingexp, levels = c("Yes", "No"))
+
+vel_diff_for_teaching_1 <- ggpaired(vel_diff_dyn_plot[subcomponent == "FtoP"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics-FtoP") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_diff_for_teaching_1.png"), plot = vel_diff_for_teaching_1, dpi = 300, width = 7, height = 5)
+img_vel_diff_for_teaching_1 <- image_read(here("paper/image", "vel_diff_for_teaching_1.png"), density = 300)
+
+vel_diff_pia_teaching_1 <- ggpaired(vel_diff_dyn_plot[subcomponent == "PtoF"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics-PtoF") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_diff_pia_teaching_1.png"), plot = vel_diff_pia_teaching_1, dpi = 300, width = 7, height = 5)
+img_vel_diff_pia_teaching_1 <- image_read(here("paper/image", "vel_diff_pia_teaching_1.png"), density = 300)
+```
+
+```{r plot-ioi-teaching-1, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-1}BAA"}
+img_ioi_teaching_1 <- image_append(c(img_ioi_art_teaching_1, img_ioi_dyn_teaching_1))
+image_write(img_ioi_teaching_1, path = here("paper/figure", "Fig13.tiff"), format = "tiff")
+img_ioi_teaching_1
+```
+
+```{r plot-art-teaching-1, out.width = "100%", fig.cap = "\\label{fig:art-teaching-1}BAA"}
+img_art_teaching_1 <- image_append(c(img_art_leg_teaching_1, img_art_sta_teaching_1))
+image_write(img_art_teaching_1, path = here("paper/figure", "Fig14.tiff"), format = "tiff")
+img_art_teaching_1
+```
+
+```{r plot-dyn-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-1}BAA"}
+img_dyn_teaching_1 <- image_append(c(img_vel_for_teaching_1, img_vel_pia_teaching_1))
+image_write(img_dyn_teaching_1, path = here("paper/figure", "Fig15.tiff"), format = "tiff")
+img_dyn_teaching_1
+```
+
+```{r plot-dyn-diff-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-1}BAA"}
+img_dyn_diff_teaching_1 <- image_append(c(img_vel_diff_for_teaching_1, img_vel_diff_pia_teaching_1))
+image_write(img_dyn_diff_teaching_1, path = here("paper/figure", "Fig16.tiff"), format = "tiff")
+img_dyn_diff_teaching_1
+```
+
+```{r ioi-teaching-desc-1, echo = FALSE}
+ioi_all <- rbind(ioi_art_all, ioi_dyn_all)
+colnames(ioi_all)[2] <- "Technique"
+ioi_all[Condition == "teaching"]$Condition <- "Teaching"
+ioi_all[Condition == "performing"]$Condition <- "Performing"
+ioi_all[Technique == "articulation"]$Technique <- "Articulation"
+ioi_all[Technique == "dynamics"]$Technique <- "Dynamics"
+
+apa_table(
+  ioi_all
+  , caption = "Caption"
+  , note = "Note"
+  , escape = TRUE
+)
+```
+
+```{r art-teaching-desc-1, echo = FALSE}
+colnames(kot_art_all)[2] <- "Technique"
+kot_art_all[Condition == "teaching"]$Condition <- "Teaching"
+kot_art_all[Condition == "performing"]$Condition <- "Performing"
+kot_art_all[Technique == "articulation"]$Technique <- "Articulation"
+
+apa_table(
+  kot_art_all
+  , caption = "Caption"
+  , note = "Note"
+  , escape = TRUE
+)
+```
+
+```{r dyn-teaching-desc-1}
+colnames(vel_dyn_all)[2] <- "Technique"
+vel_dyn_all[Condition == "teaching"]$Condition <- "Teaching"
+vel_dyn_all[Condition == "performing"]$Condition <- "Performing"
+vel_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"
+
+apa_table(
+  vel_dyn_all
+  , caption = "Caption"
+  , note = "Note"
+  , escape = TRUE
+)
+```
+
+```{r dyn-diff-teaching-desc-1}
+colnames(vel_diff_dyn_all)[2] <- "Technique"
+vel_diff_dyn_all[Condition == "teaching"]$Condition <- "Teaching"
+vel_diff_dyn_all[Condition == "performing"]$Condition <- "Performing"
+vel_diff_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"
+
+apa_table(
+  vel_diff_dyn_all
+  , caption = "Caption"
+  , note = "Note"
+  , escape = TRUE
+)
+```
+
+\newpage
+
+## Experiment 2
+```{r teaching-2, include = FALSE}
+source(here("experiment-2/analysis/teaching", "teaching_ioi.R"), chdir = TRUE)
+source(here("experiment-2/analysis/teaching", "teaching_art.R"), chdir = TRUE)
+source(here("experiment-2/analysis/teaching", "teaching_dyn.R"), chdir = TRUE)
+```
+
+```{r ioi-teaching-2, include = FALSE}
+# articulation
+ioi_art_plot$teachingexp <- factor(ioi_art_plot$teachingexp, levels = c("Yes", "No"))
+
+ioi_art_teaching_2 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
+  ylim(0.8, 1.1) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_art_teaching_2.png"), plot = ioi_art_teaching_2, dpi = 300, width = 7, height = 5)
+img_ioi_art_teaching_2 <- image_read(here("paper/image", "ioi_art_teaching_2.png"), density = 300)
+
+# dynamics
+ioi_dyn_plot$teachingexp <- factor(ioi_dyn_plot$teachingexp, levels = c("Yes", "No"))
+
+ioi_dyn_teaching_2 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Dynamics") +
+  ylim(0.8, 1.1) +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_dyn_teaching_2.png"), plot = ioi_dyn_teaching_2, dpi = 300, width = 7, height = 5)
+img_ioi_dyn_teaching_2 <- image_read(here("paper/image", "ioi_dyn_teaching_2.png"), density = 300)
+```
+
+```{r art-teaching-2, include = FALSE}
+kor_art_plot$teachingexp <- factor(kor_art_plot$teachingexp, levels = c("Yes", "No"))
+
+art_leg_teaching_2 <- ggpaired(kor_art_plot[subcomponent == "Legato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation-Legato") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "art_leg_teaching_2.png"), plot = art_leg_teaching_2, dpi = 300, width = 7, height = 5)
+img_art_leg_teaching_2 <- image_read(here("paper/image", "art_leg_teaching_2.png"), density = 300)
+
+art_sta_teaching_2 <- ggpaired(kor_art_plot[subcomponent == "Staccato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation-Staccato") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "art_sta_teaching_2.png"), plot = art_sta_teaching_2, dpi = 300, width = 7, height = 5)
+img_art_sta_teaching_2 <- image_read(here("paper/image", "art_sta_teaching_2.png"), density = 300)
+```
+
+```{r dyn-teaching-2, include = FALSE}
+vel_dyn_plot$teachingexp <- factor(vel_dyn_plot$teachingexp, levels = c("Yes", "No"))
+
+vel_for_teaching_2 <- ggpaired(vel_dyn_plot[subcomponent == "Forte"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics-Forte") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_for_teaching_2.png"), plot = vel_for_teaching_2, dpi = 300, width = 7, height = 5)
+img_vel_for_teaching_2 <- image_read(here("paper/image", "vel_for_teaching_2.png"), density = 300)
+
+vel_pia_teaching_2 <- ggpaired(vel_dyn_plot[subcomponent == "Piano"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics-Piano") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_pia_teaching_2.png"), plot = vel_pia_teaching_2, dpi = 300, width = 7, height = 5)
+img_vel_pia_teaching_2 <- image_read(here("paper/image", "vel_pia_teaching_2.png"), density = 300)
+```
+
+```{r dyn-diff-teaching-2, include = FALSE}
+vel_diff_dyn_plot$teachingexp <- factor(vel_diff_dyn_plot$teachingexp, levels = c("Yes", "No"))
+
+vel_diff_for_teaching_2 <- ggpaired(vel_diff_dyn_plot[subcomponent == "FtoP"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics-FtoP") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_diff_for_teaching_2.png"), plot = vel_diff_for_teaching_2, dpi = 300, width = 7, height = 5)
+img_vel_diff_for_teaching_2 <- image_read(here("paper/image", "vel_diff_for_teaching_2.png"), density = 300)
+
+vel_diff_pia_teaching_2 <- ggpaired(vel_diff_dyn_plot[subcomponent == "PtoF"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics-PtoF") + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_diff_pia_teaching_2.png"), plot = vel_diff_pia_teaching_2, dpi = 300, width = 7, height = 5)
+img_vel_diff_pia_teaching_2 <- image_read(here("paper/image", "vel_diff_pia_teaching_2.png"), density = 300)
+```
+
+```{r plot-ioi-teaching-2, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-2}BAA"}
+img_ioi_teaching_2 <- image_append(c(img_ioi_art_teaching_2, img_ioi_dyn_teaching_2))
+image_write(img_ioi_teaching_2, path = here("paper/figure", "Fig17.tiff"), format = "tiff")
+img_ioi_teaching_2
+```
+
+```{r plot-art-teaching-2, out.width = "100%", fig.cap = "\\label{fig:art-teaching-2}BAA"}
+img_art_teaching_2 <- image_append(c(img_art_leg_teaching_2, img_art_sta_teaching_2))
+image_write(img_art_teaching_2, path = here("paper/figure", "Fig18.tiff"), format = "tiff")
+img_art_teaching_2
+```
+
+```{r plot-dyn-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-2}BAA"}
+img_dyn_teaching_2 <- image_append(c(img_vel_for_teaching_2, img_vel_pia_teaching_2))
+image_write(img_dyn_teaching_2, path = here("paper/figure", "Fig19.tiff"), format = "tiff")
+img_dyn_teaching_2
+```
+
+```{r plot-dyn-diff-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-2}BAA"}
+img_dyn_diff_teaching_2 <- image_append(c(img_vel_diff_for_teaching_2, img_vel_diff_pia_teaching_2))
+image_write(img_dyn_diff_teaching_2, path = here("paper/figure", "Fig20.tiff"), format = "tiff")
+img_dyn_diff_teaching_2
+```
+
+```{r ioi-teaching-desc-2, echo = FALSE}
+ioi_all <- rbind(ioi_art_all, ioi_dyn_all)
+colnames(ioi_all)[2] <- "Technique"
+ioi_all[Condition == "teaching"]$Condition <- "Teaching"
+ioi_all[Condition == "performing"]$Condition <- "Performing"
+ioi_all[Technique == "articulation"]$Technique <- "Articulation"
+ioi_all[Technique == "dynamics"]$Technique <- "Dynamics"
+ioi_all$Teaching <- factor(ioi_all$Teaching, levels = c("Yes", "No"))
+setorder(ioi_all, "Technique", "Teaching")
+
+apa_table(
+  ioi_all
+  , caption = "Caption"
+  , note = "Note"
+  , escape = TRUE
+  , digits = 3
+)
+```
+
+```{r art-teaching-desc-2, echo = FALSE}
+colnames(kor_art_all)[2] <- "Technique"
+kor_art_all[Condition == "teaching"]$Condition <- "Teaching"
+kor_art_all[Condition == "performing"]$Condition <- "Performing"
+kor_art_all[Technique == "articulation"]$Technique <- "Articulation"
+kor_art_all$Teaching <- factor(kor_art_all$Teaching, levels = c("Yes", "No"))
+setorder(kor_art_all, "Technique", "Teaching")
+
+apa_table(
+  kor_art_all
+  , caption = "Caption"
+  , note = "Note"
+  , escape = TRUE
+  , digits = 3
+)
+```
+
+```{r dyn-teaching-desc-2}
+colnames(vel_dyn_all)[2] <- "Technique"
+vel_dyn_all[Condition == "teaching"]$Condition <- "Teaching"
+vel_dyn_all[Condition == "performing"]$Condition <- "Performing"
+vel_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"
+vel_dyn_all$Teaching <- factor(vel_dyn_all$Teaching, levels = c("Yes", "No"))
+setorder(vel_dyn_all, "Technique", "Teaching")
+
+apa_table(
+  vel_dyn_all
+  , caption = "Caption"
+  , note = "Note"
+  , escape = TRUE
+)
+```
+
+```{r dyn-diff-teaching-desc-2}
+colnames(vel_diff_dyn_all)[2] <- "Technique"
+vel_diff_dyn_all[Condition == "teaching"]$Condition <- "Teaching"
+vel_diff_dyn_all[Condition == "performing"]$Condition <- "Performing"
+vel_diff_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"
+vel_diff_dyn_all$Teaching <- factor(vel_diff_dyn_all$Teaching, levels = c("Yes", "No"))
+setorder(vel_diff_dyn_all, "Technique", "Teaching")
+
+apa_table(
+  vel_diff_dyn_all
+  , caption = "Caption"
+  , note = "Note"
+  , escape = TRUE
+)
+```

### 2022-08-31 paper/manuscript.tex
upload analysis for teaching experience
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 {}{\PatchFailed}
-\keywords{teaching, expertise, skill transmission, artistic expression, music}
+\keywords{teaching, skill transmission, sensorimotor communication, artistic expression, music}
 \usepackage{csquotes}
@@
   pdflang={en-EN},
-  pdfkeywords={teaching, expertise, skill transmission, artistic expression, music},
+  pdfkeywords={teaching, skill transmission, sensorimotor communication, artistic expression, music},
   hidelinks,
@@
 Experts modulate their performance of actions for teaching purposes, performing slower and exaggerated movements when demonstrating novel actions to novices.
-The present study asked whether such modulations are also used to teach techniques of artistic expression.
-One domain where subtle movement modulations are crucial for achieving artistic expression is music.
-We investigated how skilled pianists modulate their playing to demonstrate to students the techniques required for conveying piece-related musical expressions, compared to performing the piece without didactic intentions.
+The present study asked whether such modulations also occur during teaching performance of a music instrument, where subtle movement modulations are crucial for achieving artistic expression.
+While exaggerating performances of goal-directed actions outside of artistic contexts may be straightforward, it is an open question whether and how exaggeration for the purpose of teaching operates for actions that are expressive even when performed outside of a teaching context.
+Pianists were asked to demonstrate to students the techniques required for implementing notated expressions, compared to performing the piece without didactic intentions.
 Expressions in the piece concerned either articulation (i.e., legato and staccato) or dynamics (i.e., forte and piano).
@@
 When intending to teach articulation, skilled pianists produced more exaggerated staccato.
-When intending to teach dynamics, they created a larger contrast between forte and piano at structurally important points.
-We found consistent results with a simple musical scale (Experiment 1) and a more naturalistic piece of music (Experiment 2).
-These findings suggest that action modulations are performed not only to attract learners' attention but to highlight the most relevant aspects of the actions to be learnt.
+When intending to teach dynamics, they created a larger contrast between forte and piano.
+We found consistent results across a simple musical scale (Experiment 1) and a more naturalistic piece of music (Experiment 2).
+These findings show that teaching-specific action modulations generalize to expressive actions and suggest that action modulations serve to highlight the most relevant aspects of the actions to be learnt.
 }
@@
 
-Importantly, in some domains teaching requires not only demonstrating \emph{what} to do, but \emph{how} to perform actions.
-In artistic contexts, novices need to learn how exactly to perform actions by relying on specific techniques\textsuperscript{12}.
-For example, when playing a piece of music on the piano, it is not sufficient to press the keys in the correct order, but learners need to be able to implement expressive techniques\textsuperscript{13}.
-This raises the question of whether and how experts modulate their actions during didactic demonstration in artistic contexts, where expressivity is an integral part of what learners need to acquire.
-Studying action modulations that serve the teaching of artistic techniques is also of broader theoretical importance: It can show whether action modulations can be used to selectively highlight subtle aspects of actions that define expert performance.
-
-In the present study, we employed musical expressive techniques as skills to be taught because multiple properties of the performance (e.g., timing, smoothness, loudness) can be modulated at the same time.
-Specific and subtle aspects of the performance may need to be highlighted to teach a respective technique.
+Performing actions with exaggeration is straightforward for actions that are normally performed in the most efficient way possible\textsuperscript{12}.
+However, how can particular aspects of actions be highlighted when the actions themselves are expressive even outside of a teaching context?
+This is the case in music performance, where pieces are played with expression.
+Expressivity is a vital component of performance and typically the main focus of music teaching\textsuperscript{13}.
+Expressive skills are generally considered to be separate from technical skills, however, fine motor control is required to implement subtle sound modulations in expressive performance\textsuperscript{14}.
+In some music genres such as Western classical music, it is crucial to acquire the motor skills needed to perform a piece expressively.
+For example, pianists translate their own interpretations of music and convey their emotions by modulating specific parameters such as timing, smoothness and loudness of sound.
+This raises the question of whether and how musicians modulate their actions during didactic demonstration of expressive techniques.
+
+In naturalistic teaching settings, teachers have many possibilities for how to convey to a learner how to play a piece expressively.
+They may use verbal communication\textsuperscript{15,16} and non-verbal bodily cues\textsuperscript{17} as well as modulating the sounds that they are producing.
+In the present study, we focused entirely on the teaching of expressive techniques through sound in order to determine whether expert musicians would systematically exaggerate the actions necessary to implement a particular technique for musical expression.
 Pianists were asked to perform a piece to demonstrate to a learner how to implement the notated expressions (teaching condition) and to play the same piece for an audience (performing condition).
@@
 
-If experts rely on generic action modulations, it can be expected that they will play more slowly during didactic demonstration, regardless of the kind of expressive techniques to be taught.
-To the extent that experts use action modulations to support their teaching of specific expressive techniques, they should exaggerate legato (i.e., key-overlap time should be more positive) and staccato (i.e., key-overlap time should be more negative) when teaching articulation whereas they should exaggerate forte (i.e., key velocity should be higher) and piano (i.e., key velocity should be lower) when teaching dynamics.
+If experts rely on generic action modulations, it can be expected that they will play more slowly during didactic demonstration, regardless of the kind of techniques to be taught.
+To the extent that experts use action modulations to support their teaching of specific techniques, they should exaggerate legato (i.e., key-overlap time should be more positive) and staccato (i.e., key-overlap time should be more negative) when teaching articulation whereas they should exaggerate forte (i.e., key velocity should be higher) and piano (i.e., key velocity should be lower) when teaching dynamics.
 Furthermore, one could speculate that they might produce modulations specifically at structurally important points that best highlight the technique to be taught.
-Importantly, they should avoid modulating irrelevant properties of expression, e.g., the smoothness of sound while teaching dynamics.
+Importantly, they should avoid modulating irrelevant properties of expressio (e.g., the smoothness of sound while teaching dynamics).
 
@@
 In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}}).
-Participants played the piece 8 times per expressive technique per condition, so there were 32 trials in total (2 conditions x 2 expressive techniques x 8 trials).
+Participants played the piece 8 times per technique per condition, so there were 32 trials in total (2 conditions x 2 techniques x 8 trials).
 The order of the conditions was blocked and counterbalanced across participants.
-The order of the expressive techniques within each condition was also blocked and counterbalanced across participants.
+The order of the techniques within each condition was also blocked and counterbalanced across participants.
 A leading metronome (80 quarter beats per minute, 8 beats) indicated the target tempo before each trial.
@@
 Inter-onset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
-Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences\textsuperscript{14}.
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences\textsuperscript{18}.
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing tone.
@@
 For offsets, 14.08\% of the trials contained at least one pitch error (extra notes: 6.28\%, missing notes: 5.17\%, substituted notes: 2.63\%).
+We found that some participants did not precisely follow the sheet music (e.g., they held some notes longer than notated), therefore the order of offsets did not correspond to that of onsets.
+We counted these as errors and removed the erroneous notes even if the order of onsets was correct.
 As a result, less than 1 \% of total responses were corrected.
@@
 
-We performed separate analyses for the two expressive techniques (i.e., articulation and dynamics).
+We performed separate analyses for the two techniques (i.e., articulation and dynamics).
 A paired-sample \emph{t}-test or a Wilcoxon Signed-rank test (a non-parametric alternative to a paired \emph{t}-test) was performed to compare the mean IOIs in the teaching and performing condition.
@@
 We employed the same procedure as in Experiment 1 with several modifications.
-First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Fig \ref{fig:stimuli-2} A}) prior to the experiment.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}}).
-Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute).
+First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Fig \ref{fig:stimuli-2} A}) prior to the experiment so that they had enough time to familiarise themselves with the piece and perform it without pitch errors while performing it with different notated expressions in the lab.
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}}).
+Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability\textsuperscript{19}.
 The chosen tempo was again cued by a leading metronome.
@@
 Another possibility is that expression is best taught when leaving the tempo unchanged when it is irrelevant.
-This would imply that general didactic modulations like slowing down might be less useful in the context of teaching expressive techniques where timing itself can be used to add an expression to a performance.
-
-Musicians and music teachers tend to consider expressive skills as a performer's most important skill\textsuperscript{15}.
+This would imply that general didactic modulations like slowing down might be less useful in the context of teaching techniques where timing itself can be used to add an expression to a performance.
+
+One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students.
+It can be expected that teachers would also make didactic modulations in some domains other than acoustic properties, such as gestures if they are physically present in front of students.
+Moreover, given that teachers modulate their demonstration throughout ongoing interactions with learners\textsuperscript{20}, future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities.
+Also, some of the participants in the current study did not have teaching experience at all and even those who had teaching experience had taught for only a few years.
+It would be important to investigate if our findings would be observed in participants with extensive experience teaching the piano.
+
+Another limitation of the study is that it is not certain how the results might generalise to performances of more complex pieces and pieces with fewer notations.
+In more complex pieces of music, multiple expressive notations could be assigned to one note or phrase.
+Also, it is rare that almost all the notes are assigned to one expressive notation as in our stimuli.
+Future studies are needed to determine whether pianists would exaggerate specific parts of a piece or prioritise one aspect over the others given multiple concurrent expressive notations and more possibilities for giving the music different interpretations.
+
+Musicians and music teachers tend to consider expressive skills as performers' most important skills\textsuperscript{21}.
 In the present study, experts focused on the teaching of expressions that were notated.
-However, expression also has other facets\textsuperscript{13} that are not only piece-related but performer- or context-related.
-An interesting topic for future research is how experts teach not only expressive techniques as such but convey possibilities of interpretation.
-
-One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students.
-Given that teachers modulate their demonstration throughout online interactions with learners\textsuperscript{16}, future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities.
+However, expression also has other facets\textsuperscript{22} that are not only piece-related (e.g., notated expressions) but performer- or context-related (e.g., performer's interpretation, listening contexts such as recordings or concerts).
+The ultimate goal of expressive performance is considered to lie in developing performers' own interpretations of music and convey affects and emotions by using expressive tools such as articulation, dynamics, tempo and timing\textsuperscript{23}.
+The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings\textsuperscript{e.g., 15}.
+It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.
 
 The present findings extend earlier research on teaching-related action modulations.
-First, our study sheds light on the teaching of expressive techniques that are an integral part of skill acquisition in artistic domains.
+First, our study sheds light on the teaching of technical skills that are an integral part of skill acquisition in artistic domains.
 We showed that compared to an expressive performance baseline, experts made specific modulations to teach particular techniques.
 Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention.
-Rather, expert demonstrators seem to follow principles of relevance in communication\textsuperscript{17}, highlighting only those aspects they are intending to demonstrate.
+Rather, expert demonstrators seem to follow principles of relevance in communication\textsuperscript{24}, highlighting only those aspects they are intending to demonstrate.
 How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher's intentions explicitly adds to the learning success are important questions for future research.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-pezzulo_2013}{}}%
+\CSLLeftMargin{12. }%
+\CSLRightInline{Pezzulo, G., Donnarumma, F.
+\& Dindo, H.
+\href{https://doi.org/10.1371/journal.pone.0079876}{Human {Sensorimotor Communication}: {A Theory} of {Signaling} in {Online Social Interactions}}.
+\emph{PLOS ONE} \textbf{8}, e79876 (2013).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-laukka_2004}{}}%
+\CSLLeftMargin{13. }%
+\CSLRightInline{Laukka, P.
+\href{https://doi.org/10.1080/1461380032000182821}{Instrumental music teachers' views on expressivity: A report from music conservatoires}.
+\emph{Music Education Research} \textbf{6}, 45--56 (2004).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-sloboda_2000}{}}%
-\CSLLeftMargin{12. }%
+\CSLLeftMargin{14. }%
 \CSLRightInline{Sloboda, J. A. \href{https://doi.org/10.1016/S1364-6613(00)01531-X}{Individual differences in music performance}. \emph{Trends in Cognitive Sciences} \textbf{4}, 397--403 (2000).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-juslin_2003}{}}%
-\CSLLeftMargin{13. }%
-\CSLRightInline{Juslin, P.
-N.
-\href{https://doi.org/10.1177/03057356030313003}{Five {Facets} of {Musical Expression}: {A Psychologist}'s {Perspective} on {Music Performance}}.
-\emph{Psychology of Music} \textbf{31}, 273--302 (2003).}
+\leavevmode\vadjust pre{\hypertarget{ref-woody_2000}{}}%
+\CSLLeftMargin{15. }%
+\CSLRightInline{Woody, R.
+H.
+\href{https://doi.org/10.1177/1321103X0001400102}{Learning {Expressivity} in {Music Performance}: {An Exploratory Study}}.
+\emph{Research Studies in Music Education} \textbf{14}, 14--23 (2000).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-woody_2006}{}}%
+\CSLLeftMargin{16. }%
+\CSLRightInline{Woody, R.
+H.
+\href{https://doi.org/10.1177/002242940605400103}{The {Effect} of {Various Instructional Conditions} on {Expressive Music Performance}}.
+\emph{Journal of Research in Music Education} \textbf{54}, 21--36 (2006).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-simones_2015}{}}%
+\CSLLeftMargin{17. }%
+\CSLRightInline{Simones, L., Schroeder, F.
+\& Rodger, M.
+\href{https://doi.org/10.1177/0305735613498918}{Categorizations of physical gesture in piano teaching: {A} preliminary enquiry}.
+\emph{Psychology of Music} \textbf{43}, 103--121 (2015).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-bresin_2000}{}}%
-\CSLLeftMargin{14. }%
+\CSLLeftMargin{18. }%
 \CSLRightInline{Bresin, R.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-zamm_2018}{}}%
+\CSLLeftMargin{19. }%
+\CSLRightInline{Zamm, A., Wang, Y.
+\& Palmer, C.
+\href{https://doi.org/10.1177/0748730418783651}{Musicians' {Natural Frequencies} of {Performance Display Optimal Temporal Stability}}.
+\emph{J Biol Rhythms} \textbf{33}, 432--440 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-okazaki_2019}{}}%
+\CSLLeftMargin{20. }%
+\CSLRightInline{Okazaki, S., Muraoka, Y.
+\& Osu, R.
+\href{https://doi.org/10.1038/s41598-019-44049-x}{Teacher-learner interaction quantifies scaffolding behaviour in imitation learning}.
+\emph{Sci Rep} \textbf{9}, 1--13 (2019).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-lindstrom_2003}{}}%
-\CSLLeftMargin{15. }%
+\CSLLeftMargin{21. }%
 \CSLRightInline{Lindström, E., Juslin, P.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-okazaki_2019}{}}%
-\CSLLeftMargin{16. }%
-\CSLRightInline{Okazaki, S., Muraoka, Y.
-\& Osu, R.
-\href{https://doi.org/10.1038/s41598-019-44049-x}{Teacher-learner interaction quantifies scaffolding behaviour in imitation learning}.
-\emph{Sci Rep} \textbf{9}, 1--13 (2019).}
+\leavevmode\vadjust pre{\hypertarget{ref-juslin_2003}{}}%
+\CSLLeftMargin{22. }%
+\CSLRightInline{Juslin, P.
+N.
+\href{https://doi.org/10.1177/03057356030313003}{Five {Facets} of {Musical Expression}: {A Psychologist}'s {Perspective} on {Music Performance}}.
+\emph{Psychology of Music} \textbf{31}, 273--302 (2003).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-meissner_2021}{}}%
+\CSLLeftMargin{23. }%
+\CSLRightInline{Meissner, H. Theoretical {Framework} for {Facilitating Young Musicians}' {Learning} of {Expressive Performance}. \emph{Frontiers in Psychology} \textbf{11}, (2021).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-sperber_1995}{}}%
-\CSLLeftMargin{17. }%
+\CSLLeftMargin{24. }%
 \CSLRightInline{Sperber, D. \& Wilson, D. \emph{Relevance: {Communication} and cognition, 2nd ed}. ({Blackwell Publishing}, 1995).}
@@
 \hypertarget{instructions}{%
-\section{Instructions}\label{instructions}}
+\section{1. Instructions}\label{instructions}}
 
@@
 
-\hypertarget{tempo-variability}{%
-\section{Tempo Variability}\label{tempo-variability}}
+\hypertarget{temporal-variability}{%
+\section{2. Temporal Variability}\label{temporal-variability}}
 
@@
 \hypertarget{temporal-grouping}{%
-\section{Temporal grouping}\label{temporal-grouping}}
-
-In order to investigated if participants used temporal grouping to make a boundary between each subcomponent (e.g., between legato and staccato, between forte and piano) clearer for teaching purposes, we looked at IOIs only at transition points.
+\section{3. Temporal grouping}\label{temporal-grouping}}
+
+In order to investigate if participants used temporal grouping to make boundaries between subcomponents (i.e., between legato and staccato, between forte and piano) clearer for teaching purposes, we looked at IOIs only at transition points.
 
@@
 To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wilcoxon Signed-rank test revealed that IOIs at transition points were larger for the teaching condition {[}\emph{Mdn} = 195.75 (ms), \emph{IQR} = 25.75{]} than in the performing condition {[}\emph{Mdn} = 188.77 (ms), \emph{IQR} = 23.53{]} while playing the piece with the notated articulation (\emph{p} = 0.02, \emph{r} = 0.42, two-tailed, \emph{Fig \ref{fig:tra}}).
+The Wilcoxon Signed-rank test revealed that IOIs at transition points were larger in the teaching condition {[}\emph{Mdn} = 195.75 (ms), \emph{IQR} = 25.75{]} than in the performing condition {[}\emph{Mdn} = 188.77 (ms), \emph{IQR} = 23.53{]} while playing the piece with the notated articulation (\emph{p} = 0.02, \emph{r} = 0.42, two-tailed, \emph{Fig \ref{fig:tra}}).
 
@@
 
-A paired-sample \emph{t}-test showed that IOIs at transition points were larger for the teaching condition {[}\emph{M} = 200.90 (ms), \emph{SD} = 20.96{]} than the performing condition {[}\emph{M} = 197.63 (ms), \emph{SD} = 19.30{]} while playing the piece with the notated dynamics (\emph{t}(30) = 2.10, \emph{p} = 0.04, Cohen's \emph{d} = 0.38, two-tailed, \emph{Fig \ref{fig:tra}}).
+A paired-sample \emph{t}-test showed that IOIs at transition points were larger in the teaching condition {[}\emph{M} = 200.90 (ms), \emph{SD} = 20.96{]} than in the performing condition {[}\emph{M} = 197.63 (ms), \emph{SD} = 19.30{]} while playing the piece with the notated dynamics (\emph{t}(30) = 2.10, \emph{p} = 0.04, Cohen's \emph{d} = 0.38, two-tailed, \emph{Fig \ref{fig:tra}}).
 
@@
 
-A paired-sample \emph{t}-test showed that IOIs at transition points were larger for the teaching condition {[}\emph{M} = 1.00, \emph{SD} = 0.067{]} than the performing condition {[}\emph{M} = 0.98, \emph{SD} = 0.049{]} while playing the piece with the notated articulation (\emph{t}(19) = 2.27, \emph{p} = 0.03, Cohen's \emph{d} = 0.51, two-tailed, \emph{Fig \ref{fig:tra}}).
+A paired-sample \emph{t}-test showed that IOIs at transition points were larger in the teaching condition {[}\emph{M} = 1.00, \emph{SD} = 0.067{]} than in the performing condition {[}\emph{M} = 0.98, \emph{SD} = 0.049{]} while playing the piece with the notated articulation (\emph{t}(19) = 2.27, \emph{p} = 0.03, Cohen's \emph{d} = 0.51, two-tailed, \emph{Fig \ref{fig:tra}}).
 
@@
 \hypertarget{additional-analysis-for-dynamics-difference}{%
-\section{Additional analysis for dynamics difference}\label{additional-analysis-for-dynamics-difference}}
+\section{4. Additional analysis for dynamics difference}\label{additional-analysis-for-dynamics-difference}}
 
@@
 
+\clearpage
+
+\hypertarget{descriptive-statistics-depending-on-participants-teaching-experience}{%
+\section{5. Descriptive statistics depending on participants' teaching experience}\label{descriptive-statistics-depending-on-participants-teaching-experience}}
+
+\hypertarget{experiment-1-5}{%
+\subsection{Experiment 1}\label{experiment-1-5}}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-teaching-1-1} \caption{\label{fig:ioi-teaching-1}BAA}\label{fig:plot-ioi-teaching-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-1-1} \caption{\label{fig:art-teaching-1}BAA}\label{fig:plot-art-teaching-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-teaching-1-1} \caption{\label{fig:dyn-teaching-1}BAA}\label{fig:plot-dyn-teaching-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-diff-teaching-1-1} \caption{\label{fig:dyn-diff-teaching-1}BAA}\label{fig:plot-dyn-diff-teaching-1}
+\end{figure}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:ioi-teaching-desc-1}Caption}
+
+\begin{tabular}{lllllllll}
+\toprule
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{IQR}\\
+\midrule
+Performing & Articulation & Yes & 10 & 181.27 & 10.99 & 3.47 & 177.37 & 14.91\\
+Teaching & Articulation & Yes & 10 & 184.72 & 13.02 & 4.12 & 179.21 & 16.76\\
+Performing & Articulation & No & 21 & 193.49 & 18.75 & 4.09 & 185.83 & 26.43\\
+Teaching & Articulation & No & 21 & 196.67 & 20.52 & 4.48 & 192.13 & 32.33\\
+Performing & Dynamics & Yes & 10 & 182.43 & 11.52 & 3.64 & 181.67 & 13.07\\
+Teaching & Dynamics & Yes & 10 & 187.15 & 11.73 & 3.71 & 187.91 & 15.67\\
+Performing & Dynamics & No & 21 & 191.48 & 18.56 & 4.05 & 188.17 & 23.11\\
+Teaching & Dynamics & No & 21 & 193.50 & 21.22 & 4.63 & 186.28 & 25.29\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Note}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:art-teaching-desc-1}Caption}
+
+\begin{tabular}{llllllll}
+\toprule
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+\midrule
+Performing & Articulation & Legato & Yes & 10 & 21.57 & 16.18 & 5.12\\
+Performing & Articulation & Staccato & Yes & 10 & -119.66 & 22.84 & 7.22\\
+Teaching & Articulation & Legato & Yes & 10 & 25.43 & 13.11 & 4.15\\
+Teaching & Articulation & Staccato & Yes & 10 & -127.39 & 15.95 & 5.04\\
+Performing & Articulation & Legato & No & 21 & 10.40 & 16.57 & 3.62\\
+Performing & Articulation & Staccato & No & 21 & -132.48 & 19.03 & 4.15\\
+Teaching & Articulation & Legato & No & 21 & 11.36 & 14.71 & 3.21\\
+Teaching & Articulation & Staccato & No & 21 & -136.84 & 21.52 & 4.70\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Note}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:dyn-teaching-desc-1}Caption}
+
+\begin{tabular}{llllllll}
+\toprule
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+\midrule
+Performing & Dynamics & Forte & Yes & 10 & 85.20 & 7.76 & 2.45\\
+Performing & Dynamics & Piano & Yes & 10 & 60.27 & 4.57 & 1.44\\
+Teaching & Dynamics & Forte & Yes & 10 & 86.96 & 8.76 & 2.77\\
+Teaching & Dynamics & Piano & Yes & 10 & 60.11 & 5.93 & 1.87\\
+Performing & Dynamics & Forte & No & 21 & 79.97 & 6.82 & 1.49\\
+Performing & Dynamics & Piano & No & 21 & 61.33 & 4.40 & 0.96\\
+Teaching & Dynamics & Forte & No & 21 & 80.80 & 7.49 & 1.63\\
+Teaching & Dynamics & Piano & No & 21 & 60.88 & 4.76 & 1.04\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Note}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:dyn-diff-teaching-desc-1}Caption}
+
+\begin{tabular}{llllllll}
+\toprule
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+\midrule
+Performing & Dynamics & FtoP & Yes & 10 & -14.63 & 5.79 & 1.83\\
+Performing & Dynamics & PtoF & Yes & 10 & 24.32 & 9.96 & 3.15\\
+Teaching & Dynamics & FtoP & Yes & 10 & -18.81 & 7.26 & 2.30\\
+Teaching & Dynamics & PtoF & Yes & 10 & 29.07 & 8.78 & 2.78\\
+Performing & Dynamics & FtoP & No & 21 & -10.30 & 5.69 & 1.24\\
+Performing & Dynamics & PtoF & No & 21 & 19.23 & 7.71 & 1.68\\
+Teaching & Dynamics & FtoP & No & 21 & -11.87 & 6.09 & 1.33\\
+Teaching & Dynamics & PtoF & No & 21 & 21.98 & 9.56 & 2.09\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Note}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\newpage
+
+\hypertarget{experiment-2-5}{%
+\subsection{Experiment 2}\label{experiment-2-5}}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-teaching-2-1} \caption{\label{fig:ioi-teaching-2}BAA}\label{fig:plot-ioi-teaching-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-2-1} \caption{\label{fig:art-teaching-2}BAA}\label{fig:plot-art-teaching-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-teaching-2-1} \caption{\label{fig:dyn-teaching-2}BAA}\label{fig:plot-dyn-teaching-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-diff-teaching-2-1} \caption{\label{fig:dyn-diff-teaching-2}BAA}\label{fig:plot-dyn-diff-teaching-2}
+\end{figure}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:ioi-teaching-desc-2}Caption}
+
+\begin{tabular}{lllllllll}
+\toprule
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{IQR}\\
+\midrule
+Performing & Articulation & Yes & 11 & 0.954 & 0.046 & 0.014 & 0.977 & 0.064\\
+Teaching & Articulation & Yes & 11 & 0.964 & 0.049 & 0.015 & 0.976 & 0.044\\
+Performing & Articulation & No & 9 & 0.951 & 0.035 & 0.012 & 0.950 & 0.041\\
+Teaching & Articulation & No & 9 & 0.967 & 0.052 & 0.017 & 0.973 & 0.014\\
+Performing & Dynamics & Yes & 11 & 0.953 & 0.046 & 0.014 & 0.960 & 0.050\\
+Teaching & Dynamics & Yes & 11 & 0.958 & 0.047 & 0.014 & 0.972 & 0.044\\
+Performing & Dynamics & No & 9 & 0.958 & 0.041 & 0.014 & 0.959 & 0.023\\
+Teaching & Dynamics & No & 9 & 0.955 & 0.052 & 0.017 & 0.954 & 0.026\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Note}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:art-teaching-desc-2}Caption}
+
+\begin{tabular}{llllllll}
+\toprule
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+\midrule
+Performing & Articulation & Legato & Yes & 11 & 0.115 & 0.074 & 0.022\\
+Performing & Articulation & Staccato & Yes & 11 & -0.755 & 0.034 & 0.010\\
+Teaching & Articulation & Legato & Yes & 11 & 0.122 & 0.064 & 0.019\\
+Teaching & Articulation & Staccato & Yes & 11 & -0.776 & 0.034 & 0.010\\
+Performing & Articulation & Legato & No & 9 & 0.120 & 0.071 & 0.024\\
+Performing & Articulation & Staccato & No & 9 & -0.728 & 0.050 & 0.017\\
+Teaching & Articulation & Legato & No & 9 & 0.120 & 0.081 & 0.027\\
+Teaching & Articulation & Staccato & No & 9 & -0.749 & 0.073 & 0.024\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Note}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:dyn-teaching-desc-2}Caption}
+
+\begin{tabular}{llllllll}
+\toprule
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+\midrule
+Performing & Dynamics & Forte & Yes & 11 & 81.06 & 6.52 & 1.97\\
+Performing & Dynamics & Piano & Yes & 11 & 62.49 & 5.55 & 1.67\\
+Teaching & Dynamics & Forte & Yes & 11 & 83.32 & 6.48 & 1.95\\
+Teaching & Dynamics & Piano & Yes & 11 & 59.50 & 3.59 & 1.08\\
+Performing & Dynamics & Forte & No & 9 & 82.02 & 10.15 & 3.38\\
+Performing & Dynamics & Piano & No & 9 & 62.36 & 4.11 & 1.37\\
+Teaching & Dynamics & Forte & No & 9 & 83.06 & 11.54 & 3.85\\
+Teaching & Dynamics & Piano & No & 9 & 61.81 & 3.76 & 1.25\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Note}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:dyn-diff-teaching-desc-2}Caption}
+
+\begin{tabular}{llllllll}
+\toprule
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+\midrule
+Performing & Dynamics & FtoP & Yes & 11 & -14.43 & 6.20 & 1.87\\
+Performing & Dynamics & PtoF & Yes & 11 & 14.96 & 6.78 & 2.04\\
+Teaching & Dynamics & FtoP & Yes & 11 & -19.27 & 4.33 & 1.30\\
+Teaching & Dynamics & PtoF & Yes & 11 & 20.77 & 5.71 & 1.72\\
+Performing & Dynamics & FtoP & No & 9 & -13.64 & 9.46 & 3.15\\
+Performing & Dynamics & PtoF & No & 9 & 17.44 & 8.11 & 2.70\\
+Teaching & Dynamics & FtoP & No & 9 & -14.84 & 11.08 & 3.69\\
+Teaching & Dynamics & PtoF & No & 9 & 20.66 & 12.09 & 4.03\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Note}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
 \end{document}

### 2022-09-01 paper/manuscript.Rmd
add captions for figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r art-teaching-1, include = FALSE}
+# articulation
 kot_art_plot$teachingexp <- factor(kot_art_plot$teachingexp, levels = c("Yes", "No"))
@@
 ```{r dyn-teaching-1, include = FALSE}
+# dynamics
 vel_dyn_plot$teachingexp <- factor(vel_dyn_plot$teachingexp, levels = c("Yes", "No"))
@@
 ```{r dyn-diff-teaching-1, include = FALSE}
+# dynamics difference
 vel_diff_dyn_plot$teachingexp <- factor(vel_diff_dyn_plot$teachingexp, levels = c("Yes", "No"))
@@
 
-```{r plot-ioi-teaching-1, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-1}BAA"}
+```{r plot-ioi-teaching-1, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of IOIs (ms) when playing the piece with either articulation (left) or dynamics (right).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
 img_ioi_teaching_1 <- image_append(c(img_ioi_art_teaching_1, img_ioi_dyn_teaching_1))
@@
 
-```{r plot-art-teaching-1, out.width = "100%", fig.cap = "\\label{fig:art-teaching-1}BAA"}
+```{r plot-art-teaching-1, out.width = "100%", fig.cap = "\\label{fig:art-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
 img_art_teaching_1 <- image_append(c(img_art_leg_teaching_1, img_art_sta_teaching_1))
@@
 
-```{r plot-dyn-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-1}BAA"}
+```{r plot-dyn-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
 img_dyn_teaching_1 <- image_append(c(img_vel_for_teaching_1, img_vel_pia_teaching_1))
@@
 
-```{r plot-dyn-diff-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-1}BAA"}
+```{r plot-dyn-diff-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
 img_dyn_diff_teaching_1 <- image_append(c(img_vel_diff_for_teaching_1, img_vel_diff_pia_teaching_1))
@@
   ioi_all
-  , caption = "Caption"
+  , caption = "Experiment 1:"
   , note = "Note"
@@
   kot_art_all
-  , caption = "Caption"
+  , caption = "Experiment 1:"
   , note = "Note"
@@
   vel_dyn_all
-  , caption = "Caption"
+  , caption = "Experiment 1:"
   , note = "Note"
@@
   vel_diff_dyn_all
-  , caption = "Caption"
+  , caption = "Experiment 1:"
   , note = "Note"
@@
 
-```{r plot-ioi-teaching-2, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-2}BAA"}
+```{r plot-ioi-teaching-2, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of IOIs (ms) when playing the piece with either articulation (left) or dynamics (right).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
 img_ioi_teaching_2 <- image_append(c(img_ioi_art_teaching_2, img_ioi_dyn_teaching_2))
@@
 
-```{r plot-art-teaching-2, out.width = "100%", fig.cap = "\\label{fig:art-teaching-2}BAA"}
+```{r plot-art-teaching-2, out.width = "100%", fig.cap = "\\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
 img_art_teaching_2 <- image_append(c(img_art_leg_teaching_2, img_art_sta_teaching_2))
@@
 
-```{r plot-dyn-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-2}BAA"}
+```{r plot-dyn-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
 img_dyn_teaching_2 <- image_append(c(img_vel_for_teaching_2, img_vel_pia_teaching_2))
@@
 
-```{r plot-dyn-diff-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-2}BAA"}
+```{r plot-dyn-diff-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
 img_dyn_diff_teaching_2 <- image_append(c(img_vel_diff_for_teaching_2, img_vel_diff_pia_teaching_2))
@@
   ioi_all
-  , caption = "Caption"
+  , caption = "Experiment 2:"
   , note = "Note"
@@
   kor_art_all
-  , caption = "Caption"
+  , caption = "Experiment 2:"
   , note = "Note"
@@
   vel_dyn_all
-  , caption = "Caption"
+  , caption = "Experiment 2:"
   , note = "Note"
@@
   vel_diff_dyn_all
-  , caption = "Caption"
+  , caption = "Experiment 2:"
   , note = "Note"

### 2022-09-01 paper/manuscript.tex
add captions for figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-teaching-1-1} \caption{\label{fig:ioi-teaching-1}BAA}\label{fig:plot-ioi-teaching-1}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-1-1} \caption{\label{fig:art-teaching-1}BAA}\label{fig:plot-art-teaching-1}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-teaching-1-1} \caption{\label{fig:dyn-teaching-1}BAA}\label{fig:plot-dyn-teaching-1}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-diff-teaching-1-1} \caption{\label{fig:dyn-diff-teaching-1}BAA}\label{fig:plot-dyn-diff-teaching-1}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-teaching-1-1} \caption{\label{fig:ioi-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of IOIs (ms) when playing the piece with either articulation (left) or dynamics (right).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-teaching-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-1-1} \caption{\label{fig:art-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-art-teaching-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-teaching-1-1} \caption{\label{fig:dyn-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-dyn-teaching-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-diff-teaching-1-1} \caption{\label{fig:dyn-diff-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-dyn-diff-teaching-1}
 \end{figure}
@@
 
-\caption{\label{tab:ioi-teaching-desc-1}Caption}
+\caption{\label{tab:ioi-teaching-desc-1}Experiment 1:}
 
@@
 
-\caption{\label{tab:art-teaching-desc-1}Caption}
+\caption{\label{tab:art-teaching-desc-1}Experiment 1:}
 
@@
 
-\caption{\label{tab:dyn-teaching-desc-1}Caption}
+\caption{\label{tab:dyn-teaching-desc-1}Experiment 1:}
 
@@
 
-\caption{\label{tab:dyn-diff-teaching-desc-1}Caption}
+\caption{\label{tab:dyn-diff-teaching-desc-1}Experiment 1:}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-teaching-2-1} \caption{\label{fig:ioi-teaching-2}BAA}\label{fig:plot-ioi-teaching-2}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-2-1} \caption{\label{fig:art-teaching-2}BAA}\label{fig:plot-art-teaching-2}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-teaching-2-1} \caption{\label{fig:dyn-teaching-2}BAA}\label{fig:plot-dyn-teaching-2}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-diff-teaching-2-1} \caption{\label{fig:dyn-diff-teaching-2}BAA}\label{fig:plot-dyn-diff-teaching-2}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-teaching-2-1} \caption{\label{fig:ioi-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of IOIs (ms) when playing the piece with either articulation (left) or dynamics (right).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-teaching-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-2-1} \caption{\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-art-teaching-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-teaching-2-1} \caption{\label{fig:dyn-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-dyn-teaching-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-diff-teaching-2-1} \caption{\label{fig:dyn-diff-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-dyn-diff-teaching-2}
 \end{figure}
@@
 
-\caption{\label{tab:ioi-teaching-desc-2}Caption}
+\caption{\label{tab:ioi-teaching-desc-2}Experiment 2:}
 
@@
 
-\caption{\label{tab:art-teaching-desc-2}Caption}
+\caption{\label{tab:art-teaching-desc-2}Experiment 2:}
 
@@
 
-\caption{\label{tab:dyn-teaching-desc-2}Caption}
+\caption{\label{tab:dyn-teaching-desc-2}Experiment 2:}
 
@@
 
-\caption{\label{tab:dyn-diff-teaching-desc-2}Caption}
+\caption{\label{tab:dyn-diff-teaching-desc-2}Experiment 2:}
 

### 2022-09-01 paper/manuscript.Rmd
add captions for tables / correct the wrong plot for vel_dyn_1.png
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 img_vel_dyn_1 <- image_read(here("paper/image", "vel_dyn_1.png"), density = 300)
-
-# annotate * for forte
-img_vel_dyn_1 <- image_annotate(img_vel_dyn_1, "*", size = 130, color = "black", location = "+725+400")
 ```
@@
 The ultimate goal of expressive performance is considered to lie in developing performers’ own interpretations of music and convey affects and emotions by using expressive tools such as articulation, dynamics, tempo and timing [@meissner_2021].
-The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings [e.g., @woody_2000].
-It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.
+The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings [@woody_2000; @colprit_2000].
+Given that teachers demonstrate for students [@colprit_2000] and students often learn by listening to the performances of either their teachers or recordings [@volioti_2017], it is important to investigate how performance itself could affect students’ learning processes.It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.
 
@@
 apa_table(
-  ioi_all
-  , caption = "Experiment 1:"
-  , note = "Note"
+  ioi_all[, -c(8, 9)]
+  , caption = "Experiment 1: Comparison between those with and without piano teaching experience in terms of IOIs (ms)."
+  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
+Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
+Teaching indicates if participants had an experience in teaching the piano."
   , escape = TRUE
@@
   kot_art_all
-  , caption = "Experiment 1:"
-  , note = "Note"
+  , caption = "Experiment 1: Comparison between those with and without piano teaching experience in terms of KOT (ms)."
+  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
+Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
+Teaching indicates if participants had an experience in teaching the piano."
   , escape = TRUE
@@
   vel_dyn_all
-  , caption = "Experiment 1:"
-  , note = "Note"
+  , caption = "Experiment 1: Comparison between those with and without piano teaching experience in terms of KV (0-127)."
+  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
+Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
+Teaching indicates if participants had an experience in teaching the piano."
   , escape = TRUE
@@
   vel_diff_dyn_all
-  , caption = "Experiment 1:"
-  , note = "Note"
+  , caption = "Experiment 1: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points."
+  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
+Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
+Teaching indicates if participants had an experience in teaching the piano."
   , escape = TRUE
@@
 
-\newpage
+\clearpage
 
@@
 
-```{r plot-ioi-teaching-2, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of IOIs (ms) when playing the piece with either articulation (left) or dynamics (right).
+```{r plot-ioi-teaching-2, out.width = "100%", fig.cap = "\\label{fig:ioi-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
 A dashed line represents the tempo given by a metronome.
@@
 
-```{r plot-art-teaching-2, out.width = "100%", fig.cap = "\\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation.
+```{r plot-art-teaching-2, out.width = "100%", fig.cap = "\\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOR when playing the piece with articulation.
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
@@
 apa_table(
-  ioi_all
-  , caption = "Experiment 2:"
-  , note = "Note"
+  ioi_all[, -c(8, 9)]
+  , caption = "Experiment 2: Comparison between those with and without piano teaching experience in terms of normalised IOIs."
+  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
+Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
+Teaching indicates if participants had an experience in teaching the piano."
   , escape = TRUE
@@
   kor_art_all
-  , caption = "Experiment 2:"
-  , note = "Note"
+  , caption = "Experiment 2: Comparison between those with and without piano teaching experience in terms of KOR."
+  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
+Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
+Teaching indicates if participants had an experience in teaching the piano."
   , escape = TRUE
@@
   vel_dyn_all
-  , caption = "Experiment 2:"
-  , note = "Note"
+  , caption = "Experiment 2: Comparison between those with and without piano teaching experience in terms of KV (0-127)."
+  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
+Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
+Teaching indicates if participants had an experience in teaching the piano."
   , escape = TRUE
@@
   vel_diff_dyn_all
-  , caption = "Experiment 2:"
-  , note = "Note"
+  , caption = "Experiment 2: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points."
+  , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
+Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
+Teaching indicates if participants had an experience in teaching the piano."
   , escape = TRUE

### 2022-09-01 paper/manuscript.Rmd
finalise the manuscript with some minor corrections
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplementary material](#supplementary)*).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplementary material](#supplementary)* 1).
 Participants played the piece 8 times per technique per condition, so there were 32 trials in total (2 conditions x 2 techniques x 8 trials).
@@
 piano) showed that there was a significant main effect of Dynamics (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Dynamics (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-1}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`] than in the performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`) when instructed to play forte.
-There was no significant difference in KV between the teaching [*M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition [*M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] (*p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`) when instructed to play piano.
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and the performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] for each subcomponent (Forte: *p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`, Piano: *p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`).
 
@@
 First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2} A*) prior to the experiment so that they had enough time to familiarise themselves with the piece and perform it without pitch errors while performing it with different notated expressions in the lab.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in *[Supplementary material](#supplementary)*).
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in *[Supplementary material](#supplementary)* 1).
 Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability [@zamm_2018].
@@
 
-The aim of this study was to investigate how expert pianists modulate their performance to teach expressive techniques.
+The aim of this study was to investigate how expert pianists modulate their performance to teach techniques necessary to implement musical expressions.
 Overall, the findings of the two experiments showed that pianists performed one and the same piece differently depending on whether they played with the intention to teach expressive techniques or with the intention to perform the piece for an audience.
@@
 It can be expected that teachers would also make didactic modulations in some domains other than acoustic properties, such as gestures if they are physically present in front of students.
-Moreover, given that teachers modulate their demonstration throughout ongoing interactions with learners [@okazaki_2019], future studies are needed to investigate how experts dynamically adapt their performance to their students’ skill level and demonstrated abilities.
+Moreover, given that teachers modulate their demonstration throughout ongoing interactions with learners [@okazaki_2019], future studies are needed to investigate how experts dynamically adapt their performance to their students’ skill levels and demonstrated abilities.
 Also, some of the participants in the current study did not have teaching experience at all and even those who had teaching experience had taught for only a few years.
 It would be important to investigate if our findings would be observed in participants with extensive experience teaching the piano.
+We summarised descriptive statistics depending on participants' teaching experience (see *[Supplementary material](#supplementary)* 2).
 
@@
 Future studies are needed to determine whether pianists would exaggerate specific parts of a piece or prioritise one aspect over the others given multiple concurrent expressive notations and more possibilities for giving the music different interpretations.
+
+We quantified participants’ sound modulations in terms of tempo, articulation and dynamics so as to examine if and how participants modulate their sound.
+However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see *[Supplementary material](#supplementary)* 3 and 4).
+For example, participants tended to slow down at transition points regardless of the techniques they were teaching although overall, we found that participants played slower only when teaching the piece with the notated articulation.
+Future research is needed to investigate how musicians make particular modulations in relation to musical structures.
 
@@
 The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings [@woody_2000; @colprit_2000].
-Given that teachers demonstrate for students [@colprit_2000] and students often learn by listening to the performances of either their teachers or recordings [@volioti_2017], it is important to investigate how performance itself could affect students’ learning processes.It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.
+Given that students often learn by listening to recordings [@volioti_2017], it is important to investigate how performance itself could affect students’ learning processes.It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.
 
@@
 
-# Supplementary material{#supplementary}
+# Supplementary Material{#supplementary}
 
@@
 
-# 2. Temporal Variability
-
-The tempo variability of performance was assessed with the coefficient of variation (CV) of the IOIs to investigate the tempo stability of performance between the teaching and performing condition.
-CVs were calculated by dividing the standard deviation by the mean of IOIs.
-
-## Results
-## Experiment 1
-```{r variability-1, include = FALSE}
-source(here("experiment-1/analysis/stats", "variability.R"), chdir = TRUE)
-```
-
-### Articulation
-
-To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_cv_norm$p.value))}`).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$IQR)`] while playing the piece with the notated articulation (*p* `r if(ioi_art_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_cv$effsize, 2)`, two-tailed, *Fig \ref{fig:cv}*).
-
-### Dynamics
-
-To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_cv_norm$p.value))}`).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "dynamics"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "dynamics"]$IQR)`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn_cv$effsize, 2)`, two-tailed, *Fig \ref{fig:cv}*).
-
-```{r ioi-cv-1, echo = FALSE}
-ioi_cv_plot <- data.table(subject = ioi_cv_subject[Condition == "teaching"]$SubNr, teaching = ioi_cv_subject[Condition == "teaching"]$MeanCV, performing = ioi_cv_subject[Condition == "performing"]$MeanCV, skill = ioi_cv_subject[Condition == "teaching"]$Skill)
-
-ioi_cv_1 <- ggpaired(ioi_cv_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "CVs", title = "CVs: Exp 1") +
-  ylim(0, 0.20) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 0.08, method = "t.test", paired = TRUE) +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "ioi_cv_1.png"), plot = ioi_cv_1, dpi = 300, width = 7, height = 5)
-img_ioi_cv_1 <- image_read(here("paper/image", "ioi_cv_1.png"), density = 300)
-```
-
-## Experiment 2
-```{r variability-2, include = FALSE}
-source(here("experiment-2/analysis/stats", "variability.R"), chdir = TRUE)
-```
-
-### Articulation
-
-To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_cv_norm$p.value))}`).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$IQR)`] while playing the piece with the notated articulation (*p* `r if(ioi_art_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_cv$effsize, 2)`, two-tailed, *Fig \ref{fig:cv}*).
-
-### Dynamics
-
-A paired-sample *t*-test showed that the tempo variability of performance did not differ significantly between the teaching [*M* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$SD)`] and the performing condition [*M* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_cv_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_cv_ttest$statistic))`, *p* `r if(ioi_dyn_cv_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_cv_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn_cv$effsize), 3)`, two-tailed, *Fig \ref{fig:cv}*).
-
-```{r ioi-cv-2, echo = FALSE}
-ioi_cv_plot <- data.table(subject = ioi_cv_subject[Condition == "teaching"]$SubNr, teaching = ioi_cv_subject[Condition == "teaching"]$MeanCV, performing = ioi_cv_subject[Condition == "performing"]$MeanCV, skill = ioi_cv_subject[Condition == "teaching"]$Skill)
-
-ioi_cv_2 <- ggpaired(ioi_cv_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "CVs", title = "CVs: Exp 2") +
-  ylim(0, 0.20) +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 0.08, method = "t.test", paired = TRUE) +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "ioi_cv_2.png"), plot = ioi_cv_2, dpi = 300, width = 7, height = 5)
-img_ioi_cv_2 <- image_read(here("paper/image", "ioi_cv_2.png"), density = 300)
-```
-
-```{r plot-cv, out.width = "100%", fig.cap = "\\label{fig:cv}Experiment 1, 2: CVs when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
-img_ioi_cv <- image_append(c(img_ioi_cv_1, img_ioi_cv_2))
-img_ioi_cv <- image_scale(img_ioi_cv, "2200")
-image_write(img_ioi_cv, path = here("paper/figure", "Fig12.tiff"), format = "tiff")
-img_ioi_cv
-```
-
-\clearpage
-
-# 3. Temporal grouping
-
-In order to investigate if participants used temporal grouping to make boundaries between subcomponents (i.e., between legato and staccato, between forte and piano) clearer for teaching purposes, we looked at IOIs only at transition points.
-
-## Results
-## Experiment 1
-```{r variability-1-1, include = FALSE}
-source(here("experiment-1/analysis/stats", "variability.R"), chdir = TRUE)
-```
-
-### Articulation
-
-To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_tra_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_tra_norm$p.value))}`).
-The Wilcoxon Signed-rank test revealed that IOIs at transition points were larger in the teaching condition [*Mdn* = `r ioi_tra[Condition == "teaching" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_tra[Condition == "teaching" & Skill == "articulation"]$IQR`] than in the performing condition [*Mdn* = `r ioi_tra[Condition == "performing" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_tra[Condition == "performing" & Skill == "articulation"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_tra_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_tra_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_tra$effsize, 2)`, two-tailed, *Fig \ref{fig:tra}*).
-
-### Dynamics
-
-A paired-sample *t*-test showed that IOIs at transition points were larger in the teaching condition [*M* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$Mean` (ms), *SD* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$SD`] than in the performing condition [*M* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$Mean` (ms), *SD* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$SD`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_tra_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_tra_ttest$statistic))`, *p* `r if(ioi_dyn_tra_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_tra_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn_tra$effsize), 3)`, two-tailed, *Fig \ref{fig:tra}*).
-
-```{r ioi-tra-1, echo = FALSE}
-ioi_tra_plot <- data.table(subject = ioi_tra_subject[Condition == "teaching"]$SubNr, teaching = ioi_tra_subject[Condition == "teaching"]$Mean, performing = ioi_tra_subject[Condition == "performing"]$Mean, skill = ioi_tra_subject[Condition == "teaching"]$Skill)
-
-ioi_tra_1 <- ggpaired(ioi_tra_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs at transition: Exp 1") +
-  geom_hline(yintercept = 188, linetype = "dashed") +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 150, method = "t.test", paired = TRUE) +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "ioi_tra_1.png"), plot = ioi_tra_1, dpi = 300, width = 7, height = 5)
-img_ioi_tra_1 <- image_read(here("paper/image", "ioi_tra_1.png"), density = 300)
-```
-
-## Experiment 2
-### Articulation
-```{r variability-2-2, include = FALSE}
-source(here("experiment-2/analysis/stats", "variability.R"), chdir = TRUE)
-```
-
-A paired-sample *t*-test showed that IOIs at transition points were larger in the teaching condition [*M* = `r sprintf('%.2f', ioi_tra[Condition == "teaching" & Skill == "articulation"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_tra[Condition == "teaching" & Skill == "articulation"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_tra[Condition == "performing" & Skill == "articulation"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_tra[Condition == "performing" & Skill == "articulation"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_tra_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_tra_ttest$statistic))`, *p* `r if(ioi_art_tra_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_tra_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art_tra$effsize), 3)`, two-tailed, *Fig \ref{fig:tra}*).
-
-### Dynamics
-
-To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_tra_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_tra_norm$p.value))}`).
-The Wilcoxon Signed-rank test revealed that IOIs at transition points did not differ significantly between the teaching [*Mdn* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$Median`, *IQR* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$IQR`] and the performing condition [*Mdn* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$Median`, *IQR* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_tra_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_tra_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn_tra$effsize, 2)`, two-tailed, *Fig \ref{fig:tra}*).
-
-```{r ioi-tra-2, echo = FALSE}
-ioi_tra_plot <- data.table(subject = ioi_tra_subject[Condition == "teaching"]$SubNr, teaching = ioi_tra_subject[Condition == "teaching"]$Mean, performing = ioi_tra_subject[Condition == "performing"]$Mean, skill = ioi_tra_subject[Condition == "teaching"]$Skill)
-
-ioi_tra_2 <- ggpaired(ioi_tra_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOIs at transition: Exp 2") +
-  geom_hline(yintercept = 1, linetype = "dashed") +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 1.05, method = "t.test", paired = TRUE) +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "ioi_tra_2.png"), plot = ioi_tra_2, dpi = 300, width = 7, height = 5)
-img_ioi_tra_2 <- image_read(here("paper/image", "ioi_tra_2.png"), density = 300)
-```
-
-```{r plot-tra, out.width = "100%", fig.cap = "\\label{fig:tra}Experiment 1, 2: IOIs at transition points when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
-A dashed line represents the tempo given by a metronome.
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
-img_ioi_tra <- image_append(c(img_ioi_tra_1, img_ioi_tra_2))
-img_ioi_tra <- image_scale(img_ioi_tra, "2200")
-image_write(img_ioi_tra, path = here("paper/figure", "Fig10.tiff"), format = "tiff")
-img_ioi_tra
-```
-
-\clearpage
-
-# 4. Additional analysis for dynamics difference
-
-In order to examine if the dynamics difference between forte and piano at transition points is larger when the difference between forte and piano in overall performance is considered, we normalised the KV difference at transition points by dividing it by the average range between forte and piano per performance (per trial).
-
-## Results
-## Experiment 1
-```{r additional-dynamics-1, include = FALSE}
-source(here("experiment-1/analysis/stats", "dynamics.R"), chdir = TRUE)
-```
-
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Transition Type (FtoP vs.
-PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig \ref{fig:vel-diff-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (*p* `r if(vel_diff_dyn_posthoc_2$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[1]))}`) and from piano to forte (*p* `r if(vel_diff_dyn_posthoc_2$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[2]))}`) between the teaching condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] and the performing condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
-
-```{r vel-diff-dyn-normalised-1, echo = FALSE}
-vel_diff_dyn_normalised_1 <- ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas",xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff: Exp 1") + ylim(-6, 5) + stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 3, method = "t.test", paired = TRUE) + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "vel_diff_dyn_normalised_1.png"), plot = vel_diff_dyn_normalised_1, dpi = 300, width = 7, height = 5)
-img_vel_diff_dyn_normalised_1 <- image_read(here("paper/image", "vel_diff_dyn_normalised_1.png"), density = 300)
-```
-
-## Experiment 2
-```{r additional-dynamics-2, include = FALSE}
-source(here("experiment-2/analysis/stats", "dynamics.R"), chdir = TRUE)
-```
-
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
-performing) and Transition Type (FtoP vs.
-PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`).
-However, there was no significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig \ref{fig:vel-diff-2}*).
-
-```{r vel-diff-dyn-normalised-2, echo = FALSE}
-vel_diff_dyn_normalised_2 <- ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff: Exp 2") + ylim(-6, 5) + stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 3, method = "t.test", paired = TRUE) + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "vel_diff_dyn_normalised_2.png"), plot = vel_diff_dyn_normalised_2, dpi = 300, width = 7, height = 5)
-img_vel_diff_dyn_normalised_2 <- image_read(here("paper/image", "vel_diff_dyn_normalised_2.png"), density = 300)
-```
-
-```{r plot-vel-diff-2, out.width = "100%", fig.cap = "\\label{fig:vel-diff-2}Experiment 1, 2: Normalised KV Difference at transition points when playing the piece with dynamics (left: Experiment 1, right: Experiment 2).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
-img_vel_diff_dyn_normalised <- image_append(c(img_vel_diff_dyn_normalised_1, img_vel_diff_dyn_normalised_2))
-img_vel_diff_dyn_normalised <- image_scale(img_vel_diff_dyn_normalised, "2200")
-image_write(img_vel_diff_dyn_normalised, path = here("paper/figure", "Fig11.tiff"), format = "tiff")
-img_vel_diff_dyn_normalised
-```
-
-\clearpage
-
-# 5. Descriptive statistics depending on participants' teaching experience
+# 2. Descriptive statistics depending on participants' teaching experience
+
+We summarised descriptive statistics for those had an experience in teaching the piano and those who did not. Due to the lack of statistical power, we did not perform any statistical tests.
 
@@
 ```
+
+\clearpage
+
+# 3. Temporal Variability
+
+The tempo variability of performance was assessed with the coefficient of variation (CV) of the IOIs to investigate the tempo stability of performance between the teaching and performing condition.
+CVs were calculated by dividing the standard deviation by the mean of IOIs.
+
+## Results
+## Experiment 1
+```{r variability-1, include = FALSE}
+source(here("experiment-1/analysis/stats", "variability.R"), chdir = TRUE)
+```
+
+### Articulation
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_cv_norm$p.value))}`).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$IQR)`] while playing the piece with the notated articulation (*p* `r if(ioi_art_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_cv$effsize, 2)`, two-tailed, *Fig \ref{fig:cv}*).
+
+### Dynamics
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_cv_norm$p.value))}`).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "dynamics"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "dynamics"]$IQR)`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn_cv$effsize, 2)`, two-tailed, *Fig \ref{fig:cv}*).
+
+```{r ioi-cv-1, echo = FALSE}
+ioi_cv_plot <- data.table(subject = ioi_cv_subject[Condition == "teaching"]$SubNr, teaching = ioi_cv_subject[Condition == "teaching"]$MeanCV, performing = ioi_cv_subject[Condition == "performing"]$MeanCV, skill = ioi_cv_subject[Condition == "teaching"]$Skill)
+
+ioi_cv_1 <- ggpaired(ioi_cv_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "CVs", title = "CVs: Exp 1") +
+  ylim(0, 0.20) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 0.08, method = "t.test", paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_cv_1.png"), plot = ioi_cv_1, dpi = 300, width = 7, height = 5)
+img_ioi_cv_1 <- image_read(here("paper/image", "ioi_cv_1.png"), density = 300)
+```
+
+## Experiment 2
+```{r variability-2, include = FALSE}
+source(here("experiment-2/analysis/stats", "variability.R"), chdir = TRUE)
+```
+
+### Articulation
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_cv_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_cv_norm$p.value))}`).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "articulation"]$IQR)`] and performing condition [*Mdn* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$Median)`, *IQR* = `r sprintf('%.3f', ioi_cv[Condition == "performing" & Skill == "articulation"]$IQR)`] while playing the piece with the notated articulation (*p* `r if(ioi_art_cv_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_cv_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_cv$effsize, 2)`, two-tailed, *Fig \ref{fig:cv}*).
+
+### Dynamics
+
+A paired-sample *t*-test showed that the tempo variability of performance did not differ significantly between the teaching [*M* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$SD)`] and the performing condition [*M* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_cv[Condition == "teaching" & Skill == "dynamics"]$SD)`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_cv_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_cv_ttest$statistic))`, *p* `r if(ioi_dyn_cv_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_cv_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn_cv$effsize), 3)`, two-tailed, *Fig \ref{fig:cv}*).
+
+```{r ioi-cv-2, echo = FALSE}
+ioi_cv_plot <- data.table(subject = ioi_cv_subject[Condition == "teaching"]$SubNr, teaching = ioi_cv_subject[Condition == "teaching"]$MeanCV, performing = ioi_cv_subject[Condition == "performing"]$MeanCV, skill = ioi_cv_subject[Condition == "teaching"]$Skill)
+
+ioi_cv_2 <- ggpaired(ioi_cv_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "CVs", title = "CVs: Exp 2") +
+  ylim(0, 0.20) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 0.08, method = "t.test", paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_cv_2.png"), plot = ioi_cv_2, dpi = 300, width = 7, height = 5)
+img_ioi_cv_2 <- image_read(here("paper/image", "ioi_cv_2.png"), density = 300)
+```
+
+```{r plot-cv, out.width = "100%", fig.cap = "\\label{fig:cv}Experiment 1, 2: CVs when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+img_ioi_cv <- image_append(c(img_ioi_cv_1, img_ioi_cv_2))
+img_ioi_cv <- image_scale(img_ioi_cv, "2200")
+image_write(img_ioi_cv, path = here("paper/figure", "Fig12.tiff"), format = "tiff")
+img_ioi_cv
+```
+
+\clearpage
+
+# 4. Temporal grouping
+
+In order to investigate if participants used temporal grouping to make boundaries between subcomponents (i.e., between legato and staccato, between forte and piano) clearer for teaching purposes, we looked at IOIs only at transition points.
+
+## Results
+## Experiment 1
+```{r variability-1-1, include = FALSE}
+source(here("experiment-1/analysis/stats", "variability.R"), chdir = TRUE)
+```
+
+### Articulation
+
+To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_art_tra_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_tra_norm$p.value))}`).
+The Wilcoxon Signed-rank test revealed that IOIs at transition points were larger in the teaching condition [*Mdn* = `r ioi_tra[Condition == "teaching" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_tra[Condition == "teaching" & Skill == "articulation"]$IQR`] than in the performing condition [*Mdn* = `r ioi_tra[Condition == "performing" & Skill == "articulation"]$Median` (ms), *IQR* = `r ioi_tra[Condition == "performing" & Skill == "articulation"]$IQR`] while playing the piece with the notated articulation (*p* `r if(ioi_art_tra_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_tra_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art_tra$effsize, 2)`, two-tailed, *Fig \ref{fig:tra}*).
+
+### Dynamics
+
+A paired-sample *t*-test showed that IOIs at transition points were larger in the teaching condition [*M* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$Mean` (ms), *SD* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$SD`] than in the performing condition [*M* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$Mean` (ms), *SD* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$SD`] while playing the piece with the notated dynamics (*t*(`r ioi_dyn_tra_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_dyn_tra_ttest$statistic))`, *p* `r if(ioi_dyn_tra_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_tra_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_dyn_tra$effsize), 3)`, two-tailed, *Fig \ref{fig:tra}*).
+
+```{r ioi-tra-1, echo = FALSE}
+ioi_tra_plot <- data.table(subject = ioi_tra_subject[Condition == "teaching"]$SubNr, teaching = ioi_tra_subject[Condition == "teaching"]$Mean, performing = ioi_tra_subject[Condition == "performing"]$Mean, skill = ioi_tra_subject[Condition == "teaching"]$Skill)
+
+ioi_tra_1 <- ggpaired(ioi_tra_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs at transition: Exp 1") +
+  geom_hline(yintercept = 188, linetype = "dashed") +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 150, method = "t.test", paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_tra_1.png"), plot = ioi_tra_1, dpi = 300, width = 7, height = 5)
+img_ioi_tra_1 <- image_read(here("paper/image", "ioi_tra_1.png"), density = 300)
+```
+
+## Experiment 2
+### Articulation
+```{r variability-2-2, include = FALSE}
+source(here("experiment-2/analysis/stats", "variability.R"), chdir = TRUE)
+```
+
+A paired-sample *t*-test showed that IOIs at transition points were larger in the teaching condition [*M* = `r sprintf('%.2f', ioi_tra[Condition == "teaching" & Skill == "articulation"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_tra[Condition == "teaching" & Skill == "articulation"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_tra[Condition == "performing" & Skill == "articulation"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_tra[Condition == "performing" & Skill == "articulation"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_tra_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_tra_ttest$statistic))`, *p* `r if(ioi_art_tra_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_art_tra_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art_tra$effsize), 3)`, two-tailed, *Fig \ref{fig:tra}*).
+
+### Dynamics
+
+To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_tra_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_tra_norm$p.value))}`).
+The Wilcoxon Signed-rank test revealed that IOIs at transition points did not differ significantly between the teaching [*Mdn* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$Median`, *IQR* = `r ioi_tra[Condition == "teaching" & Skill == "dynamics"]$IQR`] and the performing condition [*Mdn* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$Median`, *IQR* = `r ioi_tra[Condition == "performing" & Skill == "dynamics"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_tra_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_tra_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn_tra$effsize, 2)`, two-tailed, *Fig \ref{fig:tra}*).
+
+```{r ioi-tra-2, echo = FALSE}
+ioi_tra_plot <- data.table(subject = ioi_tra_subject[Condition == "teaching"]$SubNr, teaching = ioi_tra_subject[Condition == "teaching"]$Mean, performing = ioi_tra_subject[Condition == "performing"]$Mean, skill = ioi_tra_subject[Condition == "teaching"]$Skill)
+
+ioi_tra_2 <- ggpaired(ioi_tra_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "skill", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised IOIs", title = "IOIs at transition: Exp 2") +
+  geom_hline(yintercept = 1, linetype = "dashed") +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 1.05, method = "t.test", paired = TRUE) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_tra_2.png"), plot = ioi_tra_2, dpi = 300, width = 7, height = 5)
+img_ioi_tra_2 <- image_read(here("paper/image", "ioi_tra_2.png"), density = 300)
+```
+
+```{r plot-tra, out.width = "100%", fig.cap = "\\label{fig:tra}Experiment 1, 2: IOIs at transition points when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+img_ioi_tra <- image_append(c(img_ioi_tra_1, img_ioi_tra_2))
+img_ioi_tra <- image_scale(img_ioi_tra, "2200")
+image_write(img_ioi_tra, path = here("paper/figure", "Fig10.tiff"), format = "tiff")
+img_ioi_tra
+```
+
+\clearpage
+
+# 5. Additional analysis for dynamics difference
+
+In order to examine if the dynamics difference between forte and piano at transition points is larger when the difference between forte and piano in overall performance is considered, we normalised the KV difference at transition points by dividing it by the average range between forte and piano per performance (per trial).
+
+## Results
+## Experiment 1
+```{r additional-dynamics-1, include = FALSE}
+source(here("experiment-1/analysis/stats", "dynamics.R"), chdir = TRUE)
+```
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Transition Type (FtoP vs.
+PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig \ref{fig:vel-diff-2}*).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (*p* `r if(vel_diff_dyn_posthoc_2$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[1]))}`) and from piano to forte (*p* `r if(vel_diff_dyn_posthoc_2$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[2]))}`) between the teaching condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] and the performing condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
+
+```{r vel-diff-dyn-normalised-1, echo = FALSE}
+vel_diff_dyn_normalised_1 <- ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas",xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff: Exp 1") + ylim(-6, 5) + stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 3, method = "t.test", paired = TRUE) + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_diff_dyn_normalised_1.png"), plot = vel_diff_dyn_normalised_1, dpi = 300, width = 7, height = 5)
+img_vel_diff_dyn_normalised_1 <- image_read(here("paper/image", "vel_diff_dyn_normalised_1.png"), density = 300)
+```
+
+## Experiment 2
+```{r additional-dynamics-2, include = FALSE}
+source(here("experiment-2/analysis/stats", "dynamics.R"), chdir = TRUE)
+```
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.
+performing) and Transition Type (FtoP vs.
+PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`).
+However, there was no significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig \ref{fig:vel-diff-2}*).
+
+```{r vel-diff-dyn-normalised-2, echo = FALSE}
+vel_diff_dyn_normalised_2 <- ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "Normalised Difference", title = "KV-Diff: Exp 2") + ylim(-6, 5) + stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 3, method = "t.test", paired = TRUE) + theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "vel_diff_dyn_normalised_2.png"), plot = vel_diff_dyn_normalised_2, dpi = 300, width = 7, height = 5)
+img_vel_diff_dyn_normalised_2 <- image_read(here("paper/image", "vel_diff_dyn_normalised_2.png"), density = 300)
+```
+
+```{r plot-vel-diff-2, out.width = "100%", fig.cap = "\\label{fig:vel-diff-2}Experiment 1, 2: Normalised KV Difference at transition points when playing the piece with dynamics (left: Experiment 1, right: Experiment 2).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \\textit{p} < .05, ** \\textit{p} < .01, *** \\textit{p} < .001"}
+img_vel_diff_dyn_normalised <- image_append(c(img_vel_diff_dyn_normalised_1, img_vel_diff_dyn_normalised_2))
+img_vel_diff_dyn_normalised <- image_scale(img_vel_diff_dyn_normalised, "2200")
+image_write(img_vel_diff_dyn_normalised, path = here("paper/figure", "Fig11.tiff"), format = "tiff")
+img_vel_diff_dyn_normalised
+```

### 2022-09-01 paper/manuscript.tex
finalise the manuscript with some minor corrections
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}}).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}} 1).
 Participants played the piece 8 times per technique per condition, so there were 32 trials in total (2 conditions x 2 techniques x 8 trials).
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 369, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Dynamics (\emph{F}(1,30) = 6.83, \emph{p} = 0.014, \(\eta_G^2\) = 0.003, \emph{Fig \ref{fig:vel-1}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.31{]} than in the performing condition {[}\emph{M} = 81.65, \emph{SD} = 7.43{]} (\emph{p} = 0.068) when instructed to play forte.
-There was no significant difference in KV between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.08{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.41{]} (\emph{p} = 0.24) when instructed to play piano.
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Forte: \emph{M} = 82.78, \emph{SD} = 8.31, Piano: \emph{M} = 60.63, \emph{SD} = 5.08{]} and the performing condition {[}Forte: \emph{M} = 81.65, \emph{SD} = 7.43, Piano: \emph{M} = 60.99, \emph{SD} = 4.41{]} for each subcomponent (Forte: \emph{p} = 0.068, Piano: \emph{p} = 0.24).
 
@@
 First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Fig \ref{fig:stimuli-2} A}) prior to the experiment so that they had enough time to familiarise themselves with the piece and perform it without pitch errors while performing it with different notated expressions in the lab.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}}).
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}} 1).
 Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability\textsuperscript{19}.
@@
 It can be expected that teachers would also make didactic modulations in some domains other than acoustic properties, such as gestures if they are physically present in front of students.
-Moreover, given that teachers modulate their demonstration throughout ongoing interactions with learners\textsuperscript{20}, future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities.
+Moreover, given that teachers modulate their demonstration throughout ongoing interactions with learners\textsuperscript{20}, future studies are needed to investigate how experts dynamically adapt their performance to their students' skill levels and demonstrated abilities.
 Also, some of the participants in the current study did not have teaching experience at all and even those who had teaching experience had taught for only a few years.
 It would be important to investigate if our findings would be observed in participants with extensive experience teaching the piano.
+We summarised descriptive statistics depending on participants' teaching experience (see \emph{\protect\hyperlink{supplementary}{Supplementary material}} 2).
 
@@
 
+We quantified participants' sound modulations in terms of tempo, articulation and dynamics so as to examine if and how participants modulate their sound.
+However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see \emph{\protect\hyperlink{supplementary}{Supplementary material}} 3 and 4).
+For example, participants tended to slow down at transition points regardless of the techniques they were teaching although overall, we found that participants played slower only when teaching the piece with the notated articulation.
+Future research is needed to investigate how musicians make particular modulations in relation to musical structures.
+
 Musicians and music teachers tend to consider expressive skills as performers' most important skills\textsuperscript{21}.
@@
 The ultimate goal of expressive performance is considered to lie in developing performers' own interpretations of music and convey affects and emotions by using expressive tools such as articulation, dynamics, tempo and timing\textsuperscript{23}.
-The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings\textsuperscript{e.g., 15}.
-It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.
+The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings\textsuperscript{15,24}.
+Given that students often learn by listening to recordings\textsuperscript{25}, it is important to investigate how performance itself could affect students' learning processes.It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.
 
@@
 Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention.
-Rather, expert demonstrators seem to follow principles of relevance in communication\textsuperscript{24}, highlighting only those aspects they are intending to demonstrate.
+Rather, expert demonstrators seem to follow principles of relevance in communication\textsuperscript{26}, highlighting only those aspects they are intending to demonstrate.
 How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher's intentions explicitly adds to the learning success are important questions for future research.
@@
 \href{https://doi.org/10.1177/0748730418783651}{Musicians' {Natural Frequencies} of {Performance Display Optimal Temporal Stability}}.
-\emph{J Biol Rhythms} \textbf{33}, 432--440 (2018).}
+\emph{Journal of Biological Rhythms} \textbf{33}, 432--440 (2018).}
 
@@
 \href{https://doi.org/10.1038/s41598-019-44049-x}{Teacher-learner interaction quantifies scaffolding behaviour in imitation learning}.
-\emph{Sci Rep} \textbf{9}, 1--13 (2019).}
+\emph{Scientific Reports} \textbf{9}, 1--13 (2019).}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-colprit_2000}{}}%
+\CSLLeftMargin{24. }%
+\CSLRightInline{Colprit, E.
+J.
+\href{https://doi.org/10.2307/3345394}{Observation and {Analysis} of {Suzuki String Teaching}}.
+\emph{Journal of Research in Music Education} \textbf{48}, 206--221 (2000).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-volioti_2017}{}}%
+\CSLLeftMargin{25. }%
+\CSLRightInline{Volioti, G.
+\& Williamon, A.
+\href{https://doi.org/10.1177/1029864916674048}{Recordings as learning and practising resources for performance: {Exploring} attitudes and behaviours of music students and professionals}.
+\emph{Musicae Scientiae} \textbf{21}, 499--523 (2017).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-sperber_1995}{}}%
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{Sperber, D. \& Wilson, D. \emph{Relevance: {Communication} and cognition, 2nd ed}. ({Blackwell Publishing}, 1995).}
@@
 \hypertarget{supplementary}{%
-\section{Supplementary material}\label{supplementary}}
+\section{Supplementary Material}\label{supplementary}}
 
@@
 
-\hypertarget{temporal-variability}{%
-\section{2. Temporal Variability}\label{temporal-variability}}
-
-The tempo variability of performance was assessed with the coefficient of variation (CV) of the IOIs to investigate the tempo stability of performance between the teaching and performing condition.
-CVs were calculated by dividing the standard deviation by the mean of IOIs.
-
-\hypertarget{results-2}{%
-\subsection{Results}\label{results-2}}
+\hypertarget{descriptive-statistics-depending-on-participants-teaching-experience}{%
+\section{2. Descriptive statistics depending on participants' teaching experience}\label{descriptive-statistics-depending-on-participants-teaching-experience}}
+
+We summarised descriptive statistics for those had an experience in teaching the piano and those who did not. Due to the lack of statistical power, we did not perform any statistical tests.
 
@@
 \subsection{Experiment 1}\label{experiment-1-2}}
-
-\hypertarget{articulation-2}{%
-\subsubsection{Articulation}\label{articulation-2}}
-
-To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.007).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.064, \emph{IQR} = 0.024{]} and performing condition {[}\emph{Mdn} = 0.066, \emph{IQR} = 0.016{]} while playing the piece with the notated articulation (\emph{p} = 0.06, \emph{r} = 0.34, two-tailed, \emph{Fig \ref{fig:cv}}).
-
-\hypertarget{dynamics-2}{%
-\subsubsection{Dynamics}\label{dynamics-2}}
-
-To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.024).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.070, \emph{IQR} = 0.013{]} and performing condition {[}\emph{Mdn} = 0.074, \emph{IQR} = 0.026{]} while playing the piece with the notated dynamics (\emph{p} = 0.22, \emph{r} = 0.22, two-tailed, \emph{Fig \ref{fig:cv}}).
-
-\hypertarget{experiment-2-2}{%
-\subsection{Experiment 2}\label{experiment-2-2}}
-
-\hypertarget{articulation-3}{%
-\subsubsection{Articulation}\label{articulation-3}}
-
-To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.017).
-The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.044, \emph{IQR} = 0.012{]} and performing condition {[}\emph{Mdn} = 0.047, \emph{IQR} = 0.014{]} while playing the piece with the notated articulation (\emph{p} = 0.07, \emph{r} = 0.41, two-tailed, \emph{Fig \ref{fig:cv}}).
-
-\hypertarget{dynamics-3}{%
-\subsubsection{Dynamics}\label{dynamics-3}}
-
-A paired-sample \emph{t}-test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{M} = 0.048, \emph{SD} = 0.011{]} and the performing condition {[}\emph{M} = 0.048, \emph{SD} = 0.011{]} while playing the piece with the notated dynamics (\emph{t}(19) = 1.83, \emph{p} = 0.08, Cohen's \emph{d} = 0.41, two-tailed, \emph{Fig \ref{fig:cv}}).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-cv-1} \caption{\label{fig:cv}Experiment 1, 2: CVs when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-cv}
-\end{figure}
-
-\clearpage
-
-\hypertarget{temporal-grouping}{%
-\section{3. Temporal grouping}\label{temporal-grouping}}
-
-In order to investigate if participants used temporal grouping to make boundaries between subcomponents (i.e., between legato and staccato, between forte and piano) clearer for teaching purposes, we looked at IOIs only at transition points.
-
-\hypertarget{results-3}{%
-\subsection{Results}\label{results-3}}
-
-\hypertarget{experiment-1-3}{%
-\subsection{Experiment 1}\label{experiment-1-3}}
-
-\hypertarget{articulation-4}{%
-\subsubsection{Articulation}\label{articulation-4}}
-
-To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wilcoxon Signed-rank test revealed that IOIs at transition points were larger in the teaching condition {[}\emph{Mdn} = 195.75 (ms), \emph{IQR} = 25.75{]} than in the performing condition {[}\emph{Mdn} = 188.77 (ms), \emph{IQR} = 23.53{]} while playing the piece with the notated articulation (\emph{p} = 0.02, \emph{r} = 0.42, two-tailed, \emph{Fig \ref{fig:tra}}).
-
-\hypertarget{dynamics-4}{%
-\subsubsection{Dynamics}\label{dynamics-4}}
-
-A paired-sample \emph{t}-test showed that IOIs at transition points were larger in the teaching condition {[}\emph{M} = 200.90 (ms), \emph{SD} = 20.96{]} than in the performing condition {[}\emph{M} = 197.63 (ms), \emph{SD} = 19.30{]} while playing the piece with the notated dynamics (\emph{t}(30) = 2.10, \emph{p} = 0.04, Cohen's \emph{d} = 0.38, two-tailed, \emph{Fig \ref{fig:tra}}).
-
-\hypertarget{experiment-2-3}{%
-\subsection{Experiment 2}\label{experiment-2-3}}
-
-\hypertarget{articulation-5}{%
-\subsubsection{Articulation}\label{articulation-5}}
-
-A paired-sample \emph{t}-test showed that IOIs at transition points were larger in the teaching condition {[}\emph{M} = 1.00, \emph{SD} = 0.067{]} than in the performing condition {[}\emph{M} = 0.98, \emph{SD} = 0.049{]} while playing the piece with the notated articulation (\emph{t}(19) = 2.27, \emph{p} = 0.03, Cohen's \emph{d} = 0.51, two-tailed, \emph{Fig \ref{fig:tra}}).
-
-\hypertarget{dynamics-5}{%
-\subsubsection{Dynamics}\label{dynamics-5}}
-
-To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.004).
-The Wilcoxon Signed-rank test revealed that IOIs at transition points did not differ significantly between the teaching {[}\emph{Mdn} = 1.00, \emph{IQR} = 0.06{]} and the performing condition {[}\emph{Mdn} = 0.99, \emph{IQR} = 0.05{]} while playing the piece with the notated dynamics (\emph{p} = 0.60, \emph{r} = 0.13, two-tailed, \emph{Fig \ref{fig:tra}}).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-tra-1} \caption{\label{fig:tra}Experiment 1, 2: IOIs at transition points when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
-A dashed line represents the tempo given by a metronome.
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-tra}
-\end{figure}
-
-\clearpage
-
-\hypertarget{additional-analysis-for-dynamics-difference}{%
-\section{4. Additional analysis for dynamics difference}\label{additional-analysis-for-dynamics-difference}}
-
-In order to examine if the dynamics difference between forte and piano at transition points is larger when the difference between forte and piano in overall performance is considered, we normalised the KV difference at transition points by dividing it by the average range between forte and piano per performance (per trial).
-
-\hypertarget{results-4}{%
-\subsection{Results}\label{results-4}}
-
-\hypertarget{experiment-1-4}{%
-\subsection{Experiment 1}\label{experiment-1-4}}
-
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,30) = 741, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.88) and a significant interaction between Condition and Transition Type (\emph{F}(1,30) = 7.87, \emph{p} = 0.009, \(\eta_G^2\) = 0.009, \emph{Fig \ref{fig:vel-diff-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (\emph{p} = 0.085) and from piano to forte (\emph{p} = 0.075) between the teaching condition {[}FtoP: \emph{M} = -0.62, \emph{SD} = 0.24, PtoF: \emph{M} = 1.10, \emph{SD} = 0.34{]} and the performing condition {[}FtoP: \emph{M} = -0.56, \emph{SD} = 0.25, PtoF: \emph{M} = 1.03, \emph{SD} = 0.40{]}.
-
-\hypertarget{experiment-2-4}{%
-\subsection{Experiment 2}\label{experiment-2-4}}
-
-A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,19) = 42, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.63).
-However, there was no significant interaction between Condition and Transition Type (\emph{F}(1,19) = 1.10, \emph{p} = 0.306, \(\eta_G^2\) = 0.008, \emph{Fig \ref{fig:vel-diff-2}}).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Experiment 1, 2: Normalised KV Difference at transition points when playing the piece with dynamics (left: Experiment 1, right: Experiment 2).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
-Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff-2}
-\end{figure}
-
-\clearpage
-
-\hypertarget{descriptive-statistics-depending-on-participants-teaching-experience}{%
-\section{5. Descriptive statistics depending on participants' teaching experience}\label{descriptive-statistics-depending-on-participants-teaching-experience}}
-
-\hypertarget{experiment-1-5}{%
-\subsection{Experiment 1}\label{experiment-1-5}}
 
@@
 
-\caption{\label{tab:ioi-teaching-desc-1}Experiment 1:}
-
-\begin{tabular}{lllllllll}
+\caption{\label{tab:ioi-teaching-desc-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of IOIs (ms).}
+
+\begin{tabular}{lllllll}
 \toprule
-Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{IQR}\\
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
 \midrule
-Performing & Articulation & Yes & 10 & 181.27 & 10.99 & 3.47 & 177.37 & 14.91\\
-Teaching & Articulation & Yes & 10 & 184.72 & 13.02 & 4.12 & 179.21 & 16.76\\
-Performing & Articulation & No & 21 & 193.49 & 18.75 & 4.09 & 185.83 & 26.43\\
-Teaching & Articulation & No & 21 & 196.67 & 20.52 & 4.48 & 192.13 & 32.33\\
-Performing & Dynamics & Yes & 10 & 182.43 & 11.52 & 3.64 & 181.67 & 13.07\\
-Teaching & Dynamics & Yes & 10 & 187.15 & 11.73 & 3.71 & 187.91 & 15.67\\
-Performing & Dynamics & No & 21 & 191.48 & 18.56 & 4.05 & 188.17 & 23.11\\
-Teaching & Dynamics & No & 21 & 193.50 & 21.22 & 4.63 & 186.28 & 25.29\\
+Performing & Articulation & Yes & 10 & 181.27 & 10.99 & 3.47\\
+Teaching & Articulation & Yes & 10 & 184.72 & 13.02 & 4.12\\
+Performing & Articulation & No & 21 & 193.49 & 18.75 & 4.09\\
+Teaching & Articulation & No & 21 & 196.67 & 20.52 & 4.48\\
+Performing & Dynamics & Yes & 10 & 182.43 & 11.52 & 3.64\\
+Teaching & Dynamics & Yes & 10 & 187.15 & 11.73 & 3.71\\
+Performing & Dynamics & No & 21 & 191.48 & 18.56 & 4.05\\
+Teaching & Dynamics & No & 21 & 193.50 & 21.22 & 4.63\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} Note}
+\normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
+Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
+Teaching indicates if participants had an experience in teaching the piano.}
 \end{tablenotes}
@@
 
-\caption{\label{tab:art-teaching-desc-1}Experiment 1:}
+\caption{\label{tab:art-teaching-desc-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KOT (ms).}
 
@@
 
-\caption{\label{tab:dyn-teaching-desc-1}Experiment 1:}
+\caption{\label{tab:dyn-teaching-desc-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV (0-127).}
 
@@
 
-\caption{\label{tab:dyn-diff-teaching-desc-1}Experiment 1:}
+\caption{\label{tab:dyn-diff-teaching-desc-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points.}
 
@@
 
-\newpage
-
-\hypertarget{experiment-2-5}{%
-\subsection{Experiment 2}\label{experiment-2-5}}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-teaching-2-1} \caption{\label{fig:ioi-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of IOIs (ms) when playing the piece with either articulation (left) or dynamics (right).
+\clearpage
+
+\hypertarget{experiment-2-2}{%
+\subsection{Experiment 2}\label{experiment-2-2}}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-teaching-2-1} \caption{\label{fig:ioi-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of normalised IOIs when playing the piece with either articulation (left) or dynamics (right).
 A dashed line represents the tempo given by a metronome.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-2-1} \caption{\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-2-1} \caption{\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOR when playing the piece with articulation.
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-art-teaching-2}
@@
 
-\caption{\label{tab:ioi-teaching-desc-2}Experiment 2:}
-
-\begin{tabular}{lllllllll}
+\caption{\label{tab:ioi-teaching-desc-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of normalised IOIs.}
+
+\begin{tabular}{lllllll}
 \toprule
-Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{IQR}\\
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
 \midrule
-Performing & Articulation & Yes & 11 & 0.954 & 0.046 & 0.014 & 0.977 & 0.064\\
-Teaching & Articulation & Yes & 11 & 0.964 & 0.049 & 0.015 & 0.976 & 0.044\\
-Performing & Articulation & No & 9 & 0.951 & 0.035 & 0.012 & 0.950 & 0.041\\
-Teaching & Articulation & No & 9 & 0.967 & 0.052 & 0.017 & 0.973 & 0.014\\
-Performing & Dynamics & Yes & 11 & 0.953 & 0.046 & 0.014 & 0.960 & 0.050\\
-Teaching & Dynamics & Yes & 11 & 0.958 & 0.047 & 0.014 & 0.972 & 0.044\\
-Performing & Dynamics & No & 9 & 0.958 & 0.041 & 0.014 & 0.959 & 0.023\\
-Teaching & Dynamics & No & 9 & 0.955 & 0.052 & 0.017 & 0.954 & 0.026\\
+Performing & Articulation & Yes & 11 & 0.954 & 0.046 & 0.014\\
+Teaching & Articulation & Yes & 11 & 0.964 & 0.049 & 0.015\\
+Performing & Articulation & No & 9 & 0.951 & 0.035 & 0.012\\
+Teaching & Articulation & No & 9 & 0.967 & 0.052 & 0.017\\
+Performing & Dynamics & Yes & 11 & 0.953 & 0.046 & 0.014\\
+Teaching & Dynamics & Yes & 11 & 0.958 & 0.047 & 0.014\\
+Performing & Dynamics & No & 9 & 0.958 & 0.041 & 0.014\\
+Teaching & Dynamics & No & 9 & 0.955 & 0.052 & 0.017\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} Note}
+\normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
+Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
+Teaching indicates if participants had an experience in teaching the piano.}
 \end{tablenotes}
@@
 
-\caption{\label{tab:art-teaching-desc-2}Experiment 2:}
+\caption{\label{tab:art-teaching-desc-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KOR.}
 
@@
 
-\caption{\label{tab:dyn-teaching-desc-2}Experiment 2:}
+\caption{\label{tab:dyn-teaching-desc-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV (0-127).}
 
@@
 
-\caption{\label{tab:dyn-diff-teaching-desc-2}Experiment 2:}
+\caption{\label{tab:dyn-diff-teaching-desc-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points.}
 
@@
 
+\clearpage
+
+\hypertarget{temporal-variability}{%
+\section{3. Temporal Variability}\label{temporal-variability}}
+
+The tempo variability of performance was assessed with the coefficient of variation (CV) of the IOIs to investigate the tempo stability of performance between the teaching and performing condition.
+CVs were calculated by dividing the standard deviation by the mean of IOIs.
+
+\hypertarget{results-2}{%
+\subsection{Results}\label{results-2}}
+
+\hypertarget{experiment-1-3}{%
+\subsection{Experiment 1}\label{experiment-1-3}}
+
+\hypertarget{articulation-2}{%
+\subsubsection{Articulation}\label{articulation-2}}
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.007).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.064, \emph{IQR} = 0.024{]} and performing condition {[}\emph{Mdn} = 0.066, \emph{IQR} = 0.016{]} while playing the piece with the notated articulation (\emph{p} = 0.06, \emph{r} = 0.34, two-tailed, \emph{Fig \ref{fig:cv}}).
+
+\hypertarget{dynamics-2}{%
+\subsubsection{Dynamics}\label{dynamics-2}}
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.024).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.070, \emph{IQR} = 0.013{]} and performing condition {[}\emph{Mdn} = 0.074, \emph{IQR} = 0.026{]} while playing the piece with the notated dynamics (\emph{p} = 0.22, \emph{r} = 0.22, two-tailed, \emph{Fig \ref{fig:cv}}).
+
+\hypertarget{experiment-2-3}{%
+\subsection{Experiment 2}\label{experiment-2-3}}
+
+\hypertarget{articulation-3}{%
+\subsubsection{Articulation}\label{articulation-3}}
+
+To compare the mean CVs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.017).
+The Wilcoxon Signed-rank test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{Mdn} = 0.044, \emph{IQR} = 0.012{]} and performing condition {[}\emph{Mdn} = 0.047, \emph{IQR} = 0.014{]} while playing the piece with the notated articulation (\emph{p} = 0.07, \emph{r} = 0.41, two-tailed, \emph{Fig \ref{fig:cv}}).
+
+\hypertarget{dynamics-3}{%
+\subsubsection{Dynamics}\label{dynamics-3}}
+
+A paired-sample \emph{t}-test showed that the tempo variability of performance did not differ significantly between the teaching {[}\emph{M} = 0.048, \emph{SD} = 0.011{]} and the performing condition {[}\emph{M} = 0.048, \emph{SD} = 0.011{]} while playing the piece with the notated dynamics (\emph{t}(19) = 1.83, \emph{p} = 0.08, Cohen's \emph{d} = 0.41, two-tailed, \emph{Fig \ref{fig:cv}}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-cv-1} \caption{\label{fig:cv}Experiment 1, 2: CVs when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-cv}
+\end{figure}
+
+\clearpage
+
+\hypertarget{temporal-grouping}{%
+\section{4. Temporal grouping}\label{temporal-grouping}}
+
+In order to investigate if participants used temporal grouping to make boundaries between subcomponents (i.e., between legato and staccato, between forte and piano) clearer for teaching purposes, we looked at IOIs only at transition points.
+
+\hypertarget{results-3}{%
+\subsection{Results}\label{results-3}}
+
+\hypertarget{experiment-1-4}{%
+\subsection{Experiment 1}\label{experiment-1-4}}
+
+\hypertarget{articulation-4}{%
+\subsubsection{Articulation}\label{articulation-4}}
+
+To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
+The Wilcoxon Signed-rank test revealed that IOIs at transition points were larger in the teaching condition {[}\emph{Mdn} = 195.75 (ms), \emph{IQR} = 25.75{]} than in the performing condition {[}\emph{Mdn} = 188.77 (ms), \emph{IQR} = 23.53{]} while playing the piece with the notated articulation (\emph{p} = 0.02, \emph{r} = 0.42, two-tailed, \emph{Fig \ref{fig:tra}}).
+
+\hypertarget{dynamics-4}{%
+\subsubsection{Dynamics}\label{dynamics-4}}
+
+A paired-sample \emph{t}-test showed that IOIs at transition points were larger in the teaching condition {[}\emph{M} = 200.90 (ms), \emph{SD} = 20.96{]} than in the performing condition {[}\emph{M} = 197.63 (ms), \emph{SD} = 19.30{]} while playing the piece with the notated dynamics (\emph{t}(30) = 2.10, \emph{p} = 0.04, Cohen's \emph{d} = 0.38, two-tailed, \emph{Fig \ref{fig:tra}}).
+
+\hypertarget{experiment-2-4}{%
+\subsection{Experiment 2}\label{experiment-2-4}}
+
+\hypertarget{articulation-5}{%
+\subsubsection{Articulation}\label{articulation-5}}
+
+A paired-sample \emph{t}-test showed that IOIs at transition points were larger in the teaching condition {[}\emph{M} = 1.00, \emph{SD} = 0.067{]} than in the performing condition {[}\emph{M} = 0.98, \emph{SD} = 0.049{]} while playing the piece with the notated articulation (\emph{t}(19) = 2.27, \emph{p} = 0.03, Cohen's \emph{d} = 0.51, two-tailed, \emph{Fig \ref{fig:tra}}).
+
+\hypertarget{dynamics-5}{%
+\subsubsection{Dynamics}\label{dynamics-5}}
+
+To compare the mean IOIs at transition points between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} = 0.004).
+The Wilcoxon Signed-rank test revealed that IOIs at transition points did not differ significantly between the teaching {[}\emph{Mdn} = 1.00, \emph{IQR} = 0.06{]} and the performing condition {[}\emph{Mdn} = 0.99, \emph{IQR} = 0.05{]} while playing the piece with the notated dynamics (\emph{p} = 0.60, \emph{r} = 0.13, two-tailed, \emph{Fig \ref{fig:tra}}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-tra-1} \caption{\label{fig:tra}Experiment 1, 2: IOIs at transition points when playing the piece with either articulation or dynamics (left: Experiment 1, right: Experiment 2).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-tra}
+\end{figure}
+
+\clearpage
+
+\hypertarget{additional-analysis-for-dynamics-difference}{%
+\section{5. Additional analysis for dynamics difference}\label{additional-analysis-for-dynamics-difference}}
+
+In order to examine if the dynamics difference between forte and piano at transition points is larger when the difference between forte and piano in overall performance is considered, we normalised the KV difference at transition points by dividing it by the average range between forte and piano per performance (per trial).
+
+\hypertarget{results-4}{%
+\subsection{Results}\label{results-4}}
+
+\hypertarget{experiment-1-5}{%
+\subsection{Experiment 1}\label{experiment-1-5}}
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,30) = 741, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.88) and a significant interaction between Condition and Transition Type (\emph{F}(1,30) = 7.87, \emph{p} = 0.009, \(\eta_G^2\) = 0.009, \emph{Fig \ref{fig:vel-diff-2}}).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (\emph{p} = 0.085) and from piano to forte (\emph{p} = 0.075) between the teaching condition {[}FtoP: \emph{M} = -0.62, \emph{SD} = 0.24, PtoF: \emph{M} = 1.10, \emph{SD} = 0.34{]} and the performing condition {[}FtoP: \emph{M} = -0.56, \emph{SD} = 0.25, PtoF: \emph{M} = 1.03, \emph{SD} = 0.40{]}.
+
+\hypertarget{experiment-2-5}{%
+\subsection{Experiment 2}\label{experiment-2-5}}
+
+A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,19) = 42, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.63).
+However, there was no significant interaction between Condition and Transition Type (\emph{F}(1,19) = 1.10, \emph{p} = 0.306, \(\eta_G^2\) = 0.008, \emph{Fig \ref{fig:vel-diff-2}}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Experiment 1, 2: Normalised KV Difference at transition points when playing the piece with dynamics (left: Experiment 1, right: Experiment 2).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.
+Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff-2}
+\end{figure}
+
 \end{document}

### 2022-09-01 paper/manuscript.Rmd
correct typo
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
-Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
+Technique indicates performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
 Teaching indicates if participants had an experience in teaching the piano."
@@
   , note = "N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
-Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
+Technique indicates performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
 Teaching indicates if participants had an experience in teaching the piano."

### 2022-09-01 paper/manuscript.tex
correct typo
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
-Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
+Technique indicates performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
 Teaching indicates if participants had an experience in teaching the piano.}
@@
 \normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively.
-Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
+Technique indicates performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
 Teaching indicates if participants had an experience in teaching the piano.}

### 2022-09-02 paper/manuscript.Rmd
finalise the manuscript
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title             : "The sound of teaching music: Expert pianists' performance modulations for novices"
-shorttitle        : "The sound of teaching music"
+title             : "Expert pianists make specific exaggerations for teaching"
+shorttitle        : "Expert pianists make specific exaggerations for teaching"
 
@@
 The aim of this study was to investigate how expert pianists modulate their performance to teach techniques necessary to implement musical expressions.
-Overall, the findings of the two experiments showed that pianists performed one and the same piece differently depending on whether they played with the intention to teach expressive techniques or with the intention to perform the piece for an audience.
+Overall, the findings of the two experiments showed that pianists performed one and the same piece differently depending on whether they played with the intention to teach techniques or with the intention to perform the piece for an audience.
 When playing with the intention to teach, pianists selectively highlighted relevant aspects of artistic expressions.
@@
 
-We summarised descriptive statistics for those had an experience in teaching the piano and those who did not. Due to the lack of statistical power, we did not perform any statistical tests.
+We summarised descriptive statistics for those with and without teaching experience in piano. Due to the lack of statistical power, we did not perform any statistical tests.
 
@@
 
-```{r plot-art-teaching-1, out.width = "100%", fig.cap = "\\label{fig:art-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation.
+```{r plot-art-teaching-1, out.width = "100%", fig.cap = "\\label{fig:art-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation (left: legato, right: staccato).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
@@
 
-```{r plot-dyn-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics.
+```{r plot-dyn-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics (left: forte, right: piano).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
@@
 
-```{r plot-dyn-diff-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics.
+```{r plot-dyn-diff-teaching-1, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics  (left: forte to piano, right: piano to forte).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
@@
 colnames(ioi_all)[2] <- "Technique"
+colnames(ioi_all)[3] <- "TeachingExp"
 ioi_all[Condition == "teaching"]$Condition <- "Teaching"
@@
 ioi_all[Technique == "dynamics"]$Technique <- "Dynamics"
+ioi_all$TeachingExp <- factor(ioi_all$TeachingExp, levels = c("Yes", "No"))
+setorder(ioi_all, "Technique", "TeachingExp")
 
@@
 Technique indicates performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
-Teaching indicates if participants had an experience in teaching the piano."
+TeachingExp indicates if participants had an experience in teaching the piano."
   , escape = TRUE
@@
 colnames(kot_art_all)[2] <- "Technique"
+colnames(kot_art_all)[4] <- "TeachingExp"
 kot_art_all[Condition == "teaching"]$Condition <- "Teaching"
@@
 kot_art_all[Technique == "articulation"]$Technique <- "Articulation"
+kot_art_all$TeachingExp <- factor(kot_art_all$TeachingExp, levels = c("Yes", "No"))
+setorder(kot_art_all, "Technique", "TeachingExp", "Subcomponent")
 
@@
 colnames(vel_dyn_all)[2] <- "Technique"
+colnames(vel_dyn_all)[4] <- "TeachingExp"
 vel_dyn_all[Condition == "teaching"]$Condition <- "Teaching"
@@
 vel_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"
+vel_dyn_all$TeachingExp <- factor(vel_dyn_all$TeachingExp, levels = c("Yes", "No"))
+setorder(vel_dyn_all, "Technique", "TeachingExp", "Subcomponent")
 
@@
 colnames(vel_diff_dyn_all)[2] <- "Technique"
+colnames(vel_diff_dyn_all)[4] <- "TeachingExp"
 vel_diff_dyn_all[Condition == "teaching"]$Condition <- "Teaching"
@@
 vel_diff_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"
+vel_diff_dyn_all$TeachingExp <- factor(vel_diff_dyn_all$TeachingExp, levels = c("Yes", "No"))
+setorder(vel_diff_dyn_all, "Technique", "TeachingExp", "Subcomponent")
 
@@
 
-ioi_art_teaching_2 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Articulation") +
+ioi_art_teaching_2 <- ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOIs: Articulation") +
   ylim(0.8, 1.1) +
@@
 
-ioi_dyn_teaching_2 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOIs: Dynamics") +
+ioi_dyn_teaching_2 <- ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", panel.labs = list(teachingexp = c("Teaching: Yes", "Teaching: No")), palette = "aaas", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised IOIs", title = "IOIs: Dynamics") +
   ylim(0.8, 1.1) +
@@
 
-```{r plot-art-teaching-2, out.width = "100%", fig.cap = "\\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOR when playing the piece with articulation.
+```{r plot-art-teaching-2, out.width = "100%", fig.cap = "\\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOR when playing the piece with articulation (left: legato, right: staccato).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
@@
 
-```{r plot-dyn-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics.
+```{r plot-dyn-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics  (left: forte, right: piano).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
@@
 
-```{r plot-dyn-diff-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics.
+```{r plot-dyn-diff-teaching-2, out.width = "100%", fig.cap = "\\label{fig:dyn-diff-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics (left: forte to piano, right: piano to forte).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
@@
 ioi_all[Technique == "dynamics"]$Technique <- "Dynamics"
-ioi_all$Teaching <- factor(ioi_all$Teaching, levels = c("Yes", "No"))
-setorder(ioi_all, "Technique", "Teaching")
+ioi_all$TeachingExp <- factor(ioi_all$TeachingExp, levels = c("Yes", "No"))
+setorder(ioi_all, "Technique", "TeachingExp")
 
@@
 Technique indicates performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
-Teaching indicates if participants had an experience in teaching the piano."
+TeachingExp indicates if participants had an experience in teaching the piano."
   , escape = TRUE
@@
 colnames(kor_art_all)[2] <- "Technique"
+colnames(kor_art_all)[4] <- "TeachingExp"
 kor_art_all[Condition == "teaching"]$Condition <- "Teaching"
@@
 kor_art_all[Technique == "articulation"]$Technique <- "Articulation"
-kor_art_all$Teaching <- factor(kor_art_all$Teaching, levels = c("Yes", "No"))
-setorder(kor_art_all, "Technique", "Teaching")
+kor_art_all$TeachingExp <- factor(kor_art_all$TeachingExp, levels = c("Yes", "No"))
+setorder(kor_art_all, "Technique", "TeachingExp", "Subcomponent")
 
@@
 vel_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"
-vel_dyn_all$Teaching <- factor(vel_dyn_all$Teaching, levels = c("Yes", "No"))
-setorder(vel_dyn_all, "Technique", "Teaching")
+vel_dyn_all$TeachingExp <- factor(vel_dyn_all$TeachingExp, levels = c("Yes", "No"))
+setorder(vel_dyn_all, "Technique", "TeachingExp", "Subcomponent")
 
@@
 vel_diff_dyn_all[Technique == "dynamics"]$Technique <- "Dynamics"
-vel_diff_dyn_all$Teaching <- factor(vel_diff_dyn_all$Teaching, levels = c("Yes", "No"))
-setorder(vel_diff_dyn_all, "Technique", "Teaching")
+vel_diff_dyn_all$TeachingExp <- factor(vel_diff_dyn_all$TeachingExp, levels = c("Yes", "No"))
+setorder(vel_diff_dyn_all, "Technique", "TeachingExp", "Subcomponent")
 
@@
   geom_hline(yintercept = 188, linetype = "dashed") +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 150, method = "t.test", paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 230, method = "t.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`).
-However, there was no significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig \ref{fig:vel-diff-2}*).
+However, there was no significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig \ref{fig:vel-diff-2}*).
 

### 2022-09-02 paper/manuscript.tex
finalise the manuscript
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \hypersetup{
-  pdftitle={The sound of teaching music: Expert pianists' performance modulations for novices},
+  pdftitle={Expert pianists make specific exaggerations for teaching},
   pdfauthor={Atsuko Tominaga1,*, Günther Knoblich1, \& Natalie Sebanz1},
@@
 
-\title{The sound of teaching music: Expert pianists' performance modulations for novices}
+\title{Expert pianists make specific exaggerations for teaching}
 \author{Atsuko Tominaga\textsuperscript{1,*}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
@@
 
-\shorttitle{The sound of teaching music}
+\shorttitle{Expert pianists make specific exaggerations for teaching}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-1-1} \caption{\label{fig:art-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-1-1} \caption{\label{fig:art-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation (left: legato, right: staccato).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-art-teaching-1}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-teaching-1-1} \caption{\label{fig:dyn-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-teaching-1-1} \caption{\label{fig:dyn-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics (left: forte, right: piano).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-dyn-teaching-1}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-diff-teaching-1-1} \caption{\label{fig:dyn-diff-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-diff-teaching-1-1} \caption{\label{fig:dyn-diff-teaching-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics  (left: forte to piano, right: piano to forte).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-dyn-diff-teaching-1}
@@
 \toprule
-Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{TeachingExp} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
 \midrule
@@
 Technique indicates performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
-Teaching indicates if participants had an experience in teaching the piano.}
+TeachingExp indicates if participants had an experience in teaching the piano.}
 \end{tablenotes}
@@
 \toprule
-Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{TeachingExp} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
 \midrule
 Performing & Articulation & Legato & Yes & 10 & 21.57 & 16.18 & 5.12\\
+Teaching & Articulation & Legato & Yes & 10 & 25.43 & 13.11 & 4.15\\
 Performing & Articulation & Staccato & Yes & 10 & -119.66 & 22.84 & 7.22\\
-Teaching & Articulation & Legato & Yes & 10 & 25.43 & 13.11 & 4.15\\
 Teaching & Articulation & Staccato & Yes & 10 & -127.39 & 15.95 & 5.04\\
 Performing & Articulation & Legato & No & 21 & 10.40 & 16.57 & 3.62\\
+Teaching & Articulation & Legato & No & 21 & 11.36 & 14.71 & 3.21\\
 Performing & Articulation & Staccato & No & 21 & -132.48 & 19.03 & 4.15\\
-Teaching & Articulation & Legato & No & 21 & 11.36 & 14.71 & 3.21\\
 Teaching & Articulation & Staccato & No & 21 & -136.84 & 21.52 & 4.70\\
@@
 \toprule
-Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{TeachingExp} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
 \midrule
 Performing & Dynamics & Forte & Yes & 10 & 85.20 & 7.76 & 2.45\\
+Teaching & Dynamics & Forte & Yes & 10 & 86.96 & 8.76 & 2.77\\
 Performing & Dynamics & Piano & Yes & 10 & 60.27 & 4.57 & 1.44\\
-Teaching & Dynamics & Forte & Yes & 10 & 86.96 & 8.76 & 2.77\\
 Teaching & Dynamics & Piano & Yes & 10 & 60.11 & 5.93 & 1.87\\
 Performing & Dynamics & Forte & No & 21 & 79.97 & 6.82 & 1.49\\
+Teaching & Dynamics & Forte & No & 21 & 80.80 & 7.49 & 1.63\\
 Performing & Dynamics & Piano & No & 21 & 61.33 & 4.40 & 0.96\\
-Teaching & Dynamics & Forte & No & 21 & 80.80 & 7.49 & 1.63\\
 Teaching & Dynamics & Piano & No & 21 & 60.88 & 4.76 & 1.04\\
@@
 \toprule
-Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{TeachingExp} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
 \midrule
 Performing & Dynamics & FtoP & Yes & 10 & -14.63 & 5.79 & 1.83\\
+Teaching & Dynamics & FtoP & Yes & 10 & -18.81 & 7.26 & 2.30\\
 Performing & Dynamics & PtoF & Yes & 10 & 24.32 & 9.96 & 3.15\\
-Teaching & Dynamics & FtoP & Yes & 10 & -18.81 & 7.26 & 2.30\\
 Teaching & Dynamics & PtoF & Yes & 10 & 29.07 & 8.78 & 2.78\\
 Performing & Dynamics & FtoP & No & 21 & -10.30 & 5.69 & 1.24\\
+Teaching & Dynamics & FtoP & No & 21 & -11.87 & 6.09 & 1.33\\
 Performing & Dynamics & PtoF & No & 21 & 19.23 & 7.71 & 1.68\\
-Teaching & Dynamics & FtoP & No & 21 & -11.87 & 6.09 & 1.33\\
 Teaching & Dynamics & PtoF & No & 21 & 21.98 & 9.56 & 2.09\\
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-2-1} \caption{\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOR when playing the piece with articulation.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-2-1} \caption{\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOR when playing the piece with articulation (left: legato, right: staccato).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-art-teaching-2}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-teaching-2-1} \caption{\label{fig:dyn-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-teaching-2-1} \caption{\label{fig:dyn-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV (0-127) when playing the piece with dynamics  (left: forte, right: piano).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-dyn-teaching-2}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-diff-teaching-2-1} \caption{\label{fig:dyn-diff-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics.
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-dyn-diff-teaching-2-1} \caption{\label{fig:dyn-diff-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points when playing the piece with dynamics (left: forte to piano, right: piano to forte).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-dyn-diff-teaching-2}
@@
 Technique indicates performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
-Teaching indicates if participants had an experience in teaching the piano.}
+TeachingExp indicates if participants had an experience in teaching the piano.}
 \end{tablenotes}
@@
 \toprule
-Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{TeachingExp} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
 \midrule
 Performing & Articulation & Legato & Yes & 11 & 0.115 & 0.074 & 0.022\\
+Teaching & Articulation & Legato & Yes & 11 & 0.122 & 0.064 & 0.019\\
 Performing & Articulation & Staccato & Yes & 11 & -0.755 & 0.034 & 0.010\\
-Teaching & Articulation & Legato & Yes & 11 & 0.122 & 0.064 & 0.019\\
 Teaching & Articulation & Staccato & Yes & 11 & -0.776 & 0.034 & 0.010\\
 Performing & Articulation & Legato & No & 9 & 0.120 & 0.071 & 0.024\\
+Teaching & Articulation & Legato & No & 9 & 0.120 & 0.081 & 0.027\\
 Performing & Articulation & Staccato & No & 9 & -0.728 & 0.050 & 0.017\\
-Teaching & Articulation & Legato & No & 9 & 0.120 & 0.081 & 0.027\\
 Teaching & Articulation & Staccato & No & 9 & -0.749 & 0.073 & 0.024\\
@@
 \toprule
-Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{TeachingExp} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
 \midrule
 Performing & Dynamics & Forte & Yes & 11 & 81.06 & 6.52 & 1.97\\
+Teaching & Dynamics & Forte & Yes & 11 & 83.32 & 6.48 & 1.95\\
 Performing & Dynamics & Piano & Yes & 11 & 62.49 & 5.55 & 1.67\\
-Teaching & Dynamics & Forte & Yes & 11 & 83.32 & 6.48 & 1.95\\
 Teaching & Dynamics & Piano & Yes & 11 & 59.50 & 3.59 & 1.08\\
 Performing & Dynamics & Forte & No & 9 & 82.02 & 10.15 & 3.38\\
+Teaching & Dynamics & Forte & No & 9 & 83.06 & 11.54 & 3.85\\
 Performing & Dynamics & Piano & No & 9 & 62.36 & 4.11 & 1.37\\
-Teaching & Dynamics & Forte & No & 9 & 83.06 & 11.54 & 3.85\\
 Teaching & Dynamics & Piano & No & 9 & 61.81 & 3.76 & 1.25\\
@@
 \toprule
-Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
+Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Subcomponent} & \multicolumn{1}{c}{TeachingExp} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
 \midrule
 Performing & Dynamics & FtoP & Yes & 11 & -14.43 & 6.20 & 1.87\\
+Teaching & Dynamics & FtoP & Yes & 11 & -19.27 & 4.33 & 1.30\\
 Performing & Dynamics & PtoF & Yes & 11 & 14.96 & 6.78 & 2.04\\
-Teaching & Dynamics & FtoP & Yes & 11 & -19.27 & 4.33 & 1.30\\
 Teaching & Dynamics & PtoF & Yes & 11 & 20.77 & 5.71 & 1.72\\
 Performing & Dynamics & FtoP & No & 9 & -13.64 & 9.46 & 3.15\\
+Teaching & Dynamics & FtoP & No & 9 & -14.84 & 11.08 & 3.69\\
 Performing & Dynamics & PtoF & No & 9 & 17.44 & 8.11 & 2.70\\
-Teaching & Dynamics & FtoP & No & 9 & -14.84 & 11.08 & 3.69\\
 Teaching & Dynamics & PtoF & No & 9 & 20.66 & 12.09 & 4.03\\
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,19) = 42, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.63).
-However, there was no significant interaction between Condition and Transition Type (\emph{F}(1,19) = 1.10, \emph{p} = 0.306, \(\eta_G^2\) = 0.008, \emph{Fig \ref{fig:vel-diff-2}}).
+However, there was no significant interaction between Condition and Transition Type (\emph{F}(1,19) = 1.10, \emph{p} = 0.31, \(\eta_G^2\) = 0.008, \emph{Fig \ref{fig:vel-diff-2}}).
 

### 2022-09-13 paper/manuscript.Rmd
correct some typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 We found consistent results across a simple musical scale (Experiment 1) and a more naturalistic piece of music (Experiment 2).
-These findings show that teaching-specific action modulations generalize to expressive actions and suggest that action modulations serve to highlight the most relevant aspects of the actions to be learnt.
+These findings show that teaching-specific action modulations generalise to expressive actions and suggest that action modulations serve to highlight the most relevant aspects of the actions to be learnt.
 
@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplementary material](#supplementary)* 1).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in *[Supplementary Material](#supplementary)* 1).
 Participants played the piece 8 times per technique per condition, so there were 32 trials in total (2 conditions x 2 techniques x 8 trials).
@@
 First, all participants were required to memorise the piece without expressive notations (i.e., *Fig \ref{fig:stimuli-2} A*) prior to the experiment so that they had enough time to familiarise themselves with the piece and perform it without pitch errors while performing it with different notated expressions in the lab.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in *[Supplementary material](#supplementary)* 1).
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in *[Supplementary Material](#supplementary)* 1).
 Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability [@zamm_2018].
@@
 It would be important to investigate if our findings would be observed in participants with extensive experience teaching the piano.
-We summarised descriptive statistics depending on participants' teaching experience (see *[Supplementary material](#supplementary)* 2).
+We summarised descriptive statistics depending on participants’ teaching experience (see *[Supplementary Material](#supplementary)* 2).
 
@@
 We quantified participants’ sound modulations in terms of tempo, articulation and dynamics so as to examine if and how participants modulate their sound.
-However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see *[Supplementary material](#supplementary)* 3 and 4).
+However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see *[Supplementary Material](#supplementary)* 3 and 4).
 For example, participants tended to slow down at transition points regardless of the techniques they were teaching although overall, we found that participants played slower only when teaching the piece with the notated articulation.

### 2022-09-13 paper/manuscript.tex
correct some typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 We found consistent results across a simple musical scale (Experiment 1) and a more naturalistic piece of music (Experiment 2).
-These findings show that teaching-specific action modulations generalize to expressive actions and suggest that action modulations serve to highlight the most relevant aspects of the actions to be learnt.
+These findings show that teaching-specific action modulations generalise to expressive actions and suggest that action modulations serve to highlight the most relevant aspects of the actions to be learnt.
 }
@@
 It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance.
-In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}} 1).
+In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary Material}} 1).
 Participants played the piece 8 times per technique per condition, so there were 32 trials in total (2 conditions x 2 techniques x 8 trials).
@@
 First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Fig \ref{fig:stimuli-2} A}) prior to the experiment so that they had enough time to familiarise themselves with the piece and perform it without pitch errors while performing it with different notated expressions in the lab.
-Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}} 1).
+Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in \emph{\protect\hyperlink{supplementary}{Supplementary Material}} 1).
 Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability\textsuperscript{19}.
@@
 It would be important to investigate if our findings would be observed in participants with extensive experience teaching the piano.
-We summarised descriptive statistics depending on participants' teaching experience (see \emph{\protect\hyperlink{supplementary}{Supplementary material}} 2).
+We summarised descriptive statistics depending on participants' teaching experience (see \emph{\protect\hyperlink{supplementary}{Supplementary Material}} 2).
 
@@
 We quantified participants' sound modulations in terms of tempo, articulation and dynamics so as to examine if and how participants modulate their sound.
-However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see \emph{\protect\hyperlink{supplementary}{Supplementary material}} 3 and 4).
+However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see \emph{\protect\hyperlink{supplementary}{Supplementary Material}} 3 and 4).
 For example, participants tended to slow down at transition points regardless of the techniques they were teaching although overall, we found that participants played slower only when teaching the piece with the notated articulation.

### 2022-09-26 experiment-1/analysis/demographics/questionnaire.Rmd
removed unused packages
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-# plot
-if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 ```

### 2022-12-12 paper/manuscript.Rmd
finalised manuscript (published)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 These modulations include slowing down and exaggerating sounds and actions.
-Similar findings were obtained in studies with adult learners, where exaggerations were observed when native British English speakers were talking to first language English learners [@uther_2007] and when skilled adults were teaching xylophone melodies to novices [@mcellin_2017].
+Similar findings were obtained in studies with adult learners, where exaggerations were observed when native British English speakers were talking to second language English learners [@uther_2007] and when skilled adults were teaching xylophone melodies to novices [@mcellin_2017].
 The observed modulations are thought to provide communicative signals that can facilitate learning by affecting attention and memory [@csibra_2009].
@@
 They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`).
-`r nrow(dt_included[TeachingPiano == "Yes"])` participants had an experience in teaching the piano (*M* = `r round( mean(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)` years, *SD* = `r round( sd(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`).
+`r nrow(dt_included[TeachingPiano == "Yes"])` participants had experience in teaching the piano (*M* = `r round( mean(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)` years, *SD* = `r round( sd(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`).
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
@@
 The excerpt was taken from ``A Dozen a Day - Play with Ease in Many Keys" by Edna-Mae Burnam and modified for the experiment.
-It consisted of a 6-measure isochronous melody noted in 4/4 meter.
+It consisted of a 6-measure isochronous melody noted in a 4/4 metre.
 The stimulus was composed in C major to be played with the right hand only.
@@
 To compare the mean IOIs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired *t*-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (*p* `r if(ioi_dyn_norm$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_dyn_norm$p.value))}`).
-The Wilcoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and the performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`, two-tailed, *Fig \ref{fig:ioi-1}*).
+The Wilcoxon Signed-rank test revealed no significant difference between the teaching condition [*Mdn* = `r ioi_dyn_all[Condition == "teaching"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "teaching"]$IQR`] and performing condition [*Mdn* = `r ioi_dyn_all[Condition == "performing"]$Median` (ms), *IQR* = `r ioi_dyn_all[Condition == "performing"]$IQR`] while playing the piece with the notated dynamics (*p* `r if(ioi_dyn_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_dyn_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_dyn$effsize, 2)`, two-tailed, *Fig \ref{fig:ioi-1}*).
 
@@
 piano) showed that there was a significant main effect of Dynamics (*F*(`r vel_dyn_aov$anova_table[2,1]`,`r vel_dyn_aov$anova_table[2,2]`) = `r sprintf('%.0f', vel_dyn_aov$anova_table[2,4])`, *p* `r if(vel_dyn_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_dyn_aov$anova_table[2,5])`) and a significant interaction between Condition and Dynamics (*F*(`r vel_dyn_aov$anova_table[3,1]`,`r vel_dyn_aov$anova_table[3,2]`) = `r sprintf('%.2f', vel_dyn_aov$anova_table[3,4])`, *p* `r if(vel_dyn_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_dyn_aov$anova_table[3,5])`, *Fig \ref{fig:vel-1}*).
-However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and the performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] for each subcomponent (Forte: *p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`, Piano: *p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching [Forte: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "teaching" & Subcomponent == "Piano"]$SD`] and performing condition [Forte: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Forte"]$SD`, Piano: *M* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$Mean`, *SD* = `r vel_dyn_all[Condition == "performing" & Subcomponent == "Piano"]$SD`] for each subcomponent (Forte: *p* `r if(vel_dyn_posthoc$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_dyn_posthoc$p.value[1]))}`, Piano: *p* `r if(vel_dyn_posthoc$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', vel_dyn_posthoc$p.value[2]))}`).
 
@@
 The results for KOT and KV showed that participants successfully highlighted relevant aspects of articulation and dynamics.
-Specifically, participants exaggerated staccato when teaching articulation and forte when teaching dynamics.
+Specifically, participants exaggerated staccato when teaching articulation.
 Moreover, our exploratory analysis demonstrated that participants made a larger contrast in dynamics at transition points (i.e., forte to piano or piano to forte).
@@
 They had `r round(mean(dt_2_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_2_included$PianoTotalPractice), 2)`).
-`r nrow(dt_2_included[TeachingPiano == "Yes"])` participants had an experience in teaching the piano (*M* = `r round( mean(dt_2_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)` years, *SD* = `r round( sd(dt_2_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`).
+`r nrow(dt_2_included[TeachingPiano == "Yes"])` participants had experience in teaching the piano (*M* = `r round( mean(dt_2_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)` years, *SD* = `r round( sd(dt_2_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`).
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
@@
 The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
-The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
+The modified piece consisted of a 12-measure isochronous melody notated in a 4/4 metre to be played with the right hand only.
 Original sheet music was used for the purpose of practice (*Fig \ref{fig:stimuli-2} A*).
@@
 Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in *[Supplementary Material](#supplementary)* 1).
-Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability [@zamm_2018].
+Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence shows that musicians have their own preferred tempo, which affects their temporal variability [@zamm_2018].
 The chosen tempo was again cued by a leading metronome.
@@
 While teaching dynamics, KV showed that participants exaggerated forte and piano.
-This pattern of exaggeration was more pronounced than in Experiment 1, where only higher KV for forte was observed.
+This pattern of exaggeration was more pronounced than in Experiment 1, where post-hoc comparions did not find significant differences between the two conditions for each subcomponent.
 Furthermore, we again found that when teaching dynamics, at transition points participants produced a larger contrast in dynamics between forte and piano, bidirectionally.
@@
 We quantified participants’ sound modulations in terms of tempo, articulation and dynamics so as to examine if and how participants modulate their sound.
-However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see *[Supplementary Material](#supplementary)* 3 and 4).
+However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see *[Supplementary Materials](#supplementary)* 3 and 4).
 For example, participants tended to slow down at transition points regardless of the techniques they were teaching although overall, we found that participants played slower only when teaching the piece with the notated articulation.
@@
 Musicians and music teachers tend to consider expressive skills as performers’ most important skills [@lindstrom_2003].
-In the present study, experts focused on the teaching of expressions that were notated.
+In the present study, experts focused on the teaching of specific expressions that were notated.
 However, expression also has other facets [@juslin_2003] that are not only piece-related (e.g., notated expressions) but performer- or context-related (e.g., performer’s interpretation, listening contexts such as recordings or concerts).
 The ultimate goal of expressive performance is considered to lie in developing performers’ own interpretations of music and convey affects and emotions by using expressive tools such as articulation, dynamics, tempo and timing [@meissner_2021].
-The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings [@woody_2000; @colprit_2000].
-Given that students often learn by listening to recordings [@volioti_2017], it is important to investigate how performance itself could affect students’ learning processes.It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.
+The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actual teaching settings [@woody_2000; @colprit_2000].
+Given that students often learn by listening to recordings [@volioti_2017], it is important to investigate how performance itself could affect students’ learning processes.
+It would also be interesting to examine how teachers’ playing (i.e., demonstrations) relates to the verbal instructions they give.
 
@@
 Rather, expert demonstrators seem to follow principles of relevance in communication [@sperber_1995], highlighting only those aspects they are intending to demonstrate.
-How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher's intentions explicitly adds to the learning success are important questions for future research.
+How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher’s intentions explicitly adds to the learning success are important questions for future research.
 
@@
 
+\clearpage
+
 ## Experiment 1
@@
 Technique indicates performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
-TeachingExp indicates if participants had an experience in teaching the piano."
+TeachingExp indicates if participants had experience in teaching the piano."
   , escape = TRUE
@@
 Technique indicates performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
-TeachingExp indicates if participants had an experience in teaching the piano."
+TeachingExp indicates if participants had experience in teaching the piano."
   , escape = TRUE
@@
 PtoF) showed that there was a significant main effect of Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[2,1]`,`r vel_diff_dyn_aov_3$anova_table[2,2]`) = `r sprintf('%.0f', vel_diff_dyn_aov_3$anova_table[2,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[2,5])`) and a significant interaction between Condition and Transition Type (*F*(`r vel_diff_dyn_aov_3$anova_table[3,1]`,`r vel_diff_dyn_aov_3$anova_table[3,2]`) = `r sprintf('%.2f', vel_diff_dyn_aov_3$anova_table[3,4])`, *p* `r if(vel_diff_dyn_aov_3$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', vel_diff_dyn_aov_3$anova_table[3,5])`, *Fig \ref{fig:vel-diff-2}*).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (*p* `r if(vel_diff_dyn_posthoc_2$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[1]))}`) and from piano to forte (*p* `r if(vel_diff_dyn_posthoc_2$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[2]))}`) between the teaching condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] and the performing condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (*p* `r if(vel_diff_dyn_posthoc_2$p.value[1] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[1]))}`) and from piano to forte (*p* `r if(vel_diff_dyn_posthoc_2$p.value[2] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', vel_diff_dyn_posthoc_2$p.value[2]))}`) between the teaching condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "teaching" & Subcomponent == "PtoF"]$SD`] and performing condition [FtoP: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "FtoP"]$SD`, PtoF: *M* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$Mean`, *SD* = `r vel_diff_dyn_normalised_all[Condition == "performing" & Subcomponent == "PtoF"]$SD`].
 

### 2022-12-12 paper/manuscript.tex
finalised manuscript (published)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 These modulations include slowing down and exaggerating sounds and actions.
-Similar findings were obtained in studies with adult learners, where exaggerations were observed when native British English speakers were talking to first language English learners\textsuperscript{5} and when skilled adults were teaching xylophone melodies to novices\textsuperscript{6}.
+Similar findings were obtained in studies with adult learners, where exaggerations were observed when native British English speakers were talking to second language English learners\textsuperscript{5} and when skilled adults were teaching xylophone melodies to novices\textsuperscript{6}.
 The observed modulations are thought to provide communicative signals that can facilitate learning by affecting attention and memory\textsuperscript{7}.
@@
 They had 12.45 years of practice on average (\emph{SD} = 5.66).
-10 participants had an experience in teaching the piano (\emph{M} = 3.48 years, \emph{SD} = 3.51).
+10 participants had experience in teaching the piano (\emph{M} = 3.48 years, \emph{SD} = 3.51).
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
@@
 The excerpt was taken from ``A Dozen a Day - Play with Ease in Many Keys'' by Edna-Mae Burnam and modified for the experiment.
-It consisted of a 6-measure isochronous melody noted in 4/4 meter.
+It consisted of a 6-measure isochronous melody noted in a 4/4 metre.
 The stimulus was composed in C major to be played with the right hand only.
@@
 To compare the mean IOIs between the teaching and performing condition, we conducted a Wilcoxon Signed-rank test, instead of a paired \emph{t}-test, because a Shapiro-Wilk test showed that the distribution of the mean difference was significantly different from the normal distribution (\emph{p} \textless{} 0.001).
-The Wilcoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.28 (ms), \emph{IQR} = 21.83{]} and the performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.59{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
+The Wilcoxon Signed-rank test revealed no significant difference between the teaching condition {[}\emph{Mdn} = 186.28 (ms), \emph{IQR} = 21.83{]} and performing condition {[}\emph{Mdn} = 186.27 (ms), \emph{IQR} = 19.59{]} while playing the piece with the notated dynamics (\emph{p} = 0.11, \emph{r} = 0.29, two-tailed, \emph{Fig \ref{fig:ioi-1}}).
 
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 369, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Dynamics (\emph{F}(1,30) = 6.83, \emph{p} = 0.014, \(\eta_G^2\) = 0.003, \emph{Fig \ref{fig:vel-1}}).
-However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Forte: \emph{M} = 82.78, \emph{SD} = 8.31, Piano: \emph{M} = 60.63, \emph{SD} = 5.08{]} and the performing condition {[}Forte: \emph{M} = 81.65, \emph{SD} = 7.43, Piano: \emph{M} = 60.99, \emph{SD} = 4.41{]} for each subcomponent (Forte: \emph{p} = 0.068, Piano: \emph{p} = 0.24).
+However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Forte: \emph{M} = 82.78, \emph{SD} = 8.31, Piano: \emph{M} = 60.63, \emph{SD} = 5.08{]} and performing condition {[}Forte: \emph{M} = 81.65, \emph{SD} = 7.43, Piano: \emph{M} = 60.99, \emph{SD} = 4.41{]} for each subcomponent (Forte: \emph{p} = 0.068, Piano: \emph{p} = 0.24).
 
@@
 They had 15.65 years of practice on average (\emph{SD} = 5.67).
-11 participants had an experience in teaching the piano (\emph{M} = 4.14 years, \emph{SD} = 3.65).
+11 participants had experience in teaching the piano (\emph{M} = 4.14 years, \emph{SD} = 3.65).
 All participants gave their informed consent before the experiment started and received vouchers for their participation.
@@
 The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
-The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
+The modified piece consisted of a 12-measure isochronous melody notated in a 4/4 metre to be played with the right hand only.
 Original sheet music was used for the purpose of practice (\emph{Fig \ref{fig:stimuli-2} A}).
@@
 Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in \emph{\protect\hyperlink{supplementary}{Supplementary Material}} 1).
-Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability\textsuperscript{19}.
+Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence shows that musicians have their own preferred tempo, which affects their temporal variability\textsuperscript{19}.
 The chosen tempo was again cued by a leading metronome.
@@
 We quantified participants' sound modulations in terms of tempo, articulation and dynamics so as to examine if and how participants modulate their sound.
-However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see \emph{\protect\hyperlink{supplementary}{Supplementary Material}} 3 and 4).
+However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see \emph{\protect\hyperlink{supplementary}{Supplementary Materials}} 3 and 4).
 For example, participants tended to slow down at transition points regardless of the techniques they were teaching although overall, we found that participants played slower only when teaching the piece with the notated articulation.
@@
 Musicians and music teachers tend to consider expressive skills as performers' most important skills\textsuperscript{21}.
-In the present study, experts focused on the teaching of expressions that were notated.
+In the present study, experts focused on the teaching of specific expressions that were notated.
 However, expression also has other facets\textsuperscript{22} that are not only piece-related (e.g., notated expressions) but performer- or context-related (e.g., performer's interpretation, listening contexts such as recordings or concerts).
 The ultimate goal of expressive performance is considered to lie in developing performers' own interpretations of music and convey affects and emotions by using expressive tools such as articulation, dynamics, tempo and timing\textsuperscript{23}.
-The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings\textsuperscript{15,24}.
-Given that students often learn by listening to recordings\textsuperscript{25}, it is important to investigate how performance itself could affect students' learning processes.It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.
+The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actual teaching settings\textsuperscript{15,24}.
+Given that students often learn by listening to recordings\textsuperscript{25}, it is important to investigate how performance itself could affect students' learning processes.
+It would also be interesting to examine how teachers' playing (i.e., demonstrations) relates to the verbal instructions they give.
 
@@
 \href{https://doi.org/10.1177/0748730418783651}{Musicians' {Natural Frequencies} of {Performance Display Optimal Temporal Stability}}.
-\emph{Journal of Biological Rhythms} \textbf{33}, 432--440 (2018).}
+\emph{J Biol Rhythms} \textbf{33}, 432--440 (2018).}
 
@@
 \href{https://doi.org/10.1038/s41598-019-44049-x}{Teacher-learner interaction quantifies scaffolding behaviour in imitation learning}.
-\emph{Scientific Reports} \textbf{9}, 1--13 (2019).}
+\emph{Sci Rep} \textbf{9}, 1--13 (2019).}
 
@@
 We summarised descriptive statistics for those with and without teaching experience in piano. Due to the lack of statistical power, we did not perform any statistical tests.
+
+\clearpage
 
@@
 Technique indicates performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C).
-TeachingExp indicates if participants had an experience in teaching the piano.}
+TeachingExp indicates if participants had experience in teaching the piano.}
 \end{tablenotes}
@@
 Technique indicates performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C).
-TeachingExp indicates if participants had an experience in teaching the piano.}
+TeachingExp indicates if participants had experience in teaching the piano.}
 \end{tablenotes}
@@
 A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Transition Type (FtoP vs.~PtoF) showed that there was a significant main effect of Transition Type (\emph{F}(1,30) = 741, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.88) and a significant interaction between Condition and Transition Type (\emph{F}(1,30) = 7.87, \emph{p} = 0.009, \(\eta_G^2\) = 0.009, \emph{Fig \ref{fig:vel-diff-2}}).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (\emph{p} = 0.085) and from piano to forte (\emph{p} = 0.075) between the teaching condition {[}FtoP: \emph{M} = -0.62, \emph{SD} = 0.24, PtoF: \emph{M} = 1.10, \emph{SD} = 0.34{]} and the performing condition {[}FtoP: \emph{M} = -0.56, \emph{SD} = 0.25, PtoF: \emph{M} = 1.03, \emph{SD} = 0.40{]}.
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was no significant difference when changing from forte to piano (\emph{p} = 0.085) and from piano to forte (\emph{p} = 0.075) between the teaching condition {[}FtoP: \emph{M} = -0.62, \emph{SD} = 0.24, PtoF: \emph{M} = 1.10, \emph{SD} = 0.34{]} and performing condition {[}FtoP: \emph{M} = -0.56, \emph{SD} = 0.25, PtoF: \emph{M} = 1.03, \emph{SD} = 0.40{]}.
 

### 2022-12-12 experiment-1/analysis/demographics/questionnaire.R

@@ -2,8 +2,6 @@
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-# plot
-if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
 ## ----file, include = FALSE----------------------------------------------------

### 2023-01-07 paper/manuscript.Rmd
update citation style for docx file
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
remove one space
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-#output: papaja::apa6_word
+# documentclass     : "apa6" # for word
+# classoption       : "man" # for word
+# output: papaja::apa6_word # for word
 output: 
   papaja::apa6_pdf:
-    latex_engine: xelatex
+    latex_engine: xelatex # for pdf
 ---
@@
 
-A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`
-, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art$effsize), 3)`, two-tailed, *Fig \ref{fig:ioi-2}*).
+A paired-sample *t*-test showed that participants played more slowly in the teaching condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "teaching"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "teaching"]$SD)`] than in the performing condition [*M* = `r sprintf('%.2f', ioi_art_all[Condition == "performing"]$Mean)`, *SD* = `r sprintf('%.3f', ioi_art_all[Condition == "performing"]$SD)`] while playing the piece with the notated articulation (*t*(`r ioi_art_ttest$parameter`) = `r sprintf('%.2f', abs(ioi_art_ttest$statistic))`, *p* `r if(ioi_art_ttest$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_ttest$p.value))}`, Cohen's *d* = `r round(abs(coheneff_art$effsize), 3)`, two-tailed, *Fig \ref{fig:ioi-2}*).
 A Wilcoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (*p* `r if(ioi_art_wilcoxon$p.value < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_art_wilcoxon$p.value))}`, *r* = `r round(wilcoxeff_art$effsize, 2)`, two-tailed).

### 2023-01-07 paper/manuscript.tex
remove one space
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.97, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.95, \emph{SD} = 0.040{]} while playing the piece with the notated articulation (\emph{t}(19) = 2.47
-, \emph{p} = 0.023, Cohen's \emph{d} = 0.55, two-tailed, \emph{Fig \ref{fig:ioi-2}}).
+A paired-sample \emph{t}-test showed that participants played more slowly in the teaching condition {[}\emph{M} = 0.97, \emph{SD} = 0.049{]} than in the performing condition {[}\emph{M} = 0.95, \emph{SD} = 0.040{]} while playing the piece with the notated articulation (\emph{t}(19) = 2.47, \emph{p} = 0.023, Cohen's \emph{d} = 0.55, two-tailed, \emph{Fig \ref{fig:ioi-2}}).
 A Wilcoxon Signed-rank test also confirmed that there was a significant difference between the two conditions (\emph{p} = 0.033, \emph{r} = 0.48, two-tailed).

### 2023-01-31 paper/manuscript.Rmd
supplementary figure correction (wilcox.test)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# documentclass     : "apa6" # for word
-# classoption       : "man" # for word
+documentclass     : "apa6"
+classoption       : "man"
 # output: papaja::apa6_word # for word
@@
   geom_hline(yintercept = 188, linetype = "dashed") +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 230, method = "t.test", paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 230, method = "wilcox.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   geom_hline(yintercept = 1, linetype = "dashed") +
-  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 1.05, method = "t.test", paired = TRUE) +
+  stat_compare_means(label = "p.signif", hide.ns = TRUE, size = 10, label.x = 1.5, label.y = 1.05, method = "wilcox.test", paired = TRUE) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
